### 2014-05-12 Manuscript/MissingData_Evol-SVP_abstract.tex
Abstract

\documentclass[a4paper,11pt]{article}
\usepackage[osf]{mathpazo} % palatino

\title{\large{Combining living and fossil taxa into phylogenies: the missing data issue}\vspace{-2ex}}
\author{\large{Thomas Guillerme and Natalie Cooper}\vspace{-2ex}}
\date{\vspace{-3ex}}

\begin{document}

\raggedright
 \pagestyle{empty}
\pagenumbering{gobble}

\maketitle

\begin{abstract}
%Context and bigger question
Living species represent less than 1\% of all species that have ever lived. Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology. This fact has led to an increasing consensus among scientists that both fossil and living taxa must be included in macroevolutionary studies. One approach, the Total Evidence Method, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both fossil and living taxa at the tips. Although the Total Evidence Method seems very promising, it requires a lot of data and  is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.

%Study and results
In this study we assess the effect of missing data on tree topologies inferred from total evidence supermatrices. Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the supermatrix: (1) the proportion of living taxa with no morphological data, (2) the amount of missing data in the fossil record and (3) the overall number of morphological characters in the supermatrix. We find that, in a Bayesian framework, difficulties in recovering a stable topology are mainly driven by the missing data in the molecular part of the matrix (for which fossil taxa have no data). In a Maximum Likelihood framework, however, topology is not directly affected by missing data \textit{per se}, but by the number of morphological characters shared among the taxa. Therefore, the two main drivers of incorrect topologies are the overall number of morphological characters and the number of living species with no morphological data.

%Prespectives and implications
Our results suggest that, in order to use total evidence methods, one should reduce the missing data in the morphological part of the supermatrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process. We apply our method to a comprehensive data set of both living and fossil primates. We find that using this integrative method modifies previous estimates of rates of body mass evolution within primates.

%1902/2200 characters
\end{abstract}
\end{document}

### 2014-05-12 Manuscript/MissingData_postgrad.symp._abstract.tex
Abstract

\documentclass{article}
%called preambule, \documentclass[options]{command}, \usepackage, etc...
%example \documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage{mathtools} %equationss
\usepackage{enumerate} %bullet points
%using packages: \usepackage[options]{name}

\begin{document}
\title{Combining living and fossil taxa into phylogenies: the missing data issue}
\date{Abstract}

\maketitle

Living species represent less than 1\% of all species that have ever lived. Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology. This fact has led to an increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies. One approach, known as the otal evidence method, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies. Although this approach seems very promising, it requires a lot of data. In particular it requires morphological data from both living and fossil taxa, both of which are scarce. Therefore, this approach is likely to suffer from having lots of missing data which may affect its ability to infer correct phylogenies.

Here we assess the effect of missing data on tree topologies inferred from total evidence supermatrices. Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the supermatrix: (1) the proportion of living taxa with no morphological data, (2) the amount of missing data in the fossil taxa and (3) the overall number of morphological characters for all of the taxa.

\end{document}

### 2014-05-12 Manuscript/TEM_draft.tex
Draft

%HEADER
\documentclass[12pt,letterpaper]{article}

%packages
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage[english]{babel}
\usepackage{pifont}
\usepackage{color}
\usepackage{setspace}
\usepackage{lscape}
\usepackage{indentfirst}
\usepackage[normalem]{ulem}
\usepackage{booktabs}
%\usepackage{nag}
\usepackage{natbib}
%\usepackage{bibtex}
\usepackage{float}
\usepackage{latexsym}
%\usepackage{hyperref} 
\usepackage{url}
%\usepackage{html}
\usepackage{hyperref}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
%\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{amstext} % Add and remove packages as necessary for your manuscript.

%Pagination style and stuff
\linespread{1.66} % All text should be double-spaced with occasional exceptions for tables. 
\raggedright
\setlength{\parindent}{0.5in}

\setcounter{secnumdepth}{0} % Our sections are not numbered and our papers do not have Tables of Contents. We don't present a list of figures or list of tables, either. Any common font is fine. (A common sans-serif font should be used on figures, but figures should be separate from the LaTeX document.)

\pagestyle{empty}

\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}

\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}

\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}

\renewcommand{\tableofcontents}{}

\bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib

%START
\begin{document}

%Runing head
\begin{flushright}
Version dated: \today
\end{flushright}
\bigskip
\noindent RH: Missing data in Total Evidence Supermatrices %RH = Running Head

\bigskip
\medskip
\begin{center}

%Title
\noindent{\Large \bf Effect of missing data in supermatrices containing living and fossil taxa on topological accuracy}
\bigskip

% We don't use a special title page; the author information is entered like any other text.
% FOOTNOTES: We don't allow them in the manuscript, except in tables. Don't include any footnotes in the text.

%Authors
\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$, other authors $^3$, and Natalie Cooper$^1$$^,$$^2$}\\
\noindent {\small \it 
$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Republic of Ireland;\\
$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Republic of Ireland;\\
$^3$Somewhere else }\\
\end{center}
\medskip
\noindent{\bf Corresponding author:} Thomas Guillerme, School of Natural Sciences, Trinity College Dublin, Dublin 2, Republic of Ireland; E-mail: guillert@tcd.ie.\\
\vspace{1in}

%Abstract
\subsubsection{Abstract}
Living species represent less than 1\% of all species that have ever lived. Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology. This fact has led to an increasing consensus among scientists that both fossil and living taxa must be included in macroevolutionary studies. One approach, the Total Evidence Method, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both fossil and living taxa at the tips. Although the Total Evidence Method seems very promising, it requires a lot of data and is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.

In this study we assess the effect of missing data on tree topologies inferred from total evidence supermatrices. Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the supermatrix: (1) the proportion of living taxa with no morphological data, (2) the amount of missing data in the fossil record and (3) the overall number of morphological characters in the supermatrix. We find that, in a Bayesian framework, difficulties in recovering a stable topology are mainly driven by the missing data in the molecular part of the matrix (for which fossil taxa have no data). In a Maximum Likelihood framework, however, topology is not directly affected by missing data \textit{per se}, but by the number of morphological characters shared among the taxa. Therefore, the two main drivers of incorrect topologies are the overall number of morphological characters and the number of living species with no morphological data.

Our results suggest that, in order to use total evidence methods, one should reduce the missing data in the morphological part of the supermatrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process. We apply our method to a comprehensive data set of both living and fossil primates. We find that using this integrative method modifies previous estimates of rates of body mass evolution within primates.

\noindent (Keywords: bla, bla, bla)\\

\vspace{1.5in}

\newpage

%We don't use a heading for the introduction. The nature of this section is implied by its position following the abstract and preceding the first major section heading.
Living species represent less than 1\% of all species that have ever lived \citep{novacek1992ext,raup1993extinction}. However, the majority of macroevolutionary studies focus solely on living species \citep{cooperwhat2009,meredithimpacts2011}.
Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as species richness gradients, e.g. extant clade diversity may ignore past diversification patterns \citep{shoshani1996proboscidea}; biogeographical history, e.g. extant biogeographical patterns are improved when integrating fossil data \citep{metcalfintegrating2014}; or paleoecology, e.g. extant clades niches might have differed greatly through time \citep{youngthe2010}. %Rewrite that part
These factors have led to increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}. However, to do this we need to be able to place living and fossil taxa into the same phylogenies which still remains complex.

Three main approaches have been used for combining fossil and living taxa data in phylogenies. These approaches differ in whether they treat fossil taxa as nodes or tips and how much of the available data is used (i.e. age only or age and morphology). Classical cladistic methods use morphological data from both fossil and living taxa and treat each taxon as a tip \citep{simpson1945}. This approach is commonly used by paleontologists but it ignores
the majority of %You proposed to remove that. The point here was to not get in trouble with modern paleontologist that feel "clever" using DNA as a binary character like "Presence/absence of this SINE/that microsatellite"…
the additional molecular data available from living species. Neontologists, on the other hand, more commonly use only molecular data from living species to build phylogenies. Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these analyses and only their occurrence age is used to time calibrate phylogenies \citep{zuckerkandl1965}. There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} %Should I develop this improvements here? In the end this improvements are more on the inference methods than on the type of data used which is more the point here.
as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013}. A final class of methods, known as total evidence methods use molecular data from living taxa and morphological data from both living and fossil taxa, and treats every taxon as a tip on the phylogeny \citep{eernissetaxonomic1993}.Here we focus on these total evidence methods because they have been recently
successfully developed and applied to empirical data %needs a better justification
\citep{pyrondivergence2011,ronquista2012,schragocombining2013}. Although the total evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data. In particular it requires morphological data form both living and fossil taxa, both of which are scarce. Therefore total evidence approaches are likely to suffer from having lots of missing data which may affect their ability to infer correct phyogenies. % Trevor: Missing data also influences support values that trend to decrease

The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013}. Missing molecular data has been seen by some authors as an issue because it can decrease the phylogenetic signal in some parts of the tree, especially when using large supermatrices \citep{lemmonthe2009}. However other authors do not see missing molecular data as a major issue because the phylogenetic signal is more likely to increase by having at least a 
"modest" % explain
number of highly covered genes (~50\% - \citet{rouresite-specific2011}), a higher number of taxa (especially slowly evolving taxa or taxa close to the outgroup) and by choosing more adequate models of sequence evolution rather than by reducing the amount of missing data \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}. Similarly, missing morphological data might be seen as major or minor issue for accurately infering phylogenies \citep{wiensmissing2003,sansomfossilization2013}. Because soft-tissues characters are rarely preserved in the fossil record, missing data is found in soft tissues, i.e. it is not randomly distributed, which can lead to biased placement of fossil taxa in phylogenies \citep{sansomfossilization2013}. However, the phylogenetic signal is not related to the level of missing data per se but to the number of informative characters per taxa, therefore missing data is less an issue than the number of shared informative characters \citep{wiensmissing2003}.
Although not a major problem separately \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011}, missing data in molecular and morphological supermatrices may become an issue when combined for example in a total evidence type supermatrices and no attempt has been made to study the impact of this issue until now. %Does not flow with the rest

Here we assess the effect of missing data on tree topology infered from total evidence supermatrices. The molecular part of a total evidence supermatrix contains no fossil taxa so will act like a classical molecular supermatrix \citep{ronquista2012}. The effect of missing data on such matrices is well known, therefore, we only focus on the missing data issue in the morphological part of the supermatrix. Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the supermatrix: (1) the proportion of living taxa with no morphological data, (2) the amount of missing data in the fossil taxa (i.e. the preservation quality of the fossil record) and (3) the overall number of morphological characters for both living and fossil taxa in the supermatrix. We assess how changing the values of these three parameters affects the topology of total evidence method phylogenies.

%We found that… This as an implication for…

\section{Methods} %no material?
To explore the effect of missing data on total evidence trees topologies we used the following protocol (note that we explain each step in detail below this general outline -Fig. ~\ref{Fig_Outline}):
\begin{enumerate}
\item{Generating the matrix}
\label{step:generate_matrix}
We built a randomly generated birth death tree to infer a "complete" matrix containing both molecular and morphological data for living and fossil taxa.
\item{Removing data}
\label{step:remove_data}
We removed data from the "complete" matrix to simulate the effects of missing data by modifying three parameters (1) the proportion of missing living taxa ($M_{L}$), (2) the proportion of missing data in the fossil taxa ($M_{F}$) and (3) the proportion of missing morphological characters ($M_{C}$).
\item{Building phylogenies}
\label{step:build_phylo}
We infered Bayesian phylogenetic trees from the "complete" matrix and from the matrices containing missing data.
\item{Comparing topologies}
\label{step:compare_topo}
We then compared the trees obtained from the matrices containing missing data to the ones obtained from the "complete" matrix to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on the topologies of the phylogenies we estimated.
\end{enumerate}
To measure the effect of missing data’s distribution, we repeated steps ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} with the exact same fixed parameters 50 times.

\begin{figure}
\centering
\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_outline.pdf}
\caption{Protocol outline. (1) We generated a random tree (hereafter called the "true" tree) to infer a matrix with no missing data (hereafter, the "complete" matrix). (2) We removed data fron the morphological partition of the "Complete" matrix resulting in 124 matrices with missing data. (3) We build Bayesian phylogenetic trees from each matrix. (4) We then compared the trees with missing data (from the matrices with missing data) to the tree with no missing data (hereafter, the "best" tree – infered from the "complete" matrix). We repeated step 1 to 4 50 times.}
\label{Fig_Outline}
\end{figure}

\subsection{Generating the matrix}
First we randomly generated a "true" tree of 50 taxa in R v3.0.2 \citep{R302} using the package diversitree v0.9-6 \citep{fitzjohndiversitree2012}. We generated the tree using a Birth Death process by sampling the values of the speciation events ($\lambda$) and extinction events ($\mu$) from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}. We implemented a rejection sampling algorithm to select only random trees with 25 living and 25 fossil taxa. We then added a species to the resulting Birth-Death tree as the outgroup of the tree. The mean branch length of the tree was used to separate the outgroup from the rest of the taxa and the branch length leading to the outgroup was set as the sum of the mean branch length and the longest root-to-tip length of the tree.

Next, we created a molecular and a morphological matrix from the "true" tree. The molecular matrix was infered from the "true" tree using the package phyclust v0.1-14 \citep{chen2011}. The matrix was made of 1000 characters sites for 51 taxa and generated using the seqgen algorithm \citep{ ranbaut1997seqgen}. We used the HKY model \citep{HKY85} with a random base frequencies and with the transition/transversion rate of 2 \citep{douadycomparison2003} as parameters for generating the matrix. The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \citep{yangamong-site1996}. We chose a low value of $\alpha$ to lower the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal. These parameters were selected to generate data with no special assumption about how the characters evolved as well as to reduce the computational time required if these parameters were estimated rather than defined (65 CPU years).

We infered the morphological matrix using the ape package v3.0-8 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa. We assigned the number of character states (either 2 or 3) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters. These probabilities were selected using the overall distribution of characters states extracted from 100 published empirical morphological matrices (See supplementaries). %link missing
We then ran an independent discrete character simulation for each character with the randomly selected number of states (2 or 3) and assuming an equal rate of change (i.e. evolutionary rate) from one character state to an other. This method allows us to have only two parameters per character i.e. the number of states and the evolutionary rate. For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.  All the molecular information for fossil taxa was replaced by missing data ("?"). Finally, we combined the morphological and molecular matrices obtained from the "true" tree. Hereafter we call this the "complete" matrix, i.e. the matrix with no missing data except for the molecular data of the fossil taxa.

\subsection{Removing data}
Once we obtained the "complete" matrix we modified it to get a set of matrices with missing data. We randomly replaced data with “?” in the morphological part of the matrices according to the following parameters (Fig. ~\ref{Fig_RemoveData}):

\begin{enumerate}
\item{The proportion of living taxa with no morphological data ($M_{L}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological one. Because of the increasing facility to sequence DNA for living species, the number of living species with molecular data is highly superior the the number of species with molecular and morphological data.
\item{The proportion of missing morphological data across all fossil taxa ($M_{F}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
This parameter illustrates the quality of the fossil record. 
\item{The proportion of missing morphological characters across all taxa (living and fossil - $M_{C}$): 0\%, 10\%, 25\%, 50\% or 75\%. }
This parameter illustrates the number of available morphological characters for both living and fossil taxa.
\end{enumerate}

\begin{figure}
\centering
\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_missingData.pdf}
\caption{Data removal protocol: Missing living - The proportion of living taxa with no morphological data ($M_{L}$); Missing fossil - The proportion of missing morphological data across all fossil taxa ($M_{F}$); Missing character - The proportion of missing morphological characters across all taxa (living and fossil) ($M_{C}$).}
\label{Fig_RemoveData}
\end{figure}

In practice, each parameter represent a way of removing data in the morphological part of the matrix: $M_{L}$ removes a proportion of rows from the living taxa; $M_{F}$ removes a proportion of cells from the fossil taxa; and $M_{C}$ removes a proportion of columns across both living and fossil taxa. Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the supermatrix affected: for $M_{L}$, all the morphological data of a proportion of the living taxa is removed (i.e. removing rows), as for $M_{F}$, a proportion of data is removed across the whole of the morphological matrix for fossil taxa (i.e. removing cells).

We tested all parameters combinations resulting in 125 ($5^3$) matrices. Because some parameter combinations introduce a lot of missing (e.g. $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all. When this occurred we repeated the random deletion of characters until every species had at least 5\% data.

\subsection{Building phylogenies}
From the resulting matrices we generated two types of trees, the "best" tree that is infered from the "complete" matrix and the trees with missing data infered from the 125 matrices with various amounts of missing data. The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations. The "best" tree, on the other hand, is the best tree we can build using the state-of-the-art phylogenetic methods. In real world situations, the "true" tree is never available to us because we cannot know the true evolutionnary history of a clade (except in very rare circumstances, e.g. \citet{rozen2005}). Therefore, here we focus on comparing the trees infered from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.

\subsubsection{Maximum Likelihood}
The "best" tree and the "missing-data" trees were infered using RAxML v7.0.4 \citep{stamatakisthe2008}. We used the GTR + $\Gamma_4$ model (\citet{tavare1986} – default GTRGAMMA in RAxML v7.04) as a generalisation of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data. The GTR model can be seen as a generalisation of the HKY model (the 2 parameters from the HKY model are implicitly included in the 6 from GTR model - \citet{stamatakisa2008}). We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. %Mk model seems implemented (i.e. Aprils’s slideshare http://www.slideshare.net/wrightaprilm/evolution2013). But how? Default method for standard data type (i.e. non DNA/AA). Trevor: swapping algorithm? salanin 2003

\subsubsection{Bayesian}
The "best" tree and the "missing-data" trees were infered using MrBayes v3.2.2 \citep{Ronquist2012mrbayes}. We partitioned the data treating the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition. The molecular evolutionary history was infered using the HKY model with a transition/transversion ratio \citep{douadycomparison2003} of two and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$). For the morphological data, we used the Markov \textit{k} state model \citep{lewisa2001} wich is a generalisation of the JC69 model \citep{jc69} with \textit{k} $≥$ 2, assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$). We chose these models to be consistent with the parameters used to generate the "complete" matrix.

Each Bayesian tree was estimated using two runs of four chains each for a maximum of 50$\times$$1^6$ generations. We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and implemented a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}. The effective sample size (ESS) was also checked on a random subsample of runs in each simulation to check that ESS $>>$ 200 \citep{drummond2006ess}. For each run, we removed 25\% of the iterations as burnin. We used the following priors for each tree (see supplementaries): %link missing
\begin{enumerate}
\item
the "true" tree’s topology as a starting tree (with a starting value for each branch length of 1),
\item
an exponential prior on the shape of the gamma distribution of $\alpha$=0.5 for both partitions
\item
and a transition/transversion ratio prior of 2 sampled from a strong beta distribution ($\beta$(80,40)).
\end{enumerate}

We used these priors to speed up the Bayesian process. These priors biased the way the Bayesian process calculated the branch length by giving non-random starting points and boundaries for the parameters estimation process, however, we are focusing on the effect of missing data on the topology and not on the branch length. Even using these priors, it took 65 CPU years to build 50 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).

\subsection{Comparing topologies}
We compared the topology of the "missing-data" trees infered from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$. Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees infered from supermatrices with living and fossil species and molecular and morphological characters. To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triple \citep{dobson1975triplets} and the Robison-Fould \citep{RF1981} distance. We normalised the two metrics using the Normalised Tree Similarity index \citep{Bogdanowicz2012} to generalise our results for any n number of taxa. The two metrics and the index are detailed below.

\subsubsection{Triple distance ($T_{x,y}$) \citep{dobson1975triplets}}
This metric measures the number of different subtrees of three clades between two given trees. Each triplet can be written as $I_{ijk}$=(\textit{ijk}). Where $I_{ijk}$ is equal to 0 if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to 1. For any rooted binary tree there are only three possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}. If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k});.One can calculate $S_n$, the triplet distance between two trees as:
\begin{equation}
S_n = \sum_{ijk} I_{ijk}
\end{equation}
Where:
\begin{equation}
\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
\end{equation}
And where \textit{n} is the number of taxa in both trees (modified from \citet{critchlowthe1996}). If $S_n$=0, the trees are the same (i.e. no taxa as been displaced). When $S_n$ = $\binom{n}{4}$, the trees are the most different possible (i.e. every taxa as been displaced). This metric therefore illustrates the amount of displaxed taxa but is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g. fossil taxa) than the Robinson-Foulds distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and can therefore be used as a proxy to estimate the robustness of the tree to flying taxa (see supplementaries). %link missing

\subsubsection{Robinson-Fould distance (\textit{RF}) \citep{RF1981}}
This metric measures the number of shared clades among two trees and therefore illustrates the number of exactly conserved groups among the trees. The Robison-Fould distance (also called path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
\begin{equation}
RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
\end{equation}
Where $C_{x,y}$ is the number of clades in common in the two trees. The minimal value of \textit{C} is 1 if the two trees have the same \textit{n} taxa; the maximal value in \textit{C=n-2}. This metric is more sensitive to taxa displacement than the triple metric (if one taxa gets out of a clade, then the clades are no longer considered as similar - \citet{critchlowthe1996,johnson1998,wiensmissing2003}) and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades (see supplementaries). %link missing

\subsubsection{Normalised Tree Similarity \textit{NTS} \citep{Bogdanowicz2012}}
For any tree with \textit{n} taxa compared using a tree distance metric \textit{m}, $NTS_m$ represents the similarity score between the two trees given the expected distance between two random Yule trees with \textit{n} taxa. Let $\bar{d}_{m,n}$\textit{(rand)} be the average distance between two random Yule trees with \textit{n} taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} containing each \textit{n} taxa, then:
\begin{equation}
NTS_{m,n}(x,y)=\frac{\bar{d}_{m,n}(rand) - d_{m,n}(x,y)} {\bar{d}_{m,n}(rand)}
\end{equation}
\textit{NTS} ranges from 1 to -$\infty$. For any \textit{m},\textit{n} when \textit{NTS}=1, the trees are the same; when \textit{NTS}=0 the trees are not more different than expected by chance; when \textit{NTS}$<$0, the trees are more different than expected by chance (see supplementaries). %link missing

$\newline$

We compared the "missing-data" tree to the "True" and the "Best" tree for each chain. For the Maximum Likelihood trees we performed pairwise comparisons between the "True" and "Best" tree and \textit{Y} (where \textit{Y} is one of the 125 trees resulting from the 125 super matrices including various amount of missing data) for both the Robinson-Fould and the Triple metric. We calculated the metrics using the TreeCmp java script \citep{Bogdanowicz2012}. For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparison for the same metric and the same number of taxa \textit{n}=51 (see supplementaries). %link missing
We rand the comparison for the 50 trees of each chain having the same combination of parameters values $M_L$, $M_F$ and $M_C$. We calculated the mode and the 50, 75 and 95 confidence intervals for each comparison of 50 trees using the hdrcde R package (v3.1 – \citet{hdrcde}). For the Bayesian trees, we used the exact same approach described above but instead of comparing the "True" or "Best" tree to the tree \textit{Y}, we compared 1000 trees from the "True" or "Best" tree posterior distribution to 1000 trees from the tree \textit{Y} posterior distribution resulting in 1000 pairwise comparison of the "True" or "Best" tree to the tree \textit{Y} for each 125 trees (see supplementaries - Fig. ~\ref{Fig_Compare}). %link missing

\begin{figure}
\centering
\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp3.pdf}
\caption{Tree comparison protocol. From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we infered a tree in Baysian or Maximum likelihood framework. We then compared this tree with or without fossil/living taxa to the "best/true" tree. In the Bayesian framework, 1000 trees from the posterior ditribution where used for each comparison.}
\label{Fig_Compare}
\end{figure}

\subsection{Empirical data}
We also compared the results obtained from simulated data by using \citet{ronquista2012} empirical data. The supermatrix contains 67 living species plus one outgroup and 45 fossil species of Hymenopteras with 5097 molecular characters and 354 morphological characters. From the 68 living species used in the supermatrix, only 66 had molecular data %missing???
, we therefore treated these 66 taxa as "living" taxa and all the other 47 as "fossil" taxa. We treated the matrix in the exact same way as described in step 2 and 3 resulting in 125 supermatrices with various amount of missing data and the same number of Maximum Likelihood and Bayesian trees. We used the same settings as for the simulated data in the Maximum Likelihood framework. For the Bayesian inferences however, we didn’t used any priors except that we provided a starting tree with the topology of the 68 living species (topology with the highest posterior probability from non-clock analysis - \citet{ronquista2012}). Contrary to \citet{ronquista2012} analysis, we didn’t performed any clock analysis since we were only interested in the topology of the infered tree and not the branch length.

\section{Results}
%Baysian inference doesn’t recover the topology, whatever the amount of data
%Bayesian inference with or without the fossil/living doesn’t recover the topology, whatever the amount of data
%Likelihood inference recovers the topology depending on the amount of data
%Likelihood with only living recovers the topology really well
%Likelihood with only fossil??

\section{Discussion}

\section{Supplementaries}
\label{supplementaries}
blabla
%BIBLIOGRAPHY
 % The \cite command functions as follows:
 %   \citet{key} ==>>                Jones et al. (1990)
 %   \citet*{key} ==>>               Jones, Baker, and Smith (1990)
 %   \citep{key} ==>>                (Jones et al., 1990)
 %   \citep*{key} ==>>               (Jones, Baker, and Smith, 1990)
 %   \citep[chap. 2]{key} ==>>       (Jones et al., 1990, chap. 2)
 %   \citep[e.g.][]{key} ==>>        (e.g. Jones et al., 1990)
 %   \citep[e.g.][p. 32]{key} ==>>   (e.g. Jones et al., p. 32)
 %   \citeauthor{key} ==>>           Jones et al.
 %   \citeauthor*{key} ==>>          Jones, Baker, and Smith
 %   \citeyear{key} ==>>             1990

\bibliographystyle{sysbio} %don't write the suffix
\bibliography{References} %don't write the suffix

%END
\end{document}

### 2014-05-14 Manuscript/TEM_draft.tex
Paper draft in proper latex

@@
 \bigskip
-\noindent RH: Missing data in Total Evidence Supermatrices %RH = Running Head
+\noindent RH: Missing data in Total Evidence matrices %RH = Running Head
 
@@
 %Title
-\noindent{\Large \bf Effect of missing data in supermatrices containing living and fossil taxa on topological accuracy}
+\noindent{\Large \bf Effect of missing data in matrices containing living and fossil taxa on topological accuracy}
 \bigskip
@@
 
+%IMPORTANT:
+%Note on display in the text file: sentences or sub-parts of sentences (e.g. enumerations) are delimited by a newline for clarity.
+%However, true new lines (i.e. in the pdf) are delimited by one or more empty lines.
+
 %Abstract
@@
 This fact has led to an increasing consensus among scientists that both fossil and living taxa must be included in macroevolutionary studies.
-One approach, the Total Evidence Method, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both fossil and living taxa at the tips.
-Although the Total Evidence Method seems very promising, it requires a lot of data and is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.
+One approach, the total evidence approach, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both fossil and living taxa at the tips.
+Although the total evidence approach seems very promising, it requires a lot of data and is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.
 
-In this study we assess the effect of missing data on tree topologies inferred from total evidence supermatrices.
-Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the supermatrix: (1) the proportion of living taxa with no morphological data, (2) the amount of missing data in the fossil record and (3) the overall number of morphological characters in the supermatrix.
+In this study we assess the effect of missing data on tree topologies inferred from total evidence matrices.
+Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix:
+(1) the proportion of living taxa with no morphological data,
+(2) the amount of missing data in the fossil record and
+(3) the overall number of morphological characters in the matrix.
 We find that, in a Bayesian framework, difficulties in recovering a stable topology are mainly driven by the missing data in the molecular part of the matrix (for which fossil taxa have no data).
@@
 
-Our results suggest that, in order to use total evidence methods, one should reduce the missing data in the morphological part of the supermatrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process.
-We apply our method to a comprehensive data set of both living and fossil primates.
+Our results suggest that, in order to use total evidence approaches, one should reduce the missing data in the morphological part of the matrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process.
+We apply our method to a comprehensive data set of both living and fossil primates. %Not at all!
 We find that using this integrative method modifies previous estimates of rates of body mass evolution within primates.
 
-\noindent (Keywords: bla, bla, bla)\\
+\noindent (Keywords: missing data, total evidence, bla)\\
 
@@
 
-%We don't use a heading for the introduction. The nature of this section is implied by its position following the abstract and preceding the first major section heading.
+%Syst.Biol: We don't use a heading for the introduction. The nature of this section is implied by its position following the abstract and preceding the first major section heading.
 Living species represent less than 1\% of all species that have ever lived \citep{novacek1992ext,raup1993extinction}.
-However, the majority of macroevolutionary studies focus solely on living species \citep{cooperwhat2009,meredithimpacts2011}.
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as species richness gradients, e.g.
-extant clade diversity may ignore past diversification patterns \citep{shoshani1996proboscidea}; biogeographical history, e.g.
-extant biogeographical patterns are improved when integrating fossil data \citep{metcalfintegrating2014}; or paleoecology, e.g.
-extant clades niches might have differed greatly through time \citep{youngthe2010}.
-%Rewrite that part
+However, the majority of macroevolutionary studies focus solely on living species \citep{cooperwhat2009,meredithimpacts2011,healyecology2014}.
+%or is citing of your own papers too douchebagy in the second sentence?
+Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
 These factors have led to increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-However, to do this we need to be able to place living and fossil taxa into the same phylogenies which still remains complex.
+However, to do this we need to be able to place living and fossil taxa into the same phylogenies which still remains complex despite an increasing number of significant efforts \citep{pyrondivergence2011,ronquista2012,schragocombining2013}.
+%Not a nice sentence
 
 Three main approaches have been used for combining fossil and living taxa data in phylogenies.
-These approaches differ in whether they treat fossil taxa as nodes or tips and how much of the available data is used (i.e.
-age only or age and morphology).
-Classical cladistic methods use morphological data from both fossil and living taxa and treat each taxon as a tip \citep{simpson1945}.
-This approach is commonly used by paleontologists but it ignores
-the majority of %You proposed to remove that.
-The point here was to not get in trouble with modern paleontologist that feel "clever" using DNA as a binary character like "Presence/absence of this SINE/that microsatellite"…
-the additional molecular data available from living species.
-Neontologists, on the other hand, more commonly use only molecular data from living species to build phylogenies.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these analyses and only their occurrence age is used to time calibrate phylogenies \citep{zuckerkandl1965}.
-There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} %Should I develop this improvements here?
-In the end this improvements are more on the inference methods than on the type of data used which is more the point here.
-as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013}.
-A final class of methods, known as total evidence methods use molecular data from living taxa and morphological data from both living and fossil taxa, and treats every taxon as a tip on the phylogeny \citep{eernissetaxonomic1993}.Here we focus on these total evidence methods because they have been recently
-successfully developed and applied to empirical data %needs a better justification
-\citep{pyrondivergence2011,ronquista2012,schragocombining2013}.
+These approaches differ mainly in whether they treat fossil taxa as tips or as nodes in the phylogeny and which part of the available data of the fossil is used (i.e.
+the fossil occurrence age only or both the age and the morphology).
+Classical cladistic methods uses matrices containing morphological data from both fossil and living taxa and allows to treat each taxon as a tip in the phylogeny; the relation between the taxa can be infered using optimal criteria such as the maximum parsimony criterion \citep{simpson1945}.
+This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species.
+Neontologists, on the other hand, more commonly use probabilistic methods (e.g.
+Maximum Likelihood or Bayesian) based on matrices containing only molecular data from living species to infer phylogenies.
+Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these analyses and only their occurrence age can be used to time calibrate phylogenies \citep{zuckerkandl1965}.
+There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013} but none of them uses all the available data.
+
+A final approach, known as total evidence approach use matrices containing molecular data from living taxa and morphological data from both living and fossil taxa.
+%There might be a philosophico-semantic distinction to do here.
+Is TEM a method or an approach?
+I'll say that it is an approach because it's more like an idea (put every thing together) than a way to do (how to put everything together).
+It makes more sense to me compared to a real method e.g.
+a probabilistic method such as ML where the method states precisely how to do stuff.
+It allows to use probabilistic methods, to treat every taxon as a tip on the phylogeny and to use the occurrence age of the fossil to time calibrate the phylogeny \citep{eernissetaxonomic1993}.
+Here we focus on this total evidence approach because they have been recently successfully developed and applied to empirical data \citep{pyrondivergence2011,ronquista2012,schragocombining2013}.
+%needs a better justification
 Although the total evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data.
 In particular it requires morphological data form both living and fossil taxa, both of which are scarce.
-Therefore total evidence approaches are likely to suffer from having lots of missing data which may affect their ability to infer correct phyogenies.
+Therefore total evidence approaches are likely to suffer from having lots of missing data which may affect their ability to infer correct phylogenies.
 % Trevor: Missing data also influences support values that trend to decrease
@@
 The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013}.
-Missing molecular data has been seen by some authors as an issue because it can decrease the phylogenetic signal in some parts of the tree, especially when using large supermatrices \citep{lemmonthe2009}.
-However other authors do not see missing molecular data as a major issue because the phylogenetic signal is more likely to increase by having at least a
-"modest" % explain
-number of highly covered genes (~50\% - \citet{rouresite-specific2011}), a higher number of taxa (especially slowly evolving taxa or taxa close to the outgroup) and by choosing more adequate models of sequence evolution rather than by reducing the amount of missing data \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
-Similarly, missing morphological data might be seen as major or minor issue for accurately infering phylogenies \citep{wiensmissing2003,sansomfossilization2013}.
-Because soft-tissues characters are rarely preserved in the fossil record, missing data is found in soft tissues, i.e.
-it is not randomly distributed, which can lead to biased placement of fossil taxa in phylogenies \citep{sansomfossilization2013}.
-However, the phylogenetic signal is not related to the level of missing data per se but to the number of informative characters per taxa, therefore missing data is less an issue than the number of shared informative characters \citep{wiensmissing2003}.
-Although not a major problem separately \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011}, missing data in molecular and morphological supermatrices may become an issue when combined for example in a total evidence type supermatrices and no attempt has been made to study the impact of this issue until now.
+Missing molecular data has been seen by some authors as an issue because it can decrease the phylogenetic signal in some parts of the tree, especially when using large matrices \citep{lemmonthe2009}.
+However other authors do not see missing molecular data as a major issue because the phylogenetic signal is more likely to increase by having at least a "modest" number of highly covered genes (i.e.
+approximatively half of the genes - \citet{rouresite-specific2011}), a higher number of taxa (especially slowly evolving taxa or taxa close to the outgroup) and by choosing more adequate models of sequence evolution rather than by reducing the amount of missing data \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
+Similarly, depending on the study, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study \citep{wiensmissing2003,sansomfossilization2013}.
+Because soft-tissues characters are rarely preserved in the fossil record, missing data is mainly found in soft tissues characters, and is therefore not randomly distributed, which can lead to biased placement of fossil taxa in phylogenies \citep{sansomfossilization2013}.
+However, the phylogenetic signal is not related to the level of missing data \textit{per se} but to the number of informative characters per taxa, therefore missing data is less an issue than the number of shared informative characters \citep{wiensmissing2003}.
+Although missing data as been shown to be no major problem separately in molecular and morphological matrices \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011} it may become an issue when combined in a total evidence type matrix and no attempt has been made to study the impact of this issue on phylogenetic inference until now (especially because a major part of the matrix is missing - the molecular data from the fossil taxa).
 %Does not flow with the rest
 
-Here we assess the effect of missing data on tree topology infered from total evidence supermatrices.
-The molecular part of a total evidence supermatrix contains no fossil taxa so will act like a classical molecular supermatrix \citep{ronquista2012}.
-The effect of missing data on such matrices is well known, therefore, we only focus on the missing data issue in the morphological part of the supermatrix.
-Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the supermatrix: (1) the proportion of living taxa with no morphological data, (2) the amount of missing data in the fossil taxa (i.e.
-the preservation quality of the fossil record) and (3) the overall number of morphological characters for both living and fossil taxa in the supermatrix.
-We assess how changing the values of these three parameters affects the topology of total evidence method phylogenies.
+Here we assess the effect of missing data on tree topology inferred from total evidence matrices.
+The molecular part of a total evidence matrix contains no fossil taxa so will act like a classical molecular matrix \citep{ronquista2012}.
+The effect of missing data on such matrices is well known, therefore, we only focus on the missing data issue in the morphological part of the matrix.
+Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix:
+(1) the proportion of living taxa with no morphological data,
+(2) the amount of missing data in the fossil taxa (i.e. the preservation quality of the fossil record) and
+(3) the overall number of morphological characters for both living and fossil taxa in the matrix.
+We assess how changing the values of these three parameters affects the topology of total evidence approach phylogenies.
+%This paragraph is crap
 
@@
 
-\section{Methods} %no material?
+\section{Methods} %no material needed?
 To explore the effect of missing data on total evidence trees topologies we used the following protocol (note that we explain each step in detail below this general outline -Fig. ~\ref{Fig_Outline}):
@@
 We assigned the number of character states (either 2 or 3) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of characters states extracted from 100 published empirical morphological matrices (See supplementaries).
-%link missing
+These probabilities were selected using the overall distribution of characters states extracted from 100 published empirical morphological matrices (See supplementaries). %link missing
 We then ran an independent discrete character simulation for each character with the randomly selected number of states (2 or 3) and assuming an equal rate of change (i.e.
 evolutionary rate) from one character state to an other.
-This method allows us to have only two parameters per character i.e.
-the number of states and the evolutionary rate.
+This method allows us to have only two parameters per character i.e. the number of states and the evolutionary rate.
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5. 
@@
 Finally, we combined the morphological and molecular matrices obtained from the "true" tree.
-Hereafter we call this the "complete" matrix, i.e.
-the matrix with no missing data except for the molecular data of the fossil taxa.
+Hereafter we call this the "complete" matrix, i.e. the matrix with no missing data except for the molecular data of the fossil taxa.
 
 \subsection{Removing data}
-Once we obtained the "complete" matrix we modified it to get a set of matrices with missing data.
-We randomly replaced data with “?” in the morphological part of the matrices according to the following parameters (Fig.
-~\ref{Fig_RemoveData}):
+Once we obtained the "complete" matrix we modified it to get a set of matrices with missing data
+ We randomly replaced data with "?" in the morphological part of the matrices according to the following parameters (Fig. ~\ref{Fig_RemoveData}):
 
@@
 \includegraphics[keepaspectratio=true]{Figures/TEM_Fig_missingData.pdf}
-\caption{Data removal protocol: Missing living - The proportion of living taxa with no morphological data ($M_{L}$); Missing fossil - The proportion of missing morphological data across all fossil taxa ($M_{F}$); Missing character - The proportion of missing morphological characters across all taxa (living and fossil) ($M_{C}$).}
+\caption{Data removal protocol:
+Missing living - The proportion of living taxa with no morphological data ($M_{L}$);
+Missing fossil - The proportion of missing morphological data across all fossil taxa ($M_{F}$);
+Missing character - The proportion of missing morphological characters across all taxa (living and fossil) ($M_{C}$).}
 \label{Fig_RemoveData}
@@
 In practice, each parameter represent a way of removing data in the morphological part of the matrix: $M_{L}$ removes a proportion of rows from the living taxa; $M_{F}$ removes a proportion of cells from the fossil taxa; and $M_{C}$ removes a proportion of columns across both living and fossil taxa.
-Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the supermatrix affected: for $M_{L}$, all the morphological data of a proportion of the living taxa is removed (i.e.
+Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a proportion of the living taxa is removed (i.e.
 removing rows), as for $M_{F}$, a proportion of data is removed across the whole of the morphological matrix for fossil taxa (i.e.
@@
 We tested all parameters combinations resulting in 125 ($5^3$) matrices.
-Because some parameter combinations introduce a lot of missing (e.g.
-$M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
+Because some parameter combinations introduce a lot of missing (e.g. $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
 When this occurred we repeated the random deletion of characters until every species had at least 5\% data.
@@
 The "best" tree, on the other hand, is the best tree we can build using the state-of-the-art phylogenetic methods.
-In real world situations, the "true" tree is never available to us because we cannot know the true evolutionnary history of a clade (except in very rare circumstances, e.g.
-\citet{rozen2005}).
+In real world situations, the "true" tree is never available to us because we cannot know the true evolutionnary history of a clade (except in very rare circumstances, e.g. \citet{rozen2005}).
 Therefore, here we focus on comparing the trees infered from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
@@
 The effective sample size (ESS) was also checked on a random subsample of runs in each simulation to check that ESS $>>$ 200 \citep{drummond2006ess}.
-For each run, we removed 25\% of the iterations as burnin.
-We used the following priors for each tree (see supplementaries): %link missing
+For each run, we removed 25\% of the iterations as burnin. We used the following priors for each tree (see supplementaries): %link missing
 \begin{enumerate}
@@
 We compared the topology of the "missing-data" trees infered from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
-Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees infered from supermatrices with living and fossil species and molecular and morphological characters.
+Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees infered from matrices with living and fossil species and molecular and morphological characters.
 To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triple \citep{dobson1975triplets} and the Robison-Fould \citep{RF1981} distance.
-We normalised the two metrics using the Normalised Tree Similarity index \citep{Bogdanowicz2012} to generalise our results for any n number of taxa.
-The two metrics and the index are detailed below.
+We normalised the two metrics using the Normalised Tree Similarity index \citep{Bogdanowicz2012} to generalise our results for any n number of taxa. The two metrics and the index are detailed below.
 
@@
 For any rooted binary tree there are only three possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
-If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k});.One can calculate $S_n$, the triplet distance between two trees as:
+If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k});.
+One can calculate $S_n$, the triplet distance between two trees as:
 \begin{equation}
@@
 And where \textit{n} is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
-If $S_n$=0, the trees are the same (i.e.
-no taxa as been displaced).
-When $S_n$ = $\binom{n}{4}$, the trees are the most different possible (i.e.
-every taxa as been displaced).
+If $S_n$=0, the trees are the same (i.e. no taxa as been displaced). When $S_n$ = $\binom{n}{4}$, the trees are the most different possible (i.e. every taxa as been displaced).
 This metric therefore illustrates the amount of displaxed taxa but is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g.
@@
 \textit{NTS} ranges from 1 to -$\infty$.
-For any \textit{m},\textit{n} when \textit{NTS}=1, the trees are the same; when \textit{NTS}=0 the trees are not more different than expected by chance; when \textit{NTS}$<$0, the trees are more different than expected by chance (see supplementaries).
-%link missing
+For any \textit{m},\textit{n} when \textit{NTS}=1, the trees are the same;
+when \textit{NTS}=0 the trees are not more different than expected by chance;
+when \textit{NTS}$<$0, the trees are more different than expected by chance (see supplementaries). %link missing
 
@@
 
+\subsection{Inference testing}
+How we tested the inference (vertical/horizontal/corner).
+
 \subsection{Empirical data}
 We also compared the results obtained from simulated data by using \citet{ronquista2012} empirical data.
-The supermatrix contains 67 living species plus one outgroup and 45 fossil species of Hymenopteras with 5097 molecular characters and 354 morphological characters.
-From the 68 living species used in the supermatrix, only 66 had molecular data %missing???
-, we therefore treated these 66 taxa as "living" taxa and all the other 47 as "fossil" taxa.
-We treated the matrix in the exact same way as described in step 2 and 3 resulting in 125 supermatrices with various amount of missing data and the same number of Maximum Likelihood and Bayesian trees.
+The matrix contains 67 living species plus one outgroup and 45 fossil species of Hymenopteras with 5097 molecular characters and 354 morphological characters.
+From the 68 living species used in the matrix, only 66 had molecular data, we therefore treated these 66 taxa as "living" taxa and all the other 47 as "fossil" taxa.
+We treated the matrix in the exact same way as described in step 2 and 3 resulting in 125 matrices with various amount of missing data and the same number of Maximum Likelihood and Bayesian trees.
 We used the same settings as for the simulated data in the Maximum Likelihood framework.
@@
 \section{Results}
-%Baysian inference doesn’t recover the topology, whatever the amount of data
-%Bayesian inference with or without the fossil/living doesn’t recover the topology, whatever the amount of data
-%Likelihood inference recovers the topology depending on the amount of data
-%Likelihood with only living recovers the topology really well
-%Likelihood with only fossil??
+The ability of recovering the best topology in maximum likelihood method is function of the amount of data missing.
+However, in Bayesian, topology is badly (if not at all) recovered independently of the amount of data missing.
 
 \section{Discussion}
+
+What about the "best" trees with fairly bad bootstrap support values in ML?
+This is probably due to the low number of characters used in the simulation.
+The number of morphological characters ($10^2$) is in the same order of magnitudes of empirical studies (cite) but decreases unrealistically with the $M_C$ parameter.
+%Extract the mean number of characters per matrices from the 100 random ones
+The number of molecular characters ($10^3$) is at least two orders of magnitude lower than in empirical studies that can reach $10^6$ characters (cite).
+
+Two different matrices from the "True" tree gives two different phylogenetic interpretations (i.e. seqGen and ape performs badly at generating a data matrix from a tree)?
+That seems to not really be a problem since we are interested in recovering topology with missing data, not assessing if phylogenetic matrices generation software performs correctly.
+Also, the "noise" that might be created by this potential "bad" phylogenetic matrices generation software is probably diminished by the number of replicates (i.e. 50 chains).
+
+Different phylogenetic signal between the morphological and the molecular matrix?
+That is not really a problem since the phenotypic evolution is not necessary correlated to the genotypic one.
+Also it is important to note that both information are not exactly the same.
+The main difference between molecular and morphological characters is that, because of historical and practical reasons, a morphological character matrix do no contains autapomorphies and invariant characters which are usually more common than synampomorphies in molecular character matrices (cite).
+
+What about homoplasy in the morphological matrix?
+Here, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly
+If one states that morphological characters accumulate through time in the same way as the majority of the molecular characters (neutral theory, cite) then, homoplasic characters are expected to appear randomly through time.
+Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
+
+\section{Conclusion}
+A Bayesian approach fails to recover accurate topology in a total evidence approach framework, whatever the amount of missing data.
+We think this failure is due to the intrinsic structure of the Total Evidence matrices that includes a vast amount of not randomly distributed missing data (i.e.
+the molecular part of the matrix for the fossil taxa).
+This missing data is equal to the number of fossil taxa $\times$ the number of molecular characters leading to a vast amount of likely trees to be sampled in the Bayesian mcmc tree building.
+However, one can use a Maximum Likelihood approach to obtain the likeliest topology and to use this one as a fixed.
+
+\section{Acknowledgements}
+We would like to thank Trevor Hodkinson %Also Gavin Thomas, Emmanuel Douzery and Frédéric Delsuc if not authors.
+for his useful comments on the simulation protocol and Paddy Doyle for the assistance on using the computer cluster.
+All calculations were performed on the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing.
+This cluster was funded through grants from Science Foundation Ireland.
 

### 2014-05-15 Functions/TreeBuilding/TEM_split-tasker.sh
v 0.2 - added chain number after folder name

@@ -13,9 +13,10 @@
 #########################
 #WARNING: this version only deals with nexus files generated by TEM_treesim.sh with the following parameters (25 living taxa, 25 fossil taxa, 1 outgroup - 1000 molecular characters, 100 morphological characters).
 #WARNING: this version only uses only the matrix with no missing data L00C00F00
-#version: 0.1
+#version: 0.2
+#Update: Correction in mpirun -np 8 mb ... added chain number after folder name ($folder$n)
 #----
-#guillert(at)tcd.ie - 13/05/2014
+#guillert(at)tcd.ie - 15/05/2014
 ##########################
 #Requirements:
 #-Shell script TEM_treesim_split.sh
@@ -65,8 +66,8 @@ module load cports gsl/1.16-gnu
 #TASK FILE ${n} - ${chain}
 ##########################
 
-mpirun -np 8 mb ${folder}/${chain}${n}_L00F00C00.hor.cmd ;
-mpirun -np 8 mb ${folder}/${chain}${n}_L00F00C00.ver.cmd ;
-mpirun -np 8 mb ${folder}/${chain}${n}_L00F00C00.cor.cmd ;" | sed 's/\/\#!/\#!/g' > ${chain}${n}.job
+mpirun -np 8 mb ${folder}${n}/${chain}${n}_L00F00C00.hor.cmd ;
+mpirun -np 8 mb ${folder}${n}/${chain}${n}_L00F00C00.ver.cmd ;
+mpirun -np 8 mb ${folder}${n}/${chain}${n}_L00F00C00.cor.cmd ;" | sed 's/\/\#!/\#!/g' > ${chain}${n}.job
 
 done

### 2014-05-16 Functions/TreeBuilding/TEM_split-tasker.sh
v0.3 mb is now run directly in $folder$n

@@ -13,10 +13,11 @@
 #########################
 #WARNING: this version only deals with nexus files generated by TEM_treesim.sh with the following parameters (25 living taxa, 25 fossil taxa, 1 outgroup - 1000 molecular characters, 100 morphological characters).
 #WARNING: this version only uses only the matrix with no missing data L00C00F00
-#version: 0.2
+#version: 0.3
 #Update: Correction in mpirun -np 8 mb ... added chain number after folder name ($folder$n)
+#Update: Correction in task file, mb is now run directly in $folder$n
 #----
-#guillert(at)tcd.ie - 15/05/2014
+#guillert(at)tcd.ie - 16/05/2014
 ##########################
 #Requirements:
 #-Shell script TEM_treesim_split.sh
@@ -66,8 +67,10 @@ module load cports gsl/1.16-gnu
 #TASK FILE ${n} - ${chain}
 ##########################
 
-mpirun -np 8 mb ${folder}${n}/${chain}${n}_L00F00C00.hor.cmd ;
-mpirun -np 8 mb ${folder}${n}/${chain}${n}_L00F00C00.ver.cmd ;
-mpirun -np 8 mb ${folder}${n}/${chain}${n}_L00F00C00.cor.cmd ;" | sed 's/\/\#!/\#!/g' > ${chain}${n}.job
+cd ${folder}${n}
+mpirun -np 8 mb ${chain}${n}_L00F00C00.hor.cmd ;
+mpirun -np 8 mb ${chain}${n}_L00F00C00.ver.cmd ;
+mpirun -np 8 mb ${chain}${n}_L00F00C00.cor.cmd ;
+cd .." | sed 's/\/\#!/\#!/g' > ${chain}${n}.job
 
 done

### 2014-05-16 Manuscript/TEM_draft.tex
Ignore

@@
 \end{enumerate}
-To measure the effect of missing data’s distribution, we repeated steps ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} with the exact same fixed parameters 50 times.
+To measure the effect of missing data distribution, we repeated steps ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} with the exact same fixed parameters 50 times.
+A list of all the terms used in this paper is available in table ~\ref{Tab_glossary}.
 
@@
 \end{figure}
+
+\begin{table}[ht]
+\caption{Glossary}
+\centering
+\begin{tabular}{p{5cm} p{9cm}}
+\hline
+\hfill Term & Definition \\[0.5ex]
+\hline
+\hfill living taxa & taxa with both molecular and morphological data available \\[1.5ex]
+\hfill fossil taxa & taxa with only morphological data available \\[1.5ex]
+\hfill "complete" matrix & matrix with no missing data except for the molecular data for the fossil taxa \\[1.5ex]
+\hfill "missing-data" matrix & matrix with various amount of missing data \\[1.5ex]
+\hfill $M_{L}$ & missing living taxa in the morphological part of the matrix \\[1.5ex]
+\hfill $M_{F}$ & missing data for the fossil taxa in morphological part of the matrix \\[1.5ex]
+\hfill $M_{C}$ & missing morphological characters for both living and fossil taxa \\[1.5ex]
+\hfill "true" tree & tree used to simulate the matrix \\[1.5ex]
+\hfill "best" tree & best tree inferred from the matrix \\[1.5ex]
+\hfill "missing-data" tree & tree inferred from a matrix with missing data ($M_{L}$, $M_{F}$, $M_{C}$) \\[1.5ex]
+\hfill "horizontal" sub-matrix & sub-matrix with morphological and molecular characters for living taxa \\[1.5ex]
+\hfill "vertical" sub-matrix & sub-matrix with morphological data for both living and fossil taxa \\[1.5ex]
+\hfill "corner" sub-matrix & sub-matrix with morphological data for living taxa \\[1.5ex]
+\hfill "missing-data" sub-matrix & molecular part of the matrix for the fossil taxa (no data) \\
+\end{tabular}
+\label{Tab_glossary}
+\end{table}
 
@@
 \includegraphics[keepaspectratio=true]{Figures/TEM_Fig_missingData.pdf}
-\caption{Data removal protocol:
+\caption{Different parts of the matrix. %Title?
+A: Missing data parameters:
 Missing living - The proportion of living taxa with no morphological data ($M_{L}$);
 Missing fossil - The proportion of missing morphological data across all fossil taxa ($M_{F}$);
-Missing character - The proportion of missing morphological characters across all taxa (living and fossil) ($M_{C}$).}
+Missing character - The proportion of missing morphological characters across all taxa (living and fossil) ($M_{C}$).
+B: Different parts of the matrix:
+The "horizontal" sub-matrix (orange) contains molecular and morphological data for living taxa;
+The "vertical" sub-matrix (blue) contains morphological data for living and fossil taxa;
+The "corner" sub-matrix (orange and blue striped) contains morphological data for living taxa;
+The "missing-data" sub-matrix (grey) is the molecular part of the fossil taxa and contains no data.}
 \label{Fig_RemoveData}
@@
 \subsection{Building phylogenies}
-From the resulting matrices we generated two types of trees, the "best" tree that is infered from the "complete" matrix and the trees with missing data infered from the 125 matrices with various amounts of missing data.
+From the resulting matrices we generated two types of trees, the "best" tree that is inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
 The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
@@
 In real world situations, the "true" tree is never available to us because we cannot know the true evolutionnary history of a clade (except in very rare circumstances, e.g. \citet{rozen2005}).
-Therefore, here we focus on comparing the trees infered from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
+Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
+%Maybe add a 'glossary' table with what do we call a best/true/missing tree?
 
@@
 We compared the topology of the "missing-data" trees infered from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
-Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees infered from matrices with living and fossil species and molecular and morphological characters.
+Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees infered from matrices with living and fossil taxa and molecular and morphological characters.
 To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triple \citep{dobson1975triplets} and the Robison-Fould \citep{RF1981} distance.
@@
 We compared the "missing-data" tree to the "True" and the "Best" tree for each chain.
-For the Maximum Likelihood trees we performed pairwise comparisons between the "True" and "Best" tree and \textit{Y} (where \textit{Y} is one of the 125 trees resulting from the 125 super matrices including various amount of missing data) for both the Robinson-Fould and the Triple metric.
-We calculated the metrics using the TreeCmp java script \citep{Bogdanowicz2012}.
+For the Maximum Likelihood trees we performed pairwise comparisons between the "True" and "Best" tree and the "missing-data" tree (where the "missing-data" tree is one of the 125 trees resulting from the 125 super matrices including various amount of missing data) for both the Robinson-Fould and the Triple metric.
+We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
 For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparison for the same metric and the same number of taxa \textit{n}=51 (see supplementaries).
 %link missing
-We rand the comparison for the 50 trees of each chain having the same combination of parameters values $M_L$, $M_F$ and $M_C$.
-We calculated the mode and the 50, 75 and 95 confidence intervals for each comparison of 50 trees using the hdrcde R package (v3.1 – \citet{hdrcde}).
-For the Bayesian trees, we used the exact same approach described above but instead of comparing the "True" or "Best" tree to the tree \textit{Y}, we compared 1000 trees from the "True" or "Best" tree posterior distribution to 1000 trees from the tree \textit{Y} posterior distribution resulting in 1000 pairwise comparison of the "True" or "Best" tree to the tree \textit{Y} for each 125 trees (see supplementaries - Fig.
+We ran the comparison for every "missing-data" tree being one of the 125 combination of parameters values $M_L$, $M_F$ and $M_C$ in each chain resulting in 50 comparisons for every "missing-data" trees.
+We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package (v3.1 – \citet{hdrcde}).
+
+For the Bayesian trees, we had to transform the trees' posterior distribution to a single value in order to be able to compare it to the results of the Maximum Likelihood trees comparisons.
+Traditionally people use the 50\% consensus rule in order to summarise a Bayesian tree posterior distribution \citep[e.g.][]{Ronquist2012mrbayes}.
+However, when the Bayesian inference fails to recover a stable tree topology, the resulting consensus tree may contain many polytomies up to the maximum (resulting in a star tree, i.e.
+no resolution at all).
+Because one cannot compare trees with no resolution we didn't use the consensus tree but instead we used the mode of 1000 random pairwise comparisons within the Bayesian trees posterior distribution.
+In other words, we summarized the difference between two Bayesian tree posterior distributions using the mode of the distribution of 1000 random trees drawn from the first Bayesian trees posterior distribution to 1000 random trees drawn from the the second Bayesian trees posterior distribution.
+We used the mode to summarize the distributions because it reflects more accurately what a consensus tree represents (the topology the most represented in the posterior distribution).
+A matrix with a good phylogenetic signal will lead to a posterior trees distribution containing many identical topologies.
+Using a normalised tree similarity metric to compare the trees within such a posterior distribution will lead to a high value of tree similarity, reflecting reality.
+Alternatively, a matrix with a poor phylogenetic signal will lead to a posterior distribution containing many different topologies.
+Using a normalised tree similarity metric to compare the trees within such a posterior distribution will lead to a low value of tree similarity, reflecting reality.
+%Add something in the supplementaries or a figure...
+Don't know yet what.
+We compared the pairs of Bayesian tree posterior distributions in the same way as for the Maximum Likelihood trees ("True" and "Best" tree compared to the tree "missing-data" tree).
+This resulted in a distribution of one mode per chain for each "missing-data" trees (i.e.
+50 modes) for which we calculated the mode and the 50\% and 95\% confidence intervals (see supplementaries - Fig.
 ~\ref{Fig_Compare}).
@@
 
-\begin{figure}
+\begin{figure} % Modify the figure
 \centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp3.pdf}
+\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp.pdf}
 \caption{Tree comparison protocol.
@@
 We then compared this tree with or without fossil/living taxa to the "best/true" tree.
-In the Bayesian framework, 1000 trees from the posterior ditribution where used for each comparison.}
+In the Bayesian framework, 1000 trees from the posterior distribution where used for each comparison.}
 \label{Fig_Compare}
@@
 
-\subsection{Inference testing}
-How we tested the inference (vertical/horizontal/corner).
+\subsection{Effect of missing molecular characters for fossil taxa} %Awfull name.
+%How we tested the inference (vertical/horizontal/corner).
+To assess the effect of missing molecular characters for fossil taxa in a Bayesian framework, we reran the Bayesian tree inference on different sub-matrices with no missing data in the same way as described above.
+We split the "complete-matrix" in three sub-matrices with no missing data:
+\begin{enumerate}
+\item
+A first containing both molecular and morphological data for living taxa only (hereafter called the "horizontal" sub-matrix, Table ~\ref{Tab_glossary});
+\item
+A second one containing morphological data for both living and fossil taxa (hereafter called the "vertical" sub-matrix, Table ~\ref{Tab_glossary});
+\item
+A third one containing morphological data for living taxa only (hereafter called the "corner" sub-matrix, Table ~\ref{Tab_glossary});
+\end{enumerate}
+We then compared the resulting tree posterior distribution to itself (in the same way described above) to assess the ability to recover topology for each simulation when no missing data was involved in the phylogenetic inference process.
 
-\subsection{Empirical data}
+\subsection{Empirical data} %Still not done!
 We also compared the results obtained from simulated data by using \citet{ronquista2012} empirical data.
@@
 For the Bayesian inferences however, we didn’t used any priors except that we provided a starting tree with the topology of the 68 living species (topology with the highest posterior probability from non-clock analysis - \citet{ronquista2012}).
-Contrary to \citet{ronquista2012} analysis, we didn’t performed any clock analysis since we were only interested in the topology of the infered tree and not the branch length.
+Contrary to \citet{ronquista2012} analysis, we didn’t performed any clock analysis since we were only interested in the topology of the inferred tree and not the branch length.
 

### 2014-05-16 Functions/TreeComparisons/Source.R
R sourcing function

@@ -1,0 +1,10 @@
+##########################
+#TreeCmp functions loader
+##########################
+
+#TreeCmp.Read
+source("~/PhD/Projects/Total_Evidence_Method-Missing_data/Functions/TreeComparisons/TreeCmp.Read.R")
+#TreeCmp.Plot
+source("~/PhD/Projects/Total_Evidence_Method-Missing_data/Functions/TreeComparisons/TreeCmp.Plot.R")
+#NTS
+source("~/PhD/Projects/Total_Evidence_Method-Missing_data/Functions/TreeComparisons/NTS.R")

### 2014-05-19 Manuscript/TEM_draft.tex
RAxML model comment
Added supplementaries

@@
 \label{step:generate_matrix}
-We built a randomly generated birth death tree to infer a "complete" matrix containing both molecular and morphological data for living and fossil taxa.
+We built a randomly generated birth death tree (hereafter called the "best" tree, Table ~\ref{Tab_glossary}) to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix, Table ~\ref{Tab_glossary}).
 \item{Removing data}
 \label{step:remove_data}
-We removed data from the "complete" matrix to simulate the effects of missing data by modifying three parameters (1) the proportion of missing living taxa ($M_{L}$), (2) the proportion of missing data in the fossil taxa ($M_{F}$) and (3) the proportion of missing morphological characters ($M_{C}$).
+We removed data from the "complete" matrix to simulate the effects of missing data by modifying three parameters (1) the proportion of missing living taxa ($M_{L}$), (2) the proportion of missing data in the fossil taxa ($M_{F}$) and (3) the proportion of missing morphological characters ($M_{C}$) (the resulting matrix is called hereafter the "missing-data" matrix, Table ~\ref{Tab_glossary}).
 \item{Building phylogenies}
 \label{step:build_phylo}
-We infered Bayesian phylogenetic trees from the "complete" matrix and from the matrices containing missing data.
+We inferred Bayesian phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in respectively a tree generate from a matrix containing no missing data in the morphological part (hereafter called the "best" tree, Table ~\ref{Tab_glossary}) and trees inferred from matrices with missing data in their morphological part (hereafter called the "missing-data" trees, Table ~\ref{Tab_glossary}).
 \item{Comparing topologies}
 \label{step:compare_topo}
-We then compared the trees obtained from the matrices containing missing data to the ones obtained from the "complete" matrix to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on the topologies of the phylogenies we estimated.
+We then compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on the topologies of the phylogenies we estimated.
 \end{enumerate}
@@
 \hfill "true" tree & tree used to simulate the matrix \\[1.5ex]
-\hfill "best" tree & best tree inferred from the matrix \\[1.5ex]
-\hfill "missing-data" tree & tree inferred from a matrix with missing data ($M_{L}$, $M_{F}$, $M_{C}$) \\[1.5ex]
+\hfill "best" tree & tree inferred from the "complete" matrix \\[1.5ex]
+\hfill "missing-data" tree & tree inferred the "missing-data" matrices \\[1.5ex]
 \hfill "horizontal" sub-matrix & sub-matrix with morphological and molecular characters for living taxa \\[1.5ex]
@@
 
-We infered the morphological matrix using the ape package v3.0-8 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
+We inferred the morphological matrix using the ape package v3.0-8 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
 We assigned the number of character states (either 2 or 3) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of characters states extracted from 100 published empirical morphological matrices (See supplementaries). %link missing
-We then ran an independent discrete character simulation for each character with the randomly selected number of states (2 or 3) and assuming an equal rate of change (i.e.
-evolutionary rate) from one character state to an other.
+These probabilities were selected using the overall distribution of characters states extracted from 100 published empirical morphological matrices (See \hyperref[supplementaries]{supplementaries}).
+We then ran an independent discrete character simulation for each character using the "true" tree branch length and topology with the randomly selected number of states (2 or 3) and assuming an equal rate of change (i.e.
+evolutionary rate) from one character state to an other \citep{Pagel22011994}.
 This method allows us to have only two parameters per character i.e. the number of states and the evolutionary rate.
@@
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
-%Maybe add a 'glossary' table with what do we call a best/true/missing tree?
 
 \subsubsection{Maximum Likelihood}
-The "best" tree and the "missing-data" trees were infered using RAxML v7.0.4 \citep{stamatakisthe2008}.
-We used the GTR + $\Gamma_4$ model (\citet{tavare1986} – default GTRGAMMA in RAxML v7.04) as a generalisation of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
+The "best" tree and the "missing-data" trees were infered using RAxML v8.0.20 \citep{Stamatakis21012014}.
+For the molecular data, we used the GTR + $\Gamma_4$ model (\citet{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citet{Stamatakis21012014}) as a generalisation of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
 The GTR model can be seen as a generalisation of the HKY model (the 2 parameters from the HKY model are implicitly included in the 6 from GTR model - \citet{stamatakisa2008}).
+For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} wich is a generalisation of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014})
 We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. 
-%Mk model seems implemented (i.e.
-Aprils’s slideshare http://www.slideshare.net/wrightaprilm/evolution2013).
-But how?
-Default method for standard data type (i.e.
-non DNA/AA).
-Trevor: swapping algorithm?
-salanin 2003
+The bootstrap algorithm used in RAxML is the Lazy Subtree Rearrangement (LSR) which consists in pruning one subtree from the tree and subsequently reinserting it to al neighboring branches \citep{stamatakisa2008}.
+Subtree Pruning and Reinserting methods (SPR) have been demonstrated as being better than others (e.g. Nearest Neighbouring Interchange - NNI) in recovering good bootstrap values \citep{salamin2003}.
 
@@
 The molecular evolutionary history was infered using the HKY model with a transition/transversion ratio \citep{douadycomparison2003} of two and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
-For the morphological data, we used the Markov \textit{k} state model \citep{lewisa2001} wich is a generalisation of the JC69 model \citep{jc69} with \textit{k} $≥$ 2, assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$).
+For the morphological data, we used the Markov \textit{k} state model \citep{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
 We chose these models to be consistent with the parameters used to generate the "complete" matrix.
@@
 The effective sample size (ESS) was also checked on a random subsample of runs in each simulation to check that ESS $>>$ 200 \citep{drummond2006ess}.
-For each run, we removed 25\% of the iterations as burnin. We used the following priors for each tree (see supplementaries): %link missing
+For each run, we removed 25\% of the iterations as burnin. We used the following priors for each tree (see \hyperref[supplementaries]{supplementaries}):
 \begin{enumerate}
@@
 This metric therefore illustrates the amount of displaxed taxa but is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g.
-fossil taxa) than the Robinson-Foulds distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and can therefore be used as a proxy to estimate the robustness of the tree to flying taxa (see supplementaries).
-%link missing
+fossil taxa) than the Robinson-Foulds distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and can therefore be used as a proxy to estimate the robustness of the tree to flying taxa (see \hyperref[supplementaries]{supplementaries}).
 
@@
 The minimal value of \textit{C} is 1 if the two trees have the same \textit{n} taxa; the maximal value in \textit{C=n-2}.
-This metric is more sensitive to taxa displacement than the triple metric (if one taxa gets out of a clade, then the clades are no longer considered as similar - \citet{critchlowthe1996,johnson1998,wiensmissing2003}) and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades (see supplementaries).
-%link missing
+This metric is more sensitive to taxa displacement than the triple metric (if one taxa gets out of a clade, then the clades are no longer considered as similar - \citet{critchlowthe1996,johnson1998,wiensmissing2003}) and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades (see \hyperref[supplementaries]{supplementaries}).
 
@@
 when \textit{NTS}=0 the trees are not more different than expected by chance;
-when \textit{NTS}$<$0, the trees are more different than expected by chance (see supplementaries). %link missing
+when \textit{NTS}$<$0, the trees are more different than expected by chance (see \hyperref[supplementaries]{supplementaries}).
 
@@
 We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
-For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparison for the same metric and the same number of taxa \textit{n}=51 (see supplementaries).
-%link missing
+For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparison for the same metric and the same number of taxa \textit{n}=51 (see \hyperref[supplementaries]{supplementaries}).
 We ran the comparison for every "missing-data" tree being one of the 125 combination of parameters values $M_L$, $M_F$ and $M_C$ in each chain resulting in 50 comparisons for every "missing-data" trees.
@@
 This resulted in a distribution of one mode per chain for each "missing-data" trees (i.e.
-50 modes) for which we calculated the mode and the 50\% and 95\% confidence intervals (see supplementaries - Fig.
+50 modes) for which we calculated the mode and the 50\% and 95\% confidence intervals (see \hyperref[supplementaries]{supplementaries} - Fig.
 ~\ref{Fig_Compare}).
-%link missing
 
@@
 \subsection{Effect of missing molecular characters for fossil taxa} %Awfull name.
-%How we tested the inference (vertical/horizontal/corner).
-To assess the effect of missing molecular characters for fossil taxa in a Bayesian framework, we reran the Bayesian tree inference on different sub-matrices with no missing data in the same way as described above.
-We split the "complete-matrix" in three sub-matrices with no missing data:
+To assess the effect of missing molecular characters for fossil taxa in a Bayesian framework, we split the "complete" matrix in sub-matrices containing no missing data at all:
 \begin{enumerate}
@@
 \end{enumerate}
+We reran the Bayesian tree inference on the different sub-matrices with no missing data in the same way as described above.
 We then compared the resulting tree posterior distribution to itself (in the same way described above) to assess the ability to recover topology for each simulation when no missing data was involved in the phylogenetic inference process.
@@
 \section{Results}
+%Do the check for L00F00C00
 The ability of recovering the best topology in maximum likelihood method is function of the amount of data missing.
@@
 
-\section{Supplementaries}
-\label{supplementaries}
-blabla
 %BIBLIOGRAPHY
@@
 
+\section{Supplementaries}\label{supplementaries}
+\subsection{Morphological characters states}
+In order to obtain a realistic probabilistic value for of \textit{k} characters states for each simulated morphological character, we downloaded 100 random morphological characters (with more than 100 characters each) from TreeBASE database (http://treebase.org/) published between 1985 and 2013 and covering 19 taxomomic classess (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
+We selected a total of 22563 characters ranging from 2 to 10 states.
+We calculated the proportion of characters with 2, 3, 4, 5, 6, 7, 8, 9 or 10 states.
+We then sampeled 22563 \textit{k} values between 2 and 10 with the same proportion of characters from the empirical data.
+We then used a simple t-test to check if our simulation was equal to the empirical data.
+In this study, we only simulated characters with 2 or 3 states because of the high proportion of ordered characters ecountered on characters with more than 3 states and the difficulties of simulate biologicaly sensible ordered characters.
+
+\begin{figure}
+\centering
+\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-AppendixCharacters.pdf}
+\caption{Character states distribution in empirical matrices. %Title?
+Characters states number distribution extracted from 100 random morphological matrices downloaded from RreeBase.}
+\label{Fig_AppendixCharacters}
+\end{figure}
+
+\subsection{Tree Building Software settings}
+
+\subsubsection{Maximum Likelihood - RAxML v8.0.20 \citep{Stamatakis21012014}} \\
+Model: \\
+Molecular data: \\
+GTR + $\Gamma_4$ (-m GTRGAMMA) \\
+Morphological data: \\
+Mk + $\Gamma_4$ (-K MK) \\
+Support: \\
+Rapid Boostrap algorithm (LSR), 1000 replicates \\
+
+\subsubsection{Bayesian - MrBayes v3.0.2 \citep{Ronquist2012mrbayes}} \\
+Priors: \\
+Molecular data: \\
+rates distribution shape ($\alpha$) = 0.5 \\
+Transition/Transversion ratio = 2 ($\beta$(80,40)) \\
+Morphological data: \\
+rates distribution shape ($\alpha$) = 0.5 \\
+Models: \\
+Molecular data: HKY + $\Gamma_4$ \\
+Morphological data: Mk + $\Gamma_4$ \\
+MCMC: \\
+2 runs \\
+4 chains per run \\
+generations < 50$\times$$1^6$ \\
+sample frequency = 1050$\times$$1^3$ \\
+ASDS diagnosis frequency = 50$\times$$1^3$ \\
+ASDS $<$ 0.01 \\
+ESS $>>$ 200 \\
+Burnin = 25\%
+
+\subsection{Triplets metric details ($T_{x,y}$)}
+Each triplet can be written as $I_{ijk}$=(\textit{ijk})). Where $I_{ijk}$ is equal to 0 if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to 1.
+For any rooted tree there are only four possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); and (\textit{i},\textit{j},\textit{k}); \citep{johnson1998}.
+One can calculate $S_n$, the triplet distance between two trees as:
+\begin{equation}
+S_n = \sum_{ijk} I_{ijk}
+\end{equation}
+Where:
+\begin{equation}
+\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
+\end{equation}
+And where n is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
+When all triplets across the two trees are the same, $S_n$ is equal to 0 and when all the triplets are different $S_n$ is equal to $\binom{n}{4}$.
+Because the possible number of triplets per clade is a finite number, the probability of two random trees with the same n taxa to have the same triplet is:
+\begin{equation}
+P({I_{ijk}}=0) = \frac{1}{4}
+\end{equation}
+Therefore one can calculate the probability of two random trees having the same triplets: 
+\begin{equation}
+P({S_{n}}=0) = \sum_{ijk} P_{I_{ijk}=0}
+\end{equation}
+\begin{equation}
+P({S_{n}}=0) = \frac{n!}{4(3!(n-3)!}
+\end{equation}
+And in the same way:
+\begin{equation}
+P({S_{n}}=1) = \frac{3n!}{4(3!(n-3)!}
+\end{equation}
+
+\subsection{RF metric details}
+The RF distance (or path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
+\begin{equation}
+RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
+\end{equation}
+Where $C_{x,y}$ is the number of clades in common in the two trees. 
+The minimal value of \textit{C} is equal to 1 if the two trees have the same n taxa;
+the maximal value in \textit{C}=\textit{n}-2.
+For a fully unresolved tree (star tree) \textit{N}=1 and for a fully resolved tree (binary tree) \textit{N}=\textit{n}-2.
+The minimal and maximal topological distance for \textit{n} taxa is:
+\begin{equation}
+RF_{min} = 1 + 1 - 2C_{x,y}
+\end{equation}
+And:
+\begin{equation}
+RF_{max} = 2(n-2)-2
+\end{equation}
+One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any \textit{n} taxa:
+\begin{equation}
+RF.scaled_{x,y} = \frac{RF_{x,y}-RF_{max}}{RF_{max}}
+\end{equation}
+This metric is more sensitive to taxa displacement than the Triplet distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.
+
+\subsection{Codes}
+All codes are available at: $https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/tree/master/Functions$
+
 %END

### 2014-05-20 Manuscript/TEM_draft.tex
Draft update

@@
 \subsection{Comparing topologies}
-We compared the topology of the "missing-data" trees infered from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
+We compared the topology of the "missing-data" trees inferred from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
 Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees infered from matrices with living and fossil taxa and molecular and morphological characters.
@@
 
-For the Bayesian trees, we had to transform the trees' posterior distribution to a single value in order to be able to compare it to the results of the Maximum Likelihood trees comparisons.
+For the Bayesian trees, we had to transform the trees posterior distribution to a single value in order to be able to compare it to the results of the Maximum Likelihood trees comparisons.
 Traditionally people use the 50\% consensus rule in order to summarise a Bayesian tree posterior distribution \citep[e.g.][]{Ronquist2012mrbayes}.
@@
 no resolution at all).
-Because one cannot compare trees with no resolution we didn't use the consensus tree but instead we used the mode of 1000 random pairwise comparisons within the Bayesian trees posterior distribution.
+Because one cannot compare trees with no resolution we didn't use the consensus tree but instead we used the mode of 1000 Random Pairwise Bayesian Tree Comparisons (RPBTC).
+The RPBTC allowed us to assess the difference between Bayesian trees posterior distribution by taking into account both the support value and the resolution of the consensus tree.
+The RPBTC gives a low score if comparing any Bayesian trees posterior distribution to a Bayesian trees posterior distribution with a low resolution and a low node support (if any), even if the same posterior distribution is compared to itself.
+Even though the RPBTC is dependent on 1000 random tree drawings, the output tree comparison score (i.e.
+the mode) is still significant not different between independent replicates (100 replicates of 1000 RPBTC each - see \hyperref[supplementaries]{supplementaries})
+%
+%
+%
 In other words, we summarized the difference between two Bayesian tree posterior distributions using the mode of the distribution of 1000 random trees drawn from the first Bayesian trees posterior distribution to 1000 random trees drawn from the the second Bayesian trees posterior distribution.
@@
 %Do the check for L00F00C00
+\subsection{Building the trees}
+Bayesian chains converged (SD< 0.01 and ESS >> 200 for each tree).
+\subsection{Comparing topologies}
+Did it worked? See sub-matrices
+\subsection{Effect of $M_L$}
+In ML and in Bayesian
+\subsection{Effect of $M_F$}
+In ML and in Bayesian
+\subsection{Effect of $M_C$}
+In ML and in Bayesian
+\subsection{Combined effect}
+In ML and in Bayesian
+
 The ability of recovering the best topology in maximum likelihood method is function of the amount of data missing.
@@
 \section{Discussion}
+
+\subsection{Building the trees}
+Hard to simulate morpho characters but see Pagel, April Wright, etc...
+
+\subsection{Comparing topologies}
+Hard to compare Bayesian tree posterior distributions topologies but gives a good idea.
+
+\subsection{ML vs Bayesian}
+Missing data is not a problem as long as random. However, in TEM, n.fossils*n.molecular data is missing and not random.
+Non optimal criterion approach (i.e. Bayesian) worse because of the number of considered "less"-likely trees (flat maximum likelihood landscapes).
+However, ML still good performances and can be used to fix topology.
+
+\subsection{Effect of missing data}
+Parameter blabla seems to be influencing the topology the most.
+But in practice, $M_L$ is the only one realistically solvable. Go to museums!
 
@@
 
+\subsection{Random Bayesian tree comparisons}
+We assessed the power of the Random Pairwise Bayesian Tree Comparison (RPBTC) method by comparing 1000 random trees from a posterior distribution trees set to another 1000 random trees from the same posterior distribution trees set.
+We repeated this 100 times independently using the same posterior distribution trees set each time resulting in 100 replicates of the same posterior distribution trees set compared 1000 times.
+We used an anova to test if there was no significant difference between the replicates so that the RBTC can be replicated.
+We applied this protocol on a poorly resolved tree, a resolved tree with low support value and a resolved tree with high support values.
+%Insert table with results here.
+
 \subsection{Codes}

### 2014-05-20 Analysis/Pariwise_TreeCmp.R
Pairwise comparisons

@@ -1,0 +1,139 @@
+#Script for testing the Random Pairwise Bayesian Tree Comparisons (RPBTC)
+
+#Loading the functions
+source('../Functions/TreeComparisons/NTS.R')
+source('../Functions/TreeComparisons/TreeCmp.Plot.R')
+source('../Functions/TreeComparisons/TreeCmp.Read.R')
+
+#Loading the Random tree comparisons
+Random51<-read.table("../Data/Tree_Comparisons/Random_trees/51t_Random.Cmp",header=TRUE)
+Random26<-read.table("../Data/Tree_Comparisons/Random_trees/26t_Random.Cmp",header=TRUE)
+Random25<-read.table("../Data/Tree_Comparisons/Random_trees/25t_Random.Cmp",header=TRUE)
+
+#Loading the treesets comparisons
+#Low score comparisons
+cd("../Data/Tree_Comparisons/RBTC_testing/Low_score_treesets_100/")
+tmp<-TreeCmp.Read('LowScore', verbose=TRUE)
+LowScore<-NTS(tmp, Random51)
+cd('../../../../../')
+#Medium score comparisons
+cd("../Data/Tree_Comparisons/RBTC_testing/Medium_score_treesets_100/")
+tmp<-TreeCmp.Read('MediumScore', verbose=TRUE)
+MediumScore<-NTS(tmp, Random51)
+cd('../../../../../')
+#High score comparisons
+cd("../Data/Tree_Comparisons/RBTC_testing/High_score_treesets_100/")
+tmp<-TreeCmp.Read('HighScore', verbose=TRUE)
+HighScore<-NTS(tmp, Random26)
+cd('../../../../../')
+
+#Storing each NTS score in a data.frame object
+#RF score (NTS)
+LowScore.RF<-NULL
+for (replicate in 1:100){
+    LowScore.RF[[replicate]]<-LowScore[[i]][[4]]
+}
+LowScore.RF<-as.data.frame(LowScore.RF)
+names(LowScore.RF)<-seq(1:100)
+
+#Triplets score (NTS)
+LowScore.Tr<-NULL
+for (replicate in 1:100){
+    LowScore.Tr[[replicate]]<-LowScore[[i]][[6]]
+}
+LowScore.Tr<-as.data.frame(LowScore.Tr)
+names(LowScore.Tr)<-seq(1:100)
+
[…]

### 2014-05-20 Functions/TreeComparisons/Source.R
remove

@@ -1,10 +1,0 @@
-##########################
-#TreeCmp functions loader
-##########################
-
-#TreeCmp.Read
-source("~/PhD/Projects/Total_Evidence_Method-Missing_data/Functions/TreeComparisons/TreeCmp.Read.R")
-#TreeCmp.Plot
-source("~/PhD/Projects/Total_Evidence_Method-Missing_data/Functions/TreeComparisons/TreeCmp.Plot.R")
-#NTS
-source("~/PhD/Projects/Total_Evidence_Method-Missing_data/Functions/TreeComparisons/NTS.R")

### 2014-05-20 Functions/TreeBuilding/TEM_ChainSum.sh
BurnTree variable fixed

@@ -8,12 +8,14 @@
 #<Chain Name> creates a folder named after the chain name to store the results
 ##########################
 #Create .treeset files with all the trees from two mb chains.
-#version: 0.4
+#version: 0.5
 #Update: Allows a burnin proportion value
 #Update: Error in the burnin is now fixed
 #Update: Name change and cleaning
+#Update: Name.run.t fixed
+#Update: BurnTree variable fixed
 #----
-#guillert(at)tcd.ie - 17/04/2014
+#guillert(at)tcd.ie - 19/05/2014
 ##########################
 
 
@@ -34,7 +36,7 @@ let "FirstTree += 6"
 mkdir ${name}_treesets
 
 #Create the .treeset file
-for f in *.run1.t
+for f in ${name}.run1.t
     do prefix=$(basename $f .run1.t)
 
     echo $prefix
@@ -43,7 +45,7 @@ for f in *.run1.t
     head -$header ${prefix}.run1.t > ${name}_treesets/${prefix}.treeset
 
     #Burnin
-    BurnTree=$'FirstTree'
+    BurnTree=$FirstTree
     length=$(grep '[&U]' $f | wc -l)
     let "length *= $burnin"
     let "length /= 100"

### 2014-05-20 Functions/TreeComparisons/TreeCmp.Read.R
To do sep option seems to work only with '_'

@@ -6,6 +6,7 @@
 #Update: improved input format
 #Update: output is now a 'TreeCmp' list class
 #Update: in FUN.TreeCmp.Read, read.table function as been wrapped in a try function.
+#To do: sep option seems to work only with '_'
 ##########################
 #SYNTAX :
 #<chain> the chain name to read in

### 2014-05-21 Analysis/Pariwise_TreeCmp.R
RPBTC test

@@ -12,26 +12,26 @@ Random25<-read.table("../Data/Tree_Comparisons/Random_trees/25t_Random.Cmp",head
 
 #Loading the treesets comparisons
 #Low score comparisons
-cd("../Data/Tree_Comparisons/RBTC_testing/Low_score_treesets_100/")
+cd("../Data/Tree_Comparisons/RPBTC_testing/Low_score_treesets_100/")
 tmp<-TreeCmp.Read('LowScore', verbose=TRUE)
 LowScore<-NTS(tmp, Random51)
-cd('../../../../../')
+cd('../../../../Analysis/')
 #Medium score comparisons
-cd("../Data/Tree_Comparisons/RBTC_testing/Medium_score_treesets_100/")
+cd("../Data/Tree_Comparisons/RPBTC_testing/Medium_score_treesets_100/")
 tmp<-TreeCmp.Read('MediumScore', verbose=TRUE)
 MediumScore<-NTS(tmp, Random51)
-cd('../../../../../')
+cd('../../../../Analysis/')
 #High score comparisons
-cd("../Data/Tree_Comparisons/RBTC_testing/High_score_treesets_100/")
+cd("../Data/Tree_Comparisons/RPBTC_testing/High_score_treesets_100/")
 tmp<-TreeCmp.Read('HighScore', verbose=TRUE)
 HighScore<-NTS(tmp, Random26)
-cd('../../../../../')
+cd('../../../../Analysis/')
 
 #Storing each NTS score in a data.frame object
 #RF score (NTS)
 LowScore.RF<-NULL
 for (replicate in 1:100){
-    LowScore.RF[[replicate]]<-LowScore[[i]][[4]]
+    LowScore.RF[[replicate]]<-LowScore[[replicate]][[4]]
 }
 LowScore.RF<-as.data.frame(LowScore.RF)
 names(LowScore.RF)<-seq(1:100)
@@ -39,7 +39,7 @@ names(LowScore.RF)<-seq(1:100)
 #Triplets score (NTS)
 LowScore.Tr<-NULL
 for (replicate in 1:100){
-    LowScore.Tr[[replicate]]<-LowScore[[i]][[6]]
+    LowScore.Tr[[replicate]]<-LowScore[[replicate]][[6]]
 }
[…]

### 2014-05-21 Manuscript/TEM_draft.tex
Added RPBC test

@@
 Even though the RPBTC is dependent on 1000 random tree drawings, the output tree comparison score (i.e.
-the mode) is still significant not different between independent replicates (100 replicates of 1000 RPBTC each - see \hyperref[supplementaries]{supplementaries})
+the mode) is still significant not different between independent replicates (100 replicates of 1000 RPBTC each - see \hyperref[supplementaries]{supplementaries}, table ~\ref{RPBTC_testing})
 %
@@
 We used an anova to test if there was no significant difference between the replicates so that the RBTC can be replicated.
-We applied this protocol on a poorly resolved tree, a resolved tree with low support value and a resolved tree with high support values.
-%Insert table with results here.
+We applied this protocol on a poorly resolved tree (Low Score), a resolved tree with low support value (Medium Score) and a resolved tree with high support values (High Score).
+Results are available in table ~\ref{RPBTC_testing}.
+
+\begin{table}[ht]
+\caption{Anova results: difference between 100 replicates using the RPBTC method}
+\centering
+\begin{tabular}{rllrrrr}
+  \hline
+  Tree.Type & Used.metric & Replicates & Df & F.value & p.value \\ [0.5ex]
+  \hline
+  Low Score & RF & 100.00 &   1 & 0.07 & 0.79 \\ [0.5ex]
+  Low Score & Tr & 100.00 &   1 & 0.33 & 0.56 \\ [0.5ex]
+  Medium Score & RF & 100.00 &   1 & 1.07 & 0.30 \\ [0.5ex]
+  Medium Score & Tr & 100.00 &   1 & 0.00 & 0.96 \\ [0.5ex]
+  High Score & RF & 100.00 &   1 & 0.20 & 0.66 \\ [0.5ex]
+  High Score & Tr & 100.00 &   1 & 0.61 & 0.43 \\ [0.5ex]
+  \hline
+\end{tabular}
+\label{RPBTC_testing}
+\end{table}
 

### 2014-05-22 Manuscript/TEM_draft.tex
RPBTC description

@@
 \begin{enumerate}
-\item{Generating the matrix}
+\item{Generating the matrix:} \\
 \label{step:generate_matrix}
-We built a randomly generated birth death tree (hereafter called the "best" tree, Table ~\ref{Tab_glossary}) to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix, Table ~\ref{Tab_glossary}).
-\item{Removing data}
+We built a randomly generated birth death tree (hereafter called the "true" tree, Table ~\ref{Tab_glossary}) to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix, Table ~\ref{Tab_glossary}).
+\item{Removing data:} \\
 \label{step:remove_data}
 We removed data from the "complete" matrix to simulate the effects of missing data by modifying three parameters (1) the proportion of missing living taxa ($M_{L}$), (2) the proportion of missing data in the fossil taxa ($M_{F}$) and (3) the proportion of missing morphological characters ($M_{C}$) (the resulting matrix is called hereafter the "missing-data" matrix, Table ~\ref{Tab_glossary}).
-\item{Building phylogenies}
+\item{Building phylogenies:} \\
 \label{step:build_phylo}
 We inferred Bayesian phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in respectively a tree generate from a matrix containing no missing data in the morphological part (hereafter called the "best" tree, Table ~\ref{Tab_glossary}) and trees inferred from matrices with missing data in their morphological part (hereafter called the "missing-data" trees, Table ~\ref{Tab_glossary}).
-\item{Comparing topologies}
+\item{Comparing topologies:} \\
 \label{step:compare_topo}
@@
 (4) We then compared the trees with missing data (from the matrices with missing data) to the tree with no missing data (hereafter, the "best" tree – infered from the "complete" matrix).
-We repeated step 1 to 4 50 times.}
+We repeated step 1 to 4 51 (3$\times $17) times.}
 \label{Fig_Outline}
@@
 \hfill "missing-data" tree & tree inferred the "missing-data" matrices \\[1.5ex]
+\hfill RPBTC & Random pairwise Bayesian tree comparison \\[1.5ex]
 \hfill "horizontal" sub-matrix & sub-matrix with morphological and molecular characters for living taxa \\[1.5ex]
@@
 The molecular matrix was infered from the "true" tree using the package phyclust v0.1-14 \citep{chen2011}.
-The matrix was made of 1000 characters sites for 51 taxa and generated using the seqgen algorithm \citep{ ranbaut1997seqgen}.
+The matrix was made of 1000 characters sites for 51 taxa and generated using the seqgen algorithm \citep{ranbaut1997seqgen}.
 We used the HKY model \citep{HKY85} with a random base frequencies and with the transition/transversion rate of 2 \citep{douadycomparison2003} as parameters for generating the matrix.
@@
 
-We inferred the morphological matrix using the ape package v3.0-8 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
+We inferred the morphological matrix using the ape package v3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
 We assigned the number of character states (either 2 or 3) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
@@
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
+We used a low evolution rate parameter (i.e.
+$\alpha$) in order to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
+
 All the molecular information for fossil taxa was replaced by missing data ("?").
@@
 These priors biased the way the Bayesian process calculated the branch length by giving non-random starting points and boundaries for the parameters estimation process, however, we are focusing on the effect of missing data on the topology and not on the branch length.
-Even using these priors, it took 65 CPU years to build 50 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
+Even using these priors, it took 65 CPU years to build 51 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
 
@@
 To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triple \citep{dobson1975triplets} and the Robison-Fould \citep{RF1981} distance.
-We normalised the two metrics using the Normalised Tree Similarity index \citep{Bogdanowicz2012} to generalise our results for any n number of taxa. The two metrics and the index are detailed below.
+We normalised the two metrics using the Normalised Tree Similarity index \citep{Bogdanowicz2012} to generalise our results for any \textit{n} number of taxa.
+The two metrics and the index are detailed below.
 
@@
 We compared the "missing-data" tree to the "True" and the "Best" tree for each chain.
-For the Maximum Likelihood trees we performed pairwise comparisons between the "True" and "Best" tree and the "missing-data" tree (where the "missing-data" tree is one of the 125 trees resulting from the 125 super matrices including various amount of missing data) for both the Robinson-Fould and the Triple metric.
+For the Maximum Likelihood trees we performed pairwise comparisons between the "True" and "Best" tree and the "missing-data" tree (where the "missing-data" tree is one of the 125 trees resulting from the 125 super matrices including various amount of missing data, see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triple metric.
 We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
 For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparison for the same metric and the same number of taxa \textit{n}=51 (see \hyperref[supplementaries]{supplementaries}).
-We ran the comparison for every "missing-data" tree being one of the 125 combination of parameters values $M_L$, $M_F$ and $M_C$ in each chain resulting in 50 comparisons for every "missing-data" trees.
+We ran the comparison for every "missing-data" tree being one of the 125 combination of parameters values $M_L$, $M_F$ and $M_C$ in each chain resulting in 51 comparisons for every "missing-data" trees.
 We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package (v3.1 – \citet{hdrcde}).
 
-For the Bayesian trees, we had to transform the trees posterior distribution to a single value in order to be able to compare it to the results of the Maximum Likelihood trees comparisons.
-Traditionally people use the 50\% consensus rule in order to summarise a Bayesian tree posterior distribution \citep[e.g.][]{Ronquist2012mrbayes}.
-However, when the Bayesian inference fails to recover a stable tree topology, the resulting consensus tree may contain many polytomies up to the maximum (resulting in a star tree, i.e.
-no resolution at all).
-Because one cannot compare trees with no resolution we didn't use the consensus tree but instead we used the mode of 1000 Random Pairwise Bayesian Tree Comparisons (RPBTC).
-The RPBTC allowed us to assess the difference between Bayesian trees posterior distribution by taking into account both the support value and the resolution of the consensus tree.
-The RPBTC gives a low score if comparing any Bayesian trees posterior distribution to a Bayesian trees posterior distribution with a low resolution and a low node support (if any), even if the same posterior distribution is compared to itself.
-Even though the RPBTC is dependent on 1000 random tree drawings, the output tree comparison score (i.e.
-the mode) is still significant not different between independent replicates (100 replicates of 1000 RPBTC each - see \hyperref[supplementaries]{supplementaries}, table ~\ref{RPBTC_testing})
-%
-%
-%
-In other words, we summarized the difference between two Bayesian tree posterior distributions using the mode of the distribution of 1000 random trees drawn from the first Bayesian trees posterior distribution to 1000 random trees drawn from the the second Bayesian trees posterior distribution.
-We used the mode to summarize the distributions because it reflects more accurately what a consensus tree represents (the topology the most represented in the posterior distribution).
-A matrix with a good phylogenetic signal will lead to a posterior trees distribution containing many identical topologies.
-Using a normalised tree similarity metric to compare the trees within such a posterior distribution will lead to a high value of tree similarity, reflecting reality.
-Alternatively, a matrix with a poor phylogenetic signal will lead to a posterior distribution containing many different topologies.
-Using a normalised tree similarity metric to compare the trees within such a posterior distribution will lead to a low value of tree similarity, reflecting reality.
-%Add something in the supplementaries or a figure...
-Don't know yet what.
-We compared the pairs of Bayesian tree posterior distributions in the same way as for the Maximum Likelihood trees ("True" and "Best" tree compared to the tree "missing-data" tree).
-This resulted in a distribution of one mode per chain for each "missing-data" trees (i.e.
-50 modes) for which we calculated the mode and the 50\% and 95\% confidence intervals (see \hyperref[supplementaries]{supplementaries} - Fig.
+Bayesian tree inference allows to account for the statistical uncertainty of a phylogenetic tree by not using an optimal criterion (c.f.
+Maximum Likelihood) and gives a tree posterior distribution (c.f.
+the likeliest tree in ML).
+This method has the clear advantage of better dealing with error and uncertainty but is less practical for using the results (i.e.
+the phylogeny) for any further study (but see \citet{healyecology2014}).
+To avoid this problem, people traditionally use the a consensus tree build on a majority rule in order to summarise a Bayesian tree posterior distribution leading to a single tree containing both topological and branch length information as well as support information (i.e.
+posterior probabilities, e.g.
+\citet{Ronquist2012mrbayes}).
+However, using a Bayesian consensus tree has limitation especially if the resulting consensus tree is not well supported or resolved.
+Because the metrics used in this study are used to measure variation in topology between two trees (i.e.
+taxa placement or clade position), comparing two Bayesian consensus trees is not optimal in picking topological difference signal.
+For example, if one of the trees is not resolved at all (i.e. a star tree), comparing it to any other tree (even an other star tree) will not give useful results.
+Therefore we used the entire Bayesian trees posterior distribution in order to perform the tree comparisons (i.e.
+the "Best" Bayesian tree posterior distribution vs.
+the "missing-data" Bayesian tree posterior distribution).
+
+\subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
+We compared the distribution using a Random Pairwise Bayesian Tree Comparison (RPBTC) method.
+This method consists in comparing a series of randomly selected pairs of trees from two posterior distributions (on from each distribution) and use the mode to summarize the resulting distribution as a proxy of the difference between the two trees in a Bayesian framework.
+\begin{equation}
+RPBTC_{X,Y}=Mo(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk}))
+\end{equation}
+
+Where \textit{X} and \textit{Y} are two Bayesian tree posterior distribution;
+ $d_{m,n}$ is the pairwise difference for any metric \textit{m} and \textit{n} taxa between $X_{i}$ and $Y_{j}$ which are two single trees randomly sampled respectively from \textit{X} and \textit{Y}
+ and $Mo(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk}))$ is the mode of the pairwise difference repeated \textit{k} times.
+We used the mode to summarize the distributions because the value that is the most represented in the distribution reflects more accurately what a consensus tree represents (the topology the most represented in the posterior distribution).
+Also, the mode represents a value that is present in the distribution (which, depending on the metric, can be a series of discrete values) in contrast to the other distribution summary metrics, such as the mean or the median, which can take values that are not actually present in the distribution.
+For example, if comparing two posterior tree distribution, we obtain difference values of 10, 10, 10, 5, 5 and 1, using the mean (6.8) or the median (7.5) gives a values that actually doesn't exist
+
+In this study, we calculated RPBTC for both metrics (RF and Triple distance) for each pairs of Bayesian tree distributions (i.e.
+the "Best" Bayesian tree posterior distribution vs.
+the "missing-data" Bayesian tree posterior distribution) with 1000 random pairwise comparisons (\textit{k}).
+Because the RPBTC uses random pairwise comparisons, there is a chance of comparing always trees that have the biggest or the smallest difference between two distributions, either deflating or inflating the RPBTC difference value.
+However, using 1000 random pairwise comparisons makes the difference stable (i.e. if repeated independently, no difference is detected - see \hyperref[supplementaries]{supplementaries}).
+
+$\newline$
+
+We calculated the NTS for the "Best" Bayesian tree vs.
+"missing-data" Bayesian tree comparison for both RF and Triple metric using the mode of the RPBTC for each chain resulting in a distribution of 51 modes per comparison.
+We then compared the NTS of the Bayesian trees to the NTS of the ML trees for both metrics for each comparison.
+This resulted in a distribution of 51 NTS values for the Bayesian trees and 51 NTS values for the ML trees for each tree comparison for which we calculated the mode and the 50\% and 95\% confidence intervals (see \hyperref[supplementaries]{supplementaries} - Fig.
 ~\ref{Fig_Compare}).
@@
 \subsection{Building the trees}
-Bayesian chains converged (SD< 0.01 and ESS >> 200 for each tree).
+Bayesian chains converged (SD $<$ 0.01 and ESS $>>$ 200 for each tree).
 \subsection{Comparing topologies}
@@
 Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
+After a certain amount of morphological characters, adding new ones adds homoplasy \citep{wagner2000}.
 
@@
 
-\section{Supplementaries}\label{supplementaries}
+\section{Supplementaries}
+\label{supplementaries}
 \subsection{Morphological characters states}
@@
 
-\subsection{Random Bayesian tree comparisons}
+\subsection{Tree comparisons}
+\subsubsection{Random tree comparison scaling}
+We used the comparison of 1000 random trees to obtain the mean comparison value $\bar{d}_{m,n}$\textit{(rand)} for the NTS metric.
+We randomly generated two sets of 1000 trees of \textit{n} taxa using the rmtree function of ape package (v3.0-11 \citet{paradisape:2004}) that generates a given number of random Yule trees.
+We calculated the $\bar{d}_{m,n}$\textit{(rand)} value using an approach similar to the RPCBTC (described below) by performing 1000 random pairwise comparisons using the TreeCmp java script \citep{Bogdanowicz2012}.
+
+\subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
 We assessed the power of the Random Pairwise Bayesian Tree Comparison (RPBTC) method by comparing 1000 random trees from a posterior distribution trees set to another 1000 random trees from the same posterior distribution trees set.

### 2014-05-23 Analysis/MissingDataEffect.R
MissingDataAnalysis Script

@@ -1,0 +1,79 @@
+#Script for testing the effect of missing data on tree topology.
+
+#Loading the functions
+source('../Functions/TreeComparisons/NTS.R')
+source('../Functions/TreeComparisons/TreeCmp.Plot.R')
+source('../Functions/TreeComparisons/TreeCmp.Read.R')
+
+#Loading the Random tree comparisons
+Random51<-read.table("../Data/Tree_Comparisons/Random_trees/51t_Random.Cmp",header=TRUE)
+Random26<-read.table("../Data/Tree_Comparisons/Random_trees/26t_Random.Cmp",header=TRUE)
+Random25<-read.table("../Data/Tree_Comparisons/Random_trees/25t_Random.Cmp",header=TRUE)
+
+#Loading the treesets
+#/TreeCmp_chains/
+
+#ML chains
+#/ML/
+
+#Bayesian chains
+#/Bayesian/
+
+#Plots
+
+#Setting the metric
+metric="R.F_Cluster"
+#or "Triples"
+
+#MLiving
+TreeCmp.Plot( <chain.Bayesian> , metric, probs=c(95, 50), col=palette()[1], lines=TRUE, add=FALSE, shift=0, save.details=FALSE)
+TreeCmp.Plot( <chain.ML> , metric, probs=c(95, 50), col=palette()[2], lines=TRUE, add=TRUE, shift=0.1, save.details=FALSE)
+legend( <x> , <y> , c("Bayesian", "ML"), pch=21, col=c(palette()[1:2]), pt.bg=c(palette()[1:2]), cex=0.8)
+legend( <x> , <y> , c("Mode", "50% CI", "95% CI"), col="grey", lty=c(NA, 1, 2), pch=c(21, NA, NA), pt.bg=c("grey"), cex=0.8)
+
+#MFossil
+TreeCmp.Plot( <chain.Bayesian> , metric, probs=c(95, 50), col=palette()[1], lines=TRUE, add=FALSE, shift=0, save.details=FALSE)
+TreeCmp.Plot( <chain.ML> , metric, probs=c(95, 50), col=palette()[2], lines=TRUE, add=TRUE, shift=0.1, save.details=FALSE)
[…]

### 2014-05-23 Manuscript/TEM_draft.tex
Typo updates

@@
 the fossil occurrence age only or both the age and the morphology).
-Classical cladistic methods uses matrices containing morphological data from both fossil and living taxa and allows to treat each taxon as a tip in the phylogeny; the relation between the taxa can be infered using optimal criteria such as the maximum parsimony criterion \citep{simpson1945}.
+Classical cladistic methods uses matrices containing morphological data from both fossil and living taxa and allows to treat each taxon as a tip in the phylogeny; the relation between the taxa can be inferred using optimal criteria such as the maximum parsimony criterion \citep{simpson1945}.
 This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species.
@@
 \label{step:generate_matrix}
-We built a randomly generated birth death tree (hereafter called the "true" tree, Table ~\ref{Tab_glossary}) to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix, Table ~\ref{Tab_glossary}).
+We randomly generated birth death tree (hereafter called the "true" tree, Table ~\ref{Tab_glossary}) to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix, Table ~\ref{Tab_glossary}).
 \item{Removing data:} \\
 \label{step:remove_data}
-We removed data from the "complete" matrix to simulate the effects of missing data by modifying three parameters (1) the proportion of missing living taxa ($M_{L}$), (2) the proportion of missing data in the fossil taxa ($M_{F}$) and (3) the proportion of missing morphological characters ($M_{C}$) (the resulting matrix is called hereafter the "missing-data" matrix, Table ~\ref{Tab_glossary}).
+We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (1) the proportion of missing living taxa ($M_{L}$), (2) the proportion of missing data in the fossil taxa ($M_{F}$) and (3) the proportion of missing morphological characters ($M_{C}$) (the resulting matrices are called hereafter the "missing-data" matrix, Table ~\ref{Tab_glossary}).
 \item{Building phylogenies:} \\
@@
 \end{enumerate}
-To measure the effect of missing data distribution, we repeated steps ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} with the exact same fixed parameters 50 times.
+To measure the effect of missing data distribution, we repeated steps ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} with the exact same fixed parameters 51 (3$\times$17) times.
+%that's because I had 3 cluster accounts available...
 A list of all the terms used in this paper is available in table ~\ref{Tab_glossary}.
@@
 \caption{Protocol outline.
-(1) We generated a random tree (hereafter called the "true" tree) to infer a matrix with no missing data (hereafter, the "complete" matrix).
-(2) We removed data fron the morphological partition of the "Complete" matrix resulting in 124 matrices with missing data.
-(3) We build Bayesian phylogenetic trees from each matrix.
-(4) We then compared the trees with missing data (from the matrices with missing data) to the tree with no missing data (hereafter, the "best" tree – infered from the "complete" matrix).
+(1) We generated a random tree (the "true" tree) to infer a matrix with no missing data (the "complete" matrix).
+(2) We removed data fron the morphological partition of the "complete" matrix resulting in 124 "missing-data" matrices.
+(3) We infer phylogenetic trees from each matrix in both ML and Bayesian framework.
+(4) We then compared the "missing-data" trees to the "best" tree.
 We repeated step 1 to 4 51 (3$\times $17) times.}
@@
 \hfill fossil taxa & taxa with only morphological data available \\[1.5ex]
-\hfill "complete" matrix & matrix with no missing data except for the molecular data for the fossil taxa \\[1.5ex]
+\hfill "complete" matrix & matrix with no missing data except for the molecular part of the fossil taxa \\[1.5ex]
 \hfill "missing-data" matrix & matrix with various amount of missing data \\[1.5ex]
@@
 We generated the tree using a Birth Death process by sampling the values of the speciation events ($\lambda$) and extinction events ($\mu$) from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
-We implemented a rejection sampling algorithm to select only random trees with 25 living and 25 fossil taxa.
-We then added a species to the resulting Birth-Death tree as the outgroup of the tree.
+We implemented a rejection sampling algorithm to select only random trees with 25 living and 25 fossil taxa. %Justification needed?
+We then added a taxa to the resulting Birth-Death tree as the outgroup of the tree.
 The mean branch length of the tree was used to separate the outgroup from the rest of the taxa and the branch length leading to the outgroup was set as the sum of the mean branch length and the longest root-to-tip length of the tree.
@@
 Next, we created a molecular and a morphological matrix from the "true" tree.
-The molecular matrix was infered from the "true" tree using the package phyclust v0.1-14 \citep{chen2011}.
+The molecular matrix was inferred from the "true" tree using the package phyclust v0.1-14 \citep{chen2011}.
 The matrix was made of 1000 characters sites for 51 taxa and generated using the seqgen algorithm \citep{ranbaut1997seqgen}.
@@
 The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
-We chose a low value of $\alpha$ to lower the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
-These parameters were selected to generate data with no special assumption about how the characters evolved as well as to reduce the computational time required if these parameters were estimated rather than defined (65 CPU years).
+We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
+These parameters were selected to generate data with no special assumption about how the characters evolved as well as to reduce the computational time required if these parameters were estimated rather than defined (total computational time $>$ 65 CPU years).
 
@@
 evolutionary rate) from one character state to an other \citep{Pagel22011994}.
-This method allows us to have only two parameters per character i.e. the number of states and the evolutionary rate.
+This method allows us to have only two parameters per character: the number of states and the evolutionary rate.
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
@@
 Finally, we combined the morphological and molecular matrices obtained from the "true" tree.
-Hereafter we call this the "complete" matrix, i.e. the matrix with no missing data except for the molecular data of the fossil taxa.
+Hereafter we call this the "complete" matrix: the matrix with no missing data except for the molecular data of the fossil taxa.
 
 \subsection{Removing data}
-Once we obtained the "complete" matrix we modified it to get a set of matrices with missing data
+Once we obtained the "complete" matrix we modified it to get a set of matrices with missing data. %Ugly sentence
 We randomly replaced data with "?" in the morphological part of the matrices according to the following parameters (Fig. ~\ref{Fig_RemoveData}):
@@
 This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological one.
-Because of the increasing facility to sequence DNA for living species, the number of living species with molecular data is highly superior the the number of species with molecular and morphological data.
+Because of the increasing facility to sequence DNA for living taxa, the number of living taxa with molecular data is highly superior the the number of taxa with molecular and morphological data.
 \item{The proportion of missing morphological data across all fossil taxa ($M_{F}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
@@
 
-In practice, each parameter represent a way of removing data in the morphological part of the matrix: $M_{L}$ removes a proportion of rows from the living taxa; $M_{F}$ removes a proportion of cells from the fossil taxa; and $M_{C}$ removes a proportion of columns across both living and fossil taxa.
+In practice, each parameter represent a different way of removing data in the morphological part of the matrix: $M_{L}$ removes a proportion of rows from the living taxa; $M_{F}$ removes a proportion of cells from the fossil taxa; and $M_{C}$ removes a proportion of columns across both living and fossil taxa.
 Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a proportion of the living taxa is removed (i.e.
@@
 Because some parameter combinations introduce a lot of missing (e.g. $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
-When this occurred we repeated the random deletion of characters until every species had at least 5\% data.
+When this occurred we repeated the random deletion of characters until every taxa had at least 5\% data across the whole matrix.
 
@@
 \subsubsection{Maximum Likelihood}
-The "best" tree and the "missing-data" trees were infered using RAxML v8.0.20 \citep{Stamatakis21012014}.
+The "best" tree and the "missing-data" trees were inferred using RAxML v8.0.20 \citep{Stamatakis21012014}.
 For the molecular data, we used the GTR + $\Gamma_4$ model (\citet{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citet{Stamatakis21012014}) as a generalisation of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
 The GTR model can be seen as a generalisation of the HKY model (the 2 parameters from the HKY model are implicitly included in the 6 from GTR model - \citet{stamatakisa2008}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} wich is a generalisation of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014})
+For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} wich is a generalisation of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014}).
 We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. 
-The bootstrap algorithm used in RAxML is the Lazy Subtree Rearrangement (LSR) which consists in pruning one subtree from the tree and subsequently reinserting it to al neighboring branches \citep{stamatakisa2008}.
+The bootstrap algorithm used in RAxML is the Lazy Subtree Rearrangement (LSR) which consists in pruning one subtree from the tree and subsequently reinserting it to al neighbouring branches \citep{stamatakisa2008}.
 Subtree Pruning and Reinserting methods (SPR) have been demonstrated as being better than others (e.g. Nearest Neighbouring Interchange - NNI) in recovering good bootstrap values \citep{salamin2003}.
@@
 \subsubsection{Bayesian}
-The "best" tree and the "missing-data" trees were infered using MrBayes v3.2.2 \citep{Ronquist2012mrbayes}.
-We partitioned the data treating the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
-The molecular evolutionary history was infered using the HKY model with a transition/transversion ratio \citep{douadycomparison2003} of two and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
+The "best" tree and the "missing-data" trees were inferred using MrBayes v3.2.2 \citep{Ronquist2012mrbayes}.
+We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
+The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \citep{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
 For the morphological data, we used the Markov \textit{k} state model \citep{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
@@
 Each Bayesian tree was estimated using two runs of four chains each for a maximum of 50$\times$$1^6$ generations.
-We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and implemented a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
-The effective sample size (ESS) was also checked on a random subsample of runs in each simulation to check that ESS $>>$ 200 \citep{drummond2006ess}.
-For each run, we removed 25\% of the iterations as burnin. We used the following priors for each tree (see \hyperref[supplementaries]{supplementaries}):
+We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
+The effective sample size (ESS) was also checked on a random subsample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
+For each run, we removed 25\% of the iterations as burnin.
+We used the following priors for each tree (see \hyperref[supplementaries]{supplementaries}):
 \begin{enumerate}
@@
 We used these priors to speed up the Bayesian process.
-These priors biased the way the Bayesian process calculated the branch length by giving non-random starting points and boundaries for the parameters estimation process, however, we are focusing on the effect of missing data on the topology and not on the branch length.
+These priors biased the way the Bayesian process calculated the branch length by giving non-random starting points and boundaries for the parameters estimation process, however, in this study, we focused on the effect of missing data on the topology and not on the branch length.
 Even using these priors, it took 65 CPU years to build 51 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
+
+%
+%
+%
 
@@
 We compared the topology of the "missing-data" trees inferred from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
-Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees infered from matrices with living and fossil taxa and molecular and morphological characters.
+Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees inferred from matrices with living and fossil taxa and molecular and morphological characters.
 To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triple \citep{dobson1975triplets} and the Robison-Fould \citep{RF1981} distance.
@@
 \caption{Tree comparison protocol.
-From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we infered a tree in Baysian or Maximum likelihood framework.
+From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we inferred a tree in Baysian or Maximum likelihood framework.
 We then compared this tree with or without fossil/living taxa to the "best/true" tree.
@@
 Transition/Transversion ratio = 2 ($\beta$(80,40)) \\
+Starting tree: "True" tree topology with each branch length = 1 \\
 Morphological data: \\

### 2014-05-23 Analysis/MissingDataEffect.R
MissingDataAnalysis Script

@@ -17,7 +17,11 @@ Random25<-read.table("../Data/Tree_Comparisons/Random_trees/25t_Random.Cmp",head
 #/ML/
 
 #Bayesian chains
-#/Bayesian/
+cd("../Data/Tree_Comparisons/TreeCmp_chains/Bayesian/")
+
+cd('../../../../Analysis/')
+
+
 
 #Plots
 

### 2014-05-27 Functions/TreeBuilding/TEM_split-tasker.sh
v04 Cluster Timing decreased from 4 to 1

@@ -13,9 +13,10 @@
 #########################
 #WARNING: this version only deals with nexus files generated by TEM_treesim.sh with the following parameters (25 living taxa, 25 fossil taxa, 1 outgroup - 1000 molecular characters, 100 morphological characters).
 #WARNING: this version only uses only the matrix with no missing data L00C00F00
-#version: 0.3
+#version: 0.4
 #Update: Correction in mpirun -np 8 mb ... added chain number after folder name ($folder$n)
 #Update: Correction in task file, mb is now run directly in $folder$n
+#Update: line 59 - Cluster timing is set to a maximum of one day (instead of 4)
 #----
 #guillert(at)tcd.ie - 16/05/2014
 ##########################
@@ -55,7 +56,7 @@ do
     #JOB TEMPLATE
     echo "/#!/bin/sh
 #SBATCH -n 8
-#SBATCH -t 4-00:00:00
+#SBATCH -t 1-00:00:00
 #SBATCH -p compute
 #SBATCH -J S_${n}
 source /etc/profile.d/modules.sh

### 2014-05-27 Functions/TreeComparisons/TreeCmp.Plot.R
v1.1 allows to fix ylim

@@ -3,7 +3,7 @@
 ##########################
 #Calculate the CI and the mode of a given metric using TreeCmp class objects
 #Function modified from sumBTC (guillert(at)tcd.ie - 19/02/2014)
-#v.1.0
+#v.1.1
 #Update: the three core functions have been isolated
 #Update: in FUN.hdr, allows to read data from similar tree comparison (NTS=1), setting $hdr and $alpha to NA and $mode to 1 if var(data) == 0
 #Update: in FUN.densityplot, allows to plot data from similar tree comparison (NTS=1), ignoring $hdr and $alpha and ploting only the mode (=1)
@@ -11,6 +11,7 @@
 #Update: colours are now accepted
 #Update: results can be plotted as lines
 #Update: allow to stack the plots with add=TRUE
+#Update: allow window scaling option (ylim)
 ##########################
 #SYNTAX :
 #<data> an object of the class TreeCmp
@@ -21,10 +22,11 @@
 #<lines> whether to plot the results as lines instead than as boxplots (default=FALSE). Is ignored if plot=FALSE.
 #<add> whether to add to the current plot (default=FALSE). Is ignored if plot=FALSE.
 #<shift> a numerical value between 0 and 0.6 for shifting the lines. Is ignored if plot=FALSE, lines=FALSE, add=FALSE.
+#<ylim> two numerical values to scale the y axis of the plot. If set to 'auto' (default), the window is automatically scaled with the minimum and the maximum value from the data.
 #<save.details> whether to save the details of each comparison (default=FALSE). If TRUE, saves a density plot for each comparison. The chain name will be the one given in 'data'. Is ignored if plot=FALSE.
 ##########################
 #----
-#guillert(at)tcd.ie - 08/04/2014
+#guillert(at)tcd.ie - 27/05/2014
 ##########################
 #Requirements:
 #-R 3
@@ -32,7 +34,7 @@
 #-R TreeCmp objects
 ##########################
 
-TreeCmp.Plot<-function(data, metric, probs=c(95, 75, 50), plot=TRUE, col='black', lines=FALSE, add=FALSE, shift=0, save.details=FALSE) {
+TreeCmp.Plot<-function(data, metric, probs=c(95, 75, 50), plot=TRUE, col='black', lines=FALSE, add=FALSE, shift=0, ylim='auto', save.details=FALSE) {
 
 #HEADER
 
[…]

### 2014-05-27 Functions/TreeComparisons/TreeCmp.Plot.R
Type in the help (header)

@@ -15,7 +15,7 @@
 ##########################
 #SYNTAX :
 #<data> an object of the class TreeCmp
-#<metric.number> the name of the metric of interest
+#<metric> the name of the metric of interest
 #<probs> a vector probabilities levels (default = c(95,75,50)).
 #<plot> whether to plot the results or not (default = TRUE).
 #<col> a colour for the ploting (default="black"). Is ignored if plot=FALSE.

### 2014-05-27 Functions/TreeComparisons/TreeCmp.anova.R
Runs anova and TukeyHSD on TreeCmp object

@@ -1,0 +1,150 @@
+##########################
+#Anova on TreeCmp analysis
+##########################
+#Performs an anova on a 'TreeCmp' object
+#Performs a TukeyHSD if a posthoc difference is significant
+#v.0.1
+##########################
+#SYNTAX :
+#<data> an object of the class 'TreeCmp'
+#<metric> the name of the metric of interest
+#<plot> whether to plot the distribution of the metric as a boxplot (default=FALSE)
+#<LaTeX> whether to output a LaTeX table (default=FALSE)
+##########################
+#----
+#guillert(at)tcd.ie - 27/05/2014
+##########################
+#Requirements:
+#-R 3
+#-R package 'xtable' [optional]
+##########################
+
+TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE) {
+
+#HEADER
+
+#DATA INPUT
+
+    #TreeCmp
+    if(class(TreeCmp) != 'TreeCmp') {
+        stop('The input must be a "TreeCmp" class object')
+    } else {
+        length.TreeCmp<-length(TreeCmp) #The number of comparisons
+        columns.TreeCmp<-length(TreeCmp[[1]]) #The number of columns per comparisons (i.e. the number of metrics)
+        replicates.TreeCmp<-length(TreeCmp[[1]][[1]]) #The number of replicates per comparison
+    }
+
+    #metric
+    if(class(metric) != 'character') {
+        stop('Provided metric name is not found in the "TreeCmp" object')
+    } else {
+        if(length(metric) != 1) {
+            stop('Only one metric name can be provided')
+        } else {
+            metric.column<-grep(metric, colnames(TreeCmp[[1]]))
+            if(length(metric.column) == 0) {
+                stop('Provided metric name is not found in the "TreeCmp" object')
+            } else {
+                TreeCmp.rows<-NULL
+                for (i in 1:length.TreeCmp) {
+                    TreeCmp.rows[i]<-length(TreeCmp[[i]][,metric.column])
+                }
+            }
+        }
+    }
+
+    #plot
+    if(class(plot) != 'logical'){
+        stop('Plot is not logical')
+    }
+
+    #LaTeX
+    if(class(LaTeX) != 'logical'){
+        stop('LaTeX is not logical')
+    } else {
+        if(LaTeX == TRUE){
[…]

### 2014-05-27 Analysis/Pariwise_TreeCmp.R
Added TreeCmp.anova

@@ -1,9 +1,12 @@
 #Script for testing the Random Pairwise Bayesian Tree Comparisons (RPBTC)
+#cd('../PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
 
 #Loading the functions
 source('../Functions/TreeComparisons/NTS.R')
 source('../Functions/TreeComparisons/TreeCmp.Plot.R')
 source('../Functions/TreeComparisons/TreeCmp.Read.R')
+source('../Functions/TreeComparisons/TreeCmp.anova.R')
+
 
 #Loading the Random tree comparisons
 Random51<-read.table("../Data/Tree_Comparisons/Random_trees/51t_Random.Cmp",header=TRUE)
@@ -27,123 +30,21 @@ tmp<-TreeCmp.Read('HighScore', verbose=TRUE)
 HighScore<-NTS(tmp, Random26)
 cd('../../../../Analysis/')
 
-#Storing each NTS score in a data.frame object
-#RF score (NTS)
-LowScore.RF<-NULL
-for (replicate in 1:100){
-    LowScore.RF[[replicate]]<-LowScore[[replicate]][[4]]
-}
-LowScore.RF<-as.data.frame(LowScore.RF)
-names(LowScore.RF)<-seq(1:100)
+#TreeCmp anova per metric and per score
+LS.RF<-TreeCmp.anova(LowScore, "R.F_Cluster")
+MS.RF<-TreeCmp.anova(MediumScore, "R.F_Cluster")
+HS.RF<-TreeCmp.anova(HighScore, "R.F_Cluster")
+LS.Tr<-TreeCmp.anova(LowScore, "Triples")
+MS.Tr<-TreeCmp.anova(MediumScore, "Triples")
+HS.Tr<-TreeCmp.anova(HighScore, "Triples")
 
-#Triplets score (NTS)
-LowScore.Tr<-NULL
-for (replicate in 1:100){
-    LowScore.Tr[[replicate]]<-LowScore[[replicate]][[6]]
-}
-LowScore.Tr<-as.data.frame(LowScore.Tr)
-names(LowScore.Tr)<-seq(1:100)
-
-#RF score (NTS)
-MediumScore.RF<-NULL
-for (replicate in 1:100){
[…]

### 2014-05-27 Analysis/MissingDataEffect.R
Added comparisons

@@ -7,6 +7,7 @@ source('../Functions/TreeComparisons/NTS.R')
 source('../Functions/TreeComparisons/TreeCmp.Plot.R')
 source('../Functions/TreeComparisons/TreeCmp.Read.R')
 source('../Functions/TreeComparisons/TreeCmp.Mode.R')
+source('../Functions/TreeComparisons/TreeCmp.anova.R')
 
 
 #Loading the Random tree comparisons
@@ -59,56 +60,53 @@ Bayesian.MLMFMC<-Bayesian_trees
 class(Bayesian.MLMFMC)<-class(Bayesian_trees)
 Bayesian.MLMFMC<-TreeCmp.Mode(Bayesian.MLMFMC, 1000, verbose=TRUE)
 
-
-#Transforming into lists of 51 modes.
-
-
-
 #Plots
 
 #Setting the metric
 metric="R.F_Cluster"
-#or "Triples"
+#metric="Triples"
+#fix ylim
+ylim=c(0,1)
 
 #MLiving
-TreeCmp.Plot( Bayesian.ML , metric, probs=c(95, 50), col=palette()[1], lines=TRUE, add=FALSE, shift=0, save.details=FALSE)
-TreeCmp.Plot( <chain.ML> , metric, probs=c(95, 50), col=palette()[2], lines=TRUE, add=TRUE, shift=0.1, save.details=FALSE)
+TreeCmp.Plot( Bayesian.ML , metric, probs=c(95, 50), col=palette()[1], lines=TRUE, ylim=ylim)
+TreeCmp.Plot( <chain.ML> , metric, probs=c(95, 50), col=palette()[2], lines=TRUE, add=TRUE, shift=0.1, ylim=ylim, save.details=FALSE)
 legend( <x> , <y> , c("Bayesian", "ML"), pch=21, col=c(palette()[1:2]), pt.bg=c(palette()[1:2]), cex=0.8)
 legend( <x> , <y> , c("Mode", "50% CI", "95% CI"), col="grey", lty=c(NA, 1, 2), pch=c(21, NA, NA), pt.bg=c("grey"), cex=0.8)
 
 #MFossil
-TreeCmp.Plot( Bayesian.MF , metric, probs=c(95, 50), col=palette()[1], lines=TRUE, add=FALSE, shift=0, save.details=FALSE)
[…]

### 2014-05-27 Manuscript/TEM_draft.tex
Table RPBTC corrections

@@
   \hline
-  Tree.Type & Used.metric & Replicates & Df & F.value & p.value \\ [0.5ex]
+  Tree.Type & Used.metric & Replicates & Df & F.value & p.value \\ 
   \hline
-  Low Score & RF & 100.00 &   1 & 0.07 & 0.79 \\ [0.5ex]
-  Low Score & Tr & 100.00 &   1 & 0.33 & 0.56 \\ [0.5ex]
-  Medium Score & RF & 100.00 &   1 & 1.07 & 0.30 \\ [0.5ex]
-  Medium Score & Tr & 100.00 &   1 & 0.00 & 0.96 \\ [0.5ex]
-  High Score & RF & 100.00 &   1 & 0.20 & 0.66 \\ [0.5ex]
-  High Score & Tr & 100.00 &   1 & 0.61 & 0.43 \\ [0.5ex]
+  Low Score & RF & 100.00 & 99.00 & 0.74 & 0.98 \\ 
+  Low Score & Tr & 100.00 & 99.00 & 0.97 & 0.58 \\ 
+  Medium Score & RF & 100.00 & 99.00 & 0.64 & 1.00 \\ 
+  Medium Score & Tr & 100.00 & 99.00 & 0.45 & 1.00 \\ 
+  High Score & RF & 100.00 & 99.00 & 0.20 & 1.00 \\ 
+  High Score & Tr & 100.00 & 99.00 & 0.37 & 1.00 \\ 
   \hline
 \end{tabular}
+\end{table}
 \label{RPBTC_testing}
-\end{table}
 

### 2014-05-27 Analysis/MissingDataEffect.R
Anova check updates

@@ -111,26 +111,27 @@ legend( <x> , <y> , c("Bayesian", "ML"), pch=21, col=c(palette()[1:2]), pt.bg=c(
 legend( <x> , <y> , c("Mode", "50% CI", "95% CI"), col="grey", lty=c(NA, 1, 2), pch=c(21, NA, NA), pt.bg=c("grey"), cex=0.8)
 
 
-
+bptest()
+shapiro.test()
 #Checking for differences
 #MLiving
-TreeCmp.anova(Bayesian.ML, "R.F_Cluster")
-TreeCmp.anova(Bayesian.ML, "Triples")
+TreeCmp.anova(Bayesian.ML, "R.F_Cluster", plot=TRUE) #Diff
+TreeCmp.anova(Bayesian.ML, "Triples") #No diff
 #MFossil
-TreeCmp.anova(Bayesian.MF, "R.F_Cluster")
-TreeCmp.anova(Bayesian.MF, "Triples")
+TreeCmp.anova(Bayesian.MF, "R.F_Cluster", plot=TRUE) #Diff
+TreeCmp.anova(Bayesian.MF, "Triples") #No diff
 #MCharacter
-TreeCmp.anova(Bayesian.MC, "R.F_Cluster")
-TreeCmp.anova(Bayesian.MC, "Triples")
+TreeCmp.anova(Bayesian.MC, "R.F_Cluster", plot=TRUE) #Diff
+TreeCmp.anova(Bayesian.MC, "Triples") #No diff
 #MLiving:MFossil
-TreeCmp.anova(Bayesian.MLMF, "R.F_Cluster")
-TreeCmp.anova(Bayesian.MLMF, "Triples")
+TreeCmp.anova(Bayesian.MLMF, "R.F_Cluster", plot=TRUE) #Diff
+TreeCmp.anova(Bayesian.MLMF, "Triples") #No diff
 #MLiving:MCharacter
-TreeCmp.anova(Bayesian.MLMC, "R.F_Cluster")
-TreeCmp.anova(Bayesian.MLMC, "Triples")
+TreeCmp.anova(Bayesian.MLMC, "R.F_Cluster", plot=TRUE) #Diff
+TreeCmp.anova(Bayesian.MLMC, "Triples") #No diff
 #MFossil:MCharacter
-TreeCmp.anova(Bayesian.MCMF, "R.F_Cluster")
[…]

### 2014-05-27 Functions/TreeComparisons/TreeCmp.anova.R

@@ -1,15 +1,17 @@
 ##########################
 #Anova on TreeCmp analysis
 ##########################
-#Performs an anova on a 'TreeCmp' object
-#Performs a TukeyHSD if a posthoc difference is significant
-#v.0.1
+#Performs an anova on a 'TreeCmp' object (parametric or non-parametric)
+#Performs a TukeyHSD if a posthoc difference is significant (parametric or non-parametric)
+#v.0.2
+#Update: automatic parametric/non-parametric test (and possibility to save the results)
 ##########################
 #SYNTAX :
 #<data> an object of the class 'TreeCmp'
 #<metric> the name of the metric of interest
 #<plot> whether to plot the distribution of the metric as a boxplot (default=FALSE)
 #<LaTeX> whether to output a LaTeX table (default=FALSE)
+#<save.test> whether to save the details of the various parametric/non-parametric tests (default=FALSE)
 ##########################
 #----
 #guillert(at)tcd.ie - 27/05/2014
@@ -17,9 +19,12 @@
 #Requirements:
 #-R 3
 #-R package 'xtable' [optional]
+#-R package 'pgirmess' DEVELOPEMENT VERSION ONLY
 ##########################
 
-TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE) {
+TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE, save.test=FALSE) {
+
+warning("Non-parametric test in development only!")
 
 #HEADER
 
@@ -67,6 +72,11 @@ TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE) {
         }
     }
 
+    #save.test
+    if(class(save.test) != 'logical'){
+        stop('Save.test is not logical')
+    }
+
 #FUNCTIONS
 
     FUN.data.table<-function(TreeCmp, metric, replicates.TreeCmp){
@@ -101,24 +111,107 @@ TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE) {
 
     }
 
+    FUN.parametric<-function(anova.table, save.test){
+        #Tests if the anova must be parametric or not
+         
+        #Test Normality
+        normality<-shapiro.test(anova.data[[2]])
+        p.normality<-normality[[2]]
+
+        #IN DEVELOPEMENT {
+
[…]

### 2014-05-28 Functions/TreeComparisons/TreeCmp.distribution.R
Plots the distribution of TreeCmp values

@@ -1,0 +1,87 @@
+##########################
+#TreeCmp distribution
+##########################
+#Plots the distribution of the metric of a 'TreeCmp' object
+#v.0.1
+##########################
+#SYNTAX :
+#<TreeCmp> an object of the class 'TreeCmp'
+#<metric> the name of the metric of interest
+##########################
+#----
+#guillert(at)tcd.ie - 28/05/2014
+##########################
+#Requirements:
+#-R 3
+##########################
+
+TreeCmp.distribution<-function(TreeCmp, metric) {
+
+#INPUT
+
+    #TreeCmp
+    if(class(TreeCmp) != 'TreeCmp') {
+        stop('The input must be a "TreeCmp" class object')
+    } else {
+        length.TreeCmp<-length(TreeCmp) #The number of comparisons
+        columns.TreeCmp<-length(TreeCmp[[1]]) #The number of columns per comparisons (i.e. the number of metrics)
+        replicates.TreeCmp<-length(TreeCmp[[1]][[1]]) #The number of replicates per comparison
+    }
+
+    #metric
+    if(class(metric) != 'character') {
+        stop('Provided metric name is not found in the "TreeCmp" object')
+    } else {
+        if(length(metric) != 1) {
+            stop('Only one metric name can be provided')
+        } else {
+            metric.column<-grep(metric, colnames(TreeCmp[[1]]))
+            if(length(metric.column) == 0) {
+                stop('Provided metric name is not found in the "TreeCmp" object')
+            } else {
+                TreeCmp.rows<-NULL
+                for (i in 1:length.TreeCmp) {
+                    TreeCmp.rows[i]<-length(TreeCmp[[i]][,metric.column])
+                }
+            }
+        }
+    }
+
+#FUNCTION
+
+    FUN.density.list<-function(TreeCmp, length.TreeCmp, metric.column){
+        #Calculates the density in a list
+        density.list<-NULL
+        max.den<-NULL
+        for (n in 1:length.TreeCmp){
+            density.list[[n]]<-density(TreeCmp[[n]][[metric.column]])
+            max.den[[n]]<-max(density.list[[n]]$y)
+        }
+        max.den<<-max(max.den) #Stores out of the function for the plot ylim
+
+        return(density.list)
+    }
+
+#PLOT THE DISTRIBUTION
+
+    #Calculating the density list
[…]

### 2014-05-28 Functions/TreeComparisons/TreeCmp.Plot.R
Changed input par. data > TreeCmp

@@ -5,8 +5,8 @@
 #Function modified from sumBTC (guillert(at)tcd.ie - 19/02/2014)
 #v.1.1
 #Update: the three core functions have been isolated
-#Update: in FUN.hdr, allows to read data from similar tree comparison (NTS=1), setting $hdr and $alpha to NA and $mode to 1 if var(data) == 0
-#Update: in FUN.densityplot, allows to plot data from similar tree comparison (NTS=1), ignoring $hdr and $alpha and ploting only the mode (=1)
+#Update: in FUN.hdr, allows to read TreeCmp from similar tree comparison (NTS=1), setting $hdr and $alpha to NA and $mode to 1 if var(TreeCmp) == 0
+#Update: in FUN.densityplot, allows to plot TreeCmp from similar tree comparison (NTS=1), ignoring $hdr and $alpha and ploting only the mode (=1)
 #Update: the number of probabilities level don't need to be 3 any more
 #Update: colours are now accepted
 #Update: results can be plotted as lines
@@ -14,7 +14,7 @@
 #Update: allow window scaling option (ylim)
 ##########################
 #SYNTAX :
-#<data> an object of the class TreeCmp
+#<TreeCmp> an object of the class TreeCmp
 #<metric> the name of the metric of interest
 #<probs> a vector probabilities levels (default = c(95,75,50)).
 #<plot> whether to plot the results or not (default = TRUE).
@@ -22,8 +22,8 @@
 #<lines> whether to plot the results as lines instead than as boxplots (default=FALSE). Is ignored if plot=FALSE.
 #<add> whether to add to the current plot (default=FALSE). Is ignored if plot=FALSE.
 #<shift> a numerical value between 0 and 0.6 for shifting the lines. Is ignored if plot=FALSE, lines=FALSE, add=FALSE.
-#<ylim> two numerical values to scale the y axis of the plot. If set to 'auto' (default), the window is automatically scaled with the minimum and the maximum value from the data.
-#<save.details> whether to save the details of each comparison (default=FALSE). If TRUE, saves a density plot for each comparison. The chain name will be the one given in 'data'. Is ignored if plot=FALSE.
+#<ylim> two numerical values to scale the y axis of the plot. If set to 'auto' (default), the window is automatically scaled with the minimum and the maximum value from the TreeCmp.
[…]

### 2014-05-28 Functions/TreeComparisons/TreeCmp.anova.R
v0.3 improved/corrected non-parametric test

@@ -3,8 +3,9 @@
 ##########################
 #Performs an anova on a 'TreeCmp' object (parametric or non-parametric)
 #Performs a TukeyHSD if a posthoc difference is significant (parametric or non-parametric)
-#v.0.2
+#v.0.3
 #Update: automatic parametric/non-parametric test (and possibility to save the results)
+#Update: improved/corrected parametric test
 ##########################
 #SYNTAX :
 #<data> an object of the class 'TreeCmp'
@@ -14,19 +15,15 @@
 #<save.test> whether to save the details of the various parametric/non-parametric tests (default=FALSE)
 ##########################
 #----
-#guillert(at)tcd.ie - 27/05/2014
+#guillert(at)tcd.ie - 28/05/2014
 ##########################
 #Requirements:
 #-R 3
+#-R package 'pgirmess'
 #-R package 'xtable' [optional]
-#-R package 'pgirmess' DEVELOPEMENT VERSION ONLY
 ##########################
 
 TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE, save.test=FALSE) {
-
-warning("Non-parametric test in development only!")
-
-#HEADER
 
 #DATA INPUT
 
@@ -79,7 +76,7 @@ warning("Non-parametric test in development only!")
 
 #FUNCTIONS
 
-    FUN.data.table<-function(TreeCmp, metric, replicates.TreeCmp){
+    FUN.data.table<-function(TreeCmp, metric, replicates.TreeCmp) {
         #makes a data frame of all the replicates for one metric for all the comparisons (length must be length.TreeCmp*replicates.TreeCmp)
         data.table<-NULL
         metric.number<-which(names(TreeCmp[[1]]) == metric)
@@ -111,55 +108,56 @@ warning("Non-parametric test in development only!")
 
     }
 
-    FUN.parametric<-function(anova.table, save.test){
+    FUN.parametric<-function(anova.table, save.test) {
         #Tests if the anova must be parametric or not
          
         #Test Normality
-        normality<-shapiro.test(anova.data[[2]])
-        p.normality<-normality[[2]]
+        if(length(anova.data[[2]]) > 5000){
+            #Sub-sample the list of values for shapiro test (if > 5000 values)
+            sub.sample<-sample(anova.data[[2]], 5000)
+            normality<-shapiro.test(sub.sample)
+            p.normality<-normality[[2]]
 
[…]

### 2014-05-28 Functions/TreeComparisons/TreeCmp.anova.R
Debug removed

@@ -168,10 +168,6 @@ TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE, save.test=FALS
     anova.data<-FUN.anova.data(data.table, length.TreeCmp, replicates.TreeCmp)
     parametric<-FUN.parametric(anova.data, save.test)
 
-    #DEBUG
-
-    parametric<-TRUE
-
     if(parametric == TRUE) {
         #Test is parametric
 

### 2014-05-28 Functions/TreeComparisons/TreeCmp.anova.R
Updated LaTeX output

@@ -154,7 +154,7 @@ TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE, save.test=FALS
                 cat("Test is non-parametric. \n")
             } else {
                 parametric<-TRUE
-                cat("Test is parametric. \n")               
+                cat("Test is parametric. \n")
             }
         }
 
@@ -178,6 +178,7 @@ TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE, save.test=FALS
         #If there posthoc difference
         if(pvalue < 0.05){
             posthoc<-TukeyHSD(aov)
+            posthoc<-posthoc$factors
         }
 
     } else {
@@ -196,6 +197,9 @@ TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE, save.test=FALS
             posthoc<-posthoc$dif.com
         }
 
+        #Transform the Kruskal-Wallis output for LaTeX output
+        anova<-data.frame("Kruskal-Wallis chi-squared"=anova[[1]][[1]], "df"=anova[[2]][[1]], "p-value"=anova[[3]])
+
     }
 
     #Plot (optional)
@@ -206,13 +210,13 @@ TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE, save.test=FALS
 #OUTPUT
 
     if(LaTeX == TRUE){
-        TreeCmp.xtable<<-xtable(aov)
-        cat('LaTeX anova table available in : \n TreeCmp.xtable',"\n")
-
         if(pvalue < 0.05){
-            TreeCmp.xtables<<-list(anova=xtable(aov), posthoc=xtable(posthoc$factors))
-            cat('LaTeX anova table available in : \n TreeCmp.xtables$anova',"\n")
-            cat('LaTeX TukeyHSD table available in : \n TreeCmp.xtables$TuckeyHSD',"\n")
+            TreeCmp.xtables<<-list(group=xtable(anova), posthoc=xtable(posthoc))
+            cat('LaTeX group table available in : \n TreeCmp.xtables$group',"\n")
+            cat('LaTeX posthoc table available in : \n TreeCmp.xtables$posthoc',"\n")
+        } else {
+            TreeCmp.xtable<<-xtable(anova)
+            cat('LaTeX anova table available in : \n TreeCmp.xtable',"\n")
         }
     }
 

### 2014-05-28 Analysis/MissingDataEffect.R
Updated script with TreeCmp.distribution

@@ -8,6 +8,7 @@ source('../Functions/TreeComparisons/TreeCmp.Plot.R')
 source('../Functions/TreeComparisons/TreeCmp.Read.R')
 source('../Functions/TreeComparisons/TreeCmp.Mode.R')
 source('../Functions/TreeComparisons/TreeCmp.anova.R')
+source('../Functions/TreeComparisons/TreeCmp.distribution.R')
 
 
 #Loading the Random tree comparisons
@@ -111,27 +112,28 @@ legend( <x> , <y> , c("Bayesian", "ML"), pch=21, col=c(palette()[1:2]), pt.bg=c(
 legend( <x> , <y> , c("Mode", "50% CI", "95% CI"), col="grey", lty=c(NA, 1, 2), pch=c(21, NA, NA), pt.bg=c("grey"), cex=0.8)
 
 
-bptest()
-shapiro.test()
+
+
 #Checking for differences
 #MLiving
-TreeCmp.anova(Bayesian.ML, "R.F_Cluster", plot=TRUE) #Diff
-TreeCmp.anova(Bayesian.ML, "Triples") #No diff
+#Distribution check (visual)
+TreeCmp.anova(Bayesian.ML, "R.F_Cluster", plot=TRUE, LaTeX=TRUE) #Diff
+TreeCmp.anova(Bayesian.ML, "Triples", plot=TRUE, LaTeX=TRUE) #No diff
 #MFossil
 TreeCmp.anova(Bayesian.MF, "R.F_Cluster", plot=TRUE) #Diff
-TreeCmp.anova(Bayesian.MF, "Triples") #No diff
+TreeCmp.anova(Bayesian.MF, "Triples", plot=TRUE) #No diff
 #MCharacter
 TreeCmp.anova(Bayesian.MC, "R.F_Cluster", plot=TRUE) #Diff
-TreeCmp.anova(Bayesian.MC, "Triples") #No diff
+TreeCmp.anova(Bayesian.MC, "Triples", plot=TRUE) #No diff
 #MLiving:MFossil
 TreeCmp.anova(Bayesian.MLMF, "R.F_Cluster", plot=TRUE) #Diff
-TreeCmp.anova(Bayesian.MLMF, "Triples") #No diff
+TreeCmp.anova(Bayesian.MLMF, "Triples", plot=TRUE) #No diff
 #MLiving:MCharacter
[…]

### 2014-05-28 Manuscript/TEM_draft.tex
Added statistical comparisons and pre results

@@
 
-%Abstract
+%---------------------------------------------
+%
+%       ABSTRACT
+%
+%---------------------------------------------
+
 \subsubsection{Abstract}
@@
 \newpage
+
+%---------------------------------------------
+%
+%       INTRODUCTION
+%
+%---------------------------------------------
 
@@
 
+%---------------------------------------------
+%
+%       METHODS
+%
+%---------------------------------------------
+
 \section{Methods} %no material needed?
@@
 %that's because I had 3 cluster accounts available...
+If it's a too odd number, no worries, I'll just remove the last chain.
 A list of all the terms used in this paper is available in table ~\ref{Tab_glossary}.
@@
 
+%---------------------------------------------
+%       Generating the matrix
+%---------------------------------------------
+
 \subsection{Generating the matrix}
@@
 Hereafter we call this the "complete" matrix: the matrix with no missing data except for the molecular data of the fossil taxa.
+
+%---------------------------------------------
+%       Removing data
+%---------------------------------------------
 
@@
 When this occurred we repeated the random deletion of characters until every taxa had at least 5\% data across the whole matrix.
+
+%---------------------------------------------
+%       Building phylogenies
+%---------------------------------------------
 
@@
 
-%
-%
-%
+%---------------------------------------------
+%       Comparing Topologies
+%---------------------------------------------
 
@@
 If $S_n$=0, the trees are the same (i.e. no taxa as been displaced). When $S_n$ = $\binom{n}{4}$, the trees are the most different possible (i.e. every taxa as been displaced).
-This metric therefore illustrates the amount of displaxed taxa but is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g.
+This metric therefore illustrates the amount of displayed taxa but is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g.
 fossil taxa) than the Robinson-Foulds distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and can therefore be used as a proxy to estimate the robustness of the tree to flying taxa (see \hyperref[supplementaries]{supplementaries}).
@@
 
+\subsubsection{Statistical difference between tree topologies}
+To assess the difference between the tree topologies, whether they were inferred in a Bayesian or a ML framework, we performed a group comparison test for each set of configurations (e.g.
+effect of $M_L$ in Bayesian inference, combined effect of $M_F$ and $M_C$ in ML inference, etc...).
+We used the pairwise tree comparisons as factors (i.e.
+the "Best" tree vs.
+one of each of the 125 "missing-data" trees) and the values of the 51 replicates from the pairwise comparison as a response variable.
+The 51 replicates values were either the replicates of the 51 pairwise tree comparisons in ML or the modes of the 51 RPBTC posterior distributions in Bayesian.
+When we found a significant difference between the groups, we performed a pairwise comparison test to analyse which pairs of groups differ from the others.
+Depending on whether the data was normal with variance homoscedasticity or not (controlled by performing respectively a shapiro.test and a a bartlett.test in R\{stats\}), we used respectively parametric or non-parametric tests for the group comparison and the pairwise comparisons as suggested by \citet{ruxtontime2008} (Table ~\ref{Stats_test}).
+%link broken
+
+\begin{table}[ht]
+\caption{Statistical tests}
+\centering
+\begin{tabular}{lrll}
+  \hline
+   & test & statistics & code\{package\} \\ [0.5ex]
+  \hline
+  Groups comparisons & ANOVA & parametric & anova\{stats\} \\ [0.5ex]
+                     & Kruskal-Wallis & non-parametric & kruskal.test\{stats\} \\[0.5ex]
+  \hline
+  Pairwise comparisons & Tukey HSD & parametric & TukeyHSD\{stats\} \\ [0.5ex]
+                       & Nemenyi-Damico-Wolfe-Dunn & non-parametric & kruskalmc\{pgirmess\} \\ [0.5ex]
+  \hline
+\end{tabular}
+\end{table}
+\label{Stats_test}
+
 \subsection{Effect of missing molecular characters for fossil taxa} %Awfull name.
@@
 
+%---------------------------------------------
+%       Empirical data - TO DO
+%---------------------------------------------
+
 \subsection{Empirical data} %Still not done!
@@
 
+%---------------------------------------------
+%
+%       RESULTS
+%
+%---------------------------------------------
+
 \section{Results}
@@
 Bayesian chains converged (SD $<$ 0.01 and ESS $>>$ 200 for each tree).
+Good a quick convergence when using the sub-matrices especially for the "horizontal" sub-matrices.
+Good enough phylogenetic signal.
 \subsection{Comparing topologies}
 Did it worked? See sub-matrices
-\subsection{Effect of $M_L$}
-In ML and in Bayesian
-\subsection{Effect of $M_F$}
-In ML and in Bayesian
-\subsection{Effect of $M_C$}
-In ML and in Bayesian
-\subsection{Combined effect}
-In ML and in Bayesian
+
+\subsection{Effect of the missing data parameters}
+
+\subsubsection{Effect of $M_L$}
+In ML...
+
+Topological recovery was bad in Bayesian, the NTS score was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance.
+However, it seems that there is a small difference depending on the metric used.
+For the Triple metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
+For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases.
+In other words, if living taxa are removed, clades are less conserved.
+However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the NTS score was only slightly higher than expected by chance.
+
+%Tables model...
+
+% latex table generated in R 3.0.2 by xtable 1.7-1 package
+% Wed May 28 17:54:17 2014
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrr}
+  \hline
+ & Kruskal.Wallis.chi.squared & df & p.value \\ 
+  \hline
+1 & 65.62 & 4.00 & 0.00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+% latex table generated in R 3.0.2 by xtable 1.7-1 package
+% Wed May 28 17:54:17 2014
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrl}
+  \hline
+ & obs.dif & critical.dif & difference \\ 
+  \hline
+1-2 & 45.87 & 41.00 & TRUE \\ 
+  1-3 & 60.33 & 41.00 & TRUE \\ 
+  1-4 & 76.83 & 41.00 & TRUE \\ 
+  1-5 & 99.80 & 41.00 & TRUE \\ 
+  2-3 & 14.46 & 41.00 & FALSE \\ 
+  2-4 & 30.96 & 41.00 & FALSE \\ 
+  2-5 & 53.93 & 41.00 & TRUE \\ 
+  3-4 & 16.50 & 41.00 & FALSE \\ 
+  3-5 & 39.47 & 41.00 & FALSE \\ 
+  4-5 & 22.97 & 41.00 & FALSE \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\subsubsection{Effect of $M_F$}
+In ML...
+
+Similarly than for the effect of $M_L$, Bayesian tree recovery was only slightly better than expected by chance.
+And in the same way as described above, there seems to be a small difference depending on the metric used.
+There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric.
+However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework.
+
+\subsubsection{Effect of $M_C$}
+In ML...
+
+The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Fould metric shows a significant effect of the $M_C$ parameter on tree recovery.
+\subsubsection{Combined effect}
+
+In ML...
+
+In Bayesian, same pattern as above.
+No particular effect of data combination, but more an effect of the overall amount of data.
+
+$\newline$
 
@@
 However, in Bayesian, topology is badly (if not at all) recovered independently of the amount of data missing.
+
+%---------------------------------------------
+%
+%       DISCUSSION
+%
+%---------------------------------------------
 
@@
 
+%Potential reviewers comments?
+%What about it when you vary the proportion of fossil taxa?
+
+%---------------------------------------------
+%
+%       CONCLUSION
+%
+%---------------------------------------------
+
 \section{Conclusion}
@@
 
+%---------------------------------------------
+%
+%       ...
+%
+%---------------------------------------------
+
 \section{Acknowledgements}
-We would like to thank Trevor Hodkinson %Also Gavin Thomas, Emmanuel Douzery and Frédéric Delsuc if not authors.
+We would like to thank Trevor Hodkinson and Andrew Jackson %Also Gavin Thomas, Emmanuel Douzery and Frédéric Delsuc if not authors.
 for his useful comments on the simulation protocol and Paddy Doyle for the assistance on using the computer cluster.
@@
 We applied this protocol on a poorly resolved tree (Low Score), a resolved tree with low support value (Medium Score) and a resolved tree with high support values (High Score).
-Results are available in table ~\ref{RPBTC_testing}.
+Results are available in table ~\ref{RPBTC_testing}. %link broken
 
 \begin{table}[ht]
-\caption{Anova results: difference between 100 replicates using the RPBTC method}
+\caption{Group comparison results: difference between 100 replicates using the RPBTC method} %test with TreeCmp.anova
 \centering

### 2014-05-29 Functions/TreeComparisons/TreeCmp.anova.R
v.0.4 non-parametric posthoc output improved

@@ -3,19 +3,20 @@
 ##########################
 #Performs an anova on a 'TreeCmp' object (parametric or non-parametric)
 #Performs a TukeyHSD if a posthoc difference is significant (parametric or non-parametric)
-#v.0.3
+#v.0.4
 #Update: automatic parametric/non-parametric test (and possibility to save the results)
 #Update: improved/corrected parametric test
+#Update: non-parametric posthoc output improved
 ##########################
 #SYNTAX :
-#<data> an object of the class 'TreeCmp'
+#<TreeCmp> an object of the class 'TreeCmp'
 #<metric> the name of the metric of interest
 #<plot> whether to plot the distribution of the metric as a boxplot (default=FALSE)
 #<LaTeX> whether to output a LaTeX table (default=FALSE)
 #<save.test> whether to save the details of the various parametric/non-parametric tests (default=FALSE)
 ##########################
 #----
-#guillert(at)tcd.ie - 28/05/2014
+#guillert(at)tcd.ie - 29/05/2014
 ##########################
 #Requirements:
 #-R 3
@@ -185,9 +186,10 @@ TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE, save.test=FALS
         #Test is non-parametric
 
         #Transform the factors into numerical levels
-        anova.data$factors<-as.factor(anova.data$factors)
-        levels(anova.data$factors)<-c(seq(1:length(levels(anova.data$factors))))
-        anova<-kruskal.test(values ~ factors, data=anova.data)
+        anova.data.fact<-anova.data
+        anova.data.fact$factors<-as.factor(anova.data.fact$factors)
+        levels(anova.data.fact$factors)<-c(seq(1:length(levels(anova.data.fact$factors))))
+        anova<-kruskal.test(values ~ factors, data=anova.data.fact)
         pvalue<-anova[[3]]
 
         #If there posthoc difference

### 2014-05-29 Functions/TreeComparisons/TreeCmp.pairPlot.R
v.0.1 allows to plot pairwise comparison heat maps

@@ -1,0 +1,99 @@
+##########################
+#Pairwise difference plot from an 'TreeCmp' object anova.
+##########################
+#Plots a pairwise comparison results table to summarize a "TreeCmp" object anova.
+#v.0.1
+##########################
+#SYNTAX :
+#<posthoc> the posthoc results of a TreeCmp.anova function
+#<parametric> whether the test was parametric or not (must be logical)
+##########################
+#----
+#guillert(at)tcd.ie - 29/05/2014
+##########################
+#Requirements:
+#-R 3
+##########################
+
+TreeCmp.pairPlot<-function(posthoc, parametric) {
+
+#DATA INPUT
+
+    #parametric
+    if(class(parametric) != 'logical'){
+        stop('Parametric is not logical')
+    }
+
+    #posthoc
+    if(parametric == FALSE){
+        #posthoc comes from a non-parametric Kruskal-Wallis
+        if(class(posthoc) != 'data.frame'){
+            stop('Wrong posthoc format, pothoc must be the output of TreeCmp.anova function')
+        } else {
+            critical.dif<-posthoc[[2]][1]
+            name.list<-unique(unlist(strsplit(row.names(posthoc), split="-")))
+            n.parameters<-length(name.list)
+            posthoc.scores<-posthoc[[1]]
+            #remove non.significant values
+            non.significant<-which(posthoc[[3]] == FALSE)
+            posthoc.scores[non.significant]<-NA
+        }
+    } else {
+        #posthoc comes from a parametric anova
+        if(class(posthoc) != 'matrix'){
+            stop('Wrong posthoc format, pothoc must be the output of TreeCmp.anova function')
+        } else {
+            critical.dif<-0
+            name.list<-unique(unlist(strsplit(row.names(posthoc), split="-")))
+            n.parameters<-length(name.list)
+            posthoc.scores<-abs(as.vector(posthoc[,1]))
+             #remove non.significant values
+            non.significant<-which(as.vector(posthoc[,4]) > 0.05)
+            posthoc.scores[non.significant]<-NA
+        }
+    }
+
+#FUNCTIONS
+
+    FUN.matrix<-function(posthoc.scores, n.parameters, name.list, critical.dif){
+        #transform the posthoc difference values in a squared matrix
+
[…]

### 2014-05-29 Analysis/MissingDataEffect.R
Added the new pairPlot function

@@ -9,6 +9,7 @@ source('../Functions/TreeComparisons/TreeCmp.Read.R')
 source('../Functions/TreeComparisons/TreeCmp.Mode.R')
 source('../Functions/TreeComparisons/TreeCmp.anova.R')
 source('../Functions/TreeComparisons/TreeCmp.distribution.R')
+source('../Functions/TreeComparisons/TreeCmp.pairPlot.R')
 
 
 #Loading the Random tree comparisons
@@ -117,23 +118,32 @@ legend( <x> , <y> , c("Mode", "50% CI", "95% CI"), col="grey", lty=c(NA, 1, 2), 
 #Checking for differences
 #MLiving
 #Distribution check (visual)
-TreeCmp.anova(Bayesian.ML, "R.F_Cluster", plot=TRUE, LaTeX=TRUE) #Diff
-TreeCmp.anova(Bayesian.ML, "Triples", plot=TRUE, LaTeX=TRUE) #No diff
+TreeCmp.anova(Bayesian.ML, "R.F_Cluster", plot=TRUE, LaTeX=TRUE) ;  TreeCmp.xtables$posthoc #Diff
+TreeCmp.anova(Bayesian.ML, "Triples", plot=TRUE, LaTeX=TRUE) ; TreeCmp.xtable #No diff
 #MFossil
-TreeCmp.anova(Bayesian.MF, "R.F_Cluster", plot=TRUE) #Diff
-TreeCmp.anova(Bayesian.MF, "Triples", plot=TRUE) #No diff
+TreeCmp.anova(Bayesian.MF, "R.F_Cluster", plot=TRUE, LaTeX=TRUE) ;  TreeCmp.xtables$posthoc #Diff
+TreeCmp.anova(Bayesian.MF, "Triples", plot=TRUE, LaTeX=TRUE) ; TreeCmp.xtable #No diff
 #MCharacter
-TreeCmp.anova(Bayesian.MC, "R.F_Cluster", plot=TRUE) #Diff
-TreeCmp.anova(Bayesian.MC, "Triples", plot=TRUE) #No diff
+TreeCmp.anova(Bayesian.MC, "R.F_Cluster", plot=TRUE, LaTeX=TRUE) ;  TreeCmp.xtables$posthoc #Diff
[…]

### 2014-05-29 Manuscript/TEM_draft.tex
Results tables

@@
 \end{tabular}
+\label{Stats_test}
 \end{table}
-\label{Stats_test}
 
@@
 %Tables model...
+\begin{table}
+\caption{Group difference per configuration set}
+\centering
+\begin{tabular}{rlrllll}
+    \hline
+    set & metric & framework & parametric & $stats^1$ & df & p.value \\
+    \hline
+    $M_L$ & Robinson-Fould & ML & bla & bla & bla & bla \\
+          &                & Bayesian & no & 65.62 & 4.00 & \textbf{0.00} \\
+          & Triples & ML & bla & bla & bla & bla \\
+          &         & Bayesian & no & 4.56 & 4.00 & 0.34 \\
+    $M_F$ & Robinson-Fould & ML & bla & bla & bla & bla \\
+          &                & Bayesian & no & 85.61 & 4.00 & \textbf{0.00} \\
+          & Triples & ML & bla & bla & bla & bla \\
+          &         & Bayesian & no & 3.01 & 4.00 & 0.56 \\ 
+    $M_C$ & Robinson-Fould & ML & bla & bla & bla & bla \\
+          &                & Bayesian & no & 64.23 & 4.00 & \textbf{0.00} \\
+          & Triples & ML & bla & bla & bla & bla \\
+          &         & Bayesian & no & 26.13 & 24.00 & 0.35 \\
+    $M_L+M_F$ & Robinson-Fould & ML & bla & bla & bla & bla \\
+              &                & Bayesian & no & 317.97 & 24.00 & \textbf{0.00} \\
+              & Triples & ML & bla & bla & bla & bla \\
+              &         & Bayesian & no & 3.01 & 4.00 & 0.56 \\ 
+    $M_L+M_C$ & Robinson-Fould & ML & bla & bla & bla & bla \\
+              &                & Bayesian & no & 290.10 & 24.00 & \textbf{0.00} \\
+              & Triples & ML & bla & bla & bla & bla \\
+              &         & Bayesian & no & 22.19 & 24.00 & 0.57 \\ 
+    $M_F+M_C$ & Robinson-Fould & ML & bla & bla & bla & bla \\
+              &                & Bayesian & no & 385.96 & 24.00 & \textbf{0.00} \\
+              & Triples & ML & bla & bla & bla & bla \\
+              &         & Bayesian & no & 20.23 & 24.00 & 0.68 \\
+    $M_L+M_F+M_C$ & Robinson-Fould & ML & bla & bla & bla & bla \\
+                  &                & Bayesian & no & 1167.40 & 124.00 & \textbf{0.00} \\
+                  & Triples & ML & bla & bla & bla & bla \\
+                  &         & Bayesian & no & 115.86 & 124.00 & 0.69 \\ 
+    \hline
+\label{Group_results}
+\end{tabular}
 
-% latex table generated in R 3.0.2 by xtable 1.7-1 package
-% Wed May 28 17:54:17 2014
-\begin{table}[ht]
+$^1$ F-value for parametric tests and Kruskal Wallis $chi^2$ for non parametric tests.
+\end{table}
+
+\begin{table}
+\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric}
 \centering
-\begin{tabular}{rrrr}
+\begin{tabular}{c|ccccc}
     \hline
- & Kruskal.Wallis.chi.squared & df & p.value \\ 
+              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
     \hline
-1 & 65.62 & 4.00 & 0.00 \\ 
+    $M_L$00\% & - & FALSE & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
+    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
+    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
+    $M_L$50\% & & & & - & FALSE \\
+    $M_L$75\% & & & & & - \\
     \hline
 \end{tabular}
+\centering
+\begin{tabular}{rrrl}
+ pairwise comparison & obs.dif & critical.dif & difference \\ 
+  \hline
+  $M_L$00\%-$M_L$10\% & 17.05 & 41.00 & FALSE \\ 
+  $M_L$00\%-$M_L$25\% & 43.79 & 41.00 & TRUE \\ 
+  $M_L$00\%-$M_L$50\% & 77.32 & 41.00 & TRUE \\ 
+  $M_L$00\%-$M_L$75\% & 103.99 & 41.00 & TRUE \\ 
+  $M_L$10\%-$M_L$25\% & 26.75 & 41.00 & FALSE \\ 
+  $M_L$10\%-$M_L$50\% & 60.27 & 41.00 & TRUE \\ 
+  $M_L$10\%-$M_L$75\% & 86.94 & 41.00 & TRUE \\ 
+  $M_L$25\%-$M_L$50\% & 33.53 & 41.00 & FALSE \\ 
+  $M_L$25\%-$M_L$75\% & 60.20 & 41.00 & TRUE \\ 
+  $M_L$50\%-$M_L$75\% & 26.67 & 41.00 & FALSE \\ 
+   \hline
+\end{tabular}
+\label{ML_results}
 \end{table}
 
-% latex table generated in R 3.0.2 by xtable 1.7-1 package
-% Wed May 28 17:54:17 2014
-\begin{table}[ht]
+\begin{table}
+\caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
+\centering
+\begin{tabular}{c|ccccc}
+    \hline
+              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
+    \hline
+    $M_L$00\% & - & FALSE & FALSE & \textbf{TRUE} & \textbf{TRUE}\\
+    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
+    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
+    $M_L$50\% & & & & - & FALSE \\
+    $M_L$75\% & & & & & - \\
+    \hline
+\end{tabular}
 \centering
 \begin{tabular}{rrrl}
+ pairwise comparison & obs.dif & critical.dif & difference \\ 
   \hline
- & obs.dif & critical.dif & difference \\ 
-  \hline
-1-2 & 45.87 & 41.00 & TRUE \\ 
-  1-3 & 60.33 & 41.00 & TRUE \\ 
-  1-4 & 76.83 & 41.00 & TRUE \\ 
-  1-5 & 99.80 & 41.00 & TRUE \\ 
-  2-3 & 14.46 & 41.00 & FALSE \\ 
-  2-4 & 30.96 & 41.00 & FALSE \\ 
-  2-5 & 53.93 & 41.00 & TRUE \\ 
-  3-4 & 16.50 & 41.00 & FALSE \\ 
-  3-5 & 39.47 & 41.00 & FALSE \\ 
-  4-5 & 22.97 & 41.00 & FALSE \\ 
+  $M_F$00\%-$M_F$10\% & 16.03 & 41.00 & FALSE \\ 
+  $M_F$00\%-$M_F$25\% & 38.50 & 41.00 & FALSE \\ 
+  $M_F$00\%-$M_F$50\% & 61.52 & 41.00 & TRUE \\ 
+  $M_F$00\%-$M_F$75\% & 95.52 & 41.00 & TRUE \\ 
+  $M_F$10\%-$M_F$25\% & 22.47 & 41.00 & FALSE \\ 
+  $M_F$10\%-$M_F$50\% & 45.49 & 41.00 & TRUE \\ 
+  $M_F$10\%-$M_F$75\% & 79.49 & 41.00 & TRUE \\ 
+  $M_F$25\%-$M_F$50\% & 23.02 & 41.00 & FALSE \\ 
+  $M_F$25\%-$M_F$75\% & 57.02 & 41.00 & TRUE \\ 
+  $M_F$50\%-$M_F$75\% & 34.00 & 41.00 & FALSE \\ 
    \hline
 \end{tabular}
+\label{MF_results}
+\end{table}
+
+\begin{table}
+\caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
+\centering
+\begin{tabular}{c|ccccc}
+    \hline
+              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
+    \hline
+    $M_C$00\% & - & FALSE & FALSE & \textbf{TRUE} & \textbf{TRUE}\\
+    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
+    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
+    $M_C$50\% & & & & - & FALSE \\
+    $M_C$75\% & & & & & - \\
+    \hline
+\end{tabular}
+\centering
+\begin{tabular}{rrrl}
+ pairwise comparison & obs.dif & critical.dif & difference \\ 
+  \hline
+  $M_C$00\%-$M_C$10\% & 16.03 & 41.00 & FALSE \\ 
+  $M_C$00\%-$M_C$25\% & 38.50 & 41.00 & FALSE \\ 
+  $M_C$00\%-$M_C$50\% & 61.52 & 41.00 & TRUE \\ 
+  $M_C$00\%-$M_C$75\% & 95.52 & 41.00 & TRUE \\ 
+  $M_C$10\%-$M_C$25\% & 22.47 & 41.00 & FALSE \\ 
+  $M_C$10\%-$M_C$50\% & 45.49 & 41.00 & TRUE \\ 
+  $M_C$10\%-$M_C$75\% & 79.49 & 41.00 & TRUE \\ 
+  $M_C$25\%-$M_C$50\% & 23.02 & 41.00 & FALSE \\ 
+  $M_C$25\%-$M_C$75\% & 57.02 & 41.00 & TRUE \\ 
+  $M_C$50\%-$M_C$75\% & 34.00 & 41.00 & FALSE \\ 
+   \hline
+\end{tabular}
+\label{MC_results}
 \end{table}

### 2014-05-30 Functions/TreeComparisons/TreeCmp.anova.R
minor update: print where the param test is saved

@@ -131,6 +131,7 @@ TreeCmp.anova<-function(TreeCmp, metric, plot=FALSE, LaTeX=FALSE, save.test=FALS
 
         if(save.test == TRUE) {
             parametric.test<<-list(normality=normality, homoscedasticity=homoscedasticity) #Save out of the function for output
+            cat("Parametric test saved in 'parametric.test")
         }
 
         #Choosing parametric or non-parametric test

### 2014-05-30 Functions/TreeComparisons/TreeCmp.examples.R
Examples for the TreeCmp functions

@@ -1,0 +1,33 @@
+##########################
+#TreeCmp functions examples
+##########################
+#TreeCmp.pairPlot
+##########################
+#----
+#guillert(at)tcd.ie - 30/05/2014
+##########################
+#Requirements:
+#-R 3
+##########################
+
+#TreeCmp.pairPlot
+
+    #Generating a random group comparison list (100 groups)
+    factors<-NULL ; for (i in 1:100){factors<-c(factors,rep(i, 50))}
+    factors<-as.factor(factors)
+    values<-NULL ; for (i in 1:100){values<-c(values, rnorm(50, sample(i:100,1)))}
+    example<-list("factors"=factors, "values"=values)
+
+    #non-parametric test (untested assumption, just for the example)
+    require(pgirmess)
+    posthoc<-kruskalmc(values ~ factors, data=example)
+    posthoc.nonpar<-posthoc$dif.com
+
+    #parametric test (untested assumption, just for the example)
+    aov<-aov(values ~ factors, data=example)
+    posthoc<-TukeyHSD(aov)
+    posthoc.par<-posthoc$factors
+
+    #posthoc pairwise comparisons heat plot
+    TreeCmp.pairPlot(posthoc.nonpar, parametric=FALSE)
+    TreeCmp.pairPlot(posthoc.par, parametric=TRUE)

### 2014-05-30 Functions/TreeComparisons/TreeCmp.pairPlot.R
minor update: post hoc in plot table

@@ -92,8 +92,8 @@ TreeCmp.pairPlot<-function(posthoc, parametric) {
             paste("max:",round(max(posthoc.scores, na.rm=TRUE), digit=3))
         ),
         pch=15, pt.cex=2, col=c(heat[1],heat[abs(n.parameters/2)],heat[n.parameters]))
-    title(main = "Significant pairwise differences")
+    title(main = "Significant posthoc pairwise differences")
     box()
-
+    
 #End
 }

### 2014-06-03 Manuscript/TEM_draft.tex
Improved introduction

@@
 \bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
+% We don't use a special title page; the author information is entered like any other text.
+% FOOTNOTES: We don't allow them in the manuscript, except in tables. Don't include any footnotes in the text.
+%IMPORTANT:
+%Note on display in the text file: sentences or sub-parts of sentences (e.g. enumerations) are delimited by a newline for clarity.
+%However, true new lines (i.e. in the pdf) are delimited by one or more empty lines.
+%We don't use a heading for the introduction. The nature of this section is implied by its position following the abstract and preceding the first major section heading.
 
-%START
+%---------------------------------------------
+%
+%       START
+%
+%---------------------------------------------
+
 \begin{document}
@@
 \bigskip
-\noindent RH: Missing data in Total Evidence matrices %RH = Running Head
+\noindent RH: Missing data and topology in total evidence matrices %RH = Running Head
 
@@
 \noindent{\Large \bf Effect of missing data in matrices containing living and fossil taxa on topological accuracy}
+%Alternative punch title: "Bayesian method failure to infer topology using matrices containing both living and fossil taxa"
+%Alternative punch title: "ML outperforms Bayesian for inferring topology from matrices containing both living and fossil taxa"
 \bigskip
-
-% We don't use a special title page; the author information is entered like any other text.
-% FOOTNOTES: We don't allow them in the manuscript, except in tables. Don't include any footnotes in the text.
 
@@
 \vspace{1in}
-
-%IMPORTANT:
-%Note on display in the text file: sentences or sub-parts of sentences (e.g. enumerations) are delimited by a newline for clarity.
-%However, true new lines (i.e. in the pdf) are delimited by one or more empty lines.
 
@@
 Our results suggest that, in order to use total evidence approaches, one should reduce the missing data in the morphological part of the matrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process.
-We apply our method to a comprehensive data set of both living and fossil primates. %Not at all!
-We find that using this integrative method modifies previous estimates of rates of body mass evolution within primates.
+%More results!
 
-\noindent (Keywords: missing data, total evidence, bla)\\
+\noindent (Keywords: missing data, total evidence, Bayesian, Maximum Likelihood, topology)\\
 
@@
 
-%Syst.Biol: We don't use a heading for the introduction. The nature of this section is implied by its position following the abstract and preceding the first major section heading.
 Living species represent less than 1\% of all species that have ever lived \citep{novacek1992ext,raup1993extinction}.
 However, the majority of macroevolutionary studies focus solely on living species \citep{cooperwhat2009,meredithimpacts2011,healyecology2014}.
-%or is citing of your own papers too douchebagy in the second sentence?
+%or is citing of your own papers in the too "douchebagy" for the second sentence?
 Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
@@
 
-Three main approaches have been used for combining fossil and living taxa data in phylogenies.
-These approaches differ mainly in whether they treat fossil taxa as tips or as nodes in the phylogeny and which part of the available data of the fossil is used (i.e.
+Up to now, three main approaches have been used for combining fossil and living taxa data in phylogenies.
+These approaches differ mainly in whether they treat fossil taxa as tips or as nodes in the phylogeny and on which part of the available data of the fossil is used (i.e.
 the fossil occurrence age only or both the age and the morphology).
 Classical cladistic methods uses matrices containing morphological data from both fossil and living taxa and allows to treat each taxon as a tip in the phylogeny; the relation between the taxa can be inferred using optimal criteria such as the maximum parsimony criterion \citep{simpson1945}.
-This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species.
-Neontologists, on the other hand, more commonly use probabilistic methods (e.g.
-Maximum Likelihood or Bayesian) based on matrices containing only molecular data from living species to infer phylogenies.
+This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and doesn't allow to use probabilistic methods to deal with inference uncertainty (but see \citet{spencerefficacy2013}.
+Neontologists, on the other hand, more commonly use probabilistic methods (e.g. Maximum Likelihood or Bayesian) based on matrices containing only molecular data from living species.
 Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these analyses and only their occurrence age can be used to time calibrate phylogenies \citep{zuckerkandl1965}.
-There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013} but none of them uses all the available data.
+There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013}.
+However none of them is optimal and uses all the available data.
 
@@
 a probabilistic method such as ML where the method states precisely how to do stuff.
-It allows to use probabilistic methods, to treat every taxon as a tip on the phylogeny and to use the occurrence age of the fossil to time calibrate the phylogeny \citep{eernissetaxonomic1993}.
+It allows to use probabilistic methods, to treat every taxon as a tip in the phylogeny and to use the occurrence age of the fossil to time calibrate the phylogeny \citep{eernissetaxonomic1993}.
 Here we focus on this total evidence approach because they have been recently successfully developed and applied to empirical data \citep{pyrondivergence2011,ronquista2012,schragocombining2013}.
@@
 Although the total evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data.
-In particular it requires morphological data form both living and fossil taxa, both of which are scarce.
-Therefore total evidence approaches are likely to suffer from having lots of missing data which may affect their ability to infer correct phylogenies.
+In particular it requires morphological data form both living and fossil taxa, both of which are known to be scarce.
+Therefore total evidence approach is likely to suffer from having lots of missing data which may affect their ability to infer correct topology, branch length and support values \citep{salamin2003}.
 % Trevor: Missing data also influences support values that trend to decrease
@@
 Missing molecular data has been seen by some authors as an issue because it can decrease the phylogenetic signal in some parts of the tree, especially when using large matrices \citep{lemmonthe2009}.
-However other authors do not see missing molecular data as a major issue because the phylogenetic signal is more likely to increase by having at least a "modest" number of highly covered genes (i.e.
-approximatively half of the genes - \citet{rouresite-specific2011}), a higher number of taxa (especially slowly evolving taxa or taxa close to the outgroup) and by choosing more adequate models of sequence evolution rather than by reducing the amount of missing data \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
+However other authors do not see missing molecular data as a major issue because the phylogenetic signal, rather than by reducing the amount of missing data, is more likely to increase by having:
+(i) at least a "modest" number of highly covered genes (i.e. approximatively half of the genes - \citet{rouresite-specific2011});
+(ii) a higher number of taxa (especially slowly evolving taxa or taxa close to the outgroup - \citet{rouresite-specific2011}) and
+(iii) by choosing more adequate models of sequence evolution rather than by reducing the amount of missing data \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
+
 Similarly, depending on the study, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study \citep{wiensmissing2003,sansomfossilization2013}.
@@
 However, the phylogenetic signal is not related to the level of missing data \textit{per se} but to the number of informative characters per taxa, therefore missing data is less an issue than the number of shared informative characters \citep{wiensmissing2003}.
-Although missing data as been shown to be no major problem separately in molecular and morphological matrices \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011} it may become an issue when combined in a total evidence type matrix and no attempt has been made to study the impact of this issue on phylogenetic inference until now (especially because a major part of the matrix is missing - the molecular data from the fossil taxa).
-%Does not flow with the rest
+Although missing data as been shown to be no major problem separately in molecular and morphological matrices \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011} it may become an issue when combined in a total evidence type matrices.
+Because the nature of total evidence matrices, combining both molecular and morphological information for both living and fossil taxa, a big part of the matrix contains only missing data - the molecular data for fossil taxa.
+Until now, no attempt has been made to study the impact of this issue on phylogenetic inference and especially on topological recovery.
 
 Here we assess the effect of missing data on tree topology inferred from total evidence matrices.
-The molecular part of a total evidence matrix contains no fossil taxa so will act like a classical molecular matrix \citep{ronquista2012}.
-The effect of missing data on such matrices is well known, therefore, we only focus on the missing data issue in the morphological part of the matrix.
+The molecular part for living taxa of a total evidence matrix acts like a "classical" molecular matrix \citep{ronquista2012}.
+The effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}, therefore, we only focus on the missing data issue in the morphological part of the matrix.
 Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix:
@@
 (3) the overall number of morphological characters for both living and fossil taxa in the matrix.
-We assess how changing the values of these three parameters affects the topology of total evidence approach phylogenies.
+We assess how removing data from a total evidence type matrix by changing the values of these three parameters affects the topology of the inferred tree using Maximum Likelihood and Bayesian method.
 %This paragraph is crap
 
-%We found that… This as an implication for…
+We found that whatever the amount of missing data, total evidence matrices dramatically reduce the performance of Bayesian inference.
+However, when using a Maximum Likelihood approach, missing data decreases gradually the ability of recovering the right tree topology.
+We propose that this drastic difference between Bayesian and Maximum Likelihood methods is due to a flattening of the likelihood landscape because of the intrinsic amount of missing data in a total evidence matrix (i.e.
+the missing molecular data for the fossil taxa).
+This problem leads to an inability of Bayesian chains to pick up the sub and near optimal topologies leading to high variance in the posterior tree distribution and really low support in the consensus tree.
+These results should encourage authors in not using Bayesian inference to explore the topological landscape of their phylogeny.
+Instead, we suggest that authors should first find the Maximum Likelihood tree and use it's topology as a fix prior in their Bayesian inference.
+We also propose a new method for comparing Bayesian posterior tree distribution's topology available in a wonderful R package. %Or am I just dreaming a bit too much here?
 
@@
 \centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_outline.pdf}
+\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_outline-BW.pdf}
 \caption{Protocol outline.
@@
 \centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_missingData.pdf}
+\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_missingData-BW.pdf}
 \caption{Different parts of the matrix. %Title?
@@
 \centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp.pdf}
+\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp-BW.pdf}
 \caption{Tree comparison protocol.
@@
 Good enough phylogenetic signal.
+
 \subsection{Comparing topologies}
@@
 \subsubsection{Effect of $M_L$}
-In ML...
+In ML...(see Table ~\ref{Group_results})
 
@@
 For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases.
-In other words, if living taxa are removed, clades are less conserved.
-However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the NTS score was only slightly higher than expected by chance.
+In other words, if living taxa are removed, clades are less conserved (see Table ~\ref{ML_results}).
+However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the NTS score was only slightly higher than expected by chance (see Table ~\ref{Group_results}).
 
-%Tables model...
+\subsubsection{Effect of $M_F$}
+In ML...(see Table ~\ref{Group_results})
+
+Similarly than for the effect of $M_L$, Bayesian tree recovery was only slightly better than expected by chance.
+And in the same way as described above, there seems to be a small difference depending on the metric used.
+There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric (see Table ~\ref{ML_results}).
+However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework (see Table ~\ref{Group_results}).
+
+\subsubsection{Effect of $M_C$}
+In ML...(see Table ~\ref{Group_results})
+
+The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Fould metric shows a significant effect of the $M_C$ parameter on tree recovery (see Table ~\ref{ML_results} and ~\ref{Group_results}).
+\subsubsection{Combined effect}
+
+In ML...(see Table ~\ref{Group_results})
+
+In Bayesian, same pattern as above.
+No particular effect of data combination, but more an effect of the overall amount of data (see Table ~\ref{Group_results}).
+
+$\newline$
+
+The ability of recovering the best topology in maximum likelihood method is function of the amount of data missing.
+However, in Bayesian, topology is badly (if not at all) recovered independently of the amount of data missing.
+
+%Add the global heat plot figures
 \begin{table}
@@
 \centering
-\begin{tabular}{rlrllll}
+\begin{tabular}{rlrlccc}
     \hline
@@
 \end{table}
-
-\subsubsection{Effect of $M_F$}
-In ML...
-
-Similarly than for the effect of $M_L$, Bayesian tree recovery was only slightly better than expected by chance.
-And in the same way as described above, there seems to be a small difference depending on the metric used.
-There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric.
-However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework.
-
-\subsubsection{Effect of $M_C$}
-In ML...
-
-The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Fould metric shows a significant effect of the $M_C$ parameter on tree recovery.
-\subsubsection{Combined effect}
-
-In ML...
-
-In Bayesian, same pattern as above.
-No particular effect of data combination, but more an effect of the overall amount of data.
-
-$\newline$
-
-The ability of recovering the best topology in maximum likelihood method is function of the amount of data missing.
-However, in Bayesian, topology is badly (if not at all) recovered independently of the amount of data missing.
 
@@
 \subsection{Codes}
-All codes are available at: $https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/tree/master/Functions$
+All codes are available at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions$
 

### 2014-06-03 Manuscript/TEM_draft.tex
improved methods

@@
 \section{Methods} %no material needed?
-To explore the effect of missing data on total evidence trees topologies we used the following protocol (note that we explain each step in detail below this general outline -Fig. ~\ref{Fig_Outline}):
+To explore the effect of missing data in total evidence matrices on tree topology we used the following protocol (note that we explain each step in detail below this general outline -Fig.
+~\ref{Fig_Outline}):
 \begin{enumerate}
-\item{Generating the matrix:} \\
-\label{step:generate_matrix}
+\item{Generating the matrix:} \label{step:generate_matrix} \\
 We randomly generated birth death tree (hereafter called the "true" tree, Table ~\ref{Tab_glossary}) to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix, Table ~\ref{Tab_glossary}).
-\item{Removing data:} \\
-\label{step:remove_data}
+\item{Removing data:} \label{step:remove_data} \\
 We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (1) the proportion of missing living taxa ($M_{L}$), (2) the proportion of missing data in the fossil taxa ($M_{F}$) and (3) the proportion of missing morphological characters ($M_{C}$) (the resulting matrices are called hereafter the "missing-data" matrix, Table ~\ref{Tab_glossary}).
-\item{Building phylogenies:} \\
-\label{step:build_phylo}
+\item{Building phylogenies:} \label{step:build_phylo} \\
 We inferred Bayesian phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in respectively a tree generate from a matrix containing no missing data in the morphological part (hereafter called the "best" tree, Table ~\ref{Tab_glossary}) and trees inferred from matrices with missing data in their morphological part (hereafter called the "missing-data" trees, Table ~\ref{Tab_glossary}).
-\item{Comparing topologies:} \\
-\label{step:compare_topo}
+\item{Comparing topologies:} \label{step:compare_topo} \\
 We then compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on the topologies of the phylogenies we estimated.
@@
 (1) We generated a random tree (the "true" tree) to infer a matrix with no missing data (the "complete" matrix).
-(2) We removed data fron the morphological partition of the "complete" matrix resulting in 124 "missing-data" matrices.
+(2) We removed data from the morphological partition of the "complete" matrix resulting in 125 "missing-data" matrices.
 (3) We infer phylogenetic trees from each matrix in both ML and Bayesian framework.
 (4) We then compared the "missing-data" trees to the "best" tree.
-We repeated step 1 to 4 51 (3$\times $17) times.}
+We repeated step ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} 51 (3$\times $17) times.}
 \label{Fig_Outline}
@@
 \includegraphics[keepaspectratio=true]{Figures/TEM_Fig_missingData-BW.pdf}
-\caption{Different parts of the matrix. %Title?
+\caption{Names of the different parts of the matrix.
 A: Missing data parameters:
@@
 
-In practice, each parameter represent a different way of removing data in the morphological part of the matrix: $M_{L}$ removes a proportion of rows from the living taxa; $M_{F}$ removes a proportion of cells from the fossil taxa; and $M_{C}$ removes a proportion of columns across both living and fossil taxa.
+In practice, each parameter represent a different way of removing data in the morphological part of the matrix: $M_{L}$ removes a proportion of rows from the living taxa; $M_{F}$ removes a proportion of cells from the fossil taxa; and $M_{C}$ removes a proportion of columns across both living and fossil taxa (see Fig.
+~\ref{Fig_RemoveData}).
 Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a proportion of the living taxa is removed (i.e.
@@
 The "best" tree, on the other hand, is the best tree we can build using the state-of-the-art phylogenetic methods.
-In real world situations, the "true" tree is never available to us because we cannot know the true evolutionnary history of a clade (except in very rare circumstances, e.g. \citet{rozen2005}).
+In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g. \citet{rozen2005}).
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
@@
 The GTR model can be seen as a generalisation of the HKY model (the 2 parameters from the HKY model are implicitly included in the 6 from GTR model - \citet{stamatakisa2008}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} wich is a generalisation of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014}).
-We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. 
-The bootstrap algorithm used in RAxML is the Lazy Subtree Rearrangement (LSR) which consists in pruning one subtree from the tree and subsequently reinserting it to al neighbouring branches \citep{stamatakisa2008}.
-Subtree Pruning and Reinserting methods (SPR) have been demonstrated as being better than others (e.g. Nearest Neighbouring Interchange - NNI) in recovering good bootstrap values \citep{salamin2003}.
+For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} which is a generalisation of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014}).
+We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. %The explanations of the algorithm part for the BS was proposed by Trevor
+The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists in pruning one sub-tree from the tree and subsequently reinserting it to al neighbouring branches \citep{stamatakisa2008}.
+Sub-tree Pruning and Reinserting methods (SPR) have been demonstrated as being better than others (e.g. Nearest Neighbouring Interchange - NNI) in recovering good bootstrap values \citep{salamin2003}.
 
@@
 We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
-The effective sample size (ESS) was also checked on a random subsample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
-For each run, we removed 25\% of the iterations as burnin.
-We used the following priors for each tree (see \hyperref[supplementaries]{supplementaries}):
+The effective sample size (ESS) was also checked on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
+For each run, we removed 25\% of the iterations as burn-in.
+We used the following prior for each tree (see \hyperref[supplementaries]{supplementaries}):
 \begin{enumerate}
@@
 
-We used these priors to speed up the Bayesian process.
-These priors biased the way the Bayesian process calculated the branch length by giving non-random starting points and boundaries for the parameters estimation process, however, in this study, we focused on the effect of missing data on the topology and not on the branch length.
-Even using these priors, it took 65 CPU years to build 51 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
+We used these prior to speed up the Bayesian process.
+These prior biased the way the Bayesian process calculated the branch length by giving non-random starting points and boundaries for the parameters estimation process, however, in this study, we focused on the effect of missing data on the topology and not on the branch length.
+Even using these prior, it took 65 CPU years to build 51 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
 
@@
 We compared the topology of the "missing-data" trees inferred from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
-Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees inferred from matrices with living and fossil taxa and molecular and morphological characters.
-To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triple \citep{dobson1975triplets} and the Robison-Fould \citep{RF1981} distance.
+Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees inferred from total evidence type matrices.
+To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triples \citep{dobson1975triplets} and the Robinson-Fould \citep{RF1981} distance.
 We normalised the two metrics using the Normalised Tree Similarity index \citep{Bogdanowicz2012} to generalise our results for any \textit{n} number of taxa.
-The two metrics and the index are detailed below.
+The three metrics are detailed below.
 
-\subsubsection{Triple distance ($T_{x,y}$) \citep{dobson1975triplets}}
-This metric measures the number of different subtrees of three clades between two given trees.
+\subsubsection{Triples distance ($T_{x,y}$) \citep{dobson1975triplets}}
+This metric measures the number of different sub-trees of three taxa between two given trees.
 Each triplet can be written as $I_{ijk}$=(\textit{ijk}).
@@
 And where \textit{n} is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
-If $S_n$=0, the trees are the same (i.e. no taxa as been displaced). When $S_n$ = $\binom{n}{4}$, the trees are the most different possible (i.e. every taxa as been displaced).
-This metric therefore illustrates the amount of displayed taxa but is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g.
-fossil taxa) than the Robinson-Foulds distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and can therefore be used as a proxy to estimate the robustness of the tree to flying taxa (see \hyperref[supplementaries]{supplementaries}).
+If $S_n$=0, the trees are the same (i.e. no taxa as been displaced).
+When $S_n$ = $\binom{n}{4}$, the trees are the most different possible (i.e. every taxa as been displaced).
+This metric therefore illustrates the amount of displayed taxa.
+It is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g.
+fossil taxa) than the Robinson-Foulds metric \citep{critchlowthe1996,johnson1998,wiensmissing2003}.
+Therefore we used this metric as a proxy to estimate the robustness of the tree to flying taxa (see \hyperref[supplementaries]{supplementaries}).
 
@@
 This metric measures the number of shared clades among two trees and therefore illustrates the number of exactly conserved groups among the trees.
-The Robison-Fould distance (also called path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
+The Robinson-Fould distance (also called path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
 \begin{equation}
@@
 The minimal value of \textit{C} is 1 if the two trees have the same \textit{n} taxa; the maximal value in \textit{C=n-2}.
-This metric is more sensitive to taxa displacement than the triple metric (if one taxa gets out of a clade, then the clades are no longer considered as similar - \citet{critchlowthe1996,johnson1998,wiensmissing2003}) and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades (see \hyperref[supplementaries]{supplementaries}).
+This metric is more sensitive to taxa displacement than the Triples metric (i.e.
+if one taxa gets out of a clade, then the clades are no longer considered as similar - \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
+Therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades (see \hyperref[supplementaries]{supplementaries}).
 
@@
 We compared the "missing-data" tree to the "True" and the "Best" tree for each chain.
-For the Maximum Likelihood trees we performed pairwise comparisons between the "True" and "Best" tree and the "missing-data" tree (where the "missing-data" tree is one of the 125 trees resulting from the 125 super matrices including various amount of missing data, see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triple metric.
+For the Maximum Likelihood trees we performed pairwise comparisons between the "True" and "Best" tree and the "missing-data" tree (see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triples metric.
 We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
-For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparison for the same metric and the same number of taxa \textit{n}=51 (see \hyperref[supplementaries]{supplementaries}).
-We ran the comparison for every "missing-data" tree being one of the 125 combination of parameters values $M_L$, $M_F$ and $M_C$ in each chain resulting in 51 comparisons for every "missing-data" trees.
+For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparisons for the same metric and the same number of taxa \textit{n}=51 (see \hyperref[supplementaries]{supplementaries}).
+We ran the comparison for every "missing-data" tree in each chain resulting in 51 comparisons for every "missing-data" trees. 
 We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package (v3.1 – \citet{hdrcde}).
@@
 the likeliest tree in ML).
-This method has the clear advantage of better dealing with error and uncertainty but is less practical for using the results (i.e.
-the phylogeny) for any further study (but see \citet{healyecology2014}).
-To avoid this problem, people traditionally use the a consensus tree build on a majority rule in order to summarise a Bayesian tree posterior distribution leading to a single tree containing both topological and branch length information as well as support information (i.e.
-posterior probabilities, e.g.
-\citet{Ronquist2012mrbayes}).
+This method has the clear advantage of better dealing with error and uncertainty but it's output (a distribution rather than a single tree) is less practical to use for any further study (but see \citet{healyecology2014}).
+%Awwwwww yiis!
+To avoid this problem, people traditionally use the a consensus tree build on a majority rule in order to summarise a Bayesian tree posterior distribution.
+This results into a single tree containing both topological and branch length information as well as support information (i.e. posterior probabilities, e.g. \citet{Ronquist2012mrbayes}).
 However, using a Bayesian consensus tree has limitation especially if the resulting consensus tree is not well supported or resolved.
@@
 \subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
-We compared the distribution using a Random Pairwise Bayesian Tree Comparison (RPBTC) method.
+We compared the Bayesian posterior distribution using a Random Pairwise Bayesian Tree Comparison (RPBTC) method.
 This method consists in comparing a series of randomly selected pairs of trees from two posterior distributions (on from each distribution) and use the mode to summarize the resulting distribution as a proxy of the difference between the two trees in a Bayesian framework.
 \begin{equation}
-RPBTC_{X,Y}=Mo(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk}))
+RPBTC_{X,Y}=Mo[(]d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]
 \end{equation}
 
-Where \textit{X} and \textit{Y} are two Bayesian tree posterior distribution;
- $d_{m,n}$ is the pairwise difference for any metric \textit{m} and \textit{n} taxa between $X_{i}$ and $Y_{j}$ which are two single trees randomly sampled respectively from \textit{X} and \textit{Y}
- and $Mo(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk}))$ is the mode of the pairwise difference repeated \textit{k} times.
-We used the mode to summarize the distributions because the value that is the most represented in the distribution reflects more accurately what a consensus tree represents (the topology the most represented in the posterior distribution).
-Also, the mode represents a value that is present in the distribution (which, depending on the metric, can be a series of discrete values) in contrast to the other distribution summary metrics, such as the mean or the median, which can take values that are not actually present in the distribution.
-For example, if comparing two posterior tree distribution, we obtain difference values of 10, 10, 10, 5, 5 and 1, using the mean (6.8) or the median (7.5) gives a values that actually doesn't exist
+Where \textit{X} and \textit{Y} are two Bayesian tree posterior distributions;
+ $d_{m,n}$ is the pairwise difference for any metric \textit{m} and \textit{n} taxa between $X_{i}$ and $Y_{j}$ which are two single trees randomly sampled respectively from \textit{X} and \textit{Y};
+ $Mo$ is the mode and
+ $Mo[(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]$ is the mode of the pairwise difference repeated \textit{k} times.
 
-In this study, we calculated RPBTC for both metrics (RF and Triple distance) for each pairs of Bayesian tree distributions (i.e.
+We used the mode to summarize the distributions because it is the value that is the most represented in the distribution and reflects what a consensus tree represents (the topology the most represented in the posterior distribution).
+Also, the mode is a value present in the distribution (which, depending on the metric, can be a series of discrete values) in contrast to the other distribution summary metrics, such as the mean or the median, which can take values that are not actually present in the distribution.
+For example, if comparing two posterior tree distribution, we obtain difference values of 10, 10, 10, 5, 5 and 1, using the mean (6.8) or the median (7.5) gives a values that actually doesn't exist.
+%Useful?
+
+In this study, we calculated RPBTC for both metrics (Robinson-Fould and Triples distance) for each pairs of Bayesian tree distributions (i.e.
 the "Best" Bayesian tree posterior distribution vs.
@@
 We calculated the NTS for the "Best" Bayesian tree vs.
-"missing-data" Bayesian tree comparison for both RF and Triple metric using the mode of the RPBTC for each chain resulting in a distribution of 51 modes per comparison.
+"missing-data" Bayesian tree comparison for both Robinson-Fould and Triples metric using the mode of the RPBTC for each chain resulting in a distribution of 51 modes per comparison.
 We then compared the NTS of the Bayesian trees to the NTS of the ML trees for both metrics for each comparison.
@@
 \subsection{Effect of missing molecular characters for fossil taxa} %Awfull name.
-To assess the effect of missing molecular characters for fossil taxa in a Bayesian framework, we split the "complete" matrix in sub-matrices containing no missing data at all:
+To assess the effect of missing molecular characters for fossil taxa in a Bayesian framework, we split the "complete" matrix in sub-matrices containing no missing data at all (see Fig.
+~\ref{Fig_RemoveData}):
 \begin{enumerate}
@@
 \subsection{Comparing topologies}
-Did it worked? See sub-matrices
+Did it worked? See sub-matrices.
 
@@
 However, it seems that there is a small difference depending on the metric used.
-For the Triple metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
+For the Triples metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
 For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases.

### 2014-06-09 Analysis/MissingDataEffect.R
Added draft of the ML analysis

@@ -10,6 +10,7 @@ source('../Functions/TreeComparisons/TreeCmp.Mode.R')
 source('../Functions/TreeComparisons/TreeCmp.anova.R')
 source('../Functions/TreeComparisons/TreeCmp.distribution.R')
 source('../Functions/TreeComparisons/TreeCmp.pairPlot.R')
+source('../Functions/TreeComparisons/as.TreeCmp.R')
 
 
 #Loading the Random tree comparisons
@@ -21,11 +22,32 @@ Random25<-read.table("../Data/Tree_Comparisons/Random_trees/25t_Random.Cmp",head
 #/TreeCmp_chains/
 
 #ML chains
-#/ML/
 cd("../Data/Tree_Comparisons/TreeCmp_chains/ML/")
-tmp<-TreeCmp.Read('MChain', verbose=TRUE)
+tmp<-TreeCmp.Read('MChain', verbose=FALSE)
 ML_trees<-NTS(tmp, Random51)
 cd('../../../../Analysis/')
+
+#Isolating the ML treesets
+ML.ML<-ML_trees[c(1,26,51,76,101)]
+class(ML.ML)<-class(ML_trees)
+
+ML.MF<-ML_trees[c(1,6,11,16,21)]
+class(ML.MF)<-class(ML_trees)
+
+ML.MC<-ML_trees[1:5]
+class(ML.MC)<-class(ML_trees)
+
+ML.MLMF<-ML_trees[c(1,6,11,16,21, 26,31,36,41,46, 51,56,61,66,71, 76,81,86,91,96, 101,106,111,116,121)]
+class(ML.MLMF)<-class(ML_trees)
+
+ML.MLMC<-ML_trees[c(1:5, 26:30, 51:55, 76:80, 101:105)]
+class(ML.MLMC)<-class(ML_trees)
+
+ML.MFMC<-ML_trees[c(1:25)]
+class(ML.MFMC)<-class(ML_trees)
+
+ML.MLMFMC<-ML_trees
+class(ML.MLMFMC)<-class(ML_trees)
 
 #Bayesian chains
 cd("../Data/Tree_Comparisons/TreeCmp_chains/Bayesian/")
@@ -67,57 +89,92 @@ Bayesian.MLMFMC<-TreeCmp.Mode(Bayesian.MLMFMC, 1000, verbose=TRUE)
 #Setting the metric
 metric="R.F_Cluster"
 #metric="Triples"
[…]

### 2014-06-09 Functions/TreeComparisons/TreeCmp.Plot.R
Force a list to be used as a TreeCmp object

@@ -12,6 +12,7 @@
 #Update: results can be plotted as lines
 #Update: allow to stack the plots with add=TRUE
 #Update: allow window scaling option (ylim)
+#Update: typo in as.data.frame - fixed colour of single points
 ##########################
 #SYNTAX :
 #<TreeCmp> an object of the class TreeCmp
@@ -26,7 +27,7 @@
 #<save.details> whether to save the details of each comparison (default=FALSE). If TRUE, saves a density plot for each comparison. The chain name will be the one given in 'TreeCmp'. Is ignored if plot=FALSE.
 ##########################
 #----
-#guillert(at)tcd.ie - 27/05/2014
+#guillert(at)tcd.ie - 09/06/2014
 ##########################
 #Requirements:
 #-R 3
@@ -169,7 +170,7 @@ TreeCmp.Plot<-function(TreeCmp, metric, probs=c(95, 75, 50), plot=TRUE, col='bla
 
         #Transform dat into a column format
         dat<-matrix(NA, nrow=max(TreeCmp.rows), ncol=TreeCmp.length)
-        dat<-as.TreeCmp.frame(dat)
+        dat<-as.data.frame(dat)
         for (i in 1:length(TreeCmp)) {
             #The number of rows in the TreeCmp frame is equal to the maximum number of rows in the list. If the elements in the list don't have the same number of rows, NAs are added.
             dat[,i]<-c(TreeCmp[[i]][,metric.column], rep(NA,(max(TreeCmp.rows)-length(TreeCmp[[i]][,metric.column]))))
@@ -215,7 +216,7 @@ TreeCmp.Plot<-function(TreeCmp, metric, probs=c(95, 75, 50), plot=TRUE, col='bla
             for (k in 1:length(probs)) {
                 if(is.na(temp$hdr)) {
                     #Ploting only the mode if hdr=NA
-                    points(j,temp$mode,pch=19)
+                    points(j,temp$mode,pch=19, col=border)
                 } else {
                     #Plot the probabilities distribution
                     temp2 <- temp$hdr[k, ]

### 2014-06-09 Functions/TreeComparisons/as.TreeCmp.R
Force a list to be used as a TreeCmp object

@@ -1,0 +1,61 @@
+##########################
+#as TreeCmp
+##########################
+#Transform a list or data.frame into a TreeCmp object (if possible)
+#v.0.1
+##########################
+#SYNTAX :
+#<data> the input data
+##########################
+#----
+#guillert(at)tcd.ie - 09/06/2014
+##########################
+#Requirements:
+#-R 3
+##########################
+
+as.TreeCmp<-function(data) {
+
+#DATA INPUT
+    #data
+    if(class(data) == 'TreeCmp') {
+        data.list<-TRUE
+        data.TreeCmp<-TRUE
+    } else {
+        if(class(data) == 'list') {
+            data.list<-TRUE
+            data.TreeCmp<-FALSE
+            if(class(data[[1]]) != 'data.frame') {
+                stop('The given list must be a list of data.frame objects')
+            }
+        } else {
+            if(class(data) == 'data.frame') {
+                data.list<-FALSE
+                data.TreeCmp<-FALSE
+            } else {
+                stop('Input data must be a list or a data.frame object')
+            }
+        }
+    }
+
+#TRANSFORMING THE OBJECT IN A TREECMP OBJECT
+
+    if(data.TreeCmp == TRUE){
+    #data is already a TreeCmp object
+        return(data)
+    } else {
+        if(data.list == TRUE) {
+        #Data is a a list
+            class(data)<-"TreeCmp"
+            return(data)
+        } else {
+            #Data is a data.frame
+            data<-as.list(data)
+            class(data)<-"TreeCmp"
+            return(data)
+        }
+    }
+
+
+#End
+}

### 2014-06-11 Analysis/BestTrees_NodeSupport.R
Node support analysis

@@ -1,0 +1,28 @@
+#Script for measuring the node support for the ML "best" trees.
+#cd('../PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
+
+#Loading the functions
+source('../Functions/BS.read.R')
+
+library(xtable)
+
+#Loading the trees
+setwd('../Data/Best_trees_ML')
+boostraps<-BS.read('MChain', plot=TRUE)
+setwd('../../Analysis/')
+sum.bs<-unlist(boostraps)
+sum.bs<-sum.bs[-which(is.na(sum.bs))]
+
+#Ploting the bootstrap distribution
+hist(sum.bs, xlab="Bootstraps", main="Bootstraps distribution")
+
+#Cumulate frequencies for BS thresholds
+bs.100<-length(which(sum.bs == 100))/length(sum.bs)
+bs.90<-length(which(sum.bs > 89))/length(sum.bs)
+bs.80<-length(which(sum.bs > 79))/length(sum.bs)
+bs.75<-length(which(sum.bs > 74))/length(sum.bs)
+bs.50<-length(which(sum.bs > 49))/length(sum.bs)
+s.bs<-summary(sum.bs)
+#Result table
+bs.results<-data.frame("median"=median(sum.bs), "1stQu"=s.bs[[2]], "3rdQu"=s.bs[[2]], "bs100"=bs.100, "bs.90"=bs.90, "bs.80"=bs.80, "bs.75"=bs.75, "bs.50"=bs.50)
+xtable(bs.results)

### 2014-06-11 Analysis/MissingDataEffect.R
Changed cd into setwd

@@ -1,5 +1,5 @@
 #Script for testing the effect of missing data on tree topology.
-#cd('../PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
+#setwd('../PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
 
 
 #Loading the functions
@@ -22,10 +22,10 @@ Random25<-read.table("../Data/Tree_Comparisons/Random_trees/25t_Random.Cmp",head
 #/TreeCmp_chains/
 
 #ML chains
-cd("../Data/Tree_Comparisons/TreeCmp_chains/ML/")
+setwd("../Data/Tree_Comparisons/TreeCmp_chains/ML/")
 tmp<-TreeCmp.Read('MChain', verbose=FALSE)
 ML_trees<-NTS(tmp, Random51)
-cd('../../../../Analysis/')
+setwd('../../../../Analysis/')
 
 #Isolating the ML treesets
 ML.ML<-ML_trees[c(1,26,51,76,101)]
@@ -50,10 +50,10 @@ ML.MLMFMC<-ML_trees
 class(ML.MLMFMC)<-class(ML_trees)
 
 #Bayesian chains
-cd("../Data/Tree_Comparisons/TreeCmp_chains/Bayesian/")
+setwd("../Data/Tree_Comparisons/TreeCmp_chains/Bayesian/")
 tmp<-TreeCmp.Read('MChain', verbose=TRUE)
 Bayesian_trees<-NTS(tmp, Random51)
-cd('../../../../Analysis/')
+setwd('../../../../Analysis/')
 
 #Isolating the Bayesian treesets + summarize each RPBTC by the mode  
 Bayesian.ML<-Bayesian_trees[c(1,26,51,76,101)]
@@ -88,7 +88,7 @@ Bayesian.MLMFMC<-TreeCmp.Mode(Bayesian.MLMFMC, 1000, verbose=TRUE)
 
 #Setting the metric
 metric="R.F_Cluster"
-#metric="Triples"
+metric="Triples"
 #colour figures palette("default")
 #BW figures  
 palette(c("black", "gray50"))
[…]

### 2014-06-11 Analysis/Pariwise_TreeCmp.R
Changed cd into setwd

@@ -1,5 +1,5 @@
 #Script for testing the Random Pairwise Bayesian Tree Comparisons (RPBTC)
-#cd('../PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
+#setwd('../PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
 
 #Loading the functions
 source('../Functions/TreeComparisons/NTS.R')
@@ -15,20 +15,20 @@ Random25<-read.table("../Data/Tree_Comparisons/Random_trees/25t_Random.Cmp",head
 
 #Loading the treesets comparisons
 #Low score comparisons
-cd("../Data/Tree_Comparisons/RPBTC_testing/Low_score_treesets_100/")
+setwd("../Data/Tree_Comparisons/RPBTC_testing/Low_score_treesets_100/")
 tmp<-TreeCmp.Read('LowScore', verbose=TRUE)
 LowScore<-NTS(tmp, Random51)
-cd('../../../../Analysis/')
+setwd('../../../../Analysis/')
 #Medium score comparisons
-cd("../Data/Tree_Comparisons/RPBTC_testing/Medium_score_treesets_100/")
+setwd("../Data/Tree_Comparisons/RPBTC_testing/Medium_score_treesets_100/")
 tmp<-TreeCmp.Read('MediumScore', verbose=TRUE)
 MediumScore<-NTS(tmp, Random51)
-cd('../../../../Analysis/')
+setwd('../../../../Analysis/')
 #High score comparisons
-cd("../Data/Tree_Comparisons/RPBTC_testing/High_score_treesets_100/")
+setwd("../Data/Tree_Comparisons/RPBTC_testing/High_score_treesets_100/")
 tmp<-TreeCmp.Read('HighScore', verbose=TRUE)
 HighScore<-NTS(tmp, Random26)
-cd('../../../../Analysis/')
+setwd('../../../../Analysis/')
 
 #TreeCmp anova per metric and per score
 LS.RF<-TreeCmp.anova(LowScore, "R.F_Cluster")

### 2014-06-11 Manuscript/TEM_draft.tex
Typo
Updated title
10/06 Updates (Results ML)

@@
 
+%For a colour version of the draft to a search>replace '-BW'>''
+
 %Runing head
@@
 \noindent{\Large \bf Effect of missing data in matrices containing living and fossil taxa on topological accuracy}
-%Alternative punch title: "Bayesian method failure to infer topology using matrices containing both living and fossil taxa"
-%Alternative punch title: "ML outperforms Bayesian for inferring topology from matrices containing both living and fossil taxa"
+%Alternative punch title: "Bayesian method fails to infer topology using matrices containing both living and fossil taxa"
+%Alternative punch title: "Maximum Likelihood outperforms Bayesian for inferring topology from matrices containing both living and fossil taxa"
 \bigskip
@@
 \subsection{Building the trees}
-Bayesian chains converged (SD $<$ 0.01 and ESS $>>$ 200 for each tree).
-Good a quick convergence when using the sub-matrices especially for the "horizontal" sub-matrices.
+When generating the matrices using seqgen and rTraitDisc algorithms with low evolutionary rate parameter ($\alpha$=0.5) we successfully generated evolutionary histories.
+In Maximum Likelihood, support values ranged from 0 to 100 with a median of 38 ($1^st$ quartile = 14, $3^rd$ quartile = 72 - see \hyperref[supplementaries]{supplementaries}). % This is actually bad.
+Using Bayesian inference, all the chains converged with a significant effective sample size (ASSD $<$ 0.01 and ESS $>>$ 200 for each tree).
 Good enough phylogenetic signal.
 
-\subsection{Comparing topologies}
-Did it worked? See sub-matrices.
+%\subsection{Comparing topologies}
+%Did it worked? See sub-matrices part. TO DO
 
@@
 
+%10/06 - Missing 8 ML Chains
+
 \subsubsection{Effect of $M_L$}
-In ML...(see Table ~\ref{Group_results})
-
-Topological recovery was bad in Bayesian, the NTS score was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance.
+In Maximum Likelihood framework, the topological recovery quickly drops with the amount of missing data when using both metrics (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
+However, this negative effect of missing data seems to stabilize beyond 25\% of missing data (no significant difference for trees with respectively 25\%, 50\% and 75\% missing living taxa).
+With 10\% of missing living taxa, there is a different topological recovery depending on the used metric.
+The Robinson-Fould metric shows a lower Tree similarity metric than the Triples metric meaning that individual taxa position are generally more conserved (i.e.
+low number of flying taxa) but that the full clades are less conserved (see Table ~\ref{NTS_ML_results}).
+
+Howerver, topological recovery was bad in Bayesian, the tree similarity score was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance (see Table ~\ref{NTS_Ba_results}).
 However, it seems that there is a small difference depending on the metric used.
@@
 For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases.
-In other words, if living taxa are removed, clades are less conserved (see Table ~\ref{ML_results}).
-However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the NTS score was only slightly higher than expected by chance (see Table ~\ref{Group_results}).
+In other words, if living taxa are removed, clades are less conserved (see Table ~\ref{Ba_RF-ML_results}).
+However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the NTS score was only slightly higher than expected by chance (see Table ~\ref{Group_results} - Fig.
+~\ref{Fig_Results}).
 
 \subsubsection{Effect of $M_F$}
-In ML...(see Table ~\ref{Group_results})
+In Maximum Likelihood framework, the effect of missing data in the fossil record appeared to be constant and led to a constant decrease in topological recovery when the missing data increased (see Table ~\ref{Group_results} - Fig.
+~\ref{Fig_Results}).
+The difference between the two metrics used is more important than for the $M_L$ parameter: clades are less conserved than individual taxa placement as the amount of missing data increases.
+Interestingly, 10\% or 25\% missing data for the $M_F$ parameter does not seems to affect the apparition of unstable taxa (Triples NTS modes of respectively 0.97 and 0.93 - Table ~\ref{NTS_ML_results}).
 
@@
 And in the same way as described above, there seems to be a small difference depending on the metric used.
-There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric (see Table ~\ref{ML_results}).
-However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework (see Table ~\ref{Group_results}).
+There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric (see Table ~\ref{Ba_RF-MF_results}).
+However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
 
 \subsubsection{Effect of $M_C$}
-In ML...(see Table ~\ref{Group_results})
-
-The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Fould metric shows a significant effect of the $M_C$ parameter on tree recovery (see Table ~\ref{ML_results} and ~\ref{Group_results}).
-\subsubsection{Combined effect}
-
-In ML...(see Table ~\ref{Group_results})
-
-In Bayesian, same pattern as above.
-No particular effect of data combination, but more an effect of the overall amount of data (see Table ~\ref{Group_results}).
+The number of missing morphological characters ($M_C$), however, seems to be more affecting topological recovery than $M_L$ and $M_F$ (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
+The Robinson-Fould metric shows a rapid drop in tree recovery from 10\% of missing data (Robinson-Fould NTS mode $<$ to 0.5 from 10\% missing data - Table ~\ref{NTS_ML_results}).
+This decrease is however slowlier when using the Triples metric with still a good topological recovery at 10\%  (Triples NTS modes = 0.95 - Table ~\ref{NTS_ML_results}).
+
+The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Fould metric shows a significant effect of the $M_C$ parameter on tree recovery (see Table ~\ref{Ba_RF-ML_results} and ~\ref{Group_results} - Fig.
+~\ref{Fig_Results}).
+
+%\subsubsection{Combined effect}
+%Need to do the heat plots
 
@@
 
-The ability of recovering the best topology in maximum likelihood method is function of the amount of data missing.
-However, in Bayesian, topology is badly (if not at all) recovered independently of the amount of data missing.
-
-%Add the global heat plot figures
+%To sum up:
+The ability of recovering the "best" tree's topology in maximum likelihood method is function of the amount of data missing.
+The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
+The decrease in clade conservation (low Robinson-Fould NTS score) is faster than the increase in flying taxa (low Triples NTS score).
+However, in Bayesian, topology is badly (if not at all) recovered disregarding the amount of data missing.
+
+\begin{table}
+\caption{Tree similarity values per parameter in ML framework}
+\centering
+\begin{tabular}{rrlccccc}
+    \hline
+    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
+              & missing data &        &      & lower  & upper & lower  & upper \\
+    \hline
+    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
+              &              & Triples        & 1 & NA  & NA & NA  & NA \\
+              & 10\%         & Robinson-Fould & 0.671 & 0.569  & 0.789 & 0.468  & 1.036 \\ %how many digits?
+              &              & Triples        & 0.942 & 0.868  & 0.999 & 0.612  & 1.063 \\
+              & 25\%         & Robinson-Fould & 0.516 & 0.440  & 0.630 & 0.266  & 0.892 \\
+              &              & Triples        & 0.559 & 0.708  & 0.810 & 0.291  & 0.999 \\
+              & 50\%         & Robinson-Fould & 0.432 & 0.364  & 0.521 & 0.231  & 0.642 \\
+              &              & Triples        & 0.514 & 0.419  & 0.599 & 0.293  & 0.910 \\
+              & 75\%         & Robinson-Fould & 0.372 & 0.323  & 0.434 & 0.203  & 0.577 \\
+              &              & Triples        & 0.444 & 0.338  & 0.532 & 0.192  & 0.855 \\
+    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
+              &              & Triples        & 1 & NA  & NA & NA  & NA \\
+              & 10\%         & Robinson-Fould & 0.907 & 0.766  & 1.000 & 0.407  & 1.047 \\
+              &              & Triples        & 0.976 & 0.944  & 1.003 & 0.715  & 1.104 \\
+              & 25\%         & Robinson-Fould & 0.572 & 0.488  & 0.692 & 0.325  & 0.936 \\
+              &              & Triples        & 0.937 & 0.820  & 0.993 & 0.539  & 1.086 \\
+              & 50\%         & Robinson-Fould & 0.433 & 0.302  & 0.476 & 0.170  & 0.668 \\
+              &              & Triples        & 0.732 & 0.594  & 0.873 & 0.341  & 1.025 \\
+              & 75\%         & Robinson-Fould & 0.153 & 0.115  & 0.200 & 0.068  & 0.323 \\
+              &              & Triples        & 0.458 & 0.390  & 0.611 & 0.290  & 0.856 \\   
+    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
+              &              & Triples        & 1 & NA  & NA & NA  & NA \\
+              & 10\%         & Robinson-Fould & 0.473 & 0.384  & 0.835 & 0.356  & 0.958 \\
+              &              & Triples        & 0.959 & 0.869  & 1.026 & 0.572  & 1.067 \\
+              & 25\%         & Robinson-Fould & 0.454 & 0.357  & 0.630 & 0.242  & 0.916 \\
+              &              & Triples        & 0.907 & 0.751  & 0.997 & 0.446  & 1.070 \\
+              & 50\%         & Robinson-Fould & 0.424 & 0.375  & 0.466 & 0.249  & 0.828 \\
+              &              & Triples        & 0.683 & 0.571  & 0.822 & 0.277  & 1.008 \\
+              & 75\%         & Robinson-Fould & 0.230 & 0.177  & 0.302 & 0.108  & 0.567 \\
+              &              & Triples        & 0.451 & 0.373  & 0.572 & 0.224  & 0.791 \\
+    \hline
+\label{NTS_ML_results}
+\end{tabular}
+\end{table}
+
+\begin{table}
+\caption{Tree similarity values per parameter in Bayesian framework}
+\centering
+\begin{tabular}{rrlccccc}
+    \hline
+    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
+              & missing data &      & lower  & upper & lower  & upper \\
+    \hline
+    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\ %how many digits?
+              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
+              & 10\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.028   & 0.061 \\
+              &              & Triples        & 0.033 & 0.003  & 0.049 & -0.049  & 0.117 \\
+              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.061 \\
+              &              & Triples        & 0.035 & 0.002  & 0.049 & -0.041  & 0.096 \\
+              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.056 \\
+              &              & Triples        & 0.028 & 0.003  & 0.052 & -0.038  & 0.092 \\
+              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.010   & 0.043 \\
+              &              & Triples        & -0.001 & -0.018 & 0.040 & -0.060 & 0.106 \\
+    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
+              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
+              & 10\%         & Robinson-Fould & 0.035 & 0.031  & 0.056 & 0.027   & 0.076 \\
+              &              & Triples        & 0.026 & 0.002  & 0.050 & -0.046  & 0.105 \\
+              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.0153  & 0.068 \\
+              &              & Triples        & 0.038 & 0.002  & 0.050 & -0.046  & 0.105 \\
+              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.058 \\
+              &              & Triples        & 0.014 & -0.011 & 0.039 & -0.038  & 0.124 \\
+              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.041 \\
+              &              & Triples        & 0.029 & 0.019  & 0.042 & -0.030  & 0.089 \\   
+    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
+              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
+              & 10\%         & Robinson-Fould & 0.056 & 0.035  & 0.059 & 0.0281  & 0.076 \\
+              &              & Triples        & 0.030 & 0.011  & 0.048 & -0.024  & 0.094 \\
+              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.023   & 0.076 \\
+              &              & Triples        & 0.020 & -0.001 & 0.037 & -0.060  & 0.109 \\
+              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.030   & 0.059 \\
+              &              & Triples        & 0.010 & -0.005 & 0.045 & -0.036  & 0.104 \\
+              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.011   & 0.043 \\
+              &              & Triples        & 0.024 & 0.001  & 0.042 & -0.037  & 0.104 \\
+    \hline
+\label{NTS_BA_results}
+\end{tabular}
+\end{table}
+
 \begin{table}
@@
 \centering
-\begin{tabular}{rlrlccc}
+\begin{tabular}{rllcccc}
     \hline
@@
     \hline
-    $M_L$ & Robinson-Fould & ML & bla & bla & bla & bla \\
-          &                & Bayesian & no & 65.62 & 4.00 & \textbf{0.00} \\
-          & Triples & ML & bla & bla & bla & bla \\
-          &         & Bayesian & no & 4.56 & 4.00 & 0.34 \\
-    $M_F$ & Robinson-Fould & ML & bla & bla & bla & bla \\
-          &                & Bayesian & no & 85.61 & 4.00 & \textbf{0.00} \\
-          & Triples & ML & bla & bla & bla & bla \\
-          &         & Bayesian & no & 3.01 & 4.00 & 0.56 \\ 
-    $M_C$ & Robinson-Fould & ML & bla & bla & bla & bla \\
-          &                & Bayesian & no & 64.23 & 4.00 & \textbf{0.00} \\
-          & Triples & ML & bla & bla & bla & bla \\
-          &         & Bayesian & no & 26.13 & 24.00 & 0.35 \\
-    $M_L+M_F$ & Robinson-Fould & ML & bla & bla & bla & bla \\
-              &                & Bayesian & no & 317.97 & 24.00 & \textbf{0.00} \\
-              & Triples & ML & bla & bla & bla & bla \\
-              &         & Bayesian & no & 3.01 & 4.00 & 0.56 \\ 
-    $M_L+M_C$ & Robinson-Fould & ML & bla & bla & bla & bla \\
-              &                & Bayesian & no & 290.10 & 24.00 & \textbf{0.00} \\
-              & Triples & ML & bla & bla & bla & bla \\
-              &         & Bayesian & no & 22.19 & 24.00 & 0.57 \\ 
-    $M_F+M_C$ & Robinson-Fould & ML & bla & bla & bla & bla \\
-              &                & Bayesian & no & 385.96 & 24.00 & \textbf{0.00} \\
-              & Triples & ML & bla & bla & bla & bla \\
-              &         & Bayesian & no & 20.23 & 24.00 & 0.68 \\
-    $M_L+M_F+M_C$ & Robinson-Fould & ML & bla & bla & bla & bla \\
-                  &                & Bayesian & no & 1167.40 & 124.00 & \textbf{0.00} \\
-                  & Triples & ML & bla & bla & bla & bla \\
-                  &         & Bayesian & no & 115.86 & 124.00 & 0.69 \\ 
+    $M_L$         & Robinson-Fould & ML       & no & 151.438  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.001278e-31
+                  &                & Bayesian & no & 65.62    & 4   & \textbf{0.00} \\ 
+                  & Triples        & ML       & no & 142.8329 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains  - 6.983434e-30
+                  &                & Bayesian & no & 4.56     & 4   & 0.34          \\ 
+    $M_F$         & Robinson-Fould & ML       & no & 179.9238 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 7.743067e-38
+                  &                & Bayesian & no & 85.61    & 4   & \textbf{0.00} \\
+                  & Triples        & ML       & no & 150.059  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.977386e-31
+                  &                & Bayesian & no & 3.01     & 4   & 0.56          \\
+    $M_C$         & Robinson-Fould & ML       & no & 151.3818 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.029428e-31
+                  &                & Bayesian & no & 64.23    & 4   & \textbf{0.00} \\
+                  & Triples        & ML       & no & 138.374  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 6.290979e-29
+                  &                & Bayesian & no & 26.13    & 24  & 0.35          \\
+    $M_L+M_F$     & Robinson-Fould & ML       & no & 734.9213 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.104542e-139
+                  &                & Bayesian & no & 317.97   & 24  & \textbf{0.00} \\
+                  & Triples        & ML       & no & 531.5558 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 4.581065e-97
+                  &                & Bayesian & no & 3.01     & 24  & 0.56          \\
+    $M_L+M_C$     & Robinson-Fould & ML       & no & 542.3233 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 2.619808e-99
+                  &                & Bayesian & no & 290.10   & 24  & \textbf{0.00} \\
+                  & Triples        & ML       & no & 437.7413 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains -  1.28493e-77
+                  &                & Bayesian & no & 22.19    & 24  & 0.57          \\
+    $M_F+M_C$     & Robinson-Fould & ML       & no & 812.206  & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 5.463138e-156
+                  &                & Bayesian & no & 385.96   & 24  & \textbf{0.00} \\
+                  & Triples        & ML       & no & 528.9215 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.619383e-96
+                  &                & Bayesian & no & 20.23    & 24  & 0.68          \\ 
+    $M_L+M_F+M_C$ & Robinson-Fould & ML       & no & 3603.808 & 124 & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains -3.819574e-169
+                  &                & Bayesian & no & 1167.40  & 124 & \textbf{0.00} \\
+                  & Triples        & ML       & no & 2047.923 & 124 & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains 
+                  &                & Bayesian & no & 115.86   & 124 & 0.69          \\
     \hline
@@
 \begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric}
+\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
+\centering
+\begin{tabular}{c|ccccc}
+    \hline
+              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
+    \hline
+    $M_L$00\% & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
+    $M_L$10\% & & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} \\
+    $M_L$25\% & & & - & FALSE  & \textbf{TRUE} \\
+    $M_L$50\% & & & & - & FALSE  \\
+    $M_L$75\% & & & & & - \\
+    \hline
+\end{tabular}
+\centering
+\begin{tabular}{rrrl}
+ pairwise comparison & obs.dif & critical.dif & difference \\ 
+  \hline
+  $M_L$00\%-$M_L$10\% & 56.53 & 37.66 & TRUE \\ 
+  $M_L$00\%-$M_L$25\% & 96.85 & 37.66 & TRUE \\ 
+  $M_L$00\%-$M_L$50\% & 127.03 & 37.66 & TRUE \\ 
+  $M_L$00\%-$M_L$75\% & 144.58 & 37.66 & TRUE \\ 
+  $M_L$10\%-$M_L$25\% & 40.31 & 37.66 & TRUE \\ 
+  $M_L$10\%-$M_L$50\% & 70.50 & 37.66 & TRUE \\ 
+  $M_L$10\%-$M_L$75\% & 88.05 & 37.66 & TRUE \\ 
+  $M_L$25\%-$M_L$50\% & 30.19 & 37.66 & FALSE \\ 
+  $M_L$25\%-$M_L$75\% & 47.73 & 37.66 & TRUE \\ 
+  $M_L$50\%-$M_L$75\% & 17.55 & 37.66 & FALSE \\ 
+   \hline
+\end{tabular}
+\label{ML_RF-ML_results}
+\end{table}
+
+\begin{table}
+\caption{$M_L$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
+\centering
+\begin{tabular}{c|ccccc}
+    \hline
+              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
+    \hline
+    $M_L$00\% & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
+    $M_L$10\% & & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} \\
+    $M_L$25\% & & & - & FALSE  & FALSE \\
+    $M_L$50\% & & & & - & FALSE  \\
+    $M_L$75\% & & & & & - \\
+    \hline
+\end{tabular}
+\centering
+\begin{tabular}{rrrl}
+ pairwise comparison & obs.dif & critical.dif & difference \\ 
+  \hline
+  $M_L$00\%-$M_L$10\% & 57.98 & 37.66 & TRUE \\ 
+  $M_L$00\%-$M_L$25\% & 105.83 & 37.66 & TRUE \\ 
+  $M_L$00\%-$M_L$50\% & 120.86 & 37.66 & TRUE \\ 
+  $M_L$00\%-$M_L$75\% & 140.34 & 37.66 & TRUE \\ 
+  $M_L$10\%-$M_L$25\% & 47.85 & 37.66 & TRUE \\ 
+  $M_L$10\%-$M_L$50\% & 62.88 & 37.66 & TRUE \\ 
+  $M_L$10\%-$M_L$75\% & 82.36 & 37.66 & TRUE \\ 
+  $M_L$25\%-$M_L$50\% & 15.03 & 37.66 & FALSE \\ 
+  $M_L$25\%-$M_L$75\% & 34.51 & 37.66 & FALSE \\ 
+  $M_L$50\%-$M_L$75\% & 19.48 & 37.66 & FALSE \\ 
+   \hline
+\end{tabular}
+\label{ML_Tr-ML_results}
+\end{table}
+
+\begin{table}
+\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
 \centering
@@
 \end{tabular}
-\label{ML_results}
+\label{Ba_RF-ML_results}
+\end{table}
+
+\begin{table}
+\caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
+\centering
+\begin{tabular}{c|ccccc}
+    \hline
+              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
+    \hline
+    $M_L$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
+    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
+    $M_L$25\% & & & - & \textbf{TRUE} & \textbf{TRUE} \\
+    $M_L$50\% & & & & - & \textbf{TRUE} \\
+    $M_L$75\% & & & & & - \\
+    \hline
+\end{tabular}
+\centering
+\begin{tabular}{rrrl}
+ pairwise comparison & obs.dif & critical.dif & difference \\ 
+  \hline
+  $M_F$00\%-$M_F$10\% & 59.27 & 37.66 & TRUE \\ 
+  $M_F$00\%-$M_F$25\% & 79.72 & 37.66 & TRUE \\ 
+  $M_F$00\%-$M_F$50\% & 120.70 & 37.66 & TRUE \\ 
+  $M_F$00\%-$M_F$75\% & 167.81 & 37.66 & TRUE \\ 
+  $M_F$10\%-$M_F$25\% & 20.45 & 37.66 & FALSE \\ 
+  $M_F$10\%-$M_F$50\% & 61.43 & 37.66 & TRUE \\ 
+  $M_F$10\%-$M_F$75\% & 108.55 & 37.66 & TRUE \\ 
+  $M_F$25\%-$M_F$50\% & 40.98 & 37.66 & TRUE \\ 
+  $M_F$25\%-$M_F$75\% & 88.09 & 37.66 & TRUE \\ 
+  $M_F$50\%-$M_F$75\% & 47.12 & 37.66 & TRUE \\ 
+   \hline
+\end{tabular}
+\label{ML_RF-MF_results}
+\end{table}
+
+\begin{table}
+\caption{$M_F$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
+\centering
+\begin{tabular}{c|ccccc}
+    \hline
+              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
+    \hline
+    $M_L$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
+    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
+    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
+    $M_L$50\% & & & & - & FALSE \\
+    $M_L$75\% & & & & & - \\
+    \hline
+\end{tabular}
+\centering
+\begin{tabular}{rrrl}
+ pairwise comparison & obs.dif & critical.dif & difference \\ 
+  \hline
+  $M_F$00\%-$M_F$10\% & 64.78 & 37.66 & TRUE \\ 
+  $M_F$00\%-$M_F$25\% & 89.09 & 37.66 & TRUE \\ 
+  $M_F$00\%-$M_F$50\% & 123.20 & 37.66 & TRUE \\ 
+  $M_F$00\%-$M_F$75\% & 150.43 & 37.66 & TRUE \\ 
+  $M_F$10\%-$M_F$25\% & 24.31 & 37.66 & FALSE \\ 
+  $M_F$10\%-$M_F$50\% & 58.42 & 37.66 & TRUE \\ 
+  $M_F$10\%-$M_F$75\% & 85.65 & 37.66 & TRUE \\ 
+  $M_F$25\%-$M_F$50\% & 34.10 & 37.66 & FALSE \\ 
+  $M_F$25\%-$M_F$75\% & 61.34 & 37.66 & TRUE \\ 
+  $M_F$50\%-$M_F$75\% & 27.23 & 37.66 & FALSE \\ 
+   \hline
+\end{tabular}
+\label{ML_Tr-MF_results}
 \end{table}
@@
 \end{tabular}
-\label{MF_results}
+\label{Ba_RF-MF_results}
+\end{table}
+
+\begin{table}
+\caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
+\centering
+\begin{tabular}{c|ccccc}
+    \hline
+              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
+    \hline
+    $M_C$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
+    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
+    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
+    $M_C$50\% & & & & - & \textbf{TRUE} \\
+    $M_C$75\% & & & & & - \\
+    \hline
+\end{tabular}
+\centering
+\begin{tabular}{rrrl}
+ pairwise comparison & obs.dif & critical.dif & difference \\ 
+  \hline
+  $M_C$00\%-$M_C$10\% & 67.99 & 37.66 & TRUE \\ 
+  $M_C$00\%-$M_C$25\% & 87.58 & 37.66 & TRUE \\ 
+  $M_C$00\%-$M_C$50\% & 116.16 & 37.66 & TRUE \\ 
+  $M_C$00\%-$M_C$75\% & 155.77 & 37.66 & TRUE \\ 
+  $M_C$10\%-$M_C$25\% & 19.59 & 37.66 & FALSE \\ 
+  $M_C$10\%-$M_C$50\% & 48.17 & 37.66 & TRUE \\ 
+  $M_C$10\%-$M_C$75\% & 87.78 & 37.66 & TRUE \\ 
+  $M_C$25\%-$M_C$50\% & 28.58 & 37.66 & FALSE \\ 
+  $M_C$25\%-$M_C$75\% & 68.19 & 37.66 & TRUE \\ 
+  $M_C$50\%-$M_C$75\% & 39.60 & 37.66 & TRUE \\ 
+   \hline
+\end{tabular}
+\label{ML_RF-MC_results}
+\end{table}
+
+\begin{table}
+\caption{$M_C$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
+\centering
+\begin{tabular}{c|ccccc}
+    \hline
+              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
+    \hline
+    $M_C$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
+    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
+    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
+    $M_C$50\% & & & & - & FALSE \\
+    $M_C$75\% & & & & & - \\
+    \hline
+\end{tabular}
+\centering
+\begin{tabular}{rrrl}
+ pairwise comparison & obs.dif & critical.dif & difference \\ 
+  \hline
+  $M_C$00\%-$M_C$10\% & 75.29 & 37.66 & TRUE \\ 
+  $M_C$00\%-$M_C$25\% & 88.58 & 37.66 & TRUE \\ 
+  $M_C$00\%-$M_C$50\% & 114.02 & 37.66 & TRUE \\ 
+  $M_C$00\%-$M_C$75\% & 149.60 & 37.66 & TRUE \\ 
+  $M_C$10\%-$M_C$25\% & 13.29 & 37.66 & FALSE \\ 
+  $M_C$10\%-$M_C$50\% & 38.73 & 37.66 & TRUE \\ 
+  $M_C$10\%-$M_C$75\% & 74.31 & 37.66 & TRUE \\ 
+  $M_C$25\%-$M_C$50\% & 25.44 & 37.66 & FALSE \\ 
+  $M_C$25\%-$M_C$75\% & 61.02 & 37.66 & TRUE \\ 
+  $M_C$50\%-$M_C$75\% & 35.58 & 37.66 & FALSE \\ 
+   \hline
+\end{tabular}
+\label{ML_Tr-MC_results}
 \end{table}
@@
 \end{tabular}
-\label{MC_results}
+\label{Ba_RF-MC_results}
 \end{table}
@@
 After a certain amount of morphological characters, adding new ones adds homoplasy \citep{wagner2000}.
+
+Difference between Robinson-Fould and Triples?
+Robison-Fould is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies.
+When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
+On the other hand, the Triples method is measuring the position of each taxon towards to other reference taxa.
+It will penalise only trees where taxa get removed furthest from their original clade.
+Looking at our problematic (how does missing data influence topological recovery in trees containing both living and fossil taxa) we are more interested the placement of taxa (i.e.
+where does the fossil branch) than the exact conservation of clades.
+
+What's the crucial factor?
+Regarding the fact that we are more interested in taxa placement than in clade conservation (Triples vs.
+Robinson-Fould), the parameter that affects the more the decrease in topological recovery if the number of missing living taxa in the morphological part of the matrix ($M_L$).
+This makes sense since the living taxa are bearing both the information to build the tree backbone (the molecular data) and the information used for branching the fossil on this backbone (the morphological information).
+Therefore, we advocate that an effort should be made in collecting all the available morphological information for living taxa in order to properly run Total Evidence analysis.
+%yep, we need data from living taxa.
 
@@
 However, one can use a Maximum Likelihood approach to obtain the likeliest topology and to use this one as a fixed.
+If so, an effort should be made on collecting morphological data from living taxa in order to efficiently improve the quality of the trees.
+
+%Perspectives:
+%What the effect of this missing molecular data on branch length estimation?
 
@@
 
+\subsection{Full results}
+
+\begin{table}
+\caption{Tree similarity values per parameter in ML framework}
+\centering
+\begin{tabular}{rrlccccc}
+    \hline
+    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
+              & missing data &        &      & lower  & upper & lower  & upper \\
+    \hline
+    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
+              &              & Triples        & 1 & NA  & NA & NA  & NA \\
+              & 10\%         & Robinson-Fould & 0.6714878 & 0.5692308  & 0.7894649 & 0.4689069  & 1.0361514 \\ %how many digits?
+              &              & Triples        & 0.9424133 & 0.8683742  & 0.9998559 & 0.6121399  & 1.0635250 \\
+              & 25\%         & Robinson-Fould & 0.5167595 & 0.4403147  & 0.6307692 & 0.2664429  & 0.8929104 \\
+              &              & Triples        & 0.5599494 & 0.7084874  & 0.8105781 & 0.2912950  & 0.9990703 \\
+              & 50\%         & Robinson-Fould & 0.4321852 & 0.3641026  & 0.5215627 & 0.2313246  & 0.6423846 \\
+              &              & Triples        & 0.5143009 & 0.4197127  & 0.5991498 & 0.2991432  & 0.9104526 \\
+              & 75\%         & Robinson-Fould & 0.3727090 & 0.3230769  & 0.4342249 & 0.2035474  & 0.5776286 \\
+              &              & Triples        & 0.4449563 & 0.3385128  & 0.5329149 & 0.1929395  & 0.8551570 \\
+    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
+              &              & Triples        & 1 & NA  & NA & NA  & NA \\
+              & 10\%         & Robinson-Fould & 0.9072419 & 0.7668415  & 1.0000000 & 0.4070844  & 1.0478010 \\
+              &              & Triples        & 0.9763617 & 0.9448100  & 1.0039460 & 0.7158118  & 1.1043330 \\
+              & 25\%         & Robinson-Fould & 0.5722382 & 0.4883560  & 0.6923077 & 0.3253239  & 0.9367467 \\
+              &              & Triples        & 0.9370152 & 0.8207407  & 0.9938846 & 0.5394443  & 1.0868073 \\
+              & 50\%         & Robinson-Fould & 0.4338159 & 0.3025641  & 0.4760777 & 0.1704016  & 0.6683220 \\
+              &              & Triples        & 0.7327464 & 0.5947211  & 0.8739630 & 0.3418679  & 1.0251540 \\
+              & 75\%         & Robinson-Fould & 0.1534487 & 0.1155128  & 0.2000000 & 0.0683485  & 0.3230769 \\
+              &              & Triples        & 0.4588292 & 0.3905631  & 0.6111484 & 0.2903676  & 0.8568274 \\   
+    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
+              &              & Triples        & 1 & NA  & NA & NA  & NA \\
+              & 10\%         & Robinson-Fould & 0.473285 & 0.3842965  & 0.8358974 & 0.3564753  & 0.9589744 \\
+              &              & Triples        & 0.9592557 & 0.8691580  & 1.026095 & 0.5720847  & 1.067524 \\
+              & 25\%         & Robinson-Fould & 0.4548038 & 0.3572132  & 0.6307692 & 0.2425339  & 0.9161212 \\
+              &              & Triples        & 0.9071925 & 0.7516934  & 0.9976224 & 0.4462168  & 1.0705291 \\
+              & 50\%         & Robinson-Fould & 0.4244263 & 0.3755493  & 0.4666667 & 0.2494927  & 0.8287312 \\
+              &              & Triples        & 0.6834366 & 0.5713927  & 0.8226139 & 0.2776690  & 1.0085944 \\
+              & 75\%         & Robinson-Fould & 0.2306888 & 0.1774508  & 0.3025641 & 0.1087764  & 0.5673541 \\
+              &              & Triples        & 0.4513642 & 0.3737397  & 0.5723858 & 0.2245224  & 0.7917946 \\
+    \hline
+\label{NTSML_full}
+\end{tabular}
+\end{table}
+
+\begin{table}
+\caption{Tree similarity values per parameter in Bayesian framework}
+\centering
+\begin{tabular}{rrlccccc}
+    \hline
+    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
+              & missing data &      & lower  & upper & lower  & upper \\
+    \hline
+    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\ %how many digits?
+              &              & Triples        & 0.04104231 & 0.00830962  & 0.05441173 & -0.03933377  & 0.10690054 \\
+              & 10\%         & Robinson-Fould & 0.03584309 & 03584196  & 03589744 & 02825643  & 0.06175676 \\
+              &              & Triples        & 0.03399849 & 0.003196469  & 0.04981018 & -0.049913382  & 0.11784366 \\
+              & 25\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01536739  & 0.06123106 \\
+              &              & Triples        & 0.03510953 & 0.002113358  & 0.04939228 & -0.041202264  & 0.09699332 \\
+              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01538462  & 0.05641026 \\
+              &              & Triples        & 0.02849433 & 0.003263098  & 0.05298036 & -0.038344731  & 0.09250063 \\
+              & 75\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01069992  & 0.04338811 \\
+              &              & Triples        & -0.0001940566 & -0.01801301  & 0.04058783 & -0.06030922  & 0.10665934 \\
+    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\
+              &              & Triples        & 0.04104231 & 0.008309622  & 0.05441173 & -0.039333778  & 0.10690054 \\
+              & 10\%         & Robinson-Fould & 0.03593712 & 0.03198256  & 0.05641026 & 0.02798864  & 0.07692308 \\
+              &              & Triples        & 0.02639309 & 0.002185408  & 0.05061314 & -0.046563042  & 0.10547052 \\
+              & 25\%         & Robinson-Fould & 0.03588134 & 0.03587629  & 0.03589744 & 0.01538462  & 0.0680884 \\
+              &              & Triples        & 0.03856296 & 0.002185408  & 0.05061314 & -0.046563042  & 0.10547052 \\
+              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01536289  & 0.05879269 \\
+              &              & Triples        & 0.01403716 & -0.01100608  & 0.03972324 & -0.03800324  & 0.12405207 \\
+              & 75\%         & Robinson-Fould & 0.03589206 & 0.03588825  & 0.03589744 & 0.01538462  & 0.0411843 \\
+              &              & Triples        & 0.02939363 & 0.01902351  & 0.04238907 & -0.03087905  & 0.08946917 \\   
+    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\
+              &              & Triples        & 0.04104231 & 0.008309622  & 0.05441173 & -0.039333778  & 0.10690054 \\
+              & 10\%         & Robinson-Fould & 0.05637057 & 0.03589744  & 0.05927956 & 0.02813296  & 0.07692308 \\
+              &              & Triples        & 0.03008132 & 0.01176800  & 0.04847400 & -0.02451461  & 0.09453754 \\
+              & 25\%         & Robinson-Fould & 0.03593743 & 0.03589744  & 0.03596018 & 0.02330790  & 0.07692308 \\
+              &              & Triples        & 0.02010221 & -0.001145118  & 0.0377779 & -0.060183332  & 0.1098632 \\
+              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.03050289  & 0.05950426 \\
+              &              & Triples        & 0.01053959 & -0.005595947  & 0.04510444 & -0.036278820  & 0.10478805 \\
+              & 75\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01115983  & 0.04362927 \\
+              &              & Triples        & 0.02427778 & 0.0007405737  & 0.04260522 & -0.0373004106  & 0.1048756 \\
+\label{NTSBA_full}
+\end{tabular}
+\end{table}
+
+\subsection{Bootstraps distribution}
+
+\begin{figure}
+\centering
+\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-AppendixBootstraps.pdf}
+\caption{Bootstraps distribution across the "best" trees.}                     % MISSING 8 CHAINS
+\label{Fig_AppendixBoostraps}
+\end{figure}
+
 %END

### 2014-06-11 Manuscript/TEM_draft.tex
Draft version 0.1 (Thomas)

@@
 
-We compared the "missing-data" tree to the "True" and the "Best" tree for each chain.
-For the Maximum Likelihood trees we performed pairwise comparisons between the "True" and "Best" tree and the "missing-data" tree (see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triples metric.
+We compared the "missing-data" tree to the "Best" tree for each chain.
+For the Maximum Likelihood trees we performed pairwise comparisons between the "Best" tree and the "missing-data" tree (see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triples metric.
 We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
@@
 From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we inferred a tree in Baysian or Maximum likelihood framework.
-We then compared this tree with or without fossil/living taxa to the "best/true" tree.
+We then compared this tree with or without fossil/living taxa to the "best" tree.
 In the Bayesian framework, 1000 trees from the posterior distribution where used for each comparison.}
@@
 
-\subsubsection{Statistical difference between tree topologies}
+\subsubsection{Statistical difference between tree topologies} %note that in the end, I only have non-parametric tests, the parametric explanation part can go to the supplementaries
 To assess the difference between the tree topologies, whether they were inferred in a Bayesian or a ML framework, we performed a group comparison test for each set of configurations (e.g.
@@
 \section{Results}
-%Do the check for L00F00C00
+
 \subsection{Building the trees}
-When generating the matrices using seqgen and rTraitDisc algorithms with low evolutionary rate parameter ($\alpha$=0.5) we successfully generated evolutionary histories.
+When generating the matrices using seqgen \citep{ranbaut1997seqgen} and rTraitDisc \citep{paradisape:2004} algorithms with low evolutionary rate parameter ($\alpha$=0.5) we successfully generated evolutionary histories.
 In Maximum Likelihood, support values ranged from 0 to 100 with a median of 38 ($1^st$ quartile = 14, $3^rd$ quartile = 72 - see \hyperref[supplementaries]{supplementaries}). % This is actually bad.
-Using Bayesian inference, all the chains converged with a significant effective sample size (ASSD $<$ 0.01 and ESS $>>$ 200 for each tree).
-Good enough phylogenetic signal.
+The distribution of the phylogenetic signal across the different chains provided us a good spread of evolutionary scenarios: from well resolved phylogenies to poorly resolved ones.
+This allows us to test our simulations in a theoretical as well as practical framework (i.e.
+phylogenies with low support values are rarely published but are still common in early stages of projects).
+%REF???
+- That's kind of my way to turn these bad results into something usable: even if some of our simulations are not realistic, it shows the effect of missing data in any scenario possible, not only in realistic cases.
+Also, one can advocate that this low topological resolution is actively due to the amount of missing data in the total evidence matrices since the sub-matrices (no missing data) analysis recovers way higher support value  .
+%This is something that I will had with the whole submatrices story.
+Basically, when you have all the fossil/molecular data missing, it's harder to build the tree, both in ML and in Bayesian.
+When you run the analysis without missing data, then it works way better (i.e.
+our REAL control test: can we simulate matrices in the first place).
+Using Bayesian inference, all the chains converged with a significant effective sample size by using 2 parallel runs of 4 chains each (ASSD $<$ 0.01 and ESS $>>$ 200 for inference).
 
@@
 %10/06 - Missing 8 ML Chains
+\begin{figure} 
+\centering
+\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-results-BW.pdf}
+\caption{Effect of missing data on Tree similarity.
+$M_L$ is the amount of missing living taxa with morphological data,
+$M_F$ is the amount of missing data in the fossil record,
+$M_C$ is the amount of missing morphological characters.
+The first row represent the Normalised Tree Similarity index using the Robinson-Fould metric;
+the second row represent the Normalised Tree Similarity index using the Triples metric.}
+\label{Fig_Results}
+\end{figure}
 
 \subsubsection{Effect of $M_L$}
-In Maximum Likelihood framework, the topological recovery quickly drops with the amount of missing data when using both metrics (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
-However, this negative effect of missing data seems to stabilize beyond 25\% of missing data (no significant difference for trees with respectively 25\%, 50\% and 75\% missing living taxa).
+In Maximum Likelihood framework, the topological recovery (reflected by the tree similarity) quickly drops with the increasing amount of missing data when using both metrics (see Table ~\ref{Group_results} - Fig.
+~\ref{Fig_Results}).
+However, this negative effect of missing data is stabilized beyond 25\% of missing data (no significant difference for trees with respectively 25\%, 50\% and 75\% missing living taxa).
 With 10\% of missing living taxa, there is a different topological recovery depending on the used metric.
-The Robinson-Fould metric shows a lower Tree similarity metric than the Triples metric meaning that individual taxa position are generally more conserved (i.e.
+The Robinson-Fould metric shows a lower Tree similarity index than the Triples metric meaning that individual taxa position are generally more conserved (i.e.
 low number of flying taxa) but that the full clades are less conserved (see Table ~\ref{NTS_ML_results}).
 
-Howerver, topological recovery was bad in Bayesian, the tree similarity score was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance (see Table ~\ref{NTS_Ba_results}).
-However, it seems that there is a small difference depending on the metric used.
+However, topological recovery was bad in Bayesian, regardless the amount of missing data.
+The tree similarity index was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance (see Table ~\ref{NTS_Ba_results}).
+Even if the topological recovery is really low, we detected an effect of missing data on the topology depending on the used metric.
 For the Triples metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
-For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases.
-In other words, if living taxa are removed, clades are less conserved (see Table ~\ref{Ba_RF-ML_results}).
-However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the NTS score was only slightly higher than expected by chance (see Table ~\ref{Group_results} - Fig.
+For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases (see Table ~\ref{Ba_RF-ML_results}).
+However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the tree similarity index was only slightly higher than expected by chance (see Table ~\ref{Group_results} - Fig.
 ~\ref{Fig_Results}).
@@
 Similarly than for the effect of $M_L$, Bayesian tree recovery was only slightly better than expected by chance.
-And in the same way as described above, there seems to be a small difference depending on the metric used.
+In the same way as described above, there seems to be a small difference depending on the metric used.
 There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric (see Table ~\ref{Ba_RF-MF_results}).
@@
 The Robinson-Fould metric shows a rapid drop in tree recovery from 10\% of missing data (Robinson-Fould NTS mode $<$ to 0.5 from 10\% missing data - Table ~\ref{NTS_ML_results}).
-This decrease is however slowlier when using the Triples metric with still a good topological recovery at 10\%  (Triples NTS modes = 0.95 - Table ~\ref{NTS_ML_results}).
+This decrease is however slower when using the Triples metric with still a good topological recovery at 10\%  (Triples NTS modes = 0.95 - Table ~\ref{NTS_ML_results}).
 
@@
 %To sum up:
-The ability of recovering the "best" tree's topology in maximum likelihood method is function of the amount of data missing.
+The ability of recovering the "best" tree's topology in Maximum Likelihood method is function of the amount of data missing.
 The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
 The decrease in clade conservation (low Robinson-Fould NTS score) is faster than the increase in flying taxa (low Triples NTS score).
-However, in Bayesian, topology is badly (if not at all) recovered disregarding the amount of data missing.
-
-\begin{table}
+In Bayesian framework, topology is badly (if not at all) recovered disregarding the amount of data missing.
+
+\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
 \caption{Tree similarity values per parameter in ML framework}
@@
 
-\begin{table}
+\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
 \caption{Tree similarity values per parameter in Bayesian framework}
@@
 \subsection{Building the trees}
-Hard to simulate morpho characters but see Pagel, April Wright, etc...
+Simulating evolutionary history matrices still remains a big drawback in theoretical phylogenetics. %Hahahaha.... feck.
+The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
+This configuration probably lead to globally low phylogenetic signal as well as the intrinsic difficulties to simulated characters with phylogenetic signal.
+Even though molecular characters evolution (and therefore simulation) only depends of a small number of parameters (i.e.
+the base frequencies and the substitution matrix), simulating molecular matrix with a strong evolutionary signal is still complicated when generating unrealistically small matrices.
+For morphological characters, the underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
+
+Also morphological characters studies involve many potential statistical pitfalls (e.g.
+independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy.
+%REF probably April as something about it
+\begin{enumerate}
+\item{First, morphological data can display a different signal than molecular data, especially in small matrices. %REF
+This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
+However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013}.}
+%The main difference between molecular and morphological characters is that, because of historical and practical reasons, a morphological character matrix do no contains autapomorphies and invariant characters which are usually more common than synampomorphies in molecular character matrices (cite).
+\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013}.
+However, following our simulation assumption of random character distributions, if they accumulate through time in the same way as the majority of the molecular characters then, homoplasic characters are expected to appear randomly through time \citep{davalosintegrating2014}.
+Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
+After a reaching a critical amount of morphological characters, adding new ones increases homoplasy \citep{wagner2000}.}
+\end{enumerate}
+Our simulation parameters both decrease the phylogenetic signal (difficulty to simulate morphological characters and size of the matrix) as well as it increases it (reduction of homoplasy in the morphological part of the matrix).
+Therefore, these drawbacks seems to have only a minor impact on the main results of this study: the incapacity of recovering any topology in Bayesian inference.
 
 \subsection{Comparing topologies}
-Hard to compare Bayesian tree posterior distributions topologies but gives a good idea.
-
-\subsection{ML vs Bayesian}
-Missing data is not a problem as long as random. However, in TEM, n.fossils*n.molecular data is missing and not random.
-Non optimal criterion approach (i.e. Bayesian) worse because of the number of considered "less"-likely trees (flat maximum likelihood landscapes).
-However, ML still good performances and can be used to fix topology.
+Comparing topologies is a crucial question in phylogenetics but has always been hard to normalise because of the vast amount of different metrics used as proxies for different aspects of tree similarity/dissimilarity \citep{agapowpower2002}.
+Because our global framework is to study how to include efficiently fossils into phylogenies, we chose metrics reflecting the most interesting aspect of this global question: where do individual taxa (i.e.
+the fossils) branch in the tree.
+The Robinson-Fould \citep{RF1981} and the Triples \citep{critchlowthe1996} metrics are more sensible to taxa and clade placement than other tree comparison metrics \citep[e.g.][Imbalance metric]{kirkpatricksearching1993} and where therefore favoured.
+Also by using the Normalised Tree Similarity index \citep{Bogdanowicz2012} we emphasize the aspect of "good" phylogenetic signal \textit{versus} random phylogenetic signal because it normalised the values of the metric by correcting for the expected value when comparing random trees (NTS = 0).
+
+The Robinson-Fould metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
+When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
+On the other hand, the Triples method is measuring the position of each taxon towards to other reference taxa \citep{critchlowthe1996}.
+It will penalise only trees where taxa get removed furthest from their original clade.
+Regarding our problematic (how does missing data influence topological recovery in trees containing both living and fossil taxa) we are more interested the placement of taxa (i.e.
+where does the fossil branch) than the exact conservation of clades.
+
+%Need a transition!
+
+Although the idea of comparing two single trees is straightforward, it becomes more complex when comparing trees distributions (e.g. Bayesian posterior distributions).
+The introduction of our Random Pairwise Bayesian Tree Comparison methods allows to summarize the comparison of two trees distributions by picking up the most frequent signal in the distribution of the pairwise comparisons (i.e.
+the mode).
+Even though this method is subject to randomness and might artificially increase or decrease the score of the studied metric, simulations showed that when a sufficient amount of random comparisons is performed, this method doesn't seem to be subject to randomness (e.g.
+1000 random pairwise comparisons - see ~\ref{supplementaries}).
+
+\subsection{Maximum Likelihood versus Bayesian}
+The main results of our analysis shows that Bayesian inference fails to recover any Topology in a total evidence method framework (Fig. ~\ref{Fig_Results}).
+This results is surprising regarding the behaviour of Maximum Likelihood inference (Fig.
+~\ref{Fig_Results}) as well as the Bayesian inferences performed on empirical data \citep{ronquista2012,schragocombining2013}.
+However, it is important to note that this effect was not mentioned in the aforementioned empirical tests because both studies used a Bayesian approach with fixed topology \citep{ronquista2012,schragocombining2013}.
+
+In our case, we suspect that this inability of Bayesian methods to recover topology is due to the intrinsic structure of a total evidence matrix (Fig ~\ref{Fig_RemoveData}).
+In fact, as previously studies have shown, missing data doesn't seem to be a major drawback as long as the missing data is randomly distributed ~\citep[e.g.][]{wiensmissing2003,rouresite-specific2011,sansomfossilization2013}.
+However, in total evidence matrices, the majority of the missing data is not randomly distributed but concentrated in the molecular part of the matrix for the fossil taxa (i.e.
+the missing-data sub-matrix, Fig ~\ref{Fig_RemoveData}).
+This leads to high decrease of topological recovery when using non optimal criterion approach (i.e.
+Bayesian inference) because of the high variance in the near-likeliest solutions sampled in the Bayesian posterior distribution.
+In opposition, when applying optimal criterion approach (i.e.
+Maximum Likelihood), it is still possible to sample the likeliest tree.
+%probably need to develop that way more, especially with actually analysing the likelihood landscape.
 
 \subsection{Effect of missing data}
-Parameter blabla seems to be influencing the topology the most.
-But in practice, $M_L$ is the only one realistically solvable. Go to museums!
-
-What about the "best" trees with fairly bad bootstrap support values in ML?
-This is probably due to the low number of characters used in the simulation.
-The number of morphological characters ($10^2$) is in the same order of magnitudes of empirical studies (cite) but decreases unrealistically with the $M_C$ parameter.
-%Extract the mean number of characters per matrices from the 100 random ones
-The number of molecular characters ($10^3$) is at least two orders of magnitude lower than in empirical studies that can reach $10^6$ characters (cite).
-
-Two different matrices from the "True" tree gives two different phylogenetic interpretations (i.e. seqGen and ape performs badly at generating a data matrix from a tree)?
-That seems to not really be a problem since we are interested in recovering topology with missing data, not assessing if phylogenetic matrices generation software performs correctly.
-Also, the "noise" that might be created by this potential "bad" phylogenetic matrices generation software is probably diminished by the number of replicates (i.e. 50 chains).
-
-Different phylogenetic signal between the morphological and the molecular matrix?
-That is not really a problem since the phenotypic evolution is not necessary correlated to the genotypic one.
-Also it is important to note that both information are not exactly the same.
-The main difference between molecular and morphological characters is that, because of historical and practical reasons, a morphological character matrix do no contains autapomorphies and invariant characters which are usually more common than synampomorphies in molecular character matrices (cite).
-
-What about homoplasy in the morphological matrix?
-Here, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly
-If one states that morphological characters accumulate through time in the same way as the majority of the molecular characters (neutral theory, cite) then, homoplasic characters are expected to appear randomly through time.
-Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
-After a certain amount of morphological characters, adding new ones adds homoplasy \citep{wagner2000}.
-
-Difference between Robinson-Fould and Triples?
-Robison-Fould is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies.
-When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
-On the other hand, the Triples method is measuring the position of each taxon towards to other reference taxa.
-It will penalise only trees where taxa get removed furthest from their original clade.
-Looking at our problematic (how does missing data influence topological recovery in trees containing both living and fossil taxa) we are more interested the placement of taxa (i.e.
-where does the fossil branch) than the exact conservation of clades.
-
-What's the crucial factor?
-Regarding the fact that we are more interested in taxa placement than in clade conservation (Triples vs.
-Robinson-Fould), the parameter that affects the more the decrease in topological recovery if the number of missing living taxa in the morphological part of the matrix ($M_L$).
+The three parameters we selected in this study account for three potential pitfalls in collecting the data for total evidence analysis:
+$M_L$ represents the living taxa for which there is no available morphological data,
+$M_F$ represents the quality of the fossil record,
+and $M_C$ represents the general coding effort and the overall of morphological data available.
+Ideally, the lowest possible amount of missing data is wanted in any phylogenetic analysis leading to a data collection part prior to the phylogenetic analysis.
+Each of the three parameters can be improved in a different way prior to the analysis:
+for $M_L$, one should put more effort in using natural museums history collections for coding the missing morphological data for living taxa if possible %go to the museum!;
+for the $M_F$ parameter, the amount of missing data unfortunately depends on the quality of the fossil record and can not be actively improved and depends on exceptional discoveries \citep[e.g.][]{nithe2013};
+finally for the $M_C$ parameter, improvement can be done by vast collaborative projects in order to gather as much characters as possible \citep[e.g.][]{O'Leary08022013}.
+
+Our results shows that the $M_F$ parameter have less influence on recovering the good tree topology in Maximum Likelihood framework which is fortunate since it is the parameter that is the more difficult to fix for practical reasons.
+Regarding the fact that we are more interested in taxa placement than in clade conservation (Triples \textit{versus} Robinson-Fould), the parameter that affects the more the decrease in topological recovery is the number of missing living taxa in the morphological part of the matrix ($M_L$) and the overall number of morphological characters $M_C$.
 This makes sense since the living taxa are bearing both the information to build the tree backbone (the molecular data) and the information used for branching the fossil on this backbone (the morphological information).
-Therefore, we advocate that an effort should be made in collecting all the available morphological information for living taxa in order to properly run Total Evidence analysis.
-%yep, we need data from living taxa.
+Therefore, we advocate the importance of coding morphological characters for the most living taxa possible and with the most characters possible, potentially by using collaborative projects portals such as morphobank \citep{morphobank}.
 
@@
 A Bayesian approach fails to recover accurate topology in a total evidence approach framework, whatever the amount of missing data.
-We think this failure is due to the intrinsic structure of the Total Evidence matrices that includes a vast amount of not randomly distributed missing data (i.e.
+We think this failure is due to the intrinsic structure of the total evidence matrices that includes a vast amount of non randomly distributed missing data (i.e.
 the molecular part of the matrix for the fossil taxa).
-This missing data is equal to the number of fossil taxa $\times$ the number of molecular characters leading to a vast amount of likely trees to be sampled in the Bayesian mcmc tree building.
-However, one can use a Maximum Likelihood approach to obtain the likeliest topology and to use this one as a fixed.
-If so, an effort should be made on collecting morphological data from living taxa in order to efficiently improve the quality of the trees.
+This missing data is equal to the number of fossil taxa $\times$ the number of molecular characters leading to a vast amount of trees to be sampled in Bayesian posterior distribution.
+However, one can use a Maximum Likelihood approach to fix the likeliest topology.
+If so, an effort should be made prior to the phylogenetic analysis on collecting as much morphological data as their is available from living taxa in order to efficiently improve the quality of the trees.
 
@@
 All codes are available at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions$
+The tree comparison results analysis can be repeated for more details at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Analysis$ %fully running!
+Try it at home (5-10 minutes of data loading)!
 

### 2014-06-13 Manuscript/TEM_draft.tex
Editted draft introduction

@@
 %packages
+% NC: Just to flag here how I'll comment on LaTeX files. I'll use % and then NC. Of course in git diff it'll all show up as an addition.
+% NC: You could add all the packages etc to a template .sty file to save writing it all up here. Rich's ms.sty that I provided on DropBox is a good example
 \usepackage{fixltx2e}
@@
 
+% NC: Added by NC
+\usepackage{enumerate}
+\usepackage[osf]{mathpazo} % palatino is a nicer font! 
+\pagenumbering{arabic}	% Page numbers
+
 %Pagination style and stuff
@@
 
-\pagestyle{empty}
+%\pagestyle{empty}
 
@@
 
-%---------------------------------------------
-%
-%       START
-%
-%---------------------------------------------
+% NC: You should delete this extraneous info from the Syst Biol template to tidy everything up a bit
 
@@
 %For a colour version of the draft to a search>replace '-BW'>'' 
-
-%Runing head
+% %\modulolinenumbers[1] 	% Line numbering on every line
+
+%Running head
 \begin{flushright}
@@
 \noindent{\Large \bf Effect of missing data in matrices containing living and fossil taxa on topological accuracy}
-%Alternative punch title: "Bayesian method fails to infer topology using matrices containing both living and fossil taxa"
-%Alternative punch title: "Maximum Likelihood outperforms Bayesian for inferring topology from matrices containing both living and fossil taxa"
+%Alternative punchier title: "Bayesian total evidence methods fail to recover correct topologies from matrices containing both living and fossil taxa"
+%Alternative punchier title: "Maximum Likelihood outperforms Bayesian methods for inferring topology from total evidence matrices containing both living and fossil taxa"
+
+% NC: Not sure about any of these.
+I actually think I prefer your talk title.
+So something like: The effects of missing data on the ability of total evidence approaches to recover correct tree topologies.
+No rush to decide.
+But needs total evidence in there somewhere.
 \bigskip
@@
 \noindent {\small \it 
-$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Republic of Ireland;\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Republic of Ireland;\\
+$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland;\\
+$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland;\\
 $^3$Somewhere else }\\
@@
 \medskip
-\noindent{\bf Corresponding author:} Thomas Guillerme, School of Natural Sciences, Trinity College Dublin, Dublin 2, Republic of Ireland; E-mail: guillert@tcd.ie.\\
+\noindent{\bf Corresponding author:} Thomas Guillerme, School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.\\
 \vspace{1in}
-
-%---------------------------------------------
-%
-%       ABSTRACT
-%
-%---------------------------------------------
-
-\subsubsection{Abstract}
+% NC: It's not called the Republic of, just Ireland!
+
+\newpage
+\begin{abstract}
+% NC: I'm not going to give this much work as it'll change before we submit anyway. Note though that and abstract must NOT just repeat stuff in the intro. Needs rephrasing at the very least.
 Living species represent less than 1\% of all species that have ever lived.
+% NC: I was chatting to Andy Purvis about this the other day and he said lots of palaeo folk hate that stat.
+So let's not start the abstract with it!
 Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology.
@@
 (1) the proportion of living taxa with no morphological data,
-(2) the amount of missing data in the fossil record and
+(2) the amount of missing data in the fossil record, and
 (3) the overall number of morphological characters in the matrix.
@@
 %More results!
+\end{abstract}
 
@@
 
-%---------------------------------------------
-%
-%       INTRODUCTION
-%
-%---------------------------------------------
-
-Living species represent less than 1\% of all species that have ever lived \citep{novacek1992ext,raup1993extinction}.
-However, the majority of macroevolutionary studies focus solely on living species \citep{cooperwhat2009,meredithimpacts2011,healyecology2014}.
+\section{Introduction}
+
+Although most species that have ever lived are now extinct % NC: This avoids giving the dodgy stat of 1%
+ \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species (e.g., \citealp{meredithimpacts2011}).
 %or is citing of your own papers in the too "douchebagy" for the second sentence?
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
-These factors have led to increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-However, to do this we need to be able to place living and fossil taxa into the same phylogenies which still remains complex despite an increasing number of significant efforts \citep{pyrondivergence2011,ronquista2012,schragocombining2013}.
+% NC: Yeah be careful with self citing. Fine if it's a good fit, but it's totally not here!!! A better citation would be a big biodiversity study, e.g. the bird phylogeny paper.
+Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
+This has led to increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
+However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments (e.g.
+\citealp{pyrondivergence2011,ronquista2012,schragocombining2013}).
 %Not a nice sentence
 
-Up to now, three main approaches have been used for combining fossil and living taxa data in phylogenies.
-These approaches differ mainly in whether they treat fossil taxa as tips or as nodes in the phylogeny and on which part of the available data of the fossil is used (i.e.
-the fossil occurrence age only or both the age and the morphology).
-Classical cladistic methods uses matrices containing morphological data from both fossil and living taxa and allows to treat each taxon as a tip in the phylogeny; the relation between the taxa can be inferred using optimal criteria such as the maximum parsimony criterion \citep{simpson1945}.
-This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and doesn't allow to use probabilistic methods to deal with inference uncertainty (but see \citet{spencerefficacy2013}.
-Neontologists, on the other hand, more commonly use probabilistic methods (e.g. Maximum Likelihood or Bayesian) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these analyses and only their occurrence age can be used to time calibrate phylogenies \citep{zuckerkandl1965}.
+Up to now, three main approaches have been used to place both living and fossil taxa % NC: BE consistent. Either living and fossil, or fossil and living.
+into phylogenies.
+These approaches differ mainly in whether they treat fossil taxa as tips or as nodes in the phylogeny, and in which part of the available fossil data is used (i.e.
+the age of the fossil only or both its age and morphology).
+% NC: Not sure about bit in brackets
+Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
+Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \citep{simpson1945}.
+This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty (but see \citealp{spencerefficacy2013}).
+Neontologists, on the other hand, more commonly use probabilistic approaches (e.g.
+Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
+Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates % NC: Need to check the exact phrase needed here.
+Is it range of occurrence?
+Occurrence dates?
+etc.
+ are used to time calibrate phylogenies \citep{zuckerkandl1965}.
 There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013}.
-However none of them is optimal and uses all the available data.
-
-A final approach, known as total evidence approach use matrices containing molecular data from living taxa and morphological data from both living and fossil taxa.
+However neither approach uses all the available data.
+
+A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa.
 %There might be a philosophico-semantic distinction to do here.
@@
 a probabilistic method such as ML where the method states precisely how to do stuff.
-It allows to use probabilistic methods, to treat every taxon as a tip in the phylogeny and to use the occurrence age of the fossil to time calibrate the phylogeny \citep{eernissetaxonomic1993}.
-Here we focus on this total evidence approach because they have been recently successfully developed and applied to empirical data \citep{pyrondivergence2011,ronquista2012,schragocombining2013}.
-%needs a better justification
-Although the total evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data.
-In particular it requires morphological data form both living and fossil taxa, both of which are known to be scarce.
-Therefore total evidence approach is likely to suffer from having lots of missing data which may affect their ability to infer correct topology, branch length and support values \citep{salamin2003}.
+% NC: Tricky. Don't think it's too important though. 
+% NC: I do think Total Evidence needs capital letters though, like ML and Bayesian
+This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{eernissetaxonomic1993}.
+Total Evidence methods have been successfully applied to empirical data \citep{pyrondivergence2011,ronquista2012,schragocombining2013}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
+However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data.
+In particular it requires morphological data from both living and fossil taxa, both of which are known to be scarce.
+% NC: Really?
+I know fossil data is known to be scarce, but living data?
+I think people would be surprised.
+And they are surprised.
+We were surprised!
+Needs a citation either way.
+Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
 % Trevor: Missing data also influences support values that trend to decrease
 
+% NC: You need to take a bit of time to learn about singular and plural grammar in English.
+It's something you always do wrong.
+e.g.
+The Total Evidence method *uses* this and *its* parameters are as follows.
+Versus: Total Evidence methods *use* this and *their* parameters are as follows.
+Sive and I won't be around to correct your English forever.
+
 The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013}.
-Missing molecular data has been seen by some authors as an issue because it can decrease the phylogenetic signal in some parts of the tree, especially when using large matrices \citep{lemmonthe2009}.
-However other authors do not see missing molecular data as a major issue because the phylogenetic signal, rather than by reducing the amount of missing data, is more likely to increase by having:
-(i) at least a "modest" number of highly covered genes (i.e. approximatively half of the genes - \citet{rouresite-specific2011});
-(ii) a higher number of taxa (especially slowly evolving taxa or taxa close to the outgroup - \citet{rouresite-specific2011}) and
-(iii) by choosing more adequate models of sequence evolution rather than by reducing the amount of missing data \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
-
-Similarly, depending on the study, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study \citep{wiensmissing2003,sansomfossilization2013}.
-Because soft-tissues characters are rarely preserved in the fossil record, missing data is mainly found in soft tissues characters, and is therefore not randomly distributed, which can lead to biased placement of fossil taxa in phylogenies \citep{sansomfossilization2013}.
-However, the phylogenetic signal is not related to the level of missing data \textit{per se} but to the number of informative characters per taxa, therefore missing data is less an issue than the number of shared informative characters \citep{wiensmissing2003}.
-Although missing data as been shown to be no major problem separately in molecular and morphological matrices \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011} it may become an issue when combined in a total evidence type matrices.
-Because the nature of total evidence matrices, combining both molecular and morphological information for both living and fossil taxa, a big part of the matrix contains only missing data - the molecular data for fossil taxa.
-Until now, no attempt has been made to study the impact of this issue on phylogenetic inference and especially on topological recovery.
-
-Here we assess the effect of missing data on tree topology inferred from total evidence matrices.
-The molecular part for living taxa of a total evidence matrix acts like a "classical" molecular matrix \citep{ronquista2012}.
-The effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}, therefore, we only focus on the missing data issue in the morphological part of the matrix.
-Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix:
-(1) the proportion of living taxa with no morphological data,
-(2) the amount of missing data in the fossil taxa (i.e. the preservation quality of the fossil record) and
-(3) the overall number of morphological characters for both living and fossil taxa in the matrix.
-We assess how removing data from a total evidence type matrix by changing the values of these three parameters affects the topology of the inferred tree using Maximum Likelihood and Bayesian method.
+Missing molecular data has been seen by some authors as an issue because it can decrease phylogenetic signal (i.e. % define phylogenetic signal here) 
+in some parts of the tree, especially when using large matrices \citep{lemmonthe2009}. However, this may not be a major issue because phylogenetic signal is easily increased by:
+(i) including a "modest" number of highly-covered % what do you mean by highly-covered?
+genes (i.e. approximatively half of the genes; \citealt{rouresite-specific2011});
+(ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and
+(iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
+Similarly, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study in question \citep{wiensmissing2003,sansomfossilization2013}.
+Because soft-tissue characters are rarely preserved in the fossil record, missing data is mainly found in these characters, and is therefore not randomly distributed which can lead to biased placement of fossil taxa in phylogenies \citep{sansomfossilization2013}.
+However, the phylogenetic signal is not related to the amount of missing data \textit{per se} but to the number of informative characters for each taxon, therefore missing data is less of an issue than the number of shared informative characters \citep{wiensmissing2003}.
+% NC: I'm a little dubious of what this paragraph is trying to get at in places. We should discuss at some point.
+
+Although missing data does not appear be a major problem in molecular and morphological matrices separately \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil species.
+This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data.
+Until now, no attempt has been made to study the impact of this issue on phylogenetic inference from Total Evidence methods.
+
+% NC: Somewhere here you need to flag that you're only looking at topologies and briefly explain why.
+
+Here we use simulations to assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
+The molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}.
+The effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}, therefore, we focus only on missing data in the morphological part of the matrix.
+We investigate three major parameters that directly affect the completeness of the morphological part of the matrix:
+\begin{enumerate}
+\item the proportion of living taxa with no morphological data;
+\item the amount of missing data in the fossil taxa; and
+\item the overall number of morphological characters for both living and fossil taxa in the matrix.
+\end{enumerate}
+We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
+We chose these parameters because they reflect real-life biases in data availability.
+The advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
+Missing data in fossil taxa is very common due to preservation biases (citation?), and
+% OK I'm not sure what the "real-life" reason for three is.
 %This paragraph is crap
-
-We found that whatever the amount of missing data, total evidence matrices dramatically reduce the performance of Bayesian inference.
-However, when using a Maximum Likelihood approach, missing data decreases gradually the ability of recovering the right tree topology.
-We propose that this drastic difference between Bayesian and Maximum Likelihood methods is due to a flattening of the likelihood landscape because of the intrinsic amount of missing data in a total evidence matrix (i.e.
-the missing molecular data for the fossil taxa).
-This problem leads to an inability of Bayesian chains to pick up the sub and near optimal topologies leading to high variance in the posterior tree distribution and really low support in the consensus tree.
-These results should encourage authors in not using Bayesian inference to explore the topological landscape of their phylogeny.
-Instead, we suggest that authors should first find the Maximum Likelihood tree and use it's topology as a fix prior in their Bayesian inference.
-We also propose a new method for comparing Bayesian posterior tree distribution's topology available in a wonderful R package. %Or am I just dreaming a bit too much here?
-
-%---------------------------------------------
-%
-%       METHODS
-%
-%---------------------------------------------
-
+% NC: I think it's fine!
+
+We find that when using a Maximum Likelihood approach, as missing data increases, the likelihood of recovering the correct tree topology decreases.
+However, even with no missing data, Total Evidence matrices dramatically reduce the performance of Bayesian methods for inferring tree topology.
+We propose that this drastic difference between Bayesian and Maximum Likelihood methods is due to a flattening of the likelihood landscape caused by the unavoidable amount of missing molecular data for fossil taxa in a Total Evidence matrix.
+We make suggestions for how best to deal with this issue when inferring phylogenies from Total Evidence matrices.
+
+% NC: You don't want to tell the whole story here. Just a results summary and a quick line of interpretation
+
+%This problem leads to an inability of Bayesian chains to pick up the sub and near optimal topologies leading to high variance in the posterior tree distribution and really low support in the consensus tree.
+%These results should encourage authors in not using Bayesian inference to explore the topological landscape of their phylogeny.
+%Instead, we suggest that authors should first find the Maximum Likelihood tree and use it's topology as a fix prior in their Bayesian inference.
+%We also propose a new method for comparing Bayesian posterior tree distribution's topology available in a wonderful R package. %Or am I just dreaming a bit too much here?
+
+% NC: Dreaming a bit much I think.
+There's a discussion to be had here about whether it's really a *new* method and whether you really want to make an R package per se.
+We definitely want to share the code
+
+%-----------
+% NC: It might look like I've done a lot to this but actually I think it's nearly there.
+A good thing to take from this is that writing a paper is *hard*.
+It gets easier, but it's still a deeply involved process with lots of writing and re-writing.
+Introductions are particularly difficult because if you don't set the scene, it's hard to convince people to get on board for the rest of the paper.
+This is why I said there was no way this was being submitted by the end of June!
+Still a long way to go I'm afraid in terms of time and effort.
+But that doesn't mean you're not doing well, *every* project is like this.
+2 years is about average from conception to sending out a manuscript, and you only started work on this last spring.
+%------------
+
+\newpage
 \section{Methods} %no material needed?
-To explore the effect of missing data in total evidence matrices on tree topology we used the following protocol (note that we explain each step in detail below this general outline -Fig.
-~\ref{Fig_Outline}):
+% NC: Depends on the journal. See what Syst Biol asks for as section titles.
+To explore how missing data in Total Evidence matrices influences tree topology we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
+%NC: does Syst Biol use Fig.
+or Figure?
 \begin{enumerate}
-\item{Generating the matrix:} \label{step:generate_matrix} \\
-We randomly generated birth death tree (hereafter called the "true" tree, Table ~\ref{Tab_glossary}) to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix, Table ~\ref{Tab_glossary}).
-\item{Removing data:} \label{step:remove_data} \\
-We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (1) the proportion of missing living taxa ($M_{L}$), (2) the proportion of missing data in the fossil taxa ($M_{F}$) and (3) the proportion of missing morphological characters ($M_{C}$) (the resulting matrices are called hereafter the "missing-data" matrix, Table ~\ref{Tab_glossary}).
-\item{Building phylogenies:} \label{step:build_phylo} \\
-We inferred Bayesian phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in respectively a tree generate from a matrix containing no missing data in the morphological part (hereafter called the "best" tree, Table ~\ref{Tab_glossary}) and trees inferred from matrices with missing data in their morphological part (hereafter called the "missing-data" trees, Table ~\ref{Tab_glossary}).
-\item{Comparing topologies:} \label{step:compare_topo} \\
-We then compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on the topologies of the phylogenies we estimated.
+\item{Generating the matrix} \label{step:generate_matrix} \\
+We randomly generated a birth-death tree (hereafter called the "true" tree; Table ~\ref{Tab_glossary}) and used it to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix; Table ~\ref{Tab_glossary}).
+\item{Removing data} \label{step:remove_data} \\
+We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of missing living taxa ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$) % NC: make sure to use the same descriptions here and in the introduction
+(the resulting matrices are called hereafter "missing-data" matrices; Table ~\ref{Tab_glossary}).
+\item{Building phylogenies} \label{step:build_phylo} \\
+We built phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree; Table ~\ref{Tab_glossary}) and multiple trees inferred from matrices with missing morphological data (hereafter called the "missing-data" trees; Table ~\ref{Tab_glossary}).
+Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
+\item{Comparing topologies} \label{step:compare_topo} \\
+We compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies.
 \end{enumerate}

### 2014-06-15 Manuscript/TEM_draft.tex
Editted draft methods

@@
 % NC: Depends on the journal. See what Syst Biol asks for as section titles.
-To explore how missing data in Total Evidence matrices influences tree topology we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
+To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
 %NC: does Syst Biol use Fig.
@@
 \end{enumerate}
-To measure the effect of missing data distribution, we repeated steps ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} with the exact same fixed parameters 51 (3$\times$17) times.
-%that's because I had 3 cluster accounts available...
-If it's a too odd number, no worries, I'll just remove the last chain.
-A list of all the terms used in this paper is available in table ~\ref{Tab_glossary}.
+We repeated steps 1-4 50 times. %that's because I had 3 cluster accounts available... If it's a too odd number, no worries, I'll just remove the last chain.
+% NC: Yeah 51 is odd. Let's just use 50.
+A list of all the terms used in this paper is available in Table ~\ref{Tab_glossary}.
 
@@
 \caption{Protocol outline.
-(1) We generated a random tree (the "true" tree) to infer a matrix with no missing data (the "complete" matrix).
-(2) We removed data from the morphological partition of the "complete" matrix resulting in 125 "missing-data" matrices.
-(3) We infer phylogenetic trees from each matrix in both ML and Bayesian framework.
-(4) We then compared the "missing-data" trees to the "best" tree.
-We repeated step ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} 51 (3$\times $17) times.}
+(1) We randomly generated a birth-death tree (the "true" tree) and used it to infer a matrix with no missing data (the "complete" matrix).
+(2) We removed data from the morphological part of the "complete" matrix resulting in 124 "missing-data" matrices.
+% NC: I thought we had 124 missing data matrices... I just worked this out - one of the "missing data" matrices is 0%,0%,0% so = complete matrix.
+(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
+(4) We compared the "missing-data" trees to the "best" tree.
+We repeated steps 1-4 50 times.}
 \label{Fig_Outline}
 \end{figure}
+
+% NC: One tiny edit to the figure.
+I think the ML then Bayesian story is nice, so maybe change that part of the figure so it's ML then Bayesian.
+Also let's show this figure to a couple of people to see if they get it.
+
+% NC: Not sure we need the glossary, but keep it for the thesis
 
@@
 
-%---------------------------------------------
-%       Generating the matrix
-%---------------------------------------------
+% NC: Most of the below was fine but I've tried to condense it a bit as it's currently long even for Syst Biol. The key is to explain everything, but with minimal repetition.
 
@@
 First we randomly generated a "true" tree of 50 taxa in R v3.0.2 \citep{R302} using the package diversitree v0.9-6 \citep{fitzjohndiversitree2012}.
-We generated the tree using a Birth Death process by sampling the values of the speciation events ($\lambda$) and extinction events ($\mu$) from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
-We implemented a rejection sampling algorithm to select only random trees with 25 living and 25 fossil taxa. %Justification needed?
-We then added a taxa to the resulting Birth-Death tree as the outgroup of the tree.
-The mean branch length of the tree was used to separate the outgroup from the rest of the taxa and the branch length leading to the outgroup was set as the sum of the mean branch length and the longest root-to-tip length of the tree.
-
-Next, we created a molecular and a morphological matrix from the "true" tree.
-The molecular matrix was inferred from the "true" tree using the package phyclust v0.1-14 \citep{chen2011}.
-The matrix was made of 1000 characters sites for 51 taxa and generated using the seqgen algorithm \citep{ranbaut1997seqgen}.
-We used the HKY model \citep{HKY85} with a random base frequencies and with the transition/transversion rate of 2 \citep{douadycomparison2003} as parameters for generating the matrix.
+We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
+We implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa to ensure that we had enough taxa of each type for our missing data simulations to work.
+We then added an outgroup to the tree, using the mean branch length of the tree to separate the outgroup from the rest of the taxa, and with the branch length leading to the outgroup set as the sum of the mean branch length and the longest root-to-tip length of the tree.
+
+Next, we generated a molecular and a morphological matrix from the "true" tree.
+The molecular matrix was inferred from the "true" tree using the R package phyclust v0.1-14 \citep{chen2011}.
+The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies and transition/transversion rate of 2 \citep{douadycomparison2003}.
 The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
 We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
-These parameters were selected to generate data with no special assumption about how the characters evolved as well as to reduce the computational time required if these parameters were estimated rather than defined (total computational time $>$ 65 CPU years).
-
-We inferred the morphological matrix using the ape package v3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
-We assigned the number of character states (either 2 or 3) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of characters states extracted from 100 published empirical morphological matrices (See \hyperref[supplementaries]{supplementaries}).
-We then ran an independent discrete character simulation for each character using the "true" tree branch length and topology with the randomly selected number of states (2 or 3) and assuming an equal rate of change (i.e.
+We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined (even with the parameters defined, total computational time was over 65 CPU years).
+All the molecular information for fossil taxa was replaced by missing data ("?").
+
+We inferred the morphological matrix using the R package ape v3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
+We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (See \hyperref[supplementaries]{supplementaries}).
+We then ran an independent discrete character simulation for each character using the "true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
 evolutionary rate) from one character state to an other \citep{Pagel22011994}.
-This method allows us to have only two parameters per character: the number of states and the evolutionary rate.
+This method allows us to have only two parameters for each character: the number of states and the evolutionary rate.
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used a low evolution rate parameter (i.e.
-$\alpha$) in order to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
-
-All the molecular information for fossil taxa was replaced by missing data ("?").
+We used low evolutionary rate parameters (i.e.
+$\alpha$) to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
+
 Finally, we combined the morphological and molecular matrices obtained from the "true" tree.
@@
 \subsection{Removing data}
-Once we obtained the "complete" matrix we modified it to get a set of matrices with missing data. %Ugly sentence
-We randomly replaced data with "?" in the morphological part of the matrices according to the following parameters (Fig. ~\ref{Fig_RemoveData}):
+We modified the "complete" matrix to get matrices with missing data by randomly replacing data with "?" in the morphological part of the matrices according to the following parameters (Fig.
+~\ref{Fig_RemoveData}):
 
 \begin{enumerate}
+% NC: I'm not going to check but make sure the brief explanations of the parameters here are the same as in the intro and outline above.
+Also look at my justifications in the intro too.
+I think this may need a bit more justification as it's the main part of your simulations.
 \item{The proportion of living taxa with no morphological data ($M_{L}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological one.
-Because of the increasing facility to sequence DNA for living taxa, the number of living taxa with molecular data is highly superior the the number of taxa with molecular and morphological data.
+This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological part.
+This reflects the fact that because of the increasing availability of DNA sequences for living taxa, detailed morphological data is scarce.
 \item{The proportion of missing morphological data across all fossil taxa ($M_{F}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
@@
 
+% NC: I'm not sure this figure is helpful really. Let's test it on some people.
 \begin{figure}
@@
 
-In practice, each parameter represent a different way of removing data in the morphological part of the matrix: $M_{L}$ removes a proportion of rows from the living taxa; $M_{F}$ removes a proportion of cells from the fossil taxa; and $M_{C}$ removes a proportion of columns across both living and fossil taxa (see Fig.
-~\ref{Fig_RemoveData}).
-Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a proportion of the living taxa is removed (i.e.
-removing rows), as for $M_{F}$, a proportion of data is removed across the whole of the morphological matrix for fossil taxa (i.e.
-removing cells).
+In practice, each parameter represents a different way of removing data from the matrix: $M_{L}$ removes rows from the living taxa; $M_{F}$ removes cells from the fossil taxa; and $M_{C}$ removes columns across both living and fossil taxa (see Fig.
+~\ref{Fig_RemoveData}).Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a percentage of living taxa is removed, but for $M_{F}$, a percentage of the data is removed at random from across the whole of the morphological matrix for fossil taxa.
 
 We tested all parameters combinations resulting in 125 ($5^3$) matrices.
-Because some parameter combinations introduce a lot of missing (e.g. $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
-When this occurred we repeated the random deletion of characters until every taxa had at least 5\% data across the whole matrix.
+Note that one of these combinations has no missing data so is equivalent to the "complete" matrix, thus we have 124 "missing-data" matrices.
+Because some parameter combinations introduce a lot of missing data (e.g.
+$M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
+When this occurred we repeated the random deletion of characters until every taxon had at least 5\% data across the whole matrix.
 
@@
 \subsection{Building phylogenies}
-From the resulting matrices we generated two types of trees, the "best" tree that is inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
-The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
-The "best" tree, on the other hand, is the best tree we can build using the state-of-the-art phylogenetic methods.
-In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g. \citet{rozen2005}).
+From the resulting matrices we generated two types of trees, the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 124 matrices with various amounts of missing data.The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
+The "best" tree, on the other hand, is the best tree we can build using state-of-the-art phylogenetic methods.
+In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g.
+\citet{rozen2005}).
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
@@
 The "best" tree and the "missing-data" trees were inferred using RAxML v8.0.20 \citep{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model (\citet{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citet{Stamatakis21012014}) as a generalisation of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
-The GTR model can be seen as a generalisation of the HKY model (the 2 parameters from the HKY model are implicitly included in the 6 from GTR model - \citet{stamatakisa2008}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} which is a generalisation of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014}).
-We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. %The explanations of the algorithm part for the BS was proposed by Trevor
-The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists in pruning one sub-tree from the tree and subsequently reinserting it to al neighbouring branches \citep{stamatakisa2008}.
-Sub-tree Pruning and Reinserting methods (SPR) have been demonstrated as being better than others (e.g. Nearest Neighbouring Interchange - NNI) in recovering good bootstrap values \citep{salamin2003}.
+For the molecular data, we used the GTR + $\Gamma_4$ model (\citet{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citet{Stamatakis21012014}) as a generalization of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
+The GTR model can be seen as a generalization of the HKY model (the two parameters from the HKY model are implicitly included in the six from GTR model - \citet{stamatakisa2008}).
+For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} which is a generalization of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014}).
+We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support.
+%The explanations of the algorithm part for the BS was proposed by Trevor
+The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists of pruning one sub-tree from the tree and subsequently reinserting it to all neighboring branches \citep{stamatakisa2008}.
+Sub-tree Pruning and Reinserting methods (SPR) have been demonstrated to be better than other methods (e.g.
+Nearest Neighboring Interchange - NNI) for recovering good bootstrap values \citep{salamin2003}.
 
@@
 For each run, we removed 25\% of the iterations as burn-in.
-We used the following prior for each tree (see \hyperref[supplementaries]{supplementaries}):
+We used the following priors for each tree (see \hyperref[supplementaries]{supplementaries}):
 \begin{enumerate}
-\item
-the "true" tree’s topology as a starting tree (with a starting value for each branch length of 1),
-\item
-an exponential prior on the shape of the gamma distribution of $\alpha$=0.5 for both partitions
-\item
-and a transition/transversion ratio prior of 2 sampled from a strong beta distribution ($\beta$(80,40)).
+\item the "true" tree’s topology as a starting tree (with a starting value for each branch length of 1),
+\item an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions, and
+\item a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)).
 \end{enumerate}
 
-We used these prior to speed up the Bayesian process.
-These prior biased the way the Bayesian process calculated the branch length by giving non-random starting points and boundaries for the parameters estimation process, however, in this study, we focused on the effect of missing data on the topology and not on the branch length.
-Even using these prior, it took 65 CPU years to build 51 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
+We used these prior to speed up the Bayesian estimation process.
+These priors biased the way the Bayesian process calculated branch lengths by giving non-random starting points and boundaries for parameter estimation, however, here we are focusing on the effect of missing data on tree topology and not branch lengths.
+Even using these priors, it took 65 CPU years to build 50 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
+% NC: OK so the 65 CPU is for building trees not matrices? Or is it combined time? Or does both stage take 65 CPU years? Be clearer in the earlier sections
 
@@
 
+% NC: I think this section is suffering a bit from "here is everything I know about tree comparison metrics" rather than explaining what is important for your analysis.
+You may be able to cut some more stuff out.
+
 \subsection{Comparing topologies}
-We compared the topology of the "missing-data" trees inferred from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
-Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees inferred from total evidence type matrices.
-To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triples \citep{dobson1975triplets} and the Robinson-Fould \citep{RF1981} distance.
-We normalised the two metrics using the Normalised Tree Similarity index \citep{Bogdanowicz2012} to generalise our results for any \textit{n} number of taxa.
-The three metrics are detailed below.
-
-\subsubsection{Triples distance ($T_{x,y}$) \citep{dobson1975triplets}}
-This metric measures the number of different sub-trees of three taxa between two given trees.
+We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
+We used the Triples \citep{dobson1975triplets} metric to assess the number of conserved taxa across trees, and the Robinson-Fould \citep{RF1981} distance to identify conserved clade positions.
+We then used Normalized Tree Similarity index \citep{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
+These metrics are described in detail below.
+% NC: I moved this around a bit to make it clearer what metric fits where. Remember we need American spellings.
+% Also I thought it was triplets distance? Good idea to check...
+
+\subsubsection{Triples distance}
+% NC: I've tried to remove the word displaced here, as displaced suggests that you have one correct placement and then are comparing it to other trees.
+But the general meaning is that the placements differ between trees.
+Triples distance ($T_{x,y}$; \citealp{dobson1975triplets}) measures the number of sub-trees made up of three taxa (triplets) that differ between two given trees.
 Each triplet can be written as $I_{ijk}$=(\textit{ijk}).
-Where $I_{ijk}$ is equal to 0 if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to 1.
-For any rooted binary tree there are only three possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
-If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k});.
-One can calculate $S_n$, the triplet distance between two trees as:
+Where $I_{ijk}$ is equal to zero if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to one.
+For any rooted binary tree there are only three possible combinations for each triplet: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
+If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k}).
+We can calculate the triplet distance between two trees, $S_n$, as:
 \begin{equation}
@@
 \end{equation}
-And where \textit{n} is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
-If $S_n$=0, the trees are the same (i.e. no taxa as been displaced).
-When $S_n$ = $\binom{n}{4}$, the trees are the most different possible (i.e. every taxa as been displaced).
-This metric therefore illustrates the amount of displayed taxa.
-It is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g.
+And where \textit{n} is the total number of taxa in both trees (modified from \citet{critchlowthe1996}).
+If $S_n$ = 0, the trees are identical; when $S_n$ = $\binom{n}{4}$, the trees are as different as possible (i.e.
+every taxon has a different placement in the two trees).Triples distance is less sensitive to the placement of individual taxa and to taxa with highly uncertain placement (e.g.
 fossil taxa) than the Robinson-Foulds metric \citep{critchlowthe1996,johnson1998,wiensmissing2003}.
-Therefore we used this metric as a proxy to estimate the robustness of the tree to flying taxa (see \hyperref[supplementaries]{supplementaries}).
-
-\subsubsection{Robinson-Fould distance (\textit{RF}) \citep{RF1981}}
-This metric measures the number of shared clades among two trees and therefore illustrates the number of exactly conserved groups among the trees.
-The Robinson-Fould distance (also called path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
+% NC: Not sure you need this bit
+% Therefore we used this metric as a proxy to estimate the robustness of the tree to flying taxa (see \hyperref[supplementaries]{supplementaries}).
+
+\subsubsection{Robinson-Foulds distance} % NC: Is it Robinson Fould or Robinson Fould*s*?
+Robinson-Foulds distance (\textit{RF}; \citealp{RF1981}), or "path difference"
+% NC: path difference or *distance*?
+, measures the number of shared clades across two trees.
+The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} and can be expressed as following:
 \begin{equation}
@@
 \end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees.
-The minimal value of \textit{C} is 1 if the two trees have the same \textit{n} taxa; the maximal value in \textit{C=n-2}.
-This metric is more sensitive to taxa displacement than the Triples metric (i.e.
-if one taxa gets out of a clade, then the clades are no longer considered as similar - \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
-Therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades (see \hyperref[supplementaries]{supplementaries}).
-
-\subsubsection{Normalised Tree Similarity \textit{NTS} \citep{Bogdanowicz2012}}
-For any tree with \textit{n} taxa compared using a tree distance metric \textit{m}, $NTS_m$ represents the similarity score between the two trees given the expected distance between two random Yule trees with \textit{n} taxa.
-Let $\bar{d}_{m,n}$\textit{(rand)} be the average distance between two random Yule trees with \textit{n} taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} containing each \textit{n} taxa, then:
+Where $C_{x,y}$ is the number of clades in common in the two trees. \textit{C} % NC: Why italics? Why not $C$?
+is one if the two trees have the same \textit{n} taxa; the maximal value is \textit{C = n-2}. % NC: again why italics for these bits? 
+%NC: I don't get this. Explain more clearly. also is it important?
+This metric is more sensitive to taxon displacement than Triples distance (i.e.
+if one taxon moves out of a clade, then the clades are no longer considered similar; \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
+Therefore a low value indicates that clades are similar across the two trees, whereas a high value indicates that there are few common clades (see \hyperref[supplementaries]{supplementaries}).
+%NC: See supp for what purpose exactly?
+
+\subsubsection{Normalized Tree Similarity}
+For any tree with \textit{n} taxa compared using a tree distance metric \textit{m}, Normalized Tree Similarity, $NTS_m$ (\citealp{Bogdanowicz2012}), represents the similarity score for the two trees given the expected distance between two random Yule trees with \textit{n} taxa.
+If $\bar{d}_{m,n}$\textit{(rand)} is the average distance between two random Yule trees with \textit{n} taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} containing each \textit{n} taxa, then:
 \begin{equation}
@@
 \end{equation}
-\textit{NTS} ranges from 1 to -$\infty$.
-For any \textit{m},\textit{n} when \textit{NTS}=1, the trees are the same;
-when \textit{NTS}=0 the trees are not more different than expected by chance;
-when \textit{NTS}$<$0, the trees are more different than expected by chance (see \hyperref[supplementaries]{supplementaries}).
-
-$\newline$
-
-We compared the "missing-data" tree to the "Best" tree for each chain.
-For the Maximum Likelihood trees we performed pairwise comparisons between the "Best" tree and the "missing-data" tree (see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triples metric.
-We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
-For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparisons for the same metric and the same number of taxa \textit{n}=51 (see \hyperref[supplementaries]{supplementaries}).
-We ran the comparison for every "missing-data" tree in each chain resulting in 51 comparisons for every "missing-data" trees. 
-We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package (v3.1 – \citet{hdrcde}).
-
-Bayesian tree inference allows to account for the statistical uncertainty of a phylogenetic tree by not using an optimal criterion (c.f.
+\textit{NTS} ranges from one to -$\infty$.
+For any $m,n$, when \textit{NTS} = 1, the trees are identical, when \textit{NTS} = 0 the trees are no more different than expected by chance, and when \textit{NTS} $<$ 0, the trees are more different than expected by chance (see \hyperref[supplementaries]{supplementaries}).\\
+% NC: You seem to be using italics almost at random here. Why???
+% NC: What is "chance" here?
+% NC: Why do we need to see the supp mat?
+
+\subsubsection{Tree comparisons}
+For the Maximum Likelihood trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree (see Table ~\ref{Tab_glossary}) using both the Triples and Robinson-Fould metrics % NC: be consistent in the order.
+Triples then RF.
+with the TreeCmp java script \citep{Bogdanowicz2012}.
+For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and \textit{n} = 51 taxa (see \hyperref[supplementaries]{supplementaries}).
+% NC: Why italic?
+What does the supp mat show?
+We compared each "missing-data" tree  with the "best" tree for each of our 50 simulation runs
+% NC: Chains are a MCMC thing. You mean 50 simulations not chains.
+resulting in 50 comparisons for each "missing-data" tree.
+We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}.
+%NC: But this is a Bayesian thing right?
+I think we need to check what you *actually* did here...
+
+%Bayesian tree inference allows to account for the statistical uncertainty of a phylogenetic tree by not using an optimal criterion (c.f.
 Maximum Likelihood) and gives a tree posterior distribution (c.f.
@@
 %Awwwwww yiis!
-To avoid this problem, people traditionally use the a consensus tree build on a majority rule in order to summarise a Bayesian tree posterior distribution.
-This results into a single tree containing both topological and branch length information as well as support information (i.e. posterior probabilities, e.g. \citet{Ronquist2012mrbayes}).
-However, using a Bayesian consensus tree has limitation especially if the resulting consensus tree is not well supported or resolved.
-Because the metrics used in this study are used to measure variation in topology between two trees (i.e.
+
+% NC: Sorry awwww no. This is all irrelevant. Everyone knows what a Bayesian phylogeny is, and all the reader needs to know is how you compared trees. They don't need all this waffle.
+
+%To avoid this problem, people traditionally use the a consensus tree build on a majority rule in order to summarise a Bayesian tree posterior distribution.
+%This results into a single tree containing both topological and branch length information as well as support information (i.e. posterior probabilities, e.g. \citet{Ronquist2012mrbayes}).
+%However, using a Bayesian consensus tree has limitation especially if the resulting consensus tree is not well supported or resolved.
+%Because the metrics used in this study are used to measure variation in topology between two trees (i.e.
 taxa placement or clade position), comparing two Bayesian consensus trees is not optimal in picking topological difference signal.
-For example, if one of the trees is not resolved at all (i.e. a star tree), comparing it to any other tree (even an other star tree) will not give useful results.
-Therefore we used the entire Bayesian trees posterior distribution in order to perform the tree comparisons (i.e.
+%For example, if one of the trees is not resolved at all (i.e. a star tree), comparing it to any other tree (even an other star tree) will not give useful results.
+%Therefore we used the entire Bayesian trees posterior distribution in order to perform the tree comparisons (i.e.
 the "Best" Bayesian tree posterior distribution vs.
@@
 
-\subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
-We compared the Bayesian posterior distribution using a Random Pairwise Bayesian Tree Comparison (RPBTC) method.
-This method consists in comparing a series of randomly selected pairs of trees from two posterior distributions (on from each distribution) and use the mode to summarize the resulting distribution as a proxy of the difference between the two trees in a Bayesian framework.
-\begin{equation}
-RPBTC_{X,Y}=Mo[(]d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]
-\end{equation}
-
-Where \textit{X} and \textit{Y} are two Bayesian tree posterior distributions;
- $d_{m,n}$ is the pairwise difference for any metric \textit{m} and \textit{n} taxa between $X_{i}$ and $Y_{j}$ which are two single trees randomly sampled respectively from \textit{X} and \textit{Y};
- $Mo$ is the mode and
- $Mo[(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]$ is the mode of the pairwise difference repeated \textit{k} times.
-
-We used the mode to summarize the distributions because it is the value that is the most represented in the distribution and reflects what a consensus tree represents (the topology the most represented in the posterior distribution).
+% NC: OK this whole section is far too long.
+I'm not sure how much is really needed.
+Surely all you did was randomly select pairs from two distributions and then take the mode?
+Check that what I've written below is correct.
+
+Comparing Bayesian trees was slightly harder as each analysis produced a posterior distribution of trees to compare, rather than a single tree as in Maximum Likelihood.
+Therefore, for each "missing-data" posterior distribution of trees we randomly selected one tree from the "missing-data" tree posterior distribution and one tree from the "best" tree posterior distribution and compared these trees using Triples distance and Robinson-Fould's distance as above.
+We repeated this 1000 times and then used the modal values for the metrics to summarize the difference between the two tree posterior distributions.
+For each metric, we then normalized the value using the Normalized Tree Similarity as described above.
+We used this protocol to compare each "missing-data" tree posterior distribution  with the "best" tree posterior distribution for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree.
+We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}.
+%NC: But this is a Bayesian thing right?
+I think we need to check what you *actually* did here...
+
+Finally we tested for significant differences among types of "missing-data" trees using ???
+% NC: The whole of the next section can be summarised in just a few lines like this!
+
+%\begin{equation}
+%RPBTC_{X,Y}=Mo[(]d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]
+%\end{equation}
+
+%Where \textit{X} and \textit{Y} are two Bayesian tree posterior distributions;
+ %$d_{m,n}$ is the pairwise difference for any metric \textit{m} and \textit{n} taxa between $X_{i}$ and $Y_{j}$ which are two single trees randomly sampled respectively from \textit{X} and \textit{Y};
+ %$Mo$ is the mode and
+ %$Mo[(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]$ is the mode of the pairwise difference repeated \textit{k} times.
+
+%We used the mode to summarize the distributions because it is the value that is the most represented in the distribution and reflects what a consensus tree represents (the topology the most represented in the posterior distribution).
 Also, the mode is a value present in the distribution (which, depending on the metric, can be a series of discrete values) in contrast to the other distribution summary metrics, such as the mean or the median, which can take values that are not actually present in the distribution.
@@
 
-In this study, we calculated RPBTC for both metrics (Robinson-Fould and Triples distance) for each pairs of Bayesian tree distributions (i.e.
+%In this study, we calculated RPBTC for both metrics (Robinson-Fould and Triples distance) for each pairs of Bayesian tree distributions (i.e.
 the "Best" Bayesian tree posterior distribution vs.
 the "missing-data" Bayesian tree posterior distribution) with 1000 random pairwise comparisons (\textit{k}).
-Because the RPBTC uses random pairwise comparisons, there is a chance of comparing always trees that have the biggest or the smallest difference between two distributions, either deflating or inflating the RPBTC difference value.
-However, using 1000 random pairwise comparisons makes the difference stable (i.e. if repeated independently, no difference is detected - see \hyperref[supplementaries]{supplementaries}).
-
-$\newline$
-
-We calculated the NTS for the "Best" Bayesian tree vs.
+
+%Because the RPBTC uses random pairwise comparisons, there is a chance of comparing always trees that have the biggest or the smallest difference between two distributions, either deflating or inflating the RPBTC difference value.
+However, using 1000 random pairwise comparisons makes the difference stable (i.e.
+if repeated independently, no difference is detected - see \hyperref[supplementaries]{supplementaries}).
+
+%We calculated the NTS for the "Best" Bayesian tree vs.
 "missing-data" Bayesian tree comparison for both Robinson-Fould and Triples metric using the mode of the RPBTC for each chain resulting in a distribution of 51 modes per comparison.
-We then compared the NTS of the Bayesian trees to the NTS of the ML trees for both metrics for each comparison.
-This resulted in a distribution of 51 NTS values for the Bayesian trees and 51 NTS values for the ML trees for each tree comparison for which we calculated the mode and the 50\% and 95\% confidence intervals (see \hyperref[supplementaries]{supplementaries} - Fig.
+We then compared the NTS of the Bayesian trees to the NTS of the ML trees for both metrics for each comparison.This resulted in a distribution of 51 NTS values for the Bayesian trees and 51 NTS values for the ML trees for each tree comparison for which we calculated the mode and the 50\% and 95\% confidence intervals (see \hyperref[supplementaries]{supplementaries} - Fig.
 ~\ref{Fig_Compare}).
 
-\begin{figure} % Modify the figure
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp-BW.pdf}
-\caption{Tree comparison protocol.
-From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we inferred a tree in Baysian or Maximum likelihood framework.
-We then compared this tree with or without fossil/living taxa to the "best" tree.
-In the Bayesian framework, 1000 trees from the posterior distribution where used for each comparison.}
-\label{Fig_Compare}
-\end{figure}
-
-\subsubsection{Statistical difference between tree topologies} %note that in the end, I only have non-parametric tests, the parametric explanation part can go to the supplementaries
-To assess the difference between the tree topologies, whether they were inferred in a Bayesian or a ML framework, we performed a group comparison test for each set of configurations (e.g.
+% NC: I don't think this figure helps at all
+%\begin{figure} % Modify the figure
+%\centering
+%\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp-BW.pdf}
+%\caption{Tree comparison protocol.
+%From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we inferred a tree in Baysian or Maximum likelihood framework.
+We then compared this tree with or without fossil/living taxa to the "best" tree.In the Bayesian framework, 1000 trees from the posterior distribution where used for each comparison.}
+%\label{Fig_Compare}
+%\end{figure}
+
+%\subsubsection{Statistical difference between tree topologies} %note that in the end, I only have non-parametric tests, the parametric explanation part can go to the supplementaries
+%To assess the difference between the tree topologies, whether they were inferred in a Bayesian or a ML framework, we performed a group comparison test for each set of configurations (e.g.
 effect of $M_L$ in Bayesian inference, combined effect of $M_F$ and $M_C$ in ML inference, etc...).
@@
 
-\begin{table}[ht]
-\caption{Statistical tests}
-\centering
-\begin{tabular}{lrll}
-  \hline
-   & test & statistics & code\{package\} \\ [0.5ex]
-  \hline
-  Groups comparisons & ANOVA & parametric & anova\{stats\} \\ [0.5ex]
-                     & Kruskal-Wallis & non-parametric & kruskal.test\{stats\} \\[0.5ex]
-  \hline
-  Pairwise comparisons & Tukey HSD & parametric & TukeyHSD\{stats\} \\ [0.5ex]
-                       & Nemenyi-Damico-Wolfe-Dunn & non-parametric & kruskalmc\{pgirmess\} \\ [0.5ex]
-  \hline
-\end{tabular}
-\label{Stats_test}
-\end{table}
-
-\subsection{Effect of missing molecular characters for fossil taxa} %Awfull name.
-To assess the effect of missing molecular characters for fossil taxa in a Bayesian framework, we split the "complete" matrix in sub-matrices containing no missing data at all (see Fig.
+%\begin{table}[ht]
+%\caption{Statistical tests}
+%\centering
+%\begin{tabular}{lrll}
+%  \hline
+%   & test & statistics & code\{package\} \\ [0.5ex]
+%  \hline
+%  Groups comparisons & ANOVA & parametric & anova\{stats\} \\ [0.5ex]%
+%                     & Kruskal-Wallis & non-parametric & kruskal.test\{%stats\} \\[0.5ex]
+%  \hline
+ % Pairwise comparisons & Tukey HSD & parametric & TukeyHSD\{stats\} \\ [0.5ex]
+  %                     & Nemenyi-Damico-Wolfe-Dunn & non-parametric & kruskalmc\{pgirmess\} \\ [0.5ex]
+ % \hline
+%\end{tabular}
+%\label{Stats_test}
+%\end{table}
+
+\subsection{Effect of missing molecular characters for fossil taxa} %Awful name.
+% NC: Yeah!
+% NC: OK you need to explain WHY we are doing this. What is this analysis designed to test. I think you said this was incomplete so I won't fix it I'll let you have a go first.
+To assess the ability to recover topology for each simulation when no missing data was involved in the phylogenetic inference process, we split the "complete" matrix into three sub-matrices containing no missing data (see Fig.
 ~\ref{Fig_RemoveData}):
@@
 \item
-A first containing both molecular and morphological data for living taxa only (hereafter called the "horizontal" sub-matrix, Table ~\ref{Tab_glossary});
+A sub-matrix containing both molecular and morphological data for living taxa only (hereafter called the "horizontal" sub-matrix, Table ~\ref{Tab_glossary});
+%NC: Do you think the sub-matrix names are useful here?
+We won't use this as much as the "best" tree etc.
+So maybe just write it out each time?
+Also I don't like the names horizontal, vertical and corner - they don't really describe what is going on.
+Better something like - complete living, complete morphology, morphology living?
 \item
-A second one containing morphological data for both living and fossil taxa (hereafter called the "vertical" sub-matrix, Table ~\ref{Tab_glossary});
+A sub-matrix containing morphological data for both living and fossil taxa (hereafter called the "vertical" sub-matrix, Table ~\ref{Tab_glossary});
 \item
-A third one containing morphological data for living taxa only (hereafter called the "corner" sub-matrix, Table ~\ref{Tab_glossary});
+A sub-matrix containing morphological data for living taxa only (hereafter called the "corner" sub-matrix, Table ~\ref{Tab_glossary});
 \end{enumerate}
-We reran the Bayesian tree inference on the different sub-matrices with no missing data in the same way as described above.
-We then compared the resulting tree posterior distribution to itself (in the same way described above) to assess the ability to recover topology for each simulation when no missing data was involved in the phylogenetic inference process.
+We reran the Maximum Likelihood
+%NC: Should we check ML trees too???
+and Bayesian tree inference and topology comparisons on the three sub-matrices in the same way as described above.
 
@@
 \subsection{Empirical data} %Still not done!
-We also compared the results obtained from simulated data by using \citet{ronquista2012} empirical data.
-The matrix contains 67 living species plus one outgroup and 45 fossil species of Hymenopteras with 5097 molecular characters and 354 morphological characters.
-From the 68 living species used in the matrix, only 66 had molecular data, we therefore treated these 66 taxa as "living" taxa and all the other 47 as "fossil" taxa.
-We treated the matrix in the exact same way as described in step 2 and 3 resulting in 125 matrices with various amount of missing data and the same number of Maximum Likelihood and Bayesian trees.
-We used the same settings as for the simulated data in the Maximum Likelihood framework.
-For the Bayesian inferences however, we didn’t used any priors except that we provided a starting tree with the topology of the 68 living species (topology with the highest posterior probability from non-clock analysis - \citet{ronquista2012}).
-Contrary to \citet{ronquista2012} analysis, we didn’t performed any clock analysis since we were only interested in the topology of the inferred tree and not the branch length.
+Finally, to % NC: explain WHY we would want to do this
+
+We also compared the results obtained from simulated data to results from  empirical data in \citet{ronquista2012}.
+The matrix (this would be the "complete" matrix using our terminology) contains 67 living species plus one outgroup and 45 fossil species of Hymenoptera with 5097 molecular characters and 354 morphological characters.
+Because only 66 of the 68 living species in the matrix had molecular data, we treated these 66 taxa as "living" taxa and the remaining 47 as "fossil" taxa.
+
+We removed data from the "complete" matrix in the same way as described in steps 2 and 3 above, resulting in 124 matrices with various amount of missing data.
+We used the same settings as described above to infer Maximum Likelihood trees, but for the Bayesian inferences we did not use any priors except the starting tree.
+This consisted of the topology of the 68 living species with the highest posterior probability from non-clock analysis performed in \citet{ronquista2012}.
+Note that we did not perform any of the clock analyses described in \citet{ronquista2012} because we were only interested in the topology of the inferred trees and not the branch lengths.
+
+% NC: Is there a step missing here? What did you do (or what will you do) with these trees? Will you compare them to each other? How will you compare them to the simulations?
 

### 2014-06-16 Manuscript/TEM_draft.tex
Part of intro correction (to revert)

--- 2014-06-13 Manuscript/TEM_draft.tex
@@
 \bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
-% We don't use a special title page; the author information is entered like any other text.
-% FOOTNOTES: We don't allow them in the manuscript, except in tables. Don't include any footnotes in the text.
-%IMPORTANT:
-%Note on display in the text file: sentences or sub-parts of sentences (e.g. enumerations) are delimited by a newline for clarity.
-%However, true new lines (i.e. in the pdf) are delimited by one or more empty lines.
-%We don't use a heading for the introduction. The nature of this section is implied by its position following the abstract and preceding the first major section heading.
-
-% NC: You should delete this extraneous info from the Syst Biol template to tidy everything up a bit
 
@@
 \bigskip
-\noindent RH: Missing data and topology in total evidence matrices %RH = Running Head
+\noindent RH: Missing data and topology in Total Evidence matrices %RH = Running Head
 
@@
 \noindent{\Large \bf Effect of missing data in matrices containing living and fossil taxa on topological accuracy}
-%Alternative punchier title: "Bayesian total evidence methods fail to recover correct topologies from matrices containing both living and fossil taxa"
-%Alternative punchier title: "Maximum Likelihood outperforms Bayesian methods for inferring topology from total evidence matrices containing both living and fossil taxa"
+%Alternative punchier title: "Bayesian Total Evidence methods fail to recover correct topologies from matrices containing both living and fossil taxa"
+%Alternative punchier title: "Maximum Likelihood outperforms Bayesian methods for inferring topology from Total Evidence matrices containing both living and fossil taxa"
 
@@
 I actually think I prefer your talk title.
-So something like: The effects of missing data on the ability of total evidence approaches to recover correct tree topologies.
+So something like: The effects of missing data on the ability of Total Evidence approaches to recover correct tree topologies.
 No rush to decide.
-But needs total evidence in there somewhere.
+But needs Total Evidence in there somewhere.
 \bigskip
@@
 Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology.
-This fact has led to an increasing consensus among scientists that both fossil and living taxa must be included in macroevolutionary studies.
-One approach, the total evidence approach, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both fossil and living taxa at the tips.
-Although the total evidence approach seems very promising, it requires a lot of data and is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.
-
-In this study we assess the effect of missing data on tree topologies inferred from total evidence matrices.
+This fact has led to an increasing consensus among scientists that both living and fossil taxa must be included in macroevolutionary studies.
+One approach, the Total Evidence approach, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both living and fossil taxa at the tips.
+Although the Total Evidence approach seems very promising, it requires a lot of data and is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.
+
+In this study we assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
 Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix:
@@
 
-Our results suggest that, in order to use total evidence approaches, one should reduce the missing data in the morphological part of the matrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process.
+Our results suggest that, in order to use Total Evidence approaches, one should reduce the missing data in the morphological part of the matrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process.
 %More results!
@@
 
-\noindent (Keywords: missing data, total evidence, Bayesian, Maximum Likelihood, topology)\\
+\noindent (Keywords: missing data, Total Evidence, Bayesian, Maximum Likelihood, topology)\\
 
@@
 Although most species that have ever lived are now extinct % NC: This avoids giving the dodgy stat of 1%
- \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species (e.g., \citealp{meredithimpacts2011}).
+    \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}.
 %or is citing of your own papers in the too "douchebagy" for the second sentence?
@@
 
-Up to now, three main approaches have been used to place both living and fossil taxa % NC: BE consistent. Either living and fossil, or fossil and living.
+Up to now, three main approaches have been used to place both living and fossil taxa % NC: BE consistent. Either living and fossil, or fossil and living. - TG: Let's make it living and fossil then.
     into phylogenies.
@@
 the age of the fossil only or both its age and morphology).
-% NC: Not sure about bit in brackets
+% NC: Not sure about bit in brackets - TG: it's the idea of fossil occurence (age) only when node and fossil occurence + topology when tip.
 Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
@@
 This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty (but see \citealp{spencerefficacy2013}).
-Neontologists, on the other hand, more commonly use probabilistic approaches (e.g.
-Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates % NC: Need to check the exact phrase needed here.
+Neontologists, on the other hand, more commonly use probabilistic approaches (e.g. Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
+Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence age % NC: Need to check the exact phrase needed here.
 Is it range of occurrence?
@@
 etc.
+- TG: in paleo jargon we have FAD (first apparition datum) and FOD (first occurrence datum).
+FAD is always hypothetical (we can't really exactly know when the taxa appeared) and FOD is always subjective (it depends when the fossil was found, maybe next year someone will find another one in older layers).
+So for the dating, because it's relaxed clock, people use FOD (and also probably because they don't care since it doesn't make a huge difference I guess).
  are used to time calibrate phylogenies \citep{zuckerkandl1965}.
@@
 % NC: Tricky. Don't think it's too important though. 
-% NC: I do think Total Evidence needs capital letters though, like ML and Bayesian
+% NC: I do think Total Evidence needs capital letters though, like ML and Bayesian - TG: fixed
 This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{eernissetaxonomic1993}.
@@
 However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data.
-In particular it requires morphological data from both living and fossil taxa, both of which are known to be scarce.
+%In particular it requires morphological data from both living and fossil taxa, both of which are known to be scarce.
 % NC: Really?
@@
 Needs a citation either way.
+%TG: how about this:
+The morphological data for living taxa is rarely collected when molecular data is available (e.g.
+\citealp{O'Leary08022013} vs.
+\citealp{meredithimpacts2011}), and for fossil taxa, the scarcity of the fossil record only allow to collect the data available (for example, in vertebrates, the solidest parts of the skeleton \citeal{sansomfossilization2013}).
 Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
@@
 Sive and I won't be around to correct your English forever.
+- TG: my French demons are back!
+
+%Paused here - 15/063
 
@@
 We compared the topology of the "missing-data" trees inferred from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
-Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees inferred from total evidence type matrices.
+Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees inferred from Total Evidence type matrices.
 To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triples \citep{dobson1975triplets} and the Robinson-Fould \citep{RF1981} distance.
@@
 - That's kind of my way to turn these bad results into something usable: even if some of our simulations are not realistic, it shows the effect of missing data in any scenario possible, not only in realistic cases.
-Also, one can advocate that this low topological resolution is actively due to the amount of missing data in the total evidence matrices since the sub-matrices (no missing data) analysis recovers way higher support value  .
+Also, one can advocate that this low topological resolution is actively due to the amount of missing data in the Total Evidence matrices since the sub-matrices (no missing data) analysis recovers way higher support value  .
 %This is something that I will had with the whole submatrices story.
@@
 \subsection{Maximum Likelihood versus Bayesian}
-The main results of our analysis shows that Bayesian inference fails to recover any Topology in a total evidence method framework (Fig. ~\ref{Fig_Results}).
+The main results of our analysis shows that Bayesian inference fails to recover any Topology in a Total Evidence method framework (Fig. ~\ref{Fig_Results}).
 This results is surprising regarding the behaviour of Maximum Likelihood inference (Fig.
@@
 
-In our case, we suspect that this inability of Bayesian methods to recover topology is due to the intrinsic structure of a total evidence matrix (Fig ~\ref{Fig_RemoveData}).
+In our case, we suspect that this inability of Bayesian methods to recover topology is due to the intrinsic structure of a Total Evidence matrix (Fig ~\ref{Fig_RemoveData}).
 In fact, as previously studies have shown, missing data doesn't seem to be a major drawback as long as the missing data is randomly distributed ~\citep[e.g.][]{wiensmissing2003,rouresite-specific2011,sansomfossilization2013}.
-However, in total evidence matrices, the majority of the missing data is not randomly distributed but concentrated in the molecular part of the matrix for the fossil taxa (i.e.
+However, in Total Evidence matrices, the majority of the missing data is not randomly distributed but concentrated in the molecular part of the matrix for the fossil taxa (i.e.
 the missing-data sub-matrix, Fig ~\ref{Fig_RemoveData}).
@@
 \subsection{Effect of missing data}
-The three parameters we selected in this study account for three potential pitfalls in collecting the data for total evidence analysis:
+The three parameters we selected in this study account for three potential pitfalls in collecting the data for Total Evidence analysis:
 $M_L$ represents the living taxa for which there is no available morphological data,
@@
 \section{Conclusion}
-A Bayesian approach fails to recover accurate topology in a total evidence approach framework, whatever the amount of missing data.
-We think this failure is due to the intrinsic structure of the total evidence matrices that includes a vast amount of non randomly distributed missing data (i.e.
+A Bayesian approach fails to recover accurate topology in a Total Evidence approach framework, whatever the amount of missing data.
+We think this failure is due to the intrinsic structure of the Total Evidence matrices that includes a vast amount of non randomly distributed missing data (i.e.
 the molecular part of the matrix for the fossil taxa).

### 2014-06-17 Manuscript/SupplementaryMaterial/TreeBuilding.tex
Supplementary Materials init

--- 2014-06-13 Manuscript/TEM_draft.tex
@@
-%HEADER
-\documentclass[12pt,letterpaper]{article}
-
-%packages
-% NC: Just to flag here how I'll comment on LaTeX files. I'll use % and then NC. Of course in git diff it'll all show up as an addition.
-% NC: You could add all the packages etc to a template .sty file to save writing it all up here. Rich's ms.sty that I provided on DropBox is a good example
-\usepackage{fixltx2e}
-\usepackage{textcomp}
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{verbatim}
-\usepackage[english]{babel}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{setspace}
-\usepackage{lscape}
-\usepackage{indentfirst}
-\usepackage[normalem]{ulem}
-\usepackage{booktabs}
-%\usepackage{nag}
-\usepackage{natbib}
-%\usepackage{bibtex}
-\usepackage{float}
-\usepackage{latexsym}
-%\usepackage{hyperref} 
-\usepackage{url}
-%\usepackage{html}
-\usepackage{hyperref}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{bm}
-\usepackage{array}
-\usepackage[version=3]{mhchem}
-\usepackage{ifthen}
-\usepackage{caption}
-\usepackage{hyperref}
-%\usepackage{xcolor}
-\usepackage{amsthm}
-\usepackage{amstext} % Add and remove packages as necessary for your manuscript.
-
-% NC: Added by NC
-\usepackage{enumerate}
-\usepackage[osf]{mathpazo} % palatino is a nicer font! 
-\pagenumbering{arabic}	% Page numbers
-
-%Pagination style and stuff
-\linespread{1.66} % All text should be double-spaced with occasional exceptions for tables. 
-\raggedright
-\setlength{\parindent}{0.5in}
-
-\setcounter{secnumdepth}{0} % Our sections are not numbered and our papers do not have Tables of Contents.
-We don't present a list of figures or list of tables, either.
-Any common font is fine.
-(A common sans-serif font should be used on figures, but figures should be separate from the LaTeX document.)
-
-%\pagestyle{empty}
-
-\renewcommand{\section}[1]{%
-\bigskip
-\begin{center}
-\begin{Large}
-\normalfont\scshape #1
-\medskip
-\end{Large}
-\end{center}}
-
-\renewcommand{\subsection}[1]{%
-\bigskip
-\begin{center}
-\begin{large}
-\normalfont\itshape #1
-\end{large}
-\end{center}}
-
-\renewcommand{\subsubsection}[1]{%
-\vspace{2ex}
-\noindent
-\textit{#1.}---}
-
-\renewcommand{\tableofcontents}{}
-
-\bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
-% We don't use a special title page; the author information is entered like any other text.
-% FOOTNOTES: We don't allow them in the manuscript, except in tables. Don't include any footnotes in the text.
-%IMPORTANT:
-%Note on display in the text file: sentences or sub-parts of sentences (e.g. enumerations) are delimited by a newline for clarity.
-%However, true new lines (i.e. in the pdf) are delimited by one or more empty lines.
-%We don't use a heading for the introduction. The nature of this section is implied by its position following the abstract and preceding the first major section heading.
-
-% NC: You should delete this extraneous info from the Syst Biol template to tidy everything up a bit
-
-\begin{document}
-
-%For a colour version of the draft to a search>replace '-BW'>'' 
-% %\modulolinenumbers[1] 	% Line numbering on every line
-
-%Running head
-\begin{flushright}
-Version dated: \today
-\end{flushright}
-\bigskip
-\noindent RH: Missing data and topology in total evidence matrices %RH = Running Head
-
-\bigskip
-\medskip
-\begin{center}
-
-%Title
-\noindent{\Large \bf Effect of missing data in matrices containing living and fossil taxa on topological accuracy}
-%Alternative punchier title: "Bayesian total evidence methods fail to recover correct topologies from matrices containing both living and fossil taxa"
-%Alternative punchier title: "Maximum Likelihood outperforms Bayesian methods for inferring topology from total evidence matrices containing both living and fossil taxa"
-
-% NC: Not sure about any of these.
-I actually think I prefer your talk title.
-So something like: The effects of missing data on the ability of total evidence approaches to recover correct tree topologies.
-No rush to decide.
-But needs total evidence in there somewhere.
-\bigskip
-
-%Authors
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$, other authors $^3$, and Natalie Cooper$^1$$^,$$^2$}\\
-\noindent {\small \it 
-$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland;\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland;\\
-$^3$Somewhere else }\\
-\end{center}
-\medskip
-\noindent{\bf Corresponding author:} Thomas Guillerme, School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.\\
-\vspace{1in}
-% NC: It's not called the Republic of, just Ireland!
-
-\newpage
-\begin{abstract}
-% NC: I'm not going to give this much work as it'll change before we submit anyway. Note though that and abstract must NOT just repeat stuff in the intro. Needs rephrasing at the very least.
-Living species represent less than 1\% of all species that have ever lived.
-% NC: I was chatting to Andy Purvis about this the other day and he said lots of palaeo folk hate that stat.
-So let's not start the abstract with it!
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology.
-This fact has led to an increasing consensus among scientists that both fossil and living taxa must be included in macroevolutionary studies.
-One approach, the total evidence approach, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both fossil and living taxa at the tips.
-Although the total evidence approach seems very promising, it requires a lot of data and is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.
-
-In this study we assess the effect of missing data on tree topologies inferred from total evidence matrices.
-Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix:
-(1) the proportion of living taxa with no morphological data,
-(2) the amount of missing data in the fossil record, and
-(3) the overall number of morphological characters in the matrix.
-We find that, in a Bayesian framework, difficulties in recovering a stable topology are mainly driven by the missing data in the molecular part of the matrix (for which fossil taxa have no data).
-In a Maximum Likelihood framework, however, topology is not directly affected by missing data \textit{per se}, but by the number of morphological characters shared among the taxa.
-Therefore, the two main drivers of incorrect topologies are the overall number of morphological characters and the number of living species with no morphological data.
-
-Our results suggest that, in order to use total evidence approaches, one should reduce the missing data in the morphological part of the matrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process.
-%More results!
-\end{abstract}
-
-\noindent (Keywords: missing data, total evidence, Bayesian, Maximum Likelihood, topology)\\
-
-\vspace{1.5in}
-
-\newpage
-
-\section{Introduction}
-
-Although most species that have ever lived are now extinct % NC: This avoids giving the dodgy stat of 1%
- \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species (e.g., \citealp{meredithimpacts2011}).
-%or is citing of your own papers in the too "douchebagy" for the second sentence?
-% NC: Yeah be careful with self citing. Fine if it's a good fit, but it's totally not here!!! A better citation would be a big biodiversity study, e.g. the bird phylogeny paper.
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
-This has led to increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments (e.g.
-\citealp{pyrondivergence2011,ronquista2012,schragocombining2013}).
-%Not a nice sentence
-
-Up to now, three main approaches have been used to place both living and fossil taxa % NC: BE consistent. Either living and fossil, or fossil and living.
-into phylogenies.
-These approaches differ mainly in whether they treat fossil taxa as tips or as nodes in the phylogeny, and in which part of the available fossil data is used (i.e.
-the age of the fossil only or both its age and morphology).
-% NC: Not sure about bit in brackets
-Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
-Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \citep{simpson1945}.
-This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty (but see \citealp{spencerefficacy2013}).
-Neontologists, on the other hand, more commonly use probabilistic approaches (e.g.
-Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates % NC: Need to check the exact phrase needed here.
-Is it range of occurrence?
-Occurrence dates?
-etc.
- are used to time calibrate phylogenies \citep{zuckerkandl1965}.
-There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013}.
-However neither approach uses all the available data.
-
-A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa.
-%There might be a philosophico-semantic distinction to do here.
-Is TEM a method or an approach?
-I'll say that it is an approach because it's more like an idea (put every thing together) than a way to do (how to put everything together).
-It makes more sense to me compared to a real method e.g.
-a probabilistic method such as ML where the method states precisely how to do stuff.
-% NC: Tricky. Don't think it's too important though. 
-% NC: I do think Total Evidence needs capital letters though, like ML and Bayesian
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{eernissetaxonomic1993}.
-Total Evidence methods have been successfully applied to empirical data \citep{pyrondivergence2011,ronquista2012,schragocombining2013}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
-However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data.
-In particular it requires morphological data from both living and fossil taxa, both of which are known to be scarce.
-% NC: Really?
-I know fossil data is known to be scarce, but living data?
-I think people would be surprised.
-And they are surprised.
-We were surprised!
-Needs a citation either way.
-Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
-% Trevor: Missing data also influences support values that trend to decrease
-
-% NC: You need to take a bit of time to learn about singular and plural grammar in English.
-It's something you always do wrong.
-e.g.
-The Total Evidence method *uses* this and *its* parameters are as follows.
-Versus: Total Evidence methods *use* this and *their* parameters are as follows.
-Sive and I won't be around to correct your English forever.
-
-The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013}.
-Missing molecular data has been seen by some authors as an issue because it can decrease phylogenetic signal (i.e. % define phylogenetic signal here) 
-in some parts of the tree, especially when using large matrices \citep{lemmonthe2009}. However, this may not be a major issue because phylogenetic signal is easily increased by:
-(i) including a "modest" number of highly-covered % what do you mean by highly-covered?
-genes (i.e. approximatively half of the genes; \citealt{rouresite-specific2011});
-(ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and
-(iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
-Similarly, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study in question \citep{wiensmissing2003,sansomfossilization2013}.
-Because soft-tissue characters are rarely preserved in the fossil record, missing data is mainly found in these characters, and is therefore not randomly distributed which can lead to biased placement of fossil taxa in phylogenies \citep{sansomfossilization2013}.
-However, the phylogenetic signal is not related to the amount of missing data \textit{per se} but to the number of informative characters for each taxon, therefore missing data is less of an issue than the number of shared informative characters \citep{wiensmissing2003}.
-% NC: I'm a little dubious of what this paragraph is trying to get at in places. We should discuss at some point.
-
-Although missing data does not appear be a major problem in molecular and morphological matrices separately \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil species.
-This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data.
-Until now, no attempt has been made to study the impact of this issue on phylogenetic inference from Total Evidence methods.
-
-% NC: Somewhere here you need to flag that you're only looking at topologies and briefly explain why.
-
-Here we use simulations to assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
-The molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}.
-The effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}, therefore, we focus only on missing data in the morphological part of the matrix.
-We investigate three major parameters that directly affect the completeness of the morphological part of the matrix:
-\begin{enumerate}
-\item the proportion of living taxa with no morphological data;
-\item the amount of missing data in the fossil taxa; and
-\item the overall number of morphological characters for both living and fossil taxa in the matrix.
-\end{enumerate}
-We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
-We chose these parameters because they reflect real-life biases in data availability.
-The advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
-Missing data in fossil taxa is very common due to preservation biases (citation?), and
-% OK I'm not sure what the "real-life" reason for three is.
-%This paragraph is crap
-% NC: I think it's fine!
-
-We find that when using a Maximum Likelihood approach, as missing data increases, the likelihood of recovering the correct tree topology decreases.
-However, even with no missing data, Total Evidence matrices dramatically reduce the performance of Bayesian methods for inferring tree topology.
-We propose that this drastic difference between Bayesian and Maximum Likelihood methods is due to a flattening of the likelihood landscape caused by the unavoidable amount of missing molecular data for fossil taxa in a Total Evidence matrix.
-We make suggestions for how best to deal with this issue when inferring phylogenies from Total Evidence matrices.
-
-% NC: You don't want to tell the whole story here. Just a results summary and a quick line of interpretation
-
-%This problem leads to an inability of Bayesian chains to pick up the sub and near optimal topologies leading to high variance in the posterior tree distribution and really low support in the consensus tree.
-%These results should encourage authors in not using Bayesian inference to explore the topological landscape of their phylogeny.
-%Instead, we suggest that authors should first find the Maximum Likelihood tree and use it's topology as a fix prior in their Bayesian inference.
-%We also propose a new method for comparing Bayesian posterior tree distribution's topology available in a wonderful R package. %Or am I just dreaming a bit too much here?
-
-% NC: Dreaming a bit much I think.
-There's a discussion to be had here about whether it's really a *new* method and whether you really want to make an R package per se.
-We definitely want to share the code
-
-%-----------
-% NC: It might look like I've done a lot to this but actually I think it's nearly there.
-A good thing to take from this is that writing a paper is *hard*.
-It gets easier, but it's still a deeply involved process with lots of writing and re-writing.
-Introductions are particularly difficult because if you don't set the scene, it's hard to convince people to get on board for the rest of the paper.
-This is why I said there was no way this was being submitted by the end of June!
-Still a long way to go I'm afraid in terms of time and effort.
-But that doesn't mean you're not doing well, *every* project is like this.
-2 years is about average from conception to sending out a manuscript, and you only started work on this last spring.
-%------------
-
-\newpage
-\section{Methods} %no material needed?
-% NC: Depends on the journal. See what Syst Biol asks for as section titles.
-To explore how missing data in Total Evidence matrices influences tree topology we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
-%NC: does Syst Biol use Fig.
-or Figure?
-\begin{enumerate}
-\item{Generating the matrix} \label{step:generate_matrix} \\
-We randomly generated a birth-death tree (hereafter called the "true" tree; Table ~\ref{Tab_glossary}) and used it to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix; Table ~\ref{Tab_glossary}).
-\item{Removing data} \label{step:remove_data} \\
-We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of missing living taxa ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$) % NC: make sure to use the same descriptions here and in the introduction
-(the resulting matrices are called hereafter "missing-data" matrices; Table ~\ref{Tab_glossary}).
-\item{Building phylogenies} \label{step:build_phylo} \\
-We built phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree; Table ~\ref{Tab_glossary}) and multiple trees inferred from matrices with missing morphological data (hereafter called the "missing-data" trees; Table ~\ref{Tab_glossary}).
-Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
-\item{Comparing topologies} \label{step:compare_topo} \\
-We compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies.
-\end{enumerate}
-To measure the effect of missing data distribution, we repeated steps ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} with the exact same fixed parameters 51 (3$\times$17) times.
-%that's because I had 3 cluster accounts available...
-If it's a too odd number, no worries, I'll just remove the last chain.
-A list of all the terms used in this paper is available in table ~\ref{Tab_glossary}.
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_outline-BW.pdf}
-\caption{Protocol outline.
-(1) We generated a random tree (the "true" tree) to infer a matrix with no missing data (the "complete" matrix).
-(2) We removed data from the morphological partition of the "complete" matrix resulting in 125 "missing-data" matrices.
-(3) We infer phylogenetic trees from each matrix in both ML and Bayesian framework.
-(4) We then compared the "missing-data" trees to the "best" tree.
-We repeated step ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} 51 (3$\times $17) times.}
-\label{Fig_Outline}
-\end{figure}
-
-\begin{table}[ht]
-\caption{Glossary}
-\centering
-\begin{tabular}{p{5cm} p{9cm}}
-\hline
-\hfill Term & Definition \\[0.5ex]
-\hline
-\hfill living taxa & taxa with both molecular and morphological data available \\[1.5ex]
-\hfill fossil taxa & taxa with only morphological data available \\[1.5ex]
-\hfill "complete" matrix & matrix with no missing data except for the molecular part of the fossil taxa \\[1.5ex]
-\hfill "missing-data" matrix & matrix with various amount of missing data \\[1.5ex]
-\hfill $M_{L}$ & missing living taxa in the morphological part of the matrix \\[1.5ex]
-\hfill $M_{F}$ & missing data for the fossil taxa in morphological part of the matrix \\[1.5ex]
-\hfill $M_{C}$ & missing morphological characters for both living and fossil taxa \\[1.5ex]
-\hfill "true" tree & tree used to simulate the matrix \\[1.5ex]
-\hfill "best" tree & tree inferred from the "complete" matrix \\[1.5ex]
-\hfill "missing-data" tree & tree inferred the "missing-data" matrices \\[1.5ex]
-\hfill RPBTC & Random pairwise Bayesian tree comparison \\[1.5ex]
-\hfill "horizontal" sub-matrix & sub-matrix with morphological and molecular characters for living taxa \\[1.5ex]
-\hfill "vertical" sub-matrix & sub-matrix with morphological data for both living and fossil taxa \\[1.5ex]
-\hfill "corner" sub-matrix & sub-matrix with morphological data for living taxa \\[1.5ex]
-\hfill "missing-data" sub-matrix & molecular part of the matrix for the fossil taxa (no data) \\
-\end{tabular}
-\label{Tab_glossary}
-\end{table}
-
-%---------------------------------------------
-%       Generating the matrix
-%---------------------------------------------
-
-\subsection{Generating the matrix}
-First we randomly generated a "true" tree of 50 taxa in R v3.0.2 \citep{R302} using the package diversitree v0.9-6 \citep{fitzjohndiversitree2012}.
-We generated the tree using a Birth Death process by sampling the values of the speciation events ($\lambda$) and extinction events ($\mu$) from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
-We implemented a rejection sampling algorithm to select only random trees with 25 living and 25 fossil taxa. %Justification needed?
-We then added a taxa to the resulting Birth-Death tree as the outgroup of the tree.
-The mean branch length of the tree was used to separate the outgroup from the rest of the taxa and the branch length leading to the outgroup was set as the sum of the mean branch length and the longest root-to-tip length of the tree.
-
-Next, we created a molecular and a morphological matrix from the "true" tree.
-The molecular matrix was inferred from the "true" tree using the package phyclust v0.1-14 \citep{chen2011}.
-The matrix was made of 1000 characters sites for 51 taxa and generated using the seqgen algorithm \citep{ranbaut1997seqgen}.
-We used the HKY model \citep{HKY85} with a random base frequencies and with the transition/transversion rate of 2 \citep{douadycomparison2003} as parameters for generating the matrix.
-The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
-We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
-These parameters were selected to generate data with no special assumption about how the characters evolved as well as to reduce the computational time required if these parameters were estimated rather than defined (total computational time $>$ 65 CPU years).
-
-We inferred the morphological matrix using the ape package v3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
-We assigned the number of character states (either 2 or 3) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of characters states extracted from 100 published empirical morphological matrices (See \hyperref[supplementaries]{supplementaries}).
-We then ran an independent discrete character simulation for each character using the "true" tree branch length and topology with the randomly selected number of states (2 or 3) and assuming an equal rate of change (i.e.
-evolutionary rate) from one character state to an other \citep{Pagel22011994}.
-This method allows us to have only two parameters per character: the number of states and the evolutionary rate.
-For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used a low evolution rate parameter (i.e.
-$\alpha$) in order to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
-
-All the molecular information for fossil taxa was replaced by missing data ("?").
-Finally, we combined the morphological and molecular matrices obtained from the "true" tree.
-Hereafter we call this the "complete" matrix: the matrix with no missing data except for the molecular data of the fossil taxa.
-
-%---------------------------------------------
-%       Removing data
-%---------------------------------------------
-
-\subsection{Removing data}
-Once we obtained the "complete" matrix we modified it to get a set of matrices with missing data. %Ugly sentence
-We randomly replaced data with "?" in the morphological part of the matrices according to the following parameters (Fig. ~\ref{Fig_RemoveData}):
-
-\begin{enumerate}
-\item{The proportion of living taxa with no morphological data ($M_{L}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological one.
-Because of the increasing facility to sequence DNA for living taxa, the number of living taxa with molecular data is highly superior the the number of taxa with molecular and morphological data.
-\item{The proportion of missing morphological data across all fossil taxa ($M_{F}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the quality of the fossil record. 
-\item{The proportion of missing morphological characters across all taxa (living and fossil - $M_{C}$): 0\%, 10\%, 25\%, 50\% or 75\%. }
-This parameter illustrates the number of available morphological characters for both living and fossil taxa.
-\end{enumerate}
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_missingData-BW.pdf}
-\caption{Names of the different parts of the matrix.
-A: Missing data parameters:
-Missing living - The proportion of living taxa with no morphological data ($M_{L}$);
-Missing fossil - The proportion of missing morphological data across all fossil taxa ($M_{F}$);
-Missing character - The proportion of missing morphological characters across all taxa (living and fossil) ($M_{C}$).
-B: Different parts of the matrix:
-The "horizontal" sub-matrix (orange) contains molecular and morphological data for living taxa;
-The "vertical" sub-matrix (blue) contains morphological data for living and fossil taxa;
-The "corner" sub-matrix (orange and blue striped) contains morphological data for living taxa;
-The "missing-data" sub-matrix (grey) is the molecular part of the fossil taxa and contains no data.}
-\label{Fig_RemoveData}
-\end{figure}
-
-In practice, each parameter represent a different way of removing data in the morphological part of the matrix: $M_{L}$ removes a proportion of rows from the living taxa; $M_{F}$ removes a proportion of cells from the fossil taxa; and $M_{C}$ removes a proportion of columns across both living and fossil taxa (see Fig.
-~\ref{Fig_RemoveData}).
-Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a proportion of the living taxa is removed (i.e.
-removing rows), as for $M_{F}$, a proportion of data is removed across the whole of the morphological matrix for fossil taxa (i.e.
-removing cells).
-
-We tested all parameters combinations resulting in 125 ($5^3$) matrices.
-Because some parameter combinations introduce a lot of missing (e.g. $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
-When this occurred we repeated the random deletion of characters until every taxa had at least 5\% data across the whole matrix.
-
-%---------------------------------------------
-%       Building phylogenies
-%---------------------------------------------
-
-\subsection{Building phylogenies}
-From the resulting matrices we generated two types of trees, the "best" tree that is inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
-The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
-The "best" tree, on the other hand, is the best tree we can build using the state-of-the-art phylogenetic methods.
-In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g. \citet{rozen2005}).
-Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
-
-\subsubsection{Maximum Likelihood}
-The "best" tree and the "missing-data" trees were inferred using RAxML v8.0.20 \citep{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model (\citet{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citet{Stamatakis21012014}) as a generalisation of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
-The GTR model can be seen as a generalisation of the HKY model (the 2 parameters from the HKY model are implicitly included in the 6 from GTR model - \citet{stamatakisa2008}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} which is a generalisation of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014}).
-We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. %The explanations of the algorithm part for the BS was proposed by Trevor
-The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists in pruning one sub-tree from the tree and subsequently reinserting it to al neighbouring branches \citep{stamatakisa2008}.
-Sub-tree Pruning and Reinserting methods (SPR) have been demonstrated as being better than others (e.g. Nearest Neighbouring Interchange - NNI) in recovering good bootstrap values \citep{salamin2003}.
-
-\subsubsection{Bayesian}
-The "best" tree and the "missing-data" trees were inferred using MrBayes v3.2.2 \citep{Ronquist2012mrbayes}.
-We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
-The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \citep{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
-For the morphological data, we used the Markov \textit{k} state model \citep{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
-We chose these models to be consistent with the parameters used to generate the "complete" matrix.
-
-Each Bayesian tree was estimated using two runs of four chains each for a maximum of 50$\times$$1^6$ generations.
-We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
-The effective sample size (ESS) was also checked on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
-For each run, we removed 25\% of the iterations as burn-in.
-We used the following prior for each tree (see \hyperref[supplementaries]{supplementaries}):
-\begin{enumerate}
-\item
-the "true" tree’s topology as a starting tree (with a starting value for each branch length of 1),
-\item
-an exponential prior on the shape of the gamma distribution of $\alpha$=0.5 for both partitions
-\item
-and a transition/transversion ratio prior of 2 sampled from a strong beta distribution ($\beta$(80,40)).
-\end{enumerate}
-
-We used these prior to speed up the Bayesian process.
-These prior biased the way the Bayesian process calculated the branch length by giving non-random starting points and boundaries for the parameters estimation process, however, in this study, we focused on the effect of missing data on the topology and not on the branch length.
-Even using these prior, it took 65 CPU years to build 51 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
-
-%---------------------------------------------
-%       Comparing Topologies
-%---------------------------------------------
-
-\subsection{Comparing topologies}
-We compared the topology of the "missing-data" trees inferred from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
-Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees inferred from total evidence type matrices.
-To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triples \citep{dobson1975triplets} and the Robinson-Fould \citep{RF1981} distance.
-We normalised the two metrics using the Normalised Tree Similarity index \citep{Bogdanowicz2012} to generalise our results for any \textit{n} number of taxa.
-The three metrics are detailed below.
-
-\subsubsection{Triples distance ($T_{x,y}$) \citep{dobson1975triplets}}
-This metric measures the number of different sub-trees of three taxa between two given trees.
-Each triplet can be written as $I_{ijk}$=(\textit{ijk}).
-Where $I_{ijk}$ is equal to 0 if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to 1.
-For any rooted binary tree there are only three possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
-If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k});.
-One can calculate $S_n$, the triplet distance between two trees as:
-\begin{equation}
-S_n = \sum_{ijk} I_{ijk}
-\end{equation}
-Where:
-\begin{equation}
-\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
-\end{equation}
-And where \textit{n} is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
-If $S_n$=0, the trees are the same (i.e. no taxa as been displaced).
-When $S_n$ = $\binom{n}{4}$, the trees are the most different possible (i.e. every taxa as been displaced).
-This metric therefore illustrates the amount of displayed taxa.
-It is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g.
-fossil taxa) than the Robinson-Foulds metric \citep{critchlowthe1996,johnson1998,wiensmissing2003}.
-Therefore we used this metric as a proxy to estimate the robustness of the tree to flying taxa (see \hyperref[supplementaries]{supplementaries}).
-
-\subsubsection{Robinson-Fould distance (\textit{RF}) \citep{RF1981}}
-This metric measures the number of shared clades among two trees and therefore illustrates the number of exactly conserved groups among the trees.
-The Robinson-Fould distance (also called path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
-\begin{equation}
-RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
-\end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees.
-The minimal value of \textit{C} is 1 if the two trees have the same \textit{n} taxa; the maximal value in \textit{C=n-2}.
-This metric is more sensitive to taxa displacement than the Triples metric (i.e.
-if one taxa gets out of a clade, then the clades are no longer considered as similar - \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
-Therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades (see \hyperref[supplementaries]{supplementaries}).
-
-\subsubsection{Normalised Tree Similarity \textit{NTS} \citep{Bogdanowicz2012}}
-For any tree with \textit{n} taxa compared using a tree distance metric \textit{m}, $NTS_m$ represents the similarity score between the two trees given the expected distance between two random Yule trees with \textit{n} taxa.
-Let $\bar{d}_{m,n}$\textit{(rand)} be the average distance between two random Yule trees with \textit{n} taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} containing each \textit{n} taxa, then:
-\begin{equation}
-NTS_{m,n}(x,y)=\frac{\bar{d}_{m,n}(rand) - d_{m,n}(x,y)} {\bar{d}_{m,n}(rand)}
-\end{equation}
-\textit{NTS} ranges from 1 to -$\infty$.
-For any \textit{m},\textit{n} when \textit{NTS}=1, the trees are the same;
-when \textit{NTS}=0 the trees are not more different than expected by chance;
-when \textit{NTS}$<$0, the trees are more different than expected by chance (see \hyperref[supplementaries]{supplementaries}).
-
-$\newline$
-
-We compared the "missing-data" tree to the "Best" tree for each chain.
-For the Maximum Likelihood trees we performed pairwise comparisons between the "Best" tree and the "missing-data" tree (see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triples metric.
-We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
-For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparisons for the same metric and the same number of taxa \textit{n}=51 (see \hyperref[supplementaries]{supplementaries}).
-We ran the comparison for every "missing-data" tree in each chain resulting in 51 comparisons for every "missing-data" trees. 
-We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package (v3.1 – \citet{hdrcde}).
-
-Bayesian tree inference allows to account for the statistical uncertainty of a phylogenetic tree by not using an optimal criterion (c.f.
-Maximum Likelihood) and gives a tree posterior distribution (c.f.
-the likeliest tree in ML).
-This method has the clear advantage of better dealing with error and uncertainty but it's output (a distribution rather than a single tree) is less practical to use for any further study (but see \citet{healyecology2014}).
-%Awwwwww yiis!
-To avoid this problem, people traditionally use the a consensus tree build on a majority rule in order to summarise a Bayesian tree posterior distribution.
-This results into a single tree containing both topological and branch length information as well as support information (i.e. posterior probabilities, e.g. \citet{Ronquist2012mrbayes}).
-However, using a Bayesian consensus tree has limitation especially if the resulting consensus tree is not well supported or resolved.
-Because the metrics used in this study are used to measure variation in topology between two trees (i.e.
-taxa placement or clade position), comparing two Bayesian consensus trees is not optimal in picking topological difference signal.
-For example, if one of the trees is not resolved at all (i.e. a star tree), comparing it to any other tree (even an other star tree) will not give useful results.
-Therefore we used the entire Bayesian trees posterior distribution in order to perform the tree comparisons (i.e.
-the "Best" Bayesian tree posterior distribution vs.
-the "missing-data" Bayesian tree posterior distribution).
-
-\subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
-We compared the Bayesian posterior distribution using a Random Pairwise Bayesian Tree Comparison (RPBTC) method.
-This method consists in comparing a series of randomly selected pairs of trees from two posterior distributions (on from each distribution) and use the mode to summarize the resulting distribution as a proxy of the difference between the two trees in a Bayesian framework.
-\begin{equation}
-RPBTC_{X,Y}=Mo[(]d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]
-\end{equation}
-
-Where \textit{X} and \textit{Y} are two Bayesian tree posterior distributions;
- $d_{m,n}$ is the pairwise difference for any metric \textit{m} and \textit{n} taxa between $X_{i}$ and $Y_{j}$ which are two single trees randomly sampled respectively from \textit{X} and \textit{Y};
- $Mo$ is the mode and
- $Mo[(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]$ is the mode of the pairwise difference repeated \textit{k} times.
-
-We used the mode to summarize the distributions because it is the value that is the most represented in the distribution and reflects what a consensus tree represents (the topology the most represented in the posterior distribution).
-Also, the mode is a value present in the distribution (which, depending on the metric, can be a series of discrete values) in contrast to the other distribution summary metrics, such as the mean or the median, which can take values that are not actually present in the distribution.
-For example, if comparing two posterior tree distribution, we obtain difference values of 10, 10, 10, 5, 5 and 1, using the mean (6.8) or the median (7.5) gives a values that actually doesn't exist.
-%Useful?
-
-In this study, we calculated RPBTC for both metrics (Robinson-Fould and Triples distance) for each pairs of Bayesian tree distributions (i.e.
-the "Best" Bayesian tree posterior distribution vs.
-the "missing-data" Bayesian tree posterior distribution) with 1000 random pairwise comparisons (\textit{k}).
-Because the RPBTC uses random pairwise comparisons, there is a chance of comparing always trees that have the biggest or the smallest difference between two distributions, either deflating or inflating the RPBTC difference value.
-However, using 1000 random pairwise comparisons makes the difference stable (i.e. if repeated independently, no difference is detected - see \hyperref[supplementaries]{supplementaries}).
-
-$\newline$
-
-We calculated the NTS for the "Best" Bayesian tree vs.
-"missing-data" Bayesian tree comparison for both Robinson-Fould and Triples metric using the mode of the RPBTC for each chain resulting in a distribution of 51 modes per comparison.
-We then compared the NTS of the Bayesian trees to the NTS of the ML trees for both metrics for each comparison.
-This resulted in a distribution of 51 NTS values for the Bayesian trees and 51 NTS values for the ML trees for each tree comparison for which we calculated the mode and the 50\% and 95\% confidence intervals (see \hyperref[supplementaries]{supplementaries} - Fig.
-~\ref{Fig_Compare}).
-
-\begin{figure} % Modify the figure
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp-BW.pdf}
-\caption{Tree comparison protocol.
-From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we inferred a tree in Baysian or Maximum likelihood framework.
-We then compared this tree with or without fossil/living taxa to the "best" tree.
-In the Bayesian framework, 1000 trees from the posterior distribution where used for each comparison.}
-\label{Fig_Compare}
-\end{figure}
-
-\subsubsection{Statistical difference between tree topologies} %note that in the end, I only have non-parametric tests, the parametric explanation part can go to the supplementaries
-To assess the difference between the tree topologies, whether they were inferred in a Bayesian or a ML framework, we performed a group comparison test for each set of configurations (e.g.
-effect of $M_L$ in Bayesian inference, combined effect of $M_F$ and $M_C$ in ML inference, etc...).
-We used the pairwise tree comparisons as factors (i.e.
-the "Best" tree vs.
-one of each of the 125 "missing-data" trees) and the values of the 51 replicates from the pairwise comparison as a response variable.
-The 51 replicates values were either the replicates of the 51 pairwise tree comparisons in ML or the modes of the 51 RPBTC posterior distributions in Bayesian.
-When we found a significant difference between the groups, we performed a pairwise comparison test to analyse which pairs of groups differ from the others.
-Depending on whether the data was normal with variance homoscedasticity or not (controlled by performing respectively a shapiro.test and a a bartlett.test in R\{stats\}), we used respectively parametric or non-parametric tests for the group comparison and the pairwise comparisons as suggested by \citet{ruxtontime2008} (Table ~\ref{Stats_test}).
-%link broken
-
-\begin{table}[ht]
-\caption{Statistical tests}
-\centering
-\begin{tabular}{lrll}
-  \hline
-   & test & statistics & code\{package\} \\ [0.5ex]
-  \hline
-  Groups comparisons & ANOVA & parametric & anova\{stats\} \\ [0.5ex]
-                     & Kruskal-Wallis & non-parametric & kruskal.test\{stats\} \\[0.5ex]
-  \hline
-  Pairwise comparisons & Tukey HSD & parametric & TukeyHSD\{stats\} \\ [0.5ex]
-                       & Nemenyi-Damico-Wolfe-Dunn & non-parametric & kruskalmc\{pgirmess\} \\ [0.5ex]
-  \hline
-\end{tabular}
-\label{Stats_test}
-\end{table}
-
-\subsection{Effect of missing molecular characters for fossil taxa} %Awfull name.
-To assess the effect of missing molecular characters for fossil taxa in a Bayesian framework, we split the "complete" matrix in sub-matrices containing no missing data at all (see Fig.
-~\ref{Fig_RemoveData}):
-\begin{enumerate}
-\item
-A first containing both molecular and morphological data for living taxa only (hereafter called the "horizontal" sub-matrix, Table ~\ref{Tab_glossary});
-\item
-A second one containing morphological data for both living and fossil taxa (hereafter called the "vertical" sub-matrix, Table ~\ref{Tab_glossary});
-\item
-A third one containing morphological data for living taxa only (hereafter called the "corner" sub-matrix, Table ~\ref{Tab_glossary});
-\end{enumerate}
-We reran the Bayesian tree inference on the different sub-matrices with no missing data in the same way as described above.
-We then compared the resulting tree posterior distribution to itself (in the same way described above) to assess the ability to recover topology for each simulation when no missing data was involved in the phylogenetic inference process.
-
-%---------------------------------------------
-%       Empirical data - TO DO
-%---------------------------------------------
-
-\subsection{Empirical data} %Still not done!
-We also compared the results obtained from simulated data by using \citet{ronquista2012} empirical data.
-The matrix contains 67 living species plus one outgroup and 45 fossil species of Hymenopteras with 5097 molecular characters and 354 morphological characters.
-From the 68 living species used in the matrix, only 66 had molecular data, we therefore treated these 66 taxa as "living" taxa and all the other 47 as "fossil" taxa.
-We treated the matrix in the exact same way as described in step 2 and 3 resulting in 125 matrices with various amount of missing data and the same number of Maximum Likelihood and Bayesian trees.
-We used the same settings as for the simulated data in the Maximum Likelihood framework.
-For the Bayesian inferences however, we didn’t used any priors except that we provided a starting tree with the topology of the 68 living species (topology with the highest posterior probability from non-clock analysis - \citet{ronquista2012}).
-Contrary to \citet{ronquista2012} analysis, we didn’t performed any clock analysis since we were only interested in the topology of the inferred tree and not the branch length.
-
-%---------------------------------------------
-%
-%       RESULTS
-%
-%---------------------------------------------
-
-\section{Results}
-
-\subsection{Building the trees}
-When generating the matrices using seqgen \citep{ranbaut1997seqgen} and rTraitDisc \citep{paradisape:2004} algorithms with low evolutionary rate parameter ($\alpha$=0.5) we successfully generated evolutionary histories.
-In Maximum Likelihood, support values ranged from 0 to 100 with a median of 38 ($1^st$ quartile = 14, $3^rd$ quartile = 72 - see \hyperref[supplementaries]{supplementaries}). % This is actually bad.
-The distribution of the phylogenetic signal across the different chains provided us a good spread of evolutionary scenarios: from well resolved phylogenies to poorly resolved ones.
-This allows us to test our simulations in a theoretical as well as practical framework (i.e.
-phylogenies with low support values are rarely published but are still common in early stages of projects).
-%REF???
-- That's kind of my way to turn these bad results into something usable: even if some of our simulations are not realistic, it shows the effect of missing data in any scenario possible, not only in realistic cases.
-Also, one can advocate that this low topological resolution is actively due to the amount of missing data in the total evidence matrices since the sub-matrices (no missing data) analysis recovers way higher support value  .
-%This is something that I will had with the whole submatrices story.
-Basically, when you have all the fossil/molecular data missing, it's harder to build the tree, both in ML and in Bayesian.
-When you run the analysis without missing data, then it works way better (i.e.
-our REAL control test: can we simulate matrices in the first place).
-Using Bayesian inference, all the chains converged with a significant effective sample size by using 2 parallel runs of 4 chains each (ASSD $<$ 0.01 and ESS $>>$ 200 for inference).
-
-%\subsection{Comparing topologies}
-%Did it worked? See sub-matrices part. TO DO
-
-\subsection{Effect of the missing data parameters}
-
-%10/06 - Missing 8 ML Chains
-\begin{figure} 
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-results-BW.pdf}
-\caption{Effect of missing data on Tree similarity.
-$M_L$ is the amount of missing living taxa with morphological data,
-$M_F$ is the amount of missing data in the fossil record,
-$M_C$ is the amount of missing morphological characters.
-The first row represent the Normalised Tree Similarity index using the Robinson-Fould metric;
-the second row represent the Normalised Tree Similarity index using the Triples metric.}
-\label{Fig_Results}
-\end{figure}
-
-\subsubsection{Effect of $M_L$}
-In Maximum Likelihood framework, the topological recovery (reflected by the tree similarity) quickly drops with the increasing amount of missing data when using both metrics (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-However, this negative effect of missing data is stabilized beyond 25\% of missing data (no significant difference for trees with respectively 25\%, 50\% and 75\% missing living taxa).
-With 10\% of missing living taxa, there is a different topological recovery depending on the used metric.
-The Robinson-Fould metric shows a lower Tree similarity index than the Triples metric meaning that individual taxa position are generally more conserved (i.e.
-low number of flying taxa) but that the full clades are less conserved (see Table ~\ref{NTS_ML_results}).
-
-However, topological recovery was bad in Bayesian, regardless the amount of missing data.
-The tree similarity index was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance (see Table ~\ref{NTS_Ba_results}).
-Even if the topological recovery is really low, we detected an effect of missing data on the topology depending on the used metric.
-For the Triples metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
-For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases (see Table ~\ref{Ba_RF-ML_results}).
-However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the tree similarity index was only slightly higher than expected by chance (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-
-\subsubsection{Effect of $M_F$}
-In Maximum Likelihood framework, the effect of missing data in the fossil record appeared to be constant and led to a constant decrease in topological recovery when the missing data increased (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-The difference between the two metrics used is more important than for the $M_L$ parameter: clades are less conserved than individual taxa placement as the amount of missing data increases.
-Interestingly, 10\% or 25\% missing data for the $M_F$ parameter does not seems to affect the apparition of unstable taxa (Triples NTS modes of respectively 0.97 and 0.93 - Table ~\ref{NTS_ML_results}).
-
-Similarly than for the effect of $M_L$, Bayesian tree recovery was only slightly better than expected by chance.
-In the same way as described above, there seems to be a small difference depending on the metric used.
-There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric (see Table ~\ref{Ba_RF-MF_results}).
-However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
-
-\subsubsection{Effect of $M_C$}
-The number of missing morphological characters ($M_C$), however, seems to be more affecting topological recovery than $M_L$ and $M_F$ (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
-The Robinson-Fould metric shows a rapid drop in tree recovery from 10\% of missing data (Robinson-Fould NTS mode $<$ to 0.5 from 10\% missing data - Table ~\ref{NTS_ML_results}).
-This decrease is however slower when using the Triples metric with still a good topological recovery at 10\%  (Triples NTS modes = 0.95 - Table ~\ref{NTS_ML_results}).
-
-The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Fould metric shows a significant effect of the $M_C$ parameter on tree recovery (see Table ~\ref{Ba_RF-ML_results} and ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-
-%\subsubsection{Combined effect}
-%Need to do the heat plots
-
-$\newline$
-
-%To sum up:
-The ability of recovering the "best" tree's topology in Maximum Likelihood method is function of the amount of data missing.
-The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
-The decrease in clade conservation (low Robinson-Fould NTS score) is faster than the increase in flying taxa (low Triples NTS score).
-In Bayesian framework, topology is badly (if not at all) recovered disregarding the amount of data missing.
-
-\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
-\caption{Tree similarity values per parameter in ML framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
-              & missing data &        &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.671 & 0.569  & 0.789 & 0.468  & 1.036 \\ %how many digits?
-              &              & Triples        & 0.942 & 0.868  & 0.999 & 0.612  & 1.063 \\
-              & 25\%         & Robinson-Fould & 0.516 & 0.440  & 0.630 & 0.266  & 0.892 \\
-              &              & Triples        & 0.559 & 0.708  & 0.810 & 0.291  & 0.999 \\
-              & 50\%         & Robinson-Fould & 0.432 & 0.364  & 0.521 & 0.231  & 0.642 \\
-              &              & Triples        & 0.514 & 0.419  & 0.599 & 0.293  & 0.910 \\
-              & 75\%         & Robinson-Fould & 0.372 & 0.323  & 0.434 & 0.203  & 0.577 \\
-              &              & Triples        & 0.444 & 0.338  & 0.532 & 0.192  & 0.855 \\
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.907 & 0.766  & 1.000 & 0.407  & 1.047 \\
-              &              & Triples        & 0.976 & 0.944  & 1.003 & 0.715  & 1.104 \\
-              & 25\%         & Robinson-Fould & 0.572 & 0.488  & 0.692 & 0.325  & 0.936 \\
-              &              & Triples        & 0.937 & 0.820  & 0.993 & 0.539  & 1.086 \\
-              & 50\%         & Robinson-Fould & 0.433 & 0.302  & 0.476 & 0.170  & 0.668 \\
-              &              & Triples        & 0.732 & 0.594  & 0.873 & 0.341  & 1.025 \\
-              & 75\%         & Robinson-Fould & 0.153 & 0.115  & 0.200 & 0.068  & 0.323 \\
-              &              & Triples        & 0.458 & 0.390  & 0.611 & 0.290  & 0.856 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.473 & 0.384  & 0.835 & 0.356  & 0.958 \\
-              &              & Triples        & 0.959 & 0.869  & 1.026 & 0.572  & 1.067 \\
-              & 25\%         & Robinson-Fould & 0.454 & 0.357  & 0.630 & 0.242  & 0.916 \\
-              &              & Triples        & 0.907 & 0.751  & 0.997 & 0.446  & 1.070 \\
-              & 50\%         & Robinson-Fould & 0.424 & 0.375  & 0.466 & 0.249  & 0.828 \\
-              &              & Triples        & 0.683 & 0.571  & 0.822 & 0.277  & 1.008 \\
-              & 75\%         & Robinson-Fould & 0.230 & 0.177  & 0.302 & 0.108  & 0.567 \\
-              &              & Triples        & 0.451 & 0.373  & 0.572 & 0.224  & 0.791 \\
-    \hline
-\label{NTS_ML_results}
-\end{tabular}
-\end{table}
-
-\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
-\caption{Tree similarity values per parameter in Bayesian framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
-              & missing data &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\ %how many digits?
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.028   & 0.061 \\
-              &              & Triples        & 0.033 & 0.003  & 0.049 & -0.049  & 0.117 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.061 \\
-              &              & Triples        & 0.035 & 0.002  & 0.049 & -0.041  & 0.096 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.056 \\
-              &              & Triples        & 0.028 & 0.003  & 0.052 & -0.038  & 0.092 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.010   & 0.043 \\
-              &              & Triples        & -0.001 & -0.018 & 0.040 & -0.060 & 0.106 \\
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.035 & 0.031  & 0.056 & 0.027   & 0.076 \\
-              &              & Triples        & 0.026 & 0.002  & 0.050 & -0.046  & 0.105 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.0153  & 0.068 \\
-              &              & Triples        & 0.038 & 0.002  & 0.050 & -0.046  & 0.105 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.058 \\
-              &              & Triples        & 0.014 & -0.011 & 0.039 & -0.038  & 0.124 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.041 \\
-              &              & Triples        & 0.029 & 0.019  & 0.042 & -0.030  & 0.089 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.056 & 0.035  & 0.059 & 0.0281  & 0.076 \\
-              &              & Triples        & 0.030 & 0.011  & 0.048 & -0.024  & 0.094 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.023   & 0.076 \\
-              &              & Triples        & 0.020 & -0.001 & 0.037 & -0.060  & 0.109 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.030   & 0.059 \\
-              &              & Triples        & 0.010 & -0.005 & 0.045 & -0.036  & 0.104 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.011   & 0.043 \\
-              &              & Triples        & 0.024 & 0.001  & 0.042 & -0.037  & 0.104 \\
-    \hline
-\label{NTS_BA_results}
-\end{tabular}
-\end{table}
-
-\begin{table}
-\caption{Group difference per configuration set}
-\centering
-\begin{tabular}{rllcccc}
-    \hline
-    set & metric & framework & parametric & $stats^1$ & df & p.value \\
-    \hline
-    $M_L$         & Robinson-Fould & ML       & no & 151.438  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.001278e-31
-                  &                & Bayesian & no & 65.62    & 4   & \textbf{0.00} \\ 
-                  & Triples        & ML       & no & 142.8329 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains  - 6.983434e-30
-                  &                & Bayesian & no & 4.56     & 4   & 0.34          \\ 
-    $M_F$         & Robinson-Fould & ML       & no & 179.9238 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 7.743067e-38
-                  &                & Bayesian & no & 85.61    & 4   & \textbf{0.00} \\
-                  & Triples        & ML       & no & 150.059  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.977386e-31
-                  &                & Bayesian & no & 3.01     & 4   & 0.56          \\
-    $M_C$         & Robinson-Fould & ML       & no & 151.3818 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.029428e-31
-                  &                & Bayesian & no & 64.23    & 4   & \textbf{0.00} \\
-                  & Triples        & ML       & no & 138.374  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 6.290979e-29
-                  &                & Bayesian & no & 26.13    & 24  & 0.35          \\
-    $M_L+M_F$     & Robinson-Fould & ML       & no & 734.9213 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.104542e-139
-                  &                & Bayesian & no & 317.97   & 24  & \textbf{0.00} \\
-                  & Triples        & ML       & no & 531.5558 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 4.581065e-97
-                  &                & Bayesian & no & 3.01     & 24  & 0.56          \\
-    $M_L+M_C$     & Robinson-Fould & ML       & no & 542.3233 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 2.619808e-99
-                  &                & Bayesian & no & 290.10   & 24  & \textbf{0.00} \\
-                  & Triples        & ML       & no & 437.7413 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains -  1.28493e-77
-                  &                & Bayesian & no & 22.19    & 24  & 0.57          \\
-    $M_F+M_C$     & Robinson-Fould & ML       & no & 812.206  & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 5.463138e-156
-                  &                & Bayesian & no & 385.96   & 24  & \textbf{0.00} \\
-                  & Triples        & ML       & no & 528.9215 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.619383e-96
-                  &                & Bayesian & no & 20.23    & 24  & 0.68          \\ 
-    $M_L+M_F+M_C$ & Robinson-Fould & ML       & no & 3603.808 & 124 & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains -3.819574e-169
-                  &                & Bayesian & no & 1167.40  & 124 & \textbf{0.00} \\
-                  & Triples        & ML       & no & 2047.923 & 124 & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains 
-                  &                & Bayesian & no & 115.86   & 124 & 0.69          \\
-    \hline
-\label{Group_results}
-\end{tabular}
-
-$^1$ F-value for parametric tests and Kruskal Wallis $chi^2$ for non parametric tests.
-\end{table}
-
-\begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE  & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE  \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 56.53 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$25\% & 96.85 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 127.03 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 144.58 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 40.31 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$50\% & 70.50 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 88.05 & 37.66 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 30.19 & 37.66 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 47.73 & 37.66 & TRUE \\ 
-  $M_L$50\%-$M_L$75\% & 17.55 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_RF-ML_results}
-\end{table}
-
-\begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE  & FALSE \\
-    $M_L$50\% & & & & - & FALSE  \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 57.98 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$25\% & 105.83 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 120.86 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 140.34 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 47.85 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$50\% & 62.88 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 82.36 & 37.66 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 15.03 & 37.66 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 34.51 & 37.66 & FALSE \\ 
-  $M_L$50\%-$M_L$75\% & 19.48 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_Tr-ML_results}
-\end{table}
-
-\begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & FALSE & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 17.05 & 41.00 & FALSE \\ 
-  $M_L$00\%-$M_L$25\% & 43.79 & 41.00 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 77.32 & 41.00 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 103.99 & 41.00 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 26.75 & 41.00 & FALSE \\ 
-  $M_L$10\%-$M_L$50\% & 60.27 & 41.00 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 86.94 & 41.00 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 33.53 & 41.00 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 60.20 & 41.00 & TRUE \\ 
-  $M_L$50\%-$M_L$75\% & 26.67 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{Ba_RF-ML_results}
-\end{table}
-
-\begin{table}
-\caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & \textbf{TRUE} \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 59.27 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$25\% & 79.72 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$50\% & 120.70 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 167.81 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 20.45 & 37.66 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 61.43 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 108.55 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 40.98 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$75\% & 88.09 & 37.66 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 47.12 & 37.66 & TRUE \\ 
-   \hline
-\end{tabular}
-\label{ML_RF-MF_results}
-\end{table}
-
-\begin{table}
-\caption{$M_F$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 64.78 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$25\% & 89.09 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$50\% & 123.20 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 150.43 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 24.31 & 37.66 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 58.42 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 85.65 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 34.10 & 37.66 & FALSE \\ 
-  $M_F$25\%-$M_F$75\% & 61.34 & 37.66 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 27.23 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_Tr-MF_results}
-\end{table}
-
-\begin{table}
-\caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & FALSE & FALSE & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 16.03 & 41.00 & FALSE \\ 
-  $M_F$00\%-$M_F$25\% & 38.50 & 41.00 & FALSE \\ 
-  $M_F$00\%-$M_F$50\% & 61.52 & 41.00 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 95.52 & 41.00 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 22.47 & 41.00 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 45.49 & 41.00 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 79.49 & 41.00 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 23.02 & 41.00 & FALSE \\ 
-  $M_F$25\%-$M_F$75\% & 57.02 & 41.00 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 34.00 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{Ba_RF-MF_results}
-\end{table}
-
-\begin{table}
-\caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & \textbf{TRUE} \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 67.99 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$25\% & 87.58 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$50\% & 116.16 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 155.77 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 19.59 & 37.66 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 48.17 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 87.78 & 37.66 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 28.58 & 37.66 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 68.19 & 37.66 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 39.60 & 37.66 & TRUE \\ 
-   \hline
-\end{tabular}
-\label{ML_RF-MC_results}
-\end{table}
-
-\begin{table}
-\caption{$M_C$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & FALSE \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 75.29 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$25\% & 88.58 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$50\% & 114.02 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 149.60 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 13.29 & 37.66 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 38.73 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 74.31 & 37.66 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 25.44 & 37.66 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 61.02 & 37.66 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 35.58 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_Tr-MC_results}
-\end{table}
-
-\begin{table}
-\caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & FALSE & FALSE & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & FALSE \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 16.03 & 41.00 & FALSE \\ 
-  $M_C$00\%-$M_C$25\% & 38.50 & 41.00 & FALSE \\ 
-  $M_C$00\%-$M_C$50\% & 61.52 & 41.00 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 95.52 & 41.00 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 22.47 & 41.00 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 45.49 & 41.00 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 79.49 & 41.00 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 23.02 & 41.00 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 57.02 & 41.00 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 34.00 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{Ba_RF-MC_results}
-\end{table}
-
-%---------------------------------------------
-%
-%       DISCUSSION
-%
-%---------------------------------------------
-
-\section{Discussion}
-
-\subsection{Building the trees}
-Simulating evolutionary history matrices still remains a big drawback in theoretical phylogenetics. %Hahahaha.... feck.
-The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
-This configuration probably lead to globally low phylogenetic signal as well as the intrinsic difficulties to simulated characters with phylogenetic signal.
-Even though molecular characters evolution (and therefore simulation) only depends of a small number of parameters (i.e.
-the base frequencies and the substitution matrix), simulating molecular matrix with a strong evolutionary signal is still complicated when generating unrealistically small matrices.
-For morphological characters, the underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
-
-Also morphological characters studies involve many potential statistical pitfalls (e.g.
-independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy.
-%REF probably April as something about it
-\begin{enumerate}
-\item{First, morphological data can display a different signal than molecular data, especially in small matrices. %REF
-This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
-However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013}.}
-%The main difference between molecular and morphological characters is that, because of historical and practical reasons, a morphological character matrix do no contains autapomorphies and invariant characters which are usually more common than synampomorphies in molecular character matrices (cite).
-\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013}.
-However, following our simulation assumption of random character distributions, if they accumulate through time in the same way as the majority of the molecular characters then, homoplasic characters are expected to appear randomly through time \citep{davalosintegrating2014}.
-Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
-After a reaching a critical amount of morphological characters, adding new ones increases homoplasy \citep{wagner2000}.}
-\end{enumerate}
-Our simulation parameters both decrease the phylogenetic signal (difficulty to simulate morphological characters and size of the matrix) as well as it increases it (reduction of homoplasy in the morphological part of the matrix).
-Therefore, these drawbacks seems to have only a minor impact on the main results of this study: the incapacity of recovering any topology in Bayesian inference.
-
-\subsection{Comparing topologies}
-Comparing topologies is a crucial question in phylogenetics but has always been hard to normalise because of the vast amount of different metrics used as proxies for different aspects of tree similarity/dissimilarity \citep{agapowpower2002}.
-Because our global framework is to study how to include efficiently fossils into phylogenies, we chose metrics reflecting the most interesting aspect of this global question: where do individual taxa (i.e.
-the fossils) branch in the tree.
-The Robinson-Fould \citep{RF1981} and the Triples \citep{critchlowthe1996} metrics are more sensible to taxa and clade placement than other tree comparison metrics \citep[e.g.][Imbalance metric]{kirkpatricksearching1993} and where therefore favoured.
-Also by using the Normalised Tree Similarity index \citep{Bogdanowicz2012} we emphasize the aspect of "good" phylogenetic signal \textit{versus} random phylogenetic signal because it normalised the values of the metric by correcting for the expected value when comparing random trees (NTS = 0).
-
-The Robinson-Fould metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
-When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
-On the other hand, the Triples method is measuring the position of each taxon towards to other reference taxa \citep{critchlowthe1996}.
-It will penalise only trees where taxa get removed furthest from their original clade.
-Regarding our problematic (how does missing data influence topological recovery in trees containing both living and fossil taxa) we are more interested the placement of taxa (i.e.
-where does the fossil branch) than the exact conservation of clades.
-
-%Need a transition!
-
-Although the idea of comparing two single trees is straightforward, it becomes more complex when comparing trees distributions (e.g. Bayesian posterior distributions).
-The introduction of our Random Pairwise Bayesian Tree Comparison methods allows to summarize the comparison of two trees distributions by picking up the most frequent signal in the distribution of the pairwise comparisons (i.e.
-the mode).
-Even though this method is subject to randomness and might artificially increase or decrease the score of the studied metric, simulations showed that when a sufficient amount of random comparisons is performed, this method doesn't seem to be subject to randomness (e.g.
-1000 random pairwise comparisons - see ~\ref{supplementaries}).
-
-\subsection{Maximum Likelihood versus Bayesian}
-The main results of our analysis shows that Bayesian inference fails to recover any Topology in a total evidence method framework (Fig. ~\ref{Fig_Results}).
-This results is surprising regarding the behaviour of Maximum Likelihood inference (Fig.
-~\ref{Fig_Results}) as well as the Bayesian inferences performed on empirical data \citep{ronquista2012,schragocombining2013}.
-However, it is important to note that this effect was not mentioned in the aforementioned empirical tests because both studies used a Bayesian approach with fixed topology \citep{ronquista2012,schragocombining2013}.
-
-In our case, we suspect that this inability of Bayesian methods to recover topology is due to the intrinsic structure of a total evidence matrix (Fig ~\ref{Fig_RemoveData}).
-In fact, as previously studies have shown, missing data doesn't seem to be a major drawback as long as the missing data is randomly distributed ~\citep[e.g.][]{wiensmissing2003,rouresite-specific2011,sansomfossilization2013}.
-However, in total evidence matrices, the majority of the missing data is not randomly distributed but concentrated in the molecular part of the matrix for the fossil taxa (i.e.
-the missing-data sub-matrix, Fig ~\ref{Fig_RemoveData}).
-This leads to high decrease of topological recovery when using non optimal criterion approach (i.e.
-Bayesian inference) because of the high variance in the near-likeliest solutions sampled in the Bayesian posterior distribution.
-In opposition, when applying optimal criterion approach (i.e.
-Maximum Likelihood), it is still possible to sample the likeliest tree.
-%probably need to develop that way more, especially with actually analysing the likelihood landscape.
-
-\subsection{Effect of missing data}
-The three parameters we selected in this study account for three potential pitfalls in collecting the data for total evidence analysis:
-$M_L$ represents the living taxa for which there is no available morphological data,
-$M_F$ represents the quality of the fossil record,
-and $M_C$ represents the general coding effort and the overall of morphological data available.
-Ideally, the lowest possible amount of missing data is wanted in any phylogenetic analysis leading to a data collection part prior to the phylogenetic analysis.
-Each of the three parameters can be improved in a different way prior to the analysis:
-for $M_L$, one should put more effort in using natural museums history collections for coding the missing morphological data for living taxa if possible %go to the museum!;
-for the $M_F$ parameter, the amount of missing data unfortunately depends on the quality of the fossil record and can not be actively improved and depends on exceptional discoveries \citep[e.g.][]{nithe2013};
-finally for the $M_C$ parameter, improvement can be done by vast collaborative projects in order to gather as much characters as possible \citep[e.g.][]{O'Leary08022013}.
-
-Our results shows that the $M_F$ parameter have less influence on recovering the good tree topology in Maximum Likelihood framework which is fortunate since it is the parameter that is the more difficult to fix for practical reasons.
-Regarding the fact that we are more interested in taxa placement than in clade conservation (Triples \textit{versus} Robinson-Fould), the parameter that affects the more the decrease in topological recovery is the number of missing living taxa in the morphological part of the matrix ($M_L$) and the overall number of morphological characters $M_C$.
-This makes sense since the living taxa are bearing both the information to build the tree backbone (the molecular data) and the information used for branching the fossil on this backbone (the morphological information).
-Therefore, we advocate the importance of coding morphological characters for the most living taxa possible and with the most characters possible, potentially by using collaborative projects portals such as morphobank \citep{morphobank}.
-
-%Potential reviewers comments?
-%What about it when you vary the proportion of fossil taxa?
-
-%---------------------------------------------
-%
-%       CONCLUSION
-%
-%---------------------------------------------
-
-\section{Conclusion}
-A Bayesian approach fails to recover accurate topology in a total evidence approach framework, whatever the amount of missing data.
-We think this failure is due to the intrinsic structure of the total evidence matrices that includes a vast amount of non randomly distributed missing data (i.e.
-the molecular part of the matrix for the fossil taxa).
-This missing data is equal to the number of fossil taxa $\times$ the number of molecular characters leading to a vast amount of trees to be sampled in Bayesian posterior distribution.
-However, one can use a Maximum Likelihood approach to fix the likeliest topology.
-If so, an effort should be made prior to the phylogenetic analysis on collecting as much morphological data as their is available from living taxa in order to efficiently improve the quality of the trees.
-
-%Perspectives:
-%What the effect of this missing molecular data on branch length estimation?
-
-%---------------------------------------------
-%
-%       ...
-%
-%---------------------------------------------
-
-\section{Acknowledgements}
-We would like to thank Trevor Hodkinson and Andrew Jackson %Also Gavin Thomas, Emmanuel Douzery and Frédéric Delsuc if not authors.
-for his useful comments on the simulation protocol and Paddy Doyle for the assistance on using the computer cluster.
-All calculations were performed on the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing.
-This cluster was funded through grants from Science Foundation Ireland.
-
-%BIBLIOGRAPHY
- % The \cite command functions as follows:
- %   \citet{key} ==>>                Jones et al. (1990)
- %   \citet*{key} ==>>               Jones, Baker, and Smith (1990)
- %   \citep{key} ==>>                (Jones et al., 1990)
- %   \citep*{key} ==>>               (Jones, Baker, and Smith, 1990)
- %   \citep[chap. 2]{key} ==>>       (Jones et al., 1990, chap. 2)
- %   \citep[e.g.][]{key} ==>>        (e.g. Jones et al., 1990)
- %   \citep[e.g.][p. 32]{key} ==>>   (e.g. Jones et al., p. 32)
- %   \citeauthor{key} ==>>           Jones et al.
- %   \citeauthor*{key} ==>>          Jones, Baker, and Smith
- %   \citeyear{key} ==>>             1990
-
-\bibliographystyle{sysbio} %don't write the suffix
-\bibliography{References} %don't write the suffix
-
-\section{Supplementaries}
-\label{supplementaries}
 \subsection{Morphological characters states}
@@
 \centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-AppendixCharacters.pdf}
+\includegraphics[keepaspectratio=true]{SupplementaryMaterial/Supp_Figures/TEM_Fig-AppendixCharacters.pdf}
 \caption{Character states distribution in empirical matrices. %Title?
@@
 
-\subsubsection{Maximum Likelihood - RAxML v8.0.20 \citep{Stamatakis21012014}} \\
+\subsubsection{Maximum Likelihood - RAxML v8.0.20 \citep{Stamatakis21012014}}
+
 Model: \\
@@
 
-\subsubsection{Bayesian - MrBayes v3.0.2 \citep{Ronquist2012mrbayes}} \\
+\subsubsection{Bayesian - MrBayes v3.0.2 \citep{Ronquist2012mrbayes}}
+
 Priors: \\
@@
 Burnin = 25\%
-
-\subsection{Triplets metric details ($T_{x,y}$)}
-Each triplet can be written as $I_{ijk}$=(\textit{ijk})). Where $I_{ijk}$ is equal to 0 if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to 1.
-For any rooted tree there are only four possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); and (\textit{i},\textit{j},\textit{k}); \citep{johnson1998}.
-One can calculate $S_n$, the triplet distance between two trees as:
-\begin{equation}
-S_n = \sum_{ijk} I_{ijk}
-\end{equation}
-Where:
-\begin{equation}
-\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
-\end{equation}
-And where n is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
-When all triplets across the two trees are the same, $S_n$ is equal to 0 and when all the triplets are different $S_n$ is equal to $\binom{n}{4}$.
-Because the possible number of triplets per clade is a finite number, the probability of two random trees with the same n taxa to have the same triplet is:
-\begin{equation}
-P({I_{ijk}}=0) = \frac{1}{4}
-\end{equation}
-Therefore one can calculate the probability of two random trees having the same triplets: 
-\begin{equation}
-P({S_{n}}=0) = \sum_{ijk} P_{I_{ijk}=0}
-\end{equation}
-\begin{equation}
-P({S_{n}}=0) = \frac{n!}{4(3!(n-3)!}
-\end{equation}
-And in the same way:
-\begin{equation}
-P({S_{n}}=1) = \frac{3n!}{4(3!(n-3)!}
-\end{equation}
-
-\subsection{RF metric details}
-The RF distance (or path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
-\begin{equation}
-RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
-\end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees. 
-The minimal value of \textit{C} is equal to 1 if the two trees have the same n taxa;
-the maximal value in \textit{C}=\textit{n}-2.
-For a fully unresolved tree (star tree) \textit{N}=1 and for a fully resolved tree (binary tree) \textit{N}=\textit{n}-2.
-The minimal and maximal topological distance for \textit{n} taxa is:
-\begin{equation}
-RF_{min} = 1 + 1 - 2C_{x,y}
-\end{equation}
-And:
-\begin{equation}
-RF_{max} = 2(n-2)-2
-\end{equation}
-One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any \textit{n} taxa:
-\begin{equation}
-RF.scaled_{x,y} = \frac{RF_{x,y}-RF_{max}}{RF_{max}}
-\end{equation}
-This metric is more sensitive to taxa displacement than the Triplet distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.
-
-\subsection{Tree comparisons}
-\subsubsection{Random tree comparison scaling}
-We used the comparison of 1000 random trees to obtain the mean comparison value $\bar{d}_{m,n}$\textit{(rand)} for the NTS metric.
-We randomly generated two sets of 1000 trees of \textit{n} taxa using the rmtree function of ape package (v3.0-11 \citet{paradisape:2004}) that generates a given number of random Yule trees.
-We calculated the $\bar{d}_{m,n}$\textit{(rand)} value using an approach similar to the RPCBTC (described below) by performing 1000 random pairwise comparisons using the TreeCmp java script \citep{Bogdanowicz2012}.
-
-\subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
-We assessed the power of the Random Pairwise Bayesian Tree Comparison (RPBTC) method by comparing 1000 random trees from a posterior distribution trees set to another 1000 random trees from the same posterior distribution trees set.
-We repeated this 100 times independently using the same posterior distribution trees set each time resulting in 100 replicates of the same posterior distribution trees set compared 1000 times.
-We used an anova to test if there was no significant difference between the replicates so that the RBTC can be replicated.
-We applied this protocol on a poorly resolved tree (Low Score), a resolved tree with low support value (Medium Score) and a resolved tree with high support values (High Score).
-Results are available in table ~\ref{RPBTC_testing}. %link broken
-
-\begin{table}[ht]
-\caption{Group comparison results: difference between 100 replicates using the RPBTC method} %test with TreeCmp.anova
-\centering
-\begin{tabular}{rllrrrr}
-  \hline
-  Tree.Type & Used.metric & Replicates & Df & F.value & p.value \\ 
-  \hline
-  Low Score & RF & 100.00 & 99.00 & 0.74 & 0.98 \\ 
-  Low Score & Tr & 100.00 & 99.00 & 0.97 & 0.58 \\ 
-  Medium Score & RF & 100.00 & 99.00 & 0.64 & 1.00 \\ 
-  Medium Score & Tr & 100.00 & 99.00 & 0.45 & 1.00 \\ 
-  High Score & RF & 100.00 & 99.00 & 0.20 & 1.00 \\ 
-  High Score & Tr & 100.00 & 99.00 & 0.37 & 1.00 \\ 
-  \hline
-\end{tabular}
-\end{table}
-\label{RPBTC_testing}
-
-\subsection{Codes}
-All codes are available at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions$
-The tree comparison results analysis can be repeated for more details at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Analysis$ %fully running!
-Try it at home (5-10 minutes of data loading)!
-
-\subsection{Full results}
-
-\begin{table}
-\caption{Tree similarity values per parameter in ML framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
-              & missing data &        &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.6714878 & 0.5692308  & 0.7894649 & 0.4689069  & 1.0361514 \\ %how many digits?
-              &              & Triples        & 0.9424133 & 0.8683742  & 0.9998559 & 0.6121399  & 1.0635250 \\
-              & 25\%         & Robinson-Fould & 0.5167595 & 0.4403147  & 0.6307692 & 0.2664429  & 0.8929104 \\
-              &              & Triples        & 0.5599494 & 0.7084874  & 0.8105781 & 0.2912950  & 0.9990703 \\
-              & 50\%         & Robinson-Fould & 0.4321852 & 0.3641026  & 0.5215627 & 0.2313246  & 0.6423846 \\
-              &              & Triples        & 0.5143009 & 0.4197127  & 0.5991498 & 0.2991432  & 0.9104526 \\
-              & 75\%         & Robinson-Fould & 0.3727090 & 0.3230769  & 0.4342249 & 0.2035474  & 0.5776286 \\
-              &              & Triples        & 0.4449563 & 0.3385128  & 0.5329149 & 0.1929395  & 0.8551570 \\
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.9072419 & 0.7668415  & 1.0000000 & 0.4070844  & 1.0478010 \\
-              &              & Triples        & 0.9763617 & 0.9448100  & 1.0039460 & 0.7158118  & 1.1043330 \\
-              & 25\%         & Robinson-Fould & 0.5722382 & 0.4883560  & 0.6923077 & 0.3253239  & 0.9367467 \\
-              &              & Triples        & 0.9370152 & 0.8207407  & 0.9938846 & 0.5394443  & 1.0868073 \\
-              & 50\%         & Robinson-Fould & 0.4338159 & 0.3025641  & 0.4760777 & 0.1704016  & 0.6683220 \\
-              &              & Triples        & 0.7327464 & 0.5947211  & 0.8739630 & 0.3418679  & 1.0251540 \\
-              & 75\%         & Robinson-Fould & 0.1534487 & 0.1155128  & 0.2000000 & 0.0683485  & 0.3230769 \\
-              &              & Triples        & 0.4588292 & 0.3905631  & 0.6111484 & 0.2903676  & 0.8568274 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.473285 & 0.3842965  & 0.8358974 & 0.3564753  & 0.9589744 \\
-              &              & Triples        & 0.9592557 & 0.8691580  & 1.026095 & 0.5720847  & 1.067524 \\
-              & 25\%         & Robinson-Fould & 0.4548038 & 0.3572132  & 0.6307692 & 0.2425339  & 0.9161212 \\
-              &              & Triples        & 0.9071925 & 0.7516934  & 0.9976224 & 0.4462168  & 1.0705291 \\
-              & 50\%         & Robinson-Fould & 0.4244263 & 0.3755493  & 0.4666667 & 0.2494927  & 0.8287312 \\
-              &              & Triples        & 0.6834366 & 0.5713927  & 0.8226139 & 0.2776690  & 1.0085944 \\
-              & 75\%         & Robinson-Fould & 0.2306888 & 0.1774508  & 0.3025641 & 0.1087764  & 0.5673541 \\
-              &              & Triples        & 0.4513642 & 0.3737397  & 0.5723858 & 0.2245224  & 0.7917946 \\
-    \hline
-\label{NTSML_full}
-\end{tabular}
-\end{table}
-
-\begin{table}
-\caption{Tree similarity values per parameter in Bayesian framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
-              & missing data &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\ %how many digits?
-              &              & Triples        & 0.04104231 & 0.00830962  & 0.05441173 & -0.03933377  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.03584309 & 03584196  & 03589744 & 02825643  & 0.06175676 \\
-              &              & Triples        & 0.03399849 & 0.003196469  & 0.04981018 & -0.049913382  & 0.11784366 \\
-              & 25\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01536739  & 0.06123106 \\
-              &              & Triples        & 0.03510953 & 0.002113358  & 0.04939228 & -0.041202264  & 0.09699332 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01538462  & 0.05641026 \\
-              &              & Triples        & 0.02849433 & 0.003263098  & 0.05298036 & -0.038344731  & 0.09250063 \\
-              & 75\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01069992  & 0.04338811 \\
-              &              & Triples        & -0.0001940566 & -0.01801301  & 0.04058783 & -0.06030922  & 0.10665934 \\
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\
-              &              & Triples        & 0.04104231 & 0.008309622  & 0.05441173 & -0.039333778  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.03593712 & 0.03198256  & 0.05641026 & 0.02798864  & 0.07692308 \\
-              &              & Triples        & 0.02639309 & 0.002185408  & 0.05061314 & -0.046563042  & 0.10547052 \\
-              & 25\%         & Robinson-Fould & 0.03588134 & 0.03587629  & 0.03589744 & 0.01538462  & 0.0680884 \\
-              &              & Triples        & 0.03856296 & 0.002185408  & 0.05061314 & -0.046563042  & 0.10547052 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01536289  & 0.05879269 \\
-              &              & Triples        & 0.01403716 & -0.01100608  & 0.03972324 & -0.03800324  & 0.12405207 \\
-              & 75\%         & Robinson-Fould & 0.03589206 & 0.03588825  & 0.03589744 & 0.01538462  & 0.0411843 \\
-              &              & Triples        & 0.02939363 & 0.01902351  & 0.04238907 & -0.03087905  & 0.08946917 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\
-              &              & Triples        & 0.04104231 & 0.008309622  & 0.05441173 & -0.039333778  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.05637057 & 0.03589744  & 0.05927956 & 0.02813296  & 0.07692308 \\
-              &              & Triples        & 0.03008132 & 0.01176800  & 0.04847400 & -0.02451461  & 0.09453754 \\
-              & 25\%         & Robinson-Fould & 0.03593743 & 0.03589744  & 0.03596018 & 0.02330790  & 0.07692308 \\
-              &              & Triples        & 0.02010221 & -0.001145118  & 0.0377779 & -0.060183332  & 0.1098632 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.03050289  & 0.05950426 \\
-              &              & Triples        & 0.01053959 & -0.005595947  & 0.04510444 & -0.036278820  & 0.10478805 \\
-              & 75\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01115983  & 0.04362927 \\
-              &              & Triples        & 0.02427778 & 0.0007405737  & 0.04260522 & -0.0373004106  & 0.1048756 \\
-\label{NTSBA_full}
-\end{tabular}
-\end{table}
-
-\subsection{Bootstraps distribution}
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-AppendixBootstraps.pdf}
-\caption{Bootstraps distribution across the "best" trees.}                     % MISSING 8 CHAINS
-\label{Fig_AppendixBoostraps}
-\end{figure}
-
-%END
-\end{document}

### 2014-06-17 Manuscript/SupplementaryMaterial/TreeComparison.tex
Supplementary Materials init

--- 2014-06-13 Manuscript/TEM_draft.tex
@@
-%HEADER
-\documentclass[12pt,letterpaper]{article}
-
-%packages
-% NC: Just to flag here how I'll comment on LaTeX files. I'll use % and then NC. Of course in git diff it'll all show up as an addition.
-% NC: You could add all the packages etc to a template .sty file to save writing it all up here. Rich's ms.sty that I provided on DropBox is a good example
-\usepackage{fixltx2e}
-\usepackage{textcomp}
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{verbatim}
-\usepackage[english]{babel}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{setspace}
-\usepackage{lscape}
-\usepackage{indentfirst}
-\usepackage[normalem]{ulem}
-\usepackage{booktabs}
-%\usepackage{nag}
-\usepackage{natbib}
-%\usepackage{bibtex}
-\usepackage{float}
-\usepackage{latexsym}
-%\usepackage{hyperref} 
-\usepackage{url}
-%\usepackage{html}
-\usepackage{hyperref}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{bm}
-\usepackage{array}
-\usepackage[version=3]{mhchem}
-\usepackage{ifthen}
-\usepackage{caption}
-\usepackage{hyperref}
-%\usepackage{xcolor}
-\usepackage{amsthm}
-\usepackage{amstext} % Add and remove packages as necessary for your manuscript.
-
-% NC: Added by NC
-\usepackage{enumerate}
-\usepackage[osf]{mathpazo} % palatino is a nicer font! 
-\pagenumbering{arabic}	% Page numbers
-
-%Pagination style and stuff
-\linespread{1.66} % All text should be double-spaced with occasional exceptions for tables. 
-\raggedright
-\setlength{\parindent}{0.5in}
-
-\setcounter{secnumdepth}{0} % Our sections are not numbered and our papers do not have Tables of Contents.
-We don't present a list of figures or list of tables, either.
-Any common font is fine.
-(A common sans-serif font should be used on figures, but figures should be separate from the LaTeX document.)
-
-%\pagestyle{empty}
-
-\renewcommand{\section}[1]{%
-\bigskip
-\begin{center}
-\begin{Large}
-\normalfont\scshape #1
-\medskip
-\end{Large}
-\end{center}}
-
-\renewcommand{\subsection}[1]{%
-\bigskip
-\begin{center}
-\begin{large}
-\normalfont\itshape #1
-\end{large}
-\end{center}}
-
-\renewcommand{\subsubsection}[1]{%
-\vspace{2ex}
-\noindent
-\textit{#1.}---}
-
-\renewcommand{\tableofcontents}{}
-
-\bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
-% We don't use a special title page; the author information is entered like any other text.
-% FOOTNOTES: We don't allow them in the manuscript, except in tables. Don't include any footnotes in the text.
-%IMPORTANT:
-%Note on display in the text file: sentences or sub-parts of sentences (e.g. enumerations) are delimited by a newline for clarity.
-%However, true new lines (i.e. in the pdf) are delimited by one or more empty lines.
-%We don't use a heading for the introduction. The nature of this section is implied by its position following the abstract and preceding the first major section heading.
-
-% NC: You should delete this extraneous info from the Syst Biol template to tidy everything up a bit
-
-\begin{document}
-
-%For a colour version of the draft to a search>replace '-BW'>'' 
-% %\modulolinenumbers[1] 	% Line numbering on every line
-
-%Running head
-\begin{flushright}
-Version dated: \today
-\end{flushright}
-\bigskip
-\noindent RH: Missing data and topology in total evidence matrices %RH = Running Head
-
-\bigskip
-\medskip
-\begin{center}
-
-%Title
-\noindent{\Large \bf Effect of missing data in matrices containing living and fossil taxa on topological accuracy}
-%Alternative punchier title: "Bayesian total evidence methods fail to recover correct topologies from matrices containing both living and fossil taxa"
-%Alternative punchier title: "Maximum Likelihood outperforms Bayesian methods for inferring topology from total evidence matrices containing both living and fossil taxa"
-
-% NC: Not sure about any of these.
-I actually think I prefer your talk title.
-So something like: The effects of missing data on the ability of total evidence approaches to recover correct tree topologies.
-No rush to decide.
-But needs total evidence in there somewhere.
-\bigskip
-
-%Authors
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$, other authors $^3$, and Natalie Cooper$^1$$^,$$^2$}\\
-\noindent {\small \it 
-$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland;\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland;\\
-$^3$Somewhere else }\\
-\end{center}
-\medskip
-\noindent{\bf Corresponding author:} Thomas Guillerme, School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.\\
-\vspace{1in}
-% NC: It's not called the Republic of, just Ireland!
-
-\newpage
-\begin{abstract}
-% NC: I'm not going to give this much work as it'll change before we submit anyway. Note though that and abstract must NOT just repeat stuff in the intro. Needs rephrasing at the very least.
-Living species represent less than 1\% of all species that have ever lived.
-% NC: I was chatting to Andy Purvis about this the other day and he said lots of palaeo folk hate that stat.
-So let's not start the abstract with it!
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology.
-This fact has led to an increasing consensus among scientists that both fossil and living taxa must be included in macroevolutionary studies.
-One approach, the total evidence approach, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both fossil and living taxa at the tips.
-Although the total evidence approach seems very promising, it requires a lot of data and is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.
-
-In this study we assess the effect of missing data on tree topologies inferred from total evidence matrices.
-Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix:
-(1) the proportion of living taxa with no morphological data,
-(2) the amount of missing data in the fossil record, and
-(3) the overall number of morphological characters in the matrix.
-We find that, in a Bayesian framework, difficulties in recovering a stable topology are mainly driven by the missing data in the molecular part of the matrix (for which fossil taxa have no data).
-In a Maximum Likelihood framework, however, topology is not directly affected by missing data \textit{per se}, but by the number of morphological characters shared among the taxa.
-Therefore, the two main drivers of incorrect topologies are the overall number of morphological characters and the number of living species with no morphological data.
-
-Our results suggest that, in order to use total evidence approaches, one should reduce the missing data in the morphological part of the matrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process.
-%More results!
-\end{abstract}
-
-\noindent (Keywords: missing data, total evidence, Bayesian, Maximum Likelihood, topology)\\
-
-\vspace{1.5in}
-
-\newpage
-
-\section{Introduction}
-
-Although most species that have ever lived are now extinct % NC: This avoids giving the dodgy stat of 1%
- \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species (e.g., \citealp{meredithimpacts2011}).
-%or is citing of your own papers in the too "douchebagy" for the second sentence?
-% NC: Yeah be careful with self citing. Fine if it's a good fit, but it's totally not here!!! A better citation would be a big biodiversity study, e.g. the bird phylogeny paper.
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
-This has led to increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments (e.g.
-\citealp{pyrondivergence2011,ronquista2012,schragocombining2013}).
-%Not a nice sentence
-
-Up to now, three main approaches have been used to place both living and fossil taxa % NC: BE consistent. Either living and fossil, or fossil and living.
-into phylogenies.
-These approaches differ mainly in whether they treat fossil taxa as tips or as nodes in the phylogeny, and in which part of the available fossil data is used (i.e.
-the age of the fossil only or both its age and morphology).
-% NC: Not sure about bit in brackets
-Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
-Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \citep{simpson1945}.
-This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty (but see \citealp{spencerefficacy2013}).
-Neontologists, on the other hand, more commonly use probabilistic approaches (e.g.
-Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates % NC: Need to check the exact phrase needed here.
-Is it range of occurrence?
-Occurrence dates?
-etc.
- are used to time calibrate phylogenies \citep{zuckerkandl1965}.
-There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013}.
-However neither approach uses all the available data.
-
-A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa.
-%There might be a philosophico-semantic distinction to do here.
-Is TEM a method or an approach?
-I'll say that it is an approach because it's more like an idea (put every thing together) than a way to do (how to put everything together).
-It makes more sense to me compared to a real method e.g.
-a probabilistic method such as ML where the method states precisely how to do stuff.
-% NC: Tricky. Don't think it's too important though. 
-% NC: I do think Total Evidence needs capital letters though, like ML and Bayesian
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{eernissetaxonomic1993}.
-Total Evidence methods have been successfully applied to empirical data \citep{pyrondivergence2011,ronquista2012,schragocombining2013}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
-However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data.
-In particular it requires morphological data from both living and fossil taxa, both of which are known to be scarce.
-% NC: Really?
-I know fossil data is known to be scarce, but living data?
-I think people would be surprised.
-And they are surprised.
-We were surprised!
-Needs a citation either way.
-Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
-% Trevor: Missing data also influences support values that trend to decrease
-
-% NC: You need to take a bit of time to learn about singular and plural grammar in English.
-It's something you always do wrong.
-e.g.
-The Total Evidence method *uses* this and *its* parameters are as follows.
-Versus: Total Evidence methods *use* this and *their* parameters are as follows.
-Sive and I won't be around to correct your English forever.
-
-The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013}.
-Missing molecular data has been seen by some authors as an issue because it can decrease phylogenetic signal (i.e. % define phylogenetic signal here) 
-in some parts of the tree, especially when using large matrices \citep{lemmonthe2009}. However, this may not be a major issue because phylogenetic signal is easily increased by:
-(i) including a "modest" number of highly-covered % what do you mean by highly-covered?
-genes (i.e. approximatively half of the genes; \citealt{rouresite-specific2011});
-(ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and
-(iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
-Similarly, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study in question \citep{wiensmissing2003,sansomfossilization2013}.
-Because soft-tissue characters are rarely preserved in the fossil record, missing data is mainly found in these characters, and is therefore not randomly distributed which can lead to biased placement of fossil taxa in phylogenies \citep{sansomfossilization2013}.
-However, the phylogenetic signal is not related to the amount of missing data \textit{per se} but to the number of informative characters for each taxon, therefore missing data is less of an issue than the number of shared informative characters \citep{wiensmissing2003}.
-% NC: I'm a little dubious of what this paragraph is trying to get at in places. We should discuss at some point.
-
-Although missing data does not appear be a major problem in molecular and morphological matrices separately \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil species.
-This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data.
-Until now, no attempt has been made to study the impact of this issue on phylogenetic inference from Total Evidence methods.
-
-% NC: Somewhere here you need to flag that you're only looking at topologies and briefly explain why.
-
-Here we use simulations to assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
-The molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}.
-The effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}, therefore, we focus only on missing data in the morphological part of the matrix.
-We investigate three major parameters that directly affect the completeness of the morphological part of the matrix:
-\begin{enumerate}
-\item the proportion of living taxa with no morphological data;
-\item the amount of missing data in the fossil taxa; and
-\item the overall number of morphological characters for both living and fossil taxa in the matrix.
-\end{enumerate}
-We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
-We chose these parameters because they reflect real-life biases in data availability.
-The advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
-Missing data in fossil taxa is very common due to preservation biases (citation?), and
-% OK I'm not sure what the "real-life" reason for three is.
-%This paragraph is crap
-% NC: I think it's fine!
-
-We find that when using a Maximum Likelihood approach, as missing data increases, the likelihood of recovering the correct tree topology decreases.
-However, even with no missing data, Total Evidence matrices dramatically reduce the performance of Bayesian methods for inferring tree topology.
-We propose that this drastic difference between Bayesian and Maximum Likelihood methods is due to a flattening of the likelihood landscape caused by the unavoidable amount of missing molecular data for fossil taxa in a Total Evidence matrix.
-We make suggestions for how best to deal with this issue when inferring phylogenies from Total Evidence matrices.
-
-% NC: You don't want to tell the whole story here. Just a results summary and a quick line of interpretation
-
-%This problem leads to an inability of Bayesian chains to pick up the sub and near optimal topologies leading to high variance in the posterior tree distribution and really low support in the consensus tree.
-%These results should encourage authors in not using Bayesian inference to explore the topological landscape of their phylogeny.
-%Instead, we suggest that authors should first find the Maximum Likelihood tree and use it's topology as a fix prior in their Bayesian inference.
-%We also propose a new method for comparing Bayesian posterior tree distribution's topology available in a wonderful R package. %Or am I just dreaming a bit too much here?
-
-% NC: Dreaming a bit much I think.
-There's a discussion to be had here about whether it's really a *new* method and whether you really want to make an R package per se.
-We definitely want to share the code
-
-%-----------
-% NC: It might look like I've done a lot to this but actually I think it's nearly there.
-A good thing to take from this is that writing a paper is *hard*.
-It gets easier, but it's still a deeply involved process with lots of writing and re-writing.
-Introductions are particularly difficult because if you don't set the scene, it's hard to convince people to get on board for the rest of the paper.
-This is why I said there was no way this was being submitted by the end of June!
-Still a long way to go I'm afraid in terms of time and effort.
-But that doesn't mean you're not doing well, *every* project is like this.
-2 years is about average from conception to sending out a manuscript, and you only started work on this last spring.
-%------------
-
-\newpage
-\section{Methods} %no material needed?
-% NC: Depends on the journal. See what Syst Biol asks for as section titles.
-To explore how missing data in Total Evidence matrices influences tree topology we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
-%NC: does Syst Biol use Fig.
-or Figure?
-\begin{enumerate}
-\item{Generating the matrix} \label{step:generate_matrix} \\
-We randomly generated a birth-death tree (hereafter called the "true" tree; Table ~\ref{Tab_glossary}) and used it to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix; Table ~\ref{Tab_glossary}).
-\item{Removing data} \label{step:remove_data} \\
-We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of missing living taxa ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$) % NC: make sure to use the same descriptions here and in the introduction
-(the resulting matrices are called hereafter "missing-data" matrices; Table ~\ref{Tab_glossary}).
-\item{Building phylogenies} \label{step:build_phylo} \\
-We built phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree; Table ~\ref{Tab_glossary}) and multiple trees inferred from matrices with missing morphological data (hereafter called the "missing-data" trees; Table ~\ref{Tab_glossary}).
-Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
-\item{Comparing topologies} \label{step:compare_topo} \\
-We compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies.
-\end{enumerate}
-To measure the effect of missing data distribution, we repeated steps ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} with the exact same fixed parameters 51 (3$\times$17) times.
-%that's because I had 3 cluster accounts available...
-If it's a too odd number, no worries, I'll just remove the last chain.
-A list of all the terms used in this paper is available in table ~\ref{Tab_glossary}.
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_outline-BW.pdf}
-\caption{Protocol outline.
-(1) We generated a random tree (the "true" tree) to infer a matrix with no missing data (the "complete" matrix).
-(2) We removed data from the morphological partition of the "complete" matrix resulting in 125 "missing-data" matrices.
-(3) We infer phylogenetic trees from each matrix in both ML and Bayesian framework.
-(4) We then compared the "missing-data" trees to the "best" tree.
-We repeated step ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} 51 (3$\times $17) times.}
-\label{Fig_Outline}
-\end{figure}
-
-\begin{table}[ht]
-\caption{Glossary}
-\centering
-\begin{tabular}{p{5cm} p{9cm}}
-\hline
-\hfill Term & Definition \\[0.5ex]
-\hline
-\hfill living taxa & taxa with both molecular and morphological data available \\[1.5ex]
-\hfill fossil taxa & taxa with only morphological data available \\[1.5ex]
-\hfill "complete" matrix & matrix with no missing data except for the molecular part of the fossil taxa \\[1.5ex]
-\hfill "missing-data" matrix & matrix with various amount of missing data \\[1.5ex]
-\hfill $M_{L}$ & missing living taxa in the morphological part of the matrix \\[1.5ex]
-\hfill $M_{F}$ & missing data for the fossil taxa in morphological part of the matrix \\[1.5ex]
-\hfill $M_{C}$ & missing morphological characters for both living and fossil taxa \\[1.5ex]
-\hfill "true" tree & tree used to simulate the matrix \\[1.5ex]
-\hfill "best" tree & tree inferred from the "complete" matrix \\[1.5ex]
-\hfill "missing-data" tree & tree inferred the "missing-data" matrices \\[1.5ex]
-\hfill RPBTC & Random pairwise Bayesian tree comparison \\[1.5ex]
-\hfill "horizontal" sub-matrix & sub-matrix with morphological and molecular characters for living taxa \\[1.5ex]
-\hfill "vertical" sub-matrix & sub-matrix with morphological data for both living and fossil taxa \\[1.5ex]
-\hfill "corner" sub-matrix & sub-matrix with morphological data for living taxa \\[1.5ex]
-\hfill "missing-data" sub-matrix & molecular part of the matrix for the fossil taxa (no data) \\
-\end{tabular}
-\label{Tab_glossary}
-\end{table}
-
-%---------------------------------------------
-%       Generating the matrix
-%---------------------------------------------
-
-\subsection{Generating the matrix}
-First we randomly generated a "true" tree of 50 taxa in R v3.0.2 \citep{R302} using the package diversitree v0.9-6 \citep{fitzjohndiversitree2012}.
-We generated the tree using a Birth Death process by sampling the values of the speciation events ($\lambda$) and extinction events ($\mu$) from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
-We implemented a rejection sampling algorithm to select only random trees with 25 living and 25 fossil taxa. %Justification needed?
-We then added a taxa to the resulting Birth-Death tree as the outgroup of the tree.
-The mean branch length of the tree was used to separate the outgroup from the rest of the taxa and the branch length leading to the outgroup was set as the sum of the mean branch length and the longest root-to-tip length of the tree.
-
-Next, we created a molecular and a morphological matrix from the "true" tree.
-The molecular matrix was inferred from the "true" tree using the package phyclust v0.1-14 \citep{chen2011}.
-The matrix was made of 1000 characters sites for 51 taxa and generated using the seqgen algorithm \citep{ranbaut1997seqgen}.
-We used the HKY model \citep{HKY85} with a random base frequencies and with the transition/transversion rate of 2 \citep{douadycomparison2003} as parameters for generating the matrix.
-The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
-We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
-These parameters were selected to generate data with no special assumption about how the characters evolved as well as to reduce the computational time required if these parameters were estimated rather than defined (total computational time $>$ 65 CPU years).
-
-We inferred the morphological matrix using the ape package v3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
-We assigned the number of character states (either 2 or 3) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of characters states extracted from 100 published empirical morphological matrices (See \hyperref[supplementaries]{supplementaries}).
-We then ran an independent discrete character simulation for each character using the "true" tree branch length and topology with the randomly selected number of states (2 or 3) and assuming an equal rate of change (i.e.
-evolutionary rate) from one character state to an other \citep{Pagel22011994}.
-This method allows us to have only two parameters per character: the number of states and the evolutionary rate.
-For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used a low evolution rate parameter (i.e.
-$\alpha$) in order to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
-
-All the molecular information for fossil taxa was replaced by missing data ("?").
-Finally, we combined the morphological and molecular matrices obtained from the "true" tree.
-Hereafter we call this the "complete" matrix: the matrix with no missing data except for the molecular data of the fossil taxa.
-
-%---------------------------------------------
-%       Removing data
-%---------------------------------------------
-
-\subsection{Removing data}
-Once we obtained the "complete" matrix we modified it to get a set of matrices with missing data. %Ugly sentence
-We randomly replaced data with "?" in the morphological part of the matrices according to the following parameters (Fig. ~\ref{Fig_RemoveData}):
-
-\begin{enumerate}
-\item{The proportion of living taxa with no morphological data ($M_{L}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological one.
-Because of the increasing facility to sequence DNA for living taxa, the number of living taxa with molecular data is highly superior the the number of taxa with molecular and morphological data.
-\item{The proportion of missing morphological data across all fossil taxa ($M_{F}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the quality of the fossil record. 
-\item{The proportion of missing morphological characters across all taxa (living and fossil - $M_{C}$): 0\%, 10\%, 25\%, 50\% or 75\%. }
-This parameter illustrates the number of available morphological characters for both living and fossil taxa.
-\end{enumerate}
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_missingData-BW.pdf}
-\caption{Names of the different parts of the matrix.
-A: Missing data parameters:
-Missing living - The proportion of living taxa with no morphological data ($M_{L}$);
-Missing fossil - The proportion of missing morphological data across all fossil taxa ($M_{F}$);
-Missing character - The proportion of missing morphological characters across all taxa (living and fossil) ($M_{C}$).
-B: Different parts of the matrix:
-The "horizontal" sub-matrix (orange) contains molecular and morphological data for living taxa;
-The "vertical" sub-matrix (blue) contains morphological data for living and fossil taxa;
-The "corner" sub-matrix (orange and blue striped) contains morphological data for living taxa;
-The "missing-data" sub-matrix (grey) is the molecular part of the fossil taxa and contains no data.}
-\label{Fig_RemoveData}
-\end{figure}
-
-In practice, each parameter represent a different way of removing data in the morphological part of the matrix: $M_{L}$ removes a proportion of rows from the living taxa; $M_{F}$ removes a proportion of cells from the fossil taxa; and $M_{C}$ removes a proportion of columns across both living and fossil taxa (see Fig.
-~\ref{Fig_RemoveData}).
-Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a proportion of the living taxa is removed (i.e.
-removing rows), as for $M_{F}$, a proportion of data is removed across the whole of the morphological matrix for fossil taxa (i.e.
-removing cells).
-
-We tested all parameters combinations resulting in 125 ($5^3$) matrices.
-Because some parameter combinations introduce a lot of missing (e.g. $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
-When this occurred we repeated the random deletion of characters until every taxa had at least 5\% data across the whole matrix.
-
-%---------------------------------------------
-%       Building phylogenies
-%---------------------------------------------
-
-\subsection{Building phylogenies}
-From the resulting matrices we generated two types of trees, the "best" tree that is inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
-The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
-The "best" tree, on the other hand, is the best tree we can build using the state-of-the-art phylogenetic methods.
-In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g. \citet{rozen2005}).
-Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
-
-\subsubsection{Maximum Likelihood}
-The "best" tree and the "missing-data" trees were inferred using RAxML v8.0.20 \citep{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model (\citet{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citet{Stamatakis21012014}) as a generalisation of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
-The GTR model can be seen as a generalisation of the HKY model (the 2 parameters from the HKY model are implicitly included in the 6 from GTR model - \citet{stamatakisa2008}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} which is a generalisation of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014}).
-We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. %The explanations of the algorithm part for the BS was proposed by Trevor
-The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists in pruning one sub-tree from the tree and subsequently reinserting it to al neighbouring branches \citep{stamatakisa2008}.
-Sub-tree Pruning and Reinserting methods (SPR) have been demonstrated as being better than others (e.g. Nearest Neighbouring Interchange - NNI) in recovering good bootstrap values \citep{salamin2003}.
-
-\subsubsection{Bayesian}
-The "best" tree and the "missing-data" trees were inferred using MrBayes v3.2.2 \citep{Ronquist2012mrbayes}.
-We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
-The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \citep{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
-For the morphological data, we used the Markov \textit{k} state model \citep{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
-We chose these models to be consistent with the parameters used to generate the "complete" matrix.
-
-Each Bayesian tree was estimated using two runs of four chains each for a maximum of 50$\times$$1^6$ generations.
-We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
-The effective sample size (ESS) was also checked on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
-For each run, we removed 25\% of the iterations as burn-in.
-We used the following prior for each tree (see \hyperref[supplementaries]{supplementaries}):
-\begin{enumerate}
-\item
-the "true" tree’s topology as a starting tree (with a starting value for each branch length of 1),
-\item
-an exponential prior on the shape of the gamma distribution of $\alpha$=0.5 for both partitions
-\item
-and a transition/transversion ratio prior of 2 sampled from a strong beta distribution ($\beta$(80,40)).
-\end{enumerate}
-
-We used these prior to speed up the Bayesian process.
-These prior biased the way the Bayesian process calculated the branch length by giving non-random starting points and boundaries for the parameters estimation process, however, in this study, we focused on the effect of missing data on the topology and not on the branch length.
-Even using these prior, it took 65 CPU years to build 51 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
-
-%---------------------------------------------
-%       Comparing Topologies
-%---------------------------------------------
-
-\subsection{Comparing topologies}
-We compared the topology of the "missing-data" trees inferred from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
-Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees inferred from total evidence type matrices.
-To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triples \citep{dobson1975triplets} and the Robinson-Fould \citep{RF1981} distance.
-We normalised the two metrics using the Normalised Tree Similarity index \citep{Bogdanowicz2012} to generalise our results for any \textit{n} number of taxa.
-The three metrics are detailed below.
-
-\subsubsection{Triples distance ($T_{x,y}$) \citep{dobson1975triplets}}
-This metric measures the number of different sub-trees of three taxa between two given trees.
-Each triplet can be written as $I_{ijk}$=(\textit{ijk}).
-Where $I_{ijk}$ is equal to 0 if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to 1.
-For any rooted binary tree there are only three possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
-If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k});.
-One can calculate $S_n$, the triplet distance between two trees as:
-\begin{equation}
-S_n = \sum_{ijk} I_{ijk}
-\end{equation}
-Where:
-\begin{equation}
-\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
-\end{equation}
-And where \textit{n} is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
-If $S_n$=0, the trees are the same (i.e. no taxa as been displaced).
-When $S_n$ = $\binom{n}{4}$, the trees are the most different possible (i.e. every taxa as been displaced).
-This metric therefore illustrates the amount of displayed taxa.
-It is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g.
-fossil taxa) than the Robinson-Foulds metric \citep{critchlowthe1996,johnson1998,wiensmissing2003}.
-Therefore we used this metric as a proxy to estimate the robustness of the tree to flying taxa (see \hyperref[supplementaries]{supplementaries}).
-
-\subsubsection{Robinson-Fould distance (\textit{RF}) \citep{RF1981}}
-This metric measures the number of shared clades among two trees and therefore illustrates the number of exactly conserved groups among the trees.
-The Robinson-Fould distance (also called path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
-\begin{equation}
-RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
-\end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees.
-The minimal value of \textit{C} is 1 if the two trees have the same \textit{n} taxa; the maximal value in \textit{C=n-2}.
-This metric is more sensitive to taxa displacement than the Triples metric (i.e.
-if one taxa gets out of a clade, then the clades are no longer considered as similar - \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
-Therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades (see \hyperref[supplementaries]{supplementaries}).
-
-\subsubsection{Normalised Tree Similarity \textit{NTS} \citep{Bogdanowicz2012}}
-For any tree with \textit{n} taxa compared using a tree distance metric \textit{m}, $NTS_m$ represents the similarity score between the two trees given the expected distance between two random Yule trees with \textit{n} taxa.
-Let $\bar{d}_{m,n}$\textit{(rand)} be the average distance between two random Yule trees with \textit{n} taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} containing each \textit{n} taxa, then:
-\begin{equation}
-NTS_{m,n}(x,y)=\frac{\bar{d}_{m,n}(rand) - d_{m,n}(x,y)} {\bar{d}_{m,n}(rand)}
-\end{equation}
-\textit{NTS} ranges from 1 to -$\infty$.
-For any \textit{m},\textit{n} when \textit{NTS}=1, the trees are the same;
-when \textit{NTS}=0 the trees are not more different than expected by chance;
-when \textit{NTS}$<$0, the trees are more different than expected by chance (see \hyperref[supplementaries]{supplementaries}).
-
-$\newline$
-
-We compared the "missing-data" tree to the "Best" tree for each chain.
-For the Maximum Likelihood trees we performed pairwise comparisons between the "Best" tree and the "missing-data" tree (see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triples metric.
-We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
-For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparisons for the same metric and the same number of taxa \textit{n}=51 (see \hyperref[supplementaries]{supplementaries}).
-We ran the comparison for every "missing-data" tree in each chain resulting in 51 comparisons for every "missing-data" trees. 
-We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package (v3.1 – \citet{hdrcde}).
-
-Bayesian tree inference allows to account for the statistical uncertainty of a phylogenetic tree by not using an optimal criterion (c.f.
-Maximum Likelihood) and gives a tree posterior distribution (c.f.
-the likeliest tree in ML).
-This method has the clear advantage of better dealing with error and uncertainty but it's output (a distribution rather than a single tree) is less practical to use for any further study (but see \citet{healyecology2014}).
-%Awwwwww yiis!
-To avoid this problem, people traditionally use the a consensus tree build on a majority rule in order to summarise a Bayesian tree posterior distribution.
-This results into a single tree containing both topological and branch length information as well as support information (i.e. posterior probabilities, e.g. \citet{Ronquist2012mrbayes}).
-However, using a Bayesian consensus tree has limitation especially if the resulting consensus tree is not well supported or resolved.
-Because the metrics used in this study are used to measure variation in topology between two trees (i.e.
-taxa placement or clade position), comparing two Bayesian consensus trees is not optimal in picking topological difference signal.
-For example, if one of the trees is not resolved at all (i.e. a star tree), comparing it to any other tree (even an other star tree) will not give useful results.
-Therefore we used the entire Bayesian trees posterior distribution in order to perform the tree comparisons (i.e.
-the "Best" Bayesian tree posterior distribution vs.
-the "missing-data" Bayesian tree posterior distribution).
-
-\subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
-We compared the Bayesian posterior distribution using a Random Pairwise Bayesian Tree Comparison (RPBTC) method.
-This method consists in comparing a series of randomly selected pairs of trees from two posterior distributions (on from each distribution) and use the mode to summarize the resulting distribution as a proxy of the difference between the two trees in a Bayesian framework.
-\begin{equation}
-RPBTC_{X,Y}=Mo[(]d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]
-\end{equation}
-
-Where \textit{X} and \textit{Y} are two Bayesian tree posterior distributions;
- $d_{m,n}$ is the pairwise difference for any metric \textit{m} and \textit{n} taxa between $X_{i}$ and $Y_{j}$ which are two single trees randomly sampled respectively from \textit{X} and \textit{Y};
- $Mo$ is the mode and
- $Mo[(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]$ is the mode of the pairwise difference repeated \textit{k} times.
-
-We used the mode to summarize the distributions because it is the value that is the most represented in the distribution and reflects what a consensus tree represents (the topology the most represented in the posterior distribution).
-Also, the mode is a value present in the distribution (which, depending on the metric, can be a series of discrete values) in contrast to the other distribution summary metrics, such as the mean or the median, which can take values that are not actually present in the distribution.
-For example, if comparing two posterior tree distribution, we obtain difference values of 10, 10, 10, 5, 5 and 1, using the mean (6.8) or the median (7.5) gives a values that actually doesn't exist.
-%Useful?
-
-In this study, we calculated RPBTC for both metrics (Robinson-Fould and Triples distance) for each pairs of Bayesian tree distributions (i.e.
-the "Best" Bayesian tree posterior distribution vs.
-the "missing-data" Bayesian tree posterior distribution) with 1000 random pairwise comparisons (\textit{k}).
-Because the RPBTC uses random pairwise comparisons, there is a chance of comparing always trees that have the biggest or the smallest difference between two distributions, either deflating or inflating the RPBTC difference value.
-However, using 1000 random pairwise comparisons makes the difference stable (i.e. if repeated independently, no difference is detected - see \hyperref[supplementaries]{supplementaries}).
-
-$\newline$
-
-We calculated the NTS for the "Best" Bayesian tree vs.
-"missing-data" Bayesian tree comparison for both Robinson-Fould and Triples metric using the mode of the RPBTC for each chain resulting in a distribution of 51 modes per comparison.
-We then compared the NTS of the Bayesian trees to the NTS of the ML trees for both metrics for each comparison.
-This resulted in a distribution of 51 NTS values for the Bayesian trees and 51 NTS values for the ML trees for each tree comparison for which we calculated the mode and the 50\% and 95\% confidence intervals (see \hyperref[supplementaries]{supplementaries} - Fig.
-~\ref{Fig_Compare}).
-
-\begin{figure} % Modify the figure
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp-BW.pdf}
-\caption{Tree comparison protocol.
-From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we inferred a tree in Baysian or Maximum likelihood framework.
-We then compared this tree with or without fossil/living taxa to the "best" tree.
-In the Bayesian framework, 1000 trees from the posterior distribution where used for each comparison.}
-\label{Fig_Compare}
-\end{figure}
-
-\subsubsection{Statistical difference between tree topologies} %note that in the end, I only have non-parametric tests, the parametric explanation part can go to the supplementaries
-To assess the difference between the tree topologies, whether they were inferred in a Bayesian or a ML framework, we performed a group comparison test for each set of configurations (e.g.
-effect of $M_L$ in Bayesian inference, combined effect of $M_F$ and $M_C$ in ML inference, etc...).
-We used the pairwise tree comparisons as factors (i.e.
-the "Best" tree vs.
-one of each of the 125 "missing-data" trees) and the values of the 51 replicates from the pairwise comparison as a response variable.
-The 51 replicates values were either the replicates of the 51 pairwise tree comparisons in ML or the modes of the 51 RPBTC posterior distributions in Bayesian.
-When we found a significant difference between the groups, we performed a pairwise comparison test to analyse which pairs of groups differ from the others.
-Depending on whether the data was normal with variance homoscedasticity or not (controlled by performing respectively a shapiro.test and a a bartlett.test in R\{stats\}), we used respectively parametric or non-parametric tests for the group comparison and the pairwise comparisons as suggested by \citet{ruxtontime2008} (Table ~\ref{Stats_test}).
-%link broken
-
-\begin{table}[ht]
-\caption{Statistical tests}
-\centering
-\begin{tabular}{lrll}
-  \hline
-   & test & statistics & code\{package\} \\ [0.5ex]
-  \hline
-  Groups comparisons & ANOVA & parametric & anova\{stats\} \\ [0.5ex]
-                     & Kruskal-Wallis & non-parametric & kruskal.test\{stats\} \\[0.5ex]
-  \hline
-  Pairwise comparisons & Tukey HSD & parametric & TukeyHSD\{stats\} \\ [0.5ex]
-                       & Nemenyi-Damico-Wolfe-Dunn & non-parametric & kruskalmc\{pgirmess\} \\ [0.5ex]
-  \hline
-\end{tabular}
-\label{Stats_test}
-\end{table}
-
-\subsection{Effect of missing molecular characters for fossil taxa} %Awfull name.
-To assess the effect of missing molecular characters for fossil taxa in a Bayesian framework, we split the "complete" matrix in sub-matrices containing no missing data at all (see Fig.
-~\ref{Fig_RemoveData}):
-\begin{enumerate}
-\item
-A first containing both molecular and morphological data for living taxa only (hereafter called the "horizontal" sub-matrix, Table ~\ref{Tab_glossary});
-\item
-A second one containing morphological data for both living and fossil taxa (hereafter called the "vertical" sub-matrix, Table ~\ref{Tab_glossary});
-\item
-A third one containing morphological data for living taxa only (hereafter called the "corner" sub-matrix, Table ~\ref{Tab_glossary});
-\end{enumerate}
-We reran the Bayesian tree inference on the different sub-matrices with no missing data in the same way as described above.
-We then compared the resulting tree posterior distribution to itself (in the same way described above) to assess the ability to recover topology for each simulation when no missing data was involved in the phylogenetic inference process.
-
-%---------------------------------------------
-%       Empirical data - TO DO
-%---------------------------------------------
-
-\subsection{Empirical data} %Still not done!
-We also compared the results obtained from simulated data by using \citet{ronquista2012} empirical data.
-The matrix contains 67 living species plus one outgroup and 45 fossil species of Hymenopteras with 5097 molecular characters and 354 morphological characters.
-From the 68 living species used in the matrix, only 66 had molecular data, we therefore treated these 66 taxa as "living" taxa and all the other 47 as "fossil" taxa.
-We treated the matrix in the exact same way as described in step 2 and 3 resulting in 125 matrices with various amount of missing data and the same number of Maximum Likelihood and Bayesian trees.
-We used the same settings as for the simulated data in the Maximum Likelihood framework.
-For the Bayesian inferences however, we didn’t used any priors except that we provided a starting tree with the topology of the 68 living species (topology with the highest posterior probability from non-clock analysis - \citet{ronquista2012}).
-Contrary to \citet{ronquista2012} analysis, we didn’t performed any clock analysis since we were only interested in the topology of the inferred tree and not the branch length.
-
-%---------------------------------------------
-%
-%       RESULTS
-%
-%---------------------------------------------
-
-\section{Results}
-
-\subsection{Building the trees}
-When generating the matrices using seqgen \citep{ranbaut1997seqgen} and rTraitDisc \citep{paradisape:2004} algorithms with low evolutionary rate parameter ($\alpha$=0.5) we successfully generated evolutionary histories.
-In Maximum Likelihood, support values ranged from 0 to 100 with a median of 38 ($1^st$ quartile = 14, $3^rd$ quartile = 72 - see \hyperref[supplementaries]{supplementaries}). % This is actually bad.
-The distribution of the phylogenetic signal across the different chains provided us a good spread of evolutionary scenarios: from well resolved phylogenies to poorly resolved ones.
-This allows us to test our simulations in a theoretical as well as practical framework (i.e.
-phylogenies with low support values are rarely published but are still common in early stages of projects).
-%REF???
-- That's kind of my way to turn these bad results into something usable: even if some of our simulations are not realistic, it shows the effect of missing data in any scenario possible, not only in realistic cases.
-Also, one can advocate that this low topological resolution is actively due to the amount of missing data in the total evidence matrices since the sub-matrices (no missing data) analysis recovers way higher support value  .
-%This is something that I will had with the whole submatrices story.
-Basically, when you have all the fossil/molecular data missing, it's harder to build the tree, both in ML and in Bayesian.
-When you run the analysis without missing data, then it works way better (i.e.
-our REAL control test: can we simulate matrices in the first place).
-Using Bayesian inference, all the chains converged with a significant effective sample size by using 2 parallel runs of 4 chains each (ASSD $<$ 0.01 and ESS $>>$ 200 for inference).
-
-%\subsection{Comparing topologies}
-%Did it worked? See sub-matrices part. TO DO
-
-\subsection{Effect of the missing data parameters}
-
-%10/06 - Missing 8 ML Chains
-\begin{figure} 
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-results-BW.pdf}
-\caption{Effect of missing data on Tree similarity.
-$M_L$ is the amount of missing living taxa with morphological data,
-$M_F$ is the amount of missing data in the fossil record,
-$M_C$ is the amount of missing morphological characters.
-The first row represent the Normalised Tree Similarity index using the Robinson-Fould metric;
-the second row represent the Normalised Tree Similarity index using the Triples metric.}
-\label{Fig_Results}
-\end{figure}
-
-\subsubsection{Effect of $M_L$}
-In Maximum Likelihood framework, the topological recovery (reflected by the tree similarity) quickly drops with the increasing amount of missing data when using both metrics (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-However, this negative effect of missing data is stabilized beyond 25\% of missing data (no significant difference for trees with respectively 25\%, 50\% and 75\% missing living taxa).
-With 10\% of missing living taxa, there is a different topological recovery depending on the used metric.
-The Robinson-Fould metric shows a lower Tree similarity index than the Triples metric meaning that individual taxa position are generally more conserved (i.e.
-low number of flying taxa) but that the full clades are less conserved (see Table ~\ref{NTS_ML_results}).
-
-However, topological recovery was bad in Bayesian, regardless the amount of missing data.
-The tree similarity index was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance (see Table ~\ref{NTS_Ba_results}).
-Even if the topological recovery is really low, we detected an effect of missing data on the topology depending on the used metric.
-For the Triples metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
-For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases (see Table ~\ref{Ba_RF-ML_results}).
-However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the tree similarity index was only slightly higher than expected by chance (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-
-\subsubsection{Effect of $M_F$}
-In Maximum Likelihood framework, the effect of missing data in the fossil record appeared to be constant and led to a constant decrease in topological recovery when the missing data increased (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-The difference between the two metrics used is more important than for the $M_L$ parameter: clades are less conserved than individual taxa placement as the amount of missing data increases.
-Interestingly, 10\% or 25\% missing data for the $M_F$ parameter does not seems to affect the apparition of unstable taxa (Triples NTS modes of respectively 0.97 and 0.93 - Table ~\ref{NTS_ML_results}).
-
-Similarly than for the effect of $M_L$, Bayesian tree recovery was only slightly better than expected by chance.
-In the same way as described above, there seems to be a small difference depending on the metric used.
-There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric (see Table ~\ref{Ba_RF-MF_results}).
-However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
-
-\subsubsection{Effect of $M_C$}
-The number of missing morphological characters ($M_C$), however, seems to be more affecting topological recovery than $M_L$ and $M_F$ (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
-The Robinson-Fould metric shows a rapid drop in tree recovery from 10\% of missing data (Robinson-Fould NTS mode $<$ to 0.5 from 10\% missing data - Table ~\ref{NTS_ML_results}).
-This decrease is however slower when using the Triples metric with still a good topological recovery at 10\%  (Triples NTS modes = 0.95 - Table ~\ref{NTS_ML_results}).
-
-The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Fould metric shows a significant effect of the $M_C$ parameter on tree recovery (see Table ~\ref{Ba_RF-ML_results} and ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-
-%\subsubsection{Combined effect}
-%Need to do the heat plots
-
-$\newline$
-
-%To sum up:
-The ability of recovering the "best" tree's topology in Maximum Likelihood method is function of the amount of data missing.
-The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
-The decrease in clade conservation (low Robinson-Fould NTS score) is faster than the increase in flying taxa (low Triples NTS score).
-In Bayesian framework, topology is badly (if not at all) recovered disregarding the amount of data missing.
-
-\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
-\caption{Tree similarity values per parameter in ML framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
-              & missing data &        &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.671 & 0.569  & 0.789 & 0.468  & 1.036 \\ %how many digits?
-              &              & Triples        & 0.942 & 0.868  & 0.999 & 0.612  & 1.063 \\
-              & 25\%         & Robinson-Fould & 0.516 & 0.440  & 0.630 & 0.266  & 0.892 \\
-              &              & Triples        & 0.559 & 0.708  & 0.810 & 0.291  & 0.999 \\
-              & 50\%         & Robinson-Fould & 0.432 & 0.364  & 0.521 & 0.231  & 0.642 \\
-              &              & Triples        & 0.514 & 0.419  & 0.599 & 0.293  & 0.910 \\
-              & 75\%         & Robinson-Fould & 0.372 & 0.323  & 0.434 & 0.203  & 0.577 \\
-              &              & Triples        & 0.444 & 0.338  & 0.532 & 0.192  & 0.855 \\
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.907 & 0.766  & 1.000 & 0.407  & 1.047 \\
-              &              & Triples        & 0.976 & 0.944  & 1.003 & 0.715  & 1.104 \\
-              & 25\%         & Robinson-Fould & 0.572 & 0.488  & 0.692 & 0.325  & 0.936 \\
-              &              & Triples        & 0.937 & 0.820  & 0.993 & 0.539  & 1.086 \\
-              & 50\%         & Robinson-Fould & 0.433 & 0.302  & 0.476 & 0.170  & 0.668 \\
-              &              & Triples        & 0.732 & 0.594  & 0.873 & 0.341  & 1.025 \\
-              & 75\%         & Robinson-Fould & 0.153 & 0.115  & 0.200 & 0.068  & 0.323 \\
-              &              & Triples        & 0.458 & 0.390  & 0.611 & 0.290  & 0.856 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.473 & 0.384  & 0.835 & 0.356  & 0.958 \\
-              &              & Triples        & 0.959 & 0.869  & 1.026 & 0.572  & 1.067 \\
-              & 25\%         & Robinson-Fould & 0.454 & 0.357  & 0.630 & 0.242  & 0.916 \\
-              &              & Triples        & 0.907 & 0.751  & 0.997 & 0.446  & 1.070 \\
-              & 50\%         & Robinson-Fould & 0.424 & 0.375  & 0.466 & 0.249  & 0.828 \\
-              &              & Triples        & 0.683 & 0.571  & 0.822 & 0.277  & 1.008 \\
-              & 75\%         & Robinson-Fould & 0.230 & 0.177  & 0.302 & 0.108  & 0.567 \\
-              &              & Triples        & 0.451 & 0.373  & 0.572 & 0.224  & 0.791 \\
-    \hline
-\label{NTS_ML_results}
-\end{tabular}
-\end{table}
-
-\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
-\caption{Tree similarity values per parameter in Bayesian framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
-              & missing data &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\ %how many digits?
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.028   & 0.061 \\
-              &              & Triples        & 0.033 & 0.003  & 0.049 & -0.049  & 0.117 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.061 \\
-              &              & Triples        & 0.035 & 0.002  & 0.049 & -0.041  & 0.096 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.056 \\
-              &              & Triples        & 0.028 & 0.003  & 0.052 & -0.038  & 0.092 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.010   & 0.043 \\
-              &              & Triples        & -0.001 & -0.018 & 0.040 & -0.060 & 0.106 \\
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.035 & 0.031  & 0.056 & 0.027   & 0.076 \\
-              &              & Triples        & 0.026 & 0.002  & 0.050 & -0.046  & 0.105 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.0153  & 0.068 \\
-              &              & Triples        & 0.038 & 0.002  & 0.050 & -0.046  & 0.105 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.058 \\
-              &              & Triples        & 0.014 & -0.011 & 0.039 & -0.038  & 0.124 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.041 \\
-              &              & Triples        & 0.029 & 0.019  & 0.042 & -0.030  & 0.089 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.056 & 0.035  & 0.059 & 0.0281  & 0.076 \\
-              &              & Triples        & 0.030 & 0.011  & 0.048 & -0.024  & 0.094 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.023   & 0.076 \\
-              &              & Triples        & 0.020 & -0.001 & 0.037 & -0.060  & 0.109 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.030   & 0.059 \\
-              &              & Triples        & 0.010 & -0.005 & 0.045 & -0.036  & 0.104 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.011   & 0.043 \\
-              &              & Triples        & 0.024 & 0.001  & 0.042 & -0.037  & 0.104 \\
-    \hline
-\label{NTS_BA_results}
-\end{tabular}
-\end{table}
-
-\begin{table}
-\caption{Group difference per configuration set}
-\centering
-\begin{tabular}{rllcccc}
-    \hline
-    set & metric & framework & parametric & $stats^1$ & df & p.value \\
-    \hline
-    $M_L$         & Robinson-Fould & ML       & no & 151.438  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.001278e-31
-                  &                & Bayesian & no & 65.62    & 4   & \textbf{0.00} \\ 
-                  & Triples        & ML       & no & 142.8329 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains  - 6.983434e-30
-                  &                & Bayesian & no & 4.56     & 4   & 0.34          \\ 
-    $M_F$         & Robinson-Fould & ML       & no & 179.9238 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 7.743067e-38
-                  &                & Bayesian & no & 85.61    & 4   & \textbf{0.00} \\
-                  & Triples        & ML       & no & 150.059  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.977386e-31
-                  &                & Bayesian & no & 3.01     & 4   & 0.56          \\
-    $M_C$         & Robinson-Fould & ML       & no & 151.3818 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.029428e-31
-                  &                & Bayesian & no & 64.23    & 4   & \textbf{0.00} \\
-                  & Triples        & ML       & no & 138.374  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 6.290979e-29
-                  &                & Bayesian & no & 26.13    & 24  & 0.35          \\
-    $M_L+M_F$     & Robinson-Fould & ML       & no & 734.9213 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.104542e-139
-                  &                & Bayesian & no & 317.97   & 24  & \textbf{0.00} \\
-                  & Triples        & ML       & no & 531.5558 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 4.581065e-97
-                  &                & Bayesian & no & 3.01     & 24  & 0.56          \\
-    $M_L+M_C$     & Robinson-Fould & ML       & no & 542.3233 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 2.619808e-99
-                  &                & Bayesian & no & 290.10   & 24  & \textbf{0.00} \\
-                  & Triples        & ML       & no & 437.7413 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains -  1.28493e-77
-                  &                & Bayesian & no & 22.19    & 24  & 0.57          \\
-    $M_F+M_C$     & Robinson-Fould & ML       & no & 812.206  & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 5.463138e-156
-                  &                & Bayesian & no & 385.96   & 24  & \textbf{0.00} \\
-                  & Triples        & ML       & no & 528.9215 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.619383e-96
-                  &                & Bayesian & no & 20.23    & 24  & 0.68          \\ 
-    $M_L+M_F+M_C$ & Robinson-Fould & ML       & no & 3603.808 & 124 & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains -3.819574e-169
-                  &                & Bayesian & no & 1167.40  & 124 & \textbf{0.00} \\
-                  & Triples        & ML       & no & 2047.923 & 124 & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains 
-                  &                & Bayesian & no & 115.86   & 124 & 0.69          \\
-    \hline
-\label{Group_results}
-\end{tabular}
-
-$^1$ F-value for parametric tests and Kruskal Wallis $chi^2$ for non parametric tests.
-\end{table}
-
-\begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE  & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE  \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 56.53 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$25\% & 96.85 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 127.03 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 144.58 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 40.31 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$50\% & 70.50 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 88.05 & 37.66 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 30.19 & 37.66 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 47.73 & 37.66 & TRUE \\ 
-  $M_L$50\%-$M_L$75\% & 17.55 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_RF-ML_results}
-\end{table}
-
-\begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE  & FALSE \\
-    $M_L$50\% & & & & - & FALSE  \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 57.98 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$25\% & 105.83 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 120.86 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 140.34 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 47.85 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$50\% & 62.88 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 82.36 & 37.66 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 15.03 & 37.66 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 34.51 & 37.66 & FALSE \\ 
-  $M_L$50\%-$M_L$75\% & 19.48 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_Tr-ML_results}
-\end{table}
-
-\begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & FALSE & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 17.05 & 41.00 & FALSE \\ 
-  $M_L$00\%-$M_L$25\% & 43.79 & 41.00 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 77.32 & 41.00 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 103.99 & 41.00 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 26.75 & 41.00 & FALSE \\ 
-  $M_L$10\%-$M_L$50\% & 60.27 & 41.00 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 86.94 & 41.00 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 33.53 & 41.00 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 60.20 & 41.00 & TRUE \\ 
-  $M_L$50\%-$M_L$75\% & 26.67 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{Ba_RF-ML_results}
-\end{table}
-
-\begin{table}
-\caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & \textbf{TRUE} \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 59.27 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$25\% & 79.72 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$50\% & 120.70 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 167.81 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 20.45 & 37.66 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 61.43 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 108.55 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 40.98 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$75\% & 88.09 & 37.66 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 47.12 & 37.66 & TRUE \\ 
-   \hline
-\end{tabular}
-\label{ML_RF-MF_results}
-\end{table}
-
-\begin{table}
-\caption{$M_F$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 64.78 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$25\% & 89.09 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$50\% & 123.20 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 150.43 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 24.31 & 37.66 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 58.42 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 85.65 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 34.10 & 37.66 & FALSE \\ 
-  $M_F$25\%-$M_F$75\% & 61.34 & 37.66 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 27.23 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_Tr-MF_results}
-\end{table}
-
-\begin{table}
-\caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & FALSE & FALSE & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 16.03 & 41.00 & FALSE \\ 
-  $M_F$00\%-$M_F$25\% & 38.50 & 41.00 & FALSE \\ 
-  $M_F$00\%-$M_F$50\% & 61.52 & 41.00 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 95.52 & 41.00 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 22.47 & 41.00 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 45.49 & 41.00 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 79.49 & 41.00 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 23.02 & 41.00 & FALSE \\ 
-  $M_F$25\%-$M_F$75\% & 57.02 & 41.00 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 34.00 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{Ba_RF-MF_results}
-\end{table}
-
-\begin{table}
-\caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & \textbf{TRUE} \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 67.99 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$25\% & 87.58 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$50\% & 116.16 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 155.77 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 19.59 & 37.66 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 48.17 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 87.78 & 37.66 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 28.58 & 37.66 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 68.19 & 37.66 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 39.60 & 37.66 & TRUE \\ 
-   \hline
-\end{tabular}
-\label{ML_RF-MC_results}
-\end{table}
-
-\begin{table}
-\caption{$M_C$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & FALSE \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 75.29 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$25\% & 88.58 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$50\% & 114.02 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 149.60 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 13.29 & 37.66 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 38.73 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 74.31 & 37.66 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 25.44 & 37.66 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 61.02 & 37.66 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 35.58 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_Tr-MC_results}
-\end{table}
-
-\begin{table}
-\caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & FALSE & FALSE & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & FALSE \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 16.03 & 41.00 & FALSE \\ 
-  $M_C$00\%-$M_C$25\% & 38.50 & 41.00 & FALSE \\ 
-  $M_C$00\%-$M_C$50\% & 61.52 & 41.00 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 95.52 & 41.00 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 22.47 & 41.00 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 45.49 & 41.00 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 79.49 & 41.00 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 23.02 & 41.00 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 57.02 & 41.00 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 34.00 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{Ba_RF-MC_results}
-\end{table}
-
-%---------------------------------------------
-%
-%       DISCUSSION
-%
-%---------------------------------------------
-
-\section{Discussion}
-
-\subsection{Building the trees}
-Simulating evolutionary history matrices still remains a big drawback in theoretical phylogenetics. %Hahahaha.... feck.
-The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
-This configuration probably lead to globally low phylogenetic signal as well as the intrinsic difficulties to simulated characters with phylogenetic signal.
-Even though molecular characters evolution (and therefore simulation) only depends of a small number of parameters (i.e.
-the base frequencies and the substitution matrix), simulating molecular matrix with a strong evolutionary signal is still complicated when generating unrealistically small matrices.
-For morphological characters, the underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
-
-Also morphological characters studies involve many potential statistical pitfalls (e.g.
-independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy.
-%REF probably April as something about it
-\begin{enumerate}
-\item{First, morphological data can display a different signal than molecular data, especially in small matrices. %REF
-This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
-However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013}.}
-%The main difference between molecular and morphological characters is that, because of historical and practical reasons, a morphological character matrix do no contains autapomorphies and invariant characters which are usually more common than synampomorphies in molecular character matrices (cite).
-\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013}.
-However, following our simulation assumption of random character distributions, if they accumulate through time in the same way as the majority of the molecular characters then, homoplasic characters are expected to appear randomly through time \citep{davalosintegrating2014}.
-Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
-After a reaching a critical amount of morphological characters, adding new ones increases homoplasy \citep{wagner2000}.}
-\end{enumerate}
-Our simulation parameters both decrease the phylogenetic signal (difficulty to simulate morphological characters and size of the matrix) as well as it increases it (reduction of homoplasy in the morphological part of the matrix).
-Therefore, these drawbacks seems to have only a minor impact on the main results of this study: the incapacity of recovering any topology in Bayesian inference.
-
-\subsection{Comparing topologies}
-Comparing topologies is a crucial question in phylogenetics but has always been hard to normalise because of the vast amount of different metrics used as proxies for different aspects of tree similarity/dissimilarity \citep{agapowpower2002}.
-Because our global framework is to study how to include efficiently fossils into phylogenies, we chose metrics reflecting the most interesting aspect of this global question: where do individual taxa (i.e.
-the fossils) branch in the tree.
-The Robinson-Fould \citep{RF1981} and the Triples \citep{critchlowthe1996} metrics are more sensible to taxa and clade placement than other tree comparison metrics \citep[e.g.][Imbalance metric]{kirkpatricksearching1993} and where therefore favoured.
-Also by using the Normalised Tree Similarity index \citep{Bogdanowicz2012} we emphasize the aspect of "good" phylogenetic signal \textit{versus} random phylogenetic signal because it normalised the values of the metric by correcting for the expected value when comparing random trees (NTS = 0).
-
-The Robinson-Fould metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
-When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
-On the other hand, the Triples method is measuring the position of each taxon towards to other reference taxa \citep{critchlowthe1996}.
-It will penalise only trees where taxa get removed furthest from their original clade.
-Regarding our problematic (how does missing data influence topological recovery in trees containing both living and fossil taxa) we are more interested the placement of taxa (i.e.
-where does the fossil branch) than the exact conservation of clades.
-
-%Need a transition!
-
-Although the idea of comparing two single trees is straightforward, it becomes more complex when comparing trees distributions (e.g. Bayesian posterior distributions).
-The introduction of our Random Pairwise Bayesian Tree Comparison methods allows to summarize the comparison of two trees distributions by picking up the most frequent signal in the distribution of the pairwise comparisons (i.e.
-the mode).
-Even though this method is subject to randomness and might artificially increase or decrease the score of the studied metric, simulations showed that when a sufficient amount of random comparisons is performed, this method doesn't seem to be subject to randomness (e.g.
-1000 random pairwise comparisons - see ~\ref{supplementaries}).
-
-\subsection{Maximum Likelihood versus Bayesian}
-The main results of our analysis shows that Bayesian inference fails to recover any Topology in a total evidence method framework (Fig. ~\ref{Fig_Results}).
-This results is surprising regarding the behaviour of Maximum Likelihood inference (Fig.
-~\ref{Fig_Results}) as well as the Bayesian inferences performed on empirical data \citep{ronquista2012,schragocombining2013}.
-However, it is important to note that this effect was not mentioned in the aforementioned empirical tests because both studies used a Bayesian approach with fixed topology \citep{ronquista2012,schragocombining2013}.
-
-In our case, we suspect that this inability of Bayesian methods to recover topology is due to the intrinsic structure of a total evidence matrix (Fig ~\ref{Fig_RemoveData}).
-In fact, as previously studies have shown, missing data doesn't seem to be a major drawback as long as the missing data is randomly distributed ~\citep[e.g.][]{wiensmissing2003,rouresite-specific2011,sansomfossilization2013}.
-However, in total evidence matrices, the majority of the missing data is not randomly distributed but concentrated in the molecular part of the matrix for the fossil taxa (i.e.
-the missing-data sub-matrix, Fig ~\ref{Fig_RemoveData}).
-This leads to high decrease of topological recovery when using non optimal criterion approach (i.e.
-Bayesian inference) because of the high variance in the near-likeliest solutions sampled in the Bayesian posterior distribution.
-In opposition, when applying optimal criterion approach (i.e.
-Maximum Likelihood), it is still possible to sample the likeliest tree.
-%probably need to develop that way more, especially with actually analysing the likelihood landscape.
-
-\subsection{Effect of missing data}
-The three parameters we selected in this study account for three potential pitfalls in collecting the data for total evidence analysis:
-$M_L$ represents the living taxa for which there is no available morphological data,
-$M_F$ represents the quality of the fossil record,
-and $M_C$ represents the general coding effort and the overall of morphological data available.
-Ideally, the lowest possible amount of missing data is wanted in any phylogenetic analysis leading to a data collection part prior to the phylogenetic analysis.
-Each of the three parameters can be improved in a different way prior to the analysis:
-for $M_L$, one should put more effort in using natural museums history collections for coding the missing morphological data for living taxa if possible %go to the museum!;
-for the $M_F$ parameter, the amount of missing data unfortunately depends on the quality of the fossil record and can not be actively improved and depends on exceptional discoveries \citep[e.g.][]{nithe2013};
-finally for the $M_C$ parameter, improvement can be done by vast collaborative projects in order to gather as much characters as possible \citep[e.g.][]{O'Leary08022013}.
-
-Our results shows that the $M_F$ parameter have less influence on recovering the good tree topology in Maximum Likelihood framework which is fortunate since it is the parameter that is the more difficult to fix for practical reasons.
-Regarding the fact that we are more interested in taxa placement than in clade conservation (Triples \textit{versus} Robinson-Fould), the parameter that affects the more the decrease in topological recovery is the number of missing living taxa in the morphological part of the matrix ($M_L$) and the overall number of morphological characters $M_C$.
-This makes sense since the living taxa are bearing both the information to build the tree backbone (the molecular data) and the information used for branching the fossil on this backbone (the morphological information).
-Therefore, we advocate the importance of coding morphological characters for the most living taxa possible and with the most characters possible, potentially by using collaborative projects portals such as morphobank \citep{morphobank}.
-
-%Potential reviewers comments?
-%What about it when you vary the proportion of fossil taxa?
-
-%---------------------------------------------
-%
-%       CONCLUSION
-%
-%---------------------------------------------
-
-\section{Conclusion}
-A Bayesian approach fails to recover accurate topology in a total evidence approach framework, whatever the amount of missing data.
-We think this failure is due to the intrinsic structure of the total evidence matrices that includes a vast amount of non randomly distributed missing data (i.e.
-the molecular part of the matrix for the fossil taxa).
-This missing data is equal to the number of fossil taxa $\times$ the number of molecular characters leading to a vast amount of trees to be sampled in Bayesian posterior distribution.
-However, one can use a Maximum Likelihood approach to fix the likeliest topology.
-If so, an effort should be made prior to the phylogenetic analysis on collecting as much morphological data as their is available from living taxa in order to efficiently improve the quality of the trees.
-
-%Perspectives:
-%What the effect of this missing molecular data on branch length estimation?
-
-%---------------------------------------------
-%
-%       ...
-%
-%---------------------------------------------
-
-\section{Acknowledgements}
-We would like to thank Trevor Hodkinson and Andrew Jackson %Also Gavin Thomas, Emmanuel Douzery and Frédéric Delsuc if not authors.
-for his useful comments on the simulation protocol and Paddy Doyle for the assistance on using the computer cluster.
-All calculations were performed on the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing.
-This cluster was funded through grants from Science Foundation Ireland.
-
-%BIBLIOGRAPHY
- % The \cite command functions as follows:
- %   \citet{key} ==>>                Jones et al. (1990)
- %   \citet*{key} ==>>               Jones, Baker, and Smith (1990)
- %   \citep{key} ==>>                (Jones et al., 1990)
- %   \citep*{key} ==>>               (Jones, Baker, and Smith, 1990)
- %   \citep[chap. 2]{key} ==>>       (Jones et al., 1990, chap. 2)
- %   \citep[e.g.][]{key} ==>>        (e.g. Jones et al., 1990)
- %   \citep[e.g.][p. 32]{key} ==>>   (e.g. Jones et al., p. 32)
- %   \citeauthor{key} ==>>           Jones et al.
- %   \citeauthor*{key} ==>>          Jones, Baker, and Smith
- %   \citeyear{key} ==>>             1990
-
-\bibliographystyle{sysbio} %don't write the suffix
-\bibliography{References} %don't write the suffix
-
-\section{Supplementaries}
-\label{supplementaries}
-\subsection{Morphological characters states}
-In order to obtain a realistic probabilistic value for of \textit{k} characters states for each simulated morphological character, we downloaded 100 random morphological characters (with more than 100 characters each) from TreeBASE database (http://treebase.org/) published between 1985 and 2013 and covering 19 taxomomic classess (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
-We selected a total of 22563 characters ranging from 2 to 10 states.
-We calculated the proportion of characters with 2, 3, 4, 5, 6, 7, 8, 9 or 10 states.
-We then sampeled 22563 \textit{k} values between 2 and 10 with the same proportion of characters from the empirical data.
-We then used a simple t-test to check if our simulation was equal to the empirical data.
-In this study, we only simulated characters with 2 or 3 states because of the high proportion of ordered characters ecountered on characters with more than 3 states and the difficulties of simulate biologicaly sensible ordered characters.
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-AppendixCharacters.pdf}
-\caption{Character states distribution in empirical matrices. %Title?
-Characters states number distribution extracted from 100 random morphological matrices downloaded from RreeBase.}
-\label{Fig_AppendixCharacters}
-\end{figure}
-
-\subsection{Tree Building Software settings}
-
-\subsubsection{Maximum Likelihood - RAxML v8.0.20 \citep{Stamatakis21012014}} \\
-Model: \\
-Molecular data: \\
-GTR + $\Gamma_4$ (-m GTRGAMMA) \\
-Morphological data: \\
-Mk + $\Gamma_4$ (-K MK) \\
-Support: \\
-Rapid Boostrap algorithm (LSR), 1000 replicates \\
-
-\subsubsection{Bayesian - MrBayes v3.0.2 \citep{Ronquist2012mrbayes}} \\
-Priors: \\
-Molecular data: \\
-rates distribution shape ($\alpha$) = 0.5 \\
-Transition/Transversion ratio = 2 ($\beta$(80,40)) \\
-Starting tree: "True" tree topology with each branch length = 1 \\
-Morphological data: \\
-rates distribution shape ($\alpha$) = 0.5 \\
-Models: \\
-Molecular data: HKY + $\Gamma_4$ \\
-Morphological data: Mk + $\Gamma_4$ \\
-MCMC: \\
-2 runs \\
-4 chains per run \\
-generations < 50$\times$$1^6$ \\
-sample frequency = 1050$\times$$1^3$ \\
-ASDS diagnosis frequency = 50$\times$$1^3$ \\
-ASDS $<$ 0.01 \\
-ESS $>>$ 200 \\
-Burnin = 25\%
-
 \subsection{Triplets metric details ($T_{x,y}$)}
@@
 We repeated this 100 times independently using the same posterior distribution trees set each time resulting in 100 replicates of the same posterior distribution trees set compared 1000 times.
-We used an anova to test if there was no significant difference between the replicates so that the RBTC can be replicated.
+We used an ANOVA to test if there was no significant difference between the replicates so that the RBTC can be replicated.
 We applied this protocol on a poorly resolved tree (Low Score), a resolved tree with low support value (Medium Score) and a resolved tree with high support values (High Score).
@@
   \centering
-\begin{tabular}{rllrrrr}
-  \hline
-  Tree.Type & Used.metric & Replicates & Df & F.value & p.value \\ 
-  \hline
-  Low Score & RF & 100.00 & 99.00 & 0.74 & 0.98 \\ 
-  Low Score & Tr & 100.00 & 99.00 & 0.97 & 0.58 \\ 
-  Medium Score & RF & 100.00 & 99.00 & 0.64 & 1.00 \\ 
-  Medium Score & Tr & 100.00 & 99.00 & 0.45 & 1.00 \\ 
-  High Score & RF & 100.00 & 99.00 & 0.20 & 1.00 \\ 
-  High Score & Tr & 100.00 & 99.00 & 0.37 & 1.00 \\ 
-  \hline
-\end{tabular}
+  \input{SupplementaryMaterial/Supp_Tables/RPBTC_testing}
+  \label{RPBTC_testing}
 \end{table}
-\label{RPBTC_testing}
-
-\subsection{Codes}
-All codes are available at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions$
-The tree comparison results analysis can be repeated for more details at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Analysis$ %fully running!
-Try it at home (5-10 minutes of data loading)!
-
-\subsection{Full results}
-
-\begin{table}
-\caption{Tree similarity values per parameter in ML framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
-              & missing data &        &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.6714878 & 0.5692308  & 0.7894649 & 0.4689069  & 1.0361514 \\ %how many digits?
-              &              & Triples        & 0.9424133 & 0.8683742  & 0.9998559 & 0.6121399  & 1.0635250 \\
-              & 25\%         & Robinson-Fould & 0.5167595 & 0.4403147  & 0.6307692 & 0.2664429  & 0.8929104 \\
-              &              & Triples        & 0.5599494 & 0.7084874  & 0.8105781 & 0.2912950  & 0.9990703 \\
-              & 50\%         & Robinson-Fould & 0.4321852 & 0.3641026  & 0.5215627 & 0.2313246  & 0.6423846 \\
-              &              & Triples        & 0.5143009 & 0.4197127  & 0.5991498 & 0.2991432  & 0.9104526 \\
-              & 75\%         & Robinson-Fould & 0.3727090 & 0.3230769  & 0.4342249 & 0.2035474  & 0.5776286 \\
-              &              & Triples        & 0.4449563 & 0.3385128  & 0.5329149 & 0.1929395  & 0.8551570 \\
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.9072419 & 0.7668415  & 1.0000000 & 0.4070844  & 1.0478010 \\
-              &              & Triples        & 0.9763617 & 0.9448100  & 1.0039460 & 0.7158118  & 1.1043330 \\
-              & 25\%         & Robinson-Fould & 0.5722382 & 0.4883560  & 0.6923077 & 0.3253239  & 0.9367467 \\
-              &              & Triples        & 0.9370152 & 0.8207407  & 0.9938846 & 0.5394443  & 1.0868073 \\
-              & 50\%         & Robinson-Fould & 0.4338159 & 0.3025641  & 0.4760777 & 0.1704016  & 0.6683220 \\
-              &              & Triples        & 0.7327464 & 0.5947211  & 0.8739630 & 0.3418679  & 1.0251540 \\
-              & 75\%         & Robinson-Fould & 0.1534487 & 0.1155128  & 0.2000000 & 0.0683485  & 0.3230769 \\
-              &              & Triples        & 0.4588292 & 0.3905631  & 0.6111484 & 0.2903676  & 0.8568274 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.473285 & 0.3842965  & 0.8358974 & 0.3564753  & 0.9589744 \\
-              &              & Triples        & 0.9592557 & 0.8691580  & 1.026095 & 0.5720847  & 1.067524 \\
-              & 25\%         & Robinson-Fould & 0.4548038 & 0.3572132  & 0.6307692 & 0.2425339  & 0.9161212 \\
-              &              & Triples        & 0.9071925 & 0.7516934  & 0.9976224 & 0.4462168  & 1.0705291 \\
-              & 50\%         & Robinson-Fould & 0.4244263 & 0.3755493  & 0.4666667 & 0.2494927  & 0.8287312 \\
-              &              & Triples        & 0.6834366 & 0.5713927  & 0.8226139 & 0.2776690  & 1.0085944 \\
-              & 75\%         & Robinson-Fould & 0.2306888 & 0.1774508  & 0.3025641 & 0.1087764  & 0.5673541 \\
-              &              & Triples        & 0.4513642 & 0.3737397  & 0.5723858 & 0.2245224  & 0.7917946 \\
-    \hline
-\label{NTSML_full}
-\end{tabular}
-\end{table}
-
-\begin{table}
-\caption{Tree similarity values per parameter in Bayesian framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
-              & missing data &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\ %how many digits?
-              &              & Triples        & 0.04104231 & 0.00830962  & 0.05441173 & -0.03933377  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.03584309 & 03584196  & 03589744 & 02825643  & 0.06175676 \\
-              &              & Triples        & 0.03399849 & 0.003196469  & 0.04981018 & -0.049913382  & 0.11784366 \\
-              & 25\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01536739  & 0.06123106 \\
-              &              & Triples        & 0.03510953 & 0.002113358  & 0.04939228 & -0.041202264  & 0.09699332 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01538462  & 0.05641026 \\
-              &              & Triples        & 0.02849433 & 0.003263098  & 0.05298036 & -0.038344731  & 0.09250063 \\
-              & 75\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01069992  & 0.04338811 \\
-              &              & Triples        & -0.0001940566 & -0.01801301  & 0.04058783 & -0.06030922  & 0.10665934 \\
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\
-              &              & Triples        & 0.04104231 & 0.008309622  & 0.05441173 & -0.039333778  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.03593712 & 0.03198256  & 0.05641026 & 0.02798864  & 0.07692308 \\
-              &              & Triples        & 0.02639309 & 0.002185408  & 0.05061314 & -0.046563042  & 0.10547052 \\
-              & 25\%         & Robinson-Fould & 0.03588134 & 0.03587629  & 0.03589744 & 0.01538462  & 0.0680884 \\
-              &              & Triples        & 0.03856296 & 0.002185408  & 0.05061314 & -0.046563042  & 0.10547052 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01536289  & 0.05879269 \\
-              &              & Triples        & 0.01403716 & -0.01100608  & 0.03972324 & -0.03800324  & 0.12405207 \\
-              & 75\%         & Robinson-Fould & 0.03589206 & 0.03588825  & 0.03589744 & 0.01538462  & 0.0411843 \\
-              &              & Triples        & 0.02939363 & 0.01902351  & 0.04238907 & -0.03087905  & 0.08946917 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\
-              &              & Triples        & 0.04104231 & 0.008309622  & 0.05441173 & -0.039333778  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.05637057 & 0.03589744  & 0.05927956 & 0.02813296  & 0.07692308 \\
-              &              & Triples        & 0.03008132 & 0.01176800  & 0.04847400 & -0.02451461  & 0.09453754 \\
-              & 25\%         & Robinson-Fould & 0.03593743 & 0.03589744  & 0.03596018 & 0.02330790  & 0.07692308 \\
-              &              & Triples        & 0.02010221 & -0.001145118  & 0.0377779 & -0.060183332  & 0.1098632 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.03050289  & 0.05950426 \\
-              &              & Triples        & 0.01053959 & -0.005595947  & 0.04510444 & -0.036278820  & 0.10478805 \\
-              & 75\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01115983  & 0.04362927 \\
-              &              & Triples        & 0.02427778 & 0.0007405737  & 0.04260522 & -0.0373004106  & 0.1048756 \\
-\label{NTSBA_full}
-\end{tabular}
-\end{table}
-
-\subsection{Bootstraps distribution}
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-AppendixBootstraps.pdf}
-\caption{Bootstraps distribution across the "best" trees.}                     % MISSING 8 CHAINS
-\label{Fig_AppendixBoostraps}
-\end{figure}
-
-%END
-\end{document}

### 2014-06-17 Manuscript/Tables/Group_results.tex
Tables init

--- 2014-06-13 Manuscript/TEM_draft.tex
@@
-%HEADER
-\documentclass[12pt,letterpaper]{article}
-
-%packages
-% NC: Just to flag here how I'll comment on LaTeX files. I'll use % and then NC. Of course in git diff it'll all show up as an addition.
-% NC: You could add all the packages etc to a template .sty file to save writing it all up here. Rich's ms.sty that I provided on DropBox is a good example
-\usepackage{fixltx2e}
-\usepackage{textcomp}
-\usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{verbatim}
-\usepackage[english]{babel}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{setspace}
-\usepackage{lscape}
-\usepackage{indentfirst}
-\usepackage[normalem]{ulem}
-\usepackage{booktabs}
-%\usepackage{nag}
-\usepackage{natbib}
-%\usepackage{bibtex}
-\usepackage{float}
-\usepackage{latexsym}
-%\usepackage{hyperref} 
-\usepackage{url}
-%\usepackage{html}
-\usepackage{hyperref}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{bm}
-\usepackage{array}
-\usepackage[version=3]{mhchem}
-\usepackage{ifthen}
-\usepackage{caption}
-\usepackage{hyperref}
-%\usepackage{xcolor}
-\usepackage{amsthm}
-\usepackage{amstext} % Add and remove packages as necessary for your manuscript.
-
-% NC: Added by NC
-\usepackage{enumerate}
-\usepackage[osf]{mathpazo} % palatino is a nicer font! 
-\pagenumbering{arabic}	% Page numbers
-
-%Pagination style and stuff
-\linespread{1.66} % All text should be double-spaced with occasional exceptions for tables. 
-\raggedright
-\setlength{\parindent}{0.5in}
-
-\setcounter{secnumdepth}{0} % Our sections are not numbered and our papers do not have Tables of Contents.
-We don't present a list of figures or list of tables, either.
-Any common font is fine.
-(A common sans-serif font should be used on figures, but figures should be separate from the LaTeX document.)
-
-%\pagestyle{empty}
-
-\renewcommand{\section}[1]{%
-\bigskip
-\begin{center}
-\begin{Large}
-\normalfont\scshape #1
-\medskip
-\end{Large}
-\end{center}}
-
-\renewcommand{\subsection}[1]{%
-\bigskip
-\begin{center}
-\begin{large}
-\normalfont\itshape #1
-\end{large}
-\end{center}}
-
-\renewcommand{\subsubsection}[1]{%
-\vspace{2ex}
-\noindent
-\textit{#1.}---}
-
-\renewcommand{\tableofcontents}{}
-
-\bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
-% We don't use a special title page; the author information is entered like any other text.
-% FOOTNOTES: We don't allow them in the manuscript, except in tables. Don't include any footnotes in the text.
-%IMPORTANT:
-%Note on display in the text file: sentences or sub-parts of sentences (e.g. enumerations) are delimited by a newline for clarity.
-%However, true new lines (i.e. in the pdf) are delimited by one or more empty lines.
-%We don't use a heading for the introduction. The nature of this section is implied by its position following the abstract and preceding the first major section heading.
-
-% NC: You should delete this extraneous info from the Syst Biol template to tidy everything up a bit
-
-\begin{document}
-
-%For a colour version of the draft to a search>replace '-BW'>'' 
-% %\modulolinenumbers[1] 	% Line numbering on every line
-
-%Running head
-\begin{flushright}
-Version dated: \today
-\end{flushright}
-\bigskip
-\noindent RH: Missing data and topology in total evidence matrices %RH = Running Head
-
-\bigskip
-\medskip
-\begin{center}
-
-%Title
-\noindent{\Large \bf Effect of missing data in matrices containing living and fossil taxa on topological accuracy}
-%Alternative punchier title: "Bayesian total evidence methods fail to recover correct topologies from matrices containing both living and fossil taxa"
-%Alternative punchier title: "Maximum Likelihood outperforms Bayesian methods for inferring topology from total evidence matrices containing both living and fossil taxa"
-
-% NC: Not sure about any of these.
-I actually think I prefer your talk title.
-So something like: The effects of missing data on the ability of total evidence approaches to recover correct tree topologies.
-No rush to decide.
-But needs total evidence in there somewhere.
-\bigskip
-
-%Authors
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$, other authors $^3$, and Natalie Cooper$^1$$^,$$^2$}\\
-\noindent {\small \it 
-$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland;\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland;\\
-$^3$Somewhere else }\\
-\end{center}
-\medskip
-\noindent{\bf Corresponding author:} Thomas Guillerme, School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.\\
-\vspace{1in}
-% NC: It's not called the Republic of, just Ireland!
-
-\newpage
-\begin{abstract}
-% NC: I'm not going to give this much work as it'll change before we submit anyway. Note though that and abstract must NOT just repeat stuff in the intro. Needs rephrasing at the very least.
-Living species represent less than 1\% of all species that have ever lived.
-% NC: I was chatting to Andy Purvis about this the other day and he said lots of palaeo folk hate that stat.
-So let's not start the abstract with it!
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology.
-This fact has led to an increasing consensus among scientists that both fossil and living taxa must be included in macroevolutionary studies.
-One approach, the total evidence approach, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both fossil and living taxa at the tips.
-Although the total evidence approach seems very promising, it requires a lot of data and is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.
-
-In this study we assess the effect of missing data on tree topologies inferred from total evidence matrices.
-Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix:
-(1) the proportion of living taxa with no morphological data,
-(2) the amount of missing data in the fossil record, and
-(3) the overall number of morphological characters in the matrix.
-We find that, in a Bayesian framework, difficulties in recovering a stable topology are mainly driven by the missing data in the molecular part of the matrix (for which fossil taxa have no data).
-In a Maximum Likelihood framework, however, topology is not directly affected by missing data \textit{per se}, but by the number of morphological characters shared among the taxa.
-Therefore, the two main drivers of incorrect topologies are the overall number of morphological characters and the number of living species with no morphological data.
-
-Our results suggest that, in order to use total evidence approaches, one should reduce the missing data in the morphological part of the matrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process.
-%More results!
-\end{abstract}
-
-\noindent (Keywords: missing data, total evidence, Bayesian, Maximum Likelihood, topology)\\
-
-\vspace{1.5in}
-
-\newpage
-
-\section{Introduction}
-
-Although most species that have ever lived are now extinct % NC: This avoids giving the dodgy stat of 1%
- \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species (e.g., \citealp{meredithimpacts2011}).
-%or is citing of your own papers in the too "douchebagy" for the second sentence?
-% NC: Yeah be careful with self citing. Fine if it's a good fit, but it's totally not here!!! A better citation would be a big biodiversity study, e.g. the bird phylogeny paper.
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
-This has led to increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments (e.g.
-\citealp{pyrondivergence2011,ronquista2012,schragocombining2013}).
-%Not a nice sentence
-
-Up to now, three main approaches have been used to place both living and fossil taxa % NC: BE consistent. Either living and fossil, or fossil and living.
-into phylogenies.
-These approaches differ mainly in whether they treat fossil taxa as tips or as nodes in the phylogeny, and in which part of the available fossil data is used (i.e.
-the age of the fossil only or both its age and morphology).
-% NC: Not sure about bit in brackets
-Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
-Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \citep{simpson1945}.
-This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty (but see \citealp{spencerefficacy2013}).
-Neontologists, on the other hand, more commonly use probabilistic approaches (e.g.
-Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates % NC: Need to check the exact phrase needed here.
-Is it range of occurrence?
-Occurrence dates?
-etc.
- are used to time calibrate phylogenies \citep{zuckerkandl1965}.
-There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013}.
-However neither approach uses all the available data.
-
-A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa.
-%There might be a philosophico-semantic distinction to do here.
-Is TEM a method or an approach?
-I'll say that it is an approach because it's more like an idea (put every thing together) than a way to do (how to put everything together).
-It makes more sense to me compared to a real method e.g.
-a probabilistic method such as ML where the method states precisely how to do stuff.
-% NC: Tricky. Don't think it's too important though. 
-% NC: I do think Total Evidence needs capital letters though, like ML and Bayesian
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{eernissetaxonomic1993}.
-Total Evidence methods have been successfully applied to empirical data \citep{pyrondivergence2011,ronquista2012,schragocombining2013}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
-However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data.
-In particular it requires morphological data from both living and fossil taxa, both of which are known to be scarce.
-% NC: Really?
-I know fossil data is known to be scarce, but living data?
-I think people would be surprised.
-And they are surprised.
-We were surprised!
-Needs a citation either way.
-Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
-% Trevor: Missing data also influences support values that trend to decrease
-
-% NC: You need to take a bit of time to learn about singular and plural grammar in English.
-It's something you always do wrong.
-e.g.
-The Total Evidence method *uses* this and *its* parameters are as follows.
-Versus: Total Evidence methods *use* this and *their* parameters are as follows.
-Sive and I won't be around to correct your English forever.
-
-The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013}.
-Missing molecular data has been seen by some authors as an issue because it can decrease phylogenetic signal (i.e. % define phylogenetic signal here) 
-in some parts of the tree, especially when using large matrices \citep{lemmonthe2009}. However, this may not be a major issue because phylogenetic signal is easily increased by:
-(i) including a "modest" number of highly-covered % what do you mean by highly-covered?
-genes (i.e. approximatively half of the genes; \citealt{rouresite-specific2011});
-(ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and
-(iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
-Similarly, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study in question \citep{wiensmissing2003,sansomfossilization2013}.
-Because soft-tissue characters are rarely preserved in the fossil record, missing data is mainly found in these characters, and is therefore not randomly distributed which can lead to biased placement of fossil taxa in phylogenies \citep{sansomfossilization2013}.
-However, the phylogenetic signal is not related to the amount of missing data \textit{per se} but to the number of informative characters for each taxon, therefore missing data is less of an issue than the number of shared informative characters \citep{wiensmissing2003}.
-% NC: I'm a little dubious of what this paragraph is trying to get at in places. We should discuss at some point.
-
-Although missing data does not appear be a major problem in molecular and morphological matrices separately \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil species.
-This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data.
-Until now, no attempt has been made to study the impact of this issue on phylogenetic inference from Total Evidence methods.
-
-% NC: Somewhere here you need to flag that you're only looking at topologies and briefly explain why.
-
-Here we use simulations to assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
-The molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}.
-The effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}, therefore, we focus only on missing data in the morphological part of the matrix.
-We investigate three major parameters that directly affect the completeness of the morphological part of the matrix:
-\begin{enumerate}
-\item the proportion of living taxa with no morphological data;
-\item the amount of missing data in the fossil taxa; and
-\item the overall number of morphological characters for both living and fossil taxa in the matrix.
-\end{enumerate}
-We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
-We chose these parameters because they reflect real-life biases in data availability.
-The advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
-Missing data in fossil taxa is very common due to preservation biases (citation?), and
-% OK I'm not sure what the "real-life" reason for three is.
-%This paragraph is crap
-% NC: I think it's fine!
-
-We find that when using a Maximum Likelihood approach, as missing data increases, the likelihood of recovering the correct tree topology decreases.
-However, even with no missing data, Total Evidence matrices dramatically reduce the performance of Bayesian methods for inferring tree topology.
-We propose that this drastic difference between Bayesian and Maximum Likelihood methods is due to a flattening of the likelihood landscape caused by the unavoidable amount of missing molecular data for fossil taxa in a Total Evidence matrix.
-We make suggestions for how best to deal with this issue when inferring phylogenies from Total Evidence matrices.
-
-% NC: You don't want to tell the whole story here. Just a results summary and a quick line of interpretation
-
-%This problem leads to an inability of Bayesian chains to pick up the sub and near optimal topologies leading to high variance in the posterior tree distribution and really low support in the consensus tree.
-%These results should encourage authors in not using Bayesian inference to explore the topological landscape of their phylogeny.
-%Instead, we suggest that authors should first find the Maximum Likelihood tree and use it's topology as a fix prior in their Bayesian inference.
-%We also propose a new method for comparing Bayesian posterior tree distribution's topology available in a wonderful R package. %Or am I just dreaming a bit too much here?
-
-% NC: Dreaming a bit much I think.
-There's a discussion to be had here about whether it's really a *new* method and whether you really want to make an R package per se.
-We definitely want to share the code
-
-%-----------
-% NC: It might look like I've done a lot to this but actually I think it's nearly there.
-A good thing to take from this is that writing a paper is *hard*.
-It gets easier, but it's still a deeply involved process with lots of writing and re-writing.
-Introductions are particularly difficult because if you don't set the scene, it's hard to convince people to get on board for the rest of the paper.
-This is why I said there was no way this was being submitted by the end of June!
-Still a long way to go I'm afraid in terms of time and effort.
-But that doesn't mean you're not doing well, *every* project is like this.
-2 years is about average from conception to sending out a manuscript, and you only started work on this last spring.
-%------------
-
-\newpage
-\section{Methods} %no material needed?
-% NC: Depends on the journal. See what Syst Biol asks for as section titles.
-To explore how missing data in Total Evidence matrices influences tree topology we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
-%NC: does Syst Biol use Fig.
-or Figure?
-\begin{enumerate}
-\item{Generating the matrix} \label{step:generate_matrix} \\
-We randomly generated a birth-death tree (hereafter called the "true" tree; Table ~\ref{Tab_glossary}) and used it to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix; Table ~\ref{Tab_glossary}).
-\item{Removing data} \label{step:remove_data} \\
-We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of missing living taxa ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$) % NC: make sure to use the same descriptions here and in the introduction
-(the resulting matrices are called hereafter "missing-data" matrices; Table ~\ref{Tab_glossary}).
-\item{Building phylogenies} \label{step:build_phylo} \\
-We built phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree; Table ~\ref{Tab_glossary}) and multiple trees inferred from matrices with missing morphological data (hereafter called the "missing-data" trees; Table ~\ref{Tab_glossary}).
-Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
-\item{Comparing topologies} \label{step:compare_topo} \\
-We compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies.
-\end{enumerate}
-To measure the effect of missing data distribution, we repeated steps ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} with the exact same fixed parameters 51 (3$\times$17) times.
-%that's because I had 3 cluster accounts available...
-If it's a too odd number, no worries, I'll just remove the last chain.
-A list of all the terms used in this paper is available in table ~\ref{Tab_glossary}.
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_outline-BW.pdf}
-\caption{Protocol outline.
-(1) We generated a random tree (the "true" tree) to infer a matrix with no missing data (the "complete" matrix).
-(2) We removed data from the morphological partition of the "complete" matrix resulting in 125 "missing-data" matrices.
-(3) We infer phylogenetic trees from each matrix in both ML and Bayesian framework.
-(4) We then compared the "missing-data" trees to the "best" tree.
-We repeated step ~\ref{step:generate_matrix} to ~\ref{step:compare_topo} 51 (3$\times $17) times.}
-\label{Fig_Outline}
-\end{figure}
-
-\begin{table}[ht]
-\caption{Glossary}
-\centering
-\begin{tabular}{p{5cm} p{9cm}}
-\hline
-\hfill Term & Definition \\[0.5ex]
-\hline
-\hfill living taxa & taxa with both molecular and morphological data available \\[1.5ex]
-\hfill fossil taxa & taxa with only morphological data available \\[1.5ex]
-\hfill "complete" matrix & matrix with no missing data except for the molecular part of the fossil taxa \\[1.5ex]
-\hfill "missing-data" matrix & matrix with various amount of missing data \\[1.5ex]
-\hfill $M_{L}$ & missing living taxa in the morphological part of the matrix \\[1.5ex]
-\hfill $M_{F}$ & missing data for the fossil taxa in morphological part of the matrix \\[1.5ex]
-\hfill $M_{C}$ & missing morphological characters for both living and fossil taxa \\[1.5ex]
-\hfill "true" tree & tree used to simulate the matrix \\[1.5ex]
-\hfill "best" tree & tree inferred from the "complete" matrix \\[1.5ex]
-\hfill "missing-data" tree & tree inferred the "missing-data" matrices \\[1.5ex]
-\hfill RPBTC & Random pairwise Bayesian tree comparison \\[1.5ex]
-\hfill "horizontal" sub-matrix & sub-matrix with morphological and molecular characters for living taxa \\[1.5ex]
-\hfill "vertical" sub-matrix & sub-matrix with morphological data for both living and fossil taxa \\[1.5ex]
-\hfill "corner" sub-matrix & sub-matrix with morphological data for living taxa \\[1.5ex]
-\hfill "missing-data" sub-matrix & molecular part of the matrix for the fossil taxa (no data) \\
-\end{tabular}
-\label{Tab_glossary}
-\end{table}
-
-%---------------------------------------------
-%       Generating the matrix
-%---------------------------------------------
-
-\subsection{Generating the matrix}
-First we randomly generated a "true" tree of 50 taxa in R v3.0.2 \citep{R302} using the package diversitree v0.9-6 \citep{fitzjohndiversitree2012}.
-We generated the tree using a Birth Death process by sampling the values of the speciation events ($\lambda$) and extinction events ($\mu$) from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
-We implemented a rejection sampling algorithm to select only random trees with 25 living and 25 fossil taxa. %Justification needed?
-We then added a taxa to the resulting Birth-Death tree as the outgroup of the tree.
-The mean branch length of the tree was used to separate the outgroup from the rest of the taxa and the branch length leading to the outgroup was set as the sum of the mean branch length and the longest root-to-tip length of the tree.
-
-Next, we created a molecular and a morphological matrix from the "true" tree.
-The molecular matrix was inferred from the "true" tree using the package phyclust v0.1-14 \citep{chen2011}.
-The matrix was made of 1000 characters sites for 51 taxa and generated using the seqgen algorithm \citep{ranbaut1997seqgen}.
-We used the HKY model \citep{HKY85} with a random base frequencies and with the transition/transversion rate of 2 \citep{douadycomparison2003} as parameters for generating the matrix.
-The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
-We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
-These parameters were selected to generate data with no special assumption about how the characters evolved as well as to reduce the computational time required if these parameters were estimated rather than defined (total computational time $>$ 65 CPU years).
-
-We inferred the morphological matrix using the ape package v3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
-We assigned the number of character states (either 2 or 3) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of characters states extracted from 100 published empirical morphological matrices (See \hyperref[supplementaries]{supplementaries}).
-We then ran an independent discrete character simulation for each character using the "true" tree branch length and topology with the randomly selected number of states (2 or 3) and assuming an equal rate of change (i.e.
-evolutionary rate) from one character state to an other \citep{Pagel22011994}.
-This method allows us to have only two parameters per character: the number of states and the evolutionary rate.
-For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used a low evolution rate parameter (i.e.
-$\alpha$) in order to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
-
-All the molecular information for fossil taxa was replaced by missing data ("?").
-Finally, we combined the morphological and molecular matrices obtained from the "true" tree.
-Hereafter we call this the "complete" matrix: the matrix with no missing data except for the molecular data of the fossil taxa.
-
-%---------------------------------------------
-%       Removing data
-%---------------------------------------------
-
-\subsection{Removing data}
-Once we obtained the "complete" matrix we modified it to get a set of matrices with missing data. %Ugly sentence
-We randomly replaced data with "?" in the morphological part of the matrices according to the following parameters (Fig. ~\ref{Fig_RemoveData}):
-
-\begin{enumerate}
-\item{The proportion of living taxa with no morphological data ($M_{L}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological one.
-Because of the increasing facility to sequence DNA for living taxa, the number of living taxa with molecular data is highly superior the the number of taxa with molecular and morphological data.
-\item{The proportion of missing morphological data across all fossil taxa ($M_{F}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the quality of the fossil record. 
-\item{The proportion of missing morphological characters across all taxa (living and fossil - $M_{C}$): 0\%, 10\%, 25\%, 50\% or 75\%. }
-This parameter illustrates the number of available morphological characters for both living and fossil taxa.
-\end{enumerate}
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_missingData-BW.pdf}
-\caption{Names of the different parts of the matrix.
-A: Missing data parameters:
-Missing living - The proportion of living taxa with no morphological data ($M_{L}$);
-Missing fossil - The proportion of missing morphological data across all fossil taxa ($M_{F}$);
-Missing character - The proportion of missing morphological characters across all taxa (living and fossil) ($M_{C}$).
-B: Different parts of the matrix:
-The "horizontal" sub-matrix (orange) contains molecular and morphological data for living taxa;
-The "vertical" sub-matrix (blue) contains morphological data for living and fossil taxa;
-The "corner" sub-matrix (orange and blue striped) contains morphological data for living taxa;
-The "missing-data" sub-matrix (grey) is the molecular part of the fossil taxa and contains no data.}
-\label{Fig_RemoveData}
-\end{figure}
-
-In practice, each parameter represent a different way of removing data in the morphological part of the matrix: $M_{L}$ removes a proportion of rows from the living taxa; $M_{F}$ removes a proportion of cells from the fossil taxa; and $M_{C}$ removes a proportion of columns across both living and fossil taxa (see Fig.
-~\ref{Fig_RemoveData}).
-Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a proportion of the living taxa is removed (i.e.
-removing rows), as for $M_{F}$, a proportion of data is removed across the whole of the morphological matrix for fossil taxa (i.e.
-removing cells).
-
-We tested all parameters combinations resulting in 125 ($5^3$) matrices.
-Because some parameter combinations introduce a lot of missing (e.g. $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
-When this occurred we repeated the random deletion of characters until every taxa had at least 5\% data across the whole matrix.
-
-%---------------------------------------------
-%       Building phylogenies
-%---------------------------------------------
-
-\subsection{Building phylogenies}
-From the resulting matrices we generated two types of trees, the "best" tree that is inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
-The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
-The "best" tree, on the other hand, is the best tree we can build using the state-of-the-art phylogenetic methods.
-In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g. \citet{rozen2005}).
-Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
-
-\subsubsection{Maximum Likelihood}
-The "best" tree and the "missing-data" trees were inferred using RAxML v8.0.20 \citep{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model (\citet{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citet{Stamatakis21012014}) as a generalisation of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
-The GTR model can be seen as a generalisation of the HKY model (the 2 parameters from the HKY model are implicitly included in the 6 from GTR model - \citet{stamatakisa2008}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} which is a generalisation of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014}).
-We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. %The explanations of the algorithm part for the BS was proposed by Trevor
-The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists in pruning one sub-tree from the tree and subsequently reinserting it to al neighbouring branches \citep{stamatakisa2008}.
-Sub-tree Pruning and Reinserting methods (SPR) have been demonstrated as being better than others (e.g. Nearest Neighbouring Interchange - NNI) in recovering good bootstrap values \citep{salamin2003}.
-
-\subsubsection{Bayesian}
-The "best" tree and the "missing-data" trees were inferred using MrBayes v3.2.2 \citep{Ronquist2012mrbayes}.
-We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
-The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \citep{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
-For the morphological data, we used the Markov \textit{k} state model \citep{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
-We chose these models to be consistent with the parameters used to generate the "complete" matrix.
-
-Each Bayesian tree was estimated using two runs of four chains each for a maximum of 50$\times$$1^6$ generations.
-We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
-The effective sample size (ESS) was also checked on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
-For each run, we removed 25\% of the iterations as burn-in.
-We used the following prior for each tree (see \hyperref[supplementaries]{supplementaries}):
-\begin{enumerate}
-\item
-the "true" tree’s topology as a starting tree (with a starting value for each branch length of 1),
-\item
-an exponential prior on the shape of the gamma distribution of $\alpha$=0.5 for both partitions
-\item
-and a transition/transversion ratio prior of 2 sampled from a strong beta distribution ($\beta$(80,40)).
-\end{enumerate}
-
-We used these prior to speed up the Bayesian process.
-These prior biased the way the Bayesian process calculated the branch length by giving non-random starting points and boundaries for the parameters estimation process, however, in this study, we focused on the effect of missing data on the topology and not on the branch length.
-Even using these prior, it took 65 CPU years to build 51 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
-
-%---------------------------------------------
-%       Comparing Topologies
-%---------------------------------------------
-
-\subsection{Comparing topologies}
-We compared the topology of the "missing-data" trees inferred from the matrices with missing data to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$.
-Note that we only investigate differences in topology and not in branch length because the aim of this study is to look at the effect of missing data on the topology of trees inferred from total evidence type matrices.
-To compare the topology of the resulting trees, we used two metrics to assess number of conserved taxa and clades position using respectively the Triples \citep{dobson1975triplets} and the Robinson-Fould \citep{RF1981} distance.
-We normalised the two metrics using the Normalised Tree Similarity index \citep{Bogdanowicz2012} to generalise our results for any \textit{n} number of taxa.
-The three metrics are detailed below.
-
-\subsubsection{Triples distance ($T_{x,y}$) \citep{dobson1975triplets}}
-This metric measures the number of different sub-trees of three taxa between two given trees.
-Each triplet can be written as $I_{ijk}$=(\textit{ijk}).
-Where $I_{ijk}$ is equal to 0 if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to 1.
-For any rooted binary tree there are only three possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
-If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k});.
-One can calculate $S_n$, the triplet distance between two trees as:
-\begin{equation}
-S_n = \sum_{ijk} I_{ijk}
-\end{equation}
-Where:
-\begin{equation}
-\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
-\end{equation}
-And where \textit{n} is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
-If $S_n$=0, the trees are the same (i.e. no taxa as been displaced).
-When $S_n$ = $\binom{n}{4}$, the trees are the most different possible (i.e. every taxa as been displaced).
-This metric therefore illustrates the amount of displayed taxa.
-It is less sensitive to the placement of individual taxa and to taxa of highly uncertain placement (e.g.
-fossil taxa) than the Robinson-Foulds metric \citep{critchlowthe1996,johnson1998,wiensmissing2003}.
-Therefore we used this metric as a proxy to estimate the robustness of the tree to flying taxa (see \hyperref[supplementaries]{supplementaries}).
-
-\subsubsection{Robinson-Fould distance (\textit{RF}) \citep{RF1981}}
-This metric measures the number of shared clades among two trees and therefore illustrates the number of exactly conserved groups among the trees.
-The Robinson-Fould distance (also called path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
-\begin{equation}
-RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
-\end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees.
-The minimal value of \textit{C} is 1 if the two trees have the same \textit{n} taxa; the maximal value in \textit{C=n-2}.
-This metric is more sensitive to taxa displacement than the Triples metric (i.e.
-if one taxa gets out of a clade, then the clades are no longer considered as similar - \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
-Therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades (see \hyperref[supplementaries]{supplementaries}).
-
-\subsubsection{Normalised Tree Similarity \textit{NTS} \citep{Bogdanowicz2012}}
-For any tree with \textit{n} taxa compared using a tree distance metric \textit{m}, $NTS_m$ represents the similarity score between the two trees given the expected distance between two random Yule trees with \textit{n} taxa.
-Let $\bar{d}_{m,n}$\textit{(rand)} be the average distance between two random Yule trees with \textit{n} taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} containing each \textit{n} taxa, then:
-\begin{equation}
-NTS_{m,n}(x,y)=\frac{\bar{d}_{m,n}(rand) - d_{m,n}(x,y)} {\bar{d}_{m,n}(rand)}
-\end{equation}
-\textit{NTS} ranges from 1 to -$\infty$.
-For any \textit{m},\textit{n} when \textit{NTS}=1, the trees are the same;
-when \textit{NTS}=0 the trees are not more different than expected by chance;
-when \textit{NTS}$<$0, the trees are more different than expected by chance (see \hyperref[supplementaries]{supplementaries}).
-
-$\newline$
-
-We compared the "missing-data" tree to the "Best" tree for each chain.
-For the Maximum Likelihood trees we performed pairwise comparisons between the "Best" tree and the "missing-data" tree (see Table ~\ref{Tab_glossary}) for both the Robinson-Fould and the Triples metric.
-We calculated the difference between the trees using the metrics described above by using the TreeCmp java script \citep{Bogdanowicz2012}.
-For each metric, we normalized the value using the Normalised Tree Similarity scaled with the mean value of 1000 pairwise random tree comparisons for the same metric and the same number of taxa \textit{n}=51 (see \hyperref[supplementaries]{supplementaries}).
-We ran the comparison for every "missing-data" tree in each chain resulting in 51 comparisons for every "missing-data" trees. 
-We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package (v3.1 – \citet{hdrcde}).
-
-Bayesian tree inference allows to account for the statistical uncertainty of a phylogenetic tree by not using an optimal criterion (c.f.
-Maximum Likelihood) and gives a tree posterior distribution (c.f.
-the likeliest tree in ML).
-This method has the clear advantage of better dealing with error and uncertainty but it's output (a distribution rather than a single tree) is less practical to use for any further study (but see \citet{healyecology2014}).
-%Awwwwww yiis!
-To avoid this problem, people traditionally use the a consensus tree build on a majority rule in order to summarise a Bayesian tree posterior distribution.
-This results into a single tree containing both topological and branch length information as well as support information (i.e. posterior probabilities, e.g. \citet{Ronquist2012mrbayes}).
-However, using a Bayesian consensus tree has limitation especially if the resulting consensus tree is not well supported or resolved.
-Because the metrics used in this study are used to measure variation in topology between two trees (i.e.
-taxa placement or clade position), comparing two Bayesian consensus trees is not optimal in picking topological difference signal.
-For example, if one of the trees is not resolved at all (i.e. a star tree), comparing it to any other tree (even an other star tree) will not give useful results.
-Therefore we used the entire Bayesian trees posterior distribution in order to perform the tree comparisons (i.e.
-the "Best" Bayesian tree posterior distribution vs.
-the "missing-data" Bayesian tree posterior distribution).
-
-\subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
-We compared the Bayesian posterior distribution using a Random Pairwise Bayesian Tree Comparison (RPBTC) method.
-This method consists in comparing a series of randomly selected pairs of trees from two posterior distributions (on from each distribution) and use the mode to summarize the resulting distribution as a proxy of the difference between the two trees in a Bayesian framework.
-\begin{equation}
-RPBTC_{X,Y}=Mo[(]d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]
-\end{equation}
-
-Where \textit{X} and \textit{Y} are two Bayesian tree posterior distributions;
- $d_{m,n}$ is the pairwise difference for any metric \textit{m} and \textit{n} taxa between $X_{i}$ and $Y_{j}$ which are two single trees randomly sampled respectively from \textit{X} and \textit{Y};
- $Mo$ is the mode and
- $Mo[(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]$ is the mode of the pairwise difference repeated \textit{k} times.
-
-We used the mode to summarize the distributions because it is the value that is the most represented in the distribution and reflects what a consensus tree represents (the topology the most represented in the posterior distribution).
-Also, the mode is a value present in the distribution (which, depending on the metric, can be a series of discrete values) in contrast to the other distribution summary metrics, such as the mean or the median, which can take values that are not actually present in the distribution.
-For example, if comparing two posterior tree distribution, we obtain difference values of 10, 10, 10, 5, 5 and 1, using the mean (6.8) or the median (7.5) gives a values that actually doesn't exist.
-%Useful?
-
-In this study, we calculated RPBTC for both metrics (Robinson-Fould and Triples distance) for each pairs of Bayesian tree distributions (i.e.
-the "Best" Bayesian tree posterior distribution vs.
-the "missing-data" Bayesian tree posterior distribution) with 1000 random pairwise comparisons (\textit{k}).
-Because the RPBTC uses random pairwise comparisons, there is a chance of comparing always trees that have the biggest or the smallest difference between two distributions, either deflating or inflating the RPBTC difference value.
-However, using 1000 random pairwise comparisons makes the difference stable (i.e. if repeated independently, no difference is detected - see \hyperref[supplementaries]{supplementaries}).
-
-$\newline$
-
-We calculated the NTS for the "Best" Bayesian tree vs.
-"missing-data" Bayesian tree comparison for both Robinson-Fould and Triples metric using the mode of the RPBTC for each chain resulting in a distribution of 51 modes per comparison.
-We then compared the NTS of the Bayesian trees to the NTS of the ML trees for both metrics for each comparison.
-This resulted in a distribution of 51 NTS values for the Bayesian trees and 51 NTS values for the ML trees for each tree comparison for which we calculated the mode and the 50\% and 95\% confidence intervals (see \hyperref[supplementaries]{supplementaries} - Fig.
-~\ref{Fig_Compare}).
-
-\begin{figure} % Modify the figure
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp-BW.pdf}
-\caption{Tree comparison protocol.
-From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we inferred a tree in Baysian or Maximum likelihood framework.
-We then compared this tree with or without fossil/living taxa to the "best" tree.
-In the Bayesian framework, 1000 trees from the posterior distribution where used for each comparison.}
-\label{Fig_Compare}
-\end{figure}
-
-\subsubsection{Statistical difference between tree topologies} %note that in the end, I only have non-parametric tests, the parametric explanation part can go to the supplementaries
-To assess the difference between the tree topologies, whether they were inferred in a Bayesian or a ML framework, we performed a group comparison test for each set of configurations (e.g.
-effect of $M_L$ in Bayesian inference, combined effect of $M_F$ and $M_C$ in ML inference, etc...).
-We used the pairwise tree comparisons as factors (i.e.
-the "Best" tree vs.
-one of each of the 125 "missing-data" trees) and the values of the 51 replicates from the pairwise comparison as a response variable.
-The 51 replicates values were either the replicates of the 51 pairwise tree comparisons in ML or the modes of the 51 RPBTC posterior distributions in Bayesian.
-When we found a significant difference between the groups, we performed a pairwise comparison test to analyse which pairs of groups differ from the others.
-Depending on whether the data was normal with variance homoscedasticity or not (controlled by performing respectively a shapiro.test and a a bartlett.test in R\{stats\}), we used respectively parametric or non-parametric tests for the group comparison and the pairwise comparisons as suggested by \citet{ruxtontime2008} (Table ~\ref{Stats_test}).
-%link broken
-
-\begin{table}[ht]
-\caption{Statistical tests}
-\centering
-\begin{tabular}{lrll}
-  \hline
-   & test & statistics & code\{package\} \\ [0.5ex]
-  \hline
-  Groups comparisons & ANOVA & parametric & anova\{stats\} \\ [0.5ex]
-                     & Kruskal-Wallis & non-parametric & kruskal.test\{stats\} \\[0.5ex]
-  \hline
-  Pairwise comparisons & Tukey HSD & parametric & TukeyHSD\{stats\} \\ [0.5ex]
-                       & Nemenyi-Damico-Wolfe-Dunn & non-parametric & kruskalmc\{pgirmess\} \\ [0.5ex]
-  \hline
-\end{tabular}
-\label{Stats_test}
-\end{table}
-
-\subsection{Effect of missing molecular characters for fossil taxa} %Awfull name.
-To assess the effect of missing molecular characters for fossil taxa in a Bayesian framework, we split the "complete" matrix in sub-matrices containing no missing data at all (see Fig.
-~\ref{Fig_RemoveData}):
-\begin{enumerate}
-\item
-A first containing both molecular and morphological data for living taxa only (hereafter called the "horizontal" sub-matrix, Table ~\ref{Tab_glossary});
-\item
-A second one containing morphological data for both living and fossil taxa (hereafter called the "vertical" sub-matrix, Table ~\ref{Tab_glossary});
-\item
-A third one containing morphological data for living taxa only (hereafter called the "corner" sub-matrix, Table ~\ref{Tab_glossary});
-\end{enumerate}
-We reran the Bayesian tree inference on the different sub-matrices with no missing data in the same way as described above.
-We then compared the resulting tree posterior distribution to itself (in the same way described above) to assess the ability to recover topology for each simulation when no missing data was involved in the phylogenetic inference process.
-
-%---------------------------------------------
-%       Empirical data - TO DO
-%---------------------------------------------
-
-\subsection{Empirical data} %Still not done!
-We also compared the results obtained from simulated data by using \citet{ronquista2012} empirical data.
-The matrix contains 67 living species plus one outgroup and 45 fossil species of Hymenopteras with 5097 molecular characters and 354 morphological characters.
-From the 68 living species used in the matrix, only 66 had molecular data, we therefore treated these 66 taxa as "living" taxa and all the other 47 as "fossil" taxa.
-We treated the matrix in the exact same way as described in step 2 and 3 resulting in 125 matrices with various amount of missing data and the same number of Maximum Likelihood and Bayesian trees.
-We used the same settings as for the simulated data in the Maximum Likelihood framework.
-For the Bayesian inferences however, we didn’t used any priors except that we provided a starting tree with the topology of the 68 living species (topology with the highest posterior probability from non-clock analysis - \citet{ronquista2012}).
-Contrary to \citet{ronquista2012} analysis, we didn’t performed any clock analysis since we were only interested in the topology of the inferred tree and not the branch length.
-
-%---------------------------------------------
-%
-%       RESULTS
-%
-%---------------------------------------------
-
-\section{Results}
-
-\subsection{Building the trees}
-When generating the matrices using seqgen \citep{ranbaut1997seqgen} and rTraitDisc \citep{paradisape:2004} algorithms with low evolutionary rate parameter ($\alpha$=0.5) we successfully generated evolutionary histories.
-In Maximum Likelihood, support values ranged from 0 to 100 with a median of 38 ($1^st$ quartile = 14, $3^rd$ quartile = 72 - see \hyperref[supplementaries]{supplementaries}). % This is actually bad.
-The distribution of the phylogenetic signal across the different chains provided us a good spread of evolutionary scenarios: from well resolved phylogenies to poorly resolved ones.
-This allows us to test our simulations in a theoretical as well as practical framework (i.e.
-phylogenies with low support values are rarely published but are still common in early stages of projects).
-%REF???
-- That's kind of my way to turn these bad results into something usable: even if some of our simulations are not realistic, it shows the effect of missing data in any scenario possible, not only in realistic cases.
-Also, one can advocate that this low topological resolution is actively due to the amount of missing data in the total evidence matrices since the sub-matrices (no missing data) analysis recovers way higher support value  .
-%This is something that I will had with the whole submatrices story.
-Basically, when you have all the fossil/molecular data missing, it's harder to build the tree, both in ML and in Bayesian.
-When you run the analysis without missing data, then it works way better (i.e.
-our REAL control test: can we simulate matrices in the first place).
-Using Bayesian inference, all the chains converged with a significant effective sample size by using 2 parallel runs of 4 chains each (ASSD $<$ 0.01 and ESS $>>$ 200 for inference).
-
-%\subsection{Comparing topologies}
-%Did it worked? See sub-matrices part. TO DO
-
-\subsection{Effect of the missing data parameters}
-
-%10/06 - Missing 8 ML Chains
-\begin{figure} 
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-results-BW.pdf}
-\caption{Effect of missing data on Tree similarity.
-$M_L$ is the amount of missing living taxa with morphological data,
-$M_F$ is the amount of missing data in the fossil record,
-$M_C$ is the amount of missing morphological characters.
-The first row represent the Normalised Tree Similarity index using the Robinson-Fould metric;
-the second row represent the Normalised Tree Similarity index using the Triples metric.}
-\label{Fig_Results}
-\end{figure}
-
-\subsubsection{Effect of $M_L$}
-In Maximum Likelihood framework, the topological recovery (reflected by the tree similarity) quickly drops with the increasing amount of missing data when using both metrics (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-However, this negative effect of missing data is stabilized beyond 25\% of missing data (no significant difference for trees with respectively 25\%, 50\% and 75\% missing living taxa).
-With 10\% of missing living taxa, there is a different topological recovery depending on the used metric.
-The Robinson-Fould metric shows a lower Tree similarity index than the Triples metric meaning that individual taxa position are generally more conserved (i.e.
-low number of flying taxa) but that the full clades are less conserved (see Table ~\ref{NTS_ML_results}).
-
-However, topological recovery was bad in Bayesian, regardless the amount of missing data.
-The tree similarity index was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance (see Table ~\ref{NTS_Ba_results}).
-Even if the topological recovery is really low, we detected an effect of missing data on the topology depending on the used metric.
-For the Triples metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
-For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases (see Table ~\ref{Ba_RF-ML_results}).
-However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the tree similarity index was only slightly higher than expected by chance (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-
-\subsubsection{Effect of $M_F$}
-In Maximum Likelihood framework, the effect of missing data in the fossil record appeared to be constant and led to a constant decrease in topological recovery when the missing data increased (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-The difference between the two metrics used is more important than for the $M_L$ parameter: clades are less conserved than individual taxa placement as the amount of missing data increases.
-Interestingly, 10\% or 25\% missing data for the $M_F$ parameter does not seems to affect the apparition of unstable taxa (Triples NTS modes of respectively 0.97 and 0.93 - Table ~\ref{NTS_ML_results}).
-
-Similarly than for the effect of $M_L$, Bayesian tree recovery was only slightly better than expected by chance.
-In the same way as described above, there seems to be a small difference depending on the metric used.
-There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric (see Table ~\ref{Ba_RF-MF_results}).
-However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
-
-\subsubsection{Effect of $M_C$}
-The number of missing morphological characters ($M_C$), however, seems to be more affecting topological recovery than $M_L$ and $M_F$ (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
-The Robinson-Fould metric shows a rapid drop in tree recovery from 10\% of missing data (Robinson-Fould NTS mode $<$ to 0.5 from 10\% missing data - Table ~\ref{NTS_ML_results}).
-This decrease is however slower when using the Triples metric with still a good topological recovery at 10\%  (Triples NTS modes = 0.95 - Table ~\ref{NTS_ML_results}).
-
-The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Fould metric shows a significant effect of the $M_C$ parameter on tree recovery (see Table ~\ref{Ba_RF-ML_results} and ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-
-%\subsubsection{Combined effect}
-%Need to do the heat plots
-
-$\newline$
-
-%To sum up:
-The ability of recovering the "best" tree's topology in Maximum Likelihood method is function of the amount of data missing.
-The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
-The decrease in clade conservation (low Robinson-Fould NTS score) is faster than the increase in flying taxa (low Triples NTS score).
-In Bayesian framework, topology is badly (if not at all) recovered disregarding the amount of data missing.
-
-\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
-\caption{Tree similarity values per parameter in ML framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
-              & missing data &        &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.671 & 0.569  & 0.789 & 0.468  & 1.036 \\ %how many digits?
-              &              & Triples        & 0.942 & 0.868  & 0.999 & 0.612  & 1.063 \\
-              & 25\%         & Robinson-Fould & 0.516 & 0.440  & 0.630 & 0.266  & 0.892 \\
-              &              & Triples        & 0.559 & 0.708  & 0.810 & 0.291  & 0.999 \\
-              & 50\%         & Robinson-Fould & 0.432 & 0.364  & 0.521 & 0.231  & 0.642 \\
-              &              & Triples        & 0.514 & 0.419  & 0.599 & 0.293  & 0.910 \\
-              & 75\%         & Robinson-Fould & 0.372 & 0.323  & 0.434 & 0.203  & 0.577 \\
-              &              & Triples        & 0.444 & 0.338  & 0.532 & 0.192  & 0.855 \\
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.907 & 0.766  & 1.000 & 0.407  & 1.047 \\
-              &              & Triples        & 0.976 & 0.944  & 1.003 & 0.715  & 1.104 \\
-              & 25\%         & Robinson-Fould & 0.572 & 0.488  & 0.692 & 0.325  & 0.936 \\
-              &              & Triples        & 0.937 & 0.820  & 0.993 & 0.539  & 1.086 \\
-              & 50\%         & Robinson-Fould & 0.433 & 0.302  & 0.476 & 0.170  & 0.668 \\
-              &              & Triples        & 0.732 & 0.594  & 0.873 & 0.341  & 1.025 \\
-              & 75\%         & Robinson-Fould & 0.153 & 0.115  & 0.200 & 0.068  & 0.323 \\
-              &              & Triples        & 0.458 & 0.390  & 0.611 & 0.290  & 0.856 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.473 & 0.384  & 0.835 & 0.356  & 0.958 \\
-              &              & Triples        & 0.959 & 0.869  & 1.026 & 0.572  & 1.067 \\
-              & 25\%         & Robinson-Fould & 0.454 & 0.357  & 0.630 & 0.242  & 0.916 \\
-              &              & Triples        & 0.907 & 0.751  & 0.997 & 0.446  & 1.070 \\
-              & 50\%         & Robinson-Fould & 0.424 & 0.375  & 0.466 & 0.249  & 0.828 \\
-              &              & Triples        & 0.683 & 0.571  & 0.822 & 0.277  & 1.008 \\
-              & 75\%         & Robinson-Fould & 0.230 & 0.177  & 0.302 & 0.108  & 0.567 \\
-              &              & Triples        & 0.451 & 0.373  & 0.572 & 0.224  & 0.791 \\
-    \hline
-\label{NTS_ML_results}
-\end{tabular}
-\end{table}
-
-\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
-\caption{Tree similarity values per parameter in Bayesian framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
-              & missing data &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\ %how many digits?
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.028   & 0.061 \\
-              &              & Triples        & 0.033 & 0.003  & 0.049 & -0.049  & 0.117 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.061 \\
-              &              & Triples        & 0.035 & 0.002  & 0.049 & -0.041  & 0.096 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.056 \\
-              &              & Triples        & 0.028 & 0.003  & 0.052 & -0.038  & 0.092 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.010   & 0.043 \\
-              &              & Triples        & -0.001 & -0.018 & 0.040 & -0.060 & 0.106 \\
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.035 & 0.031  & 0.056 & 0.027   & 0.076 \\
-              &              & Triples        & 0.026 & 0.002  & 0.050 & -0.046  & 0.105 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.0153  & 0.068 \\
-              &              & Triples        & 0.038 & 0.002  & 0.050 & -0.046  & 0.105 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.058 \\
-              &              & Triples        & 0.014 & -0.011 & 0.039 & -0.038  & 0.124 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.041 \\
-              &              & Triples        & 0.029 & 0.019  & 0.042 & -0.030  & 0.089 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.056 & 0.035  & 0.059 & 0.0281  & 0.076 \\
-              &              & Triples        & 0.030 & 0.011  & 0.048 & -0.024  & 0.094 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.023   & 0.076 \\
-              &              & Triples        & 0.020 & -0.001 & 0.037 & -0.060  & 0.109 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.030   & 0.059 \\
-              &              & Triples        & 0.010 & -0.005 & 0.045 & -0.036  & 0.104 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.011   & 0.043 \\
-              &              & Triples        & 0.024 & 0.001  & 0.042 & -0.037  & 0.104 \\
-    \hline
-\label{NTS_BA_results}
-\end{tabular}
-\end{table}
-
-\begin{table}
-\caption{Group difference per configuration set}
-\centering
 \begin{tabular}{rllcccc}
@@
     \hline
-\label{Group_results}
 \end{tabular}
@@
 $^1$ F-value for parametric tests and Kruskal Wallis $chi^2$ for non parametric tests.
-\end{table}
-
-\begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE  & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE  \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 56.53 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$25\% & 96.85 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 127.03 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 144.58 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 40.31 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$50\% & 70.50 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 88.05 & 37.66 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 30.19 & 37.66 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 47.73 & 37.66 & TRUE \\ 
-  $M_L$50\%-$M_L$75\% & 17.55 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_RF-ML_results}
-\end{table}
-
-\begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE  & FALSE \\
-    $M_L$50\% & & & & - & FALSE  \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 57.98 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$25\% & 105.83 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 120.86 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 140.34 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 47.85 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$50\% & 62.88 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 82.36 & 37.66 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 15.03 & 37.66 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 34.51 & 37.66 & FALSE \\ 
-  $M_L$50\%-$M_L$75\% & 19.48 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_Tr-ML_results}
-\end{table}
-
-\begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & FALSE & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 17.05 & 41.00 & FALSE \\ 
-  $M_L$00\%-$M_L$25\% & 43.79 & 41.00 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 77.32 & 41.00 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 103.99 & 41.00 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 26.75 & 41.00 & FALSE \\ 
-  $M_L$10\%-$M_L$50\% & 60.27 & 41.00 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 86.94 & 41.00 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 33.53 & 41.00 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 60.20 & 41.00 & TRUE \\ 
-  $M_L$50\%-$M_L$75\% & 26.67 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{Ba_RF-ML_results}
-\end{table}
-
-\begin{table}
-\caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & \textbf{TRUE} \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 59.27 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$25\% & 79.72 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$50\% & 120.70 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 167.81 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 20.45 & 37.66 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 61.43 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 108.55 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 40.98 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$75\% & 88.09 & 37.66 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 47.12 & 37.66 & TRUE \\ 
-   \hline
-\end{tabular}
-\label{ML_RF-MF_results}
-\end{table}
-
-\begin{table}
-\caption{$M_F$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 64.78 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$25\% & 89.09 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$50\% & 123.20 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 150.43 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 24.31 & 37.66 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 58.42 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 85.65 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 34.10 & 37.66 & FALSE \\ 
-  $M_F$25\%-$M_F$75\% & 61.34 & 37.66 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 27.23 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_Tr-MF_results}
-\end{table}
-
-\begin{table}
-\caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & FALSE & FALSE & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 16.03 & 41.00 & FALSE \\ 
-  $M_F$00\%-$M_F$25\% & 38.50 & 41.00 & FALSE \\ 
-  $M_F$00\%-$M_F$50\% & 61.52 & 41.00 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 95.52 & 41.00 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 22.47 & 41.00 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 45.49 & 41.00 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 79.49 & 41.00 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 23.02 & 41.00 & FALSE \\ 
-  $M_F$25\%-$M_F$75\% & 57.02 & 41.00 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 34.00 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{Ba_RF-MF_results}
-\end{table}
-
-\begin{table}
-\caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & \textbf{TRUE} \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 67.99 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$25\% & 87.58 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$50\% & 116.16 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 155.77 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 19.59 & 37.66 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 48.17 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 87.78 & 37.66 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 28.58 & 37.66 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 68.19 & 37.66 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 39.60 & 37.66 & TRUE \\ 
-   \hline
-\end{tabular}
-\label{ML_RF-MC_results}
-\end{table}
-
-\begin{table}
-\caption{$M_C$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & FALSE \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 75.29 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$25\% & 88.58 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$50\% & 114.02 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 149.60 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 13.29 & 37.66 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 38.73 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 74.31 & 37.66 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 25.44 & 37.66 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 61.02 & 37.66 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 35.58 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{ML_Tr-MC_results}
-\end{table}
-
-\begin{table}
-\caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & FALSE & FALSE & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & FALSE \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 16.03 & 41.00 & FALSE \\ 
-  $M_C$00\%-$M_C$25\% & 38.50 & 41.00 & FALSE \\ 
-  $M_C$00\%-$M_C$50\% & 61.52 & 41.00 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 95.52 & 41.00 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 22.47 & 41.00 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 45.49 & 41.00 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 79.49 & 41.00 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 23.02 & 41.00 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 57.02 & 41.00 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 34.00 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
-\label{Ba_RF-MC_results}
-\end{table}
-
-%---------------------------------------------
-%
-%       DISCUSSION
-%
-%---------------------------------------------
-
-\section{Discussion}
-
-\subsection{Building the trees}
-Simulating evolutionary history matrices still remains a big drawback in theoretical phylogenetics. %Hahahaha.... feck.
-The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
-This configuration probably lead to globally low phylogenetic signal as well as the intrinsic difficulties to simulated characters with phylogenetic signal.
-Even though molecular characters evolution (and therefore simulation) only depends of a small number of parameters (i.e.
-the base frequencies and the substitution matrix), simulating molecular matrix with a strong evolutionary signal is still complicated when generating unrealistically small matrices.
-For morphological characters, the underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
-
-Also morphological characters studies involve many potential statistical pitfalls (e.g.
-independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy.
-%REF probably April as something about it
-\begin{enumerate}
-\item{First, morphological data can display a different signal than molecular data, especially in small matrices. %REF
-This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
-However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013}.}
-%The main difference between molecular and morphological characters is that, because of historical and practical reasons, a morphological character matrix do no contains autapomorphies and invariant characters which are usually more common than synampomorphies in molecular character matrices (cite).
-\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013}.
-However, following our simulation assumption of random character distributions, if they accumulate through time in the same way as the majority of the molecular characters then, homoplasic characters are expected to appear randomly through time \citep{davalosintegrating2014}.
-Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
-After a reaching a critical amount of morphological characters, adding new ones increases homoplasy \citep{wagner2000}.}
-\end{enumerate}
-Our simulation parameters both decrease the phylogenetic signal (difficulty to simulate morphological characters and size of the matrix) as well as it increases it (reduction of homoplasy in the morphological part of the matrix).
-Therefore, these drawbacks seems to have only a minor impact on the main results of this study: the incapacity of recovering any topology in Bayesian inference.
-
-\subsection{Comparing topologies}
-Comparing topologies is a crucial question in phylogenetics but has always been hard to normalise because of the vast amount of different metrics used as proxies for different aspects of tree similarity/dissimilarity \citep{agapowpower2002}.
-Because our global framework is to study how to include efficiently fossils into phylogenies, we chose metrics reflecting the most interesting aspect of this global question: where do individual taxa (i.e.
-the fossils) branch in the tree.
-The Robinson-Fould \citep{RF1981} and the Triples \citep{critchlowthe1996} metrics are more sensible to taxa and clade placement than other tree comparison metrics \citep[e.g.][Imbalance metric]{kirkpatricksearching1993} and where therefore favoured.
-Also by using the Normalised Tree Similarity index \citep{Bogdanowicz2012} we emphasize the aspect of "good" phylogenetic signal \textit{versus} random phylogenetic signal because it normalised the values of the metric by correcting for the expected value when comparing random trees (NTS = 0).
-
-The Robinson-Fould metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
-When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
-On the other hand, the Triples method is measuring the position of each taxon towards to other reference taxa \citep{critchlowthe1996}.
-It will penalise only trees where taxa get removed furthest from their original clade.
-Regarding our problematic (how does missing data influence topological recovery in trees containing both living and fossil taxa) we are more interested the placement of taxa (i.e.
-where does the fossil branch) than the exact conservation of clades.
-
-%Need a transition!
-
-Although the idea of comparing two single trees is straightforward, it becomes more complex when comparing trees distributions (e.g. Bayesian posterior distributions).
-The introduction of our Random Pairwise Bayesian Tree Comparison methods allows to summarize the comparison of two trees distributions by picking up the most frequent signal in the distribution of the pairwise comparisons (i.e.
-the mode).
-Even though this method is subject to randomness and might artificially increase or decrease the score of the studied metric, simulations showed that when a sufficient amount of random comparisons is performed, this method doesn't seem to be subject to randomness (e.g.
-1000 random pairwise comparisons - see ~\ref{supplementaries}).
-
-\subsection{Maximum Likelihood versus Bayesian}
-The main results of our analysis shows that Bayesian inference fails to recover any Topology in a total evidence method framework (Fig. ~\ref{Fig_Results}).
-This results is surprising regarding the behaviour of Maximum Likelihood inference (Fig.
-~\ref{Fig_Results}) as well as the Bayesian inferences performed on empirical data \citep{ronquista2012,schragocombining2013}.
-However, it is important to note that this effect was not mentioned in the aforementioned empirical tests because both studies used a Bayesian approach with fixed topology \citep{ronquista2012,schragocombining2013}.
-
-In our case, we suspect that this inability of Bayesian methods to recover topology is due to the intrinsic structure of a total evidence matrix (Fig ~\ref{Fig_RemoveData}).
-In fact, as previously studies have shown, missing data doesn't seem to be a major drawback as long as the missing data is randomly distributed ~\citep[e.g.][]{wiensmissing2003,rouresite-specific2011,sansomfossilization2013}.
-However, in total evidence matrices, the majority of the missing data is not randomly distributed but concentrated in the molecular part of the matrix for the fossil taxa (i.e.
-the missing-data sub-matrix, Fig ~\ref{Fig_RemoveData}).
-This leads to high decrease of topological recovery when using non optimal criterion approach (i.e.
-Bayesian inference) because of the high variance in the near-likeliest solutions sampled in the Bayesian posterior distribution.
-In opposition, when applying optimal criterion approach (i.e.
-Maximum Likelihood), it is still possible to sample the likeliest tree.
-%probably need to develop that way more, especially with actually analysing the likelihood landscape.
-
-\subsection{Effect of missing data}
-The three parameters we selected in this study account for three potential pitfalls in collecting the data for total evidence analysis:
-$M_L$ represents the living taxa for which there is no available morphological data,
-$M_F$ represents the quality of the fossil record,
-and $M_C$ represents the general coding effort and the overall of morphological data available.
-Ideally, the lowest possible amount of missing data is wanted in any phylogenetic analysis leading to a data collection part prior to the phylogenetic analysis.
-Each of the three parameters can be improved in a different way prior to the analysis:
-for $M_L$, one should put more effort in using natural museums history collections for coding the missing morphological data for living taxa if possible %go to the museum!;
-for the $M_F$ parameter, the amount of missing data unfortunately depends on the quality of the fossil record and can not be actively improved and depends on exceptional discoveries \citep[e.g.][]{nithe2013};
-finally for the $M_C$ parameter, improvement can be done by vast collaborative projects in order to gather as much characters as possible \citep[e.g.][]{O'Leary08022013}.
-
-Our results shows that the $M_F$ parameter have less influence on recovering the good tree topology in Maximum Likelihood framework which is fortunate since it is the parameter that is the more difficult to fix for practical reasons.
-Regarding the fact that we are more interested in taxa placement than in clade conservation (Triples \textit{versus} Robinson-Fould), the parameter that affects the more the decrease in topological recovery is the number of missing living taxa in the morphological part of the matrix ($M_L$) and the overall number of morphological characters $M_C$.
-This makes sense since the living taxa are bearing both the information to build the tree backbone (the molecular data) and the information used for branching the fossil on this backbone (the morphological information).
-Therefore, we advocate the importance of coding morphological characters for the most living taxa possible and with the most characters possible, potentially by using collaborative projects portals such as morphobank \citep{morphobank}.
-
-%Potential reviewers comments?
-%What about it when you vary the proportion of fossil taxa?
-
-%---------------------------------------------
-%
-%       CONCLUSION
-%
-%---------------------------------------------
-
-\section{Conclusion}
-A Bayesian approach fails to recover accurate topology in a total evidence approach framework, whatever the amount of missing data.
-We think this failure is due to the intrinsic structure of the total evidence matrices that includes a vast amount of non randomly distributed missing data (i.e.
-the molecular part of the matrix for the fossil taxa).
-This missing data is equal to the number of fossil taxa $\times$ the number of molecular characters leading to a vast amount of trees to be sampled in Bayesian posterior distribution.
-However, one can use a Maximum Likelihood approach to fix the likeliest topology.
-If so, an effort should be made prior to the phylogenetic analysis on collecting as much morphological data as their is available from living taxa in order to efficiently improve the quality of the trees.
-
-%Perspectives:
-%What the effect of this missing molecular data on branch length estimation?
-
-%---------------------------------------------
-%
-%       ...
-%
-%---------------------------------------------
-
-\section{Acknowledgements}
-We would like to thank Trevor Hodkinson and Andrew Jackson %Also Gavin Thomas, Emmanuel Douzery and Frédéric Delsuc if not authors.
-for his useful comments on the simulation protocol and Paddy Doyle for the assistance on using the computer cluster.
-All calculations were performed on the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing.
-This cluster was funded through grants from Science Foundation Ireland.
-
-%BIBLIOGRAPHY
- % The \cite command functions as follows:
- %   \citet{key} ==>>                Jones et al. (1990)
- %   \citet*{key} ==>>               Jones, Baker, and Smith (1990)
- %   \citep{key} ==>>                (Jones et al., 1990)
- %   \citep*{key} ==>>               (Jones, Baker, and Smith, 1990)
- %   \citep[chap. 2]{key} ==>>       (Jones et al., 1990, chap. 2)
- %   \citep[e.g.][]{key} ==>>        (e.g. Jones et al., 1990)
- %   \citep[e.g.][p. 32]{key} ==>>   (e.g. Jones et al., p. 32)
- %   \citeauthor{key} ==>>           Jones et al.
- %   \citeauthor*{key} ==>>          Jones, Baker, and Smith
- %   \citeyear{key} ==>>             1990
-
-\bibliographystyle{sysbio} %don't write the suffix
-\bibliography{References} %don't write the suffix
-
-\section{Supplementaries}
-\label{supplementaries}
-\subsection{Morphological characters states}
-In order to obtain a realistic probabilistic value for of \textit{k} characters states for each simulated morphological character, we downloaded 100 random morphological characters (with more than 100 characters each) from TreeBASE database (http://treebase.org/) published between 1985 and 2013 and covering 19 taxomomic classess (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
-We selected a total of 22563 characters ranging from 2 to 10 states.
-We calculated the proportion of characters with 2, 3, 4, 5, 6, 7, 8, 9 or 10 states.
-We then sampeled 22563 \textit{k} values between 2 and 10 with the same proportion of characters from the empirical data.
-We then used a simple t-test to check if our simulation was equal to the empirical data.
-In this study, we only simulated characters with 2 or 3 states because of the high proportion of ordered characters ecountered on characters with more than 3 states and the difficulties of simulate biologicaly sensible ordered characters.
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-AppendixCharacters.pdf}
-\caption{Character states distribution in empirical matrices. %Title?
-Characters states number distribution extracted from 100 random morphological matrices downloaded from RreeBase.}
-\label{Fig_AppendixCharacters}
-\end{figure}
-
-\subsection{Tree Building Software settings}
-
-\subsubsection{Maximum Likelihood - RAxML v8.0.20 \citep{Stamatakis21012014}} \\
-Model: \\
-Molecular data: \\
-GTR + $\Gamma_4$ (-m GTRGAMMA) \\
-Morphological data: \\
-Mk + $\Gamma_4$ (-K MK) \\
-Support: \\
-Rapid Boostrap algorithm (LSR), 1000 replicates \\
-
-\subsubsection{Bayesian - MrBayes v3.0.2 \citep{Ronquist2012mrbayes}} \\
-Priors: \\
-Molecular data: \\
-rates distribution shape ($\alpha$) = 0.5 \\
-Transition/Transversion ratio = 2 ($\beta$(80,40)) \\
-Starting tree: "True" tree topology with each branch length = 1 \\
-Morphological data: \\
-rates distribution shape ($\alpha$) = 0.5 \\
-Models: \\
-Molecular data: HKY + $\Gamma_4$ \\
-Morphological data: Mk + $\Gamma_4$ \\
-MCMC: \\
-2 runs \\
-4 chains per run \\
-generations < 50$\times$$1^6$ \\
-sample frequency = 1050$\times$$1^3$ \\
-ASDS diagnosis frequency = 50$\times$$1^3$ \\
-ASDS $<$ 0.01 \\
-ESS $>>$ 200 \\
-Burnin = 25\%
-
-\subsection{Triplets metric details ($T_{x,y}$)}
-Each triplet can be written as $I_{ijk}$=(\textit{ijk})). Where $I_{ijk}$ is equal to 0 if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to 1.
-For any rooted tree there are only four possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); and (\textit{i},\textit{j},\textit{k}); \citep{johnson1998}.
-One can calculate $S_n$, the triplet distance between two trees as:
-\begin{equation}
-S_n = \sum_{ijk} I_{ijk}
-\end{equation}
-Where:
-\begin{equation}
-\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
-\end{equation}
-And where n is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
-When all triplets across the two trees are the same, $S_n$ is equal to 0 and when all the triplets are different $S_n$ is equal to $\binom{n}{4}$.
-Because the possible number of triplets per clade is a finite number, the probability of two random trees with the same n taxa to have the same triplet is:
-\begin{equation}
-P({I_{ijk}}=0) = \frac{1}{4}
-\end{equation}
-Therefore one can calculate the probability of two random trees having the same triplets: 
-\begin{equation}
-P({S_{n}}=0) = \sum_{ijk} P_{I_{ijk}=0}
-\end{equation}
-\begin{equation}
-P({S_{n}}=0) = \frac{n!}{4(3!(n-3)!}
-\end{equation}
-And in the same way:
-\begin{equation}
-P({S_{n}}=1) = \frac{3n!}{4(3!(n-3)!}
-\end{equation}
-
-\subsection{RF metric details}
-The RF distance (or path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
-\begin{equation}
-RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
-\end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees. 
-The minimal value of \textit{C} is equal to 1 if the two trees have the same n taxa;
-the maximal value in \textit{C}=\textit{n}-2.
-For a fully unresolved tree (star tree) \textit{N}=1 and for a fully resolved tree (binary tree) \textit{N}=\textit{n}-2.
-The minimal and maximal topological distance for \textit{n} taxa is:
-\begin{equation}
-RF_{min} = 1 + 1 - 2C_{x,y}
-\end{equation}
-And:
-\begin{equation}
-RF_{max} = 2(n-2)-2
-\end{equation}
-One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any \textit{n} taxa:
-\begin{equation}
-RF.scaled_{x,y} = \frac{RF_{x,y}-RF_{max}}{RF_{max}}
-\end{equation}
-This metric is more sensitive to taxa displacement than the Triplet distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.
-
-\subsection{Tree comparisons}
-\subsubsection{Random tree comparison scaling}
-We used the comparison of 1000 random trees to obtain the mean comparison value $\bar{d}_{m,n}$\textit{(rand)} for the NTS metric.
-We randomly generated two sets of 1000 trees of \textit{n} taxa using the rmtree function of ape package (v3.0-11 \citet{paradisape:2004}) that generates a given number of random Yule trees.
-We calculated the $\bar{d}_{m,n}$\textit{(rand)} value using an approach similar to the RPCBTC (described below) by performing 1000 random pairwise comparisons using the TreeCmp java script \citep{Bogdanowicz2012}.
-
-\subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
-We assessed the power of the Random Pairwise Bayesian Tree Comparison (RPBTC) method by comparing 1000 random trees from a posterior distribution trees set to another 1000 random trees from the same posterior distribution trees set.
-We repeated this 100 times independently using the same posterior distribution trees set each time resulting in 100 replicates of the same posterior distribution trees set compared 1000 times.
-We used an anova to test if there was no significant difference between the replicates so that the RBTC can be replicated.
-We applied this protocol on a poorly resolved tree (Low Score), a resolved tree with low support value (Medium Score) and a resolved tree with high support values (High Score).
-Results are available in table ~\ref{RPBTC_testing}. %link broken
-
-\begin{table}[ht]
-\caption{Group comparison results: difference between 100 replicates using the RPBTC method} %test with TreeCmp.anova
-\centering
-\begin{tabular}{rllrrrr}
-  \hline
-  Tree.Type & Used.metric & Replicates & Df & F.value & p.value \\ 
-  \hline
-  Low Score & RF & 100.00 & 99.00 & 0.74 & 0.98 \\ 
-  Low Score & Tr & 100.00 & 99.00 & 0.97 & 0.58 \\ 
-  Medium Score & RF & 100.00 & 99.00 & 0.64 & 1.00 \\ 
-  Medium Score & Tr & 100.00 & 99.00 & 0.45 & 1.00 \\ 
-  High Score & RF & 100.00 & 99.00 & 0.20 & 1.00 \\ 
-  High Score & Tr & 100.00 & 99.00 & 0.37 & 1.00 \\ 
-  \hline
-\end{tabular}
-\end{table}
-\label{RPBTC_testing}
-
-\subsection{Codes}
-All codes are available at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions$
-The tree comparison results analysis can be repeated for more details at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Analysis$ %fully running!
-Try it at home (5-10 minutes of data loading)!
-
-\subsection{Full results}
-
-\begin{table}
-\caption{Tree similarity values per parameter in ML framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
-              & missing data &        &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.6714878 & 0.5692308  & 0.7894649 & 0.4689069  & 1.0361514 \\ %how many digits?
-              &              & Triples        & 0.9424133 & 0.8683742  & 0.9998559 & 0.6121399  & 1.0635250 \\
-              & 25\%         & Robinson-Fould & 0.5167595 & 0.4403147  & 0.6307692 & 0.2664429  & 0.8929104 \\
-              &              & Triples        & 0.5599494 & 0.7084874  & 0.8105781 & 0.2912950  & 0.9990703 \\
-              & 50\%         & Robinson-Fould & 0.4321852 & 0.3641026  & 0.5215627 & 0.2313246  & 0.6423846 \\
-              &              & Triples        & 0.5143009 & 0.4197127  & 0.5991498 & 0.2991432  & 0.9104526 \\
-              & 75\%         & Robinson-Fould & 0.3727090 & 0.3230769  & 0.4342249 & 0.2035474  & 0.5776286 \\
-              &              & Triples        & 0.4449563 & 0.3385128  & 0.5329149 & 0.1929395  & 0.8551570 \\
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.9072419 & 0.7668415  & 1.0000000 & 0.4070844  & 1.0478010 \\
-              &              & Triples        & 0.9763617 & 0.9448100  & 1.0039460 & 0.7158118  & 1.1043330 \\
-              & 25\%         & Robinson-Fould & 0.5722382 & 0.4883560  & 0.6923077 & 0.3253239  & 0.9367467 \\
-              &              & Triples        & 0.9370152 & 0.8207407  & 0.9938846 & 0.5394443  & 1.0868073 \\
-              & 50\%         & Robinson-Fould & 0.4338159 & 0.3025641  & 0.4760777 & 0.1704016  & 0.6683220 \\
-              &              & Triples        & 0.7327464 & 0.5947211  & 0.8739630 & 0.3418679  & 1.0251540 \\
-              & 75\%         & Robinson-Fould & 0.1534487 & 0.1155128  & 0.2000000 & 0.0683485  & 0.3230769 \\
-              &              & Triples        & 0.4588292 & 0.3905631  & 0.6111484 & 0.2903676  & 0.8568274 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.473285 & 0.3842965  & 0.8358974 & 0.3564753  & 0.9589744 \\
-              &              & Triples        & 0.9592557 & 0.8691580  & 1.026095 & 0.5720847  & 1.067524 \\
-              & 25\%         & Robinson-Fould & 0.4548038 & 0.3572132  & 0.6307692 & 0.2425339  & 0.9161212 \\
-              &              & Triples        & 0.9071925 & 0.7516934  & 0.9976224 & 0.4462168  & 1.0705291 \\
-              & 50\%         & Robinson-Fould & 0.4244263 & 0.3755493  & 0.4666667 & 0.2494927  & 0.8287312 \\
-              &              & Triples        & 0.6834366 & 0.5713927  & 0.8226139 & 0.2776690  & 1.0085944 \\
-              & 75\%         & Robinson-Fould & 0.2306888 & 0.1774508  & 0.3025641 & 0.1087764  & 0.5673541 \\
-              &              & Triples        & 0.4513642 & 0.3737397  & 0.5723858 & 0.2245224  & 0.7917946 \\
-    \hline
-\label{NTSML_full}
-\end{tabular}
-\end{table}
-
-\begin{table}
-\caption{Tree similarity values per parameter in Bayesian framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
-              & missing data &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\ %how many digits?
-              &              & Triples        & 0.04104231 & 0.00830962  & 0.05441173 & -0.03933377  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.03584309 & 03584196  & 03589744 & 02825643  & 0.06175676 \\
-              &              & Triples        & 0.03399849 & 0.003196469  & 0.04981018 & -0.049913382  & 0.11784366 \\
-              & 25\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01536739  & 0.06123106 \\
-              &              & Triples        & 0.03510953 & 0.002113358  & 0.04939228 & -0.041202264  & 0.09699332 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01538462  & 0.05641026 \\
-              &              & Triples        & 0.02849433 & 0.003263098  & 0.05298036 & -0.038344731  & 0.09250063 \\
-              & 75\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01069992  & 0.04338811 \\
-              &              & Triples        & -0.0001940566 & -0.01801301  & 0.04058783 & -0.06030922  & 0.10665934 \\
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\
-              &              & Triples        & 0.04104231 & 0.008309622  & 0.05441173 & -0.039333778  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.03593712 & 0.03198256  & 0.05641026 & 0.02798864  & 0.07692308 \\
-              &              & Triples        & 0.02639309 & 0.002185408  & 0.05061314 & -0.046563042  & 0.10547052 \\
-              & 25\%         & Robinson-Fould & 0.03588134 & 0.03587629  & 0.03589744 & 0.01538462  & 0.0680884 \\
-              &              & Triples        & 0.03856296 & 0.002185408  & 0.05061314 & -0.046563042  & 0.10547052 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01536289  & 0.05879269 \\
-              &              & Triples        & 0.01403716 & -0.01100608  & 0.03972324 & -0.03800324  & 0.12405207 \\
-              & 75\%         & Robinson-Fould & 0.03589206 & 0.03588825  & 0.03589744 & 0.01538462  & 0.0411843 \\
-              &              & Triples        & 0.02939363 & 0.01902351  & 0.04238907 & -0.03087905  & 0.08946917 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\
-              &              & Triples        & 0.04104231 & 0.008309622  & 0.05441173 & -0.039333778  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.05637057 & 0.03589744  & 0.05927956 & 0.02813296  & 0.07692308 \\
-              &              & Triples        & 0.03008132 & 0.01176800  & 0.04847400 & -0.02451461  & 0.09453754 \\
-              & 25\%         & Robinson-Fould & 0.03593743 & 0.03589744  & 0.03596018 & 0.02330790  & 0.07692308 \\
-              &              & Triples        & 0.02010221 & -0.001145118  & 0.0377779 & -0.060183332  & 0.1098632 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.03050289  & 0.05950426 \\
-              &              & Triples        & 0.01053959 & -0.005595947  & 0.04510444 & -0.036278820  & 0.10478805 \\
-              & 75\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01115983  & 0.04362927 \\
-              &              & Triples        & 0.02427778 & 0.0007405737  & 0.04260522 & -0.0373004106  & 0.1048756 \\
-\label{NTSBA_full}
-\end{tabular}
-\end{table}
-
-\subsection{Bootstraps distribution}
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-AppendixBootstraps.pdf}
-\caption{Bootstraps distribution across the "best" trees.}                     % MISSING 8 CHAINS
-\label{Fig_AppendixBoostraps}
-\end{figure}
-
-%END
-\end{document}

### 2014-06-18 Manuscript/SupplementaryMaterial/TreeComparison.tex
Updated tree comparison part

@@
-\subsection{Triplets metric details ($T_{x,y}$)}
-Each triplet can be written as $I_{ijk}$=(\textit{ijk})). Where $I_{ijk}$ is equal to 0 if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to 1.
-For any rooted tree there are only four possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); and (\textit{i},\textit{j},\textit{k}); \citep{johnson1998}.
-One can calculate $S_n$, the triplet distance between two trees as:
+\subsection{Triplets distance details ($T_{x,y}$)}
+Triples distance ($T_{x,y}$; \citealp{dobson1975triplets}) measures the number of sub-trees made up of three taxa (triplets) that differ between two given trees.
+Each triplet can be written as $I_{ijk}$=(\textit{ijk}). Where $I_{ijk}$ is equal to zero if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to one.
+For any rooted binary tree there are only three possible combinations for each triplet: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
+If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k}).
+We can calculate the triplet distance between two trees, $S_n$, as:
 \begin{equation}
@@
 \end{equation}
-And where n is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
-When all triplets across the two trees are the same, $S_n$ is equal to 0 and when all the triplets are different $S_n$ is equal to $\binom{n}{4}$.
+And where \textit{n} is the total number of taxa in both trees (modified from \citet{critchlowthe1996}).
+If $S_n$ = 0, the trees are identical; when $S_n$ = $\binom{n}{4}$, the trees are as different as possible (i.e. every taxon has a different placement in the two trees).
 Because the possible number of triplets per clade is a finite number, the probability of two random trees with the same n taxa to have the same triplet is:
@@
 
-\subsection{RF metric details}
-The RF distance (or path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
+\subsection{Robinson-Foulds distance details}
+Robinson-Foulds distance (\textit{RF}; \citealp{RF1981}), or "path difference"
+% NC: path difference or *distance*? - TG: difference.
+, measures the number of shared clades across two trees.
+The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} and can be expressed as following:
 \begin{equation}
@@
 \end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees. 
-The minimal value of \textit{C} is equal to 1 if the two trees have the same n taxa;
-the maximal value in \textit{C}=\textit{n}-2.
-For a fully unresolved tree (star tree) \textit{N}=1 and for a fully resolved tree (binary tree) \textit{N}=\textit{n}-2.
-The minimal and maximal topological distance for \textit{n} taxa is:
+Where $C_{x,y}$ is the number of clades in common in the two trees. $C$ % NC: Why italics? Why not $C$?
+is one if the two trees have the same $n$ taxa; the maximal value is $C = n-2$. % NC: again why italics for these bits? 
+%NC: I don't get this. Explain more clearly. also is it important?
+This metric is more sensitive to taxon displacement than Triples distance (i.e.
+if one taxon moves out of a clade, then the clades are no longer considered similar; \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
+The minimal value of $C$ is equal to 1 if the two trees have the same n taxa;
+the maximal value in $C = n-2$.
+For a fully unresolved tree (star tree) $N$=1 and for a fully resolved tree (binary tree) $N = n-2$.
+The minimal and maximal topological distance for  taxa is:
 \begin{equation}
@@
 \end{equation}
-One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any \textit{n} taxa:
+One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any $n$ taxa:
 \begin{equation}

### 2014-06-18 Manuscript/TEM_draft.tex
Correction methods

--- 2014-06-15 Manuscript/TEM_draft.tex
@@
-%HEADER
  \documentclass[12pt,letterpaper]{article}
@@
 % NC: You could add all the packages etc to a template .sty file to save writing it all up here. Rich's ms.sty that I provided on DropBox is a good example
+
+% TG: Just a general note for the file outline: I write every sentence of a same paragraph on a new line, no real reason for that appart that I can then see when the sentence is way to long.
+Also, I kept the big hidden headers ("%----") that you don't seem to like, without them it totally wrecks my head to know how much do I still have to scroll...
+
 \usepackage{fixltx2e}
@@
 \usepackage{fullpage}
-\usepackage{amsfonts}
-\usepackage{verbatim}
-\usepackage[english]{babel}
-\usepackage{pifont}
-\usepackage{color}
-\usepackage{setspace}
-\usepackage{lscape}
-\usepackage{indentfirst}
-\usepackage[normalem]{ulem}
-\usepackage{booktabs}
 %\usepackage{nag}
@@
 \bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
-% We don't use a special title page; the author information is entered like any other text.
-% FOOTNOTES: We don't allow them in the manuscript, except in tables. Don't include any footnotes in the text.
-%IMPORTANT:
-%Note on display in the text file: sentences or sub-parts of sentences (e.g. enumerations) are delimited by a newline for clarity.
-%However, true new lines (i.e. in the pdf) are delimited by one or more empty lines.
-%We don't use a heading for the introduction. The nature of this section is implied by its position following the abstract and preceding the first major section heading.
-
-% NC: You should delete this extraneous info from the Syst Biol template to tidy everything up a bit
+
+%---------------------------------------------
+%
+%       START
+%
+%---------------------------------------------
 
@@
 %For a colour version of the draft to a search>replace '-BW'>'' 
-% %\modulolinenumbers[1] 	% Line numbering on every line
+%\modulolinenumbers[1]     % Line numbering on every line
 
@@
 \noindent{\Large \bf Effect of missing data in matrices containing living and fossil taxa on topological accuracy}
-%Alternative punchier title: "Bayesian total evidence methods fail to recover correct topologies from matrices containing both living and fossil taxa"
-%Alternative punchier title: "Maximum Likelihood outperforms Bayesian methods for inferring topology from total evidence matrices containing both living and fossil taxa"
-
+%Alternative punchier title: "Bayesian Total Evidence methods fail to recover correct topologies from matrices containing both living and fossil taxa"
+%Alternative punchier title: "Maximum Likelihood outperforms Bayesian methods for inferring topology from Total Evidence matrices containing both living and fossil taxa"
 % NC: Not sure about any of these.
 I actually think I prefer your talk title.
-So something like: The effects of missing data on the ability of total evidence approaches to recover correct tree topologies.
+So something like: The effects of missing data on the ability of Total Evidence approaches to recover correct tree topologies.
 No rush to decide.
-But needs total evidence in there somewhere.
+But needs Total Evidence in there somewhere.
+
 \bigskip
@@
 \vspace{1in}
-% NC: It's not called the Republic of, just Ireland!
+
+%---------------------------------------------
+%
+%       ABSTRACT
+%
+%---------------------------------------------
 
@@
 
+%---------------------------------------------
+%
+%       INTRODUCTION
+%
+%---------------------------------------------
+
 \section{Introduction}
@@
 This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty (but see \citealp{spencerefficacy2013}).
-Neontologists, on the other hand, more commonly use probabilistic approaches (e.g.
-Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates % NC: Need to check the exact phrase needed here.
+Neontologists, on the other hand, more commonly use probabilistic approaches (e.g. Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
+Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence date % NC: Need to check the exact phrase needed here.
 Is it range of occurrence?
@@
 etc.
+- TG: in paleo jargon we have FAD (first apparition datum) and FOD (first occurrence datum).
+FAD is always hypothetical (we can't really exactly know when the taxa appeared) and FOD is always subjective (it depends when the fossil was found, maybe next year someone will find another one in older layers).
+So for the dating, because it's relaxed clock, people use FOD (and also probably because they don't care since it doesn't make a huge difference I guess).
     are used to time calibrate phylogenies \citep{zuckerkandl1965}.
@@
 
-A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa.
+A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \citep{eernissetaxonomic1993}.
 %There might be a philosophico-semantic distinction to do here.
@@
 % NC: Tricky. Don't think it's too important though. 
-% NC: I do think Total Evidence needs capital letters though, like ML and Bayesian
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{eernissetaxonomic1993}.
+% NC: I do think Total Evidence needs capital letters though, like ML and Bayesian - TG: fixed
+This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
 Total Evidence methods have been successfully applied to empirical data \citep{pyrondivergence2011,ronquista2012,schragocombining2013}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
-However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data.
-In particular it requires morphological data from both living and fossil taxa, both of which are known to be scarce.
+However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data that can be difficult (or impossible) to collect.
+%In particular it requires morphological data from both living and fossil taxa, both of which are known to be scarce.
 % NC: Really?
@@
 Needs a citation either way.
+%TG: how about this:
+The morphological data for living taxa is rarely collected when molecular data is available (e.g.
+\citealt{O'Leary08022013} vs.
+\citealt{meredithimpacts2011}),% TG: i.e.
+in O'Leary they use meredith molecular data but only code few species from it.
+ and, for fossil taxa, the scarcity of the fossil record only allow to collect the data available (for example, in vertebrates, the hardest parts of the skeleton; \citealt{sansomfossilization2013}).
 Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
@@
 Sive and I won't be around to correct your English forever.
+- TG: my French demons are back!
 
 The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013}.
-Missing molecular data has been seen by some authors as an issue because it can decrease phylogenetic signal (i.e. % define phylogenetic signal here) 
-in some parts of the tree, especially when using large matrices \citep{lemmonthe2009}. However, this may not be a major issue because phylogenetic signal is easily increased by:
-(i) including a "modest" number of highly-covered % what do you mean by highly-covered?
-genes (i.e. approximatively half of the genes; \citealt{rouresite-specific2011});
+Missing molecular data has been seen by some authors as an issue because it can, in some part of the tree, decrease phylogenetic signal (i.e.
+the evolutionary information contained within the matrix allowing to infer topology and branch length)% NC: define phylogenetic signal here.
+    , especially when using large matrices \citep{lemmonthe2009}.
+However, this may not be a major issue because phylogenetic signal is easily increased by:
+(i) including a "modest" number of highly-covered genes (i.e.
+approximatively of the genes that are coded for most of the species; \citealt{rouresite-specific2011}) %NC: what do you mean by highly-covered?
 (ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and
@@
 However, the phylogenetic signal is not related to the amount of missing data \textit{per se} but to the number of informative characters for each taxon, therefore missing data is less of an issue than the number of shared informative characters \citep{wiensmissing2003}.
-% NC: I'm a little dubious of what this paragraph is trying to get at in places. We should discuss at some point.
+% NC: I'm a little dubious of what this paragraph is trying to get at in places.
+We should discuss at some point.
+TG: my idea was to do the "literature-review" part of the intro: missing data is already widely studied (answer to the comment: "hun...
+We already knew that") but it has never been studied in TEM framework.
+And that's kind of the point we're getting toward with our work: we don't really care about the effect of missing data per se (i.e.
+what's the threshold were it starts fucking up the phylogeny) but more the fact that Bayesian fails.
 
@@
 % NC: Somewhere here you need to flag that you're only looking at topologies and briefly explain why.
+In this study, we focused only on topology as one of the two aspects of the phylogenetic signal (topology and branch length).
+Even though both aspects are equally important, branch topology is the first and most straightforward aspect reflecting phylogenetic signal (i.e.
+topological changes are discrete opposed to branch length changes are continuous).
+Also, interestingly, the effect of Total Evidence method has not been formally assessed in previous studies using fixed topology \citep{ronquista2012,schragocombining2013}.
+%TG: or is that mean for them?
 
@@
 \item the proportion of living taxa with no morphological data;
-\item the amount of missing data in the fossil taxa; and
-\item the overall number of morphological characters for both living and fossil taxa in the matrix.
+\item the proportion of missing data in the fossil taxa; and
+\item the proportion of missing morphological characters for both living and fossil taxa in the matrix.
 \end{enumerate}
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
-We chose these parameters because they reflect real-life biases in data availability.
+We chose these parameters because they reflect empirical biases in data availability.
 The advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
-Missing data in fossil taxa is very common due to preservation biases (citation?), and
-% OK I'm not sure what the "real-life" reason for three is.
-%This paragraph is crap
+Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them.
+% NC: OK I'm not sure what the "real-life" reason for three is. - TG: fixed to "empirical" (or "practical" ?).
+% TG: This paragraph is crap
 % NC: I think it's fine!
@@
 We make suggestions for how best to deal with this issue when inferring phylogenies from Total Evidence matrices.
-
 % NC: You don't want to tell the whole story here. Just a results summary and a quick line of interpretation
@@
 %We also propose a new method for comparing Bayesian posterior tree distribution's topology available in a wonderful R package. %Or am I just dreaming a bit too much here? 
-
 % NC: Dreaming a bit much I think.
@@
 
+%---------------------------------------------
+%
+%       METHODS
+%
+%---------------------------------------------
+
 \newpage
+
 \section{Methods} %no material needed?
-% NC: Depends on the journal. See what Syst Biol asks for as section titles.
+% NC: Depends on the journal.
+See what Syst Biol asks for as section titles.
+TG: Syst Biol instructions :Elements vary, of course, depending on the type of paper.
+The first-level heading is often used for sections such as ``methods and materials,'' ``results,'' and' ``discussion;'' but other major sections are often appropriate instead.
 To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
@@
 or Figure?
+- TG: it's "Fig." when it's a reference in brackets and "Figure" when it's a reference in the text.
+Table is always refered as "Table" (not "Tab").
 \begin{enumerate}
@@
 \item{Removing data} \label{step:remove_data} \\
-We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of missing living taxa ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$) % NC: make sure to use the same descriptions here and in the introduction
+We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$) % NC: make sure to use the same descriptions here and in the introduction
 (the resulting matrices are called hereafter "missing-data" matrices; Table ~\ref{Tab_glossary}).
@@
 \item{Comparing topologies} \label{step:compare_topo} \\
-We compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies.
+We compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies
 \end{enumerate}
-We repeated steps 1-4 50 times. %that's because I had 3 cluster accounts available... If it's a too odd number, no worries, I'll just remove the last chain.
+We repeated steps 1 to 4 50 times. %that's because I had 3 cluster accounts available... If it's a too odd number, no worries, I'll just remove the last chain.
 % NC: Yeah 51 is odd. Let's just use 50.
 A list of all the terms used in this paper is available in Table ~\ref{Tab_glossary}.
+
+% NC: Not sure we need the glossary, but keep it for the thesis
+
+\begin{table}[ht]
+  \caption{Glossary}
+  \centering
+  \linespread{1.0}
+  \input{Tables/Glossary}
+  \label{Tab_glossary}
+\end{table}
 
@@
 (2) We removed data from the morphological part of the "complete" matrix resulting in 124 "missing-data" matrices.
-% NC: I thought we had 124 missing data matrices... I just worked this out - one of the "missing data" matrices is 0%,0%,0% so = complete matrix.
+% NC: I thought we had 124 missing data matrices...
+I just worked this out - one of the "missing data" matrices is 0%,0%,0% so = complete matrix.
+- TG: Yep, there is 125 matrices/trees per chain.
+1 that's L00F00C00 and 124 others.
+It's again one of these habits like "adding missing data" that I have from the coding (that actually do add missing data and do generate 125 matrices).
 (3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
@@
 Also let's show this figure to a couple of people to see if they get it.
-
-% NC: Not sure we need the glossary, but keep it for the thesis
-
-\begin{table}[ht]
-\caption{Glossary}
-\centering
-\begin{tabular}{p{5cm} p{9cm}}
-\hline
-\hfill Term & Definition \\[0.5ex]
-\hline
-\hfill living taxa & taxa with both molecular and morphological data available \\[1.5ex]
-\hfill fossil taxa & taxa with only morphological data available \\[1.5ex]
-\hfill "complete" matrix & matrix with no missing data except for the molecular part of the fossil taxa \\[1.5ex]
-\hfill "missing-data" matrix & matrix with various amount of missing data \\[1.5ex]
-\hfill $M_{L}$ & missing living taxa in the morphological part of the matrix \\[1.5ex]
-\hfill $M_{F}$ & missing data for the fossil taxa in morphological part of the matrix \\[1.5ex]
-\hfill $M_{C}$ & missing morphological characters for both living and fossil taxa \\[1.5ex]
-\hfill "true" tree & tree used to simulate the matrix \\[1.5ex]
-\hfill "best" tree & tree inferred from the "complete" matrix \\[1.5ex]
-\hfill "missing-data" tree & tree inferred the "missing-data" matrices \\[1.5ex]
-\hfill RPBTC & Random pairwise Bayesian tree comparison \\[1.5ex]
-\hfill "horizontal" sub-matrix & sub-matrix with morphological and molecular characters for living taxa \\[1.5ex]
-\hfill "vertical" sub-matrix & sub-matrix with morphological data for both living and fossil taxa \\[1.5ex]
-\hfill "corner" sub-matrix & sub-matrix with morphological data for living taxa \\[1.5ex]
-\hfill "missing-data" sub-matrix & molecular part of the matrix for the fossil taxa (no data) \\
-\end{tabular}
-\label{Tab_glossary}
-\end{table}
+-TG: Do you mean just swap the text ML/Bayesian (was: Bayesian-ML, is now: ML-Bayesian)?
+
+%---------------------------------------------
+%       Generating the matrix
+%---------------------------------------------
 
@@
 We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
-We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined (even with the parameters defined, total computational time was over 65 CPU years).
+We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, total computational time for the whole analysis was over 80 CPU years).
 All the molecular information for fossil taxa was replaced by missing data ("?").
@@
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used low evolutionary rate parameters (i.e.
-$\alpha$) to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
+We used low evolutionary rate parameters (i.e. $\alpha$) to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
 
@@
 This reflects the fact that because of the increasing availability of DNA sequences for living taxa, detailed morphological data is scarce.
-\item{The proportion of missing morphological data across all fossil taxa ($M_{F}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
+\item{The proportion of missing data in the fossil taxa ($M_{F}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
 This parameter illustrates the quality of the fossil record. 
-\item{The proportion of missing morphological characters across all taxa (living and fossil - $M_{C}$): 0\%, 10\%, 25\%, 50\% or 75\%. }
+\item{the proportion of missing morphological characters for both living and fossil taxa ($M_{C}$): 0\%, 10\%, 25\%, 50\% or 75\%. }
 This parameter illustrates the number of available morphological characters for both living and fossil taxa.
@@
 In practice, each parameter represents a different way of removing data from the matrix: $M_{L}$ removes rows from the living taxa; $M_{F}$ removes cells from the fossil taxa; and $M_{C}$ removes columns across both living and fossil taxa (see Fig.
-~\ref{Fig_RemoveData}).Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a percentage of living taxa is removed, but for $M_{F}$, a percentage of the data is removed at random from across the whole of the morphological matrix for fossil taxa.
+~\ref{Fig_RemoveData}).
+Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a percentage of living taxa is removed, but for $M_{F}$, a percentage of the data is removed at random from across the whole of the morphological matrix for fossil taxa.
 
@@
 Note that one of these combinations has no missing data so is equivalent to the "complete" matrix, thus we have 124 "missing-data" matrices.
-Because some parameter combinations introduce a lot of missing data (e.g.
-$M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
-When this occurred we repeated the random deletion of characters until every taxon had at least 5\% data across the whole matrix.
+Because some parameter combinations introduce a lot of missing data (e.g. $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
+When this occurred we repeated the random deletion of characters until every taxa had at least 5\% data across the whole morphological part of the matrix.
 
@@
 \subsection{Building phylogenies}
-From the resulting matrices we generated two types of trees, the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 124 matrices with various amounts of missing data.The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
+From the resulting matrices we generated two types of trees, the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 124 matrices with various amounts of missing data.
+The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
 The "best" tree, on the other hand, is the best tree we can build using state-of-the-art phylogenetic methods.
-In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g.
-\citet{rozen2005}).
+In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g. \citealt{rozen2005}).
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
@@
 The "best" tree and the "missing-data" trees were inferred using RAxML v8.0.20 \citep{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model (\citet{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citet{Stamatakis21012014}) as a generalization of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
-The GTR model can be seen as a generalization of the HKY model (the two parameters from the HKY model are implicitly included in the six from GTR model - \citet{stamatakisa2008}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} which is a generalization of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citet{Stamatakis21012014}).
-We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support.
-%The explanations of the algorithm part for the BS was proposed by Trevor
-The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists of pruning one sub-tree from the tree and subsequently reinserting it to all neighboring branches \citep{stamatakisa2008}.
+For the molecular data, we used the GTR + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citealt{Stamatakis21012014}) as a generalization of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
+The GTR model can be seen as a generalization of the HKY model (the two parameters from the HKY model are implicitly included in the six from GTR model - \citealt{stamatakisa2008}).
+For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} which is a generalization of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citealt{Stamatakis21012014}).
+We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. %The explanations of the algorithm part for the BS was proposed by Trevor
+The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists of pruning one sub-tree from the tree and subsequently reinserting it to all neighbouring branches \citep{stamatakisa2008}.
 Sub-tree Pruning and Reinserting methods (SPR) have been demonstrated to be better than other methods (e.g.
-Nearest Neighboring Interchange - NNI) for recovering good bootstrap values \citep{salamin2003}.
+Nearest Neighbouring Interchange - NNI) for recovering good bootstrap values \citep{salamin2003}.
+When using these parameters, it took 13 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (8 core nodes 2.30GHz clock speed).
 
@@
 For each run, we removed 25\% of the iterations as burn-in.
-We used the following priors for each tree (see \hyperref[supplementaries]{supplementaries}):
+We used the following priors for each tree (see \hyperref[SupplementaryMaterial]{Supplementary Material S1}):
 \begin{enumerate}
@@
 Even using these priors, it took 65 CPU years to build 50 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
-% NC: OK so the 65 CPU is for building trees not matrices? Or is it combined time? Or does both stage take 65 CPU years? Be clearer in the earlier sections
+% NC: OK so the 65 CPU is for building trees not matrices?
+Or is it combined time?
+Or does both stage take 65 CPU years?
+Be clearer in the earlier sections.
+- TG: I fixed that above: 65 years for the Bayesian, 13 years for the ML, 1 year for all the rest (generating, comparing, etc...)
 
@@
 You may be able to cut some more stuff out.
+- TG: since I'm just using Bogdanowicz allready published code, I maybe don't need to detail the metrics at all?
+Just say what they measure and shovel all the detail in supp?
 
@@
 % NC: I moved this around a bit to make it clearer what metric fits where. Remember we need American spellings.
-% Also I thought it was triplets distance? Good idea to check...
+% NC: Also I thought it was triplets distance?
+Good idea to check...
+- TG: It's pretty confusing, unfortunately I can't check Dobson 1975 because he was just cited in Critchlow 1996 and Johnson 1998 as the author of the method but I never got my hand on it.
+I don't remember what's the name in Johnson 1998 (Trevor's book) but it might well be 'triplets' since I started from there.
+However, it is clearly 'Triples' distance in Critchlow 1996 and in Bodganowicz 2012 (that uses QuarteT but TripleS and ref back to Critchlow 1996).
+Let's go for Triples?
+In the end if the reviewers wants us to change it, it should be a minor issue.
 
@@
 But the general meaning is that the placements differ between trees.
-Triples distance ($T_{x,y}$; \citealp{dobson1975triplets}) measures the number of sub-trees made up of three taxa (triplets) that differ between two given trees.
-Each triplet can be written as $I_{ijk}$=(\textit{ijk}).
-Where $I_{ijk}$ is equal to zero if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to one.
-For any rooted binary tree there are only three possible combinations for each triplet: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
-If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k}).
-We can calculate the triplet distance between two trees, $S_n$, as:
-\begin{equation}
-S_n = \sum_{ijk} I_{ijk}
-\end{equation}
-Where:
-\begin{equation}
-\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
-\end{equation}
-And where \textit{n} is the total number of taxa in both trees (modified from \citet{critchlowthe1996}).
-If $S_n$ = 0, the trees are identical; when $S_n$ = $\binom{n}{4}$, the trees are as different as possible (i.e.
-every taxon has a different placement in the two trees).Triples distance is less sensitive to the placement of individual taxa and to taxa with highly uncertain placement (e.g.
-fossil taxa) than the Robinson-Foulds metric \citep{critchlowthe1996,johnson1998,wiensmissing2003}.
-% NC: Not sure you need this bit
-% Therefore we used this metric as a proxy to estimate the robustness of the tree to flying taxa (see \hyperref[supplementaries]{supplementaries}).
-
-\subsubsection{Robinson-Foulds distance} % NC: Is it Robinson Fould or Robinson Fould*s*?
+The triples distance ($T_{x,y}$; \citealp{dobson1975triplets}) measures the number of sub-trees made up of three taxa that differ between two given trees (\citealt{critchlowthe1996} ; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
+This metric measures the position of each taxon and clade towards it's closest neighbours.
+It is bounded between 0 when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is not one single position of taxon/clade identical between both trees.
+Therefore this metric sensitive to the conservation of individual taxa towards the neighbouring trees.
+
+\subsubsection{Robinson-Foulds distance} % NC: Is it Robinson Fould or Robinson Fould*s*? - TG: Definitely FouldS. R. Foulds is the guys name.
 Robinson-Foulds distance (\textit{RF}; \citealp{RF1981}), or "path difference"
-% NC: path difference or *distance*?
+% NC: path difference or *distance*? - TG: difference.
     , measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} and can be expressed as following:
+The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981} ; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
+This metric is bounded between 1 when the two trees are identical and $n-2$ (for two trees with $n$ taxa) when there is not one single shared clade between both trees.
+Contrary to the Triples distance, this metric is more sensitive to the exact clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swap of two taxa will lead to a maximal score of the Robinson-Fould distance indicating a bad tree similarity.
+
+%TG: maybe draw a figure to illustrate that? Triples = not really stric, RF= really strict.
+
 \begin{equation}
@@
 \textit{NTS} ranges from one to -$\infty$.
-For any $m,n$, when \textit{NTS} = 1, the trees are identical, when \textit{NTS} = 0 the trees are no more different than expected by chance, and when \textit{NTS} $<$ 0, the trees are more different than expected by chance (see \hyperref[supplementaries]{supplementaries}).\\
+For any $m,n$, when \textit{NTS} = 1, the trees are identical, when \textit{NTS} = 0 the trees are no more different than expected by chance, and when \textit{NTS} $<$ 0, the trees are more different than expected by chance %chance.\\
 % NC: You seem to be using italics almost at random here. Why???
@@
 
-\subsection{Building the trees}
-When generating the matrices using seqgen \citep{ranbaut1997seqgen} and rTraitDisc \citep{paradisape:2004} algorithms with low evolutionary rate parameter ($\alpha$=0.5) we successfully generated evolutionary histories.
-In Maximum Likelihood, support values ranged from 0 to 100 with a median of 38 ($1^st$ quartile = 14, $3^rd$ quartile = 72 - see \hyperref[supplementaries]{supplementaries}). % This is actually bad.
-The distribution of the phylogenetic signal across the different chains provided us a good spread of evolutionary scenarios: from well resolved phylogenies to poorly resolved ones.
-This allows us to test our simulations in a theoretical as well as practical framework (i.e.
+% NC: OK it's worth thinking about what you really need to include here.
+Remember the results are about the answers to your question.
+Not necessarily everything you did.
+This section can be really short, and probably should be quite short.
+
+% NC: I may have misunderstood but I don't think you need this section at all, or the comparing topologies section you have planned. Just the actual results surely...
+
+%\subsection{Building the trees}
+%When generating the matrices using seqgen \citep{ranbaut1997seqgen} and rTraitDisc \citep{paradisape:2004} algorithms with low evolutionary rate parameter ($\alpha$ = 0.5) we successfully generated evolutionary histories.
+%In Maximum Likelihood, support values ranged from 0 to 100 with a median of 38 ($1^st$ quartile = 14, $3^rd$ quartile = 72 - see \hyperref[supplementaries]{supplementaries}). % This is actually bad.
+%The distribution of the phylogenetic signal across the different chains provided us a good spread of evolutionary scenarios: from well resolved phylogenies to poorly resolved ones.
+%This allows us to test our simulations in a theoretical as well as practical framework (i.e.
 phylogenies with low support values are rarely published but are still common in early stages of projects).
@@
 - That's kind of my way to turn these bad results into something usable: even if some of our simulations are not realistic, it shows the effect of missing data in any scenario possible, not only in realistic cases.
-Also, one can advocate that this low topological resolution is actively due to the amount of missing data in the total evidence matrices since the sub-matrices (no missing data) analysis recovers way higher support value  .
+%Also, one can advocate that this low topological resolution is actively due to the amount of missing data in the total evidence matrices since the sub-matrices (no missing data) analysis recovers way higher support value  .
 %This is something that I will had with the whole submatrices story.
@@
 our REAL control test: can we simulate matrices in the first place).
-Using Bayesian inference, all the chains converged with a significant effective sample size by using 2 parallel runs of 4 chains each (ASSD $<$ 0.01 and ESS $>>$ 200 for inference).
+%Using Bayesian inference, all the chains converged with a significant effective sample size by using 2 parallel runs of 4 chains each (ASSD $<$ 0.01 and ESS $>>$ 200 for inference).
 
@@
 
-\subsection{Effect of the missing data parameters}
+%\subsection{Effect of the missing data parameters}
+% NC: Probably don't need a subheading here
 
@@
 \caption{Effect of missing data on Tree similarity.
-$M_L$ is the amount of missing living taxa with morphological data,
-$M_F$ is the amount of missing data in the fossil record,
-$M_C$ is the amount of missing morphological characters.
-The first row represent the Normalised Tree Similarity index using the Robinson-Fould metric;
-the second row represent the Normalised Tree Similarity index using the Triples metric.}
+$M_L$ is the percentage of living taxa with no morphological data,
+$M_F$ is the percentage of missing data in the fossil record,
+$M_C$ is the percentage of missing morphological characters.
+The first row represent the Normalized Tree Similarity index using the Robinson-Foulds metric;
+the second row represent the Normalized Tree Similarity index using the Triples metric.}
 \label{Fig_Results}
@@
 
-\subsubsection{Effect of $M_L$}
-In Maximum Likelihood framework, the topological recovery (reflected by the tree similarity) quickly drops with the increasing amount of missing data when using both metrics (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-However, this negative effect of missing data is stabilized beyond 25\% of missing data (no significant difference for trees with respectively 25\%, 50\% and 75\% missing living taxa).
-With 10\% of missing living taxa, there is a different topological recovery depending on the used metric.
-The Robinson-Fould metric shows a lower Tree similarity index than the Triples metric meaning that individual taxa position are generally more conserved (i.e.
-low number of flying taxa) but that the full clades are less conserved (see Table ~\ref{NTS_ML_results}).
-
-However, topological recovery was bad in Bayesian, regardless the amount of missing data.
-The tree similarity index was only slightly above 0 which means that the mode of the RPBTC was only slightly different than expected by chance (see Table ~\ref{NTS_Ba_results}).
-Even if the topological recovery is really low, we detected an effect of missing data on the topology depending on the used metric.
-For the Triples metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
-For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases (see Table ~\ref{Ba_RF-ML_results}).
-However, in both cases, the effect of the parameter $M_L$ is still minimal compared to trees generated in a ML framework since the tree similarity index was only slightly higher than expected by chance (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
-
+%NC: OK from this point on you start referring to "Tree Similarity Index". This is not mentioned in the methods. What is it? You need to explain in the methods please.
+
+%NC: I'm not sure you need to separate these three parameters out.
+Aren't you repeating yourself a lot?
+Rewrite this with them all in together.
+You can talk about general patterns and then specific, interesting patterns.
+I've corrected the first one a bit to help out.
+\subsection{Effect of $M_L$}
+In our Maximum Likelihood analyses, topological recovery (i.e.
+tree similarity) decreased as the percentage of missing data increased, using both the Triples and Robinson-Foulds metrics (see Table ~\ref{Group_results} - Fig.~\ref{Fig_Results}).
+However, topological recovery was not significantly different for trees with 25\%, 50\% and 75\% missing morphological data for living taxa (Fig 1b???), % is this what you meant?
+And indicate the particular panel of the figure it refers to
+suggesting that although tree topology is sensitive to this missing data, once it is greater than 25\% it makes little difference.
+Overall, the Robinson-Foulds metric shows lower tree similarities than the Triples metric.
+This indicates that the positions of individual taxa are generally more conserved than the membership of entire clades (see Table ~\ref{NTS_ML_results}).
+
+Topological recovery was poor in our Bayesian analyses, regardless of the amount of missing data (Figures???), although topological recovery did decrease slightly as missing data increased.
+Overall, tree similarity index was only slightly above zero which means that trees are only slightly more similar than we would expect if we compared two completely random trees (see Table ~\ref{NTS_Ba_results}).
+%For the Triples metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
+%For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases (see Table ~\ref{Ba_RF-ML_results}).
+
+%NC: A shorter verison of this encompassing all three parameters will work here.
+
+%NC: Combine with section above
 \subsubsection{Effect of $M_F$}
@@
 
+%NC: Combine with section above
 \subsubsection{Effect of $M_C$}
@@
 
+%NC: This will be part of the section above. If it's short you may not need to sum up!
 %To sum up:
-The ability of recovering the "best" tree's topology in Maximum Likelihood method is function of the amount of data missing.
-The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
-The decrease in clade conservation (low Robinson-Fould NTS score) is faster than the increase in flying taxa (low Triples NTS score).
-In Bayesian framework, topology is badly (if not at all) recovered disregarding the amount of data missing.
-
-\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
+%The ability of recovering the "best" tree's topology in Maximum Likelihood method is function of the amount of data missing.
+%The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
+%The decrease in clade conservation (low Robinson-Fould NTS score) is faster than the increase in flying taxa (low Triples NTS score).
+%In Bayesian framework, topology is badly (if not at all) recovered disregarding the amount of data missing.
+
+% NC: You will need two sections here with subheadings - one for the sensitivity analyses you did with the submatrices, and one for the empirical results.
+
+% NC: I would make these tables wider - so have a column for Triples, and one for RF, rather than a column for metric.
+same with Bayesian vs ML, why not have a set of columns instead?
+It'll take up less space and be easier to read I think...
+
+% NC: You should do what I have done on the thesis template and have these tables as separate files rather than having them in the document.
+
+% NC: we need to sit down and look at these tables. I don't like the current format.
+
+% About the tables:
+
+% What about making this table in a longevity-results style? with every thing combined (30 columns)?
+% NC: That might be too many columns!
+
+\begin{table}
+  \centering
+  \linespread{1.0}
   \caption{Tree similarity values per parameter in ML framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
-              & missing data &        &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.671 & 0.569  & 0.789 & 0.468  & 1.036 \\ %how many digits?
-              &              & Triples        & 0.942 & 0.868  & 0.999 & 0.612  & 1.063 \\
-              & 25\%         & Robinson-Fould & 0.516 & 0.440  & 0.630 & 0.266  & 0.892 \\
-              &              & Triples        & 0.559 & 0.708  & 0.810 & 0.291  & 0.999 \\
-              & 50\%         & Robinson-Fould & 0.432 & 0.364  & 0.521 & 0.231  & 0.642 \\
-              &              & Triples        & 0.514 & 0.419  & 0.599 & 0.293  & 0.910 \\
-              & 75\%         & Robinson-Fould & 0.372 & 0.323  & 0.434 & 0.203  & 0.577 \\
-              &              & Triples        & 0.444 & 0.338  & 0.532 & 0.192  & 0.855 \\
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.907 & 0.766  & 1.000 & 0.407  & 1.047 \\
-              &              & Triples        & 0.976 & 0.944  & 1.003 & 0.715  & 1.104 \\
-              & 25\%         & Robinson-Fould & 0.572 & 0.488  & 0.692 & 0.325  & 0.936 \\
-              &              & Triples        & 0.937 & 0.820  & 0.993 & 0.539  & 1.086 \\
-              & 50\%         & Robinson-Fould & 0.433 & 0.302  & 0.476 & 0.170  & 0.668 \\
-              &              & Triples        & 0.732 & 0.594  & 0.873 & 0.341  & 1.025 \\
-              & 75\%         & Robinson-Fould & 0.153 & 0.115  & 0.200 & 0.068  & 0.323 \\
-              &              & Triples        & 0.458 & 0.390  & 0.611 & 0.290  & 0.856 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.473 & 0.384  & 0.835 & 0.356  & 0.958 \\
-              &              & Triples        & 0.959 & 0.869  & 1.026 & 0.572  & 1.067 \\
-              & 25\%         & Robinson-Fould & 0.454 & 0.357  & 0.630 & 0.242  & 0.916 \\
-              &              & Triples        & 0.907 & 0.751  & 0.997 & 0.446  & 1.070 \\
-              & 50\%         & Robinson-Fould & 0.424 & 0.375  & 0.466 & 0.249  & 0.828 \\
-              &              & Triples        & 0.683 & 0.571  & 0.822 & 0.277  & 1.008 \\
-              & 75\%         & Robinson-Fould & 0.230 & 0.177  & 0.302 & 0.108  & 0.567 \\
-              &              & Triples        & 0.451 & 0.373  & 0.572 & 0.224  & 0.791 \\
-    \hline
-\label{NTS_ML_results}
-\end{tabular}
-\end{table}
-
-\begin{table} % What about making this table in a longevity-results style? with every thing combined (30 columns)?
+  \input{Tables/NTS_ML_results}
+  \label{tab:NTS_ML_results}
+\end{table}
+
+\begin{table}
+  \centering
+  \linespread{1.0}
   \caption{Tree similarity values per parameter in Bayesian framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
-              & missing data &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\ %how many digits?
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.028   & 0.061 \\
-              &              & Triples        & 0.033 & 0.003  & 0.049 & -0.049  & 0.117 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.061 \\
-              &              & Triples        & 0.035 & 0.002  & 0.049 & -0.041  & 0.096 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.056 \\
-              &              & Triples        & 0.028 & 0.003  & 0.052 & -0.038  & 0.092 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.010   & 0.043 \\
-              &              & Triples        & -0.001 & -0.018 & 0.040 & -0.060 & 0.106 \\
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.035 & 0.031  & 0.056 & 0.027   & 0.076 \\
-              &              & Triples        & 0.026 & 0.002  & 0.050 & -0.046  & 0.105 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.0153  & 0.068 \\
-              &              & Triples        & 0.038 & 0.002  & 0.050 & -0.046  & 0.105 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.058 \\
-              &              & Triples        & 0.014 & -0.011 & 0.039 & -0.038  & 0.124 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.015   & 0.041 \\
-              &              & Triples        & 0.029 & 0.019  & 0.042 & -0.030  & 0.089 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 0.056 & 0.035  & 0.063 & 0.031   & 0.076 \\
-              &              & Triples        & 0.041 & 0.008  & 0.054 & -0.039  & 0.106 \\
-              & 10\%         & Robinson-Fould & 0.056 & 0.035  & 0.059 & 0.0281  & 0.076 \\
-              &              & Triples        & 0.030 & 0.011  & 0.048 & -0.024  & 0.094 \\
-              & 25\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.023   & 0.076 \\
-              &              & Triples        & 0.020 & -0.001 & 0.037 & -0.060  & 0.109 \\
-              & 50\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.030   & 0.059 \\
-              &              & Triples        & 0.010 & -0.005 & 0.045 & -0.036  & 0.104 \\
-              & 75\%         & Robinson-Fould & 0.035 & 0.035  & 0.035 & 0.011   & 0.043 \\
-              &              & Triples        & 0.024 & 0.001  & 0.042 & -0.037  & 0.104 \\
-    \hline
+  \input{Tables/NTS_BA_results}
   \label{NTS_BA_results}
-\end{tabular}
-\end{table}
-
-\begin{table}
+\end{table}
+
+\begin{table}
+  \centering
+  \linespread{1.0}
   \caption{Group difference per configuration set}
-\centering
-\begin{tabular}{rllcccc}
-    \hline
-    set & metric & framework & parametric & $stats^1$ & df & p.value \\
-    \hline
-    $M_L$         & Robinson-Fould & ML       & no & 151.438  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.001278e-31
-                  &                & Bayesian & no & 65.62    & 4   & \textbf{0.00} \\ 
-                  & Triples        & ML       & no & 142.8329 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains  - 6.983434e-30
-                  &                & Bayesian & no & 4.56     & 4   & 0.34          \\ 
-    $M_F$         & Robinson-Fould & ML       & no & 179.9238 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 7.743067e-38
-                  &                & Bayesian & no & 85.61    & 4   & \textbf{0.00} \\
-                  & Triples        & ML       & no & 150.059  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.977386e-31
-                  &                & Bayesian & no & 3.01     & 4   & 0.56          \\
-    $M_C$         & Robinson-Fould & ML       & no & 151.3818 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.029428e-31
-                  &                & Bayesian & no & 64.23    & 4   & \textbf{0.00} \\
-                  & Triples        & ML       & no & 138.374  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 6.290979e-29
-                  &                & Bayesian & no & 26.13    & 24  & 0.35          \\
-    $M_L+M_F$     & Robinson-Fould & ML       & no & 734.9213 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.104542e-139
-                  &                & Bayesian & no & 317.97   & 24  & \textbf{0.00} \\
-                  & Triples        & ML       & no & 531.5558 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 4.581065e-97
-                  &                & Bayesian & no & 3.01     & 24  & 0.56          \\
-    $M_L+M_C$     & Robinson-Fould & ML       & no & 542.3233 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 2.619808e-99
-                  &                & Bayesian & no & 290.10   & 24  & \textbf{0.00} \\
-                  & Triples        & ML       & no & 437.7413 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains -  1.28493e-77
-                  &                & Bayesian & no & 22.19    & 24  & 0.57          \\
-    $M_F+M_C$     & Robinson-Fould & ML       & no & 812.206  & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 5.463138e-156
-                  &                & Bayesian & no & 385.96   & 24  & \textbf{0.00} \\
-                  & Triples        & ML       & no & 528.9215 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.619383e-96
-                  &                & Bayesian & no & 20.23    & 24  & 0.68          \\ 
-    $M_L+M_F+M_C$ & Robinson-Fould & ML       & no & 3603.808 & 124 & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains -3.819574e-169
-                  &                & Bayesian & no & 1167.40  & 124 & \textbf{0.00} \\
-                  & Triples        & ML       & no & 2047.923 & 124 & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains 
-                  &                & Bayesian & no & 115.86   & 124 & 0.69          \\
-    \hline
+  \input{Tables/Group_results}
   \label{Group_results}
-\end{tabular}
-
   $^1$ F-value for parametric tests and Kruskal Wallis $chi^2$ for non parametric tests.
@@
 \begin{table}
+  \centering
+  \linespread{1.0}
   \caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE  & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE  \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 56.53 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$25\% & 96.85 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 127.03 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 144.58 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 40.31 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$50\% & 70.50 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 88.05 & 37.66 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 30.19 & 37.66 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 47.73 & 37.66 & TRUE \\ 
-  $M_L$50\%-$M_L$75\% & 17.55 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
+  \input{Tables/ML_RF-ML_results}
   \label{ML_RF-ML_results}
@@
 \begin{table}
+  \centering
+  \linespread{1.0}
   \caption{$M_L$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & \textbf{TRUE}  & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE  & FALSE \\
-    $M_L$50\% & & & & - & FALSE  \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 57.98 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$25\% & 105.83 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 120.86 & 37.66 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 140.34 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 47.85 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$50\% & 62.88 & 37.66 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 82.36 & 37.66 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 15.03 & 37.66 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 34.51 & 37.66 & FALSE \\ 
-  $M_L$50\%-$M_L$75\% & 19.48 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
+  \input{Tables/ML_Tr-ML_results}
   \label{ML_Tr-ML_results}
@@
 \begin{table}
-\caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_L$00\% & $M_L$10\% & $M_L$25\% & $M_L$50\% & $M_L$75\% \\
-    \hline
-    $M_L$00\% & - & FALSE & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_L$00\%-$M_L$10\% & 17.05 & 41.00 & FALSE \\ 
-  $M_L$00\%-$M_L$25\% & 43.79 & 41.00 & TRUE \\ 
-  $M_L$00\%-$M_L$50\% & 77.32 & 41.00 & TRUE \\ 
-  $M_L$00\%-$M_L$75\% & 103.99 & 41.00 & TRUE \\ 
-  $M_L$10\%-$M_L$25\% & 26.75 & 41.00 & FALSE \\ 
-  $M_L$10\%-$M_L$50\% & 60.27 & 41.00 & TRUE \\ 
-  $M_L$10\%-$M_L$75\% & 86.94 & 41.00 & TRUE \\ 
-  $M_L$25\%-$M_L$50\% & 33.53 & 41.00 & FALSE \\ 
-  $M_L$25\%-$M_L$75\% & 60.20 & 41.00 & TRUE \\ 
-  $M_L$50\%-$M_L$75\% & 26.67 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
+  \centering
+  \linespread{1.0}
+  \caption{$M_L$ non-parametric pairwise difference for Robinson-Foulds metric in Bayesian framework}
+  \input{Tables/Ba_RF-ML_results}
   \label{Ba_RF-ML_results}
@@
 \begin{table}
-\caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & \textbf{TRUE} \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 59.27 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$25\% & 79.72 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$50\% & 120.70 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 167.81 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 20.45 & 37.66 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 61.43 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 108.55 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 40.98 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$75\% & 88.09 & 37.66 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 47.12 & 37.66 & TRUE \\ 
-   \hline
-\end{tabular}
+  \centering
+  \linespread{1.0}
+  \caption{$M_F$ non-parametric pairwise difference for Robinson-Foulds metric in ML framework} %10/06 - Missing 8 ML Chains
+  \input{Tables/ML_RF-MF_results}
   \label{ML_RF-MF_results}
@@
 \begin{table}
+  \centering
+  \linespread{1.0}
   \caption{$M_F$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 64.78 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$25\% & 89.09 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$50\% & 123.20 & 37.66 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 150.43 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 24.31 & 37.66 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 58.42 & 37.66 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 85.65 & 37.66 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 34.10 & 37.66 & FALSE \\ 
-  $M_F$25\%-$M_F$75\% & 61.34 & 37.66 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 27.23 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
+  \input{Tables/ML_Tr-MF_results}
   \label{ML_Tr-MF_results}
@@
 \begin{table}
+  \centering
+  \linespread{1.0}
   \caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_F$00\% & $M_F$10\% & $M_F$25\% & $M_F$50\% & $M_F$75\% \\
-    \hline
-    $M_L$00\% & - & FALSE & FALSE & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_L$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_L$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_L$50\% & & & & - & FALSE \\
-    $M_L$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_F$00\%-$M_F$10\% & 16.03 & 41.00 & FALSE \\ 
-  $M_F$00\%-$M_F$25\% & 38.50 & 41.00 & FALSE \\ 
-  $M_F$00\%-$M_F$50\% & 61.52 & 41.00 & TRUE \\ 
-  $M_F$00\%-$M_F$75\% & 95.52 & 41.00 & TRUE \\ 
-  $M_F$10\%-$M_F$25\% & 22.47 & 41.00 & FALSE \\ 
-  $M_F$10\%-$M_F$50\% & 45.49 & 41.00 & TRUE \\ 
-  $M_F$10\%-$M_F$75\% & 79.49 & 41.00 & TRUE \\ 
-  $M_F$25\%-$M_F$50\% & 23.02 & 41.00 & FALSE \\ 
-  $M_F$25\%-$M_F$75\% & 57.02 & 41.00 & TRUE \\ 
-  $M_F$50\%-$M_F$75\% & 34.00 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
+  \input{Tables/Ba_RF-MF_results}
   \label{Ba_RF-MF_results}
@@
 \begin{table}
+  \centering
+  \linespread{1.0}
   \caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & \textbf{TRUE} \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 67.99 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$25\% & 87.58 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$50\% & 116.16 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 155.77 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 19.59 & 37.66 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 48.17 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 87.78 & 37.66 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 28.58 & 37.66 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 68.19 & 37.66 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 39.60 & 37.66 & TRUE \\ 
-   \hline
-\end{tabular}
+  \input{Tables/ML_RF-MC_results}
   \label{ML_RF-MC_results}
@@
 \begin{table}
+  \centering
+  \linespread{1.0}
   \caption{$M_C$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & FALSE \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 75.29 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$25\% & 88.58 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$50\% & 114.02 & 37.66 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 149.60 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 13.29 & 37.66 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 38.73 & 37.66 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 74.31 & 37.66 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 25.44 & 37.66 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 61.02 & 37.66 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 35.58 & 37.66 & FALSE \\ 
-   \hline
-\end{tabular}
+  \input{Tables/ML_Tr-MC_results}
   \label{ML_Tr-MC_results}
@@
 \begin{table}
+  \centering
+  \linespread{1.0}
   \caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-\centering
-\begin{tabular}{c|ccccc}
-    \hline
-              & $M_C$00\% & $M_C$10\% & $M_C$25\% & $M_C$50\% & $M_C$75\% \\
-    \hline
-    $M_C$00\% & - & FALSE & FALSE & \textbf{TRUE} & \textbf{TRUE}\\
-    $M_C$10\% & & - & FALSE & \textbf{TRUE} & \textbf{TRUE} \\
-    $M_C$25\% & & & - & FALSE & \textbf{TRUE} \\
-    $M_C$50\% & & & & - & FALSE \\
-    $M_C$75\% & & & & & - \\
-    \hline
-\end{tabular}
-\centering
-\begin{tabular}{rrrl}
- pairwise comparison & obs.dif & critical.dif & difference \\ 
-  \hline
-  $M_C$00\%-$M_C$10\% & 16.03 & 41.00 & FALSE \\ 
-  $M_C$00\%-$M_C$25\% & 38.50 & 41.00 & FALSE \\ 
-  $M_C$00\%-$M_C$50\% & 61.52 & 41.00 & TRUE \\ 
-  $M_C$00\%-$M_C$75\% & 95.52 & 41.00 & TRUE \\ 
-  $M_C$10\%-$M_C$25\% & 22.47 & 41.00 & FALSE \\ 
-  $M_C$10\%-$M_C$50\% & 45.49 & 41.00 & TRUE \\ 
-  $M_C$10\%-$M_C$75\% & 79.49 & 41.00 & TRUE \\ 
-  $M_C$25\%-$M_C$50\% & 23.02 & 41.00 & FALSE \\ 
-  $M_C$25\%-$M_C$75\% & 57.02 & 41.00 & TRUE \\ 
-  $M_C$50\%-$M_C$75\% & 34.00 & 41.00 & FALSE \\ 
-   \hline
-\end{tabular}
+  \input{Tables/Ba_RF-MC_results}
   \label{Ba_RF-MC_results}
@@
 \section{Discussion}
+
+% NC: Discussions usually begin with a paragraph that sums up the results.
+Then you can move on to discussing caveats, problems, extentions etc.
+You also need to show how your work fits in with the wider literature.
+So I think you need to re-write this completely with that in mind.
+Do the same as we did for the introduction - write a line for each paragraph.
+This reads like you jumped in without thinking clearly about the structure.
+It probably doesn't need to have all the same subheadings either.
 
@@
 The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
+% NC: Caveats later on
 This configuration probably lead to globally low phylogenetic signal as well as the intrinsic difficulties to simulated characters with phylogenetic signal.
@@
 independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy.
-%REF probably April as something about it
+%REF probably April has something about it
 \begin{enumerate}
@@
 
-\subsection{Comparing topologies}
+\subsection{Comparing topologies} % NC: Is any of this really relevant to the central questions of the paper?
+I don't think so...
+Maybe some of this could go in the methods where you chose these metrics.
+But it may not even be needed there.
 Comparing topologies is a crucial question in phylogenetics but has always been hard to normalise because of the vast amount of different metrics used as proxies for different aspects of tree similarity/dissimilarity \citep{agapowpower2002}.
@@
 the fossils) branch in the tree.
-The Robinson-Fould \citep{RF1981} and the Triples \citep{critchlowthe1996} metrics are more sensible to taxa and clade placement than other tree comparison metrics \citep[e.g.][Imbalance metric]{kirkpatricksearching1993} and where therefore favoured.
-Also by using the Normalised Tree Similarity index \citep{Bogdanowicz2012} we emphasize the aspect of "good" phylogenetic signal \textit{versus} random phylogenetic signal because it normalised the values of the metric by correcting for the expected value when comparing random trees (NTS = 0).
+The Robinson-Foulds \citep{RF1981} and the Triples \citep{critchlowthe1996} metrics are more sensible to taxa and clade placement than other tree comparison metrics \citep[e.g.][Imbalance metric]{kirkpatricksearching1993} and where therefore favored.
+Also by using the Normalized Tree Similarity index \citep{Bogdanowicz2012} we emphasize the aspect of "good" phylogenetic signal \textit{versus} random phylogenetic signal because it normalized the values of the metric by correcting for the expected value when comparing random trees (NTS = 0).
 
@@
 \section{Acknowledgements}
-We would like to thank Trevor Hodkinson and Andrew Jackson %Also Gavin Thomas, Emmanuel Douzery and Frédéric Delsuc if not authors.
-for his useful comments on the simulation protocol and Paddy Doyle for the assistance on using the computer cluster.
-All calculations were performed on the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing.
-This cluster was funded through grants from Science Foundation Ireland.
+% NC: Try and keep these short and to the point
+Thanks to Trevor Hodkinson and Andrew Jackson %Also Gavin Thomas, Emmanuel Douzery and Frédéric Delsuc if not authors.
+%NC: Also members of MacroJC - %TG : also Adam Kev and Sive or that's putting to many people, thinking about it, they helped at least as much as Andrew (not mentioning Trevor but that's more political)?
+for useful comments on our simulation protocol and Paddy Doyle for  assistance with the computer cluster.
+Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.
+% NC: Is this what you're supposed to cite for the cluster?
+Check.
+- TG: yep
+This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
 
@@
 
-\section{Supplementaries}
-\label{supplementaries}
-\subsection{Morphological characters states}
-In order to obtain a realistic probabilistic value for of \textit{k} characters states for each simulated morphological character, we downloaded 100 random morphological characters (with more than 100 characters each) from TreeBASE database (http://treebase.org/) published between 1985 and 2013 and covering 19 taxomomic classess (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
-We selected a total of 22563 characters ranging from 2 to 10 states.
-We calculated the proportion of characters with 2, 3, 4, 5, 6, 7, 8, 9 or 10 states.
-We then sampeled 22563 \textit{k} values between 2 and 10 with the same proportion of characters from the empirical data.
-We then used a simple t-test to check if our simulation was equal to the empirical data.
-In this study, we only simulated characters with 2 or 3 states because of the high proportion of ordered characters ecountered on characters with more than 3 states and the difficulties of simulate biologicaly sensible ordered characters.
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-AppendixCharacters.pdf}
-\caption{Character states distribution in empirical matrices. %Title?
-Characters states number distribution extracted from 100 random morphological matrices downloaded from RreeBase.}
-\label{Fig_AppendixCharacters}
-\end{figure}
-
-\subsection{Tree Building Software settings}
-
-\subsubsection{Maximum Likelihood - RAxML v8.0.20 \citep{Stamatakis21012014}} \\
-Model: \\
-Molecular data: \\
-GTR + $\Gamma_4$ (-m GTRGAMMA) \\
-Morphological data: \\
-Mk + $\Gamma_4$ (-K MK) \\
-Support: \\
-Rapid Boostrap algorithm (LSR), 1000 replicates \\
-
-\subsubsection{Bayesian - MrBayes v3.0.2 \citep{Ronquist2012mrbayes}} \\
-Priors: \\
-Molecular data: \\
-rates distribution shape ($\alpha$) = 0.5 \\
-Transition/Transversion ratio = 2 ($\beta$(80,40)) \\
-Starting tree: "True" tree topology with each branch length = 1 \\
-Morphological data: \\
-rates distribution shape ($\alpha$) = 0.5 \\
-Models: \\
-Molecular data: HKY + $\Gamma_4$ \\
-Morphological data: Mk + $\Gamma_4$ \\
-MCMC: \\
-2 runs \\
-4 chains per run \\
-generations < 50$\times$$1^6$ \\
-sample frequency = 1050$\times$$1^3$ \\
-ASDS diagnosis frequency = 50$\times$$1^3$ \\
-ASDS $<$ 0.01 \\
-ESS $>>$ 200 \\
-Burnin = 25\%
-
-\subsection{Triplets metric details ($T_{x,y}$)}
-Each triplet can be written as $I_{ijk}$=(\textit{ijk})). Where $I_{ijk}$ is equal to 0 if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to 1.
-For any rooted tree there are only four possible combinations per triplets: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); and (\textit{i},\textit{j},\textit{k}); \citep{johnson1998}.
-One can calculate $S_n$, the triplet distance between two trees as:
-\begin{equation}
-S_n = \sum_{ijk} I_{ijk}
-\end{equation}
-Where:
-\begin{equation}
-\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
-\end{equation}
-And where n is the number of taxa in both trees (modified from \citet{critchlowthe1996}).
-When all triplets across the two trees are the same, $S_n$ is equal to 0 and when all the triplets are different $S_n$ is equal to $\binom{n}{4}$.
-Because the possible number of triplets per clade is a finite number, the probability of two random trees with the same n taxa to have the same triplet is:
-\begin{equation}
-P({I_{ijk}}=0) = \frac{1}{4}
-\end{equation}
-Therefore one can calculate the probability of two random trees having the same triplets: 
-\begin{equation}
-P({S_{n}}=0) = \sum_{ijk} P_{I_{ijk}=0}
-\end{equation}
-\begin{equation}
-P({S_{n}}=0) = \frac{n!}{4(3!(n-3)!}
-\end{equation}
-And in the same way:
-\begin{equation}
-P({S_{n}}=1) = \frac{3n!}{4(3!(n-3)!}
-\end{equation}
-
-\subsection{RF metric details}
-The RF distance (or path difference) between two trees reflects the distance between the distributions of the tips among clades in the two trees \citep{RF1981} and can be expressed as following:
-\begin{equation}
-RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
-\end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees. 
-The minimal value of \textit{C} is equal to 1 if the two trees have the same n taxa;
-the maximal value in \textit{C}=\textit{n}-2.
-For a fully unresolved tree (star tree) \textit{N}=1 and for a fully resolved tree (binary tree) \textit{N}=\textit{n}-2.
-The minimal and maximal topological distance for \textit{n} taxa is:
-\begin{equation}
-RF_{min} = 1 + 1 - 2C_{x,y}
-\end{equation}
-And:
-\begin{equation}
-RF_{max} = 2(n-2)-2
-\end{equation}
-One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any \textit{n} taxa:
-\begin{equation}
-RF.scaled_{x,y} = \frac{RF_{x,y}-RF_{max}}{RF_{max}}
-\end{equation}
-This metric is more sensitive to taxa displacement than the Triplet distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.
-
-\subsection{Tree comparisons}
-\subsubsection{Random tree comparison scaling}
-We used the comparison of 1000 random trees to obtain the mean comparison value $\bar{d}_{m,n}$\textit{(rand)} for the NTS metric.
-We randomly generated two sets of 1000 trees of \textit{n} taxa using the rmtree function of ape package (v3.0-11 \citet{paradisape:2004}) that generates a given number of random Yule trees.
-We calculated the $\bar{d}_{m,n}$\textit{(rand)} value using an approach similar to the RPCBTC (described below) by performing 1000 random pairwise comparisons using the TreeCmp java script \citep{Bogdanowicz2012}.
-
-\subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
-We assessed the power of the Random Pairwise Bayesian Tree Comparison (RPBTC) method by comparing 1000 random trees from a posterior distribution trees set to another 1000 random trees from the same posterior distribution trees set.
-We repeated this 100 times independently using the same posterior distribution trees set each time resulting in 100 replicates of the same posterior distribution trees set compared 1000 times.
-We used an anova to test if there was no significant difference between the replicates so that the RBTC can be replicated.
-We applied this protocol on a poorly resolved tree (Low Score), a resolved tree with low support value (Medium Score) and a resolved tree with high support values (High Score).
-Results are available in table ~\ref{RPBTC_testing}. %link broken
-
-\begin{table}[ht]
-\caption{Group comparison results: difference between 100 replicates using the RPBTC method} %test with TreeCmp.anova
-\centering
-\begin{tabular}{rllrrrr}
-  \hline
-  Tree.Type & Used.metric & Replicates & Df & F.value & p.value \\ 
-  \hline
-  Low Score & RF & 100.00 & 99.00 & 0.74 & 0.98 \\ 
-  Low Score & Tr & 100.00 & 99.00 & 0.97 & 0.58 \\ 
-  Medium Score & RF & 100.00 & 99.00 & 0.64 & 1.00 \\ 
-  Medium Score & Tr & 100.00 & 99.00 & 0.45 & 1.00 \\ 
-  High Score & RF & 100.00 & 99.00 & 0.20 & 1.00 \\ 
-  High Score & Tr & 100.00 & 99.00 & 0.37 & 1.00 \\ 
-  \hline
-\end{tabular}
-\end{table}
-\label{RPBTC_testing}
-
-\subsection{Codes}
-All codes are available at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions$
-The tree comparison results analysis can be repeated for more details at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Analysis$ %fully running!
-Try it at home (5-10 minutes of data loading)!
-
-\subsection{Full results}
-
-\begin{table}
-\caption{Tree similarity values per parameter in ML framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & metric & mode & 50\%CI &       & 95\%CI &       \\
-              & missing data &        &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.6714878 & 0.5692308  & 0.7894649 & 0.4689069  & 1.0361514 \\ %how many digits?
-              &              & Triples        & 0.9424133 & 0.8683742  & 0.9998559 & 0.6121399  & 1.0635250 \\
-              & 25\%         & Robinson-Fould & 0.5167595 & 0.4403147  & 0.6307692 & 0.2664429  & 0.8929104 \\
-              &              & Triples        & 0.5599494 & 0.7084874  & 0.8105781 & 0.2912950  & 0.9990703 \\
-              & 50\%         & Robinson-Fould & 0.4321852 & 0.3641026  & 0.5215627 & 0.2313246  & 0.6423846 \\
-              &              & Triples        & 0.5143009 & 0.4197127  & 0.5991498 & 0.2991432  & 0.9104526 \\
-              & 75\%         & Robinson-Fould & 0.3727090 & 0.3230769  & 0.4342249 & 0.2035474  & 0.5776286 \\
-              &              & Triples        & 0.4449563 & 0.3385128  & 0.5329149 & 0.1929395  & 0.8551570 \\
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.9072419 & 0.7668415  & 1.0000000 & 0.4070844  & 1.0478010 \\
-              &              & Triples        & 0.9763617 & 0.9448100  & 1.0039460 & 0.7158118  & 1.1043330 \\
-              & 25\%         & Robinson-Fould & 0.5722382 & 0.4883560  & 0.6923077 & 0.3253239  & 0.9367467 \\
-              &              & Triples        & 0.9370152 & 0.8207407  & 0.9938846 & 0.5394443  & 1.0868073 \\
-              & 50\%         & Robinson-Fould & 0.4338159 & 0.3025641  & 0.4760777 & 0.1704016  & 0.6683220 \\
-              &              & Triples        & 0.7327464 & 0.5947211  & 0.8739630 & 0.3418679  & 1.0251540 \\
-              & 75\%         & Robinson-Fould & 0.1534487 & 0.1155128  & 0.2000000 & 0.0683485  & 0.3230769 \\
-              &              & Triples        & 0.4588292 & 0.3905631  & 0.6111484 & 0.2903676  & 0.8568274 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 1 & NA  & NA & NA  & NA \\
-              &              & Triples        & 1 & NA  & NA & NA  & NA \\
-              & 10\%         & Robinson-Fould & 0.473285 & 0.3842965  & 0.8358974 & 0.3564753  & 0.9589744 \\
-              &              & Triples        & 0.9592557 & 0.8691580  & 1.026095 & 0.5720847  & 1.067524 \\
-              & 25\%         & Robinson-Fould & 0.4548038 & 0.3572132  & 0.6307692 & 0.2425339  & 0.9161212 \\
-              &              & Triples        & 0.9071925 & 0.7516934  & 0.9976224 & 0.4462168  & 1.0705291 \\
-              & 50\%         & Robinson-Fould & 0.4244263 & 0.3755493  & 0.4666667 & 0.2494927  & 0.8287312 \\
-              &              & Triples        & 0.6834366 & 0.5713927  & 0.8226139 & 0.2776690  & 1.0085944 \\
-              & 75\%         & Robinson-Fould & 0.2306888 & 0.1774508  & 0.3025641 & 0.1087764  & 0.5673541 \\
-              &              & Triples        & 0.4513642 & 0.3737397  & 0.5723858 & 0.2245224  & 0.7917946 \\
-    \hline
-\label{NTSML_full}
-\end{tabular}
-\end{table}
-
-\begin{table}
-\caption{Tree similarity values per parameter in Bayesian framework}
-\centering
-\begin{tabular}{rrlccccc}
-    \hline
-    Parameter & amount of    & mode & 50\%CI &       & 95\%CI & \\
-              & missing data &      & lower  & upper & lower  & upper \\
-    \hline
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\ %how many digits?
-              &              & Triples        & 0.04104231 & 0.00830962  & 0.05441173 & -0.03933377  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.03584309 & 03584196  & 03589744 & 02825643  & 0.06175676 \\
-              &              & Triples        & 0.03399849 & 0.003196469  & 0.04981018 & -0.049913382  & 0.11784366 \\
-              & 25\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01536739  & 0.06123106 \\
-              &              & Triples        & 0.03510953 & 0.002113358  & 0.04939228 & -0.041202264  & 0.09699332 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01538462  & 0.05641026 \\
-              &              & Triples        & 0.02849433 & 0.003263098  & 0.05298036 & -0.038344731  & 0.09250063 \\
-              & 75\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01069992  & 0.04338811 \\
-              &              & Triples        & -0.0001940566 & -0.01801301  & 0.04058783 & -0.06030922  & 0.10665934 \\
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\
-              &              & Triples        & 0.04104231 & 0.008309622  & 0.05441173 & -0.039333778  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.03593712 & 0.03198256  & 0.05641026 & 0.02798864  & 0.07692308 \\
-              &              & Triples        & 0.02639309 & 0.002185408  & 0.05061314 & -0.046563042  & 0.10547052 \\
-              & 25\%         & Robinson-Fould & 0.03588134 & 0.03587629  & 0.03589744 & 0.01538462  & 0.0680884 \\
-              &              & Triples        & 0.03856296 & 0.002185408  & 0.05061314 & -0.046563042  & 0.10547052 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01536289  & 0.05879269 \\
-              &              & Triples        & 0.01403716 & -0.01100608  & 0.03972324 & -0.03800324  & 0.12405207 \\
-              & 75\%         & Robinson-Fould & 0.03589206 & 0.03588825  & 0.03589744 & 0.01538462  & 0.0411843 \\
-              &              & Triples        & 0.02939363 & 0.01902351  & 0.04238907 & -0.03087905  & 0.08946917 \\   
-    $M_L$     & 0\%          & Robinson-Fould & 0.05641026 & 0.03589744  & 0.06337695 & 0.03139209  & 0.07692308 \\
-              &              & Triples        & 0.04104231 & 0.008309622  & 0.05441173 & -0.039333778  & 0.10690054 \\
-              & 10\%         & Robinson-Fould & 0.05637057 & 0.03589744  & 0.05927956 & 0.02813296  & 0.07692308 \\
-              &              & Triples        & 0.03008132 & 0.01176800  & 0.04847400 & -0.02451461  & 0.09453754 \\
-              & 25\%         & Robinson-Fould & 0.03593743 & 0.03589744  & 0.03596018 & 0.02330790  & 0.07692308 \\
-              &              & Triples        & 0.02010221 & -0.001145118  & 0.0377779 & -0.060183332  & 0.1098632 \\
-              & 50\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.03050289  & 0.05950426 \\
-              &              & Triples        & 0.01053959 & -0.005595947  & 0.04510444 & -0.036278820  & 0.10478805 \\
-              & 75\%         & Robinson-Fould & 0.03589744 & 0.03589744  & 0.03589744 & 0.01115983  & 0.04362927 \\
-              &              & Triples        & 0.02427778 & 0.0007405737  & 0.04260522 & -0.0373004106  & 0.1048756 \\
-\label{NTSBA_full}
-\end{tabular}
-\end{table}
-
-\subsection{Bootstraps distribution}
-
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-AppendixBootstraps.pdf}
-\caption{Bootstraps distribution across the "best" trees.}                     % MISSING 8 CHAINS
-\label{Fig_AppendixBoostraps}
-\end{figure}
+\section{Supplementary Material}
+% NC: What do they call this in Syst Biol? Supplemental Materials is more usual...
+% NC: You probably want to give these numbers so you can refer people to supp mat section 1, section 2 etc. etc.
+    \input{SupplementaryMaterial/SupplementaryMaterial}
+    \label{SupplementaryMaterial}
 

### 2014-06-27 Manuscript/TEM_draft.tex
Updated draft (Intro+Methods) after NC comments

@@
   \linespread{1.0}
-  \input{Tables/Glossary}
+  \input{Tables/Glossary} %Remove sub-matrices
   \label{Tab_glossary}
@@
 Just say what they measure and shovel all the detail in supp?
-
 \subsection{Comparing topologies}
 We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
-We used the Triples \citep{dobson1975triplets} metric to assess the number of conserved taxa across trees, and the Robinson-Fould \citep{RF1981} distance to identify conserved clade positions.
+We used the Triples distance \citep{dobson1975triplets} to assess the number of conserved taxa across trees, and the Robinson-Foulds distance \citep{RF1981} to identify conserved clade positions.
 We then used Normalized Tree Similarity index \citep{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
@@
 But the general meaning is that the placements differ between trees.
-The triples distance ($T_{x,y}$; \citealp{dobson1975triplets}) measures the number of sub-trees made up of three taxa that differ between two given trees (\citealt{critchlowthe1996} ; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
+The Triples distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two given trees (\citealt{critchlowthe1996} ; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
 This metric measures the position of each taxon and clade towards it's closest neighbours.
@@
 \subsubsection{Robinson-Foulds distance} % NC: Is it Robinson Fould or Robinson Fould*s*? - TG: Definitely FouldS. R. Foulds is the guys name.
-Robinson-Foulds distance (\textit{RF}; \citealp{RF1981}), or "path difference"
+Robinson-Foulds distance \citep{RF1981}, or "path difference"
 % NC: path difference or *distance*? - TG: difference.
@@
 
-%TG: maybe draw a figure to illustrate that? Triples = not really stric, RF= really strict.
-
-\begin{equation}
-RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
-\end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees. \textit{C} % NC: Why italics? Why not $C$?
-is one if the two trees have the same \textit{n} taxa; the maximal value is \textit{C = n-2}. % NC: again why italics for these bits? 
-%NC: I don't get this. Explain more clearly. also is it important?
-This metric is more sensitive to taxon displacement than Triples distance (i.e.
-if one taxon moves out of a clade, then the clades are no longer considered similar; \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
-Therefore a low value indicates that clades are similar across the two trees, whereas a high value indicates that there are few common clades (see \hyperref[supplementaries]{supplementaries}).
-%NC: See supp for what purpose exactly?
+%TG: maybe draw a figure to illustrate that? Triples = not really strict, RF= really strict.
 
 \subsubsection{Normalized Tree Similarity}
-For any tree with \textit{n} taxa compared using a tree distance metric \textit{m}, Normalized Tree Similarity, $NTS_m$ (\citealp{Bogdanowicz2012}), represents the similarity score for the two trees given the expected distance between two random Yule trees with \textit{n} taxa.
-If $\bar{d}_{m,n}$\textit{(rand)} is the average distance between two random Yule trees with \textit{n} taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} containing each \textit{n} taxa, then:
-\begin{equation}
-NTS_{m,n}(x,y)=\frac{\bar{d}_{m,n}(rand) - d_{m,n}(x,y)} {\bar{d}_{m,n}(rand)}
-\end{equation}
-\textit{NTS} ranges from one to -$\infty$.
-For any $m,n$, when \textit{NTS} = 1, the trees are identical, when \textit{NTS} = 0 the trees are no more different than expected by chance, and when \textit{NTS} $<$ 0, the trees are more different than expected by chance %chance.\\
-% NC: You seem to be using italics almost at random here. Why???
-% NC: What is "chance" here?
-% NC: Why do we need to see the supp mat?
+We used the Normalized Tree Similarity index, $NTS_m$ (\citealp{Bogdanowicz2012}) to be able to compare the two metrics for any $n$ taxa.
+This index allows to scale the value of any metric $m$ (either Triples or Robinson-Foulds distance in our study) to the expected value of the metric $m$ when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
+When $NTS_m$=1, the two trees are strictly identical, when $NTS_m$=0 the trees are no more different than expected when comparing two random trees and when $NTS_m$$<$0, the difference between the two trees is greater than when comparing two random trees.
+In our study we used the $NTS_m$ index as a proxy for topological recovery: a high score of this index (i.e.
+towards 1) means that the topology is highly conserved between the two trees; on the opposite, a low score of this index (i.e.
+towards 0) means that the topological difference between the two trees is as much as expected when comparing two random trees.
 
 \subsubsection{Tree comparisons}
-For the Maximum Likelihood trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree (see Table ~\ref{Tab_glossary}) using both the Triples and Robinson-Fould metrics % NC: be consistent in the order.
+For the Maximum Likelihood trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree (see Table ~\ref{Tab_glossary}) using both the Triples and Robinson-Foulds metrics % NC: be consistent in the order.
 Triples then RF.
 with the TreeCmp java script \citep{Bogdanowicz2012}.
-For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and \textit{n} = 51 taxa (see \hyperref[supplementaries]{supplementaries}).
+For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and $n$ = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
 % NC: Why italic?
 What does the supp mat show?
+- TG: I tried to show how I generated the random trees and extracted the expected value for scaling from them.
 We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs
@@
 I think we need to check what you *actually* did here...
+- TG: That's for both Bayesian and ML, I explain how I summarize the Bayesian below.
+The hdrcde gives the Highest Density regions of the data.
 
@@
 Comparing Bayesian trees was slightly harder as each analysis produced a posterior distribution of trees to compare, rather than a single tree as in Maximum Likelihood.
-Therefore, for each "missing-data" posterior distribution of trees we randomly selected one tree from the "missing-data" tree posterior distribution and one tree from the "best" tree posterior distribution and compared these trees using Triples distance and Robinson-Fould's distance as above.
+Therefore, for each "missing-data" posterior distribution of trees we randomly selected one tree from the "missing-data" tree posterior distribution and one tree from the "best" tree posterior distribution and compared these trees using Triples distance and Robinson-Foulds distance as above.
 We repeated this 1000 times and then used the modal values for the metrics to summarize the difference between the two tree posterior distributions.
@@
 
-Finally we tested for significant differences among types of "missing-data" trees using ???
+%TG: I think I'm going to change all this session after Graham comments in Evolution.
+I probably want to compare the the 50 ML trees from RAxML to the 50 ML trees from mrBayes (in the posterior distribution) and then the 50*1000 bootstrap trees (by random pairwise tree comparison (RPTC)) from RAxML and 1000 RPTC from mrBayes.
+
+Finally we tested for significant differences among types of "missing-data" trees using non-parametric group and pairwise comparisons by using respectively Kruskal-Wallis and Nemenyi-Damico-Wolfe-Dunn test (CITE R PACKAGE)\citet{ruxtontime2008}.
 % NC: The whole of the next section can be summarised in just a few lines like this!
@@
 \item
-A sub-matrix containing both molecular and morphological data for living taxa only (hereafter called the "horizontal" sub-matrix, Table ~\ref{Tab_glossary});
+A sub-matrix containing both molecular and morphological data for living taxa only;
 %NC: Do you think the sub-matrix names are useful here?
@@
 Better something like - complete living, complete morphology, morphology living?
+%TG: agreed.
 \item
-A sub-matrix containing morphological data for both living and fossil taxa (hereafter called the "vertical" sub-matrix, Table ~\ref{Tab_glossary});
+A sub-matrix containing morphological data for both living and fossil taxa;
 \item
-A sub-matrix containing morphological data for living taxa only (hereafter called the "corner" sub-matrix, Table ~\ref{Tab_glossary});
+A sub-matrix containing morphological data for living taxa only ;
 \end{enumerate}
@@
 %NC: Should we check ML trees too???
+%TG: Not sure if it is necessary since this part (not fully developed yet) is yet another test to show that our results are significant and that the problem comes form the missing data in the fossils (or not???)
 and Bayesian tree inference and topology comparisons on the three sub-matrices in the same way as described above.
@@
 Finally, to % NC: explain WHY we would want to do this
+%TG: need to better explain an RUN this part, I'll rewrite this part properly ones the runs are over.
 
@@
 %NC: OK from this point on you start referring to "Tree Similarity Index". This is not mentioned in the methods. What is it? You need to explain in the methods please.
+%TG: stopped here 26/06.
 

### 2014-06-27 Functions/TreeComparisons/TEM_TreeCmp.sh
v0.7 - Updated Bayesian/ML -> Treesets/Singletrees

@@ -1,5 +1,5 @@
 ##########################
-#Uses TreeCmp java script to compare the distributions of Bayesian trees
+#Uses TreeCmp java script to compare either single trees or treesets
 ##########################
 #SYNTAX: TEM_TreeCmp <reference.treeset> <input.treeset> <number of species> <number of random draws> <output> <method> <trees>
 #with:
@@ -8,22 +8,26 @@
 #<number of species> number of species per trees
 #<number of random draws> number of random draws for the comparison
 #<output> a chain name for the output
-#<method> must be either Bayesian or ML
+#<method> must be either Treesets or Singletrees
 #<trees> must be either nexus or newick
+#<rooted> must be either TRUE or FALSE
 ##########################
-#version: 0.6
+#version: 0.7
 #Update: improved output format
 #Update: correction on the random tree selection from the ref tree: if the ref tree is unique, no sample is performed any more.
 #Update: cleaning improved (remove the files from the concern chain only)
 #Update: temporary results are now stored in a temporary folder
-#Update: added a Bayesian/ML method input. If the method is ML, there is no random tree drawing.
+#Update: added a Treesets/Singletrees method input. If the method is Singletrees, there is no random tree drawing.
 #Update: trees can now be nexus or newick
+#Update: changed method arguments from Bayesian/ML to Treesets/Singletrees
+#Update: improved documentation within the function
+#Update: added a rooted/unrooted option
 #----
-#guillert(at)tcd.ie - 29/04/2014
+#guillert(at)tcd.ie - 27/06/2014
 ##########################
 #Requirements:
 #-R 3.x
-#-TreeCmp java script
+#-TreeCmp java script (Bogdanowicz et al 2012)
 #-http://www.la-press.com/treecmp-comparison-of-trees-in-polynomial-time-article-a3300
 #-TreeCmp folder to be installed at the level of the analysis
 ##########################
@@ -39,12 +43,14 @@ draws=$4
 output=$5
 method=$6
 trees=$7
+rooted=$8
 
 #Creates the settings file
-echo $method > TreeCmp_settings.tmp
[…]

### 2014-06-27 Manuscript/TEM_draft.tex
Updated tables results (8 missing Chains)

@@
 
-%10/06 - Missing 8 ML Chains
 \begin{figure} 
@@
   \linespread{1.0}
-  \caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
+  \caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} 
   \input{Tables/ML_RF-ML_results}
@@
   \linespread{1.0}
-  \caption{$M_L$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
+  \caption{$M_L$ non-parametric pairwise difference for Triples metric in ML framework} 
   \input{Tables/ML_Tr-ML_results}
@@
   \linespread{1.0}
-  \caption{$M_F$ non-parametric pairwise difference for Robinson-Foulds metric in ML framework} %10/06 - Missing 8 ML Chains
+  \caption{$M_F$ non-parametric pairwise difference for Robinson-Foulds metric in ML framework} 
   \input{Tables/ML_RF-MF_results}
@@
   \linespread{1.0}
-  \caption{$M_F$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
+  \caption{$M_F$ non-parametric pairwise difference for Triples metric in ML framework} 
   \input{Tables/ML_Tr-MF_results}
@@
   \linespread{1.0}
-  \caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} %10/06 - Missing 8 ML Chains
+  \caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} 
   \input{Tables/ML_RF-MC_results}
@@
   \linespread{1.0}
-  \caption{$M_C$ non-parametric pairwise difference for Triples metric in ML framework} %10/06 - Missing 8 ML Chains
+  \caption{$M_C$ non-parametric pairwise difference for Triples metric in ML framework} 
   \input{Tables/ML_Tr-MC_results}

### 2014-06-27 Analysis/MissingDataEffect.R
Updated analysis (paper plot)

@@ -23,7 +23,7 @@ Random25<-read.table("../Data/Tree_Comparisons/Random_trees/25t_Random.Cmp",head
 
 #ML chains
 setwd("../Data/Tree_Comparisons/TreeCmp_chains/ML/")
-tmp<-TreeCmp.Read('MChain', verbose=FALSE)
+tmp<-TreeCmp.Read('MChain', verbose=TRUE)
 ML_trees<-NTS(tmp, Random51)
 setwd('../../../../Analysis/')
 
@@ -136,6 +136,29 @@ TreeCmp.Plot( Bayesian.MLMFMC , metric, probs=c(95, 50), col=palette()[1], lines
 TreeCmp.Plot( ML.MLMFMC , metric, probs=c(95, 50), col=palette()[2], lines=TRUE, add=TRUE, shift=0.1, ylim=ylim, save.details=FALSE)
 legend( 100 , 1 , c("Bayesian", "ML"), pch=21, col=c(palette()[1:2]), pt.bg=c(palette()[1:2]), cex=0.8)
 legend( 100 , 0.9 , c("Mode", "50% CI", "95% CI"), col="gray80", lty=c(NA, 1, 2), pch=c(21, NA, NA), pt.bg=c("gray80"), cex=0.8)
+
+
+#Publication style plot
+par(mfrow=c(2,3))
+palette(c("black", "gray50"))
+ylim=c(0,1)
+metric="Triples"
+TreeCmp.Plot( Bayesian.ML , metric, probs=c(95, 50), col=palette()[1], lines=TRUE, ylim=ylim)
+TreeCmp.Plot( ML.ML, metric, probs=c(95, 50), col=palette()[2], lines=TRUE, add=TRUE, shift=0.1, ylim=ylim, save.details=FALSE)
+legend( 0.5 , 0.8 , c("Bayesian", "ML"), pch=21, col=c(palette()[1:2]), pt.bg=c(palette()[1:2]), cex=0.8)
+legend( 0.5 , 0.7 , c("Mode", "50% CI", "95% CI"), col="gray80", lty=c(NA, 1, 2), pch=c(21, NA, NA), pt.bg=c("gray80"), cex=0.8)
+TreeCmp.Plot( Bayesian.MF , metric, probs=c(95, 50), col=palette()[1], lines=TRUE, ylim=ylim)
[…]

### 2014-07-24 Functions/BS.read.R
v0.2 Added requirements

@@ -2,7 +2,8 @@
 #Bootstrap reading
 ##########################
 #Extracts the Bootstrap values of a list of trees
-#v.0.1
+#v.0.2
+#Update: added require(ape)
 #TO DO: pattern should be either a list/single tree out/in of R environment
 ##########################
 #SYNTAX :
@@ -19,6 +20,9 @@
 ##########################
 
 BS.read<-function(pattern, plot=TRUE, use.tree.names=TRUE){
+
+#REQUIREMENTS
+    require(ape)
 
 #DATA INPUT
     #pattern

### 2014-07-24 Manuscript/TEM_draft.tex
Added pre-BS selection in Methods

@@
 For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} which is a generalization of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citealt{Stamatakis21012014}).
+
+In order to measure the phylogenetic signal of our simulations, we first ran a fast bootstrap analysis with 500 replicates on the "complete" matrix.
+We removed all the simulations that had a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
+We repeated this selection until we obtained 50 sets of simulations (i.e. 50 "complete" and 50*124 "missing-data" matrices) that with a relative good phylogenetic signal (median bootstrap $>$ 50).
+
 We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. %The explanations of the algorithm part for the BS was proposed by Trevor

### 2014-07-29 Manuscript/TEM_draft.tex
Acknowledgements notes

@@
 % NC: Try and keep these short and to the point
-Thanks to Trevor Hodkinson and Andrew Jackson %Also Gavin Thomas, Emmanuel Douzery and Frédéric Delsuc if not authors.
-%NC: Also members of MacroJC - %TG : also Adam Kev and Sive or that's putting to many people, thinking about it, they helped at least as much as Andrew (not mentioning Trevor but that's more political)?
-for useful comments on our simulation protocol and Paddy Doyle for  assistance with the computer cluster.
+Thanks to Frédéric Delsuc, Emmanuel Douzery, Trevor Hodkinson and Andrew Jackson and the members of the Macro Journal Club %Also Gavin Thomas if not authors - %TG : also Adam Kev and Sive or that's putting to many people, thinking about it, they helped at least as much as Andrew (not mentioning Trevor but that's more political)?
+for useful comments on our simulation protocol.
+Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
 Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.

### 2014-11-25 Analysis/MissingData_plot.R
Updated plots

@@ -30,13 +30,53 @@ if(quick.load == FALSE) {
 #Plot main figures (individual parameters in RF and Triples)
 #####################
 #windows parameters
-#quartz(width = 8.3, height = 5.8) #A5 landscape
-quartz(width = 16.6, height = 11.6) #A4 landscape
+quartz(width = 8.3, height = 5.8) #A5 landscape
+#quartz(width = 16.6, height = 11.6) #A4 landscape
 op<-par(mfrow=c(2,3), bty="l", oma=c(0,0,0,0))
 ylim=c(0,1)
 
 #data parameters
-data.sets<-c("ML_besttrees","Bayesian_contrees")#,"ML_bootstraps","Bayesian_treesets")
+data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
+metrics<-c("R.F_Cluster","Triples")
+
+#2*3 plots
+#Missing living RF
+par(mar=c(1,5,5,0) + 0.1)
+multi.TreeCmp.plot( data.sets , par_ML, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab="Robinson-Fould distance", las=2)
+#Missing fossil RF
+par(mar=c(1,3,5,2) + 0.1)
+multi.TreeCmp.plot( data.sets , par_MF, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
+#Missing character RF
+par(mar=c(1,1,5,4) + 0.1)
+multi.TreeCmp.plot( data.sets , par_MC, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
+
+xaxis.labs<-c("0%", "10%", "25%", "50%", "75%")
+#Missing living Triplets
+par(mar=c(5,5,1,0) + 0.1)
+multi.TreeCmp.plot( data.sets , par_ML, metrics[2], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=xaxis.labs, ylab="Triplets distance", las=2, xlab="Missing living taxa") 
+#Missing fossil Triplets
+par(mar=c(5,3,1,2) + 0.1)
[…]

### 2014-11-25 Functions/TreeComparisons/TreeCmp.Plot.R

@@ -412,12 +412,13 @@ multi.TreeCmp.plot<-function(data.list, parameter, metrics, col, shift='auto', .
 ##########################
 #Plot the significance levels using the data input in the plot
 ##########################
-#v.0.1
+#v.0.2
+#Update: fixed metric input
 ##########################
 #SYNTAX :
 #<data.list> a character list of the different datasets to plot
 #<parameters> a numerical list of parameters to select per dataset
-#<metrics> the list of metrics
+#<metric> the metric of interest
 #<col> the colour palette matching the list, can be default (a star) or a list of symbols or values
 #<pch> symbol for the significant comparison
 #<position> either 'above' or 'below' whether to plot the symbol above or below the data
@@ -425,15 +426,16 @@ multi.TreeCmp.plot<-function(data.list, parameter, metrics, col, shift='auto', .
 #<diff.matrix> the matrix containing the significant differences
 ##########################
 #----
-#guillert(at)tcd.ie - 23/11/2014
+#guillert(at)tcd.ie - 25/11/2014
 ##########################
 
-signi.TreeCmp.plot<-function(data.list, parameter, metrics, probs, diff.type, position, col, pch='default', shift='auto') {
+signi.TreeCmp.plot<-function(data.list, parameter, metric, probs, diff.type, position, col, pch='default', shift='auto') {
 
     debug=FALSE
     if(debug==TRUE) {
         data.list<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
         metrics<-c("R.F_Cluster","Triples")
+        metrics<-"R.F_Cluster"
         data<-1
         metric<-1
         parameter<-par_ML
@@ -509,7 +511,7 @@ signi.TreeCmp.plot<-function(data.list, parameter, metrics, probs, diff.type, po
             new.pch<-pch[1]
 
             #Calculating the hdr results
-            hdr.results<-FUN.hdr(sub.data(get(data.list[data]), parameter), grep(metrics[metric], colnames(sub.data(get(data.list[data]), parameter)[[1]])), probs)
+            hdr.results<-FUN.hdr(sub.data(get(data.list[data]), parameter), grep(metric, colnames(sub.data(get(data.list[data]), parameter)[[1]])), probs)
[…]

### 2014-11-25 Analysis/MissingData_plot.R
Revert "Updated plots"
This reverts commit 93dadd9310f6aa564a49afce04fa7b77d911aca2.

@@ -30,53 +30,13 @@ if(quick.load == FALSE) {
 #Plot main figures (individual parameters in RF and Triples)
 #####################
 #windows parameters
-quartz(width = 8.3, height = 5.8) #A5 landscape
-#quartz(width = 16.6, height = 11.6) #A4 landscape
+#quartz(width = 8.3, height = 5.8) #A5 landscape
+quartz(width = 16.6, height = 11.6) #A4 landscape
 op<-par(mfrow=c(2,3), bty="l", oma=c(0,0,0,0))
 ylim=c(0,1)
 
 #data parameters
-data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
-metrics<-c("R.F_Cluster","Triples")
-
-#2*3 plots
-#Missing living RF
-par(mar=c(1,5,5,0) + 0.1)
-multi.TreeCmp.plot( data.sets , par_ML, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab="Robinson-Fould distance", las=2)
-#Missing fossil RF
-par(mar=c(1,3,5,2) + 0.1)
-multi.TreeCmp.plot( data.sets , par_MF, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
-#Missing character RF
-par(mar=c(1,1,5,4) + 0.1)
-multi.TreeCmp.plot( data.sets , par_MC, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
-
-xaxis.labs<-c("0%", "10%", "25%", "50%", "75%")
-#Missing living Triplets
-par(mar=c(5,5,1,0) + 0.1)
-multi.TreeCmp.plot( data.sets , par_ML, metrics[2], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=xaxis.labs, ylab="Triplets distance", las=2, xlab="Missing living taxa") 
-#Missing fossil Triplets
-par(mar=c(5,3,1,2) + 0.1)
[…]

### 2014-11-25 Functions/TreeComparisons/TreeCmp.Plot.R

@@ -412,13 +412,12 @@ multi.TreeCmp.plot<-function(data.list, parameter, metrics, col, shift='auto', .
 ##########################
 #Plot the significance levels using the data input in the plot
 ##########################
-#v.0.2
-#Update: fixed metric input
+#v.0.1
 ##########################
 #SYNTAX :
 #<data.list> a character list of the different datasets to plot
 #<parameters> a numerical list of parameters to select per dataset
-#<metric> the metric of interest
+#<metrics> the list of metrics
 #<col> the colour palette matching the list, can be default (a star) or a list of symbols or values
 #<pch> symbol for the significant comparison
 #<position> either 'above' or 'below' whether to plot the symbol above or below the data
@@ -426,16 +425,15 @@ multi.TreeCmp.plot<-function(data.list, parameter, metrics, col, shift='auto', .
 #<diff.matrix> the matrix containing the significant differences
 ##########################
 #----
-#guillert(at)tcd.ie - 25/11/2014
+#guillert(at)tcd.ie - 23/11/2014
 ##########################
 
-signi.TreeCmp.plot<-function(data.list, parameter, metric, probs, diff.type, position, col, pch='default', shift='auto') {
+signi.TreeCmp.plot<-function(data.list, parameter, metrics, probs, diff.type, position, col, pch='default', shift='auto') {
 
     debug=FALSE
     if(debug==TRUE) {
         data.list<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
         metrics<-c("R.F_Cluster","Triples")
-        metrics<-"R.F_Cluster"
         data<-1
         metric<-1
         parameter<-par_ML
@@ -511,7 +509,7 @@ signi.TreeCmp.plot<-function(data.list, parameter, metric, probs, diff.type, pos
             new.pch<-pch[1]
 
             #Calculating the hdr results
-            hdr.results<-FUN.hdr(sub.data(get(data.list[data]), parameter), grep(metric, colnames(sub.data(get(data.list[data]), parameter)[[1]])), probs)
+            hdr.results<-FUN.hdr(sub.data(get(data.list[data]), parameter), grep(metrics[metric], colnames(sub.data(get(data.list[data]), parameter)[[1]])), probs)
[…]

### 2014-11-25 Analysis/MissingData_plot.R
Figures plotting code

@@ -30,59 +30,157 @@ if(quick.load == FALSE) {
 #Plot main figures (individual parameters in RF and Triples)
 #####################
 #windows parameters
-#quartz(width = 8.3, height = 5.8) #A5 landscape
-quartz(width = 16.6, height = 11.6) #A4 landscape
+quartz(width = 8.3, height = 5.8) #A5 landscape
+#quartz(width = 16.6, height = 11.6) #A4 landscape
 op<-par(mfrow=c(2,3), bty="l", oma=c(0,0,0,0))
 ylim=c(0,1)
 
 #data parameters
-data.sets<-c("ML_besttrees","Bayesian_contrees")#,"ML_bootstraps","Bayesian_treesets")
+data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
 metrics<-c("R.F_Cluster","Triples")
 
 #2*3 plots
 #Missing living RF
 par(mar=c(1,5,5,0) + 0.1)
 multi.TreeCmp.plot( data.sets , par_ML, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab="Robinson-Fould distance", las=2)
+#Missing fossil RF
+par(mar=c(1,3,5,2) + 0.1)
+multi.TreeCmp.plot( data.sets , par_MF, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
+#Missing character RF
+par(mar=c(1,1,5,4) + 0.1)
+multi.TreeCmp.plot( data.sets , par_MC, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
+
+xaxis.labs<-c("0%", "10%", "25%", "50%", "75%")
+#Missing living Triplets
+par(mar=c(5,5,1,0) + 0.1)
+multi.TreeCmp.plot( data.sets , par_ML, metrics[2], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=xaxis.labs, ylab="Triplets distance", las=2, xlab="Missing living taxa") 
+#Missing fossil Triplets
+par(mar=c(5,3,1,2) + 0.1)
[…]

### 2014-11-25 Functions/TreeComparisons/TreeCmp.Plot.R

@@ -412,12 +412,13 @@ multi.TreeCmp.plot<-function(data.list, parameter, metrics, col, shift='auto', .
 ##########################
 #Plot the significance levels using the data input in the plot
 ##########################
-#v.0.1
+#v.0.2
+#Update: fixed metric input
 ##########################
 #SYNTAX :
 #<data.list> a character list of the different datasets to plot
 #<parameters> a numerical list of parameters to select per dataset
-#<metrics> the list of metrics
+#<metric> the metric of interest
 #<col> the colour palette matching the list, can be default (a star) or a list of symbols or values
 #<pch> symbol for the significant comparison
 #<position> either 'above' or 'below' whether to plot the symbol above or below the data
@@ -425,15 +426,16 @@ multi.TreeCmp.plot<-function(data.list, parameter, metrics, col, shift='auto', .
 #<diff.matrix> the matrix containing the significant differences
 ##########################
 #----
-#guillert(at)tcd.ie - 23/11/2014
+#guillert(at)tcd.ie - 25/11/2014
 ##########################
 
-signi.TreeCmp.plot<-function(data.list, parameter, metrics, probs, diff.type, position, col, pch='default', shift='auto') {
+signi.TreeCmp.plot<-function(data.list, parameter, metric, probs, diff.type, position, col, pch='default', shift='auto') {
 
     debug=FALSE
     if(debug==TRUE) {
         data.list<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
         metrics<-c("R.F_Cluster","Triples")
+        metrics<-"R.F_Cluster"
         data<-1
         metric<-1
         parameter<-par_ML
@@ -509,7 +511,7 @@ signi.TreeCmp.plot<-function(data.list, parameter, metrics, probs, diff.type, po
             new.pch<-pch[1]
 
             #Calculating the hdr results
-            hdr.results<-FUN.hdr(sub.data(get(data.list[data]), parameter), grep(metrics[metric], colnames(sub.data(get(data.list[data]), parameter)[[1]])), probs)
+            hdr.results<-FUN.hdr(sub.data(get(data.list[data]), parameter), grep(metric, colnames(sub.data(get(data.list[data]), parameter)[[1]])), probs)            
[…]

### 2014-11-25 Manuscript/TEM_draft.tex
Added bootstrap number  cit
Removed old figures and old abstract

@@
 
-%packages
-% NC: Just to flag here how I'll comment on LaTeX files. I'll use % and then NC. Of course in git diff it'll all show up as an addition.
-% NC: You could add all the packages etc to a template .sty file to save writing it all up here. Rich's ms.sty that I provided on DropBox is a good example
-
-% TG: Just a general note for the file outline: I write every sentence of a same paragraph on a new line, no real reason for that appart that I can then see when the sentence is way to long.
-Also, I kept the big hidden headers ("%----") that you don't seem to like, without them it totally wrecks my head to know how much do I still have to scroll...
-
+%Packages
 \usepackage{fixltx2e}
@@
 \usepackage{fullpage}
-%\usepackage{nag}
 \usepackage{natbib}
-%\usepackage{bibtex}
 \usepackage{float}
 \usepackage{latexsym}
-%\usepackage{hyperref} 
+\usepackage{hyperref}
 \usepackage{url}
@@
 \usepackage{hyperref}
-%\usepackage{xcolor}
 \usepackage{amsthm}
-\usepackage{amstext} % Add and remove packages as necessary for your manuscript.
-
-% NC: Added by NC
+\usepackage{amstext}
 \usepackage{enumerate}
-\usepackage[osf]{mathpazo} % palatino is a nicer font! 
-\pagenumbering{arabic} % Page numbers
+\usepackage[osf]{mathpazo}
+\pagenumbering{arabic}
 
 %Pagination style and stuff
-\linespread{1.66} % All text should be double-spaced with occasional exceptions for tables. 
+\linespread{1.66}
 \raggedright
 \setlength{\parindent}{0.5in}
-
-\setcounter{secnumdepth}{0} % Our sections are not numbered and our papers do not have Tables of Contents.
-We don't present a list of figures or list of tables, either.
-Any common font is fine.
-(A common sans-serif font should be used on figures, but figures should be separate from the LaTeX document.)
-
-%\pagestyle{empty}
-
+\setcounter{secnumdepth}{0} 
 \renewcommand{\section}[1]{%
@@
 \end{center}}
-
 \renewcommand{\subsection}[1]{%
@@
 \end{center}}
-
 \renewcommand{\subsubsection}[1]{%
@@
 \textit{#1.}---}
-
 \renewcommand{\tableofcontents}{}
-
-\bibpunct{(}{)}{;}{a}{}{,}  % this is a citation format command for natbib
+\bibpunct{(}{)}{;}{a}{}{,}
 
@@
 \begin{document}
-
-%For a colour version of the draft to a search>replace '-BW'>'' 
 %\modulolinenumbers[1]     % Line numbering on every line
@@
 \bigskip
-\noindent RH: Missing data and topology in Total Evidence matrices %RH = Running Head
+\noindent RH: Missing data and topology in total evidence approach
 
@@
 
-%Title
-\noindent{\Large \bf Effect of missing data in matrices containing living and fossil taxa on topological accuracy}
-%Alternative punchier title: "Bayesian Total Evidence methods fail to recover correct topologies from matrices containing both living and fossil taxa"
-%Alternative punchier title: "Maximum Likelihood outperforms Bayesian methods for inferring topology from Total Evidence matrices containing both living and fossil taxa"
-% NC: Not sure about any of these.
-I actually think I prefer your talk title.
-So something like: The effects of missing data on the ability of Total Evidence approaches to recover correct tree topologies.
-No rush to decide.
-But needs Total Evidence in there somewhere.
-
+\noindent{\Large \bf Effect of missing data on topological inference using a total evidence approach}
 \bigskip
@@
 %Authors
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$, other authors $^3$, and Natalie Cooper$^1$$^,$$^2$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$, and Natalie Cooper$^1$$^,$$^2$}\\
 \noindent {\small \it 
 $^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland;\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland;\\
-$^3$Somewhere else }\\
+$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland;}\\
 \end{center}
@@
 \begin{abstract}
-% NC: I'm not going to give this much work as it'll change before we submit anyway. Note though that and abstract must NOT just repeat stuff in the intro. Needs rephrasing at the very least.
-Living species represent less than 1\% of all species that have ever lived.
-% NC: I was chatting to Andy Purvis about this the other day and he said lots of palaeo folk hate that stat.
-So let's not start the abstract with it!
+Living species represent a marginal part of all species that have ever lived.
 Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology.
@@
 In this study we assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
-Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix:
-(1) the proportion of living taxa with no morphological data,
-(2) the amount of missing data in the fossil record, and
-(3) the overall number of morphological characters in the matrix.
-We find that, in a Bayesian framework, difficulties in recovering a stable topology are mainly driven by the missing data in the molecular part of the matrix (for which fossil taxa have no data).
-In a Maximum Likelihood framework, however, topology is not directly affected by missing data \textit{per se}, but by the number of morphological characters shared among the taxa.
-Therefore, the two main drivers of incorrect topologies are the overall number of morphological characters and the number of living species with no morphological data.
+Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix: (1) the proportion of living taxa with no morphological data, (2) the amount of missing data in the fossil record, and (3) the overall number of morphological characters in the matrix.
 
-Our results suggest that, in order to use Total Evidence approaches, one should reduce the missing data in the morphological part of the matrix for living species and use a Maximum Likelihood framework to fix the topology prior to the overall Bayesian phylogenetic inference process.
-%More results!
+We find that, when using a clade conservative metric such as Robinson-Foulds distance, Bayesian method recovers the right topology better than Maximum Likelihood method.
+However, when using triplets method, there is less significant difference between both methods.
+
+-Overall missing data is doesn't affect topological recovery in Bayesian (around minimum 70\% of topological topology in the worth scenario).
+
+-However, one essential way to improve it would be to code missing living taxa.
 \end{abstract}
@@
 
-Although most species that have ever lived are now extinct % NC: This avoids giving the dodgy stat of 1%
-    \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}.
-%or is citing of your own papers in the too "douchebagy" for the second sentence?
-% NC: Yeah be careful with self citing. Fine if it's a good fit, but it's totally not here!!! A better citation would be a big biodiversity study, e.g. the bird phylogeny paper.
+Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}.
 Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
@@
 
-Up to now, three main approaches have been used to place both living and fossil taxa % NC: BE consistent. Either living and fossil, or fossil and living. - TG: Let's make it living and fossil then.
-    into phylogenies.
+Up to now, three main approaches have been used to place both living and fossil taxa into phylogenies.
 These approaches differ mainly in whether they treat fossil taxa as tips or as nodes in the phylogeny, and in which part of the available fossil data is used (i.e.
 the age of the fossil only or both its age and morphology).
-% NC: Not sure about bit in brackets - TG: it's the idea of fossil occurence (age) only when node and fossil occurence + topology when tip.
 Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
@@
 This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty (but see \citealp{spencerefficacy2013}).
-Neontologists, on the other hand, more commonly use probabilistic approaches (e.g. Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence date % NC: Need to check the exact phrase needed here.
-Is it range of occurrence?
-Occurrence dates?
-etc.
-- TG: in paleo jargon we have FAD (first apparition datum) and FOD (first occurrence datum).
-FAD is always hypothetical (we can't really exactly know when the taxa appeared) and FOD is always subjective (it depends when the fossil was found, maybe next year someone will find another one in older layers).
-So for the dating, because it's relaxed clock, people use FOD (and also probably because they don't care since it doesn't make a huge difference I guess).
-    are used to time calibrate phylogenies \citep{zuckerkandl1965}.
+Neontologists, on the other hand, more commonly use probabilistic approaches (e.g.
+Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
+Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence date are used to time calibrate phylogenies \citep{zuckerkandl1965}.
 There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013}.
@@
 A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \citep{eernissetaxonomic1993}.
-%There might be a philosophico-semantic distinction to do here.
-Is TEM a method or an approach?
-I'll say that it is an approach because it's more like an idea (put every thing together) than a way to do (how to put everything together).
-It makes more sense to me compared to a real method e.g.
-a probabilistic method such as ML where the method states precisely how to do stuff.
-% NC: Tricky. Don't think it's too important though. 
-% NC: I do think Total Evidence needs capital letters though, like ML and Bayesian - TG: fixed
 This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
@@
 However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data that can be difficult (or impossible) to collect.
-%In particular it requires morphological data from both living and fossil taxa, both of which are known to be scarce.
-% NC: Really?
-I know fossil data is known to be scarce, but living data?
-I think people would be surprised.
-And they are surprised.
-We were surprised!
-Needs a citation either way.
-%TG: how about this:
 The morphological data for living taxa is rarely collected when molecular data is available (e.g.
 \citealt{O'Leary08022013} vs.
-\citealt{meredithimpacts2011}),% TG: i.e.
-in O'Leary they use meredith molecular data but only code few species from it.
- and, for fossil taxa, the scarcity of the fossil record only allow to collect the data available (for example, in vertebrates, the hardest parts of the skeleton; \citealt{sansomfossilization2013}).
+\citealt{meredithimpacts2011}), and, for fossil taxa, the scarcity of the fossil record only allow to collect the data available (for example, in vertebrates, the hardest parts of the skeleton; \citealt{sansomfossilization2013}).
 Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
-% Trevor: Missing data also influences support values that trend to decrease
 
-% NC: You need to take a bit of time to learn about singular and plural grammar in English.
-It's something you always do wrong.
-e.g.
-The Total Evidence method *uses* this and *its* parameters are as follows.
-Versus: Total Evidence methods *use* this and *their* parameters are as follows.
-Sive and I won't be around to correct your English forever.
-- TG: my French demons are back!
-
-The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013}.
+The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014}.
 Missing molecular data has been seen by some authors as an issue because it can, in some part of the tree, decrease phylogenetic signal (i.e.
 the evolutionary information contained within the matrix allowing to infer topology and branch length)% NC: define phylogenetic signal here.
-    , especially when using large matrices \citep{lemmonthe2009}.
-However, this may not be a major issue because phylogenetic signal is easily increased by:
+    , especially when using large matrices \citep{lemmonthe2009}. However, this may not be a major issue because phylogenetic signal is easily increased by:
 (i) including a "modest" number of highly-covered genes (i.e.
 approximatively of the genes that are coded for most of the species; \citealt{rouresite-specific2011}) %NC: what do you mean by highly-covered?
-(ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and
-(iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
-Similarly, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study in question \citep{wiensmissing2003,sansomfossilization2013}.
-Because soft-tissue characters are rarely preserved in the fossil record, missing data is mainly found in these characters, and is therefore not randomly distributed which can lead to biased placement of fossil taxa in phylogenies \citep{sansomfossilization2013}.
+(ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and (iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
+Similarly, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study in question \citep{wiensmissing2003,sansomfossilization2013,pattinsonphylogeny2014}.
+Because soft-tissue characters are rarely preserved in the fossil record, missing data is mainly found in these characters, and is therefore not randomly distributed which can lead to biased placement of fossil taxa in phylogenies (e.g.
+\citeauthor{sansomfossilization2013} \citeyear{sansomfossilization2013} but see \citeauthor{pattinsonphylogeny2014} \citeyear{pattinsonphylogeny2014}).
 However, the phylogenetic signal is not related to the amount of missing data \textit{per se} but to the number of informative characters for each taxon, therefore missing data is less of an issue than the number of shared informative characters \citep{wiensmissing2003}.
@@
 Until now, no attempt has been made to study the impact of this issue on phylogenetic inference from Total Evidence methods.
+
+%ADD APRIL'S PAPER SOMEWHERE
 
@@
 Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them.
-% NC: OK I'm not sure what the "real-life" reason for three is. - TG: fixed to "empirical" (or "practical" ?).
-% TG: This paragraph is crap
-% NC: I think it's fine!
 
@@
 % NC: You don't want to tell the whole story here. Just a results summary and a quick line of interpretation
-
-%This problem leads to an inability of Bayesian chains to pick up the sub and near optimal topologies leading to high variance in the posterior tree distribution and really low support in the consensus tree.
-%These results should encourage authors in not using Bayesian inference to explore the topological landscape of their phylogeny.
-%Instead, we suggest that authors should first find the Maximum Likelihood tree and use it's topology as a fix prior in their Bayesian inference.
-%We also propose a new method for comparing Bayesian posterior tree distribution's topology available in a wonderful R package. %Or am I just dreaming a bit too much here? 
-% NC: Dreaming a bit much I think.
-There's a discussion to be had here about whether it's really a *new* method and whether you really want to make an R package per se.
-We definitely want to share the code
-
-%-----------
-% NC: It might look like I've done a lot to this but actually I think it's nearly there.
-A good thing to take from this is that writing a paper is *hard*.
-It gets easier, but it's still a deeply involved process with lots of writing and re-writing.
-Introductions are particularly difficult because if you don't set the scene, it's hard to convince people to get on board for the rest of the paper.
-This is why I said there was no way this was being submitted by the end of June!
-Still a long way to go I'm afraid in terms of time and effort.
-But that doesn't mean you're not doing well, *every* project is like this.
-2 years is about average from conception to sending out a manuscript, and you only started work on this last spring.
-%------------
 
@@
 
-\section{Methods} %no material needed?
-% NC: Depends on the journal.
-See what Syst Biol asks for as section titles.
-TG: Syst Biol instructions :Elements vary, of course, depending on the type of paper.
-The first-level heading is often used for sections such as ``methods and materials,'' ``results,'' and' ``discussion;'' but other major sections are often appropriate instead.
+\section{Methods}
 To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
-%NC: does Syst Biol use Fig.
-or Figure?
-- TG: it's "Fig." when it's a reference in brackets and "Figure" when it's a reference in the text.
-Table is always refered as "Table" (not "Tab").
 \begin{enumerate}
 \item{Generating the matrix} \label{step:generate_matrix} \\
-We randomly generated a birth-death tree (hereafter called the "true" tree; Table ~\ref{Tab_glossary}) and used it to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix; Table ~\ref{Tab_glossary}).
+We randomly generated a birth-death tree (hereafter called the "true" tree) and used it to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix).
 \item{Removing data} \label{step:remove_data} \\
-We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$) % NC: make sure to use the same descriptions here and in the introduction
-(the resulting matrices are called hereafter "missing-data" matrices; Table ~\ref{Tab_glossary}).
-\item{Building phylogenies} \label{step:build_phylo} \\
-We built phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree; Table ~\ref{Tab_glossary}) and multiple trees inferred from matrices with missing morphological data (hereafter called the "missing-data" trees; Table ~\ref{Tab_glossary}).
+We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$) (the resulting matrices are called hereafter "missing-data" matrices).
+\item{Inferring phylogenies} \label{step:build_phylo} \\
+We inffered phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree) and multiple trees inferred from matrices with missing morphological data (hereafter called the "missing-data" trees).
 Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
@@
 \end{enumerate}
-We repeated steps 1 to 4 50 times. %that's because I had 3 cluster accounts available... If it's a too odd number, no worries, I'll just remove the last chain.
-% NC: Yeah 51 is odd. Let's just use 50.
-A list of all the terms used in this paper is available in Table ~\ref{Tab_glossary}.
-
-% NC: Not sure we need the glossary, but keep it for the thesis
-
-\begin{table}[ht]
-  \caption{Glossary}
-  \centering
-  \linespread{1.0}
-  \input{Tables/Glossary} %Remove sub-matrices
-  \label{Tab_glossary}
-\end{table}
+We repeated steps 1 to 4 50 times.
 
@@
 \centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_outline-BW.pdf}
+\includegraphics[keepaspectratio=true]{Figures/BW/TEM_Fig_outline-BW.pdf}
 \caption{Protocol outline.
 (1) We randomly generated a birth-death tree (the "true" tree) and used it to infer a matrix with no missing data (the "complete" matrix).
-(2) We removed data from the morphological part of the "complete" matrix resulting in 124 "missing-data" matrices.
-% NC: I thought we had 124 missing data matrices...
-I just worked this out - one of the "missing data" matrices is 0%,0%,0% so = complete matrix.
-- TG: Yep, there is 125 matrices/trees per chain.
-1 that's L00F00C00 and 124 others.
-It's again one of these habits like "adding missing data" that I have from the coding (that actually do add missing data and do generate 125 matrices).
+(2) We removed data from the morphological part of the "complete" matrix resulting in 125 "missing-data" matrices.
 (3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
@@
 \end{figure}
-
-% NC: One tiny edit to the figure.
-I think the ML then Bayesian story is nice, so maybe change that part of the figure so it's ML then Bayesian.
-Also let's show this figure to a couple of people to see if they get it.
--TG: Do you mean just swap the text ML/Bayesian (was: Bayesian-ML, is now: ML-Bayesian)?
 
@@
 We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
-We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, total computational time for the whole analysis was over 80 CPU years).
+We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, total computational time for the whole analysis was over 150 CPU years).
 All the molecular information for fossil taxa was replaced by missing data ("?").
@@
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used low evolutionary rate parameters (i.e. $\alpha$) to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
+We used low evolutionary rate parameters (i.e.
+$\alpha$) to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
 
@@
 \subsection{Removing data}
-We modified the "complete" matrix to get matrices with missing data by randomly replacing data with "?" in the morphological part of the matrices according to the following parameters (Fig.
-~\ref{Fig_RemoveData}):
+We modified the "complete" matrix to get matrices with missing data by randomly replacing data with "?" in the morphological part of the matrices according to the following parameters:% (Fig.
+~\ref{Fig_RemoveData}): % PUT IN APPENDIX
 
@@
 
-% NC: I'm not sure this figure is helpful really. Let's test it on some people.
-\begin{figure}
-\centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_missingData-BW.pdf}
-\caption{Names of the different parts of the matrix.
-A: Missing data parameters:
-Missing living - The proportion of living taxa with no morphological data ($M_{L}$);
-Missing fossil - The proportion of missing morphological data across all fossil taxa ($M_{F}$);
-Missing character - The proportion of missing morphological characters across all taxa (living and fossil) ($M_{C}$).
-B: Different parts of the matrix:
-The "horizontal" sub-matrix (orange) contains molecular and morphological data for living taxa;
-The "vertical" sub-matrix (blue) contains morphological data for living and fossil taxa;
-The "corner" sub-matrix (orange and blue striped) contains morphological data for living taxa;
-The "missing-data" sub-matrix (grey) is the molecular part of the fossil taxa and contains no data.}
-\label{Fig_RemoveData}
-\end{figure}
-
-In practice, each parameter represents a different way of removing data from the matrix: $M_{L}$ removes rows from the living taxa; $M_{F}$ removes cells from the fossil taxa; and $M_{C}$ removes columns across both living and fossil taxa (see Fig.
-~\ref{Fig_RemoveData}).
+In practice, each parameter represents a different way of removing data from the matrix: $M_{L}$ removes rows from the living taxa; $M_{F}$ removes cells from the fossil taxa; and $M_{C}$ removes columns across both living and fossil taxa.
 Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a percentage of living taxa is removed, but for $M_{F}$, a percentage of the data is removed at random from across the whole of the morphological matrix for fossil taxa.
@@
 We tested all parameters combinations resulting in 125 ($5^3$) matrices.
-Note that one of these combinations has no missing data so is equivalent to the "complete" matrix, thus we have 124 "missing-data" matrices.
-Because some parameter combinations introduce a lot of missing data (e.g. $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
+Note that one of these combinations has no missing data so is equivalent to the "complete" matrix, thus we have one effectively complete matrix in our 125 "missing-data" matrices.
+Because some parameter combinations introduce a lot of missing data (e.g.
+$M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
 When this occurred we repeated the random deletion of characters until every taxa had at least 5\% data across the whole morphological part of the matrix.
@@
 \subsection{Building phylogenies}
-From the resulting matrices we generated two types of trees, the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 124 matrices with various amounts of missing data.
+From the resulting matrices we generated two types of trees, the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
 The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
 The "best" tree, on the other hand, is the best tree we can build using state-of-the-art phylogenetic methods.
-In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g. \citealt{rozen2005}).
+In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g.
+\citealt{rozen2005}).
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
@@
 For the molecular data, we used the GTR + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citealt{Stamatakis21012014}) as a generalization of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
-The GTR model can be seen as a generalization of the HKY model (the two parameters from the HKY model are implicitly included in the six from GTR model - \citealt{stamatakisa2008}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} which is a generalization of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citealt{Stamatakis21012014}).
+%The GTR model can be seen as a generalization of the HKY model (the two parameters from the HKY model are implicitly included in the six from GTR model - \citealt{stamatakisa2008}).
+- TG: Is that useful?
+For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} %which is a generalization of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 - TG: same?
+ assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citealt{Stamatakis21012014}).
 
@@
 We removed all the simulations that had a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
-We repeated this selection until we obtained 50 sets of simulations (i.e. 50 "complete" and 50*124 "missing-data" matrices) that with a relative good phylogenetic signal (median bootstrap $>$ 50).
+We repeated this selection until we obtained 50 sets of simulations (i.e.
+50 "complete" and 50*125 "missing-data" matrices) with a relative good phylogenetic signal (median bootstrap $>$ 50).
 
-We used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support. %The explanations of the algorithm part for the BS was proposed by Trevor
-The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists of pruning one sub-tree from the tree and subsequently reinserting it to all neighbouring branches \citep{stamatakisa2008}.
+On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support \citep{pattengale2010many}.
+%The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists of pruning one sub-tree from the tree and subsequently reinserting it to all neighbouring branches \citep{stamatakisa2008}.
 Sub-tree Pruning and Reinserting methods (SPR) have been demonstrated to be better than other methods (e.g.
 Nearest Neighbouring Interchange - NNI) for recovering good bootstrap values \citep{salamin2003}.
-When using these parameters, it took 13 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (8 core nodes 2.30GHz clock speed).
+%TG: The explanations of the algorithm part for the BS was proposed by Trevor I think it's useless
+When using these parameters, it took 6 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (8 core nodes 2.30GHz clock speed).
 
 \subsubsection{Bayesian}
-The "best" tree and the "missing-data" trees were inferred using MrBayes v3.2.2 \citep{Ronquist2012mrbayes}.
+The "best" tree and the "missing-data" trees were inferred using MrBayes v3.2.1 \citep{Ronquist2012mrbayes}.
 We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
@@
 These priors biased the way the Bayesian process calculated branch lengths by giving non-random starting points and boundaries for parameter estimation, however, here we are focusing on the effect of missing data on tree topology and not branch lengths.
-Even using these priors, it took 65 CPU years to build 50 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
-% NC: OK so the 65 CPU is for building trees not matrices?
-Or is it combined time?
-Or does both stage take 65 CPU years?
-Be clearer in the earlier sections.
-- TG: I fixed that above: 65 years for the Bayesian, 13 years for the ML, 1 year for all the rest (generating, comparing, etc...)
+Even using these priors, it took 140 CPU years to build 50 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
 
@@
 
-% NC: I think this section is suffering a bit from "here is everything I know about tree comparison metrics" rather than explaining what is important for your analysis.
-You may be able to cut some more stuff out.
-- TG: since I'm just using Bogdanowicz allready published code, I maybe don't need to detail the metrics at all?
-Just say what they measure and shovel all the detail in supp?
 \subsection{Comparing topologies}
 We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
-We used the Triples distance \citep{dobson1975triplets} to assess the number of conserved taxa across trees, and the Robinson-Foulds distance \citep{RF1981} to identify conserved clade positions.
+We used the Robinson-Foulds distance \citep{RF1981} to identify conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the number of conserved taxa across trees.
 We then used Normalized Tree Similarity index \citep{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
 These metrics are described in detail below.
-% NC: I moved this around a bit to make it clearer what metric fits where. Remember we need American spellings.
-% NC: Also I thought it was triplets distance?
-Good idea to check...
-- TG: It's pretty confusing, unfortunately I can't check Dobson 1975 because he was just cited in Critchlow 1996 and Johnson 1998 as the author of the method but I never got my hand on it.
-I don't remember what's the name in Johnson 1998 (Trevor's book) but it might well be 'triplets' since I started from there.
-However, it is clearly 'Triples' distance in Critchlow 1996 and in Bodganowicz 2012 (that uses QuarteT but TripleS and ref back to Critchlow 1996).
-Let's go for Triples?
-In the end if the reviewers wants us to change it, it should be a minor issue.
+%ADD KHUNAR PAPER SOMEWHERE
 
-\subsubsection{Triples distance}
-% NC: I've tried to remove the word displaced here, as displaced suggests that you have one correct placement and then are comparing it to other trees.
-But the general meaning is that the placements differ between trees.
-The Triples distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two given trees (\citealt{critchlowthe1996} ; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
+\subsubsection{Robinson-Foulds distance}
+Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees.
+The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981} ; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
+This metric is bounded between 1 when the two trees are identical and $n-2$ (for two trees with $n$ taxa) when there is not one single shared clade between both trees.
+This metric is sensitive to the exact clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swap of two taxa will lead to a maximal score of the Robinson-Foulds distance indicating a bad tree similarity.
+
+\subsubsection{Triplets distance}
+The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two given trees (\citealt{critchlowthe1996} ; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
 This metric measures the position of each taxon and clade towards it's closest neighbours.
@@
 
-\subsubsection{Robinson-Foulds distance} % NC: Is it Robinson Fould or Robinson Fould*s*? - TG: Definitely FouldS. R. Foulds is the guys name.
-Robinson-Foulds distance \citep{RF1981}, or "path difference"
-% NC: path difference or *distance*? - TG: difference.
-    , measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981} ; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
-This metric is bounded between 1 when the two trees are identical and $n-2$ (for two trees with $n$ taxa) when there is not one single shared clade between both trees.
-Contrary to the Triples distance, this metric is more sensitive to the exact clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swap of two taxa will lead to a maximal score of the Robinson-Fould distance indicating a bad tree similarity.
-
-%TG: maybe draw a figure to illustrate that? Triples = not really strict, RF= really strict.
-
 \subsubsection{Normalized Tree Similarity}
 We used the Normalized Tree Similarity index, $NTS_m$ (\citealp{Bogdanowicz2012}) to be able to compare the two metrics for any $n$ taxa.
-This index allows to scale the value of any metric $m$ (either Triples or Robinson-Foulds distance in our study) to the expected value of the metric $m$ when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
+This index allows to scale the value of any metric $m$ (either Robinson-Foulds or Triplets distance in our study) to the expected value of the metric $m$ when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
 When $NTS_m$=1, the two trees are strictly identical, when $NTS_m$=0 the trees are no more different than expected when comparing two random trees and when $NTS_m$$<$0, the difference between the two trees is greater than when comparing two random trees.
@@
 \subsubsection{Tree comparisons}
-For the Maximum Likelihood trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree (see Table ~\ref{Tab_glossary}) using both the Triples and Robinson-Foulds metrics % NC: be consistent in the order.
-Triples then RF.
-with the TreeCmp java script \citep{Bogdanowicz2012}.
-For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and $n$ = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
-% NC: Why italic?
-What does the supp mat show?
-- TG: I tried to show how I generated the random trees and extracted the expected value for scaling from them.
-We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs
-% NC: Chains are a MCMC thing. You mean 50 simulations not chains.
-resulting in 50 comparisons for each "missing-data" tree.
+For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Robinson-Foulds and Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012}.
+For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and n = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
+% TG: Maybe remove supplementary here
+We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree.
 We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}.
-%NC: But this is a Bayesian thing right?
-I think we need to check what you *actually* did here...
-- TG: That's for both Bayesian and ML, I explain how I summarize the Bayesian below.
-The hdrcde gives the Highest Density regions of the data.
 
-%Bayesian tree inference allows to account for the statistical uncertainty of a phylogenetic tree by not using an optimal criterion (c.f.
-Maximum Likelihood) and gives a tree posterior distribution (c.f.
-the likeliest tree in ML).
-This method has the clear advantage of better dealing with error and uncertainty but it's output (a distribution rather than a single tree) is less practical to use for any further study (but see \citet{healyecology2014}).
-%Awwwwww yiis!
+Also, to take into account the uncertainty of tree inference in both Maximum Likelihood and Bayesian (i.e node support), we ran 1000 random pairwise comparison between respectively the bootstrapped trees from the Maximum Likelihood analysis and the posterior tree distribution of the Bayesian analysis.
+In the same way that we compared a single "missing-data" tree to the "best" tree (whether the trees are Maximum Likelihood or Bayesian consensus): we randomly selected 1000 trees from the "missing-data" tree sets (either the Bootstrapped trees or the posterior tree distribution) and did a pairwise comparison with 1000 randomly selected trees from the "best" tree set.
 
-% NC: Sorry awwww no. This is all irrelevant. Everyone knows what a Bayesian phylogeny is, and all the reader needs to know is how you compared trees. They don't need all this waffle.
-
-%To avoid this problem, people traditionally use the a consensus tree build on a majority rule in order to summarise a Bayesian tree posterior distribution.
-%This results into a single tree containing both topological and branch length information as well as support information (i.e. posterior probabilities, e.g. \citet{Ronquist2012mrbayes}).
-%However, using a Bayesian consensus tree has limitation especially if the resulting consensus tree is not well supported or resolved.
-%Because the metrics used in this study are used to measure variation in topology between two trees (i.e.
-taxa placement or clade position), comparing two Bayesian consensus trees is not optimal in picking topological difference signal.
-%For example, if one of the trees is not resolved at all (i.e. a star tree), comparing it to any other tree (even an other star tree) will not give useful results.
-%Therefore we used the entire Bayesian trees posterior distribution in order to perform the tree comparisons (i.e.
-the "Best" Bayesian tree posterior distribution vs.
-the "missing-data" Bayesian tree posterior distribution).
-
-% NC: OK this whole section is far too long.
-I'm not sure how much is really needed.
-Surely all you did was randomly select pairs from two distributions and then take the mode?
-Check that what I've written below is correct.
-
-Comparing Bayesian trees was slightly harder as each analysis produced a posterior distribution of trees to compare, rather than a single tree as in Maximum Likelihood.
-Therefore, for each "missing-data" posterior distribution of trees we randomly selected one tree from the "missing-data" tree posterior distribution and one tree from the "best" tree posterior distribution and compared these trees using Triples distance and Robinson-Foulds distance as above.
-We repeated this 1000 times and then used the modal values for the metrics to summarize the difference between the two tree posterior distributions.
-For each metric, we then normalized the value using the Normalized Tree Similarity as described above.
-We used this protocol to compare each "missing-data" tree posterior distribution with the "best" tree posterior distribution for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree.
+For each of the 125 "missing-data" tree, we obtained Robinson-Foulds and Triplets distance distributions from either 50 or 50*1000 pairwise comparisons.
 We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}.
-%NC: But this is a Bayesian thing right?
-I think we need to check what you *actually* did here...
-
-%TG: I think I'm going to change all this session after Graham comments in Evolution.
-I probably want to compare the the 50 ML trees from RAxML to the 50 ML trees from mrBayes (in the posterior distribution) and then the 50*1000 bootstrap trees (by random pairwise tree comparison (RPTC)) from RAxML and 1000 RPTC from mrBayes.
-
-Finally we tested for significant differences among types of "missing-data" trees using non-parametric group and pairwise comparisons by using respectively Kruskal-Wallis and Nemenyi-Damico-Wolfe-Dunn test (CITE R PACKAGE)\citet{ruxtontime2008}.
-% NC: The whole of the next section can be summarised in just a few lines like this!
-
-%\begin{equation}
-%RPBTC_{X,Y}=Mo[(]d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]
-%\end{equation}
-
-%Where \textit{X} and \textit{Y} are two Bayesian tree posterior distributions;
- %$d_{m,n}$ is the pairwise difference for any metric \textit{m} and \textit{n} taxa between $X_{i}$ and $Y_{j}$ which are two single trees randomly sampled respectively from \textit{X} and \textit{Y};
- %$Mo$ is the mode and
- %$Mo[(d_{m,n}(X_{i1},Y_{j1}),d_{m,n}(X_{i2},Y_{j2}),d_{m,n}(X_{i3},Y_{j3}), ... ,d_{m,n}(X_{ik},Y_{jk})]$ is the mode of the pairwise difference repeated \textit{k} times.
-
-%We used the mode to summarize the distributions because it is the value that is the most represented in the distribution and reflects what a consensus tree represents (the topology the most represented in the posterior distribution).
-Also, the mode is a value present in the distribution (which, depending on the metric, can be a series of discrete values) in contrast to the other distribution summary metrics, such as the mean or the median, which can take values that are not actually present in the distribution.
-For example, if comparing two posterior tree distribution, we obtain difference values of 10, 10, 10, 5, 5 and 1, using the mean (6.8) or the median (7.5) gives a values that actually doesn't exist.
-%Useful?
-
-%In this study, we calculated RPBTC for both metrics (Robinson-Fould and Triples distance) for each pairs of Bayesian tree distributions (i.e.
-the "Best" Bayesian tree posterior distribution vs.
-the "missing-data" Bayesian tree posterior distribution) with 1000 random pairwise comparisons (\textit{k}).
-
-%Because the RPBTC uses random pairwise comparisons, there is a chance of comparing always trees that have the biggest or the smallest difference between two distributions, either deflating or inflating the RPBTC difference value.
-However, using 1000 random pairwise comparisons makes the difference stable (i.e.
-if repeated independently, no difference is detected - see \hyperref[supplementaries]{supplementaries}).
-
-%We calculated the NTS for the "Best" Bayesian tree vs.
-"missing-data" Bayesian tree comparison for both Robinson-Fould and Triples metric using the mode of the RPBTC for each chain resulting in a distribution of 51 modes per comparison.
-We then compared the NTS of the Bayesian trees to the NTS of the ML trees for both metrics for each comparison.This resulted in a distribution of 51 NTS values for the Bayesian trees and 51 NTS values for the ML trees for each tree comparison for which we calculated the mode and the 50\% and 95\% confidence intervals (see \hyperref[supplementaries]{supplementaries} - Fig.
-~\ref{Fig_Compare}).
-
-% NC: I don't think this figure helps at all
-%\begin{figure} % Modify the figure
-%\centering
-%\includegraphics[keepaspectratio=true]{Figures/TEM_Fig_TreeCmp-BW.pdf}
-%\caption{Tree comparison protocol.
-%From each of the 125 parameters combination ($M_L$, $M_F$ and $M_C$), we inferred a tree in Baysian or Maximum likelihood framework.
-We then compared this tree with or without fossil/living taxa to the "best" tree.In the Bayesian framework, 1000 trees from the posterior distribution where used for each comparison.}
-%\label{Fig_Compare}
-%\end{figure}
-
-%\subsubsection{Statistical difference between tree topologies} %note that in the end, I only have non-parametric tests, the parametric explanation part can go to the supplementaries
-%To assess the difference between the tree topologies, whether they were inferred in a Bayesian or a ML framework, we performed a group comparison test for each set of configurations (e.g.
-effect of $M_L$ in Bayesian inference, combined effect of $M_F$ and $M_C$ in ML inference, etc...).
-We used the pairwise tree comparisons as factors (i.e.
-the "Best" tree vs.
-one of each of the 125 "missing-data" trees) and the values of the 51 replicates from the pairwise comparison as a response variable.
-The 51 replicates values were either the replicates of the 51 pairwise tree comparisons in ML or the modes of the 51 RPBTC posterior distributions in Bayesian.
-When we found a significant difference between the groups, we performed a pairwise comparison test to analyse which pairs of groups differ from the others.
-Depending on whether the data was normal with variance homoscedasticity or not (controlled by performing respectively a shapiro.test and a a bartlett.test in R\{stats\}), we used respectively parametric or non-parametric tests for the group comparison and the pairwise comparisons as suggested by \citet{ruxtontime2008} (Table ~\ref{Stats_test}).
-%link broken
-
-%\begin{table}[ht]
-%\caption{Statistical tests}
-%\centering
-%\begin{tabular}{lrll}
-%  \hline
-%   & test & statistics & code\{package\} \\ [0.5ex]
-%  \hline
-%  Groups comparisons & ANOVA & parametric & anova\{stats\} \\ [0.5ex]%
-%                     & Kruskal-Wallis & non-parametric & kruskal.test\{%stats\} \\[0.5ex]
-%  \hline
- % Pairwise comparisons & Tukey HSD & parametric & TukeyHSD\{stats\} \\ [0.5ex]
-  %                     & Nemenyi-Damico-Wolfe-Dunn & non-parametric & kruskalmc\{pgirmess\} \\ [0.5ex]
- % \hline
-%\end{tabular}
-%\label{Stats_test}
-%\end{table}
-
-\subsection{Effect of missing molecular characters for fossil taxa} %Awful name.
-% NC: Yeah!
-% NC: OK you need to explain WHY we are doing this. What is this analysis designed to test. I think you said this was incomplete so I won't fix it I'll let you have a go first.
-To assess the ability to recover topology for each simulation when no missing data was involved in the phylogenetic inference process, we split the "complete" matrix into three sub-matrices containing no missing data (see Fig.
-~\ref{Fig_RemoveData}):
-\begin{enumerate}
-\item
-A sub-matrix containing both molecular and morphological data for living taxa only;
-%NC: Do you think the sub-matrix names are useful here?
-We won't use this as much as the "best" tree etc.
-So maybe just write it out each time?
-Also I don't like the names horizontal, vertical and corner - they don't really describe what is going on.
-Better something like - complete living, complete morphology, morphology living?
-%TG: agreed.
-\item
-A sub-matrix containing morphological data for both living and fossil taxa;
-\item
-A sub-matrix containing morphological data for living taxa only ;
-\end{enumerate}
-We reran the Maximum Likelihood
-%NC: Should we check ML trees too???
-%TG: Not sure if it is necessary since this part (not fully developed yet) is yet another test to show that our results are significant and that the problem comes form the missing data in the fossils (or not???)
-and Bayesian tree inference and topology comparisons on the three sub-matrices in the same way as described above.
-
-%---------------------------------------------
-%       Empirical data - TO DO
-%---------------------------------------------
-
-\subsection{Empirical data} %Still not done!
-Finally, to % NC: explain WHY we would want to do this
-%TG: need to better explain an RUN this part, I'll rewrite this part properly ones the runs are over.
-
-We also compared the results obtained from simulated data to results from empirical data in \citet{ronquista2012}.
-The matrix (this would be the "complete" matrix using our terminology) contains 67 living species plus one outgroup and 45 fossil species of Hymenoptera with 5097 molecular characters and 354 morphological characters.
-Because only 66 of the 68 living species in the matrix had molecular data, we treated these 66 taxa as "living" taxa and the remaining 47 as "fossil" taxa.
-
-We removed data from the "complete" matrix in the same way as described in steps 2 and 3 above, resulting in 124 matrices with various amount of missing data.
-We used the same settings as described above to infer Maximum Likelihood trees, but for the Bayesian inferences we did not use any priors except the starting tree.
-This consisted of the topology of the 68 living species with the highest posterior probability from non-clock analysis performed in \citet{ronquista2012}.
-Note that we did not perform any of the clock analyses described in \citet{ronquista2012} because we were only interested in the topology of the inferred trees and not the branch lengths.
-
-% NC: Is there a step missing here? What did you do (or what will you do) with these trees? Will you compare them to each other? How will you compare them to the simulations?
+Finally we tested for significant differences among the different distributions of trees using non-parametric group and pairwise comparisons by using respectively Kruskal-Wallis and Nemenyi-Damico-Wolfe-Dunn test (CITE R PACKAGE)\citet{ruxtontime2008}.
 
@@
 
-% NC: OK it's worth thinking about what you really need to include here.
-Remember the results are about the answers to your question.
-Not necessarily everything you did.
-This section can be really short, and probably should be quite short.
-
-% NC: I may have misunderstood but I don't think you need this section at all, or the comparing topologies section you have planned. Just the actual results surely...
-
-%\subsection{Building the trees}
-%When generating the matrices using seqgen \citep{ranbaut1997seqgen} and rTraitDisc \citep{paradisape:2004} algorithms with low evolutionary rate parameter ($\alpha$ = 0.5) we successfully generated evolutionary histories.
-%In Maximum Likelihood, support values ranged from 0 to 100 with a median of 38 ($1^st$ quartile = 14, $3^rd$ quartile = 72 - see \hyperref[supplementaries]{supplementaries}). % This is actually bad.
-%The distribution of the phylogenetic signal across the different chains provided us a good spread of evolutionary scenarios: from well resolved phylogenies to poorly resolved ones.
-%This allows us to test our simulations in a theoretical as well as practical framework (i.e.
-phylogenies with low support values are rarely published but are still common in early stages of projects).
-%REF???
-- That's kind of my way to turn these bad results into something usable: even if some of our simulations are not realistic, it shows the effect of missing data in any scenario possible, not only in realistic cases.
-%Also, one can advocate that this low topological resolution is actively due to the amount of missing data in the total evidence matrices since the sub-matrices (no missing data) analysis recovers way higher support value  .
-%This is something that I will had with the whole submatrices story.
-Basically, when you have all the fossil/molecular data missing, it's harder to build the tree, both in ML and in Bayesian.
-When you run the analysis without missing data, then it works way better (i.e.
-our REAL control test: can we simulate matrices in the first place).
-%Using Bayesian inference, all the chains converged with a significant effective sample size by using 2 parallel runs of 4 chains each (ASSD $<$ 0.01 and ESS $>>$ 200 for inference).
-
-%\subsection{Comparing topologies}
-%Did it worked? See sub-matrices part. TO DO
-
-%\subsection{Effect of the missing data parameters}
-% NC: Probably don't need a subheading here
-
 \begin{figure} 
 \centering
-\includegraphics[keepaspectratio=true]{Figures/TEM_Fig-results-BW.pdf}
-\caption{Effect of missing data on Tree similarity.
-$M_L$ is the percentage of living taxa with no morphological data,
-$M_F$ is the percentage of missing data in the fossil record,
-$M_C$ is the percentage of missing morphological characters.
-The first row represent the Normalized Tree Similarity index using the Robinson-Foulds metric;
-the second row represent the Normalized Tree Similarity index using the Triples metric.}
+    \includegraphics[width=1\textwidth]{Figures/colour/Parameters-AllMethods-RF+Tr-small.pdf}
+\caption{Effect of missing data on topological recovery.
+The amount of missing data for each parameter is represented on the x axis.
+The topological recovery is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
+Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
+The Maximum Likelihood trees are represented in black, the Bayesian consensus trees in red, the bootstrap trees in green and the posterior tree distribution in blue.}
 \label{Fig_Results}
@@
 
-%NC: OK from this point on you start referring to "Tree Similarity Index". This is not mentioned in the methods. What is it? You need to explain in the methods please.
-%TG: stopped here 26/06.
+\subsection{Effect of the method on topological recovery}
+-Bayesian significantly better than ML for clade conservation.
+-Bayesian consensus trees plateaus around 0.8 topological recovery at worst whatever combination of missing data.
 
-%NC: I'm not sure you need to separate these three parameters out.
-Aren't you repeating yourself a lot?
-Rewrite this with them all in together.
-You can talk about general patterns and then specific, interesting patterns.
-I've corrected the first one a bit to help out.
+\subsection{Effect of the missing data on topological recovery}
+
 \subsection{Effect of $M_L$}
-In our Maximum Likelihood analyses, topological recovery (i.e.
-tree similarity) decreased as the percentage of missing data increased, using both the Triples and Robinson-Foulds metrics (see Table ~\ref{Group_results} - Fig.~\ref{Fig_Results}).
+In our Maximum Likelihood analyses, topological recovery (i.e. tree similarity) decreased as the percentage of missing data increased, using both the Robinson-Foulds and Triplets metrics.
 However, topological recovery was not significantly different for trees with 25\%, 50\% and 75\% missing morphological data for living taxa (Fig 1b???), % is this what you meant?
@@
 suggesting that although tree topology is sensitive to this missing data, once it is greater than 25\% it makes little difference.
-Overall, the Robinson-Foulds metric shows lower tree similarities than the Triples metric.
-This indicates that the positions of individual taxa are generally more conserved than the membership of entire clades (see Table ~\ref{NTS_ML_results}).
+Overall, the Robinson-Foulds metric shows lower tree similarities than the Triplets metric.
+This indicates that the positions of individual taxa are generally more conserved than the membership of entire clades.
 
 Topological recovery was poor in our Bayesian analyses, regardless of the amount of missing data (Figures???), although topological recovery did decrease slightly as missing data increased.
-Overall, tree similarity index was only slightly above zero which means that trees are only slightly more similar than we would expect if we compared two completely random trees (see Table ~\ref{NTS_Ba_results}).
-%For the Triples metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
-%For the Robinson-Fould metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases (see Table ~\ref{Ba_RF-ML_results}).
+Overall, tree similarity index was only slightly above zero which means that trees are only slightly more similar than we would expect if we compared two completely random trees.
+%For the Triplets metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
+%For the Robinson-Foulds metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases (see Table ~\ref{Ba_RF-ML_results}).
 
@@
 \subsubsection{Effect of $M_F$}
-In Maximum Likelihood framework, the effect of missing data in the fossil record appeared to be constant and led to a constant decrease in topological recovery when the missing data increased (see Table ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
+In Maximum Likelihood framework, the effect of missing data in the fossil record appeared to be constant and led to a constant decrease in topological recovery when the missing data increase.
 The difference between the two metrics used is more important than for the $M_L$ parameter: clades are less conserved than individual taxa placement as the amount of missing data increases.
-Interestingly, 10\% or 25\% missing data for the $M_F$ parameter does not seems to affect the apparition of unstable taxa (Triples NTS modes of respectively 0.97 and 0.93 - Table ~\ref{NTS_ML_results}).
+Interestingly, 10\% or 25\% missing data for the $M_F$ parameter does not seems to affect the apparition of unstable taxa .
 
@@
 In the same way as described above, there seems to be a small difference depending on the metric used.
-There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Fould metric (see Table ~\ref{Ba_RF-MF_results}).
-However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
+There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Foulds metric .
+However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework .
 
@@
 \subsubsection{Effect of $M_C$}
-The number of missing morphological characters ($M_C$), however, seems to be more affecting topological recovery than $M_L$ and $M_F$ (see Table ~\ref{Group_results} - Fig. ~\ref{Fig_Results}).
-The Robinson-Fould metric shows a rapid drop in tree recovery from 10\% of missing data (Robinson-Fould NTS mode $<$ to 0.5 from 10\% missing data - Table ~\ref{NTS_ML_results}).
-This decrease is however slower when using the Triples metric with still a good topological recovery at 10\%  (Triples NTS modes = 0.95 - Table ~\ref{NTS_ML_results}).
+The number of missing morphological characters ($M_C$), however, seems to be more affecting topological recovery than $M_L$ and $M_F$ .
+The Robinson-Foulds metric shows a rapid drop in tree recovery from 10\% of missing data (Robinson-Foulds NTS mode $<$ to 0.5 from 10\% missing data .
+This decrease is however slower when using the Triplets metric with still a good topological recovery at 10\%  (Triplets NTS modes = 0.95 .
 
-The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Fould metric shows a significant effect of the $M_C$ parameter on tree recovery (see Table ~\ref{Ba_RF-ML_results} and ~\ref{Group_results} - Fig.
-~\ref{Fig_Results}).
+The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Foulds metric shows a significant effect of the $M_C$ parameter on tree recovery.
 
@@
 %The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
-%The decrease in clade conservation (low Robinson-Fould NTS score) is faster than the increase in flying taxa (low Triples NTS score).
+%The decrease in clade conservation (low Robinson-Foulds NTS score) is faster than the increase in flying taxa (low Triplets NTS score).
 %In Bayesian framework, topology is badly (if not at all) recovered disregarding the amount of data missing.
@@
 
-% NC: I would make these tables wider - so have a column for Triples, and one for RF, rather than a column for metric.
+% NC: I would make these tables wider - so have a column for Triplets, and one for RF, rather than a column for metric.
 same with Bayesian vs ML, why not have a set of columns instead?
@@
 % NC: That might be too many columns!
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{Tree similarity values per parameter in ML framework}
-  \input{Tables/NTS_ML_results}
-  \label{tab:NTS_ML_results}
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{Tree similarity values per parameter in Bayesian framework}
-  \input{Tables/NTS_BA_results}
-  \label{NTS_BA_results}
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{Group difference per configuration set}
-  \input{Tables/Group_results}
-  \label{Group_results}
-  $^1$ F-value for parametric tests and Kruskal Wallis $chi^2$ for non parametric tests.
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{$M_L$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} 
-  \input{Tables/ML_RF-ML_results}
-  \label{ML_RF-ML_results}
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{$M_L$ non-parametric pairwise difference for Triples metric in ML framework} 
-  \input{Tables/ML_Tr-ML_results}
-  \label{ML_Tr-ML_results}
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{$M_L$ non-parametric pairwise difference for Robinson-Foulds metric in Bayesian framework}
-  \input{Tables/Ba_RF-ML_results}
-  \label{Ba_RF-ML_results}
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{$M_F$ non-parametric pairwise difference for Robinson-Foulds metric in ML framework} 
-  \input{Tables/ML_RF-MF_results}
-  \label{ML_RF-MF_results}
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{$M_F$ non-parametric pairwise difference for Triples metric in ML framework} 
-  \input{Tables/ML_Tr-MF_results}
-  \label{ML_Tr-MF_results}
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{$M_F$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-  \input{Tables/Ba_RF-MF_results}
-  \label{Ba_RF-MF_results}
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in ML framework} 
-  \input{Tables/ML_RF-MC_results}
-  \label{ML_RF-MC_results}
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{$M_C$ non-parametric pairwise difference for Triples metric in ML framework} 
-  \input{Tables/ML_Tr-MC_results}
-  \label{ML_Tr-MC_results}
-\end{table}
-
-\begin{table}
-  \centering
-  \linespread{1.0}
-  \caption{$M_C$ non-parametric pairwise difference for Robinson-Fould metric in Bayesian framework}
-  \input{Tables/Ba_RF-MC_results}
-  \label{Ba_RF-MC_results}
-\end{table}
 
@@
 the fossils) branch in the tree.
-The Robinson-Foulds \citep{RF1981} and the Triples \citep{critchlowthe1996} metrics are more sensible to taxa and clade placement than other tree comparison metrics \citep[e.g.][Imbalance metric]{kirkpatricksearching1993} and where therefore favored.
+The Robinson-Foulds \citep{RF1981} and the Triplets \citep{critchlowthe1996} metrics are more sensible to taxa and clade placement than other tree comparison metrics \citep[e.g.][Imbalance metric]{kirkpatricksearching1993} and where therefore favored.
 Also by using the Normalized Tree Similarity index \citep{Bogdanowicz2012} we emphasize the aspect of "good" phylogenetic signal \textit{versus} random phylogenetic signal because it normalized the values of the metric by correcting for the expected value when comparing random trees (NTS = 0).
 
-The Robinson-Fould metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
+The Robinson-Foulds metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
 When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
-On the other hand, the Triples method is measuring the position of each taxon towards to other reference taxa \citep{critchlowthe1996}.
+On the other hand, the Triplets method is measuring the position of each taxon towards to other reference taxa \citep{critchlowthe1996}.
 It will penalise only trees where taxa get removed furthest from their original clade.
@@
 
-In our case, we suspect that this inability of Bayesian methods to recover topology is due to the intrinsic structure of a Total Evidence matrix (Fig ~\ref{Fig_RemoveData}).
+In our case, we suspect that this inability of Bayesian methods to recover topology is due to the intrinsic structure of a Total Evidence matrix.
 In fact, as previously studies have shown, missing data doesn't seem to be a major drawback as long as the missing data is randomly distributed ~\citep[e.g.][]{wiensmissing2003,rouresite-specific2011,sansomfossilization2013}.
 However, in Total Evidence matrices, the majority of the missing data is not randomly distributed but concentrated in the molecular part of the matrix for the fossil taxa (i.e.
-the missing-data sub-matrix, Fig ~\ref{Fig_RemoveData}).
+the missing-data sub-matrix).
 This leads to high decrease of topological recovery when using non optimal criterion approach (i.e.
@@
 Our results shows that the $M_F$ parameter have less influence on recovering the good tree topology in Maximum Likelihood framework which is fortunate since it is the parameter that is the more difficult to fix for practical reasons.
-Regarding the fact that we are more interested in taxa placement than in clade conservation (Triples \textit{versus} Robinson-Fould), the parameter that affects the more the decrease in topological recovery is the number of missing living taxa in the morphological part of the matrix ($M_L$) and the overall number of morphological characters $M_C$.
+Regarding the fact that we are more interested in taxa placement than in clade conservation (Triplets \textit{versus} Robinson-Fould), the parameter that affects the more the decrease in topological recovery is the number of missing living taxa in the morphological part of the matrix ($M_L$) and the overall number of morphological characters $M_C$.
 This makes sense since the living taxa are bearing both the information to build the tree backbone (the molecular data) and the information used for branching the fossil on this backbone (the morphological information).
@@
 \section{Acknowledgements}
-% NC: Try and keep these short and to the point
-Thanks to Frédéric Delsuc, Emmanuel Douzery, Trevor Hodkinson and Andrew Jackson and the members of the Macro Journal Club %Also Gavin Thomas if not authors - %TG : also Adam Kev and Sive or that's putting to many people, thinking about it, they helped at least as much as Andrew (not mentioning Trevor but that's more political)?
+Thanks to Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson and Andrew Jackson, Gavin Thomas, April Wright and the members of the Macro Journal Club %Also Gavin Thomas if not authors - %TG : also Adam Kev and Sive or that's putting to many people, thinking about it, they helped at least as much as Andrew (not mentioning Trevor but that's more political)?
+%Not sure how many people we have to thank but it could also include Graham Slater and Nick Matzke.
 for useful comments on our simulation protocol.
@@
 Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.
-% NC: Is this what you're supposed to cite for the cluster?
-Check.
-- TG: yep
 This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
 
-%BIBLIOGRAPHY
- % The \cite command functions as follows:
- %   \citet{key} ==>>                Jones et al. (1990)
- %   \citet*{key} ==>>               Jones, Baker, and Smith (1990)
- %   \citep{key} ==>>                (Jones et al., 1990)
- %   \citep*{key} ==>>               (Jones, Baker, and Smith, 1990)
- %   \citep[chap. 2]{key} ==>>       (Jones et al., 1990, chap. 2)
- %   \citep[e.g.][]{key} ==>>        (e.g. Jones et al., 1990)
- %   \citep[e.g.][p. 32]{key} ==>>   (e.g. Jones et al., p. 32)
- %   \citeauthor{key} ==>>           Jones et al.
- %   \citeauthor*{key} ==>>          Jones, Baker, and Smith
- %   \citeyear{key} ==>>             1990
-
-\bibliographystyle{sysbio} %don't write the suffix
-\bibliography{References} %don't write the suffix
+\bibliographystyle{sysbio}
+\bibliography{References}
 
 \section{Supplementary Material}
-% NC: What do they call this in Syst Biol? Supplemental Materials is more usual...
-% NC: You probably want to give these numbers so you can refer people to supp mat section 1, section 2 etc. etc.
+
     \input{SupplementaryMaterial/SupplementaryMaterial}

### 2014-11-25 Analysis/MissingData_plot.R
Minor plot updates

@@ -30,8 +30,8 @@ if(quick.load == FALSE) {
 #Plot main figures (individual parameters in RF and Triples)
 #####################
 #windows parameters
-quartz(width = 8.3, height = 5.8) #A5 landscape
-#quartz(width = 16.6, height = 11.6) #A4 landscape
+#quartz(width = 8.3, height = 5.8) #A5 landscape
+quartz(width = 16.6, height = 11.6) #A4 landscape
 op<-par(mfrow=c(2,3), bty="l", oma=c(0,0,0,0))
 ylim=c(0,1)
 
@@ -42,7 +42,7 @@ metrics<-c("R.F_Cluster","Triples")
 #2*3 plots
 #Missing living RF
 par(mar=c(1,5,5,0) + 0.1)
-multi.TreeCmp.plot( data.sets , par_ML, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab="Robinson-Fould distance", las=2)
+multi.TreeCmp.plot( data.sets , par_ML, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab="Robinson-Foulds distance", las=2)
 #Missing fossil RF
 par(mar=c(1,3,5,2) + 0.1)
 multi.TreeCmp.plot( data.sets , par_MF, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
@@ -69,8 +69,8 @@ par(op)
 #Plot method differences figure
 #####################
 #windows parameters
-quartz(width = 8.3, height = 5.8) #A5 landscape
-#quartz(width = 16.6, height = 11.6) #A4 landscape
+#quartz(width = 8.3, height = 5.8) #A5 landscape
+quartz(width = 16.6, height = 11.6) #A4 landscape
 op<-par(mfrow=c(2,3), bty="l", oma=c(0,0,0,0))
 ylim=c(0,1)
 
@@ -81,7 +81,7 @@ metrics<-c("R.F_Cluster","Triples")
 #2*3 Plots
 #Missing living RF
 par(mar=c(1,5,5,0) + 0.1)
[…]

### 2014-11-25 Functions/TreeComparisons/TreeCmp.Plot.R

@@ -448,7 +448,7 @@ signi.TreeCmp.plot<-function(data.list, parameter, metric, probs, diff.type, pos
         probs<-c(95,50)
         shift='auto'
         diff.type<-'method'
-        multi.TreeCmp.plot( data.sets , par_ML, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab="Robinson-Fould distance", las=2)
+        multi.TreeCmp.plot( data.sets , par_ML, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab="Robinson-Foulds distance", las=2)
     }
 
 
@@ -664,12 +664,15 @@ global.TreeCmp.plot<-function(data.list, parameter, metrics, col, col.grad=c('bl
     #FIRST PLOT
 
     #Plot the TreeCmp results
-    par(mar=c(0,4,2,2))
-    multi.TreeCmp.plot(data.list=data.list, parameter=parameter, metrics=metrics[1], col=col, shift=shift, ylab="Robinson-Fould distance", ...)
+    par(mar=c(0,4,2,2), bty="l")
+    multi.TreeCmp.plot(data.list=data.list, parameter=parameter, metrics=metrics[1], col=col, shift=shift, ylab="Robinson-Foulds distance", ...)
     
     #Add the legend bars
     par(mar=c(2,4,0,2))
-    plot(1,1, xlab='', ylab='', xlim=c(1,125), ylim=c(0,3), type='n', xaxt='n', yaxt='n', bty='n')
+    plot(1,1, xlab='', ylab='', xlim=c(1,125), ylim=c(0,3), type='n', yaxt='n', bty='n', xaxt='n') #
+    text(-2, 2.5, "ML")
+    text(-2, 1.5, "MF")
+    text(-2, 0.5, "MC")
     #ML
     col.new<-colgrad[1:5]
     for(i in 1:5) {
@@ -686,15 +689,19 @@ global.TreeCmp.plot<-function(data.list, parameter, metrics, col, col.grad=c('bl
         rect(xleft=(1*i-1), ybottom=0, xright=(1*i), ytop=1, col=col.new[i], broder=NULL, lty='blank')
     }
 
+
     #SECOND PLOT
 
     #Plot the TreeCmp results
-    par(mar=c(0,4,2,2))
[…]

### 2014-11-25 Analysis/MissingData_plot.R
Black and White figures

@@ -169,7 +169,7 @@ data.sets<-c("ML_besttrees","Bayesian_contrees")#,"ML_bootstraps","Bayesian_tree
 metrics<-c("R.F_Cluster","Triples")
 
 #Plot
-global.TreeCmp.plot( data.sets , par_MLMFMC, metrics, col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, las=2, xlab=NULL, shift=0.5, format='A5', col.grad=c("white", "black")) 
+global.TreeCmp.plot( data.sets , par_MLMFMC, metrics, col=c('black', 'grey'), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, las=2, xlab=NULL, shift=0.5, format='A5', col.grad=c("white", "black")) 
 
 
 #####################
@@ -183,4 +183,4 @@ data.sets<-c("ML_bootstraps","Bayesian_treesets")
 metrics<-c("R.F_Cluster","Triples")
 
 #Plot
-global.TreeCmp.plot( data.sets , par_MLMFMC, metrics, col=palette()[3:4], las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, las=2, xlab=NULL, shift=0.5,  format='A4',col.grad=c("white", "black")) 
+global.TreeCmp.plot( data.sets , par_MLMFMC, metrics, col=c('black', 'grey'), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, las=2, xlab=NULL, shift=0.5, format='A5', col.grad=c("white", "black")) 

### 2014-11-25 Manuscript/TEM_draft.tex
Manuscript new version

@@
 \centering
+    \includegraphics[width=1\textwidth]{Figures/BW/Global-ML+consensus-RF+Tr-small.pdf}
+\caption{Trend of the effect of missing data on topological recovery on ML and consensus trees.
+The amount of missing data per parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is represented along the x axis.
+The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
+The topological recovery is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
+Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
+The Maximum Likelihood trees are represented in black and the Bayesian consensus trees in grey.}
+\label{Fig_Global-single}
+\end{figure}
+%Differences between the parameters and between the methods (ML vs BC)
+
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{Figures/BW/Global-Bootstrap+Treeset-RF+Tr-small.pdf}
+\caption{Trend of the effect of missing data on topological recovery on ML and consensus trees.
+The amount of missing data per parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is represented along the x axis.
+The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
+The topological recovery is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
+Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
+The Bootstrap trees pairwise comparisons are represented in black and the Bayesian posterior trees distribution pairwise comparisons in grey.}
+\label{Fig_Global-treesets}
+\end{figure}
+%Differences between the parameters and between the methods (BS vs TS)
+
+\begin{figure} 
+\centering
     \includegraphics[width=1\textwidth]{Figures/colour/Parameters-AllMethods-RF+Tr-small.pdf}
-\caption{Effect of missing data on topological recovery.
+\caption{Comparison between the effect of missing data and the tree inference method on topological recovery.
 The amount of missing data for each parameter is represented on the x axis.
@@
 The Maximum Likelihood trees are represented in black, the Bayesian consensus trees in red, the bootstrap trees in green and the posterior tree distribution in blue.}
-\label{Fig_Results}
+\label{Fig_Parameter-Method}
 \end{figure}
+%Differences within the parameters, not between
 
 \subsection{Effect of the method on topological recovery}
--Bayesian significantly better than ML for clade conservation.
--Bayesian consensus trees plateaus around 0.8 topological recovery at worst whatever combination of missing data.
+-With RF Bayesian significantly better than ML for clade conservation.
+
+-With RF Bayesian consensus trees plateaus around 0.8 topological recovery at worst whatever combination of missing data.
+
+-With Tr however, ML is more performant but not significantly. When looking at uncertainty, both results tend to just overlap.
 
 \subsection{Effect of the missing data on topological recovery}
+-No big difference between the parameters individually
 
-\subsection{Effect of $M_L$}
-In our Maximum Likelihood analyses, topological recovery (i.e. tree similarity) decreased as the percentage of missing data increased, using both the Robinson-Foulds and Triplets metrics.
-However, topological recovery was not significantly different for trees with 25\%, 50\% and 75\% missing morphological data for living taxa (Fig 1b???), % is this what you meant?
-And indicate the particular panel of the figure it refers to
-suggesting that although tree topology is sensitive to this missing data, once it is greater than 25\% it makes little difference.
-Overall, the Robinson-Foulds metric shows lower tree similarities than the Triplets metric.
-This indicates that the positions of individual taxa are generally more conserved than the membership of entire clades.
-
-Topological recovery was poor in our Bayesian analyses, regardless of the amount of missing data (Figures???), although topological recovery did decrease slightly as missing data increased.
-Overall, tree similarity index was only slightly above zero which means that trees are only slightly more similar than we would expect if we compared two completely random trees.
-%For the Triplets metric there is no significant effect of the $M_L$ parameter, in other words, if living taxa are removed, there is no effect on the placement of "flying" taxa.
-%For the Robinson-Foulds metric, there is a significant effect of the $M_L$ leading small decrease in topological recovery when $M_L$ increases (see Table ~\ref{Ba_RF-ML_results}).
-
-%NC: A shorter verison of this encompassing all three parameters will work here.
-
-%NC: Combine with section above
-\subsubsection{Effect of $M_F$}
-In Maximum Likelihood framework, the effect of missing data in the fossil record appeared to be constant and led to a constant decrease in topological recovery when the missing data increase.
-The difference between the two metrics used is more important than for the $M_L$ parameter: clades are less conserved than individual taxa placement as the amount of missing data increases.
-Interestingly, 10\% or 25\% missing data for the $M_F$ parameter does not seems to affect the apparition of unstable taxa .
-
-Similarly than for the effect of $M_L$, Bayesian tree recovery was only slightly better than expected by chance.
-In the same way as described above, there seems to be a small difference depending on the metric used.
-There is no significant effect of the $M_F$ parameter when using the Triplet metric but the effect is significant when using the Robinson-Foulds metric .
-However, as mentioned previously the effect of the parameter $M_F$ is still minimal in a Bayesian framework .
-
-%NC: Combine with section above
-\subsubsection{Effect of $M_C$}
-The number of missing morphological characters ($M_C$), however, seems to be more affecting topological recovery than $M_L$ and $M_F$ .
-The Robinson-Foulds metric shows a rapid drop in tree recovery from 10\% of missing data (Robinson-Foulds NTS mode $<$ to 0.5 from 10\% missing data .
-This decrease is however slower when using the Triplets metric with still a good topological recovery at 10\%  (Triplets NTS modes = 0.95 .
-
-The effect of $M_C$ in Bayesian framework is similar as the one described for $M_L$ and $M_F$: tree recovery is only slightly better than expected by chance and only the Robinson-Foulds metric shows a significant effect of the $M_C$ parameter on tree recovery.
-
-%\subsubsection{Combined effect}
-%Need to do the heat plots
-
-$\newline$
-
-%NC: This will be part of the section above. If it's short you may not need to sum up!
-%To sum up:
-%The ability of recovering the "best" tree's topology in Maximum Likelihood method is function of the amount of data missing.
-%The parameters $M_C$ and $M_L$ have more influence than $M_F$ on decrease in topological recovery.
-%The decrease in clade conservation (low Robinson-Foulds NTS score) is faster than the increase in flying taxa (low Triplets NTS score).
-%In Bayesian framework, topology is badly (if not at all) recovered disregarding the amount of data missing.
-
-% NC: You will need two sections here with subheadings - one for the sensitivity analyses you did with the submatrices, and one for the empirical results.
-
-% NC: I would make these tables wider - so have a column for Triplets, and one for RF, rather than a column for metric.
-same with Bayesian vs ML, why not have a set of columns instead?
-It'll take up less space and be easier to read I think...
-
-% NC: You should do what I have done on the thesis template and have these tables as separate files rather than having them in the document.
-
-% NC: we need to sit down and look at these tables. I don't like the current format.
-
-% About the tables:
-
-% What about making this table in a longevity-results style? with every thing combined (30 columns)?
-% NC: That might be too many columns!
+\subsection{Effect of the amount of missing data}
 

### 2014-11-30 Functions/TreeComparisons/bhatt.coef.R
v0.2.1 verbose option properly removed

@@ -88,7 +88,7 @@ bhatt.coeff<-function(x,y, bw=bw.nrd0, ...) {
 #Pairwise Bhattacharyya Coefficient
 ##########################
 #Calculates the pairwise Bhattacharyya Coefficient for n distributions
-#v.0.2
+#v.0.2.1
 #Update: allows to give another list Y to compare to the list X.
 ##########################
 #SYNTAX :
@@ -99,7 +99,7 @@ bhatt.coeff<-function(x,y, bw=bw.nrd0, ...) {
 #<...>  any optional arguments to be passed to the given bw function.
 ##########################c
 #----
-#guillert(at)tcd.ie - 29/11/2014
+#guillert(at)tcd.ie - 30/11/2014
 ##########################
 #Requirements:
 #-R 3
@@ -133,10 +133,6 @@ pair.bhatt.coeff<-function(X, Y=NULL, bw=bw.nrd0, diag=FALSE, ...) {
         stop("'diag' must be logical.")
     }
 
-    #verbose
-    if(class(verbose) != 'logical') {
-        stop("'verbose' must be logical.")
-    }
 
     if(square.comp == TRUE) {
         #PAIRWISE COMPARISONS (ONE TO ALL)

### 2014-11-30 Analysis/MissingData_differences.R
Fixed typos in plot

@@ -65,11 +65,11 @@ load("../Data/R_data/pair.comp_within.Baytre.RF.Rda")
 load("../Data/R_data/pair.comp_within.Baytre.Tr.Rda")
 
 #Visualising the results
-pairwise.comp_within<-c("pairwise.comp_within.MLbest.RF", "pairwise.comp_within.MLbest.Tr",
-                        "pairwise.comp_within.Baycon.RF", "pairwise.comp_within.Baycon.Tr",
-                        "pairwise.comp_within.MLboot.RF", "pairwise.comp_within.MLboot.Tr",
-                        "pairwise.comp_within.Baytre.RF", "pairwise.comp_within.Baytre.Tr")
-plot.bhatt.coeff(get(pairwise.comp_within[1]), col=c("red", "green"), col.grad=c("white", "black"))
+pairwise.comp_within<-c("pair.comp_within.MLbest.RF", "pair.comp_within.MLbest.Tr",
+                        "pair.comp_within.Baycon.RF", "pair.comp_within.Baycon.Tr",
+                        "pair.comp_within.MLboot.RF", "pair.comp_within.MLboot.Tr",
+                        "pair.comp_within.Baytre.RF", "pair.comp_within.Baytre.Tr")
+plot.bhatt.coeff(get(pairwise.comp_within[8]), col=c("red", "green"), col.grad=c("white", "black"))
 
 #----------------
 #Calculating the pairwise comparison without (effect of the method)
@@ -121,5 +121,6 @@ names(pair.comp_without.Tr)<-c("MLbest vs Baycon", "MLbest vs MLboot",
                                "Baycon vs Baytre", "MLboot vs Baytre")
 
 #Plot the distributions
-plot.bhatt.coeff.dist(pair.comp_without.RF, col=palette())
-plot.bhatt.coeff.dist(pair.comp_without.Tr, col=palette())
+op<-par(bty="l")
+plot.bhatt.coeff.dist(pair.comp_without.RF, col=palette(), xlab="Bhattacharyya Coefficient", ylab="Frequency")
+plot.bhatt.coeff.dist(pair.comp_without.Tr, col=palette(), xlab="Bhattacharyya Coefficient", ylab="Frequency")

### 2014-11-30 Functions/TreeComparisons/plot.bhatt.coef.R
Removed debug

@@ -168,7 +168,7 @@ plot.bhatt.coeff.dist<-function(list, col, ...) {
     
 
     #Plotting the distributions
-    plot(1,1, xlim=c(min(ext.coord$minx), max(ext.coord$maxx)), ylim=c(min(ext.coord$miny), max(ext.coord$maxy)), type='n')#, ...)
+    plot(1,1, xlim=c(min(ext.coord$minx), max(ext.coord$maxx)), ylim=c(min(ext.coord$miny), max(ext.coord$maxy)), type='n', ...)
     for (n in 1:length(density.list)){
         lines(density.list[[n]], col=col[n])
     }

### 2014-11-30 Manuscript/TEM_draft.tex
Updated figures

@@
 \centering
+    \includegraphics[width=1\textwidth]{Figures/colour/Parameters-AllMethods-RF+Tr-small.pdf}
+\caption{Comparison between the effect of missing data and the tree inference method on topological recovery.
+The amount of missing data for each parameter is represented on the x axis.
+The topological recovery is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
+Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
+The Maximum Likelihood trees are represented in black, the Bayesian consensus trees in red, the bootstrap trees in green and the posterior tree distribution in blue.}
+\label{Fig_Results_1} %Differences within a subset of parameters and between methods
+\end{figure}
+
+\begin{figure} 
+\centering
     \includegraphics[width=1\textwidth]{Figures/BW/Global-ML+consensus-RF+Tr-small.pdf}
@@
 The Maximum Likelihood trees are represented in black and the Bayesian consensus trees in grey.}
-\label{Fig_Global-single}
+\label{Fig_Results_2} %Differences between all the parameters and between two methods (ML vs BC)
 \end{figure}
-%Differences between the parameters and between the methods (ML vs BC)
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{Figures/BW/Global-Bootstrap+Treeset-RF+Tr-small.pdf}
-\caption{Trend of the effect of missing data on topological recovery on ML and consensus trees.
-The amount of missing data per parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is represented along the x axis.
+    \includegraphics[width=1\textwidth]{Figures/colour/PairComp_within-Baycon-RF+Tr.pdf}
+\caption{Pairwise Bhattacharyya Coefficients within the Bayesian consensus trees.
+The pairwise trees comparisons are represent on both axis.
 The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
-The topological recovery is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
-Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
-The Bootstrap trees pairwise comparisons are represented in black and the Bayesian posterior trees distribution pairwise comparisons in grey.}
-\label{Fig_Global-treesets}
-\end{figure}
-%Differences between the parameters and between the methods (BS vs TS)
+The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0).
+A.
+Results for the Normalised Robinson-Foulds distance.
+B.
+Results with the for Normalised Triplets distance.}
+\label{Fig_Results_3}
+\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{Figures/colour/Parameters-AllMethods-RF+Tr-small.pdf}
-\caption{Comparison between the effect of missing data and the tree inference method on topological recovery.
-The amount of missing data for each parameter is represented on the x axis.
-The topological recovery is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
-Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
-The Maximum Likelihood trees are represented in black, the Bayesian consensus trees in red, the bootstrap trees in green and the posterior tree distribution in blue.}
-\label{Fig_Parameter-Method}
+    \includegraphics[width=1\textwidth]{Figures/colour/PairComp_without-RF+Tr.pdf}
+\caption{Distribution of the Bhattacharyya Coefficients between methods.
+Curves represents the kernel density estimations of the BC between the four different methods.
+A.
+Results for the Normalised Robinson-Foulds distance.
+B.
+Results with the for Normalised Triplets distance.
+Pikes around the two different extremes values of the BC show the two pairs of most dissimilar methods (Bayesian consensus vs.
+Bootstraps and Bayesian consensus vs.
+Bayesian tree posterior distribution) for the Normalised Robinson-Foulds distance ; and the most similar methods (Bootstraps and Bayesian tree posterior distribution) for the Normalised Triples methods.}
+\label{Fig_Results_4} %Comparisons between methods, both in RF and Triplets.(method differences)
 \end{figure}
-%Differences within the parameters, not between
 

### 2014-12-01 Analysis/MissingData_modetable.R
Table of morale values distributions

@@ -1,0 +1,82 @@
+#Plotting the results of the Total Evidence Missing Data Analysis
+
+setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
+
+#Loading the functions
+source("MissingData_fun.R")
+
+#Loading the data
+quick.load=TRUE
+if(quick.load == FALSE) {
+    source("MissingData_load.R") 
+} else {
+    #Quick load
+    load("../Data/R_data/TreeCmp-ML_bestrees.Rda")
+    load("../Data/R_data/TreeCmp-ML_bootstraps.Rda")
+    load("../Data/R_data/TreeCmp-Bayesian_contrees.Rda")
+    load("../Data/R_data/TreeCmp-Bayesian_treesets.Rda")
+
+    #Isolating the parameters per chains
+    par_ML<-c(1,26,51,76,101)
+    par_MF<-c(1,6,11,16,21)
+    par_MC<-c(1:5)
+    par_MLMF<-c(1,6,11,16,21, 26,31,36,41,46, 51,56,61,66,71, 76,81,86,91,96, 101,106,111,116,121)
+    par_MLMC<-c(1:5, 26:30, 51:55, 76:80, 101:105)
+    par_MFMC<-c(1:25)
+    par_MLMFMC<-c(1:125)
+}
+
+#Storing the sub data sets
+boot.RF<-sub.data(get(data.sets[3]), par_MLMFMC, metric=metrics[1])
+boot.Tr<-sub.data(get(data.sets[3]), par_MLMFMC, metric=metrics[2])
+bayt.RF<-sub.data(get(data.sets[4]), par_MLMFMC, metric=metrics[1])
+bayt.Tr<-sub.data(get(data.sets[4]), par_MLMFMC, metric=metrics[2])
+ML.RF<-sub.data(get(data.sets[1]), par_MLMFMC, metric=metrics[1])
+ML.Tr<-sub.data(get(data.sets[1]), par_MLMFMC, metric=metrics[2])
+Bay.RF<-sub.data(get(data.sets[2]), par_MLMFMC, metric=metrics[1])
+Bay.Tr<-sub.data(get(data.sets[2]), par_MLMFMC, metric=metrics[2])
+
+#Calculating the hdr
[…]

### 2014-12-01 Manuscript/SupplementaryMaterial/Supp_results.tex
Updated supplementary

\section{Supplementary material Section 3}

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-Global-Bootstrap+Treeset-RF+Tr-small.pdf}
\caption{Trend of the effect of missing data on topological recovery on the Bootstraps and the Bayesian posterior trees distributions. The amount of missing data per parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is represented along the x axis. The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data. The topological recovery is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row). Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions. The Bootstraps are represented in black and the Bayesian posterior trees distributions in grey.}
\label{Fig_global_BootTreesets} %Differences between all the parameters and between two methods (Boot vs treesets)
\end{figure}

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-Baytre-RF.pdf}
\caption{Pairwise Bhattacharyya Coefficients within the Bayesian posterior trees distributions. The pairwise trees comparisons are represent on both axis. The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data. The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0). Results are calculated for the Normalised Robinson-Foulds distance.}
\label{Fig_pairComp-Baytree-RF}
\end{figure} %Pairwise BC for the Bayesian treesets for RF. (parameters differences)

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-Baytre-Tr.pdf}
\caption{Pairwise Bhattacharyya Coefficients within the Bayesian posterior trees distributions. The pairwise trees comparisons are represent on both axis. The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data. The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0). Results are calculated for the Normalised Triplets distance.}
\label{Fig_pairComp-Baytree-Tr}
\end{figure} %Pairwise BC for the Bayesian treesets for Tr. (parameters differences)

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-MLbest-RF.pdf}
\caption{Pairwise Bhattacharyya Coefficients within the Maximum Likelihood trees. The pairwise trees comparisons are represent on both axis. The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data. The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0). Results are calculated for the Normalised Robinson-Foulds distance.}
\label{Fig_pairComp-MLbest-RF}
\end{figure} %Pairwise BC for the ML for RF. (parameters differences)

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-MLbest-Tr.pdf}
\caption{Pairwise Bhattacharyya Coefficients within the Maximum Likelihood trees. The pairwise trees comparisons are represent on both axis. The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data. The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0). Results are calculated for the Normalised Triplets distance.}
\label{Fig_pairComp-MLbest-Tr}
\end{figure} %Pairwise BC for the ML for Tr. (parameters differences)

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-MLboot-RF.pdf}
\caption{Pairwise Bhattacharyya Coefficients within the Bootstraps. The pairwise trees comparisons are represent on both axis. The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data. The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0). Results are calculated for the Normalised Robinson-Foulds distance.}
\label{Fig_pairComp-MLboot-RF}
\end{figure} %Pairwise BC for the Bootstraps for RF. (parameters differences)

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-MLboot-Tr.pdf}
\caption{Pairwise Bhattacharyya Coefficients within the Bootstraps. The pairwise trees comparisons are represent on both axis. The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data. The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0). Results are calculated for the Normalised Triplets distance.}
\label{Fig_pairComp-MLboot-Tr}
\end{figure} %Pairwise BC for the Bootstraps for Tr. (parameters differences)

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
 & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum likelihood-RF & 0.06 & 0.26 & 0.39 & 0.40 & 0.49 & 0.94 \\ 
  Maximum likelihood-Tr & 0.27 & 0.45 & 0.58 & 0.63 & 0.83 & 1.00 \\ 
  Bayesian consensus-RF & 0.70 & 0.71 & 0.73 & 0.76 & 0.79 & 0.96 \\ 
  Bayesian consensus-Tr & -0.25 & -0.09 & 0.18 & 0.21 & 0.40 & 0.98 \\ 
  Bootstraps-RF & 0.06 & 0.18 & 0.27 & 0.26 & 0.33 & 0.46 \\ 
  Bootstraps-Tr & 0.24 & 0.32 & 0.36 & 0.38 & 0.46 & 0.60 \\ 
  Bayesian posterior trees-RF & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
  Bayesian posterior trees-Tr & 0.24 & 0.36 & 0.40 & 0.50 & 0.68 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}

### 2014-12-01 Manuscript/SupplementaryMaterial/SupplementaryMaterial.tex
Updated supplementary

\subsection{Tree Building}
  \input{SupplementaryMaterial/TreeBuilding}
  \label{Supp_TreeBuilding}

\subsection{Tree Comparisons}
  \input{SupplementaryMaterial/TreeComparison}
  \label{Supp_TreeComparison}

\subsection{Codes}
All codes are available at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions$

The tree comparison results analysis can be repeated for more details at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Analysis$ %fully running! Try it at home (5-10 minutes of data loading)!

\subsection{Additional Results}
  \input{SupplementaryMaterial/Supp_results}
  \label{Supp_results}

### 2014-12-01 Manuscript/SupplementaryMaterial/TreeComparison.tex
Supplementary update
Updated supplementary

@@
+\section{Supplementary material Section 2}
+
 \subsection{Triplets distance details ($T_{x,y}$)}
@@
 
+\subsection{Normalised Tree Similarity}
+For any tree with \textit{n} taxa compared using a tree distance metric $m$, Normalized Tree Similarity, $NTS_m$ (\citealp{Bogdanowicz2012}), represents the similarity score for the two trees given the expected distance between two random Yule trees with $n$ taxa.
+If $\bar{d}_{m,n}$\textit{(rand)} is the average distance between two random Yule trees with $n$ taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} containing each $n$ taxa, then:
+\begin{equation}
+NTS_{m,n}(x,y)=\frac{\bar{d}_{m,n}(rand) - d_{m,n}(x,y)} {\bar{d}_{m,n}(rand)}
+\end{equation}
+\textit{NTS} ranges from one to -$\infty$.
+For any $m,n$, when $NTS$ = 1, the trees are identical, when \textit{NTS} = 0 the trees are no more different than expected by chance, and when $NTS$ $<$ 0, the trees are more different than expected when comparing two random trees.
+
 \subsection{Tree comparisons}
@@
 We calculated the $\bar{d}_{m,n}$\textit{(rand)} value using an approach similar to the RPCBTC (described below) by performing 1000 random pairwise comparisons using the TreeCmp java script \citep{Bogdanowicz2012}.
-
-\subsubsection{Random Pairwise Bayesian Tree Comparison (RPBTC)}
-We assessed the power of the Random Pairwise Bayesian Tree Comparison (RPBTC) method by comparing 1000 random trees from a posterior distribution trees set to another 1000 random trees from the same posterior distribution trees set.
-We repeated this 100 times independently using the same posterior distribution trees set each time resulting in 100 replicates of the same posterior distribution trees set compared 1000 times.
-We used an ANOVA to test if there was no significant difference between the replicates so that the RBTC can be replicated.
-We applied this protocol on a poorly resolved tree (Low Score), a resolved tree with low support value (Medium Score) and a resolved tree with high support values (High Score).
-Results are available in table ~\ref{RPBTC_testing}. %link broken
-
-\begin{table}[ht]
-  \caption{Group comparison results: difference between 100 replicates using the RPBTC method} %test with TreeCmp.anova
-  \centering
-  \input{SupplementaryMaterial/Supp_Tables/RPBTC_testing}
-  \label{RPBTC_testing}
-\end{table}

### 2014-12-01 Manuscript/TEM_draft.tex
Updated discussion plan + finished results

@@
 
--Overall missing data is doesn't affect topological recovery in Bayesian (around minimum 70\% of topological topology in the worth scenario).
+-Overall missing data is doesn't affect topology in Bayesian (around minimum 70\% of topological topology in the worth scenario).
 
@@
 
-The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014}.
+The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,wrightbayesian2014}.
 Missing molecular data has been seen by some authors as an issue because it can, in some part of the tree, decrease phylogenetic signal (i.e.
@@
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (See \hyperref[supplementaries]{supplementaries}).
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (See \hyperref[SupplementaryMaterial]{Supplementary Material Section 1}).
 We then ran an independent discrete character simulation for each character using the "true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
@@
 When $NTS_m$=1, the two trees are strictly identical, when $NTS_m$=0 the trees are no more different than expected when comparing two random trees and when $NTS_m$$<$0, the difference between the two trees is greater than when comparing two random trees.
-In our study we used the $NTS_m$ index as a proxy for topological recovery: a high score of this index (i.e.
+In our study we used the $NTS_m$ index as a proxy for topology: a high score of this index (i.e.
 towards 1) means that the topology is highly conserved between the two trees; on the opposite, a low score of this index (i.e.
@@
 \subsubsection{Tree comparisons}
+\label{tree_comparisons}
 For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Robinson-Foulds and Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012}.
-For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and n = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
+For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and n = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}).
 % TG: Maybe remove supplementary here
@@
 We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}.
-Finally we tested for significant differences among the different distributions of trees using non-parametric group and pairwise comparisons by using respectively Kruskal-Wallis and Nemenyi-Damico-Wolfe-Dunn test (CITE R PACKAGE)\citet{ruxtontime2008}.
+%Finally we tested for significant differences among the different distributions of trees using non-parametric group and pairwise comparisons by using respectively Kruskal-Wallis and Nemenyi-Damico-Wolfe-Dunn test (CITE R PACKAGE)\citet{ruxtontime2008}.
+
+In order to investigate the effect of the parameters and/or the methods used in this simulations, we measured the similarity among the different distributions using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
+The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 \citep{Bhattacharyya}.
+For each method (Maximym likelihood trees, Bayesian consensus trees, Bootstraps and Bayesian posterior trees) and metric (Robinson-Foulds and Triplets distance), we used this coefficient in two ways in order to:
+\begin{enumerate}
+\item assess the effect of the method on all parameters:
+
+We performed a pairwise comparison of each parameter between each pair of methods.
+This resulted in 125 comparisons per method and per metric (each parameter combination within one method compared to the same parameter combination in the other method).
+We then compared the distribution of these pairwise comparisons to see the global difference between two methods: if the method have really similar results, we would expect the distribution to be cluster around 1 (high probability of overlap between the distributions) and if the methods are really dissimilar around 0 (low probability of overlap - Fig.~\ref{Fig_Results-paircomp_without}).
+\item assess the effect of the parameters within a method:
+
+We performed a pairwise comparison per method and per metric for each combination of parameters.
+This resulted in 7875 pairwise comparisons per method and per metric (triangle of a 125*125 matrix).
+We represented these results as a triangular matrix with the values of each pairwise comparison coloured according to the value of the Batthacharyya Coefficient (coloured in green when the distributions overlap completely and in red when they don't - Fig.~\ref{Fig_Results-paircomp_within}).
+\end{enumerate}
 
@@
 
+\subsection{Effect of missing data on topology}
+
 \begin{figure} 
@@
     \includegraphics[width=1\textwidth]{Figures/colour/Parameters-AllMethods-RF+Tr-small.pdf}
-\caption{Comparison between the effect of missing data and the tree inference method on topological recovery.
+\caption{Comparison between the effect of missing data and the tree inference method on topology.
 The amount of missing data for each parameter is represented on the x axis.
-The topological recovery is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
+The topology is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
 Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
 The Maximum Likelihood trees are represented in black, the Bayesian consensus trees in red, the bootstrap trees in green and the posterior tree distribution in blue.}
-\label{Fig_Results_1} %Differences within a subset of parameters and between methods
+\label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
 \end{figure}
+
+As it would be expected form the literature, the amount of missing data in the morphological matrix does decrease the ability to recover the right topology, regardless of the parameter, the method or the metric used \citep{rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,wrightbayesian2014}.
+However each variable does not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam}).
+Regarding the conservation of entire clades (i.e.
+the Ronbinson-Foulds distance), the Bayesian consensus trees outperform the other methods and the amount of missing data in the living taxa ($M_{L}$) decreases the most rapidly the clade conservation when using this method.
+However, when looking at the position of wild card taxa (i.e.
+the Triplets distance), the Maximum Likelihood outperforms the other methods but with wider distributions overlap (Fig.~\ref{Fig_Results-paircomp_without}-B).
 
@@
     \includegraphics[width=1\textwidth]{Figures/BW/Global-ML+consensus-RF+Tr-small.pdf}
-\caption{Trend of the effect of missing data on topological recovery on ML and consensus trees.
+\caption{Trend of the effect of missing data on topology on ML and consensus trees.
 The amount of missing data per parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is represented along the x axis.
 The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
-The topological recovery is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
+The topology is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
 Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
 The Maximum Likelihood trees are represented in black and the Bayesian consensus trees in grey.}
-\label{Fig_Results_2} %Differences between all the parameters and between two methods (ML vs BC)
+\label{Fig_Results-global_perparam} %Differences between all the parameters and between two methods (ML vs BC)
 \end{figure}
+
+When looking at the global trend of the Bayesian consensus trees for all the parameters combinations, this method outperforms Maximum Likelihood for the Ronbinson-Foulds distance and plateaus to a minimal modal tree similarity of 0.69 regardless the amount of missing data (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+However, regarding the Triplets distance, Bayesian consensus trees seems to perform poorly but with great confidence intervals overlaps (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
+Results for the global trend comparison between the uncertainty methods (Bootstrap and Bayesian posterior tree distribution) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+
+\subsection{Effect of the method on topology}
+
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{Figures/colour/PairComp_without-RF+Tr.pdf}
+\caption{Distribution of the Bhattacharyya Coefficients between methods.
+Curves represents the kernel density estimations of the BC between the four different methods.
+A.
+Results for the Normalised Robinson-Foulds distance.
+B.
+Results with the for Normalised Triplets distance.
+Pikes around the two different extremes values of the BC show the two pairs of most dissimilar methods (Bayesian consensus vs.
+Bootstraps and Bayesian consensus vs.
+Bayesian posterior trees) for the Normalised Robinson-Foulds distance ; and the most similar methods (Bootstraps and Bayesian posterior trees) for the Normalised Triples methods.}
+\label{Fig_Results-paircomp_without} %Comparisons between methods, both in RF and Triplets.(method differences)
+\end{figure}
+
+When looking at the comparisons between the methods using the Robinson-Foulds distance, the Bayesian consensus trees have the least overlap with respectively the Bayesian posterior trees and the Bootstrap (Fig.~\ref{Fig_Results-paircomp_without}-A).
+When regarding at the Triplets distance, the two method that have the most distribution are the Bootstraps and the Bayesian posterior trees (Fig.~\ref{Fig_Results-paircomp_without}-B).
+
+\subsection{Effect of the missing data parameters on topology}
 
@@
 Results with the for Normalised Triplets distance.}
-\label{Fig_Results_3}
+\label{Fig_Results-paircomp_within}
 \end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
 
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{Figures/colour/PairComp_without-RF+Tr.pdf}
-\caption{Distribution of the Bhattacharyya Coefficients between methods.
-Curves represents the kernel density estimations of the BC between the four different methods.
-A.
-Results for the Normalised Robinson-Foulds distance.
-B.
-Results with the for Normalised Triplets distance.
-Pikes around the two different extremes values of the BC show the two pairs of most dissimilar methods (Bayesian consensus vs.
-Bootstraps and Bayesian consensus vs.
-Bayesian tree posterior distribution) for the Normalised Robinson-Foulds distance ; and the most similar methods (Bootstraps and Bayesian tree posterior distribution) for the Normalised Triples methods.}
-\label{Fig_Results_4} %Comparisons between methods, both in RF and Triplets.(method differences)
-\end{figure}
-
-\subsection{Effect of the method on topological recovery}
--With RF Bayesian significantly better than ML for clade conservation.
-
--With RF Bayesian consensus trees plateaus around 0.8 topological recovery at worst whatever combination of missing data.
-
--With Tr however, ML is more performant but not significantly. When looking at uncertainty, both results tend to just overlap.
-
-\subsection{Effect of the missing data on topological recovery}
--No big difference between the parameters individually
-
-\subsection{Effect of the amount of missing data}
+When looking at the pairwise Bhattacharyya Coefficients between all the parameters combinations for Bayesian consensus trees, the number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}-A).
+However, when looking at the Triplets distance, because of the high overlap in confidence intervals, there is no major region with low probability of overlap.
+One should note however that the bottom line of the matrix where the Bhattacharyya Coefficients are very low is due to comparing the distribution of the pairwise comparison of the best tree vs.
+the best tree, leading to a Normalised triplet score of 1 every time (with no variance, see Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+All the other pairwise comparisons are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 
@@
 
-\subsection{Building the trees}
-Simulating evolutionary history matrices still remains a big drawback in theoretical phylogenetics. %Hahahaha.... feck.
-The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
-% NC: Caveats later on
-This configuration probably lead to globally low phylogenetic signal as well as the intrinsic difficulties to simulated characters with phylogenetic signal.
-Even though molecular characters evolution (and therefore simulation) only depends of a small number of parameters (i.e.
-the base frequencies and the substitution matrix), simulating molecular matrix with a strong evolutionary signal is still complicated when generating unrealistically small matrices.
-For morphological characters, the underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
+%§1
+%Results: global
+%   -Missing data increase decreases topological recovery
+%The results of our simulations are in adequacy with recent analysis on the effect of missing data in morphological matrices \citep{rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,wrightbayesian2014}.
+Our results are showing that ability to recover the correct topology decreases with the amount of missing data, regardless the metric, the method or the parameters considered.
+In more details, our results shows however a clear superiority of the Bayesian consensus tree over other methods, at least when considering the Robinson-Foulds distance (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+Also, our results suggest that the missing data for missing taxa parameter ($M_{L}$) shows a slightly higher decrease in the ability to recover the correct topology (Fig.~\ref{Fig_Results-paircomp_within}).
+%Results 1
+%   -However, methods and metrics are unequal
+%Results 2
+%   -Additionally, parameters have different effect
+%Wrap up results
+%   -Thus, Baycon is better with loads of living
 
+%§2
+%Results from litterature
+%   -Missing data per se is not a problem, problem is in the amount of overlap
+%Proposed hypothesis
+%   -One can reduce overlap by adding living taxa
+
+%§3
+%Discussing the metrics in detail
+%   -RF is different than triplets and better.
+%However, our results can be contrasted whether one looks at the Robinson-Foulds or the Triplets distance.
+When using the second one, the confidence interval of our different parameters combinations per methods seems higher (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}) and the probability of overlap between the distribution of each parameter is high (Bhattacharyya Coefficient close to 1 - Fig.~\ref{Fig_Results-paircomp_within}-B).
+The contrast between the results of the two metrics can be explained by the actual nature of the metrics:
+%\begin{enumerate}
+%\item{The Robinson-Foulds metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
+When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
+This metric illustrates therefore the conservation of clades among the compared trees (i.e.
+when the normalised Robinson-Foulds distance is close to 1, most of the clades in the two trees are identical).}
+%\item{On the other hand, the Triplets method is measuring the position of each taxon towards two other reference taxa \citep{critchlowthe1996}.
+It will penalise only trees where taxa get removed furthest from their original clade.
+This metric illustrates therefore the amount of wildcard taxa \citep{kearneyfragmentary2002} between the to compared trees (i.e.
+when the score is closed to 1, view wildcard taxa are present in the tree).}
+%\end{enumerate}
+%Therefore, in our study, both metrics are showing two different aspects of tree topology.
+Also, disregarding their fundamental differences in accounting for topological distance, \citet{kuhnerpractical2014} have recently proposed that the Ronbinson-Foulds distance greatly outperforms the Triples distance in signalling differences in topology.
+
+%§4
+%Discussing the results 1 in detail
+%   -Why is Baycon better than the others?
+%Following \citet{kuhnerpractical2014}'s conclusions on the superiority of the Robinson-Foulds distance, the best method to use for recovering the correct topology in a Total Evidence framework is using the Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+In fact, regardless of the amount of missing data, this method always displays the highest normalised Robinson-Foulds score (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+When comparing this the probability of overlap in this method between the other onces, it is counter intuitive to note that the Bayesian consensus tree shows the highest dissimilarity with the Bayesian posterior trees from which the consensus tree is build (Fig.~\ref{Fig_Results-paircomp_within}-A).
+Note that it is also highly dissimilar to the Bootstraps.
+However that can be linked to the fact that both Bootstraps and Bayesian posterior trees are measured as 125*1000 random pairwise comparisons instead of just 125 tree comparisons for the Bayesian consensus trees.
+The fact that the 1000 trees to be compared for each of the 125 pairwise comparisons can add noise to the results for these two methods due to the random factor in the pairwise comparisons (see \hyperref[tree_comparisons]{Methods - Tree comparisons}).
+
+%§5
+%Discussing the hypothesis 2 in details
+%   -Which parameters does what?
+%Which parameter does what: for $M_L$, one should put more effort in using natural museums history collections for coding the missing morphological data for living taxa if possible %go to the museum!; for the $M_F$ parameter, the amount of missing data unfortunately depends on the quality of the fossil record and can not be actively improved and depends on exceptional discoveries \citep[e.g.][]{nithe2013}; finally for the $M_C$ parameter, improvement can be done by vast collaborative projects in order to gather as much characters as possible \citep[e.g.][]{O'Leary08022013}.
+Therefore, we advocate the importance of coding morphological characters for the most living taxa possible and with the most characters possible, potentially by using collaborative projects portals such as morphobank \citep{morphobank}.
+%   -Why is ML more important than others?
+%Our results are consistent with previous works and are showing that the amount of missing data \textit{per se} does not reduce the ability to recover the correct topology \citep[e.g.]{kearneyfragmentary2002,wiensmissing2003,pattinsonphylogeny2014}.
+In fact, as shown previously, the amount of overlap in the matrix is more crucial \citep[e.g.]{rouresite-specific2011}.
+In our case, the overlap in data is greatly reduced when different parameters are combined with a high amount of missing data (e.g.
+$M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+In his scenario the overall number of missing characters $M_{C}$ affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
+However it is important to note that in the pairwise Bhattacharyya Coefficient analysis, the area with the lowest probability of distribution overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (Right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
+
+%§6
+%Putting results into context?
+%   -Simulating morpho characters
+%The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
+% NC: Caveats later on For morphological characters, the underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
 Also morphological characters studies involve many potential statistical pitfalls (e.g.
-independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy.
-%REF probably April has something about it
-\begin{enumerate}
-\item{First, morphological data can display a different signal than molecular data, especially in small matrices. %REF
-This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
+independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy \citep{wrightbayesian2014}.
+%\begin{enumerate}
+%\item{First, morphological data can display a different signal than molecular data, especially in small matrices.
+%REF This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
 However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013}.}
-%The main difference between molecular and morphological characters is that, because of historical and practical reasons, a morphological character matrix do no contains autapomorphies and invariant characters which are usually more common than synampomorphies in molecular character matrices (cite).
-\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013}.
+%\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013}.
 However, following our simulation assumption of random character distributions, if they accumulate through time in the same way as the majority of the molecular characters then, homoplasic characters are expected to appear randomly through time \citep{davalosintegrating2014}.
@@
 After a reaching a critical amount of morphological characters, adding new ones increases homoplasy \citep{wagner2000}.}
-\end{enumerate}
-Our simulation parameters both decrease the phylogenetic signal (difficulty to simulate morphological characters and size of the matrix) as well as it increases it (reduction of homoplasy in the morphological part of the matrix).
-Therefore, these drawbacks seems to have only a minor impact on the main results of this study: the incapacity of recovering any topology in Bayesian inference.
+%\end{enumerate}
+%   -April's work (parametric models outpeforms easiest ones)
+%   -Pattinson's work (missing data in fossils is grant)
 
-\subsection{Comparing topologies} % NC: Is any of this really relevant to the central questions of the paper?
-I don't think so...
-Maybe some of this could go in the methods where you chose these metrics.
-But it may not even be needed there.
-Comparing topologies is a crucial question in phylogenetics but has always been hard to normalise because of the vast amount of different metrics used as proxies for different aspects of tree similarity/dissimilarity \citep{agapowpower2002}.
-Because our global framework is to study how to include efficiently fossils into phylogenies, we chose metrics reflecting the most interesting aspect of this global question: where do individual taxa (i.e.
-the fossils) branch in the tree.
-The Robinson-Foulds \citep{RF1981} and the Triplets \citep{critchlowthe1996} metrics are more sensible to taxa and clade placement than other tree comparison metrics \citep[e.g.][Imbalance metric]{kirkpatricksearching1993} and where therefore favored.
-Also by using the Normalized Tree Similarity index \citep{Bogdanowicz2012} we emphasize the aspect of "good" phylogenetic signal \textit{versus} random phylogenetic signal because it normalized the values of the metric by correcting for the expected value when comparing random trees (NTS = 0).
-
-The Robinson-Foulds metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
-When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
-On the other hand, the Triplets method is measuring the position of each taxon towards to other reference taxa \citep{critchlowthe1996}.
-It will penalise only trees where taxa get removed furthest from their original clade.
-Regarding our problematic (how does missing data influence topological recovery in trees containing both living and fossil taxa) we are more interested the placement of taxa (i.e.
-where does the fossil branch) than the exact conservation of clades.
-
-%Need a transition!
-
-Although the idea of comparing two single trees is straightforward, it becomes more complex when comparing trees distributions (e.g. Bayesian posterior distributions).
-The introduction of our Random Pairwise Bayesian Tree Comparison methods allows to summarize the comparison of two trees distributions by picking up the most frequent signal in the distribution of the pairwise comparisons (i.e.
-the mode).
-Even though this method is subject to randomness and might artificially increase or decrease the score of the studied metric, simulations showed that when a sufficient amount of random comparisons is performed, this method doesn't seem to be subject to randomness (e.g.
-1000 random pairwise comparisons - see ~\ref{supplementaries}).
-
-\subsection{Maximum Likelihood versus Bayesian}
-The main results of our analysis shows that Bayesian inference fails to recover any Topology in a Total Evidence method framework (Fig. ~\ref{Fig_Results}).
-This results is surprising regarding the behaviour of Maximum Likelihood inference (Fig.
-~\ref{Fig_Results}) as well as the Bayesian inferences performed on empirical data \citep{ronquista2012,schragocombining2013}.
-However, it is important to note that this effect was not mentioned in the aforementioned empirical tests because both studies used a Bayesian approach with fixed topology \citep{ronquista2012,schragocombining2013}.
-
-In our case, we suspect that this inability of Bayesian methods to recover topology is due to the intrinsic structure of a Total Evidence matrix.
-In fact, as previously studies have shown, missing data doesn't seem to be a major drawback as long as the missing data is randomly distributed ~\citep[e.g.][]{wiensmissing2003,rouresite-specific2011,sansomfossilization2013}.
-However, in Total Evidence matrices, the majority of the missing data is not randomly distributed but concentrated in the molecular part of the matrix for the fossil taxa (i.e.
-the missing-data sub-matrix).
-This leads to high decrease of topological recovery when using non optimal criterion approach (i.e.
-Bayesian inference) because of the high variance in the near-likeliest solutions sampled in the Bayesian posterior distribution.
-In opposition, when applying optimal criterion approach (i.e.
-Maximum Likelihood), it is still possible to sample the likeliest tree.
-%probably need to develop that way more, especially with actually analysing the likelihood landscape.
-
-\subsection{Effect of missing data}
-The three parameters we selected in this study account for three potential pitfalls in collecting the data for Total Evidence analysis:
-$M_L$ represents the living taxa for which there is no available morphological data,
-$M_F$ represents the quality of the fossil record,
-and $M_C$ represents the general coding effort and the overall of morphological data available.
-Ideally, the lowest possible amount of missing data is wanted in any phylogenetic analysis leading to a data collection part prior to the phylogenetic analysis.
-Each of the three parameters can be improved in a different way prior to the analysis:
-for $M_L$, one should put more effort in using natural museums history collections for coding the missing morphological data for living taxa if possible %go to the museum!;
-for the $M_F$ parameter, the amount of missing data unfortunately depends on the quality of the fossil record and can not be actively improved and depends on exceptional discoveries \citep[e.g.][]{nithe2013};
-finally for the $M_C$ parameter, improvement can be done by vast collaborative projects in order to gather as much characters as possible \citep[e.g.][]{O'Leary08022013}.
-
-Our results shows that the $M_F$ parameter have less influence on recovering the good tree topology in Maximum Likelihood framework which is fortunate since it is the parameter that is the more difficult to fix for practical reasons.
-Regarding the fact that we are more interested in taxa placement than in clade conservation (Triplets \textit{versus} Robinson-Fould), the parameter that affects the more the decrease in topological recovery is the number of missing living taxa in the morphological part of the matrix ($M_L$) and the overall number of morphological characters $M_C$.
-This makes sense since the living taxa are bearing both the information to build the tree backbone (the molecular data) and the information used for branching the fossil on this backbone (the morphological information).
-Therefore, we advocate the importance of coding morphological characters for the most living taxa possible and with the most characters possible, potentially by using collaborative projects portals such as morphobank \citep{morphobank}.
-
-%Potential reviewers comments?
-%What about it when you vary the proportion of fossil taxa?
+%§7
+%Putting results into context (back to the intro?)
+%   -No problems with using TEM and missing data
+%   -Can be improved by collecting more data.
 
@@
 
-\section{Conclusion}
-A Bayesian approach fails to recover accurate topology in a Total Evidence approach framework, whatever the amount of missing data.
-We think this failure is due to the intrinsic structure of the Total Evidence matrices that includes a vast amount of non randomly distributed missing data (i.e.
-the molecular part of the matrix for the fossil taxa).
-This missing data is equal to the number of fossil taxa $\times$ the number of molecular characters leading to a vast amount of trees to be sampled in Bayesian posterior distribution.
-However, one can use a Maximum Likelihood approach to fix the likeliest topology.
-If so, an effort should be made prior to the phylogenetic analysis on collecting as much morphological data as their is available from living taxa in order to efficiently improve the quality of the trees.
+\section{Conclusion} %Can probably just be the last paragraph but I like it to be separated as the take home message
 
-%Perspectives:
-%What the effect of this missing molecular data on branch length estimation?
+%§Conclusions: what to do
+%   -Use bayesian consensus
+%   -Go code living taxa
 
@@
 %---------------------------------------------
+\section{Supplementary material}
+Supplementary material (code, analysis and full results) can be found in the Dryad data repository at http://dx.doi.org/10.5061/dryad.XXXX. %section and sentence format asked by Syst. Biol.
 
@@
 This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
+
+%cite cheat sheet
+ % The \cite command functions as follows:
+ %   \citet{key} ==>>                Jones et al. (1990)
+ %   \citet*{key} ==>>               Jones, Baker, and Smith (1990)
+ %   \citep{key} ==>>                (Jones et al., 1990)
+ %   \citep*{key} ==>>               (Jones, Baker, and Smith, 1990)
+ %   \citep[chap. 2]{key} ==>>       (Jones et al., 1990, chap. 2)
+ %   \citep[e.g.][]{key} ==>>        (e.g. Jones et al., 1990)
+ %   \citep[e.g.][p. 32]{key} ==>>   (e.g. Jones et al., p. 32)
+ %   \citeauthor{key} ==>>           Jones et al.
+ %   \citeauthor*{key} ==>>          Jones, Baker, and Smith
+ %   \citeyear{key} ==>>             1990
 

### 2014-12-02 Manuscript/TEM_draft.tex
Updated manuscript and ref
reframed discussion + minor changes in intro/methods/results

@@
 This fact has led to an increasing consensus among scientists that both living and fossil taxa must be included in macroevolutionary studies.
-One approach, the Total Evidence approach, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both living and fossil taxa at the tips.
+One approach called Total Evidence, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both living and fossil taxa at the tips.
 Although the Total Evidence approach seems very promising, it requires a lot of data and is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.
@@
 
-We find that, when using a clade conservative metric such as Robinson-Foulds distance, Bayesian method recovers the right topology better than Maximum Likelihood method.
-However, when using triplets method, there is less significant difference between both methods.
-
--Overall missing data is doesn't affect topology in Bayesian (around minimum 70\% of topological topology in the worth scenario).
-
--However, one essential way to improve it would be to code missing living taxa.
+We find that, when using a clade conservative metric such as Robinson-Foulds distance, Bayesian consensus trees recovers the right topology better than Maximum Likelihood or than a subset of Bayesian posterior tree distributions and this regardless of the amount of missing data (minimal tree similarity of 0.69 in the worst scenario).
+Regarding the data, good topological recovery is not related to the amount of missing data \textit{per se} but to the amount of data overlap.
+The two factors that influences the most this data overlap are the parameters 1 (proportion of coded living taxa) and 3 (number of morphological characters).
 \end{abstract}
@@
 This has led to increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments (e.g.
-\citealp{pyrondivergence2011,ronquista2012,schragocombining2013}).
+However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013}.
 %Not a nice sentence
@@
 Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \citep{simpson1945}.
-This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty (but see \citealp{spencerefficacy2013}).
+This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty.
 Neontologists, on the other hand, more commonly use probabilistic approaches (e.g.
@@
 Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence date are used to time calibrate phylogenies \citep{zuckerkandl1965}.
-There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013}.
+There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
 However neither approach uses all the available data.
@@
 This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
-Total Evidence methods have been successfully applied to empirical data \citep{pyrondivergence2011,ronquista2012,schragocombining2013}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
+Total Evidence methods have been successfully applied to empirical data \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
 However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data that can be difficult (or impossible) to collect.
 The morphological data for living taxa is rarely collected when molecular data is available (e.g.
-\citealt{O'Leary08022013} vs.
-\citealt{meredithimpacts2011}), and, for fossil taxa, the scarcity of the fossil record only allow to collect the data available (for example, in vertebrates, the hardest parts of the skeleton; \citealt{sansomfossilization2013}).
+\citealt{O'Leary08022013} \textit{vs.} \citealt{meredithimpacts2011}), and, for fossil taxa, the scarcity of the fossil record only allow to collect the data available (for example, in vertebrates, the hardest parts of the skeleton; \citealt{sansomfossilization2013}).
 Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
 
-The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,wrightbayesian2014}.
+The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,sansombias2014}.
 Missing molecular data has been seen by some authors as an issue because it can, in some part of the tree, decrease phylogenetic signal (i.e.
-the evolutionary information contained within the matrix allowing to infer topology and branch length)% NC: define phylogenetic signal here.
-    , especially when using large matrices \citep{lemmonthe2009}. However, this may not be a major issue because phylogenetic signal is easily increased by:
-(i) including a "modest" number of highly-covered genes (i.e.
-approximatively of the genes that are coded for most of the species; \citealt{rouresite-specific2011}) %NC: what do you mean by highly-covered?
-(ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and (iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
+the evolutionary information contained within the matrix allowing to infer topology and branch length), especially when using large matrices \citep{lemmonthe2009}.
+However, this may not be a major issue because phylogenetic signal is easily increased by: (i) including a "modest" number (~50) of highly-covered genes (i.e.
+a number of genes that are available for all taxa ; \citealt{rouresite-specific2011}); (ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and (iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
 Similarly, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study in question \citep{wiensmissing2003,sansomfossilization2013,pattinsonphylogeny2014}.
 Because soft-tissue characters are rarely preserved in the fossil record, missing data is mainly found in these characters, and is therefore not randomly distributed which can lead to biased placement of fossil taxa in phylogenies (e.g.
-\citeauthor{sansomfossilization2013} \citeyear{sansomfossilization2013} but see \citeauthor{pattinsonphylogeny2014} \citeyear{pattinsonphylogeny2014}).
-However, the phylogenetic signal is not related to the amount of missing data \textit{per se} but to the number of informative characters for each taxon, therefore missing data is less of an issue than the number of shared informative characters \citep{wiensmissing2003}.
+\citealt{sansomfossilization2013} but see \citealt{pattinsonphylogeny2014}).
+However, the phylogenetic signal is not related to the amount of missing data \textit{per se} but to the number of informative characters for each taxon, therefore missing data is less of an issue than the number of shared informative characters \citep{kearneyfragmentary2002,wiensmissing2003,pattinsonphylogeny2014}.
 % NC: I'm a little dubious of what this paragraph is trying to get at in places.
@@
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil species.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil species.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data.
 Until now, no attempt has been made to study the impact of this issue on phylogenetic inference from Total Evidence methods.
-
-%ADD APRIL'S PAPER SOMEWHERE
-
 % NC: Somewhere here you need to flag that you're only looking at topologies and briefly explain why.
-In this study, we focused only on topology as one of the two aspects of the phylogenetic signal (topology and branch length).
-Even though both aspects are equally important, branch topology is the first and most straightforward aspect reflecting phylogenetic signal (i.e.
+In this study, we are interested in assessing the effect of missing data on the ability to recover a correct topology.
+Even though branch length is a second aspect of phylogenetic inferences equally important, topology is the first and most straightforward aspect reflecting phylogenetic signal (i.e.
 topological changes are discrete opposed to branch length changes are continuous).
 Also, interestingly, the effect of Total Evidence method has not been formally assessed in previous studies using fixed topology \citep{ronquista2012,schragocombining2013}.
-%TG: or is that mean for them?
 
-Here we use simulations to assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
-The molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}.
-The effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}, therefore, we focus only on missing data in the morphological part of the matrix.
+Here we use a simulation approach to assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
+Since the molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}, the effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
+Therefore, we focus only on missing data in the morphological part of the matrix.
 We investigate three major parameters that directly affect the completeness of the morphological part of the matrix:
@@
 The advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
-Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them.
+Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
 
-We find that when using a Maximum Likelihood approach, as missing data increases, the likelihood of recovering the correct tree topology decreases.
-However, even with no missing data, Total Evidence matrices dramatically reduce the performance of Bayesian methods for inferring tree topology.
-We propose that this drastic difference between Bayesian and Maximum Likelihood methods is due to a flattening of the likelihood landscape caused by the unavoidable amount of missing molecular data for fossil taxa in a Total Evidence matrix.
-We make suggestions for how best to deal with this issue when inferring phylogenies from Total Evidence matrices.
-% NC: You don't want to tell the whole story here. Just a results summary and a quick line of interpretation
+We find that the ability of recovering the correct topology can mainly be increased when using the Bayesian methods.
+Also minimizing the proportion of living taxa with no morphological data and the proportion of missing morphological characters improves the ability to recover the correct more than minimizing the proportion of missing data in the fossil record.
 
@@
 
-% NC: Most of the below was fine but I've tried to condense it a bit as it's currently long even for Syst Biol. The key is to explain everything, but with minimal repetition.
-
 \subsection{Generating the matrix}
@@
 \begin{enumerate}
-% NC: I'm not going to check but make sure the brief explanations of the parameters here are the same as in the intro and outline above.
-Also look at my justifications in the intro too.
-I think this may need a bit more justification as it's the main part of your simulations.
 \item{The proportion of living taxa with no morphological data ($M_{L}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
@@
 towards 0) means that the topological difference between the two trees is as much as expected when comparing two random trees.
+Note that both tree comparisons are representing different aspects of topological similarity and do not perform equaly (see \hyperref[metrics_discussion]{discussion below} and \citealt{kuhnerpractical2014}).
 
@@
 
-As it would be expected form the literature, the amount of missing data in the morphological matrix does decrease the ability to recover the right topology, regardless of the parameter, the method or the metric used \citep{rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,wrightbayesian2014}.
+As it would be expected form the literature, the amount of missing data in the morphological matrix does decrease the ability to recover the right topology, regardless of the parameter, the method or the metric used \citep[e.g.][]{rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014}.
 However each variable does not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam}).
@@
 However, regarding the Triplets distance, Bayesian consensus trees seems to perform poorly but with great confidence intervals overlaps (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
+These greater larger confidence intervals are due to the lower ability of the Triplet method to highlight precise differences in topology \citep{kuhnerpractical2014}.
 Results for the global trend comparison between the uncertainty methods (Bootstrap and Bayesian posterior tree distribution) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
@@
 When regarding at the Triplets distance, the two method that have the most distribution are the Bootstraps and the Bayesian posterior trees (Fig.~\ref{Fig_Results-paircomp_without}-B).
+These results are due to the slight difference in comparing topologies between Maximum Likelihood and Bayesian consensus trees and Bootstraps and Bayesian posterior trees: the two last ones are based on a random pairwise comparison of a sub-sample of 1000 trees.
+This is likely to add noise to the data.
 
@@
 % NC: Discussions usually begin with a paragraph that sums up the results.
-Then you can move on to discussing caveats, problems, extentions etc.
+Then you can move on to discussing caveats, problems, extensions etc.
 You also need to show how your work fits in with the wider literature.
@@
 %   -Missing data increase decreases topological recovery
-%The results of our simulations are in adequacy with recent analysis on the effect of missing data in morphological matrices \citep{rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,wrightbayesian2014}.
+The results of our simulations are in adequacy with recent analysis on the effect of missing data in morphological matrices \citep{pattinsonphylogeny2014,wrightbayesian2014,sansombias2014}.
 Our results are showing that ability to recover the correct topology decreases with the amount of missing data, regardless the metric, the method or the parameters considered.
-In more details, our results shows however a clear superiority of the Bayesian consensus tree over other methods, at least when considering the Robinson-Foulds distance (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-Also, our results suggest that the missing data for missing taxa parameter ($M_{L}$) shows a slightly higher decrease in the ability to recover the correct topology (Fig.~\ref{Fig_Results-paircomp_within}).
 %Results 1
 %   -However, methods and metrics are unequal
+However, the effect of missing data on topological recovery varies with different methods to infer topology and different metrics to compare topologies.
 %Results 2
 %   -Additionally, parameters have different effect
+Additionally, within a method and a metric, the parameters used to remove data have different effect on the ability to recover the correct topology.
 %Wrap up results
 %   -Thus, Baycon is better with loads of living
+Thus, we found that using the topologies of Bayesian consensus trees performed consistently better than other methods for recovering the correct topology (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}) and that both the amount of available characters ($M_{C}$) and the amount of data for the living taxa ($M_{L}$) are crucial missing data parameters.
 
@@
 %   -Missing data per se is not a problem, problem is in the amount of overlap
+As proposed in numerous previous studies, when recovering topology, the amount of missing data \textit{per se} is not a problem as long as enough data overlaps in the matrix \citep[e.g.][]{rouresite-specific2011,pattinsonphylogeny2014}.
 %Proposed hypothesis
-%   -One can reduce overlap by adding living taxa
+%   -One can increase overlap by adding living taxa
+To improve topological recovery in a Total Evidence framework, one can efficiently increase the amount of overlapping data by sampling a maximum of characters and by increasing the number of living data coded in the morphological part of the matrix.
 
 %§3
-%Discussing the metrics in detail
-%   -RF is different than triplets and better.
-%However, our results can be contrasted whether one looks at the Robinson-Foulds or the Triplets distance.
-When using the second one, the confidence interval of our different parameters combinations per methods seems higher (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}) and the probability of overlap between the distribution of each parameter is high (Bhattacharyya Coefficient close to 1 - Fig.~\ref{Fig_Results-paircomp_within}-B).
-The contrast between the results of the two metrics can be explained by the actual nature of the metrics:
-%\begin{enumerate}
-%\item{The Robinson-Foulds metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
-When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
-This metric illustrates therefore the conservation of clades among the compared trees (i.e.
-when the normalised Robinson-Foulds distance is close to 1, most of the clades in the two trees are identical).}
-%\item{On the other hand, the Triplets method is measuring the position of each taxon towards two other reference taxa \citep{critchlowthe1996}.
-It will penalise only trees where taxa get removed furthest from their original clade.
-This metric illustrates therefore the amount of wildcard taxa \citep{kearneyfragmentary2002} between the to compared trees (i.e.
-when the score is closed to 1, view wildcard taxa are present in the tree).}
-%\end{enumerate}
-%Therefore, in our study, both metrics are showing two different aspects of tree topology.
-Also, disregarding their fundamental differences in accounting for topological distance, \citet{kuhnerpractical2014} have recently proposed that the Ronbinson-Foulds distance greatly outperforms the Triples distance in signalling differences in topology.
+%Discussing the hypothesis 2 in details
+%   -Which parameters does what?
+For each missing data parameter used in this study, there is a different way to reduce the amount of missing data:
+\begin{enumerate}
+\item{$M_{L}$:}
+One should put more effort in using natural museums history collections for coding the missing morphological data for living taxa if possible
+\item{$M_{F}$:}
+the amount of missing data unfortunately depends on the quality of the fossil record and can not be actively improved and depends on exceptional discoveries \citep[e.g.][]{nithe2013}
+\item{$M_{C}$:}
+improvement can be done by vast collaborative projects in order to gather as much characters as possible \citep[e.g.][]{O'Leary08022013}.
+Therefore, we advocate the importance of coding morphological characters for the most living taxa possible and with the most characters possible, potentially by using collaborative projects portals such as morphobank \citep{morphobank}.
+\end{enumerate}
+%   -Why is ML more important than others?
+Our results are consistent with previous works and are showing that the amount of missing data \textit{per se} does not reduce the ability to recover the correct topology \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,pattinsonphylogeny2014}.
+In fact, as shown previously, the amount of overlap in the matrix is more crucial \citep[e.g.][]{rouresite-specific2011}.
+In our case, the overlap in data is greatly reduced when different parameters are combined with a high amount of missing data (e.g.
+$M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+In his scenario the overall number of missing characters $M_{C}$ affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
+However it is important to note that in the pairwise Bhattacharyya Coefficient analysis, the area with the lowest probability of distribution overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (Right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
 
@@
 %   -Why is Baycon better than the others?
-%Following \citet{kuhnerpractical2014}'s conclusions on the superiority of the Robinson-Foulds distance, the best method to use for recovering the correct topology in a Total Evidence framework is using the Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+Following \citet{kuhnerpractical2014}'s conclusions on the superiority of the Robinson-Foulds distance, the best method to use for recovering the correct topology in a Total Evidence framework is using the Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 In fact, regardless of the amount of missing data, this method always displays the highest normalised Robinson-Foulds score (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
@@
 %§5
-%Discussing the hypothesis 2 in details
-%   -Which parameters does what?
-%Which parameter does what: for $M_L$, one should put more effort in using natural museums history collections for coding the missing morphological data for living taxa if possible %go to the museum!; for the $M_F$ parameter, the amount of missing data unfortunately depends on the quality of the fossil record and can not be actively improved and depends on exceptional discoveries \citep[e.g.][]{nithe2013}; finally for the $M_C$ parameter, improvement can be done by vast collaborative projects in order to gather as much characters as possible \citep[e.g.][]{O'Leary08022013}.
-Therefore, we advocate the importance of coding morphological characters for the most living taxa possible and with the most characters possible, potentially by using collaborative projects portals such as morphobank \citep{morphobank}.
-%   -Why is ML more important than others?
-%Our results are consistent with previous works and are showing that the amount of missing data \textit{per se} does not reduce the ability to recover the correct topology \citep[e.g.]{kearneyfragmentary2002,wiensmissing2003,pattinsonphylogeny2014}.
-In fact, as shown previously, the amount of overlap in the matrix is more crucial \citep[e.g.]{rouresite-specific2011}.
-In our case, the overlap in data is greatly reduced when different parameters are combined with a high amount of missing data (e.g.
-$M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-In his scenario the overall number of missing characters $M_{C}$ affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
-However it is important to note that in the pairwise Bhattacharyya Coefficient analysis, the area with the lowest probability of distribution overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (Right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
+%Caveats: the metrics
+%   -RF is different and better than triplets.
+\label{metrics_discussion}
+Our results can be contrasted by looking at the normalised tree similarity when using the Robinson-Foulds or the Triplets distance.
+When using the second one, the confidence interval of our different parameters combinations per methods seems higher (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}) and the probability of overlap between the distribution of each parameter is high (Bhattacharyya Coefficient close to 1 - Fig.~\ref{Fig_Results-paircomp_within}-B).
+The contrast between the results of the two metrics can be explained by the actual nature of both metrics:
+\begin{enumerate}
+\item{The Robinson-Foulds metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
+When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
+This metric illustrates therefore the conservation of clades among the compared trees (i.e.
+when the normalised Robinson-Foulds distance is close to 1, most of the clades in the two trees are identical).}
+\item{On the other hand, the Triplets method is measuring the position of each taxon towards two other reference taxa \citep{critchlowthe1996}.
+It will penalise only trees where taxa get removed furthest from their original clade.
+This metric illustrates therefore the amount of wildcard taxa \citep{kearneyfragmentary2002} between the to compared trees (i.e.
+when the score is closed to 1, view wildcard taxa are present in the tree).}
+\end{enumerate}
+Therefore, in our study, both metrics are showing two different aspects of tree topology.
+Also, disregarding their fundamental differences in accounting for topological distance, \citet{kuhnerpractical2014} have recently proposed that the Ronbinson-Foulds distance greatly outperforms the Triples distance in signalling differences in topology.
+%Maybe also a note about the fact that in Triplets, there's more than one solution each simulation (i.e.
+wildcard taxa can fix in many places) so it increases highly the variance in the NTS.
+In RF, because the metric is more conservative (bounded between 0 and the number of nodes), the number of solutions is smaller and therefore the variance decreases.
 
 %§6
-%Putting results into context?
+%Caveats: the simulations
 %   -Simulating morpho characters
-%The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
-% NC: Caveats later on For morphological characters, the underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
+The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
+For morphological characters, the underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
 Also morphological characters studies involve many potential statistical pitfalls (e.g.
 independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy \citep{wrightbayesian2014}.
-%\begin{enumerate}
-%\item{First, morphological data can display a different signal than molecular data, especially in small matrices.
-%REF This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
+\begin{enumerate}
+\item{First, morphological data can display a different signal than molecular data, especially in small matrices. %REF
+This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
 However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013}.}
-%\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013}.
+\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013}.
 However, following our simulation assumption of random character distributions, if they accumulate through time in the same way as the majority of the molecular characters then, homoplasic characters are expected to appear randomly through time \citep{davalosintegrating2014}.
@@
 After a reaching a critical amount of morphological characters, adding new ones increases homoplasy \citep{wagner2000}.}
-%\end{enumerate}
-%   -April's work (parametric models outpeforms easiest ones)
-%   -Pattinson's work (missing data in fossils is grant)
+\end{enumerate}
 
 %§7
-%Putting results into context (back to the intro?)
-%   -No problems with using TEM and missing data
-%   -Can be improved by collecting more data.
+%Further directions - is that necessary though?
+%   -Using real data
+One way to address this issue would be to run a similar analysis using real data.
+The amount of studies using Total Evidence data type matrices is steadily increasing \citep[e.g.][]{ronquista2012,slaterphylogenetic2013,beckancient2014} and one could analysis these total evidence matrices in the same framework as this study.
+However, the true topology (obtained from a matrix with no missing data) is unlikely to be known.
+A second way to address this issue, and especially to incorporate the variation of signal in morphological character would be to simulate morphological at different rates \citep{wrightbayesian2014} or to apply patterns of missing data directly obtained from the fossil record \citep{pattinsonphylogeny2014}.
+Nevertheless, our simulation, by simplifying an ideal world, already show clear results of the effect of missing data on topology in a Total Evidence framework.
+Also, our results show how recovering the correct topology can be improved by using Bayesian consensus trees in combination with a maximum amount of characters and a of living taxa with morphological data.
+%   -Looking at branch length?
+
+%§8
+%Back to the importance of TEM
+%   -Data is not a major problem
+%   -Good and promising method.
 
@@
 
-%§Conclusions: what to do
+%§Conclusions
+%   -TEM works with missing data
 %   -Use bayesian consensus
 %   -Go code living taxa
+Using Total Evidence matrices with missing data is not a problem for recovering the correct topology as long as enough morphological data is overlapping between the fossil and living taxa.
+In order to optimize good topological recovery, with propose to use the Bayesian consensus tree over the Maximum Likelihood tree or a subset of the Bayesian posterior trees distribution.
+Also, in order to improve the overlap in data in the morphological part of the matrix, we advice to use all the morphological characters available and to code them for a maximum of living taxa present in the matrix.
 
@@
 \section{Acknowledgements}
-Thanks to Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson and Andrew Jackson, Gavin Thomas, April Wright and the members of the Macro Journal Club %Also Gavin Thomas if not authors - %TG : also Adam Kev and Sive or that's putting to many people, thinking about it, they helped at least as much as Andrew (not mentioning Trevor but that's more political)?
+Thanks to Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson and Andrew Jackson, Gavin Thomas, April Wright and the members of the Macro Journal Club %TG : also Adam, Kev and Sive or that's putting to many people, thinking about it, they helped at least as much as Andrew (not mentioning Trevor but that's more political)?
 %Not sure how many people we have to thank but it could also include Graham Slater and Nick Matzke.
@@
 
-\section{Supplementary Material}
+%Ignored for fast pdf building
+%\section{Supplementary Material}
 
-    \input{SupplementaryMaterial/SupplementaryMaterial}
-    \label{SupplementaryMaterial}
+%    \input{SupplementaryMaterial/SupplementaryMaterial}
+%    \label{SupplementaryMaterial}
 

### 2014-12-03 Manuscript/TEM_draft.tex
Discussion (results)

@@
 \subsection{Removing data}
+\label{Removing_data}
 We modified the "complete" matrix to get matrices with missing data by randomly replacing data with "?" in the morphological part of the matrices according to the following parameters:% (Fig.
@@
 %Results 1
-%   -However, methods and metrics are unequal
-However, the effect of missing data on topological recovery varies with different methods to infer topology and different metrics to compare topologies.
+%   -However, parameters have different effect
+However, within a method and a metric, the parameters used to remove data have different effect on the ability to recover the correct topology.
 %Results 2
-%   -Additionally, parameters have different effect
-Additionally, within a method and a metric, the parameters used to remove data have different effect on the ability to recover the correct topology.
+%   -Additionally, methods and metrics are unequal
+Additionally, the effect of missing data on topological recovery varies with different methods to infer topology and different metrics to compare topologies.
 %Wrap up results
 %   -Thus, Baycon is better with loads of living 
-Thus, we found that using the topologies of Bayesian consensus trees performed consistently better than other methods for recovering the correct topology (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}) and that both the amount of available characters ($M_{C}$) and the amount of data for the living taxa ($M_{L}$) are crucial missing data parameters.
+Thus, we found that using the topologies of Bayesian consensus trees performed consistently better than other methods for recovering the correct topology (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}) and that both the amount of available characters ($M_{C}$) and the amount of data for the living taxa ($M_{L}$) are crucial missing data parameters (Fig.~\ref{Fig_Results-paircomp_within}).
 
 %§2
-%Results from litterature
+%Results 1 from litterature
 %   -Missing data per se is not a problem, problem is in the amount of overlap
-As proposed in numerous previous studies, when recovering topology, the amount of missing data \textit{per se} is not a problem as long as enough data overlaps in the matrix \citep[e.g.][]{rouresite-specific2011,pattinsonphylogeny2014}.
-%Proposed hypothesis
+As proposed in numerous previous studies, when recovering topology, the amount of missing data \textit{per se} is not a problem as long as enough data overlaps in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
+In fact, as shown previously, the amount of overlap in the matrix is more crucial than the amount of missing data \citep[e.g.][]{rouresite-specific2011}.
+%Proposed hypothesis for result 1
 %   -One can increase overlap by adding living taxa
-To improve topological recovery in a Total Evidence framework, one can efficiently increase the amount of overlapping data by sampling a maximum of characters and by increasing the number of living data coded in the morphological part of the matrix.
+To improve topological recovery in a Total Evidence framework, one can efficiently increase the amount of overlapping data reducing the amount of missing data in each of our three parameters.
+In our simulations, the overlap in data is greatly reduced when different parameters are combined with a high amount of missing data (e.g.
+$M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+In this scenario the overall number of missing characters $M_{C}$ affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
+However it is important to note that in the pairwise Bhattacharyya Coefficient analysis, the area with the lowest probability of distribution overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (Right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
 
 %§3
-%Discussing the hypothesis 2 in details
+%Discussing results 1
+Since each of our parameters are illustrating one potential sources of missing data in a morphological matrix (see \hyperref[Removing_data]{Methods}), we can propose different practical ways to reduce the amount of missing data:
 %   -Which parameters does what?
-For each missing data parameter used in this study, there is a different way to reduce the amount of missing data:
 \begin{enumerate}
-\item{$M_{L}$:}
-One should put more effort in using natural museums history collections for coding the missing morphological data for living taxa if possible
-\item{$M_{F}$:}
-the amount of missing data unfortunately depends on the quality of the fossil record and can not be actively improved and depends on exceptional discoveries \citep[e.g.][]{nithe2013}
-\item{$M_{C}$:}
-improvement can be done by vast collaborative projects in order to gather as much characters as possible \citep[e.g.][]{O'Leary08022013}.
-Therefore, we advocate the importance of coding morphological characters for the most living taxa possible and with the most characters possible, potentially by using collaborative projects portals such as morphobank \citep{morphobank}.
+\item{$M_{L}$ (non coded living taxa):}
+The vast majority of known living species have specimens preserved in natural history museums.
+This is a significant resource and more research effort is needed to code morphological characters in these specimens.
+\item{$M_{F}$ (quality of the fossil record):}
+The amount of missing data in the fossil record solely depends on the preservation quality of the fossils and can not be actively improved.
+However, gaps in the matrix can be filled with the discoveries exceptional preserved fossil \citep[e.g.][]{nithe2013}.
+Fortunately, this parameter is the most difficult to improve but it also has the least influence on whether the simulations recover the correct topology (Fig.~\ref{Fig_Results-paircomp_within}).
+\item{$M_{C}$ (amount of coded characters):}
+The amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
+Improvement in both decrease the time taken to build the matrix and to increase the number of characters can be done through vast collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}.
 \end{enumerate}
-%   -Why is ML more important than others?
-Our results are consistent with previous works and are showing that the amount of missing data \textit{per se} does not reduce the ability to recover the correct topology \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,pattinsonphylogeny2014}.
-In fact, as shown previously, the amount of overlap in the matrix is more crucial \citep[e.g.][]{rouresite-specific2011}.
-In our case, the overlap in data is greatly reduced when different parameters are combined with a high amount of missing data (e.g.
-$M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-In his scenario the overall number of missing characters $M_{C}$ affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
-However it is important to note that in the pairwise Bhattacharyya Coefficient analysis, the area with the lowest probability of distribution overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (Right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
 
 %§4
-%Discussing the results 1 in detail
+%Result 2 from literature.
+%   -Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
+Nonetheless, as our results shows in Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}, the variation in the ability to recover the correct topology does not only depend on the missing data in morphological part of the matrix but also heavily on the method used to infer the topology.
+Previous studies have shown some superiority on methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014} but see \citealt{spencerefficacy2013}).
+However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} between Maximum Likelihood and Bayesian inference in a Total Evidence framework.
+%TG: I do have to check the Mk in Baysian vs Mk in ML part though but I guess it's a good selling point: we ARE the first ones to do a sensitivity analysis in a TEM framework (as far as we know).
+%Discussing results 2
 %   -Why is Baycon better than the others?
-Following \citet{kuhnerpractical2014}'s conclusions on the superiority of the Robinson-Foulds distance, the best method to use for recovering the correct topology in a Total Evidence framework is using the Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-In fact, regardless of the amount of missing data, this method always displays the highest normalised Robinson-Foulds score (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-When comparing this the probability of overlap in this method between the other onces, it is counter intuitive to note that the Bayesian consensus tree shows the highest dissimilarity with the Bayesian posterior trees from which the consensus tree is build (Fig.~\ref{Fig_Results-paircomp_within}-A).
+When using the clade conservative Robinson-Foulds distance (see \citealt{kuhnerpractical2014} and \hyperref[metrics_discussion]{discussion below}), the Bayesian consensus trees greatly outperforms the Maximum Likelihood trees in the ability to recover the correct tree topology (Fig.~\ref{Fig_Results-global_perparam}).
+In fact, regardless of the amount of missing data, this method always displays the highest normalised Robinson-Foulds score (always $>$ 0.69 - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+Also, when looking at the pairwise Bhattacharyya Coefficients within each method, the Bayesian consensus tree one display the highest probability of overlap between the results in any case (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+
+%   -Why is Baycon so different than Baytre? Since Baycon is calculated from Baytre.
+When comparing this the probability of overlap between the methods, it is counter intuitive to note that the Bayesian consensus tree shows the highest dissimilarity with the Bayesian posterior trees from which the consensus tree is build (Fig.~\ref{Fig_Results-paircomp_within}-A).
 Note that it is also highly dissimilar to the Bootstraps.
 However that can be linked to the fact that both Bootstraps and Bayesian posterior trees are measured as 125*1000 random pairwise comparisons instead of just 125 tree comparisons for the Bayesian consensus trees.
-The fact that the 1000 trees to be compared for each of the 125 pairwise comparisons can add noise to the results for these two methods due to the random factor in the pairwise comparisons (see \hyperref[tree_comparisons]{Methods - Tree comparisons}).
+The fact that the 1000 pairwise trees to be compared for each of the 125 comparisons are chosen randomly, it can add noise to the results for these two methods (see \hyperref[tree_comparisons]{methods above}).
+
+%Some sort of transition?
 
@@
 \section{Acknowledgements}
-Thanks to Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson and Andrew Jackson, Gavin Thomas, April Wright and the members of the Macro Journal Club %TG : also Adam, Kev and Sive or that's putting to many people, thinking about it, they helped at least as much as Andrew (not mentioning Trevor but that's more political)?
-%Not sure how many people we have to thank but it could also include Graham Slater and Nick Matzke.
+Thanks to Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Andrew Jackson, Gavin Thomas, April Wright and the members of the Macro Journal Club %TG : Not sure how many people we have to thank but it could also include Graham Slater, Nick Matzke and Trevor Hodkinson.
 for useful comments on our simulation protocol.
+%Thanks to Sive Finlay, Kevin Healy and Adam Kane for comments on the manuscript. %TG: or helping writing sentences does not really count.
 Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.

### 2014-12-04 Functions/TreeComparisons/plot.bhatt.coef.R
distribution plot v0.2, warn list.names

@@ -123,14 +123,14 @@ plot.bhatt.coeff<-function(matrix, col=c("blue", "red"), col.grad=c("white", "bl
 #Plot distributions of a list of Bhattacharyya Coefficients
 ##########################
 #Plots a series of kernel density estimates of Bhattacharyya Coefficients
-#v.0.1
+#v.0.2
 ##########################
 #SYNTAX :
 #<list> a list of Bhattacharyya Coefficients
 #<...>  the name of the metric of interest
 ##########################
 #----
-#guillert(at)tcd.ie - 29/11/2014
+#guillert(at)tcd.ie - 04/12/2014
 ##########################
 #Requirements:
 #-R 3
@@ -144,11 +144,14 @@ plot.bhatt.coeff.dist<-function(list, col, ...) {
     if(class(list) != 'list') {
         stop('Provided list is not a list of Bhattacharyya Coefficients.')
     }
+    if(is.null(names(list))) {
+        warning("No list names available for plotting the legend.")
+    }
 
     #list
     if(class(col) != 'character') {
         stop('col must be a vector of characters.')
-    }    
+    } 
 
 #PLOT THE DISTRIBUTION
 
@@ -174,7 +177,7 @@ plot.bhatt.coeff.dist<-function(list, col, ...) {
     }
 
     #Adding the legend
-    legend(mean(ext.coord$maxx-abs(ext.coord$minx)), max(ext.coord$maxy), names(density.list), cex=0.8, lty=1, col=col[1:length(density.list)], bty="n")
+    legend(mean(ext.coord$maxx-abs(ext.coord$minx)), max(ext.coord$maxy), names(list), cex=0.8, lty=1, col=col[1:length(density.list)], bty="n")
 
 #End
 }

### 2014-12-04 Analysis/MissingData_differences.R
Plotting the distribution of coefficients

@@ -68,8 +68,30 @@ load("../Data/R_data/pair.comp_within.Baytre.Tr.Rda")
 pairwise.comp_within<-c("pair.comp_within.MLbest.RF", "pair.comp_within.MLbest.Tr",
                         "pair.comp_within.Baycon.RF", "pair.comp_within.Baycon.Tr",
                         "pair.comp_within.MLboot.RF", "pair.comp_within.MLboot.Tr",
-                        "pair.comp_within.Baytre.RF", "pair.comp_within.Baytre.Tr")
+                        "pair.comp_within.MLboot.RF", "pair.comp_within.Baytre.Tr")
 plot.bhatt.coeff(get(pairwise.comp_within[8]), col=c("red", "green"), col.grad=c("white", "black"))
+
+#----------------
+#Plotting the distribution of coefficients per method and per metric
+#----------------
+#Plotting the coefficients distribution for RF
+Bhat.coeff.RF<-list(as.vector(pair.comp_within.MLbest.RF)[-which(is.na(as.vector(pair.comp_within.MLbest.RF)))],
+                    as.vector(pair.comp_within.Baycon.RF)[-which(is.na(as.vector(pair.comp_within.Baycon.RF)))],
+                    as.vector(pair.comp_within.MLboot.RF)[-which(is.na(as.vector(pair.comp_within.MLboot.RF)))],
+                    as.vector(pair.comp_within.Baytre.RF)[-which(is.na(as.vector(pair.comp_within.Baytre.RF)))])
+names(Bhat.coeff.RF)<-c("Maximum Likelihood - RF", "Bayesian consensus - RF", "Bootstraps - RF", "Bayesian posterior trees - RF")
+op<-par(bty="l")
+plot.bhatt.coeff.dist(Bhat.coeff.RF, col=palette(), xlab="Bhattacharyya Coefficient", ylab="Frequency")
+
+#Plotting the coefficients distribution for Tr
+Bhat.coeff.Tr<-list(as.vector(pair.comp_within.MLbest.Tr)[-which(is.na(as.vector(pair.comp_within.MLbest.Tr)))],
+                    as.vector(pair.comp_within.Baycon.Tr)[-which(is.na(as.vector(pair.comp_within.Baycon.Tr)))],
+                    as.vector(pair.comp_within.MLboot.Tr)[-which(is.na(as.vector(pair.comp_within.MLboot.Tr)))],
[…]

### 2014-12-04 Manuscript/SupplementaryMaterial/Supp_results.tex
Added BC explanations in the supplementaries

@@
 \label{Fig_global_BootTreesets} %Differences between all the parameters and between two methods (Boot vs treesets)
+\end{figure}
+
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_permeth.pdf}
+\caption{Distribution of the Pairwise Bhattacharyya Coefficients within each method.
+A- Distribution of the coefficients when comparing Ronbinson-Foulds distances.
+B- Distribution of the coefficients when comparing Triplets distances.}
+\label{Fig_Bhatt.coeff_distribution} %Differences in overlap within each method
 \end{figure}

### 2014-12-04 Manuscript/SupplementaryMaterial/TreeComparison.tex
Added BC explanations in the supplementaries

@@
 \section{Supplementary material Section 2}
+
+\subsection{Robinson-Foulds distance details}
+Robinson-Foulds distance (\textit{RF}; \citealp{RF1981}), or "path difference", measures the number of shared clades across two trees.
+The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} and can be expressed as following:
+\begin{equation}
+RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
+\end{equation}
+Where $C_{x,y}$ is the number of clades in common in the two trees.
+$C$ is equal to one if the two trees have the same $n$ taxa; and $C = n-2$ when non of the $n$ taxa are shared between the trees.This metric is more sensitive to taxon displacement than Triples distance (i.e.
+if one taxon moves out of a clade, then the clades are no longer considered similar; \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
+The minimal value of $C$ is equal to 1 if the two trees have the same n taxa; the maximal value in $C = n-2$.
+For a fully unresolved tree (star tree) $N$=1 and for a fully resolved tree (binary tree) $N = n-2$.
+The minimal and maximal topological distance for taxa is:
+\begin{equation}
+RF_{min} = 1 + 1 - 2C_{x,y}
+\end{equation}
+And:
+\begin{equation}
+RF_{max} = 2(n-2)-2
+\end{equation}
+One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any $n$ taxa:
+\begin{equation}
+RF.scaled_{x,y} = \frac{RF_{x,y}-RF_{max}}{RF_{max}}
+\end{equation}
+This metric is more sensitive to taxa displacement than the Triplet distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.
 
@@
 Triples distance ($T_{x,y}$; \citealp{dobson1975triplets}) measures the number of sub-trees made up of three taxa (triplets) that differ between two given trees.
-Each triplet can be written as $I_{ijk}$=(\textit{ijk}). Where $I_{ijk}$ is equal to zero if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to one.
+Each triplet can be written as $I_{ijk}$=(\textit{ijk}).
+Where $I_{ijk}$ is equal to zero if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to one.
 For any rooted binary tree there are only three possible combinations for each triplet: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
@@
 And where \textit{n} is the total number of taxa in both trees (modified from \citet{critchlowthe1996}).
-If $S_n$ = 0, the trees are identical; when $S_n$ = $\binom{n}{4}$, the trees are as different as possible (i.e. every taxon has a different placement in the two trees).
+If $S_n$ = 0, the trees are identical; when $S_n$ = $\binom{n}{4}$, the trees are as different as possible (i.e.
+every taxon has a different placement in the two trees).
 Because the possible number of triplets per clade is a finite number, the probability of two random trees with the same n taxa to have the same triplet is:
@@
 
-\subsection{Robinson-Foulds distance details}
-Robinson-Foulds distance (\textit{RF}; \citealp{RF1981}), or "path difference"
-% NC: path difference or *distance*? - TG: difference.
-, measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} and can be expressed as following:
-\begin{equation}
-RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
-\end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees. $C$ % NC: Why italics? Why not $C$?
-is one if the two trees have the same $n$ taxa; the maximal value is $C = n-2$. % NC: again why italics for these bits? 
-%NC: I don't get this. Explain more clearly. also is it important?
-This metric is more sensitive to taxon displacement than Triples distance (i.e.
-if one taxon moves out of a clade, then the clades are no longer considered similar; \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
-The minimal value of $C$ is equal to 1 if the two trees have the same n taxa;
-the maximal value in $C = n-2$.
-For a fully unresolved tree (star tree) $N$=1 and for a fully resolved tree (binary tree) $N = n-2$.
-The minimal and maximal topological distance for  taxa is:
-\begin{equation}
-RF_{min} = 1 + 1 - 2C_{x,y}
-\end{equation}
-And:
-\begin{equation}
-RF_{max} = 2(n-2)-2
-\end{equation}
-One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any $n$ taxa:
-\begin{equation}
-RF.scaled_{x,y} = \frac{RF_{x,y}-RF_{max}}{RF_{max}}
-\end{equation}
-This metric is more sensitive to taxa displacement than the Triplet distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.
-
 \subsection{Normalised Tree Similarity}
@@
 We calculated the $\bar{d}_{m,n}$\textit{(rand)} value using an approach similar to the RPCBTC (described below) by performing 1000 random pairwise comparisons using the TreeCmp java script \citep{Bogdanowicz2012}.
+
+\subsection{Bhattacharyya Coefficient}
+The Bhattacharyya Coefficient calculates the probability of overlap of two distributions \citep{Bhattacharyya}.
+When it is equal to 0, the probability of overlap of the distributions is null and when it is equal to 1, the two distributions are entirely overlapping.
+It is an elegant and easy to compute continuous measurement of probability of similiraty between two distributions.
+The coefficient is calculated as the sum of the square root of the relative counts shared in a \textit{n} number of bins among two distributions.
+\begin{equation}
+\text{Bhattacharyya Coefficient}=\sum_{i=1}^{n} \sqrt{{\sum{a_i}}\times{\sum{b_i}}}
+\end{equation}
+Where, for $a$ and $b$:
+\begin{equation}
+\sum{a_i}=\frac{\text{Number of counts in bin \textit{i} for the distribution \textit{a}}}{\text{Total number of counts for the distribution \textit{a}}}
+\end{equation}
+The precision of the Bhattacharyya Coefficient is directly related to the number of $n$ bins.
+If $n$ is low, the overlap will be overestimated and if $n$ is to high, the overlap will be underestimated.
+In this analysis, we determined the number of bins Silverman's rule of thumb which determines the bins to be equal to 0.9 times the minimum of the standard deviation and the interquartile range divided by 1.34 times the sample size to the negative one-fifth power (Silverman (1986, page 48, eqn (3.31))).
+
+%Supplementary ref: Silverman, B. W. (1986) Density Estimation. London: Chapman and Hall.

### 2014-12-04 Manuscript/TEM_draft.tex
Added BC explanations in the supplementaries

@@
 \noindent{\Large \bf Effect of missing data on topological inference using a total evidence approach}
+%Or do we put results in the title: Bayesian analysis and increased sampling of living taxa improves topological inference when using a total evidence approach
 \bigskip
@@
 In order to investigate the effect of the parameters and/or the methods used in this simulations, we measured the similarity among the different distributions using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 \citep{Bhattacharyya}.
+The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}).
 For each method (Maximym likelihood trees, Bayesian consensus trees, Bootstraps and Bayesian posterior trees) and metric (Robinson-Foulds and Triplets distance), we used this coefficient in two ways in order to:
@@
 %   -However, parameters have different effect
-However, within a method and a metric, the parameters used to remove data have different effect on the ability to recover the correct topology.
+However, within a method and a metric, the parameters used to remove data have different effects on the ability to recover the correct topology.
 %Results 2
@@
 %   -Missing data per se is not a problem, problem is in the amount of overlap
-As proposed in numerous previous studies, when recovering topology, the amount of missing data \textit{per se} is not a problem as long as enough data overlaps in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
-In fact, as shown previously, the amount of overlap in the matrix is more crucial than the amount of missing data \citep[e.g.][]{rouresite-specific2011}.
+As proposed in previousstudies, the amount of missing data is not a problem \textit{per se} as long as enough data overlaps in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
+%In fact, as shown previously, the amount of overlap in the matrix is more crucial than the proportion of missing data \citep[e.g.][]{rouresite-specific2011}.
 %Proposed hypothesis for result 1
 %   -One can increase overlap by adding living taxa
-To improve topological recovery in a Total Evidence framework, one can efficiently increase the amount of overlapping data reducing the amount of missing data in each of our three parameters.
+To improve topological recovery in a Total Evidence framework, one can efficiently increase the amount of overlapping data by reducing the amount of missing data in each of our three selected parameters.
 In our simulations, the overlap in data is greatly reduced when different parameters are combined with a high amount of missing data (e.g.
 $M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-In this scenario the overall number of missing characters $M_{C}$ affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
-However it is important to note that in the pairwise Bhattacharyya Coefficient analysis, the area with the lowest probability of distribution overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (Right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
+In this scenario the overall number of missing characters ($M_{C}$) affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
+However it is important to note that in the pairwise Bhattacharyya Coefficient analysis, the area with the lowest probability of distributions overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
 
@@
 %Discussing results 1
-Since each of our parameters are illustrating one potential sources of missing data in a morphological matrix (see \hyperref[Removing_data]{Methods}), we can propose different practical ways to reduce the amount of missing data:
+Each of our parameters illustrate one potential source of missing data in a morphological matrix (see \hyperref[Removing_data]{Methods}).
+We have suggestions for practical ways to reduce the missing data problem:
 %   -Which parameters does what?
@@
 The amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
-Improvement in both decrease the time taken to build the matrix and to increase the number of characters can be done through vast collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}.
+Improvement in both decreasing the time taken to build the matrix and increasing the number of characters can be done through vast collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}.
 \end{enumerate}
@@
 %§5
+
 %Caveats: the metrics
@@
 \label{metrics_discussion}
-Our results can be contrasted by looking at the normalised tree similarity when using the Robinson-Foulds or the Triplets distance.
-When using the second one, the confidence interval of our different parameters combinations per methods seems higher (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}) and the probability of overlap between the distribution of each parameter is high (Bhattacharyya Coefficient close to 1 - Fig.~\ref{Fig_Results-paircomp_within}-B).
-The contrast between the results of the two metrics can be explained by the actual nature of both metrics:
+As shown in Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}, our results can be contrasted by looking at the normalised tree similarity when using the Robinson-Foulds or the Triplets distance.
+For example, Fig.~\ref{Fig_Results-global_perparam} shows opposite results for both method when looking at the two different metrics.
+Globally, the normalised Triplets distance displays large confidence intervals than the normalised Robinson-Foulds distance and a higher probability of overlap between the distribution of each parameter (Bhattacharyya Coefficient close to 1 - Fig.~\ref{Fig_Results-paircomp_within}-B).
+The contrast between the results of the two metrics can be explained by the actual nature of both metrics (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}):
 \begin{enumerate}
-\item{The Robinson-Foulds metric is a conservative tree topology metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same number of taxa with the same topologies \citep{RF1981}.
+\item{The Robinson-Foulds metric is a conservative metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same taxa with the same topologies \citep{RF1981}.
 When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
@@
 \item{On the other hand, the Triplets method is measuring the position of each taxon towards two other reference taxa \citep{critchlowthe1996}.
-It will penalise only trees where taxa get removed furthest from their original clade.
-This metric illustrates therefore the amount of wildcard taxa \citep{kearneyfragmentary2002} between the to compared trees (i.e.
-when the score is closed to 1, view wildcard taxa are present in the tree).}
+It will penalise only trees where taxa get removed furthest from their original position.
+This metric therefore illustrates the amount of wildcard taxa \citep{kearneyfragmentary2002} between the to compared trees (i.e.
+when the score is closed to 1, few wildcard taxa are present in the tree).}
 \end{enumerate}
 Therefore, in our study, both metrics are showing two different aspects of tree topology.
-Also, disregarding their fundamental differences in accounting for topological distance, \citet{kuhnerpractical2014} have recently proposed that the Ronbinson-Foulds distance greatly outperforms the Triples distance in signalling differences in topology.
-%Maybe also a note about the fact that in Triplets, there's more than one solution each simulation (i.e.
-wildcard taxa can fix in many places) so it increases highly the variance in the NTS.
-In RF, because the metric is more conservative (bounded between 0 and the number of nodes), the number of solutions is smaller and therefore the variance decreases.
+For example, regarding the Bayesian consensus trees, completes clade are conserved most of the time (Fig.~\ref{Fig_Results-global_perparam}-top) but with increasing amount of missing data, wildcard taxa get more and more displaced (Fig.~\ref{Fig_Results-global_perparam}-bottom).
+Because we are comparing the Baysian consensus tree to a single tree topology, the probability of recovering the exact position of a wildcard taxa than in the "best" tree is really low and even reducing with the amount of missing data.
+Regarding this fundamental differences, the Robinson-Foulds metric performs better in showing a distance from the true topology than the Triplets distance \citep{kuhnerpractical2014}.
 
@@
 %   -Simulating morpho characters
-The size of our simulated matrices was at least two orders orders of magnitude lower than usual matrices, both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
-For morphological characters, the underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
+Another caveat of our analysis is the performance of our simulation protocol: in order to reduce the computational time ($~$150 CPU years) of our analysis, we had to simulate modest size matrices (1000 molecular data and 100 morphological data) that are two orders of magnitude lower than some matrices used in published analysis: both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
+%TG: note about NOT citing O'Leary here.
+1-I don't like the paper ; 2-the Ni et al paper (primates) as a big morpho matrix too (1k  characters against 4k for O'Leary, same order of magnitude?) ; 3-We might want to limit the papers we're citing?
+Also, simulating morphological characters is more complex than molecular ones.
+The underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
 Also morphological characters studies involve many potential statistical pitfalls (e.g.
-independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy \citep{wrightbayesian2014}.
+independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy \citep{wagner2000}.
 \begin{enumerate}
@@
 However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013}.}
-\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013}.
+\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013,pattinsonphylogeny2014}.
 However, following our simulation assumption of random character distributions, if they accumulate through time in the same way as the majority of the molecular characters then, homoplasic characters are expected to appear randomly through time \citep{davalosintegrating2014}.
 Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
-After a reaching a critical amount of morphological characters, adding new ones increases homoplasy \citep{wagner2000}.}
+After reaching a critical amount of morphological characters, adding new ones increases homoplasy \citep{wagner2000}.}
 \end{enumerate}
@@
 Using Total Evidence matrices with missing data is not a problem for recovering the correct topology as long as enough morphological data is overlapping between the fossil and living taxa.
-In order to optimize good topological recovery, with propose to use the Bayesian consensus tree over the Maximum Likelihood tree or a subset of the Bayesian posterior trees distribution.
+In order to optimize good topological recovery, we propose to use the Bayesian consensus tree over the Maximum Likelihood tree or over a random subset of the Bayesian posterior trees distribution.
 Also, in order to improve the overlap in data in the morphological part of the matrix, we advice to use all the morphological characters available and to code them for a maximum of living taxa present in the matrix.
@@
 
-%Ignored for fast pdf building
-%\section{Supplementary Material}
+%---------------------------------------------
+%
+%       Supplementary, separated
+%
+%---------------------------------------------
 
-%    \input{SupplementaryMaterial/SupplementaryMaterial}
-%    \label{SupplementaryMaterial}
+%Ignore for fast pdf building
+\section{Supplementary Material}
+
+    \input{SupplementaryMaterial/SupplementaryMaterial}
+    \label{SupplementaryMaterial}
 

### 2014-12-15 Manuscript/TEM_draft.tex
Discussion plan updated

@@
 \usepackage{url}
-%\usepackage{html}
 \usepackage{hyperref}
@@
 \noindent{\Large \bf Effect of missing data on topological inference using a total evidence approach}
-%Or do we put results in the title: Bayesian analysis and increased sampling of living taxa improves topological inference when using a total evidence approach
+% TG: Or "Effect of missing data and inference method on topology when using a total evidence approach"
+% TG: Or any other suggestions?
+
 \bigskip
@@
 This has led to increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013}.
-%Not a nice sentence
+However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments \citep[e.g.][]{pyrondivergence2011,ronquista2012,BEASTmaster}.
+%Or is it cheap to cite unpublished work (Nick's proto package) in the intro?
 
@@
 A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \citep{eernissetaxonomic1993}.
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
+This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (tip-dating method), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
 Total Evidence methods have been successfully applied to empirical data \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
@@
 the evolutionary information contained within the matrix allowing to infer topology and branch length), especially when using large matrices \citep{lemmonthe2009}.
-However, this may not be a major issue because phylogenetic signal is easily increased by: (i) including a "modest" number (~50) of highly-covered genes (i.e.
+However, this may not be a major issue because phylogenetic signal is easily increased by: (i) including a "modest" number ($~$50) of highly-covered genes (i.e.
 a number of genes that are available for all taxa ; \citealt{rouresite-specific2011}); (ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and (iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
@@
 And that's kind of the point we're getting toward with our work: we don't really care about the effect of missing data per se (i.e.
-what's the threshold were it starts fucking up the phylogeny) but more the fact that Bayesian fails.
+what's the threshold where it starts fucking up the phylogeny) but more the fact that Bayesian fails.
 
@@
 topological changes are discrete opposed to branch length changes are continuous).
-Also, interestingly, the effect of Total Evidence method has not been formally assessed in previous studies using fixed topology \citep{ronquista2012,schragocombining2013}.
+Also, interestingly, the effect of Total Evidence method has not been formally assessed in previous studies using fixed topology \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 
@@
 Therefore, we focus only on missing data in the morphological part of the matrix.
-We investigate three major parameters that directly affect the completeness of the morphological part of the matrix:
-\begin{enumerate}
-\item the proportion of living taxa with no morphological data;
-\item the proportion of missing data in the fossil taxa; and
-\item the proportion of missing morphological characters for both living and fossil taxa in the matrix.
-\end{enumerate} 
+We investigate three major parameters that directly affect the completeness of the morphological part of the matrix: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix.
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
@@
 Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
-
 We find that the ability of recovering the correct topology can mainly be increased when using the Bayesian methods.
@@
 \begin{enumerate}
-\item{Generating the matrix} \label{step:generate_matrix} \\
+\item{Generating the matrix:} \label{step:generate_matrix} \\
 We randomly generated a birth-death tree (hereafter called the "true" tree) and used it to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix).
-\item{Removing data} \label{step:remove_data} \\
-We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$) (the resulting matrices are called hereafter "missing-data" matrices).
-\item{Inferring phylogenies} \label{step:build_phylo} \\
-We inffered phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree) and multiple trees inferred from matrices with missing morphological data (hereafter called the "missing-data" trees).
+\item{Removing data:} \label{step:remove_data} \\
+We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$).
+The resulting matrices are called hereafter "missing-data" matrices.
+\item{Inferring phylogenies:} \label{step:build_phylo} \\
+We inferred phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree) and multiple trees inferred from matrices with missing morphological data (hereafter called the "missing-data" trees).
 Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
-\item{Comparing topologies} \label{step:compare_topo} \\
+\item{Comparing topologies:} \label{step:compare_topo} \\
 We compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies
 \end{enumerate}
-We repeated steps 1 to 4 50 times.
+We repeated the 4 steps 50 times to account for variation of our random parameters in the simulations.
 
@@
 (4) We compared the "missing-data" trees to the "best" tree.
-We repeated steps 1-4 50 times.}
+We repeated the 4 steps 50 times.}
 \label{Fig_Outline}
@@
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used low evolutionary rate parameters (i.e.
-$\alpha$) to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014}.
+We used low evolutionary rate parameters to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014,wrightbayesian2014}.
 
@@
 
-In practice, each parameter represents a different way of removing data from the matrix: $M_{L}$ removes rows from the living taxa; $M_{F}$ removes cells from the fossil taxa; and $M_{C}$ removes columns across both living and fossil taxa.
+In practice, each parameter represents a different way of removing data from the matrix: $M_{L}$ removes rows from the living taxa's data; $M_{F}$ removes cells from the fossil taxa's data; and $M_{C}$ removes columns across both living and fossil taxa's data.
 Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a percentage of living taxa is removed, but for $M_{F}$, a percentage of the data is removed at random from across the whole of the morphological matrix for fossil taxa.
@@
 The "best" tree and the "missing-data" trees were inferred using RAxML v8.0.20 \citep{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citealt{Stamatakis21012014}) as a generalization of the HKY + $\Gamma_4$ model \citep{HKY85} for the molecular data.
+For the molecular data, we used the GTR + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citealt{Stamatakis21012014})% as a generalization of the HKY + $\Gamma_4$ model \citep{HKY85}
+for the molecular data.
 %The GTR model can be seen as a generalization of the HKY model (the two parameters from the HKY model are implicitly included in the six from GTR model - \citealt{stamatakisa2008}).
@@
 %TG: The explanations of the algorithm part for the BS was proposed by Trevor I think it's useless
-When using these parameters, it took 6 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (8 core nodes 2.30GHz clock speed).
+When using these parameters, it took $~$8 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (8 core nodes 2.30GHz clock speed).
 
@@
 We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
-The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \citep{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
+The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of 2 \citep{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
 For the morphological data, we used the Markov \textit{k} state model \citep{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
@@
 For each run, we removed 25\% of the iterations as burn-in.
-We used the following priors for each tree (see \hyperref[SupplementaryMaterial]{Supplementary Material S1}):
-\begin{enumerate}
-\item the "true" tree’s topology as a starting tree (with a starting value for each branch length of 1),
-\item an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions, and
-\item a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)).
-\end{enumerate}
+We used the following priors for each tree (see \hyperref[SupplementaryMaterial]{Supplementary Material S1}): (i) the "true" tree’s topology as a starting tree (with a starting value for each branch length of 1), (ii) an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions, and (iii) a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)).
 
@@
 These priors biased the way the Bayesian process calculated branch lengths by giving non-random starting points and boundaries for parameter estimation, however, here we are focusing on the effect of missing data on tree topology and not branch lengths.
-Even using these priors, it took 140 CPU years to build 50 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
+Even using these priors, it took $~$140 CPU years to build 50 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
 
@@
 We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
-We used the Robinson-Foulds distance \citep{RF1981} to identify conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the number of conserved taxa across trees.
+We used the Robinson-Foulds distance \citep{RF1981} to assess the amount conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess amount of wildcard taxa \citep{kearneyfragmentary2002}.
 We then used Normalized Tree Similarity index \citep{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
@@
 This metric is bounded between 1 when the two trees are identical and $n-2$ (for two trees with $n$ taxa) when there is not one single shared clade between both trees.
-This metric is sensitive to the exact clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swap of two taxa will lead to a maximal score of the Robinson-Foulds distance indicating a bad tree similarity.
+This metric is sensitive to the exact clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swap of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating a bad tree similarity.
 
@@
 It is bounded between 0 when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is not one single position of taxon/clade identical between both trees.
-Therefore this metric sensitive to the conservation of individual taxa towards the neighbouring trees.
+Therefore this metric sensitive to the conservation of wildcard taxas.
 
@@
 % TG: Maybe remove supplementary here
-We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree.
+%We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree.
 We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}.
+TG: explained later on
 
@@
 
-For each of the 125 "missing-data" tree, we obtained Robinson-Foulds and Triplets distance distributions from either 50 or 50*1000 pairwise comparisons.
+For each of the 125 "missing-data" tree, we obtained Robinson-Foulds and Triplets distance distributions from either 50 or 50$\times$1000 pairwise comparisons.
 We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}.
 %Finally we tested for significant differences among the different distributions of trees using non-parametric group and pairwise comparisons by using respectively Kruskal-Wallis and Nemenyi-Damico-Wolfe-Dunn test (CITE R PACKAGE)\citet{ruxtontime2008}.
+TG: finally we didn't (Andrew's advice - see Battacharyya)
 
@@
 \begin{enumerate}
-\item assess the effect of the method on all parameters:
+\item{Assess the effect of the method on all parameters:}
 
@@
 We then compared the distribution of these pairwise comparisons to see the global difference between two methods: if the method have really similar results, we would expect the distribution to be cluster around 1 (high probability of overlap between the distributions) and if the methods are really dissimilar around 0 (low probability of overlap - Fig.~\ref{Fig_Results-paircomp_without}).
-\item assess the effect of the parameters within a method:
+\item{Assess the effect of the parameters within a method:}
 
 We performed a pairwise comparison per method and per metric for each combination of parameters.
-This resulted in 7875 pairwise comparisons per method and per metric (triangle of a 125*125 matrix).
-We represented these results as a triangular matrix with the values of each pairwise comparison coloured according to the value of the Batthacharyya Coefficient (coloured in green when the distributions overlap completely and in red when they don't - Fig.~\ref{Fig_Results-paircomp_within}).
+This resulted in 7875 pairwise comparisons per method and per metric (triangle of a 125$\times$125 matrix).
+We represented these results as a triangular matrix with the values of each pairwise comparison coloured according to the value of the Batthacharyya Coefficient (coloured in green when the distributions overlap completely and in red when they don't overlap at all- Fig.~\ref{Fig_Results-paircomp_within}).
 \end{enumerate}
@@
 
-As it would be expected form the literature, the amount of missing data in the morphological matrix does decrease the ability to recover the right topology, regardless of the parameter, the method or the metric used \citep[e.g.][]{rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014}.
+As expected form the literature, the amount of missing data in the morphological matrix does decrease the ability to recover the right topology, regardless of the parameter, the method or the metric used \citep[e.g.][]{rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014}.
 However each variable does not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam}).
@@
 the Ronbinson-Foulds distance), the Bayesian consensus trees outperform the other methods and the amount of missing data in the living taxa ($M_{L}$) decreases the most rapidly the clade conservation when using this method.
-However, when looking at the position of wild card taxa (i.e.
-the Triplets distance), the Maximum Likelihood outperforms the other methods but with wider distributions overlap (Fig.~\ref{Fig_Results-paircomp_without}-B).
+However, when looking at the position of wildcard taxa (i.e.
+using the Triplets distance), the Maximum Likelihood outperforms the other methods but with wider distributions overlap (Fig.~\ref{Fig_Results-paircomp_without}-B).
 
@@
     \includegraphics[width=1\textwidth]{Figures/BW/Global-ML+consensus-RF+Tr-small.pdf}
-\caption{Trend of the effect of missing data on topology on ML and consensus trees.
+\caption{Trend of the effect of missing data on topology on the Maximum Likelihood and the Bayesian consensus trees.
 The amount of missing data per parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is represented along the x axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
+The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data for each parameter.
 The topology is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
@@
 \caption{Distribution of the Bhattacharyya Coefficients between methods.
-Curves represents the kernel density estimations of the BC between the four different methods.
+Curves represents the kernel density estimations of the Bhattacharyya Coefficients between the four different methods.
 A.
@@
 Results with the for Normalised Triplets distance.
-Pikes around the two different extremes values of the BC show the two pairs of most dissimilar methods (Bayesian consensus vs.
+Pikes around the two different extremes values of the Bhattacharyya Coefficients show the two pairs of most dissimilar methods (Bayesian consensus vs.
 Bootstraps and Bayesian consensus vs.
@@
 When regarding at the Triplets distance, the two method that have the most distribution are the Bootstraps and the Bayesian posterior trees (Fig.~\ref{Fig_Results-paircomp_without}-B).
-These results are due to the slight difference in comparing topologies between Maximum Likelihood and Bayesian consensus trees and Bootstraps and Bayesian posterior trees: the two last ones are based on a random pairwise comparison of a sub-sample of 1000 trees.
+These results are due to the difference in comparing topologies between Maximum Likelihood and Bayesian consensus trees and Bootstraps and Bayesian posterior trees: the two last ones are based on a random pairwise comparison of a sub-sample of 1000 trees.
 This is likely to add noise to the data.
@@
 
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{Figures/colour/PairComp_within-Baycon-RF+Tr.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Bayesian consensus trees.
+%Hidden figure for 'fast' build
+%\begin{figure} 
+%\centering
+%    \includegraphics[width=1\textwidth]{Figures/colour/PairComp_within-Baycon-RF+Tr.pdf}
+%\caption{Pairwise Bhattacharyya Coefficients within the Bayesian consensus trees.
 The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0).
+The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data for each parameter.
+The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (1) to red (0).
 A.
@@
 \label{Fig_Results-paircomp_within}
-\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
+%\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
 
@@
 It probably doesn't need to have all the same subheadings either.
+% TG: Ok, from here, everything is new, I tried way more to write something proper following your "results-caveats-problems-extension" structure and I heavily based the structure of the paragraphs on the longevity paper + others (you might recognise you're style in the first paragraph if I did well).
+I left in comments the skeleton of the structure that I used to help me writing it.
+% TG: Also here's a note on my syntax notation for linking the paragraphs: "$"=end OR "^"=start of paragraph number 1.
+"a >..."=which finsih with the statement OR ">...
+b"=which follows with the statement b
+% TG: Here's the tree-style "plan" I was using (linked with the intro):
+% INTRO:
+% §1 We need fossils
+%   |
+%   \-> §2 How can we use them?
+%       |
+%       \-> §3 Total Evidence method of course!
+%           |
+%           \-> §4 But this method has a lot of missing data?
+%               |
+%               \-> §5 So what do we know about missing data from literature?
+%                   |
+%                   \-> §6 How will that effect TEM?
+%                      |
+%                      \-> §7 We use simulations to test!
+%                       
+% METHODS
+% RESULTS
+% DISCUSSION
+%       
+% §1 What do we found (Intro-§6)
+%   |
+%   \-> §2 How does that link to what we know (Intro-§5)
+%       |
+%       \-> §3 The effect of the parameters
+%       |
+%       \-> §4 The effect of the methods
+%       |
+%       \-> Caveats
+%           |
+%           \-> §5 The metrics
+%           |
+%           \-> §6 The simulations
+%           |
+%           \-> §7 How to fix them
+%           |
+%           \-> §8 How we can use fossils (Intro-§2)
+
+% CONCLUSION.
+% part 1- Fossils are important and can be used combined with living taxa (Intro-§1 & §2)
+% part 2- This is how we can improve the data to use them in the best way possible (Conclusion-§1 & §8)
 
@@
 %Results: global
-%   -Missing data increase decreases topological recovery
+%   -Missing data increase leads to decreases in recovering the correct topology
 The results of our simulations are in adequacy with recent analysis on the effect of missing data in morphological matrices \citep{pattinsonphylogeny2014,wrightbayesian2014,sansombias2014}.
-Our results are showing that ability to recover the correct topology decreases with the amount of missing data, regardless the metric, the method or the parameters considered.
+Our results are showing that the ability to recover the correct topology decreases with the increase of missing data, regardless the metric, the method or the parameters considered.
 %Results 1
@@
 
+% $§1: What we found >...
+% ^§2: >... links with previous studies
+
 %§2
@@
 %   -Missing data per se is not a problem, problem is in the amount of overlap
-As proposed in previousstudies, the amount of missing data is not a problem \textit{per se} as long as enough data overlaps in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
+As proposed in previous studies, the amount of missing data is not a problem \textit{per se} as long as enough data overlaps in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
 %In fact, as shown previously, the amount of overlap in the matrix is more crucial than the proportion of missing data \citep[e.g.][]{rouresite-specific2011}.
@@
 %   -One can increase overlap by adding living taxa
-To improve topological recovery in a Total Evidence framework, one can efficiently increase the amount of overlapping data by reducing the amount of missing data in each of our three selected parameters.
+To improve the ability to recover the "correct" topology in a Total Evidence framework, one can efficiently increase the amount of overlapping data by reducing the amount of missing data in each of our three selected parameters.
 In our simulations, the overlap in data is greatly reduced when different parameters are combined with a high amount of missing data (e.g.
@@
 
+% $§2: There is a different effect form each parameter >...
+% ^§3: >... because each parameter tells a different story.
+
 %§3
 %Discussing results 1
-Each of our parameters illustrate one potential source of missing data in a morphological matrix (see \hyperref[Removing_data]{Methods}).
-We have suggestions for practical ways to reduce the missing data problem:
+This variable effect of the missing data parameters on the ability to recover the "correct" topology (Fig.~\ref{Fig_Results-paircomp_within}) can be explained by the way our parameters are illustrating the sources of missing data in empirical matrices: (1) non coded-living  taxa ($M_{L}$), (2) the quality of the fossil record ($M_{F}$) and (3) the amount of coded characters ($M_{C}$ - see \hyperref[Removing_data]{Methods}).
+We have suggestions for practical ways to reduce this missing data problem in empirical matrices:
 %   -Which parameters does what?
@@
 \end{enumerate}
+Fixing these three major issues can lead to a sensible increase in the ability to recover the "correct" topology.
+
+% $§3: This is how we can fix the parameters >...
+% ^§4: >... but the methods also influence a lot the ability to recover the right topology.
 
@@
 %   -Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
-Nonetheless, as our results shows in Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}, the variation in the ability to recover the correct topology does not only depend on the missing data in morphological part of the matrix but also heavily on the method used to infer the topology.
+Nonetheless, the variation in the ability to recover the correct topology does not only depend on the missing data in morphological part of the matrix but also heavily on the method used to infer the topology (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 Previous studies have shown some superiority on methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014} but see \citealt{spencerefficacy2013}).
 However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} between Maximum Likelihood and Bayesian inference in a Total Evidence framework.
-%TG: I do have to check the Mk in Baysian vs Mk in ML part though but I guess it's a good selling point: we ARE the first ones to do a sensitivity analysis in a TEM framework (as far as we know).
+%TG: I guess it's a good selling point: we ARE the first ones to do a sensitivity analysis in a TEM framework (as far as I know).
 %Discussing results 2
@@
 Also, when looking at the pairwise Bhattacharyya Coefficients within each method, the Bayesian consensus tree one display the highest probability of overlap between the results in any case (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-
 %   -Why is Baycon so different than Baytre? Since Baycon is calculated from Baytre.
-When comparing this the probability of overlap between the methods, it is counter intuitive to note that the Bayesian consensus tree shows the highest dissimilarity with the Bayesian posterior trees from which the consensus tree is build (Fig.~\ref{Fig_Results-paircomp_within}-A).
+When comparing this to the probability of overlap between the methods, it is counter intuitive to note that the Bayesian consensus tree shows the highest dissimilarity with the Bayesian posterior trees from which the Bayesian consensus tree is build (Fig.~\ref{Fig_Results-paircomp_within}-A).
 Note that it is also highly dissimilar to the Bootstraps.
@@
 The fact that the 1000 pairwise trees to be compared for each of the 125 comparisons are chosen randomly, it can add noise to the results for these two methods (see \hyperref[tree_comparisons]{methods above}).
+Also, in the case of the lower performances of the Bayesian consensus trees when using the Triplets distance, one can note the high overlap of the confidence intervals (Fig.
+~\ref{Fig_Results-permeth_perparam} and Fig.
+~\ref{Fig_Results-paircomp_within}-B).
 
-%Some sort of transition?
+% $§4: Thus, baycon seems the best way to go >...
+% ^§5: >... but we have to take into account problems in our protocol.
 
 %§5
-
 %Caveats: the metrics
@@
 \label{metrics_discussion}
-As shown in Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}, our results can be contrasted by looking at the normalised tree similarity when using the Robinson-Foulds or the Triplets distance.
-For example, Fig.~\ref{Fig_Results-global_perparam} shows opposite results for both method when looking at the two different metrics.
-Globally, the normalised Triplets distance displays large confidence intervals than the normalised Robinson-Foulds distance and a higher probability of overlap between the distribution of each parameter (Bhattacharyya Coefficient close to 1 - Fig.~\ref{Fig_Results-paircomp_within}-B).
-The contrast between the results of the two metrics can be explained by the actual nature of both metrics (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}):
+This difference between the metrics can be explained by their respective ability to measure the differences between two trees \citep{kuhnerpractical2014}.
+As shown in Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}, the contrast between the results of the two metrics can be explained by the actual nature of both metrics (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}):
 \begin{enumerate}
-\item{The Robinson-Foulds metric is a conservative metric, it is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same taxa with the same topologies \citep{RF1981}.
+\item{The Robinson-Foulds metric is a conservative metric, which is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same taxa with the same topologies \citep{RF1981}.
 When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
@@
 \item{On the other hand, the Triplets method is measuring the position of each taxon towards two other reference taxa \citep{critchlowthe1996}.
-It will penalise only trees where taxa get removed furthest from their original position.
-This metric therefore illustrates the amount of wildcard taxa \citep{kearneyfragmentary2002} between the to compared trees (i.e.
+It will penalise only trees when taxa get removed furthest from their original position.
+This metric therefore illustrates the amount of wildcard taxa \citep{kearneyfragmentary2002} between the two compared trees (i.e.
 when the score is closed to 1, few wildcard taxa are present in the tree).}
@@
 Therefore, in our study, both metrics are showing two different aspects of tree topology.
-For example, regarding the Bayesian consensus trees, completes clade are conserved most of the time (Fig.~\ref{Fig_Results-global_perparam}-top) but with increasing amount of missing data, wildcard taxa get more and more displaced (Fig.~\ref{Fig_Results-global_perparam}-bottom).
+For example, regarding the Bayesian consensus trees, completes clade are conserved most of the time (Fig.~\ref{Fig_Results-global_perparam}-top) but with an increasing amount of missing data, wildcard taxa get more and more displaced (Fig.~\ref{Fig_Results-global_perparam}-bottom).
 Because we are comparing the Baysian consensus tree to a single tree topology, the probability of recovering the exact position of a wildcard taxa than in the "best" tree is really low and even reducing with the amount of missing data.
 Regarding this fundamental differences, the Robinson-Foulds metric performs better in showing a distance from the true topology than the Triplets distance \citep{kuhnerpractical2014}.
+Therefore, the results for the Bayesian consensus trees can be preferred on the Maximum Likelihood trees because - when using the Robinson-Fould distance - they display a higher ability to recover the "correct" topology (Fig.
+~\ref{Fig_Results-permeth_perparam}).
+
+% $§5: The difference in the methods is not important if you look at Kuhner >...
+% ^§6: >... but a final caveat is the way we simulate data.
 
@@
 %   -Simulating morpho characters
-Another caveat of our analysis is the performance of our simulation protocol: in order to reduce the computational time ($~$150 CPU years) of our analysis, we had to simulate modest size matrices (1000 molecular data and 100 morphological data) that are two orders of magnitude lower than some matrices used in published analysis: both for the molecular part \citep[e.g.][]{springermacroevolutionary2012} and the morphological part \citep[e.g.][]{nithe2013}.
+The variations in our results could also be explained by another methodological caveat: the performance of our simulation protocol.
+In order to reduce the computational time of our analysis ($~$150 CPU years), we had to run our sensitivity analysis on modest size matrices (1000 molecular data and 100 morphological data).
+This matrix size is two orders of magnitude lower than some matrices used in published analysis \citep[e.g.][]{springermacroevolutionary2012,nithe2013} but still ranges in the sizes of smaller matrices used in other published analysis \citep[e.g.][]{sallam2011craniodental,kellymolecular2014}.
 %TG: note about NOT citing O'Leary here.
 1-I don't like the paper ; 2-the Ni et al paper (primates) as a big morpho matrix too (1k  characters against 4k for O'Leary, same order of magnitude?) ; 3-We might want to limit the papers we're citing?
+note about citing Sean's and your paper here: 1-is it okay (regarding last time warning of being extra careful when citing you're own work)?
 Also, simulating morphological characters is more complex than molecular ones.
-The underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001}.
-Also morphological characters studies involve many potential statistical pitfalls (e.g.
+The underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001,wrightbayesian2014}.
+Morphological characters studies involve many potential statistical pitfalls (e.g.
 independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy \citep{wagner2000}.
 \begin{enumerate}
-\item{First, morphological data can display a different signal than molecular data, especially in small matrices. %REF
+\item{First, morphological data can display a different signal than molecular data, especially in small matrices \citep{wagner2000}. %REF
 This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
@@
 \end{enumerate}
+
+% $§6: Knowing the caveats, we can try to fix them >...
+% ^§7: >... by taking into account data realness (fossil templates and rate heterogeneity).
 
@@
 
+% $§7: But despite the caveats above, missing data is not a big deal >...
+% ^§8: >... We can use TEM for integrating fossils into trees.
+
 %§8
-%Back to the importance of TEM
+%Back to the importance of TEM - How we can use fossils (Intro-§2)
 %   -Data is not a major problem
+Despite the various caveats discussed above, we answered our main question on how does Total Evidence matrices perform in recovering topology with different methods and different amounts of missing data.
 %   -Good and promising method.
+Our results shows that using this method is totally possible but that a loss in topological accuracy is to be expected with a decrease of overlapping data.
+Two ways to address this eventual problem is to (i) increase the data overlap by coding living species and to have a large morphological matrix and (ii) to use a Bayesian consensus tree.
 
@@
 
-\section{Conclusion} %Can probably just be the last paragraph but I like it to be separated as the take home message
+\section{Conclusion}
 
 %§Conclusions
+% part 1- Fossils are important and can be used combined with living taxa (Intro-§1 & §2)
 %   -TEM works with missing data
+% part 2- This is how we can improve the data to use them in the best way possible (Conclusion-§1 & §8)
 %   -Use bayesian consensus
@@
 \section{Acknowledgements}
-Thanks to Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Andrew Jackson, Gavin Thomas, April Wright and the members of the Macro Journal Club %TG : Not sure how many people we have to thank but it could also include Graham Slater, Nick Matzke and Trevor Hodkinson.
+Thanks to Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Andrew Jackson, Gavin Thomas and April Wright.
+%TG : Not sure how many people we have to thank but it could also include Graham Slater, Nick Matzke and Trevor Hodkinson.
 for useful comments on our simulation protocol.
@@
 
-%Ignore for fast pdf building
-\section{Supplementary Material}
+%Ignore for 'fast' pdf building
+%\section{Supplementary Material}
 
-    \input{SupplementaryMaterial/SupplementaryMaterial}
-    \label{SupplementaryMaterial}
+%    \input{SupplementaryMaterial/SupplementaryMaterial}
+%    \label{SupplementaryMaterial}
 

### 2014-12-18 Analysis/MissingData_plot.R
Updated plot (final data input)

@@ -30,8 +30,8 @@ if(quick.load == FALSE) {
 #Plot main figures (individual parameters in RF and Triples)
 #####################
 #windows parameters
-#quartz(width = 8.3, height = 5.8) #A5 landscape
-quartz(width = 16.6, height = 11.6) #A4 landscape
+quartz(width = 8.3, height = 5.8) #A5 landscape
+#quartz(width = 16.6, height = 11.6) #A4 landscape
 op<-par(mfrow=c(2,3), bty="l", oma=c(0,0,0,0))
 ylim=c(0,1)
 
@@ -69,8 +69,8 @@ par(op)
 #Plot method differences figure
 #####################
 #windows parameters
-#quartz(width = 8.3, height = 5.8) #A5 landscape
-quartz(width = 16.6, height = 11.6) #A4 landscape
+quartz(width = 8.3, height = 5.8) #A5 landscape
+#quartz(width = 16.6, height = 11.6) #A4 landscape
 op<-par(mfrow=c(2,3), bty="l", oma=c(0,0,0,0))
 ylim=c(0,1)
 
@@ -118,8 +118,8 @@ par(op)
 #Plot parameter differences figure
 #####################
 #windows parameters
-#quartz(width = 8.3, height = 5.8) #A5 landscape
-quartz(width = 16.6, height = 11.6) #A4 landscape
+quartz(width = 8.3, height = 5.8) #A5 landscape
+#quartz(width = 16.6, height = 11.6) #A4 landscape
 op<-par(mfrow=c(2,3), bty="l", oma=c(0,0,0,0))
 ylim=c(0,1)
 

### 2014-12-18 Manuscript/SupplementaryMaterial/Supp_results.tex
Figures update

@@
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-Global-Bootstrap+Treeset-RF+Tr-small.pdf}
+    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/Boot+Baytre-AllParam-RF+Tr.pdf}
 \caption{Trend of the effect of missing data on topological recovery on the Bootstraps and the Bayesian posterior trees distributions.
@@
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_permeth.pdf}
+    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/BC-AllMethods-RF+Tr.pdf}
 \caption{Distribution of the Pairwise Bhattacharyya Coefficients within each method.
@@
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-Baytre-RF.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Bayesian posterior trees distributions.
+    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/PairwiseComp-ML-RF+Tr.pdf}
+\caption{Pairwise Bhattacharyya Coefficients within the Maximum Likelihood trees.
 The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0).
-Results are calculated for the Normalised Robinson-Foulds distance.}
+The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data for each parameter.
+The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (1) to red (0).
+A.
+Results for the Normalised Robinson-Foulds distance.
+B.
+Results with the for Normalised Triplets distance.}
 \label{Fig_pairComp-Baytree-RF}
-\end{figure} %Pairwise BC for the Bayesian treesets for RF. (parameters differences)
+\end{figure} %Pairwise BC for the ML for RF+Tr. (parameters differences)
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-Baytre-Tr.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Bayesian posterior trees distributions.
+    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/PairwiseComp-Boot-RF+Tr.pdf}
+\caption{Pairwise Bhattacharyya Coefficients within the Bootstrap trees.
 The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0).
-Results are calculated for the Normalised Triplets distance.}
+The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data for each parameter.
+The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (1) to red (0).
+A.
+Results for the Normalised Robinson-Foulds distance.
+B.
+Results with the for Normalised Triplets distance.}
 \label{Fig_pairComp-Baytree-Tr}
-\end{figure} %Pairwise BC for the Bayesian treesets for Tr. (parameters differences)
+\end{figure} %Pairwise BC for the Boot for RF+Tr. (parameters differences)
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-MLbest-RF.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Maximum Likelihood trees.
+    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/PairwiseComp-Baytre-RF+Tr.pdf}
+\caption{Pairwise Bhattacharyya Coefficients within the Bayesian posterior distribution trees.
 The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0).
-Results are calculated for the Normalised Robinson-Foulds distance.}
+The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data for each parameter.
+The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (1) to red (0).
+A.
+Results for the Normalised Robinson-Foulds distance.
+B.
+Results with the for Normalised Triplets distance.}
 \label{Fig_pairComp-MLbest-RF}
-\end{figure} %Pairwise BC for the ML for RF. (parameters differences)
-
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-MLbest-Tr.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Maximum Likelihood trees.
-The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0).
-Results are calculated for the Normalised Triplets distance.}
-\label{Fig_pairComp-MLbest-Tr}
-\end{figure} %Pairwise BC for the ML for Tr. (parameters differences)
-
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-MLboot-RF.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Bootstraps.
-The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0).
-Results are calculated for the Normalised Robinson-Foulds distance.}
-\label{Fig_pairComp-MLboot-RF}
-\end{figure} %Pairwise BC for the Bootstraps for RF. (parameters differences)
-
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/TEM_Fig-Appendix-PairComp_within-MLboot-Tr.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Bootstraps.
-The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (BC=1) to red (BC=0).
-Results are calculated for the Normalised Triplets distance.}
-\label{Fig_pairComp-MLboot-Tr}
-\end{figure} %Pairwise BC for the Bootstraps for Tr. (parameters differences)
+\end{figure} %Pairwise BC for the Baytre for RF+Tr. (parameters differences)
 
@@
   \hline
-  Maximum likelihood-RF & 0.06 & 0.26 & 0.39 & 0.40 & 0.49 & 0.94 \\ 
-  Maximum likelihood-Tr & 0.27 & 0.45 & 0.58 & 0.63 & 0.83 & 1.00 \\ 
-  Bayesian consensus-RF & 0.70 & 0.71 & 0.73 & 0.76 & 0.79 & 0.96 \\ 
-  Bayesian consensus-Tr & -0.25 & -0.09 & 0.18 & 0.21 & 0.40 & 0.98 \\ 
-  Bootstraps-RF & 0.06 & 0.18 & 0.27 & 0.26 & 0.33 & 0.46 \\ 
-  Bootstraps-Tr & 0.24 & 0.32 & 0.36 & 0.38 & 0.46 & 0.60 \\ 
+  Maximum likelihood-RF & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
+  Maximumlikelihood-Tr & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
+  Bayesian consensus-RF & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
+  Bayesian consensus-Tr & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
+  Bootstraps-RF & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
+  Bootstraps-Tr & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
   Bayesian posterior trees-RF & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
-  Bayesian posterior trees-Tr & 0.24 & 0.36 & 0.40 & 0.50 & 0.68 & 0.98 \\ 
+  Bayesian posterior trees-Tr & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
+   \hline
    \hline

### 2014-12-18 Manuscript/TEM_draft.tex
Figures update

@@
 \centering
-    \includegraphics[keepaspectratio=true]{Figures/BW/TEM_Fig_outline-BW.pdf}
+    \includegraphics[keepaspectratio=true]{Figures/In_main/Simulations_outline-BW.pdf}
 \caption{Protocol outline.
@@
 We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
-We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, total computational time for the whole analysis was over 150 CPU years).
+We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, total computational time for the whole analysis was around 150 CPU years).
 All the molecular information for fossil taxa was replaced by missing data ("?").
@@
 %TG: The explanations of the algorithm part for the BS was proposed by Trevor I think it's useless
-When using these parameters, it took $~$8 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (8 core nodes 2.30GHz clock speed).
+When using these parameters, it took $~$8 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (2.30GHz clock speed nodes).
 
@@
 These priors biased the way the Bayesian process calculated branch lengths by giving non-random starting points and boundaries for parameter estimation, however, here we are focusing on the effect of missing data on tree topology and not branch lengths.
-Even using these priors, it took $~$140 CPU years to build 50 sets of 125 Bayesian trees (8 core nodes 2.30GHz clock speed).
+Even using these priors, it took $~$140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
 
@@
 These metrics are described in detail below.
-%ADD KHUNAR PAPER SOMEWHERE
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{Figures/colour/Parameters-AllMethods-RF+Tr-small.pdf}
+    \includegraphics[width=1\textwidth]{Figures/In_main/AllMethods-RF+Tr-colour.pdf}
 \caption{Comparison between the effect of missing data and the tree inference method on topology.
@@
 \centering
-    \includegraphics[width=1\textwidth]{Figures/BW/Global-ML+consensus-RF+Tr-small.pdf}
+    \includegraphics[width=1\textwidth]{Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
 \caption{Trend of the effect of missing data on topology on the Maximum Likelihood and the Bayesian consensus trees.
@@
 \centering
-    \includegraphics[width=1\textwidth]{Figures/colour/PairComp_without-RF+Tr.pdf}
+    \includegraphics[width=1\textwidth]{Figures/In_main/BC-MethodComp-RF+Tr-colour.pdf}
 \caption{Distribution of the Bhattacharyya Coefficients between methods.
@@
 
-%Hidden figure for 'fast' build
-%\begin{figure} 
-%\centering
-%    \includegraphics[width=1\textwidth]{Figures/colour/PairComp_within-Baycon-RF+Tr.pdf}
-%\caption{Pairwise Bhattacharyya Coefficients within the Bayesian consensus trees.
+%Hide this figure for 'fast' build
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
+\caption{Pairwise Bhattacharyya Coefficients within the Bayesian consensus trees.
 The pairwise trees comparisons are represent on both axis.
@@
 \label{Fig_Results-paircomp_within}
-%\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
+\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
 
@@
 
-%Ignore for 'fast' pdf building
-%\section{Supplementary Material}
+%Ignore this for 'fast' pdf building
+\section{Supplementary Material}
 
-%    \input{SupplementaryMaterial/SupplementaryMaterial}
-%    \label{SupplementaryMaterial}
+    \input{SupplementaryMaterial/SupplementaryMaterial}
+    \label{SupplementaryMaterial}
 

### 2014-12-19 Manuscript/TEM_draft.tex
Pro perish draft version (full draft)

@@
 The two factors that influences the most this data overlap are the parameters 1 (proportion of coded living taxa) and 3 (number of morphological characters).
+
+%Probably needs to be deeeply rewritten, as said last time to Kev quoting Stuart Pim, people only read the abstract.
+
 \end{abstract}
@@
 
-As expected form the literature, the amount of missing data in the morphological matrix does decrease the ability to recover the right topology, regardless of the parameter, the method or the metric used \citep[e.g.][]{rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014}.
+As expected form the literature, the amount of missing data in the morphological part of the total evidence matrix does decrease the ability to recover the right topology, regardless of the parameter, the method or the metric used \citep[e.g.][]{rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014}.
 However each variable does not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam}).
@@
 However, when looking at the position of wildcard taxa (i.e.
-using the Triplets distance), the Maximum Likelihood outperforms the other methods but with wider distributions overlap (Fig.~\ref{Fig_Results-paircomp_without}-B).
+using the Triplets distance), the Maximum Likelihood outperforms the other methods but with wider distributions overlap (Fig.~\ref{Fig_Results-permeth_perparam}).
 
@@
 
-When looking at the global trend of the Bayesian consensus trees for all the parameters combinations, this method outperforms Maximum Likelihood for the Ronbinson-Foulds distance and plateaus to a minimal modal tree similarity of 0.69 regardless the amount of missing data (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-However, regarding the Triplets distance, Bayesian consensus trees seems to perform poorly but with great confidence intervals overlaps (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
-These greater larger confidence intervals are due to the lower ability of the Triplet method to highlight precise differences in topology \citep{kuhnerpractical2014}.
+When looking at the effect of all parameters combinations on the Bayesian consensus trees, this method outperforms Maximum Likelihood for the Ronbinson-Foulds distance and plateaus to a minimal modal tree similarity of 0.69 regardless the amount of missing data (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+However, regarding the Triplets distance, Bayesian consensus trees seems to perform poorly but with great confidence intervals overlaps (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+These larger confidence intervals are due to the lower ability of the Triplet method to highlight precise differences in topology \citep{kuhnerpractical2014}.
 Results for the global trend comparison between the uncertainty methods (Bootstrap and Bayesian posterior tree distribution) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
@@
 %Hide this figure for 'fast' build
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Bayesian consensus trees.
+%\begin{figure} 
+%\centering
+%    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
+%\caption{Pairwise Bhattacharyya Coefficients within the Bayesian consensus trees.
 The pairwise trees comparisons are represent on both axis.
@@
 However, when looking at the Triplets distance, because of the high overlap in confidence intervals, there is no major region with low probability of overlap.
-One should note however that the bottom line of the matrix where the Bhattacharyya Coefficients are very low is due to comparing the distribution of the pairwise comparison of the best tree vs.
-the best tree, leading to a Normalised triplet score of 1 every time (with no variance, see Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-All the other pairwise comparisons are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+One should note however that the bottom line of the matrix where the Bhattacharyya Coefficients are very low is due to comparing the distribution of the pairwise comparison of the "best" tree against the "missing-data" tree with no missing data ($M_{L}$=0\%, $M_{C}$=0\%, $M_{F}$=0\%) , leading to a normalised Triplets score of 1 every time (with no variance, see Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+The pairwise comparisons for the other methods (Maximum Likelihood, Bootstraps and Bayesian posterior trees) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 
@@
 However, gaps in the matrix can be filled with the discoveries exceptional preserved fossil \citep[e.g.][]{nithe2013}.
-Fortunately, this parameter is the most difficult to improve but it also has the least influence on whether the simulations recover the correct topology (Fig.~\ref{Fig_Results-paircomp_within}).
+Fortunately, this parameter is the most difficult to improve but it also has the least influence on whether the simulations recover the "correct" topology (Fig.~\ref{Fig_Results-paircomp_within}).
 \item{$M_{C}$ (amount of coded characters):}
@@
 %   -Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
-Nonetheless, the variation in the ability to recover the correct topology does not only depend on the missing data in morphological part of the matrix but also heavily on the method used to infer the topology (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+Nonetheless, the variation in the ability to recover the "correct" topology does not only depend on the missing data in morphological part of the matrix but also heavily on the method used to infer the topology (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 Previous studies have shown some superiority on methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014} but see \citealt{spencerefficacy2013}).
@@
 In order to reduce the computational time of our analysis ($~$150 CPU years), we had to run our sensitivity analysis on modest size matrices (1000 molecular data and 100 morphological data).
-This matrix size is two orders of magnitude lower than some matrices used in published analysis \citep[e.g.][]{springermacroevolutionary2012,nithe2013} but still ranges in the sizes of smaller matrices used in other published analysis \citep[e.g.][]{sallam2011craniodental,kellymolecular2014}.
+This matrix size is two orders of magnitude lower than some matrices used in published analysis \citep[e.g.][]{springermacroevolutionary2012,nithe2013} but still ranges in the sizes of smaller matrices used in other ones \citep[e.g.][]{sallam2011craniodental,kellymolecular2014}.
 %TG: note about NOT citing O'Leary here.
@@
 However, the true topology (obtained from a matrix with no missing data) is unlikely to be known.
-A second way to address this issue, and especially to incorporate the variation of signal in morphological character would be to simulate morphological at different rates \citep{wrightbayesian2014} or to apply patterns of missing data directly obtained from the fossil record \citep{pattinsonphylogeny2014}.
-Nevertheless, our simulation, by simplifying an ideal world, already show clear results of the effect of missing data on topology in a Total Evidence framework.
+A second way to address this issue, and especially to incorporate the variation of signal in morphological character would be to simulate morphological characters at different rates \citep{wrightbayesian2014} or to apply patterns of missing data directly obtained from the fossil record \citep{pattinsonphylogeny2014}.
+Nevertheless, our simulation already show clear results of the effect of missing data on topology in a Total Evidence framework.
 Also, our results show how recovering the correct topology can be improved by using Bayesian consensus trees in combination with a maximum amount of characters and a of living taxa with morphological data.
@@
 %   -Good and promising method.
-Our results shows that using this method is totally possible but that a loss in topological accuracy is to be expected with a decrease of overlapping data.
-Two ways to address this eventual problem is to (i) increase the data overlap by coding living species and to have a large morphological matrix and (ii) to use a Bayesian consensus tree.
+Our results shows that this method performance decreases mainly with a reduction in the overlap of morphological data.
+In order to address this eventual problem, we propose to increase the data overlap by coding living species and to have a large morphological matrix.
+Also, our results shows disparity among methods and a clear superiority of the Bayesian consensus tree topology over the Maximum Likelihood one.
 
@@
 \bibliography{References}
+%Needs to be cleaned a bit
 
@@
 %Ignore this for 'fast' pdf building
-\section{Supplementary Material}
+%\section{Supplementary Material}
 
-    \input{SupplementaryMaterial/SupplementaryMaterial}
-    \label{SupplementaryMaterial}
+%    \input{SupplementaryMaterial/SupplementaryMaterial}
+%    \label{SupplementaryMaterial}
 

### 2015-01-03 Manuscript/TEM_draft.tex
Comments to introduction

@@
 %Probably needs to be deeeply rewritten, as said last time to Kev quoting Stuart Pim, people only read the abstract.
+% NC: Yep, for now I'll leave this
 
@@
 Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
-This has led to increasing consensus among scientists that fossil taxa must be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
+This has led to increasing consensus among evolutionary biologists that fossil taxa should be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
 However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments \citep[e.g.][]{pyrondivergence2011,ronquista2012,BEASTmaster}.
 %Or is it cheap to cite unpublished work (Nick's proto package) in the intro?
+% NC: Check Syst Biols policies, but I think it's fine as it's an available package so it'd be like citing something from GitHub.
+You could ask Nick before publication if there's a different citation he'd prefer you to use.
 
@@
 Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence date are used to time calibrate phylogenies \citep{zuckerkandl1965}.
+Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates are used to time calibrate phylogenies \citep{zuckerkandl1965}.
 There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
@@
 A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \citep{eernissetaxonomic1993}.
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (tip-dating method), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
+This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as tip-dating; \citealt{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
 Total Evidence methods have been successfully applied to empirical data \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
 However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data that can be difficult (or impossible) to collect.
-The morphological data for living taxa is rarely collected when molecular data is available (e.g.
-\citealt{O'Leary08022013} \textit{vs.} \citealt{meredithimpacts2011}), and, for fossil taxa, the scarcity of the fossil record only allow to collect the data available (for example, in vertebrates, the hardest parts of the skeleton; \citealt{sansomfossilization2013}).
+The morphological data for living taxa are rarely collected when molecular data are available (e.g.
+\citealt{O'Leary08022013} \textit{vs.} \citealt{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record (for example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts; \citealt{sansomfossilization2013}).
 Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
@@
 The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,sansombias2014}.
-Missing molecular data has been seen by some authors as an issue because it can, in some part of the tree, decrease phylogenetic signal (i.e.
-the evolutionary information contained within the matrix allowing to infer topology and branch length), especially when using large matrices \citep{lemmonthe2009}.
-However, this may not be a major issue because phylogenetic signal is easily increased by: (i) including a "modest" number ($~$50) of highly-covered genes (i.e.
-a number of genes that are available for all taxa ; \citealt{rouresite-specific2011}); (ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and (iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
+Missing molecular data has been seen by some authors as an issue because it can, in some parts of the tree, decrease phylogenetic signal (i.e.
+the evolutionary information contained within the matrix allowing the inference of topology and branch lengths), especially when using large matrices \citep{lemmonthe2009}.
+However, this may not be a major issue because phylogenetic signal is easily increased by: (i) including a "modest" number ($~$ % NC: This symbol doesn't show up on my PDF.
+Check.
+50) of highly-covered genes (i.e.
+a number of genes that are available for all taxa; \citealt{rouresite-specific2011}); (ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and (iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
 Similarly, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study in question \citep{wiensmissing2003,sansomfossilization2013,pattinsonphylogeny2014}.
 Because soft-tissue characters are rarely preserved in the fossil record, missing data is mainly found in these characters, and is therefore not randomly distributed which can lead to biased placement of fossil taxa in phylogenies (e.g.
-\citealt{sansomfossilization2013} but see \citealt{pattinsonphylogeny2014}).
+\citealt{sansomfossilization2013}, but see \citealt{pattinsonphylogeny2014}).
 However, the phylogenetic signal is not related to the amount of missing data \textit{per se} but to the number of informative characters for each taxon, therefore missing data is less of an issue than the number of shared informative characters \citep{kearneyfragmentary2002,wiensmissing2003,pattinsonphylogeny2014}.
@@
 what's the threshold where it starts fucking up the phylogeny) but more the fact that Bayesian fails.
+% NC: OK fair enough. I think though if this is too long we can probably cut this paragraph and put a bit of it into the paragraph below.
 
 Although missing data does not appear be a major problem in molecular and morphological matrices separately \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil species.
-This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data.
+This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
 Until now, no attempt has been made to study the impact of this issue on phylogenetic inference from Total Evidence methods.
-% NC: Somewhere here you need to flag that you're only looking at topologies and briefly explain why.
-In this study, we are interested in assessing the effect of missing data on the ability to recover a correct topology.
-Even though branch length is a second aspect of phylogenetic inferences equally important, topology is the first and most straightforward aspect reflecting phylogenetic signal (i.e.
-topological changes are discrete opposed to branch length changes are continuous).
-Also, interestingly, the effect of Total Evidence method has not been formally assessed in previous studies using fixed topology \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+% NC: THIS SECTION NEEDS WORK!
+In this study, we focus on the effect of missing data on the ability to recover a correct topology, 
+% NC: I don't get the logic of this part 
+% because topology is the first and most straightforward aspect reflecting phylogenetic signal (i.e. topological changes are discrete opposed to branch length changes are continuous). 
+% NC: Need a better justification of WHY topology is interesting.
+Although correctly inferring branch lengths is also important, we do not consider the inference of branch lengths here.
+% NC: WHY??? Difficulties with simulations? Topology is more important? WHY?
+Also, the effect of Total Evidence method has not been formally assessed in previous studies using fixed topology \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+% NC: OK this last sentence could be worked in earlier on to explain WHY topology not branch lengths.
 
@@
 Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
-We find that the ability of recovering the correct topology can mainly be increased when using the Bayesian methods.
-Also minimizing the proportion of living taxa with no morphological data and the proportion of missing morphological characters improves the ability to recover the correct more than minimizing the proportion of missing data in the fossil record.
+We find that the ability of Total Evidence methods to recover the correct topology is increased when using Bayesian methods.
+Also minimizing the proportion of living taxa with no morphological data and the proportion of missing morphological characters improves the ability of Total Evidence methods to recover the correct topology more so than minimizing the proportion of missing data in the fossil record.
 

### 2015-01-03 Manuscript/TEM_draft.tex
Comments to methods up to tree comparisons

@@
 To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
+% NC: I've added numbers in here (125 matrices etc) so it matches the figure legend better. Can you check I've put the right numbers in? I think it may be 124 in some cases...
 \begin{enumerate}
@@
 We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$).
-The resulting matrices are called hereafter "missing-data" matrices.
+The resulting 125 matrices are called hereafter "missing-data" matrices.
 \item{Inferring phylogenies:} \label{step:build_phylo} \\
-We inferred phylogenetic trees from the "complete" matrix and from the "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree) and multiple trees inferred from matrices with missing morphological data (hereafter called the "missing-data" trees).
+We inferred phylogenetic trees from the "complete" matrix and from the 125 "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree) and 125 trees inferred from the matrices with missing morphological data (hereafter called the "missing-data" trees).
 Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
@@
 \end{enumerate}
-We repeated the 4 steps 50 times to account for variation of our random parameters in the simulations.
+We repeated these four steps 50 times to account for variation in our random parameters in the simulations.
 
@@
 (4) We compared the "missing-data" trees to the "best" tree.
-We repeated the 4 steps 50 times.}
+We repeated these four steps 50 times.}
 \label{Fig_Outline}
 \end{figure}
+% NC: Can you centre these around the boxes? and make the comparing topologies arrow touch the two triangles? I'll show you what I mean. 
+% NC: Also check the figure is in a format and of correct quality for Syst Biol.
 
@@
 
-\subsection{Generating the matrix}
-First we randomly generated a "true" tree of 50 taxa in R v3.0.2 \citep{R302} using the package diversitree v0.9-6 \citep{fitzjohndiversitree2012}.
+\subsection{Generating the matrix} % NC: Can you number these subsections like they are in the steps above?
+First we randomly generated a "true" tree of 50 taxa in R v.
+3.0.2 \citep{R302} using the package diversitree v.
+0.9-6 \citep{fitzjohndiversitree2012}.
 We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
@@
 Next, we generated a molecular and a morphological matrix from the "true" tree.
-The molecular matrix was inferred from the "true" tree using the R package phyclust v0.1-14 \citep{chen2011}.
-The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies and transition/transversion rate of 2 \citep{douadycomparison2003}.
+The molecular matrix was inferred from the "true" tree using the R package phyclust v.
+0.1-14 \citep{chen2011}.
+The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies and transition/transversion rate of two \citep{douadycomparison2003}.
 The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
@@
 
-We inferred the morphological matrix using the R package ape v3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
+We inferred the morphological matrix using the R package ape v.
+3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
@@
 We then ran an independent discrete character simulation for each character using the "true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
-evolutionary rate) from one character state to an other \citep{Pagel22011994}.
+evolutionary rate) from one character state to another \citep{Pagel22011994}.
 This method allows us to have only two parameters for each character: the number of states and the evolutionary rate.
@@
 Finally, we combined the morphological and molecular matrices obtained from the "true" tree.
-Hereafter we call this the "complete" matrix: the matrix with no missing data except for the molecular data of the fossil taxa.
+Hereafter we call this the "complete" matrix, i.e.
+the matrix with no missing data except for the molecular data of the fossil taxa.
 
@@
 \begin{enumerate}
-\item{The proportion of living taxa with no morphological data ($M_{L}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
+\item{$M_{L}$, the proportion of living taxa with no morphological data: 0\%, 10\%, 25\%, 50\% or 75\%.}
 This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological part.
-This reflects the fact that because of the increasing availability of DNA sequences for living taxa, detailed morphological data is scarce.
-\item{The proportion of missing data in the fossil taxa ($M_{F}$): 0\%, 10\%, 25\%, 50\% or 75\%.}
+This reflects the fact that detailed morphological data for living species are scarce.
+% NC: Not sure you can really "blame" this on DNA data availability!
+\item{$M_{F}$, the proportion of missing data in the fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%.}
 This parameter illustrates the quality of the fossil record. 
-\item{the proportion of missing morphological characters for both living and fossil taxa ($M_{C}$): 0\%, 10\%, 25\%, 50\% or 75\%. }
+\item{$M_{C}$, the proportion of missing morphological characters for both living and fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%. }
 This parameter illustrates the number of available morphological characters for both living and fossil taxa.
@@
 In practice, each parameter represents a different way of removing data from the matrix: $M_{L}$ removes rows from the living taxa's data; $M_{F}$ removes cells from the fossil taxa's data; and $M_{C}$ removes columns across both living and fossil taxa's data.
-Note that $M_{L}$ is different to $M_{F}$ not only because of the region of the matrix affected: for $M_{L}$, all the morphological data of a percentage of living taxa is removed, but for $M_{F}$, a percentage of the data is removed at random from across the whole of the morphological matrix for fossil taxa.
+Note that $M_{L}$ and $M_{F}$ differ not only because of the region of the matrix affected: for $M_{L}$ all the morphological data of a percentage of living taxa is removed, whereas for $M_{F}$ a percentage of the data is removed at random from across the whole of the morphological matrix for fossil taxa.
 
-We tested all parameters combinations resulting in 125 ($5^3$) matrices.
+We created matrices using all parameter combinations resulting in 125 ($5^3$) matrices.
 Note that one of these combinations has no missing data so is equivalent to the "complete" matrix, thus we have one effectively complete matrix in our 125 "missing-data" matrices.
@@
 $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
-When this occurred we repeated the random deletion of characters until every taxa had at least 5\% data across the whole morphological part of the matrix.
+When this occurred we repeated the random deletion of characters until each taxon had at least 5\% data across the whole morphological part of the matrix.
 
@@
 \subsection{Building phylogenies}
-From the resulting matrices we generated two types of trees, the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
+From the resulting matrices we generated two types of trees: the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
 The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
@@
 \citealt{rozen2005}).
-Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what biologists have to work with.
+Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what we have to work with.
 
 \subsubsection{Maximum Likelihood}
-The "best" tree and the "missing-data" trees were inferred using RAxML v8.0.20 \citep{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v8.0.20; \citealt{Stamatakis21012014})% as a generalization of the HKY + $\Gamma_4$ model \citep{HKY85}
-for the molecular data.
+The "best" tree and the "missing-data" trees were inferred using RAxML v.
+8.0.20 \citep{Stamatakis21012014}.
+For the molecular data, we used the GTR + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v.
+8.0.20; \citealt{Stamatakis21012014}).% as a generalization of the HKY + $\Gamma_4$ model \citep{HKY85}.
 %The GTR model can be seen as a generalization of the HKY model (the two parameters from the HKY model are implicitly included in the six from GTR model - \citealt{stamatakisa2008}).
 - TG: Is that useful?
+% NC: I think most people know that already.
 For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} %which is a generalization of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 - TG: same?
- assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v8.0.20; \citealt{Stamatakis21012014}).
+% NC: Again leave it out
+ assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
+8.0.20; \citealt{Stamatakis21012014}).
 
-In order to measure the phylogenetic signal of our simulations, we first ran a fast bootstrap analysis with 500 replicates on the "complete" matrix.
-We removed all the simulations that had a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
+To measure the phylogenetic signal of our simulations, we first ran a fast bootstrap analysis with 500 replicates on the "complete" matrix.
+We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
 We repeated this selection until we obtained 50 sets of simulations (i.e.
-50 "complete" and 50*125 "missing-data" matrices) with a relative good phylogenetic signal (median bootstrap $>$ 50).
-
-On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps per tree inference to assess the topological support \citep{pattengale2010many}.
+50 "complete" and 50 x 125 "missing-data" matrices) with a relative good phylogenetic signal (median bootstrap $>$ 50).
+On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \citep{pattengale2010many}.
 %The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists of pruning one sub-tree from the tree and subsequently reinserting it to all neighbouring branches \citep{stamatakisa2008}.
@@
 %TG: The explanations of the algorithm part for the BS was proposed by Trevor I think it's useless
-When using these parameters, it took $~$8 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (2.30GHz clock speed nodes).
+% NC: perhaps somewhere in brackets just mention it's LSR bootstrapping. No need for the rest I agree.
+Using these parameters took $~$8 % NC: again tilde isn't appearing on my PDF 
+CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (2.30GHz clock speed nodes).
 
 \subsubsection{Bayesian}
-The "best" tree and the "missing-data" trees were inferred using MrBayes v3.2.1 \citep{Ronquist2012mrbayes}.
+The "best" tree and the "missing-data" trees were inferred using MrBayes v.
+3.2.1 \citep{Ronquist2012mrbayes}.
 We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
-The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of 2 \citep{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
+The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \citep{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
 For the morphological data, we used the Markov \textit{k} state model \citep{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
@@
 
-Each Bayesian tree was estimated using two runs of four chains each for a maximum of 50$\times$$1^6$ generations.
+Each Bayesian tree was estimated using two runs of four chains each for a maximum of 50$\times$$10^6$ generations.
+% NC: Do you mean 50 x 10^6? or 5 x 10^7? It's standard to use 10^x not 1^x
  We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
@@
 For each run, we removed 25\% of the iterations as burn-in.
-We used the following priors for each tree (see \hyperref[SupplementaryMaterial]{Supplementary Material S1}): (i) the "true" tree’s topology as a starting tree (with a starting value for each branch length of 1), (ii) an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions, and (iii) a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)).
+We used the following priors for each tree (see \hyperref[SupplementaryMaterial]{Supplementary Material S1}
+ % NC: For what? What's in the SI?
+ ): (i) the "true" tree’s topology as a starting tree (with a starting value for each branch length of one), (ii) an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions, and (iii) a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)).
 
-We used these prior to speed up the Bayesian estimation process.
-These priors biased the way the Bayesian process calculated branch lengths by giving non-random starting points and boundaries for parameter estimation, however, here we are focusing on the effect of missing data on tree topology and not branch lengths.
-Even using these priors, it took $~$140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
+We used these priors to speed up the Bayesian estimation process.
+These priors biased the way the Bayesian process calculated branch lengths by giving non-random starting points and boundaries for parameter estimation however, here we are focusing on the effect of missing data on tree topology and not branch lengths.
+Even using these priors, it took $~$ % NC: Another tilde issue
+140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
 
@@
 We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
-We used the Robinson-Foulds distance \citep{RF1981} to assess the amount conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess amount of wildcard taxa \citep{kearneyfragmentary2002}.
-We then used Normalized Tree Similarity index \citep{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
+We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa \citep{kearneyfragmentary2002}.
+We then used the Normalized Tree Similarity index \citep{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
 These metrics are described in detail below.
@@
 \subsubsection{Robinson-Foulds distance}
-Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981} ; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
-This metric is bounded between 1 when the two trees are identical and $n-2$ (for two trees with $n$ taxa) when there is not one single shared clade between both trees.
-This metric is sensitive to the exact clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swap of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating a bad tree similarity.
+The Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees.
+The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}
+% NC: For what? What's in the SI?
+).
+This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
+This metric is sensitive to even minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
 
 \subsubsection{Triplets distance}
-The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two given trees (\citealt{critchlowthe1996} ; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
-This metric measures the position of each taxon and clade towards it's closest neighbours.
-It is bounded between 0 when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is not one single position of taxon/clade identical between both trees.
-Therefore this metric sensitive to the conservation of wildcard taxas.
+The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}
+% NC: For what? What's in the SI?
+).
+This metric measures the position of each taxon and clade in relation to its closest neighbours.
+It is bounded between 0 when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxon/clade position in the two trees.
+Therefore this metric sensitive to the conservation of wildcard taxa, i.e.
+taxa that frequently change position in different trees.
+% NC: You may have a better definition of wildcard taxa than mine :)
 
 \subsubsection{Normalized Tree Similarity}
-We used the Normalized Tree Similarity index, $NTS_m$ (\citealp{Bogdanowicz2012}) to be able to compare the two metrics for any $n$ taxa.
-This index allows to scale the value of any metric $m$ (either Robinson-Foulds or Triplets distance in our study) to the expected value of the metric $m$ when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
-When $NTS_m$=1, the two trees are strictly identical, when $NTS_m$=0 the trees are no more different than expected when comparing two random trees and when $NTS_m$$<$0, the difference between the two trees is greater than when comparing two random trees.
-In our study we used the $NTS_m$ index as a proxy for topology: a high score of this index (i.e.
-towards 1) means that the topology is highly conserved between the two trees; on the opposite, a low score of this index (i.e.
-towards 0) means that the topological difference between the two trees is as much as expected when comparing two random trees.
-Note that both tree comparisons are representing different aspects of topological similarity and do not perform equaly (see \hyperref[metrics_discussion]{discussion below} and \citealt{kuhnerpractical2014}).
+The Normalized Tree Similarity index, $NTS_m$ (\citealp{Bogdanowicz2012}) allows us to compare the two metrics above for any number of taxa, by scaling the value of the metric $m$ (either Robinson-Foulds or Triplets distance in our study) to the expected value of the metric $m$ when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}
+% NC: Why? For more details? Simulations etc
+).
+When $NTS_m$=1, the two trees are identical, when $NTS_m$=0 the trees are no more different than expected when comparing two random trees, and when $NTS_m$$<$0, the difference between the two trees is greater than when comparing two random trees.
+In our study we used the $NTS_m$ index as a proxy for topological similarity: a high score of this index (i.e.
+towards one) means that the topology is highly conserved between the two trees; conversely, a low score of this index (i.e.
+towards 0) means that the topological difference between the two trees is as great as expected when comparing two random trees.
+Note that the two tree comparison metrics represent different aspects of topological similarity and do not perform equally well (see \hyperref[metrics_discussion]{discussion below} and \citealt{kuhnerpractical2014}).
 
-\subsubsection{Tree comparisons}
+% NC: If I understand right, you use RF normalised and TD normalised.
+NOT RF, TD and NTS.
+So having these as three separate sections seems odd.
+Instead I'd write RF section, then explain that it is not comparable across trees with different numbers of taxa so you normalise it as suggested in Bogdanowicz's NTS method.
+Then describe TD and explain that you normalise it as discussed above for RF.
+NTS isn't really a metric, just a simple normalisation procedure.
+
+\subsubsection{Tree comparisons} % NC: Lost!!!!
 \label{tree_comparisons}
@@
 % TG: Maybe remove supplementary here
+% NC: Yeah probably. You could also fit this into the section above. Again NTS isn't really a metric in my opinion...
 %We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree.
@@
 
-Also, to take into account the uncertainty of tree inference in both Maximum Likelihood and Bayesian (i.e node support), we ran 1000 random pairwise comparison between respectively the bootstrapped trees from the Maximum Likelihood analysis and the posterior tree distribution of the Bayesian analysis.
+To take into account the uncertainty of tree inference in both Maximum Likelihood and Bayesian trees, we ran 1000 random pairwise comparisons among the bootstrapped trees from the Maximum Likelihood analysis and the posterior tree distribution of the Bayesian analysis.
 In the same way that we compared a single "missing-data" tree to the "best" tree (whether the trees are Maximum Likelihood or Bayesian consensus): we randomly selected 1000 trees from the "missing-data" tree sets (either the Bootstrapped trees or the posterior tree distribution) and did a pairwise comparison with 1000 randomly selected trees from the "best" tree set.
 
-For each of the 125 "missing-data" tree, we obtained Robinson-Foulds and Triplets distance distributions from either 50 or 50$\times$1000 pairwise comparisons.
-We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}.
+For each of the 125 "missing-data" trees, we obtained Robinson-Foulds and Triplets distance distributions from either 50 or 50$\times$1000 pairwise comparisons.
+We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v.
+3.1 \citep{hdrcde}.
 %Finally we tested for significant differences among the different distributions of trees using non-parametric group and pairwise comparisons by using respectively Kruskal-Wallis and Nemenyi-Damico-Wolfe-Dunn test (CITE R PACKAGE)\citet{ruxtontime2008}.
@@
 
-In order to investigate the effect of the parameters and/or the methods used in this simulations, we measured the similarity among the different distributions using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
+To investigate the effect of the parameters and/or the methods used in these simulations, we measured the similarity among the different distributions using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
 The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}).

### 2015-01-05 Manuscript/TEM_draft.tex
Minor comment added

@@
 
-\subsubsection{Tree comparisons} % NC: Lost!!!!
+\subsubsection{Tree comparisons} % NC: OK I'm lost in this section! We should discuss then think about how to express this in an understandable fashion.
 \label{tree_comparisons}
-For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Robinson-Foulds and Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012}.
+For the Maximum Likelihood and Bayesian consensus
+% NC: is this actually a maximum clade credibility tree
+trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Robinson-Foulds and Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012}.
 For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and n = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}).
@@
 
-To take into account the uncertainty of tree inference in both Maximum Likelihood and Bayesian trees, we ran 1000 random pairwise comparisons among the bootstrapped trees from the Maximum Likelihood analysis and the posterior tree distribution of the Bayesian analysis.
-In the same way that we compared a single "missing-data" tree to the "best" tree (whether the trees are Maximum Likelihood or Bayesian consensus): we randomly selected 1000 trees from the "missing-data" tree sets (either the Bootstrapped trees or the posterior tree distribution) and did a pairwise comparison with 1000 randomly selected trees from the "best" tree set.
+To take into account the uncertainty of tree inference in both Maximum Likelihood and Bayesian trees, % NC: instead of using the single ML or B MCC tree as above,
+we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 from the posterior tree distribution of the Bayesian analysis.
+% for the best trees and then did the same for the 125 missing data trees (resulting in 125 x 1000 missing data trees).
+For a given set of 1000 missing data trees, we sampled one missing data tree and one best tree at random and compared them using both the Robinson-Foulds and Triplets metrics as described above.
+We repeated this 1000 times for each set of missing data trees resulting in 125 x 1000 values for each metric.
 
-For each of the 125 "missing-data" trees, we obtained Robinson-Foulds and Triplets distance distributions from either 50 or 50$\times$1000 pairwise comparisons.
+We repeated all of the above for each of the 50 chains.
 We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v.
@@
 
-To investigate the effect of the parameters and/or the methods used in these simulations, we measured the similarity among the different distributions using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}).
-For each method (Maximym likelihood trees, Bayesian consensus trees, Bootstraps and Bayesian posterior trees) and metric (Robinson-Foulds and Triplets distance), we used this coefficient in two ways in order to:
+% NC: New subheading (summarising)
+
+% NC: Finally after doing all theses stupid comparisons and simulations we tested the effects of our missing data parameters (ML) on our ability to recover the correct topology. 
+To do this we measured the similarity among the different distributions % NC: OF WHAT!!! tree metrics for all of the analyses above
+using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
+The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} % FOR WHAT!!!).
+
+%For each method (Maximum Likelihood trees, Bayesian %MCC?
+%trees, Bootstraps and Bayesian posterior trees) and metric (Robinson-Foulds and Triplets distance), we used this coefficient in two ways in order to:
 \begin{enumerate}

### 2015-01-05 Manuscript/TEM_draft.tex
Edits to results

@@
 
+% Add in your three questions here?
 %---------------------------------------------
@@
 
-\subsection{Effect of missing data on topology}
+% NC: Split into three sections to answer three questions
+
+%\subsection{The effects of missing data on tree topology}
 
@@
     \includegraphics[width=1\textwidth]{Figures/In_main/AllMethods-RF+Tr-colour.pdf}
-\caption{Comparison between the effect of missing data and the tree inference method on topology.
+\caption{Comparison between the effect of missing data and the tree inference method on topological recovery.
 The amount of missing data for each parameter is represented on the x axis.
-The topology is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
-Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
-The Maximum Likelihood trees are represented in black, the Bayesian consensus trees in red, the bootstrap trees in green and the posterior tree distribution in blue.}
+The difference in topologies is represented on the y axis, using standardized Robinson-Foulds distance (upper row) and standardized Triplets distance (lower row).
+Points represent the modal value of each distribution; thick solid and thin dashed lines represents the 50\% and 95\% confidence intervals or the distributions respectively.
+Analyses using the Maximum Likelihood trees are represented in black, the Bayesian consensus %MCC?
+trees in red, the Maximum Likelihood bootstrap trees in green and the Bayesian posterior tree distribution in blue.}
 \label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
@@
 
-As expected form the literature, the amount of missing data in the morphological part of the total evidence matrix does decrease the ability to recover the right topology, regardless of the parameter, the method or the metric used \citep[e.g.][]{rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014}.
-However each variable does not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam}).
-Regarding the conservation of entire clades (i.e.
-the Ronbinson-Foulds distance), the Bayesian consensus trees outperform the other methods and the amount of missing data in the living taxa ($M_{L}$) decreases the most rapidly the clade conservation when using this method.
-However, when looking at the position of wildcard taxa (i.e.
-using the Triplets distance), the Maximum Likelihood outperforms the other methods but with wider distributions overlap (Fig.~\ref{Fig_Results-permeth_perparam}).
+% NC: Y axes should be standardised Robinson Foulds
+% NC: X-axes should have (Mc) etc in brackets after the names
+% NC: Also should probably have the X axes read : Missing character (Mc, %) and remove % from tick labels
+
+%NC: I would rewrite these figure labels as follows. I prefer the details of colours and things to be in brackets not in the text. Rewrite the others in a similar fashion.
+%NC: Remeber it's not topology, but topological recovery that you are measuring
+%The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (green) and Bayesian posterior tree distributions (blue).
+The percentage of missing data for each parameter (Mc, ML, MF) is shown on the x axis.
+Topological recovery was measured using two different tree comparison metrics: standardized Robinson-Foulds distance (upper row) and standardized Triplets distance (lower row).
+The graph shows the modal value (points), the 50\% and 95\% confidence intervals (thick solid lines) and the distributions (thin dashed lines), of the tree comparison metric for each missing data parameter and tree inference method.
+
+As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the correct topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{C}$ or $M_{F}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (standardized Robinson Foulds or standardized Triplets distance).
+% NC: Results are for YOUR results. compare it to others in the Discussion only
+% NC: I've rephrased below as it was getting confused between what is GOOD and BAD for topological recovery and the difference between the parameters and the methods. I've tried to clarify.
+% NC: Note that you need to be clear what the methods are outperforming...
+However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam}). % And figure 3 too right?
+
+%Increasing the amount of missing data in the living taxa ($M_{L}$) has the worst effect on topological recovery, though only when using the standardized Robinson Foulds distance.
+Using the standardized Robinson Foulds distance, which represents clade conservation between trees, we get the best topological recovery using Bayesian consensus trees.
+However, using standardized Triplets distance, which represents the movement of wildcard taxa, Maximum Likelihood outperforms the other methods but with wider distributional overlap (Fig.~\ref{Fig_Results-permeth_perparam}).
+%NC: What distributions?
+Do you need to say that?
 
@@
     \includegraphics[width=1\textwidth]{Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
-\caption{Trend of the effect of missing data on topology on the Maximum Likelihood and the Bayesian consensus trees.
+\caption{Trend of the effect of missing data on topological recovery on the Maximum Likelihood and the Bayesian consensus trees.
 The amount of missing data per parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is represented along the x axis.
@@
 
-When looking at the effect of all parameters combinations on the Bayesian consensus trees, this method outperforms Maximum Likelihood for the Ronbinson-Foulds distance and plateaus to a minimal modal tree similarity of 0.69 regardless the amount of missing data (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+%When looking at the effect of all parameters combinations on the Bayesian consensus trees, this method outperforms Maximum Likelihood for the Robinson-Foulds distance and plateaus to a minimal modal tree similarity of 0.69 regardless the amount of missing data (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 However, regarding the Triplets distance, Bayesian consensus trees seems to perform poorly but with great confidence intervals overlaps (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 Results for the global trend comparison between the uncertainty methods (Bootstrap and Bayesian posterior tree distribution) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+% NC: Move what you need of this to the sections below.
+% NC: Remember you're not testing the RF and TD, you're seeing that the loss of data affects clade conservation and wildcard taxa separately.
 
-\subsection{Effect of the method on topology}
-
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{Figures/In_main/BC-MethodComp-RF+Tr-colour.pdf}
-\caption{Distribution of the Bhattacharyya Coefficients between methods.
-Curves represents the kernel density estimations of the Bhattacharyya Coefficients between the four different methods.
-A.
-Results for the Normalised Robinson-Foulds distance.
-B.
-Results with the for Normalised Triplets distance.
-Pikes around the two different extremes values of the Bhattacharyya Coefficients show the two pairs of most dissimilar methods (Bayesian consensus vs.
-Bootstraps and Bayesian consensus vs.
-Bayesian posterior trees) for the Normalised Robinson-Foulds distance ; and the most similar methods (Bootstraps and Bayesian posterior trees) for the Normalised Triples methods.}
-\label{Fig_Results-paircomp_without} %Comparisons between methods, both in RF and Triplets.(method differences)
-\end{figure}
-
-When looking at the comparisons between the methods using the Robinson-Foulds distance, the Bayesian consensus trees have the least overlap with respectively the Bayesian posterior trees and the Bootstrap (Fig.~\ref{Fig_Results-paircomp_without}-A).
-When regarding at the Triplets distance, the two method that have the most distribution are the Bootstraps and the Bayesian posterior trees (Fig.~\ref{Fig_Results-paircomp_without}-B).
-These results are due to the difference in comparing topologies between Maximum Likelihood and Bayesian consensus trees and Bootstraps and Bayesian posterior trees: the two last ones are based on a random pairwise comparison of a sub-sample of 1000 trees.
-This is likely to add noise to the data.
-
-\subsection{Effect of the missing data parameters on topology}
+\subsection{Effects of the missing data parameters on topological recovery}
 
@@
 
-When looking at the pairwise Bhattacharyya Coefficients between all the parameters combinations for Bayesian consensus trees, the number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}-A).
+%Increasing the amount of missing data in the living taxa ($M_{L}$) has the worst effect on topological recovery, though only when using the standardized Robinson Foulds distance.
+
+%NC: I'd start by describing the pattern with reference to figs 2 and 3 and then the "significant" results
+
+When looking at the pairwise Bhattacharyya Coefficients between all the parameters combinations for Bayesian consensus trees, % NC: Instead just state the result.
+"We found that number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (table X)".
+In the same way as you wouldn't say "Using a linear regression we found that x ~ Y." You'd say "X and Y were significantly correlated (Table1)".
+%NC: Also what IS the result really?
+You find that ML overlaps less with the other parameters so it's significantly different in it's effect?
+And from the graph you can see it's worse?
+What does oerlap MEAN in terms of the answer to the question, i.e.
+how do the parameters affect topological recovery.
+
+the number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}-A).
 However, when looking at the Triplets distance, because of the high overlap in confidence intervals, there is no major region with low probability of overlap.
@@
 
+\subsection{Effects of the tree inference method on topological recovery}
+
+%NC: I'd start by describing the pattern with reference to figs 2 and 3 and then the "significant" results
+
+%\begin{figure} 
+%\centering
+  %  \includegraphics[width=1\textwidth]{Figures/In_main/BC-MethodComp-RF+Tr-colour.pdf}
+%\caption{Distribution of the Bhattacharyya Coefficients between methods.
+Curves represents the kernel density estimations of the Bhattacharyya Coefficients between the four different methods.
+A.
+Results for the Normalised Robinson-Foulds distance.
+B.
+Results with the for Normalised Triplets distance.
+Pikes around the two different extremes values of the Bhattacharyya Coefficients show the two pairs of most dissimilar methods (Bayesian consensus vs.
+Bootstraps and Bayesian consensus vs.
+Bayesian posterior trees) for the Normalised Robinson-Foulds distance ; and the most similar methods (Bootstraps and Bayesian posterior trees) for the Normalised Triples methods.}
+%\label{Fig_Results-paircomp_without} %Comparisons between methods, both in RF and Triplets.(method differences)
+%\end{figure}
+
+%When looking at the comparisons between the methods 
+using the Robinson-Foulds distance, the Bayesian consensus trees have the least overlap with respectively the Bayesian posterior trees and the Bootstrap (Fig.~\ref{Fig_Results-paircomp_without}-A).
+Using the Triplets distance, the two methods that have the most distribution %NC: ???
+are the Bootstraps and the Bayesian posterior trees (Fig.~\ref{Fig_Results-paircomp_without}-B).
+These results are due to the difference in comparing topologies between Maximum Likelihood and Bayesian consensus trees and Bootstraps and Bayesian posterior trees: the latter two are based on a random pairwise comparison of a sub-sample of 1000 trees.
+This is likely to add noise to the data.
+
+\subsection{Interactions among missing data parameters and the tree inference method on topological recovery}
+
+%NC: Maybe don't need this section. See how it goes.
+
+% NC: Make sure you're consistent in how you refer to things ie Maximum Likelihood bootstrap trees, not Bootstraps
 %---------------------------------------------
@@
 %---------------------------------------------
+
+% NC: Quick comment.
+You find that Bayesian consensus trees are "best".
+BUT I think you need to be careful here with what you mean by "best".
+They are most likely to recover a more similar topology to the best tree than if you do the crazy tree comparison pairs.
+But that doesn't mean they're best, just that they perform best in your simulations.
+They are only best for that.
+If you want to advise people to use a method that best recovers topology as measured using RF or Triplets, then Bayesian consensus is best.
+BUT generally that is not what people want.
+Probably need to think some more about this.
 
@@
 %   -Missing data increase leads to decreases in recovering the correct topology
-The results of our simulations are in adequacy with recent analysis on the effect of missing data in morphological matrices \citep{pattinsonphylogeny2014,wrightbayesian2014,sansombias2014}.
-Our results are showing that the ability to recover the correct topology decreases with the increase of missing data, regardless the metric, the method or the parameters considered.
+%The results of our simulations are in adequacy with recent analysis on the effect of missing data in morphological matrices \citep{pattinsonphylogeny2014,wrightbayesian2014,sansombias2014}. 
+Our results show that our ability to recover the correct topology decreases with increasing missing data, regardless of the missing data parameter ($M_{L}$, $M_{C}$ or $M_{F}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (standardized Robinson Foulds or standardized Triplets distance).
 %Results 1
 %   -However, parameters have different effect
-However, within a method and a metric, the parameters used to remove data have different effects on the ability to recover the correct topology.
+However, within a tree inference method and a metric, the parameters used to remove data have different effects on the ability to recover the correct topology.
 %Results 2

### 2015-01-05 Manuscript/TEM_draft.tex
Updated BC section

--- 2015-01-05 Manuscript/TEM_draft.tex
@@
+%LaTeX template : http://systbio.org/files/SB_LaTeX_Template_txt_extension.txt
+%Author instructions: http://www.oxfordjournals.org/our_journals/sysbio/for_authors/ms_preparation.html
+
 \documentclass[12pt,letterpaper]{article}
@@
 
-The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,sansombias2014}.
+%TG: I completely removed this paragraph
+%The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,sansombias2014}.
 Missing molecular data has been seen by some authors as an issue because it can, in some parts of the tree, decrease phylogenetic signal (i.e.
 the evolutionary information contained within the matrix allowing the inference of topology and branch lengths), especially when using large matrices \citep{lemmonthe2009}.
-However, this may not be a major issue because phylogenetic signal is easily increased by: (i) including a "modest" number ($~$ % NC: This symbol doesn't show up on my PDF.
+However, this may not be a major issue because phylogenetic signal is easily increased by: (i) including a "modest" number (\texttildelow % NC: This symbol doesn't show up on my PDF.
 Check.
-50) of highly-covered genes (i.e.
+%50) of highly-covered genes (i.e.
 a number of genes that are available for all taxa; \citealt{rouresite-specific2011}); (ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and (iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
@@
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil species.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case) \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
-Until now, no attempt has been made to study the impact of this issue on phylogenetic inference from Total Evidence methods.
+Until now, no attempt has been made to study the impact of this issue on phylogenetic inference using Total Evidence methods.
 % NC: THIS SECTION NEEDS WORK!
-In this study, we focus on the effect of missing data on the ability to recover a correct topology, 
+It is expected that missing in this framework will affect the ability to recover the correct topology as well as the correct branch length estimation \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,sansombias2014}.
+In this study, we focus only on the effect of missing data on the ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies \citep[e.g.][]{manosphylogeny2007,meredithimpacts2011,jetzthe2012}.
+%TG or at some more specific example instead of the cite: "In this study, we focus only on the effect of missing data on the ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies (e.g.
+for taxonomical focus \citep{meredithimpacts2011,jetzthe2012}, or for studying evolutionary trends %CITE?\citep[e.g.][]{})"?
+Even though branch length is another primordial aspect in phylogenies (namely for timing extinction and/or speciation events \citep[e.g.][]{ronquista2012}), previous studies have empirically assessed the effect of the Total Evidence method on branch length variation \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014} %TG: add:"by using a fixed topology"?
+Moreover, variation in branch length are more complex to simulate and to measure when topology varies.
 % NC: I don't get the logic of this part 
@@
 % NC: Need a better justification of WHY topology is interesting.
-Although correctly inferring branch lengths is also important, we do not consider the inference of branch lengths here.
+
+%Although correctly inferring branch lengths is also important, we do not consider the inference of branch lengths here.
 % NC: WHY??? Difficulties with simulations? Topology is more important? WHY?
-Also, the effect of Total Evidence method has not been formally assessed in previous studies using fixed topology \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+%Also, the effect of Total Evidence method has not been formally assessed in previous studies using fixed topology \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 % NC: OK this last sentence could be worked in earlier on to explain WHY topology not branch lengths.
@@
 % NC: Can you centre these around the boxes? and make the comparing topologies arrow touch the two triangles? I'll show you what I mean. 
+% TG: I have different versions for the figure available (you can see them opening it with inkscape I think - else I'll show you).
 % NC: Also check the figure is in a format and of correct quality for Syst Biol.
@@
 %---------------------------------------------
-%       Generating the matrix
+%       1-Generating the matrix
 %---------------------------------------------
 
-\subsection{Generating the matrix} % NC: Can you number these subsections like they are in the steps above?
+\subsection{1-Generating the matrix} % NC: Can you number these subsections like they are in the steps above? % TG: 
 First we randomly generated a "true" tree of 50 taxa in R v.
@@
 %---------------------------------------------
-%       Removing data
+%       2-Removing data
 %---------------------------------------------
 
-\subsection{Removing data}
+\subsection{2-Removing data}
 \label{Removing_data}
-We modified the "complete" matrix to get matrices with missing data by randomly replacing data with "?" in the morphological part of the matrices according to the following parameters:% (Fig.
-~\ref{Fig_RemoveData}): % PUT IN APPENDIX
+We modified the "complete" matrix to get matrices with missing data by randomly replacing data with "?" in the morphological part of the matrices according to the following parameters:
 
@@
 %---------------------------------------------
-%       Building phylogenies
+%       3-Building phylogenies
 %---------------------------------------------
 
-\subsection{Building phylogenies}
+\subsection{3-Building phylogenies}
 From the resulting matrices we generated two types of trees: the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
@@
 % NC: perhaps somewhere in brackets just mention it's LSR bootstrapping. No need for the rest I agree.
-Using these parameters took $~$8 % NC: again tilde isn't appearing on my PDF 
-CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (2.30GHz clock speed nodes).
+Using these parameters took \texttildelow 8 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (2.30GHz clock speed nodes).
 
@@
 
-Each Bayesian tree was estimated using two runs of four chains each for a maximum of 50$\times$$10^6$ generations.
-% NC: Do you mean 50 x 10^6? or 5 x 10^7? It's standard to use 10^x not 1^x
+Each Bayesian tree was estimated using two runs of four chains each for a maximum of 5$\times$$10^7$ generations.
+We used the following priors for each tree: (i) the "true" tree’s topology as a starting tree (with a starting value for each branch length of one), (ii) an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions, and (iii) a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)).
+We used these priors to speed up the Bayesian estimation process.
+These priors biased the way the Bayesian process calculated branch lengths by giving non-random starting points and boundaries for parameter estimation however, here we are focusing on the effect of missing data on tree topology and not branch lengths.
+Even using these priors, it took \texttildelow 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
+The detailed MrBayes parameters are available in the \hyperref[SupplementaryMaterial]{Supplementary Material S1}.
+
 We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
 The effective sample size (ESS) was also checked on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
-For each run, we removed 25\% of the iterations as burn-in.
-We used the following priors for each tree (see \hyperref[SupplementaryMaterial]{Supplementary Material S1}
- % NC: For what? What's in the SI?
- ): (i) the "true" tree’s topology as a starting tree (with a starting value for each branch length of one), (ii) an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions, and (iii) a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)).
-
-We used these priors to speed up the Bayesian estimation process.
-These priors biased the way the Bayesian process calculated branch lengths by giving non-random starting points and boundaries for parameter estimation however, here we are focusing on the effect of missing data on tree topology and not branch lengths.
-Even using these priors, it took $~$ % NC: Another tilde issue
-140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-
+Finally we built the Bayesian consensus tree from the combined chains excluding the 25\% first iterations (burn-in) with a strict majority rule \citep{Ronquist2012mrbayes}.
 %---------------------------------------------
-%       Comparing Topologies
+%       4-Comparing Topologies
 %---------------------------------------------
 
-\subsection{Comparing topologies}
+\subsection{4-Comparing topologies}
 We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
-We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa \citep{kearneyfragmentary2002}.
-We then used the Normalized Tree Similarity index \citep{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
+We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa  (i.e.
+taxa that frequently change position in different trees - \citealt{kearneyfragmentary2002}).
+We normalised both metrics using the Normalized Tree Similarity index \citep{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
 These metrics are described in detail below.
@@
 The Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}
-% NC: For what? What's in the SI?
-).
+The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for calculation details).
 This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
-This metric is sensitive to even minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
+This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
+We normalised this metric using the Normalized Tree Similarity index \citep{Bogdanowicz2012}.
+This index allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for the calculation details).
+Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the Normalised Robison-Foulds distance is equal to zero, the trees are no more different than expected by chance; finally if the Noramlised Robison-Foulds distance is lower than zero, the trees are more different than expected by chance.
 
 \subsubsection{Triplets distance}
-The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2}
-% NC: For what? What's in the SI?
-).
+The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for calculation details).
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
 It is bounded between 0 when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxon/clade position in the two trees.
-Therefore this metric sensitive to the conservation of wildcard taxa, i.e.
-taxa that frequently change position in different trees.
-% NC: You may have a better definition of wildcard taxa than mine :)
+Therefore this metric sensitive to the conservation of wildcard taxa.
+We normalised this metric in the same way as we did for the Robison-Foulds distance resulting in the Normalised Triplets distance.
 
-\subsubsection{Normalized Tree Similarity}
-The Normalized Tree Similarity index, $NTS_m$ (\citealp{Bogdanowicz2012}) allows us to compare the two metrics above for any number of taxa, by scaling the value of the metric $m$ (either Robinson-Foulds or Triplets distance in our study) to the expected value of the metric $m$ when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}
-% NC: Why? For more details? Simulations etc
-).
+%\subsubsection{Normalized Tree Similarity}
+%The Normalized Tree Similarity index, $NTS_m$ (\citealp{Bogdanowicz2012}) allows us to compare the two metrics above for any number of taxa, by scaling the value of the metric $m$ (either Robinson-Foulds or Triplets distance in our study) to the expected value of the metric $m$ when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
 When $NTS_m$=1, the two trees are identical, when $NTS_m$=0 the trees are no more different than expected when comparing two random trees, and when $NTS_m$$<$0, the difference between the two trees is greater than when comparing two random trees.
@@
 
-% NC: If I understand right, you use RF normalised and TD normalised.
-NOT RF, TD and NTS.
-So having these as three separate sections seems odd.
-Instead I'd write RF section, then explain that it is not comparable across trees with different numbers of taxa so you normalise it as suggested in Bogdanowicz's NTS method.
-Then describe TD and explain that you normalise it as discussed above for RF.
-NTS isn't really a metric, just a simple normalisation procedure.
-
-\subsubsection{Tree comparisons} % NC: OK I'm lost in this section! We should discuss then think about how to express this in an understandable fashion.
+\subsubsection{Paired trees comparisons} % NC: OK I'm lost in this section! We should discuss then think about how to express this in an understandable fashion. TG: or no subheading at all?
 \label{tree_comparisons}
 For the Maximum Likelihood and Bayesian consensus
-% NC: is this actually a maximum clade credibility tree
-trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Robinson-Foulds and Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012}.
-For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and n = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}).
-% TG: Maybe remove supplementary here
-% NC: Yeah probably. You could also fit this into the section above. Again NTS isn't really a metric in my opinion...
+% NC: is this actually a maximum clade credibility tree - TG: better stick with consensus as described above?
+trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting each time in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance.
+%For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and n = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}).
+% TG: That is now described in the RF distance subsection just above
 %We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree.
@@
 TG: explained later on
-
-To take into account the uncertainty of tree inference in both Maximum Likelihood and Bayesian trees, % NC: instead of using the single ML or B MCC tree as above,
-we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 from the posterior tree distribution of the Bayesian analysis.
-% for the best trees and then did the same for the 125 missing data trees (resulting in 125 x 1000 missing data trees).
-For a given set of 1000 missing data trees, we sampled one missing data tree and one best tree at random and compared them using both the Robinson-Foulds and Triplets metrics as described above.
-We repeated this 1000 times for each set of missing data trees resulting in 125 x 1000 values for each metric.
-
-We repeated all of the above for each of the 50 chains.
+Also, to take into account the uncertainty of tree inference in both method, instead of using the single Maximum Likelihood and Bayesian consensus trees as above, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 from the posterior tree distribution of the Bayesian analysis for the "best" trees and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees).
+For a given set of 1000 "missing data" trees and the 1000 "best" trees, we sampled one "missing data" tree and one "best" tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above.
+We repeated this 1000 times for each set of "missing data" trees resulting in 125$\times$1000 values for each metric.
+We repeated all the paired trees comparisons described above for each of the 50 simulations.
 We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v.
@@
 
-% NC: New subheading (summarising)
-
-% NC: Finally after doing all theses stupid comparisons and simulations we tested the effects of our missing data parameters (ML) on our ability to recover the correct topology. 
+\subsubsection{Summarizing the comparisons}
+Finally after doing all theses %stupid
+50 simulations and comparisons within, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) on the ability to recover the correct topology in a Total Evidence framework.
+% TG: the following part might be useless
+To
 To do this we measured the similarity among the different distributions % NC: OF WHAT!!! tree metrics for all of the analyses above

### 2015-01-06 Manuscript/TEM_draft.tex
Updated intro/methods/results
Results only for the table

@@
 50 simulations and comparisons within, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) on the ability to recover the correct topology in a Total Evidence framework.
-% TG: the following part might be useless
-To
-To do this we measured the similarity among the different distributions % NC: OF WHAT!!! tree metrics for all of the analyses above
-using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} % FOR WHAT!!!).
+Additionally, we also assessed the effect of the different methods (i.e.
+Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).% and the joint effect of the our missing data parameters and the methods.
+% TG: the following part might be useless (complete repetition of above)
+To assess the ability to recover the "correct" tree topology on our 50 simulations, we used the scores for both normalised metrics (Robinson-Foulds and Triplets distance).
+We looked at the following effects:
+\begin{enumerate}
+\item{The three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$);}
+\item{The methods (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution);}
+%\item{The joint effect of the missing data parameters and the methods.} actually this is implied in both already.
+\end{enumerate}
+% TG: end of the repetition
 
-%For each method (Maximum Likelihood trees, Bayesian %MCC?
-%trees, Bootstraps and Bayesian posterior trees) and metric (Robinson-Foulds and Triplets distance), we used this coefficient in two ways in order to:
-\begin{enumerate}
-\item{Assess the effect of the method on all parameters:}
+%Introducing Bhattacharyya Coefficient
+To evaluate these effects we measured the similarities among the different normalised metrics scores (Robinson-Foulds and Triplets distance) for all the analysis above using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
+The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} for calculation details).
 
-We performed a pairwise comparison of each parameter between each pair of methods.
-This resulted in 125 comparisons per method and per metric (each parameter combination within one method compared to the same parameter combination in the other method).
-We then compared the distribution of these pairwise comparisons to see the global difference between two methods: if the method have really similar results, we would expect the distribution to be cluster around 1 (high probability of overlap between the distributions) and if the methods are really dissimilar around 0 (low probability of overlap - Fig.~\ref{Fig_Results-paircomp_without}).
-\item{Assess the effect of the parameters within a method:}
+%Explaining the parameter pairwise comparisons (ANOVA equivalent=within)
+For assessing the effect of our three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$), we calculated the Bhattacharyya Coefficient between each pairwise combination for each parameter state (0\%, 10\%, 25\%, 50\%, 75\%).
+This resulted in 7875 pairwise distributions pairwise comparisons (triangular $3^5$$\times$$3^5$ matrix).
+We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient=1) and in red when they don't (Bhattacharyya Coefficient=0) - Fig.~\ref{Fig_Results-paircomp_within}).
+We repeated this procedure for all the methods (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).
 
-We performed a pairwise comparison per method and per metric for each combination of parameters.
-This resulted in 7875 pairwise comparisons per method and per metric (triangle of a 125$\times$125 matrix).
-We represented these results as a triangular matrix with the values of each pairwise comparison coloured according to the value of the Batthacharyya Coefficient (coloured in green when the distributions overlap completely and in red when they don't overlap at all- Fig.~\ref{Fig_Results-paircomp_within}).
-\end{enumerate}
+%Explaining the methods comparisons (ANOVA equivalent=among)
+Finally, for assessing the effect of the method on the ability to recover the "correct" topology, we calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g.
+Maximum Likelihood\textit{versus} Bayesian consensus).
+We then reported the minimum, the 1st quartile, the median, the mean, the 3rd quartile and the maximum of the probabilities of overlap between the methods (i.e.
+the Bhattacharyya Coefficients) in table ~\ref{Tab_Results-Difference_methods}.
+When two methods display similar normalised metrics scores for each comparison between the "best" tree and the "missing data" trees, we expect the Bhattacharyya Coefficients to be close to 1.
+On the opposite, if the two methods displays dissimilar normalised metrics scores, we expect the Bhattacharyya Coefficients to be close to 0.
 
-% Add in your three questions here?
 %---------------------------------------------
@@
 %\end{figure}
+
+\begin{table}[ht]
+\label{Tab_Results-Difference_methods}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data".}
+\centering
+\begin{tabular}{lccccccc}
+  \hline
+ Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
+    Bayesian consensus              & $Tr^2$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
+    Bootstraps                      & $Tr^2$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
+    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
+    Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
+    Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
+    bootstraps                      & $Tr^2$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
+    Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
+    Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
+    Boostraps \textit{vs.}          & $RF^1$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
+    Bayesian posterior tree         & $Tr^2$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
+   \hline
+\end{tabular}
+   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
+\end{table}
 

### 2015-01-06 Analysis/MissingData_differences.R

@@ -146,3 +146,33 @@ names(pair.comp_without.Tr)<-c("MLbest vs Baycon", "MLbest vs MLboot",
 op<-par(bty="l")
 plot.bhatt.coeff.dist(pair.comp_without.RF, col=palette(), xlab="Bhattacharyya Coefficient", ylab="Frequency")
 plot.bhatt.coeff.dist(pair.comp_without.Tr, col=palette(), xlab="Bhattacharyya Coefficient", ylab="Frequency")
+
+#Summarizing these distributions in a table
+
+
+
+#Creating the methods table
+table.tmp<-data.frame("Maximum likelihood vs. Bayesian consensus - RF"=as.vector(summary(pair.comp_without.MLbest.Baycon.RF)),
+                      "Maximum likelihood vs. Bayesian consensus - Tr"=as.vector(summary(pair.comp_without.MLbest.Baycon.Tr)),
+                      "Maximum likelihood vs. Bootstraps - RF"=as.vector(summary(pair.comp_without.MLbest.MLboot.RF)),
+                      "Maximum likelihood vs. Bootstraps - Tr"=as.vector(summary(pair.comp_without.MLbest.MLboot.Tr)),
+                      "Maximum likelihood vs. Bayesian posterior trees - RF"=as.vector(summary(pair.comp_without.MLbest.Baytre.RF)),
+                      "Maximum likelihood vs. Bayesian posterior trees - Tr"=as.vector(summary(pair.comp_without.MLbest.Baytre.Tr)),
+                      "Bayesian consensus vs. Bootstraps - RF"=as.vector(summary(pair.comp_without.Baycon.MLboot.RF)),
+                      "Bayesian consensus vs. Bootstraps - Tr"=as.vector(summary(pair.comp_without.Baycon.MLboot.Tr)),
+                      "Bayesian consensus vs. Bayesian posterior trees - RF"=as.vector(summary(pair.comp_without.Baycon.Baytre.RF)),
+                      "Bayesian consensus vs. Bayesian posterior trees - Tr"=as.vector(summary(pair.comp_without.Baycon.Baytre.Tr)),
+                      "Boostraps vs. Bayesian posterior trees - RF"=as.vector(summary(pair.comp_without.MLboot.Baytre.RF)),
+                      "Boostraps vs. Bayesian posterior trees - Tr"=as.vector(summary(pair.comp_without.MLboot.Baytre.Tr))
[…]

### 2015-01-06 Manuscript/TEM_draft.tex
Modified intro+method
Intro: updated paragraph on the why only topology
Method: updated NTS + Tree comparisons + BC explanations

@@
 Until now, no attempt has been made to study the impact of this issue on phylogenetic inference using Total Evidence methods.
+
 % NC: THIS SECTION NEEDS WORK!
-It is expected that missing in this framework will affect the ability to recover the correct topology as well as the correct branch length estimation \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,sansombias2014}.
-In this study, we focus only on the effect of missing data on the ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies \citep[e.g.][]{manosphylogeny2007,meredithimpacts2011,jetzthe2012}.
-%TG or at some more specific example instead of the cite: "In this study, we focus only on the effect of missing data on the ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies (e.g.
-for taxonomical focus \citep{meredithimpacts2011,jetzthe2012}, or for studying evolutionary trends %CITE?\citep[e.g.][]{})"?
-Even though branch length is another primordial aspect in phylogenies (namely for timing extinction and/or speciation events \citep[e.g.][]{ronquista2012}), previous studies have empirically assessed the effect of the Total Evidence method on branch length variation \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014} %TG: add:"by using a fixed topology"?
-Moreover, variation in branch length are more complex to simulate and to measure when topology varies.
-% NC: I don't get the logic of this part 
-% because topology is the first and most straightforward aspect reflecting phylogenetic signal (i.e. topological changes are discrete opposed to branch length changes are continuous). 
-% NC: Need a better justification of WHY topology is interesting.
-
-%Although correctly inferring branch lengths is also important, we do not consider the inference of branch lengths here.
-% NC: WHY??? Difficulties with simulations? Topology is more important? WHY?
-%Also, the effect of Total Evidence method has not been formally assessed in previous studies using fixed topology \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
-% NC: OK this last sentence could be worked in earlier on to explain WHY topology not branch lengths.
+In this study, we focus on the effect of missing data on our ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{}.
+%CITE-check Lloyd and Friedman or birdy stuff
+Although branch length estimation is also important (namely for timing extinction and/or speciation events - e.g.
+\citealt{ronquista2012}), we do not consider branch lengths in this study.
+This is partially due to difficulties of estimating the effect of missing data simultaneously on branch length and topology, % TG: or "This is partially due to difficulties with simulating branch lengths and topology simultaneously"
+but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework.
+To our knowledge, this question has never been formally assessed.
 
@@
 To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
-% NC: I've added numbers in here (125 matrices etc) so it matches the figure legend better. Can you check I've put the right numbers in? I think it may be 124 in some cases...
 \begin{enumerate}
@@
 This index allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for the calculation details).
-Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the Normalised Robison-Foulds distance is equal to zero, the trees are no more different than expected by chance; finally if the Noramlised Robison-Foulds distance is lower than zero, the trees are more different than expected by chance.
+Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the Normalised Robinson-Foulds distance is equal to zero, the trees are no more different than expected by chance; finally if the Normalised Robinson-Foulds distance is lower than zero, the trees are more different than expected by chance.
 
@@
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
-It is bounded between 0 when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxon/clade position in the two trees.
+It is bounded between 0 when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
 Therefore this metric sensitive to the conservation of wildcard taxa.
-We normalised this metric in the same way as we did for the Robison-Foulds distance resulting in the Normalised Triplets distance.
+We normalised this metric in the same way as we did for the Robinson-Foulds distance resulting in the Normalised Triplets distance.
 
@@
 
-\subsubsection{Paired trees comparisons} % NC: OK I'm lost in this section! We should discuss then think about how to express this in an understandable fashion. TG: or no subheading at all?
+\subsubsection{Paired trees comparisons} % NC: OK I'm lost in this section!
+We should discuss then think about how to express this in an understandable fashion.
+TG: or no subheading at all?
+See next section as well.
 \label{tree_comparisons}
@@
 TG: explained later on
-Also, to take into account the uncertainty of tree inference in both method, instead of using the single Maximum Likelihood and Bayesian consensus trees as above, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 from the posterior tree distribution of the Bayesian analysis for the "best" trees and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees).
+Also, to take into account the uncertainty of tree inference in both methods cited above, instead of using the single Maximum Likelihood and Bayesian consensus trees, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 from the posterior tree distribution of the Bayesian analysis for the "best" trees and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees).
 For a given set of 1000 "missing data" trees and the 1000 "best" trees, we sampled one "missing data" tree and one "best" tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above.
@@
 Finally after doing all theses %stupid
-50 simulations and comparisons within, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) on the ability to recover the correct topology in a Total Evidence framework.
+50 simulations and comparisons, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) on the ability to recover the correct topology in a Total Evidence framework.
 Additionally, we also assessed the effect of the different methods (i.e.
@@
 \item{The three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$);}
-\item{The methods (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution);}
+\item{The four "methods" (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution);}
 %\item{The joint effect of the missing data parameters and the methods.} actually this is implied in both already.
@@
 We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient=1) and in red when they don't (Bhattacharyya Coefficient=0) - Fig.~\ref{Fig_Results-paircomp_within}).
-We repeated this procedure for all the methods (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).
+We repeated this procedure for all the "methods" (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).
 
 %Explaining the methods comparisons (ANOVA equivalent=among)
-Finally, for assessing the effect of the method on the ability to recover the "correct" topology, we calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g.
+Finally, for assessing the effect of the method on the ability to recover the "correct" topology, we compared the similarity between the methods.
+We calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g.
 Maximum Likelihood\textit{versus} Bayesian consensus).
@@
 the Bhattacharyya Coefficients) in table ~\ref{Tab_Results-Difference_methods}.
-When two methods display similar normalised metrics scores for each comparison between the "best" tree and the "missing data" trees, we expect the Bhattacharyya Coefficients to be close to 1.
-On the opposite, if the two methods displays dissimilar normalised metrics scores, we expect the Bhattacharyya Coefficients to be close to 0.
+When two methods perform similarly in their ability to recover the "correct" tree topology (i.e.
+they display similar normalised metrics scores for each comparison between the "best" tree and the "missing data" trees), we expect the Bhattacharyya Coefficients to be close to 1.
+On the opposite, if the two methods differ, we expect the Bhattacharyya Coefficients to be close to 0.
 
@@
 \begin{table}[ht]
-\label{Tab_Results-Difference_methods}
 \caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data".}
@@
    $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
+\label{Tab_Results-Difference_methods}
 \end{table}

### 2015-01-06 Manuscript/TEM_draft.tex
Results section update

@@
 \item{The four "methods" (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution);}
-%\item{The joint effect of the missing data parameters and the methods.} actually this is implied in both already.
+%\item{The joint effect of the missing data parameters and the methods.} % actually this is implied in both already.
 \end{enumerate}
@@
 We then reported the minimum, the 1st quartile, the median, the mean, the 3rd quartile and the maximum of the probabilities of overlap between the methods (i.e.
-the Bhattacharyya Coefficients) in table ~\ref{Tab_Results-Difference_methods}.
+the Bhattacharyya Coefficients) in table~\ref{Tab_Results-Difference_methods}.
 When two methods perform similarly in their ability to recover the "correct" tree topology (i.e.
@@
 % NC: Split into three sections to answer three questions
+% TG: Ok I removed all the figures + table on top of the section in here and all the text down. We might to have a think of where to anchor the figures later on.
 
@@
     \includegraphics[width=1\textwidth]{Figures/In_main/AllMethods-RF+Tr-colour.pdf}
-\caption{Comparison between the effect of missing data and the tree inference method on topological recovery.
-The amount of missing data for each parameter is represented on the x axis.
-The difference in topologies is represented on the y axis, using standardized Robinson-Foulds distance (upper row) and standardized Triplets distance (lower row).
-Points represent the modal value of each distribution; thick solid and thin dashed lines represents the 50\% and 95\% confidence intervals or the distributions respectively.
-Analyses using the Maximum Likelihood trees are represented in black, the Bayesian consensus %MCC?
-trees in red, the Maximum Likelihood bootstrap trees in green and the Bayesian posterior tree distribution in blue.}
+\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (green) and Bayesian posterior tree distributions (blue).
+The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
+}
 \label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
@@
 % NC: Also should probably have the X axes read : Missing character (Mc, %) and remove % from tick labels
+% TG: Ok, will fix later on since it's all scripted (no illustrator cheats)
 
@@
 %The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (green) and Bayesian posterior tree distributions (blue).
-The percentage of missing data for each parameter (Mc, ML, MF) is shown on the x axis.
-Topological recovery was measured using two different tree comparison metrics: standardized Robinson-Foulds distance (upper row) and standardized Triplets distance (lower row).
-The graph shows the modal value (points), the 50\% and 95\% confidence intervals (thick solid lines) and the distributions (thin dashed lines), of the tree comparison metric for each missing data parameter and tree inference method.
-
-As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the correct topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{C}$ or $M_{F}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (standardized Robinson Foulds or standardized Triplets distance).
-% NC: Results are for YOUR results. compare it to others in the Discussion only
-% NC: I've rephrased below as it was getting confused between what is GOOD and BAD for topological recovery and the difference between the parameters and the methods. I've tried to clarify.
-% NC: Note that you need to be clear what the methods are outperforming...
-However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam}). % And figure 3 too right?
-
-%Increasing the amount of missing data in the living taxa ($M_{L}$) has the worst effect on topological recovery, though only when using the standardized Robinson Foulds distance.
-Using the standardized Robinson Foulds distance, which represents clade conservation between trees, we get the best topological recovery using Bayesian consensus trees.
-However, using standardized Triplets distance, which represents the movement of wildcard taxa, Maximum Likelihood outperforms the other methods but with wider distributional overlap (Fig.~\ref{Fig_Results-permeth_perparam}).
-%NC: What distributions?
-Do you need to say that?
+The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
 
@@
     \includegraphics[width=1\textwidth]{Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
-\caption{Trend of the effect of missing data on topological recovery on the Maximum Likelihood and the Bayesian consensus trees.
-The amount of missing data per parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is represented along the x axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data for each parameter.
-The topology is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
-Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
-The Maximum Likelihood trees are represented in black and the Bayesian consensus trees in grey.}
+\caption{Effect of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
+The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on the x axis.
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
 \label{Fig_Results-global_perparam} %Differences between all the parameters and between two methods (ML vs BC)
 \end{figure}
-
-%When looking at the effect of all parameters combinations on the Bayesian consensus trees, this method outperforms Maximum Likelihood for the Robinson-Foulds distance and plateaus to a minimal modal tree similarity of 0.69 regardless the amount of missing data (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-However, regarding the Triplets distance, Bayesian consensus trees seems to perform poorly but with great confidence intervals overlaps (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-These larger confidence intervals are due to the lower ability of the Triplet method to highlight precise differences in topology \citep{kuhnerpractical2014}.
-Results for the global trend comparison between the uncertainty methods (Bootstrap and Bayesian posterior tree distribution) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-% NC: Move what you need of this to the sections below.
-% NC: Remember you're not testing the RF and TD, you're seeing that the loss of data affects clade conservation and wildcard taxa separately.
-
-\subsection{Effects of the missing data parameters on topological recovery}
 
@@
 %    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
-%\caption{Pairwise Bhattacharyya Coefficients within the Bayesian consensus trees.
-The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data for each parameter.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (1) to red (0).
-A.
-Results for the Normalised Robinson-Foulds distance.
-B.
-Results with the for Normalised Triplets distance.}
+%\caption{Effect of missing data on topological recovery using Bayesian consensus trees.
+The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis.
+Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - red ; high probability of overlap - green).}
 \label{Fig_Results-paircomp_within}
 %\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
-
-%Increasing the amount of missing data in the living taxa ($M_{L}$) has the worst effect on topological recovery, though only when using the standardized Robinson Foulds distance.
-
-%NC: I'd start by describing the pattern with reference to figs 2 and 3 and then the "significant" results
-
-When looking at the pairwise Bhattacharyya Coefficients between all the parameters combinations for Bayesian consensus trees, % NC: Instead just state the result.
-"We found that number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (table X)".
-In the same way as you wouldn't say "Using a linear regression we found that x ~ Y." You'd say "X and Y were significantly correlated (Table1)".
-%NC: Also what IS the result really?
-You find that ML overlaps less with the other parameters so it's significantly different in it's effect?
-And from the graph you can see it's worse?
-What does oerlap MEAN in terms of the answer to the question, i.e.
-how do the parameters affect topological recovery.
-
-the number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}-A).
-However, when looking at the Triplets distance, because of the high overlap in confidence intervals, there is no major region with low probability of overlap.
-One should note however that the bottom line of the matrix where the Bhattacharyya Coefficients are very low is due to comparing the distribution of the pairwise comparison of the "best" tree against the "missing-data" tree with no missing data ($M_{L}$=0\%, $M_{C}$=0\%, $M_{F}$=0\%) , leading to a normalised Triplets score of 1 every time (with no variance, see Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-The pairwise comparisons for the other methods (Maximum Likelihood, Bootstraps and Bayesian posterior trees) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-
-\subsection{Effects of the tree inference method on topological recovery}
-
-%NC: I'd start by describing the pattern with reference to figs 2 and 3 and then the "significant" results
 
@@
     Maximum Likelihood \textit{vs.} & $RF^1$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
-    Bootstraps                      & $Tr^2$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
+    Maximum Likelihood bootstraps   & $Tr^2$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
     Maximum Likelihood \textit{vs.} & $RF^1$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
@@
     Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
-    bootstraps                      & $Tr^2$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
+    Maximum Likelihood bootstraps   & $Tr^2$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
     Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
     Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
-    Boostraps \textit{vs.}          & $RF^1$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
-    Bayesian posterior tree         & $Tr^2$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
+    Bayesian posterior tree \textit{vs.}  & $RF^1$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
+    Maximum Likelihood bootstraps   & $Tr^2$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
    \hline
@@
 \end{table}
+
+%Results plan:
+% A-Change in topology
+%   A.1-Missing data reduce topological recovery regardless the method, etc... -> figure topological recovery ~ parameters(all - no interactions) + methods(all) + metrics(all)
+%   A.2-Within a particular method, missing data reduce topological recovery   -> figure topological recovery ~ parameters(all + interaction) + methods(ML/baycon) + metrics(all)
+% B-"Significance" (i.e. Bhattacharrya Coefficients as a two-sided t-test. Or in other words, does the distributions between two stuff (within/among) differ, regardeless who's better or not:
+%   B.1-Effect of the parameters  -> figure matrix
+%   B.2-Effect if the method      -> table matrix
+
+% A-Change in topology
+% A.1
+As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the correct topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
+% NC: Results are for YOUR results. compare it to others in the Discussion only
+% NC: I've rephrased below as it was getting confused between what is GOOD and BAD for topological recovery and the difference between the parameters and the methods. I've tried to clarify.
+% NC: Note that you need to be clear what the methods are outperforming...
+However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+% And figure 3 too right?
+TG: Yep
+
+% A.2
+The Normalised Robinson-Foulds distance calculate for the trees inferred in Maximum Likelihood framework decreases with the increase of missing data in our three parameters ($M_{L}$, $M_{F}$ or $M_{C}$) with a modal Normalised Robinson-Fould score always below 0.5 when 75\% of the living taxa have no data ($M_{L}$=75\% - Fig.~\ref{Fig_Results-global_perparam}).
+However, in Bayesian inference, the ability of the method (the Bayesian consensus tree) to recover a "correct" topology plateaus at a minimal Normalised Robinson-Fould score of 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the values distribution).
+The results of the topological recovery with the parameters combination in both Normalised Robinson-Foulds and Triplets distances for the Maximum Likelihood bootstrap trees and the Bayesian posterior distribution trees are available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.
+
+% B-"Significance"
+% B.1 - within a method
+%All parameters combinations are different than the "best" tree (always red) in both metrics.
+The lowest probability of overlap (low Bhattacharrya Coefficient values) between pairwise distributions of parameters combinations ($M_{L}$, $M_{F}$ and $M_{C}$ with 0\%, 10\%, 25\%, 50\% or 75\% missing data) is always between the comparison of Normalised distance (Robinson-Foulds or Triplets) of the "missing data" tree with no missing data ($M_{L}$=0\%, $M_{F}$=0\% and $M_{C}$=0\%) and all the other "missing data" trees.
+%The parameters combinations are less dissimilar in Triplets than in RF (less red).
+However, pairwise comparisons are more dissimilar (low Bhattacharrya Coefficient values) in the Normalised Robinson-Fould distance results than in the Normalised Triplets distance results.
+%The ML parameter (0% and 10%) are more different in RF and Triplets (red stripes at the base). The ML parameter (50, 75) are the more similar (green upper part) in RF.
+The Normalised Robinson-Fould distance for the comparisons between the trees with a low amount of missing data for the living taxa ($M_{L}$=0\% and 10\%) shows the most dissimilarities with the other parameter combinations.
+%The combination of the ML parameter (0% and 10%) and the MC parameter (0, 10, 25) are the most different (details on the red stripes).
+
+%Check supplementary for the rest
+The results of the probability of overlap between the pairwise distributions of Normalised Robinson-Foulds and Triplets distances of the Maximum Likelihood trees, the Maximum Likelihood Bootstrap trees and the Bayesian posterior trees distribution is available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.
+
+% B.2 - among methods
+%The Results for the Bayesian consensus tree in RF are the more dissimilar to the results of the Maximum Likelihood bootstrap trees and to the Bayesian posterior distribution trees.
+%The Maximum Likelihood bootstrap trees and the Bayesian posterior distribution trees are the most similar for both metrics.
+
+%Increasing the amount of missing data in the living taxa ($M_{L}$) has the worst effect on topological recovery, though only when using the standardized Robinson Foulds distance.
+Using the standardized Robinson Foulds distance, which represents clade conservation between trees, we get the best topological recovery using Bayesian consensus trees.
+However, using standardized Triplets distance, which represents the movement of wildcard taxa, Maximum Likelihood outperforms the other methods but with wider distributional overlap (Fig.~\ref{Fig_Results-permeth_perparam}).
+%NC: What distributions?
+Do you need to say that?
+
+%When looking at the effect of all parameters combinations on the Bayesian consensus trees, this method outperforms Maximum Likelihood for the Robinson-Foulds distance and plateaus to a minimal modal tree similarity of 0.69 regardless the amount of missing data (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+However, regarding the Triplets distance, Bayesian consensus trees seems to perform poorly but with great confidence intervals overlaps (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+These larger confidence intervals are due to the lower ability of the Triplet method to highlight precise differences in topology \citep{kuhnerpractical2014}.
+Results for the global trend comparison between the uncertainty methods (Bootstrap and Bayesian posterior tree distribution) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+% NC: Move what you need of this to the sections below.
+% NC: Remember you're not testing the RF and TD, you're seeing that the loss of data affects clade conservation and wildcard taxa separately.
+
+%Increasing the amount of missing data in the living taxa ($M_{L}$) has the worst effect on topological recovery, though only when using the standardized Robinson Foulds distance.
+
+%NC: I'd start by describing the pattern with reference to figs 2 and 3 and then the "significant" results
+
+When looking at the pairwise Bhattacharyya Coefficients between all the parameters combinations for Bayesian consensus trees, % NC: Instead just state the result.
+"We found that number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (table X)".
+In the same way as you wouldn't say "Using a linear regression we found that x ~ Y." You'd say "X and Y were significantly correlated (Table1)".
+%NC: Also what IS the result really?
+You find that ML overlaps less with the other parameters so it's significantly different in it's effect?
+And from the graph you can see it's worse?
+What does oerlap MEAN in terms of the answer to the question, i.e.
+how do the parameters affect topological recovery.
+
+the number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}-A).
+However, when looking at the Triplets distance, because of the high overlap in confidence intervals, there is no major region with low probability of overlap.
+One should note however that the bottom line of the matrix where the Bhattacharyya Coefficients are very low is due to comparing the distribution of the pairwise comparison of the "best" tree against the "missing-data" tree with no missing data ($M_{L}$=0\%, $M_{C}$=0\%, $M_{F}$=0\%) , leading to a normalised Triplets score of 1 every time (with no variance, see Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+The pairwise comparisons for the other methods (Maximum Likelihood, Bootstraps and Bayesian posterior trees) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+
+\subsection{Effects of the tree inference method on topological recovery}
+
+%NC: I'd start by describing the pattern with reference to figs 2 and 3 and then the "significant" results
 
@@
 % NC: Make sure you're consistent in how you refer to things ie Maximum Likelihood bootstrap trees, not Bootstraps
+
 %---------------------------------------------

### 2015-01-06 Manuscript/TEM_draft.tex
Edits to discussion

--- 2015-01-06 Manuscript/TEM_draft.tex
@@
 
-% NC: Discussions usually begin with a paragraph that sums up the results.
-Then you can move on to discussing caveats, problems, extensions etc.
-You also need to show how your work fits in with the wider literature.
-So I think you need to re-write this completely with that in mind.
-Do the same as we did for the introduction - write a line for each paragraph.
-This reads like you jumped in without thinking clearly about the structure.
-It probably doesn't need to have all the same subheadings either.
 % TG: Ok, from here, everything is new, I tried way more to write something proper following your "results-caveats-problems-extension" structure and I heavily based the structure of the paragraphs on the longevity paper + others (you might recognise you're style in the first paragraph if I did well).
@@
 %   -Missing data increase leads to decreases in recovering the correct topology 
-%The results of our simulations are in adequacy with recent analysis on the effect of missing data in morphological matrices \citep{pattinsonphylogeny2014,wrightbayesian2014,sansombias2014}. 
-Our results show that our ability to recover the correct topology decreases with increasing missing data, regardless of the missing data parameter ($M_{L}$, $M_{C}$ or $M_{F}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (standardized Robinson Foulds or standardized Triplets distance).
+Our results show that the ability to recover the correct topology in a Total Evidence framework decreases with increasing missing data, regardless of how data were removed or the method of tree inference used.
+% NC: Here I've removed the tree metrics as these are just different ways of measuring topology.
 %Results 1
 %   -However, parameters have different effect
-However, within a tree inference method and a metric, the parameters used to remove data have different effects on the ability to recover the correct topology.
+However, these factors affected topological recovery in different ways and to different extents. 
 %Results 2
 %   -Additionally, methods and metrics are unequal
-Additionally, the effect of missing data on topological recovery varies with different methods to infer topology and different metrics to compare topologies.
+% Additionally, the effect of missing data on topological recovery varies with different methods to infer topology and different metrics to compare topologies.
+% NC: This is implicit in the line above
 %Wrap up results
 %   -Thus, Baycon is better with loads of living 
-Thus, we found that using the topologies of Bayesian consensus trees performed consistently better than other methods for recovering the correct topology (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}) and that both the amount of available characters ($M_{C}$) and the amount of data for the living taxa ($M_{L}$) are crucial missing data parameters (Fig.~\ref{Fig_Results-paircomp_within}).
+% NC: Make sure it's parameters then tree building methods.
+Increasing the number of characters in the matrix without data ($M_{C}$) and the amount of missing morphological data in the matrix for living taxa ($M_{L}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
+Additionally, using Bayesian consensus trees recovered the correct topology more consistently than using Maximum Likelihood trees or Bayesian posterior distributions of trees (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
+% NC: Probably needs another summing up line here. Or maybe not
 
-% $§1: What we found >...
-% ^§2: >... links with previous studies
+%NC: Make sure the order of the parameters is always the same L,F, C or whatever! THROUGHOUT the paper. Also correct RF and T to standardised OR normalised RF and T.
+
+% NC: Might want to add a couple of subheadings to match those in the results.
 
@@
 %   -Missing data per se is not a problem, problem is in the amount of overlap
-As proposed in previous studies, the amount of missing data is not a problem \textit{per se} as long as enough data overlaps in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
-%In fact, as shown previously, the amount of overlap in the matrix is more crucial than the proportion of missing data \citep[e.g.][]{rouresite-specific2011}.
+As proposed in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough data overlaps in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
+% NC: What do you mean by data overlap here?
+That species have overlapping characters?
 %Proposed hypothesis for result 1
 %   -One can increase overlap by adding living taxa
-To improve the ability to recover the "correct" topology in a Total Evidence framework, one can efficiently increase the amount of overlapping data by reducing the amount of missing data in each of our three selected parameters.
-In our simulations, the overlap in data is greatly reduced when different parameters are combined with a high amount of missing data (e.g.
-$M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-In this scenario the overall number of missing characters ($M_{C}$) affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
-However it is important to note that in the pairwise Bhattacharyya Coefficient analysis, the area with the lowest probability of distributions overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
+To improve the ability to recover the "correct" topology in a Total Evidence framework, one can efficiently increase the amount of overlapping data 
+% NC: Again what do you mean by overlapping data? And isn't this obvious???
+by reducing the missing data in each of our three missing data parameters ($M_{L}$, $M_{C}$ and $M_{F}$).
+In our simulations, the overlap in data is greatly reduced when all three missing data parameters have a lot of missing data (e.g.
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+In this scenario, the overall number of missing characters ($M_{C}$) affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
+% NC: What's the relevance of that sentence?
+However, the pairwise Bhattacharyya Coefficients show that the area with the lowest probability of distributional overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
+%NC: I don't understand what you mean here
 
@@
 %§3
-%Discussing results 1
-This variable effect of the missing data parameters on the ability to recover the "correct" topology (Fig.~\ref{Fig_Results-paircomp_within}) can be explained by the way our parameters are illustrating the sources of missing data in empirical matrices: (1) non coded-living  taxa ($M_{L}$), (2) the quality of the fossil record ($M_{F}$) and (3) the amount of coded characters ($M_{C}$ - see \hyperref[Removing_data]{Methods}).
-We have suggestions for practical ways to reduce this missing data problem in empirical matrices:
-%   -Which parameters does what?
-\begin{enumerate}
-\item{$M_{L}$ (non coded living taxa):}
-The vast majority of known living species have specimens preserved in natural history museums.
-This is a significant resource and more research effort is needed to code morphological characters in these specimens.
-\item{$M_{F}$ (quality of the fossil record):}
-The amount of missing data in the fossil record solely depends on the preservation quality of the fossils and can not be actively improved.
-However, gaps in the matrix can be filled with the discoveries exceptional preserved fossil \citep[e.g.][]{nithe2013}.
-Fortunately, this parameter is the most difficult to improve but it also has the least influence on whether the simulations recover the "correct" topology (Fig.~\ref{Fig_Results-paircomp_within}).
-\item{$M_{C}$ (amount of coded characters):}
-The amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
-Improvement in both decreasing the time taken to build the matrix and increasing the number of characters can be done through vast collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}.
-\end{enumerate}
-Fixing these three major issues can lead to a sensible increase in the ability to recover the "correct" topology.
+%Discussing results 1 % NC: This can join to the end of the previous paragraph
+This variable effect of our missing data parameters on the ability to recover the "correct" topology (Fig.~\ref{Fig_Results-paircomp_within}) can be explained by % the way our parameters are illustrating the sources of missing data in empirical matrices
+% NC: No.
+This shows how our parameters relate to the real world and give real world solutions.
+Probably first want to explain any difference in effects agnostically - i.e.
+just as parameters with no biological context.
+So maybe: ..."can be explained by the different way data is removed using each of our three missing data parameters..." THEN move on to this:
+
+Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
+Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "correct" topology.
+We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections.
+Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries.
+However, gaps in the matrix can be filled with future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
+Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "correct" topology (Fig.~\ref{Fig_Results-paircomp_within}).
+Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
+One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}, so that no one individual collects all the data.
 
@@
 %   -Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
-Nonetheless, the variation in the ability to recover the "correct" topology does not only depend on the missing data in morphological part of the matrix but also heavily on the method used to infer the topology (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-Previous studies have shown some superiority on methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014} but see \citealt{spencerefficacy2013}).
-However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} between Maximum Likelihood and Bayesian inference in a Total Evidence framework.
+Variation in our ability to recover the "correct" topology also depended heavily on the method used to infer the tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+Previous studies have shown some superiority of methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}).
+However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
 %TG: I guess it's a good selling point: we ARE the first ones to do a sensitivity analysis in a TEM framework (as far as I know).
@@
 %   -Why is Baycon better than the others?
-When using the clade conservative Robinson-Foulds distance (see \citealt{kuhnerpractical2014} and \hyperref[metrics_discussion]{discussion below}), the Bayesian consensus trees greatly outperforms the Maximum Likelihood trees in the ability to recover the correct tree topology (Fig.~\ref{Fig_Results-global_perparam}).
-In fact, regardless of the amount of missing data, this method always displays the highest normalised Robinson-Foulds score (always $>$ 0.69 - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-Also, when looking at the pairwise Bhattacharyya Coefficients within each method, the Bayesian consensus tree one display the highest probability of overlap between the results in any case (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+When using the standardised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "correct" topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}).
+In fact, regardless of the amount of missing data, Bayesian consensus trees always displays the highest normalised/standardised Robinson-Foulds score (always $>$ 0.69; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 %   -Why is Baycon so different than Baytre? Since Baycon is calculated from Baytre.
-When comparing this to the probability of overlap between the methods, it is counter intuitive to note that the Bayesian consensus tree shows the highest dissimilarity with the Bayesian posterior trees from which the Bayesian consensus tree is build (Fig.~\ref{Fig_Results-paircomp_within}-A).
-Note that it is also highly dissimilar to the Bootstraps.
-However that can be linked to the fact that both Bootstraps and Bayesian posterior trees are measured as 125*1000 random pairwise comparisons instead of just 125 tree comparisons for the Bayesian consensus trees.
-The fact that the 1000 pairwise trees to be compared for each of the 125 comparisons are chosen randomly, it can add noise to the results for these two methods (see \hyperref[tree_comparisons]{methods above}).
+We saw the poorest topological recovery using Maximum Likelihood bootstrap trees and Bayesian posterior distribution of trees (% NC: FIGURE???
+).
+This reflects the fact it is difficult to compare two distributions of trees, and each comparison between a set of missing data parameters and "best" tree in these cases involved 1000 random pairwise comparisons rather than just one.
+This will obviously add noise to the results for these two methods.
 Also, in the case of the lower performances of the Bayesian consensus trees when using the Triplets distance, one can note the high overlap of the confidence intervals (Fig.
@@
 ~\ref{Fig_Results-paircomp_within}-B).
+% NC: Don't get the relevance of the last sentence/why it fits here.
+Can you expand?
+% NC: Probably want some  statement here about bayesian being better than ML, and if you want to fix a single topology in TE we recommend a bayesian consensus.
+If you want to account for tree uncertainty we recommend a bayesian posterior distribution.
+Our results dont suggest this is bad, but more suggest that we can't compare trees very well in this case.
 
@@
 %   -RF is different and better than triplets.
+
+% NC: I think this can be condensed and slotted into the correct places above.
+So where you speak about the missing data parameters, mention differences in how they influence Rf and Triplets and why that might be.
+Then when you mention methods do the same.
+
 \label{metrics_discussion}
@@
 %   -Simulating morpho characters
-The variations in our results could also be explained by another methodological caveat: the performance of our simulation protocol.
-In order to reduce the computational time of our analysis ($~$150 CPU years), we had to run our sensitivity analysis on modest size matrices (1000 molecular data and 100 morphological data).
-This matrix size is two orders of magnitude lower than some matrices used in published analysis \citep[e.g.][]{springermacroevolutionary2012,nithe2013} but still ranges in the sizes of smaller matrices used in other ones \citep[e.g.][]{sallam2011craniodental,kellymolecular2014}.
+The variations in our results could also be explained by another methodological caveat: 
+% NC: What variations do you mean? Only an issue if it might bias results in one direction or another
+the performance of our simulation protocol.
+In order to reduce the computational time of our analysis ($~$150 CPU years), we ran our sensitivity analysis on modestly sized matrices of 1000 molecular characters and 100 morphological characters.
+These matrices are two orders of magnitude smaller than some matrices used in published analyses \citep[e.g.][]{springermacroevolutionary2012,nithe2013} but still ranges in the sizes of smaller matrices used in other ones \citep[e.g.][]{sallam2011craniodental,kellymolecular2014}.
 %TG: note about NOT citing O'Leary here.
@@
 Also, simulating morphological characters is more complex than molecular ones.
-The underlying pattern of their evolution are often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001,wrightbayesian2014}.
+The underlying pattern of their evolution is often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001,wrightbayesian2014}.
 Morphological characters studies involve many potential statistical pitfalls (e.g.

### 2015-01-06 Manuscript/TEM_draft.tex
Edits to discussion

--- 2015-01-06 Manuscript/TEM_draft.tex
@@
 %Further directions - is that necessary though?
+% NC: Cut for now, might want to put these ideas in the appropriate places above instead.
 %   -Using real data
-One way to address this issue would be to run a similar analysis using real data.
+%One way to address this issue 
+% NC: Which issue???
+%would be to run a similar analysis using real data.
 The amount of studies using Total Evidence data type matrices is steadily increasing \citep[e.g.][]{ronquista2012,slaterphylogenetic2013,beckancient2014} and one could analysis these total evidence matrices in the same framework as this study.
@@
 %   -Data is not a major problem
-Despite the various caveats discussed above, we answered our main question on how does Total Evidence matrices perform in recovering topology with different methods and different amounts of missing data.
+%Despite the various caveats discussed above, we answered our main question on how does Total Evidence matrices perform in recovering topology with different methods and different amounts of missing data.
+% NC: For a concluding paragraph, don't start by reminding people of the caveats! End strongly and confidently
+% Also no need to restate the question - instead give the answer.
 %   -Good and promising method.
-Our results shows that this method performance decreases mainly with a reduction in the overlap of morphological data.
-In order to address this eventual problem, we propose to increase the data overlap by coding living species and to have a large morphological matrix.
-Also, our results shows disparity among methods and a clear superiority of the Bayesian consensus tree topology over the Maximum Likelihood one.
+%Our results shows that this method performance decreases mainly with a reduction in the overlap of morphological data.
+To address this problem, we propose to increase the data overlap by coding living species and to have a large morphological matrix.
+Also, our results shows disparity among methods and a clear superiority %NC: Not really
+%of the Bayesian consensus tree topology over the Maximum Likelihood one.
 
@@
 Using Total Evidence matrices with missing data is not a problem for recovering the correct topology as long as enough morphological data is overlapping between the fossil and living taxa.
-In order to optimize good topological recovery, we propose to use the Bayesian consensus tree over the Maximum Likelihood tree or over a random subset of the Bayesian posterior trees distribution.
-Also, in order to improve the overlap in data in the morphological part of the matrix, we advice to use all the morphological characters available and to code them for a maximum of living taxa present in the matrix.
+%NC: Explain more clearly what you mean by overlap - of characters?
+Topological recovery is still best when there is less missing data, therefore we advise filling as many gaps in Total Evidence matrices as possible.
+Because this is difficult, if not impossible, for fossil data, we recommend coding as many
+
+using as many morphological characters as possible and to code them for a maximum of living taxa present in the matrix.
+
+we propose to use the Bayesian consensus tree over the Maximum Likelihood tree or over a random subset of the Bayesian posterior trees distribution.
+Also, in order to improve the overlap in data in the morphological part of the matrix,
+
+% NC: Never write "in order to". Just "to" will do
 
@@
 \section{Supplementary material}
-Supplementary material (code, analysis and full results) can be found in the Dryad data repository at http://dx.doi.org/10.5061/dryad.XXXX. %section and sentence format asked by Syst. Biol.
+Supplementary material (R code, analyses, supplementary methods and full results) can be found in the Dryad data repository at http://dx.doi.org/10.5061/dryad.XXXX.
+%section and sentence format asked by Syst.
+Biol.
 
@@
 %TG : Not sure how many people we have to thank but it could also include Graham Slater, Nick Matzke and Trevor Hodkinson.
+ % NC: Might be nice to thank Trevor, and it won't add too many lines to thank them all. You can also just calle people A.Wright if you want to reduce word count.
 for useful comments on our simulation protocol.
-%Thanks to Sive Finlay, Kevin Healy and Adam Kane for comments on the manuscript. %TG: or helping writing sentences does not really count.
+%Thanks to Sive Finlay, Kevin Healy and Adam Kane for comments on the manuscript.
+%TG: or helping writing sentences does not really count.
+% NC: Depends how much they helped.
+I just used to thank the lab for that stuff
 Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
@@
 %Needs to be cleaned a bit
-
+% NC: Yeah. Leave it til we send it to Gavin but it should be a priority. Almost all of them are wrong! 
 %---------------------------------------------

### 2015-01-06 Manuscript/TEM_draft.tex
Rewrote + cleaned results
-5 parts plan with 3 subparts per part

@@
 
+%Questionish §
+We use the Bhattacharyya Coefficient in a similar way as a two-sided t-test to detect significant differences between the distributions of the normalised distance between the "best" tree and the "missing data" trees (for both metrics).
+%We also take into account the effect of the methods used. REPHRASE
+This way we could investigate the effect of each our three missing data parameters and their interaction on the ability to recover the "correct" topology.
+
 %---------------------------------------------
@@
 
-%Results plan:
-% A-Change in topology
-%   A.1-Missing data reduce topological recovery regardless the method, etc... -> figure topological recovery ~ parameters(all - no interactions) + methods(all) + metrics(all)
-%   A.2-Within a particular method, missing data reduce topological recovery   -> figure topological recovery ~ parameters(all + interaction) + methods(ML/baycon) + metrics(all)
-% B-"Significance" (i.e. Bhattacharrya Coefficients as a two-sided t-test. Or in other words, does the distributions between two stuff (within/among) differ, regardeless who's better or not:
-%   B.1-Effect of the parameters  -> figure matrix
-%   B.2-Effect if the method      -> table matrix
+%Bleuah! Changing everything to stick to 5 parts as
+%Part 1: missing data decrease topology
+%Part 2: effect of ML
+%   -As missing data increase, topology decreases
+%   -This is different for the metrics
+%   -This is worst with that method
+%       (e.g. Distribution of RF values for the ML difference % missing data a substantial overlap with bla (BC=88, Fig bla))
+%Part 3: effect of MF 
+%Part 4: effect of MC
+%Part 5: combined effect
 
-% A-Change in topology
-% A.1
+%Part 1
 As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the correct topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
-% NC: Results are for YOUR results. compare it to others in the Discussion only
-% NC: I've rephrased below as it was getting confused between what is GOOD and BAD for topological recovery and the difference between the parameters and the methods. I've tried to clarify.
-% NC: Note that you need to be clear what the methods are outperforming...
 However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+In any combination of parameters or metrics, the Maximum Likelihood bootstraps trees and the Bayesian posterior distribution trees performs in a substantially similarly (Table ~\ref{Tab_Results-Difference_methods}) and worth than the Maximum Likelihood and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} - see discussion).
+%add discussion link
+
+%Part 2
+\subsection{Effect of missing living taxa ($M_{L}$)} %Or remove? Also below then
+As the amount of missing living taxa increases, the ability to recover the correct topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
+However, the decrease of the Normalised Robinson-Fould distance plateaus at minimum of 0.69 for Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the exact values) and there is no substantial difference in topological recovery for 50\% or 75\% of missing living taxa (upper triangle in Fig.~\ref{Fig_Results-paircomp_within}-A).
+Therefore, clades seems to be more conserved in the Bayesian consensus tree more than any other method (see discussion).
+%add discussion link
+On the other hand, the decrease in Normalised Triplets distance is less contrasted with a greater overlap of Normalised Triplets distance with the Maximum Likelihood tree (Fig.~\ref{Fig_Results-permeth_perparam}) - mean Bhattacharrya Coefficient=0.10 for the Normalised Robinson-Foulds distance and 0.61 for the Normalised Triplets distance - see table ~\ref{Tab_Results-Difference_methods}).
+
+%Part 3
+\subsection{Effect of missing data for fossil taxa ($M_{F}$)}
+As the amount of missing data in the fossil part of the matrix increases, the ability to recover the correct topology decreases as well (Fig.~\ref{Fig_Results-permeth_perparam}).
+For the Robinson-Fould distance, we observe the same pattern than for $M_{L}$ with a better topological recovery of the Bayesian consensus trees on any other methods (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+However, the Normalised Triplets distance drops drastically when going from 25\% of missing data to 50\% (mode) %get the modal values here
+making this parameter more sensible to wildcard taxa in the Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} - see discussion). %add discussion link
+
+%Part 4
+\subsection{Effect of missing character ($M_{C}$)}
+As described for the two previous parameters $M_{L}$ and $M_{F}$, an increase in the amount of missing characters decreases the ability to recover the correct topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
+Similarly again, the Bayesian consensus trees shows higher Normalised Robinson-Fould distance than the Maximum Likelihood tree (Fig.~\ref{Fig_Results-permeth_perparam}).
+And in a similar way as for the $M_{F}$ parameter, the Normalised Triplet distance drops drastically when going from 25\% of missing data to 50\% (mode).
+%get the modal values here
+
+%Part 5
+\subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
+The combine of the missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) reduces the ability to recover the correct topology when each parameters contains a maximum of missing data (i.e.
+$M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-global_perparam} and see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the combined effects of the parameters in both Normalised Robinson-Foulds and Normalised Triplets distance for the Maximum Likelihood Bootstraps trees and the Bayesian posterior trees distribution).
+Regarding the clade conservation (i.e.
+the Normalised Robinson-Foulds distances), the Bayesian consensus tree greatly outperforms the Maximum likelihood trees with Normalised Robinson-Foulds scores always superior than 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the exact values) regardless the combination of missing data parameters (Fig.~\ref{Fig_Results-paircomp_within}-A).
+However, for the Normalised Triplets distance, the Bayesian consensus tree often displays negative values (i.e.
+differences greater than expected by chance) but with few substantial differences (Fig.~\ref{Fig_Results-paircomp_within}-B).
+Also the differences between the Normalised scores of the Maximum Likelihood trees and Bayesian consensus trees are substantially difference for the Normalised Robinson-Foulds distance (mean Bhattacharrya Coefficient=0.10) and for the Normalised Triplets distance (mean Bhattacharrya Coefficient=0.61 - see table ~\ref{Tab_Results-Difference_methods}).
+Thus the sensitivity of the Bayesian consensus trees to wild card taxa is not highly different than the one for the Maximum likelihood trees (i.e.
+the distribution of Normalised Triplets distance have a high probability of overlap).
+The results of the probability of overlap between the pairwise distributions of Normalised Robinson-Foulds and Triplets distances of the Maximum Likelihood trees, the Maximum Likelihood Bootstrap trees and the Bayesian posterior trees distribution is available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.
+
+%Binned removed bits from former results.
+
+%As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the correct topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
+%However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 % And figure 3 too right?
 TG: Yep
-
-% A.2
-The Normalised Robinson-Foulds distance calculate for the trees inferred in Maximum Likelihood framework decreases with the increase of missing data in our three parameters ($M_{L}$, $M_{F}$ or $M_{C}$) with a modal Normalised Robinson-Fould score always below 0.5 when 75\% of the living taxa have no data ($M_{L}$=75\% - Fig.~\ref{Fig_Results-global_perparam}).
+%The Normalised Robinson-Foulds distance calculate for the trees inferred in Maximum Likelihood framework decreases with the increase of missing data in our three parameters ($M_{L}$, $M_{F}$ or $M_{C}$) with a modal Normalised Robinson-Fould score always below 0.5 when 75\% of the living taxa have no data ($M_{L}$=75\% - Fig.~\ref{Fig_Results-global_perparam}).
 However, in Bayesian inference, the ability of the method (the Bayesian consensus tree) to recover a "correct" topology plateaus at a minimal Normalised Robinson-Fould score of 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the values distribution).
 The results of the topological recovery with the parameters combination in both Normalised Robinson-Foulds and Triplets distances for the Maximum Likelihood bootstrap trees and the Bayesian posterior distribution trees are available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.
-
-% B-"Significance"
-% B.1 - within a method
 %All parameters combinations are different than the "best" tree (always red) in both metrics.
-The lowest probability of overlap (low Bhattacharrya Coefficient values) between pairwise distributions of parameters combinations ($M_{L}$, $M_{F}$ and $M_{C}$ with 0\%, 10\%, 25\%, 50\% or 75\% missing data) is always between the comparison of Normalised distance (Robinson-Foulds or Triplets) of the "missing data" tree with no missing data ($M_{L}$=0\%, $M_{F}$=0\% and $M_{C}$=0\%) and all the other "missing data" trees.
+%The lowest probability of overlap (low Bhattacharrya Coefficient values) between pairwise distributions of parameters combinations ($M_{L}$, $M_{F}$ and $M_{C}$ with 0\%, 10\%, 25\%, 50\% or 75\% missing data) is always between the comparison of Normalised distance (Robinson-Foulds or Triplets) of the "missing data" tree with no missing data ($M_{L}$=0\%, $M_{F}$=0\% and $M_{C}$=0\%) and all the other "missing data" trees.
 %The parameters combinations are less dissimilar in Triplets than in RF (less red).
-However, pairwise comparisons are more dissimilar (low Bhattacharrya Coefficient values) in the Normalised Robinson-Fould distance results than in the Normalised Triplets distance results.
+%However, pairwise comparisons are more dissimilar (low Bhattacharrya Coefficient values) in the Normalised Robinson-Fould distance results than in the Normalised Triplets distance results.
 %The ML parameter (0% and 10%) are more different in RF and Triplets (red stripes at the base). The ML parameter (50, 75) are the more similar (green upper part) in RF.
-The Normalised Robinson-Fould distance for the comparisons between the trees with a low amount of missing data for the living taxa ($M_{L}$=0\% and 10\%) shows the most dissimilarities with the other parameter combinations.
+%The Normalised Robinson-Fould distance for the comparisons between the trees with a low amount of missing data for the living taxa ($M_{L}$=0\% and 10\%) shows the most dissimilarities with the other parameter combinations.
 %The combination of the ML parameter (0% and 10%) and the MC parameter (0, 10, 25) are the most different (details on the red stripes).
-
 %Check supplementary for the rest
-The results of the probability of overlap between the pairwise distributions of Normalised Robinson-Foulds and Triplets distances of the Maximum Likelihood trees, the Maximum Likelihood Bootstrap trees and the Bayesian posterior trees distribution is available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.
-
-% B.2 - among methods
-%The Results for the Bayesian consensus tree in RF are the more dissimilar to the results of the Maximum Likelihood bootstrap trees and to the Bayesian posterior distribution trees.
-%The Maximum Likelihood bootstrap trees and the Bayesian posterior distribution trees are the most similar for both metrics.
-
-%Increasing the amount of missing data in the living taxa ($M_{L}$) has the worst effect on topological recovery, though only when using the standardized Robinson Foulds distance.
-Using the standardized Robinson Foulds distance, which represents clade conservation between trees, we get the best topological recovery using Bayesian consensus trees.
-However, using standardized Triplets distance, which represents the movement of wildcard taxa, Maximum Likelihood outperforms the other methods but with wider distributional overlap (Fig.~\ref{Fig_Results-permeth_perparam}).
-%NC: What distributions?
-Do you need to say that?
-
-%When looking at the effect of all parameters combinations on the Bayesian consensus trees, this method outperforms Maximum Likelihood for the Robinson-Foulds distance and plateaus to a minimal modal tree similarity of 0.69 regardless the amount of missing data (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-However, regarding the Triplets distance, Bayesian consensus trees seems to perform poorly but with great confidence intervals overlaps (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-These larger confidence intervals are due to the lower ability of the Triplet method to highlight precise differences in topology \citep{kuhnerpractical2014}.
-Results for the global trend comparison between the uncertainty methods (Bootstrap and Bayesian posterior tree distribution) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-% NC: Move what you need of this to the sections below.
-% NC: Remember you're not testing the RF and TD, you're seeing that the loss of data affects clade conservation and wildcard taxa separately.
-
-%Increasing the amount of missing data in the living taxa ($M_{L}$) has the worst effect on topological recovery, though only when using the standardized Robinson Foulds distance.
-
-%NC: I'd start by describing the pattern with reference to figs 2 and 3 and then the "significant" results
-
-When looking at the pairwise Bhattacharyya Coefficients between all the parameters combinations for Bayesian consensus trees, % NC: Instead just state the result.
+%The results of the probability of overlap between the pairwise distributions of Normalised Robinson-Foulds and Triplets distances of the Maximum Likelihood trees, the Maximum Likelihood Bootstrap trees and the Bayesian posterior trees distribution is available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.
+%When looking at the pairwise Bhattacharyya Coefficients between all the parameters combinations for Bayesian consensus trees, % NC: Instead just state the result.
 "We found that number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (table X)".
@@
 how do the parameters affect topological recovery.
-
-the number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}-A).
+%the number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}-A).
 However, when looking at the Triplets distance, because of the high overlap in confidence intervals, there is no major region with low probability of overlap.
@@
 The pairwise comparisons for the other methods (Maximum Likelihood, Bootstraps and Bayesian posterior trees) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-
-\subsection{Effects of the tree inference method on topological recovery}
-
 %NC: I'd start by describing the pattern with reference to figs 2 and 3 and then the "significant" results
-
 %When looking at the comparisons between the methods 
-using the Robinson-Foulds distance, the Bayesian consensus trees have the least overlap with respectively the Bayesian posterior trees and the Bootstrap (Fig.~\ref{Fig_Results-paircomp_without}-A).
+%using the Robinson-Foulds distance, the Bayesian consensus trees have the least overlap with respectively the Bayesian posterior trees and the Bootstrap (Fig.~\ref{Fig_Results-paircomp_without}-A).
 Using the Triplets distance, the two methods that have the most distribution %NC: ???
-are the Bootstraps and the Bayesian posterior trees (Fig.~\ref{Fig_Results-paircomp_without}-B).
+%are the Bootstraps and the Bayesian posterior trees (Fig.~\ref{Fig_Results-paircomp_without}-B).
 These results are due to the difference in comparing topologies between Maximum Likelihood and Bayesian consensus trees and Bootstraps and Bayesian posterior trees: the latter two are based on a random pairwise comparison of a sub-sample of 1000 trees.
 This is likely to add noise to the data.
-
-\subsection{Interactions among missing data parameters and the tree inference method on topological recovery}
-
-%NC: Maybe don't need this section. See how it goes.
-
-% NC: Make sure you're consistent in how you refer to things ie Maximum Likelihood bootstrap trees, not Bootstraps
 

### 2015-01-07 Manuscript/TEM_draft.tex
Discussion edits (draft)

@@
 Probably need to think some more about this.
+% TG: a note about the "correct" topology regarding your comment just after the discussion header: should I replace "correct" by "best" since that's how we're calling the "best" tree anyway (all in quotes)
 
@@
 
-% TG: Ok, from here, everything is new, I tried way more to write something proper following your "results-caveats-problems-extension" structure and I heavily based the structure of the paragraphs on the longevity paper + others (you might recognise you're style in the first paragraph if I did well).
-I left in comments the skeleton of the structure that I used to help me writing it.
-% TG: Also here's a note on my syntax notation for linking the paragraphs: "$"=end OR "^"=start of paragraph number 1.
-"a >..."=which finsih with the statement OR ">...
-b"=which follows with the statement b
-% TG: Here's the tree-style "plan" I was using (linked with the intro):
-% INTRO:
-% §1 We need fossils
-%   |
-%   \-> §2 How can we use them?
-%       |
-%       \-> §3 Total Evidence method of course!
-%           |
-%           \-> §4 But this method has a lot of missing data?
-%               |
-%               \-> §5 So what do we know about missing data from literature?
-%                   |
-%                   \-> §6 How will that effect TEM?
-%                      |
-%                      \-> §7 We use simulations to test!
-%                       
-% METHODS
-% RESULTS
-% DISCUSSION
-%       
-% §1 What do we found (Intro-§6)
-%   |
-%   \-> §2 How does that link to what we know (Intro-§5)
-%       |
-%       \-> §3 The effect of the parameters
-%       |
-%       \-> §4 The effect of the methods
-%       |
-%       \-> Caveats
-%           |
-%           \-> §5 The metrics
-%           |
-%           \-> §6 The simulations
-%           |
-%           \-> §7 How to fix them
-%           |
-%           \-> §8 How we can use fossils (Intro-§2)
-
-% CONCLUSION.
-% part 1- Fossils are important and can be used combined with living taxa (Intro-§1 & §2)
-% part 2- This is how we can improve the data to use them in the best way possible (Conclusion-§1 & §8)
-
-%§1
-%Results: global
-%   -Missing data increase leads to decreases in recovering the correct topology 
 Our results show that the ability to recover the correct topology in a Total Evidence framework decreases with increasing missing data, regardless of how data were removed or the method of tree inference used.
-% NC: Here I've removed the tree metrics as these are just different ways of measuring topology.
-%Results 1
-%   -However, parameters have different effect
 However, these factors affected topological recovery in different ways and to different extents.
-%Results 2
-%   -Additionally, methods and metrics are unequal
-% Additionally, the effect of missing data on topological recovery varies with different methods to infer topology and different metrics to compare topologies.
-% NC: This is implicit in the line above
-%Wrap up results
-%   -Thus, Baycon is better with loads of living 
+Decreasing the overall number of morphological characters in the matrix ($M_{C}$) and the amount of living taxa with morphological data ($M_{L}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
+Additionally, using Bayesian consensus trees recovered the correct topology more consistently than using Maximum Likelihood trees or Bayesian posterior trees distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
+% ----->
+% NC: Probably needs another summing up line here. Or maybe not
+% <-----
 % NC: Make sure it's parameters then tree building methods.
-Increasing the number of characters in the matrix without data ($M_{C}$) and the amount of missing morphological data in the matrix for living taxa ($M_{L}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
-Additionally, using Bayesian consensus trees recovered the correct topology more consistently than using Maximum Likelihood trees or Bayesian posterior distributions of trees (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
-% NC: Probably needs another summing up line here. Or maybe not
-
-%NC: Make sure the order of the parameters is always the same L,F, C or whatever! THROUGHOUT the paper. Also correct RF and T to standardised OR normalised RF and T.
-
+NC: Make sure the order of the parameters is always the same L,F, C or whatever!
+THROUGHOUT the paper.
+Also correct RF and T to standardised OR normalised RF and T.
 % NC: Might want to add a couple of subheadings to match those in the results.
-
-%§2
-%Results 1 from litterature
-%   -Missing data per se is not a problem, problem is in the amount of overlap
-As proposed in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough data overlaps in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
+As proposed in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough morphological characters overlaps for most taxa in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
 % NC: What do you mean by data overlap here?
 That species have overlapping characters?
-%Proposed hypothesis for result 1
-%   -One can increase overlap by adding living taxa
-To improve the ability to recover the "correct" topology in a Total Evidence framework, one can efficiently increase the amount of overlapping data 
-% NC: Again what do you mean by overlapping data? And isn't this obvious???
-by reducing the missing data in each of our three missing data parameters ($M_{L}$, $M_{C}$ and $M_{F}$).
-In our simulations, the overlap in data is greatly reduced when all three missing data parameters have a lot of missing data (e.g.
-$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-In this scenario, the overall number of missing characters ($M_{C}$) affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
+
+\subsection{Effect of missing living taxa ($M_{L}$)} %Or remove? Also below then
+%What is the parameter variation doing in our simulation
+When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
+Because living taxa still have 10 times more molecular characters available, even if they have no morphological data, the relation between them will always be fairly well resolved (depending on the phylogenetic signal from the molecular part of the matrix) and thus, ignoring the fossils, the living taxa clades will be always well conserved.
+However, this missing data parameter influences a lot the placement of fossil taxa because a decrease of the $M_{L}$ parameter will reduce the overlapping data between the living and the fossil taxa.
+
+%Methods
+%Part saying the difference between ML and Baycon
+When using the Normalised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "correct" topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}).
+In fact, regardless of the amount of missing data, Bayesian consensus trees always displays the highest Normalised Robinson-Foulds score (always $>$ 0.69; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for details).
+However, regarding the Normalised Triplets distance, the topological similarity with the "correct" topology decreases in the same way for both the Maximum Likelihood trees and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+%Part saying the difference between BS and baytrees
+Regardless of the metric, the Maximum Likelihood boostrap trees and the Bayesian posterior trees distribution always performed poorlier in their ability to recover the "correct" topology, even when no morphological data was missing ($M_{L}$=0\% - Fig.~\ref{Fig_Results-permeth_perparam}).
+This reflects the fact it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" tree and "best" tree in these cases involved 1000 random pairwise comparisons rather than just one.
+This will obviously add noise to the results for these two methods.
+
+%Metrics
+%Part explaining the differences between the metrics
+This difference between the metrics is linked to what they measure and how they perform at doing it (see method %add link
+, \hyperref[SupplementaryMaterial]{Supplementary Material S2} and \citealt{kuhnerpractical2014} for details on the nature of both metrics).
+In the Bayesian consensus tree topology, the clade conservation (i.e.
+Noramlised Robinson-Foulds distance) stays always high but the placement of wildcard taxa (i.e.
+the Normalised Triplets distance) decreases way faster (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+This is due to the fact that both metrics are illustrating two different aspects of tree topology (i.e.
+clade conservation for the Normalised Robinson-Foulds distance and wildcard taxa position for the Normalised Triplets distance - see method %add link!
+and \hyperref[SupplementaryMaterial]{Supplementary Material S2} for details on both metrics).
+However, it is important to note that The Normalised Robinson-Foulds distance is more sensitive than the Normalised Triplets distance: when getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
+Also, \citet{kuhnerpractical2014} have recently demonstrated that the Robinson-Foulds distance outperforms the Triplets distance for measuring topological differences between trees.
+Thus, the Bayesian consensus tree's topology seems to be the less affected by removing morphological data for living taxa (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+
+\subsection{Effect of missing data for fossil taxa ($M_{F}$)}
+%When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this reduces the probability of morphological characters for the fossil taxa to overlap with the ones for the living taxa and leads to displacement of certain taxa.
+However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
+decrease of Normalised Triplets distance), the conservation of the clades (i.e.
+Normalised Robinson-Foulds distance) is still relatively good (mode=% value
+%) when the proportion of missing data is high ($M_{F}$=75\%).
+
+%Methods
+%Decrease in clade conservation is smaller than for param ML in the Bayesian consensus tree for the clade conservation but higher than param ML for the displacement of wildcard taxa.
+This is due to the fact that the Bayesian consensus tree is a majority consensus rule and that when the fossils have less data, they kind of randomly branch all over the shop and therefore no majority consensus position (>50\% of the time) exists puting the fossil on a politomie at the base of a clade (clade conservation stays high but displacement of taxa increases).
+On the opposite in MaxLik, the topology is always dichotomous but the fossil can be missplaced in the wrong clade (with a low BS value) but is moving less.
+In other words, Bayesian consensus tree places the fossil with a higher confidence at a low taxonomic level (politomy) and the MaxLik places the fossil with a low confidence at a high taxomomic level (baldy supported dichotomies).
+We can argue that the first option (higher confidence at a low taxonomic level) is preferable (example the Darwinius debate).
+
+\subsection{Effect of missing character ($M_{C}$)}
+%When the overall number of morphological characters in the matrix ($M_{C}$) decreases
+
+%Story for MC: MC is expected to have an effect two times bigger than ML or MF, e.g.
+removing 10\% of MF removes 5\% of data (because it works only on the fossil part) and removing 10\% of MC actually removes 10\% of that.
+%Method
+%However the effect of removing characters has an effect of topology of the same order of magnitude than for ML and MF (FigRes 1): For RF, Bayesian consensus is better than ML and for Tr, MaxLik is better than Baycon and have the same drop at 50\% as for MF (different than for ML).
+
+%In this scenario, the overall number of missing characters ($M_{C}$) affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
 % NC: What's the relevance of that sentence?
-However, the pairwise Bhattacharyya Coefficients show that the area with the lowest probability of distributional overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
+%However, the pairwise Bhattacharyya Coefficients show that the area with the lowest probability of distributional overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
 %NC: I don't understand what you mean here
 
-% $§2: There is a different effect form each parameter >...
-% ^§3: >... because each parameter tells a different story.
+\subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
+%In our simulations, the overlap in data is greatly reduced when all three missing data parameters have a lot of missing data (e.g.
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+%When combining the effect, we
 
-%§3
-%Discussing results 1 % NC: This can join to the end of the previous paragraph
-This variable effect of our missing data parameters on the ability to recover the "correct" topology (Fig.~\ref{Fig_Results-paircomp_within}) can be explained by % the way our parameters are illustrating the sources of missing data in empirical matrices
-% NC: No.
-This shows how our parameters relate to the real world and give real world solutions.
-Probably first want to explain any difference in effects agnostically - i.e.
-just as parameters with no biological context.
-So maybe: ..."can be explained by the different way data is removed using each of our three missing data parameters..." THEN move on to this:
-
+%Our solutions
 Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
@@
 
-% $§3: This is how we can fix the parameters >...
-% ^§4: >... but the methods also influence a lot the ability to recover the right topology.
-
-%§4
-%Result 2 from literature.
-%   -Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
+% ----->
+% TG: Put that part somewhere in the start with the discussion of ML vs Baycon in the effect of ML part?
+%-Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
 Variation in our ability to recover the "correct" topology also depended heavily on the method used to infer the tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 %TG: I guess it's a good selling point: we ARE the first ones to do a sensitivity analysis in a TEM framework (as far as I know).
+% <-----
+
 %Discussing results 2
-%   -Why is Baycon better than the others?
-When using the standardised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "correct" topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}).
-In fact, regardless of the amount of missing data, Bayesian consensus trees always displays the highest normalised/standardised Robinson-Foulds score (always $>$ 0.69; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 %   -Why is Baycon so different than Baytre? Since Baycon is calculated from Baytre.
-We saw the poorest topological recovery using Maximum Likelihood bootstrap trees and Bayesian posterior distribution of trees (% NC: FIGURE???
-).
-This reflects the fact it is difficult to compare two distributions of trees, and each comparison between a set of missing data parameters and "best" tree in these cases involved 1000 random pairwise comparisons rather than just one.
-This will obviously add noise to the results for these two methods.
 Also, in the case of the lower performances of the Bayesian consensus trees when using the Triplets distance, one can note the high overlap of the confidence intervals (Fig.
@@
 
-%§5
-%Caveats: the metrics
-%   -RF is different and better than triplets.
-
-% NC: I think this can be condensed and slotted into the correct places above.
-So where you speak about the missing data parameters, mention differences in how they influence Rf and Triplets and why that might be.
-Then when you mention methods do the same.
-
-\label{metrics_discussion}
-This difference between the metrics can be explained by their respective ability to measure the differences between two trees \citep{kuhnerpractical2014}.
+%\label{metrics_discussion}
+%This difference between the metrics can be explained by their respective ability to measure the differences between two trees \citep{kuhnerpractical2014}.
 As shown in Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}, the contrast between the results of the two metrics can be explained by the actual nature of both metrics (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}):
-\begin{enumerate}
-\item{The Robinson-Foulds metric is a conservative metric, which is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same taxa with the same topologies \citep{RF1981}.
+%\begin{enumerate}
+%\item{The Robinson-Foulds metric is a conservative metric, which is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same taxa with the same topologies \citep{RF1981}.
 When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
@@
 when the normalised Robinson-Foulds distance is close to 1, most of the clades in the two trees are identical).}
-\item{On the other hand, the Triplets method is measuring the position of each taxon towards two other reference taxa \citep{critchlowthe1996}.
+%\item{On the other hand, the Triplets method is measuring the position of each taxon towards two other reference taxa \citep{critchlowthe1996}.
 It will penalise only trees when taxa get removed furthest from their original position.
@@
 when the score is closed to 1, few wildcard taxa are present in the tree).}
-\end{enumerate}
-Therefore, in our study, both metrics are showing two different aspects of tree topology.
+%\end{enumerate}
+%Therefore, in our study, both metrics are showing two different aspects of tree topology.
 For example, regarding the Bayesian consensus trees, completes clade are conserved most of the time (Fig.~\ref{Fig_Results-global_perparam}-top) but with an increasing amount of missing data, wildcard taxa get more and more displaced (Fig.~\ref{Fig_Results-global_perparam}-bottom).
@@
 
-% $§5: The difference in the methods is not important if you look at Kuhner >...
-% ^§6: >... but a final caveat is the way we simulate data.
-
-%§6
-%Caveats: the simulations
+%Opening/further work part - NC(meeting): Be less defensive (i.e. nothing seemed to have gone wrong THIS time).
 %   -Simulating morpho characters

### 2015-01-07 Manuscript/TEM_draft.tex
Edits to results and methods

@@
 
-% NC: THIS SECTION NEEDS WORK!
 In this study, we focus on the effect of missing data on our ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{}.
@@
 \citealt{ronquista2012}), we do not consider branch lengths in this study.
-This is partially due to difficulties of estimating the effect of missing data simultaneously on branch length and topology, % TG: or "This is partially due to difficulties with simulating branch lengths and topology simultaneously"
+This is partially due to difficulties with simulating branch lengths and topology simultaneously,
 but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
@@
 This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
-We normalised this metric using the Normalized Tree Similarity index \citep{Bogdanowicz2012}.
-This index allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for the calculation details).
-Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the Normalised Robinson-Foulds distance is equal to zero, the trees are no more different than expected by chance; finally if the Normalised Robinson-Foulds distance is lower than zero, the trees are more different than expected by chance.
+We normalised this metric by % NC: DETAILS
+\citep{Bogdanowicz2012}.
+This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for the calculation details).
+Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
 
@@
 Therefore this metric sensitive to the conservation of wildcard taxa.
-We normalised this metric in the same way as we did for the Robinson-Foulds distance resulting in the Normalised Triplets distance.
-
-%\subsubsection{Normalized Tree Similarity}
-%The Normalized Tree Similarity index, $NTS_m$ (\citealp{Bogdanowicz2012}) allows us to compare the two metrics above for any number of taxa, by scaling the value of the metric $m$ (either Robinson-Foulds or Triplets distance in our study) to the expected value of the metric $m$ when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}).
-When $NTS_m$=1, the two trees are identical, when $NTS_m$=0 the trees are no more different than expected when comparing two random trees, and when $NTS_m$$<$0, the difference between the two trees is greater than when comparing two random trees.
-In our study we used the $NTS_m$ index as a proxy for topological similarity: a high score of this index (i.e.
-towards one) means that the topology is highly conserved between the two trees; conversely, a low score of this index (i.e.
-towards 0) means that the topological difference between the two trees is as great as expected when comparing two random trees.
-Note that the two tree comparison metrics represent different aspects of topological similarity and do not perform equally well (see \hyperref[metrics_discussion]{discussion below} and \citealt{kuhnerpractical2014}).
+We normalised this metric in the same way as for the Robinson-Foulds distance resulting in the Normalised Triplets distance.
 
@@
 \subsubsection{Summarizing the comparisons}
-Finally after doing all theses %stupid
-50 simulations and comparisons, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) on the ability to recover the correct topology in a Total Evidence framework.
-Additionally, we also assessed the effect of the different methods (i.e.
-Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).% and the joint effect of the our missing data parameters and the methods.
-% TG: the following part might be useless (complete repetition of above)
- To assess the ability to recover the "correct" tree topology on our 50 simulations, we used the scores for both normalised metrics (Robinson-Foulds and Triplets distance).
-We looked at the following effects:
-\begin{enumerate}
-\item{The three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$);}
-\item{The four "methods" (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution);}
-%\item{The joint effect of the missing data parameters and the methods.} % actually this is implied in both already.
-\end{enumerate}
-% TG: end of the repetition
+Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the correct topology in a Total Evidence framework.
+We also assessed the effect of our missing data parameters jointly with the effects of different tree inference and uncertainty methods (i.e.
+Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution).
 
 %Introducing Bhattacharyya Coefficient
-To evaluate these effects we measured the similarities among the different normalised metrics scores (Robinson-Foulds and Triplets distance) for all the analysis above using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions, ranging from 0 to 1 (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} for calculation details).
+To evaluate these effects we measured similarities among different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for all the analyses above using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
+The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} for calculation details).
 
 %Explaining the parameter pairwise comparisons (ANOVA equivalent=within)
-For assessing the effect of our three missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$), we calculated the Bhattacharyya Coefficient between each pairwise combination for each parameter state (0\%, 10\%, 25\%, 50\%, 75\%).
-This resulted in 7875 pairwise distributions pairwise comparisons (triangular $3^5$$\times$$3^5$ matrix).
-We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient=1) and in red when they don't (Bhattacharyya Coefficient=0) - Fig.~\ref{Fig_Results-paircomp_within}).
+For assessing the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the Normalised Robinson-Foulds and Normalised Triplets distance distributions, for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%).
+For example, $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc.
+This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
+We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient = 1) and in red when they don't (Bhattacharyya Coefficient = 0) - Fig.~\ref{Fig_Results-paircomp_within}).
 We repeated this procedure for all the "methods" (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).
@@
 %We also take into account the effect of the methods used. REPHRASE
-This way we could investigate the effect of each our three missing data parameters and their interaction on the ability to recover the "correct" topology.
+%This way we could investigate the effect of each our three missing data parameters and their interaction on the ability to recover the "correct" topology.
 
@@
 %Part 2
-\subsection{Effect of missing living taxa ($M_{L}$)} %Or remove? Also below then
-As the amount of missing living taxa increases, the ability to recover the correct topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-However, the decrease of the Normalised Robinson-Fould distance plateaus at minimum of 0.69 for Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the exact values) and there is no substantial difference in topological recovery for 50\% or 75\% of missing living taxa (upper triangle in Fig.~\ref{Fig_Results-paircomp_within}-A).
+\subsection{Individual effects of missing data parameters} 
+As the amount of missing data increases across all three parameters, the ability to recover the correct topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
+However, the decrease of the Normalised Robinson-Fould distance plateaus at minimum of 0.69 for Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values) and there is no substantial difference in topological recovery for 50\% or 75\% of missing living taxa (upper triangle in Fig.~\ref{Fig_Results-paircomp_within}-A).
 Therefore, clades seems to be more conserved in the Bayesian consensus tree more than any other method (see discussion).

### 2015-01-07 Manuscript/TEM_draft.tex
Cleaning dealt comments

@@
 % NC: Can you centre these around the boxes? and make the comparing topologies arrow touch the two triangles? I'll show you what I mean. 
-% TG: I have different versions for the figure available (you can see them opening it with inkscape I think - else I'll show you).
 % NC: Also check the figure is in a format and of correct quality for Syst Biol.
@@
 
-\subsection{1-Generating the matrix} % NC: Can you number these subsections like they are in the steps above? % TG: 
+\subsection{1-Generating the matrix}
 First we randomly generated a "true" tree of 50 taxa in R v.
@@
 This reflects the fact that detailed morphological data for living species are scarce.
-% NC: Not sure you can really "blame" this on DNA data availability!
 \item{$M_{F}$, the proportion of missing data in the fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%.}
@@
 8.0.20; \citealt{Stamatakis21012014}).% as a generalization of the HKY + $\Gamma_4$ model \citep{HKY85}.
-%The GTR model can be seen as a generalization of the HKY model (the two parameters from the HKY model are implicitly included in the six from GTR model - \citealt{stamatakisa2008}).
-- TG: Is that useful?
-% NC: I think most people know that already.
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} %which is a generalization of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2 - TG: same?
-% NC: Again leave it out
+The GTR model can be seen as a generalization of the HKY model (the two parameters from the HKY model are implicitly included in the six from GTR model - \citealt{stamatakisa2008}).
+For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} %which is a generalization of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2
  assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
@@
 
-To measure the phylogenetic signal of our simulations, we first ran a fast bootstrap analysis with 500 replicates on the "complete" matrix.
+To measure the phylogenetic signal of our simulations, we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
 We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
@@
 Nearest Neighbouring Interchange - NNI) for recovering good bootstrap values \citep{salamin2003}.
-%TG: The explanations of the algorithm part for the BS was proposed by Trevor I think it's useless
-% NC: perhaps somewhere in brackets just mention it's LSR bootstrapping. No need for the rest I agree.
 Using these parameters took \texttildelow 8 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (2.30GHz clock speed nodes).
@@
 
-\subsubsection{Paired trees comparisons} % NC: OK I'm lost in this section!
-We should discuss then think about how to express this in an understandable fashion.
-TG: or no subheading at all?
-See next section as well.
+\subsubsection{Paired trees comparisons}
 \label{tree_comparisons}
-For the Maximum Likelihood and Bayesian consensus
-% NC: is this actually a maximum clade credibility tree - TG: better stick with consensus as described above?
-trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting each time in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance.
+For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting each time in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance.
 %For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and n = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}).
-% TG: That is now described in the RF distance subsection just above
 %We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree.
 We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}.
-TG: explained later on
 Also, to take into account the uncertainty of tree inference in both methods cited above, instead of using the single Maximum Likelihood and Bayesian consensus trees, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 from the posterior tree distribution of the Bayesian analysis for the "best" trees and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees).
@@
 3.1 \citep{hdrcde}.
-%Finally we tested for significant differences among the different distributions of trees using non-parametric group and pairwise comparisons by using respectively Kruskal-Wallis and Nemenyi-Damico-Wolfe-Dunn test (CITE R PACKAGE)\citet{ruxtontime2008}.
-TG: finally we didn't (Andrew's advice - see Battacharyya)
 
@@
 We use the Bhattacharyya Coefficient in a similar way as a two-sided t-test to detect significant differences between the distributions of the normalised distance between the "best" tree and the "missing data" trees (for both metrics).
-%We also take into account the effect of the methods used. REPHRASE
-%This way we could investigate the effect of each our three missing data parameters and their interaction on the ability to recover the "correct" topology.
 
@@
 
-% NC: Split into three sections to answer three questions
 % TG: Ok I removed all the figures + table on top of the section in here and all the text down. We might to have a think of where to anchor the figures later on.
-
-%\subsection{The effects of missing data on tree topology}
 
@@
 
+%Removed figure (table instead)
 %\begin{figure} 
@@
 
+% TG: Put this table in landscape!
 \begin{table}[ht]

### 2015-01-07 Manuscript/TEM_draft.tex
Changed correct topo into best tree's topo

@@
 
-In this study, we focus on the effect of missing data on our ability to recover a correct topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{}.
+In this study, we focus on the effect of missing data on our ability to recover a "best" tree's topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{}.
 %CITE-check Lloyd and Friedman or birdy stuff
@@
 Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
-We find that the ability of Total Evidence methods to recover the correct topology is increased when using Bayesian methods.
-Also minimizing the proportion of living taxa with no morphological data and the proportion of missing morphological characters improves the ability of Total Evidence methods to recover the correct topology more so than minimizing the proportion of missing data in the fossil record.
+We find that the ability of Total Evidence methods to recover the "best" tree's topology is increased when using Bayesian methods.
+Also minimizing the proportion of living taxa with no morphological data and the proportion of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree's topology more so than minimizing the proportion of missing data in the fossil record.
 
@@
 \subsubsection{Summarizing the comparisons}
-Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the correct topology in a Total Evidence framework.
+Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the "best" tree's topology in a Total Evidence framework.
 We also assessed the effect of our missing data parameters jointly with the effects of different tree inference and uncertainty methods (i.e.
@@
 %Explaining the methods comparisons (ANOVA equivalent=among)
-Finally, for assessing the effect of the method on the ability to recover the "correct" topology, we compared the similarity between the methods.
+Finally, for assessing the effect of the method on the ability to recover the "best" tree's topology, we compared the similarity between the methods.
 We calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g.
@@
 %Part 1
-As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the correct topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
+As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the "best" tree's topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
 However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 \subsection{Individual effects of missing data parameters} 
-As the amount of missing data increases across all three parameters, the ability to recover the correct topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
+As the amount of missing data increases across all three parameters, the ability to recover the "best" tree's topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
 However, the decrease of the Normalised Robinson-Fould distance plateaus at minimum of 0.69 for Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values) and there is no substantial difference in topological recovery for 50\% or 75\% of missing living taxa (upper triangle in Fig.~\ref{Fig_Results-paircomp_within}-A).
@@
 \subsection{Effect of missing data for fossil taxa ($M_{F}$)}
-As the amount of missing data in the fossil part of the matrix increases, the ability to recover the correct topology decreases as well (Fig.~\ref{Fig_Results-permeth_perparam}).
+As the amount of missing data in the fossil part of the matrix increases, the ability to recover the "best" tree's topology decreases as well (Fig.~\ref{Fig_Results-permeth_perparam}).
 For the Robinson-Fould distance, we observe the same pattern than for $M_{L}$ with a better topological recovery of the Bayesian consensus trees on any other methods (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 \subsection{Effect of missing character ($M_{C}$)}
-As described for the two previous parameters $M_{L}$ and $M_{F}$, an increase in the amount of missing characters decreases the ability to recover the correct topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
+As described for the two previous parameters $M_{L}$ and $M_{F}$, an increase in the amount of missing characters decreases the ability to recover the "best" tree's topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
 Similarly again, the Bayesian consensus trees shows higher Normalised Robinson-Fould distance than the Maximum Likelihood tree (Fig.~\ref{Fig_Results-permeth_perparam}).
@@
 \subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
-The combine of the missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) reduces the ability to recover the correct topology when each parameters contains a maximum of missing data (i.e.
+The combine of the missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) reduces the ability to recover the "best" tree's topology when each parameters contains a maximum of missing data (i.e.
 $M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-global_perparam} and see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the combined effects of the parameters in both Normalised Robinson-Foulds and Normalised Triplets distance for the Maximum Likelihood Bootstraps trees and the Bayesian posterior trees distribution).
@@
 
-%As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the correct topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
+%As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the "best" tree's topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
 %However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 %The Normalised Robinson-Foulds distance calculate for the trees inferred in Maximum Likelihood framework decreases with the increase of missing data in our three parameters ($M_{L}$, $M_{F}$ or $M_{C}$) with a modal Normalised Robinson-Fould score always below 0.5 when 75\% of the living taxa have no data ($M_{L}$=75\% - Fig.~\ref{Fig_Results-global_perparam}).
-However, in Bayesian inference, the ability of the method (the Bayesian consensus tree) to recover a "correct" topology plateaus at a minimal Normalised Robinson-Fould score of 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the values distribution).
+However, in Bayesian inference, the ability of the method (the Bayesian consensus tree) to recover a "best" tree's topology plateaus at a minimal Normalised Robinson-Fould score of 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the values distribution).
 The results of the topological recovery with the parameters combination in both Normalised Robinson-Foulds and Triplets distances for the Maximum Likelihood bootstrap trees and the Bayesian posterior distribution trees are available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.
@@
 Probably need to think some more about this.
-% TG: a note about the "correct" topology regarding your comment just after the discussion header: should I replace "correct" by "best" since that's how we're calling the "best" tree anyway (all in quotes)
+% TG: a note about the "best" tree's topology regarding your comment just after the discussion header: should I replace "correct" by "best" since that's how we're calling the "best" tree anyway (all in quotes)
 
@@
 
-Our results show that the ability to recover the correct topology in a Total Evidence framework decreases with increasing missing data, regardless of how data were removed or the method of tree inference used.
+Our results show that the ability to recover the "best" tree's topology in a Total Evidence framework decreases with increasing missing data, regardless of how data were removed or the method of tree inference used.
 However, these factors affected topological recovery in different ways and to different extents.
 Decreasing the overall number of morphological characters in the matrix ($M_{C}$) and the amount of living taxa with morphological data ($M_{L}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
-Additionally, using Bayesian consensus trees recovered the correct topology more consistently than using Maximum Likelihood trees or Bayesian posterior trees distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
+Additionally, using Bayesian consensus trees recovered the "best" tree's topology more consistently than using Maximum Likelihood trees or Bayesian posterior trees distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
 % ----->
@@
 %Part saying the difference between ML and Baycon
-When using the Normalised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "correct" topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}).
+When using the Normalised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "best" tree's topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}).
 In fact, regardless of the amount of missing data, Bayesian consensus trees always displays the highest Normalised Robinson-Foulds score (always $>$ 0.69; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for details).
-However, regarding the Normalised Triplets distance, the topological similarity with the "correct" topology decreases in the same way for both the Maximum Likelihood trees and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+However, regarding the Normalised Triplets distance, the topological similarity with the "best" tree's topology decreases in the same way for both the Maximum Likelihood trees and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 %Part saying the difference between BS and baytrees
-Regardless of the metric, the Maximum Likelihood boostrap trees and the Bayesian posterior trees distribution always performed poorlier in their ability to recover the "correct" topology, even when no morphological data was missing ($M_{L}$=0\% - Fig.~\ref{Fig_Results-permeth_perparam}).
+Regardless of the metric, the Maximum Likelihood boostrap trees and the Bayesian posterior trees distribution always performed poorlier in their ability to recover the "best" tree's topology, even when no morphological data was missing ($M_{L}$=0\% - Fig.~\ref{Fig_Results-permeth_perparam}).
 This reflects the fact it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" tree and "best" tree in these cases involved 1000 random pairwise comparisons rather than just one.
@@
 Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
-Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "correct" topology.
+Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "best" tree's topology.
 We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections.
@@
 However, gaps in the matrix can be filled with future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
-Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "correct" topology (Fig.~\ref{Fig_Results-paircomp_within}).
+Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "best" tree's topology (Fig.~\ref{Fig_Results-paircomp_within}).
 Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
@@
 %-Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
-Variation in our ability to recover the "correct" topology also depended heavily on the method used to infer the tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+Variation in our ability to recover the "best" tree's topology also depended heavily on the method used to infer the tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 Previous studies have shown some superiority of methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}).
@@
 Regarding this fundamental differences, the Robinson-Foulds metric performs better in showing a distance from the true topology than the Triplets distance \citep{kuhnerpractical2014}.
-Therefore, the results for the Bayesian consensus trees can be preferred on the Maximum Likelihood trees because - when using the Robinson-Fould distance - they display a higher ability to recover the "correct" topology (Fig.
+Therefore, the results for the Bayesian consensus trees can be preferred on the Maximum Likelihood trees because - when using the Robinson-Fould distance - they display a higher ability to recover the "best" tree's topology (Fig.
 ~\ref{Fig_Results-permeth_perparam}).
@@
 Nevertheless, our simulation already show clear results of the effect of missing data on topology in a Total Evidence framework.
-Also, our results show how recovering the correct topology can be improved by using Bayesian consensus trees in combination with a maximum amount of characters and a of living taxa with morphological data.
+Also, our results show how recovering the "best" tree's topology can be improved by using Bayesian consensus trees in combination with a maximum amount of characters and a of living taxa with morphological data.
 %   -Looking at branch length?
@@
 %   -Go code living taxa
-Using Total Evidence matrices with missing data is not a problem for recovering the correct topology as long as enough morphological data is overlapping between the fossil and living taxa.
+Using Total Evidence matrices with missing data is not a problem for recovering the "best" tree's topology as long as enough morphological data is overlapping between the fossil and living taxa.
 %NC: Explain more clearly what you mean by overlap - of characters?

### 2015-01-07 Manuscript/TEM_draft.tex
Updates on the method part
Structure:
-Individual effects of missing data parameters
-Combined effect of missing data parameters

@@
 %Part 1
-As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the "best" tree's topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
+%increase in missing data leads to decrease in topological recovery no matter what
+As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the "best" tree's topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
 However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-In any combination of parameters or metrics, the Maximum Likelihood bootstraps trees and the Bayesian posterior distribution trees performs in a substantially similarly (Table ~\ref{Tab_Results-Difference_methods}) and worth than the Maximum Likelihood and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} - see discussion).
-%add discussion link
+%bootstrap and bayesian treesets are always worst bad
+In any combination of missing data parameters or topological distance metrics, the Maximum Likelihood bootstraps trees and the Bayesian posterior tree distribution performs substantially similarly (Median Bhattacharrya Coefficient between both tree inference methods using Normalised Robinson-Foulds distance or Normalised Triplets distance is respectively 0.85 and 0.96 - see Table ~\ref{Tab_Results-Difference_methods}) and worth than the Maximum Likelihood trees and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} - see discussion %add discussion link
+ and also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the effect of the combined parameters on the Maximum Likelihood bootstraps trees and the Bayesian posterior tree distribution).
 
@@
 \subsection{Individual effects of missing data parameters}
-As the amount of missing data increases across all three parameters, the ability to recover the "best" tree's topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-However, the decrease of the Normalised Robinson-Fould distance plateaus at minimum of 0.69 for Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values) and there is no substantial difference in topological recovery for 50\% or 75\% of missing living taxa (upper triangle in Fig.~\ref{Fig_Results-paircomp_within}-A).
-Therefore, clades seems to be more conserved in the Bayesian consensus tree more than any other method (see discussion).
-%add discussion link
-On the other hand, the decrease in Normalised Triplets distance is less contrasted with a greater overlap of Normalised Triplets distance with the Maximum Likelihood tree (Fig.~\ref{Fig_Results-permeth_perparam}) - mean Bhattacharrya Coefficient=0.10 for the Normalised Robinson-Foulds distance and 0.61 for the Normalised Triplets distance - see table ~\ref{Tab_Results-Difference_methods}).
+%Individually, increase in missing data leads to decrease in topological recovery as well
+As the amount of missing data increases across all three parameters, our ability to recover the "best" tree's topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
+%Differences between MaxLik and Baycon in RF
+The Normalised Robinson-Foulds distance is always substantially lower for the Maximum Likelihood %(minimum Normalised Robinson-Foulds distance for $M_{L}$, $M_{F}$ and $M_{C}$ of respectively @, @ and @) %minimum RF value for each parameter individually% TG: or useless (just Bhattacharrya)?
+ trees than for the Bayesian consensus trees %(minimum Normalised Robinson-Foulds distance for $M_{L}$, $M_{F}$ and $M_{C}$ of respectively @, @ and @)  %minimum Tr value for each parameter individually% TG: or useless (just Bhattacharrya)?
+(Bhattacharrya Coefficient between the methods for $M_{L}$, $M_{F}$ and $M_{C}$ of respectively @, @ and @  %BC value for each parameter individually between both methods RF
+- Fig.~\ref{Fig_Results-permeth_perparam} upper and see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
+%Differences between MaxLik and Baycon in Tr
+However, the Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for the $M_{L}$ parameter (Bhattacharrya Coefficient between both tree inference methods for $M_{L}$ = @ %BC value for ML between both methods
+- Fig.~\ref{Fig_Results-permeth_perparam} lower and see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values); for the two other parameters $M_{F}$ and $M_{C}$, the Normalised Triplets distance is substantially higher for the Maximum Likelihood trees than for the Bayesian consensus trees (Bhattacharrya Coefficient between both tree inference methods for $M_{F}$ and $M_{C}$ of respectively @ and @  %BC value for each parameter individually between both methods Tr
+- Fig.~\ref{Fig_Results-permeth_perparam} lower and see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
 
 %Part 3
-\subsection{Effect of missing data for fossil taxa ($M_{F}$)}
-As the amount of missing data in the fossil part of the matrix increases, the ability to recover the "best" tree's topology decreases as well (Fig.~\ref{Fig_Results-permeth_perparam}).
-For the Robinson-Fould distance, we observe the same pattern than for $M_{L}$ with a better topological recovery of the Bayesian consensus trees on any other methods (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-However, the Normalised Triplets distance drops drastically when going from 25\% of missing data to 50\% (mode) %get the modal values here
-making this parameter more sensible to wildcard taxa in the Bayesian consensus tree (Fig.~\ref{Fig_Results-permeth_perparam} - see discussion). %add discussion link
+\subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
+%Combined, increase in missing data leads to decrease in topological recovery as well
+The combined effect of the missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) reduces the ability to recover the "best" tree's topology when each parameters contains a maximum of missing data (i.e.
+$M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-global_perparam} and see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the combined effects of the parameters in both Normalised Robinson-Foulds and Normalised Triplets distance for the Maximum Likelihood Bootstraps trees and the Bayesian posterior trees distribution).
+%Differences between MaxLik and Baycon in RF
+As described above, The Normalised Robinson-Foulds distance is higher for the Bayesian consensus trees topologies than for the Maximum Likelihood trees topologies with a median Normalised Robinson-Foulds distance for the Bayesian consensus trees and for the Maximum likelihood trees of respectively @ and @ % Median RF value for Baycon and MaxLik
+(Fig.~\ref{Fig_Results-global_perparam} upper - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
+%Differences between MaxLik and Baycon in Tr
+The Normalised Triplets distance is lower for the Bayesian consensus trees topologies than for the Maximum Likelihood trees topologies with a median Normalised Triplets distance for the Bayesian consensus trees and for the Maximum likelihood trees of respectively @ and @ % Median Tr value for Baycon and MaxLik
+(Fig.~\ref{Fig_Results-global_perparam} lower - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
+%Significance
+However, the Normalised Robinson-Foulds distances are substantially less similar between the Bayesian Consensus trees and the Maximum Likelihood trees than the Normalised Triplets distances (Median Bhattacharrya Coefficient between both tree inference methods for the Normalised Robinson-Foulds distance and the Normalised Triplets distance is respectively 0.10 and 0.61 - Fig.~\ref{Fig_Results-global_perparam} and table~\ref{Tab_Results-Difference_methods}).
 
-%Part 4
-\subsection{Effect of missing character ($M_{C}$)}
-As described for the two previous parameters $M_{L}$ and $M_{F}$, an increase in the amount of missing characters decreases the ability to recover the "best" tree's topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-Similarly again, the Bayesian consensus trees shows higher Normalised Robinson-Fould distance than the Maximum Likelihood tree (Fig.~\ref{Fig_Results-permeth_perparam}).
-And in a similar way as for the $M_{F}$ parameter, the Normalised Triplet distance drops drastically when going from 25\% of missing data to 50\% (mode).
-%get the modal values here
-
-%Part 5
-\subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
-The combine of the missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) reduces the ability to recover the "best" tree's topology when each parameters contains a maximum of missing data (i.e.
-$M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-global_perparam} and see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the combined effects of the parameters in both Normalised Robinson-Foulds and Normalised Triplets distance for the Maximum Likelihood Bootstraps trees and the Bayesian posterior trees distribution).
-Regarding the clade conservation (i.e.
-the Normalised Robinson-Foulds distances), the Bayesian consensus tree greatly outperforms the Maximum likelihood trees with Normalised Robinson-Foulds scores always superior than 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the exact values) regardless the combination of missing data parameters (Fig.~\ref{Fig_Results-paircomp_within}-A).
-However, for the Normalised Triplets distance, the Bayesian consensus tree often displays negative values (i.e.
-differences greater than expected by chance) but with few substantial differences (Fig.~\ref{Fig_Results-paircomp_within}-B).
-Also the differences between the Normalised scores of the Maximum Likelihood trees and Bayesian consensus trees are substantially difference for the Normalised Robinson-Foulds distance (mean Bhattacharrya Coefficient=0.10) and for the Normalised Triplets distance (mean Bhattacharrya Coefficient=0.61 - see table ~\ref{Tab_Results-Difference_methods}).
-Thus the sensitivity of the Bayesian consensus trees to wild card taxa is not highly different than the one for the Maximum likelihood trees (i.e.
-the distribution of Normalised Triplets distance have a high probability of overlap).
-The results of the probability of overlap between the pairwise distributions of Normalised Robinson-Foulds and Triplets distances of the Maximum Likelihood trees, the Maximum Likelihood Bootstrap trees and the Bayesian posterior trees distribution is available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.
-
-%Binned removed bits from former results.
-
-%As the amount of missing data in the morphological part of the total evidence matrix increases, our ability to recover the "best" tree's topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the method of tree inference (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
-%However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-% And figure 3 too right?
-TG: Yep
-%The Normalised Robinson-Foulds distance calculate for the trees inferred in Maximum Likelihood framework decreases with the increase of missing data in our three parameters ($M_{L}$, $M_{F}$ or $M_{C}$) with a modal Normalised Robinson-Fould score always below 0.5 when 75\% of the living taxa have no data ($M_{L}$=75\% - Fig.~\ref{Fig_Results-global_perparam}).
-However, in Bayesian inference, the ability of the method (the Bayesian consensus tree) to recover a "best" tree's topology plateaus at a minimal Normalised Robinson-Fould score of 0.69 (Fig.~\ref{Fig_Results-global_perparam} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the values distribution).
-The results of the topological recovery with the parameters combination in both Normalised Robinson-Foulds and Triplets distances for the Maximum Likelihood bootstrap trees and the Bayesian posterior distribution trees are available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.
-%All parameters combinations are different than the "best" tree (always red) in both metrics.
-%The lowest probability of overlap (low Bhattacharrya Coefficient values) between pairwise distributions of parameters combinations ($M_{L}$, $M_{F}$ and $M_{C}$ with 0\%, 10\%, 25\%, 50\% or 75\% missing data) is always between the comparison of Normalised distance (Robinson-Foulds or Triplets) of the "missing data" tree with no missing data ($M_{L}$=0\%, $M_{F}$=0\% and $M_{C}$=0\%) and all the other "missing data" trees.
-%The parameters combinations are less dissimilar in Triplets than in RF (less red).
-%However, pairwise comparisons are more dissimilar (low Bhattacharrya Coefficient values) in the Normalised Robinson-Fould distance results than in the Normalised Triplets distance results.
-%The ML parameter (0% and 10%) are more different in RF and Triplets (red stripes at the base). The ML parameter (50, 75) are the more similar (green upper part) in RF.
-%The Normalised Robinson-Fould distance for the comparisons between the trees with a low amount of missing data for the living taxa ($M_{L}$=0\% and 10\%) shows the most dissimilarities with the other parameter combinations.
-%The combination of the ML parameter (0% and 10%) and the MC parameter (0, 10, 25) are the most different (details on the red stripes).
-%Check supplementary for the rest
-%The results of the probability of overlap between the pairwise distributions of Normalised Robinson-Foulds and Triplets distances of the Maximum Likelihood trees, the Maximum Likelihood Bootstrap trees and the Bayesian posterior trees distribution is available in the \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}.
-%When looking at the pairwise Bhattacharyya Coefficients between all the parameters combinations for Bayesian consensus trees, % NC: Instead just state the result.
-"We found that number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (table X)".
-In the same way as you wouldn't say "Using a linear regression we found that x ~ Y." You'd say "X and Y were significantly correlated (Table1)".
-%NC: Also what IS the result really?
-You find that ML overlaps less with the other parameters so it's significantly different in it's effect?
-And from the graph you can see it's worse?
-What does oerlap MEAN in terms of the answer to the question, i.e.
-how do the parameters affect topological recovery.
-%the number of missing living taxa ($M_{L}$) results the lowest probabilities of overlap (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}-A).
-However, when looking at the Triplets distance, because of the high overlap in confidence intervals, there is no major region with low probability of overlap.
-One should note however that the bottom line of the matrix where the Bhattacharyya Coefficients are very low is due to comparing the distribution of the pairwise comparison of the "best" tree against the "missing-data" tree with no missing data ($M_{L}$=0\%, $M_{C}$=0\%, $M_{F}$=0\%) , leading to a normalised Triplets score of 1 every time (with no variance, see Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-The pairwise comparisons for the other methods (Maximum Likelihood, Bootstraps and Bayesian posterior trees) are available in the supplementary materials (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-%NC: I'd start by describing the pattern with reference to figs 2 and 3 and then the "significant" results
-%When looking at the comparisons between the methods 
-%using the Robinson-Foulds distance, the Bayesian consensus trees have the least overlap with respectively the Bayesian posterior trees and the Bootstrap (Fig.~\ref{Fig_Results-paircomp_without}-A).
-Using the Triplets distance, the two methods that have the most distribution %NC: ???
-%are the Bootstraps and the Bayesian posterior trees (Fig.~\ref{Fig_Results-paircomp_without}-B).
-These results are due to the difference in comparing topologies between Maximum Likelihood and Bayesian consensus trees and Bootstraps and Bayesian posterior trees: the latter two are based on a random pairwise comparison of a sub-sample of 1000 trees.
-This is likely to add noise to the data.
+%And now the combined effect of parameters in RF for the Baycon trees
+Regarding the effect of the combination of our missing data parameters on the ability to recover the "best" tree's topology, using both Normalised Robinson-Foulds and Normalised Triplets distances from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameters combination (red %change color
+ line at the bottom of Fig.~\ref{Fig_Results-paircomp_within} A and B).
+However, the Normalised Robinson-Foulds distance displays more dissimilarities than the Normalised Triplets distance (the median Bhattacharrya coefficient among all pairwise missing data parameters combinations from the Bayesian consensus trees using the Normalised Robinson-Foulds and the Normalised Triplets distance are respectively @ and @ %value for median BC from pairwise comp in Baycon-RF and Baycon-Tr
+ Fig.~\ref{Fig_Results-paircomp_within} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
+Additionally, when using the Normalised Robinson-Foulds distance, there is no dissimilarity between the trees with 50\% and 75\% of missing living taxa ($M_{L}$) and any other combination of parameter state for the two other missing data parameters $M_{F}$ and $M_{C}$ (e.g..
+$M_{L}$ = 50\%, $M_{F}$ = 0\%, $M_{C}$ = 10\%; or $M_{L}$ = 50\%, $M_{F}$ = 75\%, $M_{C}$ = 50\%; etc) as well as for the trees with 50\% and 75\% of missing characters ($M_{C}$) and any other combination of parameter state for the two other missing data parameters $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-global_perparam} upper and Fig.~\ref{Fig_Results-paircomp_within}-A).
 
@@
 
-Our results show that the ability to recover the "best" tree's topology in a Total Evidence framework decreases with increasing missing data, regardless of how data were removed or the method of tree inference used.
+Our results show that the ability to recover the "best" tree's topology in a Total Evidence framework decreases with increasing amount of missing data, regardless of how data were removed or the method of tree inference used.
 However, these factors affected topological recovery in different ways and to different extents.
@@
 However, regarding the Normalised Triplets distance, the topological similarity with the "best" tree's topology decreases in the same way for both the Maximum Likelihood trees and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+
+% ----->
+% TG: Put that part somewhere in the start with the discussion of ML vs Baycon in the effect of ML part?
+%-Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
+Variation in our ability to recover the "best" tree's topology also depended heavily on the method used to infer the tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+Previous studies have shown some superiority of methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}).
+However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
+%TG: I guess it's a good selling point: we ARE the first ones to do a sensitivity analysis in a TEM framework (as far as I know).
+% <-----
+
 %Part saying the difference between BS and baytrees
@@
 
-% ----->
-% TG: Put that part somewhere in the start with the discussion of ML vs Baycon in the effect of ML part?
-%-Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
-Variation in our ability to recover the "best" tree's topology also depended heavily on the method used to infer the tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-Previous studies have shown some superiority of methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}).
-However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-%TG: I guess it's a good selling point: we ARE the first ones to do a sensitivity analysis in a TEM framework (as far as I know).
-% <-----
-
-%Discussing results 2
-%   -Why is Baycon so different than Baytre? Since Baycon is calculated from Baytre.
-Also, in the case of the lower performances of the Bayesian consensus trees when using the Triplets distance, one can note the high overlap of the confidence intervals (Fig.
-~\ref{Fig_Results-permeth_perparam} and Fig.
-~\ref{Fig_Results-paircomp_within}-B).
-% NC: Don't get the relevance of the last sentence/why it fits here.
-Can you expand?
-% NC: Probably want some  statement here about bayesian being better than ML, and if you want to fix a single topology in TE we recommend a bayesian consensus.
-If you want to account for tree uncertainty we recommend a bayesian posterior distribution.
-Our results dont suggest this is bad, but more suggest that we can't compare trees very well in this case.
-
-% $§4: Thus, baycon seems the best way to go >...
-% ^§5: >... but we have to take into account problems in our protocol.
-
-%\label{metrics_discussion}
-%This difference between the metrics can be explained by their respective ability to measure the differences between two trees \citep{kuhnerpractical2014}.
-As shown in Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}, the contrast between the results of the two metrics can be explained by the actual nature of both metrics (see \hyperref[SupplementaryMaterial]{Supplementary Material S2}):
-%\begin{enumerate}
-%\item{The Robinson-Foulds metric is a conservative metric, which is more sensible to single taxon displacement because it will count clades as similar only if they are composed of the same taxa with the same topologies \citep{RF1981}.
-When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
-This metric illustrates therefore the conservation of clades among the compared trees (i.e.
-when the normalised Robinson-Foulds distance is close to 1, most of the clades in the two trees are identical).}
-%\item{On the other hand, the Triplets method is measuring the position of each taxon towards two other reference taxa \citep{critchlowthe1996}.
-It will penalise only trees when taxa get removed furthest from their original position.
-This metric therefore illustrates the amount of wildcard taxa \citep{kearneyfragmentary2002} between the two compared trees (i.e.
-when the score is closed to 1, few wildcard taxa are present in the tree).}
-%\end{enumerate}
-%Therefore, in our study, both metrics are showing two different aspects of tree topology.
-For example, regarding the Bayesian consensus trees, completes clade are conserved most of the time (Fig.~\ref{Fig_Results-global_perparam}-top) but with an increasing amount of missing data, wildcard taxa get more and more displaced (Fig.~\ref{Fig_Results-global_perparam}-bottom).
-Because we are comparing the Baysian consensus tree to a single tree topology, the probability of recovering the exact position of a wildcard taxa than in the "best" tree is really low and even reducing with the amount of missing data.
-Regarding this fundamental differences, the Robinson-Foulds metric performs better in showing a distance from the true topology than the Triplets distance \citep{kuhnerpractical2014}.
-Therefore, the results for the Bayesian consensus trees can be preferred on the Maximum Likelihood trees because - when using the Robinson-Fould distance - they display a higher ability to recover the "best" tree's topology (Fig.
-~\ref{Fig_Results-permeth_perparam}).
-
 %Opening/further work part - NC(meeting): Be less defensive (i.e. nothing seemed to have gone wrong THIS time).
@@
 \end{enumerate}
-
-% $§6: Knowing the caveats, we can try to fix them >...
-% ^§7: >... by taking into account data realness (fossil templates and rate heterogeneity).
-
-%§7
-%Further directions - is that necessary though?
-% NC: Cut for now, might want to put these ideas in the appropriate places above instead.
-%   -Using real data
-%One way to address this issue 
-% NC: Which issue???
-%would be to run a similar analysis using real data.
-The amount of studies using Total Evidence data type matrices is steadily increasing \citep[e.g.][]{ronquista2012,slaterphylogenetic2013,beckancient2014} and one could analysis these total evidence matrices in the same framework as this study.
-However, the true topology (obtained from a matrix with no missing data) is unlikely to be known.
-A second way to address this issue, and especially to incorporate the variation of signal in morphological character would be to simulate morphological characters at different rates \citep{wrightbayesian2014} or to apply patterns of missing data directly obtained from the fossil record \citep{pattinsonphylogeny2014}.
-Nevertheless, our simulation already show clear results of the effect of missing data on topology in a Total Evidence framework.
-Also, our results show how recovering the "best" tree's topology can be improved by using Bayesian consensus trees in combination with a maximum amount of characters and a of living taxa with morphological data.
-%   -Looking at branch length?
-
-% $§7: But despite the caveats above, missing data is not a big deal >...
-% ^§8: >... We can use TEM for integrating fossils into trees.
-
-%§8
-%Back to the importance of TEM - How we can use fossils (Intro-§2)
-%   -Data is not a major problem
-%Despite the various caveats discussed above, we answered our main question on how does Total Evidence matrices perform in recovering topology with different methods and different amounts of missing data.
-% NC: For a concluding paragraph, don't start by reminding people of the caveats! End strongly and confidently
-% Also no need to restate the question - instead give the answer.
-%   -Good and promising method.
-%Our results shows that this method performance decreases mainly with a reduction in the overlap of morphological data.
-To address this problem, we propose to increase the data overlap by coding living species and to have a large morphological matrix.
-Also, our results shows disparity among methods and a clear superiority %NC: Not really
-%of the Bayesian consensus tree topology over the Maximum Likelihood one.
 

### 2015-01-07 Manuscript/TEM_draft.tex
Outlined and drafted discussion

@@
 \subsection{Effect of missing data for fossil taxa ($M_{F}$)}
-%When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this reduces the probability of morphological characters for the fossil taxa to overlap with the ones for the living taxa and leads to displacement of certain taxa.
+When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this reduces the probability of morphological characters for the fossil taxa to overlap with the ones for the living taxa and leads to displacement of certain taxa.
 However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
@@
 %Methods
-%Decrease in clade conservation is smaller than for param ML in the Bayesian consensus tree for the clade conservation but higher than param ML for the displacement of wildcard taxa.
+Decrease in clade conservation is smaller than for param ML in the Bayesian consensus tree for the clade conservation but higher than param ML for the displacement of wildcard taxa.
 This is due to the fact that the Bayesian consensus tree is a majority consensus rule and that when the fossils have less data, they kind of randomly branch all over the shop and therefore no majority consensus position (>50\% of the time) exists puting the fossil on a politomie at the base of a clade (clade conservation stays high but displacement of taxa increases).
@@
 \subsection{Effect of missing character ($M_{C}$)}
-%When the overall number of morphological characters in the matrix ($M_{C}$) decreases
-
-%Story for MC: MC is expected to have an effect two times bigger than ML or MF, e.g.
+When the overall number of morphological characters in the matrix ($M_{C}$) decreases
+Story for MC: MC is expected to have an effect two times bigger than ML or MF, e.g.
 removing 10\% of MF removes 5\% of data (because it works only on the fossil part) and removing 10\% of MC actually removes 10\% of that.
 %Method
-%However the effect of removing characters has an effect of topology of the same order of magnitude than for ML and MF (FigRes 1): For RF, Bayesian consensus is better than ML and for Tr, MaxLik is better than Baycon and have the same drop at 50\% as for MF (different than for ML).
-
-%In this scenario, the overall number of missing characters ($M_{C}$) affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
+However the effect of removing characters has an effect of topology of the same order of magnitude than for ML and MF (FigRes 1): For RF, Bayesian consensus is better than ML and for Tr, MaxLik is better than Baycon and have the same drop at 50\% as for MF (different than for ML).
+In this scenario, the overall number of missing characters ($M_{C}$) affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
 % NC: What's the relevance of that sentence?
-%However, the pairwise Bhattacharyya Coefficients show that the area with the lowest probability of distributional overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
+However, the pairwise Bhattacharyya Coefficients show that the area with the lowest probability of distributional overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
 %NC: I don't understand what you mean here
@@
 \subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
-%In our simulations, the overlap in data is greatly reduced when all three missing data parameters have a lot of missing data (e.g.
+In our simulations, the overlap in data is greatly reduced when all three missing data parameters have a lot of missing data (e.g.
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-%When combining the effect, we
+When combining the effect, we
 

### 2015-01-07 Manuscript/TEM_draft.tex
Edits BC explanation

--- 2015-01-07 Manuscript/TEM_draft.tex
@@
 
-\subsubsection{Summarizing the comparisons}
-Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the "best" tree's topology in a Total Evidence framework.
+\subsection{Testing the effects of the missing data parameters on topological recovery} % NC: I think this is really a separate section to tree comparison.
+Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the "best" tree topology in a Total Evidence framework.
 We also assessed the effect of our missing data parameters jointly with the effects of different tree inference and uncertainty methods (i.e.
@@
 
-%Introducing Bhattacharyya Coefficient
-To evaluate these effects we measured similarities among different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for all the analyses above using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya} - see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} for calculation details).
+We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
+The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} for calculation details).
+Note that this is  equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
+To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%), i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc.
+(see Figure % FIGURE
+for more details).
+This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
+We performed this procedure separately for each tree inference and uncertainty method.
+When two combinations of missing data parameters have a similar ability to recover the "best" tree topology the Bhattacharyya Coefficient will be close to one.
+Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
 
-%Explaining the parameter pairwise comparisons (ANOVA equivalent=within)
-For assessing the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the Normalised Robinson-Foulds and Normalised Triplets distance distributions, for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%).
-For example, $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc.
-This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
-We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient = 1) and in red when they don't (Bhattacharyya Coefficient = 0) - Fig.~\ref{Fig_Results-paircomp_within}).
-We repeated this procedure for all the "methods" (Maximum Likelihood, Bayesian consensus, Bootstraps and Bayesian posterior trees distribution).
+% NC: Maybe put this in the results?
+% We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient = 1) and in red when they don't (Bhattacharyya Coefficient = 0) - Fig.~\ref{Fig_Results-paircomp_within}).
 
-%Explaining the methods comparisons (ANOVA equivalent=among)
-Finally, for assessing the effect of the method on the ability to recover the "best" tree's topology, we compared the similarity between the methods.
-We calculated the Bhattacharyya Coefficient between each 125 normalised metrics scores (Robinson-Foulds and Triplets distance) for the 6 pair of methods (e.g.
-Maximum Likelihood\textit{versus} Bayesian consensus).
-We then reported the minimum, the 1st quartile, the median, the mean, the 3rd quartile and the maximum of the probabilities of overlap between the methods (i.e.
-the Bhattacharyya Coefficients) in table~\ref{Tab_Results-Difference_methods}.
-When two methods perform similarly in their ability to recover the "correct" tree topology (i.e.
-they display similar normalised metrics scores for each comparison between the "best" tree and the "missing data" trees), we expect the Bhattacharyya Coefficients to be close to 1.
-On the opposite, if the two methods differ, we expect the Bhattacharyya Coefficients to be close to 0.
-
-%Questionish §
-We use the Bhattacharyya Coefficient in a similar way as a two-sided t-test to detect significant differences between the distributions of the normalised distance between the "best" tree and the "missing data" trees (for both metrics).
+To assess the effect of the different tree inference and uncertainty methods (i.e.
+Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the "best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of tree inference and uncertainty methods i.e.
+Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc.
+(see Figure % FIGURE
+for more details).
+Note that this procedure pools results from across all missing data parameter combinations so it results in just six pairwise comparisons.
+When two tree inference or uncertainty methods have a similar ability to recover the "best" tree topology the Bhattacharyya Coefficient will be close to one.
+Conversely, if the two tree inference or uncertainty methods differ, the Bhattacharyya Coefficient will be close to zero.
 

### 2015-01-07 Manuscript/TEM_draft.tex
Edits to results

@@
 However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-%bootstrap and bayesian treesets are always worst bad
-In any combination of missing data parameters or topological distance metrics, the Maximum Likelihood bootstraps trees and the Bayesian posterior tree distribution performs substantially similarly (Median Bhattacharrya Coefficient between both tree inference methods using Normalised Robinson-Foulds distance or Normalised Triplets distance is respectively 0.85 and 0.96 - see Table ~\ref{Tab_Results-Difference_methods}) and worth than the Maximum Likelihood trees and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} - see discussion %add discussion link
- and also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the effect of the combined parameters on the Maximum Likelihood bootstraps trees and the Bayesian posterior tree distribution).
 
@@
  trees than for the Bayesian consensus trees %(minimum Normalised Robinson-Foulds distance for $M_{L}$, $M_{F}$ and $M_{C}$ of respectively @, @ and @)  %minimum Tr value for each parameter individually% TG: or useless (just Bhattacharrya)?
-(Bhattacharrya Coefficient between the methods for $M_{L}$, $M_{F}$ and $M_{C}$ of respectively @, @ and @  %BC value for each parameter individually between both methods RF
-- Fig.~\ref{Fig_Results-permeth_perparam} upper and see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
+(median Bhattacharrya Coefficient = @, @, @ $M_{L}$, $M_{F}$ and $M_{C}$ respectively %BC value for each parameter individually between both methods RF
+; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 %Differences between MaxLik and Baycon in Tr
-However, the Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for the $M_{L}$ parameter (Bhattacharrya Coefficient between both tree inference methods for $M_{L}$ = @ %BC value for ML between both methods
-- Fig.~\ref{Fig_Results-permeth_perparam} lower and see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values); for the two other parameters $M_{F}$ and $M_{C}$, the Normalised Triplets distance is substantially higher for the Maximum Likelihood trees than for the Bayesian consensus trees (Bhattacharrya Coefficient between both tree inference methods for $M_{F}$ and $M_{C}$ of respectively @ and @  %BC value for each parameter individually between both methods Tr
+The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for the $M_{L}$ parameter (Bhattacharrya Coefficient between both tree inference methods for $M_{L}$ = @ %BC value for ML between both methods
+- Fig.~\ref{Fig_Results-permeth_perparam} lower and see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values), however, for the two other parameters, $M_{F}$ and $M_{C}$, the Normalised Triplets distance is substantially higher for the Maximum Likelihood trees than for the Bayesian consensus trees (Bhattacharrya Coefficient between both tree inference methods for $M_{F}$ and $M_{C}$ of respectively @ and @  %BC value for each parameter individually between both methods Tr
 - Fig.~\ref{Fig_Results-permeth_perparam} lower and see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
@@
 %Part 3
-\subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
+\subsection{Combined effect of missing data parameters}
 %Combined, increase in missing data leads to decrease in topological recovery as well
-The combined effect of the missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$) reduces the ability to recover the "best" tree's topology when each parameters contains a maximum of missing data (i.e.
-$M_{L}$=75\%, $M_{F}$=75\% and $M_{C}$=75\% - Fig.~\ref{Fig_Results-global_perparam} and see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the combined effects of the parameters in both Normalised Robinson-Foulds and Normalised Triplets distance for the Maximum Likelihood Bootstraps trees and the Bayesian posterior trees distribution).
+As expected, our ability to recover the "best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam} and see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+
+Figure @ shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (green when the distributions overlap completely, Bhattacharyya Coefficient = 1, and red when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
+
 %Differences between MaxLik and Baycon in RF
-As described above, The Normalised Robinson-Foulds distance is higher for the Bayesian consensus trees topologies than for the Maximum Likelihood trees topologies with a median Normalised Robinson-Foulds distance for the Bayesian consensus trees and for the Maximum likelihood trees of respectively @ and @ % Median RF value for Baycon and MaxLik
-(Fig.~\ref{Fig_Results-global_perparam} upper - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
+%As described above, The Normalised Robinson-Foulds distance is higher for the Bayesian consensus trees topologies than for the Maximum Likelihood trees topologies with a median Normalised Robinson-Foulds distance for the Bayesian consensus trees and for the Maximum likelihood trees of respectively @ and @ % Median RF value for Baycon and MaxLik
+%(Fig.~\ref{Fig_Results-global_perparam} upper - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
 %Differences between MaxLik and Baycon in Tr
-The Normalised Triplets distance is lower for the Bayesian consensus trees topologies than for the Maximum Likelihood trees topologies with a median Normalised Triplets distance for the Bayesian consensus trees and for the Maximum likelihood trees of respectively @ and @ % Median Tr value for Baycon and MaxLik
-(Fig.~\ref{Fig_Results-global_perparam} lower - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
+%The Normalised Triplets distance is lower for the Bayesian consensus trees topologies than for the Maximum Likelihood trees topologies with a median% Normalised Triplets distance for the Bayesian consensus trees and for the Maximum likelihood trees of respectively @ and @ % Median Tr value for Baycon and MaxLik
+%(Fig.~\ref{Fig_Results-global_perparam} lower - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
 %Significance
-However, the Normalised Robinson-Foulds distances are substantially less similar between the Bayesian Consensus trees and the Maximum Likelihood trees than the Normalised Triplets distances (Median Bhattacharrya Coefficient between both tree inference methods for the Normalised Robinson-Foulds distance and the Normalised Triplets distance is respectively 0.10 and 0.61 - Fig.~\ref{Fig_Results-global_perparam} and table~\ref{Tab_Results-Difference_methods}).
+%However, the Normalised Robinson-Foulds distances are substantially less similar between the Bayesian Consensus trees and the Maximum Likelihood trees than the Normalised Triplets distances (Median Bhattacharrya Coefficient between both tree inference methods for the Normalised Robinson-Foulds distance and the Normalised Triplets distance is respectively 0.10 and 0.61 - Fig.~\ref{Fig_Results-global_perparam} and table~\ref{Tab_Results-Difference_methods}).
 
 %And now the combined effect of parameters in RF for the Baycon trees
-Regarding the effect of the combination of our missing data parameters on the ability to recover the "best" tree's topology, using both Normalised Robinson-Foulds and Normalised Triplets distances from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameters combination (red %change color
- line at the bottom of Fig.~\ref{Fig_Results-paircomp_within} A and B).
-However, the Normalised Robinson-Foulds distance displays more dissimilarities than the Normalised Triplets distance (the median Bhattacharrya coefficient among all pairwise missing data parameters combinations from the Bayesian consensus trees using the Normalised Robinson-Foulds and the Normalised Triplets distance are respectively @ and @ %value for median BC from pairwise comp in Baycon-RF and Baycon-Tr
- Fig.~\ref{Fig_Results-paircomp_within} - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
-Additionally, when using the Normalised Robinson-Foulds distance, there is no dissimilarity between the trees with 50\% and 75\% of missing living taxa ($M_{L}$) and any other combination of parameter state for the two other missing data parameters $M_{F}$ and $M_{C}$ (e.g..
-$M_{L}$ = 50\%, $M_{F}$ = 0\%, $M_{C}$ = 10\%; or $M_{L}$ = 50\%, $M_{F}$ = 75\%, $M_{C}$ = 50\%; etc) as well as for the trees with 50\% and 75\% of missing characters ($M_{C}$) and any other combination of parameter state for the two other missing data parameters $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-global_perparam} upper and Fig.~\ref{Fig_Results-paircomp_within}-A).
+
+Using both Normalised Robinson-Foulds and Normalised Triplets distances from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Fig.~\ref{Fig_Results-paircomp_within}).
+However, the Normalised Robinson-Foulds distance (median Bhattacharrya coefficient = ; Fig @ A) displays more dissimilarities than the Normalised Triplets distance (median Bhattacharrya coefficient = ; Fig @ B).
+%value for median BC from pairwise comp in Baycon-RF and Baycon-Tr
+ Additionally, when using the Normalised Robinson-Foulds distance, once ML >= 50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure @ A).
+Likewise,  once MC >= 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Figure @ A)
+
+For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.96, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; see Table ~\ref{Tab_Results-Difference_methods}).
+However, these two methods perform worse than both the Maximum Likelihood and Bayesian consensus trees (median Bhattacharrya Coefficient = @ and @, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; see Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 

### 2015-01-07 Manuscript/TEM_draft.tex
Fixed brackets in results + removed extra comments

@@
 
-%Bleuah! Changing everything to stick to 5 parts as
-%Part 1: missing data decrease topology
-%Part 2: effect of ML
-%   -As missing data increase, topology decreases
-%   -This is different for the metrics
-%   -This is worst with that method
-%       (e.g. Distribution of RF values for the ML difference % missing data a substantial overlap with bla (BC=88, Fig bla))
-%Part 3: effect of MF 
-%Part 4: effect of MC
-%Part 5: combined effect
-
-%Part 1
-%increase in missing data leads to decrease in topological recovery no matter what
 As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the "best" tree's topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
@@
 
-%Part 2
 \subsection{Individual effects of missing data parameters}
-%Individually, increase in missing data leads to decrease in topological recovery as well
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree's topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-%Differences between MaxLik and Baycon in RF
 The Normalised Robinson-Foulds distance is always substantially lower for the Maximum Likelihood %(minimum Normalised Robinson-Foulds distance for $M_{L}$, $M_{F}$ and $M_{C}$ of respectively @, @ and @) %minimum RF value for each parameter individually% TG: or useless (just Bhattacharrya)?
  trees than for the Bayesian consensus trees %(minimum Normalised Robinson-Foulds distance for $M_{L}$, $M_{F}$ and $M_{C}$ of respectively @, @ and @)  %minimum Tr value for each parameter individually% TG: or useless (just Bhattacharrya)?
-(median Bhattacharrya Coefficient = @, @, @ $M_{L}$, $M_{F}$ and $M_{C}$ respectively %BC value for each parameter individually between both methods RF
-; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-%Differences between MaxLik and Baycon in Tr
-The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for the $M_{L}$ parameter (Bhattacharrya Coefficient between both tree inference methods for $M_{L}$ = @ %BC value for ML between both methods
-- Fig.~\ref{Fig_Results-permeth_perparam} lower and see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values), however, for the two other parameters, $M_{F}$ and $M_{C}$, the Normalised Triplets distance is substantially higher for the Maximum Likelihood trees than for the Bayesian consensus trees (Bhattacharrya Coefficient between both tree inference methods for $M_{F}$ and $M_{C}$ of respectively @ and @  %BC value for each parameter individually between both methods Tr
-- Fig.~\ref{Fig_Results-permeth_perparam} lower and see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
+(median Bhattacharrya Coefficient = @, @, @ %BC value for each parameter individually between both methods RF
+$M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for the $M_{L}$ parameter (median Bhattacharrya Coefficient = @ %BC value for ML between both methods
+; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}), however, for the two other parameters, $M_{F}$ and $M_{C}$, the Normalised Triplets distance is substantially higher for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = @, @  %BC value for each parameter individually between both methods Tr
+$M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 
-%Part 3
 \subsection{Combined effect of missing data parameters}
-%Combined, increase in missing data leads to decrease in topological recovery as well
 As expected, our ability to recover the "best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam} and see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-
-Figure @ shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (green when the distributions overlap completely, Bhattacharyya Coefficient = 1, and red when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
-
-%Differences between MaxLik and Baycon in RF
-%As described above, The Normalised Robinson-Foulds distance is higher for the Bayesian consensus trees topologies than for the Maximum Likelihood trees topologies with a median Normalised Robinson-Foulds distance for the Bayesian consensus trees and for the Maximum likelihood trees of respectively @ and @ % Median RF value for Baycon and MaxLik
-%(Fig.~\ref{Fig_Results-global_perparam} upper - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
-%Differences between MaxLik and Baycon in Tr
-%The Normalised Triplets distance is lower for the Bayesian consensus trees topologies than for the Maximum Likelihood trees topologies with a median% Normalised Triplets distance for the Bayesian consensus trees and for the Maximum likelihood trees of respectively @ and @ % Median Tr value for Baycon and MaxLik
-%(Fig.~\ref{Fig_Results-global_perparam} lower - see also \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for the table of values).
-%Significance
-%However, the Normalised Robinson-Foulds distances are substantially less similar between the Bayesian Consensus trees and the Maximum Likelihood trees than the Normalised Triplets distances (Median Bhattacharrya Coefficient between both tree inference methods for the Normalised Robinson-Foulds distance and the Normalised Triplets distance is respectively 0.10 and 0.61 - Fig.~\ref{Fig_Results-global_perparam} and table~\ref{Tab_Results-Difference_methods}).
-
-%And now the combined effect of parameters in RF for the Baycon trees
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (green when the distributions overlap completely, Bhattacharyya Coefficient = 1, and red when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
+%CHANGE COLORS
 
@@
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Fig.~\ref{Fig_Results-paircomp_within}).
-However, the Normalised Robinson-Foulds distance (median Bhattacharrya coefficient = ; Fig @ A) displays more dissimilarities than the Normalised Triplets distance (median Bhattacharrya coefficient = ; Fig @ B).
+However, the Normalised Robinson-Foulds distance (median Bhattacharrya coefficient = @ ; Fig.~\ref{Fig_Results-paircomp_within}-A) displays more dissimilarities than the Normalised Triplets distance (median Bhattacharrya coefficient = @; Fig.~\ref{Fig_Results-paircomp_within}-B).
 %value for median BC from pairwise comp in Baycon-RF and Baycon-Tr
- Additionally, when using the Normalised Robinson-Foulds distance, once ML >= 50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure @ A).
-Likewise,  once MC >= 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Figure @ A)
+ Additionally, when using the Normalised Robinson-Foulds distance, once $M_{L}$ $\geq$  50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Fig.~\ref{Fig_Results-paircomp_within}-A).
+Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-paircomp_within}-A).
 

### 2015-01-08 Manuscript/TEM_draft.tex
Discussion update

@@
 
-In this study, we focus on the effect of missing data on our ability to recover a "best" tree's topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{}.
+In this study, we focus on the effect of missing data on our ability to recover a "best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{}.
 %CITE-check Lloyd and Friedman or birdy stuff
@@
 Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
-We find that the ability of Total Evidence methods to recover the "best" tree's topology is increased when using Bayesian methods.
-Also minimizing the proportion of living taxa with no morphological data and the proportion of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree's topology more so than minimizing the proportion of missing data in the fossil record.
+We find that the ability of Total Evidence methods to recover the "best" tree topology is increased when using Bayesian methods.
+Also minimizing the proportion of living taxa with no morphological data and the proportion of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree topology more so than minimizing the proportion of missing data in the fossil record.
 
@@
 
-As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the "best" tree's topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
+As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the "best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
 However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 \subsection{Individual effects of missing data parameters}
-As the amount of missing data increases across all three parameters, our ability to recover the "best" tree's topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
+As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
 The Normalised Robinson-Foulds distance is always substantially lower for the Maximum Likelihood %(minimum Normalised Robinson-Foulds distance for $M_{L}$, $M_{F}$ and $M_{C}$ of respectively @, @ and @) %minimum RF value for each parameter individually% TG: or useless (just Bhattacharrya)?
@@
 Probably need to think some more about this.
-% TG: a note about the "best" tree's topology regarding your comment just after the discussion header: should I replace "correct" by "best" since that's how we're calling the "best" tree anyway (all in quotes)
+% TG: a note about the "best" tree topology regarding your comment just after the discussion header: should I replace "correct" by "best" since that's how we're calling the "best" tree anyway (all in quotes)
 
@@
 
-Our results show that the ability to recover the "best" tree's topology in a Total Evidence framework decreases with increasing amount of missing data, regardless of how data were removed or the method of tree inference used.
+Our results show that the ability to recover the "best" tree topology in a Total Evidence framework decreases with increasing amount of missing data, regardless of how data were removed or the method of tree inference used.
 However, these factors affected topological recovery in different ways and to different extents.
-Decreasing the overall number of morphological characters in the matrix ($M_{C}$) and the amount of living taxa with morphological data ($M_{L}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
-Additionally, using Bayesian consensus trees recovered the "best" tree's topology more consistently than using Maximum Likelihood trees or Bayesian posterior trees distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
-% ----->
-% NC: Probably needs another summing up line here. Or maybe not
-% <-----
-% NC: Make sure it's parameters then tree building methods.
-NC: Make sure the order of the parameters is always the same L,F, C or whatever!
-THROUGHOUT the paper.
-Also correct RF and T to standardised OR normalised RF and T.
-% NC: Might want to add a couple of subheadings to match those in the results.
-As proposed in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough morphological characters overlaps for most taxa in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
+Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
+Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).As proposed in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough morphological characters overlaps for most living and fossil taxa in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
 % NC: What do you mean by data overlap here?
@@
 
-\subsection{Effect of missing living taxa ($M_{L}$)} %Or remove? Also below then
-%What is the parameter variation doing in our simulation
+\subsection{Individual effects of missing data parameters}
+\subsubsection{Effect of missing data for living taxa}
 When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
-Because living taxa still have 10 times more molecular characters available, even if they have no morphological data, the relation between them will always be fairly well resolved (depending on the phylogenetic signal from the molecular part of the matrix) and thus, ignoring the fossils, the living taxa clades will be always well conserved.
+Because living taxa still have 10 times more molecular characters available (see methods) %link to methods
+, even if they have no morphological data, the relation between them will always be fairly well resolved (depending on the phylogenetic signal from the molecular part of the matrix) and thus, ignoring the fossils, the living taxa clades will be always well conserved.
 However, this missing data parameter influences a lot the placement of fossil taxa because a decrease of the $M_{L}$ parameter will reduce the overlapping data between the living and the fossil taxa.
 
-%Methods
-%Part saying the difference between ML and Baycon
-When using the Normalised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "best" tree's topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}).
-In fact, regardless of the amount of missing data, Bayesian consensus trees always displays the highest Normalised Robinson-Foulds score (always $>$ 0.69; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for details).
-However, regarding the Normalised Triplets distance, the topological similarity with the "best" tree's topology decreases in the same way for both the Maximum Likelihood trees and the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+When using the Normalised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "best" tree topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}).
+In fact, regardless of the amount of missing data, Bayesian consensus trees always displays the highest Normalised Robinson-Foulds score (always $>$ 0.69; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for all the values).
+However, regarding the Normalised Triplets distance, the similarity with the "best" tree topology decreases in the same way for both the Maximum Likelihood trees and the Bayesian consensus trees for this particular missing data parameter (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+% TG: Or is that line just repeating the results?
+However, regardless of the metric, the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution always performed poorlier in their ability to recover the "best" tree topology, even when no morphological data was missing ($M_{L}$ = 0\% - Fig.~\ref{Fig_Results-permeth_perparam}).
+This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" trees and a set of the "best" trees involved 1000 random pairwise comparisons rather than just one.
+This will obviously add noise to the results for this method of measuring the ability to recover the "best" tree topology.
 
-% ----->
-% TG: Put that part somewhere in the start with the discussion of ML vs Baycon in the effect of ML part?
-%-Probabilistic method always better (wright but see spencer) - Check McGuire[nope]/Sean[nope] literature for Bayesian vs. ML?
-Variation in our ability to recover the "best" tree's topology also depended heavily on the method used to infer the tree (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-Previous studies have shown some superiority of methods using simple probabilistic models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}).
-However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-%TG: I guess it's a good selling point: we ARE the first ones to do a sensitivity analysis in a TEM framework (as far as I know).
-% <-----
+Variation in our ability to recover the "best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}).
+However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} for recovering the "best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
+Our results shows that the topology of the Bayesian consensus tree is always closer to the "best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}).
 
-%Part saying the difference between BS and baytrees
-Regardless of the metric, the Maximum Likelihood boostrap trees and the Bayesian posterior trees distribution always performed poorlier in their ability to recover the "best" tree's topology, even when no morphological data was missing ($M_{L}$=0\% - Fig.~\ref{Fig_Results-permeth_perparam}).
-This reflects the fact it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" tree and "best" tree in these cases involved 1000 random pairwise comparisons rather than just one.
-This will obviously add noise to the results for these two methods.
+In the Bayesian consensus tree topology, the clade conservation (i.e.
+approximated by the Normalised Robinson-Foulds distance) stays always high but the placement of wildcard taxa (i.e.
+approximated by the Normalised Triplets distance) decreases way faster (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+The difference between the metrics is linked to the way each is calculated and to how they perform at underlining changes in topology (see method %add link
+, \hyperref[SupplementaryMaterial]{Supplementary Material S2} and \citealt{kuhnerpractical2014} for details on the nature of both metrics).
+In general, the Normalised Robinson-Foulds distance is more sensitive than the Normalised Triplets distance.
+%----> version one
+When getting closer to the root of the tree, the displacement of a single taxon will have a greater effect than it would in higher clades and therefore decrease the Normalised Robinson-Foulds distance.
+%----> version two
+%When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
+Also, \citet{kuhnerpractical2014} have demonstrated that the Robinson-Foulds distance outperforms the Triplets distance for measuring topological differences between trees.
+Thus, the Bayesian consensus tree topology seems to be the less affected by removing morphological data for living taxa (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 
-%Metrics
-%Part explaining the differences between the metrics
-This difference between the metrics is linked to what they measure and how they perform at doing it (see method %add link
-, \hyperref[SupplementaryMaterial]{Supplementary Material S2} and \citealt{kuhnerpractical2014} for details on the nature of both metrics).
-In the Bayesian consensus tree topology, the clade conservation (i.e.
-Noramlised Robinson-Foulds distance) stays always high but the placement of wildcard taxa (i.e.
-the Normalised Triplets distance) decreases way faster (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-This is due to the fact that both metrics are illustrating two different aspects of tree topology (i.e.
-clade conservation for the Normalised Robinson-Foulds distance and wildcard taxa position for the Normalised Triplets distance - see method %add link!
-and \hyperref[SupplementaryMaterial]{Supplementary Material S2} for details on both metrics).
-However, it is important to note that The Normalised Robinson-Foulds distance is more sensitive than the Normalised Triplets distance: when getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
-Also, \citet{kuhnerpractical2014} have recently demonstrated that the Robinson-Foulds distance outperforms the Triplets distance for measuring topological differences between trees.
-Thus, the Bayesian consensus tree's topology seems to be the less affected by removing morphological data for living taxa (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-
-\subsection{Effect of missing data for fossil taxa ($M_{F}$)}
-When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this reduces the probability of morphological characters for the fossil taxa to overlap with the ones for the living taxa and leads to displacement of certain taxa.
+\subsubsection{Effect of missing data for fossil taxa}
+When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this also reduces the probability of morphological characters for the fossil taxa to overlap with the ones for the living taxa.
+This can lead to difficulties for certain taxa to be branched in the tree (i.e.
+wildcard taxa behaviour).
 However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
 decrease of Normalised Triplets distance), the conservation of the clades (i.e.
-Normalised Robinson-Foulds distance) is still relatively good (mode=% value
-%) when the proportion of missing data is high ($M_{F}$=75\%).
+Normalised Robinson-Foulds distance) is still relatively good (mode = @% value
+) when the proportion of missing data is high ($M_{F}$ = 75\%).
 
-%Methods
-Decrease in clade conservation is smaller than for param ML in the Bayesian consensus tree for the clade conservation but higher than param ML for the displacement of wildcard taxa.
-This is due to the fact that the Bayesian consensus tree is a majority consensus rule and that when the fossils have less data, they kind of randomly branch all over the shop and therefore no majority consensus position (>50\% of the time) exists puting the fossil on a politomie at the base of a clade (clade conservation stays high but displacement of taxa increases).
-On the opposite in MaxLik, the topology is always dichotomous but the fossil can be missplaced in the wrong clade (with a low BS value) but is moving less.
-In other words, Bayesian consensus tree places the fossil with a higher confidence at a low taxonomic level (politomy) and the MaxLik places the fossil with a low confidence at a high taxomomic level (baldy supported dichotomies).
-We can argue that the first option (higher confidence at a low taxonomic level) is preferable (example the Darwinius debate).
+%----> version one
+The effect of the missing data in the fossil record ($M_{F}$) on the Normalised Robinson-Foulds distance (i.e.
+as a proxy for clade conservation) is smaller than the one of the $M_{L}$ parameter in the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+However, the effect of the $M_{F}$ parameter on the Normalised Triplets distance (i.e.
+as a proxy for displacement of wildcard taxa) is greater than the one of the $M_{L}$ parameter in the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+%----> version two
+%The effect of the missing data in the fossil record ($M_{F}$) is smaller than the one the $M_{L}$ parameter for the clade conservation (Normalised Robinson-Foulds distance) but greater for the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+This can be related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
+When a fossil taxa have less data (e.g.
+($M_{F}$ = 75\%) they will tend to branch "randomly" to any taxa in the clade sharing the most characters.
+Therefore a majority consensus position is unlikely to exist (i.e.
+every branching position is represented in $<$ 50\% of the trees) and the fossil taxa will be branched on a polytomy at the base of the clade.
+This does not decreases the Normalised Robinson-Foulds distance (clades is conserved) but increases the Normalised Triplets distance (wildcard taxa is displaced).
+On the opposite, the same scenario in a Maximum Likelihood will lead to a dichotomous branching of the fossil but with a low bootstrap support ($<$ 50).
+In other words, the Bayesian consensus tree allows a fossil taxa with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree (where the fossil will be placed with lower confidence at a higher taxonomic position).
+We can argue that the first option (the Bayesian consensus tree topology) is preferable \cite[e.g][]{}.
+%Cite the Darwinius debate or is that a bit dangerous?
 
-\subsection{Effect of missing character ($M_{C}$)}
-When the overall number of morphological characters in the matrix ($M_{C}$) decreases
-Story for MC: MC is expected to have an effect two times bigger than ML or MF, e.g.
-removing 10\% of MF removes 5\% of data (because it works only on the fossil part) and removing 10\% of MC actually removes 10\% of that.
-%Method
-However the effect of removing characters has an effect of topology of the same order of magnitude than for ML and MF (FigRes 1): For RF, Bayesian consensus is better than ML and for Tr, MaxLik is better than Baycon and have the same drop at 50\% as for MF (different than for ML).
-In this scenario, the overall number of missing characters ($M_{C}$) affects the proportion of overlapping data by reducing the overall size of the matrix (Fig.~\ref{Fig_Results-global_perparam}).
-% NC: What's the relevance of that sentence?
-However, the pairwise Bhattacharyya Coefficients show that the area with the lowest probability of distributional overlap is led by comparing the trees with no missing data for the living taxa to the trees with 75\% missing data for the living data (right lower corner in Fig.~\ref{Fig_Results-paircomp_within}).
-%NC: I don't understand what you mean here
+\subsubsection{Effect of missing character ($M_{C}$)}
+The overall number of missing morphological characters is expected to have an effect on our ability to recover the "best" tree topology two times bigger than for the $M_{L}$ and $M_{F}$ parameters.
+For example, removing 10\% of the data for the fossil taxa actually removes a total of 5\% of data in the whole matrix (because this parameter affects only half of the taxa present in the matrix); on the other hand, removing 10\% of morphological characters actually removes 10\% of data in the matrix.
+However the effect of removing characters on the ability to recover the "best" tree topology is of the same order of magnitude than for the two other parameters (Fig.~\ref{Fig_Results-permeth_perparam}).
+Reducing the overall amount of morphological characters reduces the probability of overlap of morphological characters between the taxa present in the matrix and therefore decreases our ability to recover the "best" tree topology.
 
-\subsection{Combined effect of missing data parameters ($M_{L}$, $M_{F}$ and $M_{C}$)} 
-In our simulations, the overlap in data is greatly reduced when all three missing data parameters have a lot of missing data (e.g.
-$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-When combining the effect, we
+%I moved the stuff about simulating characters over here.
+The overall number of morphological character in a Total Evidence matrix can also affect our ability to recover the "best" tree topology through two phylogenetic caveats from morphological data (in opposition to molecular data): (1) the incongruence of phylogenetic signal between morphological and molecular data and (2) homoplasy.
+The incongruence of phylogenetic signal between morphological and molecular data as previously been demonstrated as being potentially more important in small morphological matrices \citep{wagner2000}.
+Therefore, part of the decrease of the Normalised Robinson-Fould distance and the Normalised Triplets distance in our simulation could be due to conflicting phylogenetic signal between morphological and molecular data (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013,slaterphylogenetic2013,beckancient2014}.
+(2) Homoplasy, on the other hand is expected to increase with an increase of the morphological characters \citep{wagner2000}.
+%CITE
+However, the use of probabilistic methods (i.e.
+Maximum Likelihood or Bayesian) and the M\textit{k} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.
 
-%Our solutions
-Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
-Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "best" tree's topology.
+\subsection{Combined effect of missing data parameters}
+As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the effect of the missing data parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance is higher when all the missing data parameters have low missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have high missing data (i.e.
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-global_perparam}).
+However, it is important to notice that the effect of each parameter is not additive.
+The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological character ($M_{C}$) have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$); when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).
+
+\subsection{Suggested applications} %Or "Practical implications"
+Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
+Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "best" tree topology.
 We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections.
 Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries.
-However, gaps in the matrix can be filled with future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
-Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "best" tree's topology (Fig.~\ref{Fig_Results-paircomp_within}).
+However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
+% TG: I added the palaeo field work just as a kind work to the pokemon masters that could read this paper - maybe useless?
+Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "best" tree topology (Fig.~\ref{Fig_Results-paircomp_within}).
 Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
@@
 
-%Opening/further work part - NC(meeting): Be less defensive (i.e. nothing seemed to have gone wrong THIS time).
-%   -Simulating morpho characters
-The variations in our results could also be explained by another methodological caveat: 
-% NC: What variations do you mean? Only an issue if it might bias results in one direction or another
-the performance of our simulation protocol.
-In order to reduce the computational time of our analysis ($~$150 CPU years), we ran our sensitivity analysis on modestly sized matrices of 1000 molecular characters and 100 morphological characters.
-These matrices are two orders of magnitude smaller than some matrices used in published analyses \citep[e.g.][]{springermacroevolutionary2012,nithe2013} but still ranges in the sizes of smaller matrices used in other ones \citep[e.g.][]{sallam2011craniodental,kellymolecular2014}.
-%TG: note about NOT citing O'Leary here.
-1-I don't like the paper ; 2-the Ni et al paper (primates) as a big morpho matrix too (1k  characters against 4k for O'Leary, same order of magnitude?) ; 3-We might want to limit the papers we're citing?
-note about citing Sean's and your paper here: 1-is it okay (regarding last time warning of being extra careful when citing you're own work)?
+%From here I'm not quite sure how to put that and if it's actually important.
+% ---->
+Our simulation were constraint by the performance of our protocol: in order to reduce the computational time of our analysis (150 CPU years), we ran our sensitivity analysis on modestly sized matrices of 1000 molecular characters and 100 morphological characters.
+These matrices are two orders of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013} but still ranges in the sizes of more modest ones \citep[e.g.][]{sallam2011craniodental,kellymolecular2014}.
+Another point that was constrained by the performance of our protocol was the ability to assess the effect of missing data on branch length variation in a Total Evidence framework which is a crucial parameter in one of the wider applications of matrices combining both living and fossil taxa: the Tip-dating method \citep[e.g.][]{ronquista2012,BEASTmaster}.
+
 Also, simulating morphological characters is more complex than molecular ones.
@@
 Morphological characters studies involve many potential statistical pitfalls (e.g.
-independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy \citep{wagner2000}.
-\begin{enumerate}
-\item{First, morphological data can display a different signal than molecular data, especially in small matrices \citep{wagner2000}. %REF
-This might lead to a controversial phylogenetic signal in the overall matrix and lower down the support values.
-However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013}.}
-\item{Secondly, in this study, we made the assumption that theoretically, morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013,pattinsonphylogeny2014}.
-However, following our simulation assumption of random character distributions, if they accumulate through time in the same way as the majority of the molecular characters then, homoplasic characters are expected to appear randomly through time \citep{davalosintegrating2014}.
-Therefore, homoplasy is expected to be more important (by chance) in bigger morphological matrices \citep{davalosintegrating2014}.
-After reaching a critical amount of morphological characters, adding new ones increases homoplasy \citep{wagner2000}.}
-\end{enumerate}
+independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy that are discussed above.
+% add link
+Secondly, in this study, we made the assumption that morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013,pattinsonphylogeny2014}.
+Approaches such as artificial fossilisation \citep{pattinsonphylogeny2014} might solve this issue and
+% <----
 
@@
 
-%§Conclusions
-% part 1- Fossils are important and can be used combined with living taxa (Intro-§1 & §2)
-%   -TEM works with missing data
-% part 2- This is how we can improve the data to use them in the best way possible (Conclusion-§1 & §8)
-%   -Use bayesian consensus
-%   -Go code living taxa
-Using Total Evidence matrices with missing data is not a problem for recovering the "best" tree's topology as long as enough morphological data is overlapping between the fossil and living taxa.
-%NC: Explain more clearly what you mean by overlap - of characters?
-Topological recovery is still best when there is less missing data, therefore we advise filling as many gaps in Total Evidence matrices as possible.
-Because this is difficult, if not impossible, for fossil data, we recommend coding as many
-
-using as many morphological characters as possible and to code them for a maximum of living taxa present in the matrix.
-
-we propose to use the Bayesian consensus tree over the Maximum Likelihood tree or over a random subset of the Bayesian posterior trees distribution.
-Also, in order to improve the overlap in data in the morphological part of the matrix,
-
-% NC: Never write "in order to". Just "to" will do
+Using Total Evidence matrices with missing data is not a problem for recovering the "best" tree topology as long as enough morphological characters are overlapping between the fossil and living taxa.
+When missing data increases in any of our missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
+Therefore we advise filling as many gaps in Total Evidence matrices as possible.
+Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters as possible and to code them for a maximum of living taxa present in the matrix.
+Additionally, the topology of the Bayesian consensus tree, regardless the amount of missing data, was always the closer to the "best" tree topology than the Maximum likelihood trees.
+Therefore, we advise to use the topology of the Bayesian consensus tree over the topology of the Maximum Likelihood tree or of a random subset of the Bayesian posterior trees distribution
 
@@
 \section{Acknowledgements}
-Thanks to Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Andrew Jackson, Gavin Thomas and April Wright.
-%TG : Not sure how many people we have to thank but it could also include Graham Slater, Nick Matzke and Trevor Hodkinson.
+Thanks to Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Andrew Jackson, Gavin Thomas and April Wright %TG : Not sure how many people we have to thank but it could also include Graham Slater, Nick Matzke and Trevor Hodkinson.
  % NC: Might be nice to thank Trevor, and it won't add too many lines to thank them all. You can also just calle people A.Wright if you want to reduce word count.

### 2015-01-08 Manuscript/TEM_draft.tex

@@
 
-% NC: Quick comment.
-You find that Bayesian consensus trees are "best".
-BUT I think you need to be careful here with what you mean by "best".
-They are most likely to recover a more similar topology to the best tree than if you do the crazy tree comparison pairs.
-But that doesn't mean they're best, just that they perform best in your simulations.
-They are only best for that.
-If you want to advise people to use a method that best recovers topology as measured using RF or Triplets, then Bayesian consensus is best.
-BUT generally that is not what people want.
-Probably need to think some more about this.
-% TG: a note about the "best" tree topology regarding your comment just after the discussion header: should I replace "correct" by "best" since that's how we're calling the "best" tree anyway (all in quotes)
-
 \section{Discussion}
@@
 Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
-Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).As proposed in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough morphological characters overlaps for most living and fossil taxa in the matrix \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
-% NC: What do you mean by data overlap here?
-That species have overlapping characters?
+Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
+As shown in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
 
 \subsection{Individual effects of missing data parameters}
-\subsubsection{Effect of missing data for living taxa}
+\subsubsection{Missing data for living taxa ($M_{L}$)}
 When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
-Because living taxa still have 10 times more molecular characters available (see methods) %link to methods
-, even if they have no morphological data, the relation between them will always be fairly well resolved (depending on the phylogenetic signal from the molecular part of the matrix) and thus, ignoring the fossils, the living taxa clades will be always well conserved.
-However, this missing data parameter influences a lot the placement of fossil taxa because a decrease of the $M_{L}$ parameter will reduce the overlapping data between the living and the fossil taxa.
+Because living taxa still have molecular characters available for phylogenetic inference (see methods), %link to methods
+even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
+However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
 
-When using the Normalised Robinson-Foulds distance to measure topological similarity, the Bayesian consensus trees recover the "best" tree topology more often than Maximum Likelihood trees (Fig.~\ref{Fig_Results-global_perparam}).
-In fact, regardless of the amount of missing data, Bayesian consensus trees always displays the highest Normalised Robinson-Foulds score (always $>$ 0.69; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} for all the values).
-However, regarding the Normalised Triplets distance, the similarity with the "best" tree topology decreases in the same way for both the Maximum Likelihood trees and the Bayesian consensus trees for this particular missing data parameter (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-% TG: Or is that line just repeating the results?
-However, regardless of the metric, the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution always performed poorlier in their ability to recover the "best" tree topology, even when no morphological data was missing ($M_{L}$ = 0\% - Fig.~\ref{Fig_Results-permeth_perparam}).
-This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" trees and a set of the "best" trees involved 1000 random pairwise comparisons rather than just one.
-This will obviously add noise to the results for this method of measuring the ability to recover the "best" tree topology.
+\subsubsection{Missing data for fossil taxa ($M_{F}$)}
+When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this also reduces the probability of morphological characters for fossil taxa overlapping with the ones for living taxa.
+This can lead to difficulties for the placement of certain taxa in the tree.
+However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
+decrease of Normalised Triplets distance) with increasing missing data in this parameter, clade conservation (i.e.
+Normalised Robinson-Foulds distance) is still relatively good (mode = @) when the proportion of missing data is high ($M_{F}$ = 75\%).
 
-Variation in our ability to recover the "best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}).
-However, this is the first study, to our knowledge, that compares the performance of the M\textit{k} model \citep{lewisa2001} for recovering the "best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-Our results shows that the topology of the Bayesian consensus tree is always closer to the "best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}).
+The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds distance) but greater on the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
+When the fossil taxa have less data (e.g.
+$M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils.
+Therefore a majority consensus position is unlikely to exist (i.e.
+every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
+This does not decrease the Normalised Robinson-Foulds distance, because clades are conserved, but decreases the Normalised Triplets distance because the fossils act as wildcard taxa.
+Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50).
+In other words, the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree, where the fossil will be placed with lower confidence at a higher taxonomic level.
+We argue that using the Bayesian consensus tree topology is preferable because it is more conservative \cite[e.g][]{}.
+%Cite the Darwinius debate or is that a bit dangerous?
+% NC: Check pattinson
 
-In the Bayesian consensus tree topology, the clade conservation (i.e.
-approximated by the Normalised Robinson-Foulds distance) stays always high but the placement of wildcard taxa (i.e.
-approximated by the Normalised Triplets distance) decreases way faster (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-The difference between the metrics is linked to the way each is calculated and to how they perform at underlining changes in topology (see method %add link
-, \hyperref[SupplementaryMaterial]{Supplementary Material S2} and \citealt{kuhnerpractical2014} for details on the nature of both metrics).
-In general, the Normalised Robinson-Foulds distance is more sensitive than the Normalised Triplets distance.
-%----> version one
-When getting closer to the root of the tree, the displacement of a single taxon will have a greater effect than it would in higher clades and therefore decrease the Normalised Robinson-Foulds distance.
-%----> version two
-%When getting closer to the root of the tree, displacement of single taxon makes the clades not being exactly identical any more even if the clade still contains all the other taxa in both trees.
-Also, \citet{kuhnerpractical2014} have demonstrated that the Robinson-Foulds distance outperforms the Triplets distance for measuring topological differences between trees.
-Thus, the Bayesian consensus tree topology seems to be the less affected by removing morphological data for living taxa (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+\subsubsection{Missing morphological characters ($M_{C}$)}
+Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the "best" tree topology.
+We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
+Conversely, removing 10\% of morphological characters ($M_{C}$) genuinely removes 10\% of data in the matrix.
+However, the effect of removing characters on the ability to recover the "best" tree topology is of the same order of magnitude as for the other two parameters (Fig.~\ref{Fig_Results-permeth_perparam}).
+We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se} % see above or below???.
 
-\subsubsection{Effect of missing data for fossil taxa}
-When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this also reduces the probability of morphological characters for the fossil taxa to overlap with the ones for the living taxa.
-This can lead to difficulties for certain taxa to be branched in the tree (i.e.
-wildcard taxa behaviour).
-However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
-decrease of Normalised Triplets distance), the conservation of the clades (i.e.
-Normalised Robinson-Foulds distance) is still relatively good (mode = @% value
-) when the proportion of missing data is high ($M_{F}$ = 75\%).
-
-%----> version one
-The effect of the missing data in the fossil record ($M_{F}$) on the Normalised Robinson-Foulds distance (i.e.
-as a proxy for clade conservation) is smaller than the one of the $M_{L}$ parameter in the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-However, the effect of the $M_{F}$ parameter on the Normalised Triplets distance (i.e.
-as a proxy for displacement of wildcard taxa) is greater than the one of the $M_{L}$ parameter in the Bayesian consensus trees (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-%----> version two
-%The effect of the missing data in the fossil record ($M_{F}$) is smaller than the one the $M_{L}$ parameter for the clade conservation (Normalised Robinson-Foulds distance) but greater for the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-This can be related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
-When a fossil taxa have less data (e.g.
-($M_{F}$ = 75\%) they will tend to branch "randomly" to any taxa in the clade sharing the most characters.
-Therefore a majority consensus position is unlikely to exist (i.e.
-every branching position is represented in $<$ 50\% of the trees) and the fossil taxa will be branched on a polytomy at the base of the clade.
-This does not decreases the Normalised Robinson-Foulds distance (clades is conserved) but increases the Normalised Triplets distance (wildcard taxa is displaced).
-On the opposite, the same scenario in a Maximum Likelihood will lead to a dichotomous branching of the fossil but with a low bootstrap support ($<$ 50).
-In other words, the Bayesian consensus tree allows a fossil taxa with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree (where the fossil will be placed with lower confidence at a higher taxonomic position).
-We can argue that the first option (the Bayesian consensus tree topology) is preferable \cite[e.g][]{}.
-%Cite the Darwinius debate or is that a bit dangerous?
-
-\subsubsection{Effect of missing character ($M_{C}$)}
-The overall number of missing morphological characters is expected to have an effect on our ability to recover the "best" tree topology two times bigger than for the $M_{L}$ and $M_{F}$ parameters.
-For example, removing 10\% of the data for the fossil taxa actually removes a total of 5\% of data in the whole matrix (because this parameter affects only half of the taxa present in the matrix); on the other hand, removing 10\% of morphological characters actually removes 10\% of data in the matrix.
-However the effect of removing characters on the ability to recover the "best" tree topology is of the same order of magnitude than for the two other parameters (Fig.~\ref{Fig_Results-permeth_perparam}).
-Reducing the overall amount of morphological characters reduces the probability of overlap of morphological characters between the taxa present in the matrix and therefore decreases our ability to recover the "best" tree topology.
-
-%I moved the stuff about simulating characters over here.
-The overall number of morphological character in a Total Evidence matrix can also affect our ability to recover the "best" tree topology through two phylogenetic caveats from morphological data (in opposition to molecular data): (1) the incongruence of phylogenetic signal between morphological and molecular data and (2) homoplasy.
-The incongruence of phylogenetic signal between morphological and molecular data as previously been demonstrated as being potentially more important in small morphological matrices \citep{wagner2000}.
-Therefore, part of the decrease of the Normalised Robinson-Fould distance and the Normalised Triplets distance in our simulation could be due to conflicting phylogenetic signal between morphological and molecular data (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-However, regarding empirical data studies, most of the groups shows fairly congruent morphological and molecular phylogenetic signal \citep[e.g.][]{leerates2013,slaterphylogenetic2013,beckancient2014}.
-(2) Homoplasy, on the other hand is expected to increase with an increase of the morphological characters \citep{wagner2000}.
-%CITE
+Additionally, the number of morphological characters determines the size the matrix.
+This can affect our ability to recover the "best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
+The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \citep{wagner2000}.
+The size of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
+Therefore, part of the decrease of the Normalised Robinson-Fould distance and the Normalised Triplets distance in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+However, although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
+Therefore, our simulations reflect realistic parameters.
+Homoplasy, on the other hand, is expected to increase with an increase of the morphological characters \citep{}.
 However, the use of probabilistic methods (i.e.
@@
 \subsection{Combined effect of missing data parameters}
-As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the effect of the missing data parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance is higher when all the missing data parameters have low missing data (i.e.
+As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have low missing data (i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have high missing data (i.e.
@@
 However, it is important to notice that the effect of each parameter is not additive.
-The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological character ($M_{C}$) have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$); when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).
+The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$); when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).
+This has important implications for building Total Evidence phylogenies (see below).
 
-\subsection{Suggested applications} %Or "Practical implications"
+\subsection{Effects of tree inference methods}
+Variation in our ability to recover the "best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}).
+However, this is the first study, to our knowledge, to compare the performance of the M\textit{k} model \citep{lewisa2001} for recovering the "best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
+Our results show that the topology of the Bayesian consensus tree is always closer to the "best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}).
+As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
+
+It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the "best" tree topology than Maximum Likelihood and Bayesian consensus trees.
+This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-permeth_perparam}).
+This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" trees and a set of the "best" trees involved 1000 random pairwise comparisons rather than just one.
+This will obviously add noise to the results for these methods.
+
+\subsection{Practical implications}
 Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
@@
 However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
-% TG: I added the palaeo field work just as a kind work to the pokemon masters that could read this paper - maybe useless?
 Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "best" tree topology (Fig.~\ref{Fig_Results-paircomp_within}).
@@
 
-%From here I'm not quite sure how to put that and if it's actually important.
-% ---->
-Our simulation were constraint by the performance of our protocol: in order to reduce the computational time of our analysis (150 CPU years), we ran our sensitivity analysis on modestly sized matrices of 1000 molecular characters and 100 morphological characters.
-These matrices are two orders of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013} but still ranges in the sizes of more modest ones \citep[e.g.][]{sallam2011craniodental,kellymolecular2014}.
-Another point that was constrained by the performance of our protocol was the ability to assess the effect of missing data on branch length variation in a Total Evidence framework which is a crucial parameter in one of the wider applications of matrices combining both living and fossil taxa: the Tip-dating method \citep[e.g.][]{ronquista2012,BEASTmaster}.
-
-Also, simulating morphological characters is more complex than molecular ones.
-The underlying pattern of their evolution is often more complex and ruled by more parameters than molecular characters (i.e the number of character states, the states frequencies, the substitution matrix and the statistical model used) \citep{Pagel22011994,wagner2000,lewisa2001,wrightbayesian2014}.
-Morphological characters studies involve many potential statistical pitfalls (e.g.
-independent characters violation, rate variation - \citet{davalosintegrating2014}) and especially (i) incongruence with molecular signal and (ii) homoplasy that are discussed above.
-% add link
-Secondly, in this study, we made the assumption that morphological characters are randomly distributed on an organism however it seems clear that empirical morphological data does not act randomly \citep{sansomfossilization2013,pattinsonphylogeny2014}.
-Approaches such as artificial fossilisation \citep{pattinsonphylogeny2014} might solve this issue and
-% <----
+% Another practical implication of our results regards the tree building methods.
+Bayesian = better if chooisng one tree as backbone.
+However, although posterior distributions perform poorly this is more an effect of the way we compare trees (see above).
+Thus we do not suggest discarding posterior distribution.
+They are useful for comparative analyses and stuff.
 
@@
 
-Using Total Evidence matrices with missing data is not a problem for recovering the "best" tree topology as long as enough morphological characters are overlapping between the fossil and living taxa.
-When missing data increases in any of our missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
+Using Total Evidence matrices with missing data is not a problem for recovering the "best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
+When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
 Therefore we advise filling as many gaps in Total Evidence matrices as possible.
-Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters as possible and to code them for a maximum of living taxa present in the matrix.
+Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters as possible and to code them for as many living taxa present in the matrix as possible.
 Additionally, the topology of the Bayesian consensus tree, regardless the amount of missing data, was always the closer to the "best" tree topology than the Maximum likelihood trees.
-Therefore, we advise to use the topology of the Bayesian consensus tree over the topology of the Maximum Likelihood tree or of a random subset of the Bayesian posterior trees distribution
+Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed in a Total Evidence analysis, for example when using the Tip-Dating method.
+% NC: Maybe one more woolly sentence.
 

### 2015-01-08 Manuscript/TEM_draft.tex
Removed extra comments

@@
 The two factors that influences the most this data overlap are the parameters 1 (proportion of coded living taxa) and 3 (number of morphological characters).
-
-%Probably needs to be deeeply rewritten, as said last time to Kev quoting Stuart Pim, people only read the abstract.
-% NC: Yep, for now I'll leave this
-
 \end{abstract}
@@
 Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
-
-%TG: I completely removed this paragraph
-%The effect of missing data on phylogenetic inferences has been widely studied \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011,sansomfossilization2013,pattinsonphylogeny2014,sansombias2014}.
-Missing molecular data has been seen by some authors as an issue because it can, in some parts of the tree, decrease phylogenetic signal (i.e.
-the evolutionary information contained within the matrix allowing the inference of topology and branch lengths), especially when using large matrices \citep{lemmonthe2009}.
-However, this may not be a major issue because phylogenetic signal is easily increased by: (i) including a "modest" number (\texttildelow % NC: This symbol doesn't show up on my PDF.
-Check.
-%50) of highly-covered genes (i.e.
-a number of genes that are available for all taxa; \citealt{rouresite-specific2011}); (ii) adding a greater number of taxa (especially slowly-evolving taxa or taxa close to the outgroup; \citealt{rouresite-specific2011}); and (iii) choosing more appropriate models of sequence evolution \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
-Similarly, missing morphological data might be seen as either a major or minor issue for accurately inferring phylogenies depending on the study in question \citep{wiensmissing2003,sansomfossilization2013,pattinsonphylogeny2014}.
-Because soft-tissue characters are rarely preserved in the fossil record, missing data is mainly found in these characters, and is therefore not randomly distributed which can lead to biased placement of fossil taxa in phylogenies (e.g.
-\citealt{sansomfossilization2013}, but see \citealt{pattinsonphylogeny2014}).
-However, the phylogenetic signal is not related to the amount of missing data \textit{per se} but to the number of informative characters for each taxon, therefore missing data is less of an issue than the number of shared informative characters \citep{kearneyfragmentary2002,wiensmissing2003,pattinsonphylogeny2014}.
-% NC: I'm a little dubious of what this paragraph is trying to get at in places.
-We should discuss at some point.
-TG: my idea was to do the "literature-review" part of the intro: missing data is already widely studied (answer to the comment: "hun...
-We already knew that") but it has never been studied in TEM framework.
-And that's kind of the point we're getting toward with our work: we don't really care about the effect of missing data per se (i.e.
-what's the threshold where it starts fucking up the phylogeny) but more the fact that Bayesian fails.
-% NC: OK fair enough. I think though if this is too long we can probably cut this paragraph and put a bit of it into the paragraph below.
 
@@
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always substantially lower for the Maximum Likelihood %(minimum Normalised Robinson-Foulds distance for $M_{L}$, $M_{F}$ and $M_{C}$ of respectively @, @ and @) %minimum RF value for each parameter individually% TG: or useless (just Bhattacharrya)?
- trees than for the Bayesian consensus trees %(minimum Normalised Robinson-Foulds distance for $M_{L}$, $M_{F}$ and $M_{C}$ of respectively @, @ and @)  %minimum Tr value for each parameter individually% TG: or useless (just Bhattacharrya)?
-(median Bhattacharrya Coefficient = @, @, @ %BC value for each parameter individually between both methods RF
-$M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+The Normalised Robinson-Foulds distance is always substantially lower for the Maximum Likelihood  trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = @, @ and @ %BC value for each parameter individually between both methods RF
+for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for the $M_{L}$ parameter (median Bhattacharrya Coefficient = @ %BC value for ML between both methods
-; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}), however, for the two other parameters, $M_{F}$ and $M_{C}$, the Normalised Triplets distance is substantially higher for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = @, @  %BC value for each parameter individually between both methods Tr
-$M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}), however, for the two other parameters, $M_{F}$ and $M_{C}$, the Normalised Triplets distance is substantially higher for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = @ and @  %BC value for each parameter individually between both methods Tr
+for $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 
@@
 \section{Acknowledgements}
-Thanks to Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Andrew Jackson, Gavin Thomas and April Wright %TG : Not sure how many people we have to thank but it could also include Graham Slater, Nick Matzke and Trevor Hodkinson.
- % NC: Might be nice to thank Trevor, and it won't add too many lines to thank them all. You can also just calle people A.Wright if you want to reduce word count.
-for useful comments on our simulation protocol.
-%Thanks to Sive Finlay, Kevin Healy and Adam Kane for comments on the manuscript.
-%TG: or helping writing sentences does not really count.
-% NC: Depends how much they helped.
-I just used to thank the lab for that stuff
+Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Gavin Thomas and April Wright for useful comments on our simulation protocol.
 Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
@@
 \bibliography{References}
-%Needs to be cleaned a bit
-% NC: Yeah. Leave it til we send it to Gavin but it should be a priority. Almost all of them are wrong! 
+
 %---------------------------------------------

### 2015-01-09 Manuscript/SupplementaryMaterial/Supp_results.tex
Added result table per parameters

@@
 
+%Summary of metrics values for all parameters combinations
 \begin{table}[ht]
@@
 \end{table}
+
+%Summary of metrics values for the ML parameter
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+Maximum likelihood-RF & 0.44 & 0.51 & 0.63 & 0.66 & 0.78 & 0.95 \\ 
+  Maximumlikelihood-Tr & 0.45 & 0.56 & 0.76 & 0.74 & 0.93 & 0.99 \\ 
+  Bayesian consensus-RF & 0.71 & 0.73 & 0.80 & 0.82 & 0.88 & 0.95 \\ 
+  Bayesian consensus-Tr & 0.37 & 0.46 & 0.67 & 0.67 & 0.87 & 0.96 \\ 
+  Bootstraps-RF & 0.34 & 0.37 & 0.42 & 0.41 & 0.44 & 0.46 \\ 
+  Bootstraps-Tr & 0.32 & 0.40 & 0.51 & 0.46 & 0.51 & 0.57 \\ 
+  Bayesian posterior trees-RF & 0.33 & 0.41 & 0.52 & 0.50 & 0.60 & 0.65 \\ 
+  Bayesian posterior trees-Tr & 0.41 & 0.56 & 0.76 & 0.71 & 0.84 & 0.98 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+%Summary of metrics values for the ML parameter
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+Maximum likelihood-RF & 0.23 & 0.46 & 0.64 & 0.61 & 0.79 & 0.93 \\ 
+  Maximumlikelihood-Tr & 0.65 & 0.84 & 0.95 & 0.89 & 0.99 & 1.00 \\ 
+  Bayesian consensus-RF & 0.72 & 0.77 & 0.86 & 0.85 & 0.94 & 0.96 \\ 
+  Bayesian consensus-Tr & -0.16 & 0.19 & 0.63 & 0.52 & 0.96 & 0.98 \\ 
+  Bootstraps-RF & 0.14 & 0.30 & 0.40 & 0.35 & 0.45 & 0.46 \\ 
+  Bootstraps-Tr & 0.37 & 0.49 & 0.54 & 0.51 & 0.56 & 0.57 \\ 
+  Bayesian posterior trees-RF & 0.24 & 0.45 & 0.57 & 0.51 & 0.63 & 0.65 \\ 
+  Bayesian posterior trees-Tr & 0.44 & 0.81 & 0.86 & 0.82 & 0.98 & 0.98 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+%Summary of metrics values for the MC parameter
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+Maximum likelihood-RF & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
+  Maximumlikelihood-Tr & 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
+  Bayesian consensus-RF & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
+  Bayesian consensus-Tr & 0.05 & 0.16 & 0.53 & 0.50 & 0.87 & 0.92 \\ 
+  Bootstraps-RF & 0.25 & 0.34 & 0.42 & 0.38 & 0.45 & 0.46 \\ 
+  Bootstraps-Tr & 0.38 & 0.47 & 0.55 & 0.51 & 0.57 & 0.58 \\ 
+  Bayesian posterior trees-RF & 0.32 & 0.44 & 0.58 & 0.52 & 0.62 & 0.65 \\ 
+  Bayesian posterior trees-Tr & 0.39 & 0.78 & 0.82 & 0.79 & 0.98 & 0.98 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+%Summary of the BC between pairs of methods for the ML parameter
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+Maximum.likelihood.vs..Bayesian.consensus...RF & 0.30 & 0.31 & 0.69 & 0.61 & 0.77 & 1.00 \\ 
+  Maximum.likelihood.vs..Bayesian.consensus...Tr & 0.79 & 0.81 & 0.84 & 0.86 & 0.85 & 1.00 \\ 
+  Maximum.likelihood.vs..Bootstraps...RF & 0.03 & 0.22 & 0.29 & 0.36 & 0.54 & 0.69 \\ 
+  Maximum.likelihood.vs..Bootstraps...Tr & 0.08 & 0.42 & 0.53 & 0.51 & 0.74 & 0.78 \\ 
+  Maximum.likelihood.vs..Bayesian.posterior.trees...RF & 0.02 & 0.49 & 0.61 & 0.51 & 0.67 & 0.74 \\ 
+  Maximum.likelihood.vs..Bayesian.posterior.trees...Tr & 0.21 & 0.61 & 0.70 & 0.63 & 0.81 & 0.81 \\ 
+  Bayesian.consensus.vs..Bootstraps...RF & 0.01 & 0.02 & 0.02 & 0.02 & 0.03 & 0.04 \\ 
+  Bayesian.consensus.vs..Bootstraps...Tr & 0.08 & 0.69 & 0.78 & 0.64 & 0.79 & 0.84 \\ 
+  Bayesian.consensus.vs..Bayesian.posterior.trees...RF & 0.01 & 0.02 & 0.02 & 0.04 & 0.08 & 0.09 \\ 
+  Bayesian.consensus.vs..Bayesian.posterior.trees...Tr & 0.21 & 0.74 & 0.75 & 0.68 & 0.84 & 0.87 \\ 
+  Boostraps.vs..Bayesian.posterior.trees...RF & 0.69 & 0.75 & 0.85 & 0.85 & 0.95 & 1.00 \\ 
+  Boostraps.vs..Bayesian.posterior.trees...Tr & 0.91 & 0.92 & 0.96 & 0.95 & 0.97 & 0.98 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+%Summary of the BC between pairs of methods for the MF parameter
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+Maximum.likelihood.vs..Bayesian.consensus...RF & 0.00 & 0.25 & 0.48 & 0.50 & 0.76 & 1.00 \\ 
+  Maximum.likelihood.vs..Bayesian.consensus...Tr & 0.38 & 0.69 & 0.75 & 0.72 & 0.80 & 1.00 \\ 
+  Maximum.likelihood.vs..Bootstraps...RF & 0.03 & 0.18 & 0.32 & 0.36 & 0.47 & 0.77 \\ 
+  Maximum.likelihood.vs..Bootstraps...Tr & 0.08 & 0.34 & 0.40 & 0.38 & 0.53 & 0.55 \\ 
+  Maximum.likelihood.vs..Bayesian.posterior.trees...RF & 0.02 & 0.47 & 0.71 & 0.60 & 0.86 & 0.94 \\ 
+  Maximum.likelihood.vs..Bayesian.posterior.trees...Tr & 0.21 & 0.54 & 0.62 & 0.56 & 0.64 & 0.80 \\ 
+  Bayesian.consensus.vs..Bootstraps...RF & 0.00 & 0.00 & 0.01 & 0.01 & 0.01 & 0.03 \\ 
+  Bayesian.consensus.vs..Bootstraps...Tr & 0.08 & 0.38 & 0.54 & 0.49 & 0.70 & 0.75 \\ 
+  Bayesian.consensus.vs..Bayesian.posterior.trees...RF & 0.00 & 0.02 & 0.02 & 0.02 & 0.04 & 0.04 \\ 
+  Bayesian.consensus.vs..Bayesian.posterior.trees...Tr & 0.21 & 0.29 & 0.66 & 0.54 & 0.72 & 0.82 \\ 
+  Boostraps.vs..Bayesian.posterior.trees...RF & 0.69 & 0.69 & 0.72 & 0.71 & 0.72 & 0.72 \\ 
+  Boostraps.vs..Bayesian.posterior.trees...Tr & 0.91 & 0.91 & 0.91 & 0.93 & 0.92 & 0.98 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+%Summary of the BC between pairs of methods for the MC parameter
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+Maximum.likelihood.vs..Bayesian.consensus...RF & 0.03 & 0.32 & 0.66 & 0.55 & 0.75 & 1.00 \\ 
+  Maximum.likelihood.vs..Bayesian.consensus...Tr & 0.51 & 0.69 & 0.80 & 0.76 & 0.80 & 1.00 \\ 
+  Maximum.likelihood.vs..Bootstraps...RF & 0.03 & 0.17 & 0.21 & 0.31 & 0.46 & 0.68 \\ 
+  Maximum.likelihood.vs..Bootstraps...Tr & 0.08 & 0.31 & 0.39 & 0.39 & 0.56 & 0.61 \\ 
+  Maximum.likelihood.vs..Bayesian.posterior.trees...RF & 0.02 & 0.44 & 0.47 & 0.52 & 0.78 & 0.90 \\ 
+  Maximum.likelihood.vs..Bayesian.posterior.trees...Tr & 0.21 & 0.52 & 0.59 & 0.55 & 0.66 & 0.77 \\ 
+  Bayesian.consensus.vs..Bootstraps...RF & 0.00 & 0.01 & 0.01 & 0.02 & 0.02 & 0.03 \\ 
+  Bayesian.consensus.vs..Bootstraps...Tr & 0.08 & 0.47 & 0.62 & 0.51 & 0.66 & 0.73 \\ 
+  Bayesian.consensus.vs..Bayesian.posterior.trees...RF & 0.00 & 0.02 & 0.04 & 0.04 & 0.05 & 0.06 \\ 
+  Bayesian.consensus.vs..Bayesian.posterior.trees...Tr & 0.21 & 0.45 & 0.64 & 0.57 & 0.74 & 0.79 \\ 
+  Boostraps.vs..Bayesian.posterior.trees...RF & 0.69 & 0.73 & 0.73 & 0.76 & 0.81 & 0.86 \\ 
+  Boostraps.vs..Bayesian.posterior.trees...Tr & 0.91 & 0.92 & 0.93 & 0.94 & 0.96 & 0.99 \\ 
+   \hline
+\end{tabular}
+\end{table}

### 2015-01-09 Analysis/MissingData_differences.R

@@ -39,28 +39,28 @@ parameter<-par_MLMFMC
 #Calculating the pairwise comparisons within (effect of the parameters)
 #----------------
 
-#pair.comp_within.MLbest.RF<-pair.bhatt.coeff(sub.data(get(data.sets[1]), par_MLMFMC, metric=metrics[1]), diag=TRUE)
+#pair.comp_within.MLbest.RF<-pair.bhatt.coeff(sub.data(get(data.sets[1]), parameter, metric=metrics[1]), diag=TRUE)
 #save(pair.comp_within.MLbest.RF, file="../Data/R_data/pair.comp_within.MLbest.RF.Rda")
 load("../Data/R_data/pair.comp_within.MLbest.RF.Rda")
-#pair.comp_within.MLbest.Tr<-pair.bhatt.coeff(sub.data(get(data.sets[1]), par_MLMFMC, metric=metrics[2]), diag=TRUE)
+#pair.comp_within.MLbest.Tr<-pair.bhatt.coeff(sub.data(get(data.sets[1]), parameter, metric=metrics[2]), diag=TRUE)
 #save(pair.comp_within.MLbest.Tr, file="../Data/R_data/pair.comp_within.MLbest.Tr.Rda")
 load("../Data/R_data/pair.comp_within.MLbest.Tr.Rda")
-#pair.comp_within.Baycon.RF<-pair.bhatt.coeff(sub.data(get(data.sets[2]), par_MLMFMC, metric=metrics[1]), diag=TRUE)
+#pair.comp_within.Baycon.RF<-pair.bhatt.coeff(sub.data(get(data.sets[2]), parameter, metric=metrics[1]), diag=TRUE)
 #save(pair.comp_within.Baycon.RF, file="../Data/R_data/pair.comp_within.Baycon.RF.Rda")
 load("../Data/R_data/pair.comp_within.Baycon.RF.Rda")
-#pair.comp_within.Baycon.Tr<-pair.bhatt.coeff(sub.data(get(data.sets[2]), par_MLMFMC, metric=metrics[2]), diag=TRUE)
+#pair.comp_within.Baycon.Tr<-pair.bhatt.coeff(sub.data(get(data.sets[2]), parameter, metric=metrics[2]), diag=TRUE)
 #save(pair.comp_within.Baycon.Tr, file="../Data/R_data/pair.comp_within.Baycon.Tr.Rda")
 load("../Data/R_data/pair.comp_within.Baycon.Tr.Rda")
[…]

### 2015-01-09 Analysis/MissingData_modetable.R

@@ -27,14 +27,18 @@ if(quick.load == FALSE) {
 }
 
 #Storing the sub data sets
-boot.RF<-sub.data(get(data.sets[3]), par_MLMFMC, metric=metrics[1])
-boot.Tr<-sub.data(get(data.sets[3]), par_MLMFMC, metric=metrics[2])
-bayt.RF<-sub.data(get(data.sets[4]), par_MLMFMC, metric=metrics[1])
-bayt.Tr<-sub.data(get(data.sets[4]), par_MLMFMC, metric=metrics[2])
-ML.RF<-sub.data(get(data.sets[1]), par_MLMFMC, metric=metrics[1])
-ML.Tr<-sub.data(get(data.sets[1]), par_MLMFMC, metric=metrics[2])
-Bay.RF<-sub.data(get(data.sets[2]), par_MLMFMC, metric=metrics[1])
-Bay.Tr<-sub.data(get(data.sets[2]), par_MLMFMC, metric=metrics[2])
+library(hdrcde)
+data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
+metrics<-c("R.F_Cluster","Triples")
+par<-par_MLMFMC
+boot.RF<-sub.data(get(data.sets[3]), par, metric=metrics[1])
+boot.Tr<-sub.data(get(data.sets[3]), par, metric=metrics[2])
+bayt.RF<-sub.data(get(data.sets[4]), par, metric=metrics[1])
+bayt.Tr<-sub.data(get(data.sets[4]), par, metric=metrics[2])
+ML.RF<-sub.data(get(data.sets[1]), par, metric=metrics[1])
+ML.Tr<-sub.data(get(data.sets[1]), par, metric=metrics[2])
+Bay.RF<-sub.data(get(data.sets[2]), par, metric=metrics[1])
+Bay.Tr<-sub.data(get(data.sets[2]), par, metric=metrics[2])
 
 #Calculating the hdr
 ML.RF.hdr<-lapply(ML.RF[-1], hdr)
@@ -48,21 +52,21 @@ bayt.Tr.hdr<-lapply(bayt.Tr, hdr)
 
 #Extracting the modes
[…]

### 2015-01-09 Manuscript/TEM_draft.tex
Removed comments and added last §

@@
 
-%Removed figure (table instead)
-%\begin{figure} 
-%\centering
-  %  \includegraphics[width=1\textwidth]{Figures/In_main/BC-MethodComp-RF+Tr-colour.pdf}
-%\caption{Distribution of the Bhattacharyya Coefficients between methods.
-Curves represents the kernel density estimations of the Bhattacharyya Coefficients between the four different methods.
-A.
-Results for the Normalised Robinson-Foulds distance.
-B.
-Results with the for Normalised Triplets distance.
-Pikes around the two different extremes values of the Bhattacharyya Coefficients show the two pairs of most dissimilar methods (Bayesian consensus vs.
-Bootstraps and Bayesian consensus vs.
-Bayesian posterior trees) for the Normalised Robinson-Foulds distance ; and the most similar methods (Bootstraps and Bayesian posterior trees) for the Normalised Triples methods.}
-%\label{Fig_Results-paircomp_without} %Comparisons between methods, both in RF and Triplets.(method differences)
-%\end{figure}
-
 % TG: Put this table in landscape!
@@
 
-% Another practical implication of our results regards the tree building methods.
-Bayesian = better if chooisng one tree as backbone.
-However, although posterior distributions perform poorly this is more an effect of the way we compare trees (see above).
-Thus we do not suggest discarding posterior distribution.
-They are useful for comparative analyses and stuff.
+Another practical implication of our results regards the tree inference methods and their ability to recover the "best" tree topology with missing data in a Total Evidence framework.
+Because the Bayesian consensus tree consistently recovered topologies closer to the "best" tree topology than any other methods, we advice to choose this tree topology as a fixed topology for any further step in tree inference analysis such as Tip-dating \citep{BEASTmaster}.
+However, although the posterior distribution tree distribution performed poorly in recovering the "best" tree topology, this is more an effect of a caveat in our methods (sse discussion above).
+%add link
+Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed since they can be useful for further comparative analysis \citep[e.g.][]{}.
+%Bayesian PCM cite
 

### 2015-01-09 Manuscript/TEM_draft.tex
Fixed comments and figures

@@
 %Packages
+\usepackage{pdflscape}
 \usepackage{fixltx2e}
@@
 \noindent{\Large \bf Effect of missing data on topological inference using a total evidence approach}
-% TG: Or "Effect of missing data and inference method on topology when using a total evidence approach"
-% TG: Or any other suggestions?
 
@@
 
-%Authors
 \noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$, and Natalie Cooper$^1$$^,$$^2$}\\
@@
 However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments \citep[e.g.][]{pyrondivergence2011,ronquista2012,BEASTmaster}.
-%Or is it cheap to cite unpublished work (Nick's proto package) in the intro?
-% NC: Check Syst Biols policies, but I think it's fine as it's an available package so it'd be like citing something from GitHub.
-You could ask Nick before publication if there's a different citation he'd prefer you to use.
 
@@
 
-In this study, we focus on the effect of missing data on our ability to recover a "best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{}.
-%CITE-check Lloyd and Friedman or birdy stuff
+In this study, we focus on the effect of missing data on our ability to recover a "best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
 Although branch length estimation is also important (namely for timing extinction and/or speciation events - e.g.
 \citealt{ronquista2012}), we do not consider branch lengths in this study.
-This is partially due to difficulties with simulating branch lengths and topology simultaneously,
-but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+This is partially due to difficulties with simulating branch lengths and topology simultaneously, but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework.
@@
 \end{figure}
-% NC: Can you centre these around the boxes? and make the comparing topologies arrow touch the two triangles? I'll show you what I mean. 
-% NC: Also check the figure is in a format and of correct quality for Syst Biol.
 
@@
 \subsection{1-Generating the matrix}
+\label{Generating_the_matrix}
 First we randomly generated a "true" tree of 50 taxa in R v.
@@
 For the molecular data, we used the GTR + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v.
-8.0.20; \citealt{Stamatakis21012014}).% as a generalization of the HKY + $\Gamma_4$ model \citep{HKY85}.
-The GTR model can be seen as a generalization of the HKY model (the two parameters from the HKY model are implicitly included in the six from GTR model - \citealt{stamatakisa2008}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} %which is a generalization of the JC69 model \citep{jc69} with \textit{k} $\geq$ 2
- assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
+8.0.20; \citealt{Stamatakis21012014}).
+For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
 8.0.20; \citealt{Stamatakis21012014}).
@@
 On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \citep{pattengale2010many}.
-%The bootstrap algorithm used in RAxML is the Lazy Sub-tree Rearrangement (LSR) which consists of pruning one sub-tree from the tree and subsequently reinserting it to all neighbouring branches \citep{stamatakisa2008}.
-Sub-tree Pruning and Reinserting methods (SPR) have been demonstrated to be better than other methods (e.g.
-Nearest Neighbouring Interchange - NNI) for recovering good bootstrap values \citep{salamin2003}.
 Using these parameters took \texttildelow 8 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (2.30GHz clock speed nodes).
@@
 This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
-We normalised this metric by % NC: DETAILS
-\citep{Bogdanowicz2012}.
+We normalised this metric by divided the Robinson-Foulds distance by the mean Robinson-Foulds distance of 1000 random trees comparisons \citep{Bogdanowicz2012}.
 This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for the calculation details).
@@
 For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting each time in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance.
-%For each metric, we then normalized the value using the Normalized Tree Similarity scaled by the mean value of 1000 pairwise random tree comparisons for the metric in question and n = 51 taxa (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2}).
-%We compared each "missing-data" tree with the "best" tree for each of our 50 simulation runs resulting in 50 comparisons for each "missing-data" tree.
-We calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v3.1 \citep{hdrcde}.
 Also, to take into account the uncertainty of tree inference in both methods cited above, instead of using the single Maximum Likelihood and Bayesian consensus trees, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 from the posterior tree distribution of the Bayesian analysis for the "best" trees and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees).
@@
 
-\subsection{Testing the effects of the missing data parameters on topological recovery} % NC: I think this is really a separate section to tree comparison.
+\subsection{Testing the effects of the missing data parameters on topological recovery}
 Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the "best" tree topology in a Total Evidence framework.
@@
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc.
-(see Figure % FIGURE
-for more details).
+(see Fig.~\ref{Fig_Bhattacharyya_Coefficients} for more details).
 This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
@@
 
-% NC: Maybe put this in the results?
-% We represented the results in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (in green when the distribution overlap completely (Bhattacharyya Coefficient = 1) and in red when they don't (Bhattacharyya Coefficient = 0) - Fig.~\ref{Fig_Results-paircomp_within}).
-
 To assess the effect of the different tree inference and uncertainty methods (i.e.
@@
 Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc.
-(see Figure % FIGURE
-for more details).
+(see Fig.~\ref{Fig_Bhattacharyya_Coefficients} for more details).
 Note that this procedure pools results from across all missing data parameter combinations so it results in just six pairwise comparisons.
@@
 Conversely, if the two tree inference or uncertainty methods differ, the Bhattacharyya Coefficient will be close to zero.
+
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{Figures/In_main/Bhattacharyya_Coefficients.pdf}
+\caption{Bhattacharyya Coefficient calculation outline.
+A, B and C are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+\textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
+Maximum Likelihood or Bayesian).
+1: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two combination of missing data parameters.
+2: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters.}
+\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
+\end{figure}
 
@@
 \section{Results}
-
-% TG: Ok I removed all the figures + table on top of the section in here and all the text down. We might to have a think of where to anchor the figures later on.
+As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the "best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
+However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 
@@
     \includegraphics[width=1\textwidth]{Figures/In_main/AllMethods-RF+Tr-colour.pdf}
-\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (green) and Bayesian posterior tree distributions (blue).
+\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).
 The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
-}
+The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
 \label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
 \end{figure}
-
-% NC: Y axes should be standardised Robinson Foulds
-% NC: X-axes should have (Mc) etc in brackets after the names
-% NC: Also should probably have the X axes read : Missing character (Mc, %) and remove % from tick labels
-% TG: Ok, will fix later on since it's all scripted (no illustrator cheats)
-
-%NC: I would rewrite these figure labels as follows. I prefer the details of colours and things to be in brackets not in the text. Rewrite the others in a similar fashion.
-%NC: Remeber it's not topology, but topological recovery that you are measuring
-%The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (green) and Bayesian posterior tree distributions (blue).
-The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
 
@@
 
+\subsection{Individual effects of missing data parameters}
+As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} Table 5, 6 and 7).
+%add link?
+However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} Table 5, 6 and 7).
+%add link?
+
+\subsection{Combined effect of missing data parameters}
+As expected, our ability to recover the "best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
+
+\begin{figure} 
+\centering
      %Hide this figure for 'fast' build
-%\begin{figure} 
-%\centering
 %    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
-%\caption{Effect of missing data on topological recovery using Bayesian consensus trees.
+    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+\caption{Effect of missing data on topological recovery using Bayesian consensus trees.
 The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis.
-Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - red ; high probability of overlap - green).}
+Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
 \label{Fig_Results-paircomp_within}
-%\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
+\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
 
-% TG: Put this table in landscape!
+Using both Normalised Robinson-Foulds and Normalised Triplets distances from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Fig.~\ref{Fig_Results-paircomp_within}).
+However, the Normalised Robinson-Foulds distance (median Bhattacharrya coefficient = 0.79 ; Fig.~\ref{Fig_Results-paircomp_within}-A) displays more dissimilarities than the Normalised Triplets distance (median Bhattacharrya coefficient = 0.81; Fig.~\ref{Fig_Results-paircomp_within}-B).Additionally, when using the Normalised Robinson-Foulds distance, once $M_{L}$ $\geq$  50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Fig.~\ref{Fig_Results-paircomp_within}-A).
+Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-paircomp_within}-A).
+
+For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; see Table ~\ref{Tab_Results-Difference_methods}).
+However, these two methods perform worse than the Bayesian consensus trees sing Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; see Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+
+\begin{landscape}
 \begin{table}[ht]
@@
   \hline
-    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
-    Bayesian consensus              & $Tr^2$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
-    Maximum Likelihood bootstraps   & $Tr^2$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
-    Maximum Likelihood \textit{vs.} & $RF^1$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
-    Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
-    Maximum Likelihood bootstraps   & $Tr^2$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} & $RF^1$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
-    Bayesian posterior trees        & $Tr^2$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
-    Bayesian posterior tree \textit{vs.}  & $RF^1$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
-    Maximum Likelihood bootstraps   & $Tr^2$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
+                                                                       & $Tr^2$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
+                                                                       & $Tr^2$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
+                                                                       & $Tr^2$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
+                                                                       & $Tr^2$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
+                                                                       & $Tr^2$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
+                                                                       & $Tr^2$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
    \hline
@@
 \end{table}
-
-As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the "best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
-However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-
-\subsection{Individual effects of missing data parameters}
-As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always substantially lower for the Maximum Likelihood  trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = @, @ and @ %BC value for each parameter individually between both methods RF
-for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for the $M_{L}$ parameter (median Bhattacharrya Coefficient = @ %BC value for ML between both methods
-; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}), however, for the two other parameters, $M_{F}$ and $M_{C}$, the Normalised Triplets distance is substantially higher for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = @ and @  %BC value for each parameter individually between both methods Tr
-for $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-
-\subsection{Combined effect of missing data parameters}
-As expected, our ability to recover the "best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
-$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (green when the distributions overlap completely, Bhattacharyya Coefficient = 1, and red when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
-%CHANGE COLORS
-
-Using both Normalised Robinson-Foulds and Normalised Triplets distances from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Fig.~\ref{Fig_Results-paircomp_within}).
-However, the Normalised Robinson-Foulds distance (median Bhattacharrya coefficient = @ ; Fig.~\ref{Fig_Results-paircomp_within}-A) displays more dissimilarities than the Normalised Triplets distance (median Bhattacharrya coefficient = @; Fig.~\ref{Fig_Results-paircomp_within}-B).
-%value for median BC from pairwise comp in Baycon-RF and Baycon-Tr
- Additionally, when using the Normalised Robinson-Foulds distance, once $M_{L}$ $\geq$  50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Fig.~\ref{Fig_Results-paircomp_within}-A).
-Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-paircomp_within}-A).
-
-For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.96, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; see Table ~\ref{Tab_Results-Difference_methods}).
-However, these two methods perform worse than both the Maximum Likelihood and Bayesian consensus trees (median Bhattacharrya Coefficient = @ and @, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; see Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+\end{landscape}
 
@@
 When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
-Because living taxa still have molecular characters available for phylogenetic inference (see methods), %link to methods
+Because living taxa still have molecular characters available for phylogenetic inference (see \hyperref[Generating_the_matrix]{methods}), %link to methods
 even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
@@
 decrease of Normalised Triplets distance) with increasing missing data in this parameter, clade conservation (i.e.
-Normalised Robinson-Foulds distance) is still relatively good (mode = @) when the proportion of missing data is high ($M_{F}$ = 75\%).
+Normalised Robinson-Foulds distance) is still relatively good (mode = 0.72) when the proportion of missing data is high ($M_{F}$ = 75\%).
 
@@
 In other words, the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree, where the fossil will be placed with lower confidence at a higher taxonomic level.
-We argue that using the Bayesian consensus tree topology is preferable because it is more conservative \cite[e.g][]{}.
-%Cite the Darwinius debate or is that a bit dangerous?
-% NC: Check pattinson
+We argue that using the Bayesian consensus tree topology is preferable because it is more conservative \cite[e.g][]{pattinsonphylogeny2014}.
 
@@
 However, the effect of removing characters on the ability to recover the "best" tree topology is of the same order of magnitude as for the other two parameters (Fig.~\ref{Fig_Results-permeth_perparam}).
-We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se} % see above or below???.
+We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.
 
@@
 Because the Bayesian consensus tree consistently recovered topologies closer to the "best" tree topology than any other methods, we advice to choose this tree topology as a fixed topology for any further step in tree inference analysis such as Tip-dating \citep{BEASTmaster}.
-However, although the posterior distribution tree distribution performed poorly in recovering the "best" tree topology, this is more an effect of a caveat in our methods (sse discussion above).
-%add link
-Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed since they can be useful for further comparative analysis \citep[e.g.][]{}.
+However, although the posterior distribution tree distribution performed poorly in recovering the "best" tree topology, this is more an effect of a caveat in our methods (see discussion above).
+Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed since they can be useful for further comparative analysis.% \citep[e.g.][]{}.
 %Bayesian PCM cite
@@
 Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed in a Total Evidence analysis, for example when using the Tip-Dating method.
-% NC: Maybe one more woolly sentence.
+The results of this analysis are encouraging and showing that it is possible to combine both neontological and palaeontological data in the same analysis.
+Hopefully, using these approaches, will greatly improve our understanding of macroevolutionary processes through both time and space.
 
@@
 Supplementary material (R code, analyses, supplementary methods and full results) can be found in the Dryad data repository at http://dx.doi.org/10.5061/dryad.XXXX.
-%section and sentence format asked by Syst.
-Biol.
 

### 2015-01-09 Analysis/MissingData_differences.R

@@ -69,7 +69,8 @@ pairwise.comp_within<-c("pair.comp_within.MLbest.RF", "pair.comp_within.MLbest.T
                         "pair.comp_within.Baycon.RF", "pair.comp_within.Baycon.Tr",
                         "pair.comp_within.MLboot.RF", "pair.comp_within.MLboot.Tr",
                         "pair.comp_within.Baytre.RF", "pair.comp_within.Baytre.Tr")
-plot.bhatt.coeff(get(pairwise.comp_within[8]), col=c("red", "green"), col.grad=c("white", "black"))
+plot.bhatt.coeff(get(pairwise.comp_within[3]), col=c("blue", "yellow"), col.grad=c("white", "black"))
+summary(as.vector(get(pairwise.comp_within[3]))[-which(is.na(as.vector(get(pairwise.comp_within[3]))))])
 
 #----------------
 #Plotting the distribution of coefficients per method and per metric

### 2015-01-09 Analysis/MissingData_modetable.R

@@ -30,7 +30,7 @@ if(quick.load == FALSE) {
 library(hdrcde)
 data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
 metrics<-c("R.F_Cluster","Triples")
-par<-par_MLMFMC
+par<-par_MF
 boot.RF<-sub.data(get(data.sets[3]), par, metric=metrics[1])
 boot.Tr<-sub.data(get(data.sets[3]), par, metric=metrics[2])
 bayt.RF<-sub.data(get(data.sets[4]), par, metric=metrics[1])

### 2015-01-11 Manuscript/TEM_draft.tex
Minor edits to intro and methods

@@
 %Author instructions: http://www.oxfordjournals.org/our_journals/sysbio/for_authors/ms_preparation.html
+
+%NC: Supp material - add the actual table and figures for each thing. Also BE CONSISTENT. Is it Section 1 or S1? I'd go with S1 (which stands for Supp Mat 1 not section 1 really.)
 
@@
 \begin{abstract}
-Living species represent a marginal part of all species that have ever lived.
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology.
+Living species represent only a small proportion of all species that have ever lived.
+Ignoring fossil taxa may therefore lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology.
 This fact has led to an increasing consensus among scientists that both living and fossil taxa must be included in macroevolutionary studies.
@@
 
+% NC: This needs changing to reflect what we decided to focus on, i.e.
+our recommendations.
+We don't want to lead with the methods thing.
+The two previous paragraphs are fine and just need tightening up a bit
 We find that, when using a clade conservative metric such as Robinson-Foulds distance, Bayesian consensus trees recovers the right topology better than Maximum Likelihood or than a subset of Bayesian posterior tree distributions and this regardless of the amount of missing data (minimal tree similarity of 0.69 in the worst scenario).
@@
 We investigate three major parameters that directly affect the completeness of the morphological part of the matrix: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix.
+These parameters reflect empirical biases in data availability
+% NC: The next few sentences could be moved to the methods. They feel a bit clunky here.
+- the advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis, missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
-We chose these parameters because they reflect empirical biases in data availability.
-The advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
-Missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
 We find that the ability of Total Evidence methods to recover the "best" tree topology is increased when using Bayesian methods.
-Also minimizing the proportion of living taxa with no morphological data and the proportion of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree topology more so than minimizing the proportion of missing data in the fossil record.
+Also minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree topology more so than minimizing the amount of missing data in the fossil record.
 
@@
 We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$).
-The resulting 125 matrices are called hereafter "missing-data" matrices.
+We call the resulting 125 matrices "missing-data" matrices.
 \item{Inferring phylogenies:} \label{step:build_phylo} \\
@@
 
+% NC: This looks much better. Maybe put it in a box though or the shaded bits look a bit odd.
 \begin{figure}
@@
 
-\subsection{1-Generating the matrix}
+\subsection{1. Generating the matrix}
 \label{Generating_the_matrix}
@@
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (See \hyperref[SupplementaryMaterial]{Supplementary Material Section 1}).
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 1}).
 We then ran an independent discrete character simulation for each character using the "true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
@@
 
-\subsection{2-Removing data}
+\subsection{2. Removing data}
 \label{Removing_data}
@@
 In practice, each parameter represents a different way of removing data from the matrix: $M_{L}$ removes rows from the living taxa's data; $M_{F}$ removes cells from the fossil taxa's data; and $M_{C}$ removes columns across both living and fossil taxa's data.
-Note that $M_{L}$ and $M_{F}$ differ not only because of the region of the matrix affected: for $M_{L}$ all the morphological data of a percentage of living taxa is removed, whereas for $M_{F}$ a percentage of the data is removed at random from across the whole of the morphological matrix for fossil taxa.
+Note that $M_{L}$ and $M_{F}$ differ not only because of the region of the matrix affected: for $M_{L}$ all the morphological data of a percentage of living taxa are removed, whereas for $M_{F}$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
 
@@
 
-\subsection{3-Building phylogenies}
+\subsection{3. Building phylogenies}
 From the resulting matrices we generated two types of trees: the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
@@
 We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
-The effective sample size (ESS) was also checked on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
-Finally we built the Bayesian consensus tree from the combined chains excluding the 25\% first iterations (burn-in) with a strict majority rule \citep{Ronquist2012mrbayes}.
+We also checked the effective sample size (ESS) on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
+Finally we built a strict majority rule Bayesian consensus tree from the combined chains, excluding the 25\% first iterations as burn-in \citep{Ronquist2012mrbayes}.
 %---------------------------------------------
-%       4-Comparing Topologies
+%       4. Comparing Topologies
 %---------------------------------------------
 
-\subsection{4-Comparing topologies}
+\subsection{4. Comparing topologies}
 We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
 We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa  (i.e.
-taxa that frequently change position in different trees - \citealt{kearneyfragmentary2002}).
-We normalised both metrics using the Normalized Tree Similarity index \citep{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
+taxa that frequently change position in different trees; \citealt{kearneyfragmentary2002}).
+We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
 These metrics are described in detail below.
@@
 This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
-We normalised this metric by divided the Robinson-Foulds distance by the mean Robinson-Foulds distance of 1000 random trees comparisons \citep{Bogdanowicz2012}.
+We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons \citep{Bogdanowicz2012}.
 This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for the calculation details).
@@
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
-It is bounded between 0 when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
+It is bounded between zero when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
 Therefore this metric sensitive to the conservation of wildcard taxa.
@@
 
-\subsubsection{Paired trees comparisons}
+\subsubsection{Paired tree comparisons}
 \label{tree_comparisons}
-For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting each time in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance.
-Also, to take into account the uncertainty of tree inference in both methods cited above, instead of using the single Maximum Likelihood and Bayesian consensus trees, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 from the posterior tree distribution of the Bayesian analysis for the "best" trees and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees).
+For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance for each tree inference method.
+Also, to take into account the uncertainty of tree inference, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 trees from the posterior tree distribution of the Bayesian analysis for the "best" trees, and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees).
 For a given set of 1000 "missing data" trees and the 1000 "best" trees, we sampled one "missing data" tree and one "best" tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above.
 We repeated this 1000 times for each set of "missing data" trees resulting in 125$\times$1000 values for each metric.
-We repeated all the paired trees comparisons described above for each of the 50 simulations.
+We repeated all the paired tree comparisons described above for each of the 50 simulation runs.
 We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v.
@@
 To assess the effect of the different tree inference and uncertainty methods (i.e.
-Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the "best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of tree inference and uncertainty methods i.e.
+Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the "best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of tree inference and uncertainty methods, i.e.
 Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc.
@@
 
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{Figures/In_main/Bhattacharyya_Coefficients.pdf}
-\caption{Bhattacharyya Coefficient calculation outline.
+% NC: This figure isn't in the folder...
+%\begin{figure} 
+%\centering
+%    \includegraphics[width=1\textwidth]{Figures/In_main/Bhattacharyya_Coefficients.pdf}
+%\caption{Bhattacharyya Coefficient calculation outline.
 A, B and C are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
@@
 2: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters.}
-\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
-\end{figure}
+%\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
+%\end{figure}
 

### 2015-01-11 Manuscript/TEM_draft.tex
Updatted supplementaries
Tables and figures with proper format. Also BC figures in png.

--- 2015-01-09 Manuscript/TEM_draft.tex
@@
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} Table 5, 6 and 7).
-%add link?
-However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} Table 5, 6 and 7).
-%add link?
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} Table 2, 3 and 4).
+However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} Table 2, 3 and 4).
 
@@
 %Ignore this for 'fast' pdf building
-%\section{Supplementary Material}
+\section{Supplementary Material}
 
-%    \input{SupplementaryMaterial/SupplementaryMaterial}
-%    \label{SupplementaryMaterial}
+    \input{SupplementaryMaterial/SupplementaryMaterial}
+    \label{SupplementaryMaterial}
 

### 2015-01-11 Manuscript/SupplementaryMaterial/Supp_results.tex

--- 2014-12-18 Manuscript/SupplementaryMaterial/Supp_results.tex
@@
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/Boot+Baytre-AllParam-RF+Tr.pdf}
-\caption{Trend of the effect of missing data on topological recovery on the Bootstraps and the Bayesian posterior trees distributions.
-The amount of missing data per parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is represented along the x axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data.
-The topological recovery is represented on the y axis, both using Robinson-Foulds distance (upper row) and Triplets distance (lower row).
-Points represent the modal value of each distribution ; thick solid and thin dashed lines represents respectively the 50\% and 95\% confidence intervals or the distributions.
-The Bootstraps are represented in black and the Bayesian posterior trees distributions in grey.}
-\label{Fig_global_BootTreesets} %Differences between all the parameters and between two methods (Boot vs treesets)
+    \includegraphics[width=1\textwidth]{Figures/Supplementary/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
+\caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior treee distribution (grey).
+The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on the x axis.
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+\label{Fig_Supp_BootBayt_allparam} %Differences between all the parameters and between two methods (Boot vs Bayt)
 \end{figure}
@@
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/BC-AllMethods-RF+Tr.pdf}
-\caption{Distribution of the Pairwise Bhattacharyya Coefficients within each method.
-A- Distribution of the coefficients when comparing Ronbinson-Foulds distances.
-B- Distribution of the coefficients when comparing Triplets distances.}
-\label{Fig_Bhatt.coeff_distribution} %Differences in overlap within each method
-\end{figure}
+     %Hide this figure for 'fast' build
+%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.pdf}
+    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+\caption{Effect of missing data on topological recovery using the Maximum Likelihood trees.
+The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis.
+Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
+\label{Fig_Supp_paircomp_ML}
+\end{figure} %Pairwise BC for the ML trees for RF and TR. (parameters differences)
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/PairwiseComp-ML-RF+Tr.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Maximum Likelihood trees.
-The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data for each parameter.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (1) to red (0).
-A.
-Results for the Normalised Robinson-Foulds distance.
-B.
-Results with the for Normalised Triplets distance.}
-\label{Fig_pairComp-Baytree-RF}
-\end{figure} %Pairwise BC for the ML for RF+Tr. (parameters differences)
+     %Hide this figure for 'fast' build
+%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.pdf}
+    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+\caption{Effect of missing data on topological recovery using the Maximum Likelihood bootstrap trees.
+The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis.
+Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
+\label{Fig_Supp_paircomp_Boot}
+\end{figure} %Pairwise BC for the ML bootstrap trees for RF and TR. (parameters differences)
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/PairwiseComp-Boot-RF+Tr.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Bootstrap trees.
-The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data for each parameter.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (1) to red (0).
-A.
-Results for the Normalised Robinson-Foulds distance.
-B.
-Results with the for Normalised Triplets distance.}
-\label{Fig_pairComp-Baytree-Tr}
-\end{figure} %Pairwise BC for the Boot for RF+Tr. (parameters differences)
+     %Hide this figure for 'fast' build
+%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PPairwiseComp-Bayt-RF+Tr-colour.pdf}
+    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Bayt-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+\caption{Effect of missing data on topological recovery using the Bayesian posterior tree distribution.
+The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis.
+Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
+\label{Fig_Supp_paircomp_Bayt}
+\end{figure} %Pairwise BC for the Bayesian posterior trees distribution for RF and TR. (parameters differences)
 
-\begin{figure} 
+%Summary of metrics values for all parameters combinations
+\begin{table}[ht]
+\label{Tab_Supp_summary_metric_allparam}
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryMaterial/Supp_Figures/PairwiseComp-Baytre-RF+Tr.pdf}
-\caption{Pairwise Bhattacharyya Coefficients within the Bayesian posterior distribution trees.
-The pairwise trees comparisons are represent on both axis.
-The colour gradient from white to black represents respectively, 0\%, 10\%, 25\%, 50\% and 75\% of missing data for each parameter.
-The matrix represents the values of pairwise Bhattacharyya Coefficients going from green (1) to red (0).
-A.
-Results for the Normalised Robinson-Foulds distance.
-B.
-Results with the for Normalised Triplets distance.}
-\label{Fig_pairComp-MLbest-RF}
-\end{figure} %Pairwise BC for the Baytre for RF+Tr. (parameters differences)
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrr}
+\begin{tabular}{lccccccc}
   \hline
- & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+ Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
   \hline
-  Maximum likelihood-RF & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
-  Maximumlikelihood-Tr & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
-  Bayesian consensus-RF & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
-  Bayesian consensus-Tr & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
-  Bootstraps-RF & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
-  Bootstraps-Tr & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
-  Bayesian posterior trees-RF & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
-  Bayesian posterior trees-Tr & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
+  Maximum Likelihood                    & $RF^1$ & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
+                                        & $Tr^2$ & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
+  Bayesian consensus                    & $RF^1$ & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
+                                        & $Tr^2$ & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
+  Maximum Likelihood bootstraps         & $RF^1$ & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
+                                        & $Tr^2$ & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
+  Bayesian posterior tree distributions & $RF^1$ & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
+                                        & $Tr^2$ & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
    \hline
@@
 \end{tabular}
+   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
 \end{table}
+
+%Summary of metrics values for the ML parameter
+\begin{table}[ht]
+\label{Tab_Supp_summary_metric_ML}
+\centering
+\begin{tabular}{lccccccc}
+  \hline
+ Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+  Maximum Likelihood                    & $RF^1$ & 0.44 & 0.51 & 0.63 & 0.66 & 0.78 & 0.95 \\ 
+                                        & $Tr^2$ & 0.45 & 0.56 & 0.76 & 0.74 & 0.93 & 0.99 \\ 
+  Bayesian consensus                    & $RF^1$ & 0.71 & 0.73 & 0.80 & 0.82 & 0.88 & 0.95 \\ 
+                                        & $Tr^2$ & 0.37 & 0.46 & 0.67 & 0.67 & 0.87 & 0.96 \\ 
+  Maximum Likelihood bootstraps         & $RF^1$ & 0.34 & 0.37 & 0.42 & 0.41 & 0.44 & 0.46 \\ 
+                                        & $Tr^2$ & 0.32 & 0.40 & 0.51 & 0.46 & 0.51 & 0.57 \\ 
+  Bayesian posterior tree distributions & $RF^1$ & 0.33 & 0.41 & 0.52 & 0.50 & 0.60 & 0.65 \\ 
+                                        & $Tr^2$ & 0.41 & 0.56 & 0.76 & 0.71 & 0.84 & 0.98 \\ 
+   \hline
+\end{tabular}
+   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
+\end{table}
+
+%Summary of metrics values for the MF parameter
+\begin{table}[ht]
+\label{Tab_Supp_summary_metric_MF}
+\centering
+\begin{tabular}{lccccccc}
+  \hline
+ Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+  Maximum Likelihood                    & $RF^1$ & 0.23 & 0.46 & 0.64 & 0.61 & 0.79 & 0.93 \\ 
+                                        & $Tr^2$ & 0.65 & 0.84 & 0.95 & 0.89 & 0.99 & 1.00 \\ 
+  Bayesian consensus                    & $RF^1$ & 0.72 & 0.77 & 0.86 & 0.85 & 0.94 & 0.96 \\ 
+                                        & $Tr^2$ & -0.16 & 0.19 & 0.63 & 0.52 & 0.96 & 0.98 \\ 
+  Maximum Likelihood bootstraps         & $RF^1$ & 0.14 & 0.30 & 0.40 & 0.35 & 0.45 & 0.46 \\ 
+                                        & $Tr^2$ & 0.37 & 0.49 & 0.54 & 0.51 & 0.56 & 0.57 \\ 
+  Bayesian posterior tree distributions & $RF^1$ & 0.24 & 0.45 & 0.57 & 0.51 & 0.63 & 0.65 \\ 
+                                        & $Tr^2$ & 0.44 & 0.81 & 0.86 & 0.82 & 0.98 & 0.98 \\ 
+   \hline
+\end{tabular}
+   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
+\end{table}
+
+%Summary of metrics values for the MC parameter
+\begin{table}[ht]
+\label{Tab_Supp_summary_metric_MC}
+\centering
+\begin{tabular}{lccccccc}
+  \hline
+ Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+  Maximum Likelihood                    & $RF^1$ & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
+                                        & $Tr^2$ 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
+  Bayesian consensus                    & $RF^1$ & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
+                                        & $Tr^2$ & 0.05 & 0.16 & 0.53 & 0.50 & 0.87 & 0.92 \\ 
+  Maximum Likelihood bootstraps         & $RF^1$ & 0.25 & 0.34 & 0.42 & 0.38 & 0.45 & 0.46 \\ 
+                                        & $Tr^2$ & 0.38 & 0.47 & 0.55 & 0.51 & 0.57 & 0.58 \\ 
+  Bayesian posterior tree distributions & $RF^1$ & 0.32 & 0.44 & 0.58 & 0.52 & 0.62 & 0.65 \\ 
+                                        & $Tr^2$ & 0.39 & 0.78 & 0.82 & 0.79 & 0.98 & 0.98 \\ 
+   \hline
+\end{tabular}
+   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
+\end{table}
+
+%Summary of the BC between pairs of methods for the ML parameter
+\begin{landscape}
+\begin{table}[ht]
+\label{Tab_Supp_summary_BC_ML}
+\centering
+\begin{tabular}{lccccccc}
+  \hline
+ Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.30 & 0.31 & 0.69 & 0.61 & 0.77 & 1.00 \\ 
+                                                                       & $Tr^2$ & 0.79 & 0.81 & 0.84 & 0.86 & 0.85 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.22 & 0.29 & 0.36 & 0.54 & 0.69 \\ 
+                                                                       & $Tr^2$ & 0.08 & 0.42 & 0.53 & 0.51 & 0.74 & 0.78 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.49 & 0.61 & 0.51 & 0.67 & 0.74 \\ 
+                                                                       & $Tr^2$ & 0.21 & 0.61 & 0.70 & 0.63 & 0.81 & 0.81 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.01 & 0.02 & 0.02 & 0.02 & 0.03 & 0.04 \\ 
+                                                                       & $Tr^2$ & 0.08 & 0.69 & 0.78 & 0.64 & 0.79 & 0.84 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.01 & 0.02 & 0.02 & 0.04 & 0.08 & 0.09 \\ 
+                                                                       & $Tr^2$ & 0.21 & 0.74 & 0.75 & 0.68 & 0.84 & 0.87 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.69 & 0.75 & 0.85 & 0.85 & 0.95 & 1.00 \\ 
+                                                                       & $Tr^2$ & 0.91 & 0.92 & 0.96 & 0.95 & 0.97 & 0.98 \\ 
+   \hline
+\end{tabular}
+   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
+\end{table}
+\end{landscape}
+
+%Summary of the BC between pairs of methods for the MF parameter
+\begin{landscape}
+\begin{table}[ht]
+\label{Tab_Supp_summary_BC_MF}
+\centering
+\begin{tabular}{lccccccc}
+  \hline
+ Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\  
+  \hline
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.00 & 0.25 & 0.48 & 0.50 & 0.76 & 1.00 \\ 
+                                                                       & $Tr^2$ & 0.38 & 0.69 & 0.75 & 0.72 & 0.80 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.18 & 0.32 & 0.36 & 0.47 & 0.77 \\ 
+                                                                       & $Tr^2$ & 0.08 & 0.34 & 0.40 & 0.38 & 0.53 & 0.55 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.47 & 0.71 & 0.60 & 0.86 & 0.94 \\ 
+                                                                       & $Tr^2$ & 0.21 & 0.54 & 0.62 & 0.56 & 0.64 & 0.80 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.00 & 0.00 & 0.01 & 0.01 & 0.01 & 0.03 \\ 
+                                                                       & $Tr^2$ & 0.08 & 0.38 & 0.54 & 0.49 & 0.70 & 0.75 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.00 & 0.02 & 0.02 & 0.02 & 0.04 & 0.04 \\ 
+                                                                       & $Tr^2$ & 0.21 & 0.29 & 0.66 & 0.54 & 0.72 & 0.82 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.69 & 0.69 & 0.72 & 0.71 & 0.72 & 0.72 \\ 
+                                                                       & $Tr^2$ & 0.91 & 0.91 & 0.91 & 0.93 & 0.92 & 0.98 \\ 
+   \hline
+\end{tabular}
+   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
+\end{table}
+\end{landscape}
+
+%Summary of the BC between pairs of methods for the MC parameter
+\begin{landscape}
+\begin{table}[ht]
+\label{Tab_Supp_summary_BC_MC}
+\centering
+\begin{tabular}{lccccccc}
+  \hline
+ Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\  
+  \hline
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.03 & 0.32 & 0.66 & 0.55 & 0.75 & 1.00 \\ 
+                                                                       & $Tr^2$ & 0.51 & 0.69 & 0.80 & 0.76 & 0.80 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.17 & 0.21 & 0.31 & 0.46 & 0.68 \\ 
+                                                                       & $Tr^2$ & 0.08 & 0.31 & 0.39 & 0.39 & 0.56 & 0.61 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.44 & 0.47 & 0.52 & 0.78 & 0.90 \\ 
+                                                                       & $Tr^2$ & 0.21 & 0.52 & 0.59 & 0.55 & 0.66 & 0.77 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.00 & 0.01 & 0.01 & 0.02 & 0.02 & 0.03 \\ 
+                                                                       & $Tr^2$ & 0.08 & 0.47 & 0.62 & 0.51 & 0.66 & 0.73 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.00 & 0.02 & 0.04 & 0.04 & 0.05 & 0.06 \\ 
+                                                                       & $Tr^2$ & 0.21 & 0.45 & 0.64 & 0.57 & 0.74 & 0.79 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.69 & 0.73 & 0.73 & 0.76 & 0.81 & 0.86 \\ 
+                                                                       & $Tr^2$ & 0.91 & 0.92 & 0.93 & 0.94 & 0.96 & 0.99 \\ 
+   \hline
+\end{tabular}
+   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
+\end{table}
+\end{landscape}

### 2015-01-12 Manuscript/TEM_draft.tex

@@
 
-%NC: Supp material - add the actual table and figures for each thing. Also BE CONSISTENT. Is it Section 1 or S1? I'd go with S1 (which stands for Supp Mat 1 not section 1 really.)
+%NC: Supp material - add the actual table and figures for each thing.
+Also BE CONSISTENT.
+Is it Section 1 or S1?
+I'd go with S1 (which stands for Supp Mat 1 not section 1 really.).
+Normally you also refer to supplemental tables and figures as Table S1 etc.
+to distinguish between those in the main text.
+Check how Syst Biol does it.
 
@@
 Because some parameter combinations introduce a lot of missing data (e.g.
-$M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all.
+$M_L$ = 75\%, $M_F$ = 75\% and $M_C$ = 75\%), some matrices contained fossil taxa without any data at all.
 When this occurred we repeated the random deletion of characters until each taxon had at least 5\% data across the whole morphological part of the matrix.
@@
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
 \label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
@@
     \includegraphics[width=1\textwidth]{Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
-\caption{Effect of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
-The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on the x axis.
+\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
 \label{Fig_Results-global_perparam} %Differences between all the parameters and between two methods (ML vs BC)
@@
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} Table 2, 3 and 4).
-However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} Table 2, 3 and 4).
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}: Tables S2-S4).
+However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}: Tables S2-S4).
 
@@
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
+Fig.
+~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
 
@@
     \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
-\caption{Effect of missing data on topological recovery using Bayesian consensus trees.
-The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis.
-Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
+\caption{The effects of missing data on topological recovery using Bayesian consensus trees.
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
 \label{Fig_Results-paircomp_within}
@@
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Fig.~\ref{Fig_Results-paircomp_within}).
-However, the Normalised Robinson-Foulds distance (median Bhattacharrya coefficient = 0.79 ; Fig.~\ref{Fig_Results-paircomp_within}-A) displays more dissimilarities than the Normalised Triplets distance (median Bhattacharrya coefficient = 0.81; Fig.~\ref{Fig_Results-paircomp_within}-B).Additionally, when using the Normalised Robinson-Foulds distance, once $M_{L}$ $\geq$  50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Fig.~\ref{Fig_Results-paircomp_within}-A).
-Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-paircomp_within}-A).
+However, the Normalised Robinson-Foulds distance (median Bhattacharrya coefficient = 0.79; Fig.~\ref{Fig_Results-paircomp_within}A) displays more dissimilarities than the Normalised Triplets distance (median Bhattacharrya coefficient = 0.81; Fig.~\ref{Fig_Results-paircomp_within}B).
+Additionally, when using the Normalised Robinson-Foulds distance, once $M_{L}$ $\geq$  50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Fig.~\ref{Fig_Results-paircomp_within}A).
+Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-paircomp_within}A).
 
-For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; see Table ~\ref{Tab_Results-Difference_methods}).
-However, these two methods perform worse than the Bayesian consensus trees sing Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; see Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; Table ~\ref{Tab_Results-Difference_methods}).
+However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
 
+% NC: In the caption you need to explain why some of these values are in bold...
 \begin{landscape}
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data".}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
 \centering
@@
   \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
-                                                                       & $Tr^2$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
-                                                                       & $Tr^2$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
-                                                                       & $Tr^2$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
-                                                                       & $Tr^2$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
-                                                                       & $Tr^2$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
-    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
-                                                                       & $Tr^2$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
+                                                                       & $Tr$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
+                                                                       & $Tr$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
+                                                                       & $Tr$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
+                                                                       & $Tr$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
+                                                                       & $Tr$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
+                                                                       & $Tr$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
    \hline
 \end{tabular}
-   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
 \label{Tab_Results-Difference_methods}
@@
 
-Our results show that the ability to recover the "best" tree topology in a Total Evidence framework decreases with increasing amount of missing data, regardless of how data were removed or the method of tree inference used.
+Our results show that the ability to recover the "best" tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used.
 However, these factors affected topological recovery in different ways and to different extents.
 Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
-Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}).
-As shown in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
+Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam}, Fig.~\ref{Fig_Results-paircomp_within}, Table 1).
+As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
 
@@
 Therefore, our simulations reflect realistic parameters.
-Homoplasy, on the other hand, is expected to increase with an increase of the morphological characters \citep{}.
+Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{}.
+% NC: Missing a citation here.
+Also is this correct - increase and increase?
 However, the use of probabilistic methods (i.e.
@@
 \subsection{Combined effect of missing data parameters}
-As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have low missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have high missing data (i.e.
+As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have few missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a lot of missing data (i.e.
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-global_perparam}).
 However, it is important to notice that the effect of each parameter is not additive.
-The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$); when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).
+The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).
 This has important implications for building Total Evidence phylogenies (see below).
@@
 
-Another practical implication of our results regards the tree inference methods and their ability to recover the "best" tree topology with missing data in a Total Evidence framework.
-Because the Bayesian consensus tree consistently recovered topologies closer to the "best" tree topology than any other methods, we advice to choose this tree topology as a fixed topology for any further step in tree inference analysis such as Tip-dating \citep{BEASTmaster}.
-However, although the posterior distribution tree distribution performed poorly in recovering the "best" tree topology, this is more an effect of a caveat in our methods (see discussion above).
-Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed since they can be useful for further comparative analysis.% \citep[e.g.][]{}.
-%Bayesian PCM cite
+Another practical implication of our results regards the tree inference methods.
+Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in Tip-Dating \citep{BEASTmaster} % NC: Also cite ronquist paper?, and are we going with Tip-Dating?
+If yes, capitalise it throughout.
+Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see discussion above).
+Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses.% \citep[e.g.][]{}.
+%Bayesian PCM cite % Nc: Could probably cite Jetz et al again?
 
@@
 
-Using Total Evidence matrices with missing data is not a problem for recovering the "best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
+Missing data in Total Evidence matrices is not a problem for recovering the "best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
 When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
 Therefore we advise filling as many gaps in Total Evidence matrices as possible.
-Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters as possible and to code them for as many living taxa present in the matrix as possible.
-Additionally, the topology of the Bayesian consensus tree, regardless the amount of missing data, was always the closer to the "best" tree topology than the Maximum likelihood trees.
+Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.
+Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the "best" tree topology than the Maximum Likelihood trees.
 Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed in a Total Evidence analysis, for example when using the Tip-Dating method.
-The results of this analysis are encouraging and showing that it is possible to combine both neontological and palaeontological data in the same analysis.
-Hopefully, using these approaches, will greatly improve our understanding of macroevolutionary processes through both time and space.
+The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary processes through both time and space.
+% NC: Last sentence may be a bit of a stretch as we don't really talk about how to use it for macroevolutionary studies.
 
@@
 
-\section{Acknowledgements}
-Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Gavin Thomas and April Wright for useful comments on our simulation protocol.
+\section{Acknowledgments}
+Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, and April Wright for useful comments on our simulation protocol.
 Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
@@
 %---------------------------------------------
-
+% NC: I had some issues compiling this so I've commented it out.
 %Ignore this for 'fast' pdf building
@@
 
-    \input{SupplementaryMaterial/SupplementaryMaterial}
-    \label{SupplementaryMaterial}
+    %\input{SupplementaryMaterial/SupplementaryMaterial}
+    %\label{SupplementaryMaterial}
 

### 2015-01-12 Manuscript/SupplementaryMaterial/Supp_results.tex
Really minor updates on the comments #15

@@
 \label{Tab_Supp_summary_metric_allparam}
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
 \centering
@@
   \hline
-  Maximum Likelihood                    & $RF^1$ & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
-                                        & $Tr^2$ & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
-  Bayesian consensus                    & $RF^1$ & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
-                                        & $Tr^2$ & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
-  Maximum Likelihood bootstraps         & $RF^1$ & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
-                                        & $Tr^2$ & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
-  Bayesian posterior tree distributions & $RF^1$ & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
-                                        & $Tr^2$ & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
+  Maximum Likelihood                    & $RF$ & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
+                                        & $Tr$ & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
+  Bayesian consensus                    & $RF$ & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
+                                        & $Tr$ & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
+  Maximum Likelihood bootstraps         & $RF$ & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
+                                        & $Tr$ & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
+  Bayesian posterior tree distributions & $RF$ & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
+                                        & $Tr$ & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
    \hline
@@
 \label{Tab_Supp_summary_metric_ML}
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{L}$ missing data parameter only.}
 \centering
@@
   \hline
-  Maximum Likelihood                    & $RF^1$ & 0.44 & 0.51 & 0.63 & 0.66 & 0.78 & 0.95 \\ 
-                                        & $Tr^2$ & 0.45 & 0.56 & 0.76 & 0.74 & 0.93 & 0.99 \\ 
-  Bayesian consensus                    & $RF^1$ & 0.71 & 0.73 & 0.80 & 0.82 & 0.88 & 0.95 \\ 
-                                        & $Tr^2$ & 0.37 & 0.46 & 0.67 & 0.67 & 0.87 & 0.96 \\ 
-  Maximum Likelihood bootstraps         & $RF^1$ & 0.34 & 0.37 & 0.42 & 0.41 & 0.44 & 0.46 \\ 
-                                        & $Tr^2$ & 0.32 & 0.40 & 0.51 & 0.46 & 0.51 & 0.57 \\ 
-  Bayesian posterior tree distributions & $RF^1$ & 0.33 & 0.41 & 0.52 & 0.50 & 0.60 & 0.65 \\ 
-                                        & $Tr^2$ & 0.41 & 0.56 & 0.76 & 0.71 & 0.84 & 0.98 \\ 
+  Maximum Likelihood                    & $RF$ & 0.44 & 0.51 & 0.63 & 0.66 & 0.78 & 0.95 \\ 
+                                        & $Tr$ & 0.45 & 0.56 & 0.76 & 0.74 & 0.93 & 0.99 \\ 
+  Bayesian consensus                    & $RF$ & 0.71 & 0.73 & 0.80 & 0.82 & 0.88 & 0.95 \\ 
+                                        & $Tr$ & 0.37 & 0.46 & 0.67 & 0.67 & 0.87 & 0.96 \\ 
+  Maximum Likelihood bootstraps         & $RF$ & 0.34 & 0.37 & 0.42 & 0.41 & 0.44 & 0.46 \\ 
+                                        & $Tr$ & 0.32 & 0.40 & 0.51 & 0.46 & 0.51 & 0.57 \\ 
+  Bayesian posterior tree distributions & $RF$ & 0.33 & 0.41 & 0.52 & 0.50 & 0.60 & 0.65 \\ 
+                                        & $Tr$ & 0.41 & 0.56 & 0.76 & 0.71 & 0.84 & 0.98 \\ 
    \hline
@@
 \label{Tab_Supp_summary_metric_MF}
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{F}$ missing data parameter only.}
 \centering
@@
   \hline
-  Maximum Likelihood                    & $RF^1$ & 0.23 & 0.46 & 0.64 & 0.61 & 0.79 & 0.93 \\ 
-                                        & $Tr^2$ & 0.65 & 0.84 & 0.95 & 0.89 & 0.99 & 1.00 \\ 
-  Bayesian consensus                    & $RF^1$ & 0.72 & 0.77 & 0.86 & 0.85 & 0.94 & 0.96 \\ 
-                                        & $Tr^2$ & -0.16 & 0.19 & 0.63 & 0.52 & 0.96 & 0.98 \\ 
-  Maximum Likelihood bootstraps         & $RF^1$ & 0.14 & 0.30 & 0.40 & 0.35 & 0.45 & 0.46 \\ 
-                                        & $Tr^2$ & 0.37 & 0.49 & 0.54 & 0.51 & 0.56 & 0.57 \\ 
-  Bayesian posterior tree distributions & $RF^1$ & 0.24 & 0.45 & 0.57 & 0.51 & 0.63 & 0.65 \\ 
-                                        & $Tr^2$ & 0.44 & 0.81 & 0.86 & 0.82 & 0.98 & 0.98 \\ 
+  Maximum Likelihood                    & $RF$ & 0.23 & 0.46 & 0.64 & 0.61 & 0.79 & 0.93 \\ 
+                                        & $Tr$ & 0.65 & 0.84 & 0.95 & 0.89 & 0.99 & 1.00 \\ 
+  Bayesian consensus                    & $RF$ & 0.72 & 0.77 & 0.86 & 0.85 & 0.94 & 0.96 \\ 
+                                        & $Tr$ & -0.16 & 0.19 & 0.63 & 0.52 & 0.96 & 0.98 \\ 
+  Maximum Likelihood bootstraps         & $RF$ & 0.14 & 0.30 & 0.40 & 0.35 & 0.45 & 0.46 \\ 
+                                        & $Tr$ & 0.37 & 0.49 & 0.54 & 0.51 & 0.56 & 0.57 \\ 
+  Bayesian posterior tree distributions & $RF$ & 0.24 & 0.45 & 0.57 & 0.51 & 0.63 & 0.65 \\ 
+                                        & $Tr$ & 0.44 & 0.81 & 0.86 & 0.82 & 0.98 & 0.98 \\ 
    \hline
@@
 \label{Tab_Supp_summary_metric_MC}
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{C}$ missing data parameter only.}
 \centering
@@
   \hline
-  Maximum Likelihood                    & $RF^1$ & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
-                                        & $Tr^2$ 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
-  Bayesian consensus                    & $RF^1$ & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
-                                        & $Tr^2$ & 0.05 & 0.16 & 0.53 & 0.50 & 0.87 & 0.92 \\ 
-  Maximum Likelihood bootstraps         & $RF^1$ & 0.25 & 0.34 & 0.42 & 0.38 & 0.45 & 0.46 \\ 
-                                        & $Tr^2$ & 0.38 & 0.47 & 0.55 & 0.51 & 0.57 & 0.58 \\ 
-  Bayesian posterior tree distributions & $RF^1$ & 0.32 & 0.44 & 0.58 & 0.52 & 0.62 & 0.65 \\ 
-                                        & $Tr^2$ & 0.39 & 0.78 & 0.82 & 0.79 & 0.98 & 0.98 \\ 
+  Maximum Likelihood                    & $RF$ & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
+                                        & $Tr$ 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
+  Bayesian consensus                    & $RF$ & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
+                                        & $Tr$ & 0.05 & 0.16 & 0.53 & 0.50 & 0.87 & 0.92 \\ 
+  Maximum Likelihood bootstraps         & $RF$ & 0.25 & 0.34 & 0.42 & 0.38 & 0.45 & 0.46 \\ 
+                                        & $Tr$ & 0.38 & 0.47 & 0.55 & 0.51 & 0.57 & 0.58 \\ 
+  Bayesian posterior tree distributions & $RF$ & 0.32 & 0.44 & 0.58 & 0.52 & 0.62 & 0.65 \\ 
+                                        & $Tr$ & 0.39 & 0.78 & 0.82 & 0.79 & 0.98 & 0.98 \\ 
    \hline
 \end{tabular}
-   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
 \end{table}
@@
 \begin{table}[ht]
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
 \label{Tab_Supp_summary_BC_ML}
@@
   \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.30 & 0.31 & 0.69 & 0.61 & 0.77 & 1.00 \\ 
-                                                                       & $Tr^2$ & 0.79 & 0.81 & 0.84 & 0.86 & 0.85 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.22 & 0.29 & 0.36 & 0.54 & 0.69 \\ 
-                                                                       & $Tr^2$ & 0.08 & 0.42 & 0.53 & 0.51 & 0.74 & 0.78 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.49 & 0.61 & 0.51 & 0.67 & 0.74 \\ 
-                                                                       & $Tr^2$ & 0.21 & 0.61 & 0.70 & 0.63 & 0.81 & 0.81 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.01 & 0.02 & 0.02 & 0.02 & 0.03 & 0.04 \\ 
-                                                                       & $Tr^2$ & 0.08 & 0.69 & 0.78 & 0.64 & 0.79 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.01 & 0.02 & 0.02 & 0.04 & 0.08 & 0.09 \\ 
-                                                                       & $Tr^2$ & 0.21 & 0.74 & 0.75 & 0.68 & 0.84 & 0.87 \\ 
-    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.69 & 0.75 & 0.85 & 0.85 & 0.95 & 1.00 \\ 
-                                                                       & $Tr^2$ & 0.91 & 0.92 & 0.96 & 0.95 & 0.97 & 0.98 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.30 & 0.31 & 0.69 & 0.61 & 0.77 & 1.00 \\ 
+                                                                       & $Tr$ & 0.79 & 0.81 & 0.84 & 0.86 & 0.85 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.22 & 0.29 & 0.36 & 0.54 & 0.69 \\ 
+                                                                       & $Tr$ & 0.08 & 0.42 & 0.53 & 0.51 & 0.74 & 0.78 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.49 & 0.61 & 0.51 & 0.67 & 0.74 \\ 
+                                                                       & $Tr$ & 0.21 & 0.61 & 0.70 & 0.63 & 0.81 & 0.81 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.01 & 0.02 & 0.02 & 0.02 & 0.03 & 0.04 \\ 
+                                                                       & $Tr$ & 0.08 & 0.69 & 0.78 & 0.64 & 0.79 & 0.84 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.01 & 0.02 & 0.02 & 0.04 & 0.08 & 0.09 \\ 
+                                                                       & $Tr$ & 0.21 & 0.74 & 0.75 & 0.68 & 0.84 & 0.87 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.75 & 0.85 & 0.85 & 0.95 & 1.00 \\ 
+                                                                       & $Tr$ & 0.91 & 0.92 & 0.96 & 0.95 & 0.97 & 0.98 \\ 
    \hline
 \end{tabular}
-   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
 \end{table}
@@
 \begin{table}[ht]
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
 \label{Tab_Supp_summary_BC_MF}
@@
   \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.00 & 0.25 & 0.48 & 0.50 & 0.76 & 1.00 \\ 
-                                                                       & $Tr^2$ & 0.38 & 0.69 & 0.75 & 0.72 & 0.80 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.18 & 0.32 & 0.36 & 0.47 & 0.77 \\ 
-                                                                       & $Tr^2$ & 0.08 & 0.34 & 0.40 & 0.38 & 0.53 & 0.55 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.47 & 0.71 & 0.60 & 0.86 & 0.94 \\ 
-                                                                       & $Tr^2$ & 0.21 & 0.54 & 0.62 & 0.56 & 0.64 & 0.80 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.00 & 0.00 & 0.01 & 0.01 & 0.01 & 0.03 \\ 
-                                                                       & $Tr^2$ & 0.08 & 0.38 & 0.54 & 0.49 & 0.70 & 0.75 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.00 & 0.02 & 0.02 & 0.02 & 0.04 & 0.04 \\ 
-                                                                       & $Tr^2$ & 0.21 & 0.29 & 0.66 & 0.54 & 0.72 & 0.82 \\ 
-    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.69 & 0.69 & 0.72 & 0.71 & 0.72 & 0.72 \\ 
-                                                                       & $Tr^2$ & 0.91 & 0.91 & 0.91 & 0.93 & 0.92 & 0.98 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.00 & 0.25 & 0.48 & 0.50 & 0.76 & 1.00 \\ 
+                                                                       & $Tr$ & 0.38 & 0.69 & 0.75 & 0.72 & 0.80 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.18 & 0.32 & 0.36 & 0.47 & 0.77 \\ 
+                                                                       & $Tr$ & 0.08 & 0.34 & 0.40 & 0.38 & 0.53 & 0.55 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.47 & 0.71 & 0.60 & 0.86 & 0.94 \\ 
+                                                                       & $Tr$ & 0.21 & 0.54 & 0.62 & 0.56 & 0.64 & 0.80 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.00 & 0.01 & 0.01 & 0.01 & 0.03 \\ 
+                                                                       & $Tr$ & 0.08 & 0.38 & 0.54 & 0.49 & 0.70 & 0.75 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.02 & 0.02 & 0.02 & 0.04 & 0.04 \\ 
+                                                                       & $Tr$ & 0.21 & 0.29 & 0.66 & 0.54 & 0.72 & 0.82 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.69 & 0.72 & 0.71 & 0.72 & 0.72 \\ 
+                                                                       & $Tr$ & 0.91 & 0.91 & 0.91 & 0.93 & 0.92 & 0.98 \\ 
    \hline
 \end{tabular}
-   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
 \end{table}
@@
 \begin{table}[ht]
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
 \label{Tab_Supp_summary_BC_MC}
@@
   \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.03 & 0.32 & 0.66 & 0.55 & 0.75 & 1.00 \\ 
-                                                                       & $Tr^2$ & 0.51 & 0.69 & 0.80 & 0.76 & 0.80 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.17 & 0.21 & 0.31 & 0.46 & 0.68 \\ 
-                                                                       & $Tr^2$ & 0.08 & 0.31 & 0.39 & 0.39 & 0.56 & 0.61 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.44 & 0.47 & 0.52 & 0.78 & 0.90 \\ 
-                                                                       & $Tr^2$ & 0.21 & 0.52 & 0.59 & 0.55 & 0.66 & 0.77 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.00 & 0.01 & 0.01 & 0.02 & 0.02 & 0.03 \\ 
-                                                                       & $Tr^2$ & 0.08 & 0.47 & 0.62 & 0.51 & 0.66 & 0.73 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.00 & 0.02 & 0.04 & 0.04 & 0.05 & 0.06 \\ 
-                                                                       & $Tr^2$ & 0.21 & 0.45 & 0.64 & 0.57 & 0.74 & 0.79 \\ 
-    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.69 & 0.73 & 0.73 & 0.76 & 0.81 & 0.86 \\ 
-                                                                       & $Tr^2$ & 0.91 & 0.92 & 0.93 & 0.94 & 0.96 & 0.99 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.03 & 0.32 & 0.66 & 0.55 & 0.75 & 1.00 \\ 
+                                                                       & $Tr$ & 0.51 & 0.69 & 0.80 & 0.76 & 0.80 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.17 & 0.21 & 0.31 & 0.46 & 0.68 \\ 
+                                                                       & $Tr$ & 0.08 & 0.31 & 0.39 & 0.39 & 0.56 & 0.61 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.44 & 0.47 & 0.52 & 0.78 & 0.90 \\ 
+                                                                       & $Tr$ & 0.21 & 0.52 & 0.59 & 0.55 & 0.66 & 0.77 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.01 & 0.01 & 0.02 & 0.02 & 0.03 \\ 
+                                                                       & $Tr$ & 0.08 & 0.47 & 0.62 & 0.51 & 0.66 & 0.73 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.02 & 0.04 & 0.04 & 0.05 & 0.06 \\ 
+                                                                       & $Tr$ & 0.21 & 0.45 & 0.64 & 0.57 & 0.74 & 0.79 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.73 & 0.73 & 0.76 & 0.81 & 0.86 \\ 
+                                                                       & $Tr$ & 0.91 & 0.92 & 0.93 & 0.94 & 0.96 & 0.99 \\ 
    \hline
 \end{tabular}
-   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
 \end{table}

### 2015-01-12 Manuscript/TEM_draft.tex
Really minor updates on the comments #15

@@
 A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \citep{eernissetaxonomic1993}.
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as tip-dating; \citealt{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
+This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as Tip-Dating; \citealt{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
 Total Evidence methods have been successfully applied to empirical data \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
@@
 
-% NC: This looks much better. Maybe put it in a box though or the shaded bits look a bit odd.
 \begin{figure}
@@
 
-% NC: This figure isn't in the folder...
-%\begin{figure} 
-%\centering
-%    \includegraphics[width=1\textwidth]{Figures/In_main/Bhattacharyya_Coefficients.pdf}
-%\caption{Bhattacharyya Coefficient calculation outline.
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{Figures/In_main/Bhattacharyya_Coefficients.pdf}
+\caption{Bhattacharyya Coefficient calculation outline.
 A, B and C are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
@@
 2: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters.}
-%\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
-%\end{figure}
+\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
+\end{figure}
 
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
+The values in bold are the extreme values of high or low probability of overlap between two methods.}
 \centering
@@
 When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
-Because living taxa still have molecular characters available for phylogenetic inference (see \hyperref[Generating_the_matrix]{methods}), %link to methods
-even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
+Because living taxa still have molecular characters available for phylogenetic inference (see \hyperref[Generating_the_matrix]{methods}),even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
 However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
@@
 Therefore, our simulations reflect realistic parameters.
-Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{}.
+Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{wrightbayesian2014}.
 % NC: Missing a citation here.
 Also is this correct - increase and increase?
+TG: yep.
+Checking for a better cite though.
 However, the use of probabilistic methods (i.e.
@@
 Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in Tip-Dating \citep{BEASTmaster} % NC: Also cite ronquist paper?, and are we going with Tip-Dating?
-If yes, capitalise it throughout.
+Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in Tip-Dating \citep{ronquista2012,BEASTmaster}.
 Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see discussion above).
-Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses.% \citep[e.g.][]{}.
-%Bayesian PCM cite % Nc: Could probably cite Jetz et al again?
+Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses \citep[e.g.][]{jetzthe2012}.
 
@@
 Hopefully, using these approaches will greatly improve our understanding of macroevolutionary processes through both time and space.
-% NC: Last sentence may be a bit of a stretch as we don't really talk about how to use it for macroevolutionary studies.
 
@@
 
-    %\input{SupplementaryMaterial/SupplementaryMaterial}
-    %\label{SupplementaryMaterial}
+    \input{SupplementaryMaterial/SupplementaryMaterial}
+    \label{SupplementaryMaterial}
 

### 2015-01-12 Manuscript/CoverLetter.tex
Cover Letter draft

\documentclass[a4paper,11pt]{article}

\usepackage[osf]{mathpazo}
\usepackage{lastpage} 
\pagenumbering{arabic}
\linespread{1.66}

\begin{document}

\begin{flushright}
Version dated: \today
\end{flushright}

\noindent{Dear Editors,}

%introducing paragraph
A major challenge in macroevolution since the advent of the molecular clock is to accurately estimate relationships among species through time. Today, two main approaches exist to investigate this issue: the cladistics method, focusing on palaeontological data and probabilistic methods focusing on neontological data. Both approaches have their benefits but also a major issue: they do not make use of all the available data (both molecular and morphological). A third emerging approach, the Total Evidence method, allows us to address this issue. This method is extremely promising since it allows us to study macroeveolutionary patterns and processes using all the available data. However, because of the amount of data involved, the Total Evidence method is likely to suffer from problems of missing data.

%What's our paper about
Our paper, entitled "Effect of missing data on topological inference using a total evidence approach", which we submit for your consideration, is, to our knowledge, the first thorough analysis to address the effect of missing data on recovering tree topology in a Total Evidence framework. We ran extensive computational analyses ($>$ 1.5 CPU century) %or 150 CPU years?
to provide insights into the effects of missing data in a Total Evidence framework. We test the effects of three missing data parameters (the number of living taxa with morphological data; the amount of missing data in the fossil record; and the overall number of missing morphological characters) on our ability to recover the "best" tree topology by using state-of-the-art methods in tree inference and comparisons.
%Our paper, titled “Ecology and mode-of-life explain lifespan variation in birds and mammals”, which we submit for your consideration, provides insight into the selection pressures that drive lifespan in endothermic vertebrates and gives novel evolutionary context to studies of species with exceptionally long lifespans. Our analysis of 1325 birds and mammals using novel analytical methods gives the most detailed and robust insight into correlates of lifespan to-date, allowing us to test many long-standing ecological and evolutionary hypotheses in a single analysis.

%What did we find
Our key findings are that two parameters: the number of living taxa with morphological data and the overall amount of morphological characters, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology. Additionally, we found an effect of the tree inference method and we show that the Bayesian majority consensus tree always performs better than the Maximum Likelihood tree in recovering the "best" tree topology, regardless of the amount of missing data.

%What we suggest
Thus, we suggest that increasing the number of taxa with morphological data as well as the overall number of morphological characters are sufficient to improve the quality of Total Evidence tree topologies. Additionally, we suggest that fixing the topology in tree inference analysis using a Bayesian majority consensus tree could be used as a solid base for any additional analyses such as Tip-Dating or further Phylogenetic Comparative Methods.

We look forward to hearing from you soon.
\newline
\newline
\noindent{Sincerely,}
\newline
\newline
\noindent{Thomas Guillerme, on behalf of my co-author.}
\newline
\newline
Suggested reviewers: Liliana D\'{a}valos; David Pattinson; Fredrik Ronquist; Nicolas Salamin; Tanja Stadler.

\end{document}

### 2015-01-24 Manuscript/SupplementaryMaterial/TreeBuilding.tex
Updated supplementary
Tracking file edits
Edits to Supp material 1

@@
-\subsection{Morphological characters states}
-In order to obtain a realistic probabilistic value for of \textit{k} characters states for each simulated morphological character, we downloaded 100 random morphological characters (with more than 100 characters each) from TreeBASE database (http://treebase.org/) published between 1985 and 2013 and covering 19 taxomomic classess (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
-We selected a total of 22563 characters ranging from 2 to 10 states.
-We calculated the proportion of characters with 2, 3, 4, 5, 6, 7, 8, 9 or 10 states.
-We then sampeled 22563 \textit{k} values between 2 and 10 with the same proportion of characters from the empirical data.
+%\section{Supplementary material 1 : Tree building}
+
+\subsection{Morphological character states}
+To obtain a realistic value for the probability of having \textit{k} characters states for each simulated morphological character, we randomly selected 100 morphological matrices, each with more than 100 characters each, from TreeBASE (http://treebase.org/).
+We only selected matrices published between 1985 and 2013 and covering 19 taxonomic classes (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
+This resulted in a total of 22563 characters that had between two and 10 character states.
+We then extracted the proportion of characters with each number of states (two to 10) to give us an empirical estimate of the average number of character states for each character, as shown in Figure \ref{Fig_AppendixCharacters}.
+Most characters have two or three states, therefore we only simulate characters with two or three states, and sample these in proportion to their occurrence in our empirical data ()% NC: Can we put the probabilities here, I know they are in the main text)
+
+%We then sampled 22563 \textit{k} values between two and 10 with the same proportion of characters from the empirical data.
 We then used a simple t-test to check if our simulation was equal to the empirical data.
-In this study, we only simulated characters with 2 or 3 states because of the high proportion of ordered characters ecountered on characters with more than 3 states and the difficulties of simulate biologicaly sensible ordered characters.
+% NC: I don't know why this t-test or these simulations were run...
 
-\begin{figure}
+%In this study, we only simulated characters with 2 or 3 states because of the high proportion of ordered characters encountered on characters with more than 3 states and the difficulties of simulate biologicaly sensible ordered characters.
+
+\begin{figure}[h]
 \centering
-\includegraphics[keepaspectratio=true]{SupplementaryMaterial/Supp_Figures/TEM_Fig-AppendixCharacters.pdf}
-\caption{Character states distribution in empirical matrices. %Title?
-Characters states number distribution extracted from 100 random morphological matrices downloaded from RreeBase.}
+\includegraphics[keepaspectratio=true]{Figures/Supplementary/TEM_Fig-AppendixCharacters.pdf}
+\caption{The proportion of morphological characters with between two and 10 character states extracted from 100 randomly selected empirical matrices downloaded from TreeBASE.}
 \label{Fig_AppendixCharacters}
 \end{figure}
+% NC: Can you remove the title of this figure?
 
+\newpage
 \subsection{Tree Building Software settings}
 
-\subsubsection{Maximum Likelihood - RAxML v8.0.20 \citep{Stamatakis21012014}}
+For clarity we have provided the exact settings used in our tree building below.
 
-Model: \\
-Molecular data: \\
-GTR + $\Gamma_4$ (-m GTRGAMMA) \\
-Morphological data: \\
-Mk + $\Gamma_4$ (-K MK) \\
-Support: \\
-Rapid Boostrap algorithm (LSR), 1000 replicates \\
+\subsubsection{Maximum Likelihood: RAxML version 8.0.20 \citep{Stamatakis21012014}}
 
-\subsubsection{Bayesian - MrBayes v3.0.2 \citep{Ronquist2012mrbayes}}
+\begin{itemize}
+  \item Molecular data: GTR + $\Gamma_4$ (-m GTRGAMMA)
+  \item Morphological data: Mk + $\Gamma_4$ (-K MK)
+  \item Support: Rapid Boostrap algorithm (LSR), 1000 replicates
+\end{itemize}
 
-Priors: \\
-Molecular data: \\
-rates distribution shape ($\alpha$) = 0.5 \\
-Transition/Transversion ratio = 2 ($\beta$(80,40)) \\
-Starting tree: "True" tree topology with each branch length = 1 \\
-Morphological data: \\
-rates distribution shape ($\alpha$) = 0.5 \\
-Models: \\
-Molecular data: HKY + $\Gamma_4$ \\
-Morphological data: Mk + $\Gamma_4$ \\
-MCMC: \\
-2 runs \\
-4 chains per run \\
-generations < 50$\times$$1^6$ \\
-sample frequency = 1050$\times$$1^3$ \\
-ASDS diagnosis frequency = 50$\times$$1^3$ \\
-ASDS $<$ 0.01 \\
-ESS $>>$ 200 \\
-Burnin = 25\%
+\subsubsection{Bayesian: MrBayes version 3.2.1 \citep{Ronquist2012mrbayes}}
+
+\begin{itemize}
+  \item Priors
+  \begin{itemize}
+    \item Molecular data
+    \item Rates distribution shape ($\alpha$) = 0.5
+    \item Transition/Transversion ratio = 2 ($\beta$(80,40))
+    \item Starting tree: "True" tree topology with each branch length = 1
+  \end{itemize}
+  \item Morphological data
+  \begin{itemize}
+    \item rates distribution shape ($\alpha$) = 0.5
+  \end{itemize}
+  \item Models
+  \begin{itemize}
+    \item Molecular data: HKY + $\Gamma_4$
+    \item Morphological data: Mk + $\Gamma_4$
+  \end{itemize}
+  \item MCMC
+  \begin{itemize}
+    \item Two runs
+    \item Four chains per run
+    \item Generations < 50$\times$$1^6$
+    \item Sample frequency = 1050$\times$$1^3$
+    \item ASDS diagnosis frequency = 50$\times$$1^3$
+    \item ASDS $<$ 0.01
+    \item ESS $>>$ 200
+    \item Burnin = 25\%
+  \end{itemize}
+\end{itemize}

### 2015-01-24 Manuscript/SupplementaryMaterial/TreeComparison.tex
Edits to rest of supp material

--- 2014-12-01 Manuscript/SupplementaryMaterial/TreeComparison.tex
@@
-\section{Supplementary material Section 2}
+%\section{Supplementary material Section 2}
+
+\subsection{Robinson-Foulds distance}
+Robinson-Foulds distance (\textit{RF}; \citealp{RF1981}), or "path difference", measures the number of shared clades across two trees.
+The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} and can be expressed as:
+\begin{equation}
+RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
+\end{equation}
+where $C_{x,y}$ is the number of clades in common in the two trees.
+$C$ is equal to one if the two trees have the same $n$ taxa; and $C = n-2$ when none of the $n$ taxa are shared between the trees.
+This metric is more sensitive to taxon displacement than Triplets distance (i.e.
+if one taxon moves out of a clade, then the clades are no longer considered similar; \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
+The minimal value of $C$ is equal to 1 if the two trees have the same n taxa; the maximal value in $C = n-2$.
+For a fully unresolved tree (star tree) $N$=1 and for a fully resolved tree (binary tree) $N = n-2$.
+The minimal and maximal topological distance for taxa is:
+\begin{equation}
+RF_{min} = 1 + 1 - 2C_{x,y}
+\end{equation}
+and:
+\begin{equation}
+RF_{max} = 2(n-2)-2
+\end{equation}
+One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any $n$ taxa:
+\begin{equation}
+RF.scaled_{x,y} = \frac{RF_{x,y}-RF_{max}}{RF_{max}}
+\end{equation}
+This metric is more sensitive to taxa displacement than the Triplets distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.
 
 \subsection{Triplets distance details ($T_{x,y}$)}
-Triples distance ($T_{x,y}$; \citealp{dobson1975triplets}) measures the number of sub-trees made up of three taxa (triplets) that differ between two given trees.
-Each triplet can be written as $I_{ijk}$=(\textit{ijk}). Where $I_{ijk}$ is equal to zero if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to one.
+Triplets distance ($T_{x,y}$; \citealp{dobson1975triplets}) measures the number of sub-trees made up of three taxa (triplets) that differ between two given trees.
+Each triplet can be written as $I_{ijk}$=(\textit{ijk}).
+Where $I_{ijk}$ is equal to zero if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to one.
 For any rooted binary tree there are only three possible combinations for each triplet: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
@@
 \end{equation}
-Where:
+where:
 \begin{equation}
@@
 \end{equation}
-And where \textit{n} is the total number of taxa in both trees (modified from \citet{critchlowthe1996}).
-If $S_n$ = 0, the trees are identical; when $S_n$ = $\binom{n}{4}$, the trees are as different as possible (i.e. every taxon has a different placement in the two trees).
-Because the possible number of triplets per clade is a finite number, the probability of two random trees with the same n taxa to have the same triplet is:
+and where \textit{n} is the total number of taxa in both trees (modified from \citet{critchlowthe1996}).
+If $S_n$ = 0, the trees are identical; when $S_n$ = $\binom{n}{4}$, the trees are as different as possible (i.e.
+every taxon has a different placement in the two trees).
+Because the possible number of triplets per clade is a finite number, the probability of two random trees with the same $n$ taxa to have the same triplet is:
 \begin{equation}
@@
 \end{equation}
-And in the same way:
+and in the same way:
 \begin{equation}
@@
 
-\subsection{Robinson-Foulds distance details}
-Robinson-Foulds distance (\textit{RF}; \citealp{RF1981}), or "path difference"
-% NC: path difference or *distance*? - TG: difference.
-, measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} and can be expressed as following:
-\begin{equation}
-RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
-\end{equation}
-Where $C_{x,y}$ is the number of clades in common in the two trees. $C$ % NC: Why italics? Why not $C$?
-is one if the two trees have the same $n$ taxa; the maximal value is $C = n-2$. % NC: again why italics for these bits? 
-%NC: I don't get this. Explain more clearly. also is it important?
-This metric is more sensitive to taxon displacement than Triples distance (i.e.
-if one taxon moves out of a clade, then the clades are no longer considered similar; \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
-The minimal value of $C$ is equal to 1 if the two trees have the same n taxa;
-the maximal value in $C = n-2$.
-For a fully unresolved tree (star tree) $N$=1 and for a fully resolved tree (binary tree) $N = n-2$.
-The minimal and maximal topological distance for  taxa is:
-\begin{equation}
-RF_{min} = 1 + 1 - 2C_{x,y}
-\end{equation}
-And:
-\begin{equation}
-RF_{max} = 2(n-2)-2
-\end{equation}
-One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any $n$ taxa:
-\begin{equation}
-RF.scaled_{x,y} = \frac{RF_{x,y}-RF_{max}}{RF_{max}}
-\end{equation}
-This metric is more sensitive to taxa displacement than the Triplet distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.
-
 \subsection{Normalised Tree Similarity}
 For any tree with \textit{n} taxa compared using a tree distance metric $m$, Normalized Tree Similarity, $NTS_m$ (\citealp{Bogdanowicz2012}), represents the similarity score for the two trees given the expected distance between two random Yule trees with $n$ taxa.
-If $\bar{d}_{m,n}$\textit{(rand)} is the average distance between two random Yule trees with $n$ taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} containing each $n$ taxa, then:
+If $\bar{d}_{m,n}$\textit{(rand)} is the average distance between two random Yule trees with $n$ taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} each containing $n$ taxa, then:
 \begin{equation}
@@
 
-\subsection{Tree comparisons}
-\subsubsection{Random tree comparison scaling}
-We used the comparison of 1000 random trees to obtain the mean comparison value $\bar{d}_{m,n}$\textit{(rand)} for the NTS metric.
+We used the NTS method to scale all the Robinson Foulds and Triplets distances calculated in our analyses, using the TreeCmp java script \citep{Bogdanowicz2012}.
+
+% NC: Not sure this bit is needed
+%We compared 1000 randomly generated trees to obtain the mean comparison value $\bar{d}_{m,n}$\textit{(rand)}.
 We randomly generated two sets of 1000 trees of \textit{n} taxa using the rmtree function of ape package (v3.0-11 \citet{paradisape:2004}) that generates a given number of random Yule trees.
 We calculated the $\bar{d}_{m,n}$\textit{(rand)} value using an approach similar to the RPCBTC (described below) by performing 1000 random pairwise comparisons using the TreeCmp java script \citep{Bogdanowicz2012}.
+
+\subsection{Bhattacharyya Coefficient}
+The Bhattacharyya Coefficient calculates the probability of overlap of two distributions \citep{Bhattacharyya}.
+When it is equal to zero, the probability of overlap of the distributions is also zero, and when it is equal to one, the two distributions are entirely overlapping.
+It forms an elegant and easy to compute continuous measurement of the probability of similarity between two distributions.
+The coefficient is calculated as the sum of the square root of the relative counts shared in \textit{n} bins among two distributions.
+\begin{equation}
+\text{Bhattacharyya Coefficient}=\sum_{i=1}^{n} \sqrt{{\sum{a_i}}\times{\sum{b_i}}}
+\end{equation}
+where
+\begin{equation}
+\sum{a_i}=\frac{\text{Number of counts in bin \textit{i} for the distribution \textit{a}}}{\text{Total number of counts for the distribution \textit{a}}}
+\end{equation}
+and
+\begin{equation}
+\sum{b_i}=\frac{\text{Number of counts in bin \textit{i} for the distribution \textit{b}}}{\text{Total number of counts for the distribution \textit{b}}}
+\end{equation}
+The precision of the Bhattacharyya Coefficient is directly related to the number of bins, $n$.
+If $n$ is low, the overlap will be overestimated and if $n$ is too high, the overlap will be underestimated.
+In this analysis, we determined the number of bins using Silverman's rule of thumb which states that $n$ should be 0.9 times the minimum of the standard deviation and the interquartile range, % NC: of what?
+the distribution?
+divided by 1.34 times the sample size % NC: of what?
+to the negative one-fifth power (Silverman 1986, page 48, equation 3.31).
+
+%Supplementary ref: Silverman, B. W. (1986) Density Estimation. London: Chapman and Hall.

### 2015-01-25 Manuscript/SupplementaryMaterial/Supp_results.tex

@@
 \section{Supplementary material Section 3}
+
+% NC: Can you make the legends the same as the ones I already fixed in the main text please???
+% NC: Also most of the tables need to be in landscape
+% NC: Need a short paragraph here explaining that these are supplemental results for different tree building methods or whatever, just to explain why this exists!
 
@@
 \caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior treee distribution (grey).
-The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on the x axis.
+The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line)) is shown on the x axis.
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
@@
 
+% NC: Can you make the legends the same as the ones I already fixed in the main text please
 \begin{figure} 
@@
 \caption{Effect of missing data on topological recovery using the Maximum Likelihood trees.
-The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis.
-Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
+The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line)) is shown on both x and y axis.
+Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A) and Normalised Triplets distance (B)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap: blue; high probability of overlap: orange).}
 \label{Fig_Supp_paircomp_ML}
@@
 %Summary of metrics values for all parameters combinations
+% NC: Might need to be on landscape page or have first column narrower
 \begin{table}[ht]
@@
    \hline
-   \hline
 \end{tabular}
-   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
 \end{table}

### 2015-01-25 Manuscript/SupplementaryMaterial/SupplementaryMaterial.tex

@@
-\subsection{Tree Building}
+\section{Code}
+All code for performing the analyses is available at: \url{https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions} The tree comparison analysis can be repeated using code available at: \url{https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Analysis} %fully running!
+Try it at home (5-10 minutes of data loading)!
+
+\newpage
+\section{Supplementary Material 1: Tree Building}
   \input{SupplementaryMaterial/TreeBuilding}
@@
 
-\subsection{Tree Comparisons}
+\newpage
+\section{Supplementary Material 2: Tree Comparisons}
   \input{SupplementaryMaterial/TreeComparison}
@@
 
-\subsection{Codes}
-All codes are available at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions$
-
-The tree comparison results analysis can be repeated for more details at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Analysis$ %fully running!
-Try it at home (5-10 minutes of data loading)!
-
-\subsection{Additional Results}
+\newpage
+\section{Supplementary Material 3: Additional Results}
   \input{SupplementaryMaterial/Supp_results}

### 2015-01-25 Analysis/MorphologicalCharacterStates.R
Character state analysis supplementary

@@ -1,0 +1,41 @@
+#Extracting the proportion of character states from 100 matrices.
+
+#Transform the matrices using the MatRead.sh script file
+warning("UNIX based computers only.")
+system('sh ../Functions/MatRead.sh ../Data/100_Matrices')
+
+#Load the matrices from the "transformed_mat" folder (just generated from above)
+
+setwd("transformed_matrices/")
+
+#Empty list
+tab<-list() #rep('NA',length(list.files(pattern='.mat')))
+
+for (i in 1:length(list.files())) {
+	try(tmp<-read.table(list.files()[i], sep=' ',header=F), silent=TRUE)
+	try(tab.tmp<-rep(NA, ncol(tmp)), silent=TRUE)	
+	for (j in 1:ncol(tmp)) {
+		#Isolating the number of levels excluding "?" and odd ones
+		try(tab.tmp[j]<-length(grep('[0-9]', levels(as.factor(tmp[,j])))), silent=TRUE)
+	}
+	try(tab[[i]]<-tab.tmp, silent=TRUE)
+}
+	
+setwd("..")
+
+#Remove the generated folder
+system('rm -R transformed_matrices/')
+
+
+#Creating the states vector excluding chacters 0 | 1 |>10
+mat<-unlist(tab)
+matr<-mat[-c(which(mat == 0 | mat == 1 | mat > 10))]
+
+#Plotting the results
+states_probabilities<-hist(matr, breaks=c(1,2,3,4,5,6,7,8,9,10), plot=FALSE)$counts/length(matr)
+barplot(states_probabilities, xlab="Number of character states", ylab="Probabilities", names.arg=c("2","3","4","5","6","7","8","9","10"))
+
+#Saving the results
+#pdf("../Manuscript/Figures/Supplementary/TEM_Fig-AppendixCharacters.pdf")
+#barplot(states_probabilities, xlab="Number of character states", ylab="Probabilities", names.arg=c("2","3","4","5","6","7","8","9","10"), main="")
+#dev.off()

### 2015-01-25 Functions/MatRead.sh

@@ -1,0 +1,49 @@
+##########################
+#Transforming nexus morphological transformed_matrices to be read by MatRead.R
+##########################
+#Generates n transformed_matrices to be read by MatRead.R
+#Generates a citation file
+##########################
+
+Folder=$1
+
+#Creating a folder for the output and cleaning the old ones
+rm -R transformed_matrices
+rm citations.txt
+mkdir transformed_matrices
+
+#Initiating the loop
+for f in ${Folder}/*.nex
+do
+prefix=$(basename $f .nex)
+echo ${prefix}
+
+#Extracting the transformed_matrices references
+grep -A 5 'Study reference:' ${Folder}/${prefix}.nex > ${prefix}.nam.tmp
+tr '\n' ' ' < ${prefix}.nam.tmp > ${prefix}.name.tmp
+sed 's/Study reference: //g' ${prefix}.name.tmp | sed 's/  TreeBASE Study URI:.*//g' > ${prefix}.name
+
+
+#Extracting the number of taxa in order to isolate the matrix
+grep 'DIMENSIONS NTAX=' ${Folder}/${prefix}.nex > ${prefix}.ntax.tmp
+sed 's/DIMENSIONS NTAX=//g' ${prefix}.ntax.tmp | sed 's/;//g' | sed 's/ //g' > ${prefix}.ntax
+NTAX=$(sed -n '1p' ${prefix}.ntax)
+let 'NTAX += 2'
+
+#Extracting the matrix, adding a space between each characters in order to read in in R as tab.delim and collapsing the uncertainties (e.g. {01}) into missing data ("?")
+grep -A "$NTAX" 'MATRIX' ${Folder}/${prefix}.nex > ${prefix}.mat.tmp
+sed '1,3d' ${prefix}.mat.tmp | sed 's/.* //g' | sed 's/{..}/?/g' | sed 's/{...}/?/g' | sed 's/{....}/?/g' | sed 's/{.....}/?/g' | sed -e 's/\(.\)/\1 /g' > transformed_matrices/${prefix}.mat
+
+done
+
+#Writing the citation file
+for f in *.name ; do paste $f >> citations.txt ; done ; 
+
+#Cleaning temporaries files
+rm *.name.tmp
+rm *.nam.tmp
+rm *.ntax.tmp
+rm *.ntax
+rm *.mat.tmp
+rm *.name
+echo 'End'

### 2015-01-25 Manuscript/SupplementaryMaterial/Supp_results.tex
Fixed supplementary (see comment!)
There is one extra citation missing that appears only in the
supplementary (End of section 2). I didn’t manage to add an extra
reference list containing only the supplementary refs.

--- 2015-01-12 Manuscript/SupplementaryMaterial/Supp_results.tex
@@
-\section{Supplementary material Section 3}
+% NC: Can you make the legends the same as the ones I already fixed in the main text please???
+
+% NC: Need a short paragraph here explaining that these are supplemental results for different tree building methods or whatever, just to explain why this exists!
+The following section contains the supplemental results of the effect of our missing data parameters and the different tree inference methods on the the ability to recover the "correct" topology.
 
@@
     \includegraphics[width=1\textwidth]{Figures/Supplementary/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
-\caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior treee distribution (grey).
-The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on the x axis.
+    \caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey).
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
 \label{Fig_Supp_BootBayt_allparam} %Differences between all the parameters and between two methods (Boot vs Bayt)
@@
     \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
-\caption{Effect of missing data on topological recovery using the Maximum Likelihood trees.
-The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis.
-Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
+    \caption{The effects of missing data on topological recovery using Maximum Likelihood trees.
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
 \label{Fig_Supp_paircomp_ML}
@@
     \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
-\caption{Effect of missing data on topological recovery using the Maximum Likelihood bootstrap trees.
-The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis.
-Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
+    \caption{The effects of missing data on topological recovery using Maximum Likelihood bootstrap trees.
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
 \label{Fig_Supp_paircomp_Boot}
@@
     \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Bayt-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
-\caption{Effect of missing data on topological recovery using the Bayesian posterior tree distribution.
-The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis.
-Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
+    \caption{The effects of missing data on topological recovery using Bayesian posterior tree distribution.
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
 \label{Fig_Supp_paircomp_Bayt}
@@
 %Summary of metrics values for all parameters combinations
+\begin{landscape}
 \begin{table}[ht]
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
 \label{Tab_Supp_summary_metric_allparam}
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
 \centering
@@
    \hline
-   \hline
 \end{tabular}
-   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
 \end{table}
+\end{landscape}
 
 %Summary of metrics values for the ML parameter
+\begin{landscape}
 \begin{table}[ht]
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{L}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_ML}
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{L}$ missing data parameter only.}
 \centering
@@
 \end{tabular}
-   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
 \end{table}
+\end{landscape}
 
 %Summary of metrics values for the MF parameter
+\begin{landscape}
 \begin{table}[ht]
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{F}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_MF}
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{F}$ missing data parameter only.}
 \centering
@@
 \end{tabular}
-   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
 \end{table}
+\end{landscape}
 
 %Summary of metrics values for the MC parameter
+\begin{landscape}
 \begin{table}[ht]
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{C}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_MC}
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{C}$ missing data parameter only.}
 \centering
@@
   Maximum Likelihood                    & $RF$ & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
-                                        & $Tr$ 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
+                                        & $Tr$ & 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
   Bayesian consensus                    & $RF$ & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
@@
 \end{table}
+\end{landscape}
 

### 2015-01-25 Manuscript/SupplementaryMaterial/SupplementaryMaterial.tex
Fixed supplementary (see comment!)
There is one extra citation missing that appears only in the
supplementary (End of section 2). I didn’t manage to add an extra
reference list containing only the supplementary refs.

@@
 \section{Code}
-All code for performing the analyses is available at: \url{https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions} The tree comparison analysis can be repeated using code available at: \url{https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Analysis} %fully running!
-Try it at home (5-10 minutes of data loading)!
+All code for performing the analyses is available at: \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}.
 
@@
   \label{Supp_results}
+
+%\newpage
+%\section{Supplementary References}
+%    \bibliographystyle{sysbio}
+%    \bibliography{SupplementaryMaterial/Supp_Ref}

### 2015-01-25 Manuscript/SupplementaryMaterial/TreeBuilding.tex
Edits to rest of supp material
Fixed supplementary (see comment!)
There is one extra citation missing that appears only in the
supplementary (End of section 2). I didn’t manage to add an extra
reference list containing only the supplementary refs.

@@
 \subsection{Morphological character states}
-To obtain a realistic value for the probability of having \textit{k} characters states for each simulated morphological character, we randomly selected 100 morphological matrices, each with more than 100 characters each, from TreeBASE (http://treebase.org/).
+To obtain a realistic value for the probability of having \textit{k} characters states for each simulated morphological character, we randomly selected 100 morphological matrices, each with more than 100 characters each, from TreeBASE (\url{http://treebase.org/}).
 We only selected matrices published between 1985 and 2013 and covering 19 taxonomic classes (Chordata, Arthropoda, Annelida, Angiosperm, Gymnosperm and Pteridophyta).
@@
 We then extracted the proportion of characters with each number of states (two to 10) to give us an empirical estimate of the average number of character states for each character, as shown in Figure \ref{Fig_AppendixCharacters}.
-Most characters have two or three states, therefore we only simulate characters with two or three states, and sample these in proportion to their occurrence in our empirical data ()% NC: Can we put the probabilities here, I know they are in the main text)
+Most characters have two or three states, therefore we only simulate characters with two or three states, and sample these in proportion to their occurrence in our empirical data (probability of 0.85 for two states characters and 0.15 for three state characters).
+The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/MorphologicalCharacterStates.R}.
 
-%We then sampled 22563 \textit{k} values between two and 10 with the same proportion of characters from the empirical data.
-We then used a simple t-test to check if our simulation was equal to the empirical data.
-% NC: I don't know why this t-test or these simulations were run...
-
-%In this study, we only simulated characters with 2 or 3 states because of the high proportion of ordered characters encountered on characters with more than 3 states and the difficulties of simulate biologicaly sensible ordered characters.
-
-\begin{figure}[h]
+\begin{figure}
 \centering
@@
 \end{figure}
-% NC: Can you remove the title of this figure?
 
@@
 \begin{itemize}
-  \item Priors
+  \item Priors: Molecular data
   \begin{itemize}
-    \item Molecular data
     \item Rates distribution shape ($\alpha$) = 0.5
@@
   \end{itemize}
-  \item Morphological data
+  \item Priors: Morphological data
   \begin{itemize}
@@
     \item Four chains per run
-    \item Generations < 50$\times$$1^6$
-    \item Sample frequency = 1050$\times$$1^3$
-    \item ASDS diagnosis frequency = 50$\times$$1^3$
+    \item Generations $<$ 5$\times$$10^7$
+    \item Sample frequency = 1.05$\times$$10^4$
+    \item ASDS diagnosis frequency = 5$\times$$1^4$
     \item ASDS $<$ 0.01

### 2015-01-25 Manuscript/SupplementaryMaterial/TreeComparison.tex
Fixed supplementary (see comment!)
There is one extra citation missing that appears only in the
supplementary (End of section 2). I didn’t manage to add an extra
reference list containing only the supplementary refs.

@@
 \subsection{Normalised Tree Similarity}
-For any tree with \textit{n} taxa compared using a tree distance metric $m$, Normalized Tree Similarity, $NTS_m$ (\citealp{Bogdanowicz2012}), represents the similarity score for the two trees given the expected distance between two random Yule trees with $n$ taxa.
+For any tree with \textit{n} taxa compared using a tree distance metric $m$, Normalized Tree Similarity, $NTS_m$ \citep{Bogdanowicz2012}, represents the similarity score for the two trees given the expected distance between two random Yule trees with $n$ taxa.
 If $\bar{d}_{m,n}$\textit{(rand)} is the average distance between two random Yule trees with $n$ taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} each containing $n$ taxa, then:
@@
 If $n$ is low, the overlap will be overestimated and if $n$ is too high, the overlap will be underestimated.
-In this analysis, we determined the number of bins using Silverman's rule of thumb which states that $n$ should be 0.9 times the minimum of the standard deviation and the interquartile range, % NC: of what?
-the distribution?
-divided by 1.34 times the sample size % NC: of what?
-to the negative one-fifth power (Silverman 1986, page 48, equation 3.31).
-
-%Supplementary ref: Silverman, B. W. (1986) Density Estimation. London: Chapman and Hall.
+In this analysis, we determined the number of bins using Silverman's rule of thumb which states that $n$ should be 0.9 times the minimum of the standard deviation and the interquartile range of the distribution, divided by 1.34 times the sample size of the distribution to the negative one-fifth power (bw.nrd0() function in R; \citet{silverman1986density}).

### 2015-01-25 Manuscript/TEM_draft.tex
Updated references
Put supplementary on a new page
Fixed supplementary (see comment!)
There is one extra citation missing that appears only in the
supplementary (End of section 2). I didn’t manage to add an extra
reference list containing only the supplementary refs.

@@
 \usepackage{latexsym}
-\usepackage{hyperref}
 \usepackage{url}
-\usepackage{hyperref}
 \usepackage{epsfig}
@@
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see \hyperref[SupplementaryMaterial]{Supplementary Material Section 1}).
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see \hyperref[SupplementaryMaterial]{Supplementary Material 1} and Fig.\ref{Fig_AppendixCharacters}).
 We then ran an independent discrete character simulation for each character using the "true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
@@
 Even using these priors, it took \texttildelow 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-The detailed MrBayes parameters are available in the \hyperref[SupplementaryMaterial]{Supplementary Material S1}.
+The detailed MrBayes parameters are available in the \hyperref[SupplementaryMaterial]{Supplementary Material 1}.
 
@@
 The Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for calculation details).
+The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{Supplementary Material 2} for calculation details).
 This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
@@
 We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons \citep{Bogdanowicz2012}.
-This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for the calculation details).
+This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material 2} for the calculation details).
 Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
@@
 \subsubsection{Triplets distance}
-The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{Supplementary Material S2} for calculation details).
+The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{Supplementary Material 2} for calculation details).
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
@@
 We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see \hyperref[SupplementaryMaterial]{Supplementary Material Section 2} for calculation details).
+The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see \hyperref[SupplementaryMaterial]{Supplementary Material 2} for calculation details).
 Note that this is  equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
@@
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}: Tables S2-S4).
-However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}: Tables S2-S4).
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Supplementary Tables ~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}).
+However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Supplementary Tables ~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}).
 
@@
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
-Fig.
-~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; Supplementary Tables ~\ref{Tab_Supp_summary_metric_ML}, ~\ref{Tab_Supp_summary_metric_MF} and ~\ref{Tab_Supp_summary_metric_MC}).
+Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
 
@@
 For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; Table ~\ref{Tab_Results-Difference_methods}).
-However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).
+However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and Supplementary Fig.~\ref{Fig_Supp_BootBayt_allparam}).
 
-% NC: In the caption you need to explain why some of these values are in bold...
 \begin{landscape}
@@
 %
-%       Supplementary, separated
+%       Supplementary
 %
 %---------------------------------------------
-% NC: I had some issues compiling this so I've commented it out.
+
 %Ignore this for 'fast' pdf building
+\newpage
 \section{Supplementary Material}
+\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
 

### 2015-01-25 Manuscript/CoverLetter.tex
Updated cover letter

@@
 Today, two main approaches exist to investigate this issue: the cladistics method, focusing on palaeontological data and probabilistic methods focusing on neontological data.
-Both approaches have their benefits but also a major issue: they do not make use of all the available data (both molecular and morphological).
+Both approaches comes with their benefits but also with a major issue: they do not make use of all the available data (both molecular and morphological).
 A third emerging approach, the Total Evidence method, allows us to address this issue.
@@
 to provide insights into the effects of missing data in a Total Evidence framework.
-We test the effects of three missing data parameters (the number of living taxa with morphological data; the amount of missing data in the fossil record; and the overall number of missing morphological characters) on our ability to recover the "best" tree topology by using state-of-the-art methods in tree inference and comparisons.
-%Our paper, titled “Ecology and mode-of-life explain lifespan variation in birds and mammals”, which we submit for your consideration, provides insight into the selection pressures that drive lifespan in endothermic vertebrates and gives novel evolutionary context to studies of species with exceptionally long lifespans.
-Our analysis of 1325 birds and mammals using novel analytical methods gives the most detailed and robust insight into correlates of lifespan to-date, allowing us to test many long-standing ecological and evolutionary hypotheses in a single analysis.
+We test the effects of three missing data parameters (the number of living taxa with morphological data; the amount of missing data in the fossil record; and the overall number of morphological characters) on our ability to recover the "best" tree topology by using state-of-the-art methods in tree inference and tree comparisons.
 
@@
 %What we suggest
-Thus, we suggest that increasing the number of taxa with morphological data as well as the overall number of morphological characters are sufficient to improve the quality of Total Evidence tree topologies.
+Thus, we suggest that increasing the number of taxa with morphological data as well as the overall number of morphological characters is a sufficient strategy to highly improve the quality of Total Evidence tree topologies and that the problem of missing data in the fossil record is not a major issue when using this method.
 Additionally, we suggest that fixing the topology in tree inference analysis using a Bayesian majority consensus tree could be used as a solid base for any additional analyses such as Tip-Dating or further Phylogenetic Comparative Methods.

### 2015-01-25 Manuscript/TEM_draft.tex
Updated abstract

@@
 %Author instructions: http://www.oxfordjournals.org/our_journals/sysbio/for_authors/ms_preparation.html
-
-%NC: Supp material - add the actual table and figures for each thing.
-Also BE CONSISTENT.
-Is it Section 1 or S1?
-I'd go with S1 (which stands for Supp Mat 1 not section 1 really.).
-Normally you also refer to supplemental tables and figures as Table S1 etc.
-to distinguish between those in the main text.
-Check how Syst Biol does it.
 
@@
 The two previous paragraphs are fine and just need tightening up a bit
-We find that, when using a clade conservative metric such as Robinson-Foulds distance, Bayesian consensus trees recovers the right topology better than Maximum Likelihood or than a subset of Bayesian posterior tree distributions and this regardless of the amount of missing data (minimal tree similarity of 0.69 in the worst scenario).
+%We find that, when using a clade conservative metric such as Robinson-Foulds distance, Bayesian consensus trees recovers the right topology better than Maximum Likelihood or than a subset of Bayesian posterior tree distributions and this regardless of the amount of missing data (minimal tree similarity of 0.69 in the worst scenario).
 Regarding the data, good topological recovery is not related to the amount of missing data \textit{per se} but to the amount of data overlap.
 The two factors that influences the most this data overlap are the parameters 1 (proportion of coded living taxa) and 3 (number of morphological characters).
+
+We find that increasing the number of living taxa with morphological characters (1) and the overall amount of morphological data (3) have a greater impact on the ability to recover the "correct" topology than increasing the data completeness in the fossil record (2).
+Therefore, we suggest that sampling effort should be focusing on these two first aspects to increase the accuracy of topology in the Total Evidence framework.
+Additionally, we found that the Bayesian consensus tree consistently recovers topology that are closest to the "correct" topology than any other method.
+Therefore, we also suggest, after the data collection effort, that one should fix the tree topology using the Bayesian consensus tree topology for any further tree inference or macroevolutionary analysis.
+
 \end{abstract}
 
-\noindent (Keywords: missing data, Total Evidence, Bayesian, Maximum Likelihood, topology)\\
+\noindent (Keywords: missing data, Total Evidence, Bayesian, Maximum Likelihood, topology, fossil, living)\\
 

### 2015-01-26 Manuscript/TEM_draft.tex
Formatting updates + misc
Title
Key words
Italic corresponding author
Funding section

@@
 \usepackage[osf]{mathpazo}
+\usepackage{dcolumn}
+\usepackage{lineno}
 \pagenumbering{arabic}
@@
 %Pagination style and stuff
-\linespread{1.66}
+\linespread{2}
 \raggedright
@@
 \bigskip
-\noindent RH: Missing data and topology in total evidence approach
+\noindent RH: Missing data and topology in Total Evidence approach
 
@@
 
-\noindent{\Large \bf Effect of missing data on topological inference using a total evidence approach}
+\noindent{\Large \bf Effects of missing data on topological inference using a Total Evidence approach}
 
@@
 \medskip
-\noindent{\bf Corresponding author:} Thomas Guillerme, School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.\\
+\noindent{\bf Corresponding author:} \textit{Thomas Guillerme, School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\
 \vspace{1in}
+
+%Line numbering
+\modulolinenumbers[1]
+\linenumbers
 
@@
 
-\noindent (Keywords: missing data, Total Evidence, Bayesian, Maximum Likelihood, topology, fossil, living)\\
+\noindent (Keywords: morphological characters, Bayesian, Maximum Likelihood, topology, fossil, living)\\
 
@@
 We repeated these four steps 50 times to account for variation in our random parameters in the simulations.
-
-\begin{figure}
-\centering
-    \includegraphics[keepaspectratio=true]{Figures/In_main/Simulations_outline-BW.pdf}
-\caption{Protocol outline.
-(1) We randomly generated a birth-death tree (the "true" tree) and used it to infer a matrix with no missing data (the "complete" matrix).
-(2) We removed data from the morphological part of the "complete" matrix resulting in 125 "missing-data" matrices.
-(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
-(4) We compared the "missing-data" trees to the "best" tree.
-We repeated these four steps 50 times.}
-\label{Fig_Outline}
-\end{figure}
 
@@
 Even using these priors, it took \texttildelow 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-The detailed MrBayes parameters are available in the \hyperref[SupplementaryMaterial]{Supplementary Material 1}.
+The detailed MrBayes parameters are available in the \hyperref[SupplementaryMaterial]{Appendix 1}.
 
@@
 The Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{Supplementary Material 2} for calculation details).
+The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{Appendix 2} for calculation details).
 This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
@@
 We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons \citep{Bogdanowicz2012}.
-This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Supplementary Material 2} for the calculation details).
+This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Appendix 2} for the calculation details).
 Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
@@
 \subsubsection{Triplets distance}
-The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{Supplementary Material 2} for calculation details).
+The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{Appendix 2} for calculation details).
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
@@
 We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see \hyperref[SupplementaryMaterial]{Supplementary Material 2} for calculation details).
+The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see \hyperref[SupplementaryMaterial]{Appendix 2} for calculation details).
 Note that this is  equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
@@
 
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{Figures/In_main/Bhattacharyya_Coefficients.pdf}
-\caption{Bhattacharyya Coefficient calculation outline.
-A, B and C are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
-$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
-\textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
-Maximum Likelihood or Bayesian).
-1: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two combination of missing data parameters.
-2: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters.}
-\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
-\end{figure}
-
 %---------------------------------------------
@@
 
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{Figures/In_main/AllMethods-RF+Tr-colour.pdf}
-\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).
-The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-\label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
-\end{figure}
-
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
-\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
-The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-\label{Fig_Results-global_perparam} %Differences between all the parameters and between two methods (ML vs BC)
-\end{figure}
-
 \subsection{Individual effects of missing data parameters}
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Supplementary Tables ~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}).
-However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Supplementary Tables ~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}).
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Appendix Tables %~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}
+S5, S6 and S7 %Modified for build without supplementaries
+). 
+However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Appendix Tables %~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}
+S5, S6 and S7 %Modified for build without supplementaries
+).
 
@@
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; Supplementary Tables ~\ref{Tab_Supp_summary_metric_ML}, ~\ref{Tab_Supp_summary_metric_MF} and ~\ref{Tab_Supp_summary_metric_MC}).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; Appendix Tables %~\ref{Tab_Supp_summary_metric_ML}, ~\ref{Tab_Supp_summary_metric_MF} and ~\ref{Tab_Supp_summary_metric_MC}
+S2, S3 and S4 %Modified for build without supplementaries
+).
 Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
-
-\begin{figure} 
-\centering
-     %Hide this figure for 'fast' build
-%    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
-\caption{The effects of missing data on topological recovery using Bayesian consensus trees.
-Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
-\label{Fig_Results-paircomp_within}
-\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
 
@@
 For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; Table ~\ref{Tab_Results-Difference_methods}).
-However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and Supplementary Fig.~\ref{Fig_Supp_BootBayt_allparam}).
-
-\begin{landscape}
-\begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
-The values in bold are the extreme values of high or low probability of overlap between two methods.}
-\centering
-\begin{tabular}{lccccccc}
-  \hline
- Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
-  \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
-                                                                       & $Tr$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
-                                                                       & $Tr$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
-                                                                       & $Tr$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
-                                                                       & $Tr$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
-                                                                       & $Tr$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
-    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
-                                                                       & $Tr$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
-   \hline
-\end{tabular}
-\label{Tab_Results-Difference_methods}
-\end{table}
-\end{landscape}
+However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and Appendix Fig.%~\ref{Fig_Supp_BootBayt_allparam}
+S2 %Modified for build without supplementaries
+).
 
@@
 Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.
+
+\section{Funding}
 This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
@@
 
-%Ignore this for 'fast' pdf building
+%Removed to a separate file
+%\newpage
+%\section{Supplementary Material}
+%\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
+%\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
+
+%    \input{SupplementaryMaterial/SupplementaryMaterial}
+%    \label{SupplementaryMaterial}
+
+%---------------------------------------------
+%Tables
+%---------------------------------------------
 \newpage
-\section{Supplementary Material}
-\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
 
-    \input{SupplementaryMaterial/SupplementaryMaterial}
-    \label{SupplementaryMaterial}
+\begin{landscape}
+\begin{table}[ht]
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
+The values in bold are the extreme values of high or low probability of overlap between two methods.}
+\centering
+\begin{tabular}{lccccccc}
+  \hline
+ Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
+                                                                       & $Tr$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
+                                                                       & $Tr$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
+                                                                       & $Tr$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
+                                                                       & $Tr$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
+                                                                       & $Tr$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
+                                                                       & $Tr$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
+   \hline
+\end{tabular}
+\label{Tab_Results-Difference_methods}
+\end{table}
+\end{landscape}
 
+%---------------------------------------------
+%Figures Caption
+%---------------------------------------------
+
+\begin{figure}
+\caption{Protocol outline.
+(1) We randomly generated a birth-death tree (the "true" tree) and used it to infer a matrix with no missing data (the "complete" matrix).
+(2) We removed data from the morphological part of the "complete" matrix resulting in 125 "missing-data" matrices.
+(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
+(4) We compared the "missing-data" trees to the "best" tree.
+We repeated these four steps 50 times.}
+\label{Fig_Outline}
+\end{figure}
+
+\begin{figure}
+\caption{Bhattacharyya Coefficient calculation outline.
+A, B and C are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+\textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
+Maximum Likelihood or Bayesian).
+1: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two combination of missing data parameters.
+2: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters.}
+\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
+\end{figure}
+
+\begin{figure}
+\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).
+The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+\label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
+\end{figure}
+
+\begin{figure}
+\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+\label{Fig_Results-global_perparam} %Differences between all the parameters and between two methods (ML vs BC)
+\end{figure}
+
+\begin{figure}
+\caption{The effects of missing data on topological recovery using Bayesian consensus trees.
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
+\label{Fig_Results-paircomp_within}
+\end{figure}
+
+%---------------------------------------------
+%Figures
+%---------------------------------------------
+\newpage
+
+\begin{figure}[H]
+\centering
+    \includegraphics[keepaspectratio=true]{Figures/In_main/Simulations_outline-BW.pdf}
+%\caption{Protocol outline.
+%(1) We randomly generated a birth-death tree (the "true" tree) and used it to infer a matrix with no missing data (the "complete" matrix).
+%(2) We removed data from the morphological part of the "complete" matrix resulting in 125 "missing-data" matrices.
+%(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
+%(4) We compared the "missing-data" trees to the "best" tree.
+%We repeated these four steps 50 times.}
+%\label{Fig_Outline}
+\end{figure}
+Figure 1
+
+\newpage
+
+\begin{figure}[H]
+\centering
+    \includegraphics[width=1\textwidth]{Figures/In_main/Bhattacharyya_Coefficients.pdf}
+%\caption{Bhattacharyya Coefficient calculation outline.
+A, B and C are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+\textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
+Maximum Likelihood or Bayesian).
+1: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two combination of missing data parameters.
+2: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters.}
+%\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
+\end{figure}
+Figure 2
+
+\newpage
+
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{Figures/In_main/AllMethods-RF+Tr-colour.pdf}
+%\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).
+The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+%\label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
+\end{figure}
+Figure 3
+
+\newpage
+
+\begin{figure}[H]
+\centering
+    \includegraphics[width=1\textwidth]{Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
+%\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+%\label{Fig_Results-global_perparam} %Differences between all the parameters and between two methods (ML vs BC)
+\end{figure}
+Figure 4
+
+\newpage
+
+\begin{figure}[H]
+\centering
+     %Hide this figure for 'fast' build
+%    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
+    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+%\caption{The effects of missing data on topological recovery using Bayesian consensus trees.
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
+%\label{Fig_Results-paircomp_within}
+\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
+Figure 5
 %END

### 2015-01-26 Manuscript/SupplementaryMaterial/Supp_results.tex
Separated the supplementaries

@@
 % NC: Need a short paragraph here explaining that these are supplemental results for different tree building methods or whatever, just to explain why this exists!
-The following section contains the supplemental results of the effect of our missing data parameters and the different tree inference methods on the the ability to recover the "correct" topology.
+The following section contains the supplemental results of the effects of our missing data parameters and the different tree inference methods on the the ability to recover the "best" topology that are briefly discussed in the main body of the paper.
+For clarity, in the paper, we focused on the results of the effects of our missing data parameters on the Maximum Likelihood trees topology and the Bayesian consensus trees topology.
+The following additional results presented here give a greater insight into the effect of our missing data parameters on the Maximum Likelihood Bootstrapped trees topologies and the Bayesian posterior trees distribution topologies.
+Also we present here the pairwise comparisons between each parameters states for the Maximum Likelihood tree topology, the Maximum Likelihood Bootstrapped trees topologies and the Bayesian posterior trees distribution topologies.
 
@@
 \centering
-    \includegraphics[width=1\textwidth]{Figures/Supplementary/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
+    \includegraphics[width=1\textwidth]{SupplementaryFigures/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
     \caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey).
@@
 %    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-ML-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
     \caption{The effects of missing data on topological recovery using Maximum Likelihood trees.
@@
 %    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-Boot-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
     \caption{The effects of missing data on topological recovery using Maximum Likelihood bootstrap trees.
@@
 %    \includegraphics[width=1\textwidth]{Figures/Supplementary/PPairwiseComp-Bayt-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Bayt-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-Bayt-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
     \caption{The effects of missing data on topological recovery using Bayesian posterior tree distribution.

### 2015-01-26 Manuscript/CoverLetter.tex

@@
-\documentclass[a4paper,11pt]{article}
-
+\documentclass[11pt]{letter}
+\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
 \usepackage[osf]{mathpazo}
-\usepackage{lastpage} 
-\pagenumbering{arabic}
-\linespread{1.66}
-
+\signature{Thomas Guillerme \\ Natalie Cooper}
+\address{Zoology building \\ Trinity College Dublin \\ Dublin 2, Ireland \\ \\ guillert@tcd.ie}
+\longindentation=0pt
 \begin{document}
 
-\begin{flushright}
-Version dated: \today
-\end{flushright}
+\begin{letter}{}
+\opening{Dear Editors,}
 
-\noindent{Dear Editors,}
+In recent years there has been growing interest in building phylogenies that contain both living and fossil taxa (e.g.
+Fritz et al.
+2013 TREE; Heath et al.
+2014 PNAS).
+Such phylogenies could revolutionize the way we think about macroevolutionary patterns and processes, and provide a more complete understanding of trends in biodiversity through time.
+Unfortunately building such trees has proved technically difficult.
 
-%introducing paragraph
-A major challenge in macroevolution since the advent of the molecular clock is to accurately estimate relationships among species through time.
-Today, two main approaches exist to investigate this issue: the cladistics method, focusing on palaeontological data and probabilistic methods focusing on neontological data.
-Both approaches comes with their benefits but also with a major issue: they do not make use of all the available data (both molecular and morphological).
-A third emerging approach, the Total Evidence method, allows us to address this issue.
-This method is extremely promising since it allows us to study macroeveolutionary patterns and processes using all the available data.
-However, because of the amount of data involved, the Total Evidence method is likely to suffer from problems of missing data.
+One method, the Total Evidence method, allows us to use molecular and morphological data to build phylogenies with both living and fossil species as tips (Ronquist et al.
+2012 Syst Biol).
+This method is extremely promising because it allows us to use all the available data.
+However, because of the amount of data involved, the Total Evidence method is likely to be affected by missing data.
 
-%What's our paper about
-Our paper, entitled "Effect of missing data on topological inference using a total evidence approach", which we submit for your consideration, is, to our knowledge, the first thorough analysis to address the effect of missing data on recovering tree topology in a Total Evidence framework.
-We ran extensive computational analyses ($>$ 1.5 CPU century) %or 150 CPU years?
-to provide insights into the effects of missing data in a Total Evidence framework.
-We test the effects of three missing data parameters (the number of living taxa with morphological data; the amount of missing data in the fossil record; and the overall number of morphological characters) on our ability to recover the "best" tree topology by using state-of-the-art methods in tree inference and tree comparisons.
+Our paper, entitled "Effects of missing data on topological inference using a Total Evidence approach", is to our knowledge, the first to thoroughly analyze the effects of missing data on tree topology in a Total Evidence framework.
+Using simulations ($>$ 150 CPU years worth), we find that the number of living taxa with morphological data and the overall number of morphological characters, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology.
+Additionally, we show that Bayesian methods outperform Maximum Likelihood methods, regardless of the amount of missing data.
 
-%What did we find
-Our key findings are that two parameters: the number of living taxa with morphological data and the overall amount of morphological characters, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology.
-Additionally, we found an effect of the tree inference method and we show that the Bayesian majority consensus tree always performs better than the Maximum Likelihood tree in recovering the "best" tree topology, regardless of the amount of missing data.
+Our results suggest that increasing the number of taxa with morphological data and the overall number of morphological characters will greatly improve the quality of Total Evidence tree topologies.
+This has major implications for clades where detailed morphological data collection from living species is rare.
+Additionally, we recommend using a Bayesian majority consensus tree when fixing tree topology for any additional analyses such as Tip-Dating (Ronquist et al.
+2012 Syst Biol).
 
-%What we suggest
-Thus, we suggest that increasing the number of taxa with morphological data as well as the overall number of morphological characters is a sufficient strategy to highly improve the quality of Total Evidence tree topologies and that the problem of missing data in the fossil record is not a major issue when using this method.
-Additionally, we suggest that fixing the topology in tree inference analysis using a Bayesian majority consensus tree could be used as a solid base for any additional analyses such as Tip-Dating or further Phylogenetic Comparative Methods.
+We feel this paper will be of particular interest to the readers of Systematic Biology as many papers investigating methods for combining living and fossil taxa and the effects of missing data on such analyses (e.g.
+D\'{a}valos et al.
+2014; Pattinson et al.
+in press, Pyron 2011, Ronquist et al.
+2012) % Any more?
+Without going on to a second page?
+, have been recently published in the journal.
 
-We look forward to hearing from you soon.
-\newline
-\newline
-\noindent{Sincerely,}
-\newline
-\newline
-\noindent{Thomas Guillerme, on behalf of my co-author.}
-\newline
-\newline
+We look forward to hearing from you soon,
+
+\closing{Yours sincerely,}
+
 Suggested reviewers: Liliana D\'{a}valos; David Pattinson; Fredrik Ronquist; Nicolas Salamin; Tanja Stadler.
 
+\end{letter}
 \end{document}

### 2015-01-26 Manuscript/TEM_draft.tex
Shortened running head + added "online" appendix

@@
 \bigskip
-\noindent RH: Missing data and topology in Total Evidence approach
+\noindent RH: Missing data and topology in Total Evidence method
 
@@
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see \hyperref[SupplementaryMaterial]{Supplementary Material 1} and Fig.\ref{Fig_AppendixCharacters}).
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see \hyperref[SupplementaryMaterial]{online Appendix 1} and Fig.\ref{Fig_AppendixCharacters}).
 We then ran an independent discrete character simulation for each character using the "true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
@@
 Even using these priors, it took \texttildelow 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-The detailed MrBayes parameters are available in the \hyperref[SupplementaryMaterial]{Appendix 1}.
+The detailed MrBayes parameters are available in the \hyperref[SupplementaryMaterial]{online Appendix 1}.
 
@@
 The Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{Appendix 2} for calculation details).
+The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{online Appendix 2} for calculation details).
 This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
@@
 We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons \citep{Bogdanowicz2012}.
-This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{Appendix 2} for the calculation details).
+This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{online Appendix 2} for the calculation details).
 Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
@@
 \subsubsection{Triplets distance}
-The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{Appendix 2} for calculation details).
+The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{online Appendix 2} for calculation details).
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
@@
 We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see \hyperref[SupplementaryMaterial]{Appendix 2} for calculation details).
+The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see \hyperref[SupplementaryMaterial]{online Appendix 2} for calculation details).
 Note that this is  equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
@@
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Appendix Tables %~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; online Appendix Tables %~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}
 S5, S6 and S7 %Modified for build without supplementaries
 ). 
-However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Appendix Tables %~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}
+However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; online Appendix Tables %~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}
 S5, S6 and S7 %Modified for build without supplementaries
@@
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; Appendix Tables %~\ref{Tab_Supp_summary_metric_ML}, ~\ref{Tab_Supp_summary_metric_MF} and ~\ref{Tab_Supp_summary_metric_MC}
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; online Appendix Tables %~\ref{Tab_Supp_summary_metric_ML}, ~\ref{Tab_Supp_summary_metric_MF} and ~\ref{Tab_Supp_summary_metric_MC}
 S2, S3 and S4 %Modified for build without supplementaries
@@
 For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; Table ~\ref{Tab_Results-Difference_methods}).
-However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and Appendix Fig.%~\ref{Fig_Supp_BootBayt_allparam}
+However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and online Appendix Fig.%~\ref{Fig_Supp_BootBayt_allparam}
 S2 %Modified for build without supplementaries

### 2015-01-26 Manuscript/TEM_draft.tex
Edits to abstract and title page

--- 2015-01-25 Manuscript/TEM_draft.tex
@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$, and Natalie Cooper$^1$$^,$$^2$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$$^*$, and Natalie Cooper$^1$$^,$$^2$}\\
 \noindent {\small \it 
-$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland;\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland;}\\
+$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
+$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.}\\
 \end{center}
 \medskip
-\noindent{\bf Corresponding author:} Thomas Guillerme, School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.\\
+\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\
 \vspace{1in}
@@
 \begin{abstract}
-Living species represent only a small proportion of all species that have ever lived.
-Ignoring fossil taxa may therefore lead to misinterpretation of macroevolutionary patterns and processes such as trends in species richness, biogeographical history or paleoecology.
-This fact has led to an increasing consensus among scientists that both living and fossil taxa must be included in macroevolutionary studies.
-One approach called Total Evidence, uses molecular data from living taxa and morphological data from both living and fossil taxa to infer phylogenies with both living and fossil taxa at the tips.
-Although the Total Evidence approach seems very promising, it requires a lot of data and is therefore likely to suffer from missing data issues which may affect its ability to infer correct phylogenies.
+To fully understand macroevolutionary patterns and processes, we need to include both extant and extinct species in our models.
+This requires phylogenetic trees with both living and fossil taxa at the tips.
+One way to infer such phylogenies is the Total Evidence approach which uses molecular data from living taxa and morphological data from living and fossil taxa.
+Although the Total Evidence approach is very promising, it requires a lot of data and is therefore likely to suffer from missing data issues that may affect its ability to infer correct phylogenies.
 
-In this study we assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
-Using simulations we investigate three major factors that directly affect the completeness of the morphological part of the matrix: (1) the proportion of living taxa with no morphological data, (2) the amount of missing data in the fossil record, and (3) the overall number of morphological characters in the matrix.
+Here we use simulations to assess the effects of missing data on tree topologies inferred from Total Evidence matrices.
+We investigate three major factors that directly affect the completeness of the morphological part of the matrix: the proportion of living taxa with no morphological data, the amount of missing data in the fossil record, and the overall number of morphological characters in the matrix.
+We infer phylogenies from complete matrices and from matrices with various amounts of missing data, and then compare missing data topologies to the "best" tree topology inferred using the complete matrix.
 
-% NC: This needs changing to reflect what we decided to focus on, i.e.
-our recommendations.
-We don't want to lead with the methods thing.
-The two previous paragraphs are fine and just need tightening up a bit
-%We find that, when using a clade conservative metric such as Robinson-Foulds distance, Bayesian consensus trees recovers the right topology better than Maximum Likelihood or than a subset of Bayesian posterior tree distributions and this regardless of the amount of missing data (minimal tree similarity of 0.69 in the worst scenario).
-Regarding the data, good topological recovery is not related to the amount of missing data \textit{per se} but to the amount of data overlap.
-The two factors that influences the most this data overlap are the parameters 1 (proportion of coded living taxa) and 3 (number of morphological characters).
-
-We find that increasing the number of living taxa with morphological characters (1) and the overall amount of morphological data (3) have a greater impact on the ability to recover the "correct" topology than increasing the data completeness in the fossil record (2).
-Therefore, we suggest that sampling effort should be focusing on these two first aspects to increase the accuracy of topology in the Total Evidence framework.
-Additionally, we found that the Bayesian consensus tree consistently recovers topology that are closest to the "correct" topology than any other method.
-Therefore, we also suggest, after the data collection effort, that one should fix the tree topology using the Bayesian consensus tree topology for any further tree inference or macroevolutionary analysis.
+We find that the number of living taxa with morphological characters and the number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology.
+Therefore, we suggest that sampling effort should be focused on morphological data collection for living species to increase the accuracy of topological inference in a Total Evidence framework.
+Additionally, we find that Bayesian methods consistently outperform other tree inference methods.
+We therefore recommend using Bayesian consensus trees to fix the tree topology prior to further analyses such as Tip-Dating.
 

### 2015-01-26 Manuscript/TEM_draft.tex
Fixed conflict

--- 2015-01-26 Manuscript/TEM_draft.tex
@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$, and Natalie Cooper$^1$$^,$$^2$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$$^*$, and Natalie Cooper$^1$$^,$$^2$}\\
 \noindent {\small \it 
-$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland;\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland;}\\
+$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
+$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.}\\
 \end{center}
 \medskip
-\noindent{\bf Corresponding author:} \textit{Thomas Guillerme, School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\
+\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\
 \vspace{1in}

### 2015-01-27 Manuscript/CoverLetter.tex
Cover letter update
Updated cite: Jackson -> Quental (more recent)
Fixed capital letters for Tip-Dating and abstract

@@
 In recent years there has been growing interest in building phylogenies that contain both living and fossil taxa (e.g.
-Fritz et al.
+Quental and Marshall 2006 TREE; Fritz et al.
 2013 TREE; Heath et al.
@@
 This has major implications for clades where detailed morphological data collection from living species is rare.
-Additionally, we recommend using a Bayesian majority consensus tree when fixing tree topology for any additional analyses such as Tip-Dating (Ronquist et al.
+Additionally, we recommend using a Bayesian majority consensus tree when fixing tree topology for any additional analyses such as tip-dating (Ronquist et al.
 2012 Syst Biol).
@@
 2014; Pattinson et al.
-in press, Pyron 2011, Ronquist et al.
-2012) % Any more?
-Without going on to a second page?
-, have been recently published in the journal.
+in press, Sylvestro et al.
+2014, Pyron 2011, Ronquist et al.
+2012), have been recently published in the journal.
 

### 2015-01-27 Manuscript/TEM_draft.tex
Fixed capital letters for Tip-Dating and abstract

@@
 One way to infer such phylogenies is the Total Evidence approach which uses molecular data from living taxa and morphological data from living and fossil taxa.
-Although the Total Evidence approach is very promising, it requires a lot of data and is therefore likely to suffer from missing data issues that may affect its ability to infer correct phylogenies.
+%V1
+Although the Total Evidence approach is very promising, it requires many data and is therefore likely to suffer from missing data issues that may affect its ability to infer correct phylogenies.
+%V2
+Although the Total Evidence approach is very promising, morphological characters are often limited and missing, therefore this method is likely to suffer from missing data issues that may affect its ability to infer correct phylogenies.
 
-Here we use simulations to assess the effects of missing data on tree topologies inferred from Total Evidence matrices.
-We investigate three major factors that directly affect the completeness of the morphological part of the matrix: the proportion of living taxa with no morphological data, the amount of missing data in the fossil record, and the overall number of morphological characters in the matrix.
+Here we use simulations to assess the effects of missing data on tree topologies inferred from Total Evidence matrices. We investigate three major factors that directly affect the completeness 
+%How does the overall number of characters affect completeness of the matrix?
+and the size
+of the morphological part of the matrix: the proportion of living taxa with no morphological data, the amount of missing data in the fossil record, and the overall number of morphological characters in the matrix.
 We infer phylogenies from complete matrices and from matrices with various amounts of missing data, and then compare missing data topologies to the "best" tree topology inferred using the complete matrix.
 
-We find that the number of living taxa with morphological characters and the number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology.
+We find that the number of living taxa with morphological characters and the overall number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology
+%Nick's comment
+for a given number of missing characters.
 Therefore, we suggest that sampling effort should be focused on morphological data collection for living species to increase the accuracy of topological inference in a Total Evidence framework.
 Additionally, we find that Bayesian methods consistently outperform other tree inference methods.
-We therefore recommend using Bayesian consensus trees to fix the tree topology prior to further analyses such as Tip-Dating.
+We therefore recommend using Bayesian consensus trees to fix the tree topology prior to further analyses such as tip-dating.
+%Hummm Nick's comment again - That's a point to make in the discussion maybe. But let's wait for Gavin comments...
+%it's possible that adding dating information etc.
+in tip-dating makes Bayesian dated total-evidence analyses (ie tip-dating) better than non-dated Bayesian total-evidence analyses (typical total-evidence approach), because more information is being used, and because dates give some information about phylogenetic position.
+In any case, the study wasn't about tip-dating
 
@@
 A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \citep{eernissetaxonomic1993}.
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as Tip-Dating; \citealt{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
+This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as tip-dating; \citealt{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
 Total Evidence methods have been successfully applied to empirical data \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
@@
 Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in Tip-Dating \citep{ronquista2012,BEASTmaster}.
+Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \citep{ronquista2012,BEASTmaster}.
 Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see discussion above).
@@
 Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the "best" tree topology than the Maximum Likelihood trees.
-Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed in a Total Evidence analysis, for example when using the Tip-Dating method.
+Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed in a Total Evidence analysis, for example when using the tip-dating method.
 The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.

### 2015-01-27 Manuscript/TEM_draft.tex
Added Gavin's pdf comments

@@
 
+%Gavin's general comment:I have made some comments on Thomas’ paper.
+I think it generally looks good.
+The one thing that I think might be tricky to get past reviewers is the comparison between the ML and Bayesian trees.
+The problem is that you can’t tell if the poorer performance of ML is because of the inference method or the difference between the substitution models - the ML uses an overparameterised model (GTR) compared to the generating model, whereas the Bayesian analyses do not.
+It might be possible to cite your way out of this but I wouldn’t be surprised if a reviewer asks for new analyses by either using different ML software that can handle the HKY model (e.g.
+Garli) or running the Bayesian with the GTR model (although that might not be adequate because if I remember correctly RAXML uses an approximation for gamma that is not the same as in other software).
+
 %---------------------------------------------
@@
 However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data that can be difficult (or impossible) to collect.
+%G: Why does it need so much data? Is there any evidence (a citation or two) to back this up?
  The morphological data for living taxa are rarely collected when molecular data are available (e.g.
@@
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case) \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case)
+%G: Might be worthing reading and citing this: http://www.sciencemag.org/content/333/6041/448.full. Missing data can be a problem if it renders some subtrees unresolvable.
+ \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
@@
 Since the molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}, the effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
+%G: Alan Lemmon has done a fair amount of work on missingness - worth checking out the work from his lab, e.g. http://www.evotutor.org/LemmonLab/Pdfs/Lemmon-etal2009.pdf
  Therefore, we focus only on missing data in the morphological part of the matrix.
@@
 We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
-We implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa to ensure that we had enough taxa of each type for our missing data simulations to work.
+We implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa
+%G:I don’t think there is a problem with this split but I expect reviewers might ask for justification. Is this typical of existing TE studies?
+to ensure that we had enough taxa of each type for our missing data simulations to work.
 We then added an outgroup to the tree, using the mean branch length of the tree to separate the outgroup from the rest of the taxa, and with the branch length leading to the outgroup set as the sum of the mean branch length and the longest root-to-tip length of the tree.
@@
 \subsubsection{Maximum Likelihood}
-The "best" tree and the "missing-data" trees were inferred using RAxML v.
-8.0.20 \citep{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v.
+The "best" tree and the "missing-data" trees were inferred using RAxML v. 8.0.20 \citep{Stamatakis21012014}. For the molecular data, we used the GTR
+%G: I think you need to justify using the GTR model (and so also RAXML).
+The data were simulated under an HKY model so you are fitting an overparameterised model.
+It might be that the better results for the MrBayes analyses come down to model misspecification.
+You can certainly say that HKY is not an option in RAXML but then you might be asked why you don’t use another program, e.g Garli.
+ + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v.
 8.0.20; \citealt{Stamatakis21012014}).
@@
 
-To measure the phylogenetic signal of our simulations, we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
+To measure the phylogenetic signal
+%G: Do you mean you tested whether the simulated data generally had any phylogenetic signal with which to recover a tree?
+I was a bit unsure but mainly because I am hardwired to thinks of Pagel’s lambda when anyone mentions phylogenetic signal…
+ of our simulations, we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
 We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
@@
 We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
-We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa  (i.e.
+We used the Robinson-Foulds distance \citep{RF1981}
+%G: Why these two metrics? Might need a few sentences to justify them in addition to the explanation of what they are that you already have.
+to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa  (i.e.
 taxa that frequently change position in different trees; \citealt{kearneyfragmentary2002}).
@@
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; online Appendix Tables %~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood
+%G: This might be expected given the GTR rather than HKY model is used. 
+ trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; online Appendix Tables %~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}
 S5, S6 and S7 %Modified for build without supplementaries
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
+\caption{
+%G: I think this table needs to be decoded a bit.
+I wasn’t really sure what it was supposed to be getting across.
+Maybe it is because I am not familiar with the Bhatt… distance but I suspect that many readers will be similarly unfamiliar.
+Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
 The values in bold are the extreme values of high or low probability of overlap between two methods.}
@@
 \begin{figure}
-\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
+\caption{The effects of increasing missing data 
+%G: How does tree resolution influence the tree comparison metrics?
+Are the Bayesian trees always fully resolved?
+Could that bias towards shorter distances to the best tree (because there is reduced probability of conflict).
+on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
 The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
@@
 Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
+The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.
+%G: Is there a scale to the colour or is it binary? Also, I don’t get the axes. Am I missing something or are they the same?
+}
 \label{Fig_Results-paircomp_within}

### 2015-01-28 Manuscript/TEM_draft.tex
Dealt with comments

@@
 Total Evidence methods have been successfully applied to empirical data \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
-However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires a lot of data that can be difficult (or impossible) to collect.
-%G: Why does it need so much data? Is there any evidence (a citation or two) to back this up?
- The morphological data for living taxa are rarely collected when molecular data are available (e.g.
-\citealt{O'Leary08022013} \textit{vs.} \citealt{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record (for example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts; \citealt{sansomfossilization2013}).
+%G: Why does it need so much data? Is there any evidence (a citation or two) to back this up? TG: added clarity in both sentences below.
+However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological of data, both of which can be difficult (or impossible) to collect for every focal living and fossil taxa.
+For example, the morphological data for living taxa are rarely collected when molecular data are available (e.g.
+\citealt{O'Leary08022013} \textit{vs.} \citealt{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record (for example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts; \citealt{sansomfossilization2013}) and molecular data is (nearly) always unavailable.
 Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
@@
 Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case)
-%G: Might be worthing reading and citing this: http://www.sciencemag.org/content/333/6041/448.full. Missing data can be a problem if it renders some subtrees unresolvable.
- \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+%G: Might be worthing reading and citing this: http://www.sciencemag.org/content/333/6041/448.full.
+Missing data can be a problem if it renders some subtrees unresolvable.
+TG: added citation in the whole bunch.
+ \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,Sanderson22072011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
@@
 Here we use a simulation approach to assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
-Since the molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}, the effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,rouresite-specific2011}.
-%G: Alan Lemmon has done a fair amount of work on missingness - worth checking out the work from his lab, e.g. http://www.evotutor.org/LemmonLab/Pdfs/Lemmon-etal2009.pdf
+Since the molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}, the effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011}.
+%G: Alan Lemmon has done a fair amount of work on missingness - worth checking out the work from his lab, e.g.
+http://www.evotutor.org/LemmonLab/Pdfs/Lemmon-etal2009.pdf TG: added reference (that was actually in the 2014 draft)
  Therefore, we focus only on missing data in the morphological part of the matrix.
-We investigate three major parameters that directly affect the completeness of the morphological part of the matrix: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix.
-These parameters reflect empirical biases in data availability
-% NC: The next few sentences could be moved to the methods. They feel a bit clunky here.
-- the advent of molecular phylogenetics means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis, missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
+We investigate three major parameters that directly affect the completeness of the morphological part of the matrix: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix (i.e.
+the size of the matrix).
+These parameters reflect empirical biases in data availability.
+% NC: The next few sentences could be moved to the methods. They feel a bit clunky here. TG: I like the way they end the intro: "we're simulating stuff but it links to more real stuff in reality."
+The advent of molecular the molecular clock \citep{zuckerkandl1965} means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis, missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
@@
 We implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa
-%G:I don’t think there is a problem with this split but I expect reviewers might ask for justification. Is this typical of existing TE studies?
-to ensure that we had enough taxa of each type for our missing data simulations to work.
+%G:I don’t think there is a problem with this split but I expect reviewers might ask for justification. Is this typical of existing TE studies? TG: avoiding bias comment.
+to ensure that we had an equal amount of taxa of each type to avoid any bias toward a certain type of taxa (i.e.
+fossil or living).
 We then added an outgroup to the tree, using the mean branch length of the tree to separate the outgroup from the rest of the taxa, and with the branch length leading to the outgroup set as the sum of the mean branch length and the longest root-to-tip length of the tree.
@@
 \subsubsection{Maximum Likelihood}
-The "best" tree and the "missing-data" trees were inferred using RAxML v. 8.0.20 \citep{Stamatakis21012014}. For the molecular data, we used the GTR
+The "best" tree and the "missing-data" trees were inferred using RAxML v.
+8.0.20 \citep{Stamatakis21012014}.
+For the molecular data, we used the GTR + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v.
+8.0.20; \citealt{Stamatakis21012014}).
+For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
+8.0.20; \citealt{Stamatakis21012014}).
 %G: I think you need to justify using the GTR model (and so also RAXML).
@@
 You can certainly say that HKY is not an option in RAXML but then you might be asked why you don’t use another program, e.g Garli.
- + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v.
-8.0.20; \citealt{Stamatakis21012014}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
-8.0.20; \citealt{Stamatakis21012014}).
+We used RAxML because this software it is heavily used in literature for Maximum Likelihood tree inference \citep[][e.g.]{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014} and is one of the fastest available \citep{Stamatakis01102008}.However, this software has no option for using the HKY model and therefore, by using the GTR model we fitted an overparameterised model (see Discussion).
 
-To measure the phylogenetic signal
+To measure the strength of the phylogenetic signal of our simulations (i.e. the support for each branch in the phylogeny),
 %G: Do you mean you tested whether the simulated data generally had any phylogenetic signal with which to recover a tree?
 I was a bit unsure but mainly because I am hardwired to thinks of Pagel’s lambda when anyone mentions phylogenetic signal…
- of our simulations, we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
+we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
 We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
@@
 We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
-We used the Robinson-Foulds distance \citep{RF1981}
+We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa (i.e.
+taxa that frequently change position in different trees; \citealt{kearneyfragmentary2002}).
 %G: Why these two metrics? Might need a few sentences to justify them in addition to the explanation of what they are that you already have.
-to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa  (i.e.
-taxa that frequently change position in different trees; \citealt{kearneyfragmentary2002}).
-We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
-These metrics are described in detail below.
+We used these two metrics because they are widely used in literature \citep{Bogdanowicz2012}, they relative performance \citep{Kuhner04112014} and because they illustrate two different aspect of tree topology (see Discussion).
+We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa. These metrics are described in detail below.
 
@@
 Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
-Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam}, Fig.~\ref{Fig_Results-paircomp_within}, Table 1).
+Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam}, Fig.~\ref{Fig_Results-paircomp_within}, Table ~\ref{Tab_Results-Difference_methods}).
 As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
@@
 
+%^G: I think you need to justify using the GTR model (and so also RAXML).
+The data were simulated under an HKY model so you are fitting an overparameterised model.
+It might be that the better results for the MrBayes analyses come down to model misspecification.
+You can certainly say that HKY is not an option in RAXML but then you might be asked why you don’t use another program, e.g Garli.
+The Maximum Likelihood analysis model misspecification (GTR instead of HKY; see Methods) could as well explain theses differences between the Maximum Likelihood tree topologies and the Bayesian consensus tree topologies.
+However, the GTR model is a generalisation of the HKY model: a GTR model where the $\alpha$ and $\eta$ parameters are equal to the HKY $k$ parameter and when the 4 other parameters ($\beta$, $\gamma$, $\delta$ and $\epsilon$) are equal to one is effectively a HKY model with a lower degree of freedom.
+Also, the effect of overparametrisation is higher in Bayesian analysis than in Maximum Likelihood ones (see \citealt{Buckley01052002} and discussion in \citealt{douadycomparison2003}).
+Additionally, it is noticeable that the difference of the Maximum Likelihood tree topologies and the Bayesian tree topologies that could be due to model overparametrisation is not consistent with the results of the Maxmimum Likelihood bootstrap trees topologies and the Bayesian posterior tree distributions topologies that are similar (Table ~\ref{Tab_Results-Difference_methods}).
+Therefore we argue that the differences between the Maximum Likelihood tree topologies and the Bayesian consensus tree topologies is more likely due to the way the Bayesian consensus tree is built (see above) than to model misspecification.
+%$
+
 \subsubsection{Missing morphological characters ($M_{C}$)}
@@
 Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{wrightbayesian2014}.
-% NC: Missing a citation here.
-Also is this correct - increase and increase?
-TG: yep.
-Checking for a better cite though.
 However, the use of probabilistic methods (i.e.
@@
 Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \citep{ronquista2012,BEASTmaster}.
+%^Nick
+%it's possible that adding dating information etc.
+in tip-dating makes Bayesian dated total-evidence analyses (ie tip-dating) better than non-dated Bayesian total-evidence analyses (typical total-evidence approach), because more information is being used, and because dates give some information about phylogenetic position.
+In any case, the study wasn't about tip-dating
+However, also adding dating information could directly affect the ability to recover the "best" tree topology since a data Total Evidence analysis (i.e.
+tip-dating) allows to use the dates as information about phylogenetic position of the fossil clades.
+%$
 Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see discussion above).

### 2015-01-28 Manuscript/SupplementaryMaterial/Supp_results.tex
Dealt with figures comments

@@
     \caption{The effects of missing data on topological recovery using Maximum Likelihood trees.
-Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
 Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
+The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
 \label{Fig_Supp_paircomp_ML}
@@
     \caption{The effects of missing data on topological recovery using Maximum Likelihood bootstrap trees.
-Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
 Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
+The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
 \label{Fig_Supp_paircomp_Boot}
@@
     \caption{The effects of missing data on topological recovery using Bayesian posterior tree distribution.
-Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
 Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
+TThe Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
 \label{Fig_Supp_paircomp_Bayt}

### 2015-01-28 Manuscript/TEM_draft.tex
Dealt with figures comments

@@
 
+%G: How does tree resolution influence the tree comparison metrics?
+Are the Bayesian trees always fully resolved?
+Could that bias towards shorter distances to the best tree (because there is reduced probability of conflict).
+TG: This comment is kind of linking to the following.
 The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds distance) but greater on the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 Maybe it is because I am not familiar with the Bhatt… distance but I suspect that many readers will be similarly unfamiliar.
+TG: fixed clarity
 Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
-The values in bold are the extreme values of high or low probability of overlap between two methods.}
+Each lines summarize the distribution of the probability of overlap between pairs of tree inference method.
+The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.}
 \centering
@@
 \begin{figure}
-\caption{The effects of increasing missing data 
-%G: How does tree resolution influence the tree comparison metrics?
-Are the Bayesian trees always fully resolved?
-Could that bias towards shorter distances to the best tree (because there is reduced probability of conflict).
-on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
+\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
 The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
@@
 \caption{The effects of missing data on topological recovery using Bayesian consensus trees.
-Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+The x and the y axes both show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
 Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.
-%G: Is there a scale to the colour or is it binary? Also, I don’t get the axes. Am I missing something or are they the same?
+The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
+%G: Is there a scale to the colour or is it binary? Also, I don’t get the axes. Am I missing something or are they the same? TG: fixed clarity.
 }

### 2015-01-28 Manuscript/TEM_draft.tex
Updated figures (bhattacharyya split in two)

@@
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc.
-(see Fig.~\ref{Fig_Bhattacharyya_Coefficients} for more details).
+(see Fig.~\ref{Fig_Bhattacharyya_Coefficients1} for more details).
 This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
@@
 Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc.
-(see Fig.~\ref{Fig_Bhattacharyya_Coefficients} for more details).
+(see Fig.~\ref{Fig_Bhattacharyya_Coefficients2} for more details).
 Note that this procedure pools results from across all missing data parameter combinations so it results in just six pairwise comparisons.
@@
 $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two combination of missing data parameters such as BC(A,B) is the probability of overlap between the distributions A and B.}
+\label{Fig_Bhattacharyya_Coefficients1} %Bhattacharyya coefficients calculation outline
+\end{figure}
+
+\begin{figure}
+\caption{Bhattacharyya Coefficient calculation outline.
+A and B are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
 \textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
 Maximum Likelihood or Bayesian).
-1: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two combination of missing data parameters.
-2: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters.}
-\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters such as BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A between the methods $x$ and $y$.}
+\label{Fig_Bhattacharyya_Coefficients2} %Bhattacharyya coefficients calculation outline
 \end{figure}
@@
 %---------------------------------------------
+
+\setcounter{figure}{0} \renewcommand{\thefigure}{\arabic{figure}}
+
 \newpage
-
-\begin{figure}[H]
+\begin{figure}[!htbp]
 \centering
     \includegraphics[keepaspectratio=true]{Figures/In_main/Simulations_outline-BW.pdf}
-%\caption{Protocol outline.
+\caption{ }
 %(1) We randomly generated a birth-death tree (the "true" tree) and used it to infer a matrix with no missing data (the "complete" matrix).
@@
 \end{figure}
-Figure 1
 
 \newpage
-
-\begin{figure}[H]
+\begin{figure}[!htbp]
 \centering
-    \includegraphics[width=1\textwidth]{Figures/In_main/Bhattacharyya_Coefficients.pdf}
+    \includegraphics[width=0.5\textwidth]{Figures/In_main/Bhattacharyya_Coefficients1.pdf}
+\caption{ }
 %\caption{Bhattacharyya Coefficient calculation outline.
@@
 $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two combination of missing data parameters such as BC(A,B) is the probability of overlap between the distributions A and B.}
+%\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
+\end{figure}
+
+\newpage
+\begin{figure}[!htbp]
+\centering
+    \includegraphics[width=1\textwidth]{Figures/In_main/Bhattacharyya_Coefficients2.pdf}
+\caption{ }
+%\caption{Bhattacharyya Coefficient calculation outline.
+A and B are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
 \textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
 Maximum Likelihood or Bayesian).
-1: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two combination of missing data parameters.
-2: the Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters.}
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters such as BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A between the methods $x$ and $y$.}
 %\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
 \end{figure}
-Figure 2
 
 \newpage
-
-\begin{figure} 
+\begin{figure}[!htbp]
 \centering
     \includegraphics[width=1\textwidth]{Figures/In_main/AllMethods-RF+Tr-colour.pdf}
+\caption{ }
 %\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).
@@
 \end{figure}
-Figure 3
 
 \newpage
-
-\begin{figure}[H]
+\begin{figure}[!htbp]
 \centering
     \includegraphics[width=1\textwidth]{Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
+\caption{ }
 %\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
@@
 \end{figure}
-Figure 4
 
 \newpage
-
-\begin{figure}[H]
+\begin{figure}
 \centering
@@
     \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+\caption{ }
 %\caption{The effects of missing data on topological recovery using Bayesian consensus trees.
@@
 \end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
-Figure 5
 %END

### 2015-01-28 Manuscript/TEM_draft.tex
Supplementaries figure typo

@@
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see \hyperref[SupplementaryMaterial]{online Appendix 1} and Fig.\ref{Fig_AppendixCharacters}).
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see \hyperref[SupplementaryMaterial]{online Appendix 1} and Fig.
+S1).
 We then ran an independent discrete character simulation for each character using the "true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.

### 2015-01-29 Manuscript/TEM_draft.tex
Comments on the 'final' edits

@@
 One way to infer such phylogenies is the Total Evidence approach which uses molecular data from living taxa and morphological data from living and fossil taxa.
-%V1
-Although the Total Evidence approach is very promising, it requires many data and is therefore likely to suffer from missing data issues that may affect its ability to infer correct phylogenies.
-%V2
-Although the Total Evidence approach is very promising, morphological characters are often limited and missing, therefore this method is likely to suffer from missing data issues that may affect its ability to infer correct phylogenies.
 
-Here we use simulations to assess the effects of missing data on tree topologies inferred from Total Evidence matrices. We investigate three major factors that directly affect the completeness 
-%How does the overall number of characters affect completeness of the matrix?
-and the size
-of the morphological part of the matrix: the proportion of living taxa with no morphological data, the amount of missing data in the fossil record, and the overall number of morphological characters in the matrix.
+Although the Total Evidence approach is very promising, it requires a great deal of data that can be hard to collect.
+Therefore this method is likely to suffer from missing data issues that may affect its ability to infer correct phylogenies.
+
+Here we use simulations to assess the effects of missing data on tree topologies inferred from Total Evidence matrices.
+We investigate three major factors that directly affect the completeness and the size of the morphological part of the matrix: the proportion of living taxa with no morphological data, the amount of missing data in the fossil record, and the overall number of morphological characters in the matrix.
 We infer phylogenies from complete matrices and from matrices with various amounts of missing data, and then compare missing data topologies to the "best" tree topology inferred using the complete matrix.
 
-We find that the number of living taxa with morphological characters and the overall number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology
-%Nick's comment
-for a given number of missing characters.
+We find that the number of living taxa with morphological characters and the overall number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology.
+% NC: I think you misunderstood Nick's comment here.
+He meant that living > fossil but that you can't put the number of characters in here too.
+I've reverted back to the original as I'm not sure his solution is correct
 Therefore, we suggest that sampling effort should be focused on morphological data collection for living species to increase the accuracy of topological inference in a Total Evidence framework.
 Additionally, we find that Bayesian methods consistently outperform other tree inference methods.
-We therefore recommend using Bayesian consensus trees to fix the tree topology prior to further analyses such as tip-dating.
+We therefore recommend using Bayesian consensus trees to fix the tree topology prior to further analyses.
+
 %Hummm Nick's comment again - That's a point to make in the discussion maybe. But let's wait for Gavin comments...
@@
 In any case, the study wasn't about tip-dating
+% NC: Simplest thing here is to just leave out the mention of tip dating. Will need to fix in discussion
+% NC: Also Nick makes another sensible comment about *why* are living species more important. Check this is covered in discussion
 
@@
 %G: Why does it need so much data? Is there any evidence (a citation or two) to back this up? TG: added clarity in both sentences below.
-However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological of data, both of which can be difficult (or impossible) to collect for every focal living and fossil taxa.
-For example, the morphological data for living taxa are rarely collected when molecular data are available (e.g.
+However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
+Morphological data for living taxa are rarely collected when molecular data are available (e.g.
 \citealt{O'Leary08022013} \textit{vs.} \citealt{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record (for example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts; \citealt{sansomfossilization2013}) and molecular data is (nearly) always unavailable.
@@
 TG: added citation in the whole bunch.
+% NC: Did you actually *read* this paper? It makes the point that it *is* a problem if it affects certain subtrees. Might need to add a small caveat into the bit in brackets above...
  \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,Sanderson22072011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
@@
 http://www.evotutor.org/LemmonLab/Pdfs/Lemmon-etal2009.pdf TG: added reference (that was actually in the 2014 draft)
+% NC: OK.
+But be careful here.
+When people give you comments you want to take them seriously and actually read the papers they suggest rather than just adding references.
+Are you sure Lemmon is just looking at molecular stuff?
+Think Gavin's point might be that he's done stuff on missingness in morphological data too which at present we don't really cite or cover at all.
  Therefore, we focus only on missing data in the morphological part of the matrix.
-We investigate three major parameters that directly affect the completeness of the morphological part of the matrix: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix (i.e.
+We investigate three major parameters that directly affect the completeness and size of the morphological part of the matrix: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix (i.e.
 the size of the matrix).
-These parameters reflect empirical biases in data availability.
+These parameters reflect empirical biases in data availability as follows: (i)
 % NC: The next few sentences could be moved to the methods. They feel a bit clunky here. TG: I like the way they end the intro: "we're simulating stuff but it links to more real stuff in reality."
-The advent of molecular the molecular clock \citep{zuckerkandl1965} means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis, missing data in fossil taxa is very common due to preservation biases \citep{sansomfossilization2013}, and the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
+% NC: OK but hey currently don't make sense in terms of what was said perviosuly so I've numbered them.
+Still don't like this.
+It really should be integrated with the parameters.
+Basically it's poorly structured and doesn't follow very well - that's what I mean by "clunky"
+the increased availability of molecular data
+% NC: Not really the molecular clock, actually more the ease of extracting DNA. 
+means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis; (ii) missing data in fossil taxa is common due to preservation biases \citep{sansomfossilization2013}; and (iii) the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
@@
 %G:I don’t think there is a problem with this split but I expect reviewers might ask for justification. Is this typical of existing TE studies? TG: avoiding bias comment.
-to ensure that we had an equal amount of taxa of each type to avoid any bias toward a certain type of taxa (i.e.
-fossil or living).
+% NC: I think you misunderstood Gavin's point.
+He knows why you did it like this, but if usually TE studies are 95% living and 5% fossil (for example) then your results aren't really viable in real world.
+He ideally wants a statement like: this is a typical split between fossil and living tax across empirical TE studies (cite papers).
+However, I doubt this is the case.
+Actually aren't TE studeis usually dominated by fossil taxa?
+to ensure that we had enough taxa of each type for our simulations to work, and to avoid any bias toward fossil or living taxa.
 We then added an outgroup to the tree, using the mean branch length of the tree to separate the outgroup from the rest of the taxa, and with the branch length leading to the outgroup set as the sum of the mean branch length and the longest root-to-tip length of the tree.
@@
 You can certainly say that HKY is not an option in RAXML but then you might be asked why you don’t use another program, e.g Garli.
-We used RAxML because this software it is heavily used in literature for Maximum Likelihood tree inference \citep[][e.g.]{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014} and is one of the fastest available \citep{Stamatakis01102008}.However, this software has no option for using the HKY model and therefore, by using the GTR model we fitted an overparameterised model (see Discussion).
+We used RAxML because it is heavily used in the literature for Maximum Likelihood tree inference \citep[][e.g.]{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014} and is one of the fastest methods available \citep{Stamatakis01102008}.
+However, RAxML has no option for using the HKY model so we used the GTR model instead (see Discussion).
+% NC: This isn't really a justification. Also by pointing out the problem you're drawing attention to it. I actually never noticed this when reading the previous drafts. But I've never used RAxML.
 
-To measure the strength of the phylogenetic signal of our simulations (i.e. the support for each branch in the phylogeny),
+To measure the support for each branch in our simulated phylogenies
 %G: Do you mean you tested whether the simulated data generally had any phylogenetic signal with which to recover a tree?
 I was a bit unsure but mainly because I am hardwired to thinks of Pagel’s lambda when anyone mentions phylogenetic signal…
+% NC: Haha that is exactly what I've said about a million times to you.
 we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
@@
 %G: Why these two metrics? Might need a few sentences to justify them in addition to the explanation of what they are that you already have.
-We used these two metrics because they are widely used in literature \citep{Bogdanowicz2012}, they relative performance \citep{Kuhner04112014} and because they illustrate two different aspect of tree topology (see Discussion).
+We used these metrics because they are widely used in literature \citep{Bogdanowicz2012}, their relative performance is well described \citep{Kuhner04112014} % NC: Is that what you mean.
+"they relative performance" means nothing to me.
+and because they illustrate two different aspects of tree topology (see Discussion).
 We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa. These metrics are described in detail below.
@@
 TG: This comment is kind of linking to the following.
+
+% NC: Yes BUT the point is that Gavin misunderstood so you need to be clearer!
+Add a comment earlier that points out that the Bayesian consensus trees are not fully resolved (if this is true).
+If they are fully resolved more often than the ML trees you may need t oadd a paragraph to the discussion as Gavin is right, this could explain the result.
 The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds distance) but greater on the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 You can certainly say that HKY is not an option in RAXML but then you might be asked why you don’t use another program, e.g Garli.
-The Maximum Likelihood analysis model misspecification (GTR instead of HKY; see Methods) could as well explain theses differences between the Maximum Likelihood tree topologies and the Bayesian consensus tree topologies.
+The Maximum Likelihood analyses used the GTR rather than the HKY model, because HKY is not an option in RAxML.
+This misspecification of the Maximum Likelihood model could also explain the differences between the Maximum Likelihood and Bayesian consensus tree topologies.
 However, the GTR model is a generalisation of the HKY model: a GTR model where the $\alpha$ and $\eta$ parameters are equal to the HKY $k$ parameter and when the 4 other parameters ($\beta$, $\gamma$, $\delta$ and $\epsilon$) are equal to one is effectively a HKY model with a lower degree of freedom.
 Also, the effect of overparametrisation is higher in Bayesian analysis than in Maximum Likelihood ones (see \citealt{Buckley01052002} and discussion in \citealt{douadycomparison2003}).
+% NC: OK *none* of this is a justification.
+More the opposite!
+We know GTR is a generalisation of HKY, that doesn't make it better.
+That's basically what we mean when we say the ML model is overparameterised because even if all parameters = 1 it's a HKY but with fewer df.
+Thus over parameterised!
+Then yes, this would be a bigger issue for Bayesian - but so what?
+It's not an issue in Bayesian here.
+It's not like you had to choose to do it for one or the other, thus ML was a better choice.
+So this isn't a justification, it's just a series of factual statements.
+
 Additionally, it is noticeable that the difference of the Maximum Likelihood tree topologies and the Bayesian tree topologies that could be due to model overparametrisation is not consistent with the results of the Maxmimum Likelihood bootstrap trees topologies and the Bayesian posterior tree distributions topologies that are similar (Table ~\ref{Tab_Results-Difference_methods}).
 Therefore we argue that the differences between the Maximum Likelihood tree topologies and the Bayesian consensus tree topologies is more likely due to the way the Bayesian consensus tree is built (see above) than to model misspecification.
-%$
+% NC: Huh??? Not really following, nor is this a good argument.
 
@@
 Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \citep{ronquista2012,BEASTmaster}.
+Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating (\citealp{ronquista2012,BEASTmaster};
 %^Nick
@@
 In any case, the study wasn't about tip-dating
-However, also adding dating information could directly affect the ability to recover the "best" tree topology since a data Total Evidence analysis (i.e.
-tip-dating) allows to use the dates as information about phylogenetic position of the fossil clades.
-%$
-Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see discussion above).
+% NC: Note that Nick's point is that we can really only say what works best for topology only methods
+although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution).
+Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see above).
 Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses \citep[e.g.][]{jetzthe2012}.
@@
 Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the "best" tree topology than the Maximum Likelihood trees.
-Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed in a Total Evidence analysis, for example when using the tip-dating method.
+Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
 The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary processes through both time and space.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
+% NC: Why would it improve spatial methods? I've removed this.
 
@@
 \section{Acknowledgments}
-Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, and April Wright for useful comments on our simulation protocol.
+Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Nick Matzke, and April Wright for useful comments on our simulation protocol and manuscript.
 Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
@@
 TG: fixed clarity
-Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
-Each lines summarize the distribution of the probability of overlap between pairs of tree inference method.
-The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.}
+Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" trees using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
+Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
+The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
+If two methods have a high probability of overlap, they have a similar ability to recover the "correct" tree topology.
+% NC: check this is right, I think that's what Gavin wanted to be added - a way of interpreting what is going on
+}
 \centering
@@
 \caption{Bhattacharyya Coefficient calculation outline.
-A, B and C are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
 $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
-The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two combination of missing data parameters such as BC(A,B) is the probability of overlap between the distributions A and B.}
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.}
 \label{Fig_Bhattacharyya_Coefficients1} %Bhattacharyya coefficients calculation outline
@@
 \caption{Bhattacharyya Coefficient calculation outline.
-A and B are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
 $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
@@
 Maximum Likelihood or Bayesian).
-The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters such as BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A between the methods $x$ and $y$.}
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$.}
 \label{Fig_Bhattacharyya_Coefficients2} %Bhattacharyya coefficients calculation outline
 \end{figure}
+
+% NC: These are still really hard to follow but I'm at a loss as to how to explain them better.
 
@@
 \caption{The effects of missing data on topological recovery using Bayesian consensus trees.
-The x and the y axes both show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
 Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.

### 2015-01-29 Manuscript/TEM_draft.tex
Updated comments

@@
 % NC: Also Nick makes another sensible comment about *why* are living species more important. Check this is covered in discussion 
+%TG: I think that's covered in \subsection{Individual effects of missing data parameters} - \subsubsection{Missing data for living taxa ($M_{L}$)}. (If no living taxa, the fossils cannot branch.)
 
@@
 TG: added citation in the whole bunch.
-% NC: Did you actually *read* this paper? It makes the point that it *is* a problem if it affects certain subtrees. Might need to add a small caveat into the bit in brackets above...
- \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,Sanderson22072011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+% NC: Did you actually *read* this paper?
+It makes the point that it *is* a problem if it affects certain subtrees.
+Might need to add a small caveat into the bit in brackets above...
+TG: Ok, so this comment goes with the data overlap.
+I’m not entirely sure where the paper comes in.
+If the there is not enough overlap in the super matrix, then it is impossible to get a consensus clade tree (kind of supporter talked in the paper) per terrace?
+ \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
@@
 Think Gavin's point might be that he's done stuff on missingness in morphological data too which at present we don't really cite or cover at all.
+TG: I don’t like arguing on comments but this paper is about missing data in molecular matrices (still relevant though).
+The reason I removed it from the old version is that it states something kind of similar than all the others: overlap but it is specially important in bigger matrices where missing data can weaken the support in certain parts of the tree.
+The idea being probably close to the tree terrace paper.
 Therefore, we focus only on missing data in the morphological part of the matrix.
@@
 the size of the matrix).
-These parameters reflect empirical biases in data availability as follows: (i)
 % NC: The next few sentences could be moved to the methods. They feel a bit clunky here. TG: I like the way they end the intro: "we're simulating stuff but it links to more real stuff in reality."
@@
 It really should be integrated with the parameters.
-Basically it's poorly structured and doesn't follow very well - that's what I mean by "clunky"
-the increased availability of molecular data
-% NC: Not really the molecular clock, actually more the ease of extracting DNA. 
-means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis; (ii) missing data in fossil taxa is common due to preservation biases \citep{sansomfossilization2013}; and (iii) the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
+Basically it's poorly structured and doesn't follow very well - that's what I mean by "clunky"  - TG: You're right (as usual ;)) it works better without.
+%These parameters reflect empirical biases in data availability as follows: (i) the increased availability of molecular data means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis; (ii) missing data in fossil taxa is common due to preservation biases \citep{sansomfossilization2013}; and (iii) the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
-We find that the ability of Total Evidence methods to recover the "best" tree topology is increased when using Bayesian methods.
-Also minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree topology more so than minimizing the amount of missing data in the fossil record.
+%TG: I swapped the results to be consistent, ML and MC first and then Bayesian
+We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree topology more so than minimizing the amount of missing data in the fossil record.
+Additionally, we find that the ability of Total Evidence methods to recover the "best" tree topology is increased when using Bayesian methods.
 
@@
 We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
-We implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa
 %G:I don’t think there is a problem with this split but I expect reviewers might ask for justification. Is this typical of existing TE studies? TG: avoiding bias comment.
@@
 Actually aren't TE studeis usually dominated by fossil taxa?
-to ensure that we had enough taxa of each type for our simulations to work, and to avoid any bias toward fossil or living taxa.
+% TG: Really depends. In the 5 I keep citing, Pyron is close to 50/50, Ronquist and Slater are more towards living and Beck is more towards fossil
+Empirical Total Evidence matrices can be biased towards a greater number of fossils (e.g.
+88\% and 58\% fossils taxa in respectively \citealt{beckancient2014} and \citealt{schragocombining2013}) or towards more living taxa (e.g.
+69\% of living taxa in both \citealt{ronquista2012} and \citealt{slaterphylogenetic2013}) buy they can contain an even number of both type of taxa (e.g.
+52\% of living taxa in \citealt{pyrondivergence2011}).
+Therefore, to avoid biasing our simulations towards the two extreme situations (more living or fossil taxa) and to make each simulation comparable, we implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa.
 We then added an outgroup to the tree, using the mean branch length of the tree to separate the outgroup from the rest of the taxa, and with the branch length leading to the outgroup set as the sum of the mean branch length and the longest root-to-tip length of the tree.
@@
 
+%TG: this part was removed from the into: These parameters reflect empirical biases in data availability as follows: (i) the increased availability of molecular data means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis; (ii) missing data in fossil taxa is common due to preservation biases \citep{sansomfossilization2013}; and (iii) the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
+
 \begin{enumerate}
@@
 This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological part.
-This reflects the fact that detailed morphological data for living species are scarce.
+This reflects the fact that morphological data for living species are scarce.
+This can be due to the increased availability of molecular data and/or that few people have the skills to identify characters needed for detailed phylogenetic analysis.
 \item{$M_{F}$, the proportion of missing data in the fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%.}
@@
 This parameter illustrates the number of available morphological characters for both living and fossil taxa.
+This reflects the amount of effort put into identifying the characters \citep[e.g.][]{O'Leary08022013}.
 \end{enumerate}
@@
 % NC: Haha that is exactly what I've said about a million times to you.
+% TG: And after the literature review for the next paper (missing living in mammals) on community structure & Co. I start to get where you guys are comming for ;).
 we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
@@
 %G: Why these two metrics? Might need a few sentences to justify them in addition to the explanation of what they are that you already have.
-We used these metrics because they are widely used in literature \citep{Bogdanowicz2012}, their relative performance is well described \citep{Kuhner04112014} % NC: Is that what you mean.
+%We used these metrics because they are widely used in literature \citep[][e.g.]{Bogdanowicz2012}, their relative performance is well described \citep{Kuhner04112014} % NC: Is that what you mean.
 "they relative performance" means nothing to me.
-and because they illustrate two different aspects of tree topology (see Discussion).
+%and because they illustrate two different aspects of tree topology (see Discussion).
+TG: ok I've fixed that towards: "as said above, they describe two different things but also they're cool because we know how they behave and the code is available to calculate them." - a bit more honest.
+We used these two metrics because they illustrate two different aspects of tree topology (see Discussion) but also because their performance in measuring differences in topology is well described \citep{Kuhner04112014} and well implemented \citep{Bogdanowicz2012}.
 We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa. These metrics are described in detail below.
@@
 If they are fully resolved more often than the ML trees you may need t oadd a paragraph to the discussion as Gavin is right, this could explain the result.
+TG: this is now explained in the following paragraph after this one mentioning in details the polytomy stuff.
+Or should be coming before?
 The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds distance) but greater on the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 We argue that using the Bayesian consensus tree topology is preferable because it is more conservative \cite[e.g][]{pattinsonphylogeny2014}.
+
+Also, because the Bayesian consensus tree are more likely to contain polytomies when some branching are poorly resolved (in $<$ 50\% of the trees) in the "best" tree, the Bayesian consensus "missing-data" trees have a higher probability to be less distant to the "best" tree than in a Maximum Likelihood framework.
+However we minimized this effect by only selecting matrices that had a median bootstrap for all their nodes $>$ 50 (see Methods, section 3).
 
@@
 However, it is important to notice that the effect of each parameter is not additive.
-The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).
-This has important implications for building Total Evidence phylogenies (see below).
+The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and/or $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).
 
@@
 % NC: These are still really hard to follow but I'm at a loss as to how to explain them better.
+% TG: will the t-test analogy be useful here?
 

### 2015-01-29 Manuscript/TEM_draft.tex
Minor edits to your edits

@@
 %TG: I think that's covered in \subsection{Individual effects of missing data parameters} - \subsubsection{Missing data for living taxa ($M_{L}$)}. (If no living taxa, the fossils cannot branch.) 
+% NC: OK good
 
@@
 Total Evidence methods have been successfully applied to empirical data \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
-%G: Why does it need so much data? Is there any evidence (a citation or two) to back this up? TG: added clarity in both sentences below.
 However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
@@
 If the there is not enough overlap in the super matrix, then it is impossible to get a consensus clade tree (kind of supporter talked in the paper) per terrace?
+% NC: Let's chat tomorrow
  \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
@@
 The idea being probably close to the tree terrace paper.
+% NC: Again let's chat tomorrow
 Therefore, we focus only on missing data in the morphological part of the matrix.
-We investigate three major parameters that directly affect the completeness and size of the morphological part of the matrix: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix (i.e.
+We investigate three major parameters that directly affect the completeness and size of the morphological part of the matrix, and reflect empirical biases in data availability: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix (i.e.
 the size of the matrix).
-% NC: The next few sentences could be moved to the methods. They feel a bit clunky here. TG: I like the way they end the intro: "we're simulating stuff but it links to more real stuff in reality."
-% NC: OK but hey currently don't make sense in terms of what was said perviosuly so I've numbered them.
-Still don't like this.
-It really should be integrated with the parameters.
-Basically it's poorly structured and doesn't follow very well - that's what I mean by "clunky"  - TG: You're right (as usual ;)) it works better without.
-%These parameters reflect empirical biases in data availability as follows: (i) the increased availability of molecular data means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis; (ii) missing data in fossil taxa is common due to preservation biases \citep{sansomfossilization2013}; and (iii) the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
-%TG: I swapped the results to be consistent, ML and MC first and then Bayesian
+%TG: I swapped the results to be consistent, ML and MC first and then Bayesian % NC: Cool
 We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree topology more so than minimizing the amount of missing data in the fossil record.
@@
 % TG: Really depends. In the 5 I keep citing, Pyron is close to 50/50, Ronquist and Slater are more towards living and Beck is more towards fossil
-Empirical Total Evidence matrices can be biased towards a greater number of fossils (e.g.
-88\% and 58\% fossils taxa in respectively \citealt{beckancient2014} and \citealt{schragocombining2013}) or towards more living taxa (e.g.
-69\% of living taxa in both \citealt{ronquista2012} and \citealt{slaterphylogenetic2013}) buy they can contain an even number of both type of taxa (e.g.
-52\% of living taxa in \citealt{pyrondivergence2011}).
-Therefore, to avoid biasing our simulations towards the two extreme situations (more living or fossil taxa) and to make each simulation comparable, we implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa.
+% NC: Cool this is better and closer to what Gavin meant. I've condensed it a bit
+Empirical Total Evidence matrices vary in whether they have more fossil than living taxa or vice versa.
+For example, fossil taxa make up 88\% \citealt{beckancient2014}, 58\% \citealt{schragocombining2013}, 48\% \citealt{pyrondivergence2011}, 31\% \citealt{ronquista2012} and 31\% \citealt{slaterphylogenetic2013} of taxa in various studies.
+To avoid biasing our simulations towards either living or fossil taxa and to make each simulation comparable, we implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa.
 We then added an outgroup to the tree, using the mean branch length of the tree to separate the outgroup from the rest of the taxa, and with the branch length leading to the outgroup set as the sum of the mean branch length and the longest root-to-tip length of the tree.
@@
 We modified the "complete" matrix to get matrices with missing data by randomly replacing data with "?" in the morphological part of the matrices according to the following parameters:
-
-%TG: this part was removed from the into: These parameters reflect empirical biases in data availability as follows: (i) the increased availability of molecular data means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis; (ii) missing data in fossil taxa is common due to preservation biases \citep{sansomfossilization2013}; and (iii) the overall number of characters depends on the effort of the people identifying them \citep[e.g.][]{O'Leary08022013}.
-
 \begin{enumerate}
@@
 This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological part.
-This reflects the fact that morphological data for living species are scarce.
-This can be due to the increased availability of molecular data and/or that few people have the skills to identify characters needed for detailed phylogenetic analysis.
+This reflects the fact that the increased availability of molecular data means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
 \item{$M_{F}$, the proportion of missing data in the fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the quality of the fossil record. 
+This parameter illustrates the quality of the fossil record, a common problem due to preservation biases \citep{sansomfossilization2013}. 
 \item{$M_{C}$, the proportion of missing morphological characters for both living and fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%. }
@@
 % NC: Haha that is exactly what I've said about a million times to you.
-% TG: And after the literature review for the next paper (missing living in mammals) on community structure & Co. I start to get where you guys are comming for ;).
+% TG: And after the literature review for the next paper (missing living in mammals) on community structure & Co. I start to get where you guys are comming for ;). % NC :P
 we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
@@
 TG: ok I've fixed that towards: "as said above, they describe two different things but also they're cool because we know how they behave and the code is available to calculate them." - a bit more honest.
+% NC: This is better
 We used these two metrics because they illustrate two different aspects of tree topology (see Discussion) but also because their performance in measuring differences in topology is well described \citep{Kuhner04112014} and well implemented \citep{Bogdanowicz2012}.
@@
 Or should be coming before?
+%NC: It's fine down there
 The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds distance) but greater on the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 
-Also, because the Bayesian consensus tree are more likely to contain polytomies when some branching are poorly resolved (in $<$ 50\% of the trees) in the "best" tree, the Bayesian consensus "missing-data" trees have a higher probability to be less distant to the "best" tree than in a Maximum Likelihood framework.
+% NC: OK let's discuss tomorrow. I don't understand
+Also, because the Bayesian consensus trees are more likely to contain polytomies when some branches are poorly resolved (i.e., they appear in $<$ 50\% of the trees) in the "best" tree, the Bayesian consensus "missing-data" trees have a higher probability to be less distant to the "best" tree than in a Maximum Likelihood framework.
 However we minimized this effect by only selecting matrices that had a median bootstrap for all their nodes $>$ 50 (see Methods, section 3).
@@
 So this isn't a justification, it's just a series of factual statements.
-
+% NC: We can discuss tomorrow
 Additionally, it is noticeable that the difference of the Maximum Likelihood tree topologies and the Bayesian tree topologies that could be due to model overparametrisation is not consistent with the results of the Maxmimum Likelihood bootstrap trees topologies and the Bayesian posterior tree distributions topologies that are similar (Table ~\ref{Tab_Results-Difference_methods}).
@@
 % NC: Huh??? Not really following, nor is this a good argument.
-
+% NC: We can discuss tomorrow
 \subsubsection{Missing morphological characters ($M_{C}$)}

### 2015-01-29 Manuscript/SupplementaryMaterial/Supp_results.tex
Removed comments

@@
-% NC: Can you make the legends the same as the ones I already fixed in the main text please???
-
-% NC: Need a short paragraph here explaining that these are supplemental results for different tree building methods or whatever, just to explain why this exists!
 The following section contains the supplemental results of the effects of our missing data parameters and the different tree inference methods on the the ability to recover the "best" topology that are briefly discussed in the main body of the paper.
@@
 The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-\label{Fig_Supp_BootBayt_allparam} %Differences between all the parameters and between two methods (Boot vs Bayt)
+\label{Fig_Supp_BootBayt_allparam} 
 \end{figure}
@@
 \label{Fig_Supp_paircomp_ML}
-\end{figure} %Pairwise BC for the ML trees for RF and TR. (parameters differences)
+\end{figure}
 
@@
 \label{Fig_Supp_paircomp_Boot}
-\end{figure} %Pairwise BC for the ML bootstrap trees for RF and TR. (parameters differences)
+\end{figure}
 
@@
 \label{Fig_Supp_paircomp_Bayt}
-\end{figure} %Pairwise BC for the Bayesian posterior trees distribution for RF and TR. (parameters differences)
+\end{figure}
 

### 2015-01-29 Manuscript/TEM_draft.tex
Removed comments

@@
 \begin{document}
-%\modulolinenumbers[1]     % Line numbering on every line
 
@@
 
-%Gavin's general comment:I have made some comments on Thomas’ paper.
-I think it generally looks good.
-The one thing that I think might be tricky to get past reviewers is the comparison between the ML and Bayesian trees.
-The problem is that you can’t tell if the poorer performance of ML is because of the inference method or the difference between the substitution models - the ML uses an overparameterised model (GTR) compared to the generating model, whereas the Bayesian analyses do not.
-It might be possible to cite your way out of this but I wouldn’t be surprised if a reviewer asks for new analyses by either using different ML software that can handle the HKY model (e.g.
-Garli) or running the Bayesian with the GTR model (although that might not be adequate because if I remember correctly RAXML uses an approximation for gamma that is not the same as in other software).
-
 %---------------------------------------------
@@
 We find that the number of living taxa with morphological characters and the overall number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology.
-% NC: I think you misunderstood Nick's comment here.
-He meant that living > fossil but that you can't put the number of characters in here too.
-I've reverted back to the original as I'm not sure his solution is correct
 Therefore, we suggest that sampling effort should be focused on morphological data collection for living species to increase the accuracy of topological inference in a Total Evidence framework.
@@
 We therefore recommend using Bayesian consensus trees to fix the tree topology prior to further analyses.
-
-%Hummm Nick's comment again - That's a point to make in the discussion maybe. But let's wait for Gavin comments...
-%it's possible that adding dating information etc.
-in tip-dating makes Bayesian dated total-evidence analyses (ie tip-dating) better than non-dated Bayesian total-evidence analyses (typical total-evidence approach), because more information is being used, and because dates give some information about phylogenetic position.
-In any case, the study wasn't about tip-dating
-% NC: Simplest thing here is to just leave out the mention of tip dating. Will need to fix in discussion
-% NC: Also Nick makes another sensible comment about *why* are living species more important. Check this is covered in discussion 
-%TG: I think that's covered in \subsection{Individual effects of missing data parameters} - \subsubsection{Missing data for living taxa ($M_{L}$)}. (If no living taxa, the fossils cannot branch.) 
-% NC: OK good
 
@@
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
-%TG: I swapped the results to be consistent, ML and MC first and then Bayesian % NC: Cool
 We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree topology more so than minimizing the amount of missing data in the fossil record.
@@
 We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
-%G:I don’t think there is a problem with this split but I expect reviewers might ask for justification. Is this typical of existing TE studies? TG: avoiding bias comment.
-% NC: I think you misunderstood Gavin's point.
-He knows why you did it like this, but if usually TE studies are 95% living and 5% fossil (for example) then your results aren't really viable in real world.
-He ideally wants a statement like: this is a typical split between fossil and living tax across empirical TE studies (cite papers).
-However, I doubt this is the case.
-Actually aren't TE studeis usually dominated by fossil taxa?
-% TG: Really depends. In the 5 I keep citing, Pyron is close to 50/50, Ronquist and Slater are more towards living and Beck is more towards fossil
-% NC: Cool this is better and closer to what Gavin meant. I've condensed it a bit
 Empirical Total Evidence matrices vary in whether they have more fossil than living taxa or vice versa.
@@
 
-To measure the support for each branch in our simulated phylogenies
-%G: Do you mean you tested whether the simulated data generally had any phylogenetic signal with which to recover a tree?
-I was a bit unsure but mainly because I am hardwired to thinks of Pagel’s lambda when anyone mentions phylogenetic signal…
-% NC: Haha that is exactly what I've said about a million times to you.
-% TG: And after the literature review for the next paper (missing living in mammals) on community structure & Co. I start to get where you guys are comming for ;). % NC :P
-we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
+To measure the support for each branch in our simulated phylogenies we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
 We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
@@
 taxa that frequently change position in different trees; \citealt{kearneyfragmentary2002}).
-%G: Why these two metrics? Might need a few sentences to justify them in addition to the explanation of what they are that you already have.
-%We used these metrics because they are widely used in literature \citep[][e.g.]{Bogdanowicz2012}, their relative performance is well described \citep{Kuhner04112014} % NC: Is that what you mean.
-"they relative performance" means nothing to me.
-%and because they illustrate two different aspects of tree topology (see Discussion).
-TG: ok I've fixed that towards: "as said above, they describe two different things but also they're cool because we know how they behave and the code is available to calculate them." - a bit more honest.
-% NC: This is better
 We used these two metrics because they illustrate two different aspects of tree topology (see Discussion) but also because their performance in measuring differences in topology is well described \citep{Kuhner04112014} and well implemented \citep{Bogdanowicz2012}.
-We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa. These metrics are described in detail below.
+We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
+These metrics are described in detail below.
 
@@
 \begin{table}[ht]
-\caption{
-%G: I think this table needs to be decoded a bit.
-I wasn’t really sure what it was supposed to be getting across.
-Maybe it is because I am not familiar with the Bhatt… distance but I suspect that many readers will be similarly unfamiliar.
-TG: fixed clarity
-Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" trees using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" trees using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
 Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
 The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
-If two methods have a high probability of overlap, they have a similar ability to recover the "correct" tree topology.
-% NC: check this is right, I think that's what Gavin wanted to be added - a way of interpreting what is going on
-}
+If two methods have a high probability of overlap, they have a similar ability to recover the "correct" tree topology.}
 \centering

### 2015-01-30 Manuscript/OnlineAppendices-LaTeXSuppFiles/Supp_results.tex
Reorganised folders for submission

--- 2015-01-29 Manuscript/SupplementaryMaterial/Supp_results.tex
@@
 \centering
-    \includegraphics[width=1\textwidth]{SupplementaryFigures/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
+    \includegraphics[width=1\textwidth]{OnlineAppendices-LaTeXSuppFiles/SupplementaryFigures/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
     \caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey).
@@
 %    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-ML-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+    \includegraphics[width=1\textwidth]{OnlineAppendices-LaTeXSuppFiles/SupplementaryFigures/PairwiseComp-ML-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
     \caption{The effects of missing data on topological recovery using Maximum Likelihood trees.
@@
 %    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-Boot-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+    \includegraphics[width=1\textwidth]{OnlineAppendices-LaTeXSuppFiles/SupplementaryFigures/PairwiseComp-Boot-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
     \caption{The effects of missing data on topological recovery using Maximum Likelihood bootstrap trees.
@@
 %    \includegraphics[width=1\textwidth]{Figures/Supplementary/PPairwiseComp-Bayt-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-Bayt-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+    \includegraphics[width=1\textwidth]{OnlineAppendices-LaTeXSuppFiles/SupplementaryFigures/PairwiseComp-Bayt-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
     \caption{The effects of missing data on topological recovery using Bayesian posterior tree distribution.

### 2015-01-30 Manuscript/TEM_draft.tex
Reorganised folders for submission

@@
 
-\bibliographystyle{sysbio}
-\bibliography{References}
+\bibliographystyle{MainBody-LaTeXSuppFiles/sysbio}
+\bibliography{MainBody-LaTeXSuppFiles/References}
 
@@
 \centering
-    \includegraphics[keepaspectratio=true]{Figures/In_main/Simulations_outline-BW.pdf}
+    \includegraphics[keepaspectratio=true]{MainBody-LaTeXSuppFiles/Figures/In_main/Simulations_outline-BW.pdf}
 \caption{ }
@@
 \centering
-    \includegraphics[width=0.5\textwidth]{Figures/In_main/Bhattacharyya_Coefficients1.pdf}
+    \includegraphics[width=0.5\textwidth]{MainBody-LaTeXSuppFiles/Figures/In_main/Bhattacharyya_Coefficients1.pdf}
 \caption{ }
@@
 \centering
-    \includegraphics[width=1\textwidth]{Figures/In_main/Bhattacharyya_Coefficients2.pdf}
+    \includegraphics[width=1\textwidth]{MainBody-LaTeXSuppFiles/Figures/In_main/Bhattacharyya_Coefficients2.pdf}
 \caption{ }
@@
 \centering
-    \includegraphics[width=1\textwidth]{Figures/In_main/AllMethods-RF+Tr-colour.pdf}
+    \includegraphics[width=1\textwidth]{MainBody-LaTeXSuppFiles/Figures/In_main/AllMethods-RF+Tr-colour.pdf}
 \caption{ }
@@
 \centering
-    \includegraphics[width=1\textwidth]{Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
+    \includegraphics[width=1\textwidth]{MainBody-LaTeXSuppFiles/Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
 \caption{ }
@@
 %    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+    \includegraphics[width=1\textwidth]{MainBody-LaTeXSuppFiles/Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
 \caption{ }

### 2015-01-30 Manuscript/TEM_draft.tex
Edits

@@
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case)
-%G: Might be worthing reading and citing this: http://www.sciencemag.org/content/333/6041/448.full.
-Missing data can be a problem if it renders some subtrees unresolvable.
-TG: added citation in the whole bunch.
-% NC: Did you actually *read* this paper?
-It makes the point that it *is* a problem if it affects certain subtrees.
-Might need to add a small caveat into the bit in brackets above...
-TG: Ok, so this comment goes with the data overlap.
-I’m not entirely sure where the paper comes in.
-If the there is not enough overlap in the super matrix, then it is impossible to get a consensus clade tree (kind of supporter talked in the paper) per terrace?
-% NC: Let's chat tomorrow
- \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data is not biased towards particular subclades; \citet{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}) % ADD CITATION % Add Lemmon
+, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
@@
 Since the molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}, the effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011}.
-%G: Alan Lemmon has done a fair amount of work on missingness - worth checking out the work from his lab, e.g.
-http://www.evotutor.org/LemmonLab/Pdfs/Lemmon-etal2009.pdf TG: added reference (that was actually in the 2014 draft)
-% NC: OK.
-But be careful here.
-When people give you comments you want to take them seriously and actually read the papers they suggest rather than just adding references.
-Are you sure Lemmon is just looking at molecular stuff?
-Think Gavin's point might be that he's done stuff on missingness in morphological data too which at present we don't really cite or cover at all.
-TG: I don’t like arguing on comments but this paper is about missing data in molecular matrices (still relevant though).
-The reason I removed it from the old version is that it states something kind of similar than all the others: overlap but it is specially important in bigger matrices where missing data can weaken the support in certain parts of the tree.
-The idea being probably close to the tree terrace paper.
-% NC: Again let's chat tomorrow
 Therefore, we focus only on missing data in the morphological part of the matrix.
@@
 8.0.20; \citealt{Stamatakis21012014}).
-%G: I think you need to justify using the GTR model (and so also RAXML).
-The data were simulated under an HKY model so you are fitting an overparameterised model.
-It might be that the better results for the MrBayes analyses come down to model misspecification.
-You can certainly say that HKY is not an option in RAXML but then you might be asked why you don’t use another program, e.g Garli.
 We used RAxML because it is heavily used in the literature for Maximum Likelihood tree inference \citep[][e.g.]{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014} and is one of the fastest methods available \citep{Stamatakis01102008}.
-However, RAxML has no option for using the HKY model so we used the GTR model instead (see Discussion).
-% NC: This isn't really a justification. Also by pointing out the problem you're drawing attention to it. I actually never noticed this when reading the previous drafts. But I've never used RAxML.
 
@@
 We repeated this selection until we obtained 50 sets of simulations (i.e.
-50 "complete" and 50 x 125 "missing-data" matrices) with a relative good phylogenetic signal (median bootstrap $>$ 50).
+50 "complete" and 50 x 125 "missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
 On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \citep{pattengale2010many}.
 Using these parameters took \texttildelow 8 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (2.30GHz clock speed nodes).
+We performed this procedure to increase the resolution of our resulting trees.
 
@@
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood
-%G: This might be expected given the GTR rather than HKY model is used. 
- trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; online Appendix Tables %~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}
-S5, S6 and S7 %Modified for build without supplementaries
-). 
-However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; online Appendix Tables %~\ref{Tab_Supp_summary_BC_ML}, ~\ref{Tab_Supp_summary_BC_MF} and ~\ref{Tab_Supp_summary_BC_MC}
-S5, S6 and S7 %Modified for build without supplementaries
-).
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; online Appendix Tables S5, S6 and S7).
+However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; online Appendix Tables S5, S6 and S7).
 
@@
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; online Appendix Tables %~\ref{Tab_Supp_summary_metric_ML}, ~\ref{Tab_Supp_summary_metric_MF} and ~\ref{Tab_Supp_summary_metric_MC}
-S2, S3 and S4 %Modified for build without supplementaries
-).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; online Appendix Tables S2, S3 and S4).
 Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
@@
 For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; Table ~\ref{Tab_Results-Difference_methods}).
-However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and online Appendix Fig.%~\ref{Fig_Supp_BootBayt_allparam}
-S2 %Modified for build without supplementaries
-).
+However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and online Appendix Fig.
+S2).
 
@@
 
-%G: How does tree resolution influence the tree comparison metrics?
-Are the Bayesian trees always fully resolved?
-Could that bias towards shorter distances to the best tree (because there is reduced probability of conflict).
-TG: This comment is kind of linking to the following.
-
-% NC: Yes BUT the point is that Gavin misunderstood so you need to be clearer!
-Add a comment earlier that points out that the Bayesian consensus trees are not fully resolved (if this is true).
-If they are fully resolved more often than the ML trees you may need t oadd a paragraph to the discussion as Gavin is right, this could explain the result.
-TG: this is now explained in the following paragraph after this one mentioning in details the polytomy stuff.
-Or should be coming before?
-%NC: It's fine down there
 The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds distance) but greater on the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 
-% NC: OK let's discuss tomorrow. I don't understand
-Also, because the Bayesian consensus trees are more likely to contain polytomies when some branches are poorly resolved (i.e., they appear in $<$ 50\% of the trees) in the "best" tree, the Bayesian consensus "missing-data" trees have a higher probability to be less distant to the "best" tree than in a Maximum Likelihood framework.
-However we minimized this effect by only selecting matrices that had a median bootstrap for all their nodes $>$ 50 (see Methods, section 3).
-
-%^G: I think you need to justify using the GTR model (and so also RAXML).
-The data were simulated under an HKY model so you are fitting an overparameterised model.
-It might be that the better results for the MrBayes analyses come down to model misspecification.
-You can certainly say that HKY is not an option in RAXML but then you might be asked why you don’t use another program, e.g Garli.
-The Maximum Likelihood analyses used the GTR rather than the HKY model, because HKY is not an option in RAxML.
-This misspecification of the Maximum Likelihood model could also explain the differences between the Maximum Likelihood and Bayesian consensus tree topologies.
-However, the GTR model is a generalisation of the HKY model: a GTR model where the $\alpha$ and $\eta$ parameters are equal to the HKY $k$ parameter and when the 4 other parameters ($\beta$, $\gamma$, $\delta$ and $\epsilon$) are equal to one is effectively a HKY model with a lower degree of freedom.
-Also, the effect of overparametrisation is higher in Bayesian analysis than in Maximum Likelihood ones (see \citealt{Buckley01052002} and discussion in \citealt{douadycomparison2003}).
-% NC: OK *none* of this is a justification.
-More the opposite!
-We know GTR is a generalisation of HKY, that doesn't make it better.
-That's basically what we mean when we say the ML model is overparameterised because even if all parameters = 1 it's a HKY but with fewer df.
-Thus over parameterised!
-Then yes, this would be a bigger issue for Bayesian - but so what?
-It's not an issue in Bayesian here.
-It's not like you had to choose to do it for one or the other, thus ML was a better choice.
-So this isn't a justification, it's just a series of factual statements.
-% NC: We can discuss tomorrow
-Additionally, it is noticeable that the difference of the Maximum Likelihood tree topologies and the Bayesian tree topologies that could be due to model overparametrisation is not consistent with the results of the Maxmimum Likelihood bootstrap trees topologies and the Bayesian posterior tree distributions topologies that are similar (Table ~\ref{Tab_Results-Difference_methods}).
-Therefore we argue that the differences between the Maximum Likelihood tree topologies and the Bayesian consensus tree topologies is more likely due to the way the Bayesian consensus tree is built (see above) than to model misspecification.
-% NC: Huh??? Not really following, nor is this a good argument.
-% NC: We can discuss tomorrow
 \subsubsection{Missing morphological characters ($M_{C}$)}
@@
 As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
+This may also be because the "best" Bayesian consensus trees are not completely resolved, thus will always be more similar to the "missing data" trees than a completely resolved tree like the "best" Maximum Likelihood tree.
+However, we minimized the probability of unresolved "best" trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
 
@@
 Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating (\citealp{ronquista2012,BEASTmaster};
-%^Nick
-%it's possible that adding dating information etc.
-in tip-dating makes Bayesian dated total-evidence analyses (ie tip-dating) better than non-dated Bayesian total-evidence analyses (typical total-evidence approach), because more information is being used, and because dates give some information about phylogenetic position.
-In any case, the study wasn't about tip-dating
-% NC: Note that Nick's point is that we can really only say what works best for topology only methods
-although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution).
+Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating (\citealp{ronquista2012,BEASTmaster}; although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution).
 Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see above).
@@
 \section{Conclusion}
-
 Missing data in Total Evidence matrices is not a problem for recovering the "best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
@@
 Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
-% NC: Why would it improve spatial methods? I've removed this.
 
 %---------------------------------------------
-%
-%       ...
-%
-%---------------------------------------------
+
 \section{Supplementary material}
@@
 
-%cite cheat sheet
- % The \cite command functions as follows:
- %   \citet{key} ==>>                Jones et al. (1990)
- %   \citet*{key} ==>>               Jones, Baker, and Smith (1990)
- %   \citep{key} ==>>                (Jones et al., 1990)
- %   \citep*{key} ==>>               (Jones, Baker, and Smith, 1990)
- %   \citep[chap. 2]{key} ==>>       (Jones et al., 1990, chap. 2)
- %   \citep[e.g.][]{key} ==>>        (e.g. Jones et al., 1990)
- %   \citep[e.g.][p. 32]{key} ==>>   (e.g. Jones et al., p. 32)
- %   \citeauthor{key} ==>>           Jones et al.
- %   \citeauthor*{key} ==>>          Jones, Baker, and Smith
- %   \citeyear{key} ==>>             1990
-
 \bibliographystyle{MainBody-LaTeXSuppFiles/sysbio}
 \bibliography{MainBody-LaTeXSuppFiles/References}
-
-%---------------------------------------------
-%
-%       Supplementary
-%
-%---------------------------------------------
-
-%Removed to a separate file
-%\newpage
-%\section{Supplementary Material}
-%\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
-%\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
-
-%    \input{SupplementaryMaterial/SupplementaryMaterial}
-%    \label{SupplementaryMaterial}
 
@@
 Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
+Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
 The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
@@
 $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
-The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.}
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.
+Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
 \label{Fig_Bhattacharyya_Coefficients1} %Bhattacharyya coefficients calculation outline
@@
 Maximum Likelihood or Bayesian).
-The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$.}
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$.
+Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
 \label{Fig_Bhattacharyya_Coefficients2} %Bhattacharyya coefficients calculation outline
@@
 The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
-%G: Is there a scale to the colour or is it binary? Also, I don’t get the axes. Am I missing something or are they the same? TG: fixed clarity.
 }

### 2015-01-30 Manuscript/SystBiolSubmission/TEM_draft.tex
SystBiol Submission folder

--- 2015-01-30 Manuscript/TEM_draft.tex
@@
-%LaTeX template : http://systbio.org/files/SB_LaTeX_Template_txt_extension.txt
-%Author instructions: http://www.oxfordjournals.org/our_journals/sysbio/for_authors/ms_preparation.html
-
 \documentclass[12pt,letterpaper]{article}
@@
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data is not biased towards particular subclades; \citet{wiensmissing2003,wiensmissing2006,wiensmissing2008,rouresite-specific2011,pattinsonphylogeny2014}) % ADD CITATION % Add Lemmon
-, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data is not biased towards particular subclades; \citet{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}), it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
@@
 
-\bibliographystyle{MainBody-LaTeXSuppFiles/sysbio}
-\bibliography{MainBody-LaTeXSuppFiles/References}
+\bibliographystyle{sysbio}
+\bibliography{References}
 
@@
 Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
-\label{Fig_Bhattacharyya_Coefficients1} %Bhattacharyya coefficients calculation outline
+\label{Fig_Bhattacharyya_Coefficients1}
 \end{figure}
@@
 Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
-\label{Fig_Bhattacharyya_Coefficients2} %Bhattacharyya coefficients calculation outline
+\label{Fig_Bhattacharyya_Coefficients2} 
 \end{figure}
-
-% NC: These are still really hard to follow but I'm at a loss as to how to explain them better.
-% TG: will the t-test analogy be useful here?
 
@@
 The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-\label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
+\label{Fig_Results-permeth_perparam} 
 \end{figure}
@@
 The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-\label{Fig_Results-global_perparam} %Differences between all the parameters and between two methods (ML vs BC)
+\label{Fig_Results-global_perparam}
 \end{figure}
@@
 \centering
-    \includegraphics[keepaspectratio=true]{MainBody-LaTeXSuppFiles/Figures/In_main/Simulations_outline-BW.pdf}
+    \includegraphics[keepaspectratio=true]{Simulations_outline-BW.pdf}
 \caption{ }
-%(1) We randomly generated a birth-death tree (the "true" tree) and used it to infer a matrix with no missing data (the "complete" matrix).
-%(2) We removed data from the morphological part of the "complete" matrix resulting in 125 "missing-data" matrices.
-%(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
-%(4) We compared the "missing-data" trees to the "best" tree.
-%We repeated these four steps 50 times.}
-%\label{Fig_Outline}
 \end{figure}
@@
 \centering
-    \includegraphics[width=0.5\textwidth]{MainBody-LaTeXSuppFiles/Figures/In_main/Bhattacharyya_Coefficients1.pdf}
+    \includegraphics[width=0.5\textwidth]{Bhattacharyya_Coefficients1.pdf}
 \caption{ }
-%\caption{Bhattacharyya Coefficient calculation outline.
-A, B and C are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
-$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
-The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two combination of missing data parameters such as BC(A,B) is the probability of overlap between the distributions A and B.}
-%\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
 \end{figure}
@@
 \centering
-    \includegraphics[width=1\textwidth]{MainBody-LaTeXSuppFiles/Figures/In_main/Bhattacharyya_Coefficients2.pdf}
+    \includegraphics[width=1\textwidth]{Bhattacharyya_Coefficients2.pdf}
 \caption{ }
-%\caption{Bhattacharyya Coefficient calculation outline.
-A and B are distribution of tree similarity (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
-$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
-\textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
-Maximum Likelihood or Bayesian).
-The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity between two methods for the same combination of missing data parameters such as BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A between the methods $x$ and $y$.}
-%\label{Fig_Bhattacharyya_Coefficients} %Bhattacharyya coefficients calculation outline
 \end{figure}
@@
 \centering
-    \includegraphics[width=1\textwidth]{MainBody-LaTeXSuppFiles/Figures/In_main/AllMethods-RF+Tr-colour.pdf}
+    \includegraphics[width=1\textwidth]{AllMethods-RF+Tr-colour.pdf}
 \caption{ }
-%\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).
-The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-%\label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
 \end{figure}
@@
 \centering
-    \includegraphics[width=1\textwidth]{MainBody-LaTeXSuppFiles/Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
+    \includegraphics[width=1\textwidth]{ML+Baycon-AllParam-RF+Tr-BW.pdf}
 \caption{ }
-%\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
-The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-%\label{Fig_Results-global_perparam} %Differences between all the parameters and between two methods (ML vs BC)
 \end{figure}
@@
 \centering
-     %Hide this figure for 'fast' build
-%    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{MainBody-LaTeXSuppFiles/Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
+    \includegraphics[width=1\textwidth]{PairwiseComp-Baycon-RF+Tr-colour.png}
 \caption{ }
-%\caption{The effects of missing data on topological recovery using Bayesian consensus trees.
-Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
-%\label{Fig_Results-paircomp_within}
-\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)
-%END
-\end{document}
+\end{figure}

### 2015-01-30 Manuscript/SystBiolSubmission/TEM_draft.tex
removed hypref

@@
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see \hyperref[SupplementaryMaterial]{online Appendix 1} and Fig.
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see online Appendix 1 and Fig.
 S1).
@@
 Even using these priors, it took \texttildelow 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-The detailed MrBayes parameters are available in the \hyperref[SupplementaryMaterial]{online Appendix 1}.
+The detailed MrBayes parameters are available in the online Appendix 1.
 
@@
 The Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see \hyperref[SupplementaryMaterial]{online Appendix 2} for calculation details).
+The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see online Appendix 2 for calculation details).
 This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
@@
 We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons \citep{Bogdanowicz2012}.
-This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see \hyperref[SupplementaryMaterial]{online Appendix 2} for the calculation details).
+This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see online Appendix 2 for the calculation details).
 Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
@@
 \subsubsection{Triplets distance}
-The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see \hyperref[SupplementaryMaterial]{online Appendix 2} for calculation details).
+The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see online Appendix 2 for calculation details).
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
@@
 We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see \hyperref[SupplementaryMaterial]{online Appendix 2} for calculation details).
+The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see online Appendix  for calculation details).
 Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
@@
 When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
-Because living taxa still have molecular characters available for phylogenetic inference (see \hyperref[Generating_the_matrix]{methods}),even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
+Because living taxa still have molecular characters available for phylogenetic inference (see mMthods),even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
 However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
@@
 \end{figure}
+
+\end{document}

### 2015-02-06 Manuscript/PLoSONESubmission/TEM_draft.tex
PLoS One Submission init

--- 2015-01-30 Manuscript/SystBiolSubmission/TEM_draft.tex
@@
 \centering
-    \includegraphics[width=1\textwidth]{AllMethods-RF+Tr-colour.pdf}
+    \includegraphics[width=1\textwidth]{AllMethods-RF_Tr-colour.pdf}
 \caption{ }
@@
 \centering
-    \includegraphics[width=1\textwidth]{ML+Baycon-AllParam-RF+Tr-BW.pdf}
+    \includegraphics[width=1\textwidth]{ML_Baycon-AllParam-RF_Tr-BW.pdf}
 \caption{ }
@@
 \centering
-    \includegraphics[width=1\textwidth]{PairwiseComp-Baycon-RF+Tr-colour.png}
+    \includegraphics[width=1\textwidth]{PairwiseComp-Baycon-RF_Tr-colour.png}
 \caption{ }

### 2015-02-08 Manuscript/PLoSONESubmission/TEM_PLoS_draft.tex
Updated PLoSONE submission (formating)

--- 2015-02-06 Manuscript/PLoSONESubmission/TEM_draft.tex
@@
-\documentclass[12pt,letterpaper]{article}
+% Please do not include colors or graphics in the text.
+%
+% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - tabs/spacing/line breaks within cells to alter layout or alignment
+% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
+% - colors, shading, or graphic objects
+% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
+%
+% Please be sure to include all portions of an equation in the math environment.
+%
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
 
-%Packages
-\usepackage{pdflscape}
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+% Use Unicode characters when possible
+\usepackage[utf8]{inputenc}
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+% fixltx2e package for \textsubscript
 \usepackage{fixltx2e}
-\usepackage{textcomp}
-\usepackage{fullpage}
-\usepackage{natbib}
-\usepackage{float}
-\usepackage{latexsym}
-\usepackage{url}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{bm}
-\usepackage{array}
-\usepackage[version=3]{mhchem}
-\usepackage{ifthen}
-\usepackage{caption}
-\usepackage{hyperref}
-\usepackage{amsthm}
-\usepackage{amstext}
-\usepackage{enumerate}
-\usepackage[osf]{mathpazo}
-\usepackage{dcolumn}
-\usepackage{lineno}
-\pagenumbering{arabic}
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
+\usepackage{amsmath,amssymb}
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+% line numbers
+\usepackage[right]{lineno}
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+% rotating package for sideways tables
+\usepackage{rotating}
+\usepackage{setspace} 
+\doublespacing
 
-%Pagination style and stuff
-\linespread{2}
+% Text layout
 \raggedright
-\setlength{\parindent}{0.5in}
-\setcounter{secnumdepth}{0} 
-\renewcommand{\section}[1]{%
-\bigskip
-\begin{center}
-\begin{Large}
-\normalfont\scshape #1
-\medskip
-\end{Large}
-\end{center}}
-\renewcommand{\subsection}[1]{%
-\bigskip
-\begin{center}
-\begin{large}
-\normalfont\itshape #1
-\end{large}
-\end{center}}
-\renewcommand{\subsubsection}[1]{%
-\vspace{2ex}
-\noindent
-\textit{#1.}---}
-\renewcommand{\tableofcontents}{}
-\bibpunct{(}{)}{;}{a}{}{,}
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
 
-%---------------------------------------------
-%
-%       START
-%
-%---------------------------------------------
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2009}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Leave date blank
+\date{}
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+\lhead{\includegraphics[natwidth=1.3in,natheight=0.4in]{PLOSlogo.png}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\sf PLOS}
+
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
+
+%% END MACROS SECTION
 
 \begin{document}
+\vspace*{0.35in}
 
-%Running head
-\begin{flushright}
-Version dated: \today
-\end{flushright}
-\bigskip
-\noindent RH: Missing data and topology in Total Evidence method
-
-\bigskip
-\medskip
-\begin{center}
-
-\noindent{\Large \bf Effects of missing data on topological inference using a Total Evidence approach}
-
-\bigskip
-
-\noindent {\normalsize \sc Thomas Guillerme$^1$$^,$$^2$$^*$, and Natalie Cooper$^1$$^,$$^2$}\\
-\noindent {\small \it 
-$^1$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
-$^2$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.}\\
-\end{center}
-\medskip
-\noindent{*\bf Corresponding author.} \textit{Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\
-\vspace{1in}
-
-%Line numbering
-\modulolinenumbers[1]
-\linenumbers
-
-%---------------------------------------------
-%
-%       ABSTRACT
-%
-%---------------------------------------------
-
-\newpage
-\begin{abstract}
+% Title must be 150 characters or less
+\begin{flushleft}
+{\Large
+\textbf\newline{Effects of missing data on topological inference using a Total Evidence approach}
+}
+\newline
+% Insert Author names, affiliations and corresponding author email.
+\\
+Thomas Guillerme\textsuperscript{1,2,*},
+Natalie Cooper\textsuperscript{1,2}
+\\
+\bf{1} School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland
+\\
+\bf{2} Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland
+\\
+* E-mail: Corresponding guillert@tcd.ie
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
 To fully understand macroevolutionary patterns and processes, we need to include both extant and extinct species in our models.
@@
 
-\end{abstract}
-
-\noindent (Keywords: morphological characters, Bayesian, Maximum Likelihood, topology, fossil, living)\\
-
-\vspace{1.5in}
-
-\newpage
+\linenumbers
 
@@
 
-\section{Introduction}
-
-Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}.
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
-This has led to increasing consensus among evolutionary biologists that fossil taxa should be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments \citep[e.g.][]{pyrondivergence2011,ronquista2012,BEASTmaster}.
+\section*{Introduction}
+Although most species that have ever lived are now extinct \cite{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species (e.g.
+\cite{meredithimpacts2011,jetzthe2012}).
+Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events (e.g.
+\cite{pyrondivergence2011}), relationships among lineages (e.g.
+\cite{manosphylogeny2007}) or niche occupancy (e.g.
+\cite{pearmanniche2008}).
+This has led to increasing consensus among evolutionary biologists that fossil taxa should be included in macroevolutionary studies \cite{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
+However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments (e.g.
+\cite{pyrondivergence2011,ronquista2012,BEASTmaster}).
 
@@
 Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
-Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \citep{simpson1945}.
+Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \cite{simpson1945}.
 This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty.
@@
 Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates are used to time calibrate phylogenies \citep{zuckerkandl1965}.
-There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the "best" approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
+Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates are used to time calibrate phylogenies \cite{zuckerkandl1965}.
+There have been great improvements in the theory and application of these two approaches (e.g.
+\cite{bapsta2013,stadlerdating2013,heaththe2013}) as well as much debate about the "best" approach to use (e.g.
+\cite{spencerefficacy2013,wrightbayesian2014}).
 However neither approach uses all the available data.
 
-A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \citep{eernissetaxonomic1993}.
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as tip-dating; \citealt{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
-Total Evidence methods have been successfully applied to empirical data \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}, and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
+A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \cite{eernissetaxonomic1993}.
+This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as tip-dating \cite{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \cite{ronquista2012}.
+Total Evidence methods have been successfully applied to empirical data (e.g.
+\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}), and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
 However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
 Morphological data for living taxa are rarely collected when molecular data are available (e.g.
-\citealt{O'Leary08022013} \textit{vs.} \citealt{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record (for example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts; \citealt{sansomfossilization2013}) and molecular data is (nearly) always unavailable.
-Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
+\cite{O'Leary08022013} \textit{vs.} \cite{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record (for example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \cite{sansomfossilization2013}; and molecular data is (nearly) always unavailable.
+Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \cite{salamin2003}.
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data is not biased towards particular subclades; \citet{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}), it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data is not biased towards particular subclades \cite{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}), it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
@@
 
-In this study, we focus on the effect of missing data on our ability to recover a "best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
+In this study, we focus on the effect of missing data on our ability to recover a "best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species (e.g.
+\cite{meredithimpacts2011,jetzthe2012}), or for studying evolutionary transitions (e.g.
+\cite{friedmanexplosive2010}).
 Although branch length estimation is also important (namely for timing extinction and/or speciation events - e.g.
-\citealt{ronquista2012}), we do not consider branch lengths in this study.
-This is partially due to difficulties with simulating branch lengths and topology simultaneously, but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+\cite{ronquista2012}), we do not consider branch lengths in this study.
+This is partially due to difficulties with simulating branch lengths and topology simultaneously, but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \cite{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework.
@@
 Here we use a simulation approach to assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
-Since the molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \citep{ronquista2012}, the effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011}.
+Since the molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \cite{ronquista2012}, the effect of missing data on such matrices is well known \cite{wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011}.
 Therefore, we focus only on missing data in the morphological part of the matrix.
@@
 
-\newpage
-
-\section{Methods}
+\section*{Methods}
 To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
@@
 
+\begin{figure}[h]
+\caption{{\bf Protocol outline.}
+(1) We randomly generated a birth-death tree (the "true" tree) and used it to infer a matrix with no missing data (the "complete" matrix).
+(2) We removed data from the morphological part of the "complete" matrix resulting in 125 "missing-data" matrices.
+(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
+(4) We compared the "missing-data" trees to the "best" tree.
+We repeated these four steps 50 times.}
+\label{Fig_Outline}
+\end{figure}
+
 %---------------------------------------------
@@
 
-\subsection{1. Generating the matrix}
+\subsection*{1. Generating the matrix}
 \label{Generating_the_matrix}
 First we randomly generated a "true" tree of 50 taxa in R v.
-3.0.2 \citep{R302} using the package diversitree v.
-0.9-6 \citep{fitzjohndiversitree2012}.
-We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
+3.0.2 \cite{R302} using the package diversitree v.
+0.9-6 \cite{fitzjohndiversitree2012}.
+We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \cite{paradistime-dependent2011}.
 Empirical Total Evidence matrices vary in whether they have more fossil than living taxa or vice versa.
-For example, fossil taxa make up 88\% \citealt{beckancient2014}, 58\% \citealt{schragocombining2013}, 48\% \citealt{pyrondivergence2011}, 31\% \citealt{ronquista2012} and 31\% \citealt{slaterphylogenetic2013} of taxa in various studies.
+For example, fossil taxa make up 88\% \cite{beckancient2014}, 58\% \cite{schragocombining2013}, 48\% \cite{pyrondivergence2011}, 31\% \cite{ronquista2012} and 31\% \cite{slaterphylogenetic2013} of taxa in various studies.
 To avoid biasing our simulations towards either living or fossil taxa and to make each simulation comparable, we implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa.
@@
 The molecular matrix was inferred from the "true" tree using the R package phyclust v.
-0.1-14 \citep{chen2011}.
-The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies and transition/transversion rate of two \citep{douadycomparison2003}.
-The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
+0.1-14 \cite{chen2011}.
+The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \cite{ranbaut1997seqgen} and using the HKY model \cite{HKY85} with random base frequencies and transition/transversion rate of two \cite{douadycomparison2003}.
+The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \cite{yangamong-site1996}.
 We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
@@
 We inferred the morphological matrix using the R package ape v.
-3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
+3.0-11 \cite{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see online Appendix 1 and Fig.
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see Appendix S1 and Fig.
 S1).
 We then ran an independent discrete character simulation for each character using the "true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
-evolutionary rate) from one character state to another \citep{Pagel22011994}.
+evolutionary rate) from one character state to another \cite{Pagel22011994}.
 This method allows us to have only two parameters for each character: the number of states and the evolutionary rate.
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used low evolutionary rate parameters to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014,wrightbayesian2014}.
+We used low evolutionary rate parameters to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \cite{wagner2000,davalosintegrating2014,wrightbayesian2014}.
 
@@
 the matrix with no missing data except for the molecular data of the fossil taxa.
-
 %---------------------------------------------
@@
 
-\subsection{2. Removing data}
+\subsection*{2. Removing data}
 \label{Removing_data}
@@
 \item{$M_{F}$, the proportion of missing data in the fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the quality of the fossil record, a common problem due to preservation biases \citep{sansomfossilization2013}. 
+This parameter illustrates the quality of the fossil record, a common problem due to preservation biases \cite{sansomfossilization2013}. 
 \item{$M_{C}$, the proportion of missing morphological characters for both living and fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%. }
 This parameter illustrates the number of available morphological characters for both living and fossil taxa.
-This reflects the amount of effort put into identifying the characters \citep[e.g.][]{O'Leary08022013}.
+This reflects the amount of effort put into identifying the characters (e.g.
+\cite{O'Leary08022013}).
 \end{enumerate}
@@
 
-\subsection{3. Building phylogenies}
+\subsection*{3. Building phylogenies}
 From the resulting matrices we generated two types of trees: the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
@@
 In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g.
-\citealt{rozen2005}).
+\cite{rozen2005}).
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what we have to work with.
 
-\subsubsection{Maximum Likelihood}
+\paragraph*{Maximum Likelihood}
 The "best" tree and the "missing-data" trees were inferred using RAxML v.
-8.0.20 \citep{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model (\citealt{tavare1986}; default GTRGAMMA in RAxML v.
-8.0.20; \citealt{Stamatakis21012014}).
-For the morphological data, we used the implemented Markov \textit{k} state model \citep{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
-8.0.20; \citealt{Stamatakis21012014}).
-We used RAxML because it is heavily used in the literature for Maximum Likelihood tree inference \citep[][e.g.]{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014} and is one of the fastest methods available \citep{Stamatakis01102008}.
+8.0.20 \cite{Stamatakis21012014}.
+For the molecular data, we used the GTR + $\Gamma_4$ model \cite{tavare1986} (default GTRGAMMA in RAxML v.
+8.0.20  \cite{Stamatakis21012014}).
+For the morphological data, we used the implemented Markov \textit{k} state model \cite{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
+8.0.20 \cite{Stamatakis21012014}).
+We used RAxML because it is heavily used in the literature for Maximum Likelihood tree inference (e.g.
+\cite{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014}) and is one of the fastest methods available \cite{Stamatakis01102008}.
 
 To measure the support for each branch in our simulated phylogenies we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
-We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
+We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \cite{zanderminimal2004}.
 We repeated this selection until we obtained 50 sets of simulations (i.e.
 50 "complete" and 50 x 125 "missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
-On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \citep{pattengale2010many}.
+On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \cite{pattengale2010many}.
 Using these parameters took \texttildelow 8 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (2.30GHz clock speed nodes).
@@
 
-\subsubsection{Bayesian}
+\paragraph*{Bayesian}
 The "best" tree and the "missing-data" trees were inferred using MrBayes v.
-3.2.1 \citep{Ronquist2012mrbayes}.
+3.2.1 \cite{Ronquist2012mrbayes}.
 We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
-The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \citep{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
-For the morphological data, we used the Markov \textit{k} state model \citep{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
+The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \cite{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
+For the morphological data, we used the Markov \textit{k} state model \cite{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
 We chose these models to be consistent with the parameters used to generate the "complete" matrix.
@@
 Even using these priors, it took \texttildelow 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-The detailed MrBayes parameters are available in the online Appendix 1.
+The detailed MrBayes parameters are available in the Appendix S1.
 
-We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
-We also checked the effective sample size (ESS) on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
-Finally we built a strict majority rule Bayesian consensus tree from the combined chains, excluding the 25\% first iterations as burn-in \citep{Ronquist2012mrbayes}.
+We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \cite{Ronquist2012mrbayes}.
+We also checked the effective sample size (ESS) on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \cite{drummond2006ess}.
+Finally we built a strict majority rule Bayesian consensus tree from the combined chains, excluding the 25\% first iterations as burn-in \cite{Ronquist2012mrbayes}.
+
 %---------------------------------------------
@@
 
-\subsection{4. Comparing topologies}
+\subsection*{4. Comparing topologies}
 We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
-We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa (i.e.
-taxa that frequently change position in different trees; \citealt{kearneyfragmentary2002}).
-We used these two metrics because they illustrate two different aspects of tree topology (see Discussion) but also because their performance in measuring differences in topology is well described \citep{Kuhner04112014} and well implemented \citep{Bogdanowicz2012}.
-We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
+We used the Robinson-Foulds distance \cite{RF1981} to assess the amount of conserved clade positions and the Triplets distance \cite{dobson1975triplets} to assess the amount of wildcard taxa (i.e.
+taxa that frequently change position in different trees \cite{kearneyfragmentary2002}).
+We used these two metrics because they illustrate two different aspects of tree topology (see Discussion) but also because their performance in measuring differences in topology is well described \cite{Kuhner04112014} and well implemented \cite{Bogdanowicz2012}.
+We normalised both metrics using methods described in \cite{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
 These metrics are described in detail below.
 
-\subsubsection{Robinson-Foulds distance}
-The Robinson-Foulds distance \citep{RF1981}, or "path difference", measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees (\citealt{RF1981}; see online Appendix 2 for calculation details).
+\paragraph*{Robinson-Foulds distance}
+The Robinson-Foulds distance \cite{RF1981}, or "path difference", measures the number of shared clades across two trees.
+The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} (see Appendix S2 for calculation details).
 This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
 This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
-We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons \citep{Bogdanowicz2012}.
-This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see online Appendix 2 for the calculation details).
+We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons \cite{Bogdanowicz2012}.
+This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see Appendix S2 for the calculation details).
 Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
 
-\subsubsection{Triplets distance}
-The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees (\citealt{critchlowthe1996}; see online Appendix 2 for calculation details).
+\paragraph*{Triplets distance}
+The Triplets distance \cite{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees \cite{critchlowthe1996} (see Appendix S2 for calculation details).
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
@@
 
-\subsubsection{Paired tree comparisons}
+\paragraph*{Paired tree comparisons}
 \label{tree_comparisons}
-For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance for each tree inference method.
+For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \cite{Bogdanowicz2012} resulting in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance for each tree inference method.
 Also, to take into account the uncertainty of tree inference, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 trees from the posterior tree distribution of the Bayesian analysis for the "best" trees, and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees).
@@
 We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v.
-3.1 \citep{hdrcde}.
+3.1 \cite{hdrcde}.
 
-\subsection{Testing the effects of the missing data parameters on topological recovery}
+\subsection*{Testing the effects of the missing data parameters on topological recovery}
 Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the "best" tree topology in a Total Evidence framework.
@@
 
-We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions (\citealt{Bhattacharyya}; see online Appendix  for calculation details).
+We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \cite{Bhattacharyya}.
+The Bhattacharyya Coefficient is the probability of overlap between two distributions \cite{Bhattacharyya} (see Appendix S2 for calculation details).
 Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
@@
 
+\begin{figure}[h]
+\caption{{\bf Bhattacharyya Coefficient calculation outline 1.}
+ A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.
+Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
+\label{Fig_Bhattacharyya_Coefficients1}
+\end{figure}
+
 To assess the effect of the different tree inference and uncertainty methods (i.e.
@@
 
+\begin{figure}[h]
+\caption{{\bf Bhattacharyya Coefficient calculation outline 2.}
+ A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+\textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
+Maximum Likelihood or Bayesian).
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$.
+Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
+\label{Fig_Bhattacharyya_Coefficients2} 
+\end{figure}
+
 %---------------------------------------------
@@
 
-\section{Results}
+\section*{Results}
 As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the "best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
@@
 
-\subsection{Individual effects of missing data parameters}
+\begin{figure}[]
+\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).}
+ The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+\label{Fig_Results-permeth_perparam} 
+\end{figure}
+
+\begin{figure}[]
+\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).}
+ The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+\label{Fig_Results-global_perparam}
+\end{figure}
+
+\subsection*{Individual effects of missing data parameters}
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; online Appendix Tables S5, S6 and S7).
-However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; online Appendix Tables S5, S6 and S7).
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7).
+However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7).
 
-\subsection{Combined effect of missing data parameters}
+\subsection*{Combined effect of missing data parameters}
 As expected, our ability to recover the "best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; online Appendix Tables S2, S3 and S4).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; Tables S2, S3 and S4).
 Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
+
+\begin{figure}[]
+\caption{{\bf The effects of missing data on topological recovery using Bayesian consensus trees.}
+ Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
+}
+\label{Fig_Results-paircomp_within}
+\end{figure}
 
@@
 For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; Table ~\ref{Tab_Results-Difference_methods}).
-However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and online Appendix Fig.
+However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.
 S2).
 
-%---------------------------------------------
-%
-%       DISCUSSION
-%
-%---------------------------------------------
-
-\section{Discussion}
-
-Our results show that the ability to recover the "best" tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used.
-However, these factors affected topological recovery in different ways and to different extents.
-Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
-Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam}, Fig.~\ref{Fig_Results-paircomp_within}, Table ~\ref{Tab_Results-Difference_methods}).
-As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
-
-\subsection{Individual effects of missing data parameters}
-\subsubsection{Missing data for living taxa ($M_{L}$)}
-When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
-Because living taxa still have molecular characters available for phylogenetic inference (see mMthods),even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
-However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
-
-\subsubsection{Missing data for fossil taxa ($M_{F}$)}
-When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this also reduces the probability of morphological characters for fossil taxa overlapping with the ones for living taxa.
-This can lead to difficulties for the placement of certain taxa in the tree.
-However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
-decrease of Normalised Triplets distance) with increasing missing data in this parameter, clade conservation (i.e.
-Normalised Robinson-Foulds distance) is still relatively good (mode = 0.72) when the proportion of missing data is high ($M_{F}$ = 75\%).
-
-The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds distance) but greater on the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
-When the fossil taxa have less data (e.g.
-$M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils.
-Therefore a majority consensus position is unlikely to exist (i.e.
-every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
-This does not decrease the Normalised Robinson-Foulds distance, because clades are conserved, but decreases the Normalised Triplets distance because the fossils act as wildcard taxa.
-Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50).
-In other words, the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree, where the fossil will be placed with lower confidence at a higher taxonomic level.
-We argue that using the Bayesian consensus tree topology is preferable because it is more conservative \cite[e.g][]{pattinsonphylogeny2014}.
-
-\subsubsection{Missing morphological characters ($M_{C}$)}
-Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the "best" tree topology.
-We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
-Conversely, removing 10\% of morphological characters ($M_{C}$) genuinely removes 10\% of data in the matrix.
-However, the effect of removing characters on the ability to recover the "best" tree topology is of the same order of magnitude as for the other two parameters (Fig.~\ref{Fig_Results-permeth_perparam}).
-We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.
-
-Additionally, the number of morphological characters determines the size the matrix.
-This can affect our ability to recover the "best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
-The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \citep{wagner2000}.
-The size of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
-Therefore, part of the decrease of the Normalised Robinson-Fould distance and the Normalised Triplets distance in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-However, although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
-Therefore, our simulations reflect realistic parameters.
-Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{wrightbayesian2014}.
-However, the use of probabilistic methods (i.e.
-Maximum Likelihood or Bayesian) and the M\textit{k} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.
-
-\subsection{Combined effect of missing data parameters}
-As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have few missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a lot of missing data (i.e.
-$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-global_perparam}).
-However, it is important to notice that the effect of each parameter is not additive.
-The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and/or $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).
-
-\subsection{Effects of tree inference methods}
-Variation in our ability to recover the "best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{k} model \citep{lewisa2001} over cladistic methods (\citealt{wrightbayesian2014}; but see \citealt{spencerefficacy2013}).
-However, this is the first study, to our knowledge, to compare the performance of the M\textit{k} model \citep{lewisa2001} for recovering the "best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-Our results show that the topology of the Bayesian consensus tree is always closer to the "best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}).
-As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
-This may also be because the "best" Bayesian consensus trees are not completely resolved, thus will always be more similar to the "missing data" trees than a completely resolved tree like the "best" Maximum Likelihood tree.
-However, we minimized the probability of unresolved "best" trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
-
-It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the "best" tree topology than Maximum Likelihood and Bayesian consensus trees.
-This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-permeth_perparam}).
-This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" trees and a set of the "best" trees involved 1000 random pairwise comparisons rather than just one.
-This will obviously add noise to the results for these methods.
-
-\subsection{Practical implications}
-Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
-Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "best" tree topology.
-We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections.
-Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries.
-However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
-Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "best" tree topology (Fig.~\ref{Fig_Results-paircomp_within}).
-Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
-One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{morphobank} \citep{morphobank}, so that no one individual collects all the data.
-
-Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating (\citealp{ronquista2012,BEASTmaster}; although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution).
-Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see above).
-Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses \citep[e.g.][]{jetzthe2012}.
-
-%---------------------------------------------
-%
-%       CONCLUSION
-%
-%---------------------------------------------
-
-\section{Conclusion}
-Missing data in Total Evidence matrices is not a problem for recovering the "best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
-When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
-Therefore we advise filling as many gaps in Total Evidence matrices as possible.
-Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.
-Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the "best" tree topology than the Maximum Likelihood trees.
-Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
-The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
-
-%---------------------------------------------
-
-\section{Supplementary material}
-Supplementary material (R code, analyses, supplementary methods and full results) can be found in the Dryad data repository at http://dx.doi.org/10.5061/dryad.XXXX.
-
-\section{Acknowledgments}
-Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Nick Matzke, and April Wright for useful comments on our simulation protocol and manuscript.
-Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
-Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.
-
-\section{Funding}
-This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
-
-\bibliographystyle{sysbio}
-\bibliography{References}
-
-%---------------------------------------------
-%Tables
-%---------------------------------------------
-\newpage
-
-\begin{landscape}
-\begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" trees using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
+\begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+\caption{{\bf Bhattacharyya Coefficients of the pairwise method comparisons.}
+Each comparison which corresponds to the normalised distance between the "best" tree and the "missing data" trees using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
 Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
@@
 \centering
-\begin{tabular}{lccccccc}
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
   \hline
@@
 \label{Tab_Results-Difference_methods}
+\end{adjustwidth}
 \end{table}
-\end{landscape}
 
-%---------------------------------------------
-%Figures Caption
-%---------------------------------------------
+\section*{Discussion}
 
-\begin{figure}
-\caption{Protocol outline.
-(1) We randomly generated a birth-death tree (the "true" tree) and used it to infer a matrix with no missing data (the "complete" matrix).
-(2) We removed data from the morphological part of the "complete" matrix resulting in 125 "missing-data" matrices.
-(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
-(4) We compared the "missing-data" trees to the "best" tree.
-We repeated these four steps 50 times.}
-\label{Fig_Outline}
-\end{figure}
+Our results show that the ability to recover the "best" tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used.
+However, these factors affected topological recovery in different ways and to different extents.
+Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
+Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam}, Fig.~\ref{Fig_Results-paircomp_within}, Table ~\ref{Tab_Results-Difference_methods}).
+As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters (e.g.
+\cite{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}).
 
-\begin{figure}
-\caption{Bhattacharyya Coefficient calculation outline.
-A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
-$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
-The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.
-Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
-\label{Fig_Bhattacharyya_Coefficients1}
-\end{figure}
+\subsection*{Individual effects of missing data parameters}
+\paragraph*{Missing data for living taxa ($M_{L}$)}
+When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
+Because living taxa still have molecular characters available for phylogenetic inference (see mMthods),even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
+However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
 
-\begin{figure}
-\caption{Bhattacharyya Coefficient calculation outline.
-A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
-$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
-\textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
-Maximum Likelihood or Bayesian).
-The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$.
-Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
-\label{Fig_Bhattacharyya_Coefficients2} 
-\end{figure}
+\paragraph*{Missing data for fossil taxa ($M_{F}$)}
+When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this also reduces the probability of morphological characters for fossil taxa overlapping with the ones for living taxa.
+This can lead to difficulties for the placement of certain taxa in the tree.
+However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
+decrease of Normalised Triplets distance) with increasing missing data in this parameter, clade conservation (i.e.
+Normalised Robinson-Foulds distance) is still relatively good (mode = 0.72) when the proportion of missing data is high ($M_{F}$ = 75\%).
 
-\begin{figure}
-\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).
-The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-\label{Fig_Results-permeth_perparam} 
-\end{figure}
+The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds distance) but greater on the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
+When the fossil taxa have less data (e.g.
+$M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils.
+Therefore a majority consensus position is unlikely to exist (i.e.
+every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
+This does not decrease the Normalised Robinson-Foulds distance, because clades are conserved, but decreases the Normalised Triplets distance because the fossils act as wildcard taxa.
+Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50).
+In other words, the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree, where the fossil will be placed with lower confidence at a higher taxonomic level.
+We argue that using the Bayesian consensus tree topology is preferable because it is more conservative (e.g.
+\cite{pattinsonphylogeny2014}).
 
-\begin{figure}
-\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
-The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-\label{Fig_Results-global_perparam}
-\end{figure}
+\paragraph*{Missing morphological characters ($M_{C}$)}
+Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the "best" tree topology.
+We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
+Conversely, removing 10\% of morphological characters ($M_{C}$) genuinely removes 10\% of data in the matrix.
+However, the effect of removing characters on the ability to recover the "best" tree topology is of the same order of magnitude as for the other two parameters (Fig.~\ref{Fig_Results-permeth_perparam}).
+We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.
 
-\begin{figure}
-\caption{The effects of missing data on topological recovery using Bayesian consensus trees.
-Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
-}
-\label{Fig_Results-paircomp_within}
-\end{figure}
+Additionally, the number of morphological characters determines the size the matrix.
+This can affect our ability to recover the "best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
+The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \cite{wagner2000}.
+The size of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
+Therefore, part of the decrease of the Normalised Robinson-Fould distance and the Normalised Triplets distance in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+However, although these matrices are an order of magnitude smaller than some published matrices (e.g.
+\cite{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices (e.g.
+\cite{kellymolecular2014, sallam2011craniodental}).
+Therefore, our simulations reflect realistic parameters.
+Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \cite{wrightbayesian2014}.
+However, the use of probabilistic methods (i.e.
+Maximum Likelihood or Bayesian) and the M\textit{k} model \cite{lewisa2001} has been previously demonstrated to partially resolve this issue \cite{wrightbayesian2014}.
 
-%---------------------------------------------
-%Figures
-%---------------------------------------------
+\subsection*{Combined effect of missing data parameters}
+As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have few missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a lot of missing data (i.e.
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-global_perparam}).
+However, it is important to notice that the effect of each parameter is not additive.
+The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and/or $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).
 
-\setcounter{figure}{0} \renewcommand{\thefigure}{\arabic{figure}}
+\subsection*{Effects of tree inference methods}
+Variation in our ability to recover the "best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{k} model \cite{lewisa2001} over cladistic methods \cite{wrightbayesian2014} (but see \cite{spencerefficacy2013}).
+However, this is the first study, to our knowledge, to compare the performance of the M\textit{k} model \cite{lewisa2001} for recovering the "best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
+Our results show that the topology of the Bayesian consensus tree is always closer to the "best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}).
+As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
+This may also be because the "best" Bayesian consensus trees are not completely resolved, thus will always be more similar to the "missing data" trees than a completely resolved tree like the "best" Maximum Likelihood tree.
+However, we minimized the probability of unresolved "best" trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
 
-\newpage
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[keepaspectratio=true]{Simulations_outline-BW.pdf}
-\caption{ }
-\end{figure}
+It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the "best" tree topology than Maximum Likelihood and Bayesian consensus trees.
+This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-permeth_perparam}).
+This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" trees and a set of the "best" trees involved 1000 random pairwise comparisons rather than just one.
+This will obviously add noise to the results for these methods.
 
-\newpage
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=0.5\textwidth]{Bhattacharyya_Coefficients1.pdf}
-\caption{ }
-\end{figure}
+\subsection*{Practical implications}
+Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
+Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "best" tree topology.
+We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections.
+Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries.
+However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils (e.g.
+\cite{nithe2013}).
+Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "best" tree topology (Fig.~\ref{Fig_Results-paircomp_within}).
+Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
+One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{morphobank} \cite{morphobank}, so that no one individual collects all the data.
 
-\newpage
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=1\textwidth]{Bhattacharyya_Coefficients2.pdf}
-\caption{ }
-\end{figure}
+Another practical implication of our results regards the tree inference methods.
+Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \cite{ronquista2012,BEASTmaster}; although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution).
+Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see above).
+Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses (e.g.
+\cite{jetzthe2012}).
 
-\newpage
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=1\textwidth]{AllMethods-RF_Tr-colour.pdf}
-\caption{ }
-\end{figure}
+\section*{Conclusion}
+Missing data in Total Evidence matrices is not a problem for recovering the "best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
+When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
+Therefore we advise filling as many gaps in Total Evidence matrices as possible.
+Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.
+Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the "best" tree topology than the Maximum Likelihood trees.
+Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
+The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
 
-\newpage
-\begin{figure}[!htbp]
-\centering
-    \includegraphics[width=1\textwidth]{ML_Baycon-AllParam-RF_Tr-BW.pdf}
-\caption{ }
-\end{figure}
+\section*{Supporting Information}
 
-\newpage
-\begin{figure}
-\centering
-    \includegraphics[width=1\textwidth]{PairwiseComp-Baycon-RF_Tr-colour.png}
-\caption{ }
-\end{figure}
+% Include only the SI item label in the subsection heading. Use the \nameref{label} command to cite SI items in the text.
+\subsection*{Supplementary informations}
+\label{S1_Appendix1}
+\noindent{\bf Appendix S1 Tree Inference.} Supplementary information on the Tree Inference methods.
+
+\label{Figure S1}
+\noindent{\bf Figure S1 proportion of morphological characters states.}
+
+\label{S2_Appendix2}
+\noindent{\bf Appendix S2 Tree Comparisons.} Supplementary information on the Tree Comparison methods.
+
+\label{S2_Appendix3}
+\noindent{\bf Appendix S3 Additional results.}.
+
+\label{Figure S2}
+\noindent{\bf Figure S2 Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey).}
+
+\label{Figure S3}
+\noindent{\bf Figure S3 Effects of missing data on topological recovery using Maximum Likelihood trees.}
+
+\label{Figure S4}
+\noindent{\bf Figure S4 Effects of missing data on topological recovery using Maximum Likelihood bootstrap trees.}
+
+\label{Figure S5}
+\noindent{\bf Figure S5 Effects of missing data on topological recovery using Bayesian posterior tree distribution.}
+
+\label{Table S1}
+\noindent{\bf Table S1 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method.}
+
+\label{Table S2}
+\noindent{\bf Table S2 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method for the $M_{L}$ parameter.}
+
+\label{Table S3}
+\noindent{\bf Table S3 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method for the $M_{F}$ parameter.}
+
+\label{Table S4}
+\noindent{\bf Table S4 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method for the $M_{C}$ parameter.}
+
+\label{Table S5}
+\noindent{\bf Table S5 Bhattacharyya Coefficients of the pairwise method comparisons for the $M_{L}$ parameter.}
+
+\label{Table S6}
+\noindent{\bf Table S6 Bhattacharyya Coefficients of the pairwise method comparisons for the $M_{F}$ parameter..} Bhattacharyya
+
+\label{Table S7}
+\noindent{\bf Table S7 Bhattacharyya Coefficients of the pairwise method comparisons for the $M_{C}$ parameter..} Bhattacharyya
+
+% Do NOT remove this, even if you are not including acknowledgments.
+\section*{Acknowledgments}
+Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Nick Matzke, and April Wright for useful comments on our simulation protocol and manuscript.
+Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
+Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.
+%This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
+
+\section*{Author Contributions}
+Conceived and designed the experiments: TG NC. Ran the analysis: TG. Analyzed the data: TG. Wrote the paper: TG NC.
+
+\nolinenumbers
+
+\bibliographystyle{plos2009}
+\bibliography{References}
 

### 2015-04-25 Manuscript/PLoSONEresponse/response.tex
PLoS response init

\documentclass[11pt]{letter}
\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[colorlinks = true, urlcolor = blue]{hyperref} %hyperlinks
\usepackage[osf]{mathpazo}
\signature{Thomas Guillerme \\ Natalie Cooper}
\address{Zoology building \\ Trinity College Dublin \\ Dublin 2, Ireland \\ \\ guillert@tcd.ie}
\longindentation=0pt
\begin{document}

\begin{letter}{}

RE: PLOS ONE Decision: PONE-D-15-05911 - [EMID:ffa50492e76c2059]

Dear Dr Fontaneto

Please find enclosed our resubmission of our manuscript @@@[title]. We include below details of how we have dealt with the reviewer's comments.

We would however, like to draw your attention to a couple of points. Firstly, while we understand the difficulties of obtaining reviewers for papers, we were very disappointed to have the previous decision made based on just one review. As pointed out in your email, this reviewer responded immediately. We suspect they rushed their assessment because they make several comments about our paper that are completely incorrect, and would have been noticed in a thorough reading of the manuscript. For example, the reviewer says that our data are not available when we clearly state on @@@[page and lines] that all the data are deposited in figshare @@@[link]. We suspect that had a second reviewer been found the decision might have been different. %maybe recheck that

Secondly, the reviewer clearly misunderstood the meaning of the term ``statistics". Statistics does not mean that p-values need to be used. In this case we could have applied t-tests, and given the reviewer the p-values they request, but this would only tell us that the means of the distributions are significantly different, not that the distributions differ. We chose instead to use the Bhattacharya coefficient, which admittedly is an unfamiliar statistic to many but we feel we made a lot of effort to explain how this metric works in the text (and this method was suggested to us by a professional statistician!). The reviewer however appears to have refused to engage with this and instead has fixated on the lack of p-values.

\textcolor{ForestGreen}{From: PLOS ONE <em@editorialmanager.com<mailto:em@editorialmanager.com> \\
To: Thomas Guillerme <guillert@tcd.ie<mailto:guillert@tcd.ie> \\
Reply-To: PLOS ONE <plosone@plos.org<mailto:plosone@plos.org> \\
Date: 20 April 2015 12:25:39 IST \\ \\
Subject: PLOS ONE Decision: PONE-D-15-05911 - [EMID:ffa50492e76c2059] \\
PONE-D-15-05911 \\
Effects of missing data on topological inference using a Total Evidence approach \\
PLOS ONE \\ \\
Dear Mr Guillerme, \\
Thank you for submitting your manuscript to PLOS ONE. After careful consideration, we have decided that your manuscript does not meet our criteria for publication and must therefore be rejected. \\
First of all, apologies for the long delays. We immediately received the comments from the first reviewer and in the meanwhile another reviewer agreed to send us comments but failed to do so, even after several reminders. Thus, I now have to reach a decision with only one reviewer.
Unfortunately, the comments from the only reviewer we have are rather negative and the problems that have been highlighted too many. Please, go through all the comments of the reviewer: I think that they all make sense and should be addressed and solved to produce a more convincing and less ambiguous story. \\
If you think that the issues can be solved by new analyses, we will be happy to reconsider a new version of the manuscript as a re-submission. \\
I am sorry that we cannot be more positive on this occasion, but hope that you appreciate the reasons for this decision. \\
Yours sincerely, \\ \\
Diego Fontaneto \\ 
Academic Editor \\
PLOS ONE}

\textcolor{blue}{Reviewers' comments:}

\textcolor{blue}{Reviewer's Responses to Questions}

\textcolor{blue}{Comments to the Author}

\textcolor{blue}{-------------------------}

\textcolor{blue}{1. Is the manuscript technically sound, and do the data support the conclusions?}

\textcolor{blue}{The manuscript must describe a technically sound piece of scientific research with data that supports the conclusions. Experiments must have been conducted rigorously, with appropriate controls, replication, and sample sizes. The conclusions must be drawn appropriately based on the data presented.}

\textcolor{blue}{Reviewer $\#$ 1: No}

See our detailed response below.

\textcolor{blue}{-------------------------}

\textcolor{blue}{2. Has the statistical analysis been performed appropriately and rigorously?}

\textcolor{blue}{Reviewer $\#$ 1: No}

See our detailed response below.

\textcolor{blue}{-------------------------}

\textcolor{blue}{3. Does the manuscript adhere to the PLOS Data Policy?}

\textcolor{blue}{Authors must follow the PLOS Data policy, which requires authors to make all data underlying the findings described in their manuscript fully available without restriction. Please refer to the author’s Data Availability Statement in the manuscript. All data and related metadata must be deposited in an appropriate public repository, unless already provided as part of the submitted article or supporting information. If there are restrictions on the ability of authors to publicly share data—e.g. privacy or use of data from a third party— these reasons must be specified.}

\textcolor{blue}{Reviewer $\#$ 1: No}

All the data generated by the simulation is directly available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare}. Note that along with the data we share the full code used for repeating the submitted manuscript, from simulating the data to generating the figures and compiling the manuscript (available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{GitHub}. All this code is provided with a detailed description of each step that we are happy to develop if needed. However, we do apologise since we forgot to add this as a separated section in the supporting information section. We added the following part to the new manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
\textbf{Data reproducibility and availability.} All the code used in this analysis is available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{\color{blue}{GitHub}} with some information on how to use the various functions. Additionally all the simulated data is available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{\color{blue}{FigShare}}.
\end{minipage}

\textcolor{blue}{-------------------------}

\textcolor{blue}{4. Is the manuscript presented in an intelligible fashion and written in standard English?}

\textcolor{blue}{PLOS ONE does not copyedit accepted manuscripts, so the language in submitted articles must be clear, correct, and unambiguous. Any typographical or grammatical errors should be corrected at revision, so please note any specific errors here.}

\textcolor{blue}{Reviewer $\#$ 1: Yes}

\textcolor{blue}{-------------------------}

\textcolor{blue}{5. Review Comments to the Author}

\textcolor{blue}{Please use the space provided to explain your answers to the questions above. You may also include additional comments for the author, including concerns about dual publication, research ethics, or publication ethics. (Please upload your review as an attachment if it exceeds 20,000 characters)}

\textcolor{blue}{Reviewer $\#$ 1: This paper tackles an interesting questions. Despite some problems noted below, the experimental design seems adequate.}

\textcolor{blue}{Unfortunately the data analysis performed is simply the calculation of a number of summary statistics about the overlap between the estimates returned for different simulation conditions. I suspect that if they authors take some time to do statistics (calculate some P-values, confidence intervals, or take a Bayesian perspective), they'll be able to make a statistical case for most of the points that they are trying to make. But I just don't think that a simulation comparison should be publishable without *some* statistical analysis (beyond the calculation of summaries).}

We are slightly concerned by the reviewer's misconception of the term ``statistics" and suspicions (``I suspect that if they authors take some time to do statistics (calculate some P-values, confidence intervals, or take a Bayesian perspective") since: (1) t-tests and p-values are useful in some cases but have several limitations especially when one compares distributions that are not normal (in this study, we compared the modal value of topological differences, making any test comparing the mean value useless) and can be biased towards small values (i.e. significant ones) when the sample sizes are big; (2) concerning the confidence intervals, we actually did systematically calculate them; ans (3) we are not convinced by a Bayesian approach for comparing distributions.

\textcolor{blue}{The bias in the analysis in favor of Bayesian methods makes the recommendation to use Bayesian methods less satisfying.}

Our results did indeed biased us in favour of using a Bayesian approach to fix topology since our results clearly indicate that it is less likely to generate a wrong topology using a Bayesian approach (see Figure 4 and 5 and discussion line 472 to 496 in the first submission manuscript).

\textcolor{blue}{The use of Mk rather than Mkv makes the study less relevant to real data analyses.}

The Mk model assumes that some morphological characters \textbf{can} be constant and the Mkv model assumes that \textbf{all} the characters are variable. Although the reviewer is right that real data analyses do not contain invariable (constant) morphological characters, we designed our simulations to be a generalisation of real data. Because we sampled the morphological characters rate from a gamma distribution with $\alpha$=0.5, some morphological characters had only a few state changes accross the matrix. When removing the data, there was a probability for these characters to be constant (e.g. if all the taxa in the matrix appart from one had a character state 0 for a given character and only one taxa had a state 1 for the same character, if that taxa had no data available for that character, then the character appeared to be constant in the matrix). Note that both RAxML and MrBayes do not distinguish between both models since the Mkv model is just a particular case of the Mk model and are not distinguished either in the original publication that we cite in the manuscript (Lewis 2001). We've changed the mentions of the Mk model to Mkv model throughout the new manuscript.

%Add note that we changed it along the text. RAxML and MrBayes are using default.

\textcolor{blue}{Many of the equations are incorrect.}

See our detailed response below.

\textcolor{blue}{If there was a mention of the data being deposited, I missed it. I apologize if I did miss it.}

See our comment above on data availability and reproducibility.

\textcolor{blue}{-------------------------}

\textcolor{blue}{Some specific concerns:}

\textcolor{blue}{line 25 and elsewhere: incorrect smart quotes in ``best", ``classical"...}

We fixed the smart quotes in the new manuscript.

\textcolor{blue}{line 88 ``used it to infer a matrix" $->$ ``used it to simulate a matrix". ``inference" is used in place of simulation elsewhere.}

We changed ``infer" into ``simulate" both at line 88 and in the Figure 1 caption.

\textcolor{blue}{line 112: does diversitree just stop simulating when it reaches the correct number of tips, or does it deal with the biases that can result from doing that? (Stadler 2011 is the ref for a discussion of this issues, I believe)}

%TO CHECK!
Lines 116 to 118 in the previously submitted manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
To avoid biasing our simulations towards either living or fossil taxa and to make each simulation comparable, we implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa.
\end{minipage}

Lines @@@ in the new manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
To avoid biasing our simulations towards either living or fossil taxa and to make each simulation comparable, we implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa. In practice, we generated $x$ birth-death trees using the \texttt{tree.bd} function from the \texttt{diversitree R} package (FitzJohn) until we obtained $y$ trees with the right number of living and fossil taxa.
\end{minipage}

\textcolor{blue}{line 113: `` rates from a uniform distribution " with what upper and lower bounds?}

Both birth ($\lambda$) and death ($\mu$) parameters are respectively speciation and extinction probability (bounded between 0 and 1). We added this precision in the new version of the manuscript:

Lines 112 to 114 in the previously submitted manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution but maintaining $\lambda$ $>$ $\mu$.
\end{minipage}

Lines @@@112 to @@@114 in the new manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform probability distribution (between 0 and 1) but maintaining $\lambda$ $>$ $\mu$.
\end{minipage}

\textcolor{blue}{line 118: wih 25 living and 25 fossil taxa. Are the fossils just put at the extinction point for the extinct lineages?}

Yes. We clarified as follow:

Lines 118-119 in the previously submitted manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
[We] select only trees with 25 living and 25 fossil taxa.
\end{minipage}

Lines @@@112 to @@@114 in the new manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
[We] select only trees with 25 living and 25 fossil taxa. The fossil taxa where considered as unique points at the end of extinct lineages.
\end{minipage}

\textcolor{blue}{line 110 - line 120. The procedure for generating a tree with fossils is a valid procedure. But it is idiosyncratic, which will make these results harder to relate to other work. Easy of comparability with other work is one of the advantages of using formally characterized models. In this case, I think that using the TreeSim package http://cran.r-project.org/web/packages/TreeSim/TreeSim.pdf would have led to a study with higher impact. This part of the methods section would be reduced to stating the parameter values chosen, and other workers could study similar tree generation simulations. As it stands, it is unlikely that other workers will every generate trees in precisely the way that the authors do. Thus, there will always be a ``however, Guillerme and Cooper generated their trees under a different model..." caveat necessary when discussing this work in the context of other work.}

We do not agree with this reviewer's comment. Even if our study might be idiosyncratic, we beleive it is not due to the use of the \texttt{diversitree} package instead of the \texttt{TreeSim} package. Both perform in generating birth-death trees correctly and are well curated and maintained. However, we chose to use the \texttt{diversitree} package rather than the \texttt{TreeSim} package exactly for making the results of our study more applicable to other ones. In fact, the publication associated with the diversitree package (FitzJohn 2012, Methods in Ecology and Evolution) has been cited three times more (155 cites on Google Scholar) than the publication associated with the \texttt{TreeSim} package (Stadler 2011, Systematic Biology - 55 cites on Google Scholar).

\textcolor{blue}{line 126 ``random base frequencies" is too vague. Dirichlet(1, 1, 1, 1) perhaps?}

We clarified as follow:

Lines 125 to 126 in the previously submitted manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
The matrix [...] was generated [...] using the HKY model with random base frequencies[...]
\end{minipage}

Lines @@@112 to @@@114 in the new manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
The matrix [...] was generated [...] using the HKY model with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1)[...]
\end{minipage}

\textcolor{blue}{Do the simulations apply ascertainment biases as described by Lewis' Mkv (2001) model? It seems like they should... At minimum, the point should be discussed.}

We are not entirely sure if the reviewers refers to a sampling bias or to the problem link to constant characters described in Lewis (2001) (acquisition bias). If the reviewers are referring to the second (acquisition bias), both phylogenetic inference software we used (RAxML and MrBayes) are correcting for the this bias (-K MK  see Supplementary information 1 for the full list of commands). Also, the effect of the acquisition bias is to inflate branch length which is an aspect of the phylogeny we didn't explored in this study (focusing on topology). See response below to the comment \color{blue}{\textit{The use of Mk rather than Mkv makes the study less relevant to real data analyses.}}.

\textcolor{blue}{line 171 - 177: the authors need to explain the order of operations for the $M_F$ vs $M_C$ masking. It makes a difference. If you mask out $M_F$ percentage of cells, and then mask $M_C$ percentage of columns, it is possible (indeed very probable) that the percentage of missing data in fossil characters that remain will not be $M_F$ after some of the morphological columns are removed. If you mask by $M_C$ first, and then $M_F$, the resulting matrices will always have $M_F$ percentage missing data (modulo rounding error) in the simulated realizations.}

We clarified this point ($M_C$ comes second). %change in manuscript
Also, we clarified the potential confusion between the amount of data removed in one parameter ($M_F$ = 10\% for example) and the amount of missing data in the morphological matrix (5\%). %change in manuscript

\textcolor{blue}{Header for 3. should be ``building phylogenetic estimates" or ``estimating phylogenies" because the previous section described how the phylogenies and character data were ``built."}

We changed the Header 3 to infering phylogenies. %change in manuscript

\textcolor{blue}{RAxML searching: it would be nice to see some searches started from the true tree. This is clearly not an option for real data sets, but it is helpful in simulation studies because it lets the reader determine if the results could be an artifact of insufficient tree searching.}

%Probably no impact since it's not using an MCMC. But should we try it? (approx. one to two (max) month calculation).

\textcolor{blue}{Filtering the data sets for strong bootstrap support under the full data analysis surely introduces some bias in favor of the no-missing data analyses.}

We implemented this step of keeping trees with a minimum total median bootstrap to make the topological comparisons more robust. If the no-missing data tree has low node support, it is likely that the changes in topology for the missing data trees are not linked to our focal question (the effect of removing characters, living taxa or fossils) but just to the fact that low supported nodes can be just random.

\textcolor{blue}{MrBayes: using strongly informative priors for the true parameters does more than speed up the searching, it represents an ``unfair" advantage (an option that would not be availble in real data analysis in which one does not know the true values) for the Bayesian method. This makes the studies report of superior behavior of Bayesian methods less persuasive.}

We would like to emphasize that we are looking at topology and morphological data, our priors are on nucleotid frequencies and on TT ratio. This remains unchanged in all our trees. Also, we would like to emphasize on the ``speed up" part that was mandatory since the analysis took 1.5 CPU century including our ``unfair" methodological choices. We had to choose between extracting a global pattern (more faster chains) rather than some specific analysis (fewer slower chains).

\textcolor{blue}{line 254: RF measures the difference between the number of clades and the twice number of shared clades across two trees. The statement that it ``measures the number of shared clades across two trees" should be revised.}

%check

\textcolor{blue}{line 258: RF is 0 (not 1) when the trees are identical. It's maximum (for 2 rooted binary trees) is 2(n-2) not n-2.}

%check

\textcolor{blue}{The ``scaling" to produce the Normalized RF ``distance" is confusing. It is also inverting the sign of the distance, so it has become a measure of similarity.}

%check

\textcolor{blue}{line 275: why isn't the upper bound of the triples distance n choose 3 (instead of n choose 4). I think you got the formula from an unrooted quartet distance.}

%check

\textcolor{blue}{Using the BC stat may be similar in spirit to using a t-test, but it is not equivalent.}

We clarified as follow:

Lines 308 to 309 in the previously submitted manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
Note that this is equivalent to performing a two-sided t-test.
\end{minipage}

Lines @@@112 to @@@114 in the new manuscript:

\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
Note that this is similar to performing a two-sided t-test for comparing two distribution.
\end{minipage}

\textcolor{blue}{Figure 3: B is confusing. If both curves are probability distributions, then they both have to integrate to 1. So B(y) can't cover just a subset of the area covered by B(x).}

%Fix probability distributions -> "just" distributions

\textcolor{blue}{``Combined effect of missing data parameters" section ANOVA or multiple regression are the appropriate analyses to tease apart multiple interacting factors. Here the authors opt to describe results without statistical tests, which is disappointing.}

This comment links back to the reviewer's misconception of the term ``statistics" and suspicions. ANOVA or multiple regression analysis might the appropriate analyses in some cases but we believe it isn't here. %Not comparing means, non normal distributions. ANd non-parametric version (e.g. Kruskal Wallis) not approriate either.

\textcolor{blue}{line 398: ``mMthods" -> ``Methods"}

We fixed this typo.

\textcolor{blue}{line 420: contra the statement here, a consensus collapsing to a polytomy will affect the RF distance (could go up or down, but it will change).}

%Change to "affect less" or something like that.

\textcolor{blue}{-------------------------}

\textcolor{blue}{Supplement 2:}
\textcolor{blue}{more confusion about the min RF here. Also note that if you are counting the entire leaf set as a clade (e.g the star tree having N=1), then the number of clades in rooted binary tree is n-1 (not n-2). Equation 3 has an extra ``-2" As stated it implies that the max RF distance for n=3 is 0, when the correct answer is 2.}

%check

\textcolor{blue}{The definition of the scaled RF implies that it is never positive (sinc the numerator cannot be positive and the denominator is positive).}

%check

\textcolor{blue}{The decision to use Yule trees as the base line for the mean distance is not explained, and seems odd given that the true tree are not Yule trees. I'm not sure that it makes much difference, but it is odd.}

%Because we follow Bogdanowicz and use the difference between random (NULL) trees to scale.

\textcolor{blue}{The NTS can only -infinity for distances that have the property that the mean distance between random Yule trees is 0. This is not true of either distance metric here. So the real range of values is (mean-max)/mean for both the RF and triples distance.}

%check

\textcolor{blue}{Note that JavaScript is a language, so it a bit confusing to refer to a Java program as a ``Java script."}

%fix

\textcolor{blue}{Equation (6) is not correct. You shouldn't set the summation sign as the left hand side of an equation because the notation has a standard definition. More importantly, the $\#$ of triples is the number of ways of drawing 3 (unordered) leaves from the leaf set. So it is n choose 3 not n choose 4.}

%check

\textcolor{blue}{Equation (7) is true for a three-leaf tree with a polytomy occurring with probability = 1/4 and the other 3 trees being equiprobable. I don't think that is is true for any other weighting of trees for other numbers of taxa. It might be, but the authors must show this.}

%check

\textcolor{blue}{I don't think that equations 8, 9, or 10 are correct for any case other than the 3 taxon case (though I do note that the authors are using n choose 3 here).}

%check

\textcolor{blue}{Equations 13 and 14. It seems easier to just drop the summation sign here, and just define $a_i$ and $b_i$ to the right hand side of these equations.}

%check

\textcolor{blue}{-------------------------}

\textcolor{blue}{6. If you would like your identity to be revealed to the authors, please include your name here (optional).}

\textcolor{blue}{Your name and review will not be published with the manuscript.}

\textcolor{blue}{Reviewer $\#$ 1: No thanks.}

%Fair enough!

\end{letter}
\end{document}

### 2015-04-25 Manuscript/PLoSONEresponse/TEM_PLoS_response.tex
PLoS response init

--- 2015-02-08 Manuscript/PLoSONESubmission/TEM_PLoS_draft.tex
@@
-% Please do not include colors or graphics in the text.
-%
-% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
-%
-% % % % % % % % % % % % % % % % % % % % % % %
-%
-% -- FIGURES AND TABLES
-%
-% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
-%
-% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
-% - Figures should be uploaded separately from your manuscript file. 
-% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
-% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
-% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
-%
-% Tables should be cell-based and may not contain:
-% - tabs/spacing/line breaks within cells to alter layout or alignment
-% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
-% - colors, shading, or graphic objects
-% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
-%
-% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
-%
-% Below are a few tips to help format your equations and other special characters according to our specifications.
-For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
-%
-% Please be sure to include all portions of an equation in the math environment.
-%
 \documentclass[10pt,letterpaper]{article}
@@
 
-% Use adjustwidth environment to exceed column width (see example table in text)
 \usepackage{changepage}
-% Use Unicode characters when possible
+\usepackage{color}
 \usepackage[utf8]{inputenc}
-% textcomp package and marvosym package for additional characters
 \usepackage{textcomp,marvosym}
-% fixltx2e package for \textsubscript
 \usepackage{fixltx2e}
-% amsmath and amssymb packages, useful for mathematical formulas and symbols
 \usepackage{amsmath,amssymb}
-% cite package, to clean up citations in the main text. Do not remove.
 \usepackage{cite}
-% Use nameref to cite supporting information files (see Supporting Information section for more info)
 \usepackage{nameref,hyperref}
-% line numbers
 \usepackage[right]{lineno}
-% ligatures disabled
 \usepackage{microtype}
 \DisableLigatures[f]{encoding = *, family = * }
-% rotating package for sideways tables
 \usepackage{rotating}
@@
 
-% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
-% Captions will be left justified
 \usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
 
-% Use the PLoS provided BiBTeX style
 \bibliographystyle{plos2009}
 
-% Remove brackets from numbering in List of References
 \makeatletter
@@
 
-% Leave date blank
 \date{}
@@
 \lfoot{\sf PLOS}
-
-%% Include all macros below
 
@@
 \newline
-% Insert Author names, affiliations and corresponding author email.
 \\
@@
 \end{flushleft}
-% Please keep the abstract below 300 words
 \section*{Abstract}
@@
 We investigate three major factors that directly affect the completeness and the size of the morphological part of the matrix: the proportion of living taxa with no morphological data, the amount of missing data in the fossil record, and the overall number of morphological characters in the matrix.
-We infer phylogenies from complete matrices and from matrices with various amounts of missing data, and then compare missing data topologies to the "best" tree topology inferred using the complete matrix.
+We infer phylogenies from complete matrices and from matrices with various amounts of missing data, and then compare missing data topologies to the ``best" tree topology inferred using the complete matrix.
 
-We find that the number of living taxa with morphological characters and the overall number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology.
+We find that the number of living taxa with morphological characters and the overall number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the ``best" tree topology.
 Therefore, we suggest that sampling effort should be focused on morphological data collection for living species to increase the accuracy of topological inference in a Total Evidence framework.
@@
 There have been great improvements in the theory and application of these two approaches (e.g.
-\cite{bapsta2013,stadlerdating2013,heaththe2013}) as well as much debate about the "best" approach to use (e.g.
+\cite{bapsta2013,stadlerdating2013,heaththe2013}) as well as much debate about the ``best" approach to use (e.g.
 \cite{spencerefficacy2013,wrightbayesian2014}).
@@
 
-In this study, we focus on the effect of missing data on our ability to recover a "best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species (e.g.
+In this study, we focus on the effect of missing data on our ability to recover a ``best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species (e.g.
 \cite{meredithimpacts2011,jetzthe2012}), or for studying evolutionary transitions (e.g.
@@
 Here we use a simulation approach to assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
-Since the molecular part of a Total Evidence matrix acts like a "classical" molecular matrix containing only the living taxa \cite{ronquista2012}, the effect of missing data on such matrices is well known \cite{wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011}.
+Since the molecular part of a Total Evidence matrix acts like a ``classical" molecular matrix containing only the living taxa \cite{ronquista2012}, the effect of missing data on such matrices is well known \cite{wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011}.
 Therefore, we focus only on missing data in the morphological part of the matrix.
@@
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
-We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the "best" tree topology more so than minimizing the amount of missing data in the fossil record.
-Additionally, we find that the ability of Total Evidence methods to recover the "best" tree topology is increased when using Bayesian methods.
+We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the ``best" tree topology more so than minimizing the amount of missing data in the fossil record.
+Additionally, we find that the ability of Total Evidence methods to recover the ``best" tree topology is increased when using Bayesian methods.
 
@@
 \item{Generating the matrix:} \label{step:generate_matrix} \\
-We randomly generated a birth-death tree (hereafter called the "true" tree) and used it to infer a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the "complete" matrix).
+We randomly generated a birth-death tree (hereafter called the ``true" tree) and used it to simulate a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the ``complete" matrix).
 \item{Removing data:} \label{step:remove_data} \\
-We removed data from the morphological part of the "complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$).
-We call the resulting 125 matrices "missing-data" matrices.
+We removed data from the morphological part of the ``complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$).
+We call the resulting 125 matrices ``missing-data" matrices.
 \item{Inferring phylogenies:} \label{step:build_phylo} \\
-We inferred phylogenetic trees from the "complete" matrix and from the 125 "missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the "best" tree) and 125 trees inferred from the matrices with missing morphological data (hereafter called the "missing-data" trees).
+We inferred phylogenetic trees from the ``complete" matrix and from the 125 ``missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the ``best" tree) and 125 trees inferred from the matrices with missing morphological data (hereafter called the ``missing-data" trees).
 Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
 \item{Comparing topologies:} \label{step:compare_topo} \\
-We compared the "best" tree to the "missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies
+We compared the ``best" tree to the ``missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies
 \end{enumerate}
@@
 \caption{{\bf Protocol outline.}
-(1) We randomly generated a birth-death tree (the "true" tree) and used it to infer a matrix with no missing data (the "complete" matrix).
-(2) We removed data from the morphological part of the "complete" matrix resulting in 125 "missing-data" matrices.
+(1) We randomly generated a birth-death tree (the ``true" tree) and used it to simulate a matrix with no missing data (the ``complete" matrix).
+(2) We removed data from the morphological part of the ``complete" matrix resulting in 125 ``missing-data" matrices.
 (3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
-(4) We compared the "missing-data" trees to the "best" tree.
+(4) We compared the ``missing-data" trees to the ``best" tree.
 We repeated these four steps 50 times.}
@@
 \label{Generating_the_matrix}
-First we randomly generated a "true" tree of 50 taxa in R v.
+First we randomly generated a ``true" tree of 50 taxa in R v.
 3.0.2 \cite{R302} using the package diversitree v.
@@
 
-Next, we generated a molecular and a morphological matrix from the "true" tree.
-The molecular matrix was inferred from the "true" tree using the R package phyclust v.
+Next, we generated a molecular and a morphological matrix from the ``true" tree.
+The molecular matrix was inferred from the ``true" tree using the R package phyclust v.
 0.1-14 \cite{chen2011}.
@@
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see Appendix S1 and Fig.
-S1).
-We then ran an independent discrete character simulation for each character using the "true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see ~\nameref{S1_Appendix1} and Fig.
+S1 within).
+We then ran an independent discrete character simulation for each character using the ``true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
 evolutionary rate) from one character state to another \cite{Pagel22011994}.
@@
 
-Finally, we combined the morphological and molecular matrices obtained from the "true" tree.
-Hereafter we call this the "complete" matrix, i.e.
+Finally, we combined the morphological and molecular matrices obtained from the ``true" tree.
+Hereafter we call this the ``complete" matrix, i.e.
 the matrix with no missing data except for the molecular data of the fossil taxa.
@@
 \label{Removing_data}
-We modified the "complete" matrix to get matrices with missing data by randomly replacing data with "?" in the morphological part of the matrices according to the following parameters:
+We modified the ``complete" matrix to get matrices with missing data by randomly replacing data with ``?" in the morphological part of the matrices according to the following parameters:
 \begin{enumerate}
@@
 We created matrices using all parameter combinations resulting in 125 ($5^3$) matrices.
-Note that one of these combinations has no missing data so is equivalent to the "complete" matrix, thus we have one effectively complete matrix in our 125 "missing-data" matrices.
+Note that one of these combinations has no missing data so is equivalent to the ``complete" matrix, thus we have one effectively complete matrix in our 125 ``missing-data" matrices.
 Because some parameter combinations introduce a lot of missing data (e.g.
@@
 \subsection*{3. Building phylogenies}
-From the resulting matrices we generated two types of trees: the "best" tree inferred from the "complete" matrix and the "missing-data" trees inferred from the 125 matrices with various amounts of missing data.
-The "true" tree was used to generate the "complete" matrix and reflects the "true" evolutionary history in our simulations.
-The "best" tree, on the other hand, is the best tree we can build using state-of-the-art phylogenetic methods.
-In real world situations, the "true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g.
+From the resulting matrices we generated two types of trees: the ``best" tree inferred from the ``complete" matrix and the ``missing-data" trees inferred from the 125 matrices with various amounts of missing data.
+The ``true" tree was used to generate the ``complete" matrix and reflects the ``true" evolutionary history in our simulations.
+The ``best" tree, on the other hand, is the best tree we can build using state-of-the-art phylogenetic methods.
+In real world situations, the ``true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g.
 \cite{rozen2005}).
-Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the "best" tree, rather than the "true" tree, as the "best" tree is generally what we have to work with.
+Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the ``best" tree, rather than the ``true" tree, as the ``best" tree is generally what we have to work with.
 
 \paragraph*{Maximum Likelihood}
-The "best" tree and the "missing-data" trees were inferred using RAxML v.
+The ``best" tree and the ``missing-data" trees were inferred using RAxML v.
 8.0.20 \cite{Stamatakis21012014}.
@@
 
-To measure the support for each branch in our simulated phylogenies we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the "complete" matrix.
+To measure the support for each branch in our simulated phylogenies we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the ``complete" matrix.
 We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \cite{zanderminimal2004}.
 We repeated this selection until we obtained 50 sets of simulations (i.e.
-50 "complete" and 50 x 125 "missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
+50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
 On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \cite{pattengale2010many}.
@@
 \paragraph*{Bayesian}
-The "best" tree and the "missing-data" trees were inferred using MrBayes v.
+The ``best" tree and the ``missing-data" trees were inferred using MrBayes v.
 3.2.1 \cite{Ronquist2012mrbayes}.
@@
 For the morphological data, we used the Markov \textit{k} state model \cite{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
-We chose these models to be consistent with the parameters used to generate the "complete" matrix.
+We chose these models to be consistent with the parameters used to generate the ``complete" matrix.
 
 Each Bayesian tree was estimated using two runs of four chains each for a maximum of 5$\times$$10^7$ generations.
-We used the following priors for each tree: (i) the "true" tree’s topology as a starting tree (with a starting value for each branch length of one), (ii) an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions, and (iii) a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)).
+We used the following priors for each tree: (i) the ``true" tree’s topology as a starting tree (with a starting value for each branch length of one), (ii) an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions, and (iii) a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)).
 We used these priors to speed up the Bayesian estimation process.
@@
 Even using these priors, it took \texttildelow 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-The detailed MrBayes parameters are available in the Appendix S1.
+The detailed MrBayes parameters are available in the ~\nameref{S1_Appendix1}.
 
@@
 \subsection*{4. Comparing topologies}
-We compared the topology of the "missing-data" trees to the "best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
+We compared the topology of the ``missing-data" trees to the ``best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
 We used the Robinson-Foulds distance \cite{RF1981} to assess the amount of conserved clade positions and the Triplets distance \cite{dobson1975triplets} to assess the amount of wildcard taxa (i.e.
@@
 \paragraph*{Robinson-Foulds distance}
-The Robinson-Foulds distance \cite{RF1981}, or "path difference", measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} (see Appendix S2 for calculation details).
+The Robinson-Foulds distance \cite{RF1981}, or ``path difference", measures the number of shared clades across two trees.
+The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} (see ~\nameref{S2_Appendix2} for calculation details).
 This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
@@
 We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons \cite{Bogdanowicz2012}.
-This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see Appendix S2 for the calculation details).
+This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see ~\nameref{S2_Appendix2} for the calculation details).
 Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
@@
 \paragraph*{Triplets distance}
-The Triplets distance \cite{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees \cite{critchlowthe1996} (see Appendix S2 for calculation details).
+The Triplets distance \cite{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees \cite{critchlowthe1996} (see ~\nameref{S2_Appendix2} for calculation details).
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
@@
 \label{tree_comparisons}
-For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the "best" tree and each "missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \cite{Bogdanowicz2012} resulting in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance for each tree inference method.
-Also, to take into account the uncertainty of tree inference, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 trees from the posterior tree distribution of the Bayesian analysis for the "best" trees, and then did the same for the 125 "missing data" trees (resulting in 1000 "best" trees and 125$\times$1000 "missing data" trees).
-For a given set of 1000 "missing data" trees and the 1000 "best" trees, we sampled one "missing data" tree and one "best" tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above.
-We repeated this 1000 times for each set of "missing data" trees resulting in 125$\times$1000 values for each metric.
+For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the ``best" tree and each ``missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \cite{Bogdanowicz2012} resulting in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance for each tree inference method.
+Also, to take into account the uncertainty of tree inference, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 trees from the posterior tree distribution of the Bayesian analysis for the ``best" trees, and then did the same for the 125 ``missing data" trees (resulting in 1000 ``best" trees and 125$\times$1000 ``missing data" trees).
+For a given set of 1000 ``missing data" trees and the 1000 ``best" trees, we sampled one ``missing data" tree and one ``best" tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above.
+We repeated this 1000 times for each set of ``missing data" trees resulting in 125$\times$1000 values for each metric.
 We repeated all the paired tree comparisons described above for each of the 50 simulation runs.
@@
 \subsection*{Testing the effects of the missing data parameters on topological recovery}
-Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the "best" tree topology in a Total Evidence framework.
+Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the ``best" tree topology in a Total Evidence framework.
 We also assessed the effect of our missing data parameters jointly with the effects of different tree inference and uncertainty methods (i.e.
@@
 We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \cite{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions \cite{Bhattacharyya} (see Appendix S2 for calculation details).
+The Bhattacharyya Coefficient is the probability of overlap between two distributions \cite{Bhattacharyya} (see ~\nameref{S2_Appendix2} for calculation details).
 Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
@@
 We performed this procedure separately for each tree inference and uncertainty method.
-When two combinations of missing data parameters have a similar ability to recover the "best" tree topology the Bhattacharyya Coefficient will be close to one.
+When two combinations of missing data parameters have a similar ability to recover the ``best" tree topology the Bhattacharyya Coefficient will be close to one.
 Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
@@
 To assess the effect of the different tree inference and uncertainty methods (i.e.
-Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the "best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of tree inference and uncertainty methods, i.e.
+Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the ``best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of tree inference and uncertainty methods, i.e.
 Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc.
@@
 Note that this procedure pools results from across all missing data parameter combinations so it results in just six pairwise comparisons.
-When two tree inference or uncertainty methods have a similar ability to recover the "best" tree topology the Bhattacharyya Coefficient will be close to one.
+When two tree inference or uncertainty methods have a similar ability to recover the ``best" tree topology the Bhattacharyya Coefficient will be close to one.
 Conversely, if the two tree inference or uncertainty methods differ, the Bhattacharyya Coefficient will be close to zero.
@@
 \section*{Results}
-As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the "best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
+As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
 However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
 \subsection*{Individual effects of missing data parameters}
-As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7).
-However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7).
+As the amount of missing data increases across all three parameters, our ability to recover the ``best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7 in ~\nameref{S3_Appendix3}).
+However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7 in ~\nameref{S3_Appendix3}).
 
 \subsection*{Combined effect of missing data parameters}
-As expected, our ability to recover the "best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
+As expected, our ability to recover the ``best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; Tables S2, S3 and S4).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; Tables S2, S3 and S4 in ~\nameref{S3_Appendix3}).
 Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
@@
  Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the ``best" tree distribution, calculated using the Bhattacharyya Coefficient.
 The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
@@
 However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.
-S2).
+S2 in ~\nameref{S3_Appendix3}).
 
@@
 \caption{{\bf Bhattacharyya Coefficients of the pairwise method comparisons.}
-Each comparison which corresponds to the normalised distance between the "best" tree and the "missing data" trees using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
+Each comparison which corresponds to the normalised distance between the ``best" tree and the ``missing data" trees using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
 Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
@@
 The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
-If two methods have a high probability of overlap, they have a similar ability to recover the "correct" tree topology.}
+If two methods have a high probability of overlap, they have a similar ability to recover the ``correct" tree topology.}
 \centering
@@
 
-Our results show that the ability to recover the "best" tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used.
+Our results show that the ability to recover the ``best" tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used.
 However, these factors affected topological recovery in different ways and to different extents.
 Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
-Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam}, Fig.~\ref{Fig_Results-paircomp_within}, Table ~\ref{Tab_Results-Difference_methods}).
+Additionally, using Bayesian consensus trees recovered the ``best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam}, Fig.~\ref{Fig_Results-paircomp_within}, Table ~\ref{Tab_Results-Difference_methods}).
 As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters (e.g.
@@
 When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
-Because living taxa still have molecular characters available for phylogenetic inference (see mMthods),even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
+Because living taxa still have molecular characters available for phylogenetic inference (see Methods),even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
 However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
@@
 \paragraph*{Missing morphological characters ($M_{C}$)}
-Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the "best" tree topology.
+Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the ``best" tree topology.
 We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
 Conversely, removing 10\% of morphological characters ($M_{C}$) genuinely removes 10\% of data in the matrix.
-However, the effect of removing characters on the ability to recover the "best" tree topology is of the same order of magnitude as for the other two parameters (Fig.~\ref{Fig_Results-permeth_perparam}).
+However, the effect of removing characters on the ability to recover the ``best" tree topology is of the same order of magnitude as for the other two parameters (Fig.~\ref{Fig_Results-permeth_perparam}).
 We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.
@@
 Additionally, the number of morphological characters determines the size the matrix.
-This can affect our ability to recover the "best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
+This can affect our ability to recover the ``best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
 The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \cite{wagner2000}.
@@
 \subsection*{Combined effect of missing data parameters}
-As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have few missing data (i.e.
+As expected, when combining the missing data parameters, our ability to recover the ``best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have few missing data (i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a lot of missing data (i.e.
@@
 \subsection*{Effects of tree inference methods}
-Variation in our ability to recover the "best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+Variation in our ability to recover the ``best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{k} model \cite{lewisa2001} over cladistic methods \cite{wrightbayesian2014} (but see \cite{spencerefficacy2013}).
-However, this is the first study, to our knowledge, to compare the performance of the M\textit{k} model \cite{lewisa2001} for recovering the "best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-Our results show that the topology of the Bayesian consensus tree is always closer to the "best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}).
+However, this is the first study, to our knowledge, to compare the performance of the M\textit{k} model \cite{lewisa2001} for recovering the ``best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
+Our results show that the topology of the Bayesian consensus tree is always closer to the ``best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}).
 As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
-This may also be because the "best" Bayesian consensus trees are not completely resolved, thus will always be more similar to the "missing data" trees than a completely resolved tree like the "best" Maximum Likelihood tree.
-However, we minimized the probability of unresolved "best" trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
+This may also be because the ``best" Bayesian consensus trees are not completely resolved, thus will always be more similar to the ``missing data" trees than a completely resolved tree like the ``best" Maximum Likelihood tree.
+However, we minimized the probability of unresolved ``best" trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
 
-It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the "best" tree topology than Maximum Likelihood and Bayesian consensus trees.
+It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best" tree topology than Maximum Likelihood and Bayesian consensus trees.
 This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-permeth_perparam}).
-This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of "missing data" trees and a set of the "best" trees involved 1000 random pairwise comparisons rather than just one.
+This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best" trees involved 1000 random pairwise comparisons rather than just one.
 This will obviously add noise to the results for these methods.
@@
 Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
-Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the "best" tree topology.
+Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the ``best" tree topology.
 We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections.
@@
 \cite{nithe2013}).
-Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the "best" tree topology (Fig.~\ref{Fig_Results-paircomp_within}).
+Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the ``best" tree topology (Fig.~\ref{Fig_Results-paircomp_within}).
 Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
@@
 Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \cite{ronquista2012,BEASTmaster}; although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution).
-Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see above).
+Because the Bayesian consensus trees consistently recovered topologies closer to the ``best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \cite{ronquista2012,BEASTmaster}; although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution).
+Note that, although the Bayesian posterior tree distribution performed poorly in recovering the ``best" tree topology, this is due to difficulties comparing distributions of trees (see above).
 Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses (e.g.
@@
 \section*{Conclusion}
-Missing data in Total Evidence matrices is not a problem for recovering the "best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
+Missing data in Total Evidence matrices is not a problem for recovering the ``best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
 When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
@@
 Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.
-Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the "best" tree topology than the Maximum Likelihood trees.
+Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best" tree topology than the Maximum Likelihood trees.
 Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
@@
 % Include only the SI item label in the subsection heading. Use the \nameref{label} command to cite SI items in the text.
-\subsection*{Supplementary informations}
+
+\subsection*{Data reproducibility and availability}
+\noindent{\bf Data reproducibility and availability.} All the code used in this analysis is available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{\color{blue}{GitHub}} with some information on how to use the various functions.
+Additionally all the simulated data is available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{\color{blue}{FigShare}}.
+
+\subsection*{Supporting Information S1: Tree Inference}
 \label{S1_Appendix1}
-\noindent{\bf Appendix S1 Tree Inference.} Supplementary information on the Tree Inference methods.
+\noindent{\bf Supporting Information S1 Tree Inference.} Supplementary information on the Tree Inference methods. Contains Figure S1.
 
-\label{Figure S1}
-\noindent{\bf Figure S1 proportion of morphological characters states.}
+\subsection*{Supporting Information S2: Tree Comparisons}
+\label{S2_Appendix2}
+\noindent{\bf Supporting Information S2 Tree Comparisons.} Supplementary information on the Tree Comparison methods.
 
-\label{S2_Appendix2}
-\noindent{\bf Appendix S2 Tree Comparisons.} Supplementary information on the Tree Comparison methods.
-
-\label{S2_Appendix3}
-\noindent{\bf Appendix S3 Additional results.}.
-
-\label{Figure S2}
-\noindent{\bf Figure S2 Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey).}
-
-\label{Figure S3}
-\noindent{\bf Figure S3 Effects of missing data on topological recovery using Maximum Likelihood trees.}
-
-\label{Figure S4}
-\noindent{\bf Figure S4 Effects of missing data on topological recovery using Maximum Likelihood bootstrap trees.}
-
-\label{Figure S5}
-\noindent{\bf Figure S5 Effects of missing data on topological recovery using Bayesian posterior tree distribution.}
-
-\label{Table S1}
-\noindent{\bf Table S1 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method.}
-
-\label{Table S2}
-\noindent{\bf Table S2 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method for the $M_{L}$ parameter.}
-
-\label{Table S3}
-\noindent{\bf Table S3 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method for the $M_{F}$ parameter.}
-
-\label{Table S4}
-\noindent{\bf Table S4 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method for the $M_{C}$ parameter.}
-
-\label{Table S5}
-\noindent{\bf Table S5 Bhattacharyya Coefficients of the pairwise method comparisons for the $M_{L}$ parameter.}
-
-\label{Table S6}
-\noindent{\bf Table S6 Bhattacharyya Coefficients of the pairwise method comparisons for the $M_{F}$ parameter..} Bhattacharyya
-
-\label{Table S7}
-\noindent{\bf Table S7 Bhattacharyya Coefficients of the pairwise method comparisons for the $M_{C}$ parameter..} Bhattacharyya
+\subsection*{Supporting Information S3: Additional results}
+\label{S3_Appendix3}
+\noindent{\bf Supporting Information S3 Additional results.}. Contains Figures S2, S3, S4 and S5 and Tables S1, S2, S3, S4 S5, S6 and S7.
 
@@
 
-\section*{Author Contributions}
-Conceived and designed the experiments: TG NC. Ran the analysis: TG. Analyzed the data: TG. Wrote the paper: TG NC.
-
 \nolinenumbers

### 2015-04-25 Manuscript/PLoSONEresponse/response.tex
Updated response

@@
 
+%-------------
+
 \textcolor{blue}{Unfortunately the data analysis performed is simply the calculation of a number of summary statistics about the overlap between the estimates returned for different simulation conditions.
@@
 
+%-------------
+
 \textcolor{blue}{The bias in the analysis in favor of Bayesian methods makes the recommendation to use Bayesian methods less satisfying.}
@@
 Our results did indeed biased us in favour of using a Bayesian approach to fix topology since our results clearly indicate that it is less likely to generate a wrong topology using a Bayesian approach (see Figure 4 and 5 and discussion line 472 to 496 in the first submission manuscript).
+% TG: or I'm not sure I understood that one.
+%-------------
 
@@
 Note that both RAxML and MrBayes do not distinguish between both models since the Mkv model is just a particular case of the Mk model and are not distinguished either in the original publication that we cite in the manuscript (Lewis 2001).
-We've changed the mentions of the Mk model to Mkv model throughout the new manuscript.
+We've changed the mentions of the M\textit{k} model to M\textit{kv} model throughout the new manuscript.
 
 %Add note that we changed it along the text. RAxML and MrBayes are using default.
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{If there was a mention of the data being deposited, I missed it. I apologize if I did miss it.}
@@
 See our comment above on data availability and reproducibility.
+
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{line 88 ``used it to infer a matrix" $->$ ``used it to simulate a matrix". ``inference" is used in place of simulation elsewhere.}
@@
 
+%-------------
+
 \textcolor{blue}{line 112: does diversitree just stop simulating when it reaches the correct number of tips, or does it deal with the biases that can result from doing that?
@@
 
-%TO CHECK!
-Lines 116 to 118 in the previously submitted manuscript:
+See our response to the comment ``The procedure for generating a tree with fossils is a valid procedure." below.
 
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-To avoid biasing our simulations towards either living or fossil taxa and to make each simulation comparable, we implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa.
-\end{minipage}
-
-Lines @@@ in the new manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-To avoid biasing our simulations towards either living or fossil taxa and to make each simulation comparable, we implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa.
-In practice, we generated $x$ birth-death trees using the \texttt{tree.bd} function from the \texttt{diversitree R} package (FitzJohn) until we obtained $y$ trees with the right number of living and fossil taxa.
-\end{minipage}
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{line 118: wih 25 living and 25 fossil taxa. Are the fossils just put at the extinction point for the extinct lineages?}
@@
 
+%-------------
+
 \textcolor{blue}{line 110 - line 120.
@@
 In fact, the publication associated with the diversitree package (FitzJohn 2012, Methods in Ecology and Evolution) has been cited three times more (155 cites on Google Scholar) than the publication associated with the \texttt{TreeSim} package (Stadler 2011, Systematic Biology - 55 cites on Google Scholar).
+
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{Do the simulations apply ascertainment biases as described by Lewis' Mkv (2001) model? It seems like they should... At minimum, the point should be discussed.}
@@
 
+%-------------
+
 \textcolor{blue}{line 171 - 177: the authors need to explain the order of operations for the $M_F$ vs $M_C$ masking.
@@
 
-We clarified this point ($M_C$ comes second). %change in manuscript
+We clarified this point ($M_C$ comes second) in the new version of the manuscript.
 Also, we clarified the potential confusion between the amount of data removed in one parameter ($M_F$ = 10\% for example) and the amount of missing data in the morphological matrix (5\%).
-%change in manuscript
+
+Lines 171 to 178 in the previously submitted manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+In practice, each parameter represents a different way of removing data from the matrix: $M_L$ removes rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $M_C$ removes columns across both living and fossil taxa's data.
+Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
+\end{minipage}
+
+Lines @@@112 to @@@114 in the new manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+In practice, each parameter represents a different way of removing data from the matrix: $M_L$ removes rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $M_C$ removes columns across both living and fossil taxa's data.
+Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
+We first applied the parameters $M_L$ and $M_F$ on the matrix and then applied the $M_C$ parameter.
+Therefore, when 10\% data was missing for both $M_L$ and $M_F$, 10\% data was missing in the morphological part of the matrix.
+However, when applying the $M_C$ parameter with the same rate (10\%) the resulting matrix potentially had more than 10\% data missing.
+\end{minipage}
+
+%-------------
 
@@
 
-We changed the Header 3 to infering phylogenies. %change in manuscript
+We changed the Header 3 to ``estimating phylogenies". %change in manuscript
+
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{Filtering the data sets for strong bootstrap support under the full data analysis surely introduces some bias in favor of the no-missing data analyses.}
@@
 If the no-missing data tree has low node support, it is likely that the changes in topology for the missing data trees are not linked to our focal question (the effect of removing characters, living taxa or fossils) but just to the fact that low supported nodes can be just random.
+We've added the following justification in the new manuscript:
+
+Lines 208 to 210 in the previously submitted manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+We repeated this selection until we obtained 50 sets of simulations (i.e. 50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
+\end{minipage}
+
+Lines @@@112 to @@@114 in the new manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+We repeated this selection until we obtained 50 sets of simulations (i.e.
+50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
+This step was implemented to make sure that the differences we observed in topologies (see below) where due to the amount of missing data for each parameters ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
+\end{minipage}
+
+%-------------
 
@@
 
-We would like to emphasize that we are looking at topology and morphological data, our priors are on nucleotid frequencies and on TT ratio.
-This remains unchanged in all our trees.
-Also, we would like to emphasize on the ``speed up" part that was mandatory since the analysis took 1.5 CPU century including our ``unfair" methodological choices.
-We had to choose between extracting a global pattern (more faster chains) rather than some specific analysis (fewer slower chains).
+We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only.The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and where a minor part of the MCMC sampler (2.17\% of move probability for the three priors).
+Therefore we are confident that the topology is not affected allowing us to explore the effect of missing morphological data on topology as well as reducing our computational time to a realistic 1.5 CPU century.
+
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{line 258: RF is 0 (not 1) when the trees are identical. It's maximum (for 2 rooted binary trees) is 2(n-2) not n-2.}
@@
 %check
+
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{line 275: why isn't the upper bound of the triples distance n choose 3 (instead of n choose 4). I think you got the formula from an unrooted quartet distance.}
@@
 
+%-------------
+
 \textcolor{blue}{Figure 3: B is confusing. If both curves are probability distributions, then they both have to integrate to 1. So B(y) can't cover just a subset of the area covered by B(x).}
@@
 %Fix probability distributions -> "just" distributions
+
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{line 398: ``mMthods" -> ``Methods"}
@@
 %Change to "affect less" or something like that.
+
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{The definition of the scaled RF implies that it is never positive (sinc the numerator cannot be positive and the denominator is positive).}
@@
 %Because we follow Bogdanowicz and use the difference between random (NULL) trees to scale.
+
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{Note that JavaScript is a language, so it a bit confusing to refer to a Java program as a ``Java script."}
@@
 %fix
+
+%-------------
 
@@
 
+%-------------
+
 \textcolor{blue}{Equation (7) is true for a three-leaf tree with a polytomy occurring with probability = 1/4 and the other 3 trees being equiprobable.
@@
 
+%-------------
+
 \textcolor{blue}{I don't think that equations 8, 9, or 10 are correct for any case other than the 3 taxon case (though I do note that the authors are using n choose 3 here).}
@@
 
+%-------------
+
 \textcolor{blue}{Equations 13 and 14. It seems easier to just drop the summation sign here, and just define $a_i$ and $b_i$ to the right hand side of these equations.}

### 2015-04-25 Manuscript/PLoSONEresponse/TEM_PLoS_response.tex
Updated response

@@
 We call the resulting 125 matrices ``missing-data" matrices.
-\item{Inferring phylogenies:} \label{step:build_phylo} \\
+\item{Estimating phylogenies:} \label{step:build_phylo} \\
 We inferred phylogenetic trees from the ``complete" matrix and from the 125 ``missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the ``best" tree) and 125 trees inferred from the matrices with missing morphological data (hereafter called the ``missing-data" trees).
@@
 
-In practice, each parameter represents a different way of removing data from the matrix: $M_{L}$ removes rows from the living taxa's data; $M_{F}$ removes cells from the fossil taxa's data; and $M_{C}$ removes columns across both living and fossil taxa's data.
-Note that $M_{L}$ and $M_{F}$ differ not only because of the region of the matrix affected: for $M_{L}$ all the morphological data of a percentage of living taxa are removed, whereas for $M_{F}$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
+In practice, each parameter represents a different way of removing data from the matrix: $M_L$ removes rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $M_C$ removes columns across both living and fossil taxa's data.
+Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
+We first applied the parameters $M_L$ and $M_F$ on the matrix and then applied the $M_C$ parameter.
+Therefore, when 10\% data was missing for both $M_L$ and $M_F$, 10\% data was missing in the morphological part of the matrix.
+However, when applying the $M_C$ parameter with the same rate (10\%) the resulting matrix potentially had more than 10\% data missing.
 
@@
 
-\subsection*{3. Building phylogenies}
+\subsection*{3. Estimating phylogenies}
 From the resulting matrices we generated two types of trees: the ``best" tree inferred from the ``complete" matrix and the ``missing-data" trees inferred from the 125 matrices with various amounts of missing data.
@@
 8.0.20  \cite{Stamatakis21012014}).
-For the morphological data, we used the implemented Markov \textit{k} state model \cite{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
+For the morphological data, we used the implemented M\textit{kv} model \cite{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
 8.0.20 \cite{Stamatakis21012014}).
@@
 50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
+This step was implemented to make sure that the differences we observed in topologies (see below) where due to the amount of missing data for each parameters ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
 On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \cite{pattengale2010many}.
@@
 The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \cite{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
-For the morphological data, we used the Markov \textit{k} state model \cite{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{k} + $\Gamma_4$).
+For the morphological data, we used the M\textit{kv} model \cite{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{kv} + $\Gamma_4$).
 We chose these models to be consistent with the parameters used to generate the ``complete" matrix.
@@
 However, the use of probabilistic methods (i.e.
-Maximum Likelihood or Bayesian) and the M\textit{k} model \cite{lewisa2001} has been previously demonstrated to partially resolve this issue \cite{wrightbayesian2014}.
+Maximum Likelihood or Bayesian) and the M\textit{kv} model \cite{lewisa2001} has been previously demonstrated to partially resolve this issue \cite{wrightbayesian2014}.
 
@@
 Variation in our ability to recover the ``best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{k} model \cite{lewisa2001} over cladistic methods \cite{wrightbayesian2014} (but see \cite{spencerefficacy2013}).
-However, this is the first study, to our knowledge, to compare the performance of the M\textit{k} model \cite{lewisa2001} for recovering the ``best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
+For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \cite{lewisa2001} over cladistic methods \cite{wrightbayesian2014} (but see \cite{spencerefficacy2013}).
+However, this is the first study, to our knowledge, to compare the performance of the M\textit{kv} model \cite{lewisa2001} for recovering the ``best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
 Our results show that the topology of the Bayesian consensus tree is always closer to the ``best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}).

### 2015-04-26 Manuscript/PLoSONEresponse/response.tex
Updated main comments

@@
 \usepackage[osf]{mathpazo}
+\usepackage{amsmath,amssymb}
 \signature{Thomas Guillerme \\ Natalie Cooper}
@@
 Also, the effect of the acquisition bias is to inflate branch length which is an aspect of the phylogeny we didn't explored in this study (focusing on topology).
-See response below to the comment \color{blue}{\textit{The use of Mk rather than Mkv makes the study less relevant to real data analyses.}}.
+See response below to the comment ``\textit{The use of Mk rather than Mkv makes the study less relevant to real data analyses.}".
 
@@
 
-We changed the Header 3 to ``estimating phylogenies". %change in manuscript
+We changed the Header 3 to ``Estimating phylogenies".
 
@@
 
-%check
+We changed the following sentence:
+
+Lines 254 to 255 in the previously submitted manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+The Robinson-Foulds distance, or ``path difference", measures the number of shared clades across two trees.
+\end{minipage}
+
+Lines @@@112 to @@@114 in the new manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+The Robinson-Foulds distance, or ``path difference", measures the difference between the number of clades and the twice number of shared clades across two trees.
+\end{minipage}
 
@@
 
-%check
+We changed the following sentence:
+
+Lines 257 to 259 in the previously submitted manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
+\end{minipage}
+
+Lines @@@112 to @@@114 in the new manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+This metric is bounded between zero, when the two trees are identical, and $2(n-2)$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
+\end{minipage}
+
+%%%%%%%%%%%%%%%%%
+%CHECK in Soltis!
+%%%%%%%%%%%%%%%%%
 
@@
 
-%check
+We clarified the scaling method as follow:
+
+Lines 262 to 270 in the previously submitted manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons (Bogdanowicz et al.
+2012).
+This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see Supporting Information S2: Tree Comparisons for the calculation details).
+Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
+\end{minipage}
+
+Lines @@@112 to @@@114 in the new manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+We normalised this metric following Bogdanowicz \textit{et al.} (2012)'s Normalised Tree Similarity (NTS) method. This methods scales %not sure if scale here?
+any tree comparison metric using the mean distance between 1000 random trees (see Supporting Information S2: Tree Comparisons for the calculation details).
+This method is a generalisation of the topological accuracy method (Price \textit{et al.} 2010) allowing to compare topological differences between any tree with any tree comparison metric.
+In practice when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
+\end{minipage}
 
@@
 
-%check
+We changed the following sentence:
+
+Lines 274 to 275 in the previously submitted manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+It is bounded between zero when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
+\end{minipage}
+
+Lines @@@112 to @@@114 in the new manuscript:
+
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+It is bounded between zero when the two trees are identical and $\binom{n}{3}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
+\end{minipage}
+
+%%%%%%%%%%%%%%%%%
+%CHEK in Soltis!
+%%%%%%%%%%%%%%%%%
 
@@
 
-We clarified as follow:
+We clarified this point in the text as well as in the figures 2 and 3 captions (``equivalent" $->$ ``similar"):
 
@@
 
-%Fix probability distributions -> "just" distributions
+%I'm not sure how to fix that comment.
+The figure caption says (first sentence!) : "A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+ML = 10%,MF = 50%, MC = 75%)." Should I make it super clear by writing : "A and B are distributions of tree similarity metrics (not probability distributions, or roller coasters plans)" ?
 
@@
 ANOVA or multiple regression analysis might the appropriate analyses in some cases but we believe it isn't here.
-%Not comparing means, non normal distributions.
-ANd non-parametric version (e.g.
-Kruskal Wallis) not approriate either.
+In our analysis, we made the methodological choice to compare tree similarity distributions that are: not continuous distributions; bounded at 1; not always normal (both the distributions and the residuals) and sometimes multi-modal.
+Therefore, we choose not to use classical ``statistics" methods such as ANOVA or multiple regression because these methods could simply not be applied to our data set.
+%Not comparing means, non normal distributions. And non-parametric version (e.g. Kruskal Wallis) not approriate either.
 
@@
 
-\textcolor{blue}{line 398: ``mMthods" -> ``Methods"}
+\textcolor{blue}{line 398: ``mMthods" $->$ ``Methods"}
 
@@
 \textcolor{blue}{line 420: contra the statement here, a consensus collapsing to a polytomy will affect the RF distance (could go up or down, but it will change).}
+
+%To
+
+This does not decrease the 420
+Normalised Robinson-Foulds distance, because clades are conserved, but decreases the 421
+Normalised Triplets distance because the fossils act as wildcard taxa. 422
 
@@
 
-%check
+%not in this paper
 

### 2015-04-27 Manuscript/PLoSONEresponse/response.tex
Updated response (supplementaries)

@@
 
-Please find enclosed our resubmission of our manuscript @@@[title]. We include below details of how we have dealt with the reviewer's comments.
+Please find enclosed our resubmission of our manuscript ``Effects of missing data on topological inference using a Total Evidence approach".
+We include below details of how we have dealt with the reviewer's comments.
 
@@
 We suspect they rushed their assessment because they make several comments about our paper that are completely incorrect, and would have been noticed in a thorough reading of the manuscript.
-For example, the reviewer says that our data are not available when we clearly state on @@@[page and lines] that all the data are deposited in figshare @@@[link].
+
+%For example, the reviewer says that our data are not available when we clearly state on @@@[page and lines] that all the data are deposited in figshare @@@[link].
 We suspect that had a second reviewer been found the decision might have been different.
-%maybe recheck that
+% TG: ok, this part somehow was in the SystBiol manuscript as a header part of the supplementary (1 file with three appendices) and I forgot to included somewhere in the manuscript or the supplementary in last submission (3 files).
+See the "OnlineAppendices" file in the GitHub folder.
+However, references to the GitHub page is given in the 1 supplementary material, but I assume it's a bit weak...
+My bad.
+
+%We can maybe moan about the reviewer not reading carefully few things such as the comment on figure 3B "Figure 3: B is confusing.
+If both curves are probability distributions, then they both have to integrate to 1.
+So B(y) can't cover just a subset of the area covered by B(x)." where it seems he just didn't properly read the caption or about the Bhattacharya where he didn't read that we applied it because we're not comparing the means...
+But it's a bit weaker.
+
+%In general I feel actually bad for some comments (i.e. typos in the equations etc...) and will remember to check more thoroughly next time.
 
@@
 
+%So for the part about the tree comparisons (RF and Triplets, I think we can just remove the whole thing since it's not especially useful...)
+
 \textcolor{blue}{Supplement 2:}
@@
 
-%Because we follow Bogdanowicz and use the difference between random (NULL) trees to scale.
+We cited the reference (Bogdanowicz 2012) of the method.
 
@@
 
-%check
+As clarified in the new manuscript (lines @@@), the NTS is a generalisation of the topological accuracy method (Price \textit{et al.} 2010 PLoS ONE) and is therefore not specialised for the RF or the Triples distance.
 
@@
 
-%fix
+We fixed this typo.
 
@@
 
-%check
+We fixed this by removing the summation signs.
 

### 2015-04-27 Manuscript/PLoSONEresponse/OnlineAppendices-LaTeXSuppFiles/SupportingInformation2.tex
Updated response (supplementaries)

\documentclass[12pt,letterpaper]{article}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\pagenumbering{arabic}

%Pagination style and stuff
\linespread{2}
\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}

\begin{document}

\section{Supporting Information S2: Tree Comparisons}

%\subsection{Robinson-Foulds distance}
%Robinson-Foulds distance (\textit{RF} \cite{RF1981}), or "path difference", measures the number of shared clades across two trees. The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} and can be expressed as:
%\begin{equation}
%RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
%\end{equation}
%where $C_{x,y}$ is the number of clades in common in the two trees. $C$ is equal to one if the two trees have the same $n$ taxa; and $C = n-2$ when none of the $n$ taxa are shared between the trees. This metric is more sensitive to taxon displacement than Triplets distance (i.e. if one taxon moves out of a clade, then the clades are no longer considered similar \cite{critchlowthe1996,johnson1998,wiensmissing2003}). The minimal value of $C$ is equal to 1 if the two trees have the same n taxa; the maximal value in $C = n-2$. For a fully unresolved tree (star tree) $N$=1 and for a fully resolved tree (binary tree) $N = n-2$. The minimal and maximal topological distance for taxa is:
%\begin{equation}
%RF_{min} = 1 + 1 - 2C_{x,y}
%\end{equation}
%and:
%\begin{equation}
%RF_{max} = 2(n-2)-2
%\end{equation}
%One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any $n$ taxa:
%\begin{equation}
%RF.scaled_{x,y} = \frac{RF_{x,y}-RF_{max}}{RF_{max}}
%\end{equation}
%This metric is more sensitive to taxa displacement than the Triplets distance \cite{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.

%\subsection{Triplets distance details ($T_{x,y}$)}
%Triplets distance ($T_{x,y}$ \cite{dobson1975triplets}) measures the number of sub-trees made up of three taxa (triplets) that differ between two given trees. Each triplet can be written as $I_{ijk}$=(\textit{ijk}). Where $I_{ijk}$ is equal to zero if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to one. For any rooted binary tree there are only three possible combinations for each triplet: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}) \cite{johnson1998}. If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k}). We can calculate the triplet distance between two trees, $S_n$, as:
%\begin{equation}
%S_n = \sum_{ijk} I_{ijk}
%\end{equation}
%where:
%\begin{equation}
%\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
%\end{equation}
%and where \textit{n} is the total number of taxa in both trees (modified from \cite{critchlowthe1996}). If $S_n$ = 0, the trees are identical; when $S_n$ = $\binom{n}{4}$, the trees are as different as possible (i.e. every taxon has a different placement in the two trees). Because the possible number of triplets per clade is a finite number, the probability of two random trees with the same $n$ taxa to have the same triplet is:
%\begin{equation}
%P({I_{ijk}}=0) = \frac{1}{4}
%\end{equation}
%Therefore one can calculate the probability of two random trees having the same triplets: 
%\begin{equation}
%P({S_{n}}=0) = \sum_{ijk} P_{I_{ijk}=0}
%\end{equation}
%\begin{equation}
%P({S_{n}}=0) = \frac{n!}{4(3!(n-3)!}
%\end{equation}
%and in the same way:
%\begin{equation}
%P({S_{n}}=1) = \frac{3n!}{4(3!(n-3)!}
%\end{equation}

\subsection{Normalised Tree Similarity}
For any tree with \textit{n} taxa compared using a tree distance metric $m$, Normalized Tree Similarity, $NTS_m$ \cite{Bogdanowicz2012}, represents the similarity score for the two trees given the expected distance between two random Yule trees \cite{Bogdanowicz2012} with $n$ taxa. If $\bar{d}_{m,n}$\textit{(rand)} is the average distance between two random Yule trees with $n$ taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} each containing $n$ taxa, then:
\begin{equation}
NTS_{m,n}(x,y)=\frac{\bar{d}_{m,n}(rand) - d_{m,n}(x,y)} {\bar{d}_{m,n}(rand)}
\end{equation}
\textit{NTS} ranges from one to -$\infty$.
For any $m,n$, when $NTS$ = 1, the trees are identical, when \textit{NTS} = 0 the trees are no more different than expected by chance, and when $NTS$ $<$ 0, the trees are more different than expected when comparing two random trees.

We used the NTS method to scale all the Robinson Foulds and Triplets distances calculated in our analyses, using the TreeCmp JavaScript \cite{Bogdanowicz2012}.

\subsection{Bhattacharyya Coefficient}
The Bhattacharyya Coefficient calculates the probability of overlap of two distributions \cite{Bhattacharyya}. When it is equal to zero, the probability of overlap of the distributions is also zero, and when it is equal to one, the two distributions are entirely overlapping. It forms an elegant and easy to compute continuous measurement of the probability of similarity between two distributions. The coefficient is calculated as the sum of the square root of the relative counts shared in \textit{n} bins among two distributions.
\begin{equation}
\text{Bhattacharyya Coefficient}=\sum_{i=1}^{n} \sqrt{{\sum{a_i}}\times{\sum{b_i}}}
\end{equation}
where
\begin{equation}
{a_i}=\frac{\text{Number of counts in bin \textit{i} for the distribution \textit{a}}}{\text{Total number of counts for the distribution \textit{a}}}
\end{equation}
and
\begin{equation}
{b_i}=\frac{\text{Number of counts in bin \textit{i} for the distribution \textit{b}}}{\text{Total number of counts for the distribution \textit{b}}}
\end{equation}
The precision of the Bhattacharyya Coefficient is directly related to the number of bins, $n$. If $n$ is low, the overlap will be overestimated and if $n$ is too high, the overlap will be underestimated. In this analysis, we determined the number of bins using Silverman's rule of thumb which states that $n$ should be 0.9 times the minimum of the standard deviation and the interquartile range of the distribution, divided by 1.34 times the sample size of the distribution to the negative one-fifth power (\texttt{bw.nrd0()} function in R \cite{silverman1986density}).

\bibliographystyle{plos2009}
\bibliography{Supp_Ref}

%END
\end{document}

### 2015-04-28 Manuscript/PLoSONEresponse/response.tex

--- 2015-04-26 Manuscript/PLoSONEresponse/response.tex
@@
 
-Please find enclosed our resubmission of our manuscript @@@[title]. We include below details of how we have dealt with the reviewer's comments.
+Please find enclosed our resubmission of our manuscript ``Effects of missing data on topological inference using a Total Evidence approach".
+We include below details of how we have dealt with the reviewer's comments.
+Your comments and the reviewers comments are in blue and our response is in black.
 
-We would however, like to draw your attention to a couple of points.
+% NC: Just send to editor
+%We would however, like to draw your attention to a couple of points.
 Firstly, while we understand the difficulties of obtaining reviewers for papers, we were very disappointed to have the previous decision made based on just one review.
@@
 We suspect they rushed their assessment because they make several comments about our paper that are completely incorrect, and would have been noticed in a thorough reading of the manuscript.
-For example, the reviewer says that our data are not available when we clearly state on @@@[page and lines] that all the data are deposited in figshare @@@[link].
-We suspect that had a second reviewer been found the decision might have been different.
-%maybe recheck that
+We suspect that had a second reviewer been found, the decision might have been different.
 
-Secondly, the reviewer clearly misunderstood the meaning of the term ``statistics".
+%Secondly, the reviewer clearly misunderstood the meaning of the term ``statistics''.
 Statistics does not mean that p-values need to be used.
@@
 The reviewer however appears to have refused to engage with this and instead has fixated on the lack of p-values.
+% NC: Some kind of summing up here
 
-\textcolor{ForestGreen}{From: PLOS ONE <em@editorialmanager.com<mailto:em@editorialmanager.com> \\
-To: Thomas Guillerme <guillert@tcd.ie<mailto:guillert@tcd.ie> \\
-Reply-To: PLOS ONE <plosone@plos.org<mailto:plosone@plos.org> \\
-Date: 20 April 2015 12:25:39 IST \\ \\
-Subject: PLOS ONE Decision: PONE-D-15-05911 - [EMID:ffa50492e76c2059] \\
-PONE-D-15-05911 \\
-Effects of missing data on topological inference using a Total Evidence approach \\
-PLOS ONE \\ \\
+%We are slightly concerned by the reviewer's misconception of the term ``statistics" and suspicions (``I suspect that if they authors take some time to do statistics (calculate some P-values, confidence intervals, or take a Bayesian perspective") since: (1) t-tests and p-values are useful in some cases but have several limitations especially when one compares distributions that are not normal (in this study, we compared the modal value of topological differences, making any test comparing the mean value useless) and can be biased towards small values (i.e.
+significant ones) when the sample sizes are big; (2) concerning the confidence intervals, we actually did systematically calculate them; ans (3) we are not convinced by a Bayesian approach for comparing distributions.
+
+\textcolor{blue}{
 Dear Mr Guillerme, \\
@@
 
-\textcolor{blue}{Reviewer's Responses to Questions}
-
-\textcolor{blue}{Comments to the Author}
-
-\textcolor{blue}{-------------------------}
-
 \textcolor{blue}{1. Is the manuscript technically sound, and do the data support the conclusions?}
-
-\textcolor{blue}{The manuscript must describe a technically sound piece of scientific research with data that supports the conclusions.
-Experiments must have been conducted rigorously, with appropriate controls, replication, and sample sizes.
-The conclusions must be drawn appropriately based on the data presented.}
 
@@
 See our detailed response below.
-
-\textcolor{blue}{-------------------------}
 
@@
 
-\textcolor{blue}{-------------------------}
-
 \textcolor{blue}{3. Does the manuscript adhere to the PLOS Data Policy?}
-
-\textcolor{blue}{Authors must follow the PLOS Data policy, which requires authors to make all data underlying the findings described in their manuscript fully available without restriction.
-Please refer to the author’s Data Availability Statement in the manuscript.
-All data and related metadata must be deposited in an appropriate public repository, unless already provided as part of the submitted article or supporting information.
-If there are restrictions on the ability of authors to publicly share data—e.g.
-privacy or use of data from a third party— these reasons must be specified.}
 
@@
 
-All the data generated by the simulation is directly available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare}.
+We apologise for the confusion here.
+We mentioned where our data was stored in the supporting information but forgot to add this under the "Data Reproducibility and Availability" section in the manuscript.
+We have now corrected this.
+All the data generated by the simulations are directly available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare}.
+% add link!!
 Note that along with the data we share the full code used for repeating the submitted manuscript, from simulating the data to generating the figures and compiling the manuscript (available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{GitHub}.
+%add link
 All this code is provided with a detailed description of each step that we are happy to develop if needed.
-However, we do apologise since we forgot to add this as a separated section in the supporting information section.
-We added the following part to the new manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-\textbf{Data reproducibility and availability.} All the code used in this analysis is available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{\color{blue}{GitHub}} with some information on how to use the various functions.
-Additionally all the simulated data is available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{\color{blue}{FigShare}}.
-\end{minipage}
-
-\textcolor{blue}{-------------------------}
 
@@
 
-\textcolor{blue}{PLOS ONE does not copyedit accepted manuscripts, so the language in submitted articles must be clear, correct, and unambiguous.
-Any typographical or grammatical errors should be corrected at revision, so please note any specific errors here.}
-
 \textcolor{blue}{Reviewer $\#$ 1: Yes}
 
-\textcolor{blue}{-------------------------}
-
 \textcolor{blue}{5. Review Comments to the Author}
-
-\textcolor{blue}{Please use the space provided to explain your answers to the questions above.
-You may also include additional comments for the author, including concerns about dual publication, research ethics, or publication ethics.
-(Please upload your review as an attachment if it exceeds 20,000 characters)}
 
@@
 \textcolor{blue}{Unfortunately the data analysis performed is simply the calculation of a number of summary statistics about the overlap between the estimates returned for different simulation conditions.
-I suspect that if they authors take some time to do statistics (calculate some P-values, confidence intervals, or take a Bayesian perspective), they'll be able to make a statistical case for most of the points that they are trying to make.
+I suspect that if the authors take some time to do statistics (calculate some P-values, confidence intervals, or take a Bayesian perspective), they'll be able to make a statistical case for most of the points that they are trying to make.
 But I just don't think that a simulation comparison should be publishable without *some* statistical analysis (beyond the calculation of summaries).}
 
-We are slightly concerned by the reviewer's misconception of the term ``statistics" and suspicions (``I suspect that if they authors take some time to do statistics (calculate some P-values, confidence intervals, or take a Bayesian perspective") since: (1) t-tests and p-values are useful in some cases but have several limitations especially when one compares distributions that are not normal (in this study, we compared the modal value of topological differences, making any test comparing the mean value useless) and can be biased towards small values (i.e.
+% Talk to Andrew
+
+%We feel that the reviewer has slightly misunderstood the term ``statistics'' in this case.
+Firstly, p-values are useful in some cases but have several limitations especially when one compares distributions that are not normal (in this study, we compared the modal value of topological differences, making any test comparing the mean value useless) and can be biased towards small values (i.e.
 significant ones) when the sample sizes are big; (2) concerning the confidence intervals, we actually did systematically calculate them; ans (3) we are not convinced by a Bayesian approach for comparing distributions.
@@
 
+% Priors comment??? Move stuff up from below
+
+The reviewer is (I believe) referring to their comment below "".
+
 Our results did indeed biased us in favour of using a Bayesian approach to fix topology since our results clearly indicate that it is less likely to generate a wrong topology using a Bayesian approach (see Figure 4 and 5 and discussion line 472 to 496 in the first submission manuscript).
@@
 %-------------
+% add Mk = Mkv
 
@@
 
-The Mk model assumes that some morphological characters \textbf{can} be constant and the Mkv model assumes that \textbf{all} the characters are variable.
-Although the reviewer is right that real data analyses do not contain invariable (constant) morphological characters, we designed our simulations to be a generalisation of real data.
-Because we sampled the morphological characters rate from a gamma distribution with $\alpha$=0.5, some morphological characters had only a few state changes accross the matrix.
+The Mkv model assumes that all the characters are variable, whereas the Mk model is less constrained - some characters can be invariable and others variable.
+Thus the Mkv model is actually just a special case of the Mk model.
+
+The reviewer is of course correct in that real data rarely contain invariable morphological characters, because these don't contribute to phylogenetic signal.
+Our simulated data, on the other hand, will occasionally contain invariable characters merely because of missing data %
+
+However, this is not an issue because we implement opur analyses in RAxML and MrBayes.
+These programs implement the Mk do not distinguish between both models since the Mkv model is just a particular case of the Mk model and are not distinguished either in the original publication that we cite in the manuscript (Lewis 2001).
+We've changed the mentions of the M\textit{k} model to M\textit{kv} model throughout the new manuscript.
+
+%Because we sampled the morphological characters rate from a gamma distribution with $\alpha$=0.5, some morphological characters had only a few state changes accross the matrix.
 When removing the data, there was a probability for these characters to be constant (e.g.
 if all the taxa in the matrix appart from one had a character state 0 for a given character and only one taxa had a state 1 for the same character, if that taxa had no data available for that character, then the character appeared to be constant in the matrix).
-Note that both RAxML and MrBayes do not distinguish between both models since the Mkv model is just a particular case of the Mk model and are not distinguished either in the original publication that we cite in the manuscript (Lewis 2001).
-We've changed the mentions of the M\textit{k} model to M\textit{kv} model throughout the new manuscript.
 
@@
 
-See our comment above on data availability and reproducibility.
+We apologise again for the confusion here. See our comment above on Data Availability and Reproducibility.
 
@@
 
+% Number the comments
+
 \textcolor{blue}{line 25 and elsewhere: incorrect smart quotes in ``best", ``classical"...}
 
-We fixed the smart quotes in the new manuscript.
+We fixed the smart quotes in our revision. % change
 
@@
 
-\textcolor{blue}{line 113: `` rates from a uniform distribution " with what upper and lower bounds?}
+\textcolor{blue}{line 113: ``rates from a uniform distribution" with what upper and lower bounds?}
 
-Both birth ($\lambda$) and death ($\mu$) parameters are respectively speciation and extinction probability (bounded between 0 and 1). We added this precision in the new version of the manuscript:
+Both birth ($\lambda$) and death ($\mu$) parameters are probabilities bounded between 0 and 1. We have added this to the new version of the manuscript:
 
-Lines 112 to 114 in the previously submitted manuscript:
+%-------------
 
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution but maintaining $\lambda$ $>$ $\mu$.
-\end{minipage}
+\textcolor{blue}{line 118: with 25 living and 25 fossil taxa. Are the fossils just put at the extinction point for the extinct lineages?}
+
+Yes. We have clarified as follows:
 
@@
 \hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform probability distribution (between 0 and 1) but maintaining $\lambda$ $>$ $\mu$.
-\end{minipage}
-
-%-------------
-
-\textcolor{blue}{line 118: wih 25 living and 25 fossil taxa. Are the fossils just put at the extinction point for the extinct lineages?}
-
-Yes. We clarified as follow:
-
-Lines 118-119 in the previously submitted manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-[We] select only trees with 25 living and 25 fossil taxa.
-\end{minipage}
-
-Lines @@@112 to @@@114 in the new manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-[We] select only trees with 25 living and 25 fossil taxa. The fossil taxa where considered as unique points at the end of extinct lineages.
+[We] select only trees with 25 living and 25 fossil taxa. The fossil taxa were considered as unique tips at the end of extinct lineages. % correct in ms
 \end{minipage}
@@
 
-We do not agree with this reviewer's comment.
-Even if our study might be idiosyncratic, we beleive it is not due to the use of the \texttt{diversitree} package instead of the \texttt{TreeSim} package.
-Both perform in generating birth-death trees correctly and are well curated and maintained.
-However, we chose to use the \texttt{diversitree} package rather than the \texttt{TreeSim} package exactly for making the results of our study more applicable to other ones.
-In fact, the publication associated with the diversitree package (FitzJohn 2012, Methods in Ecology and Evolution) has been cited three times more (155 cites on Google Scholar) than the publication associated with the \texttt{TreeSim} package (Stadler 2011, Systematic Biology - 55 cites on Google Scholar).
+We do not agree that the procedure is idiosyncratic.
+\texttt{diversitree} and \texttt{TreeSim} both generate birth-death trees in the same way (although \texttt{diversitree} is slightly quicker) and are well curated and maintained.
+We chose to use the \texttt{diversitree} package because we are more familiar with it.
+We could also have used a number of other packages including TESS % cite etc.
+We are also by no means the only people using the diversitree package - the paper associated with it (FitzJohn 2012, Methods in Ecology and Evolution) has been cited three times more (155 cites on Google Scholar) than the publication associated with the \texttt{TreeSim} package (Stadler 2011, Systematic Biology - 55 cites on Google Scholar).
+We also disagree that using TreeSim would reduce the amount of explanation required.
+We believe that carefully outlining our methods will help people repeat our analyses.
+Thus we would describe any TreeSim analysis in just as much detail.
+Finally of course noone can repeat our analyses precisely as we used random parameters to avoid biasing the analyses towards a particular type of tree.
+We do provide all these trees in the supp mat %link
+so future workers can fully repeat our analyses.
+Thus we feel that repeating the entire set of simulations (1.5 CPU centuries worth) using TreeSim is not necessary and would not influence our results or their utility to future workers.
 
@@
 
-We clarified as follow:
-
-Lines 125 to 126 in the previously submitted manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-The matrix [...] was generated [...] using the HKY model with random base frequencies[...]
-\end{minipage}
+We have clarified as follows:
 
@@
 
-We are not entirely sure if the reviewers refers to a sampling bias or to the problem link to constant characters described in Lewis (2001) (acquisition bias).
+% NEED to fix!
+
+%We are not entirely sure if the reviewers refers to a sampling bias or to the problem link to constant characters described in Lewis (2001) (acquisition bias).
 If the reviewers are referring to the second (acquisition bias), both phylogenetic inference software we used (RAxML and MrBayes) are correcting for the this bias (-K MK  see Supplementary information 1 for the full list of commands).
@@
 
-We clarified this point ($M_C$ comes second) in the new version of the manuscript.
+We clarified this point ($M_C$ comes second) in the new version of the manuscript (see below).
 Also, we clarified the potential confusion between the amount of data removed in one parameter ($M_F$ = 10\% for example) and the amount of missing data in the morphological matrix (5\%).
-
-Lines 171 to 178 in the previously submitted manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-In practice, each parameter represents a different way of removing data from the matrix: $M_L$ removes rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $M_C$ removes columns across both living and fossil taxa's data.
-Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
-\end{minipage}
 
@@
 
-%Probably no impact since it's not using an MCMC. But should we try it? (approx. one to two (max) month calculation).
+Our simulations were designed to test the effects of missing data in as realistic a manner as possible given our constraints.
+Thus, we do not feel that this additional analysis would add much to the paper, especially since it is already rather long and complex.
 
@@
 
-Lines 208 to 210 in the previously submitted manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-We repeated this selection until we obtained 50 sets of simulations (i.e. 50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
-\end{minipage}
+% NC: expand on this a bit
 
@@
 
-\textcolor{blue}{MrBayes: using strongly informative priors for the true parameters does more than speed up the searching, it represents an ``unfair" advantage (an option that would not be availble in real data analysis in which one does not know the true values) for the Bayesian method.
+\textcolor{blue}{MrBayes: using strongly informative priors for the true parameters does more than speed up the searching, it represents an ``unfair" advantage (an option that would not be available in real data analysis in which one does not know the true values) for the Bayesian method.
 This makes the studies report of superior behavior of Bayesian methods less persuasive.}
 
-We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only.The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and where a minor part of the MCMC sampler (2.17\% of move probability for the three priors).
-Therefore we are confident that the topology is not affected allowing us to explore the effect of missing morphological data on topology as well as reducing our computational time to a realistic 1.5 CPU century.
+We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only.
+The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and were a minor part of the MCMC sampler (2.17\% of move probability% check what you call it in paper
+ for the three priors).
+
+% Most of the effects will be branch length and we don't look at that
+
+ Therefore we are confident that the topology is minimally affected by the priors, allowing us to explore the effect of missing morphological data on topology as well as reducing our computational time to a realistic 1.5 CPU centuries.
+
+% Add something to text?
 
@@
 
-We changed the following sentence:
-
-Lines 254 to 255 in the previously submitted manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-The Robinson-Foulds distance, or ``path difference", measures the number of shared clades across two trees.
-\end{minipage}
+We changed the following sentence: % correct
 
@@
 
+% Say thanks for this!
 We changed the following sentence:
-
-Lines 257 to 259 in the previously submitted manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
-\end{minipage}
 
@@
 
-We clarified the scaling method as follow:
-
-Lines 262 to 270 in the previously submitted manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons (Bogdanowicz et al.
-2012).
-This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see Supporting Information S2: Tree Comparisons for the calculation details).
-Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
-\end{minipage}
+We agree that this is confusing and thank the reviewer for pointing this inconsistency out. 
+We clarified the scaling method as follows: % change from distance to metric???
 
@@
 
+% Note that NRF distance is actually a similarity ...
+
 %-------------
@@
 
-We changed the following sentence:
-
-Lines 274 to 275 in the previously submitted manuscript:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-It is bounded between zero when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
-\end{minipage}
+We changed the following sentence: % fix
 
@@
 
+% frequency! correct.
+
 %I'm not sure how to fix that comment.
@@
 
-This comment links back to the reviewer's misconception of the term ``statistics" and suspicions.
+This comment links back to the reviewer's slight misconception of the term ``statistics".
 ANOVA or multiple regression analysis might the appropriate analyses in some cases but we believe it isn't here.
@@
 %Not comparing means, non normal distributions. And non-parametric version (e.g. Kruskal Wallis) not approriate either.
+
+% Talk to Andrew
 
@@
 
-\textcolor{blue}{-------------------------}
+%So for the part about the tree comparisons (RF and Triplets, I think we can just remove the whole thing since it's not especially useful...)
+
+We recognise that this section in our supp mat may be confusing and is actually unnecessary for the understanding the paper. Therefore we have removed this section. % fix!!!
 
@@
 
-%Because we follow Bogdanowicz and use the difference between random (NULL) trees to scale.
+We cited the reference (Bogdanowicz 2012) of the method.
+%explain in a bit more detail
 
@@
 
-%check
-
-%-------------
-
-\textcolor{blue}{-------------------------}
-
-\textcolor{blue}{6. If you would like your identity to be revealed to the authors, please include your name here (optional).}
-
-\textcolor{blue}{Your name and review will not be published with the manuscript.}
-
-\textcolor{blue}{Reviewer $\#$ 1: No thanks.}
-
-%Fair enough!
+We fixed this by removing the summation signs.
 

### 2015-04-28 Manuscript/PLoSONESubmission/OnlineAppendices-LaTeXSuppFiles/SupportingInformation2.tex

--- 2015-01-25 Manuscript/SupplementaryMaterial/TreeComparison.tex
@@
-%\section{Supplementary material Section 2}
+\documentclass[12pt,letterpaper]{article}
+
+%Packages
+\usepackage{pdflscape}
+\usepackage{fixltx2e}
+\usepackage{textcomp}
+\usepackage{fullpage}
+\usepackage{float}
+\usepackage{latexsym}
+\usepackage{url}
+\usepackage{epsfig}
+\usepackage{graphicx}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{bm}
+\usepackage{array}
+\usepackage[version=3]{mhchem}
+\usepackage{ifthen}
+\usepackage{caption}
+\usepackage{hyperref}
+\usepackage{amsthm}
+\usepackage{amstext}
+\usepackage{enumerate}
+\usepackage[osf]{mathpazo}
+\usepackage{dcolumn}
+\usepackage{lineno}
+\pagenumbering{arabic}
+
+%Pagination style and stuff
+\linespread{2}
+\raggedright
+\setlength{\parindent}{0.5in}
+\setcounter{secnumdepth}{0} 
+\renewcommand{\section}[1]{%
+\bigskip
+\begin{center}
+\begin{Large}
+\normalfont\scshape #1
+\medskip
+\end{Large}
+\end{center}}
+\renewcommand{\subsection}[1]{%
+\bigskip
+\begin{center}
+\begin{large}
+\normalfont\itshape #1
+\end{large}
+\end{center}}
+\renewcommand{\subsubsection}[1]{%
+\vspace{2ex}
+\noindent
+\textit{#1.}---}
+\renewcommand{\tableofcontents}{}
+
+\begin{document}
+
+\section{Supporting Information S2: Tree Comparisons}
 
 \subsection{Robinson-Foulds distance}
-Robinson-Foulds distance (\textit{RF}; \citealp{RF1981}), or "path difference", measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} and can be expressed as:
+Robinson-Foulds distance (\textit{RF} \cite{RF1981}), or "path difference", measures the number of shared clades across two trees.
+The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} and can be expressed as:
 \begin{equation}
@@
 This metric is more sensitive to taxon displacement than Triplets distance (i.e.
-if one taxon moves out of a clade, then the clades are no longer considered similar; \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
+if one taxon moves out of a clade, then the clades are no longer considered similar \cite{critchlowthe1996,johnson1998,wiensmissing2003}).
 The minimal value of $C$ is equal to 1 if the two trees have the same n taxa; the maximal value in $C = n-2$.
@@
 \end{equation}
-This metric is more sensitive to taxa displacement than the Triplets distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.
+This metric is more sensitive to taxa displacement than the Triplets distance \cite{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.
 
 \subsection{Triplets distance details ($T_{x,y}$)}
-Triplets distance ($T_{x,y}$; \citealp{dobson1975triplets}) measures the number of sub-trees made up of three taxa (triplets) that differ between two given trees.
+Triplets distance ($T_{x,y}$ \cite{dobson1975triplets}) measures the number of sub-trees made up of three taxa (triplets) that differ between two given trees.
 Each triplet can be written as $I_{ijk}$=(\textit{ijk}).
 Where $I_{ijk}$ is equal to zero if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to one.
-For any rooted binary tree there are only three possible combinations for each triplet: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
+For any rooted binary tree there are only three possible combinations for each triplet: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}) \cite{johnson1998}.
 If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k}).
@@
 \end{equation}
-and where \textit{n} is the total number of taxa in both trees (modified from \citet{critchlowthe1996}).
+and where \textit{n} is the total number of taxa in both trees (modified from \cite{critchlowthe1996}).
 If $S_n$ = 0, the trees are identical; when $S_n$ = $\binom{n}{4}$, the trees are as different as possible (i.e.
@@
 \subsection{Normalised Tree Similarity}
-For any tree with \textit{n} taxa compared using a tree distance metric $m$, Normalized Tree Similarity, $NTS_m$ \citep{Bogdanowicz2012}, represents the similarity score for the two trees given the expected distance between two random Yule trees with $n$ taxa.
+For any tree with \textit{n} taxa compared using a tree distance metric $m$, Normalized Tree Similarity, $NTS_m$ \cite{Bogdanowicz2012}, represents the similarity score for the two trees given the expected distance between two random Yule trees with $n$ taxa.
 If $\bar{d}_{m,n}$\textit{(rand)} is the average distance between two random Yule trees with $n$ taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} each containing $n$ taxa, then:
@@
 
-We used the NTS method to scale all the Robinson Foulds and Triplets distances calculated in our analyses, using the TreeCmp java script \citep{Bogdanowicz2012}.
-
-% NC: Not sure this bit is needed
-%We compared 1000 randomly generated trees to obtain the mean comparison value $\bar{d}_{m,n}$\textit{(rand)}.
-We randomly generated two sets of 1000 trees of \textit{n} taxa using the rmtree function of ape package (v3.0-11 \citet{paradisape:2004}) that generates a given number of random Yule trees.
-We calculated the $\bar{d}_{m,n}$\textit{(rand)} value using an approach similar to the RPCBTC (described below) by performing 1000 random pairwise comparisons using the TreeCmp java script \citep{Bogdanowicz2012}.
+We used the NTS method to scale all the Robinson Foulds and Triplets distances calculated in our analyses, using the TreeCmp java script \cite{Bogdanowicz2012}.
 
 \subsection{Bhattacharyya Coefficient}
-The Bhattacharyya Coefficient calculates the probability of overlap of two distributions \citep{Bhattacharyya}.
+The Bhattacharyya Coefficient calculates the probability of overlap of two distributions \cite{Bhattacharyya}.
 When it is equal to zero, the probability of overlap of the distributions is also zero, and when it is equal to one, the two distributions are entirely overlapping.
@@
 If $n$ is low, the overlap will be overestimated and if $n$ is too high, the overlap will be underestimated.
-In this analysis, we determined the number of bins using Silverman's rule of thumb which states that $n$ should be 0.9 times the minimum of the standard deviation and the interquartile range of the distribution, divided by 1.34 times the sample size of the distribution to the negative one-fifth power (bw.nrd0() function in R; \citet{silverman1986density}).
+In this analysis, we determined the number of bins using Silverman's rule of thumb which states that $n$ should be 0.9 times the minimum of the standard deviation and the interquartile range of the distribution, divided by 1.34 times the sample size of the distribution to the negative one-fifth power (\texttt{bw.nrd0()} function in R \cite{silverman1986density}).
+
+\bibliographystyle{plos2009}
+\bibliography{Supp_Ref}
+
+%END
+\end{document}

### 2015-04-28 Manuscript/PLoSONESubmission/OnlineAppendices-LaTeXSuppFiles/SupportingInformation3.tex

--- 2015-01-29 Manuscript/SupplementaryMaterial/Supp_results.tex
@@
+\documentclass[12pt,letterpaper]{article}
+
+%Packages
+\usepackage{pdflscape}
+\usepackage{fixltx2e}
+\usepackage{textcomp}
+\usepackage{fullpage}
+\usepackage{float}
+\usepackage{latexsym}
+\usepackage{url}
+\usepackage{epsfig}
+\usepackage{graphicx}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{bm}
+\usepackage{array}
+\usepackage[version=3]{mhchem}
+\usepackage{ifthen}
+\usepackage{caption}
+\usepackage{hyperref}
+\usepackage{amsthm}
+\usepackage{amstext}
+\usepackage{enumerate}
+\usepackage[osf]{mathpazo}
+\usepackage{dcolumn}
+\usepackage{lineno}
+\pagenumbering{arabic}
+
+%Pagination style and stuff
+\linespread{2}
+\raggedright
+\setlength{\parindent}{0.5in}
+\setcounter{secnumdepth}{0} 
+\renewcommand{\section}[1]{%
+\bigskip
+\begin{center}
+\begin{Large}
+\normalfont\scshape #1
+\medskip
+\end{Large}
+\end{center}}
+\renewcommand{\subsection}[1]{%
+\bigskip
+\begin{center}
+\begin{large}
+\normalfont\itshape #1
+\end{large}
+\end{center}}
+\renewcommand{\subsubsection}[1]{%
+\vspace{2ex}
+\noindent
+\textit{#1.}---}
+\renewcommand{\tableofcontents}{}
+
+\begin{document}
+
+\section{Supporting Information S3: Additional Results}
+
 The following section contains the supplemental results of the effects of our missing data parameters and the different tree inference methods on the the ability to recover the "best" topology that are briefly discussed in the main body of the paper.
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{L}$ missing data parameter only.}
 \label{Tab_Supp_summary_BC_ML}
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{F}$ missing data parameter only.}
 \label{Tab_Supp_summary_BC_MF}
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{C}$ missing data parameter only.}
 \label{Tab_Supp_summary_BC_MC}
@@
 \end{landscape}
+%END
+\end{document}

### 2015-04-28 Manuscript/PLoSONESubmission/TEM_PLoS_draft.tex

@@
 
-% Use adjustwidth environment to exceed column width (see example table in text)
 \usepackage{changepage}
-% Use Unicode characters when possible
 \usepackage[utf8]{inputenc}
-% textcomp package and marvosym package for additional characters
 \usepackage{textcomp,marvosym}
-% fixltx2e package for \textsubscript
 \usepackage{fixltx2e}
-% amsmath and amssymb packages, useful for mathematical formulas and symbols
 \usepackage{amsmath,amssymb}
-% cite package, to clean up citations in the main text. Do not remove.
 \usepackage{cite}
-% Use nameref to cite supporting information files (see Supporting Information section for more info)
 \usepackage{nameref,hyperref}
-% line numbers
 \usepackage[right]{lineno}
-% ligatures disabled
 \usepackage{microtype}
 \DisableLigatures[f]{encoding = *, family = * }
-% rotating package for sideways tables
 \usepackage{rotating}
@@
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see Appendix S1 and Fig.
-S1).
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see ~\nameref{S1_Appendix1} and Fig.
+S1 within).
 We then ran an independent discrete character simulation for each character using the "true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
@@
 The Robinson-Foulds distance \cite{RF1981}, or "path difference", measures the number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} (see Appendix S2 for calculation details).
+The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} (see ~\nameref{S2_Appendix2} for calculation details).
 This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
@@
 As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7).
-However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7).
+The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7 in ~\nameref{S3_Appendix3}).
+However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7 in ~\nameref{S3_Appendix3}).
 
@@
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; Tables S2, S3 and S4).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; Tables S2, S3 and S4 in ~\nameref{S3_Appendix3}).
 Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
@@
 % Include only the SI item label in the subsection heading. Use the \nameref{label} command to cite SI items in the text.
-\subsection*{Supplementary informations}
+\subsection*{Supporting Information S1: Tree Inference}
 \label{S1_Appendix1}
-\noindent{\bf Appendix S1 Tree Inference.} Supplementary information on the Tree Inference methods.
+\noindent{\bf Supporting Information S1 Tree Inference.} Supplementary information on the Tree Inference methods. Contains Figure S1.
 
-\label{Figure S1}
-\noindent{\bf Figure S1 proportion of morphological characters states.}
+\subsection*{Supporting Information S2: Tree Comparisons}
+\label{S2_Appendix2}
+\noindent{\bf Supporting Information S2 Tree Comparisons.} Supplementary information on the Tree Comparison methods.
 
-\label{S2_Appendix2}
-\noindent{\bf Appendix S2 Tree Comparisons.} Supplementary information on the Tree Comparison methods.
-
-\label{S2_Appendix3}
-\noindent{\bf Appendix S3 Additional results.}.
-
-\label{Figure S2}
-\noindent{\bf Figure S2 Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey).}
-
-\label{Figure S3}
-\noindent{\bf Figure S3 Effects of missing data on topological recovery using Maximum Likelihood trees.}
-
-\label{Figure S4}
-\noindent{\bf Figure S4 Effects of missing data on topological recovery using Maximum Likelihood bootstrap trees.}
-
-\label{Figure S5}
-\noindent{\bf Figure S5 Effects of missing data on topological recovery using Bayesian posterior tree distribution.}
-
-\label{Table S1}
-\noindent{\bf Table S1 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method.}
-
-\label{Table S2}
-\noindent{\bf Table S2 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method for the $M_{L}$ parameter.}
-
-\label{Table S3}
-\noindent{\bf Table S3 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method for the $M_{F}$ parameter.}
-
-\label{Table S4}
-\noindent{\bf Table S4 Comparisons between the "best" tree and the "missing data" trees for each different tree inference method for the $M_{C}$ parameter.}
-
-\label{Table S5}
-\noindent{\bf Table S5 Bhattacharyya Coefficients of the pairwise method comparisons for the $M_{L}$ parameter.}
-
-\label{Table S6}
-\noindent{\bf Table S6 Bhattacharyya Coefficients of the pairwise method comparisons for the $M_{F}$ parameter..} Bhattacharyya
-
-\label{Table S7}
-\noindent{\bf Table S7 Bhattacharyya Coefficients of the pairwise method comparisons for the $M_{C}$ parameter..} Bhattacharyya
+\subsection*{Supporting Information S3: Additional results}
+\label{S3_Appendix3}
+\noindent{\bf Supporting Information S3 Additional results.}. Contains Figures S2, S3, S4 and S5 and Tables S1, S2, S3, S4 S5, S6 and S7.
 

### 2015-04-28 Manuscript/PLoSONEresponse/CoverLetter.tex

--- 2015-01-27 Manuscript/CoverLetter.tex
@@
 
-Our paper, entitled "Effects of missing data on topological inference using a Total Evidence approach", is to our knowledge, the first to thoroughly analyze the effects of missing data on tree topology in a Total Evidence framework.
+Our research article, entitled "Effects of missing data on topological inference using a Total Evidence approach", is to our knowledge, the first to thoroughly analyze the effects of missing data on tree topology in a Total Evidence framework.
 Using simulations ($>$ 150 CPU years worth), we find that the number of living taxa with morphological data and the overall number of morphological characters, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology.
@@
 This has major implications for clades where detailed morphological data collection from living species is rare.
-Additionally, we recommend using a Bayesian majority consensus tree when fixing tree topology for any additional analyses such as tip-dating (Ronquist et al.
-2012 Syst Biol).
+Additionally, we recommend using a Bayesian majority consensus tree when fixing tree topology for any additional analyses.
 
-We feel this paper will be of particular interest to the readers of Systematic Biology as many papers investigating methods for combining living and fossil taxa and the effects of missing data on such analyses (e.g.
-D\'{a}valos et al.
-2014; Pattinson et al.
-in press, Sylvestro et al.
-2014, Pyron 2011, Ronquist et al.
-2012), have been recently published in the journal.
+We would like to submit this as a research article. We have had no prior interactions with PLoS regarding this article.
+
+As academic editors we suggest Matthew C. Mihlbachler, Patrick O'Grady and Nicholas Pyenson. We have no opposed reviewers or editors.
 
@@
 
-Suggested reviewers: Liliana D\'{a}valos; David Pattinson; Fredrik Ronquist; Nicolas Salamin; Tanja Stadler.
-
 \end{letter}

### 2015-04-28 Manuscript/PLoSONEresponse/OnlineAppendices-LaTeXSuppFiles/Supp_results.tex

--- 2015-04-28 Manuscript/PLoSONESubmission/OnlineAppendices-LaTeXSuppFiles/SupportingInformation3.tex
@@
-\documentclass[12pt,letterpaper]{article}
-
-%Packages
-\usepackage{pdflscape}
-\usepackage{fixltx2e}
-\usepackage{textcomp}
-\usepackage{fullpage}
-\usepackage{float}
-\usepackage{latexsym}
-\usepackage{url}
-\usepackage{epsfig}
-\usepackage{graphicx}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{bm}
-\usepackage{array}
-\usepackage[version=3]{mhchem}
-\usepackage{ifthen}
-\usepackage{caption}
-\usepackage{hyperref}
-\usepackage{amsthm}
-\usepackage{amstext}
-\usepackage{enumerate}
-\usepackage[osf]{mathpazo}
-\usepackage{dcolumn}
-\usepackage{lineno}
-\pagenumbering{arabic}
-
-%Pagination style and stuff
-\linespread{2}
-\raggedright
-\setlength{\parindent}{0.5in}
-\setcounter{secnumdepth}{0} 
-\renewcommand{\section}[1]{%
-\bigskip
-\begin{center}
-\begin{Large}
-\normalfont\scshape #1
-\medskip
-\end{Large}
-\end{center}}
-\renewcommand{\subsection}[1]{%
-\bigskip
-\begin{center}
-\begin{large}
-\normalfont\itshape #1
-\end{large}
-\end{center}}
-\renewcommand{\subsubsection}[1]{%
-\vspace{2ex}
-\noindent
-\textit{#1.}---}
-\renewcommand{\tableofcontents}{}
-
-\begin{document}
-
-\section{Supporting Information S3: Additional Results}
-
 The following section contains the supplemental results of the effects of our missing data parameters and the different tree inference methods on the the ability to recover the "best" topology that are briefly discussed in the main body of the paper.
@@
 \end{landscape}
-%END
-\end{document}

### 2015-04-28 Manuscript/PLoSONEresponse/TEM_PLoS_response.tex

@@
 \paragraph*{Robinson-Foulds distance}
-The Robinson-Foulds distance \cite{RF1981}, or ``path difference", measures the number of shared clades across two trees.
+The Robinson-Foulds distance \cite{RF1981}, or ``path difference", measures the difference between the number of clades and the twice number of shared clades across two trees.
 The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} (see ~\nameref{S2_Appendix2} for calculation details).
-This metric is bounded between one, when the two trees are identical, and $n-2$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
+This metric is bounded between zero, when the two trees are identical, and $2(n-2)$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
 This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
-We normalised this metric by dividing the Robinson-Foulds distance by the mean Robinson-Foulds distance from 1000 random tree comparisons \cite{Bogdanowicz2012}.
-This allows us to compare the metric for any number of taxa by scaling it with the expected value of the metric when comparing two random trees (see ~\nameref{S2_Appendix2} for the calculation details).
-Therefore, when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
+We normalised this metric following Bogdanowicz \textit{et al.} \cite{Bogdanowicz2012}'s Normalised Tree Similarity (NTS) method.
+This methods scales %not sure if scale here?
+any tree comparison metric using the mean distance between 1000 random trees (see ~\nameref{S2_Appendix2} for the calculation details).
+This method is a generalisation of the topological accuracy method \cite{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.
+In practice when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
 
@@
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
-It is bounded between zero when the two trees are identical and $\binom{n}{4}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
+It is bounded between zero when the two trees are identical and $\binom{n}{3}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
 Therefore this metric sensitive to the conservation of wildcard taxa.
@@
 The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.
-Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
+Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
 \label{Fig_Bhattacharyya_Coefficients1}
@@
 The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$.
-Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
+Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
 \label{Fig_Bhattacharyya_Coefficients2} 

### 2015-04-28 Manuscript/PLoSONEresponse/response.tex
Fixed some comments (incl. figures)

@@
 We apologise for the confusion here.
-We mentioned where our data was stored in the supporting information but forgot to add this under the "Data Reproducibility and Availability" section in the manuscript.
+We mentioned where our data was stored in the supporting information but forgot to add this under the ``Data Reproducibility and Availability" section in the manuscript.
 We have now corrected this.
-All the data generated by the simulations are directly available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare}.
-% add link!!
-Note that along with the data we share the full code used for repeating the submitted manuscript, from simulating the data to generating the figures and compiling the manuscript (available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{GitHub}.
-%add link
+All the data generated by the simulations are directly available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare} (goo.gl/h5Kxb5).
+Note that along with the data we share the full code used for repeating the submitted manuscript, from simulating the data to generating the figures and compiling the manuscript (available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{GitHub} (goo.gl/ewRsrm).
 All this code is provided with a detailed description of each step that we are happy to develop if needed.
@@
 
-\textcolor{blue}{Unfortunately the data analysis performed is simply the calculation of a number of summary statistics about the overlap between the estimates returned for different simulation conditions.
+\textcolor{blue}{A-Unfortunately the data analysis performed is simply the calculation of a number of summary statistics about the overlap between the estimates returned for different simulation conditions.
 I suspect that if the authors take some time to do statistics (calculate some P-values, confidence intervals, or take a Bayesian perspective), they'll be able to make a statistical case for most of the points that they are trying to make.
@@
 
-\textcolor{blue}{The bias in the analysis in favor of Bayesian methods makes the recommendation to use Bayesian methods less satisfying.}
+\textcolor{blue}{B-The bias in the analysis in favor of Bayesian methods makes the recommendation to use Bayesian methods less satisfying.}
 
@@
 
-The reviewer is (I believe) referring to their comment below "".
+The reviewer is (I believe) referring to their comment below (number 13).
+We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only.
+The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and were a minor part of the MCMC sampler (2.17\% of move probability% check what you call it in paper
+ for the three priors).
 
-Our results did indeed biased us in favour of using a Bayesian approach to fix topology since our results clearly indicate that it is less likely to generate a wrong topology using a Bayesian approach (see Figure 4 and 5 and discussion line 472 to 496 in the first submission manuscript).
+% Most of the effects will be branch length and we don't look at that
+
+Therefore we are confident that the topology is minimally affected by the priors, allowing us to explore the effect of missing morphological data on topology as well as reducing our computational time to a realistic 1.5 CPU centuries.
+
+% Add something to text?
+
+%Our results did indeed biased us in favour of using a Bayesian approach to fix topology since our results clearly indicate that it is less likely to generate a wrong topology using a Bayesian approach (see Figure 4 and 5 and discussion line 472 to 496 in the first submission manuscript).
 % TG: or I'm not sure I understood that one.
@@
 
-\textcolor{blue}{The use of Mk rather than Mkv makes the study less relevant to real data analyses.}
+\textcolor{blue}{C-The use of Mk rather than Mkv makes the study less relevant to real data analyses.}
 
-The Mkv model assumes that all the characters are variable, whereas the Mk model is less constrained - some characters can be invariable and others variable.
+Throughout our paper, we referred the M\textit{k} model as being the one described in Lewis (2001) (an equal rate Q-matrix with \textit{k} character states).
+The Mkv model is identical to the Mk model (as described in Lewis 2001) however, in practice, it assumes that all characters are variable and that therefore, one should correct for acquisition bias that can induce branch length inflation.
+This is generally by adding an algorithm in the phylogenetic software to add ``dummy'' morphological characters.
+In some software (e.g.
+GARLI or BEAST), it is important to specify whether the morphological data should be corrected for acquisition bias or not (Mk or Mkv options).
+However, in RAxML and MrBayes, this bias is corrected automatically by addition of ``dummy'' characters.
+
+The reviewer is of course correct in that real data rarely contain invariable morphological characters, because these don't contribute to phylogenetic signal.
+Our simulated data, on the other hand, will occasionally contain invariable characters merely because of missing data, therefore we made the decision to describe the model for morphological characters evolution as the M\textit{k} model (an equal rate Q-matrix with \textit{k} character states where the rate an be equal to 0).
+RAxML and MrBayes did however corrected for acquisition bias as implemented in their algorithm for treating morphological data.
+
+However, following the reviewer comment, we changed the designation of the model from M\textit{k} to M\textit{kv} to explicitly underline that we corrected for acquisition bias.
+
+%The Mkv model assumes that all the characters are variable, whereas the Mk model is less constrained - some characters can be invariable and others variable.
 Thus the Mkv model is actually just a special case of the Mk model.
 
-The reviewer is of course correct in that real data rarely contain invariable morphological characters, because these don't contribute to phylogenetic signal.
+%The reviewer is of course correct in that real data rarely contain invariable morphological characters, because these don't contribute to phylogenetic signal.
 Our simulated data, on the other hand, will occasionally contain invariable characters merely because of missing data %
 
-However, this is not an issue because we implement opur analyses in RAxML and MrBayes.
+%However, this is not an issue because we implement opur analyses in RAxML and MrBayes.
 These programs implement the Mk do not distinguish between both models since the Mkv model is just a particular case of the Mk model and are not distinguished either in the original publication that we cite in the manuscript (Lewis 2001).
-We've changed the mentions of the M\textit{k} model to M\textit{kv} model throughout the new manuscript.
+We've changed the mentions of the M\textit{k} model to M\textit{kv} model throughout our revision.
 
@@
 
-\textcolor{blue}{Many of the equations are incorrect.}
+\textcolor{blue}{D-Many of the equations are incorrect.}
 
@@
 
-\textcolor{blue}{If there was a mention of the data being deposited, I missed it. I apologize if I did miss it.}
+\textcolor{blue}{E-If there was a mention of the data being deposited, I missed it. I apologize if I did miss it.}
 
@@
 
-% Number the comments
+\textcolor{blue}{1-line 25 and elsewhere: incorrect smart quotes in ``best", ``classical"...}
 
-\textcolor{blue}{line 25 and elsewhere: incorrect smart quotes in ``best", ``classical"...}
-
-We fixed the smart quotes in our revision. % change
+We fixed the smart quotes in our revision.
 
@@
 
-\textcolor{blue}{line 88 ``used it to infer a matrix" $->$ ``used it to simulate a matrix". ``inference" is used in place of simulation elsewhere.}
+\textcolor{blue}{2-line 88 ``used it to infer a matrix" $->$ ``used it to simulate a matrix". ``inference" is used in place of simulation elsewhere.}
 
@@
 
-\textcolor{blue}{line 112: does diversitree just stop simulating when it reaches the correct number of tips, or does it deal with the biases that can result from doing that?
+\textcolor{blue}{3-line 112: does diversitree just stop simulating when it reaches the correct number of tips, or does it deal with the biases that can result from doing that?
 (Stadler 2011 is the ref for a discussion of this issues, I believe)}
@@
 
-\textcolor{blue}{line 113: ``rates from a uniform distribution" with what upper and lower bounds?}
+\textcolor{blue}{4-line 113: ``rates from a uniform distribution" with what upper and lower bounds?}
 
@@
 
-\textcolor{blue}{line 118: with 25 living and 25 fossil taxa. Are the fossils just put at the extinction point for the extinct lineages?}
+\textcolor{blue}{5-line 118: with 25 living and 25 fossil taxa. Are the fossils just put at the extinction point for the extinct lineages?}
 
@@
 
-Lines @@@112 to @@@114 in the new manuscript:
+Lines @@@112 to @@@114 in our revision:
 
@@
 
-\textcolor{blue}{line 110 - line 120.
+\textcolor{blue}{6-line 110 - line 120.
 The procedure for generating a tree with fossils is a valid procedure.
@@
 We chose to use the \texttt{diversitree} package because we are more familiar with it.
-We could also have used a number of other packages including TESS % cite etc.
-We are also by no means the only people using the diversitree package - the paper associated with it (FitzJohn 2012, Methods in Ecology and Evolution) has been cited three times more (155 cites on Google Scholar) than the publication associated with the \texttt{TreeSim} package (Stadler 2011, Systematic Biology - 55 cites on Google Scholar).
-We also disagree that using TreeSim would reduce the amount of explanation required.
+We could also have used a number of other packages including \texttt{TESS}.
+We are also by no means the only people using the \texttt{diversitree} package - the paper associated with it (FitzJohn 2012, Methods in Ecology and Evolution) has been cited three times more (155 cites on Google Scholar) than the publication associated with the \texttt{TreeSim} package (Stadler 2011, Systematic Biology - 55 cites on Google Scholar).
+
+We also disagree that using \texttt{TreeSim} would reduce the amount of explanation required.
 We believe that carefully outlining our methods will help people repeat our analyses.
-Thus we would describe any TreeSim analysis in just as much detail.
-Finally of course noone can repeat our analyses precisely as we used random parameters to avoid biasing the analyses towards a particular type of tree.
-We do provide all these trees in the supp mat %link
-so future workers can fully repeat our analyses.
-Thus we feel that repeating the entire set of simulations (1.5 CPU centuries worth) using TreeSim is not necessary and would not influence our results or their utility to future workers.
+Thus we would describe any \texttt{TreeSim} analysis in just as much detail.
+Finally of course no one can repeat our analyses precisely as we used random parameters to avoid biasing the analyses towards a particular type of tree.
+However, we do provide all these trees in the supporting information (``Data Reproducibility and Availability" section; \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare} link (goo.gl/h5Kxb5)) so future workers can fully repeat our analyses.
+Thus we feel that repeating the entire set of simulations (1.5 CPU centuries worth) using \texttt{TreeSim} is not necessary and would not influence our results or their utility to future workers.
 
@@
 
-\textcolor{blue}{line 126 ``random base frequencies" is too vague. Dirichlet(1, 1, 1, 1) perhaps?}
+\textcolor{blue}{7-line 126 ``random base frequencies" is too vague. Dirichlet(1, 1, 1, 1) perhaps?}
 
@@
 
-\textcolor{blue}{Do the simulations apply ascertainment biases as described by Lewis' Mkv (2001) model? It seems like they should... At minimum, the point should be discussed.}
+\textcolor{blue}{8-Do the simulations apply ascertainment biases as described by Lewis' Mkv (2001) model? It seems like they should... At minimum, the point should be discussed.}
 
+See our response to comment C above.
 % NEED to fix!
@@
 
-\textcolor{blue}{line 171 - 177: the authors need to explain the order of operations for the $M_F$ vs $M_C$ masking.
+\textcolor{blue}{9-line 171 - 177: the authors need to explain the order of operations for the $M_F$ vs $M_C$ masking.
 It makes a difference.
@@
 
-\textcolor{blue}{Header for 3. should be ``building phylogenetic estimates" or ``estimating phylogenies" because the previous section described how the phylogenies and character data were ``built."}
+\textcolor{blue}{10-Header for 3.
+should be ``building phylogenetic estimates" or ``estimating phylogenies" because the previous section described how the phylogenies and character data were ``built."}
 
@@
 
-\textcolor{blue}{RAxML searching: it would be nice to see some searches started from the true tree.
+\textcolor{blue}{11-RAxML searching: it would be nice to see some searches started from the true tree.
 This is clearly not an option for real data sets, but it is helpful in simulation studies because it lets the reader determine if the results could be an artifact of insufficient tree searching.}
@@
 
-\textcolor{blue}{Filtering the data sets for strong bootstrap support under the full data analysis surely introduces some bias in favor of the no-missing data analyses.}
+\textcolor{blue}{12-Filtering the data sets for strong bootstrap support under the full data analysis surely introduces some bias in favor of the no-missing data analyses.}
 
@@
 If the no-missing data tree has low node support, it is likely that the changes in topology for the missing data trees are not linked to our focal question (the effect of removing characters, living taxa or fossils) but just to the fact that low supported nodes can be just random.
-We've added the following justification in the new manuscript:
+We've added the following justification in our revision:
 
@@
 
-\textcolor{blue}{MrBayes: using strongly informative priors for the true parameters does more than speed up the searching, it represents an ``unfair" advantage (an option that would not be available in real data analysis in which one does not know the true values) for the Bayesian method.
+\textcolor{blue}{13-MrBayes: using strongly informative priors for the true parameters does more than speed up the searching, it represents an ``unfair" advantage (an option that would not be available in real data analysis in which one does not know the true values) for the Bayesian method.
 This makes the studies report of superior behavior of Bayesian methods less persuasive.}
 
-We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only.
-The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and were a minor part of the MCMC sampler (2.17\% of move probability% check what you call it in paper
- for the three priors).
-
-% Most of the effects will be branch length and we don't look at that
-
- Therefore we are confident that the topology is minimally affected by the priors, allowing us to explore the effect of missing morphological data on topology as well as reducing our computational time to a realistic 1.5 CPU centuries.
-
-% Add something to text?
+See our response to comment B above.
 
@@
 
-\textcolor{blue}{line 254: RF measures the difference between the number of clades and the twice number of shared clades across two trees.
+\textcolor{blue}{14-line 254: RF measures the difference between the number of clades and the twice number of shared clades across two trees.
 The statement that it ``measures the number of shared clades across two trees" should be revised.}
@@
 
-\textcolor{blue}{line 258: RF is 0 (not 1) when the trees are identical. It's maximum (for 2 rooted binary trees) is 2(n-2) not n-2.}
+\textcolor{blue}{15-line 258: RF is 0 (not 1) when the trees are identical. It's maximum (for 2 rooted binary trees) is 2(n-2) not n-2.}
 
-% Say thanks for this!
-We changed the following sentence:
+We thanks the reviewer for pointing at this error and we changed the following sentence:
 
-Lines @@@112 to @@@114 in the new manuscript:
+Lines @@@112 to @@@114 in our revision:
 
@@
 
-\textcolor{blue}{The ``scaling" to produce the Normalized RF ``distance" is confusing. It is also inverting the sign of the distance, so it has become a measure of similarity.}
+\textcolor{blue}{16-The ``scaling" to produce the Normalized RF ``distance" is confusing. It is also inverting the sign of the distance, so it has become a measure of similarity.}
 
 We agree that this is confusing and thank the reviewer for pointing this inconsistency out.
-We clarified the scaling method as follows: % change from distance to metric???
+We changed the mentions of ``Normalised Robinson-Foulds distance'' and ``Normalised Triplets distance'' to ``Normalised Robinson-Foulds metric'' and ``Normalised Triplets metric'' throughout the our revision.
+Additionally, we clarified the scaling method as follows:
 
-Lines @@@112 to @@@114 in the new manuscript:
+Lines @@@112 to @@@114 in our revision:
 
@@
 This method is a generalisation of the topological accuracy method (Price \textit{et al.} 2010) allowing to compare topological differences between any tree with any tree comparison metric.
-In practice when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
+In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
 \end{minipage}
-
 % Note that NRF distance is actually a similarity ...
@@
 
-\textcolor{blue}{line 275: why isn't the upper bound of the triples distance n choose 3 (instead of n choose 4). I think you got the formula from an unrooted quartet distance.}
+\textcolor{blue}{17-line 275: why isn't the upper bound of the triples distance n choose 3 (instead of n choose 4). I think you got the formula from an unrooted quartet distance.}
 
@@
 
-\textcolor{blue}{Using the BC stat may be similar in spirit to using a t-test, but it is not equivalent.}
+\textcolor{blue}{18-Using the BC stat may be similar in spirit to using a t-test, but it is not equivalent.}
 
@@
 
-\textcolor{blue}{Figure 3: B is confusing. If both curves are probability distributions, then they both have to integrate to 1. So B(y) can't cover just a subset of the area covered by B(x).}
+\textcolor{blue}{19-Figure 3: B is confusing. If both curves are probability distributions, then they both have to integrate to 1. So B(y) can't cover just a subset of the area covered by B(x).}
 
-% frequency! correct.
-
-%I'm not sure how to fix that comment.
-The figure caption says (first sentence!) : "A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
-ML = 10%,MF = 50%, MC = 75%)." Should I make it super clear by writing : "A and B are distributions of tree similarity metrics (not probability distributions, or roller coasters plans)" ?
+We thank the reviewer for pointing out this error. We fixed the shape of the distribution in Figure 3-B.
 
@@
 
-\textcolor{blue}{``Combined effect of missing data parameters" section ANOVA or multiple regression are the appropriate analyses to tease apart multiple interacting factors.
+\textcolor{blue}{20-``Combined effect of missing data parameters" section ANOVA or multiple regression are the appropriate analyses to tease apart multiple interacting factors.
 Here the authors opt to describe results without statistical tests, which is disappointing.}
@@
 
-\textcolor{blue}{line 398: ``mMthods" $->$ ``Methods"}
+\textcolor{blue}{21-line 398: ``mMthods" $->$ ``Methods"}
 
@@
 \textcolor{blue}{Supplement 2:}
-\textcolor{blue}{more confusion about the min RF here.
+\textcolor{blue}{22-more confusion about the min RF here.
 Also note that if you are counting the entire leaf set as a clade (e.g the star tree having N=1), then the number of clades in rooted binary tree is n-1 (not n-2).
@@
 
-\textcolor{blue}{The definition of the scaled RF implies that it is never positive (sinc the numerator cannot be positive and the denominator is positive).}
+\textcolor{blue}{23-The definition of the scaled RF implies that it is never positive (sinc the numerator cannot be positive and the denominator is positive).}
 
@@
 
-\textcolor{blue}{The decision to use Yule trees as the base line for the mean distance is not explained, and seems odd given that the true tree are not Yule trees.
+\textcolor{blue}{24-The decision to use Yule trees as the base line for the mean distance is not explained, and seems odd given that the true tree are not Yule trees.
 I'm not sure that it makes much difference, but it is odd.}
@@
 
-\textcolor{blue}{The NTS can only -infinity for distances that have the property that the mean distance between random Yule trees is 0.
+\textcolor{blue}{25-The NTS can only -infinity for distances that have the property that the mean distance between random Yule trees is 0.
 This is not true of either distance metric here.
@@
 
-As clarified in the new manuscript (lines @@@), the NTS is a generalisation of the topological accuracy method (Price \textit{et al.} 2010 PLoS ONE) and is therefore not specialised for the RF or the Triples distance.
+As clarified in our revision (lines @@@), the NTS is a generalisation of the topological accuracy method (Price \textit{et al.} 2010 PLoS ONE) and is therefore not specialised for the RF or the Triples distance.
 
@@
 
-\textcolor{blue}{Note that JavaScript is a language, so it a bit confusing to refer to a Java program as a ``Java script."}
+\textcolor{blue}{26-Note that JavaScript is a language, so it a bit confusing to refer to a Java program as a ``Java script."}
 
@@
 
-\textcolor{blue}{Equation (6) is not correct.
+\textcolor{blue}{27-Equation (6) is not correct.
 You shouldn't set the summation sign as the left hand side of an equation because the notation has a standard definition.
@@
 
-\textcolor{blue}{Equation (7) is true for a three-leaf tree with a polytomy occurring with probability = 1/4 and the other 3 trees being equiprobable.
+\textcolor{blue}{28-Equation (7) is true for a three-leaf tree with a polytomy occurring with probability = 1/4 and the other 3 trees being equiprobable.
 I don't think that is is true for any other weighting of trees for other numbers of taxa.
@@
 
-\textcolor{blue}{I don't think that equations 8, 9, or 10 are correct for any case other than the 3 taxon case (though I do note that the authors are using n choose 3 here).}
+\textcolor{blue}{29-I don't think that equations 8, 9, or 10 are correct for any case other than the 3 taxon case (though I do note that the authors are using n choose 3 here).}
 
@@
 
-\textcolor{blue}{Equations 13 and 14. It seems easier to just drop the summation sign here, and just define $a_i$ and $b_i$ to the right hand side of these equations.}
+\textcolor{blue}{30-Equations 13 and 14. It seems easier to just drop the summation sign here, and just define $a_i$ and $b_i$ to the right hand side of these equations.}
 

### 2015-04-28 Manuscript/PLoSONEresponse/TEM_PLoS_response.tex
Fixed some comments (incl. figures)

@@
 This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
-We normalised this metric following Bogdanowicz \textit{et al.} \cite{Bogdanowicz2012}'s Normalised Tree Similarity (NTS) method.
+We normalised this distance following Bogdanowicz \textit{et al.} \cite{Bogdanowicz2012}'s Normalised Tree Similarity (NTS) method.
 This methods scales %not sure if scale here?
@@
 This method is a generalisation of the topological accuracy method \cite{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.
-In practice when the Normalised Robinson-Foulds distance between two trees is equal to one, the trees are identical; if the distance is equal to zero, the trees are no more different than expected by chance; finally if the distance is less than zero, the trees are more different than expected by chance.
+In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
 
@@
 Therefore this metric sensitive to the conservation of wildcard taxa.
-We normalised this metric in the same way as for the Robinson-Foulds distance resulting in the Normalised Triplets distance.
+We normalised this metric in the same way as for the Robinson-Foulds distance resulting in the Normalised Triplets metric.
 
@@
 \label{tree_comparisons}
-For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the ``best" tree and each ``missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \cite{Bogdanowicz2012} resulting in 125 Normalised Robison-Foulds distances and 125 Normalised Triplets distance for each tree inference method.
+For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the ``best" tree and each ``missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \cite{Bogdanowicz2012} resulting in 125 Normalised Robinson-Foulds metrics and 125 Normalised Triplets metric for each tree inference method.
 Also, to take into account the uncertainty of tree inference, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 trees from the posterior tree distribution of the Bayesian analysis for the ``best" trees, and then did the same for the 125 ``missing data" trees (resulting in 1000 ``best" trees and 125$\times$1000 ``missing data" trees).
@@
 
-We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) using the Bhattacharyya Coefficient \cite{Bhattacharyya}.
+We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) using the Bhattacharyya Coefficient \cite{Bhattacharyya}.
 The Bhattacharyya Coefficient is the probability of overlap between two distributions \cite{Bhattacharyya} (see ~\nameref{S2_Appendix2} for calculation details).
 Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
-To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%), i.e.
+To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%), i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc.
@@
 \caption{{\bf Bhattacharyya Coefficient calculation outline 1.}
- A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+ A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
 $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
@@
 To assess the effect of the different tree inference and uncertainty methods (i.e.
-Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the ``best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets distance) for each pairwise combination of tree inference and uncertainty methods, i.e.
+Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the ``best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of tree inference and uncertainty methods, i.e.
 Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc.
@@
 \caption{{\bf Bhattacharyya Coefficient calculation outline 2.}
- A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets distances) for any combination of missing data parameters (e.g.
+ A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
 $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
@@
 \section*{Results}
-As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets distance).
+As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets metric).
 However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
@@
  The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
 The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
@@
  The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
 The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
@@
 As the amount of missing data increases across all three parameters, our ability to recover the ``best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7 in ~\nameref{S3_Appendix3}).
-However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7 in ~\nameref{S3_Appendix3}).
+The Normalised Robinson-Foulds metric is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7 in ~\nameref{S3_Appendix3}).
+However, The Normalised Triplets metric is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7 in ~\nameref{S3_Appendix3}).
 
@@
  Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the ``best" tree distribution, calculated using the Bhattacharyya Coefficient.
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the ``best" tree distribution, calculated using the Bhattacharyya Coefficient.
 The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
@@
 
-Using both Normalised Robinson-Foulds and Normalised Triplets distances from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Fig.~\ref{Fig_Results-paircomp_within}).
-However, the Normalised Robinson-Foulds distance (median Bhattacharrya coefficient = 0.79; Fig.~\ref{Fig_Results-paircomp_within}A) displays more dissimilarities than the Normalised Triplets distance (median Bhattacharrya coefficient = 0.81; Fig.~\ref{Fig_Results-paircomp_within}B).
-Additionally, when using the Normalised Robinson-Foulds distance, once $M_{L}$ $\geq$ 50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Fig.~\ref{Fig_Results-paircomp_within}A).
+However, the Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; Fig.~\ref{Fig_Results-paircomp_within}A) displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; Fig.~\ref{Fig_Results-paircomp_within}B).
+Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Fig.~\ref{Fig_Results-paircomp_within}A).
 Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-paircomp_within}A).
 
-For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; Table ~\ref{Tab_Results-Difference_methods}).
-However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.
+For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds metric or Normalised Triplets metric respectively; Table ~\ref{Tab_Results-Difference_methods}).
+However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds metric (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.
 S2 in ~\nameref{S3_Appendix3}).
@@
 \caption{{\bf Bhattacharyya Coefficients of the pairwise method comparisons.}
-Each comparison which corresponds to the normalised distance between the ``best" tree and the ``missing data" trees using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).
+Each comparison which corresponds to the normalised metric between the ``best" tree and the ``missing data" trees using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).
 Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
@@
 However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
-decrease of Normalised Triplets distance) with increasing missing data in this parameter, clade conservation (i.e.
-Normalised Robinson-Foulds distance) is still relatively good (mode = 0.72) when the proportion of missing data is high ($M_{F}$ = 75\%).
+decrease of Normalised Triplets metric) with increasing missing data in this parameter, clade conservation (i.e.
+Normalised Robinson-Foulds metric) is still relatively good (mode = 0.72) when the proportion of missing data is high ($M_{F}$ = 75\%).
 
-The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds distance) but greater on the displacement of wildcard taxa (Normalised Triplets distance; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds metric) but greater on the displacement of wildcard taxa (Normalised Triplets metric; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
@@
 every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
-This does not decrease the Normalised Robinson-Foulds distance, because clades are conserved, but decreases the Normalised Triplets distance because the fossils act as wildcard taxa.
+This does not decrease the Normalised Robinson-Foulds metric, because clades are conserved, but decreases the Normalised Triplets metric because the fossils act as wildcard taxa.
 Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50).
@@
 The size of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
-Therefore, part of the decrease of the Normalised Robinson-Fould distance and the Normalised Triplets distance in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
 However, although these matrices are an order of magnitude smaller than some published matrices (e.g.
@@
 \subsection*{Combined effect of missing data parameters}
-As expected, when combining the missing data parameters, our ability to recover the ``best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have few missing data (i.e.
+As expected, when combining the missing data parameters, our ability to recover the ``best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a lot of missing data (i.e.

### 2015-04-29 Manuscript/PLoSONEresponse/response.tex
Fixed Mk vs Mkv story

@@
 
-% Priors comment??? Move stuff up from below
+The reviewer is (I believe) referring to their comment below (number 13).
+We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only and we are ignoring the branch length.
+The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and were a minor part of the MCMC sampler (2.17\% of the moves of the MCMC sampler for the three priors combined).
+Therefore we are confident that the topology is minimally affected by the priors, allowing us to explore the effect of missing morphological data on topology as well as reducing our computational time to a realistic 1.4 CPU centuries.
+Note that the use of a starting tree that is not a prior and just ``the ``true" tree's 226 topology as a starting tree'' (line 226) is not a prior and was used to avoid the chains to be stuck in local optima.
 
-The reviewer is (I believe) referring to their comment below (number 13).
-We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only.
-The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and were a minor part of the MCMC sampler (2.17\% of move probability% check what you call it in paper
- for the three priors).
+We clarified these points in the manuscript as follow:
 
-% Most of the effects will be branch length and we don't look at that
+Lines @@@112 to @@@114 in our revision:
 
-Therefore we are confident that the topology is minimally affected by the priors, allowing us to explore the effect of missing morphological data on topology as well as reducing our computational time to a realistic 1.5 CPU centuries.
-
-% Add something to text?
-
-%Our results did indeed biased us in favour of using a Bayesian approach to fix topology since our results clearly indicate that it is less likely to generate a wrong topology using a Bayesian approach (see Figure 4 and 5 and discussion line 472 to 496 in the first submission manuscript).
-% TG: or I'm not sure I understood that one.
-%-------------
-% add Mk = Mkv
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+Each Bayesian tree was estimated using two runs of four chains each for a maximum of 5$\times$$10^7$ generations.
+For each estimation, we used the ``true" tree's topology as a starting tree (with a starting value for each branch length of one).
+We also used two priors on the molecular part of the matrix: an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5, and a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)); and one prior on the morphological part of the matrix (exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5).
+We used these priors to speed up the Bayesian estimation process.
+These priors biased the way the Bayesian process calculated branch lengths by giving non-random starting points and boundaries for parameter estimation however, here we are focusing on the effect of missing data on tree topology and not branch lengths.
+Even using these priors, it took $~$ 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
+The detailed MrBayes parameters are available in the Supporting Information S1.
+\end{minipage}
 
@@
 
-Throughout our paper, we referred the M\textit{k} model as being the one described in Lewis (2001) (an equal rate Q-matrix with \textit{k} character states).
-The Mkv model is identical to the Mk model (as described in Lewis 2001) however, in practice, it assumes that all characters are variable and that therefore, one should correct for acquisition bias that can induce branch length inflation.
-This is generally by adding an algorithm in the phylogenetic software to add ``dummy'' morphological characters.
+Throughout our paper, we referred the M\textit{k} model \textit{sensu} Lewis 2001 (a Q-matrix with \textit{k} character states and an equal substitution rate $\mu$).
+The M\textit{kv} model is identical to the M\textit{k} model (as described in Lewis 2001) however, in practice, it assumes that all characters are variable and that therefore, one should correct for acquisition bias that can induce branch length inflation.
+This is generally done by using an algorithm in the phylogenetic software that adds ``dummy'' morphological characters.
 In some software (e.g.
-GARLI or BEAST), it is important to specify whether the morphological data should be corrected for acquisition bias or not (Mk or Mkv options).
-However, in RAxML and MrBayes, this bias is corrected automatically by addition of ``dummy'' characters.
+GARLI or BEAST), it is important to specify whether the morphological data should be corrected for acquisition bias or not (M\textit{k} or M\textit{kv} options).
+However, in RAxML and MrBayes, this bias is corrected automatically by addition of ``dummy'' characters and M\textit{k} just designate the evolutionary model used with $\mu$ $\geq$ 0 (e.g.
+Nylander \textit{et al.} 2004 - Systematic Biology).
 
 The reviewer is of course correct in that real data rarely contain invariable morphological characters, because these don't contribute to phylogenetic signal.
-Our simulated data, on the other hand, will occasionally contain invariable characters merely because of missing data, therefore we made the decision to describe the model for morphological characters evolution as the M\textit{k} model (an equal rate Q-matrix with \textit{k} character states where the rate an be equal to 0).
+Our simulated data, on the other hand, will occasionally contain invariable characters merely because of missing data, therefore we made the decision to describe the model for morphological characters evolution as the M\textit{k} evolutionary model ($\mu$ $\geq$ 0 rather than the M\textit{kv} model where $\mu$ $>$ 0).
 RAxML and MrBayes did however corrected for acquisition bias as implemented in their algorithm for treating morphological data.
 
-However, following the reviewer comment, we changed the designation of the model from M\textit{k} to M\textit{kv} to explicitly underline that we corrected for acquisition bias.
+However, following the reviewer comment, we changed the designation of the model from M\textit{k} to M\textit{kv} to implicitly underline that we corrected for acquisition bias.
+We also explicitly mentioned it in the revision in the following lines:
 
-%The Mkv model assumes that all the characters are variable, whereas the Mk model is less constrained - some characters can be invariable and others variable.
-Thus the Mkv model is actually just a special case of the Mk model.
+Lines @@@112 to @@@114 in our revision:
 
-%The reviewer is of course correct in that real data rarely contain invariable morphological characters, because these don't contribute to phylogenetic signal.
-Our simulated data, on the other hand, will occasionally contain invariable characters merely because of missing data %
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+We used RAxML because it automatically corrects for acquisition bias \cite{lewisa2001}, it also is heavily used in the literature for Maximum Likelihood tree inference (e.g.
+\cite{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014}) and is one of the fastest methods available \cite{Stamatakis01102008}.
+\end{minipage}
 
-%However, this is not an issue because we implement opur analyses in RAxML and MrBayes.
-These programs implement the Mk do not distinguish between both models since the Mkv model is just a particular case of the Mk model and are not distinguished either in the original publication that we cite in the manuscript (Lewis 2001).
-We've changed the mentions of the M\textit{k} model to M\textit{kv} model throughout our revision.
+Lines @@@112 to @@@114 in our revision:
 
-%Because we sampled the morphological characters rate from a gamma distribution with $\alpha$=0.5, some morphological characters had only a few state changes accross the matrix.
-When removing the data, there was a probability for these characters to be constant (e.g.
-if all the taxa in the matrix appart from one had a character state 0 for a given character and only one taxa had a state 1 for the same character, if that taxa had no data available for that character, then the character appeared to be constant in the matrix).
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+Note that MrBayes automatically corrects for acquisition bias in the morphological data partition (Nylander \textit{et al.} 2004; Ronquist \textit{et al.} 2012).
+\end{minipage}
 
-%Add note that we changed it along the text. RAxML and MrBayes are using default.
+% To Diego:
+%Here we use M\textit{k} model \textit{sensu} Lewis 2001 (designating a Q-matrix with \textit{k} character states and an equal substitution rate $\mu$).
+In real data, the rate $\mu$ is never equal to 0 and the model is therefore designated as the M\textit{kv} model (e.g.
+Wright \& Hillis 2014).
+However, in our simulations, this rate $\mu$ can sometimes be equal to 0 due to the missing data (i.e.
+it is possible that a character can display only one state when data is missing).
+The use of the M\textit{k} model with $\mu$ $\neq$ 0 (M\textit{kv} model) was really early pinpointed as including an acquisition bias because the use of non-invariant characters could inflate branch length (Lewis 2001, Nylander \textit{et al.} 2004).
+Therefore a correcting algorithm ($\neq$ model) was proposed to fix this problem by automatically generating a proportion of invariant characters for the tree inference process.
+In some software (e.g.
+GARLI or BEAST), the use of the correcting algorithm must be enforced by specifying whether the model is M\textit{kv} ($\mu$ $\neq$ 0, so one must correct for acquisition bias).
+However, the software we used (RAxML and MrBayes) do not allow options for deciding to correct for acquisition bias or not and do automatically use a correcting algorithm that adds ``dummy'' invariant characters (whether $\mu$ $>$ 0 (M\textit{kv}) or $\mu$ $\geq$ 0 (M\textit{k})).
+
+%Again, in our paper, we use the term M\textit{k} to describe the evolutionary model because some characters could be invariant ($\mu$ $>$ 0).
+However, we do understand that this is rather a semantic point (model $\neq$ algorithm) and we are confident that the description of the model and the software used in the paper will be enough for a careful reader to understand our protocol.
+We therefore changed the mentions of M\textit{k} to M\textit{kv} in the revision to make less careful readers understand that we were aware of acquisition bias when we chose our phylogenetic inference software.
+
 %-------------
@@
 See our response to comment C above.
-% NEED to fix!
-
-%We are not entirely sure if the reviewers refers to a sampling bias or to the problem link to constant characters described in Lewis (2001) (acquisition bias).
-If the reviewers are referring to the second (acquisition bias), both phylogenetic inference software we used (RAxML and MrBayes) are correcting for the this bias (-K MK  see Supplementary information 1 for the full list of commands).
-Also, the effect of the acquisition bias is to inflate branch length which is an aspect of the phylogeny we didn't explored in this study (focusing on topology).
-See response below to the comment ``\textit{The use of Mk rather than Mkv makes the study less relevant to real data analyses.}".
 

### 2015-04-29 Manuscript/PLoSONEresponse/TEM_PLoS_response.tex
Fixed Mk vs Mkv story

@@
 8.0.20 \cite{Stamatakis21012014}).
-We used RAxML because it is heavily used in the literature for Maximum Likelihood tree inference (e.g.
+We used RAxML because it automatically corrects for acquisition bias \cite{lewisa2001}.
+It also is heavily used in the literature for Maximum Likelihood tree inference (e.g.
 \cite{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014}) and is one of the fastest methods available \cite{Stamatakis01102008}.
@@
 For the morphological data, we used the M\textit{kv} model \cite{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{kv} + $\Gamma_4$).
+Note that MrBayes automatically corrects for acquisition bias in the morphological data partition \cite{Nylander01022004,Ronquist2012mrbayes}.
 We chose these models to be consistent with the parameters used to generate the ``complete" matrix.
@@
 Each Bayesian tree was estimated using two runs of four chains each for a maximum of 5$\times$$10^7$ generations.
-We used the following priors for each tree: (i) the ``true" tree’s topology as a starting tree (with a starting value for each branch length of one), (ii) an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions, and (iii) a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)).
+For each estimation, we used the ``true" tree's topology as a starting tree (with a starting value for each branch length of one).
+We also used two priors on the molecular part of the matrix: an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5, and a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)); and one prior on the morphological part of the matrix (exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5).
 We used these priors to speed up the Bayesian estimation process.
 These priors biased the way the Bayesian process calculated branch lengths by giving non-random starting points and boundaries for parameter estimation however, here we are focusing on the effect of missing data on tree topology and not branch lengths.
-Even using these priors, it took \texttildelow 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-The detailed MrBayes parameters are available in the ~\nameref{S1_Appendix1}.
+Even using these priors, it took $~$ 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
+The detailed MrBayes parameters are available in the Supporting Information S1.
 

### 2015-05-08 Manuscript/PLoSONEresponse/response.tex
isolated the part just for the editor (describing Mk vs Mkv)

@@
 
-% To Diego:
+% To the editor:
 %Here we use M\textit{k} model \textit{sensu} Lewis 2001 (designating a Q-matrix with \textit{k} character states and an equal substitution rate $\mu$).
@@
 
-This comment links back to the reviewer's slight misconception of the term ``statistics".
-ANOVA or multiple regression analysis might the appropriate analyses in some cases but we believe it isn't here.
+This comment links back to the reviewer's slight misunderstanding of the statistics we applied in this paper.
+The purpose of ANOVA or multiple regression is to compare differences in means across groups and test whether they are different from what one would expect according to a null model.
+In contrast, in our simulation approach, we essentially have access to a representative sample of the entire population distribution of tree topology comparisons.
+Therefore, we can simply ask whether these distributions are different across XXX [AJ - treatment groups or whatever] by comparing them directly.
+One standard such approach to comparing distributions directly is to use the BHAttacadsdfasdf coefficient.
+This number has direct intepretation in a probablistic sense and is simply the probability that the two distributions are the same.
+If the two distirubions overlap completely then they have BHAVYY coefficient and probability of being the same of 1.
+In the other extreme, if the two distributions are completely disjointed, then they have a BHATTYV coefficient and hence probabilty of being the same of 0.
+Thus, the BHYVYVY is a probability, comparable to, but not the same as, a p-value that would from frequentist approaches such as t-tests, anova or glms.
+Furthermore, as a direct comparison of two known distributions, it is a very simple and test that comes with far fewer assumptions than are made in general linear models and their analogues.
+For example, the direct comparison approach using BYGYVHVHV is capable of comparing a multi-modal distributions whereas the assumption of normality required in a t-test or glm would inviolate the robustness of the resultant p-value.
+There is an additional complitation to using t-tests or glms in simulation type analyeses arising from having access to very large smaple sizes that leads to p-values shrinking to zero and resulting in comparisons being almost certain to yield "statistically significant" differences in the mean.
+We argue that in this situation, it is better to compare the entire distribution using BHCYCY rather than simply the means.
+[AJ - you could add a simple figure of two normal distributions with 10^4 samples in each that have very similar means.
+Do a t-test which will be significant, but BHVYVYVYV will be nearly 1.
+
 In our analysis, we made the methodological choice to compare tree similarity distributions that are: not continuous distributions; bounded at 1; not always normal (both the distributions and the residuals) and sometimes multi-modal.

### 2015-05-08 Analysis/Comparing_distribution_example.R
Comparing distribution example

@@ -1,0 +1,64 @@
+#This is a script to illustrate the difference between using a t-test and the Bhattacharyya Coefficient.
+
+#####################
+#Sourcing the Bhattacharyya Coefficient from GitHub
+#####################
+
+#Installing RCurl package
+install.packages("RCurl")
+
+#Source GitHub function
+source_github <- function(u) {
+  # load package
+  require(RCurl)
+  # read script lines from website and evaluate
+  script <- getURL(u, ssl.verifypeer = FALSE)
+  eval(parse(text = script))
+}  
+
+#Sourcing the Bhattacharyya Coefficient function (bhatt.coeff)
+source_github("https://raw.githubusercontent.com/TGuillerme/Total_Evidence_Method-Missing_data/master/Functions/TreeComparisons/bhatt.coef.R")
+
+#####################
+#Generating the two distributions
+#####################
+
+#Setting seed for reproducibility
+set.seed(12345)
+
+#Generating the two normal distribution with a sd of 1 and a slightly different mean (1/20 of the sd) with 50000 data points each
+#This corresponds to the number of tree comparisons between two distributions for the Maximum Likelihood boostraps trees or the Bayesian posterior distribution trees
+distribution_1<-rnorm(50000, mean=0.05, sd=1)
+distribution_2<-rnorm(50000, mean=0, sd=1)
+
+#####################
+#Calculating the differences between both distribution
+#####################
+
+#Differences between the means (classic t-test)
+t_test_results<-t.test(distribution_1, distribution_2)
+#Isolating difference
+difference<-sqrt((t_test_results$estimate[[1]])^2-(t_test_results$estimate[[2]])^2)
+#Isolating the p-value
+p_value<-t_test_results$p.value
+
+#Calculating the Bhattacharyya Coefficient (the probability of the two distributions being the same) 
+bhatt_coeff<-bhatt.coeff(distribution_1, distribution_2)
+
+#####################
+#Plotting the results
+#####################
+
+#Graphical options
+op<-par(bty="l")
+
+#Plotting the first distribution
+plot(density(distribution_1), col="red", main="", xlab="Values")
+#Plotting the second distribution
+lines(density(distribution_2), col="blue")
+
+#Adding the calculated differences
+text(3, 0.4, "Difference between means (t-test):", cex=0.8)
[…]

### 2015-05-08 Analysis/Comparing_distribution_example.R
cleaned sourcing

@@ -1,23 +1,7 @@
 #This is a script to illustrate the difference between using a t-test and the Bhattacharyya Coefficient.
 
-#####################
-#Sourcing the Bhattacharyya Coefficient from GitHub
-#####################
-
-#Installing RCurl package
-install.packages("RCurl")
-
-#Source GitHub function
-source_github <- function(u) {
-  # load package
-  require(RCurl)
-  # read script lines from website and evaluate
-  script <- getURL(u, ssl.verifypeer = FALSE)
-  eval(parse(text = script))
-}  
-
 #Sourcing the Bhattacharyya Coefficient function (bhatt.coeff)
-source_github("https://raw.githubusercontent.com/TGuillerme/Total_Evidence_Method-Missing_data/master/Functions/TreeComparisons/bhatt.coef.R")
+source("http://tguillerme.github.io/R/bhatt.coef.R")
 
 #####################
 #Generating the two distributions

### 2015-05-08 Manuscript/PLoSONEresponse/response.tex
Updated response (stats part - incl fig)

--- 2015-04-29 Manuscript/PLoSONEresponse/response.tex
@@
 \usepackage[osf]{mathpazo}
+\usepackage{graphicx}
 \usepackage{amsmath,amssymb}
@@
 
-% Talk to Andrew
+We feel that the reviewer has slightly misunderstood the statistics we applied in this paper.
+The purpose of ANOVAs or multiple regression (see suggestion from comment 20 below) is to compare differences in means across groups and test whether they are different from what one would expect according to a null model.
+In contrast, in our simulation approach, we essentially have access to a representative sample of the entire population distribution of tree topology comparisons.
+Therefore, we can simply ask whether these distributions are different when our missing data parameters varies by comparing them directly.
+One standard such approach to comparing distributions directly is to use the Bhattacharyya Coefficient.
+This number has direct interpretation in a probablistic sense and is simply the probability that the two distributions are the same.
+If the two distributions overlap completely then they have Bhattacharyya Coefficient and probability of being the same of 1.
+In the other extreme, if the two distributions are completely disjointed, then they have a Bhattacharyya Coefficient and hence probability of being the same of 0 (line 306 and 307 in the previously submitted manuscript).
+Thus, the Bhattacharyya Coefficient is a probability, comparable to, but not the same as, a p-value that could be calculated using frequentist approaches such as t-tests, ANOVAs or GLMs.
+
+Furthermore, as a direct comparison of two known distributions, using the the Bhattacharyya Coefficient comes with far fewer statistical assumptions than are required in ANOVAs, GLMs and their analogues.
+For example, the direct comparison approach using Bhattacharyya Coefficient is capable of comparing a multi-modal distributions whereas the assumption of normality required in a ANOVAs or GLMs thus using them would inviolate the robustness of the resultant p-value.
+There is an additional complication to using ANOVAs or GLMs in simulation type analyses arising from having access to very large sample sizes (e.g.
+50000 data points for the Maximum Likelihood bootstraps trees and the Bayesian posterior distribution trees) that leads to p-values shrinking to zero and resulting in comparisons being almost certain to yield "statistically significant" differences in the mean (see the attached figure ; code for reproducibility is available \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Comparing_distribution_example.R}{here} ; https://goo.gl/4djNUf).
+We argue that in this situation, it is better to compare the entire distribution using Bhattacharyya Coefficient rather than simply the means.
+Additionally, because we are comparing topologies in this paper, we made the methodological choice to use the modal topology (i.e.
+the most represented topology) instead of the mean topology (i.e.
+a hypothetical ``middle'' topology).
+
+Finally, we would like to underline that we report our 50\% and 95\% confidence intervals in Figures 4 and 5 when showing the effect of the missing data parameters (calculation detailed lines 295-295) as well as the $1^st$ and $3^rd$ quartiles when comparing the methods in Table 1.
+
+\includegraphics[scale=0.8, keepaspectratio=true]{Comparing_distributions.pdf} \\
+\textit{Different results between the Bhattacharyya Coefficient showing that the distributions are really similar (0.999 overlap probability) and a frequentist approach (simple t-test) underlining a significant difference (p-value=0).}
 
@@
 
+%Underline the difference between evolutionary models (M\textit{k}, GTR, JC, HKY) and ``models'' that are methods (usually computer algorithms) used to build trees (such as M\textit{kv})
+
 Throughout our paper, we referred the M\textit{k} model \textit{sensu} Lewis 2001 (a Q-matrix with \textit{k} character states and an equal substitution rate $\mu$).
@@
 
-\textcolor{blue}{D-Many of the equations are incorrect.}
+\textcolor{blue}{D-Many of the equations are incorrect.} %only two in the main body.
 
-See our detailed response below.
+We thank the reviewer to underline the two errors in the equations in the main body at lines 258 and 275, see our detailed response to the comments number 15 and 17 below.
+Also we recognise that the equations in our ``Supporting Information S2: Tree Comparisons'' are confusing and are actually unnecessary for the understanding of this paper.
+We therefore removed the details about the Robinson-Foulds and the Triplets distances from the ``Supporting Information S2: Tree Comparisons'' section.
 
@@
 
-%%%%%%%%%%%%%%%%%
-%CHEK in Soltis!
-%%%%%%%%%%%%%%%%%
-
 %-------------
@@
 
-This comment links back to the reviewer's slight misconception of the term ``statistics".
-ANOVA or multiple regression analysis might the appropriate analyses in some cases but we believe it isn't here.
-In our analysis, we made the methodological choice to compare tree similarity distributions that are: not continuous distributions; bounded at 1; not always normal (both the distributions and the residuals) and sometimes multi-modal.
-Therefore, we choose not to use classical ``statistics" methods such as ANOVA or multiple regression because these methods could simply not be applied to our data set.
-%Not comparing means, non normal distributions. And non-parametric version (e.g. Kruskal Wallis) not approriate either.
-
-% Talk to Andrew
+See our detailed response to comment A above.
 
@@
 
-We fixed this typo.
+We thanks the reviewer for pointing out this typo and fixed it.
 
-\textcolor{blue}{line 420: contra the statement here, a consensus collapsing to a polytomy will affect the RF distance (could go up or down, but it will change).}
+\textcolor{blue}{22-line 420: contra the statement here, a consensus collapsing to a polytomy will affect the RF distance (could go up or down, but it will change).}
 
@@
 \textcolor{blue}{Supplement 2:}
-\textcolor{blue}{22-more confusion about the min RF here.
+
+\textcolor{blue}{23-more confusion about the min RF here.
 Also note that if you are counting the entire leaf set as a clade (e.g the star tree having N=1), then the number of clades in rooted binary tree is n-1 (not n-2).
@@
 
-\textcolor{blue}{23-The definition of the scaled RF implies that it is never positive (sinc the numerator cannot be positive and the denominator is positive).}
+\textcolor{blue}{24-The definition of the scaled RF implies that it is never positive (sinc the numerator cannot be positive and the denominator is positive).}
 
@@
 
-\textcolor{blue}{24-The decision to use Yule trees as the base line for the mean distance is not explained, and seems odd given that the true tree are not Yule trees.
+\textcolor{blue}{25-The decision to use Yule trees as the base line for the mean distance is not explained, and seems odd given that the true tree are not Yule trees.
 I'm not sure that it makes much difference, but it is odd.}
@@
 
-\textcolor{blue}{25-The NTS can only -infinity for distances that have the property that the mean distance between random Yule trees is 0.
+\textcolor{blue}{26-The NTS can only -infinity for distances that have the property that the mean distance between random Yule trees is 0.
 This is not true of either distance metric here.
@@
 
-\textcolor{blue}{26-Note that JavaScript is a language, so it a bit confusing to refer to a Java program as a ``Java script."}
+\textcolor{blue}{27-Note that JavaScript is a language, so it a bit confusing to refer to a Java program as a ``Java script."}
 
@@
 
-\textcolor{blue}{27-Equation (6) is not correct.
+\textcolor{blue}{28-Equation (6) is not correct.
 You shouldn't set the summation sign as the left hand side of an equation because the notation has a standard definition.
@@
 
-\textcolor{blue}{28-Equation (7) is true for a three-leaf tree with a polytomy occurring with probability = 1/4 and the other 3 trees being equiprobable.
+\textcolor{blue}{29-Equation (7) is true for a three-leaf tree with a polytomy occurring with probability = 1/4 and the other 3 trees being equiprobable.
 I don't think that is is true for any other weighting of trees for other numbers of taxa.
@@
 
-\textcolor{blue}{29-I don't think that equations 8, 9, or 10 are correct for any case other than the 3 taxon case (though I do note that the authors are using n choose 3 here).}
+\textcolor{blue}{30-I don't think that equations 8, 9, or 10 are correct for any case other than the 3 taxon case (though I do note that the authors are using n choose 3 here).}
 
@@
 
-\textcolor{blue}{30-Equations 13 and 14. It seems easier to just drop the summation sign here, and just define $a_i$ and $b_i$ to the right hand side of these equations.}
+\textcolor{blue}{31-Equations 13 and 14. It seems easier to just drop the summation sign here, and just define $a_i$ and $b_i$ to the right hand side of these equations.}
 

### 2015-05-10 Manuscript/PLoSONEresponse/TEM_PLoS_revision.tex
fixed revisions (comment 22)

--- 2015-04-29 Manuscript/PLoSONEresponse/TEM_PLoS_response.tex
@@
 every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
-This does not decrease the Normalised Robinson-Foulds metric, because clades are conserved, but decreases the Normalised Triplets metric because the fossils act as wildcard taxa.
+In this case, the Normalised Robinson-Foulds metric will decreases when the fossil are present at a lower taxonomic level (i.e.
+separated with more nodes from the root) but affects the clade conservation less at higher taxonomic level (i.e.
+separated with less nodes from the root).
+On the opposite though, because the fossil in a high level clade has many chances to branch on different nodes within that clade, it will be more likely to act as a wildcard taxa and decrease the Normalised Triplets metric.
+Therefore, the $M_{F}$ parameter is likely to affect the Normalised Robinson-Foulds metric less than the Normalised Triplets metric for the Bayesian consensus trees.
+%This does not decrease the Normalised Robinson-Foulds metric, because clades are conserved, but decreases the Normalised Triplets metric because the fossils act as wildcard taxa.
 Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50).

### 2015-05-10 Manuscript/PLoSONEresponse/response.tex
All comment dealt with + cleaning.

@@
 \hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-We used RAxML because it automatically corrects for acquisition bias \cite{lewisa2001}, it also is heavily used in the literature for Maximum Likelihood tree inference (e.g.
-\cite{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014}) and is one of the fastest methods available \cite{Stamatakis01102008}.
+We used RAxML because it automatically corrects for acquisition bias (Lewis 2001), it also is heavily used in the literature for Maximum Likelihood tree inference (e.g.
+[...]) and is one of the fastest methods available (Stamatakis 2008).
 \end{minipage}
@@
 We implemented this step of keeping trees with a minimum total median bootstrap to make the topological comparisons more robust.
-If the no-missing data tree has low node support, it is likely that the changes in topology for the missing data trees are not linked to our focal question (the effect of removing characters, living taxa or fossils) but just to the fact that low supported nodes can be just random.
+If the no-missing data tree has low node support, it is likely that the changes in topology for the missing data trees are not linked to our focal question (the effect of removing characters, living taxa or fossils) but just to the low supported nodes.
+In fact, because we repeat the tree inference 125 times ($5^3$ parameters combinations), it is likely to have 125 different pseudo-random solution for the nodes with really low phylogentic support and that, therefore, the observed changes in topologies are just linked to the pseudo-random replication.
+Note however that we did not exculde this scenario from our protocol the trees selected with a median bootstrap $>$ 50 could well contain some unresolved nodes with a low bootstrap ($<$ 50).
 We've added the following justification in our revision:
-
-% NC: expand on this a bit
 
@@
 
-We changed the following sentence: % correct
-
-Lines @@@112 to @@@114 in our revision:
+We revised this statement lines @@@112 to @@@114 in our revision:
 
@@
 
-%%%%%%%%%%%%%%%%%
-%CHECK in Soltis!
-%%%%%%%%%%%%%%%%%
-
 %-------------
@@
 We changed the mentions of ``Normalised Robinson-Foulds distance'' and ``Normalised Triplets distance'' to ``Normalised Robinson-Foulds metric'' and ``Normalised Triplets metric'' throughout the our revision.
-Additionally, we clarified the scaling method as follows:
-
-Lines @@@112 to @@@114 in our revision:
+Additionally, we clarified the scaling method as follows (lines @@@112 to @@@114 in our revision):
 
@@
 
-We changed the following sentence: % fix
-
-Lines @@@112 to @@@114 in our revision:
+We corrected this error lines @@@112 to @@@114 in our revision:
 
@@
 
-%To
+We revised this statement at lines @@@ to @@@ in our revision
 
-This does not decrease the 420
-Normalised Robinson-Foulds distance, because clades are conserved, but decreases the 421
-Normalised Triplets distance because the fossils act as wildcard taxa. 422
-
-%Change to "affect less" or something like that.
+\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
+In this case, the Normalised Robinson-Foulds metric will decreases when the fossil are present at a lower taxonomic level (i.e.
+separated with more nodes from the root) but affects the clade conservation less at higher taxonomic level (i.e.
+separated with less nodes from the root).
+On the opposite though, because the fossil in a high level clade has many chances to branch on different nodes within that clade, it will be more likely to act as a wildcard taxa and decrease the Normalised Triplets metric.
+Therefore, the $M_{F}$ parameter is likely to affect the Normalised Robinson-Foulds metric less than the Normalised Triplets metric for the Bayesian consensus trees.
+\end{minipage}
 
 %-------------
-
-%So for the part about the tree comparisons (RF and Triplets, I think we can just remove the whole thing since it's not especially useful...)
-
-We recognise that this section in our supp mat may be confusing and is actually unnecessary for the understanding the paper. Therefore we have removed this section. % fix!!!
-
 \textcolor{blue}{Supplement 2:}
@@
 
-%check
-
+See our response to the comment D above.
 %-------------
@@
 
-%not in this paper
-
+See our response to the comment D above.
 %-------------
@@
 
-We cited the reference (Bogdanowicz 2012) of the method.
-%explain in a bit more detail
-
+We followed the protocol in the cited reference (Bogdanowicz 2012) where the normalisation of the metrics is done on the mean value of that metric when comparing to random trees.
+Using random Yule trees is the best way to generate purely random trees since they are not based on any evolutionary scenario and takes only one parameter which is the number of tips (e.g.
+as opposed to the Birth-Death tree that assumes more realistic evolutionary scenario and takes multiple parameters).
 %-------------
@@
 As clarified in our revision (lines @@@), the NTS is a generalisation of the topological accuracy method (Price \textit{et al.} 2010 PLoS ONE) and is therefore not specialised for the RF or the Triples distance.
-
 %-------------
@@
 We fixed this typo.
-
 %-------------

### 2015-05-11 Manuscript/PLoSONEresponse/TEM_PLoS_revision.tex
Minor fixed forgotten from the response

@@
 To avoid biasing our simulations towards either living or fossil taxa and to make each simulation comparable, we implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa.
+The fossil taxa were considered as unique tips at the end of extinct lineages.
 We then added an outgroup to the tree, using the mean branch length of the tree to separate the outgroup from the rest of the taxa, and with the branch length leading to the outgroup set as the sum of the mean branch length and the longest root-to-tip length of the tree.
@@
 0.1-14 \cite{chen2011}.
-The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \cite{ranbaut1997seqgen} and using the HKY model \cite{HKY85} with random base frequencies and transition/transversion rate of two \cite{douadycomparison2003}.
+The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \cite{ranbaut1997seqgen} and using the HKY model \cite{HKY85} with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1) and transition/transversion rate of two \cite{douadycomparison2003}.
 The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \cite{yangamong-site1996}.
@@
 This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
-We normalised this distance following Bogdanowicz \textit{et al.} \cite{Bogdanowicz2012}'s Normalised Tree Similarity (NTS) method.
-This methods scales %not sure if scale here?
-any tree comparison metric using the mean distance between 1000 random trees (see ~\nameref{S2_Appendix2} for the calculation details).
+We normalised this metric following Bogdanowicz's Normalised Tree Similarity (NTS) method \cite{Bogdanowicz2012}.
+This methods scales any tree comparison metric using the mean distance between 1000 random trees (see ~\nameref{S2_Appendix2} for the calculation details).
 This method is a generalisation of the topological accuracy method \cite{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.
@@
 The Bhattacharyya Coefficient is the probability of overlap between two distributions \cite{Bhattacharyya} (see ~\nameref{S2_Appendix2} for calculation details).
-Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
+Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
 To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%), i.e.

### 2015-05-11 Manuscript/PLoSONEresponse/response.tex
line number and matching with the revision

@@
 
-The reviewer is (I believe) referring to their comment below (number 13).
+The reviewer is (we believe) referring to their comment below (number 13).
 We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only and we are ignoring the branch length.
@@
 
-Lines @@@112 to @@@114 in our revision:
+Lines @234 to @247 in our revision:
 
@@
 
-Lines @@@112 to @@@114 in our revision:
+Lines @206 to @209 in our revision:
 
@@
 
-Lines @@@112 to @@@114 in our revision:
+Lines @231 to @232 in our revision:
 
@@
 
-Yes. We have clarified as follows:
-
-Lines @@@112 to @@@114 in our revision:
+Yes. We have clarified as follows in our revision (lines @118 to @119):
 
 \hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-[We] select only trees with 25 living and 25 fossil taxa. The fossil taxa were considered as unique tips at the end of extinct lineages. % correct in ms
+[We] select only trees with 25 living and 25 fossil taxa. The fossil taxa were considered as unique tips at the end of extinct lineages.
 \end{minipage}
@@
 
-Lines @@@112 to @@@114 in our revision:
+Lines @126 to @130 in our revision:
 
@@
 
-We clarified this point ($M_C$ comes second) in the new version of the manuscript (see below).
-Also, we clarified the potential confusion between the amount of data removed in one parameter ($M_F$ = 10\% for example) and the amount of missing data in the morphological matrix (5\%).
-
-Lines @@@112 to @@@114 in our revision:
+We clarified this point ($M_C$ comes second) in the new version of the manuscript (lines @173 to @183).
+We also, clarified the potential confusion between the amount of data removed in one parameter ($M_F$ = 10\% for example) and the amount of missing data in the morphological matrix (5\%) in the same revised paragraph (lines @173 to @183).
 
@@
 Note however that we did not exculde this scenario from our protocol the trees selected with a median bootstrap $>$ 50 could well contain some unresolved nodes with a low bootstrap ($<$ 50).
-We've added the following justification in our revision:
-
-Lines @@@112 to @@@114 in our revision:
+We've added the following justification in our revision (lines @215 to @220):
 
@@
 
-We revised this statement lines @@@112 to @@@114 in our revision:
+We revised this statement lines @267 to @269 in our revision:
 
@@
 
-We thanks the reviewer for pointing at this error and we changed the following sentence:
-
-Lines @@@112 to @@@114 in our revision:
+We thanks the reviewer for pointing at this error and we changed the following sentence (lines @271 to @273 in our revision):
 
@@
 We changed the mentions of ``Normalised Robinson-Foulds distance'' and ``Normalised Triplets distance'' to ``Normalised Robinson-Foulds metric'' and ``Normalised Triplets metric'' throughout the our revision.
-Additionally, we clarified the scaling method as follows (lines @@@112 to @@@114 in our revision):
+Additionally, we clarified the scaling method as follows (lines @276 to @284 in our revision):
 
@@
 
-We corrected this error lines @@@112 to @@@114 in our revision:
+We corrected this error lines @288 to @290 in our revision:
 
@@
 
-We clarified this point in the text as well as in the figures 2 and 3 captions (``equivalent" $->$ ``similar"):
-
-Lines 308 to 309 in the previously submitted manuscript:
+We clarified this point in the text as well as in the figures 2 and 3 captions (``equivalent" $->$ ``similar" ; lines @322 to @324 in our revision):
 
 \hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-Note that this is equivalent to performing a two-sided t-test.
-\end{minipage}
-
-Lines @@@112 to @@@114 in our revision:
-
-\hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-Note that this is similar to performing a two-sided t-test for comparing two distribution.
+Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
 \end{minipage}
@@
 
-We revised this statement at lines @@@ to @@@ in our revision
+We revised this statement at lines @434 to @442 in our revision:
 
@@
 
-As clarified in our revision (lines @@@), the NTS is a generalisation of the topological accuracy method (Price \textit{et al.} 2010 PLoS ONE) and is therefore not specialised for the RF or the Triples distance.
+As clarified in our revision (lines @276 to @284), the NTS is a generalisation of the topological accuracy method (Price \textit{et al.} 2010 PLoS ONE) and is therefore not specialised for the RF or the Triples distance.
 %-------------

### 2015-05-12 Manuscript/PLoSONEresponse/response.tex

@@
 \documentclass[11pt]{letter}
-\usepackage[a4paper,left=2.5cm, right=2.5cm, top=1cm, bottom=1cm]{geometry}
+\usepackage[a4paper,left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm]{geometry}
 \usepackage{color}
@@
 
+\opening{Dear Dr Fontaneto,}
+
 RE: PLOS ONE Decision: PONE-D-15-05911 - [EMID:ffa50492e76c2059]
-
-Dear Dr Fontaneto
 
@@
 We include below details of how we have dealt with the reviewer's comments.
-Your comments and the reviewers comments are in blue and our response is in black.
+Your comments and the reviewer's comments are in blue and our responses are in black.
 
-% NC: Just send to editor
-%We would however, like to draw your attention to a couple of points.
+We would however, first like to draw your attention to a couple of points.
 Firstly, while we understand the difficulties of obtaining reviewers for papers, we were very disappointed to have the previous decision made based on just one review.
-As pointed out in your email, this reviewer responded immediately.
-We suspect they rushed their assessment because they make several comments about our paper that are completely incorrect, and would have been noticed in a thorough reading of the manuscript.
-We suspect that had a second reviewer been found, the decision might have been different.
-
-%Secondly, the reviewer clearly misunderstood the meaning of the term ``statistics''.
-Statistics does not mean that p-values need to be used.
-In this case we could have applied t-tests, and given the reviewer the p-values they request, but this would only tell us that the means of the distributions are significantly different, not that the distributions differ.
-We chose instead to use the Bhattacharya coefficient, which admittedly is an unfamiliar statistic to many but we feel we made a lot of effort to explain how this metric works in the text (and this method was suggested to us by a professional statistician!).
-The reviewer however appears to have refused to engage with this and instead has fixated on the lack of p-values.
-% NC: Some kind of summing up here
-
-%We are slightly concerned by the reviewer's misconception of the term ``statistics" and suspicions (``I suspect that if they authors take some time to do statistics (calculate some P-values, confidence intervals, or take a Bayesian perspective") since: (1) t-tests and p-values are useful in some cases but have several limitations especially when one compares distributions that are not normal (in this study, we compared the modal value of topological differences, making any test comparing the mean value useless) and can be biased towards small values (i.e.
-significant ones) when the sample sizes are big; (2) concerning the confidence intervals, we actually did systematically calculate them; ans (3) we are not convinced by a Bayesian approach for comparing distributions.
+Secondly, we would like to clarify that we did use statistics (see our response to comment A below), we just used statistics that do not produce p-values.
+We therefore feel that the reviewer has slightly misunderstood the statistics we applied in the manuscript and respond to this in detail below.
 
@@
 
-\textcolor{blue}{A-Unfortunately the data analysis performed is simply the calculation of a number of summary statistics about the overlap between the estimates returned for different simulation conditions.
+\textcolor{blue}{A - Unfortunately the data analysis performed is simply the calculation of a number of summary statistics about the overlap between the estimates returned for different simulation conditions.
 I suspect that if the authors take some time to do statistics (calculate some P-values, confidence intervals, or take a Bayesian perspective), they'll be able to make a statistical case for most of the points that they are trying to make.
@@
 
-We feel that the reviewer has slightly misunderstood the statistics we applied in this paper.
+We feel that the reviewer has slightly misunderstood the statistics we applied in this manuscript.
 The purpose of ANOVAs or multiple regression (see suggestion from comment 20 below) is to compare differences in means across groups and test whether they are different from what one would expect according to a null model.
@@
 If the two distributions overlap completely then they have Bhattacharyya Coefficient and probability of being the same of 1.
-In the other extreme, if the two distributions are completely disjointed, then they have a Bhattacharyya Coefficient and hence probability of being the same of 0 (line 306 and 307 in the previously submitted manuscript).
+At the other extreme, if the two distributions are completely disjointed, then they have a Bhattacharyya Coefficient and hence probability of being the same of 0 (line 306 and 307 in the previously submitted manuscript).
 Thus, the Bhattacharyya Coefficient is a probability, comparable to, but not the same as, a p-value that could be calculated using frequentist approaches such as t-tests, ANOVAs or GLMs.
@@
 There is an additional complication to using ANOVAs or GLMs in simulation type analyses arising from having access to very large sample sizes (e.g.
-50000 data points for the Maximum Likelihood bootstraps trees and the Bayesian posterior distribution trees) that leads to p-values shrinking to zero and resulting in comparisons being almost certain to yield "statistically significant" differences in the mean (see the attached figure ; code for reproducibility is available \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Comparing_distribution_example.R}{here} ; https://goo.gl/4djNUf).
-We argue that in this situation, it is better to compare the entire distribution using Bhattacharyya Coefficient rather than simply the means.
-Additionally, because we are comparing topologies in this paper, we made the methodological choice to use the modal topology (i.e.
+50000 data points for the Maximum Likelihood bootstraps trees and the Bayesian posterior distribution trees) that leads to p-values shrinking to zero and resulting in comparisons being almost certain to yield ``statistically significant" differences in the mean (see the figure below; code for reproducing it is available from \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Comparing_distribution_example.R}{https://goo.gl/4djNUf}).
+We argue that in this situation, it is better to compare the entire distribution using the Bhattacharyya Coefficient rather than simply the means.
+Additionally, because we are comparing topologies in this manuscript, we made the methodological choice to use the modal topology (i.e.
 the most represented topology) instead of the mean topology (i.e.
@@
 
-Finally, we would like to underline that we report our 50\% and 95\% confidence intervals in Figures 4 and 5 when showing the effect of the missing data parameters (calculation detailed lines 295-295) as well as the $1^st$ and $3^rd$ quartiles when comparing the methods in Table 1.
+Finally, we would like to underline that we report our 50\% and 95\% confidence intervals in Figures 4 and 5 when showing the effect of the missing data parameters (calculation detailed lines 295-295) as well as the 1st and 3rd quartiles when comparing the methods in Table 1.
 
 \includegraphics[scale=0.8, keepaspectratio=true]{Comparing_distributions.pdf} \\
-\textit{Different results between the Bhattacharyya Coefficient showing that the distributions are really similar (0.999 overlap probability) and a frequentist approach (simple t-test) underlining a significant difference (p-value=0).}
-
-%We feel that the reviewer has slightly misunderstood the term ``statistics'' in this case.
-Firstly, p-values are useful in some cases but have several limitations especially when one compares distributions that are not normal (in this study, we compared the modal value of topological differences, making any test comparing the mean value useless) and can be biased towards small values (i.e.
-significant ones) when the sample sizes are big; (2) concerning the confidence intervals, we actually did systematically calculate them; ans (3) we are not convinced by a Bayesian approach for comparing distributions.
+\textit{The figure shows two extremely similar distributions.
+However, although the Bhattacharyya Coefficient shows that the distributions are really similar (0.999 overlap probability), a frequentist approach (simple t-test) suggests that there is a significant difference between the means (p-value \textless 0.0000001).
+This highlights the reason we chose to use the Bhattacharyya Coefficient in our manuscript.}
 
@@
 
-\textcolor{blue}{B-The bias in the analysis in favor of Bayesian methods makes the recommendation to use Bayesian methods less satisfying.}
+\textcolor{blue}{B - The bias in the analysis in favor of Bayesian methods makes the recommendation to use Bayesian methods less satisfying.}
 
-The reviewer is (we believe) referring to their comment below (number 13).
-We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only and we are ignoring the branch length.
+The reviewer is (we believe) referring to comment 13 below.
+We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only, and we are ignoring branch lengths.
 The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and were a minor part of the MCMC sampler (2.17\% of the moves of the MCMC sampler for the three priors combined).
 Therefore we are confident that the topology is minimally affected by the priors, allowing us to explore the effect of missing morphological data on topology as well as reducing our computational time to a realistic 1.4 CPU centuries.
-Note that the use of a starting tree that is not a prior and just ``the ``true" tree's 226 topology as a starting tree'' (line 226) is not a prior and was used to avoid the chains to be stuck in local optima.
+Note that the use of a starting tree that is not a prior and just ``the ``true" tree's topology'' (line 226) was used to avoid the chains getting stuck in local optima.
 
-We clarified these points in the manuscript as follow:
+We clarified these points in the manuscript as follows:
 
@@
 
-\textcolor{blue}{C-The use of Mk rather than Mkv makes the study less relevant to real data analyses.}
+\textcolor{blue}{C - The use of Mk rather than Mkv makes the study less relevant to real data analyses.}
 
-%Underline the difference between evolutionary models (M\textit{k}, GTR, JC, HKY) and ``models'' that are methods (usually computer algorithms) used to build trees (such as M\textit{kv})
-
-Throughout our paper, we referred the M\textit{k} model \textit{sensu} Lewis 2001 (a Q-matrix with \textit{k} character states and an equal substitution rate $\mu$).
+Throughout our manuscript, we referred the M\textit{k} model \textit{sensu} Lewis 2001 (a Q-matrix with \textit{k} character states and an equal substitution rate $\mu$).
 The M\textit{kv} model is identical to the M\textit{k} model (as described in Lewis 2001) however, in practice, it assumes that all characters are variable and that therefore, one should correct for acquisition bias that can induce branch length inflation.
@@
 GARLI or BEAST), it is important to specify whether the morphological data should be corrected for acquisition bias or not (M\textit{k} or M\textit{kv} options).
-However, in RAxML and MrBayes, this bias is corrected automatically by addition of ``dummy'' characters and M\textit{k} just designate the evolutionary model used with $\mu$ $\geq$ 0 (e.g.
+However, in RAxML and MrBayes, this bias is corrected automatically by addition of ``dummy'' characters and M\textit{k} just designates the evolutionary model used with $\mu$ $\geq$ 0 (e.g.
 Nylander \textit{et al.} 2004 - Systematic Biology).
@@
 The reviewer is of course correct in that real data rarely contain invariable morphological characters, because these don't contribute to phylogenetic signal.
-Our simulated data, on the other hand, will occasionally contain invariable characters merely because of missing data, therefore we made the decision to describe the model for morphological characters evolution as the M\textit{k} evolutionary model ($\mu$ $\geq$ 0 rather than the M\textit{kv} model where $\mu$ $>$ 0).
-RAxML and MrBayes did however corrected for acquisition bias as implemented in their algorithm for treating morphological data.
+Our simulated data, on the other hand, will occasionally contain invariable characters merely because of missing data, therefore we made the decision to describe the model for morphological character evolution as the M\textit{k} evolutionary model ($\mu$ $\geq$ 0 rather than the M\textit{kv} model where $\mu$ $>$ 0).
+RAxML and MrBayes do however correct for acquisition bias as implemented in their algorithm for dealing with morphological data.
 
-However, following the reviewer comment, we changed the designation of the model from M\textit{k} to M\textit{kv} to implicitly underline that we corrected for acquisition bias.
+However, following the reviewer's comment, we changed the designation of the model from M\textit{k} to M\textit{kv} in the text, to underline that we corrected for acquisition bias.
 We also explicitly mentioned it in the revision in the following lines:
@@
 
-% To the editor:
-%Here we use M\textit{k} model \textit{sensu} Lewis 2001 (designating a Q-matrix with \textit{k} character states and an equal substitution rate $\mu$).
-In real data, the rate $\mu$ is never equal to 0 and the model is therefore designated as the M\textit{kv} model (e.g.
-Wright \& Hillis 2014).
-However, in our simulations, this rate $\mu$ can sometimes be equal to 0 due to the missing data (i.e.
-it is possible that a character can display only one state when data is missing).
-The use of the M\textit{k} model with $\mu$ $\neq$ 0 (M\textit{kv} model) was really early pinpointed as including an acquisition bias because the use of non-invariant characters could inflate branch length (Lewis 2001, Nylander \textit{et al.} 2004).
-Therefore a correcting algorithm ($\neq$ model) was proposed to fix this problem by automatically generating a proportion of invariant characters for the tree inference process.
-In some software (e.g.
-GARLI or BEAST), the use of the correcting algorithm must be enforced by specifying whether the model is M\textit{kv} ($\mu$ $\neq$ 0, so one must correct for acquisition bias).
-However, the software we used (RAxML and MrBayes) do not allow options for deciding to correct for acquisition bias or not and do automatically use a correcting algorithm that adds ``dummy'' invariant characters (whether $\mu$ $>$ 0 (M\textit{kv}) or $\mu$ $\geq$ 0 (M\textit{k})).
+%-------------
 
-%Again, in our paper, we use the term M\textit{k} to describe the evolutionary model because some characters could be invariant ($\mu$ $>$ 0).
-However, we do understand that this is rather a semantic point (model $\neq$ algorithm) and we are confident that the description of the model and the software used in the paper will be enough for a careful reader to understand our protocol.
-We therefore changed the mentions of M\textit{k} to M\textit{kv} in the revision to make less careful readers understand that we were aware of acquisition bias when we chose our phylogenetic inference software.
+\textcolor{blue}{D - Many of the equations are incorrect.} %only two in the main body.
+
+We thank the reviewer for pointing out the two errors in the equations in the main body of the text at lines 258 and 275 (see our detailed response to comments 15 and 17 below).
+Also we recognise that the equations in our ``Supporting Information S2: Tree Comparisons'' are confusing and are actually unnecessary for the understanding of this manuscript.
+We therefore removed the details of the Robinson-Foulds and Triplets distances from the ``Supporting Information S2: Tree Comparisons'' section.
 
@@
 
-\textcolor{blue}{D-Many of the equations are incorrect.} %only two in the main body.
+\textcolor{blue}{E - If there was a mention of the data being deposited, I missed it. I apologize if I did miss it.}
 
-We thank the reviewer to underline the two errors in the equations in the main body at lines 258 and 275, see our detailed response to the comments number 15 and 17 below.
-Also we recognise that the equations in our ``Supporting Information S2: Tree Comparisons'' are confusing and are actually unnecessary for the understanding of this paper.
-We therefore removed the details about the Robinson-Foulds and the Triplets distances from the ``Supporting Information S2: Tree Comparisons'' section.
-
-%-------------
-
-\textcolor{blue}{E-If there was a mention of the data being deposited, I missed it. I apologize if I did miss it.}
-
-We apologise again for the confusion here. See our comment above on Data Availability and Reproducibility.
+We apologise again for the confusion here. See our comment above on ``Data Availability and Reproducibility".
 
@@
 
-\textcolor{blue}{1-line 25 and elsewhere: incorrect smart quotes in ``best", ``classical"...}
+\textcolor{blue}{1 - line 25 and elsewhere: incorrect smart quotes in ``best", ``classical"...}
 
@@
 
-\textcolor{blue}{2-line 88 ``used it to infer a matrix" $->$ ``used it to simulate a matrix". ``inference" is used in place of simulation elsewhere.}
+\textcolor{blue}{2 - line 88 ``used it to infer a matrix" $->$ ``used it to simulate a matrix". ``inference" is used in place of simulation elsewhere.}
 
-We changed ``infer" into ``simulate" both at line 88 and in the Figure 1 caption.
+We changed ``infer" to ``simulate" both at line 88 and in the Figure 1 caption.
 
@@
 
-\textcolor{blue}{3-line 112: does diversitree just stop simulating when it reaches the correct number of tips, or does it deal with the biases that can result from doing that?
+\textcolor{blue}{3 - line 112: does diversitree just stop simulating when it reaches the correct number of tips, or does it deal with the biases that can result from doing that?
 (Stadler 2011 is the ref for a discussion of this issues, I believe)}
 
-See our response to the comment ``The procedure for generating a tree with fossils is a valid procedure." below.
+See our response to comment 6 below.
 
@@
 
-\textcolor{blue}{4-line 113: ``rates from a uniform distribution" with what upper and lower bounds?}
+\textcolor{blue}{4 - line 113: ``rates from a uniform distribution" with what upper and lower bounds?}
 
-Both birth ($\lambda$) and death ($\mu$) parameters are probabilities bounded between 0 and 1. We have added this to the new version of the manuscript:
+Both birth ($\lambda$) and death ($\mu$) parameters are probabilities bounded between 0 and 1. We have added this to the new version of the manuscript.
 
@@
 
-\textcolor{blue}{5-line 118: with 25 living and 25 fossil taxa. Are the fossils just put at the extinction point for the extinct lineages?}
+\textcolor{blue}{5 - line 118: with 25 living and 25 fossil taxa. Are the fossils just put at the extinction point for the extinct lineages?}
 
@@
 
-\textcolor{blue}{6-line 110 - line 120.
+\textcolor{blue}{6 - line 110 - line 120.
 The procedure for generating a tree with fossils is a valid procedure.
@@
 Easy of comparability with other work is one of the advantages of using formally characterized models.
-In this case, I think that using the TreeSim package http://cran.r-project.org/web/packages/TreeSim/TreeSim.pdf would have led to a study with higher impact.
+In this case, I think that using the TreeSim package would have led to a study with higher impact.
 This part of the methods section would be reduced to stating the parameter values chosen, and other workers could study similar tree generation simulations.
@@
 
-\textcolor{blue}{7-line 126 ``random base frequencies" is too vague. Dirichlet(1, 1, 1, 1) perhaps?}
+\textcolor{blue}{7 - line 126 ``random base frequencies" is too vague. Dirichlet(1, 1, 1, 1) perhaps?}
 
@@
 
-\textcolor{blue}{8-Do the simulations apply ascertainment biases as described by Lewis' Mkv (2001) model? It seems like they should... At minimum, the point should be discussed.}
+\textcolor{blue}{8 - Do the simulations apply ascertainment biases as described by Lewis' Mkv (2001) model? It seems like they should... At minimum, the point should be discussed.}
 
@@
 
-\textcolor{blue}{9-line 171 - 177: the authors need to explain the order of operations for the $M_F$ vs $M_C$ masking.
+\textcolor{blue}{9 - line 171 - 177: the authors need to explain the order of operations for the $M_F$ vs $M_C$ masking.
 It makes a difference.
@@
 We clarified this point ($M_C$ comes second) in the new version of the manuscript (lines @173 to @183).
-We also, clarified the potential confusion between the amount of data removed in one parameter ($M_F$ = 10\% for example) and the amount of missing data in the morphological matrix (5\%) in the same revised paragraph (lines @173 to @183).
+We also clarified the potential confusion between the amount of data removed in one parameter ($M_F$ = 10\% for example) and the amount of missing data in the morphological matrix (5\%) in the same revised paragraph (lines @173 to @183).
 
@@
 Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
-We first applied the parameters $M_L$ and $M_F$ on the matrix and then applied the $M_C$ parameter.
+We first applied the parameters $M_L$ and $M_F$ to the matrix and then applied the $M_C$ parameter.
 Therefore, when 10\% data was missing for both $M_L$ and $M_F$, 10\% data was missing in the morphological part of the matrix.
-However, when applying the $M_C$ parameter with the same rate (10\%) the resulting matrix potentially had more than 10\% data missing.
+However, when applying the $M_C$ parameter with the same percentage (10\%) the resulting matrix potentially had more than 10\% data missing.
 \end{minipage}
@@
 
-\textcolor{blue}{10-Header for 3.
+\textcolor{blue}{10 - Header for 3.
 should be ``building phylogenetic estimates" or ``estimating phylogenies" because the previous section described how the phylogenies and character data were ``built."}
@@
 
-\textcolor{blue}{11-RAxML searching: it would be nice to see some searches started from the true tree.
+\textcolor{blue}{11 - RAxML searching: it would be nice to see some searches started from the true tree.
 This is clearly not an option for real data sets, but it is helpful in simulation studies because it lets the reader determine if the results could be an artifact of insufficient tree searching.}
 
-Our simulations were designed to test the effects of missing data in as realistic a manner as possible given our constraints.
-Thus, we do not feel that this additional analysis would add much to the paper, especially since it is already rather long and complex.
+Our simulations were designed to test the effects of missing data in as realistic a manner as possible given computational constraints.
+Thus, we do not feel that this additional analysis would add much to the manuscript, especially since it is already rather long and complex.
 
@@
 
-\textcolor{blue}{12-Filtering the data sets for strong bootstrap support under the full data analysis surely introduces some bias in favor of the no-missing data analyses.}
+\textcolor{blue}{12 - Filtering the data sets for strong bootstrap support under the full data analysis surely introduces some bias in favor of the no-missing data analyses.}
 
@@
 If the no-missing data tree has low node support, it is likely that the changes in topology for the missing data trees are not linked to our focal question (the effect of removing characters, living taxa or fossils) but just to the low supported nodes.
-In fact, because we repeat the tree inference 125 times ($5^3$ parameters combinations), it is likely to have 125 different pseudo-random solution for the nodes with really low phylogentic support and that, therefore, the observed changes in topologies are just linked to the pseudo-random replication.
-Note however that we did not exculde this scenario from our protocol the trees selected with a median bootstrap $>$ 50 could well contain some unresolved nodes with a low bootstrap ($<$ 50).
+In fact, because we repeat the tree inference 125 times ($5^3$ parameters combinations), it is likely that we have 125 different pseudo-random solution for the nodes with really low phylogenetic support and that, therefore, the observed changes in topologies are just linked to the pseudo-random replication.
+Note however that we did not exclude this scenario from our protocol; the trees selected with a median bootstrap $>$ 50 could well contain some unresolved nodes with a low bootstrap ($<$ 50).
 We've added the following justification in our revision (lines @215 to @220):
@@
 50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
-This step was implemented to make sure that the differences we observed in topologies (see below) where due to the amount of missing data for each parameters ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
+This step was implemented to make sure that the differences we observed in topologies (see below) were due to the amount of missing data for each parameter ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
 \end{minipage}
@@
 
-\textcolor{blue}{13-MrBayes: using strongly informative priors for the true parameters does more than speed up the searching, it represents an ``unfair" advantage (an option that would not be available in real data analysis in which one does not know the true values) for the Bayesian method.
+\textcolor{blue}{13 - MrBayes: using strongly informative priors for the true parameters does more than speed up the searching, it represents an ``unfair" advantage (an option that would not be available in real data analysis in which one does not know the true values) for the Bayesian method.
 This makes the studies report of superior behavior of Bayesian methods less persuasive.}
@@
 
-\textcolor{blue}{14-line 254: RF measures the difference between the number of clades and the twice number of shared clades across two trees.
+\textcolor{blue}{14 - line 254: RF measures the difference between the number of clades and the twice number of shared clades across two trees.
 The statement that it ``measures the number of shared clades across two trees" should be revised.}
@@
 
-\textcolor{blue}{15-line 258: RF is 0 (not 1) when the trees are identical. It's maximum (for 2 rooted binary trees) is 2(n-2) not n-2.}
+\textcolor{blue}{15 - line 258: RF is 0 (not 1) when the trees are identical. It's maximum (for 2 rooted binary trees) is 2(n-2) not n-2.}
 
-We thanks the reviewer for pointing at this error and we changed the following sentence (lines @271 to @273 in our revision):
+We thanks the reviewer for pointing out this error and we changed the following sentence (lines @271 to @273 in our revision):
 
@@
 
-\textcolor{blue}{16-The ``scaling" to produce the Normalized RF ``distance" is confusing. It is also inverting the sign of the distance, so it has become a measure of similarity.}
+\textcolor{blue}{16 - The ``scaling" to produce the Normalized RF ``distance" is confusing. It is also inverting the sign of the distance, so it has become a measure of similarity.}
 
-We agree that this is confusing and thank the reviewer for pointing this inconsistency out.
+We agree that this is confusing and thank the reviewer for pointing this out.
 We changed the mentions of ``Normalised Robinson-Foulds distance'' and ``Normalised Triplets distance'' to ``Normalised Robinson-Foulds metric'' and ``Normalised Triplets metric'' throughout the our revision.
@@
 We normalised this metric following Bogdanowicz \textit{et al.} (2012)'s Normalised Tree Similarity (NTS) method. This methods scales %not sure if scale here?
-any tree comparison metric using the mean distance between 1000 random trees (see Supporting Information S2: Tree Comparisons for the calculation details).
+any tree comparison metric using the mean distance between 1000 random trees (see Supporting Information S2: Tree Comparisons for the calculation details)
+% NC: Haven't you got rid of this now????
+.
 This method is a generalisation of the topological accuracy method (Price \textit{et al.} 2010) allowing to compare topological differences between any tree with any tree comparison metric.
 In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
+Note that once rescaled, the Normalised Robinson-Foulds metric is a measure of similarity, rather than of distance like the original Robinson-Foulds metric.
 \end{minipage}
-% Note that NRF distance is actually a similarity ...
-
+% NC: Does triplets need to be in here too???
 %-------------
 
-\textcolor{blue}{17-line 275: why isn't the upper bound of the triples distance n choose 3 (instead of n choose 4). I think you got the formula from an unrooted quartet distance.}
+\textcolor{blue}{17 - line 275: why isn't the upper bound of the triples distance n choose 3 (instead of n choose 4). I think you got the formula from an unrooted quartet distance.}
 
@@
 
-\textcolor{blue}{18-Using the BC stat may be similar in spirit to using a t-test, but it is not equivalent.}
+\textcolor{blue}{18 - Using the BC stat may be similar in spirit to using a t-test, but it is not equivalent.}
 
-We clarified this point in the text as well as in the figures 2 and 3 captions (``equivalent" $->$ ``similar" ; lines @322 to @324 in our revision):
+We clarified this point in the text as well as in the figures 2 and 3 captions (``equivalent" $->$ ``comparable" ; lines @322 to @324 in our revision):
 
 \hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
+Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
 \end{minipage}
@@
 
-\textcolor{blue}{19-Figure 3: B is confusing. If both curves are probability distributions, then they both have to integrate to 1. So B(y) can't cover just a subset of the area covered by B(x).}
+\textcolor{blue}{19 - Figure 3: B is confusing. If both curves are probability distributions, then they both have to integrate to 1. So B(y) can't cover just a subset of the area covered by B(x).}
 
-We thank the reviewer for pointing out this error. We fixed the shape of the distribution in Figure 3-B.
+We thank the reviewer for pointing out this error. We fixed the shape of the distribution in Figure 3B.
 
@@
 
-\textcolor{blue}{20-``Combined effect of missing data parameters" section ANOVA or multiple regression are the appropriate analyses to tease apart multiple interacting factors.
+\textcolor{blue}{20 - ``Combined effect of missing data parameters" section ANOVA or multiple regression are the appropriate analyses to tease apart multiple interacting factors.
 Here the authors opt to describe results without statistical tests, which is disappointing.}
@@
 
-\textcolor{blue}{21-line 398: ``mMthods" $->$ ``Methods"}
+\textcolor{blue}{21 - line 398: ``mMthods" $->$ ``Methods"}
 
@@
 
-\textcolor{blue}{22-line 420: contra the statement here, a consensus collapsing to a polytomy will affect the RF distance (could go up or down, but it will change).}
+\textcolor{blue}{22 - line 420: contra the statement here, a consensus collapsing to a polytomy will affect the RF distance (could go up or down, but it will change).}
 
@@
 \hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-In this case, the Normalised Robinson-Foulds metric will decreases when the fossil are present at a lower taxonomic level (i.e.
+In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present at a lower taxonomic level (i.e.
 separated with more nodes from the root) but affects the clade conservation less at higher taxonomic level (i.e.
 separated with less nodes from the root).
-On the opposite though, because the fossil in a high level clade has many chances to branch on different nodes within that clade, it will be more likely to act as a wildcard taxa and decrease the Normalised Triplets metric.
+Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxa and decrease the Normalised Triplets metric.
 Therefore, the $M_{F}$ parameter is likely to affect the Normalised Robinson-Foulds metric less than the Normalised Triplets metric for the Bayesian consensus trees.
@@
 
-\textcolor{blue}{23-more confusion about the min RF here.
+\textcolor{blue}{23 - more confusion about the min RF here.
 Also note that if you are counting the entire leaf set as a clade (e.g the star tree having N=1), then the number of clades in rooted binary tree is n-1 (not n-2).
@@
 
-See our response to the comment D above.
+See our response to comment D above.
 %-------------
 
-\textcolor{blue}{24-The definition of the scaled RF implies that it is never positive (sinc the numerator cannot be positive and the denominator is positive).}
+\textcolor{blue}{24 - The definition of the scaled RF implies that it is never positive (sinc the numerator cannot be positive and the denominator is positive).}
 
-See our response to the comment D above.
+See our response to comment D above.
 %-------------
 
-\textcolor{blue}{25-The decision to use Yule trees as the base line for the mean distance is not explained, and seems odd given that the true tree are not Yule trees.
+\textcolor{blue}{25 - The decision to use Yule trees as the base line for the mean distance is not explained, and seems odd given that the true tree are not Yule trees.
 I'm not sure that it makes much difference, but it is odd.}
 
-We followed the protocol in the cited reference (Bogdanowicz 2012) where the normalisation of the metrics is done on the mean value of that metric when comparing to random trees.
-Using random Yule trees is the best way to generate purely random trees since they are not based on any evolutionary scenario and takes only one parameter which is the number of tips (e.g.
-as opposed to the Birth-Death tree that assumes more realistic evolutionary scenario and takes multiple parameters).
+We followed the protocol in the cited reference (Bogdanowicz 2012) where the normalisation of the metrics is done on the mean value of that metric when compared to random trees.
+Using random Yule trees is the best way to generate purely random trees since they are not based on any evolutionary model and take only one parameter, the number of tips, as opposed to the Birth-Death tree that use a more realistic evolutionary model and take multiple parameters.
 %-------------
 
-\textcolor{blue}{26-The NTS can only -infinity for distances that have the property that the mean distance between random Yule trees is 0.
+\textcolor{blue}{26 - The NTS can only -infinity for distances that have the property that the mean distance between random Yule trees is 0.
 This is not true of either distance metric here.
@@
 
-\textcolor{blue}{27-Note that JavaScript is a language, so it a bit confusing to refer to a Java program as a ``Java script."}
+\textcolor{blue}{27 - Note that JavaScript is a language, so it a bit confusing to refer to a Java program as a ``Java script."}
 
@@
 
-\textcolor{blue}{28-Equation (6) is not correct.
+\textcolor{blue}{28 - Equation (6) is not correct.
 You shouldn't set the summation sign as the left hand side of an equation because the notation has a standard definition.
@@
 
-\textcolor{blue}{29-Equation (7) is true for a three-leaf tree with a polytomy occurring with probability = 1/4 and the other 3 trees being equiprobable.
+\textcolor{blue}{29 - Equation (7) is true for a three-leaf tree with a polytomy occurring with probability = 1/4 and the other 3 trees being equiprobable.
 I don't think that is is true for any other weighting of trees for other numbers of taxa.
@@
 
-See our response to the comment D above.
+See our response to comment D above.
 %-------------
 
-\textcolor{blue}{30-I don't think that equations 8, 9, or 10 are correct for any case other than the 3 taxon case (though I do note that the authors are using n choose 3 here).}
+\textcolor{blue}{30 - I don't think that equations 8, 9, or 10 are correct for any case other than the 3 taxon case (though I do note that the authors are using n choose 3 here).}
 
-See our response to the comment D above.
+See our response to comment D above.
 %-------------
 
-\textcolor{blue}{31-Equations 13 and 14. It seems easier to just drop the summation sign here, and just define $a_i$ and $b_i$ to the right hand side of these equations.}
+\textcolor{blue}{31 - Equations 13 and 14. It seems easier to just drop the summation sign here, and just define $a_i$ and $b_i$ to the right hand side of these equations.}
 
@@
 
+\bigskip
+
+We hope we have dealt with all these comments satisfactorily.
+
+\closing{Yours sincerely,}
+
 \end{letter}

### 2015-05-12 Manuscript/PLoSONEresponse/OnlineAppendices-LaTeXSuppFiles/SupportingInformation3.tex
Updated final response

--- 2015-04-28 Manuscript/PLoSONESubmission/OnlineAppendices-LaTeXSuppFiles/SupportingInformation3.tex
@@
 The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row).
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
 The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
@@
 The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
 The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
@@
 The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
 TThe Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
@@
 \begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).}
 \label{Tab_Supp_summary_metric_allparam}
@@
 \begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{L}$ missing data parameter only.}
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_ML}
@@
 \begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{F}$ missing data parameter only.}
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_MF}
@@
 \begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{C}$ missing data parameter only.}
+\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_MC}
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{L}$ missing data parameter only.}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
 \label{Tab_Supp_summary_BC_ML}
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{F}$ missing data parameter only.}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
 \label{Tab_Supp_summary_BC_MF}
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr) for the $M_{C}$ missing data parameter only.}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
 \label{Tab_Supp_summary_BC_MC}

### 2015-05-12 Manuscript/PLoSONEresponse/TEM_PLoS_revision.tex
Updated final response

@@
 0.9-6 \cite{fitzjohndiversitree2012}.
-We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution but maintaining $\lambda$ $>$ $\mu$ \cite{paradistime-dependent2011}.
+We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution (bounded between 0 and 1) but maintaining $\lambda$ $>$ $\mu$ \cite{paradistime-dependent2011}.
 Empirical Total Evidence matrices vary in whether they have more fossil than living taxa or vice versa.
@@
 Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
-We first applied the parameters $M_L$ and $M_F$ on the matrix and then applied the $M_C$ parameter.
+We first applied the parameters $M_L$ and $M_F$ to the matrix and then applied the $M_C$ parameter.
 Therefore, when 10\% data was missing for both $M_L$ and $M_F$, 10\% data was missing in the morphological part of the matrix.
-However, when applying the $M_C$ parameter with the same rate (10\%) the resulting matrix potentially had more than 10\% data missing.
+However, when applying the $M_C$ parameter with the same percentage (10\%) the resulting matrix potentially had more than 10\% data missing.
 
@@
 We used RAxML because it automatically corrects for acquisition bias \cite{lewisa2001}.
-It also is heavily used in the literature for Maximum Likelihood tree inference (e.g.
+It is also heavily used in the literature for Maximum Likelihood tree inference (e.g.
 \cite{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014}) and is one of the fastest methods available \cite{Stamatakis01102008}.
@@
 50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
-This step was implemented to make sure that the differences we observed in topologies (see below) where due to the amount of missing data for each parameters ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
+This step was implemented to make sure that the differences we observed in topologies (see below) were due to the amount of missing data for each parameter ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
 On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \cite{pattengale2010many}.
@@
 This methods scales any tree comparison metric using the mean distance between 1000 random trees (see ~\nameref{S2_Appendix2} for the calculation details).
-This method is a generalisation of the topological accuracy method \cite{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.
-In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
+This method is a generalisation of the topological accuracy method \cite{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
+Note that once rescaled, the Normalised Robinson-Foulds metric is a measure of similarity, rather than of distance like the original Robinson-Foulds metric.
 
@@
 The Bhattacharyya Coefficient is the probability of overlap between two distributions \cite{Bhattacharyya} (see ~\nameref{S2_Appendix2} for calculation details).
-Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
+Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
 To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%), i.e.
@@
 every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
-In this case, the Normalised Robinson-Foulds metric will decreases when the fossil are present at a lower taxonomic level (i.e.
+In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present at a lower taxonomic level (i.e.
 separated with more nodes from the root) but affects the clade conservation less at higher taxonomic level (i.e.
 separated with less nodes from the root).
-On the opposite though, because the fossil in a high level clade has many chances to branch on different nodes within that clade, it will be more likely to act as a wildcard taxa and decrease the Normalised Triplets metric.
+Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxa and decrease the Normalised Triplets metric.
 Therefore, the $M_{F}$ parameter is likely to affect the Normalised Robinson-Foulds metric less than the Normalised Triplets metric for the Bayesian consensus trees.
-%This does not decrease the Normalised Robinson-Foulds metric, because clades are conserved, but decreases the Normalised Triplets metric because the fossils act as wildcard taxa.
 Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50).

### 2015-06-08 Manuscript/MPESubmission/response.tex
Initial commit of MPE version

--- 2015-05-12 Manuscript/PLoSONEresponse/response.tex
@@
 All the data generated by the simulations are directly available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare} (goo.gl/h5Kxb5).
-Note that along with the data we share the full code used for repeating the submitted manuscript, from simulating the data to generating the figures and compiling the manuscript (available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{GitHub} (goo.gl/ewRsrm).
+Note that along with the data we share the full code used for repeating the submitted manuscript, from simulating the data to generating the figures and compiling the manuscript available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{GitHub} (goo.gl/ewRsrm).
 All this code is provided with a detailed description of each step that we are happy to develop if needed.
@@
 If the two distributions overlap completely then they have Bhattacharyya Coefficient and probability of being the same of 1.
-At the other extreme, if the two distributions are completely disjointed, then they have a Bhattacharyya Coefficient and hence probability of being the same of 0 (line 306 and 307 in the previously submitted manuscript).
+At the other extreme, if the two distributions are completely disjointed, then they have a Bhattacharyya Coefficient and hence probability of being the same of 0 (line 306 and 307 in the previously submitted manuscript and lines 320 and 319 in our revision).
 Thus, the Bhattacharyya Coefficient is a probability, comparable to, but not the same as, a p-value that could be calculated using frequentist approaches such as t-tests, ANOVAs or GLMs.
@@
 \textit{The figure shows two extremely similar distributions.
-However, although the Bhattacharyya Coefficient shows that the distributions are really similar (0.999 overlap probability), a frequentist approach (simple t-test) suggests that there is a significant difference between the means (p-value \textless 0.0000001).
+However, although the Bhattacharyya Coefficient shows that the distributions are really similar (0.999 overlap probability), a frequentist approach (simple t-test) suggests that there is a significant difference between the means (p-value $<$ 0.0000001).
 This highlights the reason we chose to use the Bhattacharyya Coefficient in our manuscript.}
@@
 We would like to emphasize that in this study we focused on the effect of missing data in the morphological part of the matrix on topology only, and we are ignoring branch lengths.
-The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and were a minor part of the MCMC sampler (2.17\% of the moves of the MCMC sampler for the three priors combined).
+The choice of the following priors (``exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5 for both partitions" (line 228 in the previously submitted manuscript and line 240 in our revision) and the ``transition/transversion ratio prior of two sampled from a strong beta 229 distribution ($\beta$(80,40))" (line 229 in the previously submitted manuscript and line 241 in our revision)) affected mainly the molecular part of the matrix that remains constant (gamma distribution and transition/transversion ratio) and were a minor part of the MCMC sampler (2.17\% of the moves of the MCMC sampler for the three priors combined).
 Therefore we are confident that the topology is minimally affected by the priors, allowing us to explore the effect of missing morphological data on topology as well as reducing our computational time to a realistic 1.4 CPU centuries.
-Note that the use of a starting tree that is not a prior and just ``the ``true" tree's topology'' (line 226) was used to avoid the chains getting stuck in local optima.
+Note that the use of the ``the ``true" tree's topology'' (line 226 in the previously submitted manuscript and line 237 in our revision) as a starting tree is not a prior and was used to avoid the chains getting stuck in local optima.
 
@@
 
-Lines @234 to @247 in our revision:
+Lines 248 to 249 in our revision:
 
@@
 
-Lines @206 to @209 in our revision:
+Lines 208 to 209 in our revision:
 
 \hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-We used RAxML because it automatically corrects for acquisition bias (Lewis 2001), it also is heavily used in the literature for Maximum Likelihood tree inference (e.g.
+We used RAxML because it automatically corrects for acquisition bias (Lewis 2001).
+It is also heavily used in the literature for Maximum Likelihood tree inference (e.g.
 [...]) and is one of the fastest methods available (Stamatakis 2008).
@@
 
-Lines @231 to @232 in our revision:
+Lines 233 to 234 in our revision:
 
@@
 
-\textcolor{blue}{D - Many of the equations are incorrect.} %only two in the main body.
+\textcolor{blue}{D - Many of the equations are incorrect.}
 
@@
 
-Yes. We have clarified as follows in our revision (lines @118 to @119):
+Yes. We have clarified as follows in our revision (lines 118 to 119):
 
@@
 
-Lines @126 to @130 in our revision:
+Lines 126 to 129 in our revision:
 
@@
 
-We clarified this point ($M_C$ comes second) in the new version of the manuscript (lines @173 to @183).
-We also clarified the potential confusion between the amount of data removed in one parameter ($M_F$ = 10\% for example) and the amount of missing data in the morphological matrix (5\%) in the same revised paragraph (lines @173 to @183).
+We clarified this point ($M_C$ comes second) in the new version of the manuscript (lines 173 to 183).
+We also clarified the potential confusion between the amount of data removed in one parameter ($M_F$ = 10\% for example) and the amount of missing data in the morphological matrix (5\%) in the same revised paragraph (lines 173 to 183).
 
@@
 Note however that we did not exclude this scenario from our protocol; the trees selected with a median bootstrap $>$ 50 could well contain some unresolved nodes with a low bootstrap ($<$ 50).
-We've added the following justification in our revision (lines @215 to @220):
+We've added the following justification in our revision (lines 215 to 220):
 
@@
 
-We revised this statement lines @267 to @269 in our revision:
+We revised this statement lines 267 to 269 in our revision:
 
@@
 
-We thanks the reviewer for pointing out this error and we changed the following sentence (lines @271 to @273 in our revision):
+We thanks the reviewer for pointing out this error and we changed the following sentence (lines 271 to 273 in our revision):
 
@@
 We changed the mentions of ``Normalised Robinson-Foulds distance'' and ``Normalised Triplets distance'' to ``Normalised Robinson-Foulds metric'' and ``Normalised Triplets metric'' throughout the our revision.
-Additionally, we clarified the scaling method as follows (lines @276 to @284 in our revision):
+Additionally, we clarified the scaling method as follows (lines 276 to 284 in our revision):
 
 \hfill\begin{minipage}{\dimexpr\textwidth-1cm}
-We normalised this metric following Bogdanowicz \textit{et al.} (2012)'s Normalised Tree Similarity (NTS) method. This methods scales %not sure if scale here?
-any tree comparison metric using the mean distance between 1000 random trees (see Supporting Information S2: Tree Comparisons for the calculation details)
-% NC: Haven't you got rid of this now????
-.
+We normalised this metric following Bogdanowicz \textit{et al.} (2012)'s Normalised Tree Similarity (NTS) method.
+This methods scales any tree comparison metric using the mean distance between 1000 random trees (see Supporting Information S2: Tree Comparisons for the calculation details).
 This method is a generalisation of the topological accuracy method (Price \textit{et al.} 2010) allowing to compare topological differences between any tree with any tree comparison metric.
@@
 \end{minipage}
-% NC: Does triplets need to be in here too???
 %-------------
@@
 
-We corrected this error lines @288 to @290 in our revision:
+We corrected this error lines 290 to 291 in our revision:
 
@@
 
-We clarified this point in the text as well as in the figures 2 and 3 captions (``equivalent" $->$ ``comparable" ; lines @322 to @324 in our revision):
+We clarified this point in the text as well as in the figures 2 and 3 captions (``equivalent" $->$ ``comparable" ; lines 324 to 326 in our revision):
 
@@
 
-We revised this statement at lines @434 to @442 in our revision:
+We revised this statement at lines 436 to 444 in our revision:
 

### 2015-06-08 Manuscript/MPESubmission/OnlineAppendices-LaTeXSuppFiles/SupportingInformation2.tex
Initial commit of MPE version

--- 2015-04-27 Manuscript/PLoSONEresponse/OnlineAppendices-LaTeXSuppFiles/SupportingInformation2.tex
@@
 \subsection{Normalised Tree Similarity}
-For any tree with \textit{n} taxa compared using a tree distance metric $m$, Normalized Tree Similarity, $NTS_m$ \cite{Bogdanowicz2012}, represents the similarity score for the two trees given the expected distance between two random Yule trees \cite{Bogdanowicz2012} with $n$ taxa.
-If $\bar{d}_{m,n}$\textit{(rand)} is the average distance between two random Yule trees with $n$ taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} each containing $n$ taxa, then:
+For any tree with \textit{n} taxa compared using a tree difference metric $m$, Normalized Tree Similarity, $NTS_m$ \cite{Bogdanowicz2012}, represents the similarity score for the two trees given the expected difference between two random Yule trees \cite{Bogdanowicz2012} with $n$ taxa.
+If $\bar{d}_{m,n}$\textit{(rand)} is the average difference between two random Yule trees with $n$ taxa and $d_{m,n}$\textit{(x,y)} the difference between the two trees \textit{x} and \textit{y} each containing $n$ taxa, then:
 \begin{equation}
@@
 
-We used the NTS method to scale all the Robinson Foulds and Triplets distances calculated in our analyses, using the TreeCmp JavaScript \cite{Bogdanowicz2012}.
+We used the NTS method to scale all the Robinson Foulds and Triplets metrics calculated in our analyses, using the TreeCmp JavaScript \cite{Bogdanowicz2012}.
 

### 2015-06-08 Manuscript/MPESubmission/TEM_draft.tex
Initial commit of cover letter for MPE

--- 2015-05-12 Manuscript/PLoSONEresponse/TEM_PLoS_revision.tex
@@
 \subsection*{Data reproducibility and availability}
-\noindent{\bf Data reproducibility and availability.} All the code used in this analysis is available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{\color{blue}{GitHub}} with some information on how to use the various functions.
-Additionally all the simulated data is available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{\color{blue}{FigShare}}.
+\noindent{\bf Data reproducibility and availability.} All the code used in this analysis is available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{\color{blue}{GitHub}} (\href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{
+goo.gl/4djNUf}) with some information on how to use the various functions.
+Additionally all the simulated data is available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{\color{blue}{FigShare}} (dx.doi.org/10.6084/m9.figshare.1306861).
 

### 2015-06-08 Manuscript/MPESubmission/TEM_draft.tex
Reformatted to MPE format

--- 2015-05-12 Manuscript/PLoSONEresponse/TEM_PLoS_revision.tex
@@
-\documentclass[10pt,letterpaper]{article}
-\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+\documentclass[12pt,letterpaper]{article}
 
+%Packages
+\usepackage{xcolor}
 \usepackage{changepage}
-\usepackage{color}
-\usepackage[utf8]{inputenc}
-\usepackage{textcomp,marvosym}
+\usepackage{pdflscape}
 \usepackage{fixltx2e}
-\usepackage{amsmath,amssymb}
-\usepackage{cite}
-\usepackage{nameref,hyperref}
-\usepackage[right]{lineno}
-\usepackage{microtype}
-\DisableLigatures[f]{encoding = *, family = * }
-\usepackage{rotating}
-\usepackage{setspace} 
-\doublespacing
+\usepackage{textcomp}
+\usepackage{fullpage}
+\usepackage{natbib}
+\usepackage{float}
+\usepackage{latexsym}
+\usepackage{url}
+\usepackage{epsfig}
+\usepackage{graphicx}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{bm}
+\usepackage{array}
+\usepackage[version=3]{mhchem}
+\usepackage{ifthen}
+\usepackage{caption}
+\usepackage{hyperref}
+\usepackage{amsthm}
+\usepackage{amstext}
+\usepackage{enumerate}
+\usepackage[osf]{mathpazo}
+\usepackage{dcolumn}
+\usepackage{lineno}
+\usepackage{longtable}
+\pagenumbering{arabic}
 
-% Text layout
+%Pagination style and stuff
+\linespread{2}
 \raggedright
-\setlength{\parindent}{0.5cm}
-\textwidth 5.25in 
-\textheight 8.75in
-
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
-
-\bibliographystyle{plos2009}
-
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-\date{}
-
-% Header and Footer with logo
-\usepackage{lastpage,fancyhdr,graphicx}
-\pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-\lhead{\includegraphics[natwidth=1.3in,natheight=0.4in]{PLOSlogo.png}}
-\rfoot{\thepage/\pageref{LastPage}}
-\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
-\fancyheadoffset[L]{2.25in}
-\fancyfootoffset[L]{2.25in}
-\lfoot{\sf PLOS}
-
-\newcommand{\lorem}{{\bf LOREM}}
-\newcommand{\ipsum}{{\bf IPSUM}}
-
-%% END MACROS SECTION
+\setlength{\parindent}{0.5in}
+%\setcounter{secnumdepth}{0}
 
@@
 
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf\newline{Effects of missing data on topological inference using a Total Evidence approach}
-}
-\newline
-\\
-Thomas Guillerme\textsuperscript{1,2,*},
-Natalie Cooper\textsuperscript{1,2}
-\\
-\bf{1} School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland
-\\
-\bf{2} Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland
-\\
-* E-mail: Corresponding guillert@tcd.ie
-\end{flushleft}
-\section*{Abstract}
+%Running head
+\begin{flushright}
+%Version dated: \today
+\end{flushright}
+\bigskip
+\noindent Running head: Missing data and topology in Total Evidence method
+
+\bigskip
+\medskip
+\begin{center}
+
+\noindent{\Large \bf Effects of missing data on topological inference using a Total Evidence approach}
+
+\bigskip
+
+\noindent {\normalsize \sc Thomas Guillerme$^a$$^,$$^b$$^*$, and Natalie Cooper$^a$$^,$$^b$$^,$$^1$}\\
+\noindent {\small \it 
+$^a$School of Natural Sciences, Trinity College Dublin, Dublin 2, Ireland.\\
+$^b$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
+$^*$Corresponding author. Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.\\
+$^1$Present address: Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK. E-mail: nhcooper12@gmail.com}\\
+\end{center}
+
+\vspace{1in}
+
+%Line numbering
+\modulolinenumbers[1]
+\linenumbers
+
+\newpage
+\section* {Abstract}
 To fully understand macroevolutionary patterns and processes, we need to include both extant and extinct species in our models.
@@
 
-\linenumbers
+\bigskip
+\noindent
+\textbf{Keywords:} morphological characters, Bayesian, Maximum Likelihood, topology, fossil, living.
 
@@
 
-\section*{Introduction}
-Although most species that have ever lived are now extinct \cite{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species (e.g.
-\cite{meredithimpacts2011,jetzthe2012}).
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events (e.g.
-\cite{pyrondivergence2011}), relationships among lineages (e.g.
-\cite{manosphylogeny2007}) or niche occupancy (e.g.
-\cite{pearmanniche2008}).
-This has led to increasing consensus among evolutionary biologists that fossil taxa should be included in macroevolutionary studies \cite{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
+\newpage
+\section{Introduction}
+Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species (e.g.
+\citealp{meredithimpacts2011,jetzthe2012}).
+Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages (e.g.
+\citealp{manosphylogeny2007}) or niche occupancy (e.g.
+\citealp{pearmanniche2008}).
+This has led to increasing consensus among evolutionary biologists that fossil taxa should be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
 However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments (e.g.
-\cite{pyrondivergence2011,ronquista2012,BEASTmaster}).
+\citealp{pyrondivergence2011,ronquista2012,BEASTmaster}).
 
@@
 Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
-Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \cite{simpson1945}.
+Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \citep{simpson1945}.
 This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty.
@@
 Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates are used to time calibrate phylogenies \cite{zuckerkandl1965}.
+Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates are used to time calibrate phylogenies \citep{zuckerkandl1965}.
 There have been great improvements in the theory and application of these two approaches (e.g.
-\cite{bapsta2013,stadlerdating2013,heaththe2013}) as well as much debate about the ``best" approach to use (e.g.
-\cite{spencerefficacy2013,wrightbayesian2014}).
+\citealp{bapsta2013,stadlerdating2013,heaththe2013}) as well as much debate about the ``best" approach to use (e.g.
+\citealp{spencerefficacy2013,wrightbayesian2014}).
 However neither approach uses all the available data.
 
-A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \cite{eernissetaxonomic1993}.
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as tip-dating \cite{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \cite{ronquista2012}.
+A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \citep{eernissetaxonomic1993}.
+This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as tip-dating \citealp{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
 Total Evidence methods have been successfully applied to empirical data (e.g.
-\cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}), and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
+\citealp{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}), and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
 However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
 Morphological data for living taxa are rarely collected when molecular data are available (e.g.
-\cite{O'Leary08022013} \textit{vs.} \cite{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record (for example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \cite{sansomfossilization2013}; and molecular data is (nearly) always unavailable.
-Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \cite{salamin2003}.
+\citealp{O'Leary08022013} \textit{vs.} \citealp{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record (for example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data is (nearly) always unavailable.
+Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data is not biased towards particular subclades \cite{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}), it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data is not biased towards particular subclades \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
@@
 In this study, we focus on the effect of missing data on our ability to recover a ``best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species (e.g.
-\cite{meredithimpacts2011,jetzthe2012}), or for studying evolutionary transitions (e.g.
-\cite{friedmanexplosive2010}).
+\citealp{meredithimpacts2011,jetzthe2012}), or for studying evolutionary transitions (e.g.
+\citealp{friedmanexplosive2010}).
 Although branch length estimation is also important (namely for timing extinction and/or speciation events - e.g.
-\cite{ronquista2012}), we do not consider branch lengths in this study.
-This is partially due to difficulties with simulating branch lengths and topology simultaneously, but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \cite{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+\citealp{ronquista2012}), we do not consider branch lengths in this study.
+This is partially due to difficulties with simulating branch lengths and topology simultaneously, but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework.
@@
 Here we use a simulation approach to assess the effect of missing data on tree topologies inferred from Total Evidence matrices.
-Since the molecular part of a Total Evidence matrix acts like a ``classical" molecular matrix containing only the living taxa \cite{ronquista2012}, the effect of missing data on such matrices is well known \cite{wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011}.
+Since the molecular part of a Total Evidence matrix acts like a ``classical" molecular matrix containing only the living taxa \citep{ronquista2012}, the effect of missing data on such matrices is well known \citep{wiensmissing2006,wiensmissing2008,lemmonthe2009,rouresite-specific2011}.
 Therefore, we focus only on missing data in the morphological part of the matrix.
@@
 
-\section*{Methods}
-To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Fig.~\ref{Fig_Outline}).
+\section{Materials and Methods}
+To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Figure 1).
 \begin{enumerate}
@@
 
-\begin{figure}[h]
-\caption{{\bf Protocol outline.}
-(1) We randomly generated a birth-death tree (the ``true" tree) and used it to simulate a matrix with no missing data (the ``complete" matrix).
-(2) We removed data from the morphological part of the ``complete" matrix resulting in 125 ``missing-data" matrices.
-(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
-(4) We compared the ``missing-data" trees to the ``best" tree.
-We repeated these four steps 50 times.}
-\label{Fig_Outline}
-\end{figure}
+%\begin{figure}[h]
+%\caption{{\bf Protocol outline.}
+%(1) We randomly generated a birth-death tree (the ``true" tree) and used it to simulate a matrix with no missing data (the ``complete" matrix).
+%(2) We removed data from the morphological part of the ``complete" matrix resulting in 125 ``missing-data" matrices.
+%(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
+%(4) We compared the ``missing-data" trees to the ``best" tree.
+%We repeated these four steps 50 times.}
+%\label{Fig_Outline}
+%\end{figure}
 
@@
 
-\subsection*{1. Generating the matrix}
+\subsection{Generating the matrix}
 \label{Generating_the_matrix}
 First we randomly generated a ``true" tree of 50 taxa in R v.
-3.0.2 \cite{R302} using the package diversitree v.
-0.9-6 \cite{fitzjohndiversitree2012}.
-We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution (bounded between 0 and 1) but maintaining $\lambda$ $>$ $\mu$ \cite{paradistime-dependent2011}.
+3.0.2 \citep{R302} using the package diversitree v.
+0.9-6 \citep{fitzjohndiversitree2012}.
+We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution (bounded between 0 and 1) but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
 Empirical Total Evidence matrices vary in whether they have more fossil than living taxa or vice versa.
-For example, fossil taxa make up 88\% \cite{beckancient2014}, 58\% \cite{schragocombining2013}, 48\% \cite{pyrondivergence2011}, 31\% \cite{ronquista2012} and 31\% \cite{slaterphylogenetic2013} of taxa in various studies.
+For example, fossil taxa make up 88\% \citep{beckancient2014}, 58\% \citep{schragocombining2013}, 48\% \citep{pyrondivergence2011}, 31\% \citep{ronquista2012} and 31\% \citep{slaterphylogenetic2013} of taxa in various studies.
 To avoid biasing our simulations towards either living or fossil taxa and to make each simulation comparable, we implemented a rejection sampling algorithm to select only trees with 25 living and 25 fossil taxa.
@@
 The molecular matrix was inferred from the ``true" tree using the R package phyclust v.
-0.1-14 \cite{chen2011}.
-The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \cite{ranbaut1997seqgen} and using the HKY model \cite{HKY85} with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1) and transition/transversion rate of two \cite{douadycomparison2003}.
-The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \cite{yangamong-site1996}.
+0.1-14 \citep{chen2011}.
+The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1) and transition/transversion rate of two \citep{douadycomparison2003}.
+The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
 We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
@@
 We inferred the morphological matrix using the R package ape v.
-3.0-11 \cite{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
+3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see ~\nameref{S1_Appendix1} and Fig.
-S1 within).
+These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see Appendix A and Fig.
+A1 within).
 We then ran an independent discrete character simulation for each character using the ``true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
-evolutionary rate) from one character state to another \cite{Pagel22011994}.
+evolutionary rate) from one character state to another \citep{Pagel22011994}.
 This method allows us to have only two parameters for each character: the number of states and the evolutionary rate.
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used low evolutionary rate parameters to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \cite{wagner2000,davalosintegrating2014,wrightbayesian2014}.
+We used low evolutionary rate parameters to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014,wrightbayesian2014}.
 
@@
 
-\subsection*{2. Removing data}
+\subsection{Removing data}
 \label{Removing_data}
@@
 \item{$M_{F}$, the proportion of missing data in the fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the quality of the fossil record, a common problem due to preservation biases \cite{sansomfossilization2013}. 
+This parameter illustrates the quality of the fossil record, a common problem due to preservation biases \citep{sansomfossilization2013}. 
 \item{$M_{C}$, the proportion of missing morphological characters for both living and fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%. }
@@
 This reflects the amount of effort put into identifying the characters (e.g.
-\cite{O'Leary08022013}).
+\citealp{O'Leary08022013}).
 \end{enumerate}
@@
 
-\subsection*{3. Estimating phylogenies}
+\subsection{Estimating phylogenies}
 From the resulting matrices we generated two types of trees: the ``best" tree inferred from the ``complete" matrix and the ``missing-data" trees inferred from the 125 matrices with various amounts of missing data.
@@
 In real world situations, the ``true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g.
-\cite{rozen2005}).
+\citealp{rozen2005}).
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the ``best" tree, rather than the ``true" tree, as the ``best" tree is generally what we have to work with.
 
-\paragraph*{Maximum Likelihood}
+\subsubsection{Maximum Likelihood}
 The ``best" tree and the ``missing-data" trees were inferred using RAxML v.
-8.0.20 \cite{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model \cite{tavare1986} (default GTRGAMMA in RAxML v.
-8.0.20  \cite{Stamatakis21012014}).
-For the morphological data, we used the implemented M\textit{kv} model \cite{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
-8.0.20 \cite{Stamatakis21012014}).
-We used RAxML because it automatically corrects for acquisition bias \cite{lewisa2001}.
+8.0.20 \citep{Stamatakis21012014}.
+For the molecular data, we used the GTR + $\Gamma_4$ model \citep{tavare1986} (default GTRGAMMA in RAxML v.
+8.0.20  \citep{Stamatakis21012014}).
+For the morphological data, we used the implemented M\textit{kv} model \citep{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
+8.0.20 \citep{Stamatakis21012014}).
+We used RAxML because it automatically corrects for acquisition bias \citep{lewisa2001}.
 It is also heavily used in the literature for Maximum Likelihood tree inference (e.g.
-\cite{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014}) and is one of the fastest methods available \cite{Stamatakis01102008}.
+\citealp{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014}) and is one of the fastest methods available \citep{Stamatakis01102008}.
 
 To measure the support for each branch in our simulated phylogenies we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the ``complete" matrix.
-We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \cite{zanderminimal2004}.
+We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
 We repeated this selection until we obtained 50 sets of simulations (i.e.
@@
 This step was implemented to make sure that the differences we observed in topologies (see below) were due to the amount of missing data for each parameter ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
-On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \cite{pattengale2010many}.
+On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \citep{pattengale2010many}.
 Using these parameters took \texttildelow 8 CPU years to build 50 sets of 125 bootstrapped Maximum Likelihood trees (2.30GHz clock speed nodes).
@@
 
-\paragraph*{Bayesian}
+\subsubsection{Bayesian}
 The ``best" tree and the ``missing-data" trees were inferred using MrBayes v.
-3.2.1 \cite{Ronquist2012mrbayes}.
+3.2.1 \citep{Ronquist2012mrbayes}.
 We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
-The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \cite{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
-For the morphological data, we used the M\textit{kv} model \cite{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{kv} + $\Gamma_4$).
-Note that MrBayes automatically corrects for acquisition bias in the morphological data partition \cite{Nylander01022004,Ronquist2012mrbayes}.
+The molecular evolutionary history was inferred using the HKY model with a transition/transversion ratio of two \citep{douadycomparison2003} and a gamma distribution for the rate variation with four distinct categories (HKY + $\Gamma_4$).
+For the morphological data, we used the M\textit{kv} model \citep{lewisa2001}, with equal state frequency and a unique overall substitution rate ($\mu$) with four distinct rates categories (M\textit{kv} + $\Gamma_4$).
+Note that MrBayes automatically corrects for acquisition bias in the morphological data partition \citep{Nylander01022004,Ronquist2012mrbayes}.
 We chose these models to be consistent with the parameters used to generate the ``complete" matrix.
@@
 Even using these priors, it took $~$ 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-The detailed MrBayes parameters are available in the Supporting Information S1.
+The detailed MrBayes parameters are available in Appendix A.
 
-We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \cite{Ronquist2012mrbayes}.
-We also checked the effective sample size (ESS) on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \cite{drummond2006ess}.
-Finally we built a strict majority rule Bayesian consensus tree from the combined chains, excluding the 25\% first iterations as burn-in \cite{Ronquist2012mrbayes}.
+We used the average standard deviation of split frequencies (ASDS) as a proxy to estimate the convergence of the chains and used a stop rule when the ASDS went below 0.01 \citep{Ronquist2012mrbayes}.
+We also checked the effective sample size (ESS) on a random sub-sample of runs in each simulation to ensure that ESS $>>$ 200 \citep{drummond2006ess}.
+Finally we built a strict majority rule Bayesian consensus tree from the combined chains, excluding the 25\% first iterations as burn-in \citep{Ronquist2012mrbayes}.
 
@@
 
-\subsection*{4. Comparing topologies}
+\subsection{Comparing topologies}
 We compared the topology of the ``missing-data" trees to the ``best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
-We used the Robinson-Foulds distance \cite{RF1981} to assess the amount of conserved clade positions and the Triplets distance \cite{dobson1975triplets} to assess the amount of wildcard taxa (i.e.
-taxa that frequently change position in different trees \cite{kearneyfragmentary2002}).
-We used these two metrics because they illustrate two different aspects of tree topology (see Discussion) but also because their performance in measuring differences in topology is well described \cite{Kuhner04112014} and well implemented \cite{Bogdanowicz2012}.
-We normalised both metrics using methods described in \cite{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
+We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa (i.e.
+taxa that frequently change position in different trees \citep{kearneyfragmentary2002}).
+We used these two metrics because they illustrate two different aspects of tree topology (see Discussion) but also because their performance in measuring differences in topology is well described \citep{Kuhner04112014} and well implemented \citep{Bogdanowicz2012}.
+We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
 These metrics are described in detail below.
 
-\paragraph*{Robinson-Foulds distance}
-The Robinson-Foulds distance \cite{RF1981}, or ``path difference", measures the difference between the number of clades and the twice number of shared clades across two trees.
-The metric reflects the distance between the distributions of tips among clades in the two trees \cite{RF1981} (see ~\nameref{S2_Appendix2} for calculation details).
+\subsubsection{Robinson-Foulds distance}
+The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and the twice number of shared clades across two trees.
+The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} (see Appendix B for calculation details).
 This metric is bounded between zero, when the two trees are identical, and $2(n-2)$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
 This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
-We normalised this metric following Bogdanowicz's Normalised Tree Similarity (NTS) method \cite{Bogdanowicz2012}.
-This methods scales any tree comparison metric using the mean distance between 1000 random trees (see ~\nameref{S2_Appendix2} for the calculation details).
-This method is a generalisation of the topological accuracy method \cite{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
+We normalised this metric following Bogdanowicz's Normalised Tree Similarity (NTS) method \citep{Bogdanowicz2012}.
+This methods scales any tree comparison metric using the mean distance between 1000 random trees (see Appendix B for the calculation details).
+This method is a generalisation of the topological accuracy method \citep{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
 Note that once rescaled, the Normalised Robinson-Foulds metric is a measure of similarity, rather than of distance like the original Robinson-Foulds metric.
 
-\paragraph*{Triplets distance}
-The Triplets distance \cite{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees \cite{critchlowthe1996} (see ~\nameref{S2_Appendix2} for calculation details).
+\subsubsection{Triplets distance}
+The Triplets distance \citep{dobson1975triplets} measures the number of sub-trees made up of three taxa that differ between two trees \citep{critchlowthe1996} (see Appendix B for calculation details).
 This metric measures the position of each taxon and clade in relation to its closest neighbours.
@@
 
-\paragraph*{Paired tree comparisons}
+\subsubsection{Paired tree comparisons}
 \label{tree_comparisons}
-For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the ``best" tree and each ``missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \cite{Bogdanowicz2012} resulting in 125 Normalised Robinson-Foulds metrics and 125 Normalised Triplets metric for each tree inference method.
+For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the ``best" tree and each ``missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting in 125 Normalised Robinson-Foulds metrics and 125 Normalised Triplets metric for each tree inference method.
 Also, to take into account the uncertainty of tree inference, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 trees from the posterior tree distribution of the Bayesian analysis for the ``best" trees, and then did the same for the 125 ``missing data" trees (resulting in 1000 ``best" trees and 125$\times$1000 ``missing data" trees).
@@
 We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v.
-3.1 \cite{hdrcde}.
+3.1 \citep{hdrcde}.
 
-\subsection*{Testing the effects of the missing data parameters on topological recovery}
+\subsection{Testing the effects of the missing data parameters on topological recovery}
 Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the ``best" tree topology in a Total Evidence framework.
@@
 
-We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) using the Bhattacharyya Coefficient \cite{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions \cite{Bhattacharyya} (see ~\nameref{S2_Appendix2} for calculation details).
+We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
+The Bhattacharyya Coefficient is the probability of overlap between two distributions \citep{Bhattacharyya} (see Appendix B for calculation details).
 Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
@@
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc.
-(see Fig.~\ref{Fig_Bhattacharyya_Coefficients1} for more details).
+(see Figure 2 for more details).
 This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
@@
 
-\begin{figure}[h]
-\caption{{\bf Bhattacharyya Coefficient calculation outline 1.}
- A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
+%\begin{figure}[h]
+%\caption{{\bf Bhattacharyya Coefficient calculation outline 1.}
+% A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
 $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
@@
 Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
-\label{Fig_Bhattacharyya_Coefficients1}
-\end{figure}
+%\label{Fig_Bhattacharyya_Coefficients1}
+%\end{figure}
 
@@
 Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc.
-(see Fig.~\ref{Fig_Bhattacharyya_Coefficients2} for more details).
+(see Figure 3 for more details).
 Note that this procedure pools results from across all missing data parameter combinations so it results in just six pairwise comparisons.
@@
 
-\begin{figure}[h]
-\caption{{\bf Bhattacharyya Coefficient calculation outline 2.}
- A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
+%\begin{figure}[h]
+%\caption{{\bf Bhattacharyya Coefficient calculation outline 2.}
+ %A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
 $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
@@
 Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.}
-\label{Fig_Bhattacharyya_Coefficients2} 
-\end{figure}
+%\label{Fig_Bhattacharyya_Coefficients2} 
+%\end{figure}
 
@@
 
-\section*{Results}
+\section{Results}
 As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets metric).
-However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
+However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Figure 4 and Figure 5).
 
-\begin{figure}[]
-\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).}
+%\begin{figure}[]
+%\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).}
+ %The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+%\label{Fig_Results-permeth_perparam} 
+%\end{figure}
+
+%\begin{figure}[]
+%\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).}
+ %The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+%\label{Fig_Results-global_perparam}
+%\end{figure}
+
+\subsection{Individual effects of missing data parameters}
+As the amount of missing data increases across all three parameters, our ability to recover the ``best" tree topology decreases (Figure 4).
+The Normalised Robinson-Foulds metric is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
+However, The Normalised Triplets metric is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
+
+\subsection{Combined effect of missing data parameters}
+As expected, our ability to recover the ``best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Figure 5; Tables C2, C3 and C4 in Appendix C).
+Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Figure 6).
+
+%\begin{figure}[]
+%\caption{{\bf The effects of missing data on topological recovery using Bayesian consensus trees.}
+% Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the ``best" tree distribution, calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
+%}
+%\label{Fig_Results-paircomp_within}
+%\end{figure}
+
+Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Figure 6).
+However, the Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; Figure 6A) displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; Figure 6B).
+Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Figure 6A).
+
+For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds metric or Normalised Triplets metric respectively; Table ~\ref{Tab_Results-Difference_methods}).
+However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds metric (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Figure 4 and Figure C2 in Appendix C).
+
+\section{Discussion}
+
+Our results show that the ability to recover the ``best" tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used.
+However, these factors affected topological recovery in different ways and to different extents.
+Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Figure 6).
+Additionally, using Bayesian consensus trees recovered the ``best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Figure 5, Figure 6, Table ~\ref{Tab_Results-Difference_methods}).
+As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters (e.g.
+\citealp{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}).
+
+\subsection{Individual effects of missing data parameters}
+\subsubsection{Missing data for living taxa ($M_{L}$)}
+When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
+Because living taxa still have molecular characters available for phylogenetic inference (see Methods),even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
+However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
+
+\subsubsection{Missing data for fossil taxa ($M_{F}$)}
+When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this also reduces the probability of morphological characters for fossil taxa overlapping with the ones for living taxa.
+This can lead to difficulties for the placement of certain taxa in the tree.
+However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
+decrease of Normalised Triplets metric) with increasing missing data in this parameter, clade conservation (i.e.
+Normalised Robinson-Foulds metric) is still relatively good (mode = 0.72) when the proportion of missing data is high ($M_{F}$ = 75\%).
+
+The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds metric) but greater on the displacement of wildcard taxa (Normalised Triplets metric; Figure 4 and Figure 5).
+This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
+When the fossil taxa have less data (e.g.
+$M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils.
+Therefore a majority consensus position is unlikely to exist (i.e.
+every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
+In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present at a lower taxonomic level (i.e.
+separated with more nodes from the root) but affects the clade conservation less at higher taxonomic level (i.e.
+separated with less nodes from the root).
+Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxa and decrease the Normalised Triplets metric.
+Therefore, the $M_{F}$ parameter is likely to affect the Normalised Robinson-Foulds metric less than the Normalised Triplets metric for the Bayesian consensus trees.
+Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50).
+In other words, the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree, where the fossil will be placed with lower confidence at a higher taxonomic level.
+We argue that using the Bayesian consensus tree topology is preferable because it is more conservative (e.g.
+\citealp{pattinsonphylogeny2014}).
+
+\subsubsection{Missing morphological characters ($M_{C}$)}
+Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the ``best" tree topology.
+We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
+Conversely, removing 10\% of morphological characters ($M_{C}$) genuinely removes 10\% of data in the matrix.
+However, the effect of removing characters on the ability to recover the ``best" tree topology is of the same order of magnitude as for the other two parameters (Figure 4).
+We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.
+
+Additionally, the number of morphological characters determines the size the matrix.
+This can affect our ability to recover the ``best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
+The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \citep{wagner2000}.
+The size of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
+Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
+However, although these matrices are an order of magnitude smaller than some published matrices (e.g.
+\citealp{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices (e.g.
+\citealp{kellymolecular2014, sallam2011craniodental}).
+Therefore, our simulations reflect realistic parameters.
+Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{wrightbayesian2014}.
+However, the use of probabilistic methods (i.e.
+Maximum Likelihood or Bayesian) and the M\textit{kv} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.
+
+\subsection{Combined effect of missing data parameters}
+As expected, when combining the missing data parameters, our ability to recover the ``best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a lot of missing data (i.e.
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Figure 5).
+However, it is important to notice that the effect of each parameter is not additive.
+The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and/or $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6).
+
+\subsection{Effects of tree inference methods}
+Variation in our ability to recover the ``best" tree topology depends heavily on the tree inference method (Figure 4 and Figure 5).
+For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over cladistic methods \citep{wrightbayesian2014} (but see \citealp{spencerefficacy2013}).
+However, this is the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
+Our results show that the topology of the Bayesian consensus tree is always closer to the ``best" tree topology than the one of the Maximum Likelihood tree (Figure 5).
+As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
+This may also be because the ``best" Bayesian consensus trees are not completely resolved, thus will always be more similar to the ``missing data" trees than a completely resolved tree like the ``best" Maximum Likelihood tree.
+However, we minimized the probability of unresolved ``best" trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
+
+It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best" tree topology than Maximum Likelihood and Bayesian consensus trees.
+This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $M_{C}$ = 0\%; Figure 4).
+This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best" trees involved 1000 random pairwise comparisons rather than just one.
+This will obviously add noise to the results for these methods.
+
+\subsection{Practical implications}
+Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
+Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the ``best" tree topology.
+We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections.
+Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries.
+However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils (e.g.
+\citealp{nithe2013}).
+Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the ``best" tree topology (Figure 6).
+Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
+One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}, so that no one individual collects all the data.
+
+Another practical implication of our results regards the tree inference methods.
+Because the Bayesian consensus trees consistently recovered topologies closer to the ``best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \citep{ronquista2012,BEASTmaster}; although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution).
+Note that, although the Bayesian posterior tree distribution performed poorly in recovering the ``best" tree topology, this is due to difficulties comparing distributions of trees (see above).
+Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses (e.g.
+\citealp{jetzthe2012}).
+
+\section{Conclusions}
+Missing data in Total Evidence matrices is not a problem for recovering the ``best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
+When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
+Therefore we advise filling as many gaps in Total Evidence matrices as possible.
+Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.
+Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best" tree topology than the Maximum Likelihood trees.
+Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
+The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
+
+% Do NOT remove this, even if you are not including acknowledgments.
+\section{Acknowledgments}
+Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Nick Matzke, and April Wright for useful comments on our simulation protocol and manuscript.
+Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
+Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.
+This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
+
+\nolinenumbers
+
+\bibliographystyle{sysbio}
+\bibliography{References}
+
+\newpage
+\noindent
+\textbf{Figure captions}\\
+\bigskip
+
+\noindent
+\textbf{Figure 1:} Protocol outline.
+(1) We randomly generated a birth-death tree (the ``true" tree) and used it to simulate a matrix with no missing data (the ``complete" matrix).
+(2) We removed data from the morphological part of the ``complete" matrix resulting in 125 ``missing-data" matrices.
+(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
+(4) We compared the ``missing-data" trees to the ``best" tree.
+We repeated these four steps 50 times.\\
+\bigskip
+
+\noindent
+\textbf{Figure 2:} Bhattacharyya Coefficient calculation outline 1.
+A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.
+Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.\\
+
+\noindent
+\textbf{Figure 3:} Bhattacharyya Coefficient calculation outline 2.
+A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+\textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
+Maximum Likelihood or Bayesian).
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$.
+Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.\\
+
+\bigskip
+\noindent
+\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).
 The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-\label{Fig_Results-permeth_perparam} 
-\end{figure}
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
 
-\begin{figure}[]
-\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).}
+\bigskip
+\noindent
+\textbf{Figure 5:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
 The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-\label{Fig_Results-global_perparam}
-\end{figure}
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
 
-\subsection*{Individual effects of missing data parameters}
-As the amount of missing data increases across all three parameters, our ability to recover the ``best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
-The Normalised Robinson-Foulds metric is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7 in ~\nameref{S3_Appendix3}).
-However, The Normalised Triplets metric is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; Tables S5, S6 and S7 in ~\nameref{S3_Appendix3}).
-
-\subsection*{Combined effect of missing data parameters}
-As expected, our ability to recover the ``best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
-$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; Tables S2, S3 and S4 in ~\nameref{S3_Appendix3}).
-Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).
-
-\begin{figure}[]
-\caption{{\bf The effects of missing data on topological recovery using Bayesian consensus trees.}
+\bigskip
+\noindent
+\textbf{Figure 6:} The effects of missing data on topological recovery using Bayesian consensus trees.
 Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
@@
 The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
-}
-\label{Fig_Results-paircomp_within}
-\end{figure}
 
-Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Fig.~\ref{Fig_Results-paircomp_within}).
-However, the Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; Fig.~\ref{Fig_Results-paircomp_within}A) displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; Fig.~\ref{Fig_Results-paircomp_within}B).
-Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Fig.~\ref{Fig_Results-paircomp_within}A).
-Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-paircomp_within}A).
+\newpage
+\begin{landscape}
 
-For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds metric or Normalised Triplets metric respectively; Table ~\ref{Tab_Results-Difference_methods}).
-However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds metric (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.
-S2 in ~\nameref{S3_Appendix3}).
+\noindent
+\textbf{Tables}\\
+\bigskip
 
-\begin{table}[!ht]
-\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
-\caption{{\bf Bhattacharyya Coefficients of the pairwise method comparisons.}
+\noindent
+\textbf{Table 1:} Bhattacharyya Coefficients of the pairwise method comparisons.
 Each comparison which corresponds to the normalised metric between the ``best" tree and the ``missing data" trees using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).
@@
 The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
-If two methods have a high probability of overlap, they have a similar ability to recover the ``correct" tree topology.}
+If two methods have a high probability of overlap, they have a similar ability to recover the ``correct" tree topology.\\
+
+\begin{table}[!ht]
+%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+\caption{}
 \centering
@@
 \label{Tab_Results-Difference_methods}
-\end{adjustwidth}
+%\end{adjustwidth}
 \end{table}
-
-\section*{Discussion}
-
-Our results show that the ability to recover the ``best" tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used.
-However, these factors affected topological recovery in different ways and to different extents.
-Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}).
-Additionally, using Bayesian consensus trees recovered the ``best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam}, Fig.~\ref{Fig_Results-paircomp_within}, Table ~\ref{Tab_Results-Difference_methods}).
-As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters (e.g.
-\cite{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}).
-
-\subsection*{Individual effects of missing data parameters}
-\paragraph*{Missing data for living taxa ($M_{L}$)}
-When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
-Because living taxa still have molecular characters available for phylogenetic inference (see Methods),even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
-However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
-
-\paragraph*{Missing data for fossil taxa ($M_{F}$)}
-When the overall proportion of data for the fossil taxa ($M_{F}$) decreases, this also reduces the probability of morphological characters for fossil taxa overlapping with the ones for living taxa.
-This can lead to difficulties for the placement of certain taxa in the tree.
-However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
-decrease of Normalised Triplets metric) with increasing missing data in this parameter, clade conservation (i.e.
-Normalised Robinson-Foulds metric) is still relatively good (mode = 0.72) when the proportion of missing data is high ($M_{F}$ = 75\%).
-
-The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds metric) but greater on the displacement of wildcard taxa (Normalised Triplets metric; Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
-When the fossil taxa have less data (e.g.
-$M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils.
-Therefore a majority consensus position is unlikely to exist (i.e.
-every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
-In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present at a lower taxonomic level (i.e.
-separated with more nodes from the root) but affects the clade conservation less at higher taxonomic level (i.e.
-separated with less nodes from the root).
-Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxa and decrease the Normalised Triplets metric.
-Therefore, the $M_{F}$ parameter is likely to affect the Normalised Robinson-Foulds metric less than the Normalised Triplets metric for the Bayesian consensus trees.
-Conversely, the same scenario in a Maximum Likelihood framework will lead to a dichotomous branching of the fossils but with low bootstrap support ($<$ 50).
-In other words, the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree, where the fossil will be placed with lower confidence at a higher taxonomic level.
-We argue that using the Bayesian consensus tree topology is preferable because it is more conservative (e.g.
-\cite{pattinsonphylogeny2014}).
-
-\paragraph*{Missing morphological characters ($M_{C}$)}
-Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the ``best" tree topology.
-We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
-Conversely, removing 10\% of morphological characters ($M_{C}$) genuinely removes 10\% of data in the matrix.
-However, the effect of removing characters on the ability to recover the ``best" tree topology is of the same order of magnitude as for the other two parameters (Fig.~\ref{Fig_Results-permeth_perparam}).
-We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.
-
-Additionally, the number of morphological characters determines the size the matrix.
-This can affect our ability to recover the ``best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
-The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \cite{wagner2000}.
-The size of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
-Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-However, although these matrices are an order of magnitude smaller than some published matrices (e.g.
-\cite{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices (e.g.
-\cite{kellymolecular2014, sallam2011craniodental}).
-Therefore, our simulations reflect realistic parameters.
-Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \cite{wrightbayesian2014}.
-However, the use of probabilistic methods (i.e.
-Maximum Likelihood or Bayesian) and the M\textit{kv} model \cite{lewisa2001} has been previously demonstrated to partially resolve this issue \cite{wrightbayesian2014}.
-
-\subsection*{Combined effect of missing data parameters}
-As expected, when combining the missing data parameters, our ability to recover the ``best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a lot of missing data (i.e.
-$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-global_perparam}).
-However, it is important to notice that the effect of each parameter is not additive.
-The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and/or $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}).
-
-\subsection*{Effects of tree inference methods}
-Variation in our ability to recover the ``best" tree topology depends heavily on the tree inference method (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}).
-For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \cite{lewisa2001} over cladistic methods \cite{wrightbayesian2014} (but see \cite{spencerefficacy2013}).
-However, this is the first study, to our knowledge, to compare the performance of the M\textit{kv} model \cite{lewisa2001} for recovering the ``best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-Our results show that the topology of the Bayesian consensus tree is always closer to the ``best" tree topology than the one of the Maximum Likelihood tree (Fig.~\ref{Fig_Results-global_perparam}).
-As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
-This may also be because the ``best" Bayesian consensus trees are not completely resolved, thus will always be more similar to the ``missing data" trees than a completely resolved tree like the ``best" Maximum Likelihood tree.
-However, we minimized the probability of unresolved ``best" trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
-
-It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best" tree topology than Maximum Likelihood and Bayesian consensus trees.
-This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-permeth_perparam}).
-This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best" trees involved 1000 random pairwise comparisons rather than just one.
-This will obviously add noise to the results for these methods.
-
-\subsection*{Practical implications}
-Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
-Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the ``best" tree topology.
-We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections.
-Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries.
-However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils (e.g.
-\cite{nithe2013}).
-Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the ``best" tree topology (Fig.~\ref{Fig_Results-paircomp_within}).
-Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
-One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{morphobank} \cite{morphobank}, so that no one individual collects all the data.
-
-Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the ``best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \cite{ronquista2012,BEASTmaster}; although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution).
-Note that, although the Bayesian posterior tree distribution performed poorly in recovering the ``best" tree topology, this is due to difficulties comparing distributions of trees (see above).
-Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses (e.g.
-\cite{jetzthe2012}).
-
-\section*{Conclusion}
-Missing data in Total Evidence matrices is not a problem for recovering the ``best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
-When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
-Therefore we advise filling as many gaps in Total Evidence matrices as possible.
-Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.
-Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best" tree topology than the Maximum Likelihood trees.
-Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
-The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
-
-\section*{Supporting Information}
-
-% Include only the SI item label in the subsection heading. Use the \nameref{label} command to cite SI items in the text.
-
-\subsection*{Data reproducibility and availability}
-\noindent{\bf Data reproducibility and availability.} All the code used in this analysis is available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{\color{blue}{GitHub}} with some information on how to use the various functions.
-Additionally all the simulated data is available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{\color{blue}{FigShare}}.
-
-\subsection*{Supporting Information S1: Tree Inference}
-\label{S1_Appendix1}
-\noindent{\bf Supporting Information S1 Tree Inference.} Supplementary information on the Tree Inference methods. Contains Figure S1.
-
-\subsection*{Supporting Information S2: Tree Comparisons}
-\label{S2_Appendix2}
-\noindent{\bf Supporting Information S2 Tree Comparisons.} Supplementary information on the Tree Comparison methods.
-
-\subsection*{Supporting Information S3: Additional results}
-\label{S3_Appendix3}
-\noindent{\bf Supporting Information S3 Additional results.}. Contains Figures S2, S3, S4 and S5 and Tables S1, S2, S3, S4 S5, S6 and S7.
-
-% Do NOT remove this, even if you are not including acknowledgments.
-\section*{Acknowledgments}
-Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Nick Matzke, and April Wright for useful comments on our simulation protocol and manuscript.
-Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
-Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.
-%This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).
-
-\nolinenumbers
-
-\bibliographystyle{plos2009}
-\bibliography{References}
+\end{landscape}
 

### 2015-06-22 Manuscript/MPESubmission/OnlineAppendices-LaTeXSuppFiles/AppendixD.tex
MPE submission

--- 2015-05-12 Manuscript/PLoSONEresponse/OnlineAppendices-LaTeXSuppFiles/SupportingInformation3.tex
@@
 
+\renewcommand\thefigure{C.\arabic{figure}}
+\renewcommand\thetable{C.\arabic{table}}
+
 \begin{document}
 
-\section{Supporting Information S3: Additional Results}
-
-The following section contains the supplemental results of the effects of our missing data parameters and the different tree inference methods on the the ability to recover the "best" topology that are briefly discussed in the main body of the paper.
-For clarity, in the paper, we focused on the results of the effects of our missing data parameters on the Maximum Likelihood trees topology and the Bayesian consensus trees topology.
-The following additional results presented here give a greater insight into the effect of our missing data parameters on the Maximum Likelihood Bootstrapped trees topologies and the Bayesian posterior trees distribution topologies.
-Also we present here the pairwise comparisons between each parameters states for the Maximum Likelihood tree topology, the Maximum Likelihood Bootstrapped trees topologies and the Bayesian posterior trees distribution topologies.
-
-\begin{figure} 
-\centering
-    \includegraphics[width=1\textwidth]{SupplementaryFigures/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
-    \caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey).
-The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
-\label{Fig_Supp_BootBayt_allparam} 
-\end{figure}
-
-\begin{figure} 
-\centering
-     %Hide this figure for 'fast' build
-%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-ML-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
-    \caption{The effects of missing data on topological recovery using Maximum Likelihood trees.
-The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
-\label{Fig_Supp_paircomp_ML}
-\end{figure}
-
-\begin{figure} 
-\centering
-     %Hide this figure for 'fast' build
-%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-Boot-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
-    \caption{The effects of missing data on topological recovery using Maximum Likelihood bootstrap trees.
-The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
-\label{Fig_Supp_paircomp_Boot}
-\end{figure}
-
-\begin{figure} 
-\centering
-     %Hide this figure for 'fast' build
-%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PPairwiseComp-Bayt-RF+Tr-colour.pdf}
-    \includegraphics[width=1\textwidth]{SupplementaryFigures/PairwiseComp-Bayt-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
-    \caption{The effects of missing data on topological recovery using Bayesian posterior tree distribution.
-The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
-TThe Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.}
-\label{Fig_Supp_paircomp_Bayt}
-\end{figure}
-
-%Summary of metrics values for all parameters combinations
-\begin{landscape}
-\begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).}
-\label{Tab_Supp_summary_metric_allparam}
-\centering
-\begin{tabular}{lccccccc}
-  \hline
- Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
-  \hline
-  Maximum Likelihood                    & $RF$ & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
-                                        & $Tr$ & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
-  Bayesian consensus                    & $RF$ & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
-                                        & $Tr$ & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
-  Maximum Likelihood bootstraps         & $RF$ & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
-                                        & $Tr$ & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
-  Bayesian posterior tree distributions & $RF$ & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
-                                        & $Tr$ & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
-   \hline
-\end{tabular}
-\end{table}
-\end{landscape}
-
-%Summary of metrics values for the ML parameter
-\begin{landscape}
-\begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
-\label{Tab_Supp_summary_metric_ML}
-\centering
-\begin{tabular}{lccccccc}
-  \hline
- Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
-  \hline
-  Maximum Likelihood                    & $RF$ & 0.44 & 0.51 & 0.63 & 0.66 & 0.78 & 0.95 \\ 
-                                        & $Tr$ & 0.45 & 0.56 & 0.76 & 0.74 & 0.93 & 0.99 \\ 
-  Bayesian consensus                    & $RF$ & 0.71 & 0.73 & 0.80 & 0.82 & 0.88 & 0.95 \\ 
-                                        & $Tr$ & 0.37 & 0.46 & 0.67 & 0.67 & 0.87 & 0.96 \\ 
-  Maximum Likelihood bootstraps         & $RF$ & 0.34 & 0.37 & 0.42 & 0.41 & 0.44 & 0.46 \\ 
-                                        & $Tr$ & 0.32 & 0.40 & 0.51 & 0.46 & 0.51 & 0.57 \\ 
-  Bayesian posterior tree distributions & $RF$ & 0.33 & 0.41 & 0.52 & 0.50 & 0.60 & 0.65 \\ 
-                                        & $Tr$ & 0.41 & 0.56 & 0.76 & 0.71 & 0.84 & 0.98 \\ 
-   \hline
-\end{tabular}
-\end{table}
-\end{landscape}
-
-%Summary of metrics values for the MF parameter
-\begin{landscape}
-\begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
-\label{Tab_Supp_summary_metric_MF}
-\centering
-\begin{tabular}{lccccccc}
-  \hline
- Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
-  \hline
-  Maximum Likelihood                    & $RF$ & 0.23 & 0.46 & 0.64 & 0.61 & 0.79 & 0.93 \\ 
-                                        & $Tr$ & 0.65 & 0.84 & 0.95 & 0.89 & 0.99 & 1.00 \\ 
-  Bayesian consensus                    & $RF$ & 0.72 & 0.77 & 0.86 & 0.85 & 0.94 & 0.96 \\ 
-                                        & $Tr$ & -0.16 & 0.19 & 0.63 & 0.52 & 0.96 & 0.98 \\ 
-  Maximum Likelihood bootstraps         & $RF$ & 0.14 & 0.30 & 0.40 & 0.35 & 0.45 & 0.46 \\ 
-                                        & $Tr$ & 0.37 & 0.49 & 0.54 & 0.51 & 0.56 & 0.57 \\ 
-  Bayesian posterior tree distributions & $RF$ & 0.24 & 0.45 & 0.57 & 0.51 & 0.63 & 0.65 \\ 
-                                        & $Tr$ & 0.44 & 0.81 & 0.86 & 0.82 & 0.98 & 0.98 \\ 
-   \hline
-\end{tabular}
-\end{table}
-\end{landscape}
-
-%Summary of metrics values for the MC parameter
-\begin{landscape}
-\begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
-\label{Tab_Supp_summary_metric_MC}
-\centering
-\begin{tabular}{lccccccc}
-  \hline
- Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
-  \hline
-  Maximum Likelihood                    & $RF$ & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
-                                        & $Tr$ & 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
-  Bayesian consensus                    & $RF$ & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
-                                        & $Tr$ & 0.05 & 0.16 & 0.53 & 0.50 & 0.87 & 0.92 \\ 
-  Maximum Likelihood bootstraps         & $RF$ & 0.25 & 0.34 & 0.42 & 0.38 & 0.45 & 0.46 \\ 
-                                        & $Tr$ & 0.38 & 0.47 & 0.55 & 0.51 & 0.57 & 0.58 \\ 
-  Bayesian posterior tree distributions & $RF$ & 0.32 & 0.44 & 0.58 & 0.52 & 0.62 & 0.65 \\ 
-                                        & $Tr$ & 0.39 & 0.78 & 0.82 & 0.79 & 0.98 & 0.98 \\ 
-   \hline
-\end{tabular}
-\end{table}
-\end{landscape}
-
-%Summary of the BC between pairs of methods for the ML parameter
-\begin{landscape}
-\begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
-\label{Tab_Supp_summary_BC_ML}
-\centering
-\begin{tabular}{lccccccc}
-  \hline
- Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
-  \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.30 & 0.31 & 0.69 & 0.61 & 0.77 & 1.00 \\ 
-                                                                       & $Tr$ & 0.79 & 0.81 & 0.84 & 0.86 & 0.85 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.22 & 0.29 & 0.36 & 0.54 & 0.69 \\ 
-                                                                       & $Tr$ & 0.08 & 0.42 & 0.53 & 0.51 & 0.74 & 0.78 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.49 & 0.61 & 0.51 & 0.67 & 0.74 \\ 
-                                                                       & $Tr$ & 0.21 & 0.61 & 0.70 & 0.63 & 0.81 & 0.81 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.01 & 0.02 & 0.02 & 0.02 & 0.03 & 0.04 \\ 
-                                                                       & $Tr$ & 0.08 & 0.69 & 0.78 & 0.64 & 0.79 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.01 & 0.02 & 0.02 & 0.04 & 0.08 & 0.09 \\ 
-                                                                       & $Tr$ & 0.21 & 0.74 & 0.75 & 0.68 & 0.84 & 0.87 \\ 
-    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.75 & 0.85 & 0.85 & 0.95 & 1.00 \\ 
-                                                                       & $Tr$ & 0.91 & 0.92 & 0.96 & 0.95 & 0.97 & 0.98 \\ 
-   \hline
-\end{tabular}
-\end{table}
-\end{landscape}
-
-%Summary of the BC between pairs of methods for the MF parameter
-\begin{landscape}
-\begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
-\label{Tab_Supp_summary_BC_MF}
-\centering
-\begin{tabular}{lccccccc}
-  \hline
- Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\  
-  \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.00 & 0.25 & 0.48 & 0.50 & 0.76 & 1.00 \\ 
-                                                                       & $Tr$ & 0.38 & 0.69 & 0.75 & 0.72 & 0.80 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.18 & 0.32 & 0.36 & 0.47 & 0.77 \\ 
-                                                                       & $Tr$ & 0.08 & 0.34 & 0.40 & 0.38 & 0.53 & 0.55 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.47 & 0.71 & 0.60 & 0.86 & 0.94 \\ 
-                                                                       & $Tr$ & 0.21 & 0.54 & 0.62 & 0.56 & 0.64 & 0.80 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.00 & 0.01 & 0.01 & 0.01 & 0.03 \\ 
-                                                                       & $Tr$ & 0.08 & 0.38 & 0.54 & 0.49 & 0.70 & 0.75 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.02 & 0.02 & 0.02 & 0.04 & 0.04 \\ 
-                                                                       & $Tr$ & 0.21 & 0.29 & 0.66 & 0.54 & 0.72 & 0.82 \\ 
-    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.69 & 0.72 & 0.71 & 0.72 & 0.72 \\ 
-                                                                       & $Tr$ & 0.91 & 0.91 & 0.91 & 0.93 & 0.92 & 0.98 \\ 
-   \hline
-\end{tabular}
-\end{table}
-\end{landscape}
-
-%Summary of the BC between pairs of methods for the MC parameter
-\begin{landscape}
-\begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
-\label{Tab_Supp_summary_BC_MC}
-\centering
-\begin{tabular}{lccccccc}
-  \hline
- Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\  
-  \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.03 & 0.32 & 0.66 & 0.55 & 0.75 & 1.00 \\ 
-                                                                       & $Tr$ & 0.51 & 0.69 & 0.80 & 0.76 & 0.80 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.17 & 0.21 & 0.31 & 0.46 & 0.68 \\ 
-                                                                       & $Tr$ & 0.08 & 0.31 & 0.39 & 0.39 & 0.56 & 0.61 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.44 & 0.47 & 0.52 & 0.78 & 0.90 \\ 
-                                                                       & $Tr$ & 0.21 & 0.52 & 0.59 & 0.55 & 0.66 & 0.77 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.01 & 0.01 & 0.02 & 0.02 & 0.03 \\ 
-                                                                       & $Tr$ & 0.08 & 0.47 & 0.62 & 0.51 & 0.66 & 0.73 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.02 & 0.04 & 0.04 & 0.05 & 0.06 \\ 
-                                                                       & $Tr$ & 0.21 & 0.45 & 0.64 & 0.57 & 0.74 & 0.79 \\ 
-    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.73 & 0.73 & 0.76 & 0.81 & 0.86 \\ 
-                                                                       & $Tr$ & 0.91 & 0.92 & 0.93 & 0.94 & 0.96 & 0.99 \\ 
-   \hline
-\end{tabular}
-\end{table}
-\end{landscape}
+\section{Data reproducibility and availability}
+\noindent{\bf Data reproducibility and availability.} All the code used in this analysis is available on \href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{GitHub} (\href{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}{
+goo.gl/4djNUf}) with some information on how to use the various functions.
+Additionally all the simulated data is available on \href{http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861}{FigShare} (dx.doi.org/10.6084/m9.figshare.1306861).
 %END

### 2015-06-22 Analysis/MissingDataEffect.R
removed path

@@ -1,5 +1,5 @@
 #Script for testing the effect of missing data on tree topology.
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
+setwd('Analysis/')
 
 #Loading the functions
 sourceDir <- function(path, trace = TRUE, ...) {
@@ -14,8 +14,8 @@ library(ape)
 
 #Loading the Random tree comparisons
 Random51<-read.table("../Data/Tree_Comparisons/RandomTrees/51t_Random.Cmp",header=TRUE)
-Random26<-read.table("../Data/Tree_Comparisons/RandomTrees/26t_Random.Cmp",header=TRUE)
-Random25<-read.table("../Data/Tree_Comparisons/RandomTrees/25t_Random.Cmp",header=TRUE)
+#Random26<-read.table("../Data/Tree_Comparisons/RandomTrees/26t_Random.Cmp",header=TRUE)
+#Random25<-read.table("../Data/Tree_Comparisons/RandomTrees/25t_Random.Cmp",header=TRUE)
 
 #Loading the treesets
 

### 2015-06-22 Manuscript/Talk_abstract.tex
Talk abstract (correct number of words)

\documentclass[a4paper,11pt]{article}

\usepackage{natbib}
\usepackage{enumerate}
\usepackage{lastpage} 
\usepackage[utf8]{inputenc}
\usepackage{times}
\pagenumbering{arabic}
\linespread{1.5}
\begin{document}
\fontsize{12pt}{1.5}
\begin{flushright}
Version dated: \today
\end{flushright}
\begin{center}

%Font style Times New Roman. Font size 12. Title in bold. Maximum word limit 300 which includes name and affiliation details.

%Title
\noindent{\Large{\bf{Missing data and topology in Total Evidence method}}}\\
\bigskip
%Author
\noindent{Thomas Guillerme - guillert@tcd.ie}\\
\noindent{Zoology Department, Trinity College Dublin, Ireland}\\

\end{center}

To understand macroevolutionary patterns and processes, we need to include extant and extinct species in our models. This requires phylogenetic trees with living and fossil taxa at the tips. One way to infer such phylogenies is the Total Evidence approach which uses molecular data from living taxa and morphological data from living and fossil taxa.

Although the Total Evidence approach is very promising, it requires a great deal of data that can be hard to collect. Therefore this method would likely suffer from missing data issues that may affect its ability to infer correct phylogenies.

We use simulations to assess the effects of missing data on tree topologies inferred from Total Evidence matrices. We investigate three factors that affect the completeness and the size of the morphological part of the matrix: the proportion of living taxa with no morphological data, the amount of missing data in the fossil record, and the overall number of morphological characters in the matrix. We infer phylogenies from complete matrices and from matrices with various amounts of missing data, and then compare missing data topologies to the ``best" tree topology inferred using the complete matrix.

We find that the number of living taxa with morphological characters and the number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the ``best" tree topology. We suggest that sampling effort should be focused on morphological data collection for living species to increase the accuracy of topological inference in a Total Evidence framework. Additionally, we find that Bayesian methods consistently outperform other tree inference methods. We recommend using Bayesian consensus trees to fix the tree topology prior to further analyses.

\end{document}

### 2015-06-28 Manuscript/MPESubmission/Revision/response.tex
Revision init

\documentclass[12pt,letterpaper]{article}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\pagenumbering{arabic}

%Pagination style and stuff
\linespread{2}
\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}

\begin{document}

\section{Reviewer 1:}

\begin{enumerate}
\item{\textbf{Novelty:}} The novelty of this study should be spelled out more clearly by more a more thorough review of the literature, not just by referring to a few relevant works, but by pointing out what is so new about this study.
The authors assert that this is the first study to examine the effects of missing morphological data from a total evidence dataset, but that is what Pattinson et al. 2014 did (as well as Wiens 2005).
This study examines it in a different way with a study design that is unique and adds to the field. For example, one feature of this study that is unique is that a tree with fossils is simulated and the data simulated on that tree.
In other studies of missing data, data were removed from living taxa to determine if their positions could still be inferred (Pattinson et al. 2015; Wiens and Tiu 2012).
This study will have bearing on an old assertion that including fossils specifically may help alleviate long-branch attraction, not only additional taxa with varying amounts of missing data, as suggested by others (e.g., Wiens 2005).
Similarly, in the discussion, the authors need to point out how their results fit in with the results of previous research more clearly.

\item{\textbf{Graphical abstract:}} Current graphical abstract does not capture the content of the article for readers in a single glance.
The authors should not use red and green together.
Figure 5 illustrates topology distance with different inference methods and varying proportions of missing data in rows vs columns and would make a better graphical abstract.
For the graphical abstract, Fig 5 needs to be made clearer in one glance what is going on.
More details are given in the comments on that figure below.

\item{\textbf{Highlights:}} wording seems awkward

\item{\textbf{Introduction:}} Needs stronger emphasis on what specifically is unique about this study compared to what the many other studies on missing data have done and found.
Clearly outlining the questions in this study that are left open from previous research is important because some of the findings in this study suggest that there may be more error associated with missing data than previous research has suggested.

\item{\textbf{Abbreviation:}} One suggestion is to consider abbreviating Maximum Likelihood and Bayesian Inference as ML and BI respectively early in the methods and use throughout.

\item{\textbf{Methodology - True tree:}} I suggest it warrants comparing at least the 'best' tree to the 'true' tree.

\item{\textbf{Methodology - Starting tree:}} The authors should at least include some analysis in which no starting tree is given, the max. likelihood 'best' tree is given, and / or the 'best' tree is given but with some perturbation.
I fear that one of the main results may be biased due to this one methodological detail - a strong prior on tree topology.
The justification given was 'to speed up the Bayesian analysis' which does not seem adequate, given these datasets are small.

\item{\textbf{Results - Bhattacharyya meaning:}} Some descriptive statistics are given, but it would help to tell us if these differences are meaningful.
For example, on line 414 the Bhattacharrya Coefficients are given suggesting RF distances are lower for ML than BI trees at 0.69, 0.48, and 0.66 - but if this coefficient represents the probability of overlap between distributions then shouldn't the coefficients be <0.05 to say that one is lower than the other?
In the table, mean and median values of <0.01 and >0.85 are highlighted as being the most extreme.
Later the focus on Bhattacharyya Coefficients is 1 when distributions overlap completely and 0 when they do not - should this be the criterion?
Just make it clear early on so the readers knows what to think when they see these numbers.

\item{\textbf{Results - Bayesian consensus vs. Bayesian posterior:}} One major finding that was not fully addressed was that the distances between the BI posterior distributions of trees do not overlap the BI best or consensus trees.
It seems from the tables in Appendix C that the topology distance metrics for Bayesian posterior distribution are similar to those for the Bayesian consensus, albeit usually lower for RF distance than the Tr distance.
In Figure 4, it seems like the distributions of metrics based on CIs for Bayesian consensus and posterior overlap up until >25\% missing data.
How then in comparisons of the distributions are they always so low (Table 1 max prob. of overlapping distributions = 0.11)?

\item{\textbf{Results - Bayesian Triplets:}} Another important result that is not mentioned is that using the wildcard taxon metric (Tr), Bayesian consensus trees do quite poorly with missing data.
Values in Figures 4 and 5 clearly drop below zero, although negative values are not illustrated on the axis, and this indicates that the tree topologies are no better than, or even worse than, random trees.
It is especially true when fossils are missing data, and that is going to be an important finding with practical application but must be addressed.

\item{\textbf{Conclusion:}} The authors should clarify how their results compare to those of previous studies more directly to emphasize the nuances that make this study unique.

\item{\textbf{Figure 4:}} Figure 4 has a color scheme issue, the colors in the caption do not match up to the figure.

\item{\textbf{Figure 4-5:}} Am I wrong or shouldn't the y-axis for Normalised Triplets metric go below 0 - it looks those metrics for many missing data combinations have values below zero indicating trees are more different from the 'best' tree than expected by chance (Appendix B 1.1., see Appendix C Tables 1-3) and that is important to show.

\item{\textbf{Table caption:}} Table caption should include a line about 'pooling across missing data schemes' to clarify what is presented.

\item{\textbf{Previous research:}} The missing-data problem has a long history in the literature and there are references to some of those older and recent papers.
But I feel it is necessary to summarize the following questions: what did those studies find, what are the questions those studies left open, and how does this study address them?
\end{enumerate}

\section{Reviewer 2:}

\begin{enumerate}
\item{\textbf{Compare the trees to the "true" tree:}} I don't see how the "best" tree approach is helpful in this case. It adds an additional layer of complexity that can't be accounted for via simulation. The point of simulations is to compare empirical realities to the generating process, to see how much results diverge. I think the most helpful and straightforward thing to do is eliminate the "best" tree part of the MS, and compare the inferred trees directly to the real topology to quantify variance.
\item{\textbf{Better explanation of the missing data:}} I found the explanation of the missing data strategy very difficult to follow. First, they oddly use the verb "to infer" for matrices; are they inferring matrices? Second, it's not clear anywhere that the authors actually removed the molecular data from the fossil taxa, though I assume they did. I hope the authors can re-write in simpler language ("explain like I'm 5") for easier comprehension.
\item{\textbf{$M_C$ is not missing data per se:}} If Mc is just removing columns, then they are shrinking the total matrix. This is not missing data per se. They are instead comparing matrices of different sizes, and the quantity is "data present." Any matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large. A matrix with 10,000 characters is tiny if 10,000,000 could be scored. This is semantic.
\item{\textbf{Specific variation of data per fossil:}} More importantly, it doesn't seem like missing data varied among fossil taxa, unless I'm misreading. What we really want to know is what happens when some fossil taxa are fragmentary, as this is the most common empirical situation. Thus, there are morphological characters scored for some living taxa, some fossil taxa, and missing from others. It's unclear if this was simulated. It would be good if the authors could detail the case where, say, all living taxa had morphological data (0 Ml), the matrix was large (0 Mc), and Mf varied *among* fossil taxa, such that some fossils had 100\% coverage, and varied down to, say, 5\%. That's the crucial case here.
\end{enumerate}

\section{Specific comments:}
\begin{enumerate}
\item{\textbf{line 31:}} has it really been shown that the majority of macroevolutionary studies focuses on extant species only? The cited references are examples of large scale extant-only phylogenies. To make such a pointed statement, the authors would need to find support from some kind of literature review or meta-analysis comparing the number of paleontological to neontological macroevolutionary studies. Otherwise, rephrase. 
\item{\textbf{line 37 and throughout:}} Rephrase so that 'however' does not start the sentence, such as 'To do this, however, we need…'. Sentences should not start with however if the meaning is to contrast with the previous statement (Strunk, W. and White, E.B. 2009. The Elements of Style. NY: Pearson Education. p 48-49).
\item{\textbf{line 43:}} sentence choppy, consider rephrasing or splitting into two sentences.
\item{\textbf{line 44:}} three alternatives: age only, morph only, or age and morph jointly. 
\item{\textbf{line 47:}} Simpson 1944 doesn't seem the appropriate reference for maximum parsimony. Consider (Felsenstein 2004; Hennig 1966) 
\item{\textbf{line 53:}} omit 'are used as nodes rather than tips in these phylogenies and their'. Fossils aren't really used as nodes in node calibration, the second part of the sentence is more accurate.
\item{\textbf{line 66:}} omit 'have been successfully applied to empirical data' and just say 'becoming increasingly popular…list of references'. 
\item{\textbf{line 76 and throughout the manuscript:}} watch out for inconsistent parentheses.
\item{\textbf{line 76 and throughout:}} correct use of the word 'data' from singular to plural throughout
\item{\textbf{line 77:}} change 'a lot' to 'large proportions of' or something more accurate and less colloquial
\item{\textbf{line 82:}} change to '…missing data are not phylogenetically biased, Wiens…'
\item{\textbf{line 99:}} unclear - what is meant by 'with a fixed topology approach'? The studies references investigated divergence times using total evidence datasets and in some cases used constraints but not completely fixed topologies
\item{\textbf{lines 80-103:}} the authors need to clearly justify the importance of this study compared the many studies that have investigated the effects of missing data on topology inference. The justification currently given is not particularly strong and the specific unique questions asked in this study that are currently unanswered should be made clear. The current argument does not always seem justified by the literature review. For example, line 81 the authors argue that papers by Wiens and Pattinson examined morphological or molecular matrices separately, but in Pattinson, for example, combined datasets for living species were used to simulate extinct species by deleting the molecular and parts of the morphological datasets from living species. The authors could argue that a limitation of the Pattinson study, which they address here, was the limited scheme for generating missing data from the datasets. In this study, the authors perturbed the datasets more thoroughly in a clearly structured experimental design to investigate the effects of eliminating living versus fossil data or all data overall. That is one aspect of this study that I perceived as new and unique. Another aspect of this study design that could be new but is not explored here is the effect of the missing data on the morphological substitution rate parameter estimates. \textbf{Can you recover the true values from the perturbed datasets?} This study also used a unique way of evaluating the tree inference under missing data scenarios. These and other aspects of the study uniqueness must be clearer.
\item{\textbf{line 124:}} replace 'sections' with 'partition'
\item{\textbf{line 125:}} omit 'note that we explain each step in detail below this general outline', replace with something like 'Our general protocol included the following steps (Figure 1):'…
\item{\textbf{line 167 and throughout manuscript:}} replace usage of 'inferred' when referring to simulating data. Replace with 'generated', 'simulated', or 'created', for example. 'Infer' should only be used when a model is fit to data to infer some output such as a tree, transition rates, etc. 
\item{\textbf{line 173:}} streamline to "The substitution rates were selected from a gamma distribution with…" or something similar. 
\item{\textbf{line 173 to 176:}} need some justification for why the substitution rate with alpha shape of 0.5  will avoid homoplasy. Yang 1996 is cited but clarify the justification of what is 'too much' homoplasy. 
\item{\textbf{Line 176 to 180:}} could be streamlined. It is not clear what is meant by 'no special assumption about how the characters evolved', what constitutes a special assumption? Could potentially be reduced to something like 'this model and these parameter settings strike a balance between realism for empirical datasets (citations) and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations'. Just for example.
\item{\textbf{line 182:}} give the ape function used
\item{\textbf{line 185 to 188:}} streamline, e.g. '…sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A).'
\item{\textbf{line 193 to 196:}} again, justify why gamma with shape of 0.5 is appropriate for minimizing homoplasy. This distribution places highest probability on substation rates less than 0.5, when one of the cited references potentially suggests that the lowest topological error was found at rates of ~0.5-1.5, and that very low substitution rates incurred error similar to very high rates ((Wright and Hillis 2014) 2014 Figure 3). Other cited references do not clearly state that this distribution and shape setting is optimal, so the authors really must justify. 
\item{\textbf{line 201:}} change 'get' for 'obtain' or something less colloquial. 
\item{\textbf{line 231-235:}} streamline, e.g., "The minimal dataset included in this study was 5% of the morphological matrix for any taxon".
\item{\textbf{line 237-246:}} I understand the rational presented by the authors for comparing missing data trees to the best tree inferred from max. likelihood/Bayesian techniques rather than compare missing data trees to the true tree used to simulate the data. I still feel it would be an important contribution to this paper, however, to compare the missing data trees to the true tree to evaluate the effects of missing data on the tree inference. Also it may lend insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it. 
\item{\textbf{line 249:}} condense citations to one parenthetical statement, 'GTR model (Tavare, 1986, default setting in RAx…)'. Also watch for citations in parentheses inside of other parenthetical statements on line 250 and throughout the manuscript. 
\item{\textbf{line 251:}} omit 'implemented'
\item{\textbf{line 273:}} 'Bayesian Inference'
\item{\textbf{line 286:}} Why was the true tree used as a starting tree? Earlier the authors argued that the true tree is almost never known, so using it here detracts from the generalizability of these results. If the only justification is 'to speed up the Bayesian estimation process' as stated later, then it seems like this may bias one of the main results, which is that Bayesian analyses outperformed max. likelihood, in which no starting tree was used. I understand the computational limitations, but the authors should present some analysis with complete and missing data in which no starting tree is given compared to analyses in which a starting tree is given to evaluate whether using a starting tree significantly affects the resulting trees, perhaps in an Appendix. Or at least include some perturbation to the starting tree to avoid overly influencing the resulting topology. It would seem to me that this is illustrated in figure 5, where-in with the Robinson-Foulds metric, the distance to the best tree in Bayesian analysis seems to have a lower bound at ~0.7, even under the worst cases of missing data. This may be the worst the analysis can do given that it started out with the true tree. 
\item{\textbf{line 318:}} replace 'the twice number' with 'twice the number'
\item{\textbf{line 327:}} 'This methods' omit 's'
\item{\textbf{line 343:}} 'this metric is sensitive' add 'is'
\item{\textbf{line 424:}} 'Figure ??' should be Figure 6, but then do not repeat Figure 6 in the parentheses. 
\item{\textbf{line 436:}} 'effect'
\item{\textbf{line 489:}} Streamline to something like: '…when the fossil is present near the tips but affects the clade conservation less when fossils are near the root'. 
\item{\textbf{line 494:}} 'taxon'
\item{\textbf{line 515:}} '…the size of the matrix' missing 'of'
\item{\textbf{line 518 to 520:}} I am not sure that (Wagner 2000) states that incongruence between molecular and morphological data is 'more important in small morphological matrices'. 
\item{\textbf{line 520:}} Either change 'size' to plural or change 'were' to 'was', the former seems to flow better. 
\item{\textbf{line 528:}} missing close parenthesis after Ni et al.
\item{\textbf{line 532:}} Wright and Hillis 2014 found that increasing the size of the dataset improves topological accuracy, and they do not argue that simply increasing the number of characters increases the amount of homoplasy. Rather they found that the fastest evolving characters exhibited homoplasy, while the slowest evolving lacked phylogenetic signal. The citation seems misrepresented. Further, there was higher phylogenetic accuracy of simulated datasets with more characters than fewer characters, for slow or fast evolving characters (Wiens 2005).
\item{\textbf{line 540:}} replace 'a lot' with 'larger proportions of missing data', a lot seems too colloquial for this kind of article. 
\item{\textbf{lines 542-547:}} These seem like the most important practical results and should be more strongly emphasized. These are very important results for practitioners and the authors should try to make it more clear to the reader that this is a key take-home message.
\item{\textbf{line 552:}} replace 'cladistic' with 'parsimony'. While one has come to be associated with the other in the literature, they are not synonymous. 
\item{\textbf{line 557-558:}} '…than the "best" maximum likelihood tree'
\item{\textbf{line 549-565}}: again, I am left wondering how much the differences in performance of max. likelihood and Bayesian inference techniques are driven by the use of the 'true' tree as a starting tree in the Bayesian MCMC search, while the likelihood analysis had no such guide tree to begin with. The authors at least need to make this point clear, if not re-run some subset of analyses without introducing a starting tree to test the effects of using the 'true' tree as a starting tree or not. I understand the computational limitations, but the practical limitation that empiricists will never have the 'true' tree to start with may (or may not) have a large impact on our interpretation and the generality of these findings. 
\item{\textbf{line 593:}} 'MorphoBank' - the 'b' should be capitalized in keeping with the authors' usage (O'Leary and Kaufman 2011)
\item{\textbf{line 599:}} remove extra close parenthesis inside parenthetical statement
\item{\textbf{line 595-606:}} this paragraph seems unsatisfactory advice for practitioners. First, advising that the tree topology be fixed using the Bayesian consensus tree before conducting analyses such as dating strongly conditions the results on only a single tree and as the authors point out in the parenthetical statement, the dating information could improve accuracy. Indeed one of the many advantages of tip-dating is the joint inference of tree topology and divergence times. The following statements that the posterior distribution should not be discarded does not leave the reader with a clear path that they should follow - if only the fixed tree topology is used to estimate divergence times, how can the substitution rate-scaled posterior distribution of highly probable trees be used for most comparative analyses which require time-scaled branch lengths? This paragraph requires further justification and this study does have important bearing on a subject that gets little attention. 
\item{\textbf{line 612-613:}} too obvious and not especially helpful. We should aim to have as little missing data as possible…more helpful, and something especially discovered in this study, is concentrating on having overlapping morphological characters between the living and fossil taxa to be able to place the living taxa with strong support via large datasets (combined data) and being able to position the fossils based on their shared derived characters with living taxa. This is an important point to make clearer. 
\item{\textbf{line 613-615:}} rephrase - sentence is choppy with all the commas. 
\item{\textbf{line 615-618:}} It seems more important to me that the authors emphasize what implications it may have for previous and future studies that Bayesian analyses outperform max. likelihood, other than to re-hash that the authors feel the Bayesian consensus should be used to fix topologies for other downstream analyses. For example, can the authors cite some combined analyses that used only likelihood analyses and came up with controversial results? Or what about other studies that have done combined analyses with different optimality criteria and found different topologies? What about implications for studies that have used molecular scaffold techniques to fix the topology of living taxa based on analyses of only DNA analyses and then fit the fossils into this fixed tree? Also 'further' is misspelled.  
\item{\textbf{line 628:}} use of computer cluster may be better placed in the methods, depending on the norm for MPE.

\item{\textbf{Figure 3:}} Do not repeat the aside to similarity with t-test again from Fig 2 and text.
\item{\textbf{Figure 4:}} color codes in legend do not match up to the colors in the PDF I received. I don't know if this is an error in conversion, but the colors are black, red, green and blue in my PDF. The legend states the following: 'Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue)' which doesn't seem correct since blue is repeated. This makes interpreting the figure difficult for me at this time, but I assume given the text in the Results that the Bayesian methods are black and red in the figure I received, indicating they were closer to the best tree than max likelihood trees. Red and green should not be used together.
\item{\textbf{Figure 5:}} This figure best summarizes the results, and a version of this should be the Graphical Abstract, with legends for the color cods and spelling out the missing data parameters, and the Y axis could just be 'clade similarity index' or 'rouge taxon index' so that the reader doesn't need to know what the names of the metrics are to grasp the graphic. But concerning the Normalized Triplets metric, according to the text and appendix, if this index is 0, the similarity of the missing data tree to the best tree is no more than random, and < 0 it is more different than random. For the Bayesian analyses, many results had values < 0 but the axis is set to 0. Why not illustrate the distribution of values that is < 0? From this result, it appears the Bayesian analysis does quite poorly with missing data in terms of rouge taxa, and worse than max. likelihood. This figure also suggests to me that the methods perform pretty badly with missing data, in contrast to previous findings that missing data are not overly deleterious. For example, when all data are complete except Mf=50\%, the normalized RF metric is <0.8. Missing data for up to 50\% of characters has been suggested to still have high accuracy in uncovering the true tree (Wiens 2003; Wiens 2005), although for small datasets such as the one used in this study, the results show a similar decline in accuracy. This is a point that should be expanded upon in the discussion - how these results compare to other missing data studies and what do they tell us about the performance of the inference methods with missing data?
\item{\textbf{Figure 6:}} I think this may be the most important figure because it is demonstrating the probability of overlap between the distributions of the best trees and the missing data trees. It took me a long time to figure it out, however, and to figure out why the top of the triangle is all orange. I would have expected something more like in Appendix C Figure C2 or Fig. C3. At first I understood this to indicate high probability of overlap of the missing data trees which had the most missing data with the best tree. But after re-reading the legend and methods several times, I am interpreting it as follows: the top of the triangle indicates a similarly poor fit of the distributions of missing data trees to the best tree under all parameter combinations once there is a high proportion of missing data. The point is raised in the Results that once missing data in the living portion is > 50\% then they all perform poorly, but it was not easy to grasp. Somehow this figure needs to be spelled out better in the caption and results so the reader can grasp it more quickly and easily without having to go back to the methods section to completely figure out what is being represented.

\item{\textbf{Table 1:}} The wording of the caption is off - the second paragraph is awkward. 
Are these values the probabilities of overlap between the distributions of the 'best' tree versus trees from each inference method pooled across all combinations of missing data? That is what is described in methods line 397 but it doesn't say so in the caption. Two things:

1)  Comparing to the tables in Appendix C, the distribution of probabilities of overlap between max. likelihood and Bayesian trees are higher in most missing data schemes in tables in the Appendices than the overall values given in Table 1. How could the max. likelihood and Bayesian trees be so different?

2)  Pooling across all missing data schemes seems to wash out the effect of tree-inference method. I am not sure what a better way to present these would be, but perhaps some kind of matrix where the cells are the joint probabilities of overlap of each method with each missing data category, and then the marginal probabilities would be the probabilities of overlap for each method, averaged over missing data categories, as well as the marginal probabilities of overlap for each missing data category, averaged across methods. 
e.g, RF distance:

Missing data:   0   25  50  75  Marginal \\
BI  1   0.9 0.8 0.75    0.863 \\
ML  1   0.9 0.75    0.25    0.725 \\
Marginal    1   0.9 0.78    0.5  \\

This may not be feasible in the text due to space limitations, maybe tables like these could be added to an appendix.

\item{\textbf{Appendix C:}} Tables 1-3 repeats 'of the'. Should label the first two rows as 'Best maximum likelihood tree' and 'Best Bayesian inference tree'
\item{\textbf{Appendix D:}} I find the online code repository thorough, but the repository is loaded with many files, especially for graphing. If it is possible to condense these files it may be easier for a reader to duplicate the study.

\end{enumerate}

\end{document}

### 2015-06-28 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixB.tex
Revision init

--- 2015-06-08 Manuscript/MPESubmission/OnlineAppendices-LaTeXSuppFiles/SupportingInformation2.tex
@@
 %Packages
+\usepackage{xcolor}
+\usepackage{changepage}
 \usepackage{pdflscape}
@@
 \usepackage{fullpage}
+\usepackage{natbib}
 \usepackage{float}
@@
 \usepackage{lineno}
+\usepackage{longtable}
 \pagenumbering{arabic}
@@
 \setlength{\parindent}{0.5in}
-\setcounter{secnumdepth}{0} 
-\renewcommand{\section}[1]{%
-\bigskip
-\begin{center}
-\begin{Large}
-\normalfont\scshape #1
-\medskip
-\end{Large}
-\end{center}}
-\renewcommand{\subsection}[1]{%
-\bigskip
-\begin{center}
-\begin{large}
-\normalfont\itshape #1
-\end{large}
-\end{center}}
-\renewcommand{\subsubsection}[1]{%
-\vspace{2ex}
-\noindent
-\textit{#1.}---}
-\renewcommand{\tableofcontents}{}
-
+%\setcounter{secnumdepth}{0} 
 \begin{document}
 
-\section{Supporting Information S2: Tree Comparisons}
+\section{Appendix B: Tree Comparisons}
 
@@
 \subsection{Normalised Tree Similarity}
-For any tree with \textit{n} taxa compared using a tree difference metric $m$, Normalized Tree Similarity, $NTS_m$ \cite{Bogdanowicz2012}, represents the similarity score for the two trees given the expected difference between two random Yule trees \cite{Bogdanowicz2012} with $n$ taxa.
+For any tree with \textit{n} taxa compared using a tree difference metric $m$, Normalized Tree Similarity, $NTS_m$ \citep{Bogdanowicz2012}, represents the similarity score for the two trees given the expected difference between two random Yule trees \citep{Bogdanowicz2012} with $n$ taxa.
 If $\bar{d}_{m,n}$\textit{(rand)} is the average difference between two random Yule trees with $n$ taxa and $d_{m,n}$\textit{(x,y)} the difference between the two trees \textit{x} and \textit{y} each containing $n$ taxa, then:
@@
 
-We used the NTS method to scale all the Robinson Foulds and Triplets metrics calculated in our analyses, using the TreeCmp JavaScript \cite{Bogdanowicz2012}.
+We used the NTS method to scale all the Robinson Foulds and Triplets metrics calculated in our analyses, using the TreeCmp JavaScript \citep{Bogdanowicz2012}.
 
 \subsection{Bhattacharyya Coefficient}
-The Bhattacharyya Coefficient calculates the probability of overlap of two distributions \cite{Bhattacharyya}.
+The Bhattacharyya Coefficient calculates the probability of overlap of two distributions \citep{Bhattacharyya}.
 When it is equal to zero, the probability of overlap of the distributions is also zero, and when it is equal to one, the two distributions are entirely overlapping.
@@
 If $n$ is low, the overlap will be overestimated and if $n$ is too high, the overlap will be underestimated.
-In this analysis, we determined the number of bins using Silverman's rule of thumb which states that $n$ should be 0.9 times the minimum of the standard deviation and the interquartile range of the distribution, divided by 1.34 times the sample size of the distribution to the negative one-fifth power (\texttt{bw.nrd0()} function in R \cite{silverman1986density}).
+In this analysis, we determined the number of bins using Silverman's rule of thumb which states that $n$ should be 0.9 times the minimum of the standard deviation and the interquartile range of the distribution, divided by 1.34 times the sample size of the distribution to the negative one-fifth power (\texttt{bw.nrd0()} function in R \citep{silverman1986density}).
 
-\bibliographystyle{plos2009}
+\bibliographystyle{sysbio}
 \bibliography{Supp_Ref}

### 2015-06-29 Manuscript/MPESubmission/Revision/response.tex
Revision init

@@
 \usepackage{lineno}
+\usepackage{color}
+\usepackage[usenames,dvipsnames]{xcolor}
 \pagenumbering{arabic}
@@
 \begin{enumerate}
-\item{\textbf{Novelty:}} The novelty of this study should be spelled out more clearly by more a more thorough review of the literature, not just by referring to a few relevant works, but by pointing out what is so new about this study.
+\item{\textcolor{blue}{\textbf{Novelty:} The novelty of this study should be spelled out more clearly by more a more thorough review of the literature, not just by referring to a few relevant works, but by pointing out what is so new about this study.
 The authors assert that this is the first study to examine the effects of missing morphological data from a total evidence dataset, but that is what Pattinson et al. 2014 did (as well as Wiens 2005).
@@
 This study will have bearing on an old assertion that including fossils specifically may help alleviate long-branch attraction, not only additional taxa with varying amounts of missing data, as suggested by others (e.g., Wiens 2005).
-Similarly, in the discussion, the authors need to point out how their results fit in with the results of previous research more clearly.
+Similarly, in the discussion, the authors need to point out how their results fit in with the results of previous research more clearly. }}
 
-\item{\textbf{Graphical abstract:}} Current graphical abstract does not capture the content of the article for readers in a single glance.
+\item{\textcolor{blue}{\textbf{Graphical abstract:} Current graphical abstract does not capture the content of the article for readers in a single glance.
 The authors should not use red and green together.
@@
 For the graphical abstract, Fig 5 needs to be made clearer in one glance what is going on.
-More details are given in the comments on that figure below.
+More details are given in the comments on that figure below.}}
 
-\item{\textbf{Highlights:}} wording seems awkward
+\item{\textcolor{blue}{\textbf{Highlights:} wording seems awkward}}
 
-\item{\textbf{Introduction:}} Needs stronger emphasis on what specifically is unique about this study compared to what the many other studies on missing data have done and found.
-Clearly outlining the questions in this study that are left open from previous research is important because some of the findings in this study suggest that there may be more error associated with missing data than previous research has suggested.
+\item{\textcolor{blue}{\textbf{Introduction:} Needs stronger emphasis on what specifically is unique about this study compared to what the many other studies on missing data have done and found.
+Clearly outlining the questions in this study that are left open from previous research is important because some of the findings in this study suggest that there may be more error associated with missing data than previous research has suggested.}}
 
-\item{\textbf{Abbreviation:}} One suggestion is to consider abbreviating Maximum Likelihood and Bayesian Inference as ML and BI respectively early in the methods and use throughout.
+\item{\textcolor{blue}{\textbf{Abbreviation:} One suggestion is to consider abbreviating Maximum Likelihood and Bayesian Inference as ML and BI respectively early in the methods and use throughout.}}
 
-\item{\textbf{Methodology - True tree:}} I suggest it warrants comparing at least the 'best' tree to the 'true' tree.
+\item{\textcolor{blue}{\textbf{Methodology - True tree:} I suggest it warrants comparing at least the 'best' tree to the 'true' tree.}}
 
-\item{\textbf{Methodology - Starting tree:}} The authors should at least include some analysis in which no starting tree is given, the max.
+\item{\textcolor{blue}{\textbf{Methodology - Starting tree:} The authors should at least include some analysis in which no starting tree is given, the max.
 likelihood 'best' tree is given, and / or the 'best' tree is given but with some perturbation.
 I fear that one of the main results may be biased due to this one methodological detail - a strong prior on tree topology.
-The justification given was 'to speed up the Bayesian analysis' which does not seem adequate, given these datasets are small.
+The justification given was 'to speed up the Bayesian analysis' which does not seem adequate, given these datasets are small.}}
 
-\item{\textbf{Results - Bhattacharyya meaning:}} Some descriptive statistics are given, but it would help to tell us if these differences are meaningful.
+\item{\textcolor{blue}{\textbf{Results - Bhattacharyya meaning:} Some descriptive statistics are given, but it would help to tell us if these differences are meaningful.
 For example, on line 414 the Bhattacharrya Coefficients are given suggesting RF distances are lower for ML than BI trees at 0.69, 0.48, and 0.66 - but if this coefficient represents the probability of overlap between distributions then shouldn't the coefficients be <0.05 to say that one is lower than the other?
@@
 Later the focus on Bhattacharyya Coefficients is 1 when distributions overlap completely and 0 when they do not - should this be the criterion?
-Just make it clear early on so the readers knows what to think when they see these numbers.
+Just make it clear early on so the readers knows what to think when they see these numbers. }}
 
-\item{\textbf{Results - Bayesian consensus vs.
-Bayesian posterior:}} One major finding that was not fully addressed was that the distances between the BI posterior distributions of trees do not overlap the BI best or consensus trees.
+\item{\textcolor{blue}{\textbf{Results - Bayesian consensus vs.
+Bayesian posterior:} One major finding that was not fully addressed was that the distances between the BI posterior distributions of trees do not overlap the BI best or consensus trees.
 It seems from the tables in Appendix C that the topology distance metrics for Bayesian posterior distribution are similar to those for the Bayesian consensus, albeit usually lower for RF distance than the Tr distance.
 In Figure 4, it seems like the distributions of metrics based on CIs for Bayesian consensus and posterior overlap up until >25\% missing data.
-How then in comparisons of the distributions are they always so low (Table 1 max prob. of overlapping distributions = 0.11)?
+How then in comparisons of the distributions are they always so low (Table 1 max prob. of overlapping distributions = 0.11)?}}
 
-\item{\textbf{Results - Bayesian Triplets:}} Another important result that is not mentioned is that using the wildcard taxon metric (Tr), Bayesian consensus trees do quite poorly with missing data.
+\item{\textcolor{blue}{\textbf{Results - Bayesian Triplets:} Another important result that is not mentioned is that using the wildcard taxon metric (Tr), Bayesian consensus trees do quite poorly with missing data.
 Values in Figures 4 and 5 clearly drop below zero, although negative values are not illustrated on the axis, and this indicates that the tree topologies are no better than, or even worse than, random trees.
-It is especially true when fossils are missing data, and that is going to be an important finding with practical application but must be addressed.
+It is especially true when fossils are missing data, and that is going to be an important finding with practical application but must be addressed. }}
 
-\item{\textbf{Conclusion:}} The authors should clarify how their results compare to those of previous studies more directly to emphasize the nuances that make this study unique.
+\item{\textcolor{blue}{\textbf{Conclusion:} The authors should clarify how their results compare to those of previous studies more directly to emphasize the nuances that make this study unique. }}
 
-\item{\textbf{Figure 4:}} Figure 4 has a color scheme issue, the colors in the caption do not match up to the figure.
+\item{\textcolor{blue}{\textbf{Figure 4:} Figure 4 has a color scheme issue, the colors in the caption do not match up to the figure.}}
 
-\item{\textbf{Figure 4-5:}} Am I wrong or shouldn't the y-axis for Normalised Triplets metric go below 0 - it looks those metrics for many missing data combinations have values below zero indicating trees are more different from the 'best' tree than expected by chance (Appendix B 1.1., see Appendix C Tables 1-3) and that is important to show.
+\item{\textcolor{blue}{\textbf{Figure 4-5:} Am I wrong or shouldn't the y-axis for Normalised Triplets metric go below 0 - it looks those metrics for many missing data combinations have values below zero indicating trees are more different from the 'best' tree than expected by chance (Appendix B 1.1., see Appendix C Tables 1-3) and that is important to show.}}
 
-\item{\textbf{Table caption:}} Table caption should include a line about 'pooling across missing data schemes' to clarify what is presented.
+\item{\textcolor{blue}{\textbf{Table caption:} Table caption should include a line about 'pooling across missing data schemes' to clarify what is presented.}}
 
-\item{\textbf{Previous research:}} The missing-data problem has a long history in the literature and there are references to some of those older and recent papers.
-But I feel it is necessary to summarize the following questions: what did those studies find, what are the questions those studies left open, and how does this study address them?
+\item{\textcolor{blue}{\textbf{Previous research:} The missing-data problem has a long history in the literature and there are references to some of those older and recent papers.
+But I feel it is necessary to summarize the following questions: what did those studies find, what are the questions those studies left open, and how does this study address them?}}
 \end{enumerate}
@@
 \begin{enumerate}
-\item{\textbf{Compare the trees to the "true" tree:}} I don't see how the "best" tree approach is helpful in this case.
+\item{\textcolor{blue}{\textbf{Compare the trees to the "true" tree:} I don't see how the "best" tree approach is helpful in this case.
 It adds an additional layer of complexity that can't be accounted for via simulation.
 The point of simulations is to compare empirical realities to the generating process, to see how much results diverge.
-I think the most helpful and straightforward thing to do is eliminate the "best" tree part of the MS, and compare the inferred trees directly to the real topology to quantify variance.
-\item{\textbf{Better explanation of the missing data:}} I found the explanation of the missing data strategy very difficult to follow.
+I think the most helpful and straightforward thing to do is eliminate the "best" tree part of the MS, and compare the inferred trees directly to the real topology to quantify variance.}}
+
+\item{\textcolor{blue}{\textbf{Better explanation of the missing data:} I found the explanation of the missing data strategy very difficult to follow.
 First, they oddly use the verb "to infer" for matrices; are they inferring matrices?
 Second, it's not clear anywhere that the authors actually removed the molecular data from the fossil taxa, though I assume they did.
-I hope the authors can re-write in simpler language ("explain like I'm 5") for easier comprehension.
-\item{\textbf{$M_C$ is not missing data per se:}} If Mc is just removing columns, then they are shrinking the total matrix.
+I hope the authors can re-write in simpler language ("explain like I'm 5") for easier comprehension.}}
+
+\item{\textcolor{blue}{\textbf{$M_C$ is not missing data per se:} If Mc is just removing columns, then they are shrinking the total matrix.
 This is not missing data per se.
@@
 A matrix with 10,000 characters is tiny if 10,000,000 could be scored.
-This is semantic.
-\item{\textbf{Specific variation of data per fossil:}} More importantly, it doesn't seem like missing data varied among fossil taxa, unless I'm misreading.
+This is semantic.}}
+
+\item{\textcolor{blue}{\textbf{Specific variation of data per fossil:} More importantly, it doesn't seem like missing data varied among fossil taxa, unless I'm misreading.
 What we really want to know is what happens when some fossil taxa are fragmentary, as this is the most common empirical situation.
@@
 It would be good if the authors could detail the case where, say, all living taxa had morphological data (0 Ml), the matrix was large (0 Mc), and Mf varied *among* fossil taxa, such that some fossils had 100\% coverage, and varied down to, say, 5\%.
-That's the crucial case here.
+That's the crucial case here.}}
 \end{enumerate}
@@
 \begin{enumerate}
-\item{\textbf{line 31:}} has it really been shown that the majority of macroevolutionary studies focuses on extant species only?
+
+\item{\textcolor{blue}{\textbf{line 31:} has it really been shown that the majority of macroevolutionary studies focuses on extant species only?
 The cited references are examples of large scale extant-only phylogenies.
@@
 Otherwise, rephrase.
-\item{\textbf{line 37 and throughout:}} Rephrase so that 'however' does not start the sentence, such as 'To do this, however, we need…'.
+}}
+\item{\textcolor{blue}{\textbf{line 37 and throughout:} Rephrase so that 'however' does not start the sentence, such as 'To do this, however, we need…'.
 Sentences should not start with however if the meaning is to contrast with the previous statement (Strunk, W.
@@
 NY: Pearson Education.
-p 48-49).
-\item{\textbf{line 43:}} sentence choppy, consider rephrasing or splitting into two sentences.
-\item{\textbf{line 44:}} three alternatives: age only, morph only, or age and morph jointly. 
-\item{\textbf{line 47:}} Simpson 1944 doesn't seem the appropriate reference for maximum parsimony. Consider (Felsenstein 2004; Hennig 1966) 
-\item{\textbf{line 53:}} omit 'are used as nodes rather than tips in these phylogenies and their'.
-Fossils aren't really used as nodes in node calibration, the second part of the sentence is more accurate.
-\item{\textbf{line 66:}} omit 'have been successfully applied to empirical data' and just say 'becoming increasingly popular…list of references'. 
-\item{\textbf{line 76 and throughout the manuscript:}} watch out for inconsistent parentheses.
-\item{\textbf{line 76 and throughout:}} correct use of the word 'data' from singular to plural throughout
-\item{\textbf{line 77:}} change 'a lot' to 'large proportions of' or something more accurate and less colloquial
-\item{\textbf{line 82:}} change to '…missing data are not phylogenetically biased, Wiens…'
-\item{\textbf{line 99:}} unclear - what is meant by 'with a fixed topology approach'?
-The studies references investigated divergence times using total evidence datasets and in some cases used constraints but not completely fixed topologies
-\item{\textbf{lines 80-103:}} the authors need to clearly justify the importance of this study compared the many studies that have investigated the effects of missing data on topology inference.
+p 48-49).}}
+\item{\textcolor{blue}{\textbf{line 43:} sentence choppy, consider rephrasing or splitting into two sentences.}}
+\item{\textcolor{blue}{\textbf{line 44:} three alternatives: age only, morph only, or age and morph jointly. }}
+\item{\textcolor{blue}{\textbf{line 47:} Simpson 1944 doesn't seem the appropriate reference for maximum parsimony. Consider (Felsenstein 2004; Hennig 1966) }}
+\item{\textcolor{blue}{\textbf{line 53:} omit 'are used as nodes rather than tips in these phylogenies and their'.
+Fossils aren't really used as nodes in node calibration, the second part of the sentence is more accurate.}}
+\item{\textcolor{blue}{\textbf{line 66:} omit 'have been successfully applied to empirical data' and just say 'becoming increasingly popular…list of references'. }}
+\item{\textcolor{blue}{\textbf{line 76 and throughout the manuscript:} watch out for inconsistent parentheses.}}
+\item{\textcolor{blue}{\textbf{line 76 and throughout:} correct use of the word 'data' from singular to plural throughout}}
+\item{\textcolor{blue}{\textbf{line 77:} change 'a lot' to 'large proportions of' or something more accurate and less colloquial}}
+\item{\textcolor{blue}{\textbf{line 82:} change to '…missing data are not phylogenetically biased, Wiens…'}}
+\item{\textcolor{blue}{\textbf{line 99:} unclear - what is meant by 'with a fixed topology approach'?
+The studies references investigated divergence times using total evidence datasets and in some cases used constraints but not completely fixed topologies}}
+\item{\textcolor{blue}{\textbf{lines 80-103:} the authors need to clearly justify the importance of this study compared the many studies that have investigated the effects of missing data on topology inference.
 The justification currently given is not particularly strong and the specific unique questions asked in this study that are currently unanswered should be made clear.
@@
 That is one aspect of this study that I perceived as new and unique.
-Another aspect of this study design that could be new but is not explored here is the effect of the missing data on the morphological substitution rate parameter estimates.
-\textbf{Can you recover the true values from the perturbed datasets?} This study also used a unique way of evaluating the tree inference under missing data scenarios.
-These and other aspects of the study uniqueness must be clearer.
-\item{\textbf{line 124:}} replace 'sections' with 'partition'
-\item{\textbf{line 125:}} omit 'note that we explain each step in detail below this general outline', replace with something like 'Our general protocol included the following steps (Figure 1):'…
-\item{\textbf{line 167 and throughout manuscript:}} replace usage of 'inferred' when referring to simulating data.
+Another aspect of this study design that could be new but is not explored here is the effect of the missing data on the morphological substitution rate parameter estimates.\textbf{Can you recover the true values from the perturbed datasets?} This study also used a unique way of evaluating the tree inference under missing data scenarios.
+These and other aspects of the study uniqueness must be clearer.}}
+\item{\textcolor{blue}{\textbf{line 124:} replace 'sections' with 'partition'}}
+\item{\textcolor{blue}{\textbf{line 125:} omit 'note that we explain each step in detail below this general outline', replace with something like 'Our general protocol included the following steps (Figure 1):'…}}
+\item{\textcolor{blue}{\textbf{line 167 and throughout manuscript:} replace usage of 'inferred' when referring to simulating data.
 Replace with 'generated', 'simulated', or 'created', for example.
 'Infer' should only be used when a model is fit to data to infer some output such as a tree, transition rates, etc.
-\item{\textbf{line 173:}} streamline to "The substitution rates were selected from a gamma distribution with…" or something similar. 
-\item{\textbf{line 173 to 176:}} need some justification for why the substitution rate with alpha shape of 0.5  will avoid homoplasy.
+}}
+\item{\textcolor{blue}{\textbf{line 173:} streamline to "The substitution rates were selected from a gamma distribution with…" or something similar. }}
+\item{\textcolor{blue}{\textbf{line 173 to 176:} need some justification for why the substitution rate with alpha shape of 0.5  will avoid homoplasy.
 Yang 1996 is cited but clarify the justification of what is 'too much' homoplasy.
-\item{\textbf{Line 176 to 180:}} could be streamlined.
+}}
+\item{\textcolor{blue}{\textbf{Line 176 to 180:} could be streamlined.
 It is not clear what is meant by 'no special assumption about how the characters evolved', what constitutes a special assumption?
 Could potentially be reduced to something like 'this model and these parameter settings strike a balance between realism for empirical datasets (citations) and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations'.
-Just for example.
-\item{\textbf{line 182:}} give the ape function used
-\item{\textbf{line 185 to 188:}} streamline, e.g.
-'…sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A).'
-\item{\textbf{line 193 to 196:}} again, justify why gamma with shape of 0.5 is appropriate for minimizing homoplasy.
+Just for example.}}
+\item{\textcolor{blue}{\textbf{line 182:} give the ape function used}}
+\item{\textcolor{blue}{\textbf{line 185 to 188:} streamline, e.g.
+'…sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A).'}}
+\item{\textcolor{blue}{\textbf{line 193 to 196:} again, justify why gamma with shape of 0.5 is appropriate for minimizing homoplasy.
 This distribution places highest probability on substation rates less than 0.5, when one of the cited references potentially suggests that the lowest topological error was found at rates of ~0.5-1.5, and that very low substitution rates incurred error similar to very high rates ((Wright and Hillis 2014) 2014 Figure 3).
 Other cited references do not clearly state that this distribution and shape setting is optimal, so the authors really must justify.
-\item{\textbf{line 201:}} change 'get' for 'obtain' or something less colloquial. 
-\item{\textbf{line 231-235:}} streamline, e.g., "The minimal dataset included in this study was 5% of the morphological matrix for any taxon".
-\item{\textbf{line 237-246:}} I understand the rational presented by the authors for comparing missing data trees to the best tree inferred from max.
+}}
+\item{\textcolor{blue}{\textbf{line 201:} change 'get' for 'obtain' or something less colloquial.}}
+\item{\textcolor{blue}{\textbf{line 231-235:} streamline, e.g., 'The minimal dataset included in this study was 5\% of the morphological matrix for any taxon'.}}
+\item{\textcolor{blue}{\textbf{line 237-246:} I understand the rational presented by the authors for comparing missing data trees to the best tree inferred from max.
 likelihood/Bayesian techniques rather than compare missing data trees to the true tree used to simulate the data.
@@
 Also it may lend insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it.
-\item{\textbf{line 249:}} condense citations to one parenthetical statement, 'GTR model (Tavare, 1986, default setting in RAx…)'.
+}}
+\item{\textcolor{blue}{\textbf{line 249:} condense citations to one parenthetical statement, 'GTR model (Tavare, 1986, default setting in RAx…)'.
 Also watch for citations in parentheses inside of other parenthetical statements on line 250 and throughout the manuscript.
-\item{\textbf{line 251:}} omit 'implemented'
-\item{\textbf{line 273:}} 'Bayesian Inference'
-\item{\textbf{line 286:}} Why was the true tree used as a starting tree?
+}}
+\item{\textcolor{blue}{\textbf{line 251:} omit 'implemented'}}
+\item{\textcolor{blue}{\textbf{line 273:} 'Bayesian Inference'}}
+\item{\textcolor{blue}{\textbf{line 286:} Why was the true tree used as a starting tree?
 Earlier the authors argued that the true tree is almost never known, so using it here detracts from the generalizability of these results.
@@
 This may be the worst the analysis can do given that it started out with the true tree.
-\item{\textbf{line 318:}} replace 'the twice number' with 'twice the number'
-\item{\textbf{line 327:}} 'This methods' omit 's'
-\item{\textbf{line 343:}} 'this metric is sensitive' add 'is'
-\item{\textbf{line 424:}} 'Figure ??' should be Figure 6, but then do not repeat Figure 6 in the parentheses. 
-\item{\textbf{line 436:}} 'effect'
-\item{\textbf{line 489:}} Streamline to something like: '…when the fossil is present near the tips but affects the clade conservation less when fossils are near the root'. 
-\item{\textbf{line 494:}} 'taxon'
-\item{\textbf{line 515:}} '…the size of the matrix' missing 'of'
-\item{\textbf{line 518 to 520:}} I am not sure that (Wagner 2000) states that incongruence between molecular and morphological data is 'more important in small morphological matrices'. 
-\item{\textbf{line 520:}} Either change 'size' to plural or change 'were' to 'was', the former seems to flow better. 
-\item{\textbf{line 528:}} missing close parenthesis after Ni et al.
-\item{\textbf{line 532:}} Wright and Hillis 2014 found that increasing the size of the dataset improves topological accuracy, and they do not argue that simply increasing the number of characters increases the amount of homoplasy.
+}}
+\item{\textcolor{blue}{\textbf{line 318:} replace 'the twice number' with 'twice the number'}}
+\item{\textcolor{blue}{\textbf{line 327:} 'This methods' omit 's'}}
+\item{\textcolor{blue}{\textbf{line 343:} 'this metric is sensitive' add 'is'}}
+\item{\textcolor{blue}{\textbf{line 424:} 'Figure ??' should be Figure 6, but then do not repeat Figure 6 in the parentheses. }}
+\item{\textcolor{blue}{\textbf{line 436:} 'effect'}}
+\item{\textcolor{blue}{\textbf{line 489:} Streamline to something like: '…when the fossil is present near the tips but affects the clade conservation less when fossils are near the root'. }}
+\item{\textcolor{blue}{\textbf{line 494:} 'taxon'}}
+\item{\textcolor{blue}{\textbf{line 515:} '…the size of the matrix' missing 'of'}}
+\item{\textcolor{blue}{\textbf{line 518 to 520:} I am not sure that (Wagner 2000) states that incongruence between molecular and morphological data is 'more important in small morphological matrices'.
+}}
+\item{\textcolor{blue}{\textbf{line 520:} Either change 'size' to plural or change 'were' to 'was', the former seems to flow better. }}
+\item{\textcolor{blue}{\textbf{line 528:} missing close parenthesis after Ni et al.}}
+\item{\textcolor{blue}{\textbf{line 532:} Wright and Hillis 2014 found that increasing the size of the dataset improves topological accuracy, and they do not argue that simply increasing the number of characters increases the amount of homoplasy.
 Rather they found that the fastest evolving characters exhibited homoplasy, while the slowest evolving lacked phylogenetic signal.
 The citation seems misrepresented.
-Further, there was higher phylogenetic accuracy of simulated datasets with more characters than fewer characters, for slow or fast evolving characters (Wiens 2005).
-\item{\textbf{line 540:}} replace 'a lot' with 'larger proportions of missing data', a lot seems too colloquial for this kind of article. 
-\item{\textbf{lines 542-547:}} These seem like the most important practical results and should be more strongly emphasized.
-These are very important results for practitioners and the authors should try to make it more clear to the reader that this is a key take-home message.
-\item{\textbf{line 552:}} replace 'cladistic' with 'parsimony'. While one has come to be associated with the other in the literature, they are not synonymous. 
-\item{\textbf{line 557-558:}} '…than the "best" maximum likelihood tree'
-\item{\textbf{line 549-565}}: again, I am left wondering how much the differences in performance of max.
+Further, there was higher phylogenetic accuracy of simulated datasets with more characters than fewer characters, for slow or fast evolving characters (Wiens 2005).}}
+\item{\textcolor{blue}{\textbf{line 540:} replace 'a lot' with 'larger proportions of missing data', a lot seems too colloquial for this kind of article. }}
+\item{\textcolor{blue}{\textbf{lines 542-547:} These seem like the most important practical results and should be more strongly emphasized.
+These are very important results for practitioners and the authors should try to make it more clear to the reader that this is a key take-home message.}}
+\item{\textcolor{blue}{\textbf{line 552:} replace 'cladistic' with 'parsimony'. While one has come to be associated with the other in the literature, they are not synonymous. }}
+\item{\textcolor{blue}{\textbf{line 557-558:} '…than the "best" maximum likelihood tree'}}
+\item{\textcolor{blue}{\textbf{line 549-565:} again, I am left wondering how much the differences in performance of max.
 likelihood and Bayesian inference techniques are driven by the use of the 'true' tree as a starting tree in the Bayesian MCMC search, while the likelihood analysis had no such guide tree to begin with.
@@
 I understand the computational limitations, but the practical limitation that empiricists will never have the 'true' tree to start with may (or may not) have a large impact on our interpretation and the generality of these findings.
-\item{\textbf{line 593:}} 'MorphoBank' - the 'b' should be capitalized in keeping with the authors' usage (O'Leary and Kaufman 2011)
-\item{\textbf{line 599:}} remove extra close parenthesis inside parenthetical statement
-\item{\textbf{line 595-606:}} this paragraph seems unsatisfactory advice for practitioners.
+}}
+\item{\textcolor{blue}{\textbf{line 593:} 'MorphoBank' - the 'b' should be capitalized in keeping with the authors' usage (O'Leary and Kaufman 2011)}}
+\item{\textcolor{blue}{\textbf{line 599:} remove extra close parenthesis inside parenthetical statement}}
+\item{\textcolor{blue}{\textbf{line 595-606:} this paragraph seems unsatisfactory advice for practitioners.
 First, advising that the tree topology be fixed using the Bayesian consensus tree before conducting analyses such as dating strongly conditions the results on only a single tree and as the authors point out in the parenthetical statement, the dating information could improve accuracy.
@@
 This paragraph requires further justification and this study does have important bearing on a subject that gets little attention.
-\item{\textbf{line 612-613:}} too obvious and not especially helpful.
+}}
+\item{\textcolor{blue}{\textbf{line 612-613:} too obvious and not especially helpful.
 We should aim to have as little missing data as possible…more helpful, and something especially discovered in this study, is concentrating on having overlapping morphological characters between the living and fossil taxa to be able to place the living taxa with strong support via large datasets (combined data) and being able to position the fossils based on their shared derived characters with living taxa.
 This is an important point to make clearer.
-\item{\textbf{line 613-615:}} rephrase - sentence is choppy with all the commas. 
-\item{\textbf{line 615-618:}} It seems more important to me that the authors emphasize what implications it may have for previous and future studies that Bayesian analyses outperform max.
+}}
+\item{\textcolor{blue}{\textbf{line 613-615:} rephrase - sentence is choppy with all the commas. }}
+\item{\textcolor{blue}{\textbf{line 615-618:} It seems more important to me that the authors emphasize what implications it may have for previous and future studies that Bayesian analyses outperform max.
 likelihood, other than to re-hash that the authors feel the Bayesian consensus should be used to fix topologies for other downstream analyses.
@@
 Also 'further' is misspelled.
-\item{\textbf{line 628:}} use of computer cluster may be better placed in the methods, depending on the norm for MPE.
+}}
+\item{\textcolor{blue}{\textbf{line 628:} use of computer cluster may be better placed in the methods, depending on the norm for MPE.}}
 
-\item{\textbf{Figure 3:}} Do not repeat the aside to similarity with t-test again from Fig 2 and text.
-\item{\textbf{Figure 4:}} color codes in legend do not match up to the colors in the PDF I received.
+\item{\textcolor{blue}{\textbf{Figure 3:} Do not repeat the aside to similarity with t-test again from Fig 2 and text.}}
+\item{\textcolor{blue}{\textbf{Figure 4:} color codes in legend do not match up to the colors in the PDF I received.
 I don't know if this is an error in conversion, but the colors are black, red, green and blue in my PDF.
@@
 This makes interpreting the figure difficult for me at this time, but I assume given the text in the Results that the Bayesian methods are black and red in the figure I received, indicating they were closer to the best tree than max likelihood trees.
-Red and green should not be used together.
-\item{\textbf{Figure 5:}} This figure best summarizes the results, and a version of this should be the Graphical Abstract, with legends for the color cods and spelling out the missing data parameters, and the Y axis could just be 'clade similarity index' or 'rouge taxon index' so that the reader doesn't need to know what the names of the metrics are to grasp the graphic.
+Red and green should not be used together.}}
+\item{\textcolor{blue}{\textbf{Figure 5:} This figure best summarizes the results, and a version of this should be the Graphical Abstract, with legends for the color cods and spelling out the missing data parameters, and the Y axis could just be 'clade similarity index' or 'rouge taxon index' so that the reader doesn't need to know what the names of the metrics are to grasp the graphic.
 But concerning the Normalized Triplets metric, according to the text and appendix, if this index is 0, the similarity of the missing data tree to the best tree is no more than random, and < 0 it is more different than random.
@@
 Missing data for up to 50\% of characters has been suggested to still have high accuracy in uncovering the true tree (Wiens 2003; Wiens 2005), although for small datasets such as the one used in this study, the results show a similar decline in accuracy.
-This is a point that should be expanded upon in the discussion - how these results compare to other missing data studies and what do they tell us about the performance of the inference methods with missing data?
-\item{\textbf{Figure 6:}} I think this may be the most important figure because it is demonstrating the probability of overlap between the distributions of the best trees and the missing data trees.
+This is a point that should be expanded upon in the discussion - how these results compare to other missing data studies and what do they tell us about the performance of the inference methods with missing data?}}
+\item{\textcolor{blue}{\textbf{Figure 6:} I think this may be the most important figure because it is demonstrating the probability of overlap between the distributions of the best trees and the missing data trees.
 It took me a long time to figure it out, however, and to figure out why the top of the triangle is all orange.
@@
 The point is raised in the Results that once missing data in the living portion is > 50\% then they all perform poorly, but it was not easy to grasp.
-Somehow this figure needs to be spelled out better in the caption and results so the reader can grasp it more quickly and easily without having to go back to the methods section to completely figure out what is being represented.
+Somehow this figure needs to be spelled out better in the caption and results so the reader can grasp it more quickly and easily without having to go back to the methods section to completely figure out what is being represented.}}
 
-\item{\textbf{Table 1:}} The wording of the caption is off - the second paragraph is awkward. 
+\item{\textcolor{blue}{\textbf{Table 1:} The wording of the caption is off - the second paragraph is awkward. 
 Are these values the probabilities of overlap between the distributions of the 'best' tree versus trees from each inference method pooled across all combinations of missing data?
 That is what is described in methods line 397 but it doesn't say so in the caption.
-Two things:
-
+Two things:\\
 1)  Comparing to the tables in Appendix C, the distribution of probabilities of overlap between max.
@@
 How could the max.
-likelihood and Bayesian trees be so different?
-
+likelihood and Bayesian trees be so different?\\
 2)  Pooling across all missing data schemes seems to wash out the effect of tree-inference method.
 I am not sure what a better way to present these would be, but perhaps some kind of matrix where the cells are the joint probabilities of overlap of each method with each missing data category, and then the marginal probabilities would be the probabilities of overlap for each method, averaged over missing data categories, as well as the marginal probabilities of overlap for each missing data category, averaged across methods.
-e.g, RF distance:
-
+e.g, RF distance:\\
 Missing data:   0   25  50  75  Marginal \\
@@
 Marginal    1   0.9 0.78    0.5  \\
+This may not be feasible in the text due to space limitations, maybe tables like these could be added to an appendix.}}
 
-This may not be feasible in the text due to space limitations, maybe tables like these could be added to an appendix.
-
-\item{\textbf{Appendix C:}} Tables 1-3 repeats 'of the'. Should label the first two rows as 'Best maximum likelihood tree' and 'Best Bayesian inference tree'
-\item{\textbf{Appendix D:}} I find the online code repository thorough, but the repository is loaded with many files, especially for graphing.
-If it is possible to condense these files it may be easier for a reader to duplicate the study.
+\item{\textcolor{blue}{\textbf{Appendix C:} Tables 1-3 repeats 'of the'. Should label the first two rows as 'Best maximum likelihood tree' and 'Best Bayesian inference tree'}}
+\item{\textcolor{blue}{\textbf{Appendix D:} I find the online code repository thorough, but the repository is loaded with many files, especially for graphing.
+If it is possible to condense these files it may be easier for a reader to duplicate the study.}}
 

### 2015-06-29 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Revision init

--- 2015-06-08 Manuscript/MPESubmission/TEM_draft.tex
@@
 8.0.20  \citep{Stamatakis21012014}).
-For the morphological data, we used the implemented M\textit{kv} model \citep{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{k} + $\Gamma_4$; -K MK option in RAxML v.
+For the morphological data, we used the implemented M\textit{kv} model \citep{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{kv} + $\Gamma_4$; -K MK option in RAxML v.
 8.0.20 \citep{Stamatakis21012014}).

### 2015-06-29 Manuscript/MPESubmission/Revision/response.tex
minor comments

@@
 }}
+
+We revised this statement as follow: "many large-scale macroevolutionary studies focus solely on living species" (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 37 and throughout:} Rephrase so that 'however' does not start the sentence, such as 'To do this, however, we need…'.
@@
 p 48-49).}}
+
+%TG: Adam suggests there is a response to the Strunk and White book saying it's ok to say however. However (see what I did here? ;)) I don't mind changing it.
+
 \item{\textcolor{blue}{\textbf{line 43:} sentence choppy, consider rephrasing or splitting into two sentences.}}
+
+We modified the sentence as follow: "These approaches differ mainly in how they treat the fossil taxa and it's data.
+One can use the fossil as tips or as nodes in the phylogeny and use it's age only, it's morphology only or it's age and morphology jointly." (line @@@)
+
 \item{\textcolor{blue}{\textbf{line 44:} three alternatives: age only, morph only, or age and morph jointly. }}
+
+See previous change (Specific comment number 3).
+
 \item{\textcolor{blue}{\textbf{line 47:} Simpson 1944 doesn't seem the appropriate reference for maximum parsimony. Consider (Felsenstein 2004; Hennig 1966) }}
+
+We modified the citation as suggested by reviewer 1 (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 53:} omit 'are used as nodes rather than tips in these phylogenies and their'.
 Fossils aren't really used as nodes in node calibration, the second part of the sentence is more accurate.}}
+
+We modified the sentence to "Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies" (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 66:} omit 'have been successfully applied to empirical data' and just say 'becoming increasingly popular…list of references'. }}
+
+We modified the sentence as follow: "Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies (e.g. ...)" (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 76 and throughout the manuscript:} watch out for inconsistent parentheses.}}
+
+We fixed this issue.
+
 \item{\textcolor{blue}{\textbf{line 76 and throughout:} correct use of the word 'data' from singular to plural throughout}}
+
+%TG: Is that data/datum? molecular data or molecular datum?
 \item{\textcolor{blue}{\textbf{line 77:} change 'a lot' to 'large proportions of' or something more accurate and less colloquial}}

### 2015-06-29 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
minor comments

@@
 \section{Introduction}
-Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, the majority of macroevolutionary studies focus solely on living species (e.g.
+Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, the many large-scale macroevolutionary studies focus solely on living species (e.g.
 \citealp{meredithimpacts2011,jetzthe2012}).
@@
 Up to now, three main approaches have been used to place both living and fossil taxa into phylogenies.
-These approaches differ mainly in whether they treat fossil taxa as tips or as nodes in the phylogeny, and in which part of the available fossil data is used (i.e.
-the age of the fossil only or both its age and morphology).
+These approaches differ mainly in how they treat the fossil taxa and it's data.
+One can use the fossil as tips or as nodes in the phylogeny and use it's age only, it's morphology only or it's age and morphology jointly.
 Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
-Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \citep{simpson1945}.
+Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \citep{Hennig1966,felsenstein2004}.
 This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty.
-Neontologists, on the other hand, more commonly use probabilistic approaches (e.g.
-Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, fossils are used as nodes rather than tips in these phylogenies and their occurrence dates are used to time calibrate phylogenies \citep{zuckerkandl1965}.
+Neontologists, on the other hand, more commonly use probabilistic approaches (e.g. Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
+Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies \citep{zuckerkandl1965}.
 There have been great improvements in the theory and application of these two approaches (e.g.
@@
 This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as tip-dating \citealp{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
-Total Evidence methods have been successfully applied to empirical data (e.g.
-\citealp{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}), and are becoming an increasingly popular way of adding fossil taxa to phylogenies.
+Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies (e.g.
+\citealp{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}).
 However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
 Morphological data for living taxa are rarely collected when molecular data are available (e.g.
-\citealp{O'Leary08022013} \textit{vs.} \citealp{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record (for example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data is (nearly) always unavailable.
+\citealp{O'Leary08022013} \textit{vs.} \citealp{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record.
+For example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data is (nearly) always unavailable.
 Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.

### 2015-06-29 Manuscript/MPESubmission/Revision/response.tex
lines + cites changes

@@
 \begin{document}
+
+%Add introduction (dear Dr blablalbal, here are our revisions)
+
+%Resubmit the new version of the manuscript normaly but also add a pdf containing the track changes (highlithing in yellow the bit that have changed): "We are aware on the length of the manuscript.
+To facilitate the reviewers task, we attached to our revised manuscript, a pdf of the same manuscript but with the changes highlighted in yellow."
+
+%TG: Add a section "Additional 'fancy' changes" (not fancy obviously)
+% For clarity, we replaced the font of all the functions in R language in courier font (R, diversitree, etc...). \texttt{} - ACTUALLY DO THE CHANGE IF RELEVAN
 
@@
 
-\section{Reviewer 2:}
-
-\begin{enumerate}
-\item{\textcolor{blue}{\textbf{Compare the trees to the "true" tree:} I don't see how the "best" tree approach is helpful in this case.
-It adds an additional layer of complexity that can't be accounted for via simulation.
-The point of simulations is to compare empirical realities to the generating process, to see how much results diverge.
-I think the most helpful and straightforward thing to do is eliminate the "best" tree part of the MS, and compare the inferred trees directly to the real topology to quantify variance.}}
-
-\item{\textcolor{blue}{\textbf{Better explanation of the missing data:} I found the explanation of the missing data strategy very difficult to follow.
-First, they oddly use the verb "to infer" for matrices; are they inferring matrices?
-Second, it's not clear anywhere that the authors actually removed the molecular data from the fossil taxa, though I assume they did.
-I hope the authors can re-write in simpler language ("explain like I'm 5") for easier comprehension.}}
-
-\item{\textcolor{blue}{\textbf{$M_C$ is not missing data per se:} If Mc is just removing columns, then they are shrinking the total matrix.
-This is not missing data per se.
-They are instead comparing matrices of different sizes, and the quantity is "data present." Any matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
-A matrix with 10,000 characters is tiny if 10,000,000 could be scored.
-This is semantic.}}
-
-\item{\textcolor{blue}{\textbf{Specific variation of data per fossil:} More importantly, it doesn't seem like missing data varied among fossil taxa, unless I'm misreading.
-What we really want to know is what happens when some fossil taxa are fragmentary, as this is the most common empirical situation.
-Thus, there are morphological characters scored for some living taxa, some fossil taxa, and missing from others.
-It's unclear if this was simulated.
-It would be good if the authors could detail the case where, say, all living taxa had morphological data (0 Ml), the matrix was large (0 Mc), and Mf varied *among* fossil taxa, such that some fossils had 100\% coverage, and varied down to, say, 5\%.
-That's the crucial case here.}}
-\end{enumerate}
-
 \section{Specific comments:}
@@
 
-We modified the sentence to "Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies" (line @@@).
+We modified the following sentence as suggested: "Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies" (line @@@).
 
@@
 
-We modified the sentence as follow: "Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies (e.g. ...)" (line @@@).
+We modified the following sentence as suggested: "Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies (e.g. ...)" (line @@@).
 
@@
 %TG: Is that data/datum? molecular data or molecular datum?
+
 \item{\textcolor{blue}{\textbf{line 77:} change 'a lot' to 'large proportions of' or something more accurate and less colloquial}}
+
+We modified the following sentence as suggested: "Total Evidence matrices are likely to contain a large proportions of missing data" (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 82:} change to '…missing data are not phylogenetically biased, Wiens…'}}
+
+We modified the following sentence as suggested: "as long as enough data overlaps in each case, and missing data are not phylogenetically biased" (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 99:} unclear - what is meant by 'with a fixed topology approach'?
 The studies references investigated divergence times using total evidence datasets and in some cases used constraints but not completely fixed topologies}}
+
+We modified the previous statement to: "but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints" (line @@@).
+
 \item{\textcolor{blue}{\textbf{lines 80-103:} the authors need to clearly justify the importance of this study compared the many studies that have investigated the effects of missing data on topology inference.
@@
 That is one aspect of this study that I perceived as new and unique.
-Another aspect of this study design that could be new but is not explored here is the effect of the missing data on the morphological substitution rate parameter estimates.\textbf{Can you recover the true values from the perturbed datasets?} This study also used a unique way of evaluating the tree inference under missing data scenarios.
+Another aspect of this study design that could be new but is not explored here is the effect of the missing data on the morphological substitution rate parameter estimates.
+\textbf{Can you recover the true values from the perturbed datasets?} This study also used a unique way of evaluating the tree inference under missing data scenarios.
 These and other aspects of the study uniqueness must be clearer.}}
+
+%Emphasize on the questions and the uniqueness
+%The authors could argue that a limitation of the Pattinson study, which they address here, was the limited scheme for generating missing data from the datasets.
+%In this study, the authors perturbed the datasets more thoroughly in a clearly structured experimental design to investigate the effects of eliminating living versus fossil data or all data overall.
+%That is one aspect of this study that I perceived as new and unique.
+%This study also used a unique way of evaluating the tree inference under missing data scenarios.
+
+%Another aspect of this study design that could be new but is not explored here is the effect of the missing data on the morphological substitution rate parameter estimates.
+Can you recover the true values from the perturbed datasets?
+%We did not explored any effect of the missing data on parameters estimation in this study because we thought that it was already a complex/long one and focusing solely on topology made it easier to read.
+%We think, however, that the reviewer underlines a really interesting question especially since we have the data available.
+%It seems that parameters vary blablablabla.
+%And say something like we did the analysis but still consider it should not go in the manuscript.
+
 \item{\textcolor{blue}{\textbf{line 124:} replace 'sections' with 'partition'}}
+
+We modified the following sentence as suggested: "the morphological partition of Total Evidence matrices" (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 125:} omit 'note that we explain each step in detail below this general outline', replace with something like 'Our general protocol included the following steps (Figure 1):'…}}
+
+We removed the part suggested by the reviewer as follow: "we used the following protocol (Figure 1):" (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 167 and throughout manuscript:} replace usage of 'inferred' when referring to simulating data.
@@
 }}
+
+%To do!
+
 \item{\textcolor{blue}{\textbf{line 173:} streamline to "The substitution rates were selected from a gamma distribution with…" or something similar. }}
+
+We modified the following sentence as suggested: "The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5." (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 173 to 176:} need some justification for why the substitution rate with alpha shape of 0.5  will avoid homoplasy.
@@
 }}
+
+%To do
+
 \item{\textcolor{blue}{\textbf{Line 176 to 180:} could be streamlined.
@@
 Just for example.}}
+
+%To do
+
 \item{\textcolor{blue}{\textbf{line 182:} give the ape function used}}
+
+%To do
+
 \item{\textcolor{blue}{\textbf{line 185 to 188:} streamline, e.g.
 '…sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A).'}}
+
+We modified the following sentence as suggested: "sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A and Fig.
+A1 within)." (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 193 to 196:} again, justify why gamma with shape of 0.5 is appropriate for minimizing homoplasy.
@@
 }}
+
+%To do
+
 \item{\textcolor{blue}{\textbf{line 201:} change 'get' for 'obtain' or something less colloquial.}}
+
+We modified the following sentence as suggested: "We modified the complete matrix to obtain matrices with missing data" (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 231-235:} streamline, e.g., 'The minimal dataset included in this study was 5\% of the morphological matrix for any taxon'.}}
+
+We modified the sentence as follows: "To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon." (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 237-246:} I understand the rational presented by the authors for comparing missing data trees to the best tree inferred from max.
@@
 }}
+
+%To do (MAJOR)
+
 \item{\textcolor{blue}{\textbf{line 249:} condense citations to one parenthetical statement, 'GTR model (Tavare, 1986, default setting in RAx…)'.
@@
 }}
+
+We fixed these issues throughout the manuscript.
+
 \item{\textcolor{blue}{\textbf{line 251:} omit 'implemented'}}
@@
 
+\section{Reviewer 2:}
+
+\begin{enumerate}
+\item{\textcolor{blue}{\textbf{Compare the trees to the "true" tree:} I don't see how the "best" tree approach is helpful in this case.
+It adds an additional layer of complexity that can't be accounted for via simulation.
+The point of simulations is to compare empirical realities to the generating process, to see how much results diverge.
+I think the most helpful and straightforward thing to do is eliminate the "best" tree part of the MS, and compare the inferred trees directly to the real topology to quantify variance.}}
+
+\item{\textcolor{blue}{\textbf{Better explanation of the missing data:} I found the explanation of the missing data strategy very difficult to follow.
+First, they oddly use the verb "to infer" for matrices; are they inferring matrices?
+Second, it's not clear anywhere that the authors actually removed the molecular data from the fossil taxa, though I assume they did.
+I hope the authors can re-write in simpler language ("explain like I'm 5") for easier comprehension.}}
+
+\item{\textcolor{blue}{\textbf{$M_C$ is not missing data per se:} If Mc is just removing columns, then they are shrinking the total matrix.
+This is not missing data per se.
+They are instead comparing matrices of different sizes, and the quantity is "data present." Any matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
+A matrix with 10,000 characters is tiny if 10,000,000 could be scored.
+This is semantic.}}
+
+\item{\textcolor{blue}{\textbf{Specific variation of data per fossil:} More importantly, it doesn't seem like missing data varied among fossil taxa, unless I'm misreading.
+What we really want to know is what happens when some fossil taxa are fragmentary, as this is the most common empirical situation.
+Thus, there are morphological characters scored for some living taxa, some fossil taxa, and missing from others.
+It's unclear if this was simulated.
+It would be good if the authors could detail the case where, say, all living taxa had morphological data (0 Ml), the matrix was large (0 Mc), and Mf varied *among* fossil taxa, such that some fossils had 100\% coverage, and varied down to, say, 5\%.
+That's the crucial case here.}}
+\end{enumerate}
+
 \end{document}

### 2015-06-29 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
lines + cites changes

@@
 \citealp{meredithimpacts2011,jetzthe2012}).
-Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages (e.g.
-\citealp{manosphylogeny2007}) or niche occupancy (e.g.
-\citealp{pearmanniche2008}).
+Ignoring fossil taxa may lead to misinterpretation of macroevolutionary patterns and processes such as the timing of diversification events \citep[e.g.][]{pyrondivergence2011}, relationships among lineages \citep[e.g.][]{manosphylogeny2007} or niche occupancy \citep[e.g.][]{pearmanniche2008}.
 This has led to increasing consensus among evolutionary biologists that fossil taxa should be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments (e.g.
-\citealp{pyrondivergence2011,ronquista2012,BEASTmaster}).
+However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments \citep[e.g.][]{pyrondivergence2011,ronquista2012,BEASTmaster}.
 
@@
 Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies \citep{zuckerkandl1965}.
-There have been great improvements in the theory and application of these two approaches (e.g.
-\citealp{bapsta2013,stadlerdating2013,heaththe2013}) as well as much debate about the ``best" approach to use (e.g.
-\citealp{spencerefficacy2013,wrightbayesian2014}).
+There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the ``best" approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
 However neither approach uses all the available data.
@@
 A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \citep{eernissetaxonomic1993}.
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny (known as tip-dating \citealp{ronquista2012}), and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
-Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies (e.g.
-\citealp{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}).
+This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny \citep[known as tip-dating;][]{ronquista2012}, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
+Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
@@
 For example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data is (nearly) always unavailable.
-Therefore Total Evidence matrices are likely to contain a lot of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
+Therefore Total Evidence matrices are likely to contain a large proportions of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data is not biased towards particular subclades \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data are not phylogenetically biased \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
@@
 
-In this study, we focus on the effect of missing data on our ability to recover a ``best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species (e.g.
-\citealp{meredithimpacts2011,jetzthe2012}), or for studying evolutionary transitions (e.g.
-\citealp{friedmanexplosive2010}).
-Although branch length estimation is also important (namely for timing extinction and/or speciation events - e.g.
-\citealp{ronquista2012}), we do not consider branch lengths in this study.
-This is partially due to difficulties with simulating branch lengths and topology simultaneously, but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but with a fixed topology approach \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+In this study, we focus on the effect of missing data on our ability to recover a ``best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
+Although branch length estimation is also important \citep[namely for timing extinction and/or speciation events; e.g.][]{ronquista2012}, we do not consider branch lengths in this study.
+This is partially due to difficulties with simulating branch lengths and topology simultaneously, but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework.
@@
 \section{Materials and Methods}
-To explore how missing data in the morphological sections of Total Evidence matrices influences tree topology, we used the following protocol (note that we explain each step in detail below this general outline; Figure 1).
+To explore how missing data in the morphological partition of Total Evidence matrices influences tree topology, we used the following protocol (Figure 1):
 \begin{enumerate}
@@
 \label{Generating_the_matrix}
-First we randomly generated a ``true" tree of 50 taxa in R v.
-3.0.2 \citep{R302} using the package diversitree v.
-0.9-6 \citep{fitzjohndiversitree2012}.
+First we randomly generated a ``true" tree of 50 taxa in R v. 3.0.2 \citep{R302} using the package diversitree v. 0.9-6 \citep{fitzjohndiversitree2012}.
 We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution (bounded between 0 and 1) but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
@@
 Next, we generated a molecular and a morphological matrix from the ``true" tree.
-The molecular matrix was inferred from the ``true" tree using the R package phyclust v.
-0.1-14 \citep{chen2011}.
+The molecular matrix was inferred from the ``true" tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
 The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1) and transition/transversion rate of two \citep{douadycomparison2003}.
-The substitution rates were distributed following a gamma distribution with an alpha ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
+The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
 We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
@@
 
-We inferred the morphological matrix using the R package ape v.
-3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
-We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters.
-These probabilities were selected using the overall distribution of character states extracted from 100 published empirical morphological matrices (see Appendix A and Fig.
+We inferred the morphological matrix using the R package ape v. 3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
+We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A and Fig.
 A1 within).
@@
 \label{Removing_data}
-We modified the ``complete" matrix to get matrices with missing data by randomly replacing data with ``?" in the morphological part of the matrices according to the following parameters:
+We modified the ``complete" matrix to obtain matrices with missing data by randomly replacing data with ``?" in the morphological part of the matrices according to the following parameters:
 \begin{enumerate}
@@
 This parameter illustrates the number of available morphological characters for both living and fossil taxa.
-This reflects the amount of effort put into identifying the characters (e.g.
-\citealp{O'Leary08022013}).
+This reflects the amount of effort put into identifying the characters \citep[e.g][]{O'Leary08022013}.
 \end{enumerate}
@@
 Note that one of these combinations has no missing data so is equivalent to the ``complete" matrix, thus we have one effectively complete matrix in our 125 ``missing-data" matrices.
-Because some parameter combinations introduce a lot of missing data (e.g.
-$M_L$ = 75\%, $M_F$ = 75\% and $M_C$ = 75\%), some matrices contained fossil taxa without any data at all.
-When this occurred we repeated the random deletion of characters until each taxon had at least 5\% data across the whole morphological part of the matrix.
+To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
 
@@
 The ``best" tree, on the other hand, is the best tree we can build using state-of-the-art phylogenetic methods.
-In real world situations, the ``true" tree is never available to us because we cannot know the true evolutionary history of a clade (except in very rare circumstances, e.g.
-\citealp{rozen2005}).
+In real world situations, the ``true" tree is never available to us because we cannot know the true evolutionary history of a clade \citep[except in very rare circumstances, e.g.][]{rozen2005}.
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the ``best" tree, rather than the ``true" tree, as the ``best" tree is generally what we have to work with.
@@
 8.0.20 \citep{Stamatakis21012014}.
-For the molecular data, we used the GTR + $\Gamma_4$ model \citep{tavare1986} (default GTRGAMMA in RAxML v.
-8.0.20  \citep{Stamatakis21012014}).
+For the molecular data, we used the GTR + $\Gamma_4$ model (\citealp{tavare1986}; default GTRGAMMA in RAxML v.
+8.0.20; \citealp{Stamatakis21012014}).
 For the morphological data, we used the implemented M\textit{kv} model \citep{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{kv} + $\Gamma_4$; -K MK option in RAxML v.
-8.0.20 \citep{Stamatakis21012014}).
+8.0.20; \citealp{Stamatakis21012014}).
 We used RAxML because it automatically corrects for acquisition bias \citep{lewisa2001}.
-It is also heavily used in the literature for Maximum Likelihood tree inference (e.g.
-\citealp{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014}) and is one of the fastest methods available \citep{Stamatakis01102008}.
+It is also heavily used in the literature for Maximum Likelihood tree inference \citep[e.g.][]{rouresite-specific2011,Bogdanowicz2012,springermacroevolutionary2012,O'Leary08022013,kellymolecular2014} and is one of the fastest methods available \citep{Stamatakis01102008}.
 
@@
 We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
-We repeated this selection until we obtained 50 sets of simulations (i.e.
-50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
+We repeated this selection until we obtained 50 sets of simulations (i.e. 50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
 This step was implemented to make sure that the differences we observed in topologies (see below) were due to the amount of missing data for each parameter ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
@@
 \subsubsection{Bayesian}
-The ``best" tree and the ``missing-data" trees were inferred using MrBayes v.
-3.2.1 \citep{Ronquist2012mrbayes}.
+The ``best" tree and the ``missing-data" trees were inferred using MrBayes v. 3.2.1 \citep{Ronquist2012mrbayes}.
 We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
@@
 We compared the topology of the ``missing-data" trees to the ``best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
-We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa (i.e.
-taxa that frequently change position in different trees \citep{kearneyfragmentary2002}).
+We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa \citep[i.e.
+taxa that frequently change position in different trees][]{kearneyfragmentary2002}.
 We used these two metrics because they illustrate two different aspects of tree topology (see Discussion) but also because their performance in measuring differences in topology is well described \citep{Kuhner04112014} and well implemented \citep{Bogdanowicz2012}.
@@
 This methods scales any tree comparison metric using the mean distance between 1000 random trees (see Appendix B for the calculation details).
-This method is a generalisation of the topological accuracy method \citep{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
+This method is a generalisation of the topological accuracy method \citep{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.
+In practice when the Normalised Robinson-Foulds metric between two trees is equal to one, the trees are identical; if the metric is equal to zero, the trees are no more different than expected by chance; finally if the metric is less than zero, the trees are more different than expected by chance.
 Note that once rescaled, the Normalised Robinson-Foulds metric is a measure of similarity, rather than of distance like the original Robinson-Foulds metric.
@@
 We repeated all the paired tree comparisons described above for each of the 50 simulation runs.
-We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v.
-3.1 \citep{hdrcde}.
+We then calculated the mode and the 50\% and 95\% confidence intervals from the resulting distribution using the hdrcde R package v. 3.1 \citep{hdrcde}.
 
@@
 Additionally, using Bayesian consensus trees recovered the ``best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Figure 5, Figure 6, Table ~\ref{Tab_Results-Difference_methods}).
-As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters (e.g.
-\citealp{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}).
+As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
 
@@
 When the number of living taxa with morphological data ($M_{L}$) decreases, entire rows of data are being removed from the living taxa part of the matrix.
-Because living taxa still have molecular characters available for phylogenetic inference (see Methods),even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
+Because living taxa still have molecular characters available for phylogenetic inference (see Methods), even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
 However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
@@
 This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
-When the fossil taxa have less data (e.g.
-$M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils.
+When the fossil taxa have less data (e.g. $M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils.
 Therefore a majority consensus position is unlikely to exist (i.e.
@@
 In other words, the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree, where the fossil will be placed with lower confidence at a higher taxonomic level.
-We argue that using the Bayesian consensus tree topology is preferable because it is more conservative (e.g.
-\citealp{pattinsonphylogeny2014}).
+We argue that using the Bayesian consensus tree topology is preferable because it is more conservative \citep[e.g.][]{pattinsonphylogeny2014}.
 
@@
 Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
-However, although these matrices are an order of magnitude smaller than some published matrices (e.g.
-\citealp{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices (e.g.
-\citealp{kellymolecular2014, sallam2011craniodental}).
-Therefore, our simulations reflect realistic parameters.
-Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{wrightbayesian2014}.
+However, although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
+Therefore, our simulations reflect realistic parameters. Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{wrightbayesian2014}. 
 However, the use of probabilistic methods (i.e.
@@
 Variation in our ability to recover the ``best" tree topology depends heavily on the tree inference method (Figure 4 and Figure 5).
-For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over cladistic methods \citep{wrightbayesian2014} (but see \citealp{spencerefficacy2013}).
+For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over cladistic methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
 However, this is the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
@@
 Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries.
-However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils (e.g.
-\citealp{nithe2013}).
+However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
 Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the ``best" tree topology (Figure 6).
 Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
-One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{Morphobank} \citep{morphobank}, so that no one individual collects all the data.
+One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{MorphoBank} \citep{morphobank}, so that no one individual collects all the data.
 
@@
 Note that, although the Bayesian posterior tree distribution performed poorly in recovering the ``best" tree topology, this is due to difficulties comparing distributions of trees (see above).
-Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses (e.g.
-\citealp{jetzthe2012}).
+Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses \citep[e.g.][]{jetzthe2012}.
 
@@
 Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
+Thanks to the two anonymous reviewers for their useful and enthusiastic comments.
 Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.

### 2015-06-29 Manuscript/MPESubmission/Revision/response.tex
Updated response

@@
 \item{\textcolor{blue}{\textbf{Figure 4:} Figure 4 has a color scheme issue, the colors in the caption do not match up to the figure.}}
+
+%We apologies for this silly mistake.
 
@@
 \item{\textcolor{blue}{\textbf{line 251:} omit 'implemented'}}
+
+We removed this word.
+
 \item{\textcolor{blue}{\textbf{line 273:} 'Bayesian Inference'}}
+
+We changed the title as suggested.
+
 \item{\textcolor{blue}{\textbf{line 286:} Why was the true tree used as a starting tree?
@@
 }}
+
+%To do (MAJOR)
+
 \item{\textcolor{blue}{\textbf{line 318:} replace 'the twice number' with 'twice the number'}}
+
+We fixed this typo.
+% TG: should I not add something like "and thank the reviewer for pointing it out."?
+
 \item{\textcolor{blue}{\textbf{line 327:} 'This methods' omit 's'}}
+
+We fixed this typo.
+
 \item{\textcolor{blue}{\textbf{line 343:} 'this metric is sensitive' add 'is'}}
+
+We fixed this typo.
+
 \item{\textcolor{blue}{\textbf{line 424:} 'Figure ??' should be Figure 6, but then do not repeat Figure 6 in the parentheses. }}
+
+We fixed this typo.
+
 \item{\textcolor{blue}{\textbf{line 436:} 'effect'}}
+
+We fixed this typo as well as in the following sentence (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 489:} Streamline to something like: '…when the fossil is present near the tips but affects the clade conservation less when fossils are near the root'. }}
+
+We modified the following sentence as suggested: "In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root." (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 494:} 'taxon'}}
+
+We fixed this typo.
+
 \item{\textcolor{blue}{\textbf{line 515:} '…the size of the matrix' missing 'of'}}
+
+We fixed this typo.
+
 \item{\textcolor{blue}{\textbf{line 518 to 520:} I am not sure that (Wagner 2000) states that incongruence between molecular and morphological data is 'more important in small morphological matrices'.
 }}
+
+%To do
+
 \item{\textcolor{blue}{\textbf{line 520:} Either change 'size' to plural or change 'were' to 'was', the former seems to flow better. }}
+
+We modified the following sentence: "The sizes of our data matrices were constrained by the performance of our protocol" (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 528:} missing close parenthesis after Ni et al.}}
+
+We fixed this typo.
+
 \item{\textcolor{blue}{\textbf{line 532:} Wright and Hillis 2014 found that increasing the size of the dataset improves topological accuracy, and they do not argue that simply increasing the number of characters increases the amount of homoplasy.
@@
 Further, there was higher phylogenetic accuracy of simulated datasets with more characters than fewer characters, for slow or fast evolving characters (Wiens 2005).}}
+
+%To do
+
 \item{\textcolor{blue}{\textbf{line 540:} replace 'a lot' with 'larger proportions of missing data', a lot seems too colloquial for this kind of article. }}
+
+We modified the sentence as suggested: "they have a larger proportion of missing data" (line @@@).
+
 \item{\textcolor{blue}{\textbf{lines 542-547:} These seem like the most important practical results and should be more strongly emphasized.
 These are very important results for practitioners and the authors should try to make it more clear to the reader that this is a key take-home message.}}
+
+%To do
+
 \item{\textcolor{blue}{\textbf{line 552:} replace 'cladistic' with 'parsimony'. While one has come to be associated with the other in the literature, they are not synonymous. }}
+
+We replaced 'cladistic' by 'parsimony' as suggested (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 557-558:} '…than the "best" maximum likelihood tree'}}
+
+We modified the sentence as suggested : "than the ``best'' Maximum Likelihood tree" (line @@@).
+
 \item{\textcolor{blue}{\textbf{line 549-565:} again, I am left wondering how much the differences in performance of max.
@@
 }}
+
+%To do (MAJOR)
+
 \item{\textcolor{blue}{\textbf{line 593:} 'MorphoBank' - the 'b' should be capitalized in keeping with the authors' usage (O'Leary and Kaufman 2011)}}
+
+We fixed this typo.
+
 \item{\textcolor{blue}{\textbf{line 599:} remove extra close parenthesis inside parenthetical statement}}
+
+We fixed this typo.
+
 \item{\textcolor{blue}{\textbf{line 595-606:} this paragraph seems unsatisfactory advice for practitioners.
@@
 }}
+
+%To do (rewrite)
+
 \item{\textcolor{blue}{\textbf{line 612-613:} too obvious and not especially helpful.
@@
 }}
+
 \item{\textcolor{blue}{\textbf{line 613-615:} rephrase - sentence is choppy with all the commas. }}

### 2015-06-29 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Updated response

@@
 8.0.20; \citealp{Stamatakis21012014}).
-For the morphological data, we used the implemented M\textit{kv} model \citep{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{kv} + $\Gamma_4$; -K MK option in RAxML v.
+For the morphological data, we used the M\textit{kv} model \citep{lewisa2001} assuming an equal state frequency and a unique overall substitution rate ($\mu$) following a gamma distribution of the rate variation with four distinct categories (M\textit{kv} + $\Gamma_4$; -K MK option in RAxML v.
 8.0.20; \citealp{Stamatakis21012014}).
@@
 
-\subsubsection{Bayesian}
+\subsubsection{Bayesian Inference}
 The ``best" tree and the ``missing-data" trees were inferred using MrBayes v. 3.2.1 \citep{Ronquist2012mrbayes}.
@@
 \subsubsection{Robinson-Foulds distance}
-The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and the twice number of shared clades across two trees.
+The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and twice the number of shared clades across two trees.
 The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} (see Appendix B for calculation details).
@@
 We normalised this metric following Bogdanowicz's Normalised Tree Similarity (NTS) method \citep{Bogdanowicz2012}.
-This methods scales any tree comparison metric using the mean distance between 1000 random trees (see Appendix B for the calculation details).
+This method scales any tree comparison metric using the mean distance between 1000 random trees (see Appendix B for the calculation details).
 This method is a generalisation of the topological accuracy method \citep{Price2010} allowing to compare topological differences between any tree with any tree comparison metric.
@@
 It is bounded between zero when the two trees are identical and $\binom{n}{3}$ (for two trees with $n$ taxa) when there is no shared taxa/clade position in the two trees.
-Therefore this metric sensitive to the conservation of wildcard taxa.
+Therefore this metric is sensitive to the conservation of wildcard taxa.
 We normalised this metric in the same way as for the Robinson-Foulds distance resulting in the Normalised Triplets metric.
@@
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Figure 5; Tables C2, C3 and C4 in Appendix C).
-Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Figure 6).
+Figure 6 shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0).
 
@@
 However, the Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; Figure 6A) displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; Figure 6B).
-Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Figure 6A).
+Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ (Figure 6A).
 
@@
 every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
-In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present at a lower taxonomic level (i.e.
-separated with more nodes from the root) but affects the clade conservation less at higher taxonomic level (i.e.
-separated with less nodes from the root).
-Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxa and decrease the Normalised Triplets metric.
+In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.
+Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxon and decrease the Normalised Triplets metric.
 Therefore, the $M_{F}$ parameter is likely to affect the Normalised Robinson-Foulds metric less than the Normalised Triplets metric for the Bayesian consensus trees.
@@
 
-Additionally, the number of morphological characters determines the size the matrix.
+Additionally, the number of morphological characters determines the size the of matrix.
 This can affect our ability to recover the ``best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
 The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \citep{wagner2000}.
-The size of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
+The sizes of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
 Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
@@
 As expected, when combining the missing data parameters, our ability to recover the ``best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a lot of missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a larger proportion of missing data (i.e.
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Figure 5).
@@
 Variation in our ability to recover the ``best" tree topology depends heavily on the tree inference method (Figure 4 and Figure 5).
-For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over cladistic methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
+For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over parsimony methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
 However, this is the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-Our results show that the topology of the Bayesian consensus tree is always closer to the ``best" tree topology than the one of the Maximum Likelihood tree (Figure 5).
+Our results show that the topology of the Bayesian consensus tree is always closer to the ``best" tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
 As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
@@
 Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the ``best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \citep{ronquista2012,BEASTmaster}; although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution).
+Because the Bayesian consensus trees consistently recovered topologies closer to the ``best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \citep[][; although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution]{ronquista2012,BEASTmaster}.
 Note that, although the Bayesian posterior tree distribution performed poorly in recovering the ``best" tree topology, this is due to difficulties comparing distributions of trees (see above).

### 2015-06-30 Manuscript/MPESubmission/Revision/response.tex
Del with all the minor changes

@@
 
+%To do (rewrite)
+
 \item{\textcolor{blue}{\textbf{line 613-615:} rephrase - sentence is choppy with all the commas. }}
+
+We modified the sentence as follow: "Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa." (line @@@),
+
 \item{\textcolor{blue}{\textbf{line 615-618:} It seems more important to me that the authors emphasize what implications it may have for previous and future studies that Bayesian analyses outperform max.
@@
 }}
+
+%To do (rewrite) - typo ok
+
 \item{\textcolor{blue}{\textbf{line 628:} use of computer cluster may be better placed in the methods, depending on the norm for MPE.}}
 
+The Lonsdale cluster requires this specific line in the acknowledgements of any study using it's infrastructure.
+
 \item{\textcolor{blue}{\textbf{Figure 3:} Do not repeat the aside to similarity with t-test again from Fig 2 and text.}}
+
+We removed the sentence about the t-test similarity from both figures captions. %TG: not sure if that's what the reviewer wanted.
+
 \item{\textcolor{blue}{\textbf{Figure 4:} color codes in legend do not match up to the colors in the PDF I received.
@@
 Red and green should not be used together.}}
+
+We apologies for this silly mistake and thanks the reviewer for pointing it out.
+We changed the color scheme to "Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
+%To do: change the graphs colours
+
 \item{\textcolor{blue}{\textbf{Figure 5:} This figure best summarizes the results, and a version of this should be the Graphical Abstract, with legends for the color cods and spelling out the missing data parameters, and the Y axis could just be 'clade similarity index' or 'rouge taxon index' so that the reader doesn't need to know what the names of the metrics are to grasp the graphic.
@@
 This is a point that should be expanded upon in the discussion - how these results compare to other missing data studies and what do they tell us about the performance of the inference methods with missing data?}}
+
+We thank the reviewer for this excellent suggestion and modified our graphical abstract.
+%To do (writing)
+%Missing data for up to 50\% of characters has been suggested to still have high accuracy in uncovering the true tree (Wiens 2003; Wiens 2005)
+%how these results compare to other missing data studies and what do they tell us about the performance of the inference methods with missing data?
+
 \item{\textcolor{blue}{\textbf{Figure 6:} I think this may be the most important figure because it is demonstrating the probability of overlap between the distributions of the best trees and the missing data trees.
@@
 Somehow this figure needs to be spelled out better in the caption and results so the reader can grasp it more quickly and easily without having to go back to the methods section to completely figure out what is being represented.}}
+
+%To do (writing)
 
@@
 \item{\textcolor{blue}{\textbf{Appendix C:} Tables 1-3 repeats 'of the'. Should label the first two rows as 'Best maximum likelihood tree' and 'Best Bayesian inference tree'}}
+
+We removed the repetition in the caption. %Not sure I understand the second part of the comment.
+
 \item{\textcolor{blue}{\textbf{Appendix D:} I find the online code repository thorough, but the repository is loaded with many files, especially for graphing.
 If it is possible to condense these files it may be easier for a reader to duplicate the study.}}
+
+%To do
 

### 2015-06-30 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixC.tex
Del with all the minor changes

--- 2015-05-12 Manuscript/PLoSONEresponse/OnlineAppendices-LaTeXSuppFiles/SupportingInformation3.tex
@@
 
+\renewcommand\thefigure{C.\arabic{figure}}
+\renewcommand\thetable{C.\arabic{table}}
+
 \begin{document}
 
-\section{Supporting Information S3: Additional Results}
+\section{Appendix C: Additional Results}
 
@@
 \begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).}
+\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).}
 \label{Tab_Supp_summary_metric_allparam}
@@
 \begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
+\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_ML}
@@
 \begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
+\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_MF}
@@
 \begin{table}[ht]
-\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
+\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_MC}

### 2015-06-30 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Del with all the minor changes

@@
 Therefore we advise filling as many gaps in Total Evidence matrices as possible.
-Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.
+Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa.
 Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best" tree topology than the Maximum Likelihood trees.
-Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
+Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for further analyses.
 The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
@@
 $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
-The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.
-Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.\\
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.\\
 
@@
 Maximum Likelihood or Bayesian).
-The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$.
-Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.\\
+The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$.\\
 
@@
 \noindent
-\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).
+\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange).
 The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.

### 2015-07-08 Manuscript/MPESubmission/Revision/response.tex
Response updates

@@
 
-%TG: Adam suggests there is a response to the Strunk and White book saying it's ok to say however. However (see what I did here? ;)) I don't mind changing it.
+We thank the reviewer for this stylistic suggestion and modified the placement of 'however' throughout the manuscript.
 
@@
 
-%TG: Is that data/datum? molecular data or molecular datum?
+%We corrected the plural/singular of data throughout the text. 
+% TG: actually I'm not going to change it:
+We thank the reviewer for pointing out this inconsistency, however, 'data' is now commonly used for the singular case (Pinker, S.
+The Sense of Style: The Thinking Person's Guide to Writing in the 21st Century.
+Penguin, 2014.
+p 271 and Guardian and Observer style guide (2015): \url{http://goo.gl/cmKjNe}).
 
@@
 
-\item{\textcolor{blue}{\textbf{lines 80-103:} the authors need to clearly justify the importance of this study compared the many studies that have investigated the effects of missing data on topology inference.
+\item{\textcolor{blue}{\textbf{lines 80-103:} (1) the authors need to clearly justify the importance of this study compared the many studies that have investigated the effects of missing data on topology inference.
 The justification currently given is not particularly strong and the specific unique questions asked in this study that are currently unanswered should be made clear.
@@
 That is one aspect of this study that I perceived as new and unique.
-Another aspect of this study design that could be new but is not explored here is the effect of the missing data on the morphological substitution rate parameter estimates.
-\textbf{Can you recover the true values from the perturbed datasets?} This study also used a unique way of evaluating the tree inference under missing data scenarios.
+(2) Another aspect of this study design that could be new but is not explored here is the effect of the missing data on the morphological substitution rate parameter estimates.
+\textbf{Can you recover the true values from the perturbed datasets?}
+This study also used a unique way of evaluating the tree inference under missing data scenarios.
 These and other aspects of the study uniqueness must be clearer.}}
+
+(1) - We thank the reviewer for this judicious comment and emphasized on both the novelty aspect of our study and on the questions we are actually testing.
+
+"Until now, only few attempts have been made to study the impact of this issue on phylogenetic inference using Total Evidence methods (e.g.
+Wiens et al., 2005; Pattinson et al., 2014).%\citep[e.g.][]{Wiens01102005,pattinsonphylogeny2014}
+For example, Wiens et al. (2005) %\cite{Wiens01102005}
+ only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
+Furthermore, Pattinson et al. (2014) %\cite{pattinsonphylogeny2014}
+ assess the effect of removing both molecular and morphological data (``artificial extinction'') on the topology.
+These studies, however, are limited by their empirical approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014).
+
+In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
+We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa."
+
+Pattinson: artificial fossil - removing molecular data from living species to see how topology changes.
+Pattinson use combined analysis but only to understand the patterns of fossilisation (via artificial fossil).
+Wiens only focuses on a special case
+Limitation of Pattinson was the limited scheme for generating missing data from the datasets (specific to primates and observed patterns).
+In our case, we (a) don't focus on any group or any pattern, our approach is purely theoretical.
+The difference from Wiens or Pattinson is that we thoroughly investigate the effects of eliminating living versus fossil data or all data overall.
+(b) We also propose a thorough comparison between the ML and Bayesian inference and a measure two clear effects that missing data could have on topology: displacing wild-card taxa and not conserving clades structure.
 

### 2015-07-08 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Response updates

@@
 This has led to increasing consensus among evolutionary biologists that fossil taxa should be included in macroevolutionary studies \citep{jacksonwhat2006,quentaldiversity2010,dietlconservation2011,slaterunifying2013,fritzdiversity2013}.
-However, to do this we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments \citep[e.g.][]{pyrondivergence2011,ronquista2012,BEASTmaster}.
+To do this, however, we need to be able to place living and fossil taxa into the same phylogenies; a task that remains difficult despite recent methodological developments \citep[e.g.][]{pyrondivergence2011,ronquista2012,BEASTmaster}.
 
@@
 There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the ``best" approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
-However neither approach uses all the available data.
+Neither approach, however, uses all the available data.
 
@@
 Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
-However, although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
+Although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
 Morphological data for living taxa are rarely collected when molecular data are available (e.g.
@@
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data are not phylogenetically biased \citep{wiensmissing2003,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data are not phylogenetically biased \citep{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
-Until now, no attempt has been made to study the impact of this issue on phylogenetic inference using Total Evidence methods.
+Until now, only few attempts have been made to study study the impact of this issue on phylogenetic inference using Total Evidence methods \citep[e.g.][]{Wiens01102005,pattinsonphylogeny2014}.
+For example, \cite{Wiens01102005} only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
+Furthermore, \cite{pattinsonphylogeny2014} assess the effect of removing both molecular and morphological data (``artificial extinction'') on the topology.
+These studies, however, are limited by their empirical approach \citep{Wiens01102005} and the scheme of generating missing data \citep{pattinsonphylogeny2014}.
 
-In this study, we focus on the effect of missing data on our ability to recover a ``best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
+In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
+We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa.
+We focus on the effect of missing data on our ability to recover a ``best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
 Although branch length estimation is also important \citep[namely for timing extinction and/or speciation events; e.g.][]{ronquista2012}, we do not consider branch lengths in this study.
@@
 Therefore, when 10\% data was missing for both $M_L$ and $M_F$, 10\% data was missing in the morphological part of the matrix.
-However, when applying the $M_C$ parameter with the same percentage (10\%) the resulting matrix potentially had more than 10\% data missing.
+When applying the $M_C$ parameter with the same percentage (10\%), however, the resulting matrix potentially had more than 10\% data missing.
 
@@
 As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets metric).
-However, the different missing data parameters and tree inference methods do not affect the topology in the same way (Figure 4 and Figure 5).
+Nonetheless, the different missing data parameters and tree inference methods do not affect the topology in the same way (Figure 4 and Figure 5).
 
@@
 The Normalised Robinson-Foulds metric is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
-However, The Normalised Triplets metric is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
+The Normalised Triplets metric, however, is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
 
@@
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Figure 6).
-However, the Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; Figure 6A) displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; Figure 6B).
+The Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; Figure 6A), however, displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; Figure 6B).
 Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
@@
 For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds metric or Normalised Triplets metric respectively; Table ~\ref{Tab_Results-Difference_methods}).
-However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds metric (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Figure 4 and Figure C2 in Appendix C).
+These two methods, however, perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds metric (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Figure 4 and Figure C2 in Appendix C).
 
@@
 Our results show that the ability to recover the ``best" tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used.
-However, these factors affected topological recovery in different ways and to different extents.
+These factors, however, affected topological recovery in different ways and to different extents.
 Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Figure 6).
@@
 Because living taxa still have molecular characters available for phylogenetic inference (see Methods), even if they have no morphological data, the relationships among them will always be fairly well-resolved (depending on the phylogenetic signal from the molecular part of the matrix).
-However, this missing data parameter has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
+This missing data parameter, however, has a huge influence on the placement of fossil taxa because a decrease in the $M_{L}$ parameter reduces the amount of overlapping data among the living and fossil taxa, meaning there is no part of the living taxa tree that the fossils can branch off.
 
@@
 This can lead to difficulties for the placement of certain taxa in the tree.
-However, it is important to note that even though the number of displaced wildcard taxa increases (i.e.
+It is important, however, to note that even though the number of displaced wildcard taxa increases (i.e.
 decrease of Normalised Triplets metric) with increasing missing data in this parameter, clade conservation (i.e.
@@
 Conversely, removing 10\% of morphological characters ($M_{C}$) genuinely removes 10\% of data in the matrix.
-However, the effect of removing characters on the ability to recover the ``best" tree topology is of the same order of magnitude as for the other two parameters (Figure 4).
+Nonetheless, the effect of removing characters on the ability to recover the ``best" tree topology is of the same order of magnitude as for the other two parameters (Figure 4).
 We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.
@@
 Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
-However, although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
+Although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
 Therefore, our simulations reflect realistic parameters. Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{wrightbayesian2014}. 
-However, the use of probabilistic methods (i.e.
+Nonetheless, the use of probabilistic methods (i.e.
 Maximum Likelihood or Bayesian) and the M\textit{kv} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.
@@
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Figure 5).
-However, it is important to notice that the effect of each parameter is not additive.
+It is important, however, to notice that the effect of each parameter is not additive.
 The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and/or $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6).
@@
 For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over parsimony methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
-However, this is the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
+This is, however, the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
 Our results show that the topology of the Bayesian consensus tree is always closer to the ``best" tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
@@
 This may also be because the ``best" Bayesian consensus trees are not completely resolved, thus will always be more similar to the ``missing data" trees than a completely resolved tree like the ``best" Maximum Likelihood tree.
-However, we minimized the probability of unresolved ``best" trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
+Nonetheless, we minimized the probability of unresolved ``best" trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
 
@@
 Increasing data for fossil species is harder, since it depends on fossil preservation biases and new fossil discoveries.
-However, gaps in the matrix can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
+Gaps in the matrix, however, can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
 Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the ``best" tree topology (Figure 6).

### 2015-07-08 Analysis/Rates_estimates.R
Rates estimates analysis

@@ -1,0 +1,137 @@
+#Script for post-analysing the morphological rates estimates
+
+#Shell part (DO NOT RUN)
+do.not.run=TRUE
+if(do.not.run == FALSE) {
+system('
+mkdir morpho_rates
+mkdir morpho_rates/TrueRates
+mkdir morpho_rates/EstimatesRates
+
+for n in $(seq -w 01 50)
+do
+    tar xvf True_trees/Chain${n}_TrueTree/Chain${n}-randoms.tar MorphoGammaRates.txt
+    cat MorphoGammaRates.txt > morpho_rates/TrueRates/Chain${n}_TrueRates.txt
+    rm MorphoGammaRates.txt
+    grep "alpha{2}" Bayesian_trees/Chain${n}_Bayesian/MCMC/Chain${n}_*.pstat > morpho_rates/EstimatesRates/Chain${n}_EstRates.txt
+    printf .
+done
+')
+
+    #Reading the True Rates
+    files<-list.files("morpho_rates/TrueRates/", pattern="Chain")
+    True_Rates<-read.table(paste("morpho_rates/TrueRates/",files[n], sep=""))
+    for (n in 2:50) {
+        tmp<-read.table(paste("morpho_rates/TrueRates/",files[n], sep=""))
+        True_Rates<-cbind(True_Rates, tmp)
+    }
+    names(True_Rates)<-seq(1:50)
+    save(True_Rates, file="True_morpho.rates.Rda")
+
+    #Reading the Estimates rates
+    Estimates_rates<-list()
+    files<-list.files("morpho_rates/EstimatesRates/", pattern="Chain")
+    for (n in 1:50) {
+        tmp<-read.table(paste("morpho_rates/EstimatesRates/",files[n], sep=""))
+        tmp<-tmp[,c(2,3)]
+        names(tmp)<-c("Mean", "Variance")
+        Estimates_rates[[n]]<-tmp
+    }
+    names(Estimates_rates)<-seq(1:50)
+    #Removed bugged chains
+    Estimates_rates<-Estimates_rates[-c(3, 31)]
+    save(Estimates_rates, file="Estimate_morpho.rates.Rda")
+
+} else {
+    load("../Data/R_data/True_morpho.rates.Rda")
+    load("../Data/R_data/Estimate_morpho.rates.Rda")
+}
+
+
[…]

### 2015-07-08 Manuscript/MPESubmission/Revision/response.tex
Added rates analysis to appendix A

@@
 (2) Another aspect of this study design that could be new but is not explored here is the effect of the missing data on the morphological substitution rate parameter estimates.
-\textbf{Can you recover the true values from the perturbed datasets?}
+Can you recover the true values from the perturbed datasets?
 This study also used a unique way of evaluating the tree inference under missing data scenarios.
@@
 
-(1) - We thank the reviewer for this judicious comment and emphasized on both the novelty aspect of our study and on the questions we are actually testing.
+(1) - We thank the reviewer for this judicious comment and emphasized on both the novelty aspect of our study and on the questions we are actually testing as follow:
 
@@
 These studies, however, are limited by their empirical approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014).
-
 In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
 We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa."
-
-Pattinson: artificial fossil - removing molecular data from living species to see how topology changes.
+%Pattinson: artificial fossil - removing molecular data from living species to see how topology changes.
 Pattinson use combined analysis but only to understand the patterns of fossilisation (via artificial fossil).
-Wiens only focuses on a special case
-Limitation of Pattinson was the limited scheme for generating missing data from the datasets (specific to primates and observed patterns).
-In our case, we (a) don't focus on any group or any pattern, our approach is purely theoretical.
-The difference from Wiens or Pattinson is that we thoroughly investigate the effects of eliminating living versus fossil data or all data overall.
-(b) We also propose a thorough comparison between the ML and Bayesian inference and a measure two clear effects that missing data could have on topology: displacing wild-card taxa and not conserving clades structure.
+%Wiens only focuses on a special case
+%Limitation of Pattinson was the limited scheme for generating missing data from the datasets (specific to primates and observed patterns).
+%In our case, we (a) don't focus on any group or any pattern, our approach is purely theoretical.
+%The difference from Wiens or Pattinson is that we thoroughly investigate the effects of eliminating living versus fossil data or all data overall.
+%(b) We also propose a thorough comparison between the ML and Bayesian inference and a measure two clear effects that missing data could have on topology: displacing wild-card taxa and not conserving clades structure.
 
@@
 
-%Another aspect of this study design that could be new but is not explored here is the effect of the missing data on the morphological substitution rate parameter estimates.
-Can you recover the true values from the perturbed datasets?
-%We did not explored any effect of the missing data on parameters estimation in this study because we thought that it was already a complex/long one and focusing solely on topology made it easier to read.
-%We think, however, that the reviewer underlines a really interesting question especially since we have the data available.
-%It seems that parameters vary blablablabla.
-%And say something like we did the analysis but still consider it should not go in the manuscript.
+(2) - the second point underlined by the reviewer in this specific comment strikes use as really interesting.
+However, we considered this study to be already long and complex.
+We therefore think that adding an extra question to this study (i.e.
+what is the effect of missing data on the morphological substitution rate parameter estimates?) might loose the reader.
+Nonetheless, we added the result of such analysis in the \textbf{Appendix-A Morphological rates estimations results} section.
+% TG: what do you think?
+I don't want to attract the reviewer's attention too much (i.e.
+that's another shit load of analysis to do to do it properly) and certainly not the reader (it's not relevant to the whole paper) but sticking that to the supplementary kind of does what the reviewers want to know.
 

### 2015-07-08 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
Revision init
Added rates analysis to appendix A

--- 2015-01-25 Manuscript/SupplementaryMaterial/TreeBuilding.tex
@@
-%\section{Supplementary material 1 : Tree building}
+\documentclass[12pt,letterpaper]{article}
+
+%Packages
+\usepackage{xcolor}
+\usepackage{changepage}
+\usepackage{pdflscape}
+\usepackage{fixltx2e}
+\usepackage{textcomp}
+\usepackage{fullpage}
+\usepackage{natbib}
+\usepackage{float}
+\usepackage{latexsym}
+\usepackage{url}
+\usepackage{epsfig}
+\usepackage{graphicx}
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{bm}
+\usepackage{array}
+\usepackage[version=3]{mhchem}
+\usepackage{ifthen}
+\usepackage{caption}
+\usepackage{hyperref}
+\usepackage{amsthm}
+\usepackage{amstext}
+\usepackage{enumerate}
+\usepackage[osf]{mathpazo}
+\usepackage{dcolumn}
+\usepackage{lineno}
+\usepackage{longtable}
+\pagenumbering{arabic}
+
+%Pagination style and stuff
+\linespread{2}
+\raggedright
+\setlength{\parindent}{0.5in}
+%\setcounter{secnumdepth}{0}
+
+\renewcommand\thefigure{A.\arabic{figure}}
+
+\begin{document}
+
+\section{Appendix A: Tree Inference}
 
@@
 \centering
-\includegraphics[keepaspectratio=true]{Figures/Supplementary/TEM_Fig-AppendixCharacters.pdf}
+\includegraphics[keepaspectratio=true]{SupplementaryFigures/TEM_Fig-AppendixCharacters.pdf}
 \caption{The proportion of morphological characters with between two and 10 character states extracted from 100 randomly selected empirical matrices downloaded from TreeBASE.}
@@
 
+\subsection{Morphological rates estimations results}
+To test the sensitivity of the Bayesian prior on the morphological rates distribution shape ($\alpha$ = 0.5), we compared the shape of the distribution of our generated morphological characters ($\alpha_{TRUE}$) to the shape of the distribution of morphological characters rate estimated by the Bayesian inference ($\alpha_{EST}$).
+The $\alpha_{TRUE}$ parameter was estimated by fitting a gamma distribution to the 100 individual morphological rates (randomly sampled from a gamma distribution with $\alpha$ = 0.5) using the \texttt{fitdistr} function from the \texttt{MASS R} package \citep{MASS}.
+We measured the difference as:
+\begin{equation}
+\Delta\alpha=\alpha_{TRUE} - \alpha_{EST}
+\end{equation}
+We then calculated the mode and the 95\% and 50\% confidence intervals over the 50 chains using the \texttt{hdr} function from the \texttt{hdrcde R} package \citep{hdrcde}.
+The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Rates_estimates.R}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth,keepaspectratio]{SupplementaryFigures/Rates_estimates.pdf}
+\caption{Differences between the simulated rate of each morphological character ($\alpha$ parameter) and the Bayesian estimates for the trees with missing data.
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Positive values shows underestimation of the parameter and negative values shows overestimation.
+One unit of difference means that the estimated rate is twice as big/small than the true rate.
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the rates estimates.}
+\label{Fig_AppendixCharacters}
+\end{figure}
+
 \newpage
-\subsection{Tree Building Software settings}
+\subsection{Tree Inference Software settings}
 
@@
 
-\subsubsection{Maximum Likelihood: RAxML version 8.0.20 \citep{Stamatakis21012014}}
+\subsubsection{Maximum Likelihood: RAxML version 8.0.20 \cite{Stamatakis21012014}}
 
@@
   \item Molecular data: GTR + $\Gamma_4$ (-m GTRGAMMA)
-  \item Morphological data: Mk + $\Gamma_4$ (-K MK)
+  \item Morphological data: M\textit{kv} + $\Gamma_4$ (-K MK)
   \item Support: Rapid Boostrap algorithm (LSR), 1000 replicates
@@
 
-\subsubsection{Bayesian: MrBayes version 3.2.1 \citep{Ronquist2012mrbayes}}
+\subsubsection{Bayesian: MrBayes version 3.2.1 \cite{Ronquist2012mrbayes}}
 
@@
     \item Molecular data: HKY + $\Gamma_4$
-    \item Morphological data: Mk + $\Gamma_4$
+    \item Morphological data: M\textit{kv} + $\Gamma_4$
   \end{itemize}
@@
 \end{itemize}
+
+\bibliographystyle{sysbio}
+\bibliography{Supp_Ref}
+
+%END
+\end{document}

### 2015-07-09 Manuscript/MPESubmission/Revision/response.tex
updated response (namely the true tree choice justification to reviewers)

@@
 
-%To do!
+We replace 'inferred' by 'simulated' on lines @@@ and @@@ %LaTeX lines 192 'The molecular matrix was simulated' and 199 'We simulated the morphological matrix using the R'
 
@@
 
-%To do
+We changed the sentence to:
+"In practice, such a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal (Hassanin et al., 1998; Estoup et al., 2002).%\citep{Hassanin1998611,EstoupHomoplasy}.
+Also, we chose this low value to be consistent with our protocol for simulating morphological characters (see below)." (line @@@).
 
@@
 
-%To do
+We modified the sentence as follows: "We selected this model and parameters settings to be a compromise between realism from empirical datasets (e.g.
+Kelly et al., 2014) %\citep[e.g.][]{kellymolecular2014}
+and models with few parameters (cf. Springer et. al., 2012) %\citep[cf.][]{springermacroevolutionary2012}
+to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)."
 
@@
 
-%To do
+We modified the sentence as follows: "We simulated the morphological matrix using the rTraitDisc function from the R package ape v.
+3.0-11 (Paradis et al., 2004) to generate a matrix of 100 character sites for 51 taxa."
 
@@
 
-%To do
+% To do!
+% TG: a fair point... Can I just say: "Coz' April told me it was fine?"
+% TG: one way to justify it should be to say that we tried to avoid homoplasy but we actually wanted to add a little bit of it to be kind of realistic.
+% TG: in Wright & Hillis, low rate characters make the tree as "shitty" as the "shittiest" ML estimation.
+Now this "shittiness" is in quotes because it's actually a relative "shittiness".
+Using the Mk model with no missing data, the Topological difference is actually very low (so it's not "shitty") for low rate characters.
+% TG: then a low rate character with 75% missing gets the maximum topological difference with the Mk model (which is still all right).
+% TG: also, in our case, the actual character rate is not that low.
+We sample from a low distribution but force them to be non-invariable (therefore, our pratical minimum rate is 0.02 because for one character, at least one species must have a different state (1/51)).
 
@@
 
+We do agree with the point raised with both reviewers on comparing the true tree to the best/missing-data trees.
+As reviewer 1 states: "[this analysis can give] insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it.".
+We've added the results of these comparisons to the supplementary materials (Appendix 3 - @@@).
+However, even though we agree with the comments of both reviewer that knowing the difference between the true and best/missing-data trees is important, we do not think that it "adds an additional layer of complexity that can't be accounted for via simulation" (reviewer 2, comment 1).
+In this simulations, we had to make a choice between (1) simulating larger matrices but with less chains (i.e. less variance) or (2) smaller matrices with more variance.
+We deliberately made the second choice because our main question was "what is the effect of our parameters on recovering the topology".
+In fact, regarding our problematic, we maintain that comparing the best tree to the missing-data trees is the best approach to answer this specific question for two reasons:
+\begin{enumerate}
+\item{The true tree is always unknown:} as we state in the manuscript: "In real world situations, the ``true'' tree is never available" (line 242 in the first manuscript and @@@ in the revision).
+\item{The true tree is just the starting point of the simulations:} in fact, we use the true tree only as a starting point for our simulations.
+Once generated, the true tree is fed to a cascade of algorithms and softwares (i.e.
+seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes) described in section 2.1 and 2.3 in order to get the best tree and the missing-data trees.
+Comparing the latests to the true tree will answer to the problematic "what is the effect of our missing-data parameters and the cascade of algorithms and softwares on recovering the true topology?".
+\end{enumerate}
+Therefore we added the second justification to the main text (see below this response) and added these analysis (true vs.
+best/missing-data trees) to the supplementary for potential readers that might be interested by this specific problematic.
+
+We added the following justification in the main text:
+"blalbalbal" (line @@@).
+
 %To do (MAJOR)
+% Differences between True tree and best/missing-data trees.
+%"insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it" -> no, our parameters were not adequate.
+However this can be due to the size of the matrices.
+It is likely that phylo signal (ability to get close to the true tree) will increase with bigger matrices.
+Here we had to make a choice between larger matrices and less variance or smaller matrices and higher variances.
+We chose the second point because we were mainly interested in the effect of our parameters on recovering the topology.
+Because the true tree is just a starting tree for our simulations, it has not a lot of influence on the generation of the best tree.
+In other words, the best/true tree difference is more an assessment of the ability of the cascade of algorithms and softwares used (tree.bd{diversitree}, seqGen{phyclust}, rTraitDisc{ape}, RAxML, MrBayes) to perform with small size matrices.
 

### 2015-07-09 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
updated response (namely the true tree choice justification to reviewers)

@@
 Next, we generated a molecular and a morphological matrix from the ``true" tree.
-The molecular matrix was inferred from the ``true" tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
+The molecular matrix was simulated from the ``true" tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
 The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1) and transition/transversion rate of two \citep{douadycomparison2003}.
 The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
-We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal.
-We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, total computational time for the whole analysis was around 150 CPU years).
+In practice, such a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal \citep{Hassanin1998611,EstoupHomoplasy}.
+Also, we chose this low value to be consistent with our protocol for simulating morphological characters (see below).
+%We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal. %Justify
+%We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, total computational time for the whole analysis was around 150 CPU years).
+We selected this model and parameters settings to be a compromise between realism from empirical datasets \citep[e.g.][]{kellymolecular2014} and models with few parameters \citep[cf.][]{springermacroevolutionary2012} to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).
 All the molecular information for fossil taxa was replaced by missing data ("?").
 
-We inferred the morphological matrix using the R package ape v. 3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
+We simulated the morphological matrix using the rTraitDisc function from the R package ape v. 3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
 We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A and Fig.

### 2015-07-11 Manuscript/MPESubmission/Revision/response.tex
response update

@@
 
+We added an extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains.
+We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or using a random tree (default in MrBayes) as a starting tree.
+We see no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
+We made these data available in the supplementary materials Appendix A @@@ and we also added a precision in the manuscript:
+
+blabalbalbal. line @@@ 
 %To do (MAJOR)
@@
 
-%To do
+%To do - Probably needs to find a better cite.
 

### 2015-07-11 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
response update

@@
 
+\subsection{Effect of the starting tree in Bayesian inference}
+We used the ``true'' tree as a starting tree for the two MCMC chains in the Bayesian inference for decreasing the computing time.
+To assess if this had an effect on the topology of the ``best'' tree, we ran a sub-sample of trees using a different random tree for the two MCMC chains \citep[default MrBayes option;][]{Ronquist2012mrbayes}.
+We tested this effect on five trees with the five levels of missing data (i.e.
+first tree: $M_L$=0\%, $M_F$=0\% and $M_C$=0\%; second tree: $M_L$=10\%, $M_F$=10\% and $M_C$=10\%, etc.) on the first 20 simulation chains.
+We then compared the trees inferred using the ``true'' tree as a starting tree to the trees using a random tree as a starting tree using the normalised Robinson-Foulds and Triplets metrics.
+Additionally, we compared the trees inferred using a random tree as a starting tree to the ``best'' tree (inferred using the ``true'' tree as a starting tree) and to the ``true'' tree to see whether our results varied if we didn't used the ``true'' tree as a starting tree.
+
+\begin{table}[!ht]
+\caption{Normalised topological metric between trees using the ``true'' or a random tree as a starting tree for the MCMC chains.}
+\centering
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
+  \hline
+ Tree &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+    $M_L$=0\%; $M_F$=0\%; $M_C$=0\%          & $RF$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
+                                             & $Tr$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
+    $M_L$=10\%; $M_F$=10\%; $M_C$=10\%       & $RF$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
+                                             & $Tr$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
+    $M_L$=25\%; $M_F$=25\%; $M_C$=25\%       & $RF$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
+                                             & $Tr$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
+    $M_L$=50\%; $M_F$=50\%; $M_C$=50\%       & $RF$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
+                                             & $Tr$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
+    $M_L$=75\%; $M_F$=75\%; $M_C$=75\%       & $RF$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
+                                             & $Tr$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
+   \hline
+\end{tabular}
+\label{Tab_Results-Difference_methods}
+\end{table}
+
+%Incl table + incl figure
+
 \newpage

### 2015-07-13 Manuscript/MPESubmission/Revision/response.tex
Daily revision update

@@
 
-%To do
+We modified the sentence to:
+
+"Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
+It seems therefore more important to integrate a maximum of living taxa with morphological data as well as a maximum number of characters.
+Additionally, when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6)."
+
+%Original: The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and/or $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6).
 
@@
 
+We do understand the reviewers concern and added an analysis in the Appendix A (@@@) where we tested the effect of using the ``true'' tree as a starting tree.
+We found that there is no significant effect between using the ``true'' tree or a random tree (MrBayes default) as a starting tree on topological recovery.
+We refer to this supplementary analysis in the main text as follows:
+
+"Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A - @@@ for details on the analysis of the effect of the starting tree in Bayesian inference)."
+
 %To do (MAJOR)
@@
 The following statements that the posterior distribution should not be discarded does not leave the reader with a clear path that they should follow - if only the fixed tree topology is used to estimate divergence times, how can the substitution rate-scaled posterior distribution of highly probable trees be used for most comparative analyses which require time-scaled branch lengths?
-This paragraph requires further justification and this study does have important bearing on a subject that gets little attention.
-}}
+This paragraph requires further justification and this study does have important bearing on a subject that gets little attention. }}
 
+We revised the statements in the whole paragraph as follows:
+"Another practical implication of our results regards the tree inference methods.
+Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advice using the Bayesian consensus trees as a topological constraint for tree inferences using the Total-Evidence method such as tip-dating (e.g.
+Ronquist et al., 2012a; Wood et al., 2013; Matzke, 2014; although it is possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution).
+Additionally, using the Bayesian consensus tree rather than the Maximum Likelihood can reduce the amount of ``false positive'' topologies.
+As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve nodes weakly supported due to missing data than the Maximum Likelihood tree that is more likely to wrongly resolve such nodes (i.e.
+creating a ``false positive'' node).
+Note, however, that we do not suggest discarding the Bayesian posterior tree distribution once one chose to use the Bayesian consensus tree topology even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
+This can be particularly important because these trees will be invaluable for phylogenetic comparative analyses.
+For example a sub-sample of posterior trees distributions can be used to asses macroecological questions while better taking into account topological uncertainty (e.g.
+Fritz et al.
+2009 and Jetz et al.
+2012 trees used in Healy et al.
+2014)."
 %To do (rewrite)
@@
 }}
-
-%To do (rewrite)
 
@@
 What about implications for studies that have used molecular scaffold techniques to fix the topology of living taxa based on analyses of only DNA analyses and then fit the fossils into this fixed tree?
-Also 'further' is misspelled.
-}}
+Also 'further' is misspelled.  }}
+
+%Check:
+%Wood
+%Schrago
+%Pyron
+%Schrago
+%Beck,
+%Slater
+%Ronquist
+%-Wittenberg.et.al 2015 - Org Divers Evol (TEM)
+%-Carrizo&Catalano 2015 - Cladistics (TEM)
+%-Djernæs et al 2015 MPE (TEM)
+%-Arcila et al 2015 - MPE (TEM)
 
 %To do (rewrite) - typo ok
+
+%Original: Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the “best” tree topology than the Maximum Likelihood trees.
+Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
+The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
 

### 2015-07-13 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
response update
Daily revision update

@@
 We investigate three major factors that directly affect the completeness and the size of the morphological part of the matrix: the proportion of living taxa with no morphological data, the amount of missing data in the fossil record, and the overall number of morphological characters in the matrix.
-We infer phylogenies from complete matrices and from matrices with various amounts of missing data, and then compare missing data topologies to the ``best" tree topology inferred using the complete matrix.
+We infer phylogenies from complete matrices and from matrices with various amounts of missing data, and then compare missing data topologies to the ``best'' tree topology inferred using the complete matrix.
 
-We find that the number of living taxa with morphological characters and the overall number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the ``best" tree topology.
+We find that the number of living taxa with morphological characters and the overall number of morphological characters in the matrix, are more important than the amount of missing data in the fossil record for recovering the ``best'' tree topology.
 Therefore, we suggest that sampling effort should be focused on morphological data collection for living species to increase the accuracy of topological inference in a Total Evidence framework.
@@
 Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies \citep{zuckerkandl1965}.
-There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the ``best" approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
+There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the ``best'' approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
 Neither approach, however, uses all the available data.
@@
 We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa.
-We focus on the effect of missing data on our ability to recover a ``best" tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
+We focus on the effect of missing data on our ability to recover a ``best'' tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
 Although branch length estimation is also important \citep[namely for timing extinction and/or speciation events; e.g.][]{ronquista2012}, we do not consider branch lengths in this study.
@@
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
-We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the ``best" tree topology more so than minimizing the amount of missing data in the fossil record.
-Additionally, we find that the ability of Total Evidence methods to recover the ``best" tree topology is increased when using Bayesian methods.
+We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the ``best'' tree topology more so than minimizing the amount of missing data in the fossil record.
+Additionally, we find that the ability of Total Evidence methods to recover the ``best'' tree topology is increased when using Bayesian methods.
 
@@
 \item{Generating the matrix:} \label{step:generate_matrix} \\
-We randomly generated a birth-death tree (hereafter called the ``true" tree) and used it to simulate a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the ``complete" matrix).
+We randomly generated a birth-death tree (hereafter called the ``true'' tree) and used it to simulate a matrix containing both molecular and morphological data for living and fossil taxa (hereafter called the ``complete'' matrix).
 \item{Removing data:} \label{step:remove_data} \\
-We removed data from the morphological part of the ``complete" matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$).
-We call the resulting 125 matrices ``missing-data" matrices.
+We removed data from the morphological part of the ``complete'' matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$).
+We call the resulting 125 matrices ``missing-data'' matrices.
 \item{Estimating phylogenies:} \label{step:build_phylo} \\
-We inferred phylogenetic trees from the ``complete" matrix and from the 125 ``missing-data" matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the ``best" tree) and 125 trees inferred from the matrices with missing morphological data (hereafter called the ``missing-data" trees).
+We inferred phylogenetic trees from the ``complete'' matrix and from the 125 ``missing-data'' matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the ``best'' tree) and 125 trees inferred from the matrices with missing morphological data (hereafter called the ``missing-data'' trees).
 Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
 \item{Comparing topologies:} \label{step:compare_topo} \\
-We compared the ``best" tree to the ``missing-data" trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies
+We compared the ``best'' tree to the ``missing-data'' trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies
 \end{enumerate}
@@
 %\caption{{\bf Protocol outline.}
-%(1) We randomly generated a birth-death tree (the ``true" tree) and used it to simulate a matrix with no missing data (the ``complete" matrix).
-%(2) We removed data from the morphological part of the ``complete" matrix resulting in 125 ``missing-data" matrices.
+%(1) We randomly generated a birth-death tree (the ``true'' tree) and used it to simulate a matrix with no missing data (the ``complete'' matrix).
+%(2) We removed data from the morphological part of the ``complete'' matrix resulting in 125 ``missing-data'' matrices.
 %(3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
-%(4) We compared the ``missing-data" trees to the ``best" tree.
+%(4) We compared the ``missing-data'' trees to the ``best'' tree.
 %We repeated these four steps 50 times.}
@@
 \label{Generating_the_matrix}
-First we randomly generated a ``true" tree of 50 taxa in R v. 3.0.2 \citep{R302} using the package diversitree v. 0.9-6 \citep{fitzjohndiversitree2012}.
+First we randomly generated a ``true'' tree of 50 taxa in R v. 3.0.2 \citep{R302} using the package diversitree v. 0.9-6 \citep{fitzjohndiversitree2012}.
 We generated the tree using a birth death process by sampling speciation ($\lambda$) and extinction ($\mu$) rates from a uniform distribution (bounded between 0 and 1) but maintaining $\lambda$ $>$ $\mu$ \citep{paradistime-dependent2011}.
@@
 
-Next, we generated a molecular and a morphological matrix from the ``true" tree.
-The molecular matrix was simulated from the ``true" tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
+Next, we generated a molecular and a morphological matrix from the ``true'' tree.
+The molecular matrix was simulated from the ``true'' tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
 The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1) and transition/transversion rate of two \citep{douadycomparison2003}.
@@
 A1 within).
-We then ran an independent discrete character simulation for each character using the ``true" tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
+We then ran an independent discrete character simulation for each character using the ``true'' tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
 evolutionary rate) from one character state to another \citep{Pagel22011994}.
@@
 
-Finally, we combined the morphological and molecular matrices obtained from the ``true" tree.
-Hereafter we call this the ``complete" matrix, i.e.
+Finally, we combined the morphological and molecular matrices obtained from the ``true'' tree.
+Hereafter we call this the ``complete'' matrix, i.e.
 the matrix with no missing data except for the molecular data of the fossil taxa.
@@
 \label{Removing_data}
-We modified the ``complete" matrix to obtain matrices with missing data by randomly replacing data with ``?" in the morphological part of the matrices according to the following parameters:
+We modified the ``complete'' matrix to obtain matrices with missing data by randomly replacing data with ``?" in the morphological part of the matrices according to the following parameters:
 \begin{enumerate}
@@
 We created matrices using all parameter combinations resulting in 125 ($5^3$) matrices.
-Note that one of these combinations has no missing data so is equivalent to the ``complete" matrix, thus we have one effectively complete matrix in our 125 ``missing-data" matrices.
+Note that one of these combinations has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
 To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
@@
 \subsection{Estimating phylogenies}
-From the resulting matrices we generated two types of trees: the ``best" tree inferred from the ``complete" matrix and the ``missing-data" trees inferred from the 125 matrices with various amounts of missing data.
-The ``true" tree was used to generate the ``complete" matrix and reflects the ``true" evolutionary history in our simulations.
-The ``best" tree, on the other hand, is the best tree we can build using state-of-the-art phylogenetic methods.
-In real world situations, the ``true" tree is never available to us because we cannot know the true evolutionary history of a clade \citep[except in very rare circumstances, e.g.][]{rozen2005}.
-Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the ``best" tree, rather than the ``true" tree, as the ``best" tree is generally what we have to work with.
+From the resulting matrices we generated two types of trees: the ``best'' tree inferred from the ``complete'' matrix and the ``missing-data'' trees inferred from the 125 matrices with various amounts of missing data.
+The ``true'' tree was used to generate the ``complete'' matrix and reflects the ``true'' evolutionary history in our simulations.
+The ``best'' tree, on the other hand, is the best tree we can build using state-of-the-art phylogenetic methods.
+In real world situations, the ``true'' tree is never available to us because we cannot know the true evolutionary history of a clade \citep[except in very rare circumstances, e.g.][]{rozen2005}.
+Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the ``best'' tree, rather than the ``true'' tree, as the ``best'' tree is generally what we have to work with.
 
 \subsubsection{Maximum Likelihood}
-The ``best" tree and the ``missing-data" trees were inferred using RAxML v.
+The ``best'' tree and the ``missing-data'' trees were inferred using RAxML v.
 8.0.20 \citep{Stamatakis21012014}.
@@
 
-To measure the support for each branch in our simulated phylogenies we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the ``complete" matrix.
+To measure the support for each branch in our simulated phylogenies we first ran a fast bootstrap analysis (Lazy Sub-tree Rearrangement) with 500 replicates on the ``complete'' matrix.
 We removed all the simulations with a median bootstrap support lower than 50 as a proxy for weak phylogenetic signal \citep{zanderminimal2004}.
-We repeated this selection until we obtained 50 sets of simulations (i.e. 50 ``complete" and 50 x 125 ``missing-data" matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
+We repeated this selection until we obtained 50 sets of simulations (i.e.
+50 ``complete'' and 50 x 125 ``missing-data'' matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
 This step was implemented to make sure that the differences we observed in topologies (see below) were due to the amount of missing data for each parameter ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
@@
 \subsubsection{Bayesian Inference}
-The ``best" tree and the ``missing-data" trees were inferred using MrBayes v. 3.2.1 \citep{Ronquist2012mrbayes}.
+The ``best'' tree and the ``missing-data'' trees were inferred using MrBayes v. 3.2.1 \citep{Ronquist2012mrbayes}.
 We partitioned the data to treat the molecular part as a non-codon DNA partition and the morphological part as a multi-state morphological partition.
@@
 Note that MrBayes automatically corrects for acquisition bias in the morphological data partition \citep{Nylander01022004,Ronquist2012mrbayes}.
-We chose these models to be consistent with the parameters used to generate the ``complete" matrix.
+We chose these models to be consistent with the parameters used to generate the ``complete'' matrix.
 
 Each Bayesian tree was estimated using two runs of four chains each for a maximum of 5$\times$$10^7$ generations.
-For each estimation, we used the ``true" tree's topology as a starting tree (with a starting value for each branch length of one).
+For each estimation, we used the ``true'' tree's topology as a starting tree (with a starting value for each branch length of one).
 We also used two priors on the molecular part of the matrix: an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5, and a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)); and one prior on the morphological part of the matrix (exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5).
@@
 \subsection{Comparing topologies}
-We compared the topology of the ``missing-data" trees to the ``best" tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
+We compared the topology of the ``missing-data'' trees to the ``best'' tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
 We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa \citep[i.e.
@@
 \label{tree_comparisons}
-For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the ``best" tree and each ``missing-data" tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting in 125 Normalised Robinson-Foulds metrics and 125 Normalised Triplets metric for each tree inference method.
-Also, to take into account the uncertainty of tree inference, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 trees from the posterior tree distribution of the Bayesian analysis for the ``best" trees, and then did the same for the 125 ``missing data" trees (resulting in 1000 ``best" trees and 125$\times$1000 ``missing data" trees).
-For a given set of 1000 ``missing data" trees and the 1000 ``best" trees, we sampled one ``missing data" tree and one ``best" tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above.
+For the Maximum Likelihood and Bayesian consensus trees we performed pairwise comparisons between the ``best'' tree and each ``missing-data'' tree using both the Normalised Robinson-Foulds and Normalised Triplets metrics with the TreeCmp java script \citep{Bogdanowicz2012} resulting in 125 Normalised Robinson-Foulds metrics and 125 Normalised Triplets metric for each tree inference method.
+Also, to take into account the uncertainty of tree inference, we extracted 1000 random bootstrapped trees from the Maximum Likelihood analysis and 1000 trees from the posterior tree distribution of the Bayesian analysis for the ``best'' trees, and then did the same for the 125 ``missing data" trees (resulting in 1000 ``best'' trees and 125$\times$1000 ``missing data" trees).
+For a given set of 1000 ``missing data" trees and the 1000 ``best'' trees, we sampled one ``missing data" tree and one ``best'' tree at random and compared them using both the Normalised Robinson-Foulds and Normalised Triplets metrics as described above.
 We repeated this 1000 times for each set of ``missing data" trees resulting in 125$\times$1000 values for each metric.
@@
 \subsection{Testing the effects of the missing data parameters on topological recovery}
-Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the ``best" tree topology in a Total Evidence framework.
+Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the ``best'' tree topology in a Total Evidence framework.
 We also assessed the effect of our missing data parameters jointly with the effects of different tree inference and uncertainty methods (i.e.
@@
 We performed this procedure separately for each tree inference and uncertainty method.
-When two combinations of missing data parameters have a similar ability to recover the ``best" tree topology the Bhattacharyya Coefficient will be close to one.
+When two combinations of missing data parameters have a similar ability to recover the ``best'' tree topology the Bhattacharyya Coefficient will be close to one.
 Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
@@
 To assess the effect of the different tree inference and uncertainty methods (i.e.
-Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the ``best" tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of tree inference and uncertainty methods, i.e.
+Maximum Likelihood, Bayesian consensus, Maximum Likelihood bootstrap trees and Bayesian posterior tree distribution) on our ability to recover the ``best'' tree topology, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of tree inference and uncertainty methods, i.e.
 Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc.
@@
 Note that this procedure pools results from across all missing data parameter combinations so it results in just six pairwise comparisons.
-When two tree inference or uncertainty methods have a similar ability to recover the ``best" tree topology the Bhattacharyya Coefficient will be close to one.
+When two tree inference or uncertainty methods have a similar ability to recover the ``best'' tree topology the Bhattacharyya Coefficient will be close to one.
 Conversely, if the two tree inference or uncertainty methods differ, the Bhattacharyya Coefficient will be close to zero.
@@
 \section{Results}
-As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best" tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets metric).
+As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best'' tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets metric).
 Nonetheless, the different missing data parameters and tree inference methods do not affect the topology in the same way (Figure 4 and Figure 5).
@@
 \subsection{Individual effects of missing data parameters}
-As the amount of missing data increases across all three parameters, our ability to recover the ``best" tree topology decreases (Figure 4).
+As the amount of missing data increases across all three parameters, our ability to recover the ``best'' tree topology decreases (Figure 4).
 The Normalised Robinson-Foulds metric is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
@@
 \subsection{Combined effect of missing data parameters}
-As expected, our ability to recover the ``best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
+As expected, our ability to recover the ``best'' tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
@@
 % Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the ``best" tree distribution, calculated using the Bhattacharyya Coefficient.
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the ``best'' tree distribution, calculated using the Bhattacharyya Coefficient.
 The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
@@
 
-Our results show that the ability to recover the ``best" tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used.
+Our results show that the ability to recover the ``best'' tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used.
 These factors, however, affected topological recovery in different ways and to different extents.
 Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Figure 6).
-Additionally, using Bayesian consensus trees recovered the ``best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Figure 5, Figure 6, Table ~\ref{Tab_Results-Difference_methods}).
+Additionally, using Bayesian consensus trees recovered the ``best'' tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Figure 5, Figure 6, Table ~\ref{Tab_Results-Difference_methods}).
 As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
@@
 \subsubsection{Missing morphological characters ($M_{C}$)}
-Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the ``best" tree topology.
+Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the ``best'' tree topology.
 We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
 Conversely, removing 10\% of morphological characters ($M_{C}$) genuinely removes 10\% of data in the matrix.
-Nonetheless, the effect of removing characters on the ability to recover the ``best" tree topology is of the same order of magnitude as for the other two parameters (Figure 4).
+Nonetheless, the effect of removing characters on the ability to recover the ``best'' tree topology is of the same order of magnitude as for the other two parameters (Figure 4).
 We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.
@@
 Additionally, the number of morphological characters determines the size the of matrix.
-This can affect our ability to recover the ``best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
+This can affect our ability to recover the ``best'' tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
 The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \citep{wagner2000}.
@@
 \subsection{Combined effect of missing data parameters}
-As expected, when combining the missing data parameters, our ability to recover the ``best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e.
+As expected, when combining the missing data parameters, our ability to recover the ``best'' tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a larger proportion of missing data (i.e.
@@
 It is important, however, to notice that the effect of each parameter is not additive.
-The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and/or $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6).
+Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
+It seems therefore more important to integrate a maximum of living taxa with morphological data as well as a maximum number of characters.
+Additionally, when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6).
 
 \subsection{Effects of tree inference methods}
-Variation in our ability to recover the ``best" tree topology depends heavily on the tree inference method (Figure 4 and Figure 5).
+Variation in our ability to recover the ``best'' tree topology depends heavily on the tree inference method (Figure 4 and Figure 5).
 For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over parsimony methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
-This is, however, the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best" tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-Our results show that the topology of the Bayesian consensus tree is always closer to the ``best" tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
+This is, however, the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best'' tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
+Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A - @@@ for details on the analysis of the effect of the starting tree in Bayesian inference).
 As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
-This may also be because the ``best" Bayesian consensus trees are not completely resolved, thus will always be more similar to the ``missing data" trees than a completely resolved tree like the ``best" Maximum Likelihood tree.
-Nonetheless, we minimized the probability of unresolved ``best" trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
+This may also be because the ``best'' Bayesian consensus trees are not completely resolved, thus will always be more similar to the ``missing data" trees than a completely resolved tree like the ``best'' Maximum Likelihood tree.
+Nonetheless, we minimized the probability of unresolved ``best'' trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
 
-It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best" tree topology than Maximum Likelihood and Bayesian consensus trees.
+It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best'' tree topology than Maximum Likelihood and Bayesian consensus trees.
 This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $M_{C}$ = 0\%; Figure 4).
-This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best" trees involved 1000 random pairwise comparisons rather than just one.
+This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best'' trees involved 1000 random pairwise comparisons rather than just one.
 This will obviously add noise to the results for these methods.
@@
 Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
-Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the ``best" tree topology.
+Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the ``best'' tree topology.
 We can increase the number of living taxa with coded morphological characters by increasing research efforts in this area, and encouraging use of our vast natural history collections.
@@
 Gaps in the matrix, however, can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
-Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the ``best" tree topology (Figure 6).
+Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the ``best'' tree topology (Figure 6).
 Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
@@
 Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the ``best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in tip-dating \citep[][; although it is possible that including dating information during tree inference could improve the accuracy of the Bayesian posterior tree distribution]{ronquista2012,BEASTmaster}.
-Note that, although the Bayesian posterior tree distribution performed poorly in recovering the ``best" tree topology, this is due to difficulties comparing distributions of trees (see above).
-Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses \citep[e.g.][]{jetzthe2012}.
+Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advice using the Bayesian consensus trees as a topological constraint for tree inferences using the Total-Evidence method such as tip-dating \citep[e.g.][; although it is possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution]{ronquista2012,Wood01032013,BEASTmaster}.
+Additionally, using the Bayesian consensus tree rather than the Maximum Likelihood can reduce the amount of ``false positive'' topologies.
+As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve nodes weakly supported due to missing data than the Maximum Likelihood tree that is more likely to wrongly resolve such nodes (i.e.
+creating a ``false positive'' node).
+Note, however, that we do not suggest discarding the Bayesian posterior tree distribution once one chose to use the Bayesian consensus tree topology even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
+This can be particularly important because these trees will be invaluable for phylogenetic comparative analyses.
+For example a sub-sample of posterior trees distributions can be used to asses macroecological questions while better taking into account topological uncertainty (e.g.
+\citealt{FritzTree} and \citealt{jetzthe2012} trees used in \citealt{healy2014}).
 
 \section{Conclusions}
-Missing data in Total Evidence matrices is not a problem for recovering the ``best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
+Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
 When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
@@
 Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa.
-Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best" tree topology than the Maximum Likelihood trees.
+Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
 Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for further analyses.
@@
 Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
-Thanks to the two anonymous reviewers for their useful and enthusiastic comments.
+Thanks to the two anonymous reviewers for their useful and enthusiastic %or "encouraging"
+comments.
 Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.
@@
 \textbf{Figure 1:} Protocol outline.
-(1) We randomly generated a birth-death tree (the ``true" tree) and used it to simulate a matrix with no missing data (the ``complete" matrix).
-(2) We removed data from the morphological part of the ``complete" matrix resulting in 125 ``missing-data" matrices.
+(1) We randomly generated a birth-death tree (the ``true'' tree) and used it to simulate a matrix with no missing data (the ``complete'' matrix).
+(2) We removed data from the morphological part of the ``complete'' matrix resulting in 125 ``missing-data'' matrices.
 (3) We built phylogenetic trees from each matrix using both Maximum Likelihood and Bayesian methods.
-(4) We compared the ``missing-data" trees to the ``best" tree.
+(4) We compared the ``missing-data'' trees to the ``best'' tree.
 We repeated these four steps 50 times.\\
@@
 Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the ``best" tree distribution, calculated using the Bhattacharyya Coefficient.
+Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the ``best'' tree distribution, calculated using the Bhattacharyya Coefficient.
 The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
@@
 \textbf{Table 1:} Bhattacharyya Coefficients of the pairwise method comparisons.
-Each comparison which corresponds to the normalised metric between the ``best" tree and the ``missing data" trees using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).
+Each comparison which corresponds to the normalised metric between the ``best'' tree and the ``missing data" trees using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).
 Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.

### 2015-07-13 Analysis/Rates_estimates.R
flipped over/under estimation

@@ -64,12 +64,12 @@ True_tab<-True_tab[-c(3,31),]
 #Calculating the difference between the estimate the true rate
 differences<-matrix(data=NA, ncol=48, nrow=125)
 for(n in 1:48) {
-    differences[,n]<-True_tab[n,1]-Estimates_rates[[n]][,1]
+    differences[,n]<-Estimates_rates[[n]][,1]-True_tab[n,1]
 }
 
 differences_theoretical<-matrix(data=NA, ncol=48, nrow=125)
 for(n in 1:48) {
-    differences_theoretical[,n]<-0.5-Estimates_rates[[n]][,1]
+    differences_theoretical[,n]<-Estimates_rates[[n]][,1]-0.5
 }
 
 #Calculate the hdr for each tree
@@ -80,7 +80,7 @@ for (i in 1:125) {
 
 
 xlab=""
-ylabels="Gamma shape difference (True - Estimate)"
+ylabels="Gamma shape difference (Estimate - True)"
 ylim=c(-1,1)
 probs=c(95,50)
 

### 2015-07-16 Manuscript/MPESubmission/Revision/response.tex
Main text changes

@@
 
+Unfortunately, we didn't found many studies formally testing the important questions asked by the reviewer. However, there exist few empirical analysis that contain results supporting our claims.
+Firstly, for the question about the topological differences between Maximum Likelihood and the Bayesian consensus tree, in Arcila et al 2015 (MPE)'s supplementary analysis, the authors compare the topology inferred from a Total Evidence matrix in both Likelihood (RAxML) and Bayesian (MrBayes) and the topology of the Bayesian consensus tree is much better supported than the one of the best ML tree.
+Specifically, the ML tree has 7/28 nodes have a "fair" bootstrap support ($>$ 75) and 3/28 with a good support ($>$ 90) as the Bayesian consensus tree has 17/25 nodes with a "fair" posterior probability ($>$80) and 14/25 with a good posterior probability ($>$95).
+We therefore used this reference to strengthen our claim of the superiority of the Bayesian consensus tree over the Maximum Likelihood one.
+Secondly, for the question about molecular scaffolds, from the few studies clearly specifying their topological constraint method, only Beck and Lee 2013 (Proceedings B) seems to compare an analysis with and without taxonomic constraint on divergence time and rates estimation (the constraint seems to have no effect) but the branch length aspect is not covered in our study.
+
+We modified the mentioned section of the conclusion as follows:
+"Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+This has also been observed in empirical data (e.g. Arcila et al., 2015) were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds (e.g.
+Slater, 2013), taxonomic constraints (e.g.
+Slater, 2013; Beck and Lee, 2014) or even by fixing the topology (e.g.
+Ronquist et al., 2012a).
+Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed."
 %Check:
+% Study: method 
 %Wood 
@@
 %Pyron
-%Schrago
 %Beck,
@@
 %-Djernæs et al 2015 MPE (TEM)
-%-Arcila et al 2015 - MPE (TEM)
-
-%To do (rewrite) - typo ok
+%-Arcila et al 2015 - MPE (TEM) This study suggests that current implementations of tip dating may overestimate ages of divergence in calibrated phylogenies + cool figure 1
 

### 2015-07-16 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Main text changes

@@
 Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa.
+
 Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
-Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for further analyses.
+This has also been observed in empirical data \cite[e.g.][]{Arcila2015131} were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds \citep[e.g.][]{slaterphylogenetic2013}, taxonomic constraints \citep[e.g.][]{slaterphylogenetic2013,beckancient2014} or even by fixing the topology \cite[e.g.][]{ronquista2012}.
+Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.
+
 The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.

### 2015-07-16 Analysis/MissingData_plot.R
Updated figures

@@ -37,29 +37,30 @@ ylim=c(0,1)
 
 #data parameters
 data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
+pal<-c("black", "grey", "blue", "orange")
 metrics<-c("R.F_Cluster","Triples")
 
 #2*3 plots
 #Missing living RF
 par(mar=c(1,5,5,0) + 0.1)
-multi.TreeCmp.plot( data.sets , par_ML, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab="Robinson-Foulds distance", las=2)
+multi.TreeCmp.plot( data.sets , par_ML, metrics[1], col=pal, las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab="Normalised Robinson-Foulds metric", las=2)
 #Missing fossil RF
 par(mar=c(1,3,5,2) + 0.1)
-multi.TreeCmp.plot( data.sets , par_MF, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
+multi.TreeCmp.plot( data.sets , par_MF, metrics[1], col=pal, las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
 #Missing character RF
 par(mar=c(1,1,5,4) + 0.1)
-multi.TreeCmp.plot( data.sets , par_MC, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
+multi.TreeCmp.plot( data.sets , par_MC, metrics[1], col=pal, las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
 
 xaxis.labs<-c("0%", "10%", "25%", "50%", "75%")
 #Missing living Triplets
 par(mar=c(5,5,1,0) + 0.1)
-multi.TreeCmp.plot( data.sets , par_ML, metrics[2], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=xaxis.labs, ylab="Triplets distance", las=2, xlab="Missing living taxa") 
[…]

### 2015-07-16 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Updated figures

@@
 \noindent
-\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange).
+\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange).
 The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.

### 2015-07-16 Analysis/MissingData_plot.R

@@ -65,100 +65,6 @@ multi.TreeCmp.plot( data.sets , par_MC, metrics[2], col=pal, las=2, probs=c(95, 
 #Back to default
 par(op)
 
-
-#####################
-#Plot method differences figure
-#####################
-#windows parameters
-quartz(width = 8.3, height = 5.8) #A5 landscape
-#quartz(width = 16.6, height = 11.6) #A4 landscape
-op<-par(mfrow=c(2,3), bty="l", oma=c(0,0,0,0))
-ylim=c(0,1)
-
-#data parameters
-data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
-metrics<-c("R.F_Cluster","Triples")
-
-#2*3 Plots
-#Missing living RF
-par(mar=c(1,5,5,0) + 0.1)
-multi.TreeCmp.plot( data.sets , par_ML, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab="Robinson-Foulds distance", las=2)
-signi.TreeCmp.plot( data.sets , par_ML, metrics[1], probs=c(95, 50), diff.type='method', position='below', col=palette(), pch='default', shift='auto')
-
-#Missing fossil RF
-par(mar=c(1,3,5,2) + 0.1)
-multi.TreeCmp.plot( data.sets , par_MF, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
-signi.TreeCmp.plot( data.sets , par_MF, metrics[1], probs=c(95, 50), diff.type='method', position='below', col=palette(), pch='default', shift='auto')
-
-#Missing character RF
-par(mar=c(1,1,5,4) + 0.1)
-multi.TreeCmp.plot( data.sets , par_MC, metrics[1], col=palette(), las=2, probs=c(95, 50), ylim=ylim, xaxis=NULL, xlab=NULL, ylab=NULL, las=2)
-signi.TreeCmp.plot( data.sets , par_MC, metrics[1], probs=c(95, 50), diff.type='method', position='below', col=palette(), pch='default', shift='auto')
-
-xaxis.labs<-c("0%", "10%", "25%", "50%", "75%")
-#Missing living Triplets
-par(mar=c(5,5,1,0) + 0.1)
[…]

### 2015-07-16 Functions/TreeComparisons/TreeCmp.Plot.R

@@ -707,7 +707,7 @@ global.TreeCmp.plot<-function(data.list, parameter, metrics, col, col.grad=c('bl
 
     #Plot the TreeCmp results
     par(mar=c(0,4,2,2), bty="l")
-    multi.TreeCmp.plot(data.list=data.list, parameter=parameter, metrics=metrics[2], col=col, shift=shift, ylab="Normalised Triplets distance metric", ylim=ylim2, ...)
+    multi.TreeCmp.plot(data.list=data.list, parameter=parameter, metrics=metrics[2], col=col, shift=shift, ylab="Normalised Triplets metric", ylim=ylim2, ...)
     #Add the 0 line if ylim1 < 0
     if(min(ylim2) < 0 ) {
         abline(h=0, lty=3)

### 2015-07-16 Manuscript/MPESubmission/Revision/response.tex
To do: update the figure 6 explanations (in methods+results)

@@
 We apologies for this silly mistake and thanks the reviewer for pointing it out.
-We changed the color scheme to "Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
-%To do: change the graphs colours
+We changed the colors to match with the paper color scheme: "Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
 
@@
 For the Bayesian analyses, many results had values < 0 but the axis is set to 0.
-Why not illustrate the distribution of values that is < 0?
-From this result, it appears the Bayesian analysis does quite poorly with missing data in terms of rouge taxa, and worse than max.
-likelihood.
+Why not illustrate the distribution of values that is < 0? %Fixed
+From this result, it appears the Bayesian analysis does quite poorly with missing data in terms of rouge taxa, and worse than max. likelihood.
 This figure also suggests to me that the methods perform pretty badly with missing data, in contrast to previous findings that missing data are not overly deleterious.
@@
 We thank the reviewer for this excellent suggestion and modified our graphical abstract.
+We also fixed the Normalised Triplets metrics axis to range from -0.5 to 1.
+We added a horizontal doted line at Triplets=0 to make sure the reader take notes is this change of scale.
+
 %To do (writing)
@@
 It took me a long time to figure it out, however, and to figure out why the top of the triangle is all orange.
-I would have expected something more like in Appendix C Figure C2 or Fig.
-C3.
+I would have expected something more like in Appendix C Figure C2 or Fig. C3.
 At first I understood this to indicate high probability of overlap of the missing data trees which had the most missing data with the best tree.
@@
 
+We thank the reviewer for pointing out this difficulty and acknowledge that such graphical representations of so many pairwise comparisons might be confusing.
+We emphasized the figure caption, the method section explaining how we calculated this figure and the results.
+
+We modified the figure caption as follows:
+"The effects of missing data on topological recovery using Bayesian consensus trees.
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
+Blue regions denotes a poor overlap in Normalised metric between the different parameters combinations (i.e.
+the parameters have a strong effect on the Normalised metric and thus the topological recovery).
+Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
+the parameters have a weak effect on the Normalised metric and thus the topological recovery)."
+
+We modified the methods section as follows:
+""
+
+We modified the results section as follows:
+""
 %To do (writing)

### 2015-07-16 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
To do: update the figure 6 explanations (in methods+results)

@@
 
+%Add figure explanation
+The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
+Blue regions denotes a poor overlap in Normalised metric between the different parameters combinations (i.e.
+the parameters have a strong effect on the Normalised metric and thus the topological recovery).
+Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
+the parameters have a weak effect on the Normalised metric and thus the topological recovery).
+For example, for the Normalised Robinson-Foulds metric, the tree with all the missing data parameters at 0\% have a low probability of have it's Normalised Robinson-Foulds metric overlapping with any other tree (deep blue line at the base of the graph).
+Conversely, each parameter combination compared to itself display a complete overlap of Normalised Robinson-Foulds metric (Bhattacharyya Coefficient = 1).
+
 %\begin{figure}[h]
@@
 Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ (Figure 6A).
+
+%Add figure explanation
+The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.
+The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
+Blue regions denotes a poor overlap in Normalised metric between the different parameters combinations (i.e.
+the parameters have a strong effect on the Normalised metric and thus the topological recovery).
+Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
+the parameters have a weak effect on the Normalised metric and thus the topological recovery).
+For example, for the Normalised Robinson-Foulds metric, the tree with all the missing data parameters at 0\% have a low probability of have it's Normalised Robinson-Foulds metric overlapping with any other tree (deep blue line at the base of the graph).
+Conversely, each parameter combination compared to itself display a complete overlap of Normalised Robinson-Foulds metric (Bhattacharyya Coefficient = 1).
 
@@
 Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the ``best'' tree distribution, calculated using the Bhattacharyya Coefficient.
+The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.
 The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
+Blue regions denotes a poor overlap in Normalised metric between the different parameters combinations (i.e.
+the parameters have a strong effect on the Normalised metric and thus the topological recovery).
+Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
+the parameters have a weak effect on the Normalised metric and thus the topological recovery).
 

### 2015-07-17 Manuscript/MPESubmission/Revision/response.tex
Updated revisions

@@
 Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a weak effect on the Normalised metric and thus the topological recovery)."
+the parameters have a weak effect on the Normalised metric and thus the topological recovery)." (lines @@@)
 
 We modified the methods section as follows:
-""
+"Because of the difficulties to represent so many pairwise comparisons in a meaningful way, we summarized as a heat map of Bhattacharyya Coefficients (see Figure 6).
+In this type of figures, parameters that have similar or different effects on recovering the ``best'' topology (rather good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6)." (lines @@@)
 
 We modified the results section as follows:
-""
+"Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
+The Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; blue regions in Figure 6A), however, displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; blue regions in Figure 6B).
+The orange upper triangle in Figure 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Figure 6A).
+There is no additional effect (i.e. no dissimilarities) of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A)." lines(@@)
 %To do (writing)

### 2015-07-17 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Updated revisions

@@
 Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
+Because of the difficulties to represent so many pairwise comparisons in a meaningful way, we summarized as a heat map of Bhattacharyya Coefficients (see Figure 6).
+In this type of figures, parameters that have similar or different effects on recovering the ``best'' topology (rather good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6).
 
@@
 Conversely, if the two tree inference or uncertainty methods differ, the Bhattacharyya Coefficient will be close to zero.
-
-%Add figure explanation
-The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
-Blue regions denotes a poor overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a strong effect on the Normalised metric and thus the topological recovery).
-Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a weak effect on the Normalised metric and thus the topological recovery).
-For example, for the Normalised Robinson-Foulds metric, the tree with all the missing data parameters at 0\% have a low probability of have it's Normalised Robinson-Foulds metric overlapping with any other tree (deep blue line at the base of the graph).
-Conversely, each parameter combination compared to itself display a complete overlap of Normalised Robinson-Foulds metric (Bhattacharyya Coefficient = 1).
 
@@
 Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Figure 6).
-The Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; Figure 6A), however, displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; Figure 6B).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
+The Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; blue regions in Figure 6A), however, displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; blue regions in Figure 6B).
+The orange upper triangle in Figure 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Figure 6A).
+There is no additional effect (i.e. no dissimilarities) of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
 Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ (Figure 6A).
+Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
 
-%Add figure explanation
-The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
-Blue regions denotes a poor overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a strong effect on the Normalised metric and thus the topological recovery).
-Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a weak effect on the Normalised metric and thus the topological recovery).
-For example, for the Normalised Robinson-Foulds metric, the tree with all the missing data parameters at 0\% have a low probability of have it's Normalised Robinson-Foulds metric overlapping with any other tree (deep blue line at the base of the graph).
-Conversely, each parameter combination compared to itself display a complete overlap of Normalised Robinson-Foulds metric (Bhattacharyya Coefficient = 1).
+%Adds that it correspond to the "line" of distributions in figure 5 (upper grey distributions)
 

### 2015-07-22 Manuscript/MPESubmission/Revision/response.tex
Updated revisions

@@
 
+% TG: I don't understand this comment at all. What does "The wording of the caption is off" means? And which pararagraph is he talking about?
+We fixed the highlithing in bold to the classical values of 0.05 or 0.95.
+We changed the caption to:
+"Bhattacharyya Coefficients of the pairwise method comparisons.
+Each comparisons corresponds to the comparison of the full (125) distributions between the ``best'' tree and the ``missing data'' trees for all the four methods (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) and the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
+Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
+The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
+If two methods have a high probability of overlap, they have a similar ability to recover the ``correct" tree topology"
+
+1) We are not entirely sure if the reviewer means (i) why the two methods display such differences ; or (ii) why, within each method, the combination of parameters displays a difference between each single parameters ($M_LM_FM_C$ vs.
+$M_L$, $M_F$ and $M_C$).
+In the first case, we think that the difference between the methods is due to the fact that the Maximum Likelihood trees are more likely to display ``false positive'' topologies (i.e.
+clades that are incorrect but can be well supported; or not) as the Bayesian consensus tree, when many data is missing, are more likely to collapse nodes, resulting in trees with less resolution but more ``conserved clades'' (i.e.
+the clade ((a;b);c)) could be collapsed in (a;b;c) which is still a corrected representation of the relations within the clades rather than ((b;c);a)).
+In the second case, the fact that the missing data parameters display better scores alone is because in practice, when this occured, the two other parameters where set a 0\%, therefore resulting in matrices with a maximum of 75\%/2 missing data for $M_L$ and $M_F$ an 75\% maximum for $M_C$.
+When combined, the maximum amount of missing data can be higher, especially when each parameter is set a 75\% missing data.
+
+2) %TG: not entirely sure what the reviewer means by marginal probabilities.
+I made to extra summarizing tables in the supplementaries that may be actually easier to read than most of the others so we might want to add them to the main text.
+
 \item{\textcolor{blue}{\textbf{Appendix C:} Tables 1-3 repeats 'of the'. Should label the first two rows as 'Best maximum likelihood tree' and 'Best Bayesian inference tree'}}

### 2015-07-22 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixC.tex
Updated revisions

@@
 \end{landscape}
+
+\begin{table}[ht]
+\caption{Summary of the median Bhattacharyya Coefficients comparisons between the missing data parameters.
+Each value corresponds to the median Bhattacharyya Coefficient between the Normalised metrics (Robinson-Foulds - RF; Triplets - Tr) between the ``best'' and the ``missing data'' trees (as outlined in Figure 2 in the main text) for the the trees with: Missing Living taxa only ($M_L$); Missing Fossil data only ($M_F$); Missing characters only ($M_C$); and the combination of the three parameters ($M_LM_FM_C$).}
+\begin{tabular}{llcccc}  
+  \hline
+ Method & Metric & $M_L$ & $M_F$ & $M_C$ & $M_LM_FM_C$ \\ 
+  \hline
+Maximum Likelihood & RF & 0.54 & 0.37 & 0.56 & 0.54 \\ 
+                   & Tr & 0.80 & 0.75 & 0.80 & 0.79 \\ 
+Bayesian consensus & RF & 0.73 & 0.59 & 0.60 & 0.79 \\ 
+   & Tr & 0.86 & 0.67 & 0.81 & 0.81 \\ 
+Maximum Likelihood bootstraps & RF & 0.98 & 0.96 & 0.97 & 0.78 \\ 
+   & Tr & 0.99 & 1.00 & 1.00 & 0.96 \\ 
+Bayesian posterior trees & RF & 0.86 & 0.93 & 0.94 & 0.68 \\ 
+  & Tr & 0.98 & 0.99 & 0.99 & 0.91 \\ 
+ \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\caption{Summary of the median Bhattacharyya Coefficients comparisons between the missing data parameters and the methods.
+Each value corresponds to the median Bhattacharyya Coefficient between the Normalised metrics (Robinson-Foulds - RF; Triplets - Tr) between the `methods (as outlined in Figure 3 in the main text) for the the trees with: Missing Living taxa only ($M_L$); Missing Fossil data only ($M_F$); Missing characters only ($M_C$); and the combination of the three parameters ($M_LM_FM_C$).}
+\begin{tabular}{llcccc}  
+  \hline
+ Method & Metric & $M_L$ & $M_F$ & $M_C$ & $M_LM_FM_C$ \\ 
+  \hline
+Maximum likelihood vs. Bayesian consensus & RF & 0.69 & 0.48 & 0.66 & 0.10 \\ 
+   & Tr & 0.84 & 0.75 & 0.80 & 0.61 \\ 
+Maximum likelihood vs. Bootstraps & RF & 0.29 & 0.32 & 0.21 & 0.69 \\ 
+   & Tr & 0.53 & 0.40 & 0.39 & 0.65 \\ 
+Maximum likelihood vs. Bayesian posterior trees & RF & 0.61 & 0.71 & 0.47 & 0.80 \\ 
+  & Tr & 0.70 & 0.62 & 0.59 & 0.73 \\ 
+Bayesian consensus vs. Bootstraps & RF & 0.02 & 0.01 & 0.01 & 0.00 \\ 
+  & Tr & 0.78 & 0.54 & 0.62 & 0.59 \\ 
+Bayesian consensus vs. Bayesian posterior trees & RF & 0.02 & 0.02 & 0.04 & 0.01 \\ 
+   & Tr & 0.75 & 0.66 & 0.64 & 0.56 \\ 
+Boostraps vs. Bayesian posterior trees & RF & 0.85 & 0.72 & 0.73 & 0.85 \\ 
+   & Tr & 0.96 & 0.91 & 0.93 & 0.98 \\ 
+ \hline
+\end{tabular}
+\end{table}
 %END

### 2015-07-22 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Updated revisions

@@
 \textbf{Table 1:} Bhattacharyya Coefficients of the pairwise method comparisons.
-Each comparison which corresponds to the normalised metric between the ``best'' tree and the ``missing data" trees using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).
+Each comparisons corresponds to the comparison of the full (125) distributions between the ``best'' tree and the ``missing data'' trees for all the four methods (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) and the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
 Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
-Note that this is equivalent to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.
 The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
@@
   \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
-                                                                       & $Tr$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & \textbf{0.00} & \textbf{0.00} & 0.10 & 0.20 & 0.32 & \textbf{1.00} \\ 
+                                                                       & $Tr$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & \textbf{1.00} \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & \textbf{0.03} & 0.54 & 0.69 & 0.64 & 0.77 & \textbf{0.98} \\ 
                                                                        & $Tr$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & \textbf{0.02} & 0.74 & 0.80 & 0.79 & 0.89 & \textbf{0.98} \\ 
                                                                        & $Tr$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & \textbf{0.00} & \textbf{0.00} & \textbf{0.00} & \textbf{0.01} & \textbf{0.01} & \textbf{0.04} \\ 
                                                                        & $Tr$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & \textbf{0.00} & \textbf{0.00} & \textbf{0.01} & \textbf{0.02} & \textbf{0.04} & 0.11 \\ 
                                                                        & $Tr$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
-    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
-                                                                       & $Tr$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.50 & 0.77 & 0.85 & 0.85 & \textbf{0.96} & \textbf{1.00} \\ 
+                                                                       & $Tr$ & 0.91 & \textbf{0.96} & \textbf{0.98} & \textbf{0.97} & \textbf{0.99} & \textbf{1.00} \\ 
    \hline

### 2015-07-22 Analysis/MissingData_differences.R

@@ -33,44 +33,48 @@ if(quick.load == FALSE) {
 #Setting the variables
 data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
 metrics<-c("R.F_Cluster","Triples")
-parameter<-par_MLMFMC
+parameter<-par_MC
 
 #----------------
 #Calculating the pairwise comparisons within (effect of the parameters)
 #----------------
 
-#pair.comp_within.MLbest.RF<-pair.bhatt.coeff(sub.data(get(data.sets[1]), parameter, metric=metrics[1]), diag=TRUE)
+pair.comp_within.MLbest.RF<-pair.bhatt.coeff(sub.data(get(data.sets[1]), parameter, metric=metrics[1]), diag=TRUE)
 #save(pair.comp_within.MLbest.RF, file="../Data/R_data/pair.comp_within.MLbest.RF.Rda")
-load("../Data/R_data/pair.comp_within.MLbest.RF.Rda")
-#pair.comp_within.MLbest.Tr<-pair.bhatt.coeff(sub.data(get(data.sets[1]), parameter, metric=metrics[2]), diag=TRUE)
+#load("../Data/R_data/pair.comp_within.MLbest.RF.Rda")
+pair.comp_within.MLbest.Tr<-pair.bhatt.coeff(sub.data(get(data.sets[1]), parameter, metric=metrics[2]), diag=TRUE)
 #save(pair.comp_within.MLbest.Tr, file="../Data/R_data/pair.comp_within.MLbest.Tr.Rda")
-load("../Data/R_data/pair.comp_within.MLbest.Tr.Rda")
-#pair.comp_within.Baycon.RF<-pair.bhatt.coeff(sub.data(get(data.sets[2]), parameter, metric=metrics[1]), diag=TRUE)
+#load("../Data/R_data/pair.comp_within.MLbest.Tr.Rda")
+pair.comp_within.Baycon.RF<-pair.bhatt.coeff(sub.data(get(data.sets[2]), parameter, metric=metrics[1]), diag=TRUE)
 #save(pair.comp_within.Baycon.RF, file="../Data/R_data/pair.comp_within.Baycon.RF.Rda")
-load("../Data/R_data/pair.comp_within.Baycon.RF.Rda")
-#pair.comp_within.Baycon.Tr<-pair.bhatt.coeff(sub.data(get(data.sets[2]), parameter, metric=metrics[2]), diag=TRUE)
[…]

### 2015-07-22 Analysis/MissingData_modetable.R

@@ -30,7 +30,7 @@ if(quick.load == FALSE) {
 library(hdrcde)
 data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
 metrics<-c("R.F_Cluster","Triples")
-par<-par_MF
+par<-par_MLMFMC
 boot.RF<-sub.data(get(data.sets[3]), par, metric=metrics[1])
 boot.Tr<-sub.data(get(data.sets[3]), par, metric=metrics[2])
 bayt.RF<-sub.data(get(data.sets[4]), par, metric=metrics[1])

### 2015-07-22 Analysis/MissingDataEffect.R
Removed old analysis

@@ -1,557 +1,0 @@
-#Script for testing the effect of missing data on tree topology.
-setwd('Analysis/')
-
-#Loading the functions
-sourceDir <- function(path, trace = TRUE, ...) {
-    for (nm in list.files(path, pattern = "[.][RrSsQq]$")) {
-    if(trace) cat(nm,":")
-        source(file.path(path, nm), ...)
-        if(trace) cat("\n")
-    }
-}
-sourceDir(path="../Functions/TreeComparisons/")
-library(ape)
-
-#Loading the Random tree comparisons
-Random51<-read.table("../Data/Tree_Comparisons/RandomTrees/51t_Random.Cmp",header=TRUE)
-#Random26<-read.table("../Data/Tree_Comparisons/RandomTrees/26t_Random.Cmp",header=TRUE)
-#Random25<-read.table("../Data/Tree_Comparisons/RandomTrees/25t_Random.Cmp",header=TRUE)
-
-#Loading the treesets
-
-#ML chains
-
-#Best trees
-setwd("../Data/Tree_Comparisons/WithRFL/ML/besttree")
-tmp<-TreeCmp.Read('Chain', verbose=TRUE)
-ML_besttrees<-NTS(tmp, Random51)
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
-
-#Bootstraps
-setwd("../Data/Tree_Comparisons/ML/bootstraps")
-tmp<-TreeCmp.Read('Chain', verbose=TRUE)
-ML_bootstraps<-NTS(tmp, Random51[-7])
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
-
-#Bayesian chains
-
-#Consensus
-setwd("../Data/Tree_Comparisons/WithRFL/Bayesian/consensus")
-tmp<-TreeCmp.Read('Chain', verbose=TRUE)
-Bayesian_contrees<-NTS(tmp, Random51)
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
-
-#Treesets
-setwd("../Data/Tree_Comparisons/WithRFL/Bayesian/treesets")
-tmp<-TreeCmp.Read('Chain', verbose=TRUE)
-Bayesian_treesets<-NTS(tmp, Random51)
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
-
[…]

### 2015-07-22 Analysis/BestTrees_NodeSupport.R
Removed absolute path

@@ -1,5 +1,4 @@
 #Script for measuring the node support for the ML "best" trees.
-#cd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
 
 #Loading the functions
 source('../Functions/BS.read.R')

### 2015-07-22 Analysis/MissingData_differences.R

@@ -1,7 +1,5 @@
 #Measuring the differences between methods/parameters
 
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
-
 #Loading the functions
 source("MissingData_fun.R")
 

### 2015-07-22 Analysis/MissingData_fun.R

@@ -1,6 +1,4 @@
 #Data loading for Total Evidence Missing Data Analysis
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
-
 
 #######################
 #Loading the functions

### 2015-07-22 Analysis/MissingData_load.R

@@ -1,6 +1,4 @@
 #Data loading for Total Evidence Missing Data Analysis
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
-
 
 #######################
 #Loading the functions
@@ -25,14 +23,14 @@ Random25<-Random25[,-7]
 setwd("../Data/Tree_Comparisons/ML/besttree")
 tmp<-TreeCmp.Read('Chain', verbose=TRUE)
 ML_besttrees<-NTS(tmp, Random51)
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
+setwd("../../../../Analysis")
 save(ML_besttrees, file="../Data/R_data/TreeCmp-ML_bestrees.Rda")
 
 #Bootstraps
 setwd("../Data/Tree_Comparisons/ML/bootstraps")
 tmp<-TreeCmp.Read('Chain', verbose=TRUE)
 ML_bootstraps<-NTS(tmp, Random51)
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
+setwd("../../../../Analysis")
 save(ML_bootstraps, file="../Data/R_data/TreeCmp-ML_bootstraps.Rda")
 
 #Bayesian chains
@@ -41,14 +39,14 @@ save(ML_bootstraps, file="../Data/R_data/TreeCmp-ML_bootstraps.Rda")
 setwd("../Data/Tree_Comparisons/Bayesian/consensus")
 tmp<-TreeCmp.Read('Chain', verbose=TRUE)
 Bayesian_contrees<-NTS(tmp, Random51)
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
+setwd("../../../../Analysis")
 save(Bayesian_contrees, file="../Data/R_data/TreeCmp-Bayesian_contrees.Rda")
 
 #Treesets
 setwd("../Data/Tree_Comparisons/Bayesian/treesets")
 tmp<-TreeCmp.Read('Chain', verbose=TRUE)
 Bayesian_treesets<-NTS(tmp, Random51)
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
+setwd("../../../../Analysis")
 save(Bayesian_treesets, file="../Data/R_data/TreeCmp-Bayesian_treesets.Rda")
 
 #Isolating the parameters per chains

### 2015-07-22 Analysis/MissingData_modetable.R

@@ -1,7 +1,5 @@
 #Plotting the results of the Total Evidence Missing Data Analysis
 
-setwd('~/PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
-
 #Loading the functions
 source("MissingData_fun.R")
 

### 2015-07-22 Analysis/Pariwise_TreeCmp.R

@@ -1,5 +1,4 @@
 #Script for testing the Random Pairwise Bayesian Tree Comparisons (RPBTC)
-#setwd('../PhD/Projects/Total_Evidence_Method-Missing_data/Analysis/')
 
 #Loading the functions
 source('../Functions/TreeComparisons/NTS.R')

### 2015-07-22 Analysis/Pariwise_TreeCmp.R
Removed old anova analysis

@@ -1,49 +1,0 @@
-#Script for testing the Random Pairwise Bayesian Tree Comparisons (RPBTC)
-
-#Loading the functions
-source('../Functions/TreeComparisons/NTS.R')
-source('../Functions/TreeComparisons/TreeCmp.Plot.R')
-source('../Functions/TreeComparisons/TreeCmp.Read.R')
-source('../Functions/TreeComparisons/TreeCmp.anova.R')
-
-
-#Loading the Random tree comparisons
-Random51<-read.table("../Data/Tree_Comparisons/Random_trees/51t_Random.Cmp",header=TRUE)
-Random26<-read.table("../Data/Tree_Comparisons/Random_trees/26t_Random.Cmp",header=TRUE)
-Random25<-read.table("../Data/Tree_Comparisons/Random_trees/25t_Random.Cmp",header=TRUE)
-
-#Loading the treesets comparisons
-#Low score comparisons
-setwd("../Data/Tree_Comparisons/RPBTC_testing/Low_score_treesets_100/")
-tmp<-TreeCmp.Read('LowScore', verbose=TRUE)
-LowScore<-NTS(tmp, Random51)
-setwd('../../../../Analysis/')
-#Medium score comparisons
-setwd("../Data/Tree_Comparisons/RPBTC_testing/Medium_score_treesets_100/")
-tmp<-TreeCmp.Read('MediumScore', verbose=TRUE)
-MediumScore<-NTS(tmp, Random51)
-setwd('../../../../Analysis/')
-#High score comparisons
-setwd("../Data/Tree_Comparisons/RPBTC_testing/High_score_treesets_100/")
-tmp<-TreeCmp.Read('HighScore', verbose=TRUE)
-HighScore<-NTS(tmp, Random26)
-setwd('../../../../Analysis/')
-
-#TreeCmp anova per metric and per score
-LS.RF<-TreeCmp.anova(LowScore, "R.F_Cluster")
-MS.RF<-TreeCmp.anova(MediumScore, "R.F_Cluster")
-HS.RF<-TreeCmp.anova(HighScore, "R.F_Cluster")
-LS.Tr<-TreeCmp.anova(LowScore, "Triples")
-MS.Tr<-TreeCmp.anova(MediumScore, "Triples")
[…]

### 2015-07-22 Analysis/Analysis_Demo.R
Demo init

@@ -1,0 +1,15 @@
+# This is a cleaned up a explained version on how the results where obtained for the paper "Effects of missing data on topological inference using a Total Evidence approach"
+# Please email me any questions or query (guillert@tcd.ie).
+
+# The whole analysis is divided in four parts
+# A. Generating the matrix
+# B. Removing data
+# C. Estimating phylogenies
+# D. Comparing topologies
+# The three first parts of the analysis are time consuming (~150 CPU years) and therefore not covered in this demo.
+# However you can find explanations and details about the code used in these parts in the README.md file link to this repository.
+# This part only covers the analysis ran in R and will be outlined as follows
+# 1. Extracting the tree comparisons from the TreeCmp java script.
+# 2. Calculating the normalised metrics and visualising the results (Figure 5 in the paper)
+# 3. Calculating the pairwise Bhattacharyya Coefficients between the "best" trees and the "missing data" trees (as outlined in figure 2 in the paper)
+# 4. Calculating the pairwise Bhattacharyya Coefficients between the methods and the metrics (as outlined in figure 3 in the paper)

### 2015-07-23 Analysis/Analysis_Demo.R
Updated analysis demo

@@ -1,4 +1,5 @@
-# This is a cleaned up a explained version on how the results where obtained for the paper "Effects of missing data on topological inference using a Total Evidence approach"
+# This is a cleaned up a explained version on how the results where obtained for the paper "Effects of missing data on topological 
+# inference using a Total Evidence approach"
 # Please email me any questions or query (guillert@tcd.ie).
 
 # The whole analysis is divided in four parts
@@ -11,5 +12,121 @@
 # This part only covers the analysis ran in R and will be outlined as follows
 # 1. Extracting the tree comparisons from the TreeCmp java script.
 # 2. Calculating the normalised metrics and visualising the results (Figure 5 in the paper)
-# 3. Calculating the pairwise Bhattacharyya Coefficients between the "best" trees and the "missing data" trees (as outlined in figure 2 in the paper)
+# 3. Calculating the pairwise Bhattacharyya Coefficients between the "best" trees and the "missing data" trees (as outlined in
+# figure 2 in the paper)
 # 4. Calculating the pairwise Bhattacharyya Coefficients between the methods and the metrics (as outlined in figure 3 in the paper)
+
+# Each function described below has an associated manual at the start of the function files in Functions/TreeComparisons/
+# The information contains:
+# -The function name
+# -A brief function description
+# -A version log
+# -And a list of requirements
+
+
+#--------------
+# Data
+#--------------
+
+# The tree comparisons are available on the following link
+browseURL("http://figshare.com/articles/Effect_of_missing_data_on_topological_inference_using_a_total_evidence_approach/1306861")
+# The folder containing the tree comparisons is the "Tree_Comparisons.zip" folder.
+
+# The various user functions used for this analysis can be load by sourcing the following file
+source("MissingData_fun.R")
+
+# Data can be automatically loaded by running the following script
+source("MissingData_load.R") 
+# However, note that this function loads all the data available and can take some time.
+# The following indications explain in more details what how the data is loaded and how the metrics can be normalised
+
+#--------------
+# 1 - Extracting the tree comparisons from the TreeCmp java script
+#--------------
+
+# The tree comparisons are stored in .Cmp files that contain several informations such as the trees that where compared
+# and the raw differences between these trees using various metrics.
+# All the trees here are compared to the best tree of each chain that is entitled Chain<number>_L00F00C00 where number
[…]

### 2015-07-23 Analysis/MissingData_load.R

@@ -8,10 +8,10 @@ source("MissingData_fun.R")
 #Loading the Random tree comparisons
 Random51<-read.table("../Data/Tree_Comparisons/RandomTrees/51t_Random.Cmp",header=TRUE)
 Random51<-Random51[,-7]
-Random26<-read.table("../Data/Tree_Comparisons/RandomTrees/26t_Random.Cmp",header=TRUE)
-Random26<-Random26[,-7]
-Random25<-read.table("../Data/Tree_Comparisons/RandomTrees/25t_Random.Cmp",header=TRUE)
-Random25<-Random25[,-7]
+#Random26<-read.table("../Data/Tree_Comparisons/RandomTrees/26t_Random.Cmp",header=TRUE)
+#Random26<-Random26[,-7]
+#Random25<-read.table("../Data/Tree_Comparisons/RandomTrees/25t_Random.Cmp",header=TRUE)
+#Random25<-Random25[,-7]
 
 #######################
 #Loading the treesets

### 2015-07-24 Manuscript/MPESubmission/Revision/response.tex
Revisions updates
MC -> NC (in main + response)

@@
 %TG: Add a section "Additional 'fancy' changes" (not fancy obviously)
-% For clarity, we replaced the font of all the functions in R language in courier font (R, diversitree, etc...). \texttt{} - ACTUALLY DO THE CHANGE IF RELEVAN
+% For clarity, we replaced the font of all the functions in R language in courier font (R, diversitree, etc...). \texttt{} - ACTUALLY DO THE CHANGE IF RELEVANT
+%We are very grateful to both referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper.
+We have taken all of their comments on board, and respond to their points below.
 
@@
 
-%To do
+%Working on it.
 
@@
 
+%TG: what does MS means? Manuscript?
+%We do not agree blabalbalbal
+
 \item{\textcolor{blue}{\textbf{Better explanation of the missing data:} I found the explanation of the missing data strategy very difficult to follow.
@@
 I hope the authors can re-write in simpler language ("explain like I'm 5") for easier comprehension.}}
+
+%To do
 
@@
 
+% TG: Ok i've changed it every where but personally I don't like it.
+It makes it even more messy and hard to read.
+However, because I didn't answered the first and last comment of this reviewer, I felt that at least 2/4 is a minimum.
+We entirely agree with this comment and hope it will help to the global comprehension of the manuscript.
+We replaced the $M_C$ (missing characters) parameter by $N_C$ (number of characters with five states: 100, 90, 75, 50 and 25) in the revised manuscript, the appendices and the figures.
+We also slightly modified some explanations related to the change of missing characters to number of characters:
+
+"$N_{C}$, the number of morphological characters for both living and fossil taxa: 100, 90, 75, 50 or 25 characters." line @@@
+
+"When applying the $N_C$ parameter with the second parameter state (90 characters), however, the resulting matrix potentially had more than 10\% data missing.
+Not that the $N_C$ parameter states can be seen in the same way as the $M_L$ and $M_F$ ones: the first state (100 characters) corresponds to 0\% less characters than the ``complete'' matrix, the second (90 characters), corresponds to 10\% less than the ``complete'' matrix, etc." line @@@
+
+"To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $N_{C}$ = 100 etc.
+(see Figure 2 for more details)." line @@@
+
+"Number of morphological characters ($N_{C}$)" line @@@ %Header in discussion
+
+"Additionally, when both $M_{L}$ reach 50\% missing data and $N_{C}$ reach 50 characters, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6)." line @@@
+
 \item{\textcolor{blue}{\textbf{Specific variation of data per fossil:} More importantly, it doesn't seem like missing data varied among fossil taxa, unless I'm misreading.
@@
 That's the crucial case here.}}
+
+%Artifical fossilisation thing?
 \end{enumerate}

### 2015-07-24 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
MC -> NC (in main + response)

@@
 \item{Removing data:} \label{step:remove_data} \\
-We removed data from the morphological part of the ``complete'' matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the proportion of missing morphological characters ($M_{C}$).
+We removed data from the morphological part of the ``complete'' matrix to simulate the effects of missing data by modifying three parameters (i) the proportion of living taxa with no morphological data ($M_{L}$), (ii) the proportion of missing data in the fossil taxa ($M_{F}$) and (iii) the number of morphological characters ($N_{C}$).
 We call the resulting 125 matrices ``missing-data'' matrices.
@@
 \item{Comparing topologies:} \label{step:compare_topo} \\
-We compared the ``best'' tree to the ``missing-data'' trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $M_{C}$) and their interactions on the topologies of our phylogenies
+We compared the ``best'' tree to the ``missing-data'' trees to assess the influence of each parameter ($M_{L}$, $M_{F}$, $N_{C}$) and their interactions on the topologies of our phylogenies
 \end{enumerate}
@@
 This parameter illustrates the quality of the fossil record, a common problem due to preservation biases \citep{sansomfossilization2013}. 
-\item{$M_{C}$, the proportion of missing morphological characters for both living and fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%. }
+\item{$N_{C}$, the number of morphological characters for both living and fossil taxa: 100, 90, 75, 50 or 25 characters. }
 This parameter illustrates the number of available morphological characters for both living and fossil taxa.
@@
 
-In practice, each parameter represents a different way of removing data from the matrix: $M_L$ removes rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $M_C$ removes columns across both living and fossil taxa's data.
+In practice, each parameter represents a different way of removing data from the matrix: $M_L$ removes rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $N_C$ removes columns across both living and fossil taxa's data.
 Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
-We first applied the parameters $M_L$ and $M_F$ to the matrix and then applied the $M_C$ parameter.
+We first applied the parameters $M_L$ and $M_F$ to the matrix and then applied the $N_C$ parameter.
 Therefore, when 10\% data was missing for both $M_L$ and $M_F$, 10\% data was missing in the morphological part of the matrix.
-When applying the $M_C$ parameter with the same percentage (10\%), however, the resulting matrix potentially had more than 10\% data missing.
+When applying the $N_C$ parameter with the second parameter state (90 characters), however, the resulting matrix potentially had more than 10\% data missing.
+Not that the $N_C$ parameter states can be seen in the same way as the $M_L$ and $M_F$ ones: the first state (100 characters) corresponds to 0\% less characters than the ``complete'' matrix, the second (90 characters), corresponds to 10\% less than the ``complete'' matrix, etc.
 
@@
 50 ``complete'' and 50 x 125 ``missing-data'' matrices) with a relatively strong phylogenetic signal (median bootstrap $>$ 50).
-This step was implemented to make sure that the differences we observed in topologies (see below) were due to the amount of missing data for each parameter ($M_L$, $M_F$ and $M_C$) and not simply to low branch support that is likely to lead to different topologies.
+This step was implemented to make sure that the differences we observed in topologies (see below) were due to the amount of missing data for each parameter ($M_L$, $M_F$ and $N_C$) and not simply to low branch support that is likely to lead to different topologies.
 On these selected simulations, we used the fast bootstrap algorithm and performed 1000 bootstraps for each tree inference to assess topological support \citep{pattengale2010many}.
@@
 \subsection{Comparing topologies}
-We compared the topology of the ``missing-data'' trees to the ``best'' tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $M_{C}$ on tree topology.
+We compared the topology of the ``missing-data'' trees to the ``best'' tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $N_{C}$ on tree topology.
 We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa \citep[i.e.
@@
 \subsection{Testing the effects of the missing data parameters on topological recovery}
-Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$ and their interactions) on our ability to recover the ``best'' tree topology in a Total Evidence framework.
+Finally, we tested the effects of our missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$ and their interactions) on our ability to recover the ``best'' tree topology in a Total Evidence framework.
 We also assessed the effect of our missing data parameters jointly with the effects of different tree inference and uncertainty methods (i.e.
@@
 Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
-To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $M_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\%), i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $M_{C}$ = 0\% etc.
+To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $N_{C}$ = 100 etc.
 (see Figure 2 for more details).
@@
 % A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
-$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $N_{C}$ = 75\%).
 The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.
@@
  %A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
-$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $N_{C}$ = 75\%).
 \textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
@@
 \section{Results}
-As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best'' tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $M_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets metric).
+As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best'' tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $N_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets metric).
 Nonetheless, the different missing data parameters and tree inference methods do not affect the topology in the same way (Figure 4 and Figure 5).
@@
 %\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue).}
- %The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
+ %The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $N_{C}$) is shown on the x axis.
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
@@
 %\caption{{\bf The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).}
- %The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+ %The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $N_{C}$ (lower line).
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
@@
 As the amount of missing data increases across all three parameters, our ability to recover the ``best'' tree topology decreases (Figure 4).
-The Normalised Robinson-Foulds metric is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
-The Normalised Triplets metric, however, is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
+The Normalised Robinson-Foulds metric is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $N_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
+The Normalised Triplets metric, however, is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $N_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $N_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
 
@@
 As expected, our ability to recover the ``best'' tree topology is worst when each parameter contains the maximum amount of missing data (i.e.
-$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Figure 5; Tables C2, C3 and C4 in Appendix C).
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $N_{C}$ = 75\%), and best when there is no missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 0\%; Figure 5; Tables C2, C3 and C4 in Appendix C).
 Figure 6 shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0).
@@
 %\caption{{\bf The effects of missing data on topological recovery using Bayesian consensus trees.}
-% Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+% Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $N_{C}$ (lower line).
 Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the ``best'' tree distribution, calculated using the Bhattacharyya Coefficient.
@@
 Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
 The Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; blue regions in Figure 6A), however, displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; blue regions in Figure 6B).
 The orange upper triangle in Figure 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Figure 6A).
-There is no additional effect (i.e. no dissimilarities) of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
+There is no additional effect (i.e. no dissimilarities) of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
 
@@
 These factors, however, affected topological recovery in different ways and to different extents.
-Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Figure 6).
+Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($N_{C}$) had worst effects on topological recovery (Figure 6).
 Additionally, using Bayesian consensus trees recovered the ``best'' tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Figure 5, Figure 6, Table ~\ref{Tab_Results-Difference_methods}).
@@
 
-\subsubsection{Missing morphological characters ($M_{C}$)}
+\subsubsection{Number of morphological characters ($N_{C}$)}
 Reducing the overall number of morphological characters reduces the probability of their overlap among the taxa in the matrix, and therefore decreases our ability to recover the ``best'' tree topology.
 We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
-Conversely, removing 10\% of morphological characters ($M_{C}$) genuinely removes 10\% of data in the matrix.
+Conversely, removing 10\% of morphological characters ($N_{C}$) genuinely removes 10\% of data in the matrix.
 Nonetheless, the effect of removing characters on the ability to recover the ``best'' tree topology is of the same order of magnitude as for the other two parameters (Figure 4).
@@
 As expected, when combining the missing data parameters, our ability to recover the ``best'' tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a larger proportion of missing data (i.e.
-$M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Figure 5).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) and lower when they have a larger proportion of missing data (i.e.
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $N_{C}$ = 25; Figure 5).
 It is important, however, to notice that the effect of each parameter is not additive.
-Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
+Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
 It seems therefore more important to integrate a maximum of living taxa with morphological data as well as a maximum number of characters.
-Additionally, when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6).
+Additionally, when both $M_{L}$ reach 50\% missing data and $N_{C}$ reach 50 characters, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6).
 
@@
 It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best'' tree topology than Maximum Likelihood and Bayesian consensus trees.
-This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $M_{C}$ = 0\%; Figure 4).
+This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $N_{C}$ = 0\%; Figure 4).
 This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best'' trees involved 1000 random pairwise comparisons rather than just one.
@@
 \subsection{Practical implications}
-Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($M_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
+Our missing data parameters illustrate different sources of missing data in empirical matrices as follows: ($M_{L}$) the paucity of coded morphological characters for living taxa; ($M_{F}$) the missing data for fossils (or parts of fossils) that have not been preserved in the fossil record; and ($N_{C}$) characters that have not been coded across living and fossil species, perhaps due to difficulties in coding or poor preservation of the feature in collections.
 Filling these gaps in empirical Total Evidence matrices should lead to a substantial increase in our ability to recover the ``best'' tree topology.
@@
 Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
-When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
+When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
 Therefore we advise filling as many gaps in Total Evidence matrices as possible.
@@
 A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
-$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $N_{C}$ = 25).
 The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.\\
@@
 A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g.
-$M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%).
+$M_{L}$ = 10\%, $M_{F}$ = 50\%, $N_{C}$ = 25).
 \textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g.
@@
 \textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange).
-The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis.
+The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $N_{C}$) is shown on the x axis.
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
@@
 \textbf{Figure 5:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
-The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
@@
 \textbf{Figure 6:} The effects of missing data on topological recovery using Bayesian consensus trees.
-Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
 The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.

### 2015-07-24 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
flipped over/under estimation
changed  MC-NC in the supp

@@
 \begin{equation}
-\Delta\alpha=\alpha_{TRUE} - \alpha_{EST}
+\Delta\alpha=\alpha_{EST} - \alpha_{TRUE}
 \end{equation}
@@
 \caption{Differences between the simulated rate of each morphological character ($\alpha$ parameter) and the Bayesian estimates for the trees with missing data.
-The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Positive values shows underestimation of the parameter and negative values shows overestimation.
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
+Positive values shows overestimation of the parameter and negative values shows underestimation.
 One unit of difference means that the estimated rate is twice as big/small than the true rate.
@@
 We tested this effect on five trees with the five levels of missing data (i.e.
-first tree: $M_L$=0\%, $M_F$=0\% and $M_C$=0\%; second tree: $M_L$=10\%, $M_F$=10\% and $M_C$=10\%, etc.) on the first 20 simulation chains.
+first tree: $M_L$=0\%, $M_F$=0\% and $N_C$=100; second tree: $M_L$=10\%, $M_F$=10\% and $N_C$=90, etc.) on the first 20 simulation chains.
 We then compared the trees inferred using the ``true'' tree as a starting tree to the trees using a random tree as a starting tree using the normalised Robinson-Foulds and Triplets metrics.
@@
   \hline
-    $M_L$=0\%; $M_F$=0\%; $M_C$=0\%          & $RF$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
+    $M_L$=0\%; $M_F$=0\%; $N_C$=0\%          & $RF$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
                                              & $Tr$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
-    $M_L$=10\%; $M_F$=10\%; $M_C$=10\%       & $RF$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
+    $M_L$=10\%; $M_F$=10\%; $N_C$=10\%       & $RF$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
                                              & $Tr$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
-    $M_L$=25\%; $M_F$=25\%; $M_C$=25\%       & $RF$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
+    $M_L$=25\%; $M_F$=25\%; $N_C$=25\%       & $RF$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
                                              & $Tr$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
-    $M_L$=50\%; $M_F$=50\%; $M_C$=50\%       & $RF$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
+    $M_L$=50\%; $M_F$=50\%; $N_C$=50\%       & $RF$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
                                              & $Tr$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
-    $M_L$=75\%; $M_F$=75\%; $M_C$=75\%       & $RF$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
+    $M_L$=75\%; $M_F$=75\%; $N_C$=75\%       & $RF$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
                                              & $Tr$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 

### 2015-07-24 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixC.tex
changed  MC-NC in the supp

@@
     \caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey).
-The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
@@
     \caption{The effects of missing data on topological recovery using Maximum Likelihood trees.
-The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
 Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
@@
     \caption{The effects of missing data on topological recovery using Maximum Likelihood bootstrap trees.
-The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+The x and the y axes both show show the percentage ofmissing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
 Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
@@
     \caption{The effects of missing data on topological recovery using Bayesian posterior tree distribution.
-The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+The x and the y axes both show show the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
 Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds metric and (B) Normalised Triplets metric distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient.
@@
 \begin{table}[ht]
-\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
+\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $N_{C}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_MC}
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $N_{C}$ missing data parameter only.}
 \label{Tab_Supp_summary_BC_MC}
@@
 \caption{Summary of the median Bhattacharyya Coefficients comparisons between the missing data parameters.
-Each value corresponds to the median Bhattacharyya Coefficient between the Normalised metrics (Robinson-Foulds - RF; Triplets - Tr) between the ``best'' and the ``missing data'' trees (as outlined in Figure 2 in the main text) for the the trees with: Missing Living taxa only ($M_L$); Missing Fossil data only ($M_F$); Missing characters only ($M_C$); and the combination of the three parameters ($M_LM_FM_C$).}
+Each value corresponds to the median Bhattacharyya Coefficient between the Normalised metrics (Robinson-Foulds - RF; Triplets - Tr) between the ``best'' and the ``missing data'' trees (as outlined in Figure 2 in the main text) for the the trees with: Missing Living taxa only ($M_L$); Missing Fossil data only ($M_F$); Number of characters only ($N_C$); and the combination of the three parameters (combined).}
 \begin{tabular}{llcccc}  
   \hline
- Method & Metric & $M_L$ & $M_F$ & $M_C$ & $M_LM_FM_C$ \\ 
+ Method & Metric & $M_L$ & $M_F$ & $N_C$ & combined \\ 
   \hline
@@
 \caption{Summary of the median Bhattacharyya Coefficients comparisons between the missing data parameters and the methods.
-Each value corresponds to the median Bhattacharyya Coefficient between the Normalised metrics (Robinson-Foulds - RF; Triplets - Tr) between the `methods (as outlined in Figure 3 in the main text) for the the trees with: Missing Living taxa only ($M_L$); Missing Fossil data only ($M_F$); Missing characters only ($M_C$); and the combination of the three parameters ($M_LM_FM_C$).}
+Each value corresponds to the median Bhattacharyya Coefficient between the Normalised metrics (Robinson-Foulds - RF; Triplets - Tr) between the `methods (as outlined in Figure 3 in the main text) for the the trees with: Missing Living taxa only ($M_L$); Missing Fossil data only ($M_F$); Number of characters only ($N_C$); and the combination of the three parameters (combined).}
 \begin{tabular}{llcccc}  
   \hline
- Method & Metric & $M_L$ & $M_F$ & $M_C$ & $M_LM_FM_C$ \\ 
+ Method & Metric & $M_L$ & $M_F$ & $N_C$ & combined \\ 
   \hline

### 2015-07-24 Analysis/Analysis_Demo.R
changed MC -> NC in demo

@@ -36,7 +36,7 @@ browseURL("http://figshare.com/articles/Effect_of_missing_data_on_topological_in
 source("MissingData_fun.R")
 
 # Data can be automatically loaded by running the following script
-source("MissingData_load.R") 
+#source("MissingData_load.R") 
 # However, note that this function loads all the data available and can take some time.
 # The following indications explain in more details what how the data is loaded and how the metrics can be normalised
 
@@ -51,7 +51,7 @@ source("MissingData_load.R")
 # In the case of the best tree:
 # the parameter ML, corresponding to the number of missing living taxa is set at 0%
 # the parameter MF, corresponding to the number of missing data in the fossil record is set at 0%
-# the parameter MC, corresponding to the number of missing characters is set at 0%
+# the parameter MC, corresponding to the number of missing characters from the complete matrix is set at 0%.
 # For example, the comparison Chain18_L10F25C50 is the results of the comparison between the best tree of the chain 18
 # (Chain18_L00F00C00) and the missing-data tree with the parameters ML=10%, MF=25% and MC=50%.
 # These files can be easily converted in a R object called "treeCmp" object that is a concatenation of all the metrics

### 2015-07-24 Analysis/MissingData_differences.R
Updated figures (MC -> NC)

@@ -31,36 +31,36 @@ if(quick.load == FALSE) {
 #Setting the variables
 data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
 metrics<-c("R.F_Cluster","Triples")
-parameter<-par_MC
+parameter<-par_MLMFMC
 
 #----------------
 #Calculating the pairwise comparisons within (effect of the parameters)
 #----------------
 
-pair.comp_within.MLbest.RF<-pair.bhatt.coeff(sub.data(get(data.sets[1]), parameter, metric=metrics[1]), diag=TRUE)
+#pair.comp_within.MLbest.RF<-pair.bhatt.coeff(sub.data(get(data.sets[1]), parameter, metric=metrics[1]), diag=TRUE)
 #save(pair.comp_within.MLbest.RF, file="../Data/R_data/pair.comp_within.MLbest.RF.Rda")
-#load("../Data/R_data/pair.comp_within.MLbest.RF.Rda")
-pair.comp_within.MLbest.Tr<-pair.bhatt.coeff(sub.data(get(data.sets[1]), parameter, metric=metrics[2]), diag=TRUE)
+load("../Data/R_data/pair.comp_within.MLbest.RF.Rda")
+#pair.comp_within.MLbest.Tr<-pair.bhatt.coeff(sub.data(get(data.sets[1]), parameter, metric=metrics[2]), diag=TRUE)
 #save(pair.comp_within.MLbest.Tr, file="../Data/R_data/pair.comp_within.MLbest.Tr.Rda")
-#load("../Data/R_data/pair.comp_within.MLbest.Tr.Rda")
-pair.comp_within.Baycon.RF<-pair.bhatt.coeff(sub.data(get(data.sets[2]), parameter, metric=metrics[1]), diag=TRUE)
+load("../Data/R_data/pair.comp_within.MLbest.Tr.Rda")
+#pair.comp_within.Baycon.RF<-pair.bhatt.coeff(sub.data(get(data.sets[2]), parameter, metric=metrics[1]), diag=TRUE)
 #save(pair.comp_within.Baycon.RF, file="../Data/R_data/pair.comp_within.Baycon.RF.Rda")
-#load("../Data/R_data/pair.comp_within.Baycon.RF.Rda")
[…]

### 2015-07-24 Analysis/MissingData_plot.R

@@ -58,7 +58,7 @@ par(mar=c(5,3,1,2) + 0.1)
 multi.TreeCmp.plot( data.sets , par_MF, metrics[2], col=pal, las=2, probs=c(95, 50), ylim=ylim, xaxis=xaxis.labs, ylab=NULL, las=2, xlab="Missing data in the fossil record (MF %)")
 #Missing character Triplets
 par(mar=c(5,2,1,4) + 0.1)
-multi.TreeCmp.plot( data.sets , par_MC, metrics[2], col=pal, las=2, probs=c(95, 50), ylim=ylim, xaxis=xaxis.labs, ylab=NULL, las=2, xlab="Missing characters (MC %)") 
+multi.TreeCmp.plot( data.sets , par_MC, metrics[2], col=pal, las=2, probs=c(95, 50), ylim=ylim, xaxis=c("100","90","75","50","25"), ylab=NULL, las=2, xlab="Number of characters (NC)") 
 
 #Back to default
 par(op)

### 2015-07-24 Functions/TreeComparisons/TreeCmp.Plot.R

@@ -685,7 +685,7 @@ global.TreeCmp.plot<-function(data.list, parameter, metrics, col, col.grad=c('bl
     plot(1,1, xlab='', ylab='', xlim=c(1,125), ylim=c(0,3), type='n', yaxt='n', bty='n', xaxt='n') #
     text(-2, 2.5, "ML")
     text(-2, 1.5, "MF")
-    text(-2, 0.5, "MC")
+    text(-2, 0.5, "NC")
     #ML
     col.new<-colgrad[1:5]
     for(i in 1:5) {
@@ -718,7 +718,7 @@ global.TreeCmp.plot<-function(data.list, parameter, metrics, col, col.grad=c('bl
     plot(1,1, xlab='', ylab='', xlim=c(1,125), ylim=c(0,3), type='n', xaxt='n', yaxt='n', bty='n')
     text(-2, 2.5, "ML")
     text(-2, 1.5, "MF")
-    text(-2, 0.5, "MC")
+    text(-2, 0.5, "NC")
     #ML
     col.new<-colgrad[1:5]
     for(i in 1:5) {

### 2015-07-24 Functions/TreeComparisons/TreeCmp.pairPlot.R

@@ -109,7 +109,7 @@ TreeCmp.pairPlot<-function(posthoc, parametric, binary=FALSE, col=c("yellow", "r
             plot(1,1, xlab='', ylab='', xlim=c(1,125), ylim=c(0,3), type='n', yaxt='n', bty='n', xaxt='n')
             text(-2, 2.5, "ML")
             text(-2, 1.5, "MF")
-            text(-2, 0.5, "MC")
+            text(-2, 0.5, "NC")
             #ML
             col.new<-colgrad[1:5]
             for(i in 1:5) {
@@ -130,7 +130,7 @@ TreeCmp.pairPlot<-function(posthoc, parametric, binary=FALSE, col=c("yellow", "r
             plot(1,1, xlab='', ylab='', xlim=c(0,3), ylim=c(1,125), type='n', yaxt='n', bty='n', xaxt='n')
             text(2.5, -2, "ML")
             text(1.5, -2, "MF")
-            text(0.5, -2, "MC")
+            text(0.5, -2, "NC")
             #ML
             col.new<-colgrad[1:5]
             for(i in 1:5) {

### 2015-07-24 Functions/TreeComparisons/plot.bhatt.coef.R

@@ -33,7 +33,7 @@ plot.bhatt.coeff<-function(matrix, col=c("blue", "red"), col.grad=c("white", "bl
             plot(1,1, xlab='', ylab='', xlim=c(1,125), ylim=c(0,3), type='n', yaxt='n', bty='n', xaxt='n')
             text(-2, 2.5, "ML")
             text(-2, 1.5, "MF")
-            text(-2, 0.5, "MC")
+            text(-2, 0.5, "NC")
             #ML
             col.new<-colgrad[1:5]
             for(i in 1:5) {
@@ -54,7 +54,7 @@ plot.bhatt.coeff<-function(matrix, col=c("blue", "red"), col.grad=c("white", "bl
             plot(1,1, xlab='', ylab='', xlim=c(0,3), ylim=c(1,125), type='n', yaxt='n', bty='n', xaxt='n')
             text(2.5, -2, "ML")
             text(1.5, -2, "MF")
-            text(0.5, -2, "MC")
+            text(0.5, -2, "NC")
             #ML
             col.new<-colgrad[1:5]
             for(i in 1:5) {

### 2015-07-24 Manuscript/MPESubmission/Revision/response.tex
Added justification why not to do Rev2 comment1

@@
 
-%TG: what does MS means? Manuscript?
-%We do not agree blabalbalbal
+A similar point/suggestion was also raised by reviewer 1 (see Reviewer 1 General comment number 6 and specific comment number 25).
+We do understand the concerns of this reviewer as well but, again, we do not agree that comparing the "missing-data" trees to the "best" trees adds a layer of complexity to our simulations.
+In fact, we argue that doing so actually adds a layer of complexity to the analysis.
+The problematic we are tackling in this manuscript is to understand "what is the effect of our three parameters on recovering topology ?"
+To answer this question, however, we have to add a layer between the data (i.e. the "complete" or "missing-data" matrices) and the actual topology which is the tree inference.
+We test the effect of this layer by using different tree inference methods.
+
+Because we use a simulation approach to test this problematic, we have to start from a random starting point (the "true" tree) and then do the following steps: (1) generate and remove the data (2) infer the topologies and (3) compare the topologies.
+Note that in this schematic representation, the "true" tree is the step (0) of our analysis and involves a series of algorithms and softwares (i.e.
+seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes).
+To compare the "true" tree's topology to the "missing-data" topologies will be equivalent to explore the effect of our three parameters on recovering topology through 3 layers (1- generate the data; 2- infer the topologies; and 3- compare the topologies).
+By comparing the "best" tree's topology to the "missing-data" ones, we explore the effect of our three parameters through only 2 layers (2- infer the topologies; and 3- compare the topologies).
+Note that we explore the effect of these 2 layers by using two different tree inference methods (effectively 4 since we also use the Maximum Likelihood bootstraps trees and the Bayesian posterior trees) and 2 different Normalised topological metrics (Robinson-Foulds and Triplets).
+Using as main results the comparisons between the "true" and the "missing-data" trees would also need the exploration of the effect of the cascade of algorithms and softwares (maybe by modifying the algorithms used to generate the "true" tree and the matrices?).
+
+However, we have added all the comparisons between the "true" tree and the "missing-data" trees in the supplementary.
+It is interesting to notice that the algorithms and softwares as well as probably the size of our matrices are actually not ideal an the recovery of the "true" tree is rather low and scales down the our results (see figure @).
+However, even though these comparisons are worst, it is encouraging to see that the patterns are conserved (i.e.
+Bayesian consensus trees $>$ Maximum Likelihood trees $>$ both bootstraps trees and Bayesian posterior trees).
 

### 2015-07-26 Analysis/Analysis_Demo.R
Added bhattacharya coeff part to the demo

@@ -130,3 +130,34 @@ TreeCmp.Plot(TreeCmp=MC_parameter_only, metric=chosen_metric, probs=CI, lines=FA
 # 3. Calculating the pairwise Bhattacharyya Coefficients between the "best" trees and the "missing data" trees (as outlined in
 # figure 2 in the paper)
 #--------------
+
+# The Bhattacharyya Coefficients is the probability of overlap between two distributions, it can be calculated using the
+# bhatt.coeff function
+bhatt.coeff(rnorm(1000),rnorm(1000))
+
+# Here we want to know the differences between the normalised tree comparisons metrics of each set of parameters.
+# We can use the pair.bhatt.coeff function that calculates the pairwise Bhattacharyya Coefficients.
+# First, to facilitate the task, we can transform the TreeCmp object directly into a list of distributions for a given metric and
+# a given set of parameters using the sub.data function.
+MC_parameter_RF_metric<-sub.data(TreeCmp_object, parameter=c(1:5), metric="R.F_Cluster")
+
+# We can then feed this list of distributions to the pair.bhatt.coeff function
+pair.bhatt.coeff(MC_parameter_RF_metric, diag=TRUE)
+
+# This outputs a matrix with the probability of overlap between the distributions.
+# Note that the diagonal is always equals to 1 and that, in this example, there is a null probability of overlap between
+# the normalised Robinson-Foulds metrics between the "best" tree and the two "missing-data" trees with MC=50% and 75%.
+
+# This matrix representation of the probabilities overlaps is understandable when few pairwise comparisons are studied
+# but gets more tricky when the number of pairwise comparisons increases.
+# For example, if we want to calculate all the pairwise comparisons (7875) the matrix representation is not really helpful.
+All_parameter_RF_metric<-sub.data(TreeCmp_object, parameter=c(1:125), metric="R.F_Cluster")
+All_parameters_pairwise<-pair.bhatt.coeff(All_parameter_RF_metric, diag=TRUE)
+head(All_parameters_pairwise)
+
+# We can use the plot.bhatt.coeff function to represent these results as a heat map.
+plot.bhatt.coeff(All_parameters_pairwise, col=c("blue", "orange"), col.grad=c("white", "black"))
+
+#--------------
+# 4. Calculating the pairwise Bhattacharyya Coefficients between the methods and the metrics (as outlined in figure 3 in the paper)
+#--------------

### 2015-07-27 Manuscript/MPESubmission/Revision/response.tex
Response edit

@@
 
-%To do
+The lines 180 and 181 states "All the molecular information for fossil taxa was replaced by missing data ("?")" in the former manuscript.
+We changed the verb "infer" (see specific comment number 16 from reviewer 1).
+
+We assumed that this reviewer is mentioning the section "2.2 Removing data" as the "explanation of the missing data strategy". We modified the whole section emphasising on clarity and simplicity.
+
+"To explore the effect of missing data on topological recovery, we removed various amounts of the “complete” matrix to obtain matrices with missing morphological data.Hereafter, we call these matrices with missing morphological data the “missing-data” matrices.
+We removed morphological following three main data incompleteness parameters:
+1.
+The proportion of missing living taxa ($M_L$).
+This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
+It represents the number of living taxa that are present in the matrix but have only molecular data available.
+This reflects the fact that, because of the increasing facility of collecting molecular data, morphological data for living species is rarely collected (Guillerme and Cooper, 2015).
+Therefore, many living species will have only molecular data available.
+In practice, we removed all the morphological data from randomly chosen living with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.
+2.
+The proportion of missing data in the fossil record ($M_F$).
+This missing data parameter represents the completeness of the fossil record.
+In fact, due to preservation biases, missing data for fossil taxa are common (Sansom and Wills, 2013).
+In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
+3.
+The number of morphological characters for both living and fossil taxa ($N_C$).
+This parameter is not a missing data parameter per se but rather an indication of the size of the matrix.
+Any morphological matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
+Therefore, this parameter correspond to the overall number of characters available for both living and fossil taxa.
+In practice, we randomly removed whole characters from the morphological matrix shrinking it to: 100, 90,75, 50 or 25 characters.
+Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of “missing” morphological characters).
+
+In practice, each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living taxa’s data; $M_F$ removes cells from the fossil taxa’s data; and $N_C$ removes columns across both living and fossil taxa’s data.
+Note that $M_L$  and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
+
+We created matrices using all parameter combinations resulting in 125 ($5^3$) “missing-data” matrices.
+Note that one of these combinations ($M_L$=0\%; $M_F$=0\% and $N_C$=100) has no missing data so is equivalent to the “complete” matrix, thus we have one effectively complete matrix in our 125 “missing-data” matrices.
+In practice, we first removed the data following the two missing data parameters $M_L$ and $M_F$ and then removed data following the $N_C$ parameters.
+To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
+Note that the living taxa always had at least 90\% of data (the 1000 molecular characters)." (line @@@)
 
@@
 We entirely agree with this comment and hope it will help to the global comprehension of the manuscript.
-We replaced the $M_C$ (missing characters) parameter by $N_C$ (number of characters with five states: 100, 90, 75, 50 and 25) in the revised manuscript, the appendices and the figures.
+We replaced the $M_C$ (missing characters) parameter by $N_C$ (number of characters with five states: 100, 90, 75, 50 and 25) in the revised manuscript (see reviewer 2 comment 2 above), the appendices and the figures.
 We also slightly modified some explanations related to the change of missing characters to number of characters:
-
-"$N_{C}$, the number of morphological characters for both living and fossil taxa: 100, 90, 75, 50 or 25 characters." line @@@
-
-"When applying the $N_C$ parameter with the second parameter state (90 characters), however, the resulting matrix potentially had more than 10\% data missing.
-Not that the $N_C$ parameter states can be seen in the same way as the $M_L$ and $M_F$ ones: the first state (100 characters) corresponds to 0\% less characters than the ``complete'' matrix, the second (90 characters), corresponds to 10\% less than the ``complete'' matrix, etc." line @@@
 

### 2015-07-27 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Response edit

@@
 Finally, we combined the morphological and molecular matrices obtained from the ``true'' tree.
-Hereafter we call this the ``complete'' matrix, i.e.
-the matrix with no missing data except for the molecular data of the fossil taxa.
+Hereafter we call this the ``complete'' matrix, i.e. the matrix with no missing data except for the molecular data of the fossil taxa.
 %---------------------------------------------
@@
 \subsection{Removing data}
-\label{Removing_data}
-We modified the ``complete'' matrix to obtain matrices with missing data by randomly replacing data with ``?" in the morphological part of the matrices according to the following parameters:
+% \label{Removing_data}
+% We modified the ``complete'' matrix to obtain matrices with missing data by randomly replacing data with ``?" in the morphological part of the matrices according to the following parameters:
+% \begin{enumerate}
+% \item{$M_{L}$, the proportion of living taxa with no morphological data: 0\%, 10\%, 25\%, 50\% or 75\%.}
+% This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological part.
+This reflects the fact that the increased availability of molecular data means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
+% \item{$M_{F}$, the proportion of missing data in the fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%.}
+% This parameter illustrates the quality of the fossil record, a common problem due to preservation biases \citep{sansomfossilization2013}. 
+% \item{$N_{C}$, the number of morphological characters for both living and fossil taxa: 100, 90, 75, 50 or 25 characters. }
+% This parameter illustrates the number of available morphological characters for both living and fossil taxa.
+This reflects the amount of effort put into identifying the characters \citep[e.g][]{O'Leary08022013}.
+% \end{enumerate}
+
+% In practice, each parameter represents a different way of removing data from the matrix: $M_L$ removes rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $N_C$ removes columns across both living and fossil taxa's data.
+% Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
+% We first applied the parameters $M_L$ and $M_F$ to the matrix and then applied the $N_C$ parameter.
+% Therefore, when 10\% data was missing for both $M_L$ and $M_F$, 10\% data was missing in the morphological part of the matrix.
+% When applying the $N_C$ parameter with the second parameter state (90 characters), however, the resulting matrix potentially had more than 10\% data missing.
+
+% We created matrices using all parameter combinations resulting in 125 ($5^3$) matrices.
+% Note that one of these combinations has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
+% To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
+
+% TG: this part is entirely re-written.
+To explore the effect of missing data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
+Hereafter, we call these matrices with missing morphological data the ``missing-data'' matrices.
+We removed morphological following three main data incompleteness parameters:
 \begin{enumerate}
-\item{$M_{L}$, the proportion of living taxa with no morphological data: 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological part.
-This reflects the fact that the increased availability of molecular data means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
-\item{$M_{F}$, the proportion of missing data in the fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%.}
-This parameter illustrates the quality of the fossil record, a common problem due to preservation biases \citep{sansomfossilization2013}. 
-\item{$N_{C}$, the number of morphological characters for both living and fossil taxa: 100, 90, 75, 50 or 25 characters. }
-This parameter illustrates the number of available morphological characters for both living and fossil taxa.
-This reflects the amount of effort put into identifying the characters \citep[e.g][]{O'Leary08022013}.
+\item{The proportion of missing living taxa ($M_{L}$).}
+This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
+It represents the number of living taxa that are present in the matrix but have only molecular data available.
+This reflects the fact that, because of the increasing facility of collecting molecular data, morphological data for living species is rarely collected \citep{GuillermeCooperMissing}.
+% TG: see what I did here?
+;)
+Therefore, many living species will have only molecular data available.
+In practice, we removed all the morphological data from randomly chosen living with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.
+\item{The proportion of missing data in the fossil record ($M_{F}$).}
+This missing data parameter represents the completeness of the fossil record.
+In fact, due to preservation biases, missing data for fossil taxa are common \citep{sansomfossilization2013}. 
+In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
+\item{The number of morphological characters for both living and fossil taxa ($N_{C}$).}
+This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
+Any morphological matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
+% TG: this is word for word what the reviewer wrote.
+I didn't like doing that first but after seeing the Ineterface Focus review and how they copy pasted the discussion suggestion, I think it's really nice for the reviewer to see a bit of it's work directly in the paper (at least it was for me!).
+Therefore, this parameter correspond to the overall number of characters available for both living and fossil taxa.
+In practice, we randomly removed whole characters from the morphological matrix shrinking it to: 100, 90, 75, 50 or 25 characters.
+Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).
 \end{enumerate}
 
-In practice, each parameter represents a different way of removing data from the matrix: $M_L$ removes rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $N_C$ removes columns across both living and fossil taxa's data.
+% TG: this part is nearly exactly the same as before.
+In practice, each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $N_C$ removes columns across both living and fossil taxa's data.
 Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
-We first applied the parameters $M_L$ and $M_F$ to the matrix and then applied the $N_C$ parameter.
-Therefore, when 10\% data was missing for both $M_L$ and $M_F$, 10\% data was missing in the morphological part of the matrix.
-When applying the $N_C$ parameter with the second parameter state (90 characters), however, the resulting matrix potentially had more than 10\% data missing.
-Not that the $N_C$ parameter states can be seen in the same way as the $M_L$ and $M_F$ ones: the first state (100 characters) corresponds to 0\% less characters than the ``complete'' matrix, the second (90 characters), corresponds to 10\% less than the ``complete'' matrix, etc.
 
-We created matrices using all parameter combinations resulting in 125 ($5^3$) matrices.
-Note that one of these combinations has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
+% TG: and this one is slightly modified only.
+We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data''matrices.
+Note that one of these combinations ($M_{L}$=0\%; $M_{F}$=0\% and $N_{C}$=100) has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
+In practice, we first removed the data following the two missing data parameters $M_L$ and $M_F$ and then removed data following the $N_C$ parameters.
 To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
+Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).
 

### 2015-07-27 Manuscript/MPESubmission/Revision/response.tex
reviewer 2  updates (draft finished)

@@
 
-%Artifical fossilisation thing?
+In this study, we applied a similar approach to what the reviewer 2 suggests (i.e.
+no missing data for living taxa, and maximal number of characters but various amount of missing data for the fossil taxa).
+This results are present in figure 4, central row.
+However, the amount of changes among the fossil taxa was not set to any specific amount of missing data per fossil (e.g.
+like the "artificial fossil" approach in Pattinson et al 2014, Syst.
+Biol.) but we rather treated the missing data in the "whole" fossil record.
+Because our fossil taxa and their characters where randomly simulated, we didn't had any \textit{a priori} assumption on how to remove the data.
+We therefore decided to randomly remove a certain amount of data among all the fossil.
+For example, when removing 75\% of the data for the fossil taxa ($M_F$=75\%) different scenarios could happen, ranging from a "homogeneous" incompleteness data distribution (each specific fossil taxa had 75\% of missing data) to a "heterogeneous" incompleteness data distribution  (80\% of the specific fossil taxa had 5\% available data and 21\% of the specific fossil taxa had 100\% available data).
+In empirical matrices, the incompleteness data distribution probably lies in between both models.
+
+Also, a second argument for using this "simplistic" way to deal with the missing data in the fossil record follows our main recommendation from this study: the best way to improve the quality of Total Evidence trees is to have a maximum number of living species with available morphological data.
+In fact lines 584 to 589 in the first version of the manuscript (line @@@ in the revised version), we argue that even if the effect of the $M_F$ parameter is important, there is unfortunately not much we can do about it since the completeness of this part of the matrix solely depends on exceptional palaeontological discoveries and excavation effort.
+We feel that these parameters, however crucial for our understanding of macroevolution, are unfortunately not predicable and recommendations such as "we need more fossil data" are empty and less practical than "we need more living species data".
+
 \end{enumerate}

### 2015-07-27 Manuscript/MPESubmission/Revision/response.tex
Graphical abstract revision
Standardised Bhattacharyya coefficient across whole work

@@
 
+We gladly followed reviewer 1's graphical abstract suggestions and have changed it to be similar to the figure 5.
+% TG: It's freaking not catchy at all but at this point I give up.
+I've tried many ways but the flippin' paper IS NOT attractive.
+It might be useful, well written, interesting, citable, etc...
+But it won't be sexy.
+
 \item{\textcolor{blue}{\textbf{Highlights:} wording seems awkward}}
+
+We rephrased the highlights of the paper as follows: %3 to 5 bullet points (maximum 85 characters, including spaces, per bullet point). TG: probably need improvement here.
+
+1. The Total Evidence methods allows to combine living and fossil data.
+2. This study test the effect of missing morphological data on topology.
+3. The number of living taxa with data as the most effect on recovering topology.
+4. Bayesian inference outperforms Maximum Likelihood for correct clade recovery.
+5. We suggest that the sampling effort should be focused on data for living taxa.
 
@@
 
+% To do (see reviewer 1 specific comment 13)
+
 \item{\textcolor{blue}{\textbf{Abbreviation:} One suggestion is to consider abbreviating Maximum Likelihood and Bayesian Inference as ML and BI respectively early in the methods and use throughout.}}
 
+We do understand the reviewers concern are agree that ML and BI are two commonly used abbreviations.
+Nonetheless, we would rather prefer not abbreviate these terms: in fact, even though they might add some "smoothness" % TG: that's not the word
+to the sentences where Maximum Likelihood and/or Bayesian Inference are stated, we think they might add confusion or difficulties to the whole paper.
+Spelling out the abbreviation might be locally "heavy" (i.e.
+in the sentence) but at least are completely clear and does not ask the reader to look back for the actual meaning of such or such abbreviation.
+However, we leave this decision to the editor if she judges it better to abbreviate those terms. % TG: or maybe this statement is useless? Of course we leave it to the editor...
+
 \item{\textcolor{blue}{\textbf{Methodology - True tree:} I suggest it warrants comparing at least the 'best' tree to the 'true' tree.}}
+
+We compared both the ``best'' tree and the ``missing-data'' trees to the ``true'' tree as suggested by both reviewer.
+We did not, however, include the results to the main text (available in appendix @@@).
+We justify this decision to both reviewers specific concerns (see reviewer 1 specific comment 25 and reviewer 2 comment 1).
 
@@
 
+We dealt with this legitimate suggestion and ran an extra analysis on a subset of matrices (20 chains and only the trees with the same number of parameters; i.e.
+$M_L$=0\%, $M_F$=0\% and $M_C$=0\%; $M_L$=10\%, $M_F$=10\% and $M_C$=10\%, etc...) were we reran the Bayesian inference using a random tree rather than the ``true'' tree as a starting tree (default MrBayes procedure).
+We found no effect of the starting tree (random or ``true'') on the resulting topology.
+We added this analysis to the appendix @@@ and added a short note in the main text.
+See details at reviewer 1 specific comment 46 below.
+
 \item{\textcolor{blue}{\textbf{Results - Bhattacharyya meaning:} Some descriptive statistics are given, but it would help to tell us if these differences are meaningful.
-For example, on line 414 the Bhattacharrya Coefficients are given suggesting RF distances are lower for ML than BI trees at 0.69, 0.48, and 0.66 - but if this coefficient represents the probability of overlap between distributions then shouldn't the coefficients be <0.05 to say that one is lower than the other?
-In the table, mean and median values of <0.01 and >0.85 are highlighted as being the most extreme.
+For example, on line 414 the Bhattacharrya Coefficients are given suggesting RF distances are lower for ML than BI trees at 0.69, 0.48, and 0.66 - but if this coefficient represents the probability of overlap between distributions then shouldn't the coefficients be $<$0.05 to say that one is lower than the other?
+In the table, mean and median values of $<$0.01 and $>$0.85 are highlighted as being the most extreme.
 Later the focus on Bhattacharyya Coefficients is 1 when distributions overlap completely and 0 when they do not - should this be the criterion?
 Just make it clear early on so the readers knows what to think when they see these numbers. }}
+
+We clarified and unified % TG: or standardised?
+the interpretation of the Bhattacharyya Coefficient throughout the manuscript.
+We advocate that two distributions are significantly different when their probability of overlapping (the Bhattacharyya Coefficient) is $>$0.05 and that they are significantly similar when this probability is $>$0.95.
+Even though other values (e.g.
+0.85) are not conclusive, we still reported them to show the relative overlap of the distributions (i.e.
+two distributions with a coefficient of 0.85 are not significantly similar but are closer to each other than two distributions with a coefficient of 0.15; which are themselves not significantly different but still less closer to each other than the previous ones).
+We added the following statement to the main text and to the appendix B:
+"When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
+When this coefficient is $>$0.95 both distributions are significantly similar.
+Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions." line (@@@)
 

### 2015-07-27 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixB.tex
Standardised Bhattacharyya coefficient across whole work

@@
 
+When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
+When this coefficient is $>$0.95 both distributions are significantly similar.
+Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions.
+
 \bibliographystyle{sysbio}

### 2015-07-27 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixC.tex
Standardised Bhattacharyya coefficient across whole work

@@
   \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.30 & 0.31 & 0.69 & 0.61 & 0.77 & 1.00 \\ 
-                                                                       & $Tr$ & 0.79 & 0.81 & 0.84 & 0.86 & 0.85 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.22 & 0.29 & 0.36 & 0.54 & 0.69 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.30 & 0.31 & 0.69 & 0.61 & 0.77 & \textbf{1.00} \\ 
+                                                                       & $Tr$ & 0.79 & 0.81 & 0.84 & 0.86 & 0.85 & \textbf{1.00} \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & \textbf{0.03} & 0.22 & 0.29 & 0.36 & 0.54 & 0.69 \\ 
                                                                        & $Tr$ & 0.08 & 0.42 & 0.53 & 0.51 & 0.74 & 0.78 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.49 & 0.61 & 0.51 & 0.67 & 0.74 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & \textbf{0.02} & 0.49 & 0.61 & 0.51 & 0.67 & 0.74 \\ 
                                                                        & $Tr$ & 0.21 & 0.61 & 0.70 & 0.63 & 0.81 & 0.81 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.01 & 0.02 & 0.02 & 0.02 & 0.03 & 0.04 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & \textbf{0.01} & \textbf{0.02} & \textbf{0.02} & \textbf{0.02} & \textbf{0.03} & \textbf{0.04} \\ 
                                                                        & $Tr$ & 0.08 & 0.69 & 0.78 & 0.64 & 0.79 & 0.84 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.01 & 0.02 & 0.02 & 0.04 & 0.08 & 0.09 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & \textbf{0.01} & \textbf{0.02} & \textbf{0.02} & \textbf{0.04} & 0.08 & 0.09 \\ 
                                                                        & $Tr$ & 0.21 & 0.74 & 0.75 & 0.68 & 0.84 & 0.87 \\ 
-    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.75 & 0.85 & 0.85 & 0.95 & 1.00 \\ 
-                                                                       & $Tr$ & 0.91 & 0.92 & 0.96 & 0.95 & 0.97 & 0.98 \\ 
+    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.75 & 0.85 & 0.85 & \textbf{0.95} & \textbf{1.00} \\ 
+                                                                       & $Tr$ & 0.91 & 0.92 & \textbf{0.96} & \textbf{0.95} & \textbf{0.97} & \textbf{0.98} \\ 
    \hline
@@
   \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.00 & 0.25 & 0.48 & 0.50 & 0.76 & 1.00 \\ 
-                                                                       & $Tr$ & 0.38 & 0.69 & 0.75 & 0.72 & 0.80 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.18 & 0.32 & 0.36 & 0.47 & 0.77 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & \textbf{0.00} & 0.25 & 0.48 & 0.50 & 0.76 & \textbf{1.00} \\ 
+                                                                       & $Tr$ & 0.38 & 0.69 & 0.75 & 0.72 & 0.80 & \textbf{1.00} \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & \textbf{0.03} & 0.18 & 0.32 & 0.36 & 0.47 & 0.77 \\ 
                                                                        & $Tr$ & 0.08 & 0.34 & 0.40 & 0.38 & 0.53 & 0.55 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.47 & 0.71 & 0.60 & 0.86 & 0.94 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & \textbf{0.02} & 0.47 & 0.71 & 0.60 & 0.86 & 0.94 \\ 
                                                                        & $Tr$ & 0.21 & 0.54 & 0.62 & 0.56 & 0.64 & 0.80 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.00 & 0.01 & 0.01 & 0.01 & 0.03 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & \textbf{0.00} & \textbf{0.00} & \textbf{0.01} & \textbf{0.01} & \textbf{0.01} & \textbf{0.03} \\ 
                                                                        & $Tr$ & 0.08 & 0.38 & 0.54 & 0.49 & 0.70 & 0.75 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.02 & 0.02 & 0.02 & 0.04 & 0.04 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & \textbf{0.00} & \textbf{0.02} & \textbf{0.02} & \textbf{0.02} & \textbf{0.04} & \textbf{0.04} \\ 
                                                                        & $Tr$ & 0.21 & 0.29 & 0.66 & 0.54 & 0.72 & 0.82 \\ 
     Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.69 & 0.72 & 0.71 & 0.72 & 0.72 \\ 
-                                                                       & $Tr$ & 0.91 & 0.91 & 0.91 & 0.93 & 0.92 & 0.98 \\ 
+                                                                       & $Tr$ & 0.91 & 0.91 & 0.91 & 0.93 & 0.92 & \textbf{0.98} \\ 
    \hline
@@
   \hline
-    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.03 & 0.32 & 0.66 & 0.55 & 0.75 & 1.00 \\ 
-                                                                       & $Tr$ & 0.51 & 0.69 & 0.80 & 0.76 & 0.80 & 1.00 \\ 
-    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.17 & 0.21 & 0.31 & 0.46 & 0.68 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & \textbf{0.03} & 0.32 & 0.66 & 0.55 & 0.75 & \textbf{1.00} \\ 
+                                                                       & $Tr$ & 0.51 & 0.69 & 0.80 & 0.76 & 0.80 & \textbf{1.00} \\ 
+    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & \textbf{0.03} & 0.17 & 0.21 & 0.31 & 0.46 & 0.68 \\ 
                                                                        & $Tr$ & 0.08 & 0.31 & 0.39 & 0.39 & 0.56 & 0.61 \\ 
-    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.44 & 0.47 & 0.52 & 0.78 & 0.90 \\ 
+    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & \textbf{0.02} & 0.44 & 0.47 & 0.52 & 0.78 & 0.90 \\ 
                                                                        & $Tr$ & 0.21 & 0.52 & 0.59 & 0.55 & 0.66 & 0.77 \\ 
-    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.01 & 0.01 & 0.02 & 0.02 & 0.03 \\ 
+    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & \textbf{0.00} & \textbf{0.01} & \textbf{0.01} & \textbf{0.02} & \textbf{0.02} & \textbf{0.03} \\ 
                                                                        & $Tr$ & 0.08 & 0.47 & 0.62 & 0.51 & 0.66 & 0.73 \\ 
-    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.02 & 0.04 & 0.04 & 0.05 & 0.06 \\ 
+    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & \textbf{0.00} & \textbf{0.02} & \textbf{0.04} & \textbf{0.04} & 0.05 & 0.06 \\ 
                                                                        & $Tr$ & 0.21 & 0.45 & 0.64 & 0.57 & 0.74 & 0.79 \\ 
     Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.69 & 0.73 & 0.73 & 0.76 & 0.81 & 0.86 \\ 
-                                                                       & $Tr$ & 0.91 & 0.92 & 0.93 & 0.94 & 0.96 & 0.99 \\ 
+                                                                       & $Tr$ & 0.91 & 0.92 & 0.93 & 0.94 & \textbf{0.96} & \textbf{0.99} \\ 
    \hline
@@
    & Tr & 0.86 & 0.67 & 0.81 & 0.81 \\ 
-Maximum Likelihood bootstraps & RF & 0.98 & 0.96 & 0.97 & 0.78 \\ 
-   & Tr & 0.99 & 1.00 & 1.00 & 0.96 \\ 
+Maximum Likelihood bootstraps & RF & \textbf{0.98} & \textbf{0.96} & \textbf{0.97} & 0.78 \\ 
+   & Tr & \textbf{0.99} & \textbf{1.00} & \textbf{1.00} & \textbf{0.96} \\ 
 Bayesian posterior trees & RF & 0.86 & 0.93 & 0.94 & 0.68 \\ 
-  & Tr & 0.98 & 0.99 & 0.99 & 0.91 \\ 
+  & Tr & \textbf{0.98} & \textbf{0.99} & \textbf{0.99} & 0.91 \\ 
  \hline
@@
   & Tr & 0.70 & 0.62 & 0.59 & 0.73 \\ 
-Bayesian consensus vs. Bootstraps & RF & 0.02 & 0.01 & 0.01 & 0.00 \\ 
+Bayesian consensus vs. Bootstraps & RF & \textbf{0.02} & \textbf{0.01} & \textbf{0.01} & \textbf{0.00} \\ 
   & Tr & 0.78 & 0.54 & 0.62 & 0.59 \\ 
-Bayesian consensus vs. Bayesian posterior trees & RF & 0.02 & 0.02 & 0.04 & 0.01 \\ 
+Bayesian consensus vs. Bayesian posterior trees & RF & \textbf{0.02} & \textbf{0.02} & \textbf{0.04} & \textbf{0.01} \\ 
    & Tr & 0.75 & 0.66 & 0.64 & 0.56 \\ 
 Boostraps vs. Bayesian posterior trees & RF & 0.85 & 0.72 & 0.73 & 0.85 \\ 
-   & Tr & 0.96 & 0.91 & 0.93 & 0.98 \\ 
+   & Tr & \textbf{0.96} & 0.91 & 0.93 & \textbf{0.98} \\ 
  \hline

### 2015-07-27 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Standardised Bhattacharyya coefficient across whole work

@@
 Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
+When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
+When this coefficient is $>$0.95 both distributions are significantly similar.
+Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions.
 To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.

### 2015-07-27 Analysis/MissingData_plot.R
Updated figure 4 (with Tr below 0)

@@ -52,13 +52,16 @@ multi.TreeCmp.plot( data.sets , par_MC, metrics[1], col=pal, las=2, probs=c(95, 
 xaxis.labs<-c("0%", "10%", "25%", "50%", "75%")
 #Missing living Triplets
 par(mar=c(5,5,1,0) + 0.1)
-multi.TreeCmp.plot( data.sets , par_ML, metrics[2], col=pal, las=2, probs=c(95, 50), ylim=ylim, xaxis=xaxis.labs, ylab="Normalised Triplets metric", las=2, xlab="Missing living taxa (ML %)") 
+multi.TreeCmp.plot( data.sets , par_ML, metrics[2], col=pal, las=2, probs=c(95, 50), ylim=c(-0.5,1), xaxis=xaxis.labs, ylab="Normalised Triplets metric", las=2, xlab="Missing living taxa (ML %)")
+abline(h=0, lty=3)
 #Missing fossil Triplets
 par(mar=c(5,3,1,2) + 0.1)
-multi.TreeCmp.plot( data.sets , par_MF, metrics[2], col=pal, las=2, probs=c(95, 50), ylim=ylim, xaxis=xaxis.labs, ylab=NULL, las=2, xlab="Missing data in the fossil record (MF %)")
+multi.TreeCmp.plot( data.sets , par_MF, metrics[2], col=pal, las=2, probs=c(95, 50), ylim=c(-0.5,1), xaxis=xaxis.labs, ylab=NULL, las=2, xlab="Missing data in the fossil record (MF %)")
+abline(h=0, lty=3)
 #Missing character Triplets
 par(mar=c(5,2,1,4) + 0.1)
-multi.TreeCmp.plot( data.sets , par_MC, metrics[2], col=pal, las=2, probs=c(95, 50), ylim=ylim, xaxis=c("100","90","75","50","25"), ylab=NULL, las=2, xlab="Number of characters (NC)") 
+multi.TreeCmp.plot( data.sets , par_MC, metrics[2], col=pal, las=2, probs=c(95, 50), ylim=c(-0.5,1), xaxis=c("100","90","75","50","25"), ylab=NULL, las=2, xlab="Number of characters (NC)") 
+abline(h=0, lty=3)
 
 #Back to default
 par(op)

### 2015-07-28 Manuscript/MPESubmission/Revision/response.tex
Updated figure 4 (with Tr below 0)
Updated response (figure 4 story)

@@
 
+The difference between the Bayesian consensus and the Bayesian posterior trees can be due to two factors:
+\begin{enumerate}
+\item first, the way we assess topological recovery is slightly different in each case: for the Bayesian consensus trees, we simply compared the "missin-data" consensus tree to the "best" consensus tree; for the Bayesian posterior trees however, we randomly compared 1000 "missing-data" Bayesian posterior trees to 1000 "best" Bayesian posterior trees.
+Therefore, even if the comparison methodology is the same, the fact that the second methods draws random trees, we expect the topological recovery score to be "blurred" in comparison to other comparisons that do not involve random comparisons.
+\item secondly, the Bayesian posterior trees are always resolved and thus, in the same way as the Maximum Likelihood tree, more likely to contain "false-positive" nodes.
+On the other hand, because the consensus tree is a strict majority consensus rule, only nodes with a a good support will be represented in the tree and all other nodes will be collapsed.
+This reduces the amount of "false-positive" nodes (since an unresolved tree is always better than a wrongly resolved one) and increases clade conservation but also increases the placement of wild-card taxa (since it is likely that some Triplets will just not exist in a poorly resolved consensus tree).
+\end{enumerate}
+Both points were discussed in the former manuscript lines 566 and 573 (or lines @@@ and @@@ in the revised version). %TG: maybe expand this bit then?
+
+% TG: I'm confused for the second part of the question:
+% TG: In figure 4 and in Table 1, if you look at the Bayesian consensus (grey) trees and the Bayesian posterior trees (orange) for RF, they do not really overlap.
+Ever.
+The 0.11 probability must come from the parameters combinations that, when is really shitty must show a little bit of overlap.
+But for each parameter one by one: nope.
+No overlap.
+% TG: On the other side, in figure 4 and Table 1, for Bayesian consensus and Bayesian posterior for Tr, the do overlap much more.
+And that's what's in the Table 1 (min 0.21; median 0.56; and max 0.87).
+% TG: Did the reviewer just miss read the labels of the Table 1 or am I miles away from what he means?
+
 \item{\textcolor{blue}{\textbf{Results - Bayesian Triplets:} Another important result that is not mentioned is that using the wildcard taxon metric (Tr), Bayesian consensus trees do quite poorly with missing data.
@@
 
+We modified the figures 4 and 5 to show Normalised Triplets metric values below 0 and added a dashed line at 0 to underline the difference in scale between both metrics.
+% TG: a a paragraph in the results about the wild-card fossils
+As mentioned in the previous comment (reviewer 1 comment 9), this is due to the fact that the Bayesian consensus tree will be less and less resolved when missing data increases.
+This has two opposite effects.
+The first one being to retain clades in a be
+
 \item{\textcolor{blue}{\textbf{Conclusion:} The authors should clarify how their results compare to those of previous studies more directly to emphasize the nuances that make this study unique. }}
+
+We modified the conclusion as follows (see reviewer 1 specific comments 52)
 
@@
 
-%We apologies for this silly mistake.
+We apologies for this silly mistake and fixed the colors in the figure. We also changed the color scheme to match with the general color scheme of the paper (black, grey, orange and blue).
 
 \item{\textcolor{blue}{\textbf{Figure 4-5:} Am I wrong or shouldn't the y-axis for Normalised Triplets metric go below 0 - it looks those metrics for many missing data combinations have values below zero indicating trees are more different from the 'best' tree than expected by chance (Appendix B 1.1., see Appendix C Tables 1-3) and that is important to show.}}
+
+We fixed this graphical mistake as state above in reviewer 1 comment 13.
 
@@
 But I feel it is necessary to summarize the following questions: what did those studies find, what are the questions those studies left open, and how does this study address them?}}
+
+% specifically
+% what did those studies find: missing data is ok but must overlap
+% what are the questions those studies left open: is it the same when different type of data are used (molecular + morphological)
+% how does this study address them: we test whether m
 \end{enumerate}

### 2015-07-29 Manuscript/MPESubmission/Revision/response.tex
Added highlights + morph rate justification

@@
 As mentioned in the previous comment (reviewer 1 comment 9), this is due to the fact that the Bayesian consensus tree will be less and less resolved when missing data increases.
-This has two opposite effects.
-The first one being to retain clades in a be
+This has two related effects.
+The first one being to retain "good" clade conservation (i.e. avoiding "false-positive" nodes by collapsing unsupported ones).
+And the second one is that when these clades are collapsed, it reduces the amount of available triplets and thus increases the "wild-card taxa" score (Normalised Triplets metric).
+In fact, when a clade ((a;b);c); is collapsed to (a;b;c), the whole clade is conserved (a, b and c are still in a common clade) however they also all become wild-card taxa: the position of a regarding b and c is now incorrect.
 
@@
 
-% specifically
-% what did those studies find: missing data is ok but must overlap
-% what are the questions those studies left open: is it the same when different type of data are used (molecular + morphological)
-% how does this study address them: we test whether m
+% specifically state line by line where we dealt with which and which question.
+
 \end{enumerate}
@@
 We thank the reviewer for pointing out this inconsistency, however, 'data' is now commonly used for the singular case (Pinker, S.
-The Sense of Style: The Thinking Person's Guide to Writing in the 21st Century.
+\textit{The Sense of Style: The Thinking Person's Guide to Writing in the 21st Century}.
 Penguin, 2014.
-p 271 and Guardian and Observer style guide (2015): \url{http://goo.gl/cmKjNe}).
+p 271; and Guardian and Observer style guide (2015): \url{http://goo.gl/cmKjNe}).
 
@@
 
-"Until now, only few attempts have been made to study the impact of this issue on phylogenetic inference using Total Evidence methods (e.g.
-Wiens et al., 2005; Pattinson et al., 2014).%\citep[e.g.][]{Wiens01102005,pattinsonphylogeny2014}
-For example, Wiens et al. (2005) %\cite{Wiens01102005}
- only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
-Furthermore, Pattinson et al. (2014) %\cite{pattinsonphylogeny2014}
- assess the effect of removing both molecular and morphological data (``artificial extinction'') on the topology.
-These studies, however, are limited by their empirical approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014).
-In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
+"Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
+using combined molecular and morphological data; e.g.
+Wiens et al., 2005; Pattinson et al., 2014).
+For example, Wiens et al.
+(2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
+Furthermore, Pattinson et al.
+(2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
+These studies, however, are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014)."
+
+"In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
 We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa."
@@
 Nonetheless, we added the result of such analysis in the \textbf{Appendix-A Morphological rates estimations results} section.
+We linked to this analysis in the methods part as follows:
+"The detailed MrBayes parameters are available in Appendix A along with details on the $\alpha$ parameter estimation."
+
 % TG: what do you think?
 I don't want to attract the reviewer's attention too much (i.e.
-that's another shit load of analysis to do to do it properly) and certainly not the reader (it's not relevant to the whole paper) but sticking that to the supplementary kind of does what the reviewers want to know.
+that's another shit load of analysis to do to do it properly) and certainly not the reader's attention either (it's not relevant to the whole paper) but sticking that to the supplementary kind of does what the reviewers wants to know.
 
@@
 
-We modified the sentence as follows: "We selected this model and parameters settings to be a compromise between realism from empirical datasets (e.g.
-Kelly et al., 2014) %\citep[e.g.][]{kellymolecular2014}
+We modified the sentence as follows:
+"We selected this model and parameters settings to be a compromise between realism from empirical datasets (e.g. Kelly et al., 2014) %\citep[e.g.][]{kellymolecular2014}
 and models with few parameters (cf. Springer et. al., 2012) %\citep[cf.][]{springermacroevolutionary2012}
@@
 
-% To do!
-% TG: a fair point... Can I just say: "Coz' April told me it was fine?"
-% TG: one way to justify it should be to say that we tried to avoid homoplasy but we actually wanted to add a little bit of it to be kind of realistic.
-% TG: in Wright & Hillis, low rate characters make the tree as "shitty" as the "shittiest" ML estimation.
-Now this "shittiness" is in quotes because it's actually a relative "shittiness".
-Using the Mk model with no missing data, the Topological difference is actually very low (so it's not "shitty") for low rate characters.
-% TG: then a low rate character with 75% missing gets the maximum topological difference with the Mk model (which is still all right).
-% TG: also, in our case, the actual character rate is not that low.
-We sample from a low distribution but force them to be non-invariable (therefore, our pratical minimum rate is 0.02 because for one character, at least one species must have a different state (1/51)).
+We initially chose this parameter from personal communication with April Wright and to be consistent with the molecular rate parameter.
+We like to point out that, in figure 3 from Wright and Hillis 2014, even if the Topological error is high when rates are near 0, it is mainly the case for parsimony and both parsimony and Mk with missing data, the fitted model for the Mk model with no missing data (light purple line) is still at lowest when the rates are near 0.
+Topological error, however, becomes minimal for all scenari (Parsimony and Mk, with or without missing data) at a rate of 0.5 (see modified figure 3 from Wright and Hillis (2014) in this response; we added the tick marks and the 0.5 rate limit by using the ruler in Adobe Illustrator).
+
+\begin{figure}[]
+\centering
+    \includegraphics[keepaspectratio=true]{response_fig1.pdf}
+\caption{Figure 3 from Wright and Hillis 2014 with added x axis ticks. The red line represents the 0.5 rate.}
+\end{figure}
+
+We do agree, however, that if rates get lower than 0.5, variance in topological error increases (especially for the Mk model with missing data).
+Therefore, we implemented the \textit{a posteriori} matrix selection step to our analysis (i.e.
+keeping only the ``complete'' matrices that led to trees with a median bootstrap $>$ 50; explained lines 259-268 in the submitted manuscript and lines @@@ and @@@ in the revised one).
+
+We rephrased the section underlined by the reviewer as follows:
+
+"We used low evolutionary rate parameters to be consistent with the molecular rates parameters and to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal (Wright and Hillis, 2014).
+Note that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increases variance in topological error but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004)."
 

### 2015-07-29 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Added highlights + morph rate justification

@@
 \documentclass[12pt,letterpaper]{article}
+
+%---------------------------------------------
+%
+%       THIS IS THE HIGHLIGHTED VERSION.
+%       For the final version, carefully remove:
+%       '\protect'
+%       '\hl{' (don't forget the '}'
+%
+%---------------------------------------------
 
@@
 \usepackage{xcolor}
+\usepackage{color,soul}
 \usepackage{changepage}
@@
 \section{Introduction}
-Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, the many large-scale macroevolutionary studies focus solely on living species (e.g.
+Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, the \hl{many large-scale macroevolutionary studies focus solely on living species} (e.g.
 \citealp{meredithimpacts2011,jetzthe2012}).
@@
 Up to now, three main approaches have been used to place both living and fossil taxa into phylogenies.
-These approaches differ mainly in how they treat the fossil taxa and it's data.
-One can use the fossil as tips or as nodes in the phylogeny and use it's age only, it's morphology only or it's age and morphology jointly.
+\hl{These approaches differ mainly in how they treat the fossil taxa and it's data.
+One can use the fossil as tips or as nodes in the phylogeny and use it's age only, it's morphology only or it's age and morphology jointly.}
 Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
@@
 Neontologists, on the other hand, more commonly use probabilistic approaches (e.g. Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies \citep{zuckerkandl1965}.
+\hl{Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies} \citep{zuckerkandl1965}.
 There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the ``best'' approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
@@
 This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny \citep[known as tip-dating;][]{ronquista2012}, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
-Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+\hl{Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies} \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014,Arcila2015131}.
 Although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
@@
 For example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data is (nearly) always unavailable.
-Therefore Total Evidence matrices are likely to contain a large proportions of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
+Therefore \hl{Total Evidence matrices are likely to contain a large proportions of missing data} that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately (as long as enough data overlaps in each case, and missing data are not phylogenetically biased \citep{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately (\hl{as long as enough data overlaps in each case, and missing data are not phylogeneticaly biased} \citep{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
-Until now, only few attempts have been made to study study the impact of this issue on phylogenetic inference using Total Evidence methods \citep[e.g.][]{Wiens01102005,pattinsonphylogeny2014}.
-For example, \cite{Wiens01102005} only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
-Furthermore, \cite{pattinsonphylogeny2014} assess the effect of removing both molecular and morphological data (``artificial extinction'') on the topology.
-These studies, however, are limited by their empirical approach \citep{Wiens01102005} and the scheme of generating missing data \citep{pattinsonphylogeny2014}.
+\hl{Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework \protect\citep[\hl{i.e.
+using combined molecular and morphological data}; e.g.][]{Wiens01102005,pattinsonphylogeny2014}.}
+\hl{For example, \protect\cite{Wiens01102005} }\hl{only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.}
+\hl{Furthermore, \protect\cite{pattinsonphylogeny2014} }\hl{assess the effect of removing both molecular and morphological data (``artificial extinction'') on the topology.}
+\hl{These studies, however, are limited by their empirical-only approach \protect\citep{Wiens01102005} }\hl{and the scheme of generating missing data \protect\citep{pattinsonphylogeny2014}.}
 
-In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
-We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa.
+\hl{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
+We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa.}
 We focus on the effect of missing data on our ability to recover a ``best'' tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
 Although branch length estimation is also important \citep[namely for timing extinction and/or speciation events; e.g.][]{ronquista2012}, we do not consider branch lengths in this study.
-This is partially due to difficulties with simulating branch lengths and topology simultaneously, but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+This is partially due to difficulties with simulating branch lengths and topology simultaneously, \hl{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints} \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework.
@@
 \section{Materials and Methods}
-To explore how missing data in the morphological partition of Total Evidence matrices influences tree topology, we used the following protocol (Figure 1):
+To explore how missing data in \hl{the morphological partition of Total Evidence matrices} influences tree topology, \hl{we used the following protocol (Figure 1):}
 \begin{enumerate}
@@
 Next, we generated a molecular and a morphological matrix from the ``true'' tree.
-The molecular matrix was simulated from the ``true'' tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
+The molecular matrix was \hl{simulated} from the ``true'' tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
 The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1) and transition/transversion rate of two \citep{douadycomparison2003}.
-The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
-In practice, such a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal \citep{Hassanin1998611,EstoupHomoplasy}.
-Also, we chose this low value to be consistent with our protocol for simulating morphological characters (see below).
+\hl{The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5} \citep{yangamong-site1996}.
+\hl{In practice, such a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal} \citep{Hassanin1998611,EstoupHomoplasy}.
+\hl{Also, we chose this low value to be consistent with our protocol for simulating morphological characters (see below).}
 %We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal. %Justify
 %We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, total computational time for the whole analysis was around 150 CPU years).
-We selected this model and parameters settings to be a compromise between realism from empirical datasets \citep[e.g.][]{kellymolecular2014} and models with few parameters \citep[cf.][]{springermacroevolutionary2012} to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).
+\hl{We selected this model and parameters settings to be a compromise between realism from empirical datasets} \citep[e.g.][]{kellymolecular2014} \hl{and models with few parameters} \citep[cf.][]{springermacroevolutionary2012} \hl{to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).}
 All the molecular information for fossil taxa was replaced by missing data ("?").
 
-We simulated the morphological matrix using the rTraitDisc function from the R package ape v. 3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
-We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A and Fig.
-A1 within).
+\hl{We simulated the morphological matrix using the rTraitDisc function from the R package ape v. 3.0-11} \citep{paradisape:2004}\hl{ to generate a matrix of 100 character sites for 51 taxa.}
+\hl{We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A and Fig.
+A1 within).}
 We then ran an independent discrete character simulation for each character using the ``true'' tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
@@
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used low evolutionary rate parameters to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wagner2000,davalosintegrating2014,wrightbayesian2014}.
+\hl{We used low evolutionary rate parameters to be consistent with the molecular rates parameters and to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal} \citep{wrightbayesian2014}.
+\hl{Note that} \cite{wrightbayesian2014} \hl{have shown that low morphological rates ($<0.5$) increases variance in topological error but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal} \citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}.
 
@@
 Even using these priors, it took $~$ 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-The detailed MrBayes parameters are available in Appendix A.
+\hl{The detailed MrBayes parameters are available in Appendix A along with details on the $\alpha$ parameter estimation.}
 

### 2015-07-29 Manuscript/MPESubmission/Revision/response.tex
True trees analysis

@@
 
-%We corrected the plural/singular of data throughout the text. 
 % TG: actually I'm not going to change it:
@@
 We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa."
-%Pattinson: artificial fossil - removing molecular data from living species to see how topology changes.
-Pattinson use combined analysis but only to understand the patterns of fossilisation (via artificial fossil).
-%Wiens only focuses on a special case
-%Limitation of Pattinson was the limited scheme for generating missing data from the datasets (specific to primates and observed patterns).
-%In our case, we (a) don't focus on any group or any pattern, our approach is purely theoretical.
-%The difference from Wiens or Pattinson is that we thoroughly investigate the effects of eliminating living versus fossil data or all data overall.
-%(b) We also propose a thorough comparison between the ML and Bayesian inference and a measure two clear effects that missing data could have on topology: displacing wild-card taxa and not conserving clades structure.
-
-%Emphasize on the questions and the uniqueness
-%The authors could argue that a limitation of the Pattinson study, which they address here, was the limited scheme for generating missing data from the datasets.
-%In this study, the authors perturbed the datasets more thoroughly in a clearly structured experimental design to investigate the effects of eliminating living versus fossil data or all data overall.
-%That is one aspect of this study that I perceived as new and unique.
-%This study also used a unique way of evaluating the tree inference under missing data scenarios.
 
@@
 We modified the sentence as follows:
-"We selected this model and parameters settings to be a compromise between realism from empirical datasets (e.g. Kelly et al., 2014) %\citep[e.g.][]{kellymolecular2014}
-and models with few parameters (cf. Springer et. al., 2012) %\citep[cf.][]{springermacroevolutionary2012}
-to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)."
+"We selected this model and parameters settings to be a compromise between realism from empirical datasets (e.g.
+Kelly et al., 2014) and models with few parameters (cf.
+Springer et.
+al., 2012) to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)."
 
@@
 
-We modified the following sentence as suggested: "We modified the complete matrix to obtain matrices with missing data" (line @@@).
+We modified this whole section following reviewer 2 comment 2.
 
@@
 As reviewer 1 states: "[this analysis can give] insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it.".
-We've added the results of these comparisons to the supplementary materials (Appendix 3 - @@@).
+We've added the results of these comparisons to the supplementary materials (Appendix C - figures C5 and C6).
 However, even though we agree with the comments of both reviewer that knowing the difference between the true and best/missing-data trees is important, we do not think that it "adds an additional layer of complexity that can't be accounted for via simulation" (reviewer 2, comment 1).
@@
 \section{Reviewer 2:}
+Note that we split the second comment from reviewer 2 in three different comments (comments 2,3 and 4 below) to improve the readability of our response to the reviewers.
 

### 2015-07-29 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixB.tex
True trees analysis

@@
 
+\subsection{Differences between the ``true'' and the ``best'' trees.}
+We compared every ``true'' trees (i.e.
+the trees \textbf{used to create} the ``complete'' matrices) to the ``best'' trees (i.e.
+the trees \textbf{inferred from} the ``complete'' matrices) in figure ~\ref{Fig_Supp_True_Best}.
+Note that the difference between the ``true'' and the ``best'' trees represents the effect of the parameters choice and the algorithms used to create the ``complete'' matrix as well the as the capacity of RAxML and MrBayes to infer phylogenies from this particular matrices (i.e.
+small sized and generated using specific algorithms).
+This does not affect, however, the results of our analysis since we deliberately compared the the ``missing-data'' trees to the ``best'' tree rather than to the ``true'' tree.
+
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{SupplementaryFigures/True_vs_Best_trees.pdf}
+    \caption{Pairwise comparisons between the 50 ``true'' trees and the 50 ``best'' trees from the Maximum Likelihood and Bayesian inference methods.
+The horizontal blue and red lines represents respectively the 95\% and the 50\% confidence intervals.}
+\label{Fig_Supp_True_Best} 
+\end{figure}
+
 \bibliographystyle{sysbio}

### 2015-07-29 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixC.tex
Updated figure 4 (with Tr below 0)
True trees analysis

@@
 \label{Fig_Supp_paircomp_Bayt}
+\end{figure}
+
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{SupplementaryFigures/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
+    \caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey).
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+\label{Fig_Supp_BootBayt_allparam} 
+\end{figure}
+
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{SupplementaryFigures/Singles_True.pdf}
+    \caption{Effect of increasing missing data on recovering the ``true'' tree topology (the tree used for starting our simulations) for the Maximum Likelihood trees (black) and Bayesian consensus trees (grey).
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+\label{Fig_Supp_Singles_true} 
+\end{figure}
+
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{SupplementaryFigures/Treesets_True.pdf}
+    \caption{Effect of increasing missing data on topological recovering the ``true'' tree topology (the tree used for starting our simulations) for the Maximum Likelihood Bootstrap trees (black) and Bayesian posterior tree distribution (grey).
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
+Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
+\label{Fig_Supp_Treesets_true} 
 \end{figure}

### 2015-07-29 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
True trees analysis

@@
 % TG: this part is entirely re-written.
-To explore the effect of missing data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
+\hl{To explore the effect of missing data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
 Hereafter, we call these matrices with missing morphological data the ``missing-data'' matrices.
-We removed morphological following three main data incompleteness parameters:
+We removed morphological following three main data incompleteness parameters:}
 \begin{enumerate}
-\item{The proportion of missing living taxa ($M_{L}$).}
-This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
+\item{\hl{The proportion of missing living taxa ($M_{L}$).}}
+\hl{This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
 It represents the number of living taxa that are present in the matrix but have only molecular data available.
-This reflects the fact that, because of the increasing facility of collecting molecular data, morphological data for living species is rarely collected \citep{GuillermeCooperMissing}.
+This reflects the fact that, because of the increasing facility of collecting molecular data, morphological data for living species is rarely collected} \citep{GuillermeCooperMissing}.
 % TG: see what I did here?
 ;)
-Therefore, many living species will have only molecular data available.
-In practice, we removed all the morphological data from randomly chosen living with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.
-\item{The proportion of missing data in the fossil record ($M_{F}$).}
-This missing data parameter represents the completeness of the fossil record.
-In fact, due to preservation biases, missing data for fossil taxa are common \citep{sansomfossilization2013}. 
-In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
-\item{The number of morphological characters for both living and fossil taxa ($N_{C}$).}
-This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
+\hl{Therefore, many living species will have only molecular data available.
+In practice, we removed all the morphological data from randomly chosen living with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.}
+\item{\hl{The proportion of missing data in the fossil record ($M_{F}$).}}
+\hl{This missing data parameter represents the completeness of the fossil record.
+In fact, due to preservation biases, missing data for fossil taxa are common }\citep{sansomfossilization2013}. 
+\hl{In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.}
+\item{\hl{The number of morphological characters for both living and fossil taxa ($N_{C}$).}}
+\hl{This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
 Any morphological matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
@@
 In practice, we randomly removed whole characters from the morphological matrix shrinking it to: 100, 90, 75, 50 or 25 characters.
-Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).
+Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).}
 \end{enumerate}
@@
 % TG: this part is nearly exactly the same as before.
-In practice, each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $N_C$ removes columns across both living and fossil taxa's data.
+In practice, each parameter represents a different way of removing data from the \hl{morphological part of} the matrix: $M_L$ removes \hl{entire} rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $N_C$ removes columns across both living and fossil taxa's data.
 Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
@@
 % TG: and this one is slightly modified only.
-We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data''matrices.
+\hl{We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data''matrices.
 Note that one of these combinations ($M_{L}$=0\%; $M_{F}$=0\% and $N_{C}$=100) has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
@@
 To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
-Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).
+Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).}
 

### 2015-07-29 Analysis/True_vs_best_tree.R

@@ -1,5 +1,4 @@
 #Script for testing the effect of missing data on tree topology.
-setwd('Analysis/')
 
 #Loading the functions
 sourceDir <- function(path, trace = TRUE, ...) {

### 2015-07-29 Manuscript/MPESubmission/Revision/response.tex
Starting tree effect (tmp - missing tables)

@@
 
-We added the following justification in the main text:
-"blalbalbal" (line @@@).
+We added the following sentence to the main text to clarify and link to the supplementary analysis:
+"In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices (i.e.
+diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
+The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." (line @@@).
 
-%To do (MAJOR)
-% Differences between True tree and best/missing-data trees.
-%"insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it" -> no, our parameters were not adequate.
-However this can be due to the size of the matrices.
-It is likely that phylo signal (ability to get close to the true tree) will increase with bigger matrices.
-Here we had to make a choice between larger matrices and less variance or smaller matrices and higher variances.
-We chose the second point because we were mainly interested in the effect of our parameters on recovering the topology.
-Because the true tree is just a starting tree for our simulations, it has not a lot of influence on the generation of the best tree.
-In other words, the best/true tree difference is more an assessment of the ability of the cascade of algorithms and softwares used (tree.bd{diversitree}, seqGen{phyclust}, rTraitDisc{ape}, RAxML, MrBayes) to perform with small size matrices.
+% TG: not sure how this justifications sounds...
 

### 2015-07-29 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Starting tree effect (tmp - missing tables)

@@
 In real world situations, the ``true'' tree is never available to us because we cannot know the true evolutionary history of a clade \citep[except in very rare circumstances, e.g.][]{rozen2005}.
+\hl{In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices (i.e.
+diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).}
+\hl{The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery.}
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the ``best'' tree, rather than the ``true'' tree, as the ``best'' tree is generally what we have to work with.

### 2015-07-29 Analysis/Starting_tree.R

@@ -1,5 +1,4 @@
 #Script for testing the effect of the starting tree in the Bayesian inference analysis.
-setwd('Analysis/')
 
 #Loading the functions
 sourceDir <- function(path, trace = TRUE, ...) {
@@ -45,17 +44,12 @@ True_to_true<-NTS(tmp, Random51)
 setwd('../../../../Analysis/')
 
 #########################
-#Load the CPU timer
-#########################
-timer<-read.csv("../Data/Starting_tree/timing.csv")
-paste("Using the true tree makes it ", round(sum(timer$random.tree, na.rm=TRUE)/sum(timer$true.tree, na.rm=TRUE), digit=3), " times faster.", sep="")
-
-
-#########################
 #Plotting the differences
 #########################
 
 #Plotting the results (RF)
+quartz(width = 8.3, height = 5.8) #A5 landscape
+op<-par(mfrow=c(1,2), bty="l", oma=c(0,0,0,0))
 TreeCmp.Plot(True_to_best, "R.F_Cluster", probs=c(95, 50), plot=TRUE, col=palette()[1], lines=TRUE, add=FALSE, shift=0, ylim=c(0,1), save.details=FALSE, xaxis='auto', ylab='Normalised Robinson-Foulds metric', xlab=NULL)
 TreeCmp.Plot(Rand_to_best, "R.F_Cluster", probs=c(95, 50), plot=TRUE, col=palette()[2], lines=TRUE, add=TRUE, shift=0.1, ylim=c(0,1), save.details=FALSE, xaxis='auto', ylab='Normalised Robinson-Foulds metric', xlab=NULL)
 TreeCmp.Plot(True_to_true, "R.F_Cluster", probs=c(95, 50), plot=TRUE, col=palette()[3], lines=TRUE, add=TRUE, shift=0.2, ylim=c(0,1), save.details=FALSE, xaxis='auto', ylab='Normalised Robinson-Foulds metric', xlab=NULL)
@@ -68,6 +62,7 @@ TreeCmp.Plot(Rand_to_best, "Triples", probs=c(95, 50), plot=TRUE, col=palette()[
 TreeCmp.Plot(True_to_true, "Triples", probs=c(95, 50), plot=TRUE, col=palette()[3], lines=TRUE, add=TRUE, shift=0.2, ylim=c(-0.5,1), save.details=FALSE, xaxis='auto', ylab='Normalised Triplets metric', xlab=NULL)
[…]

### 2015-07-30 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
Standardised Bhattacharyya coefficient across whole work
Added highlights + morph rate justification
Starting tree effect (tmp - missing tables)
Added starting tree results

--- 2015-07-08 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
@@
 \renewcommand\thefigure{A.\arabic{figure}}
+\renewcommand\thetable{A.\arabic{table}}
 
@@
 
+\subsection{Effect of the starting tree in Bayesian inference}
+We used the ``true'' tree as a starting tree for the two MCMC chains in the Bayesian inference for decreasing the computing time.
+To assess if this had an effect on the topology of the ``best'' tree, we ran a sub-sample of trees using a different random tree for the two MCMC chains \citep[default MrBayes option;][]{Ronquist2012mrbayes}.
+We tested this effect on five trees with the five levels of missing data (i.e.
+first tree: $M_L$=0\%, $M_F$=0\% and $N_C$=100 (i.e.
+0\% ``missing''); second tree: $M_L$=10\%, $M_F$=10\% and $N_C$=90, etc.) on the first 20 simulation chains.
+We then compared the trees inferred using a random starting tree to the ``best'' tree using the normalised Robinson-Foulds and Triplets metrics in an identical way as described in the main text (figure ~\ref{Fig_startingTree}).
+We then compared the normalised Robinson-Foulds and Triplets metrics of the ``missing-data'' trees using the ``true'' tree as a starting tree or a random tree as a starting tree using simple linear regression and nested linear regression (\texttt{lm(metric $\sim$ starting tree)}) and \texttt{lmList(metric $\sim$ starting tree | missing data}) in \texttt{R}).
+The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Starting_tree.R}.
+We found no significant effect of using the ``true'' or a random tree as a starting tree on the ability to recover the ``best'' tree (tables ~\ref{Tab_lm_res} and ~\ref{Tab_lmList_res}).
+
+\begin{table}[ht]
+\caption{Effect of the using the ``true'' tree as a starting tree on topological recovery for the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics using simple linear regression.}
+\label{Tab_lm_res}
+\centering
+\begin{tabular}{rlcccc}
+  \hline
+ & & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
+  \hline
+RF & (Intercept) & 0.8100 & 0.0122 & 66.16 & \textbf{0.0000} \\ 
+   & starting:true & 0.0019 & 0.0173 & 0.11 & 0.9125 \\
+Tr & (Intercept) & 0.4424 & 0.0418 & 10.59 & \textbf{0.0000} \\ 
+   & starting:true & 0.0159 & 0.0591 & 0.27 & 0.7887 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\caption{Effect of the using the ``true'' tree as a starting tree on topological recovery for the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics for each missing data parameter using nested linear regression.}
+\label{Tab_lmList_res}
+\centering
+\begin{tabular}{rccccc}
+  \hline
+ & & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
+  \hline
+RF & 0\%  & 0.02 & 0.02 & 1.02 & 0.31 \\ 
+   & 10\% & -0.01 & 0.02 & -0.50 & 0.62 \\ 
+   & 25\% & 0.00 & 0.02 & 0.00 & 1.00 \\ 
+   & 50\% & 0.00 & 0.02 & 0.03 & 0.98 \\ 
+   & 75\% & -0.00 & 0.02 & -0.03 & 0.98 \\
+Tr & 0\%  & 0.09 & 0.08 & 1.06 & 0.29 \\ 
+   & 10\% & -0.00 & 0.08 & -0.05 & 0.96 \\ 
+   & 25\% & 0.00 & 0.08 & 0.01 & 0.99 \\ 
+   & 50\% & -0.00 & 0.08 & -0.04 & 0.97 \\ 
+   & 75\% & -0.00 & 0.08 & -0.00 & 1.00 \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics[keepaspectratio=true, width=\textwidth]{SupplementaryFigures/Starting_tree.pdf}
+\caption{Effect of using the ``true'' tree (black) or a random tree (red) as the starting tree for the Bayesian inference. The x axis, represent the levels of missing data ()}
+\label{Fig_startingTree}
+\end{figure}
+
 \subsection{Morphological rates estimations results}
@@
 \caption{Differences between the simulated rate of each morphological character ($\alpha$ parameter) and the Bayesian estimates for the trees with missing data.
-The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
-Positive values shows underestimation of the parameter and negative values shows overestimation.
+The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
+Positive values shows overestimation of the parameter and negative values shows underestimation.
 One unit of difference means that the estimated rate is twice as big/small than the true rate.
 The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the rates estimates.}
-\label{Fig_AppendixCharacters}
+\label{Fig_AppendixRates}
 \end{figure}

### 2015-07-30 Analysis/Starting_tree.R

@@ -50,18 +50,18 @@ setwd('../../../../Analysis/')
 #Plotting the results (RF)
 quartz(width = 8.3, height = 5.8) #A5 landscape
 op<-par(mfrow=c(1,2), bty="l", oma=c(0,0,0,0))
-TreeCmp.Plot(True_to_best, "R.F_Cluster", probs=c(95, 50), plot=TRUE, col=palette()[1], lines=TRUE, add=FALSE, shift=0, ylim=c(0,1), save.details=FALSE, xaxis='auto', ylab='Normalised Robinson-Foulds metric', xlab=NULL)
+TreeCmp.Plot(True_to_best, "R.F_Cluster", probs=c(95, 50), plot=TRUE, col=palette()[1], lines=TRUE, add=FALSE, shift=0, ylim=c(0,1), save.details=FALSE, xaxis='auto', ylab='Normalised Robinson-Foulds metric', xlab=c("0%","10%","25%","50%","75%"))
 TreeCmp.Plot(Rand_to_best, "R.F_Cluster", probs=c(95, 50), plot=TRUE, col=palette()[2], lines=TRUE, add=TRUE, shift=0.1, ylim=c(0,1), save.details=FALSE, xaxis='auto', ylab='Normalised Robinson-Foulds metric', xlab=NULL)
-TreeCmp.Plot(True_to_true, "R.F_Cluster", probs=c(95, 50), plot=TRUE, col=palette()[3], lines=TRUE, add=TRUE, shift=0.2, ylim=c(0,1), save.details=FALSE, xaxis='auto', ylab='Normalised Robinson-Foulds metric', xlab=NULL)
-TreeCmp.Plot(Rand_to_true, "R.F_Cluster", probs=c(95, 50), plot=TRUE, col=palette()[4], lines=TRUE, add=TRUE, shift=0.3, ylim=c(0,1), save.details=FALSE, xaxis='auto', ylab='Normalised Robinson-Foulds metric', xlab=NULL)
-legend(3.8,0.4, legend=c("True tree (start) vs best tree","Random tree (start) vs best tree", "True tree (start) vs true tree", "Random tree (start) vs true tree"), col=palette()[1:4], pch=19, cex=0.7)
[…]

### 2015-07-30 Manuscript/MPESubmission/Revision/response.tex
Revisions updates

@@
 We see no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
-We made these data available in the supplementary materials Appendix A @@@ and we also added a precision in the manuscript:
+We made these data available in the supplementary materials Appendix A (figure A.2 and tables A.1 and A.2) and we also added a precision in the manuscript:
 
-blabalbalbal. line @@@ 
-%To do (MAJOR)
+"We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
+Note that a starting is not a Bayesian prior on topology \textit{per se} and did not significantly affected topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line @@@
 

### 2015-07-30 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
Revisions updates

@@
 
-\subsection{Effect of the starting tree in Bayesian inference}
+\subsection{Effect of the starting tree on Bayesian inference}
 We used the ``true'' tree as a starting tree for the two MCMC chains in the Bayesian inference for decreasing the computing time.

### 2015-07-30 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Revisions updates

@@
 For each estimation, we used the ``true'' tree's topology as a starting tree (with a starting value for each branch length of one).
+\hl{We used a fixed starting tree rather than a random starting tree} \citep[\hl{default MrBayes;}][]{Ronquist2012mrbayes} \hl{to speed up our Bayesian inferences.
+Note that a starting is not a Bayesian prior on topology \textit{per se} and did not significantly affected topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').}
 We also used two priors on the molecular part of the matrix: an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5, and a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)); and one prior on the morphological part of the matrix (exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5).
@@
 \subsubsection{Robinson-Foulds distance}
-The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and twice the number of shared clades across two trees.
+The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and \ul{twice the number} of shared clades across two trees.
 The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} (see Appendix B for calculation details).
@@
 every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
-In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.
+\ul{In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.}
 Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxon and decrease the Normalised Triplets metric.

### 2015-07-30 Manuscript/MPESubmission/Revision/response.tex
Updated revisions

@@
 % TG: a a paragraph in the results about the wild-card fossils
-As mentioned in the previous comment (reviewer 1 comment 9), this is due to the fact that the Bayesian consensus tree will be less and less resolved when missing data increases.
+As mentioned in the previous comment (reviewer 1 specific comment 9), this is due to the fact that the Bayesian consensus tree will be less and less resolved when missing data increases.
 This has two related effects.
@@
 
-We fixed this graphical mistake as state above in reviewer 1 comment 13.
+We fixed this graphical mistake as state above in reviewer 1 specific comment 13.
 
@@
 
-%To do - Probably needs to find a better cite.
+We now refer to two more classical citations (Bremer and Struwe 1992 (American Journal of Botany) and Patterson et al.
+1993 (Annual Review of Ecology and Systematics)) on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to small sized matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters).
 
@@
 
-%To do
+We retracted the entire statement concerning homoplasy and matrix size ("Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters (Wright and Hillis 2014).")
 
@@
 
-We modified the sentence to:
+We modified the paragraph to:
 
-"Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
-It seems therefore more important to integrate a maximum of living taxa with morphological data as well as a maximum number of characters.
-Additionally, when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6)."
+"Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
+For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combinations of the other parameters ($M_F$ and $N_C$; Figure 6).
+In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) doesn't matter and and increase/decrease in both does not affect topology.
+A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
+This has important practical implications, especially on the best strategy to improve topology by collecting morphological characters (see below)." lines (@@@)
 
@@
 
-We do understand the reviewers concern and added an analysis in the Appendix A (@@@) where we tested the effect of using the ``true'' tree as a starting tree.
+As developed above (reviewer 1, specific comment 29), we do understand the reviewers concern and added an analysis in the Appendix A (figure A.2 and tables A.1 and A.2) where we tested the effect of using the ``true'' tree as a starting tree.
 We found that there is no significant effect between using the ``true'' tree or a random tree (MrBayes default) as a starting tree on topological recovery.
+
 We refer to this supplementary analysis in the main text as follows:
@@
 "Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A - @@@ for details on the analysis of the effect of the starting tree in Bayesian inference)."
-
-%To do (MAJOR)
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'') for details on the analysis of the effect of the starting tree in Bayesian inference)." lines @@@
 
@@
 2014)."
-%To do (rewrite)
 
@@
 
+We are thankful for this comment as, indeed, our suggestion was a bit to obvious and not really helpful. We modified the end of this paragraph as follows:
+
+"Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e.
+50\%) for accurately combining both living and fossil species in phylogenies.
+Doing so grants a good overlap of morphological characters between living and fossil taxa, allowing the fossil taxa to be positioned relatively to the living taxa based on their actual shared derived characters rather than on simply available data."
+
+%Original: Therefore we advise filling as many gaps in Total Evidence matrices as possible.
+Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.
+
 \item{\textcolor{blue}{\textbf{line 613-615:} rephrase - sentence is choppy with all the commas. }}
 
-We modified the sentence as follow: "Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa." (line @@@),
+We removed this sentence and replaced it by the two new sentences outlined above (reviewer 1 specific comment 50)
+%We modified the sentence as follow: "Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa." (line @@@), TG: actually we removed the sentence.
 
@@
 Ronquist et al., 2012a).
-Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed."
-%Check:
-% Study: method 
-%Wood 
-%Schrago
-%Pyron
-%Beck,
-%Slater
-%Ronquist
-%-Wittenberg.et.al 2015 - Org Divers Evol (TEM)
-%-Carrizo&Catalano 2015 - Cladistics (TEM)
-%-Djernæs et al 2015 MPE (TEM)
-%-Arcila et al 2015 - MPE (TEM) This study suggests that current implementations of tip dating may overestimate ages of divergence in calibrated phylogenies + cool figure 1
-
+Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.
+To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological data in the same phylogeny as long as both types of data sufficiently overlap.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes."
 %Original: Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the “best” tree topology than the Maximum Likelihood trees.
@@
 
-The Lonsdale cluster requires this specific line in the acknowledgements of any study using it's infrastructure.
+The Lonsdale cluster requires this specific line in the acknowledgements of any study using it's infrastructure (\url{http://www.tchpc.tcd.ie/resources/acknowledgementpolicy}).
+If requested, however, we can also give additional details on the cluster used in the methods section.
+The current version of the manuscript only contains information on the nodes clock speed (2.30GHz clock speed nodes) lines @@@.
 
@@
 We also fixed the Normalised Triplets metrics axis to range from -0.5 to 1.
-We added a horizontal doted line at Triplets=0 to make sure the reader take notes is this change of scale.
+We added a horizontal doted line at Triplets = 0 to make sure the reader take notes is this change of scale.
 
-%To do (writing)
+%how these results compare to other missing data studies
 %Missing data for up to 50\% of characters has been suggested to still have high accuracy in uncovering the true tree (Wiens 2003; Wiens 2005)
-%how these results compare to other missing data studies and what do they tell us about the performance of the inference methods with missing data?
+%In Wiens when missing data ~50\%, accuracy is ~70/60
+
+%what do they tell us about the performance of the inference methods with missing data?
+%Bayesian seems more conservative than Maximum Likelihood.
+%Better to be more conservative?
+
+% STILL TO DO
+
+% TG: maybe put that somewhere?
+% In this study we chose to differentiate both metrics (RF and Triplets) because they account for two important aspects of topology in macroevolution.
+% The RF metric is a proxy for clade conservation which is important for macroevolutionary or macroecological methods such as phylogenetic comparative methods (rather than the specific position of each taxa, the conservation and support of clades is essential).
+% For example, for a tree (((a;b);c)((d;e);f);, swapping the taxa a and c (the clade is conserved) as less impact on comparative methods than swapping a and d (the clades are less conserved).
+%TG: is this kind of right?
+For PGLS for example...
+% % Or, other example with the same tree but for palaebioography: if ((a;b);c); and ((d;e);f); have different geographical origin; then a swap a and d will have important implications.
+% On the other hand the Triplets metric might be more important from macroevolutionary methods such as dating.
+% For example on the same tree (((a;b);c)((d;e);f); if a is a younger taxa than c, the swap of a and c will have a bigger effect on the age of the clade.
 

### 2015-07-30 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Updated revisions

@@
 This can affect our ability to recover the ``best'' tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
-The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \citep{wagner2000}.
-The sizes of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
+The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \hl{(}\citealt{bremer1992phylogeny,patterson1993congruence}\hl{; see }\citealt{masters2002lack}\hl{ for a empirical example)}.
+\ul{The sizes of our data matrices were constrained by the performance of our protocol}: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
 Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
 Although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
-Therefore, our simulations reflect realistic parameters. Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{wrightbayesian2014}. 
+Therefore, our simulations reflect realistic parameters.
 Nonetheless, the use of probabilistic methods (i.e.
@@
 As expected, when combining the missing data parameters, our ability to recover the ``best'' tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) and lower when they have a larger proportion of missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) and lower when \hl{they have a larger proportion of missing data} (i.e.
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $N_{C}$ = 25; Figure 5).
 It is important, however, to notice that the effect of each parameter is not additive.
-Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
-It seems therefore more important to integrate a maximum of living taxa with morphological data as well as a maximum number of characters.
-Additionally, when both $M_{L}$ reach 50\% missing data and $N_{C}$ reach 50 characters, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6).
+\hl{Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
+For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combinations of the other parameters ($M_F$ and $N_C$; Figure 6).
+In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) doesn't matter and and increase/decrease in both does not affect topology.
+A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
+This has important practical implications, especially on the best strategy to improve topology by collecting morphological characters (see below).}
 
@@
 Variation in our ability to recover the ``best'' tree topology depends heavily on the tree inference method (Figure 4 and Figure 5).
-For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over parsimony methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
+For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over \hl{parsimony} methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
 This is, however, the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best'' tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A - @@@ for details on the analysis of the effect of the starting tree in Bayesian inference).
+\hl{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'') for details on the analysis of the effect of the starting tree in Bayesian inference).}
 As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
@@
 
-Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advice using the Bayesian consensus trees as a topological constraint for tree inferences using the Total-Evidence method such as tip-dating \citep[e.g.][; although it is possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution]{ronquista2012,Wood01032013,BEASTmaster}.
-Additionally, using the Bayesian consensus tree rather than the Maximum Likelihood can reduce the amount of ``false positive'' topologies.
+\hl{Another practical implication of our results regards the tree inference methods.
+Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advice using the Bayesian consensus trees as a topological constraint for tree inferences using the Total-Evidence method such as tip-dating} \citep[\hl{e.g.}][\hl{; although it is possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution}]{ronquista2012,Wood01032013,BEASTmaster}.
+\hl{Additionally, using the Bayesian consensus tree rather than the Maximum Likelihood can reduce the amount of ``false positive'' topologies.
 As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve nodes weakly supported due to missing data than the Maximum Likelihood tree that is more likely to wrongly resolve such nodes (i.e.
@@
 For example a sub-sample of posterior trees distributions can be used to asses macroecological questions while better taking into account topological uncertainty (e.g.
-\citealt{FritzTree} and \citealt{jetzthe2012} trees used in \citealt{healy2014}).
+}\citealt{FritzTree}\hl{ and }\citealt{jetzthe2012}\hl{ trees used in }\citealt{healy2014}\hl{).}
 
@@
 When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
-Therefore we advise filling as many gaps in Total Evidence matrices as possible.
-Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa.
+\hl{Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e.
+50\%) for accurately combining both living and fossil species in phylogenies.
+Doing so grants a good overlap of morphological characters between living and fossil taxa, allowing the fossil taxa to be positioned relatively to the living taxa based on their actual shared derived characters rather than on simply available data.}
 
-Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
-This has also been observed in empirical data \cite[e.g.][]{Arcila2015131} were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
-This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds \citep[e.g.][]{slaterphylogenetic2013}, taxonomic constraints \citep[e.g.][]{slaterphylogenetic2013,beckancient2014} or even by fixing the topology \cite[e.g.][]{ronquista2012}.
-Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.
+\hl{Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+This has also been observed in empirical data} \cite[\hl{e.g.}][]{Arcila2015131}\hl{ were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds }\citep[\hl{e.g.}][]{slaterphylogenetic2013}\hl{, taxonomic constraints }\citep[\hl{e.g.}][]{slaterphylogenetic2013,beckancient2014}\hl{ or even by fixing the topology }\cite[\hl{e.g.}][]{ronquista2012}.
+\hl{Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.}
 
-The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
+\hl{To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological data in the same phylogeny as long as both types of data sufficiently overlap.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.}
 
-% Do NOT remove this, even if you are not including acknowledgments.
 \section{Acknowledgments}
@@
 Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
-Thanks to the two anonymous reviewers for their useful and enthusiastic %or "encouraging"
+Thanks to the two anonymous reviewers for their useful and enthusiastic % TG: or "useful and encouraging"?
+I feel just useful does gives justice to them, especially to the work of reviewer 1...
 comments.
@@
 \noindent
-\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange).
+\textbf{Figure 4:} The effects of increasing missing data on topological recovery using \hl{Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)}.
 The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $N_{C}$) is shown on the x axis.

### 2015-07-30 Manuscript/MPESubmission/Revision/response.tex
Still one minor comment to fix

@@
 We see no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
-We made these data available in the supplementary materials Appendix A (figure A.2 and tables A.1 and A.2) and we also added a precision in the manuscript:
+We made these data available in the supplementary materials Appendix A (figure A2 and tables A1 and A2) and we also added a precision in the manuscript:
 
@@
 
-As developed above (reviewer 1, specific comment 29), we do understand the reviewers concern and added an analysis in the Appendix A (figure A.2 and tables A.1 and A.2) where we tested the effect of using the ``true'' tree as a starting tree.
+As developed above (reviewer 1, specific comment 29), we do understand the reviewers concern and added an analysis in the Appendix A (figure A2 and tables A1 and A2) where we tested the effect of using the ``true'' tree as a starting tree.
 We found that there is no significant effect between using the ``true'' tree or a random tree (MrBayes default) as a starting tree on topological recovery.
@@
 Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A)." lines(@@)
-%To do (writing)
 
@@
 I made to extra summarizing tables in the supplementaries that may be actually easier to read than most of the others so we might want to add them to the main text.
+We added two extra tables in the Appendix C (Table C8 and C9) that summarized the median Bhattacharyya Coefficients within each methods and each parameters individually and combined (Table C8; as illustrated in Figure 2) as well as the median Bhattacharyya Coefficients within each methods and each parameters individually and combined (Table C9; as illustrated in Figure 3).
+% TG: If that's actually what the reviewer 1 wanted, I think we can propose to add it to the main text rather than the appendices.
+They can always decide to move it back to the appendices if it doesn't bring much.
 
@@
 
-%Working on it.
+% TG: Working on it.
+See https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Analysis_Demo.R .
+I've finished 75% of it but because it's coding, I'll finish it one day when I have no more energy (or hangover).
 
@@
 
-However, we have added all the comparisons between the "true" tree and the "missing-data" trees in the supplementary.
+However, we have added all the comparisons between the "true" tree and the "missing-data" trees in the Appendix C (Figures C6 and C7).
 It is interesting to notice that the algorithms and softwares as well as probably the size of our matrices are actually not ideal an the recovery of the "true" tree is rather low and scales down the our results (see figure @).

### 2015-07-30 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Still one minor comment to fix

@@
 Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
-Because of the difficulties to represent so many pairwise comparisons in a meaningful way, we summarized as a heat map of Bhattacharyya Coefficients (see Figure 6).
-In this type of figures, parameters that have similar or different effects on recovering the ``best'' topology (rather good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6).
+\hl{Because of the difficulties to represent so many pairwise comparisons in a meaningful way, we summarized as a heat map of Bhattacharyya Coefficients (see Figure 6).
+In this type of figures, parameters that have similar or different effects on recovering the ``best'' topology (rather good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6).}
 
@@
 
-Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+\hl{Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
@@
 Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
+Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).}
 
@@
 \noindent
-\textbf{Figure 6:} The effects of missing data on topological recovery using Bayesian consensus trees.
+\textbf{Figure 6:} \hl{The effects of missing data on topological recovery using Bayesian consensus trees.
 Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
@@
 Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a weak effect on the Normalised metric and thus the topological recovery).
+the parameters have a weak effect on the Normalised metric and thus the topological recovery)}.
 

### 2015-07-31 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
revision updates

@@
 Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions.
-To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
+\hl{To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $N_{C}$ = 100 etc.
-(see Figure 2 for more details).
+(see Figure 2 for more details).}
 This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).

### 2015-07-31 Manuscript/MPESubmission/Revision/response.tex
revision updates
updated response

@@
 
-%Add introduction (dear Dr blablalbal, here are our revisions)
+Dear Dr D\'{a}valos,
 
-%Resubmit the new version of the manuscript normaly but also add a pdf containing the track changes (highlithing in yellow the bit that have changed): "We are aware on the length of the manuscript.
-To facilitate the reviewers task, we attached to our revised manuscript, a pdf of the same manuscript but with the changes highlighted in yellow."
+MPE-15-254: Interim Decision
 
-%TG: Add a section "Additional 'fancy' changes" (not fancy obviously)
-% For clarity, we replaced the font of all the functions in R language in courier font (R, diversitree, etc...). \texttt{} - ACTUALLY DO THE CHANGE IF RELEVANT
-%We are very grateful to both referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper.
+We are very grateful to both referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper.
 We have taken all of their comments on board, and respond to their points below.
+For improving clarity in this document, we sorted the reviewers comments as follows:
+\begin{enumerate}
+\item \textbf{Reviewer 1:} containing reviewer's 1 major comments
+\item \textbf{Reviewer 1 Specific comments:} containing reviewer's 1 specific comments
+\item \textbf{Reviewer 2:} containing reviewer's 2 comments
+\end{enumerate}
+Also, we uploaded two versions of the revised manuscript. Both have the exact same content but one has all the changes and additions to the main text highlighted in yellow.
+
+% TG: actually, is it possible to send them the response as a pdf or will I have to copy paste all that in a mail?
+If the second, it's not that a big problem (just annoying) but I think we should try sending the pdf, it's way clearer like that.
 
@@
 Similarly, in the discussion, the authors need to point out how their results fit in with the results of previous research more clearly. }}
+
+% The novelty of this study should be spelled out more clearly by more a more thorough review of the literature, not just by referring to a few relevant works, but by pointing out what is so new about this study.
+
+% The authors assert that this is the first study to examine the effects of missing morphological data from a total evidence dataset, but that is what Pattinson et al.
+2014 did (as well as Wiens 2005).
+
+% This study will have bearing on an old assertion that including fossils specifically may help alleviate long-branch attraction, not only additional taxa with varying amounts of missing data, as suggested by others (e.g., Wiens 2005).
+
+% Similarly, in the discussion, the authors need to point out how their results fit in with the results of previous research more clearly.
 
@@
 
-1. The Total Evidence methods allows to combine living and fossil data.
-2. This study test the effect of missing morphological data on topology.
-3. The number of living taxa with data as the most effect on recovering topology.
-4. Bayesian inference outperforms Maximum Likelihood for correct clade recovery.
-5. We suggest that the sampling effort should be focused on data for living taxa.
+\begin{enumerate}[(1)]
+\item The Total Evidence methods allows to combine living and fossil data.
+\item This study test the effect of missing morphological data on topology.
+\item The number of living taxa with data as the most effect on recovering topology.
+\item Bayesian inference outperforms Maximum Likelihood for correct clade recovery.
+\item We suggest that the sampling effort should be focused on data for living taxa.
+\end{enumerate}
 
@@
 
-% To do (see reviewer 1 specific comment 13)
+% Needs stronger emphasis on what specifically is unique about this study compared to what the many other studies on missing data have done and found.
+% To do (see reviewer 1 specific comment 13) - links to major comment one on novelty
+
+% Clearly outlining the questions in this study that are left open from previous research is important because some of the findings in this study suggest that there may be more error associated with missing data than previous research has suggested.
 
@@
 We do understand the reviewers concern are agree that ML and BI are two commonly used abbreviations.
-Nonetheless, we would rather prefer not abbreviate these terms: in fact, even though they might add some "smoothness" % TG: that's not the word
-to the sentences where Maximum Likelihood and/or Bayesian Inference are stated, we think they might add confusion or difficulties to the whole paper.
-Spelling out the abbreviation might be locally "heavy" (i.e.
-in the sentence) but at least are completely clear and does not ask the reader to look back for the actual meaning of such or such abbreviation.
+Nonetheless, we would rather prefer not abbreviate these terms: in fact, even though they might add some concision to the sentences where Maximum Likelihood and/or Bayesian Inference are stated, we think they might add confusion or difficulties to the whole paper.
+Spelling out the abbreviation might be locally less concise (i.e.
+in the sentence) but at least are completely clear and does not ask the reader to look back for the actual meaning of such or such abbreviation (Berlin 2013; Radiology; \url{http://pubs.rsna.org/doi/abs/10.1148/radiol.12121776}).
 However, we leave this decision to the editor if she judges it better to abbreviate those terms. % TG: or maybe this statement is useless? Of course we leave it to the editor...
@@
 We compared both the ``best'' tree and the ``missing-data'' trees to the ``true'' tree as suggested by both reviewer.
-We did not, however, include the results to the main text (available in appendix @@@).
-We justify this decision to both reviewers specific concerns (see reviewer 1 specific comment 25 and reviewer 2 comment 1).
+We did not, however, include the results to the main text (available in appendix B; figure B1 and appendix C; figures C6 and C7).
+We do agree with the point raised with both reviewers on comparing the true tree to the best/missing-data trees.
+As reviewer 1 states: "[this analysis can give] insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it.".
+However, even though we agree with the comments of both reviewer that knowing the difference between the true and best/missing-data trees is important, we do not think that it "adds an additional layer of complexity that can't be accounted for via simulation" (reviewer 2, comment 1).
+In this simulations, we had to make a choice between (1) simulating larger matrices but with less chains (i.e. less variance) or (2) smaller matrices with more variance.
+We deliberately made the second choice because our main question was "what is the effect of our parameters on recovering the topology".
+In fact, regarding our problematic, we maintain that comparing the best tree to the missing-data trees is the best approach to answer this specific question for two reasons:
+\begin{enumerate}
+\item{The true tree is always unknown:} as we state in the manuscript: "In real world situations, the ``true'' tree is never available" (line 242 in the first manuscript and @@@ in the revision).
+\item{The true tree is just the starting point of the simulations:} in fact, we use the true tree only as a starting point for our simulations.
+Once generated, the true tree is fed to a cascade of algorithms and softwares (i.e.
+seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes) described in section 2.1 and 2.3 in order to get the best tree and the missing-data trees.
+Comparing the latests to the true tree will answer to the problematic "what is the effect of our missing-data parameters and the cascade of algorithms and softwares on recovering the true topology?".
+\end{enumerate}
+Therefore we added the second justification to the main text (see below this response) and added these analysis (true vs.
+best/missing-data trees) to the supplementary for potential readers that might be interested by this specific problematic.
+
+We added the following sentence to the main text to clarify and link to the supplementary analysis:
+"In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices (i.e.
+diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
+The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." (line @@@).
 
@@
 
-We dealt with this legitimate suggestion and ran an extra analysis on a subset of matrices (20 chains and only the trees with the same number of parameters; i.e.
-$M_L$=0\%, $M_F$=0\% and $M_C$=0\%; $M_L$=10\%, $M_F$=10\% and $M_C$=10\%, etc...) were we reran the Bayesian inference using a random tree rather than the ``true'' tree as a starting tree (default MrBayes procedure).
-We found no effect of the starting tree (random or ``true'') on the resulting topology.
-We added this analysis to the appendix @@@ and added a short note in the main text.
-See details at reviewer 1 specific comment 46 below.
+We dealt with this legitimate suggestion an extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains. 
+We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or using a random tree (default in MrBayes) as a starting tree.
+We found no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
+We made these data available in the supplementary materials Appendix A (figure A2 and tables A1 and A2) and we also added a precision in the manuscript:
+
+"We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
+Note that a starting is not a Bayesian prior on topology \textit{per se} and did not significantly affected topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line @@@
 
@@
 
-We clarified and unified % TG: or standardised?
-the interpretation of the Bhattacharyya Coefficient throughout the manuscript.
+We clarified the interpretation of the Bhattacharyya Coefficient throughout the manuscript.
 We advocate that two distributions are significantly different when their probability of overlapping (the Bhattacharyya Coefficient) is $>$0.05 and that they are significantly similar when this probability is $>$0.95.
@@
 We added the following statement to the main text and to the appendix B:
+
 "When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
@@
 \begin{enumerate}
-\item first, the way we assess topological recovery is slightly different in each case: for the Bayesian consensus trees, we simply compared the "missin-data" consensus tree to the "best" consensus tree; for the Bayesian posterior trees however, we randomly compared 1000 "missing-data" Bayesian posterior trees to 1000 "best" Bayesian posterior trees.
+\item first, the way we assess topological recovery is slightly different in each case: for the Bayesian consensus trees, we simply compared the "missing-data" consensus tree to the "best" consensus tree; for the Bayesian posterior trees however, we randomly compared 1000 "missing-data" Bayesian posterior trees to 1000 "best" Bayesian posterior trees.
 Therefore, even if the comparison methodology is the same, the fact that the second methods draws random trees, we expect the topological recovery score to be "blurred" in comparison to other comparisons that do not involve random comparisons.
@@
 \end{enumerate}
-Both points were discussed in the former manuscript lines 566 and 573 (or lines @@@ and @@@ in the revised version). %TG: maybe expand this bit then?
+Both points were discussed in the former manuscript lines 566 and 573 (or lines @@@ and @@@ in the revised version).
+We expanded the paragraph to:
+
+"Additionally, the Bayesian posterior trees performed way worth than the Bayesian consensus tree (figure 4, table 1 and appendix C figure C5 and tables C5, C6 and C7).
+This can be due to the fact that the Bayesian posterior trees are always resolved and thus more likely to contain wrongly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
+Conversely, the Bayesian consensus tree might not resolve nodes that are badly supported and thus more likely to contain only correctly resolved nodes (i.e.
+increasing the Normalised Robinson-Foulds metric)." lines (@@@)
+
+Concerning the last question of this comment: "How then in comparisons of the distributions are they always so low (Table 1 max prob.
+of overlapping distributions = 0.11)?" when looking at all parameters separately (figure 4), there is no clear overlap between the Bayesian consensus trees and the Bayesian posterior trees (median Bhattacharyya Coefficient of 0.01; table 1).
+The maximum observed Bhattacharyya Coefficient (0.11; table 1) is probably due to the combination of parameters: when all the parameters have many missing data, the probability of overlap between trees was slightly higher (0.11) but still really low.
+For the Normalised Triplets distance however, the overlap is higher (median Bhattacharyya Coefficient of 0.59; table 1) as illustrated in figure 4 (i.e. much more overlap between the distributions).
 
@@
 We modified the figures 4 and 5 to show Normalised Triplets metric values below 0 and added a dashed line at 0 to underline the difference in scale between both metrics.
-% TG: a a paragraph in the results about the wild-card fossils
-As mentioned in the previous comment (reviewer 1 specific comment 9), this is due to the fact that the Bayesian consensus tree will be less and less resolved when missing data increases.
+As mentioned above (reviewer 1 comment 9), this is due to the fact that the Bayesian consensus tree will be less and less resolved when missing data increases.
 This has two related effects.
@@
 In fact, when a clade ((a;b);c); is collapsed to (a;b;c), the whole clade is conserved (a, b and c are still in a common clade) however they also all become wild-card taxa: the position of a regarding b and c is now incorrect.
+
+We added a precision on this point in the...
+% TG: add a paragraph in the results about the wild-card fossils
+
+% TG: STOPPED HERE.
 
@@
 
-\section{Specific comments:}
+\section{Reviewer 1 Specific comments:}
 \begin{enumerate}
@@
 
-We do agree with the point raised with both reviewers on comparing the true tree to the best/missing-data trees.
-As reviewer 1 states: "[this analysis can give] insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it.".
-We've added the results of these comparisons to the supplementary materials (Appendix C - figures C5 and C6).
-However, even though we agree with the comments of both reviewer that knowing the difference between the true and best/missing-data trees is important, we do not think that it "adds an additional layer of complexity that can't be accounted for via simulation" (reviewer 2, comment 1).
-In this simulations, we had to make a choice between (1) simulating larger matrices but with less chains (i.e. less variance) or (2) smaller matrices with more variance.
-We deliberately made the second choice because our main question was "what is the effect of our parameters on recovering the topology".
-In fact, regarding our problematic, we maintain that comparing the best tree to the missing-data trees is the best approach to answer this specific question for two reasons:
-\begin{enumerate}
-\item{The true tree is always unknown:} as we state in the manuscript: "In real world situations, the ``true'' tree is never available" (line 242 in the first manuscript and @@@ in the revision).
-\item{The true tree is just the starting point of the simulations:} in fact, we use the true tree only as a starting point for our simulations.
-Once generated, the true tree is fed to a cascade of algorithms and softwares (i.e.
-seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes) described in section 2.1 and 2.3 in order to get the best tree and the missing-data trees.
-Comparing the latests to the true tree will answer to the problematic "what is the effect of our missing-data parameters and the cascade of algorithms and softwares on recovering the true topology?".
-\end{enumerate}
-Therefore we added the second justification to the main text (see below this response) and added these analysis (true vs.
-best/missing-data trees) to the supplementary for potential readers that might be interested by this specific problematic.
-
-We added the following sentence to the main text to clarify and link to the supplementary analysis:
-"In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices (i.e.
-diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
-The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." (line @@@).
-
-% TG: not sure how this justifications sounds...
+We dealt with this comment above (Reviewer 1, major comment 6) and below (Reviewer 2 comment 1).
 
@@
 
-We added an extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains.
-We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or using a random tree (default in MrBayes) as a starting tree.
-We see no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
-We made these data available in the supplementary materials Appendix A (figure A2 and tables A1 and A2) and we also added a precision in the manuscript:
-
-"We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
-Note that a starting is not a Bayesian prior on topology \textit{per se} and did not significantly affected topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line @@@
+We dealt with this comment above (see Reviewer 1 major comment 7).
 
@@
 We do understand the concerns of this reviewer as well but, again, we do not agree that comparing the "missing-data" trees to the "best" trees adds a layer of complexity to our simulations.
-In fact, we argue that doing so actually adds a layer of complexity to the analysis.
+In fact, we argue that comparing the "missing-data" trees to the "true" tree actually adds a layer of complexity to the analysis.
 The problematic we are tackling in this manuscript is to understand "what is the effect of our three parameters on recovering topology ?"
@@
 
-"Number of morphological characters ($N_{C}$)" line @@@ %Header in discussion
-
-"Additionally, when both $M_{L}$ reach 50\% missing data and $N_{C}$ reach 50 characters, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6)." line @@@
+"i.e. $N_{C}$ = 90" line (@@@)
 
@@
 In fact lines 584 to 589 in the first version of the manuscript (line @@@ in the revised version), we argue that even if the effect of the $M_F$ parameter is important, there is unfortunately not much we can do about it since the completeness of this part of the matrix solely depends on exceptional palaeontological discoveries and excavation effort.
-We feel that these parameters, however crucial for our understanding of macroevolution, are unfortunately not predicable and recommendations such as "we need more fossil data" are empty and less practical than "we need more living species data".
+We feel that these parameters, however crucial for our understanding of macroevolution, are unfortunately not predicable and a recommendation such as "we need more fossil data" is less practical than our main recommendation: "we need more living species data".
 

### 2015-07-31 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
updated response

@@
 \subsubsection{Robinson-Foulds distance}
-The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and \ul{twice the number} of shared clades across two trees.
+The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and \hl{twice the number} of shared clades across two trees.
 The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} (see Appendix B for calculation details).
@@
 Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
-When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
+\hl{When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
 When this coefficient is $>$0.95 both distributions are significantly similar.
-Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions.
+Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions.}
 \hl{To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
@@
 every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
-\ul{In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.}
+\hl{In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.}
 Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxon and decrease the Normalised Triplets metric.
@@
 We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
-Conversely, removing 10\% of morphological characters ($N_{C}$) genuinely removes 10\% of data in the matrix.
+Conversely, removing 10\% of morphological characters (\hl{i.e. $N_{C}$ = 90}) genuinely removes 10\% of data in the matrix.
 Nonetheless, the effect of removing characters on the ability to recover the ``best'' tree topology is of the same order of magnitude as for the other two parameters (Figure 4).
@@
 The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \hl{(}\citealt{bremer1992phylogeny,patterson1993congruence}\hl{; see }\citealt{masters2002lack}\hl{ for a empirical example)}.
-\ul{The sizes of our data matrices were constrained by the performance of our protocol}: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
+\hl{The sizes of our data matrices were constrained by the performance of our protocol}: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
 Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
@@
 
+The Bayesian consensus trees, however, performs badly for the Normalised Triplets metric: some parameters combinations, especially when the $M_F$ parameter reaches 75\% missing data, leads to negative values (figure 5).
+Normalised Triplets metric values below 0 means that the placement of some taxa is worth than expected by just randomly placing this taxa in the tree.
+This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees [CONTINUE LATER ON].
+
 It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best'' tree topology than Maximum Likelihood and Bayesian consensus trees.
-This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $N_{C}$ = 0\%; Figure 4).
+This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $N_{C}$ = 100; Figure 4).
 This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best'' trees involved 1000 random pairwise comparisons rather than just one.
-This will obviously add noise to the results for these methods.
+\hl{Additionally, the Bayesian posterior trees performed way worth than the Bayesian consensus tree (figure 4, table 1 and appendix C figure C5 and tables C5, C6 and C7).
+This can be due to the fact that the Bayesian posterior trees are always resolved and thus more likely to contain wrongly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
+Conversely, the Bayesian consensus tree might not resolve nodes that are badly supported and thus more likely to contain only correctly resolved nodes (i.e.
+increasing the Normalised Robinson-Foulds metric).}
 

### 2015-07-31 Manuscript/MPESubmission/Revision/response.tex
One more comment to deal with (last)

@@
 
-We added a precision on this point in the...
-% TG: add a paragraph in the results about the wild-card fossils
+We added a precision on this point in the about the effects of the tree inference method:
 
-% TG: STOPPED HERE.
+"The Bayesian consensus trees, however, performs badly for the Normalised Triplets metric: some parameters combinations, especially when the $M_F$ parameter reaches 75\% missing data, leads to negative values (figure 5).
+Normalised Triplets metric values below 0 means that the placement of some taxa is worth than expected by just randomly placing this taxa in the tree.
+This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
+In fact, even if clades are conserved (figure 5), the resolutions within them can be poor to non-existent when data is poor.
+In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters." lines (@@@)
 
@@
 
-We modified the conclusion as follows (see reviewer 1 specific comments 52)
+We also modified the conclusion to answer the specific comment 52 from reviewer 1 (see below).
+
+%what did those studies find
+
+%what are the questions those studies left open
+
+%and how does this study address them?
+
+% Previous (13)
+"Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
+using combined molecular and morphological data; e.g.
+Wiens et al., 2005; Pattinson et al., 2014).
+For example, Wiens et al.
+(2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
+Furthermore, Pattinson et al.
+(2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
+These studies, however, are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014)."
+
+"In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
+We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa."
+
+% What was left open (56)
+
+% Conclusion (52)
+Unfortunately, we didn't found many studies formally testing the important questions asked by the reviewer. However, there exist few empirical analysis that contain results supporting our claims.
+Firstly, for the question about the topological differences between Maximum Likelihood and the Bayesian consensus tree, in Arcila et al 2015 (MPE)'s supplementary analysis, the authors compare the topology inferred from a Total Evidence matrix in both Likelihood (RAxML) and Bayesian (MrBayes) and the topology of the Bayesian consensus tree is much better supported than the one of the best ML tree.
+Specifically, the ML tree has 7/28 nodes have a "fair" bootstrap support ($>$ 75) and 3/28 with a good support ($>$ 90) as the Bayesian consensus tree has 17/25 nodes with a "fair" posterior probability ($>$80) and 14/25 with a good posterior probability ($>$95).
+We therefore used this reference to strengthen our claim of the superiority of the Bayesian consensus tree over the Maximum Likelihood one.
+Secondly, for the question about molecular scaffolds, from the few studies clearly specifying their topological constraint method, only Beck and Lee 2013 (Proceedings B) seems to compare an analysis with and without taxonomic constraint on divergence time and rates estimation (the constraint seems to have no effect) but the branch length aspect is not covered in our study.
+
+We modified the mentioned section of the conclusion as follows:
+"Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+This has also been observed in empirical data (e.g. Arcila et al., 2015) were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds (e.g.
+Slater, 2013), taxonomic constraints (e.g.
+Slater, 2013; Beck and Lee, 2014) or even by fixing the topology (e.g.
+Ronquist et al., 2012a).
+Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.
+To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological data in the same phylogeny as long as both types of data sufficiently overlap.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes."
+%Original: Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the “best” tree topology than the Maximum Likelihood trees.
+Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
+The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
 
@@
 
-We fixed this graphical mistake as state above in reviewer 1 specific comment 13.
+We changed the axis of the graphs showing Normalised Triplets metric to show values below 0 throughout the manuscript and the appendices.
 
 \item{\textcolor{blue}{\textbf{Table caption:} Table caption should include a line about 'pooling across missing data schemes' to clarify what is presented.}}
+
+%TG: WUT? Didn't understood that one at all.
 

### 2015-07-31 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
One more comment to deal with (last)

@@
 
-The Bayesian consensus trees, however, performs badly for the Normalised Triplets metric: some parameters combinations, especially when the $M_F$ parameter reaches 75\% missing data, leads to negative values (figure 5).
+\hl{The Bayesian consensus trees, however, performs badly for the Normalised Triplets metric: some parameters combinations, especially when the $M_F$ parameter reaches 75\% missing data, leads to negative values (figure 5).
 Normalised Triplets metric values below 0 means that the placement of some taxa is worth than expected by just randomly placing this taxa in the tree.
-This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees [CONTINUE LATER ON].
+This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
+In fact, even if clades are conserved (figure 5), the resolutions within them can be poor to non-existent when data is poor.
+In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.}
 

### 2015-08-02 Manuscript/MPESubmission/Revision/response.tex
all comments dealt with (kind of)

@@
 This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
-In fact, even if clades are conserved (figure 5), the resolutions within them can be poor to non-existent when data is poor.
-In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters." lines (@@@)
+In fact, even if clades are conserved (figure 5), the resolutions within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
+In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
+These results are in agreement with previous studies that have showed that missing data can be deleterious for recovering ``correct'' topologies, especially for small matrices of 100 characters (Wiens 2003).
+It is important to note, however, that this effect can be reduced by just increasing the number of characters (Wiens 2003)." lines (@@@)
 
@@
 
-%how these results compare to other missing data studies
-%Missing data for up to 50\% of characters has been suggested to still have high accuracy in uncovering the true tree (Wiens 2003; Wiens 2005)
-%In Wiens when missing data ~50\%, accuracy is ~70/60
-
-%what do they tell us about the performance of the inference methods with missing data?
-%Bayesian seems more conservative than Maximum Likelihood.
-%Better to be more conservative?
-
-% STILL TO DO
-
-% TG: maybe put that somewhere?
-% In this study we chose to differentiate both metrics (RF and Triplets) because they account for two important aspects of topology in macroevolution.
-% The RF metric is a proxy for clade conservation which is important for macroevolutionary or macroecological methods such as phylogenetic comparative methods (rather than the specific position of each taxa, the conservation and support of clades is essential).
-% For example, for a tree (((a;b);c)((d;e);f);, swapping the taxa a and c (the clade is conserved) as less impact on comparative methods than swapping a and d (the clades are less conserved).
-%TG: is this kind of right?
-For PGLS for example...
-% % Or, other example with the same tree but for palaebioography: if ((a;b);c); and ((d;e);f); have different geographical origin; then a swap a and d will have important implications.
-% On the other hand the Triplets metric might be more important from macroevolutionary methods such as dating.
-% For example on the same tree (((a;b);c)((d;e);f); if a is a younger taxa than c, the swap of a and c will have a bigger effect on the age of the clade.
+We also linked these results to previous studies (Wiens 2003) and discussed the negative values of the Normalised Triplets metric in the reviewer 1 comment 10 (see above).
 
@@
 Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A)." lines(@@)
+Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
+This can be interpreted as, in figure 5 for the Normalised Robinson-Foulds metric, the overlap between the distributions once $M_L$=50\%." lines(@@)
 

### 2015-08-02 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
all comments dealt with (kind of)

@@
 Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).}
-
-%Adds that it correspond to the "line" of distributions in figure 5 (upper grey distributions)
+Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
+This can be interpreted as, in figure 5 for the Normalised Robinson-Foulds metric, the overlap between the distributions once $M_L$=50\%.}
 
@@
 These two methods, however, perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds metric (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Figure 4 and Figure C2 in Appendix C).
+
+%---------------------------------------------
+%
+%       Discussion
+%
+%---------------------------------------------
 
@@
 This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
-In fact, even if clades are conserved (figure 5), the resolutions within them can be poor to non-existent when data is poor.
-In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.}
+In fact, even if clades are conserved (figure 5), the resolutions within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
+In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
+These results are in agreement with previous studies that have showed that missing data can be deleterious for recovering ``correct'' topologies, especially for small matrices of 100 characters }\citep{wiensmissing2003}.
+\hl{It is important to note, however, that this effect can be reduced by just increasing the number of characters }\citep{wiensmissing2003}.
 

### 2015-08-02 Manuscript/MPESubmission/Revision/response.tex
Added comments number in latex file
Conclusion comment (11)

@@
 \begin{enumerate}
+% -----------------
+% Comment 1
+% -----------------
 \item{\textcolor{blue}{\textbf{Novelty:} The novelty of this study should be spelled out more clearly by more a more thorough review of the literature, not just by referring to a few relevant works, but by pointing out what is so new about this study.
@@
 
+% -----------------
+% Comment 2
+% -----------------
 \item{\textcolor{blue}{\textbf{Graphical abstract:} Current graphical abstract does not capture the content of the article for readers in a single glance.
@@
 
+% -----------------
+% Comment 3
+% -----------------
 \item{\textcolor{blue}{\textbf{Highlights:} wording seems awkward}}
@@
 
+% -----------------
+% Comment 4
+% -----------------
 \item{\textcolor{blue}{\textbf{Introduction:} Needs stronger emphasis on what specifically is unique about this study compared to what the many other studies on missing data have done and found.
@@
 
+% -----------------
+% Comment 5
+% -----------------
 \item{\textcolor{blue}{\textbf{Abbreviation:} One suggestion is to consider abbreviating Maximum Likelihood and Bayesian Inference as ML and BI respectively early in the methods and use throughout.}}
@@
 
+% -----------------
+% Comment 6
+% -----------------
 \item{\textcolor{blue}{\textbf{Methodology - True tree:} I suggest it warrants comparing at least the 'best' tree to the 'true' tree.}}
@@
 
+% -----------------
+% Comment 7
+% -----------------
 \item{\textcolor{blue}{\textbf{Methodology - Starting tree:} The authors should at least include some analysis in which no starting tree is given, the max.
@@
 
+% -----------------
+% Comment 8
+% -----------------
 \item{\textcolor{blue}{\textbf{Results - Bhattacharyya meaning:} Some descriptive statistics are given, but it would help to tell us if these differences are meaningful.
@@
 
+% -----------------
+% Comment 9
+% -----------------
 \item{\textcolor{blue}{\textbf{Results - Bayesian consensus vs.
@@
 
+% -----------------
+% Comment 10
+% -----------------
 \item{\textcolor{blue}{\textbf{Results - Bayesian Triplets:} Another important result that is not mentioned is that using the wildcard taxon metric (Tr), Bayesian consensus trees do quite poorly with missing data.
@@
 
+% -----------------
+% Comment 11
+% -----------------
 \item{\textcolor{blue}{\textbf{Conclusion:} The authors should clarify how their results compare to those of previous studies more directly to emphasize the nuances that make this study unique. }}
@@
 % Previous (13)
+
 "Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
@@
 Wiens et al., 2005; Pattinson et al., 2014).
-For example, Wiens et al.
-(2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
-Furthermore, Pattinson et al.
-(2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
+For example, Wiens et al. (2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
+Furthermore, Pattinson et al. (2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
 These studies, however, are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014)."
@@
 % Conclusion (52)
-Unfortunately, we didn't found many studies formally testing the important questions asked by the reviewer. However, there exist few empirical analysis that contain results supporting our claims.
-Firstly, for the question about the topological differences between Maximum Likelihood and the Bayesian consensus tree, in Arcila et al 2015 (MPE)'s supplementary analysis, the authors compare the topology inferred from a Total Evidence matrix in both Likelihood (RAxML) and Bayesian (MrBayes) and the topology of the Bayesian consensus tree is much better supported than the one of the best ML tree.
-Specifically, the ML tree has 7/28 nodes have a "fair" bootstrap support ($>$ 75) and 3/28 with a good support ($>$ 90) as the Bayesian consensus tree has 17/25 nodes with a "fair" posterior probability ($>$80) and 14/25 with a good posterior probability ($>$95).
-We therefore used this reference to strengthen our claim of the superiority of the Bayesian consensus tree over the Maximum Likelihood one.
-Secondly, for the question about molecular scaffolds, from the few studies clearly specifying their topological constraint method, only Beck and Lee 2013 (Proceedings B) seems to compare an analysis with and without taxonomic constraint on divergence time and rates estimation (the constraint seems to have no effect) but the branch length aspect is not covered in our study.
-
-We modified the mentioned section of the conclusion as follows:
-"Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+
+Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e.
+50\%) for accurately combining both living and fossil species in phylogenies.
+Doing so grants a good overlap of morphological characters between living and fossil taxa, allowing the fossil taxa to be positioned relatively to the living taxa based on their actual shared derived characters rather than on simply available data."
+
+Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
 This has also been observed in empirical data (e.g. Arcila et al., 2015) were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
@@
 
+See also reviewer 1 specific comment 52 for details on the choice of citing Arcial et al 2015.
+
+% -----------------
+% Comment 12
+% -----------------
 \item{\textcolor{blue}{\textbf{Figure 4:} Figure 4 has a color scheme issue, the colors in the caption do not match up to the figure.}}
@@
 
+% -----------------
+% Comment 13
+% -----------------
 \item{\textcolor{blue}{\textbf{Figure 4-5:} Am I wrong or shouldn't the y-axis for Normalised Triplets metric go below 0 - it looks those metrics for many missing data combinations have values below zero indicating trees are more different from the 'best' tree than expected by chance (Appendix B 1.1., see Appendix C Tables 1-3) and that is important to show.}}
@@
 
+% -----------------
+% Comment 14
+% -----------------
 \item{\textcolor{blue}{\textbf{Table caption:} Table caption should include a line about 'pooling across missing data schemes' to clarify what is presented.}}
@@
 
+% -----------------
+% Comment 15
+% -----------------
 \item{\textcolor{blue}{\textbf{Previous research:} The missing-data problem has a long history in the literature and there are references to some of those older and recent papers.
@@
 
+% -----------------
+% Specific Comment 1
+% -----------------
 \item{\textcolor{blue}{\textbf{line 31:} has it really been shown that the majority of macroevolutionary studies focuses on extant species only?
@@
 
+% -----------------
+% Specific Comment 2
+% -----------------
 \item{\textcolor{blue}{\textbf{line 37 and throughout:} Rephrase so that 'however' does not start the sentence, such as 'To do this, however, we need…'.
@@
 
+% -----------------
+% Specific Comment 3
+% -----------------
 \item{\textcolor{blue}{\textbf{line 43:} sentence choppy, consider rephrasing or splitting into two sentences.}}
@@
 
+% -----------------
+% Specific Comment 4
+% -----------------
 \item{\textcolor{blue}{\textbf{line 44:} three alternatives: age only, morph only, or age and morph jointly. }}
@@
 
+% -----------------
+% Specific Comment 5
+% -----------------
 \item{\textcolor{blue}{\textbf{line 47:} Simpson 1944 doesn't seem the appropriate reference for maximum parsimony. Consider (Felsenstein 2004; Hennig 1966) }}
@@
 
+% -----------------
+% Specific Comment 6
+% -----------------
 \item{\textcolor{blue}{\textbf{line 53:} omit 'are used as nodes rather than tips in these phylogenies and their'.
@@
 
+% -----------------
+% Specific Comment 7
+% -----------------
 \item{\textcolor{blue}{\textbf{line 66:} omit 'have been successfully applied to empirical data' and just say 'becoming increasingly popular…list of references'. }}
@@
 
+% -----------------
+% Specific Comment 8
+% -----------------
 \item{\textcolor{blue}{\textbf{line 76 and throughout the manuscript:} watch out for inconsistent parentheses.}}
@@
 
+% -----------------
+% Specific Comment 9
+% -----------------
 \item{\textcolor{blue}{\textbf{line 76 and throughout:} correct use of the word 'data' from singular to plural throughout}}
@@
 
+% -----------------
+% Specific Comment 10
+% -----------------
 \item{\textcolor{blue}{\textbf{line 77:} change 'a lot' to 'large proportions of' or something more accurate and less colloquial}}
@@
 
+% -----------------
+% Specific Comment 11
+% -----------------
 \item{\textcolor{blue}{\textbf{line 82:} change to '…missing data are not phylogenetically biased, Wiens…'}}
@@
 
+% -----------------
+% Specific Comment 12
+% -----------------
 \item{\textcolor{blue}{\textbf{line 99:} unclear - what is meant by 'with a fixed topology approach'?
@@
 
+% -----------------
+% Specific Comment 13
+% -----------------
 \item{\textcolor{blue}{\textbf{lines 80-103:} (1) the authors need to clearly justify the importance of this study compared the many studies that have investigated the effects of missing data on topology inference.
@@
 
+% -----------------
+% Specific Comment 14
+% -----------------
 \item{\textcolor{blue}{\textbf{line 124:} replace 'sections' with 'partition'}}
@@
 
+% -----------------
+% Specific Comment 15
+% -----------------
 \item{\textcolor{blue}{\textbf{line 125:} omit 'note that we explain each step in detail below this general outline', replace with something like 'Our general protocol included the following steps (Figure 1):'…}}
@@
 
+% -----------------
+% Specific Comment 16
+% -----------------
 \item{\textcolor{blue}{\textbf{line 167 and throughout manuscript:} replace usage of 'inferred' when referring to simulating data.
@@
 
+% -----------------
+% Specific Comment 17
+% -----------------
 \item{\textcolor{blue}{\textbf{line 173:} streamline to "The substitution rates were selected from a gamma distribution with…" or something similar. }}
@@
 
+% -----------------
+% Specific Comment 18
+% -----------------
 \item{\textcolor{blue}{\textbf{line 173 to 176:} need some justification for why the substitution rate with alpha shape of 0.5  will avoid homoplasy.
@@
 
+% -----------------
+% Specific Comment 19
+% -----------------
 \item{\textcolor{blue}{\textbf{Line 176 to 180:} could be streamlined.
@@
 
+% -----------------
+% Specific Comment 20
+% -----------------
 \item{\textcolor{blue}{\textbf{line 182:} give the ape function used}}
@@
 
+% -----------------
+% Specific Comment 21
+% -----------------
 \item{\textcolor{blue}{\textbf{line 185 to 188:} streamline, e.g.
@@
 
+% -----------------
+% Specific Comment 22
+% -----------------
 \item{\textcolor{blue}{\textbf{line 193 to 196:} again, justify why gamma with shape of 0.5 is appropriate for minimizing homoplasy.
@@
 
+% -----------------
+% Specific Comment 23
+% -----------------
 \item{\textcolor{blue}{\textbf{line 201:} change 'get' for 'obtain' or something less colloquial.}}
@@
 
+% -----------------
+% Specific Comment 24
+% -----------------
 \item{\textcolor{blue}{\textbf{line 231-235:} streamline, e.g., 'The minimal dataset included in this study was 5\% of the morphological matrix for any taxon'.}}
@@
 
+% -----------------
+% Specific Comment 25
+% -----------------
 \item{\textcolor{blue}{\textbf{line 237-246:} I understand the rational presented by the authors for comparing missing data trees to the best tree inferred from max.
@@
 
+% -----------------
+% Specific Comment 26
+% -----------------
 \item{\textcolor{blue}{\textbf{line 249:} condense citations to one parenthetical statement, 'GTR model (Tavare, 1986, default setting in RAx…)'.
@@
 
+% -----------------
+% Specific Comment 27
+% -----------------
 \item{\textcolor{blue}{\textbf{line 251:} omit 'implemented'}}
@@
 
+% -----------------
+% Specific Comment 28
+% -----------------
 \item{\textcolor{blue}{\textbf{line 273:} 'Bayesian Inference'}}
@@
 
+% -----------------
+% Specific Comment 29
+% -----------------
 \item{\textcolor{blue}{\textbf{line 286:} Why was the true tree used as a starting tree?
@@
 
+% -----------------
+% Specific Comment 30
+% -----------------
 \item{\textcolor{blue}{\textbf{line 318:} replace 'the twice number' with 'twice the number'}}
@@
 
+% -----------------
+% Specific Comment 31
+% -----------------
 \item{\textcolor{blue}{\textbf{line 327:} 'This methods' omit 's'}}
@@
 
+% -----------------
+% Specific Comment 32
+% -----------------
 \item{\textcolor{blue}{\textbf{line 343:} 'this metric is sensitive' add 'is'}}
@@
 
+% -----------------
+% Specific Comment 33
+% -----------------
 \item{\textcolor{blue}{\textbf{line 424:} 'Figure ??' should be Figure 6, but then do not repeat Figure 6 in the parentheses. }}
@@
 
+% -----------------
+% Specific Comment 34
+% -----------------
 \item{\textcolor{blue}{\textbf{line 436:} 'effect'}}
@@
 
+% -----------------
+% Specific Comment 35
+% -----------------
 \item{\textcolor{blue}{\textbf{line 489:} Streamline to something like: '…when the fossil is present near the tips but affects the clade conservation less when fossils are near the root'. }}
@@
 
+% -----------------
+% Specific Comment 36
+% -----------------
 \item{\textcolor{blue}{\textbf{line 494:} 'taxon'}}
@@
 
+% -----------------
+% Specific Comment 37
+% -----------------
 \item{\textcolor{blue}{\textbf{line 515:} '…the size of the matrix' missing 'of'}}
@@
 
+% -----------------
+% Specific Comment 38
+% -----------------
 \item{\textcolor{blue}{\textbf{line 518 to 520:} I am not sure that (Wagner 2000) states that incongruence between molecular and morphological data is 'more important in small morphological matrices'.
@@
 
+% -----------------
+% Specific Comment 39
+% -----------------
 \item{\textcolor{blue}{\textbf{line 520:} Either change 'size' to plural or change 'were' to 'was', the former seems to flow better. }}
@@
 
+% -----------------
+% Specific Comment 40
+% -----------------
 \item{\textcolor{blue}{\textbf{line 528:} missing close parenthesis after Ni et al.}}
@@
 
+% -----------------
+% Specific Comment 41
+% -----------------
 \item{\textcolor{blue}{\textbf{line 532:} Wright and Hillis 2014 found that increasing the size of the dataset improves topological accuracy, and they do not argue that simply increasing the number of characters increases the amount of homoplasy.
@@
 
+% -----------------
+% Specific Comment 42
+% -----------------
 \item{\textcolor{blue}{\textbf{line 540:} replace 'a lot' with 'larger proportions of missing data', a lot seems too colloquial for this kind of article. }}
@@
 
+% -----------------
+% Specific Comment 43
+% -----------------
 \item{\textcolor{blue}{\textbf{lines 542-547:} These seem like the most important practical results and should be more strongly emphasized.
@@
 
+% -----------------
+% Specific Comment 44
+% -----------------
 \item{\textcolor{blue}{\textbf{line 552:} replace 'cladistic' with 'parsimony'. While one has come to be associated with the other in the literature, they are not synonymous. }}
@@
 
+% -----------------
+% Specific Comment 45
+% -----------------
 \item{\textcolor{blue}{\textbf{line 557-558:} '…than the "best" maximum likelihood tree'}}
@@
 
+% -----------------
+% Specific Comment 46
+% -----------------
 \item{\textcolor{blue}{\textbf{line 549-565:} again, I am left wondering how much the differences in performance of max.
@@
 
+% -----------------
+% Specific Comment 47
+% -----------------
 \item{\textcolor{blue}{\textbf{line 593:} 'MorphoBank' - the 'b' should be capitalized in keeping with the authors' usage (O'Leary and Kaufman 2011)}}
@@
 
+% -----------------
+% Specific Comment 48
+% -----------------
 \item{\textcolor{blue}{\textbf{line 599:} remove extra close parenthesis inside parenthetical statement}}
@@
 
+% -----------------
+% Specific Comment 49
+% -----------------
 \item{\textcolor{blue}{\textbf{line 595-606:} this paragraph seems unsatisfactory advice for practitioners.
@@
 
+% -----------------
+% Specific Comment 50
+% -----------------
 \item{\textcolor{blue}{\textbf{line 612-613:} too obvious and not especially helpful.
@@
 
-We are thankful for this comment as, indeed, our suggestion was a bit to obvious and not really helpful. We modified the end of this paragraph as follows:
-
-"Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e.
-50\%) for accurately combining both living and fossil species in phylogenies.
-Doing so grants a good overlap of morphological characters between living and fossil taxa, allowing the fossil taxa to be positioned relatively to the living taxa based on their actual shared derived characters rather than on simply available data."
-
-%Original: Therefore we advise filling as many gaps in Total Evidence matrices as possible.
-Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.
-
+We are thankful for this comment as, indeed, our suggestion was a bit to obvious and not really helpful. See reviewer 1 comment 11 for the changes to the conclusion.
+
+% -----------------
+% Specific Comment 51
+% -----------------
 \item{\textcolor{blue}{\textbf{line 613-615:} rephrase - sentence is choppy with all the commas. }}
@@
 
+% -----------------
+% Specific Comment 52
+% -----------------
 \item{\textcolor{blue}{\textbf{line 615-618:} It seems more important to me that the authors emphasize what implications it may have for previous and future studies that Bayesian analyses outperform max.
@@
 
-We modified the mentioned section of the conclusion as follows:
-"Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
-This has also been observed in empirical data (e.g. Arcila et al., 2015) were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
-This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds (e.g.
-Slater, 2013), taxonomic constraints (e.g.
-Slater, 2013; Beck and Lee, 2014) or even by fixing the topology (e.g.
-Ronquist et al., 2012a).
-Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.
-To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological data in the same phylogeny as long as both types of data sufficiently overlap.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes."
-%Original: Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the “best” tree topology than the Maximum Likelihood trees.
-Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
-The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
-
+See reviewer's comment 11 above for the effective changes in the revised manuscript
+
+% -----------------
+% Specific Comment 53
+% -----------------
 \item{\textcolor{blue}{\textbf{line 628:} use of computer cluster may be better placed in the methods, depending on the norm for MPE.}}
@@
 
+% -----------------
+% Specific Comment 54
+% -----------------
 \item{\textcolor{blue}{\textbf{Figure 3:} Do not repeat the aside to similarity with t-test again from Fig 2 and text.}}
@@
 
+% -----------------
+% Specific Comment 55
+% -----------------
 \item{\textcolor{blue}{\textbf{Figure 4:} color codes in legend do not match up to the colors in the PDF I received.
@@
 
+% -----------------
+% Specific Comment 56
+% -----------------
 \item{\textcolor{blue}{\textbf{Figure 5:} This figure best summarizes the results, and a version of this should be the Graphical Abstract, with legends for the color cods and spelling out the missing data parameters, and the Y axis could just be 'clade similarity index' or 'rouge taxon index' so that the reader doesn't need to know what the names of the metrics are to grasp the graphic.
@@
 
+% -----------------
+% Specific Comment 57
+% -----------------
 \item{\textcolor{blue}{\textbf{Figure 6:} I think this may be the most important figure because it is demonstrating the probability of overlap between the distributions of the best trees and the missing data trees.
@@
 
+% -----------------
+% Specific Comment 58
+% -----------------
 \item{\textcolor{blue}{\textbf{Table 1:} The wording of the caption is off - the second paragraph is awkward. 
@@
 
+% -----------------
+% Specific Comment 59
+% -----------------
 \item{\textcolor{blue}{\textbf{Appendix C:} Tables 1-3 repeats 'of the'. Should label the first two rows as 'Best maximum likelihood tree' and 'Best Bayesian inference tree'}}
@@
 
+% -----------------
+% Specific Comment 60
+% -----------------
 \item{\textcolor{blue}{\textbf{Appendix D:} I find the online code repository thorough, but the repository is loaded with many files, especially for graphing.
@@
 
+% -----------------
+% Comment 1
+% -----------------
 \begin{enumerate}
@@
 
+% -----------------
+% Comment 2
+% -----------------
 \item{\textcolor{blue}{\textbf{Better explanation of the missing data:} I found the explanation of the missing data strategy very difficult to follow.
@@
 
+% -----------------
+% Comment 3
+% -----------------
 \item{\textcolor{blue}{\textbf{$M_C$ is not missing data per se:} If Mc is just removing columns, then they are shrinking the total matrix.
@@
 
-"i.e. $N_{C}$ = 90" line (@@@)
-
+% -----------------
+% Comment 4
+% -----------------
 \item{\textcolor{blue}{\textbf{Specific variation of data per fossil:} More importantly, it doesn't seem like missing data varied among fossil taxa, unless I'm misreading.

### 2015-08-02 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Conclusion comment (11)

@@
 \section{Conclusions}
-Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
+\hl{Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data }\citep{Wiens01102005,pattinsonphylogeny2014}.
+\hl{This study proposes a more in depth approach where missing data is generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
+This allowed use to confirm previous results that missing data can be specially deleterious in small sized matrices }\citep{wiensmissing2003}\hl{ but also relieved a yet unexplored aspect: the crucial importance of coding morphological data for living species.}
+\hl{In fact, missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
 When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
-\hl{Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e.
+Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e.
 50\%) for accurately combining both living and fossil species in phylogenies.

### 2015-08-02 Manuscript/MPESubmission/Revision/response.tex
Revisions draft finished

@@
 
-% TG: actually, is it possible to send them the response as a pdf or will I have to copy paste all that in a mail?
+% TG: actually, is it possible to send them the response as a pdf?
+or will I have to copy paste all that in a mail?
 If the second, it's not that a big problem (just annoying) but I think we should try sending the pdf, it's way clearer like that.
@@
 
-% The novelty of this study should be spelled out more clearly by more a more thorough review of the literature, not just by referring to a few relevant works, but by pointing out what is so new about this study.
-
-% The authors assert that this is the first study to examine the effects of missing morphological data from a total evidence dataset, but that is what Pattinson et al.
-2014 did (as well as Wiens 2005).
-
-% This study will have bearing on an old assertion that including fossils specifically may help alleviate long-branch attraction, not only additional taxa with varying amounts of missing data, as suggested by others (e.g., Wiens 2005).
-
-% Similarly, in the discussion, the authors need to point out how their results fit in with the results of previous research more clearly.
+We thank the reviewer for this helpful comment and emphasized on the novelty of this paper through the introduction, the discussion and the conclusion.
+Please see specific changes outlined at reviewer 1 comment 4, 10 and 11 below.
 
@@
 
-% Needs stronger emphasis on what specifically is unique about this study compared to what the many other studies on missing data have done and found.
-% To do (see reviewer 1 specific comment 13) - links to major comment one on novelty
-
-% Clearly outlining the questions in this study that are left open from previous research is important because some of the findings in this study suggest that there may be more error associated with missing data than previous research has suggested.
+We emphasised on novelty throughout the introduction by more clearly stating the difference between the present study and the literature and more clearly outlining the question of this study.
+
+"Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
+using combined molecular and morphological data; e.g.
+Wiens et al., 2005; Pattinson et al., 2014).
+For example, Wiens et al. (2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
+Furthermore, Pattinson et al. (2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
+As for morphological and molecular data separately, these studies shows that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
+These studies, however, are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014).
+Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data)."
+
+"In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
+We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa."
+
+"We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods by using two different topological metrics, based on the Robinson-Foulds (Robinson and Foulds, 1981) and Triplets (Critchlow et al., 1996) distances.
+Thus, in the present study, we propose a new approach on the missing-data question focus on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological metrics."
 
@@
 
-We also modified the conclusion to answer the specific comment 52 from reviewer 1 (see below).
-
-%what did those studies find
-
-%what are the questions those studies left open
-
-%and how does this study address them?
-
-% Previous (13)
-
-"Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
-using combined molecular and morphological data; e.g.
-Wiens et al., 2005; Pattinson et al., 2014).
-For example, Wiens et al. (2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
-Furthermore, Pattinson et al. (2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
-These studies, however, are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014)."
-
-"In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
-We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa."
-
-% What was left open (56)
-
-% Conclusion (52)
-
+We rewrote the conclusion to take aboard this comment and the specific comments 13, 52 (see especially this comment for details on the choice of citing Arcial et al 2015.) and 56.
+
+"Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data (Wiens et al 2005; Pattinson et al 2014)
+This study proposes a more in depth approach where missing data is generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
+This allowed use to confirm previous results that missing data can be specially deleterious in small sized matrices (Wiens 2003) but also relieved a yet unexplored aspect: the crucial importance of coding morphological data for living species.
+In fact, missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
+When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
 Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e.
@@
 Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.
+
 To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological data in the same phylogeny as long as both types of data sufficiently overlap.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes."
-%Original: Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the “best” tree topology than the Maximum Likelihood trees.
-Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
-The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.
-
-See also reviewer 1 specific comment 52 for details on the choice of citing Arcial et al 2015.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes." lines (@@@)
 
@@
 
-% specifically state line by line where we dealt with which and which question.
+We emphasized more on the recent findings from the missing-data papers in the introduction, and discussion (see comment 4 and 10 above).
 
@@
 
-(1) - We thank the reviewer for this judicious comment and emphasized on both the novelty aspect of our study and on the questions we are actually testing as follow:
-
-"Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
-using combined molecular and morphological data; e.g.
-Wiens et al., 2005; Pattinson et al., 2014).
-For example, Wiens et al.
-(2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
-Furthermore, Pattinson et al.
-(2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
-These studies, however, are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014)."
-
-"In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
-We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa."
+(1) - We thank the reviewer for this judicious comment and emphasized on both the novelty aspect of our study and on the questions we are actually testing (see reviewer 1 comment 4 above).
 
@@
 This has important practical implications, especially on the best strategy to improve topology by collecting morphological characters (see below)." lines (@@@)
-
-%Original: The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and/or $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Figure 6).
 
@@
 
-We removed this sentence and replaced it by the two new sentences outlined above (reviewer 1 specific comment 50)
-%We modified the sentence as follow: "Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa." (line @@@), TG: actually we removed the sentence.
+We removed this sentence and replaced it by the two new sentences outlined above (reviewer 1 specific comment 50).
 
@@
 We replaced the $M_C$ (missing characters) parameter by $N_C$ (number of characters with five states: 100, 90, 75, 50 and 25) in the revised manuscript (see reviewer 2 comment 2 above), the appendices and the figures.
-We also slightly modified some explanations related to the change of missing characters to number of characters:
-
-"To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $N_{C}$ = 100 etc.
-(see Figure 2 for more details)." line @@@
+We also slightly modified some explanations related to the change of missing characters to number of characters throughout the manuscript.
 

### 2015-08-02 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Revisions draft finished

@@
 \hl{Furthermore, \protect\cite{pattinsonphylogeny2014} }\hl{assess the effect of removing both molecular and morphological data (``artificial extinction'') on the topology.}
+\hl{As for morphological and molecular data separately, these studies shows that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.}
 \hl{These studies, however, are limited by their empirical-only approach \protect\citep{Wiens01102005} }\hl{and the scheme of generating missing data \protect\citep{pattinsonphylogeny2014}.}
+\hl{Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data).}
 
@@
 the size of the matrix).
-We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods.
+\hl{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods by using two different topological metrics, based on the Robinson-Foulds }\citep{RF1981}\hl{ and Triplets }\citep{critchlowthe1996}\hl{ distances.
+Thus, in the present study, we propose a new approach on the missing-data question focus on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological metrics.}
 We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the ``best'' tree topology more so than minimizing the amount of missing data in the fossil record.

### 2015-08-02 Manuscript/MPESubmission/Revision/response.tex
Added line numbering

@@
 These studies, however, are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014).
-Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data)."
+Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data)." lines 85-98 @@@
 
 "In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
-We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa."
+We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa." lines 99-103 @@@
 
 "We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods by using two different topological metrics, based on the Robinson-Foulds (Robinson and Foulds, 1981) and Triplets (Critchlow et al., 1996) distances.
-Thus, in the present study, we propose a new approach on the missing-data question focus on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological metrics."
+Thus, in the present study, we propose a new approach on the missing-data question focus on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological metrics." lines 127-134 @@@
 
@@
 diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
-The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." (line @@@).
+The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." line 276-280 @@@.
+See also appendix B section "Differences between the ``true'' and the ``best'' trees."
 
@@
 "We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
-Note that a starting is not a Bayesian prior on topology \textit{per se} and did not significantly affected topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line @@@
+Note that a starting is not a Bayesian prior on topology \textit{per se} and did not significantly affected topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line 323-327 @@@.
+See also Appendix A section "Effect of the starting tree on Bayesian inference".
 
@@
 When this coefficient is $>$0.95 both distributions are significantly similar.
-Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions." line (@@@)
+Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions." line 419-424 @@@.
 
@@
 \end{enumerate}
-Both points were discussed in the former manuscript lines 566 and 573 (or lines @@@ and @@@ in the revised version).
+Both points were discussed in the former manuscript lines 566-573 (or lines 645-652 @@@ in the revised version).
 We expanded the paragraph to:
@@
 Conversely, the Bayesian consensus tree might not resolve nodes that are badly supported and thus more likely to contain only correctly resolved nodes (i.e.
-increasing the Normalised Robinson-Foulds metric)." lines (@@@)
+increasing the Normalised Robinson-Foulds metric)." lines 652-658.
+@@@
 
@@
 These results are in agreement with previous studies that have showed that missing data can be deleterious for recovering ``correct'' topologies, especially for small matrices of 100 characters (Wiens 2003).
-It is important to note, however, that this effect can be reduced by just increasing the number of characters (Wiens 2003)." lines (@@@)
+It is important to note, however, that this effect can be reduced by just increasing the number of characters (Wiens 2003)." lines 632-644. @@@
 
@@
 To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological data in the same phylogeny as long as both types of data sufficiently overlap.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes." lines (@@@)
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes." lines 602-634. @@@
 
@@
 
-We revised this statement as follow: "many large-scale macroevolutionary studies focus solely on living species" (line @@@).
+We revised this statement as follow: "many large-scale macroevolutionary studies focus solely on living species" (line 30-31 @@@).
 
@@
 We modified the sentence as follow: "These approaches differ mainly in how they treat the fossil taxa and it's data.
-One can use the fossil as tips or as nodes in the phylogeny and use it's age only, it's morphology only or it's age and morphology jointly." (line @@@)
+One can use the fossil as tips or as nodes in the phylogeny and use it's age only, it's morphology only or it's age and morphology jointly." (line 42-44@@@)
 
@@
 
-We modified the citation as suggested by reviewer 1 (line @@@).
+We modified the citation as suggested by reviewer 1 (line 47-48@@@).
 
@@
 
-We modified the following sentence as suggested: "Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies" (line @@@).
+We modified the following sentence as suggested: "Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies" (line 53-54@@@).
 
@@
 
-We modified the following sentence as suggested: "Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies (e.g. ...)" (line @@@).
+We modified the following sentence as suggested: "Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies (e.g. ...)" (line 64-66@@@).
 
@@
 
-We modified the following sentence as suggested: "Total Evidence matrices are likely to contain a large proportions of missing data" (line @@@).
+We modified the following sentence as suggested: "Total Evidence matrices are likely to contain a large proportions of missing data" (line 75-76@@@).
 
@@
 
-We modified the following sentence as suggested: "as long as enough data overlaps in each case, and missing data are not phylogenetically biased" (line @@@).
+We modified the following sentence as suggested: "as long as enough data overlaps in each case, and missing data are not phylogenetically biased" (line 79-80@@@).
 
@@
 
-We modified the previous statement to: "but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints" (line @@@).
+We modified the previous statement to: "but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints" (line 111-113@@@).
 
@@
 We linked to this analysis in the methods part as follows:
-"The detailed MrBayes parameters are available in Appendix A along with details on the $\alpha$ parameter estimation."
-
+
+"The detailed MrBayes parameters are available in Appendix A along with details on the $\alpha$ parameter estimation." lines 336-338 @@@
+See also Appendix A "Morphological rates estimations results".
 % TG: what do you think?
@@
 
-We modified the following sentence as suggested: "the morphological partition of Total Evidence matrices" (line @@@).
+We modified the following sentence as suggested: "the morphological partition of Total Evidence matrices" (line 141@@@).
 
@@
 
-We removed the part suggested by the reviewer as follow: "we used the following protocol (Figure 1):" (line @@@).
+We removed the part suggested by the reviewer as follow: "we used the following protocol (Figure 1):" (line 142@@@).
 
@@
 
-We replace 'inferred' by 'simulated' on lines @@@ and @@@ %LaTeX lines 192 'The molecular matrix was simulated' and 199 'We simulated the morphological matrix using the R'
+We replace 'inferred' by 'simulated' when referring to simulating data throughout the manuscript.
 
@@
 
-We modified the following sentence as suggested: "The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5." (line @@@).
+We modified the following sentence as suggested: "The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5." (line 188@@@).
 
@@
 "In practice, such a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal (Hassanin et al., 1998; Estoup et al., 2002).%\citep{Hassanin1998611,EstoupHomoplasy}.
-Also, we chose this low value to be consistent with our protocol for simulating morphological characters (see below)." (line @@@).
+Also, we chose this low value to be consistent with our protocol for simulating morphological characters (see below)." (line 189-192@@@).
 
@@
 Springer et.
-al., 2012) to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)."
+al., 2012) to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)." lines 192-198 @@@
 
@@
 We modified the sentence as follows: "We simulated the morphological matrix using the rTraitDisc function from the R package ape v.
-3.0-11 (Paradis et al., 2004) to generate a matrix of 100 character sites for 51 taxa."
+3.0-11 (Paradis et al., 2004) to generate a matrix of 100 character sites for 51 taxa." lines 200-202.
 
@@
 "We used low evolutionary rate parameters to be consistent with the molecular rates parameters and to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal (Wright and Hillis, 2014).
-Note that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increases variance in topological error but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004)."
+Note that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increases variance in topological error but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004)." lines 211-217 @@@
 
@@
 
-We modified the sentence as follows: "To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon." (line @@@).
+We modified the sentence as follows: "To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon." (line 264-267@@@).
 
@@
 
-We modified the following sentence as suggested: "In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root." (line @@@).
+We modified the following sentence as suggested: "In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root." (line 547-549@@@).
 
@@
 We now refer to two more classical citations (Bremer and Struwe 1992 (American Journal of Botany) and Patterson et al.
-1993 (Annual Review of Ecology and Systematics)) on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to small sized matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters).
+1993 (Annual Review of Ecology and Systematics)) on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to small sized matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters) lines 578-579@@@.
 
@@
 
-We modified the following sentence: "The sizes of our data matrices were constrained by the performance of our protocol" (line @@@).
+We modified the following sentence: "The sizes of our data matrices were constrained by the performance of our protocol" (line 579-580@@@).
 
@@
 
-We modified the sentence as suggested: "they have a larger proportion of missing data" (line @@@).
+We modified the sentence as suggested: "they have a larger proportion of missing data" (line 597@@@).
 
@@
 A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
-This has important practical implications, especially on the best strategy to improve topology by collecting morphological characters (see below)." lines (@@@)
+This has important practical implications, especially on the best strategy to improve topology by collecting morphological characters (see below)." lines 589-609@@@
 
@@
 
-We replaced 'cladistic' by 'parsimony' as suggested (line @@@).
+We replaced 'cladistic' by 'parsimony' as suggested (line 614@@@).
 
@@
 
-We modified the sentence as suggested : "than the ``best'' Maximum Likelihood tree" (line @@@).
+We modified the sentence as suggested : "than the ``best'' Maximum Likelihood tree". lines 619-620@@@).
 
@@
 "Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'') for details on the analysis of the effect of the starting tree in Bayesian inference)." lines @@@
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'') for details on the analysis of the effect of the starting tree in Bayesian inference)." lines 618-624@@@
 
@@
 2012 trees used in Healy et al.
-2014)."
+2014)." lines 680-700@@@
 
@@
 We changed the colors to match with the paper color scheme: "Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
+See figure 4 caption.
 
@@
 Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a weak effect on the Normalised metric and thus the topological recovery)." (lines @@@)
+the parameters have a weak effect on the Normalised metric and thus the topological recovery)." See figure 6 caption.
 
@@
 "Because of the difficulties to represent so many pairwise comparisons in a meaningful way, we summarized as a heat map of Bhattacharyya Coefficients (see Figure 6).
-In this type of figures, parameters that have similar or different effects on recovering the ``best'' topology (rather good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6)." (lines @@@)
+In this type of figures, parameters that have similar or different effects on recovering the ``best'' topology (rather good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6)." lines 434-439@@@
 
@@
 Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
-This can be interpreted as, in figure 5 for the Normalised Robinson-Foulds metric, the overlap between the distributions once $M_L$=50\%." lines(@@)
+This can be interpreted as, in figure 5 for the Normalised Robinson-Foulds metric, the overlap between the distributions once $M_L$=50\%." lines 480-497@@@
 
@@
 To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
-Note that the living taxa always had at least 90\% of data (the 1000 molecular characters)." (line @@@)
+Note that the living taxa always had at least 90\% of data (the 1000 molecular characters)." lines 222-267@@@
 
@@
 Also, a second argument for using this "simplistic" way to deal with the missing data in the fossil record follows our main recommendation from this study: the best way to improve the quality of Total Evidence trees is to have a maximum number of living species with available morphological data.
-In fact lines 584 to 589 in the first version of the manuscript (line @@@ in the revised version), we argue that even if the effect of the $M_F$ parameter is important, there is unfortunately not much we can do about it since the completeness of this part of the matrix solely depends on exceptional palaeontological discoveries and excavation effort.
+In fact lines 584 to 589 in the first version of the manuscript (line 669-674@@@ in the revised version), we argue that even if the effect of the $M_F$ parameter is important, there is unfortunately not much we can do about it since the completeness of this part of the matrix solely depends on exceptional palaeontological discoveries and excavation effort.
 We feel that these parameters, however crucial for our understanding of macroevolution, are unfortunately not predicable and a recommendation such as "we need more fossil data" is less practical than our main recommendation: "we need more living species data".

### 2015-08-02 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Added line numbering

@@
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately (\hl{as long as enough data overlaps in each case, and missing data are not phylogeneticaly biased} \citep{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data does not appear be a major problem in molecular and morphological matrices separately \citep[\hl{as long as enough data overlaps in each case, and missing data are not phylogeneticaly biased;}][]{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
@@
 \noindent
-\textbf{Table 1:} Bhattacharyya Coefficients of the pairwise method comparisons.
-Each comparisons corresponds to the comparison of the full (125) distributions between the ``best'' tree and the ``missing data'' trees for all the four methods (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) and the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
+\textbf{Table 1:} \hl{Bhattacharyya Coefficients of the pairwise method comparisons.}
+\hl{Each comparisons corresponds to the comparison of the full (125) distributions between the ``best'' tree and the ``missing data'' trees for all the four methods (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) and the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
 Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
 The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
-If two methods have a high probability of overlap, they have a similar ability to recover the ``correct" tree topology.\\
+If two methods have a high probability of overlap, they have a similar ability to recover the ``correct" tree topology.}\\
 

### 2015-08-04 Manuscript/MPESubmission/Revision/response.tex
Added appendix D R demo script

@@
 
-% TG: Working on it.
-See https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Analysis_Demo.R .
-I've finished 75% of it but because it's coding, I'll finish it one day when I have no more energy (or hangover).
+We updated the readme page of the github repository linking to the various R codes.
+We also now provided a version of the analysis in R ran thoughout the manuscript as a step by step tutorial for a subset of data.
+See \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Analysis_Demo.R} and Appendix D.
 

### 2015-08-04 Manuscript/MPESubmission/Revision/response.tex
added Manos cites + discussion sentence

@@
 using combined molecular and morphological data; e.g.
-Wiens et al., 2005; Pattinson et al., 2014).
+Wiens et al., 2005; Manos et al., 2007; Pattinson et al., 2014).
 For example, Wiens et al. (2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
@@
 As for morphological and molecular data separately, these studies shows that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
-These studies, however, are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Pattinson et al., 2014).
+These studies, however, are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Manos et al., 2007; Pattinson et al., 2014).
 Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data)." lines 85-98 @@@
@@
 We emphasized more on the recent findings from the missing-data papers in the introduction, and discussion (see comment 4 and 10 above).
+We also included that our results for the $M_F$ parameter for the Normalised Robinson-Foulds metric where in accordance with Manos et al. 2007 in the discussion about the $M_F$ parameter:
+
+"These results are in agreement with Manos et al. (2007) where as few as 16 characters where sufficient for correctly assigning ``artificial fossils'' to their ``correct'' clade." lines @@@
 

### 2015-08-04 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
added Manos cites + discussion sentence

@@
 \hl{Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework \protect\citep[\hl{i.e.
-using combined molecular and morphological data}; e.g.][]{Wiens01102005,pattinsonphylogeny2014}.}
+using combined molecular and morphological data}; e.g.][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.}
 \hl{For example, \protect\cite{Wiens01102005} }\hl{only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.}
@@
 \hl{As for morphological and molecular data separately, these studies shows that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.}
-\hl{These studies, however, are limited by their empirical-only approach \protect\citep{Wiens01102005} }\hl{and the scheme of generating missing data \protect\citep{pattinsonphylogeny2014}.}
+\hl{These studies, however, are limited by their empirical-only approach \protect\citep{Wiens01102005} }\hl{and the scheme of generating missing data \protect\citep{manosphylogeny2007,pattinsonphylogeny2014}.}
 \hl{Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data).}
@@
 Normalised Robinson-Foulds metric) is still relatively good (mode = 0.72) when the proportion of missing data is high ($M_{F}$ = 75\%).
+\hl{These results are in agreement with \cite{manosphylogeny2007} where as few as 16 characters where sufficient for correctly assigning ``artificial fossils'' to their ``correct'' clade.}
 

### 2015-08-04 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
error in highlighting (fixed)

@@
 Normalised Robinson-Foulds metric) is still relatively good (mode = 0.72) when the proportion of missing data is high ($M_{F}$ = 75\%).
-\hl{These results are in agreement with \cite{manosphylogeny2007} where as few as 16 characters where sufficient for correctly assigning ``artificial fossils'' to their ``correct'' clade.}
+\hl{These results are in agreement with }\cite{manosphylogeny2007}\hl{ where as few as 16 characters where sufficient for correctly assigning ``artificial fossils'' to their ``correct'' clade.}
 

### 2015-08-06 Manuscript/MPESubmission/Revision/response.tex

@@
 %Packages
+\setlength\parindent{0pt} %NC: To get rid of automatic indenting
 \usepackage{pdflscape}
@@
 
-MPE-15-254: Interim Decision
+RE: MPE-15-254
 
@@
 \begin{enumerate}
-\item \textbf{Reviewer 1:} containing reviewer's 1 major comments
-\item \textbf{Reviewer 1 Specific comments:} containing reviewer's 1 specific comments
-\item \textbf{Reviewer 2:} containing reviewer's 2 comments
+\item \textbf{Reviewer 1:} containing reviewer 1's major comments
+\item \textbf{Reviewer 1 specific comments:} containing reviewer 1's specific comments
+\item \textbf{Reviewer 2:} containing reviewer 2's comments
 \end{enumerate}
-Also, we uploaded two versions of the revised manuscript. Both have the exact same content but one has all the changes and additions to the main text highlighted in yellow.
+Also, we uploaded two versions of the revised manuscript. Both have the exact same content but one has all the changes and additions to the main text highlighted in yellow to help with reviewing.
 
@@
 If the second, it's not that a big problem (just annoying) but I think we should try sending the pdf, it's way clearer like that.
+% NC: You can usually upload a file for the cover letter/response so pdf should be fine.
 
@@
 
-We thank the reviewer for this helpful comment and emphasized on the novelty of this paper through the introduction, the discussion and the conclusion.
-Please see specific changes outlined at reviewer 1 comment 4, 10 and 11 below.
+We thank the reviewer for this helpful comment and emphasize the novelty of this paper through the introduction, the discussion and the conclusion.
+Please see specific changes outlined at reviewer 1 comments 4, 10 and 11 below.
 
@@
 
-We gladly followed reviewer 1's graphical abstract suggestions and have changed it to be similar to the figure 5.
+We gladly followed reviewer 1's graphical abstract suggestions and have changed it to be similar to Figure 5.
 % TG: It's freaking not catchy at all but at this point I give up.
@@
 But it won't be sexy.
+% NC: Agreed! :)
 
@@
 
-We rephrased the highlights of the paper as follows: %3 to 5 bullet points (maximum 85 characters, including spaces, per bullet point). TG: probably need improvement here.
+We rephrased the highlights as follows: %3 to 5 bullet points (maximum 85 characters, including spaces, per bullet point). TG: probably need improvement here.
+% NC: Yeah I never even looked at these first time round. Mine aren't great but do fit in 85 characters!
 
 \begin{enumerate}[(1)]
-\item The Total Evidence methods allows to combine living and fossil data.
-\item This study test the effect of missing morphological data on topology.
-\item The number of living taxa with data as the most effect on recovering topology.
+\item The Total Evidence method places living and fossil taxa on the same phylogeny.
+\item Here we test the effect of missing morphological data on Total Evidence topologies.
+\item The number of living taxa with morphological data has greatest effects on topology.
 \item Bayesian inference outperforms Maximum Likelihood for correct clade recovery.
-\item We suggest that the sampling effort should be focused on data for living taxa.
+\item We recommend increased sampling effort for morphological data of living taxa.
 \end{enumerate}
@@
 
-We emphasised on novelty throughout the introduction by more clearly stating the difference between the present study and the literature and more clearly outlining the question of this study.
-
-"Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
-using combined molecular and morphological data; e.g.
+We have increased our emphasis on novelty throughout the introduction by more clearly stating the difference between the present study and the literature and more clearly outlining the question of this study, as follows:
+
+% NC: Don't forget to change stuff in the manuscript if I change it here and vice versa.
+
+"Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
+using both molecular and morphological data; e.g.
 Wiens et al., 2005; Manos et al., 2007; Pattinson et al., 2014).
+% NC: Argh this is really confusing. It doesn't logically flow
 For example, Wiens et al. (2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
 Furthermore, Pattinson et al. (2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
-As for morphological and molecular data separately, these studies shows that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
-These studies, however, are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Manos et al., 2007; Pattinson et al., 2014).
+For morphological and molecular data separately, both studies show that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
+However, they are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Manos et al., 2007; Pattinson et al., 2014).
 Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data)." lines 85-98 @@@
-
-"In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
+%%% NC: Paragraph above needs work aka a bit of logic to it.
+
+"In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living or fossil data, or all data for certain characters.
+% NC: All data overall suggests an empty matrix
 We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa." lines 99-103 @@@
 
-"We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods by using two different topological metrics, based on the Robinson-Foulds (Robinson and Foulds, 1981) and Triplets (Critchlow et al., 1996) distances.
-Thus, in the present study, we propose a new approach on the missing-data question focus on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological metrics." lines 127-134 @@@
+"We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred % NC: Didn't one reviewer have issues with the word inferred?
+using Maximum Likelihood and Bayesian methods by using two different topological distance metrics, based on the Robinson-Foulds (Robinson and Foulds, 1981) and Triplets (Critchlow et al., 1996) distances.
+Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological distance metrics." lines 127-134 @@@
+% NC: Why are the distinct topological metrics novel?
 
@@
 
-We do understand the reviewers concern are agree that ML and BI are two commonly used abbreviations.
-Nonetheless, we would rather prefer not abbreviate these terms: in fact, even though they might add some concision to the sentences where Maximum Likelihood and/or Bayesian Inference are stated, we think they might add confusion or difficulties to the whole paper.
-Spelling out the abbreviation might be locally less concise (i.e.
-in the sentence) but at least are completely clear and does not ask the reader to look back for the actual meaning of such or such abbreviation (Berlin 2013; Radiology; \url{http://pubs.rsna.org/doi/abs/10.1148/radiol.12121776}).
-However, we leave this decision to the editor if she judges it better to abbreviate those terms. % TG: or maybe this statement is useless? Of course we leave it to the editor...
+We understand the reviewer's concern.
+Nonetheless, we would prefer not abbreviate these terms.
+Although they might make sentences where Maximum Likelihood and/or Bayesian Inference are stated more concise, we think spelling out the terms makes our paper clearer and does not ask the reader to look back for the actual meaning of the abbreviation (Berlin 2013; Radiology; \url{http://pubs.rsna.org/doi/abs/10.1148/radiol.12121776}).
+However, we can change this at a later stage if necessary.
 
@@
 
-We compared both the ``best'' tree and the ``missing-data'' trees to the ``true'' tree as suggested by both reviewer.
-We did not, however, include the results to the main text (available in appendix B; figure B1 and appendix C; figures C6 and C7).
-We do agree with the point raised with both reviewers on comparing the true tree to the best/missing-data trees.
-As reviewer 1 states: "[this analysis can give] insights to whether the parameter settings used in the simulations and in the analyses were adequate for generating datasets and inferring the true tree, or something close to it.".
-However, even though we agree with the comments of both reviewer that knowing the difference between the true and best/missing-data trees is important, we do not think that it "adds an additional layer of complexity that can't be accounted for via simulation" (reviewer 2, comment 1).
-In this simulations, we had to make a choice between (1) simulating larger matrices but with less chains (i.e. less variance) or (2) smaller matrices with more variance.
-We deliberately made the second choice because our main question was "what is the effect of our parameters on recovering the topology".
-In fact, regarding our problematic, we maintain that comparing the best tree to the missing-data trees is the best approach to answer this specific question for two reasons:
+We have now compared both the ``best'' tree and the ``missing-data'' trees to the ``true'' tree as suggested by both reviewers.
+The results are available in appendix B; figure B1 and appendix C; figures C6 and C7).
+% NC: Worth describing the results here.
+
+Although though we agree with the comments of both reviewers that knowing the difference between the true and best/missing-data trees is important, we do not think that these results warrant inclusion in the main text.
+In these simulations, we had to make a choice between (1) simulating larger matrices but with fewer chains (i.e. less variance) or (2) smaller matrices with more variance.
+We deliberately made the second choice because our main question was "what is the effect of our parameters on recovering topology".
+To answer this question, we maintain that comparing the best tree to the missing-data trees is the best approach for two reasons:
 \begin{enumerate}
 \item{The true tree is always unknown:} as we state in the manuscript: "In real world situations, the ``true'' tree is never available" (line 242 in the first manuscript and @@@ in the revision).
-\item{The true tree is just the starting point of the simulations:} in fact, we use the true tree only as a starting point for our simulations.
-Once generated, the true tree is fed to a cascade of algorithms and softwares (i.e.
-seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes) described in section 2.1 and 2.3 in order to get the best tree and the missing-data trees.
-Comparing the latests to the true tree will answer to the problematic "what is the effect of our missing-data parameters and the cascade of algorithms and softwares on recovering the true topology?".
+\item{The true tree is just the starting point of the simulations:} we use the true tree only as a starting point for our simulations.
+Once generated, the true tree is fed to a cascade of algorithms and software (i.e.
+seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes) described in section 2.1 and 2.3 to get the best tree and the missing-data trees.
+Comparing the latter to the true tree will answer our question "what is the effect of our missing-data parameters and the cascade of algorithms and software on recovering the true topology?".
 \end{enumerate}
-Therefore we added the second justification to the main text (see below this response) and added these analysis (true vs.
-best/missing-data trees) to the supplementary for potential readers that might be interested by this specific problematic.
+Therefore we added a second justification to the main text (see below this response) and added these analysis (true vs.
+best/missing-data trees) to the supplementary for readers that might be interested by this specific problem.
 
 We added the following sentence to the main text to clarify and link to the supplementary analysis:
-"In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices (i.e.
-diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
+"In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices %NC: What dod you mean by R packages choices?
+Do you mean the way the trees were simulated?
+ (i.e. diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
 The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." line 276-280 @@@.
 See also appendix B section "Differences between the ``true'' and the ``best'' trees."
+% NC: Needs a little more explanation, and a line about what analysis you did in appendix b.
+So something along the lines of "We compare best trees to missing data trees, but could also compare true trees to the missing data trees.
+We do not do this because...
+However, results of analyses comparing the two can be found in appendix B.
+This makes little difference to our overall results." That might need to be better integrated with the paragraph(s) before.
 
@@
 
-We dealt with this legitimate suggestion an extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains. 
+We performed an extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains. 
 We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or using a random tree (default in MrBayes) as a starting tree.
+% NC: Yes but the reviewer is asking for something different - they ask you to have no starting tree. If this is impossible in MrBayes then say so, and say you used a random tree instead. 
 We found no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
-We made these data available in the supplementary materials Appendix A (figure A2 and tables A1 and A2) and we also added a precision in the manuscript:
+Therefore, we do not believe our results are merely the result of strong priors on the tree topology.
+% NC: The reviewer also says that saving time is not a good justification.
+It'd be worth explaining here how long it actually takes even though these datasets are small.
+These results are available in the supplementary materials Appendix A (figure A2 and tables A1 and A2) and we also added a note to the manuscript:
 
 "We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
-Note that a starting is not a Bayesian prior on topology \textit{per se} and did not significantly affected topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line 323-327 @@@.
+Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly % NC: Significantly as in statisically? if not change this word.
+affect topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line 323-327 @@@.
 See also Appendix A section "Effect of the starting tree on Bayesian inference".
@@
 We clarified the interpretation of the Bhattacharyya Coefficient throughout the manuscript.
-We advocate that two distributions are significantly different when their probability of overlapping (the Bhattacharyya Coefficient) is $>$0.05 and that they are significantly similar when this probability is $>$0.95.
+We say that two distributions are significantly different when their probability of overlapping (the Bhattacharyya Coefficient) is $>$0.05 and that they are significantly similar when this probability is $>$0.95.
 Even though other values (e.g.
 0.85) are not conclusive, we still reported them to show the relative overlap of the distributions (i.e.
-two distributions with a coefficient of 0.85 are not significantly similar but are closer to each other than two distributions with a coefficient of 0.15; which are themselves not significantly different but still less closer to each other than the previous ones).
+two distributions with a coefficient of 0.85 are not significantly similar but are closer to each other than two distributions with a coefficient of 0.15; which are themselves not significantly different but still less close to each other than the previous ones).
 We added the following statement to the main text and to the appendix B:
@@
 "When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
-When this coefficient is $>$0.95 both distributions are significantly similar.
-Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions." line 419-424 @@@.
+When this coefficient is $>$0.95, the distributions are significantly similar.
+Values in between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions." line 419-424 @@@.
+
+% NC: Make sure this is consistent throughout, add this to where you mention the BC being between 0 and 1 too?
 
@@
 The difference between the Bayesian consensus and the Bayesian posterior trees can be due to two factors:
-\begin{enumerate}
-\item first, the way we assess topological recovery is slightly different in each case: for the Bayesian consensus trees, we simply compared the "missing-data" consensus tree to the "best" consensus tree; for the Bayesian posterior trees however, we randomly compared 1000 "missing-data" Bayesian posterior trees to 1000 "best" Bayesian posterior trees.
-Therefore, even if the comparison methodology is the same, the fact that the second methods draws random trees, we expect the topological recovery score to be "blurred" in comparison to other comparisons that do not involve random comparisons.
-\item secondly, the Bayesian posterior trees are always resolved and thus, in the same way as the Maximum Likelihood tree, more likely to contain "false-positive" nodes.
+\begin{enumerate}[I]
+\item the way we assess topological recovery is slightly different in each case: for the Bayesian consensus trees, we simply compared the missing-data consensus tree to the best consensus tree; for the Bayesian posterior trees however, we randomly compared 1000 missing-data Bayesian posterior trees to 1000 best Bayesian posterior trees.
+Therefore, even if the comparison methodology is the same, the fact that the second methods draws random trees, we expect the topological recovery score to be blurred in comparison to other comparisons that do not involve random comparisons.
+\item the Bayesian posterior trees are always resolved and thus, in the same way as the Maximum Likelihood tree, more likely to contain false-positive nodes.
 On the other hand, because the consensus tree is a strict majority consensus rule, only nodes with a a good support will be represented in the tree and all other nodes will be collapsed.
-This reduces the amount of "false-positive" nodes (since an unresolved tree is always better than a wrongly resolved one) and increases clade conservation but also increases the placement of wild-card taxa (since it is likely that some Triplets will just not exist in a poorly resolved consensus tree).
+This reduces the number of false-positive nodes (since an unresolved tree is always better than a wrongly resolved one) and increases clade conservation but also increases the displacement of wild-card taxa (since it is likely that some Triplets will just not exist in a poorly-resolved consensus tree).
 \end{enumerate}
-Both points were discussed in the former manuscript lines 566-573 (or lines 645-652 @@@ in the revised version).
+Both points were discussed in the original manuscript lines 566-573 (or lines 645-652 @@@ in the revised version).
 We expanded the paragraph to:
 
-"Additionally, the Bayesian posterior trees performed way worth than the Bayesian consensus tree (figure 4, table 1 and appendix C figure C5 and tables C5, C6 and C7).
-This can be due to the fact that the Bayesian posterior trees are always resolved and thus more likely to contain wrongly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
-Conversely, the Bayesian consensus tree might not resolve nodes that are badly supported and thus more likely to contain only correctly resolved nodes (i.e.
+"Additionally, the Bayesian posterior trees performed more poorly than the Bayesian consensus tree (figure 4, table 1 and appendix C figure C5 and tables C5, C6 and C7).
+This may be because the Bayesian posterior trees are always resolved and thus more likely to contain incorrectly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
+Conversely, the Bayesian consensus tree might not resolve nodes that are poorly supported and thus is more likely to contain only correctly resolved nodes (i.e.
 increasing the Normalised Robinson-Foulds metric)." lines 652-658.
@@
 
+% NC: It's more likely an issue with reading figure 4.
+Check your figure legends.
+I think they might be incorrect.
+Then the reviewer might be seeing overlap of the wrong two distributions...
+But if they do overlap all the way to 25% then it is weird that the highest BC is 0.11 isn't it?
+
 % -----------------
@@
 This has two related effects.
-The first one being to retain "good" clade conservation (i.e. avoiding "false-positive" nodes by collapsing unsupported ones).
-And the second one is that when these clades are collapsed, it reduces the amount of available triplets and thus increases the "wild-card taxa" score (Normalised Triplets metric).
+The first one being to retain "good" clade conservation (i.e. avoiding false-positive nodes by collapsing unsupported ones).
+And the second one is that when these clades are collapsed, it reduces the number of available triplets and thus increases the wild-card taxa score (Normalised Triplets metric).
 In fact, when a clade ((a;b);c); is collapsed to (a;b;c), the whole clade is conserved (a, b and c are still in a common clade) however they also all become wild-card taxa: the position of a regarding b and c is now incorrect.
 
-We added a precision on this point in the about the effects of the tree inference method:
-
-"The Bayesian consensus trees, however, performs badly for the Normalised Triplets metric: some parameters combinations, especially when the $M_F$ parameter reaches 75\% missing data, leads to negative values (figure 5).
-Normalised Triplets metric values below 0 means that the placement of some taxa is worth than expected by just randomly placing this taxa in the tree.
+We added a note on this in the section about the effects of the tree inference method:
+
+"The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (figure 5).
+A Normalised Triplets metric value below 0 means that the placement of some taxa is worse % NC: It's worse not worth!
+than expected by just randomly placing this taxon in the tree.
 This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
-In fact, even if clades are conserved (figure 5), the resolutions within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
+Even if clades are conserved (figure 5), the resolution within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
 In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
-These results are in agreement with previous studies that have showed that missing data can be deleterious for recovering ``correct'' topologies, especially for small matrices of 100 characters (Wiens 2003).
-It is important to note, however, that this effect can be reduced by just increasing the number of characters (Wiens 2003)." lines 632-644. @@@
+These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters (Wiens 2003).
+It is important to note, however, that this effect can be reduced by increasing the number of characters (Wiens 2003)." lines 632-644. @@@
 
@@
 
-We rewrote the conclusion to take aboard this comment and the specific comments 13, 52 (see especially this comment for details on the choice of citing Arcial et al 2015.) and 56.
-
-"Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data (Wiens et al 2005; Pattinson et al 2014)
+We rewrote the conclusion to deal with this comment and the specific comments 13, 52 (see comment 52 for details on the choice of citing Arcial et al 2015.) and 56.
+
+"Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data (Wiens et al 2005; Pattinson et al 2014).
+% NC: I still don't get these limitations.
+How are they are issue?
+What aspects?
 This study proposes a more in depth approach where missing data is generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
-This allowed use to confirm previous results that missing data can be specially deleterious in small sized matrices (Wiens 2003) but also relieved a yet unexplored aspect: the crucial importance of coding morphological data for living species.
-In fact, missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
+This allowed us to confirm previous results that missing data can be especially problematic in small matrices (Wiens 2003) but also revealed the crucial importance of coding morphological data for living species.
+Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
 When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
-Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e.
+Therefore we advise that one should focus on coding morphological characters for a large number of the living taxa present in the matrix (i.e.
 50\%) for accurately combining both living and fossil species in phylogenies.
-Doing so grants a good overlap of morphological characters between living and fossil taxa, allowing the fossil taxa to be positioned relatively to the living taxa based on their actual shared derived characters rather than on simply available data."
-
-Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
-This has also been observed in empirical data (e.g. Arcila et al., 2015) were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
-This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds (e.g.
+Doing so grants a good overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their actual shared derived characters rather than simply on available data."
+
+Additionally, the topology of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+This has also been observed in empirical data (e.g. Arcila et al., 2015) where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds (e.g.
 Slater, 2013), taxonomic constraints (e.g.
@@
 Ronquist et al., 2012a).
-Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.
+Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.
 
@@
 
-We apologies for this silly mistake and fixed the colors in the figure. We also changed the color scheme to match with the general color scheme of the paper (black, grey, orange and blue).
+We apologise for this silly mistake and fixed the colors in the figure. We also changed the color scheme to match with the general color scheme of the paper (black, grey, orange and blue).
 
@@
 
-We changed the axis of the graphs showing Normalised Triplets metric to show values below 0 throughout the manuscript and the appendices.
+We changed the axes of the graphs showing Normalised Triplets metric to show values below 0 throughout the manuscript and the appendices.
 
@@
 %TG: WUT? Didn't understood that one at all.
+% NC: The reviewer means that you're presenting results from across all your different parameters, and all your different amounts of missing data in the table.
+You need to clarify this, especially as in figures you separate these out.
+Pooling just means combining.
+I noticed this on my re-read just now too.
 
@@
 
-We emphasized more on the recent findings from the missing-data papers in the introduction, and discussion (see comment 4 and 10 above).
+We now emphasize more of the recent findings from the missing-data papers in the introduction, and discussion (see comment 4 and 10 above).
 We also included that our results for the $M_F$ parameter for the Normalised Robinson-Foulds metric where in accordance with Manos et al. 2007 in the discussion about the $M_F$ parameter:

### 2015-08-06 Manuscript/MPESubmission/Revision/response.tex
Edits up to reviwer 1 specific comments

@@
 
-We rewrote the conclusion to deal with this comment and the specific comments 13, 52 (see comment 52 for details on the choice of citing Arcial et al 2015.) and 56.
+We rewrote the conclusion to deal with this comment and the specific comments 13, 52 (see comment 52 for details on the choice of citing Arcial et al 2015.) 
+% NC: Is it Arcial or Arcila?
+and 56.
 
@@
 Additionally, the topology of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
-This has also been observed in empirical data (e.g. Arcila et al., 2015) where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This has also been observed in empirical data (e.g. Arcila et al., 2015) 
+% NC: Arcial or arcila?
+where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
 This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds (e.g.
@@
 
-We revised this statement as follow: "many large-scale macroevolutionary studies focus solely on living species" (line 30-31 @@@).
+We revised this statement as follows: "many large-scale macroevolutionary studies focus solely on living species" (line 30-31 @@@).
 
@@
 
-We modified the sentence as follow: "These approaches differ mainly in how they treat the fossil taxa and it's data.
-One can use the fossil as tips or as nodes in the phylogeny and use it's age only, it's morphology only or it's age and morphology jointly." (line 42-44@@@)
+We modified the sentence as follows: "These approaches differ mainly in how they treat fossil taxa and their data.
+One can use fossils as tips or as nodes in the phylogeny, and can use only the age of the fossils, only the morphology of the fossils, or age and morphology jointly." (line 42-44@@@)
 
@@
 
-We modified the following sentence as suggested: "Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies" (line 53-54@@@).
+We modified the following sentence as suggested: "Because fossil taxa do not usually have available DNA, only fossil occurrence dates are used to time calibrate phylogenies" (line 53-54@@@).
 
@@
 
-We modified the following sentence as suggested: "Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies (e.g. ...)" (line 64-66@@@).
+We modified the sentence as suggested.
+% NC: If it's just you doing exactly what they ask for there is no need to re-write the sentence.
 
@@
 
+% NC: Hmmm I reckon this is probably inconsistent throughout then as I usually correct it to plural. I'd just change this tbh. Or at least check you've been consistent.
+
 % -----------------
@@
 
-We modified the following sentence as suggested: "Total Evidence matrices are likely to contain a large proportions of missing data" (line 75-76@@@).
+We modified the sentence as suggested.
 
@@
 
-We modified the following sentence as suggested: "as long as enough data overlaps in each case, and missing data are not phylogenetically biased" (line 79-80@@@).
+We modified the sentence as suggested.
 
@@
 
-We modified the previous statement to: "but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints" (line 111-113@@@).
+We modified the statement to: "but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints" (line 111-113@@@).
 
@@
 
-(1) - We thank the reviewer for this judicious comment and emphasized on both the novelty aspect of our study and on the questions we are actually testing (see reviewer 1 comment 4 above).
+(1) - We thank the reviewer for this comment and now emphasize both the novel aspects of our study and the questions we are actually testing (see reviewer 1 comment 4 above).
 
 (2) - the second point underlined by the reviewer in this specific comment strikes use as really interesting.
-However, we considered this study to be already long and complex.
+However, we considered this study to already be long and complex.
 We therefore think that adding an extra question to this study (i.e.
-what is the effect of missing data on the morphological substitution rate parameter estimates?) might loose the reader.
-Nonetheless, we added the result of such analysis in the \textbf{Appendix-A Morphological rates estimations results} section.
-We linked to this analysis in the methods part as follows:
+what is the effect of missing data on the morphological substitution rate parameter estimates?) might lose the reader and make an already long paper even longer.
+Nonetheless, we added the result of these analyses to \textbf{Appendix-A Morphological rates estimations results}.
+We link to this analysis in the methods as follows:
 
@@
 
+% NC: I'm a bit confused - did you do what they wanted or not?
+If not then the whole "this will make the paper longer" argument holds.
+If you did do what they wanted then it might need a little more explanation somewhere.
+You can't just throw random stuff into the appendices for no reason.
+
 % -----------------
@@
 
-We modified the following sentence as suggested: "the morphological partition of Total Evidence matrices" (line 141@@@).
+We modified the sentence as suggested.
 
@@
 
-We removed the part suggested by the reviewer as follow: "we used the following protocol (Figure 1):" (line 142@@@).
+We removed the part suggested by the reviewer.
 
@@
 We replace 'inferred' by 'simulated' when referring to simulating data throughout the manuscript.
+% NC: Check this - I'm sure I caught one up above in the comments here.
 
@@
 
-We modified the following sentence as suggested: "The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5." (line 188@@@).
+We modified the sentence as suggested.
 
@@
 We changed the sentence to:
-"In practice, such a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal (Hassanin et al., 1998; Estoup et al., 2002).%\citep{Hassanin1998611,EstoupHomoplasy}.
+"In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal (Hassanin et al., 1998; Estoup et al., 2002).%\citep{Hassanin1998611,EstoupHomoplasy}.
 Also, we chose this low value to be consistent with our protocol for simulating morphological characters (see below)." (line 189-192@@@).
+
+% NC: I think what is needed here is a definition of what is "too much". Is it 50%? 20%? etc. Too much is arbitrary.
 
@@
 We modified the sentence as follows:
-"We selected this model and parameters settings to be a compromise between realism from empirical datasets (e.g.
+"We selected this model and these parameters to strike a balance between realism for empirical datasets (e.g.
 Kelly et al., 2014) and models with few parameters (cf.
@@
 
+% NC: Why not just use the phrasing given by the reviewer for the first part?
+
 % -----------------
@@
 
-We modified the sentence as follows: "We simulated the morphological matrix using the rTraitDisc function from the R package ape v.
-3.0-11 (Paradis et al., 2004) to generate a matrix of 100 character sites for 51 taxa." lines 200-202.
+We now state that we used the rTraitDisc function.
 
@@
 
-We modified the following sentence as suggested: "sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A and Fig.
-A1 within)." (line @@@).
+We modified the sentence as suggested.
 
@@
 We initially chose this parameter from personal communication with April Wright and to be consistent with the molecular rate parameter.
-We like to point out that, in figure 3 from Wright and Hillis 2014, even if the Topological error is high when rates are near 0, it is mainly the case for parsimony and both parsimony and Mk with missing data, the fitted model for the Mk model with no missing data (light purple line) is still at lowest when the rates are near 0.
-Topological error, however, becomes minimal for all scenari (Parsimony and Mk, with or without missing data) at a rate of 0.5 (see modified figure 3 from Wright and Hillis (2014) in this response; we added the tick marks and the 0.5 rate limit by using the ruler in Adobe Illustrator).
+In figure 3 from Wright and Hillis 2014, even if the topological error is high when rates are near 0, it is mainly the case for parsimony and both parsimony and Mk with missing data, the fitted model for the Mk model with no missing data (light purple line) is still lowest when the rates are near 0.
+Topological error, however, becomes minimal for all scenarios (Parsimony and Mk, with or without missing data) at a rate of 0.5 (see modified figure 3 from Wright and Hillis (2014) in this response; we added the tick marks and the 0.5 rate limit by using the ruler in Adobe Illustrator).
 
@@
 
-"We used low evolutionary rate parameters to be consistent with the molecular rates parameters and to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal (Wright and Hillis, 2014).
-Note that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increases variance in topological error but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004)." lines 211-217 @@@
+"We used low evolutionary rate parameters to be consistent with the molecular rate parameters and to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal (Wright and Hillis, 2014).
+Note that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increase variance in topological error but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004)." lines 211-217 @@@
+
+% NC: Might need a bit more explanation.
+For example mentioning that "Topological error, however, becomes minimal for all scenarios (Parsimony and Mk, with or without missing data) at a rate of 0.5 (see figure 3 from Wright and Hillis (2014)" as you've mentioned above.
 
@@
 % TG: should I not add something like "and thank the reviewer for pointing it out."?
+% NC: No, only thank them if it's something really stupid - like the mislabeling of the figure axes - or really insightful.
 
@@
 
-We modified the following sentence as suggested: "In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root." (line 547-549@@@).
-
+We modified the sentence as suggested.
 % -----------------
@@
 
-We now refer to two more classical citations (Bremer and Struwe 1992 (American Journal of Botany) and Patterson et al.
-1993 (Annual Review of Ecology and Systematics)) on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to small sized matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters) lines 578-579@@@.
+We now refer to two more classical citations Bremer and Struwe 1992 (American Journal of Botany) and Patterson et al.
+1993 (Annual Review of Ecology and Systematics) on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to having a small matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters) lines 578-579@@@.
 
@@
 
-We modified the following sentence: "The sizes of our data matrices were constrained by the performance of our protocol" (line 579-580@@@).
+We modified the sentence as follows: "The sizes of our data matrices were constrained by the performance of our protocol" (line 579-580@@@).
 
@@
 
-We retracted the entire statement concerning homoplasy and matrix size ("Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters (Wright and Hillis 2014).")
+We removed the entire statement concerning homoplasy and matrix size ("Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters (Wright and Hillis 2014).")
 
@@
 "Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
-For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combinations of the other parameters ($M_F$ and $N_C$; Figure 6).
-In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) doesn't matter and and increase/decrease in both does not affect topology.
+For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combination of the other parameters ($M_F$ and $N_C$; Figure 6).
+In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) does not matter and an increase/decrease in both does not affect topology.
 A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
-This has important practical implications, especially on the best strategy to improve topology by collecting morphological characters (see below)." lines 589-609@@@
+This has important practical implications, especially for the best strategy to improve topology by collecting morphological characters (see below)." lines 589-609@@@
 
@@
 
-We replaced 'cladistic' by 'parsimony' as suggested (line 614@@@).
+We replaced 'cladistic' by 'parsimony' as suggested.
 
@@
 
-We modified the sentence as suggested : "than the ``best'' Maximum Likelihood tree". lines 619-620@@@).
+We modified the sentence as suggested.
 
@@
 
-As developed above (reviewer 1, specific comment 29), we do understand the reviewers concern and added an analysis in the Appendix A (figure A2 and tables A1 and A2) where we tested the effect of using the ``true'' tree as a starting tree.
-We found that there is no significant effect between using the ``true'' tree or a random tree (MrBayes default) as a starting tree on topological recovery.
+As discussed above (reviewer 1, specific comment 29), we understand the reviewer's concern and have added an analysis in Appendix A (figure A2 and tables A1 and A2) where we tested the effect of using the ``true'' tree as a starting tree.
+We found that there is no significant effect of using the ``true'' tree or a random tree (MrBayes default) as a starting tree on topological recovery.
 
@@
 
-We revised the statements in the whole paragraph as follows:
+We revised the paragraph as follows:
 "Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advice using the Bayesian consensus trees as a topological constraint for tree inferences using the Total-Evidence method such as tip-dating (e.g.
-Ronquist et al., 2012a; Wood et al., 2013; Matzke, 2014; although it is possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution).
-Additionally, using the Bayesian consensus tree rather than the Maximum Likelihood can reduce the amount of ``false positive'' topologies.
-As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve nodes weakly supported due to missing data than the Maximum Likelihood tree that is more likely to wrongly resolve such nodes (i.e.
-creating a ``false positive'' node).
-Note, however, that we do not suggest discarding the Bayesian posterior tree distribution once one chose to use the Bayesian consensus tree topology even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
-This can be particularly important because these trees will be invaluable for phylogenetic comparative analyses.
-For example a sub-sample of posterior trees distributions can be used to asses macroecological questions while better taking into account topological uncertainty (e.g.
+Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advise that where a topological constraint is needed, Bayesian consensus trees should be used.
+This may apply to tree inferences using the Total-Evidence method such as tip-dating (e.g.
+Ronquist et al., 2012a; Wood et al., 2013; Matzke, 2014).
+It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
+Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies.
+As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve poorly supported nodes due to missing data than the Maximum Likelihood tree that is more likely to incorrectly resolve such nodes (i.e.
+creating a false positive node).
+Note, however, that we do not suggest discarding the Bayesian posterior tree distributions even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
+These trees will be invaluable for phylogenetic comparative analyses.
+For example a sub-sample of posterior trees distributions can be used to assess macroecological questions while better taking into account topological uncertainty (e.g.
 Fritz et al.
@@
 
+% NC: I think you misunderstood this comment so I have rephrased.
+Their issue was you were recommending people use constraints for tip dating, but then saying not using a constraint might make tip dating better.
+Instead I've made it so you're saying IF you need a constraint, use this, but be careful.
+
 % -----------------
@@
 
-We are thankful for this comment as, indeed, our suggestion was a bit to obvious and not really helpful. See reviewer 1 comment 11 for the changes to the conclusion.
+Agreed. See reviewer 1 comment 11 for the changes to the conclusion.
 
@@
 We removed this sentence and replaced it by the two new sentences outlined above (reviewer 1 specific comment 50).
+
+%NC: Ummm but specific comment 50 refers to comment 11. Should this also refer to comment 11?
 
@@
 
-Unfortunately, we didn't found many studies formally testing the important questions asked by the reviewer. However, there exist few empirical analysis that contain results supporting our claims.
-Firstly, for the question about the topological differences between Maximum Likelihood and the Bayesian consensus tree, in Arcila et al 2015 (MPE)'s supplementary analysis, the authors compare the topology inferred from a Total Evidence matrix in both Likelihood (RAxML) and Bayesian (MrBayes) and the topology of the Bayesian consensus tree is much better supported than the one of the best ML tree.
+Unfortunately, we did not find many studies formally testing the important questions asked by the reviewer. However, there are a few empirical analyses that contain results supporting our claims.
+Firstly, for the question about the topological differences between Maximum Likelihood and the Bayesian consensus tree, in Arcila %NC: Arcial or arcila
+et al 2015 (MPE)'s supplementary analysis, the authors compare the topology inferred from a Total Evidence matrix in both Likelihood (RAxML) and Bayesian (MrBayes) and the topology of the Bayesian consensus tree is much better supported than the one of the best ML tree.
 Specifically, the ML tree has 7/28 nodes have a "fair" bootstrap support ($>$ 75) and 3/28 with a good support ($>$ 90) as the Bayesian consensus tree has 17/25 nodes with a "fair" posterior probability ($>$80) and 14/25 with a good posterior probability ($>$95).
 We therefore used this reference to strengthen our claim of the superiority of the Bayesian consensus tree over the Maximum Likelihood one.
-Secondly, for the question about molecular scaffolds, from the few studies clearly specifying their topological constraint method, only Beck and Lee 2013 (Proceedings B) seems to compare an analysis with and without taxonomic constraint on divergence time and rates estimation (the constraint seems to have no effect) but the branch length aspect is not covered in our study.
-
-See reviewer's comment 11 above for the effective changes in the revised manuscript
+Secondly, for the question about molecular scaffolds, from the few studies clearly specifying their topological constraint method, only Beck and Lee 2013 (Proceedings B) seems to compare an analysis with and without taxonomic constraint on divergence time and rate estimation (the constraint seems to have no effect) but the branch length aspect is not covered in our study.
+
+See our respone to reviewer 1's comment 11 above for changes in the revised manuscript.
 
@@
 
-The Lonsdale cluster requires this specific line in the acknowledgements of any study using it's infrastructure (\url{http://www.tchpc.tcd.ie/resources/acknowledgementpolicy}).
+The Lonsdale cluster requires this specific line in the acknowledgements of any study using its infrastructure (\url{http://www.tchpc.tcd.ie/resources/acknowledgementpolicy}).
 If requested, however, we can also give additional details on the cluster used in the methods section.
@@
 
-We removed the sentence about the t-test similarity from both figures captions. %TG: not sure if that's what the reviewer wanted.
+We removed the sentence about the t-test similarity from both figure captions. %TG: not sure if that's what the reviewer wanted. % NC: Yeah I think that's what they meant
 
@@
 
-We apologies for this silly mistake and thanks the reviewer for pointing it out.
+We apologies for this silly mistake and thank the reviewer for pointing it out.
 We changed the colors to match with the paper color scheme: "Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
@@
 We also fixed the Normalised Triplets metrics axis to range from -0.5 to 1.
-We added a horizontal doted line at Triplets = 0 to make sure the reader take notes is this change of scale.
+We added a horizontal dotted line at Triplets = 0 to make sure the reader notes this change of scale.
 
@@
 We thank the reviewer for pointing out this difficulty and acknowledge that such graphical representations of so many pairwise comparisons might be confusing.
-We emphasized the figure caption, the method section explaining how we calculated this figure and the results.
+We have made the figure caption clearer, and the methods section explaining how we calculated this figure and the results.
 
@@
 The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
-Blue regions denotes a poor overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a strong effect on the Normalised metric and thus the topological recovery).
-Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a weak effect on the Normalised metric and thus the topological recovery)." See figure 6 caption.
+The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to high probability of overlap in orange.
+Blue regions denote a poor overlap in Normalised metric values between the different parameter combinations (i.e.
+the parameters have a strong effect on the metric and thus the topological recovery).
+Conversely, orange regions denote a high overlap in Normalised metric values between the different parameter combinations (i.e.
+the parameters have a weak effect on the metric and thus the topological recovery)." See figure 6 caption.
+
+% NC: Sometimes in this you have N_C, sometimes M_C. FIX IT!!!
 
 We modified the methods section as follows:
-"Because of the difficulties to represent so many pairwise comparisons in a meaningful way, we summarized as a heat map of Bhattacharyya Coefficients (see Figure 6).
-In this type of figures, parameters that have similar or different effects on recovering the ``best'' topology (rather good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6)." lines 434-439@@@
+"Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Figure 6).
+In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6)." lines 434-439@@@
 
@@
 The orange upper triangle in Figure 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Figure 6A).
-There is no additional effect (i.e. no dissimilarities) of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+There is no additional effect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
 Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+% NC: This sentence and the one before it seem to be the same thing?
 Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
-This can be interpreted as, in figure 5 for the Normalised Robinson-Foulds metric, the overlap between the distributions once $M_L$=50\%." lines 480-497@@@
+In figure 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%." lines 480-497@@@
 
@@
 % TG: I don't understand this comment at all. What does "The wording of the caption is off" means? And which pararagraph is he talking about? 
-We fixed the highlithing in bold to the classical values of 0.05 or 0.95.
+% NC: The "wording is off" means the wording is incorrect. Probably means the second sentence not paragraph. I've editted the caption as mentioned further up in the comments.
+
+We fixed the highlighting in bold to the classical values of 0.05 or 0.95.
 We changed the caption to:
 "Bhattacharyya Coefficients of the pairwise method comparisons.
-Each comparisons corresponds to the comparison of the full (125) distributions between the ``best'' tree and the ``missing data'' trees for all the four methods (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) and the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
-Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
-The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
+Each line summarizes the probabilities of overlap between the distributions of the 'best' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
+Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
 If two methods have a high probability of overlap, they have a similar ability to recover the ``correct" tree topology"
@@
 
+% NC: OK I think the point here is that Table 1 has a column saying the "Max." BC score. So the reviewer is saying that this is happening:
+% In Appendix C, B posterior vs B consensus for some parameter value is 0.5 (for example), but the max value when you compare this in Table 1 is 0.11. How is this possible???????
+% NC: So your response here is completely irrelevant.
+
 2) %TG: not entirely sure what the reviewer means by marginal probabilities.
@@
 
+% NC: https://en.wikipedia.org/wiki/Marginal_distribution
+% NC: Marginal probabilities are basically the probability of one thing given you're controlling for another thing.
+So they are looking to see the effect of tree inference method independent of parameter differences.
+Otherwise they suggest you are hiding important differences via this summarizing over all parameter combinations.
+% NC: I think you can argue that yes, table 1 hides some important effects, but that it was difficult to summarize these results and more in depth tables can be seen in the appendix.
+
 % -----------------
@@
 
-We removed the repetition in the caption. %Not sure I understand the second part of the comment.
+We removed the repetition in the caption.
+If we understand the second point correctly, we have not added "best" to these as the Maximum Likelihood tree is by definition the tree with the highest likelihood, and the Bayesian consensus is the consensus tree of the Bayesian inference trees.
+So adding the word best here is not entirely accurate.
+
+%Not sure I understand the second part of the comment.
+% NC: I think they are a bit confused and think you need to explain it is the best ML and best bayesian tree.
+But that's what ML and consensus means.
+Maybe they want clarification that it is only one tree in each of those and multiple trees in the others?
+I've added a bit above.
 
@@
 
-We updated the readme page of the github repository linking to the various R codes.
-We also now provided a version of the analysis in R ran thoughout the manuscript as a step by step tutorial for a subset of data.
+We updated the README page of the GitHub repository linking to the various R code to make it clearer what files are needed.
+We also now provide a version of the analysis in R run thoughout the manuscript as a step by step tutorial for a subset of data.
 See \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Analysis_Demo.R} and Appendix D.
+% NC: Maybe add something like this if this is your intention. It's probably worth it. No need to keep lots of un-necessary files.
+We will also clean up this repo once the paper has been accepted. At present it is functioning both as our working repo for code as well as the repo for this paper.
 

### 2015-08-06 Manuscript/MPESubmission/Revision/response.tex
Edits up to reviwer 2 comments

@@
 \section{Reviewer 2:}
-Note that we split the second comment from reviewer 2 in three different comments (comments 2,3 and 4 below) to improve the readability of our response to the reviewers.
+Note that we split the second comment from reviewer 2 in three different comments (comments 2,3 and 4 below) to improve the readability of our response to the reviewers
 

### 2015-08-06 Manuscript/MPESubmission/Revision/response.tex
Comment 4 update

--- 2015-08-06 Manuscript/MPESubmission/Revision/response.tex
@@
 If the second, it's not that a big problem (just annoying) but I think we should try sending the pdf, it's way clearer like that.
-% NC: You can usually upload a file for the cover letter/response so pdf should be fine.
+% NC: You can usually upload a file for the cover letter/response so pdf should be fine. TG: Cool!
 
@@
 We gladly followed reviewer 1's graphical abstract suggestions and have changed it to be similar to Figure 5.
-% TG: It's freaking not catchy at all but at this point I give up.
-I've tried many ways but the flippin' paper IS NOT attractive.
-It might be useful, well written, interesting, citable, etc...
-But it won't be sexy.
-% NC: Agreed! :)
 
@@
 We rephrased the highlights as follows: %3 to 5 bullet points (maximum 85 characters, including spaces, per bullet point). TG: probably need improvement here.
-% NC: Yeah I never even looked at these first time round. Mine aren't great but do fit in 85 characters!
+% NC: Yeah I never even looked at these first time round. Mine aren't great but do fit in 85 characters! TG: that looks way better, thanks!
 
@@
 
-"Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
+% "Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
 using both molecular and morphological data; e.g.
@@
 % NC: Argh this is really confusing. It doesn't logically flow
-For example, Wiens et al. (2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
-Furthermore, Pattinson et al. (2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
-For morphological and molecular data separately, both studies show that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
-However, they are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Manos et al., 2007; Pattinson et al., 2014).
-Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data)." lines 85-98 @@@
+% For example, Wiens et al. (2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
+% Furthermore, Pattinson et al. (2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
+% For morphological and molecular data separately, both studies show that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
+% However, they are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Manos et al., 2007; Pattinson et al., 2014).
+% Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data)." lines 85-98 @@@
 %%% NC: Paragraph above needs work aka a bit of logic to it.
+
+% TG: how's that (+idea flow):
+% 0(before this section)-Missing data can be a problem
+% 1-And until now, not many studies
+Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
+using both molecular and morphological data; e.g.
+Wiens et al., 2005; Manos et al., 2007; Pattinson et al., 2014).
+% 2-These studies have done X and Y to see if missing data's a problem
+These studies have assessed the effect of missing data on topology either:(1) by comparing a dataset with missing data to sub-sets without missing data (Wiens et al 2005); or (2) by removing both molecular and some morphological data from living taxa (``artificial fossil'') (Manos et al., 2007; Pattinson et al., 2014).
+% 3-They show that it's not a problem
+Both types of approaches have shown that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
+% 4-But they've done it a certain way
+However, these studies were limited by their empirical-only approach and the use of only living taxa (Wiens et al., 2005) or the scheme of generating missing data (i.e.
+using the patterns from the fossil record; Manos et al., 2007; Pattinson et al., 2014).
+Besides, the studies using fossil data, mainly on the palaeontological aspect of the question (i.e.
+the effect of missing data in fossils in Total Evidence matrices; Manos et al., 2007; Pattinson et al.
+2014).
+% (next §)5-We're doing it this way
 
@@
 
-"We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred % NC: Didn't one reviewer have issues with the word inferred?
-using Maximum Likelihood and Bayesian methods by using two different topological distance metrics, based on the Robinson-Foulds (Robinson and Foulds, 1981) and Triplets (Critchlow et al., 1996) distances.
-Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological distance metrics." lines 127-134 @@@
+% "We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred % NC: Didn't one reviewer have issues with the word inferred?
+TG: for the matrices only (and he/she's right); trees are still inferred.
+% using Maximum Likelihood and Bayesian methods by using two different topological distance metrics, based on the Robinson-Foulds (Robinson and Foulds, 1981) and Triplets (Critchlow et al., 1996) distances.
+% Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological distance metrics." lines 127-134 @@@
 % NC: Why are the distinct topological metrics novel?
+
+% TG: I actually rewrote this bit as well:
+We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
+We inferred the topology from the matrices using both Maximum Likelihood and Bayesian inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and for wild-card taxa placement (based on the Triplets distance; Critchlow et al., 1996).
+Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring difference on two aspects of topology (c.f.
+Wiens et al., 2005; Pattinson et al., 2014 measuring respectively only the clades support or the number of shared clades).
 
@@
 We have now compared both the ``best'' tree and the ``missing-data'' trees to the ``true'' tree as suggested by both reviewers.
-The results are available in appendix B; figure B1 and appendix C; figures C6 and C7).
+The results are available in appendix B; figure B1 and appendix C; figures C6 and C7.
 % NC: Worth describing the results here.
@@
 We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or using a random tree (default in MrBayes) as a starting tree.
-% NC: Yes but the reviewer is asking for something different - they ask you to have no starting tree. If this is impossible in MrBayes then say so, and say you used a random tree instead. 
+% NC: Yes but the reviewer is asking for something different - they ask you to have no starting tree.
+If this is impossible in MrBayes then say so, and say you used a random tree instead.
+TG: Maybe there's a phylogeny confusion here.
+In both ML or Bayesian, you need to start from a tree, then calculate the likelihood of all the parameters, then change the tree, calculate again, etc...
 We found no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
-Therefore, we do not believe our results are merely the result of strong priors on the tree topology.
+Therefore, we do not believe our results are merely the result of using a specific starting tree topology %strong priors on the tree topology.
+TG: That's the big difference that the PLoS ONE reviewer also didn't got.
+A starting tree is not a Bayesian prior (in P(A|B) = P(B|A)*P(A)/P(B; P(A) is the prior).
+It's just a user fixed starting point for the MCMC walk rather than a random one.
 % NC: The reviewer also says that saving time is not a good justification.
 It'd be worth explaining here how long it actually takes even though these datasets are small.
+TG: about that: it using a random tree or the true tree doesn't change the timing significantly.
+The whole idea was suggested by Rob or Gavin (don't remember exactly) at lunch time when I was visiting Gavin in Sheffield.
+To speed up the search in the tree space they suggested I already start from a sensible tree in the tree space, that was meant to speed up the analysis.
 These results are available in the supplementary materials Appendix A (figure A2 and tables A1 and A2) and we also added a note to the manuscript:
@@
 "We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
-Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly % NC: Significantly as in statisically? if not change this word.
+Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly % NC: Significantly as in statisically?
+if not change this word.
+TG: yap, p-value all all the mumbo-jumbo
 affect topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line 323-327 @@@.
@@
 But if they do overlap all the way to 25% then it is weird that the highest BC is 0.11 isn't it?
+TG: yes definitely.
+Just doubled check, it might be because the colors where miss labelled and the reviewer miss read the graph (my fault).
 

### 2015-08-06 Manuscript/MPESubmission/Revision/response.tex
Edits end of reviewer comments

--- 2015-08-06 Manuscript/MPESubmission/Revision/response.tex
@@
 \section{Reviewer 2:}
-Note that we split the second comment from reviewer 2 in three different comments (comments 2,3 and 4 below) to improve the readability of our response to the reviewers
+Note that we split the second comment from reviewer 2 in three different comments (comments 2,3 and 4 below) to improve the readability of our response to the reviewers.
 
@@
 A similar point/suggestion was also raised by reviewer 1 (see Reviewer 1 General comment number 6 and specific comment number 25).
-We do understand the concerns of this reviewer as well but, again, we do not agree that comparing the "missing-data" trees to the "best" trees adds a layer of complexity to our simulations.
+We understand the concerns of the reviewer but, again, we do not agree that comparing the "missing-data" trees to the "best" trees adds a layer of complexity to our simulations.
 In fact, we argue that comparing the "missing-data" trees to the "true" tree actually adds a layer of complexity to the analysis.
-The problematic we are tackling in this manuscript is to understand "what is the effect of our three parameters on recovering topology ?"
-To answer this question, however, we have to add a layer between the data (i.e. the "complete" or "missing-data" matrices) and the actual topology which is the tree inference.
+The problem we are tackling in this manuscript is to understand "what is the effect of our three parameters on recovering topology ?"
+To answer this question, we have to add a layer between the data (i.e. the "complete" or "missing-data" matrices) and the actual topology which is the tree inference.
 We test the effect of this layer by using different tree inference methods.
 
-Because we use a simulation approach to test this problematic, we have to start from a random starting point (the "true" tree) and then do the following steps: (1) generate and remove the data (2) infer the topologies and (3) compare the topologies.
-Note that in this schematic representation, the "true" tree is the step (0) of our analysis and involves a series of algorithms and softwares (i.e.
+% NC: Note: problematic does not mean what you think it does... Something can be problematic (it's an adjective not a noun).
+
+Because we use a simulation approach to test this problem, we have to start from a random starting point (the "true" tree) and then do the following steps: (1) generate and remove the data (2) infer the topologies and (3) compare the topologies.
+Note that in this schematic representation, the "true" tree is the step (0) of our analysis and involves a series of algorithms and software (i.e.
 seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes).
-To compare the "true" tree's topology to the "missing-data" topologies will be equivalent to explore the effect of our three parameters on recovering topology through 3 layers (1- generate the data; 2- infer the topologies; and 3- compare the topologies).
+To compare the "true" tree's topology to the "missing-data" topologies will be equivalent of exploring the effect of our three parameters on recovering topology through 3 layers (1- generate the data; 2- infer the topologies; and 3- compare the topologies).
 By comparing the "best" tree's topology to the "missing-data" ones, we explore the effect of our three parameters through only 2 layers (2- infer the topologies; and 3- compare the topologies).
@@
 
-However, we have added all the comparisons between the "true" tree and the "missing-data" trees in the Appendix C (Figures C6 and C7).
-It is interesting to notice that the algorithms and softwares as well as probably the size of our matrices are actually not ideal an the recovery of the "true" tree is rather low and scales down the our results (see figure @).
-However, even though these comparisons are worst, it is encouraging to see that the patterns are conserved (i.e.
+% NC: This argument makes no sense. The point the reviewer is making is that we go:
+% True tree > Best tree > comparisons when we could just go True tree > comparisons.
+So our approach *is* more complex than their suggestion.
+Instead the justification here is the same as used for reviewer one - that in real life scenarios we never know the true tree and we want it to be as close to real life as possible.
+BUT that we have tested their idea and find it makes no difference but all the results are in the appendix.
+
+However, we have added analyses with comparisons between the "true" tree and the "missing-data" trees in the Appendix C (Figures C6 and C7).
+It is interesting to note that the algorithms and software as well as probably the size of our matrices are actually not ideal for the recovery of the "true" tree is rather low and scales down the our results (see figure @).% NC: HUH????
+However, even though these comparisons are worse than comparing the best tree to the missing data trees, it is encouraging to see that the patterns in the results are conserved (i.e.
 Bayesian consensus trees $>$ Maximum Likelihood trees $>$ both bootstraps trees and Bayesian posterior trees).
@@
 
-We assumed that this reviewer is mentioning the section "2.2 Removing data" as the "explanation of the missing data strategy". We modified the whole section emphasising on clarity and simplicity.
-
-"To explore the effect of missing data on topological recovery, we removed various amounts of the “complete” matrix to obtain matrices with missing morphological data.Hereafter, we call these matrices with missing morphological data the “missing-data” matrices.
-We removed morphological following three main data incompleteness parameters:
+We assume the reviewer is referring to section "2.2 Removing data". We therefore modified the whole section to improve its clarity and simplicity as follows:
+
+"To explore the effect of missing morphological data on topological recovery, we removed various amounts of the “complete” matrix to obtain matrices with missing morphological data.
+Hereafter, we call these matrices with missing morphological data the “missing-data” matrices.
+% NC: add clarification here about molecular data
+Note that the amount of molecular data remained constant throughout our simulations - XX in living taxa and no molecular data for fossil taxa (see above).
+% NC: Is this correct?
+ We removed morphological data using three data incompleteness parameters:
 1.
@@
 It represents the number of living taxa that are present in the matrix but have only molecular data available.
-This reflects the fact that, because of the increasing facility of collecting molecular data, morphological data for living species is rarely collected (Guillerme and Cooper, 2015).
+This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species is rarely collected (Guillerme and Cooper, 2015).
 Therefore, many living species will have only molecular data available.
-In practice, we removed all the morphological data from randomly chosen living with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.
+In practice, we removed all the morphological data from randomly chosen living taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.
 2.
@@
 This missing data parameter represents the completeness of the fossil record.
-In fact, due to preservation biases, missing data for fossil taxa are common (Sansom and Wills, 2013).
+Due to preservation biases, missing data for fossil taxa are common (Sansom and Wills, 2013).
 In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
@@
 The number of morphological characters for both living and fossil taxa ($N_C$).
-This parameter is not a missing data parameter per se but rather an indication of the size of the matrix.
+This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
 Any morphological matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
-Therefore, this parameter correspond to the overall number of characters available for both living and fossil taxa.
-In practice, we randomly removed whole characters from the morphological matrix shrinking it to: 100, 90,75, 50 or 25 characters.
+Therefore, this parameter corresponds to the overall number of characters available for both living and fossil taxa.
+In practice, we randomly removed whole characters from the morphological matrix reducing it to: 100, 90,75, 50 or 25 characters.
 Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of “missing” morphological characters).
 
-In practice, each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living taxa’s data; $M_F$ removes cells from the fossil taxa’s data; and $N_C$ removes columns across both living and fossil taxa’s data.
+In practice, each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
 Note that $M_L$  and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
@@
 
+% NC: NC_C here again not M_C...
+
 % -----------------
@@
 However, because I didn't answered the first and last comment of this reviewer, I felt that at least 2/4 is a minimum.
-We entirely agree with this comment and hope it will help to the global comprehension of the manuscript.
+We agree with this comment and hope it will help to the global comprehension of the manuscript.
 We replaced the $M_C$ (missing characters) parameter by $N_C$ (number of characters with five states: 100, 90, 75, 50 and 25) in the revised manuscript (see reviewer 2 comment 2 above), the appendices and the figures.
 We also slightly modified some explanations related to the change of missing characters to number of characters throughout the manuscript.
+
+% NC: Ah OK. You haven't corrected this everywhere in these comments I don't think...
 
@@
 no missing data for living taxa, and maximal number of characters but various amount of missing data for the fossil taxa).
-This results are present in figure 4, central row.
-However, the amount of changes among the fossil taxa was not set to any specific amount of missing data per fossil (e.g.
+These results are present in figure 4, central row.
+However, we think we may have not explained ourselves entirely clearly.
+The amount of changes among the fossil taxa was not set to any specific amount of missing data per fossil (e.g.
 like the "artificial fossil" approach in Pattinson et al 2014, Syst.
@@
 Because our fossil taxa and their characters where randomly simulated, we didn't had any \textit{a priori} assumption on how to remove the data.
-We therefore decided to randomly remove a certain amount of data among all the fossil.
+We therefore decided to randomly remove a certain amount of data among all the fossils.
 For example, when removing 75\% of the data for the fossil taxa ($M_F$=75\%) different scenarios could happen, ranging from a "homogeneous" incompleteness data distribution (each specific fossil taxa had 75\% of missing data) to a "heterogeneous" incompleteness data distribution  (80\% of the specific fossil taxa had 5\% available data and 21\% of the specific fossil taxa had 100\% available data).
@@
 
-Also, a second argument for using this "simplistic" way to deal with the missing data in the fossil record follows our main recommendation from this study: the best way to improve the quality of Total Evidence trees is to have a maximum number of living species with available morphological data.
+% NC: OK - I'd add a line somewhere in the manuscript to make this clear. something like:
+
+%"Note that 55\% missing data for fossil taxa does not mean that each fossil is missing 50\% of it's morphological data.
+Instead this 50\% refers to the whole matrix.
+Some fossils may reatin 100% of their data and others may lose all their data at this parameter value"
+
+Also, a second argument for using this slightly simplistic way to deal with the missing data in the fossil record follows our main recommendation from this study: the best way to improve the quality of Total Evidence trees is to have a maximum number of living species with available morphological data.
 In fact lines 584 to 589 in the first version of the manuscript (line 669-674@@@ in the revised version), we argue that even if the effect of the $M_F$ parameter is important, there is unfortunately not much we can do about it since the completeness of this part of the matrix solely depends on exceptional palaeontological discoveries and excavation effort.
@@
 
+% NC: Not sure the second paragraph is needed.
+The reviewer is saying that they want to know what happens if living taxa have all their data but fossils have variable amounts.
+And I think this is what you've done for some parameter combinations.
+Just need to apologise for not being clear and clarify in the text how data were missing from fossils.
+Doesn't need so much waffle and defensiveness - they are asking you to do what you've already done!
+:)
+
 \end{enumerate}

### 2015-08-06 Manuscript/MPESubmission/Revision/response.tex
edits on revision

@@
 The results are available in appendix B; figure B1 and appendix C; figures C6 and C7.
+The ability to recover the ``true'' tree is poor in both methods and using both metrics (Maximum Likelihood 50\% CI approximatively between 0.2 and 0.3 Normalised Robinson-Foulds and Triplets metric; Bayesian consensus tree 50\% CI approximatively between 0.3 and 0.4 Normalised Robinson-Foulds metric and between 0 and 0.2 Normalised Triplets metric).
+This can be related to the size of our matrices or to the ability of the algorithms to generate the ``complete matrix'' but also to our decision to put our minimum median bootstrap acceptance at 50 rather than higher values (line @@@ in the former manuscript and @@@ in our revision).
+When comparing the ``true'' tree to the ``missing data'' trees, the results are \textit{scaled down} due to the low scores of the ``true'' \textit{vs.} ``best'' trees scores.
+However, we can observe the same patterns between the methods as when using the ``best'' tree (i.e. Bayesian always outperforms ML in clade conservation but does poorly in wild-card taxa placement).
+An interesting feature is that the Normalised Robinson-Foulds metrics actually \textit{increases} with missing data.
+This is due to the fact that the difference between the ``best'' and the ``true'' tree is high making trees with enough data more different.
+When data is missing, however, more clades are unresolved and thus closer to clades from the ``true'' trees.
+
 % NC: Worth describing the results here.
@@
 We added the following sentence to the main text to clarify and link to the supplementary analysis:
-"In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices %NC: What dod you mean by R packages choices?
+% "In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices %NC: What dod you mean by R packages choices?
 Do you mean the way the trees were simulated?
- (i.e. diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
-The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." line 276-280 @@@.
+%  (i.e. diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
+% The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." line 276-280 @@@.
+
+% TG: changed to:
+"We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
+In practice, the difference between the ``best'' trees and the ``missing data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing data'' trees.
+The difference between the ``true'' and the ``missing data'' tree, however, represents the effect of our parameters used to generate the ``true'' tree and the algorithms used to generated the ``complete'' matrix as well as the effect of our missing data parameters and the phylogenetic methods used.
+Because the main aim of this study, is to look at the effect the missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees to ``missing data'' trees.
+The results of the comparisons of the ``true'' and the ``missing data'' tree are available in appendix B.
+Note that these results makes little difference to our overall results." line 276-280 @@@.
+
 See also appendix B section "Differences between the ``true'' and the ``best'' trees."
@@
 We found no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
-Therefore, we do not believe our results are merely the result of using a specific starting tree topology %strong priors on the tree topology.
+Therefore, we do not believe our results are merely the result of using a specific starting tree topology.
+%strong priors on the tree topology.
 TG: That's the big difference that the PLoS ONE reviewer also didn't got.
@@
 It's just a user fixed starting point for the MCMC walk rather than a random one.
-% NC: The reviewer also says that saving time is not a good justification.
-It'd be worth explaining here how long it actually takes even though these datasets are small.
-TG: about that: it using a random tree or the true tree doesn't change the timing significantly.
+% NC: The reviewer also says that saving time is not a good justification. It'd be worth explaining here how long it actually takes even though these datasets are small.
+%
+% TO DISCUSS @@@
+% TG: about that: it using a random tree or the true tree doesn't change the timing significantly.
 The whole idea was suggested by Rob or Gavin (don't remember exactly) at lunch time when I was visiting Gavin in Sheffield.
 To speed up the search in the tree space they suggested I already start from a sensible tree in the tree space, that was meant to speed up the analysis.
+%
 These results are available in the supplementary materials Appendix A (figure A2 and tables A1 and A2) and we also added a note to the manuscript:
@@
 % NC: Make sure this is consistent throughout, add this to where you mention the BC being between 0 and 1 too?
+
+% TO DO IN THE MANUSCRIPT @@@
 
@@
 But if they do overlap all the way to 25% then it is weird that the highest BC is 0.11 isn't it?
-TG: yes definitely.
-Just doubled check, it might be because the colors where miss labelled and the reviewer miss read the graph (my fault).
+% TO CHECK @@@
+% TG: yes definitely. Just doubled check, it might be because the colors where miss labelled and the reviewer miss read the graph (my fault).
 
@@
 "The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (figure 5).
-A Normalised Triplets metric value below 0 means that the placement of some taxa is worse % NC: It's worse not worth!
-than expected by just randomly placing this taxon in the tree.
+A Normalised Triplets metric value below 0 means that the placement of some taxa is worse than expected by just randomly placing this taxon in the tree.
 This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
@@
 
-We rewrote the conclusion to deal with this comment and the specific comments 13, 52 (see comment 52 for details on the choice of citing Arcial et al 2015.) 
-% NC: Is it Arcial or Arcila?
-and 56.
+We rewrote the conclusion to deal with this comment and the specific comments 13, 52 (see comment 52 for details on the choice of citing Arcila et al 2015.) and 56.
+% NC: Is it Arcial or Arcila? TG: Dahiana Arcila
 
@@
 What aspects?
+
+% TO DO: develop the limitations @@@
+
 This study proposes a more in depth approach where missing data is generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
@@
 Additionally, the topology of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
-This has also been observed in empirical data (e.g. Arcila et al., 2015) 
-% NC: Arcial or arcila?
-where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This has also been observed in empirical data (e.g. Arcila et al., 2015) where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
 This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds (e.g.
@@
 Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.
+% TG: maybe add in a non tip-dating analysis?
 
@@
 I noticed this on my re-read just now too.
+
+% TO DO @@@
 
@@
 We modified the sentence as suggested.
+
 % -----------------
@@
 
-We removed this sentence and replaced it by the two new sentences outlined above (reviewer 1 specific comment 50).
+We removed this sentence and replaced it by the two new sentences outlined above (reviewer 1 comment 11).
 
@@
 Unfortunately, we did not find many studies formally testing the important questions asked by the reviewer. However, there are a few empirical analyses that contain results supporting our claims.
-Firstly, for the question about the topological differences between Maximum Likelihood and the Bayesian consensus tree, in Arcila %NC: Arcial or arcila
-et al 2015 (MPE)'s supplementary analysis, the authors compare the topology inferred from a Total Evidence matrix in both Likelihood (RAxML) and Bayesian (MrBayes) and the topology of the Bayesian consensus tree is much better supported than the one of the best ML tree.
+Firstly, for the question about the topological differences between Maximum Likelihood and the Bayesian consensus tree, in Arcila et al 2015 (MPE)'s supplementary analysis, the authors compare the topology inferred from a Total Evidence matrix in both Likelihood (RAxML) and Bayesian (MrBayes) and the topology of the Bayesian consensus tree is much better supported than the one of the best ML tree.
 Specifically, the ML tree has 7/28 nodes have a "fair" bootstrap support ($>$ 75) and 3/28 with a good support ($>$ 90) as the Bayesian consensus tree has 17/25 nodes with a "fair" posterior probability ($>$80) and 14/25 with a good posterior probability ($>$95).
@@
 
-See our respone to reviewer 1's comment 11 above for changes in the revised manuscript.
+See our response to reviewer 1's comment 11 above for changes in the revised manuscript.
 
@@
 % NC: Sometimes in this you have N_C, sometimes M_C. FIX IT!!!
+
+% TO DO: thorough check through. @@@
 

### 2015-08-07 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
updated Natalie's email
Updated introduction revisions (rev 1)

@@
 $^*$Corresponding author. Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.\\
-$^1$Present address: Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK. E-mail: nhcooper12@gmail.com}\\
+$^1$Present address: Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK. E-mail: natalie.cooper@nhm.ac.uk}\\
 \end{center}
@@
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
-\hl{Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework \protect\citep[\hl{i.e.
+\hl{Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework }\citep[\hl{i.e.
+using both molecular and morphological data; e.g.}][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{These studies have assessed the effect of missing data on topology either:(1) by comparing a dataset with missing data to sub-sets without missing data }\citep{Wiens01102005}\hl{; or (2) by removing both molecular and some morphological data from living taxa (``artificial fossil'') }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{Both types of approaches have shown that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
+However, these studies were limited by their empirical-only approach and the use of only living taxa }\citep{Wiens01102005}\hl{ or the scheme of generating missing data }\citep[\hl{i.e.
+using the patterns from the fossil record;}][]{manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{Besides, the studies using fossil data, mainly on the palaeontological aspect of the question }\citep[\hl{i.e.
+the effect of missing data in fossils in Total Evidence matrices;}][]{manosphylogeny2007,pattinsonphylogeny2014}.
+%Former revised version
+% \hl{Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework \protect\citep[\hl{i.e.
 using combined molecular and morphological data}; e.g.][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.}
-\hl{For example, \protect\cite{Wiens01102005} }\hl{only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.}
-\hl{Furthermore, \protect\cite{pattinsonphylogeny2014} }\hl{assess the effect of removing both molecular and morphological data (``artificial extinction'') on the topology.}
-\hl{As for morphological and molecular data separately, these studies shows that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.}
-\hl{These studies, however, are limited by their empirical-only approach \protect\citep{Wiens01102005} }\hl{and the scheme of generating missing data \protect\citep{manosphylogeny2007,pattinsonphylogeny2014}.}
-\hl{Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data).}
+% \hl{For example, \protect\cite{Wiens01102005} }\hl{only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.}
+% \hl{Furthermore, \protect\cite{pattinsonphylogeny2014} }\hl{assess the effect of removing both molecular and morphological data (``artificial extinction'') on the topology.}
+% \hl{As for morphological and molecular data separately, these studies shows that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.}
+% \hl{These studies, however, are limited by their empirical-only approach \protect\citep{Wiens01102005} }\hl{and the scheme of generating missing data \protect\citep{manosphylogeny2007,pattinsonphylogeny2014}.}
+% \hl{Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data).}
 
-\hl{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by thoroughly eliminating living or fossil data or all data overall.
+\hl{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living or fossil data, or all data for certain characters.
 We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa.}
@@
 the size of the matrix).
-\hl{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods by using two different topological metrics, based on the Robinson-Foulds }\citep{RF1981}\hl{ and Triplets }\citep{critchlowthe1996}\hl{ distances.
-Thus, in the present study, we propose a new approach on the missing-data question focus on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological metrics.}
-We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the ``best'' tree topology more so than minimizing the amount of missing data in the fossil record.
-Additionally, we find that the ability of Total Evidence methods to recover the ``best'' tree topology is increased when using Bayesian methods.
+%in
+\hl{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
+We inferred the topology from the matrices using both Maximum Likelihood and Bayesian inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation } \citep[\hl{based on the Robinson-Foulds distance;}][]{RF1981}\hl{ and for wild-card taxa placement }\citep[\hl{based on the Triplets distance;}][]{critchlowthe1996}.
+\hl{Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring difference on two aspects of topology }\citep[c.f.][\hl{measuring respectively only the clades support or the number of shared clades}]{Wiens01102005,pattinsonphylogeny2014}.
+%Former revised version
+% \hl{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods by using two different topological metrics, based on the Robinson-Foulds }\citep{RF1981}\hl{ and Triplets }\citep{critchlowthe1996}\hl{ distances.
+% Thus, in the present study, we propose a new approach on the missing-data question focus on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological metrics.}
+% We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the ``best'' tree topology more so than minimizing the amount of missing data in the fossil record.
+% Additionally, we find that the ability of Total Evidence methods to recover the ``best'' tree topology is increased when using Bayesian methods.
 

### 2015-08-07 Manuscript/MPESubmission/Revision/response.tex
Updated introduction revisions (rev 1)
Updated revision (major rev 1)

@@
 % Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological distance metrics." lines 127-134 @@@
-% NC: Why are the distinct topological metrics novel?
+% NC: Why are the distinct topological metrics novel? TG: not novel but novel to use two aspects (people used only clade related metrics before).
 
@@
 
-% NC: Worth describing the results here.
+% NC: Worth describing the results here. TG: Done.
 
@@
 
+"The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 (full overlap; Bhattacharyya, 1943, see Appendix B for calculation details)." lines @@@
+
 "When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
@@
 
-% NC: Make sure this is consistent throughout, add this to where you mention the BC being between 0 and 1 too?
-
-% TO DO IN THE MANUSCRIPT @@@
+% NC: Make sure this is consistent throughout, add this to where you mention the BC being between 0 and 1 too? TG: Done.
 
@@
 But if they do overlap all the way to 25% then it is weird that the highest BC is 0.11 isn't it?
-% TO CHECK @@@
-% TG: yes definitely. Just doubled check, it might be because the colors where miss labelled and the reviewer miss read the graph (my fault).
+% TG: yes definitely.
+Just doubled and tripled checked, it might be because the colors where miss labelled and the reviewer miss read the graph (my fault).
+In the tables 1, C5, C6 and C7, RF overlap are always low (max < 0.11) and Tr overlap are always high (max ~80+).
+And that's what's visible in figure 4.
+Grey and orange lines do not overlap in RF and do overlap in Tr.
 
@@
 
-"Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data (Wiens et al 2005; Pattinson et al 2014).
+% "Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data (Wiens et al 2005; Pattinson et al 2014).
 % NC: I still don't get these limitations.
@@
 What aspects?
-
-% TO DO: develop the limitations @@@
-
+% TG: Limitations are:
+% Empirical data (Wiens et al., 2005; Manos et al.,2007, Pattinson et al., 2014)
+% No fossils (Wiens et al., 2005)
+% Scheme of generating missing data (Manos et al.,2007, Pattinson et al., 2014)
+% Focus on palaeontological aspect of missing data (Manos et al.,2007, Pattinson et al., 2014)
+% Using clade conservation metrics only (Wiens et al., 2005; Pattinson et al., 2014)
+
+% TG: I rephrased as follows.
+Previous studies have explored the effect of missing morphological data in Total Evidence matrices (Wiens et al., 2005; Manos et al.,2007, Pattinson et al., 2014).
+The conclusions of theses studies, however, were limited by their empirical approach as well as several aspects of their experimental design (e.g.
+not including fossil taxa, Wiens et al., 2005; or focusing on the palaeontological aspect, Manos et al.,2007, Pattinson et al., 2014).
+%
 This study proposes a more in depth approach where missing data is generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
@@
 Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.
-% TG: maybe add in a non tip-dating analysis?
+% TG: maybe add in a non tip-dating analysis way (i.e. Nick Matzke comment)?
 

### 2015-08-07 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Updated revision (major rev 1)

@@
 In real world situations, the ``true'' tree is never available to us because we cannot know the true evolutionary history of a clade \citep[except in very rare circumstances, e.g.][]{rozen2005}.
-\hl{In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices (i.e.
+\hl{We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
+In practice, the difference between the ``best'' trees and the ``missing data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing data'' trees.
+The difference between the ``true'' and the ``missing data'' tree, however, represents the effect of our parameters used to generate the ``true'' tree and the algorithms used to generated the ``complete'' matrix as well as the effect of our missing data parameters and the phylogenetic methods used.
+Because the main aim of this study, is to look at the effect the missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees to ``missing data'' trees.
+The results of the comparisons of the ``true'' and the ``missing data'' tree are available in appendix B and C.
+Note that these results makes little difference to our overall results.}
+% Former revised version
+% \hl{In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices (i.e.
 diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).}
-\hl{The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery.}
+% \hl{The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery.}
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the ``best'' tree, rather than the ``true'' tree, as the ``best'' tree is generally what we have to work with.
@@
 For each estimation, we used the ``true'' tree's topology as a starting tree (with a starting value for each branch length of one).
-\hl{We used a fixed starting tree rather than a random starting tree} \citep[\hl{default MrBayes;}][]{Ronquist2012mrbayes} \hl{to speed up our Bayesian inferences.
-Note that a starting is not a Bayesian prior on topology \textit{per se} and did not significantly affected topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').}
+\hl{We used a fixed starting tree rather than a random starting tree }\citep[\hl{default MrBayes;}][]{Ronquist2012mrbayes} \hl{to speed up our Bayesian inferences.
+Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').}
 We also used two priors on the molecular part of the matrix: an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5, and a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)); and one prior on the morphological part of the matrix (exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5).
@@
 We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions \citep{Bhattacharyya} (see Appendix B for calculation details).
+\hl{The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 }\citep[\hl{full overlap;}][\hl{see Appendix B for calculation details}]{Bhattacharyya}.
 Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
 \hl{When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
-When this coefficient is $>$0.95 both distributions are significantly similar.
-Values in between these two threshold just show the probability of overlap between the distributions but are not conclusive to assess the similarity or differences between the distributions.}
+When this coefficient is $>$0.95, the distributions are significantly similar.
+Values in between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.}
 \hl{To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
@@
 
-\hl{The Bayesian consensus trees, however, performs badly for the Normalised Triplets metric: some parameters combinations, especially when the $M_F$ parameter reaches 75\% missing data, leads to negative values (figure 5).
-Normalised Triplets metric values below 0 means that the placement of some taxa is worth than expected by just randomly placing this taxa in the tree.
+\hl{The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (figure 5).
+A Normalised Triplets metric value below 0 means that the placement of some taxa is worse than expected by just randomly placing this taxon in the tree.
 This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
-In fact, even if clades are conserved (figure 5), the resolutions within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
+Even if clades are conserved (figure 5), the resolution within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
 In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
-These results are in agreement with previous studies that have showed that missing data can be deleterious for recovering ``correct'' topologies, especially for small matrices of 100 characters }\citep{wiensmissing2003}.
-\hl{It is important to note, however, that this effect can be reduced by just increasing the number of characters }\citep{wiensmissing2003}.
+These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters }\citep{wiensmissing2003}.
+\hl{It is important to note, however, that this effect can be reduced by increasing the number of characters }\citep{wiensmissing2003}.
 
@@
 \section{Conclusions}
-\hl{Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data }\citep{Wiens01102005,pattinsonphylogeny2014}.
+% \hl{Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data }\citep{Wiens01102005,pattinsonphylogeny2014}.
+
+\hl{Previous studies have explored the effect of missing morphological data in Total Evidence matrices }\citep{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{The conclusions of theses studies, however, were limited by their empirical approach as well as several aspects of their experimental design (e.g.
+not including fossil taxa, }\citealt{Wiens01102005}\hl{; or focusing on the palaeontological aspect, }\citealt{manosphylogeny2007,pattinsonphylogeny2014}.
 \hl{This study proposes a more in depth approach where missing data is generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
-This allowed use to confirm previous results that missing data can be specially deleterious in small sized matrices }\citep{wiensmissing2003}\hl{ but also relieved a yet unexplored aspect: the crucial importance of coding morphological data for living species.}
-\hl{In fact, missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
+This allowed us to confirm previous results that missing data can be especially problematic in small matrices }\citep{wiensmissing2003}\hl{ but also revealed the crucial importance of coding morphological data for living species.
+Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
 When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
-Therefore we advise that one should focus on having morphological characters coded for a large amount of living taxa present in the matrix (i.e.
+Therefore we advise that one should focus on coding morphological characters for a large number of the living taxa present in the matrix (i.e.
 50\%) for accurately combining both living and fossil species in phylogenies.
-Doing so grants a good overlap of morphological characters between living and fossil taxa, allowing the fossil taxa to be positioned relatively to the living taxa based on their actual shared derived characters rather than on simply available data.}
+Doing so grants a good overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their actual shared derived characters rather than simply on available data.}
 
-\hl{Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
-This has also been observed in empirical data} \cite[\hl{e.g.}][]{Arcila2015131}\hl{ were Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
-This might have an important impact on estimating topologies in the Total Evidence framework since previous studies had to rely either on molecular scaffolds }\citep[\hl{e.g.}][]{slaterphylogenetic2013}\hl{, taxonomic constraints }\citep[\hl{e.g.}][]{slaterphylogenetic2013,beckancient2014}\hl{ or even by fixing the topology }\cite[\hl{e.g.}][]{ronquista2012}.
-\hl{Therefore, we suggest extracting such topological frames from the Bayesian consensus tree if needed.}
+\hl{Additionally, the topology of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+This has also been observed in empirical data} \cite[\hl{e.g.}][]{Arcila2015131}\hl{ where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds }\citep[\hl{e.g.}][]{slaterphylogenetic2013}\hl{, taxonomic constraints }\citep[\hl{e.g.}][]{slaterphylogenetic2013,beckancient2014}\hl{ or even by fixing the topology }\cite[\hl{e.g.}][]{ronquista2012}.
+\hl{Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.}
 

### 2015-08-07 Manuscript/MPESubmission/Revision/response.tex
Updated revisions

@@
 We inferred the topology from the matrices using both Maximum Likelihood and Bayesian inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and for wild-card taxa placement (based on the Triplets distance; Critchlow et al., 1996).
-Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring difference on two aspects of topology (c.f.
+Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring difference on two aspects of topology (cf.
 Wiens et al., 2005; Pattinson et al., 2014 measuring respectively only the clades support or the number of shared clades).
@@
 
-% NC: Hmmm I reckon this is probably inconsistent throughout then as I usually correct it to plural. I'd just change this tbh. Or at least check you've been consistent.
+% NC: Hmmm I reckon this is probably inconsistent throughout then as I usually correct it to plural.
+I'd just change this tbh.
+Or at least check you've been consistent.
+TG: ok maybe be I have a plural/singular problem (I definitely have).
+I actually never know when to use it a singular or plural: "One data point", that's ok, but "morphological data for one taxa"?
+Would that be singular or plural?
 
@@
 You can't just throw random stuff into the appendices for no reason.
+% TO DISCUSS @@@
+% TG: Ok, maybe we should remove it. I've done the analysis because it's interesting and easy to do but maybe we can just add the results to this response, not to the revised manuscript?
 
@@
 % NC: Check this - I'm sure I caught one up above in the comments here.
+TG: I think the idea is that we didn't "infer" any matrices, we simulated/generated/created them.
+The trees are still inferred though.
 
@@
 % NC: I think what is needed here is a definition of what is "too much". Is it 50%? 20%? etc. Too much is arbitrary.
+% TO DISCUSS @@@
+% TG: That's why I removed the "too much".
+I'm actually not sure how Homoplasy is measured (as in units of homoplasy).
+Wagner did a lot of work on it but he usually refers to homoplasy through proxies like characters retention, etc...
 
@@
 We modified the sentence as follows:
-"We selected this model and these parameters to strike a balance between realism for empirical datasets (e.g.
+% "We selected this model and these parameters to strike a balance between realism for empirical datasets (e.g.
 Kelly et al., 2014) and models with few parameters (cf.
@@
 
-% NC: Why not just use the phrasing given by the reviewer for the first part?
+"This model and these parameter settings strike a balance between realism for empirical datasets  (e.g.
+Douady, Kelly et al., 2014) and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)".
+lines @@@
+
+% NC: Why not just use the phrasing given by the reviewer for the first part? TG: Done.
 
@@
 "We used low evolutionary rate parameters to be consistent with the molecular rate parameters and to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal (Wright and Hillis, 2014).
-Note that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increase variance in topological error but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004)." lines 211-217 @@@
+In fact, topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model (Lewis, 2001; Wright and Hillis, 2014).
+Note, however, that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increase variance in topological error but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004)." lines 211-217 @@@
 
@@
 For example mentioning that "Topological error, however, becomes minimal for all scenarios (Parsimony and Mk, with or without missing data) at a rate of 0.5 (see figure 3 from Wright and Hillis (2014)" as you've mentioned above.
+% TG: added a sentence.
 
@@
 We modified the paragraph to:
-
+%Stopped here
 "Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).

### 2015-08-07 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Updated revisions

@@
 Up to now, three main approaches have been used to place both living and fossil taxa into phylogenies.
-\hl{These approaches differ mainly in how they treat the fossil taxa and it's data.
-One can use the fossil as tips or as nodes in the phylogeny and use it's age only, it's morphology only or it's age and morphology jointly.}
+\hl{These approaches differ mainly in how they treat fossil taxa and their data.
+One can use fossils as tips or as nodes in the phylogeny, and can use only the age of the fossils, only the morphology of the fossils, or age and morphology jointly.}
 Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
@@
 Neontologists, on the other hand, more commonly use probabilistic approaches (e.g. Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-\hl{Because fossil taxa do not usually have available DNA, only fossils' occurrence dates are used to time calibrate phylogenies} \citep{zuckerkandl1965}.
+\hl{Because fossil taxa do not usually have available DNA, only fossil occurrence dates are used to time calibrate phylogenies }\citep{zuckerkandl1965}.
 There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the ``best'' approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
@@
 We inferred the topology from the matrices using both Maximum Likelihood and Bayesian inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation } \citep[\hl{based on the Robinson-Foulds distance;}][]{RF1981}\hl{ and for wild-card taxa placement }\citep[\hl{based on the Triplets distance;}][]{critchlowthe1996}.
-\hl{Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring difference on two aspects of topology }\citep[c.f.][\hl{measuring respectively only the clades support or the number of shared clades}]{Wiens01102005,pattinsonphylogeny2014}.
+\hl{Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring difference on two aspects of topology }\citep[cf.][\hl{measuring respectively only the clades support or the number of shared clades}]{Wiens01102005,pattinsonphylogeny2014}.
 %Former revised version
@@
 %We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, total computational time for the whole analysis was around 150 CPU years).
-\hl{We selected this model and parameters settings to be a compromise between realism from empirical datasets} \citep[e.g.][]{kellymolecular2014} \hl{and models with few parameters} \citep[cf.][]{springermacroevolutionary2012} \hl{to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).}
+\hl{This model and these parameter settings strike a balance between realism for empirical datasets }\citep[e.g.][]{douadycomparison2003,kellymolecular2014}\hl{ and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).}
 All the molecular information for fossil taxa was replaced by missing data ("?").
@@
 \hl{We used low evolutionary rate parameters to be consistent with the molecular rates parameters and to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal} \citep{wrightbayesian2014}.
-\hl{Note that} \cite{wrightbayesian2014} \hl{have shown that low morphological rates ($<0.5$) increases variance in topological error but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal} \citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}.
+\hl{In fact, topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model }\citep{lewisa2001,wrightbayesian2014}.
+\hl{Note, however, that } \cite{wrightbayesian2014}\hl{ have shown that low morphological rates (< 0.5) increase variance in topological error but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal }\citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}.
 

### 2015-08-07 Manuscript/MPESubmission/Revision/response.tex
Dealt with comments on reviewer 2's comments

@@
 We modified the paragraph to:
-%Stopped here
+
 "Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
@@
 Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advise that where a topological constraint is needed, Bayesian consensus trees should be used.
-This may apply to tree inferences using the Total-Evidence method such as tip-dating (e.g.
+This may apply to tree inferences using the Total Evidence method such as tip-dating (e.g.
 Ronquist et al., 2012a; Wood et al., 2013; Matzke, 2014).
 It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
-Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies.
+Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies (Swofford et al., 2001).
 As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve poorly supported nodes due to missing data than the Maximum Likelihood tree that is more likely to incorrectly resolve such nodes (i.e.
@@
 Instead I've made it so you're saying IF you need a constraint, use this, but be careful.
+TG: Deadly!
+Thanks!
 
@@
 
-We also linked these results to previous studies (Wiens 2003) and discussed the negative values of the Normalised Triplets metric in the reviewer 1 comment 10 (see above).
+We also linked these results to previous studies (Wiens 2003; Systematic Biology) and discussed the negative values of the Normalised Triplets metric in the reviewer 1 comment 10 (see above).
 
@@
 "The effects of missing data on topological recovery using Bayesian consensus trees.
-Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line).
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $N_{C}$ (lower line).
 The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.
@@
 
-% NC: Sometimes in this you have N_C, sometimes M_C. FIX IT!!!
+% NC: Sometimes in this you have N_C, sometimes M_C. FIX IT!!! TG: Following reviewer 2's comment, it's now N_C rather than M_C.
 
@@
 "Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
 The Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; blue regions in Figure 6A), however, displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; blue regions in Figure 6B).
 The orange upper triangle in Figure 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Figure 6A).
-There is no additional effect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-% NC: This sentence and the one before it seem to be the same thing?
-Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
+% There is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+% Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+% NC: This sentence and the one before it seem to be the same thing? TG: Yep, I modified to:
+Once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
 In figure 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%." lines 480-497@@@
@@
 See \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Analysis_Demo.R} and Appendix D.
-% NC: Maybe add something like this if this is your intention. It's probably worth it. No need to keep lots of un-necessary files.
-We will also clean up this repo once the paper has been accepted. At present it is functioning both as our working repo for code as well as the repo for this paper.
+% NC: Maybe add something like this if this is your intention. It's probably worth it. No need to keep lots of un-necessary files. TG: Agree!
+We will also clean up this repository once the paper has been accepted.
+At present it is functioning both as our working repository for code as well as the repository for this paper.
 
@@
 
-"To explore the effect of missing morphological data on topological recovery, we removed various amounts of the “complete” matrix to obtain matrices with missing morphological data.
-Hereafter, we call these matrices with missing morphological data the “missing-data” matrices.
+"To explore the effect of missing morphological data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
+Hereafter, we call these matrices with missing morphological data the ``missing-data'' matrices.
 % NC: add clarification here about molecular data
-Note that the amount of molecular data remained constant throughout our simulations - XX in living taxa and no molecular data for fossil taxa (see above).
-% NC: Is this correct?
+Note that the amount of molecular data remained constant throughout our simulations: 1000 molecular characters for each living taxa and no molecular data for fossil taxa (see above).
+% NC: Is this correct? TG: yap ;).
 We removed morphological data using three data incompleteness parameters:
-1.
-The proportion of missing living taxa ($M_L$).
-This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
+\begin{enumerate}
+\item{The proportion of missing living taxa ($M_L$).} This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
 It represents the number of living taxa that are present in the matrix but have only molecular data available.
@@
 In practice, we removed all the morphological data from randomly chosen living taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.
-2.
-The proportion of missing data in the fossil record ($M_F$).
-This missing data parameter represents the completeness of the fossil record.
+\item{The proportion of missing data in the fossil record ($M_F$).} This missing data parameter represents the completeness of the fossil record.
 Due to preservation biases, missing data for fossil taxa are common (Sansom and Wills, 2013).
 In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
-3.
-The number of morphological characters for both living and fossil taxa ($N_C$).
-This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
+%Added this part following comment 4.
+Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
+Instead this 50\% refers to the whole matrix.
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a minimum of 5\%, see below).
+\item{The number of morphological characters for both living and fossil taxa ($N_C$).} This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
 Any morphological matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
@@
 In practice, we randomly removed whole characters from the morphological matrix reducing it to: 100, 90,75, 50 or 25 characters.
-Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of “missing” morphological characters).
+Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).
+\end{enumerate}
 
@@
 
-% NC: Ah OK. You haven't corrected this everywhere in these comments I don't think...
+% NC: Ah OK. You haven't corrected this everywhere in these comments I don't think... TG: Should be good now.
 
@@
 % NC: OK - I'd add a line somewhere in the manuscript to make this clear. something like:
-
-%"Note that 55\% missing data for fossil taxa does not mean that each fossil is missing 50\% of it's morphological data.
+% TG: I've added the following to the new removing data section:
+
+We added the following precision to the explanation on removing data (See our response to reviewer 2's comment 1):
+"Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
 Instead this 50\% refers to the whole matrix.
-Some fossils may reatin 100% of their data and others may lose all their data at this parameter value"
-
-Also, a second argument for using this slightly simplistic way to deal with the missing data in the fossil record follows our main recommendation from this study: the best way to improve the quality of Total Evidence trees is to have a maximum number of living species with available morphological data.
-In fact lines 584 to 589 in the first version of the manuscript (line 669-674@@@ in the revised version), we argue that even if the effect of the $M_F$ parameter is important, there is unfortunately not much we can do about it since the completeness of this part of the matrix solely depends on exceptional palaeontological discoveries and excavation effort.
-We feel that these parameters, however crucial for our understanding of macroevolution, are unfortunately not predicable and a recommendation such as "we need more fossil data" is less practical than our main recommendation: "we need more living species data".
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a minimum of 5\%, see below)."
+
+% Also, a second argument for using this slightly simplistic way to deal with the missing data in the fossil record follows our main recommendation from this study: the best way to improve the quality of Total Evidence trees is to have a maximum number of living species with available morphological data.
+% In fact lines 584 to 589 in the first version of the manuscript (line 669-674@@@ in the revised version), we argue that even if the effect of the $M_F$ parameter is important, there is unfortunately not much we can do about it since the completeness of this part of the matrix solely depends on exceptional palaeontological discoveries and excavation effort.
+% We feel that these parameters, however crucial for our understanding of macroevolution, are unfortunately not predicable and a recommendation such as "we need more fossil data" is less practical than our main recommendation: "we need more living species data".
 

### 2015-08-07 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Dealt with comments on reviewer 2's comments

@@
 % TG: this part is entirely re-written.
-\hl{To explore the effect of missing data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
+\hl{To explore the effect of missing morphological data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
 Hereafter, we call these matrices with missing morphological data the ``missing-data'' matrices.
-We removed morphological following three main data incompleteness parameters:}
+Note that the amount of molecular data remained constant throughout our simulations: 1000 molecular characters for each living taxa and no molecular data for fossil taxa (see above).
+We removed morphological data using three data incompleteness parameters:}
 \begin{enumerate}
-\item{\hl{The proportion of missing living taxa ($M_{L}$).}}
+\item{\hl{The proportion of missing living taxa ($M_L$).}}
 \hl{This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
 It represents the number of living taxa that are present in the matrix but have only molecular data available.
-This reflects the fact that, because of the increasing facility of collecting molecular data, morphological data for living species is rarely collected} \citep{GuillermeCooperMissing}.
-% TG: see what I did here?
-;)
+This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species is rarely collected }\citep{GuillermeCooperMissing}.
 \hl{Therefore, many living species will have only molecular data available.
-In practice, we removed all the morphological data from randomly chosen living with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.}
-\item{\hl{The proportion of missing data in the fossil record ($M_{F}$).}}
+In practice, we removed all the morphological data from randomly chosen living taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.}
+\item{\hl{The proportion of missing data in the fossil record ($M_F$).}}
 \hl{This missing data parameter represents the completeness of the fossil record.
-In fact, due to preservation biases, missing data for fossil taxa are common }\citep{sansomfossilization2013}. 
-\hl{In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.}
-\item{\hl{The number of morphological characters for both living and fossil taxa ($N_{C}$).}}
+Due to preservation biases, missing data for fossil taxa are common }\citep{sansomfossilization2013}.
+\hl{In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
+Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
+Instead this 50\% refers to the whole matrix.
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a minimum of 5\%, see below).}
+\item{\hl{TThe number of morphological characters for both living and fossil taxa ($N_C$).}}
 \hl{This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
 Any morphological matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
-% TG: this is word for word what the reviewer wrote.
-I didn't like doing that first but after seeing the Ineterface Focus review and how they copy pasted the discussion suggestion, I think it's really nice for the reviewer to see a bit of it's work directly in the paper (at least it was for me!).
-Therefore, this parameter correspond to the overall number of characters available for both living and fossil taxa.
-In practice, we randomly removed whole characters from the morphological matrix shrinking it to: 100, 90, 75, 50 or 25 characters.
+Therefore, this parameter corresponds to the overall number of characters available for both living and fossil taxa.
+In practice, we randomly removed whole characters from the morphological matrix reducing it to: 100 (i.e. 0 characters removed), 90,75, 50 or 25 characters.
 Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).}
@@
 
-% TG: this part is nearly exactly the same as before.
-In practice, each parameter represents a different way of removing data from the \hl{morphological part of} the matrix: $M_L$ removes \hl{entire} rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $N_C$ removes columns across both living and fossil taxa's data.
+\hl{In practice, each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
 Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
 
-% TG: and this one is slightly modified only.
-\hl{We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data''matrices.
-Note that one of these combinations ($M_{L}$=0\%; $M_{F}$=0\% and $N_{C}$=100) has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
+We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data'' matrices.
+Note that one of these combinations ($M_L$=0\%; $M_F$=0\% and $N_C$=100) has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
 In practice, we first removed the data following the two missing data parameters $M_L$ and $M_F$ and then removed data following the $N_C$ parameters.
@@
 In real world situations, the ``true'' tree is never available to us because we cannot know the true evolutionary history of a clade \citep[except in very rare circumstances, e.g.][]{rozen2005}.
-\hl{We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
-In practice, the difference between the ``best'' trees and the ``missing data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing data'' trees.
-The difference between the ``true'' and the ``missing data'' tree, however, represents the effect of our parameters used to generate the ``true'' tree and the algorithms used to generated the ``complete'' matrix as well as the effect of our missing data parameters and the phylogenetic methods used.
-Because the main aim of this study, is to look at the effect the missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees to ``missing data'' trees.
-The results of the comparisons of the ``true'' and the ``missing data'' tree are available in appendix B and C.
+\hl{We compare ``best'' trees to ``missing-data'' trees but could also compare ``true'' trees to the ``missing-data'' trees.
+In practice, the difference between the ``best'' trees and the ``missing-data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing-data'' trees.
+The difference between the ``true'' and the ``missing-data'' tree, however, represents the effect of our parameters used to generate the ``true'' tree and the algorithms used to generated the ``complete'' matrix as well as the effect of our missing data parameters and the phylogenetic methods used.
+Because the main aim of this study, is to look at the effect the missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees to ``missing-data'' trees.
+The results of the comparisons of the ``true'' and the ``missing-data'' tree are available in appendix B and C.
 Note that these results makes little difference to our overall results.}
@@
 Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
-\hl{Because of the difficulties to represent so many pairwise comparisons in a meaningful way, we summarized as a heat map of Bhattacharyya Coefficients (see Figure 6).
-In this type of figures, parameters that have similar or different effects on recovering the ``best'' topology (rather good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6).}
+\hl{Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Figure 6).
+In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6).}
 
@@
 
-\hl{Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+\hl{UUsing both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
@@
 The orange upper triangle in Figure 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Figure 6A).
-There is no additional effect (i.e. no dissimilarities) of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
+Once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
 Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
-This can be interpreted as, in figure 5 for the Normalised Robinson-Foulds metric, the overlap between the distributions once $M_L$=50\%.}
+In figure 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.}
 
@@
 \hl{Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
-For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combinations of the other parameters ($M_F$ and $N_C$; Figure 6).
-In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) doesn't matter and and increase/decrease in both does not affect topology.
+For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combination of the other parameters ($M_F$ and $N_C$; Figure 6).
+In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) does not matter and an increase/decrease in both does not affect topology.
 A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
-This has important practical implications, especially on the best strategy to improve topology by collecting morphological characters (see below).}
+This has important practical implications, especially for the best strategy to improve topology by collecting morphological characters (see below).}
 
@@
 
+% \hl{Another practical implication of our results regards the tree inference methods.
+% Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advice using the Bayesian consensus trees as a topological constraint for tree inferences using the Total-Evidence method such as tip-dating} \citep[\hl{e.g.}][\hl{; although it is possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution}]{ronquista2012,Wood01032013,BEASTmaster}.
+% \hl{Additionally, using the Bayesian consensus tree rather than the Maximum Likelihood can reduce the amount of ``false positive'' topologies.
+% As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve nodes weakly supported due to missing data than the Maximum Likelihood tree that is more likely to wrongly resolve such nodes (i.e.
+creating a ``false positive'' node).
+% Note, however, that we do not suggest discarding the Bayesian posterior tree distribution once one chose to use the Bayesian consensus tree topology even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
+% This can be particularly important because these trees will be invaluable for phylogenetic comparative analyses.
+% For example a sub-sample of posterior trees distributions can be used to asses macroecological questions while better taking into account topological uncertainty (e.g.
+}\citealt{FritzTree}\hl{ and }\citealt{jetzthe2012}\hl{ trees used in }\citealt{healy2014}\hl{).}
 \hl{Another practical implication of our results regards the tree inference methods.
-Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advice using the Bayesian consensus trees as a topological constraint for tree inferences using the Total-Evidence method such as tip-dating} \citep[\hl{e.g.}][\hl{; although it is possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution}]{ronquista2012,Wood01032013,BEASTmaster}.
-\hl{Additionally, using the Bayesian consensus tree rather than the Maximum Likelihood can reduce the amount of ``false positive'' topologies.
-As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve nodes weakly supported due to missing data than the Maximum Likelihood tree that is more likely to wrongly resolve such nodes (i.e.
-creating a ``false positive'' node).
-Note, however, that we do not suggest discarding the Bayesian posterior tree distribution once one chose to use the Bayesian consensus tree topology even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
-This can be particularly important because these trees will be invaluable for phylogenetic comparative analyses.
-For example a sub-sample of posterior trees distributions can be used to asses macroecological questions while better taking into account topological uncertainty (e.g.
+Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advise that where a topological constraint is needed, Bayesian consensus trees should be used.
+This may apply to tree inferences using the Total Evidence method such as tip-dating }\citep[e.g.][]{ronquista2012,Wood01032013,BEASTmaster}.
+\hl{It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
+Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies }\citep{Swofford2001}.
+\hl{As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve poorly supported nodes due to missing data than the Maximum Likelihood tree that is more likely to incorrectly resolve such nodes (i.e.
+creating a false positive node).
+Note, however, that we do not suggest discarding the Bayesian posterior tree distributions even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
+These trees will be invaluable for phylogenetic comparative analyses.
+For example a sub-sample of posterior trees distributions can be used to assess macroecological questions while better taking into account topological uncertainty (e.g.
 }\citealt{FritzTree}\hl{ and }\citealt{jetzthe2012}\hl{ trees used in }\citealt{healy2014}\hl{).}
@@
 % \hl{Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data }\citep{Wiens01102005,pattinsonphylogeny2014}.
-
 \hl{Previous studies have explored the effect of missing morphological data in Total Evidence matrices }\citep{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
@@
 \textbf{Figure 6:} \hl{The effects of missing data on topological recovery using Bayesian consensus trees.
-Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the two parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and number of characters from 100 to 25 for the parameter $N_{C}$ (lower line).
+Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $N_{C}$ (lower line).
 The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to a high probability of overlap in orange.
-Blue regions denotes a poor overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a strong effect on the Normalised metric and thus the topological recovery).
-Conversely, orange regions denotes a high overlap in Normalised metric between the different parameters combinations (i.e.
-the parameters have a weak effect on the Normalised metric and thus the topological recovery)}.
+The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to high probability of overlap in orange.
+Blue regions denote a poor overlap in Normalised metric values between the different parameter combinations (i.e.
+the parameters have a strong effect on the metric and thus the topological recovery).
+Conversely, orange regions denote a high overlap in Normalised metric values between the different parameter combinations (i.e.
+the parameters have a weak effect on the metric and thus the topological recovery)}.
 
@@
 \textbf{Table 1:} \hl{Bhattacharyya Coefficients of the pairwise method comparisons.}
-\hl{Each comparisons corresponds to the comparison of the full (125) distributions between the ``best'' tree and the ``missing data'' trees for all the four methods (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) and the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
-Each line summarizes the distribution of the probability of overlap between pairs of tree inference methods.
-The values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
-If two methods have a high probability of overlap, they have a similar ability to recover the ``correct" tree topology.}\\
+\hlEach line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
+Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
+If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology.}\\
 

### 2015-08-09 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
Appendices figures/tables notations corrected

@@
 
-\renewcommand\thefigure{A.\arabic{figure}}
-\renewcommand\thetable{A.\arabic{table}}
+\renewcommand\thefigure{A\arabic{figure}}
+\renewcommand\thetable{A\arabic{table}}
 
@@
 \subsection{Effect of the starting tree on Bayesian inference}
-We used the ``true'' tree as a starting tree for the two MCMC chains in the Bayesian inference for decreasing the computing time.
-To assess if this had an effect on the topology of the ``best'' tree, we ran a sub-sample of trees using a different random tree for the two MCMC chains \citep[default MrBayes option;][]{Ronquist2012mrbayes}.
+We used the ``true'' tree as a starting tree for the two MCMC chains in the Bayesian inference to decrease the computing time needed.
+To assess if this had an effect on the topology of the ``best'' tree, we ran a sub-sample of trees using a different random starting tree for the two MCMC chains \citep[default MrBayes option;][]{Ronquist2012mrbayes}.
 We tested this effect on five trees with the five levels of missing data (i.e.
@@
 We then compared the trees inferred using a random starting tree to the ``best'' tree using the normalised Robinson-Foulds and Triplets metrics in an identical way as described in the main text (figure ~\ref{Fig_startingTree}).
-We then compared the normalised Robinson-Foulds and Triplets metrics of the ``missing-data'' trees using the ``true'' tree as a starting tree or a random tree as a starting tree using simple linear regression and nested linear regression (\texttt{lm(metric $\sim$ starting tree)}) and \texttt{lmList(metric $\sim$ starting tree | missing data}) in \texttt{R}).
+We also compared the normalised Robinson-Foulds and Triplets metrics of the ``missing-data'' trees using the ``true'' tree as a starting tree or a random tree as a starting tree using simple linear regression and nested linear regression (\texttt{lm(metric $\sim$ starting tree)}) and \texttt{lmList(metric $\sim$ starting tree | missing data}) in \texttt{R}).
+% NC: Why the regressions? What are they telling you? Why simple and nested?
 The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Starting_tree.R}.
-We found no significant effect of using the ``true'' or a random tree as a starting tree on the ability to recover the ``best'' tree (tables ~\ref{Tab_lm_res} and ~\ref{Tab_lmList_res}).
+We found no significant effect of using the ``true'' instead of a random tree as a starting tree on our ability to recover the ``best'' tree (tables ~\ref{Tab_lm_res} and ~\ref{Tab_lmList_res}).
 
 \begin{table}[ht]
-\caption{Effect of the using the ``true'' tree as a starting tree on topological recovery for the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics using simple linear regression.}
+\caption{Effect of the using the ``true'' tree as a starting tree on topological recovery for the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics using simple linear regression.} % NC: I'm not sure what these two tables are telling you.
+Clarify.
+The caption says nothing about the random starting trees.
 \label{Tab_lm_res}
@@
 \includegraphics[keepaspectratio=true, width=\textwidth]{SupplementaryFigures/Starting_tree.pdf}
-\caption{Effect of using the ``true'' tree (black) or a random tree (red) as the starting tree for the Bayesian inference. The x axis, represent the levels of missing data ()}
+\caption{Effect of using the ``true'' tree (black) or a random tree (red) as the starting tree for the Bayesian inference.
+The x axis, represents the amount of missing data ()} % NC: What's missing from the brackets here or are they a typo?
 \label{Fig_startingTree}
@@
 
-\subsection{Morphological rates estimations results}
-To test the sensitivity of the Bayesian prior on the morphological rates distribution shape ($\alpha$ = 0.5), we compared the shape of the distribution of our generated morphological characters ($\alpha_{TRUE}$) to the shape of the distribution of morphological characters rate estimated by the Bayesian inference ($\alpha_{EST}$).
-The $\alpha_{TRUE}$ parameter was estimated by fitting a gamma distribution to the 100 individual morphological rates (randomly sampled from a gamma distribution with $\alpha$ = 0.5) using the \texttt{fitdistr} function from the \texttt{MASS R} package \citep{MASS}.
+\subsection{Morphological rates estimates results}
+To test the sensitivity of the Bayesian prior on the morphological rates distribution shape ($\alpha$ = 0.5), we compared the shape of the distribution of our generated morphological characters ($\alpha_{TRUE}$) to the shape of the distribution of morphological character rates estimated by the Bayesian inference ($\alpha_{EST}$).
+The $\alpha_{TRUE}$ parameter was estimated by fitting a gamma distribution to the 100 individual morphological rates (randomly sampled from a gamma distribution with $\alpha$ = 0.5) using the \texttt{fitdistr} function from the \texttt{MASS} R package \citep{MASS}.
 We measured the difference as:
@@
 \end{equation}
-We then calculated the mode and the 95\% and 50\% confidence intervals over the 50 chains using the \texttt{hdr} function from the \texttt{hdrcde R} package \citep{hdrcde}.
+We then calculated the mode and the 95\% and 50\% confidence intervals over the 50 chains using the \texttt{hdr} function from the \texttt{hdrcde} R package \citep{hdrcde}.
 The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Rates_estimates.R}.

### 2015-08-09 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixB.tex
Added appendix tables and giros prefix
Appendices figures/tables notations corrected

--- 2015-07-27 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixB.tex
@@
 \setlength{\parindent}{0.5in}
+
+\renewcommand\thefigure{B\arabic{figure}}
+\renewcommand\thetable{B\arabic{table}}
+
 %\setcounter{secnumdepth}{0} 
@@
 
+\subsection{Differences between the ``true'' and the ``best'' trees.}
+% NC: Need a line or two to explain the motivation of this analysis. Why did you do it? (not just because reviewers asked you to!)
+We compared our ``true'' trees (i.e.
+the trees \textbf{used to create} the ``complete'' matrices) to the ``best'' trees (i.e.
+the trees \textbf{inferred from} % NC: I thought we weren't using the word inferred
+ the ``complete'' matrices) in figure ~\ref{Fig_Supp_True_Best}.
+Note that the difference between the ``true'' and the ``best'' trees represents the effect of the parameters choice and the algorithms used to create the ``complete'' matrix as well the as the capacity of RAxML and MrBayes to infer % NC: again
+phylogenies from this particular matrices (i.e. small sized and generated using specific algorithms).
+This does not affect, however, the results of our analysis since we deliberately compared the the ``missing-data'' trees to the ``best'' tree rather than to the ``true'' tree.
+
+\begin{figure} 
+\centering
+    \includegraphics[width=1\textwidth]{SupplementaryFigures/True_vs_Best_trees.pdf}
+    \caption{Pairwise comparisons among the 50 ``true'' trees and the 50 ``best'' trees from the Maximum Likelihood and Bayesian inference methods.
+The horizontal blue and red lines represent, respectively, the 95\% and 50\% confidence intervals.}
+\label{Fig_Supp_True_Best} 
+\end{figure}
+
 \bibliographystyle{sysbio}

### 2015-08-09 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixC.tex
Appendices figures/tables notations corrected

@@
 
-\renewcommand\thefigure{C.\arabic{figure}}
-\renewcommand\thetable{C.\arabic{table}}
+\renewcommand\thefigure{C\arabic{figure}}
+\renewcommand\thetable{C\arabic{table}}
 
@@
 For clarity, in the paper, we focused on the results of the effects of our missing data parameters on the Maximum Likelihood trees topology and the Bayesian consensus trees topology.
-The following additional results presented here give a greater insight into the effect of our missing data parameters on the Maximum Likelihood Bootstrapped trees topologies and the Bayesian posterior trees distribution topologies.
-Also we present here the pairwise comparisons between each parameters states for the Maximum Likelihood tree topology, the Maximum Likelihood Bootstrapped trees topologies and the Bayesian posterior trees distribution topologies.
+The following additional results presented here give a greater insight into the effect of our missing data parameters on the Maximum Likelihood bootstrapped tree topologies and the Bayesian posterior tree distribution topologies.
+Also we present here the pairwise comparisons between each parameters states for the Maximum Likelihood tree topology, the Maximum Likelihood bootstrapped trees topologies and the Bayesian posterior trees distribution topologies.
 

### 2015-08-10 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
dealt with comments in the supplementaries

@@
 We then compared the trees inferred using a random starting tree to the ``best'' tree using the normalised Robinson-Foulds and Triplets metrics in an identical way as described in the main text (figure ~\ref{Fig_startingTree}).
-We also compared the normalised Robinson-Foulds and Triplets metrics of the ``missing-data'' trees using the ``true'' tree as a starting tree or a random tree as a starting tree using simple linear regression and nested linear regression (\texttt{lm(metric $\sim$ starting tree)}) and \texttt{lmList(metric $\sim$ starting tree | missing data}) in \texttt{R}).
+We used a two-way ANOVA to test any significant effect of the starting tree (``true'' or random) on the normalised Robinson-Foulds and Triplets metrics.
 % NC: Why the regressions? What are they telling you? Why simple and nested?
+% TG: Ok, made it totally plain easier. Anova.
 The code used for this section is available at \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Starting_tree.R}.
-We found no significant effect of using the ``true'' instead of a random tree as a starting tree on our ability to recover the ``best'' tree (tables ~\ref{Tab_lm_res} and ~\ref{Tab_lmList_res}).
+We found no significant effect of using the ``true'' instead of a random tree as a starting tree on our ability to recover the ``best'' tree (table ~\ref{Tab_lm_res}).
 
 \begin{table}[ht]
-\caption{Effect of the using the ``true'' tree as a starting tree on topological recovery for the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics using simple linear regression.} % NC: I'm not sure what these two tables are telling you.
-Clarify.
-The caption says nothing about the random starting trees.
+\caption{Test of the effect of using either a random tree or the ``true'' tree as a starting tree on two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics using a two-way ANOVA.}
 \label{Tab_lm_res}
 \centering
-\begin{tabular}{rlcccc}
+\begin{tabular}{rlcccccc}
   \hline
- & & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
+ metric & terms & Df & Sum Sq & Mean Sq & F value & Pr($>$F) \\ 
   \hline
-RF & (Intercept) & 0.8100 & 0.0122 & 66.16 & \textbf{0.0000} \\ 
-   & starting:true & 0.0019 & 0.0173 & 0.11 & 0.9125 \\
-Tr & (Intercept) & 0.4424 & 0.0418 & 10.59 & \textbf{0.0000} \\ 
-   & starting:true & 0.0159 & 0.0591 & 0.27 & 0.7887 \\ 
-   \hline
-\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-\caption{Effect of the using the ``true'' tree as a starting tree on topological recovery for the two Normalised Robins-Foulds (RF) and Triplets (Tr) metrics for each missing data parameter using nested linear regression.}
-\label{Tab_lmList_res}
-\centering
-\begin{tabular}{rccccc}
-  \hline
- & & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
-  \hline
-RF & 0\%  & 0.02 & 0.02 & 1.02 & 0.31 \\ 
-   & 10\% & -0.01 & 0.02 & -0.50 & 0.62 \\ 
-   & 25\% & 0.00 & 0.02 & 0.00 & 1.00 \\ 
-   & 50\% & 0.00 & 0.02 & 0.03 & 0.98 \\ 
-   & 75\% & -0.00 & 0.02 & -0.03 & 0.98 \\
-Tr & 0\%  & 0.09 & 0.08 & 1.06 & 0.29 \\ 
-   & 10\% & -0.00 & 0.08 & -0.05 & 0.96 \\ 
-   & 25\% & 0.00 & 0.08 & 0.01 & 0.99 \\ 
-   & 50\% & -0.00 & 0.08 & -0.04 & 0.97 \\ 
-   & 75\% & -0.00 & 0.08 & -0.00 & 1.00 \\ 
+RF & starting    & 1 & 0.00 & 0.00 & 0.01 & 0.9125 \\ 
+   & Residuals   & 198 & 2.97 & 0.01 &  &  \\ 
+Tr & starting    & 1 & 0.01 & 0.01 & 0.07 & 0.7887 \\ 
+   & Residuals   & 198 & 34.57 & 0.17 &  &  \\ 
    \hline

### 2015-08-10 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixB.tex
dealt with comments in the supplementaries

@@
 % NC: Need a line or two to explain the motivation of this analysis. Why did you do it? (not just because reviewers asked you to!)
-We compared our ``true'' trees (i.e.
+% TG:  How's that?
+In our simulation protocol, we used the ``true'' tree to generate the molecular characters and the morphological characters for the living and fossil taxa (i.e. the ``complete'' matrix).
+The following analysis measures the performance of our parameter and algorithms choices to generate the ``complete'' matrix.
+%
+To asses the performance of our simulation protocol, we compared our ``true'' trees (i.e.
 the trees \textbf{used to create} the ``complete'' matrices) to the ``best'' trees (i.e.
-the trees \textbf{inferred from} % NC: I thought we weren't using the word inferred
+the trees \textbf{inferred from} % NC: I thought we weren't using the word inferred TG: no, it's just the matrices that are not inferred anymore.
  the ``complete'' matrices) in figure ~\ref{Fig_Supp_True_Best}.
-Note that the difference between the ``true'' and the ``best'' trees represents the effect of the parameters choice and the algorithms used to create the ``complete'' matrix as well the as the capacity of RAxML and MrBayes to infer % NC: again
-phylogenies from this particular matrices (i.e. small sized and generated using specific algorithms).
+Note that the difference between the ``true'' and the ``best'' trees represents the effect of the parameters choice and the algorithms used to create the ``complete'' matrix as well the as the capacity of RAxML and MrBayes to infer phylogenies from this particular matrices (i.e.
+small sized and generated using specific algorithms).
 This does not affect, however, the results of our analysis since we deliberately compared the the ``missing-data'' trees to the ``best'' tree rather than to the ``true'' tree.

### 2015-08-10 Analysis/Starting_tree.R

@@ -96,26 +96,13 @@ best_RF_test<-data.frame("RF"=c(True_RF, Rand_RF), "starting"=as.factor(c(rep("t
 best_Tr_test<-data.frame("Tr"=c(True_Tr, Rand_Tr), "starting"=as.factor(c(rep("true", 100), rep("rand", 100))), "miss.data"=rep(trees,2))
 #Comparisons to best true
 
-# True_RF<-c(True_to_true[[1]][[4]])
-# True_Tr<-c(True_to_true[[1]][[6]])
-# Rand_RF<-c(Rand_to_true[[1]][[4]], mean(Rand_to_true[[2]][[4]])) #One value is missing
-# Rand_Tr<-c(Rand_to_true[[1]][[6]], mean(Rand_to_true[[2]][[6]])) #One value is missing
-# for (tree in 2:length(True_to_true)) {
-#     True_RF<-c(True_RF, True_to_true[[tree]][[4]])
-#     True_Tr<-c(True_Tr, True_to_true[[tree]][[6]])
-#     Rand_RF<-c(Rand_RF, Rand_to_true[[tree]][[4]])
-#     Rand_Tr<-c(Rand_Tr, Rand_to_true[[tree]][[6]])
-# }
-# true_RF_test<-data.frame("RF"=c(True_RF, Rand_RF), "starting"=as.factor(c(rep("true", 100), rep("rand", 100))), "miss.data"=rep(trees,2))
-# true_Tr_test<-data.frame("Tr"=c(True_Tr, Rand_Tr), "starting"=as.factor(c(rep("true", 100), rep("rand", 100))), "miss.data"=rep(trees,2))
-
 
 #Running the tests
 library(xtable)
 library(lme4)
 #General effect
-xtable(summary(lm(RF~starting, data=best_RF_test)))
-xtable(summary(lm(Tr~starting, data=best_Tr_test)))
+xtable(anova(lm(RF~starting, data=best_RF_test)))
+xtable(anova(lm(Tr~starting, data=best_Tr_test)))
 
 
 #Nested effect

### 2015-08-10 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Unified figures (Fig.) and tables (Table)

@@
 \section{Materials and Methods}
-To explore how missing data in \hl{the morphological partition of Total Evidence matrices} influences tree topology, \hl{we used the following protocol (Figure 1):}
+To explore how missing data in \hl{the morphological partition of Total Evidence matrices} influences tree topology, \hl{we used the following protocol (Fig. 1):}
 \begin{enumerate}
@@
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $N_{C}$ = 100 etc.
-(see Figure 2 for more details).}
+(see Fig.
+2 for more details).}
 This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
@@
 Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
-\hl{Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Figure 6).
-In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6).}
+\hl{Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Fig. 6).
+In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Fig.
+6).}
 
@@
 Maximum Likelihood \textit{versus} Bayesian consensus; Maximum Likelihood \textit{versus} Maximum Likelihood bootstrap trees etc.
-(see Figure 3 for more details).
+(see Fig.
+3 for more details).
 Note that this procedure pools results from across all missing data parameter combinations so it results in just six pairwise comparisons.
@@
 As the amount of missing data in the morphological part of the Total Evidence matrix increases, our ability to recover the ``best'' tree topology decreases, regardless of the missing data parameter ($M_{L}$, $M_{F}$ or $N_{C}$), the tree inference method (Maximum Likelihood or Bayesian) or the tree comparison metric used (Normalised Robinson-Foulds or Normalised Triplets metric).
-Nonetheless, the different missing data parameters and tree inference methods do not affect the topology in the same way (Figure 4 and Figure 5).
+Nonetheless, the different missing data parameters and tree inference methods do not affect the topology in the same way (Fig. 4 and Fig. 5).
 
@@
 \subsection{Individual effects of missing data parameters}
-As the amount of missing data increases across all three parameters, our ability to recover the ``best'' tree topology decreases (Figure 4).
-The Normalised Robinson-Foulds metric is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $N_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
-The Normalised Triplets metric, however, is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $N_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $N_{C}$ respectively; Figure 4; Tables C5, C6 and C7 in Appendix C).
+As the amount of missing data increases across all three parameters, our ability to recover the ``best'' tree topology decreases (Fig. 4).
+The Normalised Robinson-Foulds metric is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $N_{C}$ respectively; Fig.
+4; Tables C5, C6 and C7 in Appendix C).
+The Normalised Triplets metric, however, is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $N_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $N_{C}$ respectively; Fig.
+4; Tables C5, C6 and C7 in Appendix C).
 
@@
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $N_{C}$ = 75\%), and best when there is no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 0\%; Figure 5; Tables C2, C3 and C4 in Appendix C).
-Figure 6 shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0).
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 0\%; Fig.
+5; Tables C2, C3 and C4 in Appendix C).
+Fig.
+6 shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0).
 
@@
 
-\hl{UUsing both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
-The Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; blue regions in Figure 6A), however, displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; blue regions in Figure 6B).
-The orange upper triangle in Figure 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Figure 6A).
-Once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
-In figure 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.}
+\hl{Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Fig.
+6).
+The Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; blue regions in Fig.
+6A), however, displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; blue regions in Fig.
+6B).
+The orange upper triangle in Fig. 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Fig. 6A).
+Once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Fig. 6A).
+Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Fig.
+6A).
+In Fig. 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.}
 
 For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds metric or Normalised Triplets metric respectively; Table ~\ref{Tab_Results-Difference_methods}).
-These two methods, however, perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds metric (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Figure 4 and Figure C2 in Appendix C).
+These two methods, however, perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds metric (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.
+4 and Fig.
+C2 in Appendix C).
 
@@
 These factors, however, affected topological recovery in different ways and to different extents.
-Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($N_{C}$) had worst effects on topological recovery (Figure 6).
-Additionally, using Bayesian consensus trees recovered the ``best'' tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Figure 5, Figure 6, Table ~\ref{Tab_Results-Difference_methods}).
+Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($N_{C}$) had worst effects on topological recovery (Fig. 6).
+Additionally, using Bayesian consensus trees recovered the ``best'' tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.
+5, Fig.
+6, Table ~\ref{Tab_Results-Difference_methods}).
 As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
@@
 
-The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds metric) but greater on the displacement of wildcard taxa (Normalised Triplets metric; Figure 4 and Figure 5).
+The effect of the missing data in the fossil record ($M_{F}$) is less than the effect of the $M_{L}$ parameter on clade conservation (Normalised Robinson-Foulds metric) but greater on the displacement of wildcard taxa (Normalised Triplets metric; Fig.
+4 and Fig.
+5).
 This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
@@
 Conversely, removing 10\% of morphological characters (\hl{i.e. $N_{C}$ = 90}) genuinely removes 10\% of data in the matrix.
-Nonetheless, the effect of removing characters on the ability to recover the ``best'' tree topology is of the same order of magnitude as for the other two parameters (Figure 4).
+Nonetheless, the effect of removing characters on the ability to recover the ``best'' tree topology is of the same order of magnitude as for the other two parameters (Fig. 4).
 We suspect this again reflects the importance of overlapping characters, as opposed to the number of characters \textit{per se}.
@@
 \hl{The sizes of our data matrices were constrained by the performance of our protocol}: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
-Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Figure 4 and Figure 5).
+Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.
+4 and Fig.
+5).
 Although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}.
@@
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) and lower when \hl{they have a larger proportion of missing data} (i.e.
-$M_{L}$ = 75\%, $M_{F}$ = 75\% and $N_{C}$ = 25; Figure 5).
+$M_{L}$ = 75\%, $M_{F}$ = 75\% and $N_{C}$ = 25; Fig.
+5).
 It is important, however, to notice that the effect of each parameter is not additive.
 \hl{Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
-For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combination of the other parameters ($M_F$ and $N_C$; Figure 6).
+For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combination of the other parameters ($M_F$ and $N_C$; Fig. 6).
 In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) does not matter and an increase/decrease in both does not affect topology.
-A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
+A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Fig. 6).
 This has important practical implications, especially for the best strategy to improve topology by collecting morphological characters (see below).}
@@
 \subsection{Effects of tree inference methods}
-Variation in our ability to recover the ``best'' tree topology depends heavily on the tree inference method (Figure 4 and Figure 5).
+Variation in our ability to recover the ``best'' tree topology depends heavily on the tree inference method (Fig. 4 and Fig. 5).
 For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over \hl{parsimony} methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
 This is, however, the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best'' tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-\hl{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
+\hl{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
 Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'') for details on the analysis of the effect of the starting tree in Bayesian inference).}
@@
 
-\hl{The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (figure 5).
+\hl{The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (Fig.
+5).
 A Normalised Triplets metric value below 0 means that the placement of some taxa is worse than expected by just randomly placing this taxon in the tree.
 This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
-Even if clades are conserved (figure 5), the resolution within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
+Even if clades are conserved (Fig. 5), the resolution within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
 In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
@@
 It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best'' tree topology than Maximum Likelihood and Bayesian consensus trees.
-This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $N_{C}$ = 100; Figure 4).
+This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $N_{C}$ = 100; Fig. 4).
 This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best'' trees involved 1000 random pairwise comparisons rather than just one.
-\hl{Additionally, the Bayesian posterior trees performed way worth than the Bayesian consensus tree (figure 4, table 1 and appendix C figure C5 and tables C5, C6 and C7).
+\hl{Additionally, the Bayesian posterior trees performed way worth than the Bayesian consensus tree (Fig. 4, Table 1 and appendix C Fig. C5 and Tables C5, C6 and C7).
 This can be due to the fact that the Bayesian posterior trees are always resolved and thus more likely to contain wrongly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
@@
 Gaps in the matrix, however, can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
-Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the ``best'' tree topology (Figure 6).
+Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the ``best'' tree topology (Fig. 6).
 Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
@@
 Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies }\citep{Swofford2001}.
-\hl{As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve poorly supported nodes due to missing data than the Maximum Likelihood tree that is more likely to incorrectly resolve such nodes (i.e.
+\hl{As shown in Fig.
+5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve poorly supported nodes due to missing data than the Maximum Likelihood tree that is more likely to incorrectly resolve such nodes (i.e.
 creating a false positive node).

### 2015-08-10 Manuscript/MPESubmission/Revision/response.tex
Outlined points to discuss for tomorrow.
Comments on revisions

@@
 
+%%%%
+We take on board the coment and agree that this would be a bad bias if it was the case. However we ran extra analysis and there is no bias.
+%%%%
+
 We performed an extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains. 
-We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or using a random tree (default in MrBayes) as a starting tree.
+We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or not using a starting tree (i.e. using the default in MrBayes which is using a random tree).
+
 % NC: Yes but the reviewer is asking for something different - they ask you to have no starting tree.
@@
 %
+% %%%%%%%%%%%%%%%%%%%
 % TO DISCUSS @@@
+% %%%%%%%%%%%%%%%%%%%
+%
 % TG: about that: it using a random tree or the true tree doesn't change the timing significantly.
@@
 
-% TO DO @@@
+% TG: Add that to the caption
+% Pooled the data across the different missing data parameters
+
+% %%%%%%%%%%%%%%%%%%%
+% TO DISCUSS @@@
+% %%%%%%%%%%%%%%%%%%%
 
@@
 
+% TG: "data" ARE <- change
+
 % NC: Hmmm I reckon this is probably inconsistent throughout then as I usually correct it to plural.
@@
 Would that be singular or plural?
+
+% %%%%%%%%%%%%%%%%%%%
+% TO DISCUSS @@@
+% %%%%%%%%%%%%%%%%%%%
 
@@
 See also Appendix A "Morphological rates estimations results".
+
+We did it
+but we put in appendix
+because it's too long.
+
 % TG: what do you think?
@@
 You can't just throw random stuff into the appendices for no reason.
+%
+% %%%%%%%%%%%%%%%%%%%
 % TO DISCUSS @@@
+% %%%%%%%%%%%%%%%%%%%
+%
 % TG: Ok, maybe we should remove it. I've done the analysis because it's interesting and easy to do but maybe we can just add the results to this response, not to the revised manuscript?
@@
 
+Homoplasy is a difficult concept, not really a value so we went for...
 We changed the sentence to:
@@
 % NC: I think what is needed here is a definition of what is "too much". Is it 50%? 20%? etc. Too much is arbitrary.
+% %%%%%%%%%%%%%%%%%%%
 % TO DISCUSS @@@
+% %%%%%%%%%%%%%%%%%%%
 % TG: That's why I removed the "too much".
@@
 
-% TO DO: thorough check through. @@@
-
 We modified the methods section as follows:
@@
 likelihood and Bayesian trees be so different?\\
+% TG: we really apologize but we think there's an error. We don't know where their coming from but it's probably our fault because of mislabelling. Hopefully this is fixed.
+
+% Check again table. Difference between tables and figures.
+
 2)  Pooling across all missing data schemes seems to wash out the effect of tree-inference method.
@@
 
+% Move that where it's relevant !!!!!!!!
 We fixed the highlighting in bold to the classical values of 0.05 or 0.95.
@@
 
+% %%%%%%%%%%%%%%%%%%%
+% TO DISCUSS @@@
+% %%%%%%%%%%%%%%%%%%%
+
 2) %TG: not entirely sure what the reviewer means by marginal probabilities.
@@
 
+% And fix this one with adding the tables.
+% Marginal probabilities checking. Do it?
+
 % -----------------
@@
 We removed the repetition in the caption.
-If we understand the second point correctly, we have not added "best" to these as the Maximum Likelihood tree is by definition the tree with the highest likelihood, and the Bayesian consensus is the consensus tree of the Bayesian inference trees.
+% If we understand the second point correctly, we have not added "best" to these as the Maximum Likelihood tree is by definition the tree with the highest likelihood, and the Bayesian consensus is the consensus tree of the Bayesian inference trees.
 So adding the word best here is not entirely accurate.
+
+Change caption in the "best" ML tree and "best" Bay con tree to blablabla.
 
@@
 I've added a bit above.
+
+% %%%%%%%%%%%%%%%%%%%
+% TO DISCUSS @@@
+% %%%%%%%%%%%%%%%%%%%
 

### 2015-08-10 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Comments on revisions

@@
 \textbf{Table 1:} \hl{Bhattacharyya Coefficients of the pairwise method comparisons.}
-\hlEach line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
+\hl{Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
 Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.

### 2015-08-11 Manuscript/MPESubmission/Revision/response.tex
Nearly all comments done (still the marginal probabilities to deal with)

--- 2015-08-07 Manuscript/MPESubmission/Revision/response.tex
@@
 
+% TG: I got rid of all your minor comments for cleaning up a bit.
+
 \section{Reviewer 1:}
@@
 We have increased our emphasis on novelty throughout the introduction by more clearly stating the difference between the present study and the literature and more clearly outlining the question of this study, as follows:
-
-% NC: Don't forget to change stuff in the manuscript if I change it here and vice versa.
 
@@
 "In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living or fossil data, or all data for certain characters.
-% NC: All data overall suggests an empty matrix
 We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa." lines 99-103 @@@
-
-% "We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred % NC: Didn't one reviewer have issues with the word inferred?
-TG: for the matrices only (and he/she's right); trees are still inferred.
-% using Maximum Likelihood and Bayesian methods by using two different topological distance metrics, based on the Robinson-Foulds (Robinson and Foulds, 1981) and Triplets (Critchlow et al., 1996) distances.
-% Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological distance metrics." lines 127-134 @@@
-% NC: Why are the distinct topological metrics novel? TG: not novel but novel to use two aspects (people used only clade related metrics before).
 
@@
 Although they might make sentences where Maximum Likelihood and/or Bayesian Inference are stated more concise, we think spelling out the terms makes our paper clearer and does not ask the reader to look back for the actual meaning of the abbreviation (Berlin 2013; Radiology; \url{http://pubs.rsna.org/doi/abs/10.1148/radiol.12121776}).
+Additionally this could lead to a confusion between ML (Maximum Likelihood) and $M_L$ (Proportion of living taxa with no morphological data).
 However, we can change this at a later stage if necessary.
@@
 
+% NC: Needs a little more explanation, and a line about what analysis you did in appendix b.
+So something along the lines of "We compare best trees to missing data trees, but could also compare true trees to the missing data trees.
+We do not do this because...
+However, results of analyses comparing the two can be found in appendix B.
+This makes little difference to our overall results." That might need to be better integrated with the paragraph(s) before.
 % TG: changed to:
@@
 See also appendix B section "Differences between the ``true'' and the ``best'' trees."
-% NC: Needs a little more explanation, and a line about what analysis you did in appendix b.
-So something along the lines of "We compare best trees to missing data trees, but could also compare true trees to the missing data trees.
-We do not do this because...
-However, results of analyses comparing the two can be found in appendix B.
-This makes little difference to our overall results." That might need to be better integrated with the paragraph(s) before.
 
@@
 
-We performed an extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains. 
-We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or using a random tree (default in MrBayes) as a starting tree.
-% NC: Yes but the reviewer is asking for something different - they ask you to have no starting tree.
-If this is impossible in MrBayes then say so, and say you used a random tree instead.
-TG: Maybe there's a phylogeny confusion here.
-In both ML or Bayesian, you need to start from a tree, then calculate the likelihood of all the parameters, then change the tree, calculate again, etc...
+We took on board this comment and agree with the reviewer that if using the ``true'' tree as a starting had an effect it would bias our results.
+We ran, however, an extra analysis to test if that was the case and found that it had no effect of the inferred tree topology.
+We performed this extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains. 
+We compared the topology of these trees in Bayesian inference when using the ``true'' tree as a starting tree or without using a starting tree (i.e.
+using the default in MrBayes which is using a different random tree for each MCMC).
 We found no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
 Therefore, we do not believe our results are merely the result of using a specific starting tree topology.
-%strong priors on the tree topology.
-TG: That's the big difference that the PLoS ONE reviewer also didn't got.
-A starting tree is not a Bayesian prior (in P(A|B) = P(B|A)*P(A)/P(B; P(A) is the prior).
-It's just a user fixed starting point for the MCMC walk rather than a random one.
-% NC: The reviewer also says that saving time is not a good justification. It'd be worth explaining here how long it actually takes even though these datasets are small.
-%
-% TO DISCUSS @@@
-% TG: about that: it using a random tree or the true tree doesn't change the timing significantly.
-The whole idea was suggested by Rob or Gavin (don't remember exactly) at lunch time when I was visiting Gavin in Sheffield.
-To speed up the search in the tree space they suggested I already start from a sensible tree in the tree space, that was meant to speed up the analysis.
-%
-These results are available in the supplementary materials Appendix A (figure A2 and tables A1 and A2) and we also added a note to the manuscript:
+These results are available in the supplementary materials Appendix A (Figure A2 and Table A1) and we also added a note to the manuscript:
 
@@
 Values in between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions." line 419-424 @@@.
-
-% NC: Make sure this is consistent throughout, add this to where you mention the BC being between 0 and 1 too? TG: Done.
 
@@
 Concerning the last question of this comment: "How then in comparisons of the distributions are they always so low (Table 1 max prob.
-of overlapping distributions = 0.11)?" when looking at all parameters separately (figure 4), there is no clear overlap between the Bayesian consensus trees and the Bayesian posterior trees (median Bhattacharyya Coefficient of 0.01; table 1).
-The maximum observed Bhattacharyya Coefficient (0.11; table 1) is probably due to the combination of parameters: when all the parameters have many missing data, the probability of overlap between trees was slightly higher (0.11) but still really low.
+of overlapping distributions = 0.11)?", we apologise for the confusion here and assume it is due to our mislabelling of the figure 4 (see reviewer 1's comment 12).
+In fact, when looking at all parameters separately (figure 4), there is no clear overlap between the Bayesian consensus trees and the Bayesian posterior trees (median Bhattacharyya Coefficient of 0.01; Table 1).
+The maximum observed Bhattacharyya Coefficient (0.11; Table 1) is probably due to the combination of parameters: when all the parameters have many missing data, the probability of overlap between trees was slightly higher (0.11) but still really low.
 For the Normalised Triplets distance however, the overlap is higher (median Bhattacharyya Coefficient of 0.59; table 1) as illustrated in figure 4 (i.e. much more overlap between the distributions).
-
-% TG: I'm confused for the second part of the question:
-% TG: In figure 4 and in Table 1, if you look at the Bayesian consensus (grey) trees and the Bayesian posterior trees (orange) for RF, they do not really overlap.
-Ever.
-The 0.11 probability must come from the parameters combinations that, when is really shitty must show a little bit of overlap.
-But for each parameter one by one: nope.
-No overlap.
-% TG: On the other side, in figure 4 and Table 1, for Bayesian consensus and Bayesian posterior for Tr, the do overlap much more.
-And that's what's in the Table 1 (min 0.21; median 0.56; and max 0.87).
-% TG: Did the reviewer just miss read the labels of the Table 1 or am I miles away from what he means?
-
-% NC: It's more likely an issue with reading figure 4.
-Check your figure legends.
-I think they might be incorrect.
-Then the reviewer might be seeing overlap of the wrong two distributions...
-But if they do overlap all the way to 25% then it is weird that the highest BC is 0.11 isn't it?
-% TG: yes definitely.
-Just doubled and tripled checked, it might be because the colors where miss labelled and the reviewer miss read the graph (my fault).
-In the tables 1, C5, C6 and C7, RF overlap are always low (max < 0.11) and Tr overlap are always high (max ~80+).
-And that's what's visible in figure 4.
-Grey and orange lines do not overlap in RF and do overlap in Tr.
 
@@
 
-We apologise for this silly mistake and fixed the colors in the figure. We also changed the color scheme to match with the general color scheme of the paper (black, grey, orange and blue).
+We apologies for this silly mistake and thank the reviewer for pointing it out.
+We changed the colors to match with the paper color scheme: "Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
+See figure 4 caption.
 
@@
 
-%TG: WUT? Didn't understood that one at all.
-% NC: The reviewer means that you're presenting results from across all your different parameters, and all your different amounts of missing data in the table.
-You need to clarify this, especially as in figures you separate these out.
-Pooling just means combining.
-I noticed this on my re-read just now too.
-
-% TO DO @@@
+We clarified the caption of the Table 1 as follow:
+
+"Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
+Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
+If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology." see Table 1 caption.
 
@@
 
-% TG: actually I'm not going to change it:
-We thank the reviewer for pointing out this inconsistency, however, 'data' is now commonly used for the singular case (Pinker, S.
-\textit{The Sense of Style: The Thinking Person's Guide to Writing in the 21st Century}.
-Penguin, 2014.
-p 271; and Guardian and Observer style guide (2015): \url{http://goo.gl/cmKjNe}).
-
-% NC: Hmmm I reckon this is probably inconsistent throughout then as I usually correct it to plural.
-I'd just change this tbh.
-Or at least check you've been consistent.
-TG: ok maybe be I have a plural/singular problem (I definitely have).
-I actually never know when to use it a singular or plural: "One data point", that's ok, but "morphological data for one taxa"?
-Would that be singular or plural?
+We thank the reviewer for pointing out this inconsistency and changed the use of ``data'' from singular to plural throughout the manuscript.
 
@@
 (2) - the second point underlined by the reviewer in this specific comment strikes use as really interesting.
-However, we considered this study to already be long and complex.
+We ran this additional analysis by comparing the estimated $\alpha$ parameter for the morphological rates distribution to the ``true'' $\alpha$ parameter.
+We found results that the $\alpha$ parameter is usually slightly underestimated for the various ``missing-data'' trees unless the number of living species with no morphological data is high ($M_L$ = 75\%), then, the $\alpha$ are overestimated.
+These results are somehow in agreement with our topological differences results: the number of living species seems to be a driving missing data parameter for inferring Total Evidence trees.
+Nonetheless, we considered this study to already be long and complex.
 We therefore think that adding an extra question to this study (i.e.
 what is the effect of missing data on the morphological substitution rate parameter estimates?) might lose the reader and make an already long paper even longer.
-Nonetheless, we added the result of these analyses to \textbf{Appendix-A Morphological rates estimations results}.
+Therefore, we added the result of these analyses to \textbf{Appendix-A Morphological rates estimations results}.
 We link to this analysis in the methods as follows:
@@
 See also Appendix A "Morphological rates estimations results".
-% TG: what do you think?
-I don't want to attract the reviewer's attention too much (i.e.
-that's another shit load of analysis to do to do it properly) and certainly not the reader's attention either (it's not relevant to the whole paper) but sticking that to the supplementary kind of does what the reviewers wants to know.
-
-% NC: I'm a bit confused - did you do what they wanted or not?
-If not then the whole "this will make the paper longer" argument holds.
-If you did do what they wanted then it might need a little more explanation somewhere.
-You can't just throw random stuff into the appendices for no reason.
-% TO DISCUSS @@@
-% TG: Ok, maybe we should remove it. I've done the analysis because it's interesting and easy to do but maybe we can just add the results to this response, not to the revised manuscript?
 
@@
 
-We changed the sentence to:
+Since homoplasy is a difficult concept to measure (e.g. with character exhaustion; Wagner 2000, Evolution) so we didn't really add a value for ``too much'' homoplasy.
+We therefore removed this mention and changed the sentence to:
 "In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal (Hassanin et al., 1998; Estoup et al., 2002).%\citep{Hassanin1998611,EstoupHomoplasy}.
 Also, we chose this low value to be consistent with our protocol for simulating morphological characters (see below)." (line 189-192@@@).
-
-% NC: I think what is needed here is a definition of what is "too much". Is it 50%? 20%? etc. Too much is arbitrary.
-% TO DISCUSS @@@
-% TG: That's why I removed the "too much".
-I'm actually not sure how Homoplasy is measured (as in units of homoplasy).
-Wagner did a lot of work on it but he usually refers to homoplasy through proxies like characters retention, etc...
 
@@
 We modified the sentence as follows:
-% "We selected this model and these parameters to strike a balance between realism for empirical datasets (e.g.
-Kelly et al., 2014) and models with few parameters (cf.
-Springer et.
-al., 2012) to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)." lines 192-198 @@@
 
@@
 lines @@@
-
-% NC: Why not just use the phrasing given by the reviewer for the first part? TG: Done.
 
@@
 We fixed this typo.
-% TG: should I not add something like "and thank the reviewer for pointing it out."?
-% NC: No, only thank them if it's something really stupid - like the mislabeling of the figure axes - or really insightful.
 
@@
 
-% NC: I think you misunderstood this comment so I have rephrased.
-Their issue was you were recommending people use constraints for tip dating, but then saying not using a constraint might make tip dating better.
-Instead I've made it so you're saying IF you need a constraint, use this, but be careful.
-TG: Deadly!
-Thanks!
-
 % -----------------
@@
 We removed this sentence and replaced it by the two new sentences outlined above (reviewer 1 comment 11).
-
-%NC: Ummm but specific comment 50 refers to comment 11. Should this also refer to comment 11?
 
@@
 
-We apologies for this silly mistake and thank the reviewer for pointing it out.
-We changed the colors to match with the paper color scheme: "Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
-See figure 4 caption.
+We apologies again for this silly mistake. We dealt with this comment above (see reviewer 1's comment 12).
 
@@
 the parameters have a weak effect on the metric and thus the topological recovery)." See figure 6 caption.
-
-% NC: Sometimes in this you have N_C, sometimes M_C. FIX IT!!! TG: Following reviewer 2's comment, it's now N_C rather than M_C.
-
-% TO DO: thorough check through. @@@
 
@@
 
-% TG: I don't understand this comment at all. What does "The wording of the caption is off" means? And which pararagraph is he talking about? 
-% NC: The "wording is off" means the wording is incorrect. Probably means the second sentence not paragraph. I've editted the caption as mentioned further up in the comments.
-
-We fixed the highlighting in bold to the classical values of 0.05 or 0.95.
-We changed the caption to:
-"Bhattacharyya Coefficients of the pairwise method comparisons.
-Each line summarizes the probabilities of overlap between the distributions of the 'best' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
-Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
-If two methods have a high probability of overlap, they have a similar ability to recover the ``correct" tree topology"
-
 1) We are not entirely sure if the reviewer means (i) why the two methods display such differences ; or (ii) why, within each method, the combination of parameters displays a difference between each single parameters ($M_LM_FM_C$ vs.
@@
 the clade ((a;b);c)) could be collapsed in (a;b;c) which is still a corrected representation of the relations within the clades rather than ((b;c);a)).
-In the second case, the fact that the missing data parameters display better scores alone is because in practice, when this occured, the two other parameters where set a 0\%, therefore resulting in matrices with a maximum of 75\%/2 missing data for $M_L$ and $M_F$ an 75\% maximum for $M_C$.
-When combined, the maximum amount of missing data can be higher, especially when each parameter is set a 75\% missing data.
-
-% NC: OK I think the point here is that Table 1 has a column saying the "Max." BC score. So the reviewer is saying that this is happening:
-% In Appendix C, B posterior vs B consensus for some parameter value is 0.5 (for example), but the max value when you compare this in Table 1 is 0.11. How is this possible???????
-% NC: So your response here is completely irrelevant.
+
+In the second case, the fact that probabilities of overlap between the methods when the missing data parameters were taken separately (i.e.
+only the focal missing data parameter varies) was higher than the parameters combinations is probably just due to the overall amount of missing data.
+In practice, when only one missing data parameter varies the two other parameters where set a 0\%, therefore resulting in matrices with a maximum of 75\%/2 missing data for $M_L$ and $M_F$ an 75\% maximum for $M_C$.
+When combined, the maximum amount of missing data can be much higher, especially when each parameter is set a 75\% missing data, thus decreasing the probability of the ``missing-data'' trees to be similar to the ``best'' tree (see Table 1 in this response).
+
+\begin{table}[!ht]
+%\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+\caption{Distribution overlap between Maximum Likelihood (ML) and Bayesian consensus tree (BI) across all missing data parameters combinations (pooled) and every single missing data parameter individually ($M_L$, $M_F$, $N_C$)}
+\centering
+\begin{tabular}{|l|c|c|c|c|c|c|c|}
+  \hline
+ Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
+  \hline
+    ML \textit{vs.} BI (pooled) & $RF$ & \textbf{0.00} & \textbf{0.00} & 0.10 & 0.20 & 0.32 & \textbf{1.00} \\ 
+    ML \textit{vs.} BI ($M_L$)  & $RF$ & 0.30          & 0.31          & 0.69 & 0.61 & 0.77 & \textbf{1.00} \\ 
+    ML \textit{vs.} BI ($M_F$)  & $RF$ & \textbf{0.00} & 0.25          & 0.48 & 0.50 & 0.76 & \textbf{1.00} \\ 
+    ML \textit{vs.} BI ($N_C$)  & $RF$ & \textbf{0.03} & 0.32          & 0.66 & 0.55 & 0.75 & \textbf{1.00} \\
+    ML \textit{vs.} BI (pooled) & $Tr$ & 0.34          & 0.49          & 0.61 & 0.62 & 0.75 & \textbf{1.00} \\
+    ML \textit{vs.} BI ($M_L$)  & $Tr$ & 0.79          & 0.81          & 0.84 & 0.86 & 0.85 & \textbf{1.00} \\  
+    ML \textit{vs.} BI ($N_C$)  & $Tr$ & 0.51          & 0.69          & 0.80 & 0.76 & 0.80 & \textbf{1.00} \\
+    ML \textit{vs.} BI ($M_F$)  & $Tr$ & 0.38          & 0.69          & 0.75 & 0.72 & 0.80 & \textbf{1.00} \\  
+   \hline
+\end{tabular}
+\label{Tab_Results-Difference_methods}
+%\end{adjustwidth}
+\end{table}
+
+% %%%%%%%%%%%%%%%%%%%
+% TO DISCUSS @@@
+% %%%%%%%%%%%%%%%%%%%
 
@@
 We removed the repetition in the caption.
-If we understand the second point correctly, we have not added "best" to these as the Maximum Likelihood tree is by definition the tree with the highest likelihood, and the Bayesian consensus is the consensus tree of the Bayesian inference trees.
-So adding the word best here is not entirely accurate.
-
-%Not sure I understand the second part of the comment.
-% NC: I think they are a bit confused and think you need to explain it is the best ML and best bayesian tree.
-But that's what ML and consensus means.
-Maybe they want clarification that it is only one tree in each of those and multiple trees in the others?
-I've added a bit above.
+If we understood the second point correctly, the reviewer suggests we add ``\textit{best}'' to the two first rows of the tables, and by ``\textit{best}'', the reviewer means ``optimal'' (as opposed to the ``best'' tree in our protocol that refers to any tree inferred by any method from the ``complete matrix''.
+We don't think it is necessary to follow this suggestion as the Maximum Likelihood tree is by definition the tree with the highest likelihood, and the Bayesian consensus is the consensus tree of the Bayesian inference trees.
+So adding the word ``\textit{best}'' here is not entirely accurate.
 

### 2015-08-11 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Nearly all comments done (still the marginal probabilities to deal with)

@@
 \citealp{O'Leary08022013} \textit{vs.} \citealp{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record.
-For example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data is (nearly) always unavailable.
+For example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data \hl{are} (nearly) always unavailable.
 Therefore \hl{Total Evidence matrices are likely to contain a large proportions of missing data} that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
@@
 \hl{These studies have assessed the effect of missing data on topology either:(1) by comparing a dataset with missing data to sub-sets without missing data }\citep{Wiens01102005}\hl{; or (2) by removing both molecular and some morphological data from living taxa (``artificial fossil'') }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{Both types of approaches have shown that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
+\hl{Both types of approaches have shown that missing data \hl{are} not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
 However, these studies were limited by their empirical-only approach and the use of only living taxa }\citep{Wiens01102005}\hl{ or the scheme of generating missing data }\citep[\hl{i.e.
@@
 % \hl{Furthermore, \protect\cite{pattinsonphylogeny2014} }\hl{assess the effect of removing both molecular and morphological data (``artificial extinction'') on the topology.}
-% \hl{As for morphological and molecular data separately, these studies shows that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.}
+% \hl{As for morphological and molecular data separately, these studies shows that missing data are not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.}
 % \hl{These studies, however, are limited by their empirical-only approach \protect\citep{Wiens01102005} }\hl{and the scheme of generating missing data \protect\citep{manosphylogeny2007,pattinsonphylogeny2014}.}
@@
 This is partially due to difficulties with simulating branch lengths and topology simultaneously, \hl{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints} \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
-Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework.
+Thus understanding the sensitivity of topology to missing \hl{data} are important for assessing the accuracy of tree estimation in the Total Evidence framework.
 To our knowledge, this question has never been formally assessed.
@@
 6, Table ~\ref{Tab_Results-Difference_methods}).
-As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
+As seen in previous studies, our results show that the amount of missing data are not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.
 
@@
 decrease of Normalised Triplets metric) with increasing missing data in this parameter, clade conservation (i.e.
-Normalised Robinson-Foulds metric) is still relatively good (mode = 0.72) when the proportion of missing data is high ($M_{F}$ = 75\%).
+Normalised Robinson-Foulds metric) is still relatively good (mode = 0.72) when the proportion of missing data are high ($M_{F}$ = 75\%).
 \hl{These results are in agreement with }\cite{manosphylogeny2007}\hl{ where as few as 16 characters where sufficient for correctly assigning ``artificial fossils'' to their ``correct'' clade.}
@@
 This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
-Even if clades are conserved (Fig. 5), the resolution within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
+Even if clades are conserved (Fig. 5), the resolution within them can be poor to non-existent when a lot of data are missing (i.e. 75\%).
 In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
@@
 Gaps in the matrix, however, can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
-Fortunately, although this data is the most difficult to collect, it also has the least influence on whether our simulations recover the ``best'' tree topology (Fig. 6).
+Fortunately, although this data are the most difficult to collect, it also has the least influence on whether our simulations recover the ``best'' tree topology (Fig. 6).
 Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
@@
 not including fossil taxa, }\citealt{Wiens01102005}\hl{; or focusing on the palaeontological aspect, }\citealt{manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{This study proposes a more in depth approach where missing data is generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
+\hl{This study proposes a more in depth approach where missing data are generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
 This allowed us to confirm previous results that missing data can be especially problematic in small matrices }\citep{wiensmissing2003}\hl{ but also revealed the crucial importance of coding morphological data for living species.

### 2015-08-11 Analysis/Marginal_probabilities.R
Marginal probabilities calculation

@@ -1,0 +1,100 @@
+#Loading the functions
+source("MissingData_fun.R")
+
+#Loading the data
+quick.load=TRUE
+if(quick.load == FALSE) {
+    source("MissingData_load.R") 
+} else {
+    #Quick load
+    load("../Data/R_data/TreeCmp-ML_bestrees.Rda")
+    load("../Data/R_data/TreeCmp-ML_bootstraps.Rda")
+    load("../Data/R_data/TreeCmp-Bayesian_contrees.Rda")
+    load("../Data/R_data/TreeCmp-Bayesian_treesets.Rda")
+
+    #Isolating the parameters per chains
+    par_ML<-c(1,26,51,76,101)
+    par_MF<-c(1,6,11,16,21)
+    par_MC<-c(1:5)
+    par_MLMF<-c(1,6,11,16,21, 26,31,36,41,46, 51,56,61,66,71, 76,81,86,91,96, 101,106,111,116,121)
+    par_MLMC<-c(1:5, 26:30, 51:55, 76:80, 101:105)
+    par_MFMC<-c(1:25)
+    par_MLMFMC<-c(1:125)
+}
+
+
+#Setting the variables
+data.sets<-c("ML_besttrees","Bayesian_contrees","ML_bootstraps","Bayesian_treesets")
+metrics<-c("R.F_Cluster","Triples")
+par<-par_ML
+
+#----------------
+# Extracting the mode value for each tree
+#----------------
+missing_data_param<-c("0%","10%","25%","50%","75%")
+#missing_data_param<-c(100,90,75,50,25) #For NC
+
+#Only for each parameters individually
+mar_dist<-matrix(NA, nrow=(4+1), ncol=(length(par)+1))
+colnames(mar_dist)<-c(missing_data_param, "Marginal")
+rownames(mar_dist)<-c("Maximum Likelihood","Bayesian consensus", "Maximum Likelihood Bootstrap", "Bayesian posterior distribution", "Marginal")
+
+#Creating the matrices for RF and Tr
+mar_dist_Tr<-mar_dist_RF<-mar_dist
+
+#Get the mode in there!
+mode.val<-function (x) {
+    as.numeric(names(sort(-table(x))[1]))
+}
+for (method in 1:length(data.sets)) {
+    #RF
[…]

### 2015-08-11 Analysis/Marginal_probabilities.R
Updated marginal prob

@@ -69,32 +69,15 @@ raw_Tr<-mar_dist_Tr[c(1:4), c(1:5)]
 #----------------
 #sum of row
 for (method in 1:length(data.sets)) {
-    mar_dist_RF[method,6]<-sum(mar_dist_RF[method,c(1:5)])
-    mar_dist_Tr[method,6]<-sum(mar_dist_Tr[method,c(1:5)])
+    mar_dist_RF[method,6]<-sum(mar_dist_RF[method,c(1:5)])/length(par)
+    mar_dist_Tr[method,6]<-sum(mar_dist_Tr[method,c(1:5)])/length(par)
 }
 #sum of column
 for (param in 1:length(par)) {
-    mar_dist_RF[5,param]<-sum(mar_dist_RF[c(1:4),param])
-    mar_dist_Tr[5,param]<-sum(mar_dist_Tr[c(1:4),param])
+    mar_dist_RF[5,param]<-sum(mar_dist_RF[c(1:4),param])/length(data.sets)
+    mar_dist_Tr[5,param]<-sum(mar_dist_Tr[c(1:4),param])/length(data.sets)
 }
 
 library(xtable)
 xtable(mar_dist_RF)
 xtable(mar_dist_Tr)
-
-#Calculate the probabilities
-#mar_dist_RF[5,6]<-total_RF<-sum(mar_dist_RF[5,], na.rm=T)
-#mar_dist_Tr[5,6]<-total_Tr<-sum(mar_dist_Tr[5,], na.rm=T)
-
-#round(mar_dist_RF/total_RF, digit=2)
-#round(mar_dist_Tr/total_Tr, digit=2)
-
-#----------------
-# Same but with the non-marginal values being the mode values
-#----------------
-
-#RF_tmp<-round(mar_dist_RF/total_RF, digit=2)
-#RF_tmp[c(1:4), c(1:5)]<-raw_RF
-
-#Tr_tmp<-round(mar_dist_Tr/total_Tr, digit=2)
-#Tr_tmp[c(1:4), c(1:5)]<-raw_Tr

### 2015-08-11 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixC.tex
Marginal prob tables

@@
 \begin{table}[ht]
-\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).}
+\caption{Summary of the comparisons between the "best" tree and the "missing-data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).}
 \label{Tab_Supp_summary_metric_allparam}
@@
 \begin{table}[ht]
-\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
+\caption{Summary of the comparisons between the "best" tree and the "missing-data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_ML}
@@
 \begin{table}[ht]
-\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
+\caption{Summary of the comparisons between the "best" tree and the "missing-data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_MF}
@@
 \begin{table}[ht]
-\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $N_{C}$ missing data parameter only.}
+\caption{Summary of the comparisons between the "best" tree and the "missing-data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $N_{C}$ missing data parameter only.}
 \label{Tab_Supp_summary_metric_MC}
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing-data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
 \label{Tab_Supp_summary_BC_ML}
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing-data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
 \label{Tab_Supp_summary_BC_MF}
@@
 \begin{table}[ht]
-\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $N_{C}$ missing data parameter only.}
+\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised metric between the "best" tree and the "missing-data" using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $N_{C}$ missing data parameter only.}
 \label{Tab_Supp_summary_BC_MC}
@@
 
+% Marginal distributions tables
 \begin{table}[ht]
-\caption{Summary of the median Bhattacharyya Coefficients comparisons between the missing data parameters.
-Each value corresponds to the median Bhattacharyya Coefficient between the Normalised metrics (Robinson-Foulds - RF; Triplets - Tr) between the ``best'' and the ``missing data'' trees (as outlined in Figure 2 in the main text) for the the trees with: Missing Living taxa only ($M_L$); Missing Fossil data only ($M_F$); Number of characters only ($N_C$); and the combination of the three parameters (combined).}
-\begin{tabular}{llcccc}  
+\caption{Marginal distribution of the modal Normalised Robinson-Foulds metric for the comparison between the ``best'' tree and the ``missing-data'' trees for each different tree inference method for the $M_{L}$ missing data parameter only.}
+\centering
+\begin{tabular}{rrrrrrr}
   \hline
- Method & Metric & $M_L$ & $M_F$ & $N_C$ & combined \\ 
+ & 0\% & 10\% & 25\% & 50\% & 75\% & Marginal \\ 
   \hline
-Maximum Likelihood & RF & 0.54 & 0.37 & 0.56 & 0.54 \\ 
-                   & Tr & 0.80 & 0.75 & 0.80 & 0.79 \\ 
-Bayesian consensus & RF & 0.73 & 0.59 & 0.60 & 0.79 \\ 
-                   & Tr & 0.86 & 0.67 & 0.81 & 0.81 \\ 
-Maximum Likelihood bootstraps & RF & \textbf{0.98} & \textbf{0.96} & \textbf{0.97} & 0.78 \\ 
-                   & Tr & \textbf{0.99} & \textbf{1.00} & \textbf{1.00} & \textbf{0.96} \\ 
-Bayesian posterior trees & RF & 0.86 & 0.93 & 0.94 & 0.68 \\ 
-                   & Tr & \textbf{0.98} & \textbf{0.99} & \textbf{0.99} & 0.91 \\ 
+Maximum Likelihood & 1.00 & 0.94 & 0.71 & 0.59 & 0.45 & 0.74 \\ 
+  Bayesian consensus & 1.00 & 0.95 & 0.87 & 0.74 & 0.71 & 0.85 \\ 
+  Maximum Likelihood Bootstrap & 0.47 & 0.45 & 0.43 & 0.36 & 0.34 & 0.41 \\ 
+  Bayesian posterior distribution & 0.67 & 0.61 & 0.53 & 0.41 & 0.32 & 0.51 \\ 
+  Marginal & 0.79 & 0.74 & 0.64 & 0.53 & 0.46 &  \\ 
    \hline
@@
 \begin{table}[ht]
-\caption{Summary of the median Bhattacharyya Coefficients comparisons between the missing data parameters and the methods.
-Each value corresponds to the median Bhattacharyya Coefficient between the Normalised metrics (Robinson-Foulds - RF; Triplets - Tr) between the `methods (as outlined in Figure 3 in the main text) for the the trees with: Missing Living taxa only ($M_L$); Missing Fossil data only ($M_F$); Number of characters only ($N_C$); and the combination of the three parameters (combined).}
-\begin{tabular}{llcccc}  
+\caption{Marginal distribution of the modal Normalised Triplets metric for the comparison between the ``best'' tree and the ``missing-data'' trees for each different tree inference method for the $M_{L}$ missing data parameter only.}
+\centering
+\begin{tabular}{rrrrrrr}
   \hline
- Method & Metric & $M_L$ & $M_F$ & $N_C$ & combined \\ 
+ & 0\% & 10\% & 25\% & 50\% & 75\% & Marginal \\ 
   \hline
-Maximum likelihood vs. Bayesian consensus & RF & 0.69 & 0.48 & 0.66 & 0.10 \\ 
-   & Tr & 0.84 & 0.75 & 0.80 & 0.61 \\ 
-Maximum likelihood vs. Bootstraps & RF & 0.29 & 0.32 & 0.21 & 0.69 \\ 
-   & Tr & 0.53 & 0.40 & 0.39 & 0.65 \\ 
-Maximum likelihood vs. Bayesian posterior trees & RF & 0.61 & 0.71 & 0.47 & 0.80 \\ 
-  & Tr & 0.70 & 0.62 & 0.59 & 0.73 \\ 
-Bayesian consensus vs. Bootstraps & RF & \textbf{0.02} & \textbf{0.01} & \textbf{0.01} & \textbf{0.00} \\ 
-  & Tr & 0.78 & 0.54 & 0.62 & 0.59 \\ 
-Bayesian consensus vs. Bayesian posterior trees & RF & \textbf{0.02} & \textbf{0.02} & \textbf{0.04} & \textbf{0.01} \\ 
-   & Tr & 0.75 & 0.66 & 0.64 & 0.56 \\ 
-Boostraps vs. Bayesian posterior trees & RF & 0.85 & 0.72 & 0.73 & 0.85 \\ 
-   & Tr & \textbf{0.96} & 0.91 & 0.93 & \textbf{0.98} \\ 
+Maximum Likelihood & 1.00 & 1.00 & 0.82 & 0.64 & 0.33 & 0.76 \\ 
+  Bayesian consensus & 1.00 & 0.19 & 0.19 & -0.05 & -0.39 & 0.19 \\ 
+  Maximum Likelihood Bootstrap & 1.00 & 0.51 & 0.44 & 0.41 & 0.37 & 0.55 \\ 
+  Bayesian posterior distribution & 1.00 & 0.84 & 0.56 & 0.37 & 0.36 & 0.63 \\ 
+  Marginal & 1.00 & 0.64 & 0.50 & 0.34 & 0.17 &  \\ 
    \hline
@@
 \end{table}
+
+\begin{table}[ht]
+\caption{Marginal distribution of the modal Normalised Robinson-Foulds metric for the comparison between the ``best'' tree and the ``missing-data'' trees for each different tree inference method for the $M_{F}$ missing data parameter only.}
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & 0\% & 10\% & 25\% & 50\% & 75\% & Marginal \\ 
+  \hline
+Maximum Likelihood & 1.00 & 0.90 & 0.75 & 0.47 & 0.24 & 0.67 \\ 
+  Bayesian consensus & 1.00 & 0.97 & 0.93 & 0.77 & 0.72 & 0.88 \\ 
+  Maximum Likelihood Bootstrap & 0.47 & 0.47 & 0.41 & 0.30 & 0.14 & 0.36 \\ 
+  Bayesian posterior distribution & 0.67 & 0.61 & 0.57 & 0.45 & 0.24 & 0.51 \\ 
+  Marginal & 0.79 & 0.74 & 0.67 & 0.50 & 0.33 &  \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\caption{Marginal distribution of the modal Normalised Triples metric for the comparison between the ``best'' tree and the ``missing-data'' trees for each different tree inference method for the $M_{F}$ missing data parameter only.}
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & 0\% & 10\% & 25\% & 50\% & 75\% & Marginal \\ 
+  \hline
+Maximum Likelihood & 1.00 & 1.00 & 0.88 & 0.74 & 0.62 & 0.85 \\ 
+  Bayesian consensus & 1.00 & 1.00 & 0.10 & -0.09 & -0.44 & 0.31 \\ 
+  Maximum Likelihood Bootstrap & 1.00 & 0.68 & 0.63 & 0.48 & 0.38 & 0.63 \\ 
+  Bayesian posterior distribution & 1.00 & 0.98 & 0.98 & 0.72 & 0.36 & 0.81 \\ 
+  Marginal & 1.00 & 0.92 & 0.65 & 0.46 & 0.23 &  \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\caption{Marginal distribution of the modal Normalised Robinson-Foulds metric for the comparison between the ``best'' tree and the ``missing-data'' trees for each different tree inference method for the $N_{C}$ missing data parameter only.}
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & 100 & 90 & 75 & 50 & 25 & Marginal \\ 
+  \hline
+Maximum Likelihood & 1.00 & 0.94 & 0.73 & 0.57 & 0.36 & 0.72 \\ 
+  Bayesian consensus & 1.00 & 0.90 & 0.88 & 0.83 & 0.75 & 0.87 \\ 
+  Maximum Likelihood Bootstrap & 0.47 & 0.45 & 0.41 & 0.34 & 0.24 & 0.38 \\ 
+  Bayesian posterior distribution & 0.67 & 0.63 & 0.57 & 0.45 & 0.32 & 0.53 \\ 
+  Marginal & 0.79 & 0.73 & 0.65 & 0.55 & 0.42 &  \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+\caption{Marginal distribution of the modal Normalised Triples metric for the comparison between the ``best'' tree and the ``missing-data'' trees for each different tree inference method for the $N_{C}$ missing data parameter only.}
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ & 100 & 90 & 75 & 50 & 25 & Marginal \\ 
+  \hline
+Maximum Likelihood & 1.00 & 1.00 & 0.93 & 0.69 & 0.51 & 0.83 \\ 
+  Bayesian consensus & 1.00 & 0.70 & 1.00 & -0.19 & -0.41 & 0.42 \\ 
+  Maximum Likelihood Bootstrap & 1.00 & 0.55 & 0.48 & 0.59 & 0.30 & 0.58 \\ 
+  Bayesian posterior distribution & 1.00 & 0.98 & 0.82 & 0.64 & 0.34 & 0.76 \\ 
+  Marginal & 1.00 & 0.81 & 0.81 & 0.43 & 0.18 &  \\ 
+   \hline
+\end{tabular}
+\end{table}
+
+% \begin{table}[ht]
+% \caption{Summary of the median Bhattacharyya Coefficients comparisons between the missing data parameters.
+Each value corresponds to the median Bhattacharyya Coefficient between the Normalised metrics (Robinson-Foulds - RF; Triplets - Tr) between the ``best'' and the ``missing data'' trees (as outlined in Figure 2 in the main text) for the the trees with: Missing Living taxa only ($M_L$); Missing Fossil data only ($M_F$); Number of characters only ($N_C$); and the combination of the three parameters (combined).}
+% \begin{tabular}{llcccc}  
+%   \hline
+%  Method & Metric & $M_L$ & $M_F$ & $N_C$ & combined \\ 
+%   \hline
+% Maximum Likelihood & RF & 0.54 & 0.37 & 0.56 & 0.54 \\ 
+%                    & Tr & 0.80 & 0.75 & 0.80 & 0.79 \\ 
+% Bayesian consensus & RF & 0.73 & 0.59 & 0.60 & 0.79 \\ 
+%                    & Tr & 0.86 & 0.67 & 0.81 & 0.81 \\ 
+% Maximum Likelihood bootstraps & RF & \textbf{0.98} & \textbf{0.96} & \textbf{0.97} & 0.78 \\ 
+%                    & Tr & \textbf{0.99} & \textbf{1.00} & \textbf{1.00} & \textbf{0.96} \\ 
+% Bayesian posterior trees & RF & 0.86 & 0.93 & 0.94 & 0.68 \\ 
+%                    & Tr & \textbf{0.98} & \textbf{0.99} & \textbf{0.99} & 0.91 \\ 
+%  \hline
+% \end{tabular}
+% \end{table}
+
+% \begin{table}[ht]
+% \caption{Summary of the median Bhattacharyya Coefficients comparisons between the missing data parameters and the methods.
+Each value corresponds to the median Bhattacharyya Coefficient between the Normalised metrics (Robinson-Foulds - RF; Triplets - Tr) between the `methods (as outlined in Figure 3 in the main text) for the the trees with: Missing Living taxa only ($M_L$); Missing Fossil data only ($M_F$); Number of characters only ($N_C$); and the combination of the three parameters (combined).}
+% \begin{tabular}{llcccc}  
+%   \hline
+%  Method & Metric & $M_L$ & $M_F$ & $N_C$ & combined \\ 
+%   \hline
+% Maximum likelihood vs. Bayesian consensus & RF & 0.69 & 0.48 & 0.66 & 0.10 \\ 
+%    & Tr & 0.84 & 0.75 & 0.80 & 0.61 \\ 
+% Maximum likelihood vs. Bootstraps & RF & 0.29 & 0.32 & 0.21 & 0.69 \\ 
+%    & Tr & 0.53 & 0.40 & 0.39 & 0.65 \\ 
+% Maximum likelihood vs. Bayesian posterior trees & RF & 0.61 & 0.71 & 0.47 & 0.80 \\ 
+%   & Tr & 0.70 & 0.62 & 0.59 & 0.73 \\ 
+% Bayesian consensus vs. Bootstraps & RF & \textbf{0.02} & \textbf{0.01} & \textbf{0.01} & \textbf{0.00} \\ 
+%   & Tr & 0.78 & 0.54 & 0.62 & 0.59 \\ 
+% Bayesian consensus vs. Bayesian posterior trees & RF & \textbf{0.02} & \textbf{0.02} & \textbf{0.04} & \textbf{0.01} \\ 
+%    & Tr & 0.75 & 0.66 & 0.64 & 0.56 \\ 
+% Boostraps vs. Bayesian posterior trees & RF & 0.85 & 0.72 & 0.73 & 0.85 \\ 
+%    & Tr & \textbf{0.96} & 0.91 & 0.93 & \textbf{0.98} \\ 
+%  \hline
+% \end{tabular}
+% \end{table}
 %END

### 2015-08-11 Manuscript/MPESubmission/Revision/response.tex
Edited response

@@
 
-See also appendix B section "Differences between the ``true'' and the ``best'' trees."
+See also appendix B section "Differences between the ``true'' and the ``best'' trees." and reviewer 2's comment 1 below.
 
@@
 
-% %%%%%%%%%%%%%%%%%%%
-% TO DISCUSS @@@
-% %%%%%%%%%%%%%%%%%%%
-
-2) %TG: not entirely sure what the reviewer means by marginal probabilities.
-I made to extra summarizing tables in the supplementaries that may be actually easier to read than most of the others so we might want to add them to the main text.
-We added two extra tables in the Appendix C (Table C8 and C9) that summarized the median Bhattacharyya Coefficients within each methods and each parameters individually and combined (Table C8; as illustrated in Figure 2) as well as the median Bhattacharyya Coefficients within each methods and each parameters individually and combined (Table C9; as illustrated in Figure 3).
-% TG: If that's actually what the reviewer 1 wanted, I think we can propose to add it to the main text rather than the appendices.
-They can always decide to move it back to the appendices if it doesn't bring much.
-
-% NC: https://en.wikipedia.org/wiki/Marginal_distribution
-% NC: Marginal probabilities are basically the probability of one thing given you're controlling for another thing.
-So they are looking to see the effect of tree inference method independent of parameter differences.
-Otherwise they suggest you are hiding important differences via this summarizing over all parameter combinations.
-% NC: I think you can argue that yes, table 1 hides some important effects, but that it was difficult to summarize these results and more in depth tables can be seen in the appendix.
-
-% And fix this one with adding the tables.
-% Marginal probabilities checking. Do it?
+2) We followed the reviewer's suggestion and added 6 extra tables in the Appendix C (Table C8 to C13) containing the marginal probability of each missing data parameter individually and each tree inference method.
+We considered running the same type of analysis for the combination of the parameter, however, this involved a table of 5*126 cells which was highly impractical to read.
+% TG: Not sure if the last bit is necessary...
 
@@
 See \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Analysis_Demo.R} and Appendix D.
-% NC: Maybe add something like this if this is your intention. It's probably worth it. No need to keep lots of un-necessary files. TG: Agree!
+% NC: Maybe add something like this if this is your intention. It's probably worth it. No need to keep lots of un-necessary files. TG: Done and agree
 We will also clean up this repository once the paper has been accepted.
@@
 
-A similar point/suggestion was also raised by reviewer 1 (see Reviewer 1 General comment number 6 and specific comment number 25).
-We understand the concerns of the reviewer but, again, we do not agree that comparing the "missing-data" trees to the "best" trees adds a layer of complexity to our simulations.
-In fact, we argue that comparing the "missing-data" trees to the "true" tree actually adds a layer of complexity to the analysis.
-The problem we are tackling in this manuscript is to understand "what is the effect of our three parameters on recovering topology ?"
-To answer this question, we have to add a layer between the data (i.e. the "complete" or "missing-data" matrices) and the actual topology which is the tree inference.
-We test the effect of this layer by using different tree inference methods.
-
-% NC: Note: problematic does not mean what you think it does... Something can be problematic (it's an adjective not a noun).
-
-Because we use a simulation approach to test this problem, we have to start from a random starting point (the "true" tree) and then do the following steps: (1) generate and remove the data (2) infer the topologies and (3) compare the topologies.
-Note that in this schematic representation, the "true" tree is the step (0) of our analysis and involves a series of algorithms and software (i.e.
-seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes).
-To compare the "true" tree's topology to the "missing-data" topologies will be equivalent of exploring the effect of our three parameters on recovering topology through 3 layers (1- generate the data; 2- infer the topologies; and 3- compare the topologies).
-By comparing the "best" tree's topology to the "missing-data" ones, we explore the effect of our three parameters through only 2 layers (2- infer the topologies; and 3- compare the topologies).
-Note that we explore the effect of these 2 layers by using two different tree inference methods (effectively 4 since we also use the Maximum Likelihood bootstraps trees and the Bayesian posterior trees) and 2 different Normalised topological metrics (Robinson-Foulds and Triplets).
-Using as main results the comparisons between the "true" and the "missing-data" trees would also need the exploration of the effect of the cascade of algorithms and softwares (maybe by modifying the algorithms used to generate the "true" tree and the matrices?).
-
-% NC: This argument makes no sense. The point the reviewer is making is that we go:
+A similar point/suggestion was also raised by reviewer 1 and we ran the suggested extra analysis (see Reviewer 1 General comment number 6 and specific comment number 25).
+We understand the concerns of the reviewer but, we do not agree that comparing the "missing-data" trees to the "best" trees adds a layer of complexity to our simulations.
+In practice, the ``true'' tree can be perceived as a ``random'' seed to generate the ``complete'' matrix.
+The ``true'' tree is simply a birth-death tree with $\lambda$ $>$ $\mu$ and a set number of living and fossil taxa (see details in the manuscript section "2.1 Generating the matrix").
+All these parameters, were only used to create a realistic ``complete'' matrix in order to make our results relevant to biologists.
+The aim of this study was to assess how good will a topology inferred from a matrix with missing data (which is likely in biology) compared to the ones inferred form the ``complete'' matrix (the ``best'' tree).
+
+% TG: maybe add a sentence in the manuscript?
+
+% NC: This argument [TG: the one commented out below] makes no sense. The point the reviewer is making is that we go:
 % True tree > Best tree > comparisons when we could just go True tree > comparisons.
@@
 BUT that we have tested their idea and find it makes no difference but all the results are in the appendix.
-
-However, we have added analyses with comparisons between the "true" tree and the "missing-data" trees in the Appendix C (Figures C6 and C7).
-It is interesting to note that the algorithms and software as well as probably the size of our matrices are actually not ideal for the recovery of the "true" tree is rather low and scales down the our results (see figure @).% NC: HUH????
-However, even though these comparisons are worse than comparing the best tree to the missing data trees, it is encouraging to see that the patterns in the results are conserved (i.e.
+% TG: you're right, after reflection, I think we should have never mentioned the ``true'' tree and just say in the methods that we used a random Birth-Death tree with such and such parameters to generate the matrix.
+It's kind of too late to backup by now but the argument was just kind of saying: "we don't care if we can recover the "true" tree because we don't give a flip on how the algorithms perform at making a matrix".
+It's all random anyway...
+% TG : I've commented out the non-sense justification below and wrote something shorter above (and referring to reviewer's 1 comment).
+
+% Because we use a simulation approach to test this problem, we have to start from a random starting point (the "true" tree) and then do the following steps: (1) generate and remove the data (2) infer the topologies and (3) compare the topologies.
+% Note that in this schematic representation, the "true" tree is the step (0) of our analysis and involves a series of algorithms and software (i.e.
+seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes).
+% To compare the "true" tree's topology to the "missing-data" topologies will be equivalent of exploring the effect of our three parameters on recovering topology through 3 layers (1- generate the data; 2- infer the topologies; and 3- compare the topologies).
+% By comparing the "best" tree's topology to the "missing-data" ones, we explore the effect of our three parameters through only 2 layers (2- infer the topologies; and 3- compare the topologies).
+% Note that we explore the effect of these 2 layers by using two different tree inference methods (effectively 4 since we also use the Maximum Likelihood bootstraps trees and the Bayesian posterior trees) and 2 different Normalised topological metrics (Robinson-Foulds and Triplets).
+% Using as main results the comparisons between the "true" and the "missing-data" trees would also need the exploration of the effect of the cascade of algorithms and softwares (maybe by modifying the algorithms used to generate the "true" tree and the matrices?).
+
+% However, we have added analyses with comparisons between the "true" tree and the "missing-data" trees in the Appendix C (Figures C6 and C7).
+% It is interesting to note that the algorithms and software as well as probably the size of our matrices are actually not ideal for the recovery of the "true" tree is rather low and scales down the our results (see figure @).% NC: HUH????
+% However, even though these comparisons are worse than comparing the best tree to the missing data trees, it is encouraging to see that the patterns in the results are conserved (i.e.
 Bayesian consensus trees $>$ Maximum Likelihood trees $>$ both bootstraps trees and Bayesian posterior trees).
@@
 
-% NC: NC_C here again not M_C...
-
 % -----------------
@@
 
-% TG: Ok i've changed it every where but personally I don't like it.
-It makes it even more messy and hard to read.
-However, because I didn't answered the first and last comment of this reviewer, I felt that at least 2/4 is a minimum.
 We agree with this comment and hope it will help to the global comprehension of the manuscript.
@@
 We also slightly modified some explanations related to the change of missing characters to number of characters throughout the manuscript.
-
-% NC: Ah OK. You haven't corrected this everywhere in these comments I don't think... TG: Should be good now.
 

### 2015-08-11 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Edited response

@@
 using both molecular and morphological data; e.g.}][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{These studies have assessed the effect of missing data on topology either:(1) by comparing a dataset with missing data to sub-sets without missing data }\citep{Wiens01102005}\hl{; or (2) by removing both molecular and some morphological data from living taxa (``artificial fossil'') }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{These studies have assessed the effect of missing data on topology either: (1) by comparing a dataset with missing data to sub-sets without missing data }\citep{Wiens01102005}\hl{; or (2) by removing both molecular and some morphological data from living taxa (``artificial fossil'') }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
 \hl{Both types of approaches have shown that missing data \hl{are} not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.

### 2015-08-11 Manuscript/MPESubmission/Revision/response.tex
Minor edits to formatting

@@
 %Packages
-\setlength\parindent{0pt} %NC: To get rid of automatic indenting
 \usepackage{pdflscape}
@@
 %Pagination style and stuff
-\linespread{2}
+%\linespread{2} 
+% NC: Cover letters don't need to be double spaced and by removing this we reduce the page count substantially! i.e. to non "losing the will to live" lengths...
+
 \raggedright
@@
 
+\setlength\parindent{0pt}
+
 \begin{document}
 
-Dear Dr D\'{a}valos,
-
-RE: MPE-15-254
-
+\textbf{RE: MPE-15-254}\\
+\bigskip
+Dear Dr D\'{a}valos,\\
+\bigskip
 We are very grateful to both referees for their helpful and constructive comments, which we believe have helped us to significantly improve our paper.
@@
 
+Please let us know if you require any further information,\\
+\bigskip
+
+Thomas Guillerme (guillert@tcd.ie)\\
+Natalie Cooper
+
 \end{document}

### 2015-08-11 Manuscript/MPESubmission/Revision/response.tex
Minor edits to formatting
Edits to comment 10

@@
 
+% NC: Probably want to change figure to Figure, table to Table, appendix to Appendix to be consistent.
+
 \textbf{RE: MPE-15-254}\\
@@
 
-% TG: actually, is it possible to send them the response as a pdf?
-or will I have to copy paste all that in a mail?
-If the second, it's not that a big problem (just annoying) but I think we should try sending the pdf, it's way clearer like that.
-% NC: You can usually upload a file for the cover letter/response so pdf should be fine. TG: Cool!
-
-% TG: I got rid of all your minor comments for cleaning up a bit.
-
 \section{Reviewer 1:}
@@
 
-We rephrased the highlights as follows: %3 to 5 bullet points (maximum 85 characters, including spaces, per bullet point). TG: probably need improvement here.
-% NC: Yeah I never even looked at these first time round. Mine aren't great but do fit in 85 characters! TG: that looks way better, thanks!
+We rephrased the highlights as follows:
 
@@
 
-% "Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
-using both molecular and morphological data; e.g.
-Wiens et al., 2005; Manos et al., 2007; Pattinson et al., 2014).
-% NC: Argh this is really confusing. It doesn't logically flow
-% For example, Wiens et al. (2005) only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.
-% Furthermore, Pattinson et al. (2014) assess the effect of removing both molecular and morphological data (“artificial extinction”) on the topology.
-% For morphological and molecular data separately, both studies show that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
-% However, they are limited by their empirical-only approach (Wiens et al., 2005) and the scheme of generating missing data (Manos et al., 2007; Pattinson et al., 2014).
-% Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data)." lines 85-98 @@@
-%%% NC: Paragraph above needs work aka a bit of logic to it.
-
-% TG: how's that (+idea flow):
 % 0(before this section)-Missing data can be a problem
 % 1-And until now, not many studies
+
+% NC: Line numbers?
 Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
@@
 % 2-These studies have done X and Y to see if missing data's a problem
-These studies have assessed the effect of missing data on topology either:(1) by comparing a dataset with missing data to sub-sets without missing data (Wiens et al 2005); or (2) by removing both molecular and some morphological data from living taxa (``artificial fossil'') (Manos et al., 2007; Pattinson et al., 2014).
+These studies have assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data (Wiens et al.
+2005); or (2) removing both molecular and some morphological data from living taxa to create artificial fossils (Manos et al., 2007; Pattinson et al., 2014).
 % 3-They show that it's not a problem
-Both types of approaches have shown that missing data is not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
+Both approaches have shown that missing data are not a major problem and should not be an obstacle to combining both living and fossil species in the same phylogenies.
 % 4-But they've done it a certain way
@@
 using the patterns from the fossil record; Manos et al., 2007; Pattinson et al., 2014).
-Besides, the studies using fossil data, mainly on the palaeontological aspect of the question (i.e.
+% NC: This is still a problem (it's ok up to 4). WHY is an empirical-only approach limiting? WHY is it limiting to generate missing data using fossil patterns? 
+Besides, the studies using fossil data, mainly focus on the palaeontological aspect of the question (i.e.
 the effect of missing data in fossils in Total Evidence matrices; Manos et al., 2007; Pattinson et al.
 2014).
+% NC: I don't like all these i.e.s here. Why not be specific in the text rather than throwing everything into brackets? If you're only citing manos and patterson then just say what they did.
 % (next §)5-We're doing it this way
 
-"In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living or fossil data, or all data for certain characters.
-We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa." lines 99-103 @@@
+[lines 99-103 @@@] In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living or fossil data, or all data for certain characters.
+% NC: Could be more specific here and have a sentence that says. "This is an advance on previous studies because..."
+We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades and (ii) the displacement of wild-card taxa.
+% NC: Again how is this an advance. I guess it's related to this sentence from below:
+% (cf. Wiens et al., 2005; Pattinson et al., 2014 measuring respectively only the clades support or the number of shared clades).
 
 % TG: I actually rewrote this bit as well:
+% NC: line numbers?
 We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
-We inferred the topology from the matrices using both Maximum Likelihood and Bayesian inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and for wild-card taxa placement (based on the Triplets distance; Critchlow et al., 1996).
-Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring difference on two aspects of topology (cf.
+We inferred the topology from the matrices using both Maximum Likelihood and Bayesian Inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and for wild-card taxa placement (based on the Triplets distance; Critchlow et al., 1996).
+Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring the effects of missing data on two aspects of topology (cf.
 Wiens et al., 2005; Pattinson et al., 2014 measuring respectively only the clades support or the number of shared clades).
+% NC: This paragraph seems to be repeating what you've already said above but in slightly more detail. Might make more sense to combine this paragraph and the one above it?
 
@@
 The results are available in appendix B; figure B1 and appendix C; figures C6 and C7.
-The ability to recover the ``true'' tree is poor in both methods and using both metrics (Maximum Likelihood 50\% CI approximatively between 0.2 and 0.3 Normalised Robinson-Foulds and Triplets metric; Bayesian consensus tree 50\% CI approximatively between 0.3 and 0.4 Normalised Robinson-Foulds metric and between 0 and 0.2 Normalised Triplets metric).
-This can be related to the size of our matrices or to the ability of the algorithms to generate the ``complete matrix'' but also to our decision to put our minimum median bootstrap acceptance at 50 rather than higher values (line @@@ in the former manuscript and @@@ in our revision).
+The ability to recover the ``true'' tree is poor for both methods and using both metrics (Maximum Likelihood 50\% CI between approximately 0.2 and 0.3 Normalised Robinson-Foulds and Triplets metric; Bayesian consensus tree 50\% CI between approximately 0.3 and 0.4 Normalised Robinson-Foulds metric and between 0 and 0.2 Normalised Triplets metric).
+% NC: What are these results comparing?
+All missing data trees  to the true tree?
+True trees to best trees?
+Clarify.
+This may be related to the size of our matrices or to the ability of the algorithms to generate the ``complete matrix'' but also to our decision to put our minimum median bootstrap acceptance at 50 rather than higher values (line @@@ in the former manuscript and @@@ in our revision).
 When comparing the ``true'' tree to the ``missing data'' trees, the results are \textit{scaled down} due to the low scores of the ``true'' \textit{vs.} ``best'' trees scores.
 However, we can observe the same patterns between the methods as when using the ``best'' tree (i.e. Bayesian always outperforms ML in clade conservation but does poorly in wild-card taxa placement).
-An interesting feature is that the Normalised Robinson-Foulds metrics actually \textit{increases} with missing data.
-This is due to the fact that the difference between the ``best'' and the ``true'' tree is high making trees with enough data more different.
-When data is missing, however, more clades are unresolved and thus closer to clades from the ``true'' trees.
-
-% NC: Worth describing the results here. TG: Done.
+An interesting feature is that the Normalised Robinson-Foulds metric actually \textit{increases} with missing data % NC: specify what is being compared, e.g.
+"when we compare the missing data trees to the ``true'' trees.""
+This is because the difference between the ``best'' and the ``true'' tree is high, making trees with more data more different than those with lots of missing data.
+When data are missing, however, more clades are unresolved and thus closer to clades from the ``true'' trees.
 
@@
 We added the following sentence to the main text to clarify and link to the supplementary analysis:
-% "In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices %NC: What dod you mean by R packages choices?
-Do you mean the way the trees were simulated?
-%  (i.e. diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
-% The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." line 276-280 @@@.
-
-% NC: Needs a little more explanation, and a line about what analysis you did in appendix b.
-So something along the lines of "We compare best trees to missing data trees, but could also compare true trees to the missing data trees.
-We do not do this because...
-However, results of analyses comparing the two can be found in appendix B.
-This makes little difference to our overall results." That might need to be better integrated with the paragraph(s) before.
-% TG: changed to:
+
 "We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
 In practice, the difference between the ``best'' trees and the ``missing data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing data'' trees.
-The difference between the ``true'' and the ``missing data'' tree, however, represents the effect of our parameters used to generate the ``true'' tree and the algorithms used to generated the ``complete'' matrix as well as the effect of our missing data parameters and the phylogenetic methods used.
-Because the main aim of this study, is to look at the effect the missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees to ``missing data'' trees.
+The difference between the ``true'' and the ``missing data'' trees, however, represents the effect of our parameters used to generate the ``true'' tree and the algorithms used to generate the ``complete'' matrix as well as the effect of our missing data parameters and the phylogenetic methods used.
+Because the main aim of this study is to look at the effect our missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees and ``missing data'' trees.
 The results of the comparisons of the ``true'' and the ``missing data'' tree are available in appendix B.
-Note that these results makes little difference to our overall results." line 276-280 @@@.
+Note that this makes little difference to our overall results." line 276-280 @@@.
 
@@
 
-We took on board this comment and agree with the reviewer that if using the ``true'' tree as a starting had an effect it would bias our results.
-We ran, however, an extra analysis to test if that was the case and found that it had no effect of the inferred tree topology.
+We took on board this comment and agree with the reviewer that if using the ``true'' tree as a starting tree had an effect it would bias our results.
+We therefore ran an extra analysis to test if that was the case and found that it had no major effect of the inferred tree topologies.
 We performed this extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains. 
 We compared the topology of these trees in Bayesian inference when using the ``true'' tree as a starting tree or without using a starting tree (i.e.
-using the default in MrBayes which is using a different random tree for each MCMC).
-We found no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
+using the default in MrBayes which is to use a different random tree for each MCMC chain).
+We found no significant difference in topologies between the Bayesian inferences starting with the true tree or a random tree.
 Therefore, we do not believe our results are merely the result of using a specific starting tree topology.
@@
 "We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
-Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly % NC: Significantly as in statisically?
-if not change this word.
-TG: yap, p-value all all the mumbo-jumbo
-affect topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line 323-327 @@@.
+Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line 323-327 @@@.
 See also Appendix A section "Effect of the starting tree on Bayesian inference".
@@
 When this coefficient is $>$0.95, the distributions are significantly similar.
-Values in between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions." line 419-424 @@@.
+Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions." line 419-424 @@@.
 
@@
 \item the way we assess topological recovery is slightly different in each case: for the Bayesian consensus trees, we simply compared the missing-data consensus tree to the best consensus tree; for the Bayesian posterior trees however, we randomly compared 1000 missing-data Bayesian posterior trees to 1000 best Bayesian posterior trees.
-Therefore, even if the comparison methodology is the same, the fact that the second methods draws random trees, we expect the topological recovery score to be blurred in comparison to other comparisons that do not involve random comparisons.
+Therefore, even if the comparison methodology is the same, the fact that the second methods draws random trees means we expect the topological recovery score to be blurred in comparison to other comparisons that do not involve random comparisons.
 \item the Bayesian posterior trees are always resolved and thus, in the same way as the Maximum Likelihood tree, more likely to contain false-positive nodes.
@@
 Concerning the last question of this comment: "How then in comparisons of the distributions are they always so low (Table 1 max prob.
-of overlapping distributions = 0.11)?", we apologise for the confusion here and assume it is due to our mislabelling of the figure 4 (see reviewer 1's comment 12).
+of overlapping distributions = 0.11)?", we apologise for the confusion here and assume it is due to our mislabeling of the figure 4 (see reviewer 1's comment 12).
 In fact, when looking at all parameters separately (figure 4), there is no clear overlap between the Bayesian consensus trees and the Bayesian posterior trees (median Bhattacharyya Coefficient of 0.01; Table 1).

### 2015-08-11 Manuscript/MPESubmission/Revision/response.tex
Edits to end of comments

@@
 \documentclass[12pt,letterpaper]{article}
+
+%%%% NC: FOR CLARITY - can you make the bits that are just copied from the text italic? Sometimes it jumps quickly between a comment to a piece of text. Italics will help make it clearer.%%%%
 
@@
 A Normalised Triplets metric value below 0 means that the placement of some taxa is worse than expected by just randomly placing this taxon in the tree.
-This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
-Even if clades are conserved (figure 5), the resolution within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
+This can be interpreted as the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
+Even if clades are conserved (figure 5), the resolution within them can be poor to non-existent when a lot of data are missing (i.e. 75\%).
 In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
@@
 
-We rewrote the conclusion to deal with this comment and the specific comments 13, 52 (see comment 52 for details on the choice of citing Arcila et al 2015.) and 56.
-% NC: Is it Arcial or Arcila? TG: Dahiana Arcila
-
-% "Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data (Wiens et al 2005; Pattinson et al 2014).
-% NC: I still don't get these limitations.
-How are they are issue?
-What aspects?
-% TG: Limitations are:
-% Empirical data (Wiens et al., 2005; Manos et al.,2007, Pattinson et al., 2014)
-% No fossils (Wiens et al., 2005)
-% Scheme of generating missing data (Manos et al.,2007, Pattinson et al., 2014)
-% Focus on palaeontological aspect of missing data (Manos et al.,2007, Pattinson et al., 2014)
-% Using clade conservation metrics only (Wiens et al., 2005; Pattinson et al., 2014)
+We rewrote the conclusion to deal with this comment (and the specific comments 13, 52 - see comment 52 for details on the choice of citing Arcila et al 2015. - and 56), as follows:
 
@@
 not including fossil taxa, Wiens et al., 2005; or focusing on the palaeontological aspect, Manos et al.,2007, Pattinson et al., 2014).
-%
-This study proposes a more in depth approach where missing data is generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
-This allowed us to confirm previous results that missing data can be especially problematic in small matrices (Wiens 2003) but also revealed the crucial importance of coding morphological data for living species.
+% NC: Again needs a bit more explanation. WHY is empirical approach a problem? Because it only gives answers to one specific scenario or clade?
+Here we used a more in depth approach where missing data were generated from simulated data and according to three clearly defined missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
+This allowed us to confirm previous results that missing data can be especially problematic in small matrices (Wiens 2003), but also revealed the crucial importance of coding morphological data for living species in Total Evidence phylogenies.
 Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
 When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
-Therefore we advise that one should focus on coding morphological characters for a large number of the living taxa present in the matrix (i.e.
-50\%) for accurately combining both living and fossil species in phylogenies.
-Doing so grants a good overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their actual shared derived characters rather than simply on available data."
+Therefore we advise increased focus on coding morphological characters for a large number of the living taxa present in the matrix (i.e.
+50\%) if the goal is to accurately combine both living and fossil species in phylogenies.
+Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data."
 
@@
 Slater, 2013), taxonomic constraints (e.g.
-Slater, 2013; Beck and Lee, 2014) or even by fixing the topology (e.g.
+Slater, 2013; Beck and Lee, 2014) or even on fixing the topology (e.g.
 Ronquist et al., 2012a).
@@
 % TG: maybe add in a non tip-dating analysis way (i.e. Nick Matzke comment)? 
-
-To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological data in the same phylogeny as long as both types of data sufficiently overlap.
+% NC: Sure if it fits. If you want to "cite" Nick you can use (N.Matzke pers.comm.)
+
+To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological morphological data in the same phylogeny as long as there is sufficient overlap.
 Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes." lines 602-634. @@@
+% NC: I don't like this last bit.
+It's really weak.
+I'd cut it to be honest.
+The first sentence is not very clear (suddenly you revert back to neo and paleo) and the second sentence has no real explanation or link to the rest of the paper.
 
@@
 We apologies for this silly mistake and thank the reviewer for pointing it out.
-We changed the colors to match with the paper color scheme: "Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
+We changed the colors to match with the paper's color scheme: "Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
 See figure 4 caption.
@@
 
-We clarified the caption of the Table 1 as follow:
+We clarified the caption of the Table 1 as follows:
 
@@
 Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
-If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology." see Table 1 caption.
+If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology."
+% NC: Do you want to say values > 0.95 are significantly similar, values < 0.05 are significantly different?
 
@@
 We now emphasize more of the recent findings from the missing-data papers in the introduction, and discussion (see comment 4 and 10 above).
-We also included that our results for the $M_F$ parameter for the Normalised Robinson-Foulds metric where in accordance with Manos et al. 2007 in the discussion about the $M_F$ parameter:
-
-"These results are in agreement with Manos et al. (2007) where as few as 16 characters where sufficient for correctly assigning ``artificial fossils'' to their ``correct'' clade." lines @@@
+We also included that our results for the $M_F$ parameter for the Normalised Robinson-Foulds metric were in accordance with Manos et al. 2007 in the discussion about the $M_F$ parameter:
+
+"These results are in agreement with Manos et al. (2007) where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade." lines @@@
+
+% NC: Try not to have too many " ".
 
@@
 
-We modified the following sentence as suggested: "Because fossil taxa do not usually have available DNA, only fossil occurrence dates are used to time calibrate phylogenies" (line 53-54@@@).
+We modified the sentence as suggested.
 
@@
 We modified the sentence as suggested.
-% NC: If it's just you doing exactly what they ask for there is no need to re-write the sentence.
 
@@
 We found results that the $\alpha$ parameter is usually slightly underestimated for the various ``missing-data'' trees unless the number of living species with no morphological data is high ($M_L$ = 75\%), then, the $\alpha$ are overestimated.
-These results are somehow in agreement with our topological differences results: the number of living species seems to be a driving missing data parameter for inferring Total Evidence trees.
+These results are somewhat in agreement with our topological differences results: the number of living species seems to be a driving missing data parameter for inferring Total Evidence trees.
 Nonetheless, we considered this study to already be long and complex.
@@
 what is the effect of missing data on the morphological substitution rate parameter estimates?) might lose the reader and make an already long paper even longer.
-Therefore, we added the result of these analyses to \textbf{Appendix-A Morphological rates estimations results}.
+Therefore, we added the results of these analyses to \textbf{Appendix-A Morphological rates estimations results}.
 We link to this analysis in the methods as follows:
@@
 See also Appendix A "Morphological rates estimations results".
+
+% NC: Needs a bit more explanation. Something like, we also include in the appendix an analysis showing ... though it is beyond the scope of this paper to discuss these results further here.
 
@@
 We replace 'inferred' by 'simulated' when referring to simulating data throughout the manuscript.
-% NC: Check this - I'm sure I caught one up above in the comments here.
-TG: I think the idea is that we didn't "infer" any matrices, we simulated/generated/created them.
-The trees are still inferred though.
 
@@
 
-Since homoplasy is a difficult concept to measure (e.g. with character exhaustion; Wagner 2000, Evolution) so we didn't really add a value for ``too much'' homoplasy.
-We therefore removed this mention and changed the sentence to:
+Since homoplasy is a difficult concept to measure (e.g. with character exhaustion; Wagner 2000, Evolution) it is difficult to assign a value for ``too much'' homoplasy.
+We therefore removed this comment and changed the sentence to:
 "In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal (Hassanin et al., 1998; Estoup et al., 2002).%\citep{Hassanin1998611,EstoupHomoplasy}.
-Also, we chose this low value to be consistent with our protocol for simulating morphological characters (see below)." (line 189-192@@@).
+Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below)." (line 189-192@@@).
 
@@
 We initially chose this parameter from personal communication with April Wright and to be consistent with the molecular rate parameter.
-In figure 3 from Wright and Hillis 2014, even if the topological error is high when rates are near 0, it is mainly the case for parsimony and both parsimony and Mk with missing data, the fitted model for the Mk model with no missing data (light purple line) is still lowest when the rates are near 0.
+In figure 3 from Wright and Hillis 2014, even if the topological error is high when rates are near 0, as is mainly the case for parsimony and both parsimony and Mk with missing data, the fitted model for the Mk model with no missing data (light purple line) is still lowest when the rates are near 0.
 Topological error, however, becomes minimal for all scenarios (Parsimony and Mk, with or without missing data) at a rate of 0.5 (see modified figure 3 from Wright and Hillis (2014) in this response; we added the tick marks and the 0.5 rate limit by using the ruler in Adobe Illustrator).
@@
 
-We rephrased the section underlined by the reviewer as follows:
-
-"We used low evolutionary rate parameters to be consistent with the molecular rate parameters and to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal (Wright and Hillis, 2014).
-In fact, topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model (Lewis, 2001; Wright and Hillis, 2014).
-Note, however, that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increase variance in topological error but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004)." lines 211-217 @@@
-
-% NC: Might need a bit more explanation.
-For example mentioning that "Topological error, however, becomes minimal for all scenarios (Parsimony and Mk, with or without missing data) at a rate of 0.5 (see figure 3 from Wright and Hillis (2014)" as you've mentioned above.
-% TG: added a sentence.
+We rephrased the section highlighted by the reviewer as follows:
+
+"We used low evolutionary rate parameters to be consistent with the molecular rate parameters, to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal (Wright and Hillis, 2014).
+Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model (Lewis, 2001; Wright and Hillis, 2014).
+Note, however, that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004) so this should not influence our results." lines 211-217 @@@
 
@@
 
-We now refer to two more classical citations Bremer and Struwe 1992 (American Journal of Botany) and Patterson et al.
-1993 (Annual Review of Ecology and Systematics) on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to having a small matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters) lines 578-579@@@.
+We now refer to two more classical citations: Bremer and Struwe 1992 (American Journal of Botany) and Patterson et al.
+1993 (Annual Review of Ecology and Systematics), on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to having a small matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters) lines 578-579@@@.
 
@@
 
-We modified the sentence as suggested: "they have a larger proportion of missing data" (line 597@@@).
+We modified the sentence as suggested.
 
@@
 "Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
-For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combination of the other parameters ($M_F$ and $N_C$; Figure 6).
-In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) does not matter and an increase/decrease in both does not affect topology.
+For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference among trees with any combination of the other parameters ($M_F$ and $N_C$; Figure 6).
+In other words, when the number of missing living taxa reaches 50\%, neither the amount of missing data in the fossil record ($M_F$), nor the number of characters in the matrix ($N_C$) affect topology.
 A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
-This has important practical implications, especially for the best strategy to improve topology by collecting morphological characters (see below)." lines 589-609@@@
+This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below)." lines 589-609@@@
 
@@
 As discussed above (reviewer 1, specific comment 29), we understand the reviewer's concern and have added an analysis in Appendix A (figure A2 and tables A1 and A2) where we tested the effect of using the ``true'' tree as a starting tree.
-We found that there is no significant effect of using the ``true'' tree or a random tree (MrBayes default) as a starting tree on topological recovery.
+We found that there is no significant effect of using the ``true'' tree instead of a random tree (MrBayes default) as a starting tree on topological recovery.
 
@@
 "Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'') for details on the analysis of the effect of the starting tree in Bayesian inference)." lines 618-624@@@
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details on the analysis of the effect of the starting tree in Bayesian inference)." lines 618-624@@@
 
@@
 We therefore used this reference to strengthen our claim of the superiority of the Bayesian consensus tree over the Maximum Likelihood one.
+
 Secondly, for the question about molecular scaffolds, from the few studies clearly specifying their topological constraint method, only Beck and Lee 2013 (Proceedings B) seems to compare an analysis with and without taxonomic constraint on divergence time and rate estimation (the constraint seems to have no effect) but the branch length aspect is not covered in our study.
@@
 
-We removed the sentence about the t-test similarity from both figure captions. %TG: not sure if that's what the reviewer wanted. % NC: Yeah I think that's what they meant
+We removed the sentence about the t-test similarity from both figure captions.
 
@@
 Conversely, orange regions denote a high overlap in Normalised metric values between the different parameter combinations (i.e.
-the parameters have a weak effect on the metric and thus the topological recovery)." See figure 6 caption.
+the parameters have a weak effect on the metric and thus the topological recovery)."
 
@@
 "Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Figure 6).
-In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6)." lines 434-439@@@
+In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6)." lines 434-439@@@
 
@@
 The orange upper triangle in Figure 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Figure 6A).
-% There is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-% Additionally, when using the Normalised Robinson-Foulds metric, once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-% NC: This sentence and the one before it seem to be the same thing? TG: Yep, I modified to:
 Once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
 Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
-In figure 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%." lines 480-497@@@
+In Figure 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%." lines 480-497@@@
 
@@
 
+We dealt with the caption of the Table 1 in reviewer 1 comment 14 above.
+
+For the other two points:
 1) We are not entirely sure if the reviewer means (i) why the two methods display such differences ; or (ii) why, within each method, the combination of parameters displays a difference between each single parameters ($M_LM_FM_C$ vs.
@@
 
-2) We followed the reviewer's suggestion and added 6 extra tables in the Appendix C (Table C8 to C13) containing the marginal probability of each missing data parameter individually and each tree inference method.
-We considered running the same type of analysis for the combination of the parameter, however, this involved a table of 5*126 cells which was highly impractical to read.
-% TG: Not sure if the last bit is necessary...
+2) We followed the reviewer's suggestion and added six extra tables to Appendix C (Table C8 to C13) containing the marginal probability of each missing data parameter individually and for each tree inference method.
 
@@
 
-We removed the repetition in the caption.
-If we understood the second point correctly, the reviewer suggests we add ``\textit{best}'' to the two first rows of the tables, and by ``\textit{best}'', the reviewer means ``optimal'' (as opposed to the ``best'' tree in our protocol that refers to any tree inferred by any method from the ``complete matrix''.
-We don't think it is necessary to follow this suggestion as the Maximum Likelihood tree is by definition the tree with the highest likelihood, and the Bayesian consensus is the consensus tree of the Bayesian inference trees.
-So adding the word ``\textit{best}'' here is not entirely accurate.
+We removed the repetition in the caption and changed the first two rows as suggested.
+%If we understood the second point correctly, the reviewer suggests we add ``\textit{best}'' to the two first rows of the tables, and by ``\textit{best}'', the reviewer means ``optimal'' (as opposed to the ``best'' tree in our protocol that refers to any tree inferred by any method from the ``complete matrix''.
+% We don't think it is necessary to follow this suggestion as the Maximum Likelihood tree is by definition the tree with the highest likelihood, and the Bayesian consensus is the consensus tree of the Bayesian inference trees.
+%So adding the word ``\textit{best}'' here is not entirely accurate.
+
+% NC: No we discussed this last night - I think they *do* mean "best" as in our protocol. So just make the changes to the first two rows.
 
@@
 We updated the README page of the GitHub repository linking to the various R code to make it clearer what files are needed.
-We also now provide a version of the analysis in R run thoughout the manuscript as a step by step tutorial for a subset of data.
+We also now provide a version of the analysis in R run throughout the manuscript as a step by step tutorial for a subset of data.
 See \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data/blob/master/Analysis/Analysis_Demo.R} and Appendix D.
-% NC: Maybe add something like this if this is your intention. It's probably worth it. No need to keep lots of un-necessary files. TG: Done and agree
 We will also clean up this repository once the paper has been accepted.
@@
 
-A similar point/suggestion was also raised by reviewer 1 and we ran the suggested extra analysis (see Reviewer 1 General comment number 6 and specific comment number 25).
-We understand the concerns of the reviewer but, we do not agree that comparing the "missing-data" trees to the "best" trees adds a layer of complexity to our simulations.
-In practice, the ``true'' tree can be perceived as a ``random'' seed to generate the ``complete'' matrix.
-The ``true'' tree is simply a birth-death tree with $\lambda$ $>$ $\mu$ and a set number of living and fossil taxa (see details in the manuscript section "2.1 Generating the matrix").
-All these parameters, were only used to create a realistic ``complete'' matrix in order to make our results relevant to biologists.
-The aim of this study was to assess how good will a topology inferred from a matrix with missing data (which is likely in biology) compared to the ones inferred form the ``complete'' matrix (the ``best'' tree).
-
-% TG: maybe add a sentence in the manuscript?
-
-% NC: This argument [TG: the one commented out below] makes no sense. The point the reviewer is making is that we go:
-% True tree > Best tree > comparisons when we could just go True tree > comparisons.
-So our approach *is* more complex than their suggestion.
-Instead the justification here is the same as used for reviewer one - that in real life scenarios we never know the true tree and we want it to be as close to real life as possible.
-BUT that we have tested their idea and find it makes no difference but all the results are in the appendix.
-% TG: you're right, after reflection, I think we should have never mentioned the ``true'' tree and just say in the methods that we used a random Birth-Death tree with such and such parameters to generate the matrix.
-It's kind of too late to backup by now but the argument was just kind of saying: "we don't care if we can recover the "true" tree because we don't give a flip on how the algorithms perform at making a matrix".
-It's all random anyway...
-% TG : I've commented out the non-sense justification below and wrote something shorter above (and referring to reviewer's 1 comment).
-
-% Because we use a simulation approach to test this problem, we have to start from a random starting point (the "true" tree) and then do the following steps: (1) generate and remove the data (2) infer the topologies and (3) compare the topologies.
-% Note that in this schematic representation, the "true" tree is the step (0) of our analysis and involves a series of algorithms and software (i.e.
-seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes).
-% To compare the "true" tree's topology to the "missing-data" topologies will be equivalent of exploring the effect of our three parameters on recovering topology through 3 layers (1- generate the data; 2- infer the topologies; and 3- compare the topologies).
-% By comparing the "best" tree's topology to the "missing-data" ones, we explore the effect of our three parameters through only 2 layers (2- infer the topologies; and 3- compare the topologies).
-% Note that we explore the effect of these 2 layers by using two different tree inference methods (effectively 4 since we also use the Maximum Likelihood bootstraps trees and the Bayesian posterior trees) and 2 different Normalised topological metrics (Robinson-Foulds and Triplets).
-% Using as main results the comparisons between the "true" and the "missing-data" trees would also need the exploration of the effect of the cascade of algorithms and softwares (maybe by modifying the algorithms used to generate the "true" tree and the matrices?).
-
-% However, we have added analyses with comparisons between the "true" tree and the "missing-data" trees in the Appendix C (Figures C6 and C7).
-% It is interesting to note that the algorithms and software as well as probably the size of our matrices are actually not ideal for the recovery of the "true" tree is rather low and scales down the our results (see figure @).% NC: HUH????
-% However, even though these comparisons are worse than comparing the best tree to the missing data trees, it is encouraging to see that the patterns in the results are conserved (i.e.
-Bayesian consensus trees $>$ Maximum Likelihood trees $>$ both bootstraps trees and Bayesian posterior trees).
+A similar point/suggestion was also raised by reviewer 1 and we ran the suggested extra analysis. See reviewer 1 general comment number 6 and specific comment number 25 for a detailed response.
+
+We understand the concerns of the reviewer but, we would prefer to stick with our "best" tree approach.
+In practice, the ``true'' tree can be perceived as a ``random'' seed to generate the ``complete'' matrix, as it simply a birth-death tree with $\lambda$ $>$ $\mu$ and a set number of living and fossil taxa (see details in the manuscript section "2.1 Generating the matrix").
+All these parameters were only used to create a realistic ``complete'' matrix in order to make our results relevant to biologists.
+
+% NC: Can you rephrase a little here so it sounds less like you saying "no, and you're stupid".
+Doesn't the response to reviewer 1 cover this already?
+I've had a go but you may need to check what I've written.
 
@@
 Hereafter, we call these matrices with missing morphological data the ``missing-data'' matrices.
-% NC: add clarification here about molecular data
-Note that the amount of molecular data remained constant throughout our simulations: 1000 molecular characters for each living taxa and no molecular data for fossil taxa (see above).
-% NC: Is this correct? TG: yap ;).
+Note that the amount of molecular data remained constant throughout our simulations: 1000 molecular characters for living taxa and no molecular data for fossil taxa (see above).
 We removed morphological data using three data incompleteness parameters:
@@
 Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
-Instead this 50\% refers to the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a minimum of 5\%, see below).
+Instead this 50\% refers to missing fossil data across the whole matrix.
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a maximum of 95\% missing data, see below).
+% NC: normally you'd say up to a maximum rather than up to a minimum.
+Is what I've written correct?
 \item{The number of morphological characters for both living and fossil taxa ($N_C$).} This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
@@
 
-In practice, each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
+Each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
 Note that $M_L$  and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
@@
 Biol.) but we rather treated the missing data in the "whole" fossil record.
-Because our fossil taxa and their characters where randomly simulated, we didn't had any \textit{a priori} assumption on how to remove the data.
+Because our fossil taxa and their characters where randomly simulated, we didn't have any \textit{a priori} assumption on how to remove the data.
 We therefore decided to randomly remove a certain amount of data among all the fossils.
@@
 
-% NC: OK - I'd add a line somewhere in the manuscript to make this clear. something like:
-% TG: I've added the following to the new removing data section:
-
-We added the following precision to the explanation on removing data (See our response to reviewer 2's comment 1):
+We added the following comment to the explanation on removing data (See our response to reviewer 2's comment 1):
 "Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
@@
 
-% Also, a second argument for using this slightly simplistic way to deal with the missing data in the fossil record follows our main recommendation from this study: the best way to improve the quality of Total Evidence trees is to have a maximum number of living species with available morphological data.
-% In fact lines 584 to 589 in the first version of the manuscript (line 669-674@@@ in the revised version), we argue that even if the effect of the $M_F$ parameter is important, there is unfortunately not much we can do about it since the completeness of this part of the matrix solely depends on exceptional palaeontological discoveries and excavation effort.
-% We feel that these parameters, however crucial for our understanding of macroevolution, are unfortunately not predicable and a recommendation such as "we need more fossil data" is less practical than our main recommendation: "we need more living species data".
-
-% NC: Not sure the second paragraph is needed.
-The reviewer is saying that they want to know what happens if living taxa have all their data but fossils have variable amounts.
-And I think this is what you've done for some parameter combinations.
-Just need to apologise for not being clear and clarify in the text how data were missing from fossils.
-Doesn't need so much waffle and defensiveness - they are asking you to do what you've already done!
-:)
-
 \end{enumerate}
 
-Please let us know if you require any further information,\\
+We hope we have responded to all these comments appropriately. Please let us know if you require any further information,\\
 \bigskip

### 2015-08-11 Manuscript/MPESubmission/Revision/response.tex

@@
 
+=======
+%\linespread{2}
 \raggedright
@@
 
-We added the following comment to the explanation on removing data (See our response to reviewer 2's comment 1):
-"Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
-Instead this 50\% refers to the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a minimum of 5\%, see below)."
+We added the following comment to the explanation on removing data (See our response to reviewer 2's comment 1).
+
+Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
+Instead this 50\% refers to missing fossil data across the whole matrix.
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a maximum of 95\% missing data, see below).
 

### 2015-08-11 Manuscript/MPESubmission/Revision/response.tex
Added quotes from manuscript in italic

@@
 
-=======
 %\linespread{2}
@@
 
-% NC: Line numbers?
-Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
+% NC: Line numbers? TG: I'll do all line numbering once I've incorporated all the edits to the manuscript.
+\textit{Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
 using both molecular and morphological data; e.g.
@@
 the effect of missing data in fossils in Total Evidence matrices; Manos et al., 2007; Pattinson et al.
-2014).
+2014).} lines @@@.
 % NC: I don't like all these i.e.s here. Why not be specific in the text rather than throwing everything into brackets? If you're only citing manos and patterson then just say what they did.
@@
 
-[lines 99-103 @@@] In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living or fossil data, or all data for certain characters.
+\textit{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living or fossil data, or all data for certain characters.
 % NC: Could be more specific here and have a sentence that says. "This is an advance on previous studies because..."
-We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades and (ii) the displacement of wild-card taxa.
+We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades and (ii) the displacement of wild-card taxa.} lines @@@
 % NC: Again how is this an advance. I guess it's related to this sentence from below:
@@
 % NC: line numbers?
-We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
+\textit{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
 We inferred the topology from the matrices using both Maximum Likelihood and Bayesian Inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and for wild-card taxa placement (based on the Triplets distance; Critchlow et al., 1996).
 Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring the effects of missing data on two aspects of topology (cf.
-Wiens et al., 2005; Pattinson et al., 2014 measuring respectively only the clades support or the number of shared clades).
+Wiens et al., 2005; Pattinson et al., 2014 measuring respectively only the clades support or the number of shared clades).} lines @@@
 % NC: This paragraph seems to be repeating what you've already said above but in slightly more detail. Might make more sense to combine this paragraph and the one above it?
@@
 
-"We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
+\textit{We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
 In practice, the difference between the ``best'' trees and the ``missing data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing data'' trees.
@@
 The results of the comparisons of the ``true'' and the ``missing data'' tree are available in appendix B.
-Note that this makes little difference to our overall results." line 276-280 @@@.
+Note that this makes little difference to our overall results.} line 276-280 @@@.
 
@@
 
-"We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
-Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'')." line 323-327 @@@.
+\textit{We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
+Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').} line 323-327 @@@.
 See also Appendix A section "Effect of the starting tree on Bayesian inference".
@@
 
-"The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 (full overlap; Bhattacharyya, 1943, see Appendix B for calculation details)." lines @@@
-
-"When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
+\textit{The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 (full overlap; Bhattacharyya, 1943, see Appendix B for calculation details).} lines @@@
+
+\textit{When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
 When this coefficient is $>$0.95, the distributions are significantly similar.
-Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions." line 419-424 @@@.
+Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.} line 419-424 @@@.
 
@@
 
-"Additionally, the Bayesian posterior trees performed more poorly than the Bayesian consensus tree (figure 4, table 1 and appendix C figure C5 and tables C5, C6 and C7).
+\textit{Additionally, the Bayesian posterior trees performed more poorly than the Bayesian consensus tree (figure 4, table 1 and appendix C figure C5 and tables C5, C6 and C7).
 This may be because the Bayesian posterior trees are always resolved and thus more likely to contain incorrectly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
 Conversely, the Bayesian consensus tree might not resolve nodes that are poorly supported and thus is more likely to contain only correctly resolved nodes (i.e.
-increasing the Normalised Robinson-Foulds metric)." lines 652-658.
+increasing the Normalised Robinson-Foulds metric).} lines 652-658.
 @@@
@@
 
-"The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (figure 5).
+\textit{The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (figure 5).
 A Normalised Triplets metric value below 0 means that the placement of some taxa is worse than expected by just randomly placing this taxon in the tree.
@@
 These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters (Wiens 2003).
-It is important to note, however, that this effect can be reduced by increasing the number of characters (Wiens 2003)." lines 632-644. @@@
+It is important to note, however, that this effect can be reduced by increasing the number of characters (Wiens 2003).} lines 632-644. @@@
 
@@
 % TG: I rephrased as follows.
-Previous studies have explored the effect of missing morphological data in Total Evidence matrices (Wiens et al., 2005; Manos et al.,2007, Pattinson et al., 2014).
+\textit{Previous studies have explored the effect of missing morphological data in Total Evidence matrices (Wiens et al., 2005; Manos et al.,2007, Pattinson et al., 2014).
 The conclusions of theses studies, however, were limited by their empirical approach as well as several aspects of their experimental design (e.g.
@@
 50\%) if the goal is to accurately combine both living and fossil species in phylogenies.
-Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data."
-
-Additionally, the topology of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data.} lines @@@
+
+\textit{Additionally, the topology of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
 This has also been observed in empirical data (e.g. Arcila et al., 2015) where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
@@
 Ronquist et al., 2012a).
-Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.
+Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.} lines @@@
 % TG: maybe add in a non tip-dating analysis way (i.e. Nick Matzke comment)? 
@@
 
-To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological morphological data in the same phylogeny as long as there is sufficient overlap.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes." lines 602-634. @@@
+\textit{To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological morphological data in the same phylogeny as long as there is sufficient overlap.
+Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.} lines 602-634. @@@
 % NC: I don't like this last bit.
@@
 We apologies for this silly mistake and thank the reviewer for pointing it out.
-We changed the colors to match with the paper's color scheme: "Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
+We changed the colors to match with the paper's color scheme: \textit{Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)}.
 See figure 4 caption.
@@
 
-"Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
+\textit{Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
 Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
-If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology."
+If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology."}
 % NC: Do you want to say values > 0.95 are significantly similar, values < 0.05 are significantly different?
@@
 
-"These results are in agreement with Manos et al. (2007) where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade." lines @@@
+\textit{These results are in agreement with Manos et al. (2007) where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.} lines @@@
 
@@
 
-We revised this statement as follows: "many large-scale macroevolutionary studies focus solely on living species" (line 30-31 @@@).
+We revised this statement as follows: \textit{many large-scale macroevolutionary studies focus solely on living species} (line 30-31 @@@).
 
@@
 
-We modified the sentence as follows: "These approaches differ mainly in how they treat fossil taxa and their data.
-One can use fossils as tips or as nodes in the phylogeny, and can use only the age of the fossils, only the morphology of the fossils, or age and morphology jointly." (line 42-44@@@)
+We modified the sentence as follows: \textit{These approaches differ mainly in how they treat fossil taxa and their data.
+One can use fossils as tips or as nodes in the phylogeny, and can use only the age of the fossils, only the morphology of the fossils, or age and morphology jointly.} (line 42-44@@@)
 
@@
 
-We modified the statement to: "but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints" (line 111-113@@@).
+We modified the statement to: \textit{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints} (line 111-113@@@).
 
@@
 
-"The detailed MrBayes parameters are available in Appendix A along with details on the $\alpha$ parameter estimation." lines 336-338 @@@
+\textit{The detailed MrBayes parameters are available in Appendix A along with details on the $\alpha$ parameter estimation.} lines 336-338 @@@
 See also Appendix A "Morphological rates estimations results".
@@
 We therefore removed this comment and changed the sentence to:
-"In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal (Hassanin et al., 1998; Estoup et al., 2002).%\citep{Hassanin1998611,EstoupHomoplasy}.
-Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below)." (line 189-192@@@).
+\textit{In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal (Hassanin et al., 1998; Estoup et al., 2002).%\citep{Hassanin1998611,EstoupHomoplasy}.
+Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below).} (line 189-192@@@).
 
@@
 
-"This model and these parameter settings strike a balance between realism for empirical datasets  (e.g.
-Douady, Kelly et al., 2014) and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)".
+\textit{This model and these parameter settings strike a balance between realism for empirical datasets  (e.g.
+Douady, Kelly et al., 2014) and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)}.
 lines @@@
@@
 
-"We used low evolutionary rate parameters to be consistent with the molecular rate parameters, to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal (Wright and Hillis, 2014).
+\textit{We used low evolutionary rate parameters to be consistent with the molecular rate parameters, to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal (Wright and Hillis, 2014).
 Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model (Lewis, 2001; Wright and Hillis, 2014).
-Note, however, that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004) so this should not influence our results." lines 211-217 @@@
+Note, however, that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004) so this should not influence our results.} lines 211-217 @@@
 
@@
 
-We modified the sentence as follows: "To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon." (line 264-267@@@).
+We modified the sentence as follows: \textit{To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.} (line 264-267@@@).
 
@@
 
-We modified the sentence as follows: "The sizes of our data matrices were constrained by the performance of our protocol" (line 579-580@@@).
+We modified the sentence as follows: \textit{The sizes of our data matrices were constrained by the performance of our protocol} (line 579-580@@@).
 
@@
 
-"Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
+\textit{Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
 For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference among trees with any combination of the other parameters ($M_F$ and $N_C$; Figure 6).
@@
 A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
-This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below)." lines 589-609@@@
+This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).} lines 589-609@@@
 
@@
 
-"Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details on the analysis of the effect of the starting tree in Bayesian inference)." lines 618-624@@@
+\textit{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details on the analysis of the effect of the starting tree in Bayesian inference).} lines 618-624@@@
 
@@
 We revised the paragraph as follows:
-"Another practical implication of our results regards the tree inference methods.
+\textit{Another practical implication of our results regards the tree inference methods.
 Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advise that where a topological constraint is needed, Bayesian consensus trees should be used.
@@
 2012 trees used in Healy et al.
-2014)." lines 680-700@@@
+2014).} lines 680-700@@@
 
@@
 We modified the figure caption as follows:
-"The effects of missing data on topological recovery using Bayesian consensus trees.
+\textit{The effects of missing data on topological recovery using Bayesian consensus trees.
 Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $N_{C}$ (lower line).
@@
 Conversely, orange regions denote a high overlap in Normalised metric values between the different parameter combinations (i.e.
-the parameters have a weak effect on the metric and thus the topological recovery)."
+the parameters have a weak effect on the metric and thus the topological recovery).}
 
 We modified the methods section as follows:
-"Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Figure 6).
-In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6)." lines 434-439@@@
+\textit{Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Figure 6).
+In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6).} lines 434-439@@@
 
 We modified the results section as follows:
-"Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+\textit{Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
@@
 Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
-In Figure 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%." lines 480-497@@@
+In Figure 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.} lines 480-497@@@
 
@@
 
-"To explore the effect of missing morphological data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
+\textit{To explore the effect of missing morphological data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
 Hereafter, we call these matrices with missing morphological data the ``missing-data'' matrices.
@@
 Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).
-\end{enumerate}
-
-Each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
-Note that $M_L$  and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
-
-We created matrices using all parameter combinations resulting in 125 ($5^3$) “missing-data” matrices.
+\end{enumerate}}
+
+\textit{Each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
+Note that $M_L$  and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.}
+
+\textit{We created matrices using all parameter combinations resulting in 125 ($5^3$) “missing-data” matrices.
 Note that one of these combinations ($M_L$=0\%; $M_F$=0\% and $N_C$=100) has no missing data so is equivalent to the “complete” matrix, thus we have one effectively complete matrix in our 125 “missing-data” matrices.
@@
 To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
-Note that the living taxa always had at least 90\% of data (the 1000 molecular characters)." lines 222-267@@@
+Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).} lines 222-267@@@
 
@@
 
-Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
+\textit{Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
 Instead this 50\% refers to missing fossil data across the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a maximum of 95\% missing data, see below).
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a maximum of 95\% missing data, see below).}
 

### 2015-08-11 Manuscript/MPESubmission/Revision/response.tex
updated response

@@
 % 4-But they've done it a certain way
-However, these studies were limited by their empirical-only approach and the use of only living taxa (Wiens et al., 2005) or the scheme of generating missing data (i.e.
-using the patterns from the fossil record; Manos et al., 2007; Pattinson et al., 2014).
 % NC: This is still a problem (it's ok up to 4). WHY is an empirical-only approach limiting? WHY is it limiting to generate missing data using fossil patterns?
-Besides, the studies using fossil data, mainly focus on the palaeontological aspect of the question (i.e.
-the effect of missing data in fossils in Total Evidence matrices; Manos et al., 2007; Pattinson et al.
-2014).} lines @@@.
-% NC: I don't like all these i.e.s here. Why not be specific in the text rather than throwing everything into brackets? If you're only citing manos and patterson then just say what they did.
+These studies, however, were limited by because of their empirical approach limiting the conclusions of the studies to similar scenarios. %TG: or scenari? Reviewer 1 is picky with these things
+Additionally, Wiens et al.
+(2005) didn't include fossil taxa in their analyse thus not allowing to test the placement of fossil taxa; and Manos et al.
+(2007) and Pattinson et al.
+(2014) used only patters of the fossil record to generate missing data thus not allowing to test the effect of missing data for other patterns observed for living taxa (e.g.
+the complete absence of morphological data; Guillerme & Cooper 2015).
+% TG: Does that sound better?
+Besides, Manos et al. (2007) and Pattinson et al. (2014), mainly focused on the palaeontological aspect of the question.} lines @@@.
+% NC: I don't like all these i.e.s here.
+Why not be specific in the text rather than throwing everything into brackets?
+If you're only citing manos and patterson then just say what they did.
+TG: no more i.es then ;).
 % (next §)5-We're doing it this way
 
-\textit{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living or fossil data, or all data for certain characters.
-% NC: Could be more specific here and have a sentence that says. "This is an advance on previous studies because..."
-We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades and (ii) the displacement of wild-card taxa.} lines @@@
+\textit{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living taxa with morphological data, fossil data, all data for certain characters and the combination of these three aspects.
+This is an advance on previous studies because we use large scale simulations and and analyse the effect of three distinct aspect of missing data thus focusing on both neontological and palaeontological parts of the matrix.
+Besides, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and (ii) the displacement of wild-card taxa (based on the Triplets distance; Critchlow et al., 1996) rather than just the clade conservation or support (cf.
+Wiens et al., 2005; Pattinson et al., 2014).} lines @@@
 % NC: Again how is this an advance. I guess it's related to this sentence from below:
@@
 % NC: line numbers?
-\textit{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
-We inferred the topology from the matrices using both Maximum Likelihood and Bayesian Inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and for wild-card taxa placement (based on the Triplets distance; Critchlow et al., 1996).
-Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring the effects of missing data on two aspects of topology (cf.
+% \textit{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
+% We inferred the topology from the matrices using both Maximum Likelihood and Bayesian Inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and for wild-card taxa placement (based on the Triplets distance; Critchlow et al., 1996).
+% Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring the effects of missing data on two aspects of topology (cf.
 Wiens et al., 2005; Pattinson et al., 2014 measuring respectively only the clades support or the number of shared clades).} lines @@@
-% NC: This paragraph seems to be repeating what you've already said above but in slightly more detail. Might make more sense to combine this paragraph and the one above it?
+% NC: This paragraph seems to be repeating what you've already said above but in slightly more detail. Might make more sense to combine this paragraph and the one above it? TG: Ok, I've combined them
 
@@
 The results are available in appendix B; figure B1 and appendix C; figures C6 and C7.
-The ability to recover the ``true'' tree is poor for both methods and using both metrics (Maximum Likelihood 50\% CI between approximately 0.2 and 0.3 Normalised Robinson-Foulds and Triplets metric; Bayesian consensus tree 50\% CI between approximately 0.3 and 0.4 Normalised Robinson-Foulds metric and between 0 and 0.2 Normalised Triplets metric).
+The ability to recover the ``true'' tree is poor for both methods and using both metrics.
+When comparing the ``true'' tree to the ``best'' trees, in Maximum Likelihood the 50\% CI lays between approximately 0.2 and 0.3 for the Normalised Robinson-Foulds and Triplets metric; and, in Bayesian consensus tree  the 50\% CI lays between approximately 0.3 and 0.4 for the Normalised Robinson-Foulds metric and between 0 and 0.2 for the Normalised Triplets metric).
 % NC: What are these results comparing?
@@
 Clarify.
+TG: is that ok?
 This may be related to the size of our matrices or to the ability of the algorithms to generate the ``complete matrix'' but also to our decision to put our minimum median bootstrap acceptance at 50 rather than higher values (line @@@ in the former manuscript and @@@ in our revision).
@@
 However, we can observe the same patterns between the methods as when using the ``best'' tree (i.e. Bayesian always outperforms ML in clade conservation but does poorly in wild-card taxa placement).
-An interesting feature is that the Normalised Robinson-Foulds metric actually \textit{increases} with missing data % NC: specify what is being compared, e.g.
+An peculiar feature is that the Normalised Robinson-Foulds metric actually \textit{increases} when we compare the ``missing data'' trees to the ``true'' tree.
+% NC: specify what is being compared, e.g.
 "when we compare the missing data trees to the ``true'' trees.""
@@
 \textit{Previous studies have explored the effect of missing morphological data in Total Evidence matrices (Wiens et al., 2005; Manos et al.,2007, Pattinson et al., 2014).
-The conclusions of theses studies, however, were limited by their empirical approach as well as several aspects of their experimental design (e.g.
-not including fossil taxa, Wiens et al., 2005; or focusing on the palaeontological aspect, Manos et al.,2007, Pattinson et al., 2014).
+The conclusions of theses studies, however, were limited by their empirical approach making their results applicable only to similar scenarios. %TG: or scenari
+Additionally, these studies focused either only on living taxa (Wiens et al., 2005) or on the patterns of missing data from the fossil record only (Manos et al.,2007, Pattinson et al., 2014).
+The conclusions of these studies were therefore restricted to these specific aspect of missing data and, besides their interesting conclusions, % TG: seems unfair to me to just bash them, they are actually nice papers
+they didn't provide any insight of the combined effect of different missing data parameters.
+% as well as several aspects of their experimental design (e.g. not including fossil taxa, Wiens et al., 2005; or focusing on the palaeontological aspect, Manos et al.,2007, Pattinson et al., 2014).
 % NC: Again needs a bit more explanation. WHY is empirical approach a problem? Because it only gives answers to one specific scenario or clade?
@@
 % NC: Sure if it fits. If you want to "cite" Nick you can use (N.Matzke pers.comm.)
-
-\textit{To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological morphological data in the same phylogeny as long as there is sufficient overlap.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.} lines 602-634. @@@
+% TG: ok, actually leave it, I've tried something but the cite wasn't that convincing and we're already being carefull enough here.
+
+% \textit{To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological morphological data in the same phylogeny as long as there is sufficient overlap.
+% Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.} lines 602-634. @@@
 % NC: I don't like this last bit.
@@
 The first sentence is not very clear (suddenly you revert back to neo and paleo) and the second sentence has no real explanation or link to the rest of the paper.
+TG: let's just remove it then...
+You're right it's totally useless and after reading the whole paper, the reader won't mind saving a few more second of their life.
 
@@
 Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
-If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology."}
-% NC: Do you want to say values > 0.95 are significantly similar, values < 0.05 are significantly different?
+If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology.
+Values $>$ 0.95 denote significantly similar distributions and values $<$ 0.05 denote significantly different distributions."}
+% NC: Do you want to say values > 0.95 are significantly similar, values < 0.05 are significantly different?  TG: yes.
 
@@
 \textit{These results are in agreement with Manos et al. (2007) where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.} lines @@@
-
-% NC: Try not to have too many " ".
 
@@
 
-\textit{The detailed MrBayes parameters are available in Appendix A along with details on the $\alpha$ parameter estimation.} lines 336-338 @@@
+\textit{The detailed MrBayes parameters are available in Appendix A.
+We also included analysis showing details on the $\alpha$ parameter estimation though it is beyond the scope of this paper and are not discussed further here.} lines 336-338 @@@
 See also Appendix A "Morphological rates estimations results".
 
-% NC: Needs a bit more explanation. Something like, we also include in the appendix an analysis showing ... though it is beyond the scope of this paper to discuss these results further here.
+% NC: Needs a bit more explanation.
+Something like, we also include in the appendix an analysis showing ...
+though it is beyond the scope of this paper to discuss these results further here.
+TG: is that ok?
 
@@
 
-% NC: No we discussed this last night - I think they *do* mean "best" as in our protocol. So just make the changes to the first two rows.
+% NC: No we discussed this last night - I think they *do* mean "best" as in our protocol.
+So just make the changes to the first two rows.
+TG: Erf, deeply sorry, I probably forgot to take the not here while chatting.
+Ok after reflection I either don't remember at all what I said or either don't agree.
+The tables show the "Summary of the comparisons between the "best" tree and the "missing-data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr)."
+%
+% Tree inference method       Metric   Min.  1st Qu.  etc...
+% Maximum Likelihood tree       RF      1.1     2.2     3.3
+%
+% And this reads: the comparison between the "best" <insert what's in the Tree inference method column> tree compared to the "missing-data" trees in the same method for the <insert metric> is minimum <insert value> etc...
 
@@
 
-We understand the concerns of the reviewer but, we would prefer to stick with our "best" tree approach.
+We understand the concerns of the reviewer and added the comparisons between the ``true'' and the ``missing-data'' trees to the Appendix C.
+Nonetheless, we would prefer to stick with our "best" tree approach in the main manuscript.
 In practice, the ``true'' tree can be perceived as a ``random'' seed to generate the ``complete'' matrix, as it simply a birth-death tree with $\lambda$ $>$ $\mu$ and a set number of living and fossil taxa (see details in the manuscript section "2.1 Generating the matrix").
@@
 I've had a go but you may need to check what I've written.
+TG: I've added the bit at the start saying "no but look, we did it in the appendix".
+I didn't really get the "and you're stupid" (I'm not the best person to judge these things ;)).
+Maybe add something along the line "And we're sorry for the confusion, we now the manuscript is long and complicated but we made a effort of describing each step in great details, even the random seed generation..."
 
@@
 Instead this 50\% refers to missing fossil data across the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a maximum of 95\% missing data, see below).
-% NC: normally you'd say up to a maximum rather than up to a minimum.
-Is what I've written correct?
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum of 5\% available data; see below)
+%up to a maximum of 95\% missing data, see below).
+% NC: normally you'd say up to a maximum rather than up to a minimum. Is what I've written correct? TG: maybe change it "down to a minimum"?
 \item{The number of morphological characters for both living and fossil taxa ($N_C$).} This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.

### 2015-08-12 Manuscript/MPESubmission/Revision/response.tex
Unified appendix/figure/table

@@
 
-% NC: Probably want to change figure to Figure, table to Table, appendix to Appendix to be consistent.
+% NC: Probably want to change figure to Figure, table to Table, appendix to Appendix to be consistent. TG: I've changed to Fig. and Table as in MPE guidelines.
 
@@
 
-We gladly followed reviewer 1's graphical abstract suggestions and have changed it to be similar to Figure 5.
+We gladly followed reviewer 1's graphical abstract suggestions and have changed it to be similar to Fig. 5.
 
@@
 We have now compared both the ``best'' tree and the ``missing-data'' trees to the ``true'' tree as suggested by both reviewers.
-The results are available in appendix B; figure B1 and appendix C; figures C6 and C7.
+The results are available in Appendix B; Fig. B1 and Appendix C; Figs. C6 and C7.
 The ability to recover the ``true'' tree is poor for both methods and using both metrics.
@@
 Because the main aim of this study is to look at the effect our missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees and ``missing data'' trees.
-The results of the comparisons of the ``true'' and the ``missing data'' tree are available in appendix B.
+The results of the comparisons of the ``true'' and the ``missing data'' tree are available in Appendix B.
 Note that this makes little difference to our overall results.} line 276-280 @@@.
 
-See also appendix B section "Differences between the ``true'' and the ``best'' trees." and reviewer 2's comment 1 below.
+See also Appendix B section "Differences between the ``true'' and the ``best'' trees." and reviewer 2's comment 1 below.
 
@@
 Therefore, we do not believe our results are merely the result of using a specific starting tree topology.
-These results are available in the supplementary materials Appendix A (Figure A2 and Table A1) and we also added a note to the manuscript:
+These results are available in the supplementary materials Appendix A (Fig. A2 and Table A1) and we also added a note to the manuscript:
 
@@
 two distributions with a coefficient of 0.85 are not significantly similar but are closer to each other than two distributions with a coefficient of 0.15; which are themselves not significantly different but still less close to each other than the previous ones).
-We added the following statement to the main text and to the appendix B:
+We added the following statement to the main text and to the Appendix B:
 
@@
 
-\textit{Additionally, the Bayesian posterior trees performed more poorly than the Bayesian consensus tree (figure 4, table 1 and appendix C figure C5 and tables C5, C6 and C7).
+\textit{Additionally, the Bayesian posterior trees performed more poorly than the Bayesian consensus tree (Fig. 4, Table 1 and Appendix C Fig. C5 and Tables C5, C6 and C7).
 This may be because the Bayesian posterior trees are always resolved and thus more likely to contain incorrectly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
@@
 Concerning the last question of this comment: "How then in comparisons of the distributions are they always so low (Table 1 max prob.
-of overlapping distributions = 0.11)?", we apologise for the confusion here and assume it is due to our mislabeling of the figure 4 (see reviewer 1's comment 12).
-In fact, when looking at all parameters separately (figure 4), there is no clear overlap between the Bayesian consensus trees and the Bayesian posterior trees (median Bhattacharyya Coefficient of 0.01; Table 1).
+of overlapping distributions = 0.11)?", we apologise for the confusion here and assume it is due to our mislabeling of the Fig.
+4 (see reviewer 1's comment 12).
+In fact, when looking at all parameters separately (Fig.
+4), there is no clear overlap between the Bayesian consensus trees and the Bayesian posterior trees (median Bhattacharyya Coefficient of 0.01; Table 1).
 The maximum observed Bhattacharyya Coefficient (0.11; Table 1) is probably due to the combination of parameters: when all the parameters have many missing data, the probability of overlap between trees was slightly higher (0.11) but still really low.
-For the Normalised Triplets distance however, the overlap is higher (median Bhattacharyya Coefficient of 0.59; table 1) as illustrated in figure 4 (i.e. much more overlap between the distributions).
+For the Normalised Triplets distance however, the overlap is higher (median Bhattacharyya Coefficient of 0.59; Table 1) as illustrated in Fig. 4 (i.e. much more overlap between the distributions).
 
@@
 
-We modified the figures 4 and 5 to show Normalised Triplets metric values below 0 and added a dashed line at 0 to underline the difference in scale between both metrics.
+We modified the Figs. 4 and 5 to show Normalised Triplets metric values below 0 and added a dashed line at 0 to underline the difference in scale between both metrics.
 As mentioned above (reviewer 1 comment 9), this is due to the fact that the Bayesian consensus tree will be less and less resolved when missing data increases.
@@
 
-\textit{The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (figure 5).
+\textit{The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (Fig.
+5).
 A Normalised Triplets metric value below 0 means that the placement of some taxa is worse than expected by just randomly placing this taxon in the tree.
 This can be interpreted as the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
-Even if clades are conserved (figure 5), the resolution within them can be poor to non-existent when a lot of data are missing (i.e. 75\%).
+Even if clades are conserved (Fig. 5), the resolution within them can be poor to non-existent when a lot of data are missing (i.e. 75\%).
 In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
@@
 We changed the colors to match with the paper's color scheme: \textit{Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)}.
-See figure 4 caption.
+See Fig.
+4 caption.
 
@@
 \item{\textcolor{blue}{\textbf{line 193 to 196:} again, justify why gamma with shape of 0.5 is appropriate for minimizing homoplasy.
-This distribution places highest probability on substation rates less than 0.5, when one of the cited references potentially suggests that the lowest topological error was found at rates of ~0.5-1.5, and that very low substitution rates incurred error similar to very high rates ((Wright and Hillis 2014) 2014 Figure 3).
+This distribution places highest probability on substation rates less than 0.5, when one of the cited references potentially suggests that the lowest topological error was found at rates of ~0.5-1.5, and that very low substitution rates incurred error similar to very high rates ((Wright and Hillis 2014) 2014 Fig.
+3).
 Other cited references do not clearly state that this distribution and shape setting is optimal, so the authors really must justify.
@@
 We initially chose this parameter from personal communication with April Wright and to be consistent with the molecular rate parameter.
-In figure 3 from Wright and Hillis 2014, even if the topological error is high when rates are near 0, as is mainly the case for parsimony and both parsimony and Mk with missing data, the fitted model for the Mk model with no missing data (light purple line) is still lowest when the rates are near 0.
-Topological error, however, becomes minimal for all scenarios (Parsimony and Mk, with or without missing data) at a rate of 0.5 (see modified figure 3 from Wright and Hillis (2014) in this response; we added the tick marks and the 0.5 rate limit by using the ruler in Adobe Illustrator).
+In Fig.
+3 from Wright and Hillis 2014, even if the topological error is high when rates are near 0, as is mainly the case for parsimony and both parsimony and Mk with missing data, the fitted model for the Mk model with no missing data (light purple line) is still lowest when the rates are near 0.
+Topological error, however, becomes minimal for all scenarios (Parsimony and Mk, with or without missing data) at a rate of 0.5 (see modified Fig.
+3 from Wright and Hillis (2014) in this response; we added the tick marks and the 0.5 rate limit by using the ruler in Adobe Illustrator).
 
@@
     \includegraphics[keepaspectratio=true]{response_fig1.pdf}
-\caption{Figure 3 from Wright and Hillis 2014 with added x axis ticks. The red line represents the 0.5 rate.}
+\caption{Fig. 3 from Wright and Hillis 2014 with added x axis ticks. The red line represents the 0.5 rate.}
 \end{figure}
@@
 Or at least include some perturbation to the starting tree to avoid overly influencing the resulting topology.
-It would seem to me that this is illustrated in figure 5, where-in with the Robinson-Foulds metric, the distance to the best tree in Bayesian analysis seems to have a lower bound at ~0.7, even under the worst cases of missing data.
+It would seem to me that this is illustrated in Fig.
+5, where-in with the Robinson-Foulds metric, the distance to the best tree in Bayesian analysis seems to have a lower bound at ~0.7, even under the worst cases of missing data.
 This may be the worst the analysis can do given that it started out with the true tree.
@@
 \textit{Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
-For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference among trees with any combination of the other parameters ($M_F$ and $N_C$; Figure 6).
+For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference among trees with any combination of the other parameters ($M_F$ and $N_C$; Fig. 6).
 In other words, when the number of missing living taxa reaches 50\%, neither the amount of missing data in the fossil record ($M_F$), nor the number of characters in the matrix ($N_C$) affect topology.
-A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Figure 6).
+A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Fig. 6).
 This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).} lines 589-609@@@
@@
 
-As discussed above (reviewer 1, specific comment 29), we understand the reviewer's concern and have added an analysis in Appendix A (figure A2 and tables A1 and A2) where we tested the effect of using the ``true'' tree as a starting tree.
+As discussed above (reviewer 1, specific comment 29), we understand the reviewer's concern and have added an analysis in Appendix A (Fig.
+A2 and Tables A1 and A2) where we tested the effect of using the ``true'' tree as a starting tree.
 We found that there is no significant effect of using the ``true'' tree instead of a random tree (MrBayes default) as a starting tree on topological recovery.
@@
 
-\textit{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Figure 5).
+\textit{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
 Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details on the analysis of the effect of the starting tree in Bayesian inference).} lines 618-624@@@
@@
 Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies (Swofford et al., 2001).
-As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve poorly supported nodes due to missing data than the Maximum Likelihood tree that is more likely to incorrectly resolve such nodes (i.e.
+As shown in Fig.
+5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve poorly supported nodes due to missing data than the Maximum Likelihood tree that is more likely to incorrectly resolve such nodes (i.e.
 creating a false positive node).
@@
 It took me a long time to figure it out, however, and to figure out why the top of the triangle is all orange.
-I would have expected something more like in Appendix C Figure C2 or Fig. C3.
+I would have expected something more like in Appendix C figure C2 or figure C3.
 At first I understood this to indicate high probability of overlap of the missing data trees which had the most missing data with the best tree.
@@
 We modified the methods section as follows:
-\textit{Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Figure 6).
-In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Figure 6).} lines 434-439@@@
+\textit{Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Fig. 6).
+In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Fig.
+6).} lines 434-439@@@
 
@@
 \textit{Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Figure 6).
-The Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; blue regions in Figure 6A), however, displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; blue regions in Figure 6B).
-The orange upper triangle in Figure 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Figure 6A).
-Once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Figure 6A).
-Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Figure 6A).
-In Figure 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.} lines 480-497@@@
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Fig.
+6).
+The Normalised Robinson-Foulds metric (median Bhattacharrya coefficient = 0.79; blue regions in Fig.
+6A), however, displays more dissimilarities than the Normalised Triplets metric (median Bhattacharrya coefficient = 0.81; blue regions in Fig.
+6B).
+The orange upper triangle in Fig. 6A shows a high probability of overlap of the Normalised Robinson-Foulds metric for the trees with the $M_{L}$ parameter $\geq$ 50\% (Fig. 6A).
+Once $M_{L}$ $\geq$ 50\%, there is no additional effect of $M_{F}$ and $N_{C}$, regardless of the amount of missing data in these parameters (Fig. 6A).
+Likewise, once $N_{C}$ $<$ 50, there is no additional effect of $M_{L}$ and $M_{F}$ as denoted by the high probability of Normalised Robinson-Foulds metric overlap (horizontal orange stripes between the blue regions Fig.
+6A).
+In Fig. 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.} lines 480-497@@@
 
@@
 
-2) We followed the reviewer's suggestion and added six extra tables to Appendix C (Table C8 to C13) containing the marginal probability of each missing data parameter individually and for each tree inference method.
+2) We followed the reviewer's suggestion and added six extra tables to Appendix C (Tables C8 to C13) containing the marginal probability of each missing data parameter individually and for each tree inference method.
 
@@
 no missing data for living taxa, and maximal number of characters but various amount of missing data for the fossil taxa).
-These results are present in figure 4, central row.
+These results are present in Fig. 4, central row.
 However, we think we may have not explained ourselves entirely clearly.

### 2015-08-12 Manuscript/MPESubmission/Revision/response.tex
Edits to response - minor

@@
 \documentclass[12pt,letterpaper]{article}
-
-%%%% NC: FOR CLARITY - can you make the bits that are just copied from the text italic? Sometimes it jumps quickly between a comment to a piece of text. Italics will help make it clearer.%%%%
 
@@
 %\linespread{2}
-% NC: Cover letters don't need to be double spaced and by removing this we reduce the page count substantially! i.e. to non "losing the will to live" lengths...
-
-%\linespread{2}
+
 \raggedright
@@
 
-% NC: Probably want to change figure to Figure, table to Table, appendix to Appendix to be consistent. TG: I've changed to Fig. and Table as in MPE guidelines.
-
 \textbf{RE: MPE-15-254}\\
@@
 % 1-And until now, not many studies
-
-% NC: Line numbers? TG: I'll do all line numbering once I've incorporated all the edits to the manuscript.
 \textit{Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
@@
 % 2-These studies have done X and Y to see if missing data's a problem
-These studies have assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data (Wiens et al.
+These studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data (Wiens et al.
 2005); or (2) removing both molecular and some morphological data from living taxa to create artificial fossils (Manos et al., 2007; Pattinson et al., 2014).
@@
 % 4-But they've done it a certain way
-% NC: This is still a problem (it's ok up to 4). WHY is an empirical-only approach limiting? WHY is it limiting to generate missing data using fossil patterns?
-These studies, however, were limited by because of their empirical approach limiting the conclusions of the studies to similar scenarios. %TG: or scenari? Reviewer 1 is picky with these things
-Additionally, Wiens et al.
-(2005) didn't include fossil taxa in their analyse thus not allowing to test the placement of fossil taxa; and Manos et al.
-(2007) and Pattinson et al.
-(2014) used only patters of the fossil record to generate missing data thus not allowing to test the effect of missing data for other patterns observed for living taxa (e.g.
-the complete absence of morphological data; Guillerme & Cooper 2015).
-% TG: Does that sound better?
-Besides, Manos et al. (2007) and Pattinson et al. (2014), mainly focused on the palaeontological aspect of the question.} lines @@@.
-% NC: I don't like all these i.e.s here.
-Why not be specific in the text rather than throwing everything into brackets?
-If you're only citing manos and patterson then just say what they did.
-TG: no more i.es then ;).
-% (next §)5-We're doing it this way
+The way these studies were conducted, however, means that their conclusions are not generally applicable across all scenarios involving missing data in Total Evidence phylogenies.
+For example, using an empirical (rather than simulation based) approach limits their conclusions to studies with similar distributions of data across species in the phylogeny.
+Additionally, Wiens et al. (2005) did not include fossil taxa in their analyses, so their results cannot be used to make conclusions about how missing data may influence the placement of fossils. 
+Manos et al. (2007) and Pattinson et al. (2014) include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices. 
+Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is likely (Guillerme \& Cooper 2015).
+Finally, Manos et al. (2007) and Pattinson et al. (2014), mainly focused on how missing data in fossil taxa affects the placement of fossils, ignoring the effects of missing data in living species.}
 
 \textit{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living taxa with morphological data, fossil data, all data for certain characters and the combination of these three aspects.
-This is an advance on previous studies because we use large scale simulations and and analyse the effect of three distinct aspect of missing data thus focusing on both neontological and palaeontological parts of the matrix.
-Besides, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and (ii) the displacement of wild-card taxa (based on the Triplets distance; Critchlow et al., 1996) rather than just the clade conservation or support (cf.
+This is an advance on previous studies because we use large-scale simulations and analyze the effect of three distinct aspect of missing data thus focusing on both neontological and paleontological parts of the matrix.
+In addition, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and (ii) the displacement of wild-card taxa (based on the Triplets distance; Critchlow et al., 1996) rather than just a single measure of clade conservation or clade support (cf.
 Wiens et al., 2005; Pattinson et al., 2014).} lines @@@
-% NC: Again how is this an advance. I guess it's related to this sentence from below:
-% (cf. Wiens et al., 2005; Pattinson et al., 2014 measuring respectively only the clades support or the number of shared clades).
-
-% TG: I actually rewrote this bit as well:
-% NC: line numbers?
-% \textit{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
-% We inferred the topology from the matrices using both Maximum Likelihood and Bayesian Inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and for wild-card taxa placement (based on the Triplets distance; Critchlow et al., 1996).
-% Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring the effects of missing data on two aspects of topology (cf.
-Wiens et al., 2005; Pattinson et al., 2014 measuring respectively only the clades support or the number of shared clades).} lines @@@
-% NC: This paragraph seems to be repeating what you've already said above but in slightly more detail. Might make more sense to combine this paragraph and the one above it? TG: Ok, I've combined them
 
@@
 TG: is that ok?
+
+% NC: ok true trees versus best trees, but for all missing data trees?
+
 This may be related to the size of our matrices or to the ability of the algorithms to generate the ``complete matrix'' but also to our decision to put our minimum median bootstrap acceptance at 50 rather than higher values (line @@@ in the former manuscript and @@@ in our revision).
@@
 However, we can observe the same patterns between the methods as when using the ``best'' tree (i.e. Bayesian always outperforms ML in clade conservation but does poorly in wild-card taxa placement).
-An peculiar feature is that the Normalised Robinson-Foulds metric actually \textit{increases} when we compare the ``missing data'' trees to the ``true'' tree.
-% NC: specify what is being compared, e.g.
-"when we compare the missing data trees to the ``true'' trees.""
+An peculiar feature is that the Normalised Robinson-Foulds metric actually \textit{increases} when we compare the ``missing data'' trees to the ``true'' tree. % 
 This is because the difference between the ``best'' and the ``true'' tree is high, making trees with more data more different than those with lots of missing data.
@@
 
-% TG: I rephrased as follows.
 \textit{Previous studies have explored the effect of missing morphological data in Total Evidence matrices (Wiens et al., 2005; Manos et al.,2007, Pattinson et al., 2014).
-The conclusions of theses studies, however, were limited by their empirical approach making their results applicable only to similar scenarios. %TG: or scenari
-Additionally, these studies focused either only on living taxa (Wiens et al., 2005) or on the patterns of missing data from the fossil record only (Manos et al.,2007, Pattinson et al., 2014).
-The conclusions of these studies were therefore restricted to these specific aspect of missing data and, besides their interesting conclusions, % TG: seems unfair to me to just bash them, they are actually nice papers
-they didn't provide any insight of the combined effect of different missing data parameters.
-% as well as several aspects of their experimental design (e.g. not including fossil taxa, Wiens et al., 2005; or focusing on the palaeontological aspect, Manos et al.,2007, Pattinson et al., 2014).
-% NC: Again needs a bit more explanation. WHY is empirical approach a problem? Because it only gives answers to one specific scenario or clade?
-Here we used a more in depth approach where missing data were generated from simulated data and according to three clearly defined missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
+The conclusions of theses studies, however, were limited by their empirical approach making their results applicable only to similar missing data scenarios. %TG: or scenari
+% NC: I think scenarios is the correct form
+Additionally, these studies focused either only on living taxa (Wiens et al., 2005) or on the patterns of missing data from the fossil record only (Manos et al., 2007, Pattinson et al., 2014).
+Here we instead used an approach where missing data were generated from simulated data and according to three clearly defined missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$) that removed data from both the living and fossil taxa.
 This allowed us to confirm previous results that missing data can be especially problematic in small matrices (Wiens 2003), but also revealed the crucial importance of coding morphological data for living species in Total Evidence phylogenies.
@@
 Therefore we advise increased focus on coding morphological characters for a large number of the living taxa present in the matrix (i.e.
-50\%) if the goal is to accurately combine both living and fossil species in phylogenies.
+at least 50\%) if the goal is to accurately combine both living and fossil species in phylogenies.
 Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data.} lines @@@
@@
 Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.} lines @@@
-% TG: maybe add in a non tip-dating analysis way (i.e. Nick Matzke comment)? 
-% NC: Sure if it fits. If you want to "cite" Nick you can use (N.Matzke pers.comm.)
-% TG: ok, actually leave it, I've tried something but the cite wasn't that convincing and we're already being carefull enough here.
-
-% \textit{To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological morphological data in the same phylogeny as long as there is sufficient overlap.
-% Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.} lines 602-634. @@@
-% NC: I don't like this last bit.
-It's really weak.
-I'd cut it to be honest.
-The first sentence is not very clear (suddenly you revert back to neo and paleo) and the second sentence has no real explanation or link to the rest of the paper.
-TG: let's just remove it then...
-You're right it's totally useless and after reading the whole paper, the reader won't mind saving a few more second of their life.
 
@@
 Values $>$ 0.95 denote significantly similar distributions and values $<$ 0.05 denote significantly different distributions."}
-% NC: Do you want to say values > 0.95 are significantly similar, values < 0.05 are significantly different?  TG: yes.
 
@@
 \textit{The detailed MrBayes parameters are available in Appendix A.
-We also included analysis showing details on the $\alpha$ parameter estimation though it is beyond the scope of this paper and are not discussed further here.} lines 336-338 @@@
+We also included an analysis showing the effect of missing data on the morphological substitution rate parameter ($\alpha$) estimates, although it is beyond the scope of this paper so the results are not discussed further here.} lines 336-338 @@@
 See also Appendix A "Morphological rates estimations results".
-
-% NC: Needs a bit more explanation.
-Something like, we also include in the appendix an analysis showing ...
-though it is beyond the scope of this paper to discuss these results further here.
-TG: is that ok?
 
@@
 
-We removed the repetition in the caption and changed the first two rows as suggested.
-%If we understood the second point correctly, the reviewer suggests we add ``\textit{best}'' to the two first rows of the tables, and by ``\textit{best}'', the reviewer means ``optimal'' (as opposed to the ``best'' tree in our protocol that refers to any tree inferred by any method from the ``complete matrix''.
-% We don't think it is necessary to follow this suggestion as the Maximum Likelihood tree is by definition the tree with the highest likelihood, and the Bayesian consensus is the consensus tree of the Bayesian inference trees.
-%So adding the word ``\textit{best}'' here is not entirely accurate.
-
-% NC: No we discussed this last night - I think they *do* mean "best" as in our protocol.
-So just make the changes to the first two rows.
-TG: Erf, deeply sorry, I probably forgot to take the not here while chatting.
-Ok after reflection I either don't remember at all what I said or either don't agree.
-The tables show the "Summary of the comparisons between the "best" tree and the "missing-data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr)."
-%
-% Tree inference method       Metric   Min.  1st Qu.  etc...
-% Maximum Likelihood tree       RF      1.1     2.2     3.3
-%
-% And this reads: the comparison between the "best" <insert what's in the Tree inference method column> tree compared to the "missing-data" trees in the same method for the <insert metric> is minimum <insert value> etc...
+We removed the repetition in the caption. However, we are not sure why the reviewer thinks we should change the labels of the first two rows so we have left them as is for now. 
+% NC: OK I think you're looking at the wrong tables but never mind let's just leave it and say we don't understand.
 
@@
 
-% NC: Can you rephrase a little here so it sounds less like you saying "no, and you're stupid".
-Doesn't the response to reviewer 1 cover this already?
-I've had a go but you may need to check what I've written.
-TG: I've added the bit at the start saying "no but look, we did it in the appendix".
-I didn't really get the "and you're stupid" (I'm not the best person to judge these things ;)).
-Maybe add something along the line "And we're sorry for the confusion, we now the manuscript is long and complicated but we made a effort of describing each step in great details, even the random seed generation..."
+% NC: Let's just leave this as is for now then. As long as what is written is correct (this was the last thing I did rushing out the door last night).
 
@@
 Instead this 50\% refers to missing fossil data across the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum of 5\% available data; see below)
-%up to a maximum of 95\% missing data, see below).
-% NC: normally you'd say up to a maximum rather than up to a minimum. Is what I've written correct? TG: maybe change it "down to a minimum"?
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data; see below)
+% NC: Down to a minimum also sound weird. I've added threshold which makes it sound better.
 \item{The number of morphological characters for both living and fossil taxa ($N_C$).} This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
@@
 like the "artificial fossil" approach in Pattinson et al 2014, Syst.
-Biol.) but we rather treated the missing data in the "whole" fossil record.
+Biol.) but we rather treated the missing data in the whole fossil record.
 Because our fossil taxa and their characters where randomly simulated, we didn't have any \textit{a priori} assumption on how to remove the data.
 We therefore decided to randomly remove a certain amount of data among all the fossils.
-For example, when removing 75\% of the data for the fossil taxa ($M_F$=75\%) different scenarios could happen, ranging from a "homogeneous" incompleteness data distribution (each specific fossil taxa had 75\% of missing data) to a "heterogeneous" incompleteness data distribution  (80\% of the specific fossil taxa had 5\% available data and 21\% of the specific fossil taxa had 100\% available data).
+For example, when removing 75\% of the data for the fossil taxa ($M_F$=75\%) different scenarios could happen, ranging from a homogeneous incompleteness data distribution (each specific fossil taxa had 75\% of missing data) to a heterogeneous incompleteness data distribution  (80\% of the specific fossil taxa had 5\% available data and 21\% of the specific fossil taxa had 100\% available data).
 In empirical matrices, the incompleteness data distribution probably lies in between both models.
@@
 Instead this 50\% refers to missing fossil data across the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a maximum of 95\% missing data, see below).}
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data, see below).}
 

### 2015-08-12 Manuscript/MPESubmission/Revision/response.tex
Final edits to response

@@
 
-% 0(before this section)-Missing data can be a problem
-% 1-And until now, not many studies
 \textit{Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework (i.e.
@@
 Wiens et al., 2005; Manos et al., 2007; Pattinson et al., 2014).
-% 2-These studies have done X and Y to see if missing data's a problem
 These studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data (Wiens et al.
 2005); or (2) removing both molecular and some morphological data from living taxa to create artificial fossils (Manos et al., 2007; Pattinson et al., 2014).
-% 3-They show that it's not a problem
 Both approaches have shown that missing data are not a major problem and should not be an obstacle to combining both living and fossil species in the same phylogenies.
-% 4-But they've done it a certain way
 The way these studies were conducted, however, means that their conclusions are not generally applicable across all scenarios involving missing data in Total Evidence phylogenies.
@@
 When comparing the ``true'' tree to the ``best'' trees, in Maximum Likelihood the 50\% CI lays between approximately 0.2 and 0.3 for the Normalised Robinson-Foulds and Triplets metric; and, in Bayesian consensus tree  the 50\% CI lays between approximately 0.3 and 0.4 for the Normalised Robinson-Foulds metric and between 0 and 0.2 for the Normalised Triplets metric).
-% NC: What are these results comparing?
-All missing data trees  to the true tree?
-True trees to best trees?
-Clarify.
-TG: is that ok?
-
-% NC: ok true trees versus best trees, but for all missing data trees?
-
 This may be related to the size of our matrices or to the ability of the algorithms to generate the ``complete matrix'' but also to our decision to put our minimum median bootstrap acceptance at 50 rather than higher values (line @@@ in the former manuscript and @@@ in our revision).
+
 When comparing the ``true'' tree to the ``missing data'' trees, the results are \textit{scaled down} due to the low scores of the ``true'' \textit{vs.} ``best'' trees scores.
 However, we can observe the same patterns between the methods as when using the ``best'' tree (i.e. Bayesian always outperforms ML in clade conservation but does poorly in wild-card taxa placement).
-An peculiar feature is that the Normalised Robinson-Foulds metric actually \textit{increases} when we compare the ``missing data'' trees to the ``true'' tree. % 
+An peculiar feature is that the Normalised Robinson-Foulds metric actually \textit{increases} when we compare the ``missing data'' trees to the ``true'' tree.
 This is because the difference between the ``best'' and the ``true'' tree is high, making trees with more data more different than those with lots of missing data.
@@
 
-Although though we agree with the comments of both reviewers that knowing the difference between the true and best/missing-data trees is important, we do not think that these results warrant inclusion in the main text.
+Although though we agree with the comments of both reviewers that knowing the difference between the true and ``best''/``missing-data'' trees is important, we do not think that these results warrant inclusion in the main text.
 In these simulations, we had to make a choice between (1) simulating larger matrices but with fewer chains (i.e. less variance) or (2) smaller matrices with more variance.
@@
 \end{enumerate}
-Therefore we added a second justification to the main text (see below this response) and added these analysis (true vs.
-best/missing-data trees) to the supplementary for readers that might be interested by this specific problem.
+Therefore we added a second justification to the main text (see below this response) and added these analysis (``true'' \textit{vs.} ``best''/``missing-data'' trees) to the supplementary for readers that might be interested by this specific problem.
 
@@
 \textit{Previous studies have explored the effect of missing morphological data in Total Evidence matrices (Wiens et al., 2005; Manos et al.,2007, Pattinson et al., 2014).
-The conclusions of theses studies, however, were limited by their empirical approach making their results applicable only to similar missing data scenarios. %TG: or scenari
-% NC: I think scenarios is the correct form
+The conclusions of theses studies, however, were limited by their empirical approach making their results applicable only to similar missing data scenarios.
 Additionally, these studies focused either only on living taxa (Wiens et al., 2005) or on the patterns of missing data from the fossil record only (Manos et al., 2007, Pattinson et al., 2014).
@@
 \textit{The detailed MrBayes parameters are available in Appendix A.
-We also included an analysis showing the effect of missing data on the morphological substitution rate parameter ($\alpha$) estimates, although it is beyond the scope of this paper so the results are not discussed further here.} lines 336-338 @@@
+We also included an analysis showing the effect of missing data on the estimation of the shape parameter ($\alpha$) of the morphological substitution rate distribution.
+This extra analysis, however, is beyond the scope of this paper so the results are not discussed further here.} lines 336-338 @@@
 See also Appendix A "Morphological rates estimations results".
@@
 We removed the repetition in the caption. However, we are not sure why the reviewer thinks we should change the labels of the first two rows so we have left them as is for now.
-% NC: OK I think you're looking at the wrong tables but never mind let's just leave it and say we don't understand.
 
@@
 We understand the concerns of the reviewer and added the comparisons between the ``true'' and the ``missing-data'' trees to the Appendix C.
-Nonetheless, we would prefer to stick with our "best" tree approach in the main manuscript.
+Nonetheless, we would prefer to stick with our ``best'' tree approach in the main manuscript.
 In practice, the ``true'' tree can be perceived as a ``random'' seed to generate the ``complete'' matrix, as it simply a birth-death tree with $\lambda$ $>$ $\mu$ and a set number of living and fossil taxa (see details in the manuscript section "2.1 Generating the matrix").
 All these parameters were only used to create a realistic ``complete'' matrix in order to make our results relevant to biologists.
-
-% NC: Let's just leave this as is for now then. As long as what is written is correct (this was the last thing I did rushing out the door last night).
 
@@
 In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
-%Added this part following comment 4.
 Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
@@
 Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data; see below)
-% NC: Down to a minimum also sound weird. I've added threshold which makes it sound better.
 \item{The number of morphological characters for both living and fossil taxa ($N_C$).} This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.

### 2015-08-12 Manuscript/MPESubmission/Revision/response.tex
Added revisions to the manuscript

@@
 Manos et al. (2007) and Pattinson et al. (2014) include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices. 
-Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is likely (Guillerme \& Cooper 2015).
+Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is likely (Guillerme and Cooper 2015).
 Finally, Manos et al. (2007) and Pattinson et al. (2014), mainly focused on how missing data in fossil taxa affects the placement of fossils, ignoring the effects of missing data in living species.}
@@
 In addition, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and (ii) the displacement of wild-card taxa (based on the Triplets distance; Critchlow et al., 1996) rather than just a single measure of clade conservation or clade support (cf.
-Wiens et al., 2005; Pattinson et al., 2014).} lines @@@
+Wiens et al., 2005; Pattinson et al., 2014).} lines 85-118@@@
 
@@
 The results of the comparisons of the ``true'' and the ``missing data'' tree are available in Appendix B.
-Note that this makes little difference to our overall results.} line 276-280 @@@.
+Note that this makes little difference to our overall results.} line 297-308 @@@.
 
@@
 \textit{We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
-Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').} line 323-327 @@@.
+Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').} line 352-357 @@@.
 See also Appendix A section "Effect of the starting tree on Bayesian inference".
@@
 
-\textit{The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 (full overlap; Bhattacharyya, 1943, see Appendix B for calculation details).} lines @@@
+\textit{The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 (full overlap; Bhattacharyya, 1943, see Appendix B for calculation details).} lines 445-447@@@
 
@@
 When this coefficient is $>$0.95, the distributions are significantly similar.
-Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.} line 419-424 @@@.
+Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.} line 449-454 @@@.
 
@@
 Conversely, the Bayesian consensus tree might not resolve nodes that are poorly supported and thus is more likely to contain only correctly resolved nodes (i.e.
-increasing the Normalised Robinson-Foulds metric).} lines 652-658.
+increasing the Normalised Robinson-Foulds metric).} lines 681-687.
 @@@
@@
 These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters (Wiens 2003).
-It is important to note, however, that this effect can be reduced by increasing the number of characters (Wiens 2003).} lines 632-644. @@@
+It is important to note, however, that this effect can be reduced by increasing the number of characters (Wiens 2003).} lines 661-673. @@@
 
@@
 at least 50\%) if the goal is to accurately combine both living and fossil species in phylogenies.
-Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data.} lines @@@
+Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data.}
 
@@
 Ronquist et al., 2012a).
-Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.} lines @@@
+Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.} lines 731-62@@@
 
@@
 If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology.
-Values $>$ 0.95 denote significantly similar distributions and values $<$ 0.05 denote significantly different distributions."}
+Values $>$ 0.95 denote significantly similar distributions and values $<$ 0.05 denote significantly different distributions.}
 
@@
 
-\textit{These results are in agreement with Manos et al. (2007) where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.} lines @@@
+\textit{These results are in agreement with Manos et al. (2007) where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.} lines 566-567@@@
 
@@
 
-We modified the citation as suggested by reviewer 1 (line 47-48@@@).
+We modified the citation as suggested by reviewer 1 (line 48@@@).
 
@@
 
-We modified the statement to: \textit{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints} (line 111-113@@@).
+We modified the statement to: \textit{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints} (line 126-128@@@).
 
@@
 We also included an analysis showing the effect of missing data on the estimation of the shape parameter ($\alpha$) of the morphological substitution rate distribution.
-This extra analysis, however, is beyond the scope of this paper so the results are not discussed further here.} lines 336-338 @@@
+This extra analysis, however, is beyond the scope of this paper so the results are not discussed further here.} lines 366-370 @@@
 See also Appendix A "Morphological rates estimations results".
@@
 \textit{In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal (Hassanin et al., 1998; Estoup et al., 2002).%\citep{Hassanin1998611,EstoupHomoplasy}.
-Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below).} (line 189-192@@@).
+Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below).} (line 204-206@@@).
 
@@
 \textit{This model and these parameter settings strike a balance between realism for empirical datasets (e.g.
-Douady, Kelly et al., 2014) and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)}.
-lines @@@
+Douady et al., 2003; Kelly et al., 2014) and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)}.
+lines 206-213@@@
 
@@
 
-We now state that we used the rTraitDisc function.
+We now state that we used the rTraitDisc function on line 214 @@@.
 
@@
 Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model (Lewis, 2001; Wright and Hillis, 2014).
-Note, however, that Wright and Hillis (2014) have shown that low morphological rates (< 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004) so this should not influence our results.} lines 211-217 @@@
+Note, however, that Wright and Hillis (2014) have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004) so this should not influence our results.} lines 225-233 @@@
 
@@
 
-We modified the sentence as follows: \textit{To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.} (line 264-267@@@).
+We modified the sentence as follows: \textit{To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.} (line 285-287@@@).
 
@@
 We now refer to two more classical citations: Bremer and Struwe 1992 (American Journal of Botany) and Patterson et al.
-1993 (Annual Review of Ecology and Systematics), on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to having a small matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters) lines 578-579@@@.
+1993 (Annual Review of Ecology and Systematics), on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to having a small matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters) lines 609-610@@@.
 
@@
 
-We removed the entire statement concerning homoplasy and matrix size ("Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters (Wright and Hillis 2014).")
+We removed the entire statement concerning homoplasy and matrix size (\textit{Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters (Wright and Hillis 2014).})
 
@@
 A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Fig. 6).
-This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).} lines 589-609@@@
+This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).} lines 630-640@@@
 
@@
 \textit{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details on the analysis of the effect of the starting tree in Bayesian inference).} lines 618-624@@@
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details on the analysis of the effect of the starting tree in Bayesian inference).} lines 649-655@@@
 
@@
 It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
-Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies (Swofford et al., 2001).
+Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies (sensu Swofford et al., 2001).
 As shown in Fig.
@@
 2012 trees used in Healy et al.
-2014).} lines 680-700@@@
+2014).} lines 711-731@@@
 
@@
 In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Fig.
-6).} lines 434-439@@@
+6).} lines 467-471@@@
 
@@
 6A).
-In Fig. 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.} lines 480-497@@@
+In Fig. 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.} lines 512-526@@@
 
@@
 Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data; see below)
-\item{The number of morphological characters for both living and fossil taxa ($N_C$).} This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
+\item{The number of morphological characters for both living and fossil taxa ($N_C$).}
+This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
 Any morphological matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
@@
 
-\textit{We created matrices using all parameter combinations resulting in 125 ($5^3$) “missing-data” matrices.
-Note that one of these combinations ($M_L$=0\%; $M_F$=0\% and $N_C$=100) has no missing data so is equivalent to the “complete” matrix, thus we have one effectively complete matrix in our 125 “missing-data” matrices.
+\textit{We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data'' matrices.
+Note that one of these combinations ($M_L$=0\%; $M_F$=0\% and $N_C$=100) has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
 In practice, we first removed the data following the two missing data parameters $M_L$ and $M_F$ and then removed data following the $N_C$ parameters.
 To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
-Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).} lines 222-267@@@
+Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).} lines 238-289@@@
 
@@
 Instead this 50\% refers to missing fossil data across the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data, see below).}
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data, see below).} lines 259-263 @@@
 

### 2015-08-12 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Unified appendix/figure/table
Added revisions to the manuscript

--- 2015-08-11 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
@@
 \hl{Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework }\citep[\hl{i.e.
-using both molecular and morphological data; e.g.}][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{These studies have assessed the effect of missing data on topology either:(1) by comparing a dataset with missing data to sub-sets without missing data }\citep{Wiens01102005}\hl{; or (2) by removing both molecular and some morphological data from living taxa (``artificial fossil'') }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{Both types of approaches have shown that missing data \hl{are} not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.
-However, these studies were limited by their empirical-only approach and the use of only living taxa }\citep{Wiens01102005}\hl{ or the scheme of generating missing data }\citep[\hl{i.e.
-using the patterns from the fossil record;}][]{manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{Besides, the studies using fossil data, mainly on the palaeontological aspect of the question }\citep[\hl{i.e.
-the effect of missing data in fossils in Total Evidence matrices;}][]{manosphylogeny2007,pattinsonphylogeny2014}.
-%Former revised version
-% \hl{Until now, only few attempts have been made to study study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework \protect\citep[\hl{i.e.
-using combined molecular and morphological data}; e.g.][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.}
-% \hl{For example, \protect\cite{Wiens01102005} }\hl{only assess the effect of missing data by comparing the node support between the complete dataset that contains missing data and sub-datasets with no missing data.}
-% \hl{Furthermore, \protect\cite{pattinsonphylogeny2014} }\hl{assess the effect of removing both molecular and morphological data (``artificial extinction'') on the topology.}
-% \hl{As for morphological and molecular data separately, these studies shows that missing data are not a major problem and should not be an obstacle to combine both living and fossil species in the same phylogenies.}
-% \hl{These studies, however, are limited by their empirical-only approach \protect\citep{Wiens01102005} }\hl{and the scheme of generating missing data \protect\citep{manosphylogeny2007,pattinsonphylogeny2014}.}
-% \hl{Also both studies focus on the palaeontological aspect of the question (i.e. the fossil data).}
+using both molecular and morphological data;}][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{These studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data }\citep{Wiens01102005}\hl{; or (2) removing both molecular and some morphological data from living taxa to create artificial fossils }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{Both approaches have shown that missing data are not a major problem and should not be an obstacle to combining both living and fossil species in the same phylogenies.
+The way these studies were conducted, however, means that their conclusions are not generally applicable across all scenarios involving missing data in Total Evidence phylogenies.
+For example, using an empirical (rather than simulation based) approach limits their conclusions to studies with similar distributions of data across species in the phylogeny.
+Additionally, }\cite{Wiens01102005}\hl{ did not include fossil taxa in their analyses, so their results cannot be used to make conclusions about how missing data may influence the placement of fossils.}
+\cite{manosphylogeny2007,pattinsonphylogeny2014}\hl{ include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices. 
+Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is likely }\citep{GuillermeCooperMissing}.
+\hl{Finally, }\cite{manosphylogeny2007,pattinsonphylogeny2014}\hl{ mainly focused on how missing data in fossil taxa affects the placement of fossils, ignoring the effects of missing data in living species.}
 
-\hl{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living or fossil data, or all data for certain characters.
-We test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian Inference methods: (i) the conservation of clades and (ii) the displacement of wild-card taxa.}
+\hl{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living taxa with morphological data, fossil data, all data for certain characters and the combination of these three aspects.
+This is an advance on previous studies because we use large-scale simulations and analyse the effect of three distinct aspect of missing data thus focusing on both neontological and paleontological parts of the matrix.
+In addition, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades }\citep[\hl{based on the Robinson-Foulds distance;}][]{RF1981}\hl{ and (ii) the displacement of wild-card taxa }\citep[\hl{based on the Triplets distance;}][]{critchlowthe1996}\hl{ rather than just a single measure of clade conservation or clade support }\citep[\hl{cf.}][]{Wiens01102005,pattinsonphylogeny2014}.
+
 We focus on the effect of missing data on our ability to recover a ``best'' tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
 Although branch length estimation is also important \citep[namely for timing extinction and/or speciation events; e.g.][]{ronquista2012}, we do not consider branch lengths in this study.
-This is partially due to difficulties with simulating branch lengths and topology simultaneously, \hl{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints} \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+This is partially due to difficulties with simulating branch lengths and topology simultaneously, \hl{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints }\citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 Thus understanding the sensitivity of topology to missing \hl{data} are important for assessing the accuracy of tree estimation in the Total Evidence framework.
@@
 \hl{The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5} \citep{yangamong-site1996}.
-\hl{In practice, such a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal} \citep{Hassanin1998611,EstoupHomoplasy}.
-\hl{Also, we chose this low value to be consistent with our protocol for simulating morphological characters (see below).}
-%We chose a low value of $\alpha$ to reduce the number of sites with high substitution rates, thus avoiding too much homoplasy and a decrease in phylogenetic signal. %Justify
-%We selected the parameters above to generate data with no special assumption about how the characters evolved, and to reduce the computational time required if these parameters were estimated rather than defined in the tree building part of the analysis (even with the parameters defined, total computational time for the whole analysis was around 150 CPU years).
-\hl{This model and these parameter settings strike a balance between realism for empirical datasets }\citep[e.g.][]{douadycomparison2003,kellymolecular2014}\hl{ and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).}
+\hl{In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal }\citep{Hassanin1998611,EstoupHomoplasy}.
+\hl{Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below).}
+\hl{This model and these parameter settings strike a balance between realism for empirical datasets }\citep[\hl{e.g.}][]{douadycomparison2003,kellymolecular2014}\hl{ and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).}
 All the molecular information for fossil taxa was replaced by missing data ("?").
@@
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-\hl{We used low evolutionary rate parameters to be consistent with the molecular rates parameters and to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal} \citep{wrightbayesian2014}.
-\hl{In fact, topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model }\citep{lewisa2001,wrightbayesian2014}.
-\hl{Note, however, that } \cite{wrightbayesian2014}\hl{ have shown that low morphological rates (< 0.5) increase variance in topological error but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal }\citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}.
+\hl{We used low evolutionary rate parameters to be consistent with the molecular rate parameters, to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal }\citep{wrightbayesian2014}.
+\hl{Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model }\citep{lewisa2001,wrightbayesian2014}.
+hl{Note, however, that } \cite{wrightbayesian2014}\hl{ have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal }(\citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}\hl{ so this should not influence our results.}
 
@@
 \subsection{Removing data}
-% \label{Removing_data}
-% We modified the ``complete'' matrix to obtain matrices with missing data by randomly replacing data with ``?" in the morphological part of the matrices according to the following parameters:
-% \begin{enumerate}
-% \item{$M_{L}$, the proportion of living taxa with no morphological data: 0\%, 10\%, 25\%, 50\% or 75\%.}
-% This parameter illustrates the number of living taxa that are present in the molecular part of the matrix but not in the morphological part.
-This reflects the fact that the increased availability of molecular data means that morphological data for living species is rarely collected, and few people have the skills to identify characters needed for detailed phylogenetic analysis.
-% \item{$M_{F}$, the proportion of missing data in the fossil taxa: 0\%, 10\%, 25\%, 50\% or 75\%.}
-% This parameter illustrates the quality of the fossil record, a common problem due to preservation biases \citep{sansomfossilization2013}. 
-% \item{$N_{C}$, the number of morphological characters for both living and fossil taxa: 100, 90, 75, 50 or 25 characters. }
-% This parameter illustrates the number of available morphological characters for both living and fossil taxa.
-This reflects the amount of effort put into identifying the characters \citep[e.g][]{O'Leary08022013}.
-% \end{enumerate}
-
-% In practice, each parameter represents a different way of removing data from the matrix: $M_L$ removes rows from the living taxa's data; $M_F$ removes cells from the fossil taxa's data; and $N_C$ removes columns across both living and fossil taxa's data.
-% Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
-% We first applied the parameters $M_L$ and $M_F$ to the matrix and then applied the $N_C$ parameter.
-% Therefore, when 10\% data was missing for both $M_L$ and $M_F$, 10\% data was missing in the morphological part of the matrix.
-% When applying the $N_C$ parameter with the second parameter state (90 characters), however, the resulting matrix potentially had more than 10\% data missing.
-
-% We created matrices using all parameter combinations resulting in 125 ($5^3$) matrices.
-% Note that one of these combinations has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
-% To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
-
-% TG: this part is entirely re-written.
 \hl{To explore the effect of missing morphological data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
 Hereafter, we call these matrices with missing morphological data the ``missing-data'' matrices.
-Note that the amount of molecular data remained constant throughout our simulations: 1000 molecular characters for each living taxa and no molecular data for fossil taxa (see above).
+Note that the amount of molecular data remained constant throughout our simulations: 1000 molecular characters for living taxa and no molecular data for fossil taxa (see above).
 We removed morphological data using three data incompleteness parameters:}
@@
 \item{\hl{The proportion of missing data in the fossil record ($M_F$).}}
-\hl{This missing data parameter represents the completeness of the fossil record.
+\hl{TThis missing data parameter represents the completeness of the fossil record.
 Due to preservation biases, missing data for fossil taxa are common }\citep{sansomfossilization2013}.
@@
 Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
-Instead this 50\% refers to the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (up to a minimum of 5\%, see below).}
-\item{\hl{TThe number of morphological characters for both living and fossil taxa ($N_C$).}}
+Instead this 50\% refers to missing fossil data across the whole matrix.
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data; see below).}
+\item{\hl{The number of morphological characters for both living and fossil taxa ($N_C$).}}
 \hl{This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
@@
 Therefore, this parameter corresponds to the overall number of characters available for both living and fossil taxa.
-In practice, we randomly removed whole characters from the morphological matrix reducing it to: 100 (i.e. 0 characters removed), 90,75, 50 or 25 characters.
+In practice, we randomly removed whole characters from the morphological matrix reducing it to: 100, 90,75, 50 or 25 characters.
 Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).}
@@
 
-\hl{In practice, each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
+\hl{Each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
 Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
@@
 In real world situations, the ``true'' tree is never available to us because we cannot know the true evolutionary history of a clade \citep[except in very rare circumstances, e.g.][]{rozen2005}.
-\hl{We compare ``best'' trees to ``missing-data'' trees but could also compare ``true'' trees to the ``missing-data'' trees.
-In practice, the difference between the ``best'' trees and the ``missing-data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing-data'' trees.
-The difference between the ``true'' and the ``missing-data'' tree, however, represents the effect of our parameters used to generate the ``true'' tree and the algorithms used to generated the ``complete'' matrix as well as the effect of our missing data parameters and the phylogenetic methods used.
-Because the main aim of this study, is to look at the effect the missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees to ``missing-data'' trees.
-The results of the comparisons of the ``true'' and the ``missing-data'' tree are available in appendix B and C.
-Note that these results makes little difference to our overall results.}
-% Former revised version
-% \hl{In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices (i.e.
-diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).}
-% \hl{The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery.}
+\hl{We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
+In practice, the difference between the ``best'' trees and the ``missing data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing data'' trees.
+The difference between the ``true'' and the ``missing data'' trees, however, represents the effect of our parameters used to generate the ``true'' tree and the algorithms used to generate the ``complete'' matrix as well as the effect of our missing data parameters and the phylogenetic methods used.
+Because the main aim of this study is to look at the effect our missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees and ``missing data'' trees.
+The results of the comparisons of the ``true'' and the ``missing data'' tree are available in Appendix B.
+Note that this makes little difference to our overall results.}
 Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the ``best'' tree, rather than the ``true'' tree, as the ``best'' tree is generally what we have to work with.
@@
 \hl{We used a fixed starting tree rather than a random starting tree }\citep[\hl{default MrBayes;}][]{Ronquist2012mrbayes} \hl{to speed up our Bayesian inferences.
-Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').}
+Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').}
 We also used two priors on the molecular part of the matrix: an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5, and a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)); and one prior on the morphological part of the matrix (exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5).
@@
 Even using these priors, it took $~$ 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-\hl{The detailed MrBayes parameters are available in Appendix A along with details on the $\alpha$ parameter estimation.}
+\hl{The detailed MrBayes parameters are available in Appendix A.
+We also included an analysis showing the effect of missing data on the estimation of the shape parameter ($\alpha$) of the morphological substitution rate distribution.
+This extra analysis, however, is beyond the scope of this paper so the results are not discussed further here.}
 
@@
 When this coefficient is $>$0.95, the distributions are significantly similar.
-Values in between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.}
+Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.}
 \hl{To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
@@
 \hl{Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Fig. 6).
-In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either good or bad effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Fig.
+In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Fig.
 6).}
@@
 Normalised Robinson-Foulds metric) is still relatively good (mode = 0.72) when the proportion of missing data are high ($M_{F}$ = 75\%).
-\hl{These results are in agreement with }\cite{manosphylogeny2007}\hl{ where as few as 16 characters where sufficient for correctly assigning ``artificial fossils'' to their ``correct'' clade.}
+\hl{These results are in agreement with }\cite{manosphylogeny2007}\hl{ where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.}
 
@@
 \hl{Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
-For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference between trees with any combination of the other parameters ($M_F$ and $N_C$; Fig. 6).
-In other words, when the number of missing living taxa reaches 50\%, the amount of missing data in the fossil record ($M_F$) or the number of characters ($N_C$) does not matter and an increase/decrease in both does not affect topology.
+For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference among trees with any combination of the other parameters ($M_F$ and $N_C$; Fig. 6).
+In other words, when the number of missing living taxa reaches 50\%, neither the amount of missing data in the fossil record ($M_F$), nor the number of characters in the matrix ($N_C$) affect topology.
 A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Fig. 6).
-This has important practical implications, especially for the best strategy to improve topology by collecting morphological characters (see below).}
+This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).}
 
@@
 \hl{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference (see Methods) had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'') for details on the analysis of the effect of the starting tree in Bayesian inference).}
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details on the analysis of the effect of the starting tree in Bayesian inference).}
 As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
@@
 A Normalised Triplets metric value below 0 means that the placement of some taxa is worse than expected by just randomly placing this taxon in the tree.
-This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
+This can be interpreted as the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
 Even if clades are conserved (Fig. 5), the resolution within them can be poor to non-existent when a lot of data are missing (i.e. 75\%).
@@
 This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best'' trees involved 1000 random pairwise comparisons rather than just one.
-\hl{Additionally, the Bayesian posterior trees performed way worth than the Bayesian consensus tree (Fig. 4, Table 1 and appendix C Fig. C5 and Tables C5, C6 and C7).
-This can be due to the fact that the Bayesian posterior trees are always resolved and thus more likely to contain wrongly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
-Conversely, the Bayesian consensus tree might not resolve nodes that are badly supported and thus more likely to contain only correctly resolved nodes (i.e.
+\hl{Additionally, the Bayesian posterior trees performed more poorly than the Bayesian consensus tree (Fig. 4, Table 1 and Appendix C Fig. C5 and Tables C5, C6 and C7).
+This may be because the Bayesian posterior trees are always resolved and thus more likely to contain incorrectly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
+Conversely, the Bayesian consensus tree might not resolve nodes that are poorly supported and thus is more likely to contain only correctly resolved nodes (i.e.
 increasing the Normalised Robinson-Foulds metric).}
@@
 
-% \hl{Another practical implication of our results regards the tree inference methods.
-% Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advice using the Bayesian consensus trees as a topological constraint for tree inferences using the Total-Evidence method such as tip-dating} \citep[\hl{e.g.}][\hl{; although it is possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution}]{ronquista2012,Wood01032013,BEASTmaster}.
-% \hl{Additionally, using the Bayesian consensus tree rather than the Maximum Likelihood can reduce the amount of ``false positive'' topologies.
-% As shown in figure 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve nodes weakly supported due to missing data than the Maximum Likelihood tree that is more likely to wrongly resolve such nodes (i.e.
-creating a ``false positive'' node).
-% Note, however, that we do not suggest discarding the Bayesian posterior tree distribution once one chose to use the Bayesian consensus tree topology even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
-% This can be particularly important because these trees will be invaluable for phylogenetic comparative analyses.
-% For example a sub-sample of posterior trees distributions can be used to asses macroecological questions while better taking into account topological uncertainty (e.g.
-}\citealt{FritzTree}\hl{ and }\citealt{jetzthe2012}\hl{ trees used in }\citealt{healy2014}\hl{).}
 \hl{Another practical implication of our results regards the tree inference methods.
 Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advise that where a topological constraint is needed, Bayesian consensus trees should be used.
-This may apply to tree inferences using the Total Evidence method such as tip-dating }\citep[e.g.][]{ronquista2012,Wood01032013,BEASTmaster}.
+This may apply to tree inferences using the Total Evidence method such as tip-dating }\citep[\hl{e.g.}][]{ronquista2012,Wood01032013,BEASTmaster}.
 \hl{It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
-Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies }\citep{Swofford2001}.
+Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies }\citep[\hl{sensu}][]{Swofford2001}.
 \hl{As shown in Fig.
@@
 These trees will be invaluable for phylogenetic comparative analyses.
-For example a sub-sample of posterior trees distributions can be used to assess macroecological questions while better taking into account topological uncertainty (e.g.
-}\citealt{FritzTree}\hl{ and }\citealt{jetzthe2012}\hl{ trees used in }\citealt{healy2014}\hl{).}
+For example a sub-sample of posterior trees distributions can be used to assess macroecological questions while better taking into account topological uncertainty }(\hl{e.g.
+}\citealt{fritzdiversity2013} and \citealt{jetzthe2012}\hl{ used in }\citealt{healy2014}).
 
 \section{Conclusions}
-% \hl{Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data }\citep{Wiens01102005,pattinsonphylogeny2014}.
 \hl{Previous studies have explored the effect of missing morphological data in Total Evidence matrices }\citep{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{The conclusions of theses studies, however, were limited by their empirical approach as well as several aspects of their experimental design (e.g.
-not including fossil taxa, }\citealt{Wiens01102005}\hl{; or focusing on the palaeontological aspect, }\citealt{manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{This study proposes a more in depth approach where missing data are generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
-This allowed us to confirm previous results that missing data can be especially problematic in small matrices }\citep{wiensmissing2003}\hl{ but also revealed the crucial importance of coding morphological data for living species.
+\hl{The conclusions of theses studies, however, were limited by their empirical approach making their results applicable only to similar missing data scenarios.
+Additionally, these studies focused either only on living taxa }\citep{Wiens01102005}\hl{ or on the patterns of missing data from the fossil record only }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{Here we instead used an approach where missing data were generated from simulated data and according to three clearly defined missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$) that removed data from both the living and fossil taxa.
+This allowed us to confirm previous results that missing data can be especially problematic in small matrices }\citep{wiensmissing2003}\hl{, but also revealed the crucial importance of coding morphological data for living species in Total Evidence phylogenies.
 Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
 When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
-Therefore we advise that one should focus on coding morphological characters for a large number of the living taxa present in the matrix (i.e.
-50\%) for accurately combining both living and fossil species in phylogenies.
-Doing so grants a good overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their actual shared derived characters rather than simply on available data.}
+Therefore we advise increased focus on coding morphological characters for a large number of the living taxa present in the matrix (i.e.
+at least 50\%) if the goal is to accurately combine both living and fossil species in phylogenies.
+Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data.}
 
 \hl{Additionally, the topology of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
-This has also been observed in empirical data} \cite[\hl{e.g.}][]{Arcila2015131}\hl{ where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
-This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds }\citep[\hl{e.g.}][]{slaterphylogenetic2013}\hl{, taxonomic constraints }\citep[\hl{e.g.}][]{slaterphylogenetic2013,beckancient2014}\hl{ or even by fixing the topology }\cite[\hl{e.g.}][]{ronquista2012}.
+This has also been observed in empirical data }\citep[\hl{e.g.}][]{Arcila2015131}\hl{ where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds }\citep[\hl{e.g.}][]{slaterphylogenetic2013}\hl{, taxonomic constraints }\citep[\hl{e.g.}][]{slaterphylogenetic2013,beckancient2014}\hl{ or even on fixing the topology }\citep[\hl{e.g.}][]{ronquista2012}.
 \hl{Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.}
-
-\hl{To conclude, the results of our analyses are encouraging and show that it is possible to accurately combine both neontological and palaeontological data in the same phylogeny as long as both types of data sufficiently overlap.
-Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.}
 
@@
 Conversely, orange regions denote a high overlap in Normalised metric values between the different parameter combinations (i.e.
-the parameters have a weak effect on the metric and thus the topological recovery)}.
+the parameters have a weak effect on the metric and thus the topological recovery).}
 
@@
 Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
-If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology.}\\
+If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology.
+Values $>$ 0.95 denote significantly similar distributions and values $<$ 0.05 denote significantly different distributions.}\\
 

### 2015-08-12 Manuscript/MPESubmission/Revision/response.tex
Checked manuscript

@@
 Manos et al. (2007) and Pattinson et al. (2014) include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices. 
-Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is likely (Guillerme and Cooper 2015).
+Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data are likely (Guillerme and Cooper 2015).
 Finally, Manos et al. (2007) and Pattinson et al. (2014), mainly focused on how missing data in fossil taxa affects the placement of fossils, ignoring the effects of missing data in living species.}
@@
 We ran this additional analysis by comparing the estimated $\alpha$ parameter for the morphological rates distribution to the ``true'' $\alpha$ parameter.
-We found results that the $\alpha$ parameter is usually slightly underestimated for the various ``missing-data'' trees unless the number of living species with no morphological data is high ($M_L$ = 75\%), then, the $\alpha$ are overestimated.
+We found results that the $\alpha$ parameter is usually slightly underestimated for the various ``missing-data'' trees unless the number of living species with no morphological data are high ($M_L$ = 75\%), then, the $\alpha$ are overestimated.
 These results are somewhat in agreement with our topological differences results: the number of living species seems to be a driving missing data parameter for inferring Total Evidence trees.
@@
 In the first case, we think that the difference between the methods is due to the fact that the Maximum Likelihood trees are more likely to display ``false positive'' topologies (i.e.
-clades that are incorrect but can be well supported; or not) as the Bayesian consensus tree, when many data is missing, are more likely to collapse nodes, resulting in trees with less resolution but more ``conserved clades'' (i.e.
+clades that are incorrect but can be well supported; or not) as the Bayesian consensus tree, when many data are missing, are more likely to collapse nodes, resulting in trees with less resolution but more ``conserved clades'' (i.e.
 the clade ((a;b);c)) could be collapsed in (a;b;c) which is still a corrected representation of the relations within the clades rather than ((b;c);a)).

### 2015-08-12 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Checked manuscript

@@
 \cite{manosphylogeny2007,pattinsonphylogeny2014}\hl{ include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices. 
-Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is likely }\citep{GuillermeCooperMissing}.
-\hl{Finally, }\cite{manosphylogeny2007,pattinsonphylogeny2014}\hl{ mainly focused on how missing data in fossil taxa affects the placement of fossils, ignoring the effects of missing data in living species.}
+Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data are likely }\citep{GuillermeCooperMissing}.
+\hl{Finally, }\cite{manosphylogeny2007}\hl{ and }\cite{pattinsonphylogeny2014}\hl{ mainly focused on how missing data in fossil taxa affects the placement of fossils, ignoring the effects of missing data in living species.}
 
@@
 This is partially due to difficulties with simulating branch lengths and topology simultaneously, \hl{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints }\citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
-Thus understanding the sensitivity of topology to missing \hl{data} are important for assessing the accuracy of tree estimation in the Total Evidence framework.
+Thus understanding the sensitivity of topology to missing data are important for assessing the accuracy of tree estimation in the Total Evidence framework.
 To our knowledge, this question has never been formally assessed.
@@
 \hl{Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model }\citep{lewisa2001,wrightbayesian2014}.
-hl{Note, however, that } \cite{wrightbayesian2014}\hl{ have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal }(\citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}\hl{ so this should not influence our results.}
+hl{Note, however, that } \cite{wrightbayesian2014}\hl{ have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal }\citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}\hl{ so this should not influence our results.}
 
@@
 \item{\hl{The proportion of missing data in the fossil record ($M_F$).}}
-\hl{TThis missing data parameter represents the completeness of the fossil record.
+\hl{This missing data parameter represents the completeness of the fossil record.
 Due to preservation biases, missing data for fossil taxa are common }\citep{sansomfossilization2013}.
@@
 \hl{Each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
-Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
+Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.}
 
-We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data'' matrices.
+\hl{We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data'' matrices.
 Note that one of these combinations ($M_L$=0\%; $M_F$=0\% and $N_C$=100) has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
@@
 This metric is bounded between zero, when the two trees are identical, and $2(n-2)$ (for two trees with $n$ taxa) when there is no shared clade in the two trees.
-This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa (\textit{(((a,b),c),((d,e),f))}), the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
+This metric is sensitive to minor changes in clade conservation: if the trees are composed of two clades of three taxa \textit{(((a,b),c),((d,e),f))}, the swapping of any two taxa will lead to a maximal score of the Robinson-Foulds distance indicating poor tree similarity.
 We normalised this metric following Bogdanowicz's Normalised Tree Similarity (NTS) method \citep{Bogdanowicz2012}.

### 2015-08-12 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Typos and one major comment on intro

@@
 \section{Introduction}
-Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, the \hl{many large-scale macroevolutionary studies focus solely on living species} (e.g.
+Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, \hl{many large-scale macroevolutionary studies focus solely on living species} (e.g.
 \citealp{meredithimpacts2011,jetzthe2012}).
@@
 This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny \citep[known as tip-dating;][]{ronquista2012}, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
-\hl{Total Evidence methods is becoming an increasingly popular way of adding fossil taxa to phylogenies} \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014,Arcila2015131}.
+\hl{The Total Evidence method is becoming an increasingly popular way of adding fossil taxa to phylogenies} \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014,Arcila2015131}.
 Although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
@@
 using both molecular and morphological data;}][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{These studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data }\citep{Wiens01102005}\hl{; or (2) removing both molecular and some morphological data from living taxa to create artificial fossils }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{These previous studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data }\citep{Wiens01102005}\hl{; or (2) removing both molecular and some morphological data from living taxa to create artificial fossils }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
 \hl{Both approaches have shown that missing data are not a major problem and should not be an obstacle to combining both living and fossil species in the same phylogenies.
@@
 \cite{manosphylogeny2007,pattinsonphylogeny2014}\hl{ include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices. 
-Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data are likely }\citep{GuillermeCooperMissing}.
-\hl{Finally, }\cite{manosphylogeny2007}\hl{ and }\cite{pattinsonphylogeny2014}\hl{ mainly focused on how missing data in fossil taxa affects the placement of fossils, ignoring the effects of missing data in living species.}
+Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is a likely pattern}\citep{GuillermeCooperMissing}.
+\hl{Finally, }\cite{manosphylogeny2007}\hl{ and }\cite{pattinsonphylogeny2014}\hl{ mainly focused on how missing data in fossil taxa affect the placement of fossils, ignoring the effects of missing data in living species.}
 
 \hl{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living taxa with morphological data, fossil data, all data for certain characters and the combination of these three aspects.
-This is an advance on previous studies because we use large-scale simulations and analyse the effect of three distinct aspect of missing data thus focusing on both neontological and paleontological parts of the matrix.
+This is an advance on previous studies because we use large-scale simulations and analyse the effects of three distinct aspects of missing data thus focusing on both neontological and paleontological parts of the matrix.
 In addition, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades }\citep[\hl{based on the Robinson-Foulds distance;}][]{RF1981}\hl{ and (ii) the displacement of wild-card taxa }\citep[\hl{based on the Triplets distance;}][]{critchlowthe1996}\hl{ rather than just a single measure of clade conservation or clade support }\citep[\hl{cf.}][]{Wiens01102005,pattinsonphylogeny2014}.
 
-We focus on the effect of missing data on our ability to recover a ``best'' tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
+We focus on the effects of missing data on our ability to recover tree topology because it is a crucial aspect of a phylogeny in many macroevolutionary studies, for example when trying to elucidate the evolutionary relationships among species \citep[e.g.][]{meredithimpacts2011,jetzthe2012}, or for studying evolutionary transitions \citep[e.g.][]{friedmanexplosive2010}.
 Although branch length estimation is also important \citep[namely for timing extinction and/or speciation events; e.g.][]{ronquista2012}, we do not consider branch lengths in this study.
 This is partially due to difficulties with simulating branch lengths and topology simultaneously, \hl{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints }\citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
-Thus understanding the sensitivity of topology to missing data are important for assessing the accuracy of tree estimation in the Total Evidence framework.
+Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework.
 To our knowledge, this question has never been formally assessed.
@@
 \hl{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
-We inferred the topology from the matrices using both Maximum Likelihood and Bayesian inference methods and measured the differences in topology using two different topological distance metrics as proxies for clade conservation } \citep[\hl{based on the Robinson-Foulds distance;}][]{RF1981}\hl{ and for wild-card taxa placement }\citep[\hl{based on the Triplets distance;}][]{critchlowthe1996}.
-\hl{Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring difference on two aspects of topology }\citep[cf.][\hl{measuring respectively only the clades support or the number of shared clades}]{Wiens01102005,pattinsonphylogeny2014}.
+We infer the topology from the matrices using both Maximum Likelihood and Bayesian inference methods and measure the differences in topology using two different topological distance metrics as proxies for clade conservation } \citep[\hl{based on the Robinson-Foulds distance;}][]{RF1981}\hl{ and for wild-card taxa placement }\citep[\hl{based on the Triplets distance;}][]{critchlowthe1996}.
+%\hl{Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring two aspects of topological recovery }\citep[cf.][\hl{measuring respectively only the clades support or the number of shared clades}]{Wiens01102005,pattinsonphylogeny2014}.
+
+% NC: Reading this in full it seems like the last bit repeats what you've already said.
+I know the reviewer asked for novelty to be spelled out but I think you do that well above here so maybe leave this bit out to avoid repetition.
+
+% NC: You've also cut out the "here we find" sentences. Did the reviewers ask for this? If not why not put them back and mention novelty in there instead subtly.
+
 %Former revised version

### 2015-08-13 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Minor edits to Methods

@@
 \hl{Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model }\citep{lewisa2001,wrightbayesian2014}.
-hl{Note, however, that } \cite{wrightbayesian2014}\hl{ have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal }\citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}\hl{ so this should not influence our results.}
+\hl{Note, however, that } \cite{wrightbayesian2014}\hl{ have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal }\citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}\hl{ so this should not influence our results.}
 
@@
 It represents the number of living taxa that are present in the matrix but have only molecular data available.
-This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species is rarely collected }\citep{GuillermeCooperMissing}.
+This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species are rarely collected }\citep{GuillermeCooperMissing}.
 \hl{Therefore, many living species will have only molecular data available.
@@
 Due to preservation biases, missing data for fossil taxa are common }\citep{sansomfossilization2013}.
-\hl{In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
+\hl{In practice, we randomly removed a proportion of data from across the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
 Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
 Instead this 50\% refers to missing fossil data across the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data; see below).}
+Some fossils may retain 100\% of their data and others may lose most of their data at this parameter value (down to a minimum threshold of 5\% available data; see below).}
 \item{\hl{The number of morphological characters for both living and fossil taxa ($N_C$).}}
@@
 Therefore, this parameter corresponds to the overall number of characters available for both living and fossil taxa.
-In practice, we randomly removed whole characters from the morphological matrix reducing it to: 100, 90,75, 50 or 25 characters.
+In practice, we randomly removed entire characters from the morphological matrix reducing it to: 100, 90, 75, 50 or 25 characters.
 Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).}
@@
 Because the main aim of this study is to look at the effect our missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees and ``missing data'' trees.
-The results of the comparisons of the ``true'' and the ``missing data'' tree are available in Appendix B.
+The results of the comparisons of the ``true'' tree and the ``missing data'' trees are available in Appendix B.
 Note that this makes little difference to our overall results.}
-Therefore, here we focus on comparing the trees inferred from the matrices with missing data to the ``best'' tree, rather than the ``true'' tree, as the ``best'' tree is generally what we have to work with.
 
@@
 We compared the topology of the ``missing-data'' trees to the ``best'' tree to measure the effect of the three parameters $M_{L}$, $M_{F}$ and $N_{C}$ on tree topology.
-We used the Robinson-Foulds distance \citep{RF1981} to assess the amount of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the amount of wildcard taxa \citep[i.e.
+We used the Robinson-Foulds distance \citep{RF1981} to assess the number of conserved clade positions and the Triplets distance \citep{dobson1975triplets} to assess the number of wildcard taxa \citep[i.e.
 taxa that frequently change position in different trees][]{kearneyfragmentary2002}.

### 2015-08-13 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Minor edits to Results

@@
 4; Tables C5, C6 and C7 in Appendix C).
-The Normalised Triplets metric, however, is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $N_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $N_{C}$ respectively; Fig.
+The Normalised Triplets metric, however, is similar when comparing the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $N_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $N_{C}$ respectively; Fig.
 4; Tables C5, C6 and C7 in Appendix C).

### 2015-08-13 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Minor edits to Discussion and captions

@@
 These factors, however, affected topological recovery in different ways and to different extents.
-Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($N_{C}$) had worst effects on topological recovery (Fig. 6).
+Decreasing the number of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($N_{C}$) had worst effects on topological recovery (Fig. 6).
 Additionally, using Bayesian consensus trees recovered the ``best'' tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.
@@
 
-Additionally, the number of morphological characters determines the size the of matrix.
+Additionally, the number of morphological characters determines the size of the matrix.
 This can affect our ability to recover the ``best'' tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
@@
 \hl{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details on the analysis of the effect of the starting tree in Bayesian inference).}
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details).}
 As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
@@
 
-\hl{The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (Fig.
+\hl{The Bayesian consensus trees, however, perform poorly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (Fig.
 5).
@@
 Even if clades are conserved (Fig. 5), the resolution within them can be poor to non-existent when a lot of data are missing (i.e. 75\%).
-In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
+In such cases, the fossil taxa are equally likely to be placed in any of the clades that they share the most characters with.
 These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters }\citep{wiensmissing2003}.
@@
 It is also worth noting that across all our analyses, the topologies of the Maximum Likelihood bootstrap trees and the Bayesian posterior trees distribution were always further from the ``best'' tree topology than Maximum Likelihood and Bayesian consensus trees.
-This was true even when no morphological data was missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $N_{C}$ = 100; Fig. 4).
+This was true even when no morphological data were missing ($M_{L}$ = 0\%; $M_{F}$ = 0\%, $N_{C}$ = 100; Fig. 4).
 This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best'' trees involved 1000 random pairwise comparisons rather than just one.
@@
 This may be because the Bayesian posterior trees are always resolved and thus more likely to contain incorrectly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
-Conversely, the Bayesian consensus tree might not resolve nodes that are poorly supported and thus is more likely to contain only correctly resolved nodes (i.e.
+Conversely, the Bayesian consensus trees might not resolve nodes that are poorly supported and thus are more likely to contain only correctly resolved nodes (i.e.
 increasing the Normalised Robinson-Foulds metric).}
@@
 Gaps in the matrix, however, can be filled with efforts in palaeontological field work that can potentially lead to future discoveries of exceptionally preserved fossils \citep[e.g.][]{nithe2013}.
-Fortunately, although this data are the most difficult to collect, it also has the least influence on whether our simulations recover the ``best'' tree topology (Fig. 6).
+Fortunately, although these data are the most difficult to collect, they also have the least influence on whether our simulations recover the ``best'' tree topology (Fig. 6).
 Finally, although increasing the number of coded characters is relatively straightforward, the amount of time it takes to build a morphological matrix increases directly with the number of characters involved.
-One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{MorphoBank} \citep{morphobank}, so that no one individual collects all the data.
+One solution to this problem may be to engage with collaborative data collection projects through web portals such as \textit{MorphoBank} \citep{morphobank}, so that no single individual collects all the data.
 
@@
 \hl{It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
-Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies }\citep[\hl{sensu}][]{Swofford2001}.
+Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies }\citep[\hl{\textit{sensu}}][]{Swofford2001}.
 \hl{As shown in Fig.
@@
 creating a false positive node).
-Note, however, that we do not suggest discarding the Bayesian posterior tree distributions even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
+Note, however, that we do not suggest discarding the Bayesian posterior tree distributions even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be traced to the difficulties comparing distributions of trees; see above).
 These trees will be invaluable for phylogenetic comparative analyses.
-For example a sub-sample of posterior trees distributions can be used to assess macroecological questions while better taking into account topological uncertainty }(\hl{e.g.
+For example a sub-sample of posterior tree distributions can be used to assess macroecological questions while better taking into account topological uncertainty }(\hl{e.g.
 }\citealt{fritzdiversity2013} and \citealt{jetzthe2012}\hl{ used in }\citealt{healy2014}).
@@
 
-\hl{Additionally, the topology of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+\hl{Additionally, the topologies of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
 This has also been observed in empirical data }\citep[\hl{e.g.}][]{Arcila2015131}\hl{ where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
@@
 Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
-Thanks to the two anonymous reviewers for their useful and enthusiastic % TG: or "useful and encouraging"?
-I feel just useful does gives justice to them, especially to the work of reviewer 1...
-comments.
+Thanks to two anonymous reviewers for their useful and encouraging comments.
 Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.
@@
 
+% NC: Do you think we should add to any of these that BC <0.5 = significantly different and BC > 95% is significantly similar? Up to you.
+
 \noindent
@@
 \textbf{Table 1:} \hl{Bhattacharyya Coefficients of the pairwise method comparisons.}
-\hl{Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
+\hl{Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robinson-Foulds (RF) and Triplets (Tr) metrics.
 Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.

### 2015-08-13 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex

@@
 
-Although missing data does not appear be a major problem in molecular and morphological matrices separately \citep[\hl{as long as enough data overlaps in each case, and missing data are not phylogeneticaly biased;}][]{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data do not appear be a major problem in molecular and morphological matrices separately \citep[\hl{as long as enough data overlaps in each case, and missing data are not phylogeneticaly biased;}][]{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
@@
 \cite{manosphylogeny2007,pattinsonphylogeny2014}\hl{ include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices. 
-Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is a likely pattern}\citep{GuillermeCooperMissing}.
+Data for living species are unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is a likely pattern}\citep{GuillermeCooperMissing}.
 \hl{Finally, }\cite{manosphylogeny2007}\hl{ and }\cite{pattinsonphylogeny2014}\hl{ mainly focused on how missing data in fossil taxa affect the placement of fossils, ignoring the effects of missing data in living species.}
@@
 \item{Estimating phylogenies:} \label{step:build_phylo} \\
-We inferred phylogenetic trees from the ``complete'' matrix and from the 125 ``missing-data'' matrices resulting in one tree generated from a matrix containing no missing data (hereafter called the ``best'' tree) and 125 trees inferred from the matrices with missing morphological data (hereafter called the ``missing-data'' trees).
+We inferred phylogenetic trees from the ``complete'' matrix and from the 125 ``missing-data'' matrices resulting in one tree generated from a matrix with no missing data (hereafter called the ``best'' tree) and 125 trees inferred from the matrices with missing morphological data (hereafter called the ``missing-data'' trees).
 Phylogenies were inferred via both Maximum Likelihood and Bayesian approaches.
@@
 This can be interpreted as the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
-Even if clades are conserved (Fig. 5), the resolution within them can be poor to non-existent when a lot of data are missing (i.e. 75\%).
+Even if clades are conserved (Fig. 5), the resolution within them can be poor to non-existent when a large proportion of data are missing (i.e. 75\%).
 In such cases, the fossil taxa are equally likely to be placed in any of the clades that they share the most characters with.

### 2015-08-13 Manuscript/MPESubmission/Revision/response.tex
Matched response and manuscript

@@
 Wiens et al., 2005; Manos et al., 2007; Pattinson et al., 2014).
-These studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data (Wiens et al.
+These previous studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data (Wiens et al.
 2005); or (2) removing both molecular and some morphological data from living taxa to create artificial fossils (Manos et al., 2007; Pattinson et al., 2014).
@@
 Manos et al. (2007) and Pattinson et al. (2014) include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices. 
-Data for living species is unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data are likely (Guillerme and Cooper 2015).
-Finally, Manos et al. (2007) and Pattinson et al. (2014), mainly focused on how missing data in fossil taxa affects the placement of fossils, ignoring the effects of missing data in living species.}
+Data for living species are unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is a likely pattern (Guillerme and Cooper 2015).
+Finally, Manos et al. (2007) and Pattinson et al. (2014), mainly focused on how missing data in fossil taxa affect the placement of fossils, ignoring the effects of missing data in living species.}
 
 \textit{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living taxa with morphological data, fossil data, all data for certain characters and the combination of these three aspects.
-This is an advance on previous studies because we use large-scale simulations and analyze the effect of three distinct aspect of missing data thus focusing on both neontological and paleontological parts of the matrix.
+This is an advance on previous studies because we use large-scale simulations and analyze the effects of three distinct aspects of missing data thus focusing on both neontological and paleontological parts of the matrix.
 In addition, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades (based on the Robinson-Foulds distance; Robinson and Foulds, 1981) and (ii) the displacement of wild-card taxa (based on the Triplets distance; Critchlow et al., 1996) rather than just a single measure of clade conservation or clade support (cf.
-Wiens et al., 2005; Pattinson et al., 2014).} lines 85-118@@@
+Wiens et al., 2005; Pattinson et al., 2014).} lines 85-118
 
@@
 Because the main aim of this study is to look at the effect our missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees and ``missing data'' trees.
-The results of the comparisons of the ``true'' and the ``missing data'' tree are available in Appendix B.
+The results of the comparisons of the ``true'' tree and the ``missing data'' trees are available in Appendix B.
 Note that this makes little difference to our overall results.} line 297-308 @@@.
@@
 This may be because the Bayesian posterior trees are always resolved and thus more likely to contain incorrectly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
-Conversely, the Bayesian consensus tree might not resolve nodes that are poorly supported and thus is more likely to contain only correctly resolved nodes (i.e.
+Conversely, the Bayesian consensus trees might not resolve nodes that are poorly supported and thus are more likely to contain only correctly resolved nodes (i.e.
 increasing the Normalised Robinson-Foulds metric).} lines 681-687.
@@
 
-\textit{The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (Fig.
+\textit{The Bayesian consensus trees, however, perform poorly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (Fig.
 5).
@@
 This can be interpreted as the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
-Even if clades are conserved (Fig. 5), the resolution within them can be poor to non-existent when a lot of data are missing (i.e. 75\%).
-In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
+Even if clades are conserved (Fig. 5), the resolution within them can be poor to non-existent when a large proportion of data are missing (i.e. 75\%).
+In such cases, the fossil taxa are equally likely to be placed in any of the clades that they share the most characters with.
 These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters (Wiens 2003).
@@
 
-\textit{Additionally, the topology of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+\textit{Additionally, the topologies of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
 This has also been observed in empirical data (e.g. Arcila et al., 2015) where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
@@
 
-\textit{Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robins-Foulds (RF) and Triplets (Tr) metrics.
+\textit{Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robinson-Foulds (RF) and Triplets (Tr) metrics.
 Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
@@
 
-We revised this statement as follows: \textit{many large-scale macroevolutionary studies focus solely on living species} (line 30-31 @@@).
+We revised this statement as follows: \textit{many large-scale macroevolutionary studies focus solely on living species} (line 30-31).
 
@@
 
+We modified the sentence as suggested lines (64-66).
+
+% -----------------
+% Specific Comment 8
+% -----------------
+\item{\textcolor{blue}{\textbf{line 76 and throughout the manuscript:} watch out for inconsistent parentheses.}}
+
+We fixed this issue.
+
+% -----------------
+% Specific Comment 9
+% -----------------
+\item{\textcolor{blue}{\textbf{line 76 and throughout:} correct use of the word 'data' from singular to plural throughout}}
+
+We thank the reviewer for pointing out this inconsistency and changed the use of ``data'' from singular to plural throughout the manuscript.
+
+% -----------------
+% Specific Comment 10
+% -----------------
+\item{\textcolor{blue}{\textbf{line 77:} change 'a lot' to 'large proportions of' or something more accurate and less colloquial}}
+
 We modified the sentence as suggested.
@@
 % -----------------
-% Specific Comment 8
-% -----------------
-\item{\textcolor{blue}{\textbf{line 76 and throughout the manuscript:} watch out for inconsistent parentheses.}}
-
-We fixed this issue.
-
-% -----------------
-% Specific Comment 9
-% -----------------
-\item{\textcolor{blue}{\textbf{line 76 and throughout:} correct use of the word 'data' from singular to plural throughout}}
-
-We thank the reviewer for pointing out this inconsistency and changed the use of ``data'' from singular to plural throughout the manuscript.
-
-% -----------------
-% Specific Comment 10
-% -----------------
-\item{\textcolor{blue}{\textbf{line 77:} change 'a lot' to 'large proportions of' or something more accurate and less colloquial}}
-
-We modified the sentence as suggested.
-
-% -----------------
 % Specific Comment 11
@@
 
-We modified the sentence as suggested.
+We modified the sentence as suggested (lines 79-80).
 
@@
 \textit{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details on the analysis of the effect of the starting tree in Bayesian inference).} lines 649-655@@@
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details).} lines 649-655@@@
 
@@
 It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
-Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies (sensu Swofford et al., 2001).
+Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies (\textit{sensu} Swofford et al., 2001).
 As shown in Fig.
@@
 creating a false positive node).
-Note, however, that we do not suggest discarding the Bayesian posterior tree distributions even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be imputed to the difficulties comparing distributions of trees; see above).
+Note, however, that we do not suggest discarding the Bayesian posterior tree distributions even though they performed poorly in recovering the ``best'' tree topology in our simulations (this can probably be traced to the difficulties comparing distributions of trees; see above).
 These trees will be invaluable for phylogenetic comparative analyses.
-For example a sub-sample of posterior trees distributions can be used to assess macroecological questions while better taking into account topological uncertainty (e.g.
+For example a sub-sample of posterior tree distributions can be used to assess macroecological questions while better taking into account topological uncertainty (e.g.
 Fritz et al.
@@
 It represents the number of living taxa that are present in the matrix but have only molecular data available.
-This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species is rarely collected (Guillerme and Cooper, 2015).
+This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species are rarely collected (Guillerme and Cooper, 2015).
 Therefore, many living species will have only molecular data available.
@@
 Due to preservation biases, missing data for fossil taxa are common (Sansom and Wills, 2013).
-In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
+In practice, we randomly removed a proportion of data from across the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
 Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
 Instead this 50\% refers to missing fossil data across the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data; see below)
+Some fossils may retain 100\% of their data and others may lose most of their data at this parameter value (down to a minimum threshold of 5\% available data; see below)
 \item{The number of morphological characters for both living and fossil taxa ($N_C$).}
@@
 Therefore, this parameter corresponds to the overall number of characters available for both living and fossil taxa.
-In practice, we randomly removed whole characters from the morphological matrix reducing it to: 100, 90,75, 50 or 25 characters.
+In practice, we randomly removed entire characters from the morphological matrix reducing it to: 100, 90,75, 50 or 25 characters.
 Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).

### 2015-08-13 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Matched response and manuscript

@@
 Therefore, we focus only on missing data in the morphological part of the matrix.
-We investigate three major parameters that directly affect the completeness and size of the morphological part of the matrix, and reflect empirical biases in data availability: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the proportion of missing morphological characters for both living and fossil taxa in the matrix (i.e.
+We investigate three major parameters that directly affect the completeness and size of the morphological part of the matrix, and reflect empirical biases in data availability: (i) the proportion of living taxa with no morphological data; (ii) the proportion of missing data in the fossil taxa; and (iii) the amount of morphological characters for both living and fossil taxa in the matrix (i.e.
 the size of the matrix).
-%in
-\hl{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
-We infer the topology from the matrices using both Maximum Likelihood and Bayesian inference methods and measure the differences in topology using two different topological distance metrics as proxies for clade conservation } \citep[\hl{based on the Robinson-Foulds distance;}][]{RF1981}\hl{ and for wild-card taxa placement }\citep[\hl{based on the Triplets distance;}][]{critchlowthe1996}.
-%\hl{Thus, in the present study, we propose a new approach to the missing-data question, focusing on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and measuring two aspects of topological recovery }\citep[cf.][\hl{measuring respectively only the clades support or the number of shared clades}]{Wiens01102005,pattinsonphylogeny2014}.
-
-% NC: Reading this in full it seems like the last bit repeats what you've already said.
-I know the reviewer asked for novelty to be spelled out but I think you do that well above here so maybe leave this bit out to avoid repetition.
-
-% NC: You've also cut out the "here we find" sentences. Did the reviewers ask for this? If not why not put them back and mention novelty in there instead subtly.
-
-%Former revised version
-% \hl{We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the topology of trees inferred using Maximum Likelihood and Bayesian methods by using two different topological metrics, based on the Robinson-Foulds }\citep{RF1981}\hl{ and Triplets }\citep{critchlowthe1996}\hl{ distances.
-% Thus, in the present study, we propose a new approach on the missing-data question focus on both palaeontological and neontological data (rather than just palaeontological data), using only simulated data and two distinct topological metrics.}
-% We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the ``best'' tree topology more so than minimizing the amount of missing data in the fossil record.
-% Additionally, we find that the ability of Total Evidence methods to recover the ``best'' tree topology is increased when using Bayesian methods.
+We remove data from a Total Evidence matrix by changing the values of these three parameters and then assess how this affects the resulting tree topology.
+We infer the topology from the matrices using both Maximum Likelihood and Bayesian inference methods and measure the differences in topology using two different topological distance metrics as proxies for clade conservation and for wild-card taxa placement.
+We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the ``best'' tree topology more so than minimizing the amount of missing data in the fossil record.
+Additionally, we find that the ability of Total Evidence methods to recover the ``best'' tree topology is increased when using Bayesian methods.
 

### 2015-08-13 Manuscript/MPESubmission/Revision/response.tex
Added line numbers to repsonse

@@
 When comparing the ``true'' tree to the ``best'' trees, in Maximum Likelihood the 50\% CI lays between approximately 0.2 and 0.3 for the Normalised Robinson-Foulds and Triplets metric; and, in Bayesian consensus tree  the 50\% CI lays between approximately 0.3 and 0.4 for the Normalised Robinson-Foulds metric and between 0 and 0.2 for the Normalised Triplets metric).
-This may be related to the size of our matrices or to the ability of the algorithms to generate the ``complete matrix'' but also to our decision to put our minimum median bootstrap acceptance at 50 rather than higher values (line @@@ in the former manuscript and @@@ in our revision).
+This may be related to the size of our matrices or to the ability of the algorithms to generate the ``complete matrix'' but also to our decision to put our minimum median bootstrap acceptance at 50 rather than higher values.
 
@@
 \begin{enumerate}
-\item{The true tree is always unknown:} as we state in the manuscript: "In real world situations, the ``true'' tree is never available" (line 242 in the first manuscript and @@@ in the revision).
+\item{The true tree is always unknown:} as we state in the manuscript: "In real world situations, the ``true'' tree is never available" (line 242 in the first manuscript and 295 in the revision).
 \item{The true tree is just the starting point of the simulations:} we use the true tree only as a starting point for our simulations.
@@
 The results of the comparisons of the ``true'' tree and the ``missing data'' trees are available in Appendix B.
-Note that this makes little difference to our overall results.} line 297-308 @@@.
+Note that this makes little difference to our overall results.} lines 297-308.
 
@@
 \textit{We used a fixed starting tree rather than a random starting tree (default MrBayes; Ronquist et al., 2012b) to speed up our Bayesian inferences.
-Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').} line 352-357 @@@.
+Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').} line 349-354.
 See also Appendix A section "Effect of the starting tree on Bayesian inference".
@@
 
-\textit{The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 (full overlap; Bhattacharyya, 1943, see Appendix B for calculation details).} lines 445-447@@@
+\textit{The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 (full overlap; Bhattacharyya, 1943, see Appendix B for calculation details).} lines 444-446.
 
@@
 When this coefficient is $>$0.95, the distributions are significantly similar.
-Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.} line 449-454 @@@.
+Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.} line 448-453.
 
@@
 \end{enumerate}
-Both points were discussed in the original manuscript lines 566-573 (or lines 645-652 @@@ in the revised version).
+Both points were discussed in the original manuscript lines 566-573 (or lines 672-678 in the revised version).
 We expanded the paragraph to:
@@
 Conversely, the Bayesian consensus trees might not resolve nodes that are poorly supported and thus are more likely to contain only correctly resolved nodes (i.e.
-increasing the Normalised Robinson-Foulds metric).} lines 681-687.
-@@@
+increasing the Normalised Robinson-Foulds metric).} lines 679-685.
 
@@
 These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters (Wiens 2003).
-It is important to note, however, that this effect can be reduced by increasing the number of characters (Wiens 2003).} lines 661-673. @@@
+It is important to note, however, that this effect can be reduced by increasing the number of characters (Wiens 2003).} lines 659-671.
 
@@
 Ronquist et al., 2012a).
-Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.} lines 731-62@@@
+Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.} lines 751-760.
 
@@
 
-\textit{These results are in agreement with Manos et al. (2007) where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.} lines 566-567@@@
+\textit{These results are in agreement with Manos et al. (2007) where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.} lines 565-566.
 
@@
 We modified the sentence as follows: \textit{These approaches differ mainly in how they treat fossil taxa and their data.
-One can use fossils as tips or as nodes in the phylogeny, and can use only the age of the fossils, only the morphology of the fossils, or age and morphology jointly.} (line 42-44@@@)
+One can use fossils as tips or as nodes in the phylogeny, and can use only the age of the fossils, only the morphology of the fossils, or age and morphology jointly.} (line 42-44)
 
@@
 
-We modified the citation as suggested by reviewer 1 (line 48@@@).
+We modified the citation as suggested by reviewer 1 (line 48).
 
@@
 
-We modified the statement to: \textit{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints} (line 126-128@@@).
+We modified the statement to: \textit{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints} (line 126-128).
 
@@
 We also included an analysis showing the effect of missing data on the estimation of the shape parameter ($\alpha$) of the morphological substitution rate distribution.
-This extra analysis, however, is beyond the scope of this paper so the results are not discussed further here.} lines 366-370 @@@
+This extra analysis, however, is beyond the scope of this paper so the results are not discussed further here.} lines 363-367.
 See also Appendix A "Morphological rates estimations results".
@@
 \textit{In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal (Hassanin et al., 1998; Estoup et al., 2002).%\citep{Hassanin1998611,EstoupHomoplasy}.
-Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below).} (line 204-206@@@).
+Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below).} (line 203-206).
 
@@
 Douady et al., 2003; Kelly et al., 2014) and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years)}.
-lines 206-213@@@
+lines 206-212.
 
@@
 
-We now state that we used the rTraitDisc function on line 214 @@@.
+We now state that we used the rTraitDisc function on line 213.
 
@@
 Therefore, we implemented the \textit{a posteriori} matrix selection step to our analysis (i.e.
-keeping only the ``complete'' matrices that led to trees with a median bootstrap $>$ 50; explained lines 259-268 in the submitted manuscript and lines @@@ and @@@ in the revised one).
+keeping only the ``complete'' matrices that led to trees with a median bootstrap $>$ 50; explained lines 259-268 in the submitted manuscript and lines 321 and 327 in the revised one).
 
@@
 Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model (Lewis, 2001; Wright and Hillis, 2014).
-Note, however, that Wright and Hillis (2014) have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004) so this should not influence our results.} lines 225-233 @@@
+Note, however, that Wright and Hillis (2014) have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a “fair” phylogenetic signal (see Estimating phylogenies section below; Zander, 2004) so this should not influence our results.} lines 224-232.
 
@@
 
-We modified the sentence as follows: \textit{To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.} (line 285-287@@@).
+We modified the sentence as follows: \textit{To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.} (line 285-287).
 
@@
 
-We fixed this typo as well as in the following sentence (line @@@).
+We fixed this typo.
 
@@
 We now refer to two more classical citations: Bremer and Struwe 1992 (American Journal of Botany) and Patterson et al.
-1993 (Annual Review of Ecology and Systematics), on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to having a small matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters) lines 609-610@@@.
+1993 (Annual Review of Ecology and Systematics), on phylogenetic signal and congruence between morphological and molecular data and also refer to a classical example of conflict probably due to having a small matrix (Masters and Brothers 2002; American Journal of Physical Anthropology; 36 characters) lines 606-607.
 
@@
 
-We modified the sentence as follows: \textit{The sizes of our data matrices were constrained by the performance of our protocol} (line 579-580@@@).
+We modified the sentence as follows: \textit{The sizes of our data matrices were constrained by the performance of our protocol} (line 607-608).
 
@@
 A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Fig. 6).
-This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).} lines 630-640@@@
+This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).} lines 627-637.
 
@@
 \textit{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details).} lines 649-655@@@
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details).} lines 646-651.
 
@@
 2012 trees used in Healy et al.
-2014).} lines 711-731@@@
+2014).} lines 707-727.
 
@@
 If requested, however, we can also give additional details on the cluster used in the methods section.
-The current version of the manuscript only contains information on the nodes clock speed (2.30GHz clock speed nodes) lines @@@.
+The current version of the manuscript only contains information on the nodes clock speed (2.30GHz clock speed nodes) line 333.
 
@@
 In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Fig.
-6).} lines 467-471@@@
+6).} lines 464-468.
 
@@
 6A).
-In Fig. 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.} lines 512-526@@@
+In Fig. 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.} lines 509-523.
 
@@
 To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
-Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).} lines 238-289@@@
+Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).} lines 237-288.
 
@@
 Instead this 50\% refers to missing fossil data across the whole matrix.
-Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data, see below).} lines 259-263 @@@
+Some fossils may retain 100\% of their data and others may lose all their data at this parameter value (down to a minimum threshold of 5\% available data, see below).} lines 258-262.
 

### 2015-08-24 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Rev1 edits

@@
 \documentclass[12pt,letterpaper]{article}
-
-%---------------------------------------------
-%
-%       THIS IS THE HIGHLIGHTED VERSION.
-%       For the final version, carefully remove:
-%       '\protect'
-%       '\hl{' (don't forget the '}'
-%
-%---------------------------------------------
 
@@
 
-\noindent {\normalsize \sc Thomas Guillerme$^a$$^,$$^b$$^*$, and Natalie Cooper$^a$$^,$$^b$$^,$$^1$}\\
+\noindent {\normalsize \sc Thomas Guillerme$^a$$^,$$^b$$^*$, and Natalie Cooper$^a$$^,$$^b$$^,$$^c$}\\
 \noindent {\small \it 
@@
 $^b$Trinity Centre for Biodiversity Research, Trinity College Dublin, Dublin 2, Ireland.\\
-$^*$Corresponding author. Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.\\
-$^1$Present address: Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK. E-mail: natalie.cooper@nhm.ac.uk}\\
+$^c$Department of Life Sciences, Natural History Museum, Cromwell Road, London, SW7 5BD, UK.\\% E-mail: natalie.cooper@nhm.ac.uk}\\
+$^*$Corresponding author. Zoology Building, Trinity College Dublin, Dublin 2, Ireland; E-mail: guillert@tcd.ie; Fax: +353 1 6778094; Tel: +353 1 896 2571.}\\
+
 \end{center}
@@
 \section{Introduction}
-Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, \hl{many large-scale macroevolutionary studies focus solely on living species} (e.g.
+Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, many large-scale macroevolutionary studies focus solely on living species (e.g.
 \citealp{meredithimpacts2011,jetzthe2012}).
@@
 Up to now, three main approaches have been used to place both living and fossil taxa into phylogenies.
-\hl{These approaches differ mainly in how they treat fossil taxa and their data.
-One can use fossils as tips or as nodes in the phylogeny, and can use only the age of the fossils, only the morphology of the fossils, or age and morphology jointly.}
+These approaches differ mainly in how they treat fossil taxa and their data.
+One can use fossils as tips or as nodes in the phylogeny, and can use only the age of the fossils, only the morphology of the fossils, or age and morphology jointly.
 Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
@@
 Neontologists, on the other hand, more commonly use probabilistic approaches (e.g. Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-\hl{Because fossil taxa do not usually have available DNA, only fossil occurrence dates are used to time calibrate phylogenies }\citep{zuckerkandl1965}.
+Because fossil taxa do not usually have available DNA, only fossil occurrence dates are used to time calibrate phylogenies \citep{zuckerkandl1965}.
 There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the ``best'' approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
@@
 This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny \citep[known as tip-dating;][]{ronquista2012}, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
-\hl{The Total Evidence method is becoming an increasingly popular way of adding fossil taxa to phylogenies} \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014,Arcila2015131}.
+The Total Evidence method is becoming an increasingly popular way of adding fossil taxa to phylogenies \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014,Arcila2015131}.
 Although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
@@
 \citealp{O'Leary08022013} \textit{vs.} \citealp{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record.
-For example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data \hl{are} (nearly) always unavailable.
-Therefore \hl{Total Evidence matrices are likely to contain a large proportions of missing data} that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
+For example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data are (nearly) always unavailable.
+Therefore Total Evidence matrices are likely to contain a large proportion of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
 
-Although missing data do not appear be a major problem in molecular and morphological matrices separately \citep[\hl{as long as enough data overlaps in each case, and missing data are not phylogeneticaly biased;}][]{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data do not appear be a major problem in molecular and morphological matrices separately \citep[as long as enough data overlap in each case, and missing data are not phylogenetically biased;][]{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
-\hl{Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework }\citep[\hl{i.e.
-using both molecular and morphological data;}][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{These previous studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data }\citep{Wiens01102005}\hl{; or (2) removing both molecular and some morphological data from living taxa to create artificial fossils }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{Both approaches have shown that missing data are not a major problem and should not be an obstacle to combining both living and fossil species in the same phylogenies.
+Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework \citep[i.e.
+using both molecular and morphological data;][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
+These previous studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data \citep{Wiens01102005}; or (2) removing both molecular and some morphological data from living taxa to create artificial fossils \citep{manosphylogeny2007,pattinsonphylogeny2014}.
+Both approaches have shown that missing data are not a major problem and should not be an obstacle to combining both living and fossil species in the same phylogenies.
 The way these studies were conducted, however, means that their conclusions are not generally applicable across all scenarios involving missing data in Total Evidence phylogenies.
 For example, using an empirical (rather than simulation based) approach limits their conclusions to studies with similar distributions of data across species in the phylogeny.
-Additionally, }\cite{Wiens01102005}\hl{ did not include fossil taxa in their analyses, so their results cannot be used to make conclusions about how missing data may influence the placement of fossils.}
-\cite{manosphylogeny2007,pattinsonphylogeny2014}\hl{ include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices. 
+Additionally, one of the three previous studies did not include fossil taxa in their analyses, so their results cannot be used to make conclusions about how missing data may influence the placement of fossils \citep{wiensmissing2003}.
+The other two studies did include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices \citep{manosphylogeny2007,pattinsonphylogeny2014}.
 Data for living species are unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is a likely pattern}\citep{GuillermeCooperMissing}.
-\hl{Finally, }\cite{manosphylogeny2007}\hl{ and }\cite{pattinsonphylogeny2014}\hl{ mainly focused on how missing data in fossil taxa affect the placement of fossils, ignoring the effects of missing data in living species.}
+Finally, these previous studies mainly focused on how missing data in fossil taxa affect the placement of fossils, ignoring the effects of missing data in living species \citep{manosphylogeny2007,pattinsonphylogeny2014}.
 
-\hl{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living taxa with morphological data, fossil data, all data for certain characters and the combination of these three aspects.
+In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living taxa with morphological data, fossil data, all data for certain characters and the combination of these three aspects.
 This is an advance on previous studies because we use large-scale simulations and analyse the effects of three distinct aspects of missing data thus focusing on both neontological and paleontological parts of the matrix.
-In addition, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades }\citep[\hl{based on the Robinson-Foulds distance;}][]{RF1981}\hl{ and (ii) the displacement of wild-card taxa }\citep[\hl{based on the Triplets distance;}][]{critchlowthe1996}\hl{ rather than just a single measure of clade conservation or clade support }\citep[\hl{cf.}][]{Wiens01102005,pattinsonphylogeny2014}.
+In addition, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades \citep[based on the Robinson-Foulds distance;][]{RF1981} and (ii) the displacement of wild-card taxa \citep[based on the Triplets distance;][]{critchlowthe1996} rather than just a single measure of clade conservation or clade support \citep[cf.][]{Wiens01102005,pattinsonphylogeny2014}.
 
@@
 Although branch length estimation is also important \citep[namely for timing extinction and/or speciation events; e.g.][]{ronquista2012}, we do not consider branch lengths in this study.
-This is partially due to difficulties with simulating branch lengths and topology simultaneously, \hl{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints }\citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+This is partially due to difficulties with simulating branch lengths and topology simultaneously, but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework.
@@
 \section{Materials and Methods}
-To explore how missing data in \hl{the morphological partition of Total Evidence matrices} influences tree topology, \hl{we used the following protocol (Fig. 1):}
+To explore how missing data in the morphological partition of Total Evidence matrices influences tree topology, we used the following protocol (Fig. 1):
 \begin{enumerate}
@@
 Next, we generated a molecular and a morphological matrix from the ``true'' tree.
-The molecular matrix was \hl{simulated} from the ``true'' tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
+The molecular matrix was simulated from the ``true'' tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
 The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1) and transition/transversion rate of two \citep{douadycomparison2003}.
-\hl{The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5} \citep{yangamong-site1996}.
-\hl{In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal }\citep{Hassanin1998611,EstoupHomoplasy}.
-\hl{Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below).}
-\hl{This model and these parameter settings strike a balance between realism for empirical datasets }\citep[\hl{e.g.}][]{douadycomparison2003,kellymolecular2014}\hl{ and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).}
+The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
+In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal \citep{Hassanin1998611,EstoupHomoplasy}.
+Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below).
+This model and these parameter settings strike a balance between realism for empirical datasets \citep[e.g.][]{douadycomparison2003,kellymolecular2014} and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).
 All the molecular information for fossil taxa was replaced by missing data ("?").
 
-\hl{We simulated the morphological matrix using the rTraitDisc function from the R package ape v. 3.0-11} \citep{paradisape:2004}\hl{ to generate a matrix of 100 character sites for 51 taxa.}
-\hl{We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A and Fig.
-A1 within).}
+We simulated the morphological matrix using the rTraitDisc function from the R package ape v. 3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
+We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A and Fig.
+A1 within).
 We then ran an independent discrete character simulation for each character using the ``true'' tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
@@
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-\hl{We used low evolutionary rate parameters to be consistent with the molecular rate parameters, to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal }\citep{wrightbayesian2014}.
-\hl{Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model }\citep{lewisa2001,wrightbayesian2014}.
-\hl{Note, however, that } \cite{wrightbayesian2014}\hl{ have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal }\citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}\hl{ so this should not influence our results.}
+We used low evolutionary rate parameters to be consistent with the molecular rate parameters, to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wrightbayesian2014}.
+Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model \citep{lewisa2001,wrightbayesian2014}.
+Note, however, that \cite{wrightbayesian2014} have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal \citep[see Estimating phylogenies section below;][]{zanderminimal2004} so this should not influence our results.
 
@@
 \subsection{Removing data}
-\hl{To explore the effect of missing morphological data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
+To explore the effect of missing morphological data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
 Hereafter, we call these matrices with missing morphological data the ``missing-data'' matrices.
 Note that the amount of molecular data remained constant throughout our simulations: 1000 molecular characters for living taxa and no molecular data for fossil taxa (see above).
-We removed morphological data using three data incompleteness parameters:}
+We removed morphological data using three data incompleteness parameters:
 \begin{enumerate}
-\item{\hl{The proportion of missing living taxa ($M_L$).}}
-\hl{This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
+\item{The proportion of missing living taxa ($M_L$).}
+This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
 It represents the number of living taxa that are present in the matrix but have only molecular data available.
-This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species are rarely collected }\citep{GuillermeCooperMissing}.
-\hl{Therefore, many living species will have only molecular data available.
-In practice, we removed all the morphological data from randomly chosen living taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.}
-\item{\hl{The proportion of missing data in the fossil record ($M_F$).}}
-\hl{This missing data parameter represents the completeness of the fossil record.
-Due to preservation biases, missing data for fossil taxa are common }\citep{sansomfossilization2013}.
-\hl{In practice, we randomly removed a proportion of data from across the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
+This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species are rarely collected \citep{GuillermeCooperMissing}.
+Therefore, many living species will have only molecular data available.
+In practice, we removed all the morphological data from randomly chosen living taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.
+\item{The proportion of missing data in the fossil record ($M_F$).}
+This missing data parameter represents the completeness of the fossil record.
+Due to preservation biases, missing data for fossil taxa are common \citep{sansomfossilization2013}.
+In practice, we randomly removed a proportion of data from across the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
 Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
 Instead this 50\% refers to missing fossil data across the whole matrix.
-Some fossils may retain 100\% of their data and others may lose most of their data at this parameter value (down to a minimum threshold of 5\% available data; see below).}
-\item{\hl{The number of morphological characters for both living and fossil taxa ($N_C$).}}
-\hl{This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
+Some fossils may retain 100\% of their data and others may lose most of their data at this parameter value (down to a minimum threshold of 5\% available data; see below).
+\item{The number of morphological characters for both living and fossil taxa ($N_C$).}
+This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
 Any morphological matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
@@
 In practice, we randomly removed entire characters from the morphological matrix reducing it to: 100, 90, 75, 50 or 25 characters.
-Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).}
+Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).
 \end{enumerate}
 
-\hl{Each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
-Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.}
+Each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
+Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
 
-\hl{We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data'' matrices.
+We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data'' matrices.
 Note that one of these combinations ($M_L$=0\%; $M_F$=0\% and $N_C$=100) has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
@@
 To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
-Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).}
+Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).
 
@@
 In real world situations, the ``true'' tree is never available to us because we cannot know the true evolutionary history of a clade \citep[except in very rare circumstances, e.g.][]{rozen2005}.
-\hl{We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
+We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
 In practice, the difference between the ``best'' trees and the ``missing data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing data'' trees.
@@
 The results of the comparisons of the ``true'' tree and the ``missing data'' trees are available in Appendix B.
-Note that this makes little difference to our overall results.}
+Note that this makes little difference to our overall results.
 
@@
 For each estimation, we used the ``true'' tree's topology as a starting tree (with a starting value for each branch length of one).
-\hl{We used a fixed starting tree rather than a random starting tree }\citep[\hl{default MrBayes;}][]{Ronquist2012mrbayes} \hl{to speed up our Bayesian inferences.
-Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').}
+We used a fixed starting tree rather than a random starting tree \citep[default MrBayes;][]{Ronquist2012mrbayes} to speed up our Bayesian inferences.
+Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').
 We also used two priors on the molecular part of the matrix: an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5, and a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)); and one prior on the morphological part of the matrix (exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5).
@@
 Even using these priors, it took $~$ 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-\hl{The detailed MrBayes parameters are available in Appendix A.
+The detailed MrBayes parameters are available in Appendix A.
 We also included an analysis showing the effect of missing data on the estimation of the shape parameter ($\alpha$) of the morphological substitution rate distribution.
-This extra analysis, however, is beyond the scope of this paper so the results are not discussed further here.}
+This extra analysis, however, is beyond the scope of this paper so the results are not discussed further here.
 
@@
 \subsubsection{Robinson-Foulds distance}
-The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and \hl{twice the number} of shared clades across two trees.
+The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and twice the number of shared clades across two trees.
 The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} (see Appendix B for calculation details).
@@
 We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-\hl{The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 }\citep[\hl{full overlap;}][\hl{see Appendix B for calculation details}]{Bhattacharyya}.
+The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 \citep[full overlap;][see Appendix B for calculation details]{Bhattacharyya}.
 Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
-\hl{When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
+When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
 When this coefficient is $>$0.95, the distributions are significantly similar.
-Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.}
-\hl{To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
+Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.
+To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $N_{C}$ = 100 etc.
 (see Fig.
-2 for more details).}
+2 for more details).
 This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
@@
 Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
-\hl{Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Fig. 6).
+Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Fig. 6).
 In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Fig.
-6).}
+6).
 
@@
 
-\hl{Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Fig.
@@
 6A).
-In Fig. 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.}
+In Fig. 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.
 
@@
 Normalised Robinson-Foulds metric) is still relatively good (mode = 0.72) when the proportion of missing data are high ($M_{F}$ = 75\%).
-\hl{These results are in agreement with }\cite{manosphylogeny2007}\hl{ where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.}
+These results are in agreement with \cite{manosphylogeny2007} where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.
 
@@
 every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
-\hl{In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.}
+In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.
 Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxon and decrease the Normalised Triplets metric.
@@
 We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
-Conversely, removing 10\% of morphological characters (\hl{i.e. $N_{C}$ = 90}) genuinely removes 10\% of data in the matrix.
+Conversely, removing 10\% of morphological characters (i.e. $N_{C}$ = 90) genuinely removes 10\% of data in the matrix.
 Nonetheless, the effect of removing characters on the ability to recover the ``best'' tree topology is of the same order of magnitude as for the other two parameters (Fig. 4).
@@
 This can affect our ability to recover the ``best'' tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
-The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \hl{(}\citealt{bremer1992phylogeny,patterson1993congruence}\hl{; see }\citealt{masters2002lack}\hl{ for a empirical example)}.
-\hl{The sizes of our data matrices were constrained by the performance of our protocol}: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
+The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices (\citealt{bremer1992phylogeny,patterson1993congruence}; see \citealt{masters2002lack} for an empirical example).
+The sizes of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
 Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.
@@
 As expected, when combining the missing data parameters, our ability to recover the ``best'' tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) and lower when \hl{they have a larger proportion of missing data} (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) and lower when they have a larger proportion of missing data (i.e.
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $N_{C}$ = 25; Fig.
@@
 It is important, however, to notice that the effect of each parameter is not additive.
-\hl{Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
+Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
 For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference among trees with any combination of the other parameters ($M_F$ and $N_C$; Fig. 6).
@@
 A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Fig. 6).
-This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).}
+This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).
 
@@
 Variation in our ability to recover the ``best'' tree topology depends heavily on the tree inference method (Fig. 4 and Fig. 5).
-For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over \hl{parsimony} methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
+For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over parsimony methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
 This is, however, the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best'' tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-\hl{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details).}
+Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details).
 As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
@@
 
-\hl{The Bayesian consensus trees, however, perform poorly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (Fig.
+The Bayesian consensus trees, however, perform poorly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (Fig.
 5).
@@
 In such cases, the fossil taxa are equally likely to be placed in any of the clades that they share the most characters with.
-These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters }\citep{wiensmissing2003}.
-\hl{It is important to note, however, that this effect can be reduced by increasing the number of characters }\citep{wiensmissing2003}.
+These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters \citep{wiensmissing2003}.
+It is important to note, however, that this effect can be reduced by increasing the number of characters \citep{wiensmissing2003}.
 
@@
 This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best'' trees involved 1000 random pairwise comparisons rather than just one.
-\hl{Additionally, the Bayesian posterior trees performed more poorly than the Bayesian consensus tree (Fig. 4, Table 1 and Appendix C Fig. C5 and Tables C5, C6 and C7).
+Additionally, the Bayesian posterior trees performed more poorly than the Bayesian consensus tree (Fig. 4, Table 1 and Appendix C Fig. C5 and Tables C5, C6 and C7).
 This may be because the Bayesian posterior trees are always resolved and thus more likely to contain incorrectly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
 Conversely, the Bayesian consensus trees might not resolve nodes that are poorly supported and thus are more likely to contain only correctly resolved nodes (i.e.
-increasing the Normalised Robinson-Foulds metric).}
+increasing the Normalised Robinson-Foulds metric).
 
@@
 
-\hl{Another practical implication of our results regards the tree inference methods.
+Another practical implication of our results regards the tree inference methods.
 Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advise that where a topological constraint is needed, Bayesian consensus trees should be used.
-This may apply to tree inferences using the Total Evidence method such as tip-dating }\citep[\hl{e.g.}][]{ronquista2012,Wood01032013,BEASTmaster}.
-\hl{It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
-Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies }\citep[\hl{\textit{sensu}}][]{Swofford2001}.
-\hl{As shown in Fig.
+This may apply to tree inferences using the Total Evidence method such as tip-dating \citep[e.g.][]{ronquista2012,Wood01032013,BEASTmaster}.
+It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
+Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies \citep[\textit{sensu}][]{Swofford2001}.
+As shown in Fig.
 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve poorly supported nodes due to missing data than the Maximum Likelihood tree that is more likely to incorrectly resolve such nodes (i.e.
@@
 These trees will be invaluable for phylogenetic comparative analyses.
-For example a sub-sample of posterior tree distributions can be used to assess macroecological questions while better taking into account topological uncertainty }(\hl{e.g.
-}\citealt{fritzdiversity2013} and \citealt{jetzthe2012}\hl{ used in }\citealt{healy2014}).
+For example a sub-sample of posterior tree distributions can be used to assess macroecological questions while better taking into account topological uncertainty (e.g.
+\citealt{fritzdiversity2013} and \citealt{jetzthe2012} used in \citealt{healy2014}).
 
 \section{Conclusions}
-\hl{Previous studies have explored the effect of missing morphological data in Total Evidence matrices }\citep{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{The conclusions of theses studies, however, were limited by their empirical approach making their results applicable only to similar missing data scenarios.
-Additionally, these studies focused either only on living taxa }\citep{Wiens01102005}\hl{ or on the patterns of missing data from the fossil record only }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
-\hl{Here we instead used an approach where missing data were generated from simulated data and according to three clearly defined missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$) that removed data from both the living and fossil taxa.
-This allowed us to confirm previous results that missing data can be especially problematic in small matrices }\citep{wiensmissing2003}\hl{, but also revealed the crucial importance of coding morphological data for living species in Total Evidence phylogenies.
+Previous studies have explored the effect of missing morphological data in Total Evidence matrices \citep{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
+The conclusions of theses studies, however, were limited by their empirical approach making their results applicable only to similar missing data scenarios.
+Additionally, these studies focused either only on living taxa \citep{Wiens01102005} or on the patterns of missing data from the fossil record only \citep{manosphylogeny2007,pattinsonphylogeny2014}.
+Here we instead used an approach where missing data were generated from simulated data and according to three clearly defined missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$) that removed data from both the living and fossil taxa.
+This allowed us to confirm previous results that missing data can be especially problematic in small matrices \citep{wiensmissing2003}, but also revealed the crucial importance of coding morphological data for living species in Total Evidence phylogenies.
 Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
@@
 at least 50\%) if the goal is to accurately combine both living and fossil species in phylogenies.
-Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data.}
+Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data.
 
-\hl{Additionally, the topologies of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
-This has also been observed in empirical data }\citep[\hl{e.g.}][]{Arcila2015131}\hl{ where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
-This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds }\citep[\hl{e.g.}][]{slaterphylogenetic2013}\hl{, taxonomic constraints }\citep[\hl{e.g.}][]{slaterphylogenetic2013,beckancient2014}\hl{ or even on fixing the topology }\citep[\hl{e.g.}][]{ronquista2012}.
-\hl{Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.}
+Additionally, the topologies of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+This has also been observed in empirical data \citep[e.g.][]{Arcila2015131} where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds \citep[e.g.][]{slaterphylogenetic2013}, taxonomic constraints \citep[e.g.][]{slaterphylogenetic2013,beckancient2014} or even on fixing the topology \citep[e.g.][]{ronquista2012}.
+Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.
 
@@
 \noindent
-\textbf{Figure 4:} The effects of increasing missing data on topological recovery using \hl{Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)}.
+\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange).
 The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $N_{C}$) is shown on the x axis.
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.\\
 
@@
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.\\
 
@@
 \noindent
-\textbf{Figure 6:} \hl{The effects of missing data on topological recovery using Bayesian consensus trees.
+\textbf{Figure 6:} The effects of missing data on topological recovery using Bayesian consensus trees.
 Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $N_{C}$ (lower line).
@@
 Conversely, orange regions denote a high overlap in Normalised metric values between the different parameter combinations (i.e.
-the parameters have a weak effect on the metric and thus the topological recovery).}
+the parameters have a weak effect on the metric and thus the topological recovery).\\
 
@@
 \begin{landscape}
-
-% NC: Do you think we should add to any of these that BC <0.5 = significantly different and BC > 95% is significantly similar? Up to you.
 
@@
 \noindent
-\textbf{Table 1:} \hl{Bhattacharyya Coefficients of the pairwise method comparisons.}
-\hl{Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robinson-Foulds (RF) and Triplets (Tr) metrics.
+\textbf{Table 1:} Bhattacharyya Coefficients of the pairwise method comparisons.
+Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robinson-Foulds (RF) and Triplets (Tr) metrics.
 Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
 If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology.
-Values $>$ 0.95 denote significantly similar distributions and values $<$ 0.05 denote significantly different distributions.}\\
+Values $>$ 0.95 denote significantly similar distributions and values $<$ 0.05 denote significantly different distributions.\\
 

### 2015-08-24 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
few typos intro

@@
 Relationships among the taxa are then inferred using optimality criteria such as maximum parsimony \citep{Hennig1966,felsenstein2004}.
-This approach is commonly used by paleontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty.
+This approach is commonly used by palaeontologists but it ignores the additional molecular data available from living species and does not allow use of probabilistic methods for dealing with phylogenetic uncertainty.
 Neontologists, on the other hand, more commonly use probabilistic approaches (e.g. Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
@@
 A final approach, known as the Total Evidence method, uses matrices containing molecular data from living taxa and morphological data from both living and fossil taxa \citep{eernissetaxonomic1993}.
-This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny \citep[known as tip-dating;][]{ronquista2012}, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
+This approach treats every taxa as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny \citep[known as tip-dating;][]{ronquista2012}, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
 The Total Evidence method is becoming an increasingly popular way of adding fossil taxa to phylogenies \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014,Arcila2015131}.
@@
 The other two studies did include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices \citep{manosphylogeny2007,pattinsonphylogeny2014}.
-Data for living species are unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is a likely pattern}\citep{GuillermeCooperMissing}.
+Data for living species are unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is a likely pattern \citep{GuillermeCooperMissing}.
 Finally, these previous studies mainly focused on how missing data in fossil taxa affect the placement of fossils, ignoring the effects of missing data in living species \citep{manosphylogeny2007,pattinsonphylogeny2014}.
@@
 In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living taxa with morphological data, fossil data, all data for certain characters and the combination of these three aspects.
-This is an advance on previous studies because we use large-scale simulations and analyse the effects of three distinct aspects of missing data thus focusing on both neontological and paleontological parts of the matrix.
+This is an advance on previous studies because we use large-scale simulations and analyse the effects of three distinct aspects of missing data thus focusing on both neontological and palaeontological parts of the matrix.
 In addition, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades \citep[based on the Robinson-Foulds distance;][]{RF1981} and (ii) the displacement of wild-card taxa \citep[based on the Triplets distance;][]{critchlowthe1996} rather than just a single measure of clade conservation or clade support \citep[cf.][]{Wiens01102005,pattinsonphylogeny2014}.

### 2015-08-24 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
taxon -> taxa + aknowledgements

@@
 In practice, we first removed the data following the two missing data parameters $M_L$ and $M_F$ and then removed data following the $N_C$ parameters.
-To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon.
+To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxa.
 Note that the living taxa always had at least 90\% of data (the 1000 molecular characters).
@@
 This is related to the fact that the Bayesian consensus tree is built using a majority consensus rule.
-When the fossil taxa have less data (e.g. $M_{F}$ = 75\%) they will tend to branch with any taxon in the clade that shares most characters with the fossils.
+When the fossil taxa have less data (e.g. $M_{F}$ = 75\%) they will tend to branch with any taxa in the clade that shares most characters with the fossils.
 Therefore a majority consensus position is unlikely to exist (i.e.
@@
 Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster.
-Thanks to two anonymous reviewers for their useful and encouraging comments.
+Thanks to one anonymous reviewer and Alex Pyron for their useful and encouraging comments.
 Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland.

### 2015-08-24 Manuscript/MPESubmission/Revision/TEM_revised_manuscript2.tex
revisions second round

--- 2015-08-24 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
@@
 We infer the topology from the matrices using both Maximum Likelihood and Bayesian inference methods and measure the differences in topology using two different topological distance metrics as proxies for clade conservation and for wild-card taxa placement.
-We find that minimizing the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the ``best'' tree topology more so than minimizing the amount of missing data in the fossil record.
+We find that minimising the number of living taxa with no morphological data and the number of missing morphological characters improves the ability of Total Evidence methods to recover the ``best'' tree topology more so than minimising the amount of missing data in the fossil record.
 Additionally, we find that the ability of Total Evidence methods to recover the ``best'' tree topology is increased when using Bayesian methods.
@@
 We used these two metrics because they illustrate two different aspects of tree topology (see Discussion) but also because their performance in measuring differences in topology is well described \citep{Kuhner04112014} and well implemented \citep{Bogdanowicz2012}.
-We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalize our results for any \textit{n} number of taxa.
+We normalised both metrics using methods described in \citet{Bogdanowicz2012} to generalise our results for any \textit{n} number of taxa.
 These metrics are described in detail below.
@@
 Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
-Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Fig. 6).
+Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarised these results as a heat map of Bhattacharyya Coefficients (see Fig. 6).
 In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Fig.
@@
 This may also be because the ``best'' Bayesian consensus trees are not completely resolved, thus will always be more similar to the ``missing data" trees than a completely resolved tree like the ``best'' Maximum Likelihood tree.
-Nonetheless, we minimized the probability of unresolved ``best'' trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
+Nonetheless, we minimised the probability of unresolved ``best'' trees in our Bayesian analyses by only using datasets with strong phylogenetic signal (see Methods).
 
@@
 The topological recovery is measured as (A) the Normalised Robinson-Foulds metric and (B) the Normalised Triplets metric calculated using the Bhattacharyya Coefficient.
-The Bhattacharyya Coefficient values are indicated using a color gradient ranging from low probability of overlap in blue, to high probability of overlap in orange.
+The Bhattacharyya Coefficient values are indicated using a colour gradient ranging from low probability of overlap in blue, to high probability of overlap in orange.
 Blue regions denote a poor overlap in Normalised metric values between the different parameter combinations (i.e.
@@
 \textbf{Table 1:} Bhattacharyya Coefficients of the pairwise method comparisons.
-Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robinson-Foulds (RF) and Triplets (Tr) metrics.
+Each line summarises the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robinson-Foulds (RF) and Triplets (Tr) metrics.
 Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.

### 2015-09-03 Manuscript/CoverLetter.tex
Cleaned up manuscript folder

@@
 
-Our paper, entitled "Effects of missing data on topological inference using a Total Evidence approach", is to our knowledge, the first to thoroughly analyze the effects of missing data on tree topology in a Total Evidence framework.
-Using simulations ($>$ 150 CPU years worth), we find that the number of living taxa with morphological data and the overall number of morphological characters, are more important than the amount of missing data in the fossil record for recovering the "best" tree topology.
+Our research article, entitled ``Effects of missing data on topological inference using a Total Evidence approach'', is to our knowledge, the first to thoroughly analyze the effects of missing data on tree topology in a Total Evidence framework.
+Using simulations ($>$ 150 CPU years worth), we find that the number of living taxa with morphological data and the overall number of morphological characters, are more important than the amount of missing data in the fossil record for recovering the ``best'' tree topology.
 Additionally, we show that Bayesian methods outperform Maximum Likelihood methods, regardless of the amount of missing data.
@@
 This has major implications for clades where detailed morphological data collection from living species is rare.
-Additionally, we recommend using a Bayesian majority consensus tree when fixing tree topology for any additional analyses such as tip-dating (Ronquist et al.
-2012 Syst Biol).
-
-We feel this paper will be of particular interest to the readers of Systematic Biology as many papers investigating methods for combining living and fossil taxa and the effects of missing data on such analyses (e.g.
-D\'{a}valos et al.
-2014; Pattinson et al.
-in press, Sylvestro et al.
-2014, Pyron 2011, Ronquist et al.
-2012), have been recently published in the journal.
+Additionally, we recommend using a Bayesian majority consensus tree when fixing tree topology for any additional analyses.
 

### 2015-09-03 Manuscript/TEM_revised_manuscrip_highlithed.tex
Cleaned up manuscript folder

--- 2015-08-24 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
@@
 \section{Introduction}
-Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, many large-scale macroevolutionary studies focus solely on living species (e.g.
+Although most species that have ever lived are now extinct \citep{novacek1992ext,raup1993extinction}, \hl{many large-scale macroevolutionary studies focus solely on living species} (e.g.
 \citealp{meredithimpacts2011,jetzthe2012}).
@@
 Up to now, three main approaches have been used to place both living and fossil taxa into phylogenies.
-These approaches differ mainly in how they treat fossil taxa and their data.
-One can use fossils as tips or as nodes in the phylogeny, and can use only the age of the fossils, only the morphology of the fossils, or age and morphology jointly.
+\hl{These approaches differ mainly in how they treat fossil taxa and their data.
+One can use fossils as tips or as nodes in the phylogeny, and can use only the age of the fossils, only the morphology of the fossils, or age and morphology jointly.}
 Classical cladistic methods use matrices containing morphological data from both living and fossil taxa and treat each taxon as a tip in the phylogeny.
@@
 Neontologists, on the other hand, more commonly use probabilistic approaches (e.g. Maximum Likelihood or Bayesian methods) based on matrices containing only molecular data from living species.
-Because fossil taxa do not usually have available DNA, only fossil occurrence dates are used to time calibrate phylogenies \citep{zuckerkandl1965}.
+\hl{Because fossil taxa do not usually have available DNA, only fossil occurrence dates are used to time calibrate phylogenies }\citep{zuckerkandl1965}.
 There have been great improvements in the theory and application of these two approaches \citep[e.g.][]{bapsta2013,stadlerdating2013,heaththe2013} as well as much debate about the ``best'' approach to use \citep[e.g.][]{spencerefficacy2013,wrightbayesian2014}.
@@
 This approach treats every taxon as a tip in the phylogeny, uses the occurrence age of the fossils to time calibrate the phylogeny \citep[known as tip-dating;][]{ronquista2012}, and allows the use of probabilistic methods for estimating phylogenetic uncertainty \citep{ronquista2012}.
-The Total Evidence method is becoming an increasingly popular way of adding fossil taxa to phylogenies \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014,Arcila2015131}.
+\hl{The Total Evidence method is becoming an increasingly popular way of adding fossil taxa to phylogenies} \citep[e.g.][]{pyrondivergence2011,ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014,Arcila2015131}.
 Although the Total Evidence approach seems very promising, there is one big drawback in using this approach: it requires both molecular and morphological data, both of which can be difficult (or impossible) to collect for every living and fossil taxon in the tree.
@@
 \citealp{O'Leary08022013} \textit{vs.} \citealp{meredithimpacts2011}), and for fossil taxa, data can only be collected from features preserved in the fossil record.
-For example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data are (nearly) always unavailable.
-Therefore Total Evidence matrices are likely to contain a large proportion of missing data that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
+For example, in vertebrates, the hardest parts of the skeleton are more often preserved than soft parts \citep{sansomfossilization2013}; and molecular data \hl{are} (nearly) always unavailable.
+Therefore \hl{Total Evidence matrices are likely to contain a large proportions of missing data} that may affect the method's ability to infer correct topologies, branch lengths and support values \citep{salamin2003}.
 
-Although missing data do not appear be a major problem in molecular and morphological matrices separately \citep[as long as enough data overlap in each case, and missing data are not phylogenetically biased;][]{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
+Although missing data do not appear be a major problem in molecular and morphological matrices separately \citep[\hl{as long as enough data overlaps in each case, and missing data are not phylogeneticaly biased;}][]{wiensmissing2003,Wiens01102005,wiensmissing2006,wiensmissing2008,lemmonthe2009,Sanderson22072011,rouresite-specific2011,pattinsonphylogeny2014}, it may become more of an issue in Total Evidence matrices containing both molecular and morphological data for living and fossil taxa.
 This may be particularly problematic as fossil taxa (generally) do not have molecular data, resulting in a large section of missing data in Total Evidence matrices.
-Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework \citep[i.e.
-using both molecular and morphological data;][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
-These previous studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data \citep{Wiens01102005}; or (2) removing both molecular and some morphological data from living taxa to create artificial fossils \citep{manosphylogeny2007,pattinsonphylogeny2014}.
-Both approaches have shown that missing data are not a major problem and should not be an obstacle to combining both living and fossil species in the same phylogenies.
+\hl{Until now, few attempts have been made to study the impact of this missing data issue on phylogenetic inference in a Total Evidence framework }\citep[\hl{i.e.
+using both molecular and morphological data;}][]{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{These previous studies assessed the effect of missing data on topology by either (1) comparing a dataset with missing data to subsets without missing data }\citep{Wiens01102005}\hl{; or (2) removing both molecular and some morphological data from living taxa to create artificial fossils }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{Both approaches have shown that missing data are not a major problem and should not be an obstacle to combining both living and fossil species in the same phylogenies.
 The way these studies were conducted, however, means that their conclusions are not generally applicable across all scenarios involving missing data in Total Evidence phylogenies.
 For example, using an empirical (rather than simulation based) approach limits their conclusions to studies with similar distributions of data across species in the phylogeny.
-Additionally, one of the three previous studies did not include fossil taxa in their analyses, so their results cannot be used to make conclusions about how missing data may influence the placement of fossils \citep{wiensmissing2003}.
-The other two studies did include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices \citep{manosphylogeny2007,pattinsonphylogeny2014}.
+Additionally, }\cite{Wiens01102005}\hl{ did not include fossil taxa in their analyses, so their results cannot be used to make conclusions about how missing data may influence the placement of fossils.}
+\cite{manosphylogeny2007,pattinsonphylogeny2014}\hl{ include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices. 
 Data for living species are unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is a likely pattern}\citep{GuillermeCooperMissing}.
-Finally, these previous studies mainly focused on how missing data in fossil taxa affect the placement of fossils, ignoring the effects of missing data in living species \citep{manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{Finally, }\cite{manosphylogeny2007}\hl{ and }\cite{pattinsonphylogeny2014}\hl{ mainly focused on how missing data in fossil taxa affect the placement of fossils, ignoring the effects of missing data in living species.}
 
-In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living taxa with morphological data, fossil data, all data for certain characters and the combination of these three aspects.
+\hl{In this study, we propose a theoretical assessment of the effect of missing data in the Total Evidence method by removing living taxa with morphological data, fossil data, all data for certain characters and the combination of these three aspects.
 This is an advance on previous studies because we use large-scale simulations and analyse the effects of three distinct aspects of missing data thus focusing on both neontological and paleontological parts of the matrix.
-In addition, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades \citep[based on the Robinson-Foulds distance;][]{RF1981} and (ii) the displacement of wild-card taxa \citep[based on the Triplets distance;][]{critchlowthe1996} rather than just a single measure of clade conservation or clade support \citep[cf.][]{Wiens01102005,pattinsonphylogeny2014}.
+In addition, we test the effect of missing data by measuring two crucial aspects of topology in both Maximum Likelihood and Bayesian phylogenies: (i) the conservation of clades }\citep[\hl{based on the Robinson-Foulds distance;}][]{RF1981}\hl{ and (ii) the displacement of wild-card taxa }\citep[\hl{based on the Triplets distance;}][]{critchlowthe1996}\hl{ rather than just a single measure of clade conservation or clade support }\citep[\hl{cf.}][]{Wiens01102005,pattinsonphylogeny2014}.
 
@@
 Although branch length estimation is also important \citep[namely for timing extinction and/or speciation events; e.g.][]{ronquista2012}, we do not consider branch lengths in this study.
-This is partially due to difficulties with simulating branch lengths and topology simultaneously, but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints \citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
+This is partially due to difficulties with simulating branch lengths and topology simultaneously, \hl{but also because previous studies have already empirically assessed the effect of the Total Evidence method on branch length variation but using topological constraints }\citep{ronquista2012,schragocombining2013,slaterphylogenetic2013,beckancient2014}.
 Thus understanding the sensitivity of topology to missing data is important for assessing the accuracy of tree estimation in the Total Evidence framework.
@@
 \section{Materials and Methods}
-To explore how missing data in the morphological partition of Total Evidence matrices influences tree topology, we used the following protocol (Fig. 1):
+To explore how missing data in \hl{the morphological partition of Total Evidence matrices} influences tree topology, \hl{we used the following protocol (Fig. 1):}
 \begin{enumerate}
@@
 Next, we generated a molecular and a morphological matrix from the ``true'' tree.
-The molecular matrix was simulated from the ``true'' tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
+The molecular matrix was \hl{simulated} from the ``true'' tree using the R package phyclust v. 0.1-14 \citep{chen2011}.
 The matrix contained 1000 character sites for 51 taxa and was generated using the seqgen algorithm \citep{ranbaut1997seqgen} and using the HKY model \citep{HKY85} with random base frequencies (sampled from a uniform probability distribution bounded between 0 and 1 with the total frequency for the four bases equal to 1) and transition/transversion rate of two \citep{douadycomparison2003}.
-The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5 \citep{yangamong-site1996}.
-In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal \citep{Hassanin1998611,EstoupHomoplasy}.
-Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below).
-This model and these parameter settings strike a balance between realism for empirical datasets \citep[e.g.][]{douadycomparison2003,kellymolecular2014} and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).
+\hl{The substitution rates were selected from a gamma distribution with an ($\alpha$) shape of 0.5} \citep{yangamong-site1996}.
+\hl{In practice, a value of $\alpha$ $<$ 1 decreases the number of sites with high substitution rates, thus reducing homoplasic sites and increasing the phylogenetic signal }\citep{Hassanin1998611,EstoupHomoplasy}.
+\hl{Also, we chose this $\alpha$ value to be consistent with our protocol for simulating morphological characters (see below).}
+\hl{This model and these parameter settings strike a balance between realism for empirical datasets }\citep[\hl{e.g.}][]{douadycomparison2003,kellymolecular2014}\hl{ and parameter richness with more complex models (e.g., GTR, multiple partitions with independent models), making them more suitable for our computational limitations (even with the parameters defined, the total computational time for the whole analysis was around 150 CPU years).}
 All the molecular information for fossil taxa was replaced by missing data ("?").
 
-We simulated the morphological matrix using the rTraitDisc function from the R package ape v. 3.0-11 \citep{paradisape:2004} to generate a matrix of 100 character sites for 51 taxa.
-We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A and Fig.
-A1 within).
+\hl{We simulated the morphological matrix using the rTraitDisc function from the R package ape v. 3.0-11} \citep{paradisape:2004}\hl{ to generate a matrix of 100 character sites for 51 taxa.}
+\hl{We assigned the number of character states (either two or three) for each morphological character by sampling with a probability of 0.85 for two states characters and 0.15 for three state characters (based on an empirical review of published matrices, see Appendix A and Fig.
+A1 within).}
 We then ran an independent discrete character simulation for each character using the ``true'' tree with the character's randomly selected number of states (two or three) and assuming an equal rate of change (i.e.
@@
 For each character, the evolutionary rate was sampled from a gamma distribution with $\alpha$ = 0.5.
-We used low evolutionary rate parameters to be consistent with the molecular rate parameters, to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal \citep{wrightbayesian2014}.
-Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model \citep{lewisa2001,wrightbayesian2014}.
-Note, however, that \cite{wrightbayesian2014} have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal \citep[see Estimating phylogenies section below;][]{zanderminimal2004} so this should not influence our results.
+\hl{We used low evolutionary rate parameters to be consistent with the molecular rate parameters, to avoid homoplasy in the morphological part of the matrix and create a clear phylogenetic signal }\citep{wrightbayesian2014}.
+\hl{Topological error has been shown to be minimal at a morphological rate of 0.5 when using the M\textit{kv} model }\citep{lewisa2001,wrightbayesian2014}.
+\hl{Note, however, that } \cite{wrightbayesian2014}\hl{ have shown that low morphological rates ($<$ 0.5) increase variance in topological error, but we discarded simulations with such topological error by selecting only matrices with a ``fair'' phylogenetic signal }\citep[\hl{see Estimating phylogenies section below;}][]{zanderminimal2004}\hl{ so this should not influence our results.}
 
@@
 \subsection{Removing data}
-To explore the effect of missing morphological data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
+\hl{To explore the effect of missing morphological data on topological recovery, we removed various amounts of the ``complete'' matrix to obtain matrices with missing morphological data.
 Hereafter, we call these matrices with missing morphological data the ``missing-data'' matrices.
 Note that the amount of molecular data remained constant throughout our simulations: 1000 molecular characters for living taxa and no molecular data for fossil taxa (see above).
-We removed morphological data using three data incompleteness parameters:
+We removed morphological data using three data incompleteness parameters:}
 \begin{enumerate}
-\item{The proportion of missing living taxa ($M_L$).}
-This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
+\item{\hl{The proportion of missing living taxa ($M_L$).}}
+\hl{This first missing-data parameter corresponds to the proportion of living taxa with no morphological data.
 It represents the number of living taxa that are present in the matrix but have only molecular data available.
-This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species are rarely collected \citep{GuillermeCooperMissing}.
-Therefore, many living species will have only molecular data available.
-In practice, we removed all the morphological data from randomly chosen living taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.
-\item{The proportion of missing data in the fossil record ($M_F$).}
-This missing data parameter represents the completeness of the fossil record.
-Due to preservation biases, missing data for fossil taxa are common \citep{sansomfossilization2013}.
-In practice, we randomly removed a proportion of data from across the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
+This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species are rarely collected }\citep{GuillermeCooperMissing}.
+\hl{Therefore, many living species will have only molecular data available.
+In practice, we removed all the morphological data from randomly chosen living taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.}
+\item{\hl{The proportion of missing data in the fossil record ($M_F$).}}
+\hl{This missing data parameter represents the completeness of the fossil record.
+Due to preservation biases, missing data for fossil taxa are common }\citep{sansomfossilization2013}.
+\hl{In practice, we randomly removed a proportion of data from across the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
 Note that 50\% missing data for fossil taxa does not mean that each fossil is missing 50\% of its morphological data.
 Instead this 50\% refers to missing fossil data across the whole matrix.
-Some fossils may retain 100\% of their data and others may lose most of their data at this parameter value (down to a minimum threshold of 5\% available data; see below).
-\item{The number of morphological characters for both living and fossil taxa ($N_C$).}
-This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
+Some fossils may retain 100\% of their data and others may lose most of their data at this parameter value (down to a minimum threshold of 5\% available data; see below).}
+\item{\hl{The number of morphological characters for both living and fossil taxa ($N_C$).}}
+\hl{This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix.
 Any morphological matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large.
@@
 In practice, we randomly removed entire characters from the morphological matrix reducing it to: 100, 90, 75, 50 or 25 characters.
-Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).
+Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of ``missing'' morphological characters).}
 \end{enumerate}
 
-Each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
-Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.
+\hl{Each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data.
+Note that $M_L$ and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.}
 
-We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data'' matrices.
+\hl{We created matrices using all parameter combinations resulting in 125 ($5^3$) ``missing-data'' matrices.
 Note that one of these combinations ($M_L$=0\%; $M_F$=0\% and $N_C$=100) has no missing data so is equivalent to the ``complete'' matrix, thus we have one effectively complete matrix in our 125 ``missing-data'' matrices.
@@
 In real world situations, the ``true'' tree is never available to us because we cannot know the true evolutionary history of a clade \citep[except in very rare circumstances, e.g.][]{rozen2005}.
-We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
+\hl{We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
 In practice, the difference between the ``best'' trees and the ``missing data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing data'' trees.
@@
 The results of the comparisons of the ``true'' tree and the ``missing data'' trees are available in Appendix B.
-Note that this makes little difference to our overall results.
+Note that this makes little difference to our overall results.}
 
@@
 For each estimation, we used the ``true'' tree's topology as a starting tree (with a starting value for each branch length of one).
-We used a fixed starting tree rather than a random starting tree \citep[default MrBayes;][]{Ronquist2012mrbayes} to speed up our Bayesian inferences.
-Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').
+\hl{We used a fixed starting tree rather than a random starting tree }\citep[\hl{default MrBayes;}][]{Ronquist2012mrbayes} \hl{to speed up our Bayesian inferences.
+Note that a starting tree is not a Bayesian prior on topology \textit{per se} and using a fixed starting tree did not significantly affect topology compared to using random starting trees (see Appendix A, section ``Effect of the starting tree on Bayesian inference'').}
 We also used two priors on the molecular part of the matrix: an exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5, and a transition/transversion ratio prior of two sampled from a strong beta distribution ($\beta$(80,40)); and one prior on the morphological part of the matrix (exponential prior on the shape of the gamma distribution of $\alpha$ = 0.5).
@@
 Even using these priors, it took $~$ 140 CPU years to build 50 sets of 125 Bayesian trees (2.30GHz clock speed nodes).
-The detailed MrBayes parameters are available in Appendix A.
+\hl{The detailed MrBayes parameters are available in Appendix A.
 We also included an analysis showing the effect of missing data on the estimation of the shape parameter ($\alpha$) of the morphological substitution rate distribution.
-This extra analysis, however, is beyond the scope of this paper so the results are not discussed further here.
+This extra analysis, however, is beyond the scope of this paper so the results are not discussed further here.}
 
@@
 \subsubsection{Robinson-Foulds distance}
-The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and twice the number of shared clades across two trees.
+The Robinson-Foulds distance \citep{RF1981}, or ``path difference", measures the difference between the number of clades and \hl{twice the number} of shared clades across two trees.
 The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} (see Appendix B for calculation details).
@@
 We measured similarities among the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) using the Bhattacharyya Coefficient \citep{Bhattacharyya}.
-The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 \citep[full overlap;][see Appendix B for calculation details]{Bhattacharyya}.
+\hl{The Bhattacharyya Coefficient is the probability of overlap between two distributions bounded between 0 (no overlap) and 1 }\citep[\hl{full overlap;}][\hl{see Appendix B for calculation details}]{Bhattacharyya}.
 Note that this is comparable to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing whole distributions not just their means.
-When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
+\hl{When the Bhattacharyya Coefficient between two distributions is $<$0.05, the distributions are significantly different.
 When this coefficient is $>$0.95, the distributions are significantly similar.
-Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.
-To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
+Values between these two thresholds show the probability of overlap between the distributions but do not allow us to define the significance of the similarity or differences between distributions.}
+\hl{To assess the effect of our missing data parameters, we calculated the Bhattacharyya Coefficient between the distributions of the different metrics scores (Normalised Robinson-Foulds and Normalised Triplets metric) for each pairwise combination of missing data parameters ($M_{L}$, $M_{F}$, $N_{C}$) and parameter states (0\%, 10\%, 25\%, 50\%, 75\% and 100, 90, 75, 50, 25 characters), i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100; $M_{L}$ = 10\%, $M_{F}$ = 0\%, $N_{C}$ = 100 etc.
 (see Fig.
-2 for more details).
+2 for more details).}
 This resulted in 7875 pairwise comparisons (a triangular matrix with $3^5$$\times$$3^5$ cells).
@@
 Conversely, if the two combinations of missing data parameters differ, the Bhattacharyya Coefficient will be close to zero.
-Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Fig. 6).
+\hl{Because of the difficulties in representing so many pairwise comparisons in a meaningful way, we summarized these results as a heat map of Bhattacharyya Coefficients (see Fig. 6).
 In this type of figure, parameters that have similar effects on recovering the ``best'' topology (either positive or negative effects) will be denoted by similar colour patches in the heat map representation of these comparisons (see Fig.
-6).
+6).}
 
@@
 
-Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
+\hl{Using both Normalised Robinson-Foulds and Normalised Triplets metrics from the Bayesian consensus trees, the parameter combination with no missing data (i.e.
 $M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) is always the most dissimilar to all the other parameter combinations (thin deep blue line at the base of Fig.
@@
 6A).
-In Fig. 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.
+In Fig. 5 for the Normalised Robinson-Foulds metric, this can be interpreted as the overlap between the distributions once $M_L$=50\%.}
 
@@
 Normalised Robinson-Foulds metric) is still relatively good (mode = 0.72) when the proportion of missing data are high ($M_{F}$ = 75\%).
-These results are in agreement with \cite{manosphylogeny2007} where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.
+\hl{These results are in agreement with }\cite{manosphylogeny2007}\hl{ where as few as 16 characters were sufficient for correctly assigning artificial fossils to their correct clade.}
 
@@
 every branching position is represented in $<$ 50\% of the trees in the Bayesian posterior distribution) and the fossil taxa will form a polytomy at the base of the clade.
-In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.
+\hl{In this case, the Normalised Robinson-Foulds metric will decrease when the fossil is present near the tips but affects the clade conservation less when fossils are near the root.}
 Conversely, because a fossil in a high taxonomic level clade has many chances to branch on different nodes within the clade, it will be more likely to act as a wildcard taxon and decrease the Normalised Triplets metric.
@@
 We expected the decrease in this parameter to have an effect twice as large as that for the $M_{L}$ and $M_{F}$ parameters, because removing 10\% of the data for the fossil or living taxa only removes 5\% of data from the whole matrix (because this parameter affects only half of the taxa present in the matrix).
-Conversely, removing 10\% of morphological characters (i.e. $N_{C}$ = 90) genuinely removes 10\% of data in the matrix.
+Conversely, removing 10\% of morphological characters (\hl{i.e. $N_{C}$ = 90}) genuinely removes 10\% of data in the matrix.
 Nonetheless, the effect of removing characters on the ability to recover the ``best'' tree topology is of the same order of magnitude as for the other two parameters (Fig. 4).
@@
 This can affect our ability to recover the ``best'' tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy.
-The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices (\citealt{bremer1992phylogeny,patterson1993congruence}; see \citealt{masters2002lack} for an empirical example).
-The sizes of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
+The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \hl{(}\citealt{bremer1992phylogeny,patterson1993congruence}\hl{; see }\citealt{masters2002lack}\hl{ for a empirical example)}.
+\hl{The sizes of our data matrices were constrained by the performance of our protocol}: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters.
 Therefore, part of the decrease of the Normalised Robinson-Foulds metric and the Normalised Triplets metric in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.
@@
 As expected, when combining the missing data parameters, our ability to recover the ``best'' tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds metric and the Normalised Triplets metric are higher when all the missing data parameters have few missing data (i.e.
-$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) and lower when they have a larger proportion of missing data (i.e.
+$M_{L}$ = 0\%, $M_{F}$ = 0\%, $N_{C}$ = 100) and lower when \hl{they have a larger proportion of missing data} (i.e.
 $M_{L}$ = 75\%, $M_{F}$ = 75\% and $N_{C}$ = 25; Fig.
@@
 It is important, however, to notice that the effect of each parameter is not additive.
-Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
+\hl{Surprisingly, the number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($N_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$).
 For any additional missing living taxa with morphological data ($M_L$) beyond 50\%, there is no difference among trees with any combination of the other parameters ($M_F$ and $N_C$; Fig. 6).
@@
 A similar effect can be observed when the $N_C$ parameter reaches 50 characters (Fig. 6).
-This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).
+This has important practical implications, especially for the best strategy to improve topology by collecting more morphological data (see below).}
 
@@
 Variation in our ability to recover the ``best'' tree topology depends heavily on the tree inference method (Fig. 4 and Fig. 5).
-For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over parsimony methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
+For morphological data, previous studies have shown some superiority of probabilistic tree inference methods with simple evolutionary models such as the M\textit{kv} model \citep{lewisa2001} over \hl{parsimony} methods (\citealp{wrightbayesian2014}; but see \citealp{spencerefficacy2013}).
 This is, however, the first study, to our knowledge, to compare the performance of the M\textit{kv} model \citep{lewisa2001} for recovering the ``best'' tree topology using Maximum Likelihood and Bayesian methods in a Total Evidence framework.
-Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
-Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details).
+\hl{Our results show that the topology of the Bayesian consensus tree is always closer to the ``best'' tree topology than the ``best'' Maximum Likelihood tree (Fig. 5).
+Note that the methodological choice of using the ``true'' tree as a starting tree for the Bayesian Inference rather than a random starting tree (see Methods), had no significant effect on topological recovery (see Appendix A, section ``Effect of the starting tree on Bayesian inference'' for details).}
 As described above, this is because the Bayesian consensus tree allows a fossil taxon with few data to be placed with a higher confidence at a lower taxonomic level than the Maximum Likelihood tree.
@@
 
-The Bayesian consensus trees, however, perform poorly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (Fig.
+\hl{The Bayesian consensus trees, however, perform poorly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (Fig.
 5).
@@
 In such cases, the fossil taxa are equally likely to be placed in any of the clades that they share the most characters with.
-These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters \citep{wiensmissing2003}.
-It is important to note, however, that this effect can be reduced by increasing the number of characters \citep{wiensmissing2003}.
+These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters }\citep{wiensmissing2003}.
+\hl{It is important to note, however, that this effect can be reduced by increasing the number of characters }\citep{wiensmissing2003}.
 
@@
 This reflects the fact that it is difficult to compare two distributions of trees, and each comparison between a set of ``missing data" trees and a set of the ``best'' trees involved 1000 random pairwise comparisons rather than just one.
-Additionally, the Bayesian posterior trees performed more poorly than the Bayesian consensus tree (Fig. 4, Table 1 and Appendix C Fig. C5 and Tables C5, C6 and C7).
+\hl{Additionally, the Bayesian posterior trees performed more poorly than the Bayesian consensus tree (Fig. 4, Table 1 and Appendix C Fig. C5 and Tables C5, C6 and C7).
 This may be because the Bayesian posterior trees are always resolved and thus more likely to contain incorrectly resolved nodes (i.e. decreasing the Normalised Robinson-Foulds metric).
 Conversely, the Bayesian consensus trees might not resolve nodes that are poorly supported and thus are more likely to contain only correctly resolved nodes (i.e.
-increasing the Normalised Robinson-Foulds metric).
+increasing the Normalised Robinson-Foulds metric).}
 
@@
 
-Another practical implication of our results regards the tree inference methods.
+\hl{Another practical implication of our results regards the tree inference methods.
 Because the Bayesian consensus trees consistently recovered topologies closer to the ``best'' tree topology than the Maximum Likelihood trees, we advise that where a topological constraint is needed, Bayesian consensus trees should be used.
-This may apply to tree inferences using the Total Evidence method such as tip-dating \citep[e.g.][]{ronquista2012,Wood01032013,BEASTmaster}.
-It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
-Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies \citep[\textit{sensu}][]{Swofford2001}.
-As shown in Fig.
+This may apply to tree inferences using the Total Evidence method such as tip-dating }\citep[\hl{e.g.}][]{ronquista2012,Wood01032013,BEASTmaster}.
+\hl{It is, however, possible that including dating information during tree inference could also improve the accuracy of the Bayesian posterior tree distribution, so a fixed topology should be used with caution.
+Using the Bayesian consensus tree rather than the Maximum Likelihood can also reduce the number of false positive topologies }\citep[\hl{\textit{sensu}}][]{Swofford2001}.
+\hl{As shown in Fig.
 5 and discussed in the section above (Effects of tree inference methods), the Bayesian consensus tree is more likely to not resolve poorly supported nodes due to missing data than the Maximum Likelihood tree that is more likely to incorrectly resolve such nodes (i.e.
@@
 These trees will be invaluable for phylogenetic comparative analyses.
-For example a sub-sample of posterior tree distributions can be used to assess macroecological questions while better taking into account topological uncertainty (e.g.
-\citealt{fritzdiversity2013} and \citealt{jetzthe2012} used in \citealt{healy2014}).
+For example a sub-sample of posterior tree distributions can be used to assess macroecological questions while better taking into account topological uncertainty }(\hl{e.g.
+}\citealt{fritzdiversity2013} and \citealt{jetzthe2012}\hl{ used in }\citealt{healy2014}).
 
 \section{Conclusions}
-Previous studies have explored the effect of missing morphological data in Total Evidence matrices \citep{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
-The conclusions of theses studies, however, were limited by their empirical approach making their results applicable only to similar missing data scenarios.
-Additionally, these studies focused either only on living taxa \citep{Wiens01102005} or on the patterns of missing data from the fossil record only \citep{manosphylogeny2007,pattinsonphylogeny2014}.
-Here we instead used an approach where missing data were generated from simulated data and according to three clearly defined missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$) that removed data from both the living and fossil taxa.
-This allowed us to confirm previous results that missing data can be especially problematic in small matrices \citep{wiensmissing2003}, but also revealed the crucial importance of coding morphological data for living species in Total Evidence phylogenies.
+\hl{Previous studies have explored the effect of missing morphological data in Total Evidence matrices }\citep{Wiens01102005,manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{The conclusions of theses studies, however, were limited by their empirical approach making their results applicable only to similar missing data scenarios.
+Additionally, these studies focused either only on living taxa }\citep{Wiens01102005}\hl{ or on the patterns of missing data from the fossil record only }\citep{manosphylogeny2007,pattinsonphylogeny2014}.
+\hl{Here we instead used an approach where missing data were generated from simulated data and according to three clearly defined missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$) that removed data from both the living and fossil taxa.
+This allowed us to confirm previous results that missing data can be especially problematic in small matrices }\citep{wiensmissing2003}\hl{, but also revealed the crucial importance of coding morphological data for living species in Total Evidence phylogenies.
 Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
@@
 at least 50\%) if the goal is to accurately combine both living and fossil species in phylogenies.
-Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data.
+Doing so will increase overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their shared derived characters rather than simply on available data.}
 
-Additionally, the topologies of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
-This has also been observed in empirical data \citep[e.g.][]{Arcila2015131} where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
-This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds \citep[e.g.][]{slaterphylogenetic2013}, taxonomic constraints \citep[e.g.][]{slaterphylogenetic2013,beckancient2014} or even on fixing the topology \citep[e.g.][]{ronquista2012}.
-Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.
+\hl{Additionally, the topologies of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
+This has also been observed in empirical data }\citep[\hl{e.g.}][]{Arcila2015131}\hl{ where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
+This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds }\citep[\hl{e.g.}][]{slaterphylogenetic2013}\hl{, taxonomic constraints }\citep[\hl{e.g.}][]{slaterphylogenetic2013,beckancient2014}\hl{ or even on fixing the topology }\citep[\hl{e.g.}][]{ronquista2012}.
+\hl{Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.}
 
@@
 \noindent
-\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange).
+\textbf{Figure 4:} The effects of increasing missing data on topological recovery using \hl{Maximum Likelihood trees (black), Bayesian consensus trees (grey), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)}.
 The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $N_{C}$) is shown on the x axis.
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.\\
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
 
@@
 Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row).
-The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.\\
+The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
 
@@
 \noindent
-\textbf{Figure 6:} The effects of missing data on topological recovery using Bayesian consensus trees.
+\textbf{Figure 6:} \hl{The effects of missing data on topological recovery using Bayesian consensus trees.
 Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $N_{C}$ (lower line).
@@
 Conversely, orange regions denote a high overlap in Normalised metric values between the different parameter combinations (i.e.
-the parameters have a weak effect on the metric and thus the topological recovery).\\
+the parameters have a weak effect on the metric and thus the topological recovery).}
 
@@
 \begin{landscape}
+
+% NC: Do you think we should add to any of these that BC <0.5 = significantly different and BC > 95% is significantly similar? Up to you.
 
@@
 \noindent
-\textbf{Table 1:} Bhattacharyya Coefficients of the pairwise method comparisons.
-Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robinson-Foulds (RF) and Triplets (Tr) metrics.
+\textbf{Table 1:} \hl{Bhattacharyya Coefficients of the pairwise method comparisons.}
+\hl{Each line summarizes the probabilities of overlap between the distributions of the ``best'' tree versus trees from each inference method (Maximum Likelihood; Bayesian consensus; Maximum Likelihood Bootstraps and Bayesian posterior trees) pooled across all combinations of missing data parameter values, using the Normalised Robinson-Foulds (RF) and Triplets (Tr) metrics.
 Values highlighted in bold are the extreme values of high or low probability of overlap between two methods.
 If two methods have a high probability of overlap, they have a similar ability to recover the ``correct'' tree topology.
-Values $>$ 0.95 denote significantly similar distributions and values $<$ 0.05 denote significantly different distributions.\\
+Values $>$ 0.95 denote significantly similar distributions and values $<$ 0.05 denote significantly different distributions.}\\
 

### 2015-09-03 Manuscript/response2.tex
Cleaned up manuscript folder

\documentclass[12pt,letterpaper]{article}

%Packages
\usepackage{pdflscape}
\usepackage{fixltx2e}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{float}
\usepackage{latexsym}
\usepackage{url}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{array}
\usepackage[version=3]{mhchem}
\usepackage{ifthen}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsthm}
\usepackage{amstext}
\usepackage{enumerate}
\usepackage[osf]{mathpazo}
\usepackage{dcolumn}
\usepackage{lineno}
\usepackage{color}
\usepackage[usenames,dvipsnames]{xcolor}
\pagenumbering{arabic}

%Pagination style and stuff
%\linespread{2}

\raggedright
\setlength{\parindent}{0.5in}
\setcounter{secnumdepth}{0} 
\renewcommand{\section}[1]{%
\bigskip
\begin{center}
\begin{Large}
\normalfont\scshape #1
\medskip
\end{Large}
\end{center}}
\renewcommand{\subsection}[1]{%
\bigskip
\begin{center}
\begin{large}
\normalfont\itshape #1
\end{large}
\end{center}}
\renewcommand{\subsubsection}[1]{%
\vspace{2ex}
\noindent
\textit{#1.}---}
\renewcommand{\tableofcontents}{}

\setlength\parindent{0pt}

\begin{document}

\textbf{RE: MPE-15-254}\\
\bigskip
Dear Dr D\'{a}valos,\\
\bigskip
We thank both reviewers for their additional comments.
We dealt with the two minor comments of reviewer 1.

\section{Reviewer 1:}

\begin{enumerate}
% -----------------
% Comment 1
% -----------------
\item{\textcolor{blue}{Some typos persist, e.g., line 76: "...likely to contain a large proportions of missing data"; line 79: "(as long as enough data overlaps)", line 80: "...phylogeneticaly"}}

We fixed these typos and checked thoroughly for other potential typos throughout the manuscript.
% -----------------
% Comment 2
% -----------------
\item{\textcolor{blue}{Somewhat choppy style on line 101. In general it would be a bit more stream-lined and flow better if the style of citing references were more like this: "Some previous research on the effects of missing data did not include fossils (e.g., Wiens 2003). Others used the patchy distribution of missing data in fossils from empirical datasets to create artificial fossils from extant species (Manos, Pattinson)" rather than "Wiens (2003) did not include fossils. Manos (2007) and Pattinson (2015) used the patchy distribution of missingness in fossils to create artificial fossils from extant species". Make each sentence a statement summarizing the results of a study and cite the study at the end, rather than make the statement about the specific study and what they found. This is how the authors usually write, but in this section of literature review, it gets choppy.}}

We rewrote this paragraph as follows:

\textit{Additionally, one of the three previous studies did not include fossil taxa in their analyses, so their results cannot be used to make conclusions about how missing data may influence the placement of fossils (Wiens, 2003). The other two studies did include fossil taxa, but used the patchiness of the fossil record to determine how to remove data from their matrices (Manos et al., 2007; Pattinson et al., 2014). Data for living species are unlikely to be missing in this patchy way, instead full molecular data with the complete absence of morphological data is a likely pattern (Guillerme and Cooper, 2015).
Finally, these previous studies mainly focused on how missing data in fossil taxa affect the placement of fossils, ignoring the effects of missing data in living species (Manos et al., 2007; Pattinson et al., 2014).}. lines 99-108.

\end{enumerate}

Please let us know if you require any further information,\\
\bigskip

Thomas Guillerme (guillert@tcd.ie)\\
Natalie Cooper

\end{document}
